### 2020-01-21 ICLM/miaoyan/ordinalT.tex
upload draft

%%%%%%%% ICML 2020 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%

\documentclass{article}

% Recommended, but optional, packages for figures and better typesetting:
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{booktabs} % for professional tables

% hyperref makes hyperlinks in the resulting PDF.
% If your build breaks (sometimes temporarily if a hyperlink spans a page)
% please comment out the following usepackage line and replace
% \usepackage{icml2020} with \usepackage[nohyperref]{icml2020} above.
\usepackage{hyperref}

% Attempt to make hyperref and algorithmic work together better:
\newcommand{\theHalgorithm}{\arabic{algorithm}}

% Use the following line for the initial blind version submitted for review:
\usepackage{icml2020}

% If accepted, instead use the following line for the camera-ready submission:
%\usepackage[accepted]{icml2020}

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Ordinal tensor denoising and completion}

\usepackage{multirow}
\usepackage{graphicx}
%\usepackage[utf8]{inputenc} % allow utf-8 input
%\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
%\usepackage{booktabs}       % professional-quality tables
\usepackage{amsmath,amssymb} 
\usepackage{amsthm}    % blackboard math symbols
%\usepackage{nicefrac}       % compact symbols for 1/2, etc.
%\usepackage{microtype}      % microtypography
\usepackage{bm}
%\usepackage{subfig}
%\usepackage[english]{babel}
%\usepackage{algorithm}
%\usepackage{appendix}
\usepackage{mathtools}
\mathtoolsset{showonlyrefs}

\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{pro}{Property}
\newtheorem{assumption}{Assumption}

\theoremstyle{definition}
\newtheorem{defn}{Definition}
\newtheorem{cor}{Corollary}
\newtheorem{example}{Example}
\newtheorem{rmk}{Remark}

\usepackage{dsfont}
%\usepackage{algpseudocode,algorithm}
%\algnewcommand\algorithmicinput{\textbf{Input:}}
%\algnewcommand\algorithmicoutput{\textbf{Output:}}
%\algnewcommand\INPUT{\item[\algorithmicinput]}
%\algnewcommand\OUTPUT{\item[\algorithmicoutput]}
%\DeclareMathOperator*{\minimize}{minimize}

\newcommand*{\KeepStyleUnderBrace}[1]{%f
  \mathop{%
    \mathchoice
    {\underbrace{\displaystyle#1}}%
    {\underbrace{\textstyle#1}}%
    {\underbrace{\scriptstyle#1}}%
    {\underbrace{\scriptscriptstyle#1}}%
  }\limits
}

\input macros.tex

\begin{document}

\twocolumn[
\icmltitle{Tensor denoising and completion based on ordinal observations}

% It is OKAY to include author information, even for blind
% submissions: the style file will automatically remove it for you
% unless you've provided the [accepted] option to the icml2020
% package.

% List of affiliations: The first argument should be a (short)
% identifier you will use later to specify author affiliations
% Academic affiliations should list Department, University, City, Region, Country
% Industry affiliations should list Company, City, Region, Country

% You can specify symbols, otherwise they are numbered in order.
% Ideally, you should not use this facility. Affiliations will be numbered
% in order of appearance and this is the preferred way.
\icmlsetsymbol{equal}{*}

\begin{icmlauthorlist}
\icmlauthor{Chanwoo Lee}{equal,to}
\icmlauthor{Miaoyan Wang}{equal,to}
\end{icmlauthorlist}

\icmlaffiliation{to}{Department of Statistics, University of Wisconsin at Madison, USA}

\icmlcorrespondingauthor{Miaoyan Wang}{miaoyan.wang@wisc.edu.com}

% You may provide any keywords that you
% find helpful for describing your paper; these are used to populate
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{Higher-order tensors, ordinal data, tensor denoising, tensor completion}

\vskip 0.3in
]

% this must go after the closing bracket ] following \twocolumn[ ...

% This command actually creates the footnote in the first column
% listing the affiliations and the copyright notice.
% The command takes one argument, which is text to display at the start of the footnote.
% The \icmlEqualContribution command is standard text for equal contribution.
% Remove it (just {}) if you do not need this facility.

%\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.

\begin{abstract}
Higher-order tensors arise frequently in applications such as neuroimaging, recommendation system, social network analysis, and psychological studies. We consider the problem of low-rank tensor estimation from possibly incomplete, ordinal-valued observations. Two related problems are studied, one on tensor denoising and another on tensor completion. 
%{\color{red}In contrast to the continuous-valued problems, problems with ordinal tensors exhibit an intrinsic phase transition depending on the signal-to-noise ratio. (Not sure whether this should be mentioned)}
We propose a multi-linear cumulative link model, develop a rank-constrained M-estimator, and obtain theoretical accuracy guarantees. 
Our error bound enjoys a faster convergence rate than existing results, and we show that the proposed estimator is minimax optimal under the considered model. 
An efficient completion method is further developed, which guarantees the consistent recovery of an order-$K$ $(d,\ldots,d)$-dimensional tensor using only $O(Kd)$ noisy, quantized observations. We demonstrate the outperformance of our approach over previous methods through applications to the tasks of clustering and collaborative filtering.

\end{abstract}

\section{Introduction {\color{red}(needs rewriting)}}
Higher-order tensors commonly arise in modern applications. In a context-aware recommendation system, for example, the survey results can be organized as a three-way tensor, where the $(i,j,k)$-th entry indicates the rating from user $i$ to the item $j$ at context $k$. While there are numerous methods for continuous-valued tensors, fewer methods are available for learning tensors based on qualitative observations. Qualitative observations usually take values in a limited set of categories which may be on an ordinal or on a purely nominal scale. The present work focuses on ordinal tensors; namely, the tensor entries are categorical variables with a natural ordering among the categories. 
{\color{red}(Prior work and our contribution. Make a table of comparison.)}

\section{Preliminaries}
Let $\tY\in\mathbb{R}^{d_1\times \cdots \times d_K}$ denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional tensor. We use $y_\omega$ to denote the tensor entry indexed by $\omega$, where $\omega\in[d_1]\times\cdots\times[d_K]$.  The Frobenius norm of $\tY$ is defined as $\FnormSize{}{\tY}=\sum_\omega y^2_\omega$ and the infinity norm of $\tY$ is defined as $\mnormSize{}{\tY}=\max_{\omega}|y_\omega|$. We use $\tY_{(k)}$ to denote the unfolded matrix of size $d_k$-by-$\prod_{i\neq k}d_k$, obtained by reshaping the tensor along the mode-$k$. The Tucker rank of $\tY$ is defined as a length-$K$ vector $\mr = (r_1,\ldots,r_K)$, where $r_k$ is the rank of matrix $\tY_{(k)}$ for all $k \in[K]$. We say that an event $A$ occurs ``with very high probability'' if $\mathbb{P}(A)$ tends to 1 faster than any polynomial of tensor dimension $d_{\min}=\min\{d_1,\ldots,d_K\}$ as $d_{\min}\to\infty$.

We use lower-case letters (e.g., $a$, $b$, $c$) for scalars/vectors, upper-case boldface letters (e.g., $\mA$, $\mB$, $\mC$) for matrices, and calligraphy letters (e.g., $\tA$, $\tB$, $\tC$) for tensors of order three or greater. For ease of notation, we allow the basic arithmetic operators (e.g., $\leq, +, -$) to be applied to pairs of vectors in an element-wise manner. We use the shorthand $[n]$ to denote the $n$-set $\{1,\ldots,n\}$ for $n \in N_{+}$.

\section{Model formulation and motivation}
\subsection{Observation model}
Let $\tY$ denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional tensor. Suppose the entries of $\tY$ are ordinal-valued, and the observation space consists of $L$ ordered categories, denoted by $[L]:=\{1,\ldots,L\}$. We propose a cumulative link model for the ordinal tensor $\tY=\entry{y_\omega}\in[L]^{d_1\times \cdots\times d_K}$. Specifically, assume the entries $y_\omega$ are (conditionally) independently distributed with cumulative probabilities:
\begin{equation}\label{eq:model}
\mathbb{P}(y_\omega\leq \ell)=f(\theta_\omega+b_\ell),\ \text{for all}\ \ell\in[L-1],
\end{equation}
where $\Theta=\entry{\theta_\omega}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is a continuous-valued parameter tensor satisfying certain low-dimensional structure (to be specified later), $\mb=(b_1,\ldots,b_{L-1})$ is a set of unknown scalars satisfying $b_1<\cdots <b_{L-1}$, and $f(\cdot):\mathbb{R}\mapsto[0,1]$ is a known, strictly increasing function. We refer to $\mb$ as the cut-off points and $f$ the link function.

The formulation~\eqref{eq:model} imposes an additive model to the transformed probability of cumulative categories, rather than that of each single category. This modeling choice is to respect the ordering structure among the categories. For example, if we choose the inverse link $f^{-1}(x)=\log {x\over 1-x}$ to be the log odds, then the model~\eqref{eq:model} assumes linear spacing between the proportional odds:
\begin{equation}\label{eq:logodd}
\log {\mathbb{P}(y_{\omega}\leq \ell) \over \mathbb{P}(y_\omega >  \ell) } - \log {\mathbb{P}(y_{\omega}\leq {\ell-1}) \over \mathbb{P}(y_\omega >  {\ell-1}) } = b_\ell-b_{\ell-1},
\end{equation}
for all tensor entries $y_\omega$. When there are only two categories in the observation space (e.g. binary tensors), the cumulative model~\eqref{eq:model} is equivalent to the usual multinomial link model. In general, however, when the number of categories $L\geq 3$, the proportional odds assumption~\eqref{eq:logodd} is more parsimonious, in that, the ordered categories can be envisaged as contiguous intervals on the continuous scale, where the points of division are exactly $b_1<\cdots <b_{L-1}$. This interpretation will be made more explicit in the next section.

\subsection{Latent-variable interpretation.}

We show that the ordinal tensor model~\eqref{eq:model} is equivalent to an $L$-level quantization model on $\tY=\entry{y_\omega}$:
\begin{align}\label{eq:quantization}
y_\omega&=
\begin{cases}
1,& \text{if $y^*_\omega\in(-\infty, b_1]$},\\
2,& \text{if $y^*_\omega\in(b_1, b_2]$},\\
\vdots  &\vdots\\
L,& \text{if $y^*_\omega\in(b_{L-1}, \infty)$},\\
\end{cases}
\end{align}
for all $\omega\in[d_1]\times \cdots \times [d_k]$. Here, $\tY^*=\entry{y^*_{i_1,\ldots,i_K}}$ is a latent continuous-valued tensor following an additive noise model:
\begin{align}\label{eq:latent}
\KeepStyleUnderBrace{\tY^*}_{\text{latent continuous-valued tensor}}=\KeepStyleUnderBrace{\Theta}_{\text{signal tensor}}+\KeepStyleUnderBrace{\tE}_{\text{i.i.d.\ noise}},
\end{align}
where $\tE=\entry{\varepsilon_\omega}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is a noise tensor with i.i.d.\ entries following distribution $\tF(\varepsilon)$. From the viewpoint of~\eqref{eq:latent}, the parameter tensor $\Theta$ can be interpreted as the latent signal tensor prior to contamination and quantization.

The equivalence between the latent-variable model~\eqref{eq:quantization} and the cumulative link model~\eqref{eq:model} is established if the link $f$ behaves like a cumulative distribution function of noise $\varepsilon$, i.e., $f(\theta)=\mathbb{P}(\varepsilon\geq -\theta)$. We describe several common choices of link $f$, or equivalently, the distribution of $\tE$.

\begin{example}[Logistic model] The logistic model is characterized by~\eqref{eq:model} with $f(\theta)=(1+e^{-\theta/\sigma})^{-1}$, where $\sigma>0$ is the scale parameter. Equivalently, the noise $\varepsilon_\omega$ in~\eqref{eq:quantization} follows i.i.d.\ logistic distribution with scale parameter $\sigma$. 
\end{example}
\begin{example}[Probit model] The probit model is characterized by~\eqref{eq:model} with
$f(\theta)=\mathbb{P}(z\leq \theta/\sigma)$, where $z\sim N(0,1)$. Equivalently, the noise $\varepsilon_\omega$ in~\eqref{eq:quantization} follows i.i.d.\ $N(0,\sigma^2)$.
\end{example}
\begin{example}[Laplace model] The Laplacian model is characterized by~\eqref{eq:model} with
\[
f(\theta)=\begin{cases}
{1\over 2}e^{\theta\over \sigma}, & \text{if $\theta<0$},\\
1-{1\over 2}e^{-{\theta\over \sigma}}, & \text{if $\theta\geq 0$},
\end{cases}
\]
where $\sigma>0$ is the scale parameter. Equivalently, the noise $\varepsilon_\omega$ in~\eqref{eq:quantization} follows i.i.d.\ Laplace distribution with scale parameter $\sigma$. 
\end{example}
Other link functions are also possible, such as Cauchy function, inverse log-log, etc~\cite{mccullagh1980regression}. All the models share the property that the ordered categories can be thought of as contiguous interval on some continuous scale. We should point out that, although the latent-variable interpretation is incisive, our estimation procedure does not require the existence of $\tY^*$. Therefore, our model~\eqref{eq:model} is general and still valid in the absence of quantization process. More generally, we make the following assumptions to the link function $f$.

\begin{assumption}\label{ass:link}
The link function is assumed to satisfy:
\vspace{-.4cm}
\begin{enumerate}
\item $f(\theta)$ is strictly increasing, strictly log-concave, and twice-differentiable in $\theta\in\mathbb{R}/\{0\}$.
\vspace{-.2cm}
\item $f'(\theta)$ is unimodal and symmetric with respect to $\theta=0$. 
\end{enumerate}
\end{assumption}

\subsection{Problem 1: Tensor denoising.}~\label{sec:denoising}
The first question we aim to address is tensor denoising:

(P1) Given the quantization process induced by $f$ and the cut-off points $\mb$, how accurate can we recover the latent signal tensor $\Theta$ from the ordinal observation $\tY$?

Clearly, the problem (P1) cannot be solved uniformly for all possible $\Theta$. We focus on a class of ``low-rank'' and ``flat'' signal tensors, which is a plausible assumption in practical applications~\cite{zhou2013tensor,bhaskar20151}. Specifically, we consider the parameter space:
\begin{equation}\label{eq:space}
\tP=\left\{\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K} \colon \text{rank}(\Theta)\leq \mr,\ \mnormSize{}{\Theta}\leq \alpha\right\}.
\end{equation}
where $\mr=(r_1,\ldots,r_K)$ denotes the Tucker rank of $\Theta$.

The candidate tensor of our interest satisfies two constraints. The first is that $\Theta$ is a low-rank tensor, with $r_k=O(1)$ for all $k\in[K]$. The Tucker low-rankness is popularly used in tensor data analysis, and is shown to provide a reasonable tradeoff between model complexity and model flexibility. Note that, unlike matrices, there are various notations of tensor low-rankness, such as CP representation~\cite{hitchcock1927expression} and train representation~\cite{oseledets2011tensor}. Some notation of low-rankness may lead to mathematically ill-defined optimization; for example, the set of low-rank CP tensors is not necessarily compact~\cite{de2008tensor}. We choose Tucker representation for parsimony and easier interpretation.

The second constraint is that the entries of $\Theta$ are uniformly bounded in magnitude by a constant $\alpha \in \mathbb{R}_{+}$. In view of~\eqref{eq:latent}, we refer to $\alpha$ as the signal level. The entry-wise bound assumption is a technical condition that ensures the well-posedness for the problems we considered. In the context of denoising with ordinal observations, the entry-wise bound helps avoid the degeneracy in probability estimation.

\subsection{Problem 2: Tensor completion}
Motivated by applications in collaborative filtering and recommendation system, we also also consider a more general setup when only a subset of tensor entries $y_\omega$ are observed. Let $\Omega\subset[d_1]\times \cdots\times[d_K]$ denote the set of observed indices. The second question we aim to address is stated as follows:

(P2) Given an incomplete set of ordinal observation $\{y_{\omega}\}_{\omega\in\Omega}$ based on the model~\eqref{eq:model}, how many sampled entries do we need to consistently estimate $\Theta$?

The answer to (P2) depends on the choice of $\Omega$. We propose a general model on $\Omega$ that allows both uniform and non-uniform sampling. Specifically, let $\Pi=\{\pi_{i_1,\ldots,i_K}\}$ denote a predefine probability distribution over the index set such that $\sum_{\omega\in[d_1]\times \cdots \times [d_K]} \pi_\omega =1$. We assume that each index in $\Omega$ are drawn with replacement using distribution $\Pi$. This model relaxes the the commonly-used uniform sampling and is arguably a better fit in practical applications.

We consider the same parameter space~\eqref{eq:space} for the completion problem. In addition to the reasons mentioned in Section~\ref{sec:denoising}, the entrywise bound assumption also serves as the incoherence condition here. In classical matrix completion, the incoherence is often imposed on the singular vectors. This assumption was relaxed by considering ``flat'' matrices with bounded magnitude~\cite{negahban2011estimation,cai2013max,bhaskar20151}. We adopt the same assumption for higher-order tensors.

\section{Rank-constrained M-estimator}
We present a general treatment to both problems mentioned above. With a little abuse of notation, we use $\Omega$ to denote either the full index set $\Omega=[d_1]\times \cdots \times [d_K]$ (for the tensor denoising) or a random subset induced from the sampling distribution $\Pi$ (for the tensor completion). Define $b_0=-\infty$, $b_L=\infty$, $f(-\infty)=0$ and $f(\infty)=1$. The log-likelihood associated with the observed entries is
\begin{align}\label{eq:objective}
\logl(\Theta, \mb)&=\sum_{\omega\in\Omega}\sum_{\ell\in[L]} \Big\{\mathds{1}_{\{y_\omega=\ell\}} \log \big[f(\theta_\omega+b_\ell)- \notag \\
&\quad \quad \quad \quad \quad \quad \quad  f(\theta_\omega+b_{\ell-1})\big]\Big\}.
\end{align}
We propose a rank-constrained maximum likelihood estimator (M-estimator) for $\Theta$:
\begin{align}\label{eq:estimator}
\hat \Theta&=\argmax_{\Theta\in \tP}\logl(\Theta, \mb),\ \text{where}\notag \\
\tP&=\left\{\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K} \colon \text{rank}(\Theta)\leq \mr,\ \mnormSize{}{\Theta}\leq \alpha\right\}.
 \end{align}
 In practice, the cut-off points $\mb$ are unknown and should be jointly estimated with $\Theta$. For technical convenience, we assume that, in this section, the cut-off points $\mb$ are known, or otherwise have been consistently estimated {\color{red}(??)}. The adaptation of unknown $\mb$ will be addressed in Section~\ref{sec:algorithm}. Similar treatment applies to the hyper-parameter $\mr$.

We define a few key quantities that will be used in our theory. Let $g_\ell=f(\theta+b_\ell)-f(\theta+b_{\ell-1})$ for all $\ell\in[L]$, and
\begin{align}\label{eq:regular}
A_\alpha &= \min_{\ell\in[L], |\theta|\leq \alpha} g_\ell(\theta),\quad  U_\alpha=\max_{\ell\in[L], |\theta|\leq \alpha } {\dot{g}_\ell(\theta)\over g_\ell(\theta)},\\
 L_\alpha &= \min_{\ell\in[L], |\theta|\leq \alpha }\left[{\dot{g}_\ell^2 (\theta)\over g_\ell^2(\theta)} -{\ddot{g}_\ell(\theta)\over g_\ell(\theta)}\right],
\end{align}
where $\dot{g}(\theta)=dg(\theta)/d\theta$, and $\alpha$ is the entrywise bound of $\Theta$. In view of equation~\eqref{eq:latent}, these quantities characterize the geometry (such as flatness and convexity) of the latent noise distribution. Under the assumption~\ref{ass:link}, all these quantities are strictly positive and independent of the tensor dimension.

\subsection{Estimation error for tensor denoising}
For the tensor denosing problem, we assume the full set of tensor entries are observed. In such a case, we assess the estimation accuracy using the mean squared error (MSE):
\[
\text{MSE}(\hat \Theta, \trueT)={1\over \prod_k d_k}\FnormSize{}{\Theta-\trueT}^2.
\]
The next theorem establishes the upper bound for the proposed $\hat \Theta$ in~\eqref{eq:estimator}.

\begin{thm}[Statistical convergence] \label{thm:rate}
Consider an ordinal tensor $\tY\in[L]^{d_1\times\dots\times d_K}$ generated from model~\eqref{eq:model}, with the link function $f$ and the true coefficient tensor $\trueT\in\tP$. Define $r_{\max}=\max_k r_k$. Then, with very high probability, the estimator in~\eqref{eq:estimator} satisfies
\begin{equation}\label{eq:rate}
\text{MSE}(\hat \Theta, \trueT) \leq \min\left( 4\alpha^2,\ {c_2  U^2_\alpha r_{\max}^{K-1}  \over  L^2_\alpha } {\sum_kd_k\over  \prod_k d_k} \right),
\end{equation}
where $c_1, c_2>0$ are two constants that depend only on $K$. 
\end{thm}
Theorem~\ref{thm:rate} provides the statistical converge for the estimator~\eqref{eq:estimator}. Its proof is given in the Supplements. In fact, the proof of this theorem shows that the same statistical rate holds not only for the global optimizer of~\eqref{eq:estimator}, but also for any local optimizer $\check \Theta$ in the level set $\{\check\Theta\in\tP: \logl(\check\Theta)\geq \logl(\trueT)\}$. This suggests that the local optimality itself is not necessarily a severe concern in our context, as long as the objective value at the estimator is large enough. {\color{red}In Section~\ref{algorithm}, we will provide an algorithm-dependent rate quantifying the interplay between the statistical and computational errors.}

A similar conclusion can be obtained for the prediction error, measured in KL divergence, between the two generative distributions. 
\begin{cor}[Prediction error]~\label{cor:prediction}
Assume the same set-up as in Theorem~\ref{thm:rate}. Let $\mathbb{P}_{\tY}$ and $\hat{\mathbb{P}}_{\tY}$ denote the distributions generating the $L$-level ($L\geq 2$) ordinal tensor $\tY$, given the true parameter $\Theta$ and its estimator $\hat \Theta$, respectively. Then, with very high probability, 
\begin{equation}\label{eq:KLrate}
\text{KL}(\mathbb{P}_{\tY} || \hat{\mathbb{P}}_{\tY}) \leq  {c_2 U^2_\alpha r_{\max}^{K-1}  \over L^2_\alpha } {(4L-6)\dot{f}^2(0)  \over A_\alpha} {\sum_kd_k\over  \prod_k d_k},
\end{equation}
where $c_1, c_2>0$ are the same constants as in Theorem~\ref{thm:rate}.  
\end{cor}
To gain insight into these bounds, we consider a special setting where the dimensions are the same in all modes, i.e., $d_1=\cdots=d_K=d$. In such a case, our bound \eqref{eq:rate} reduces to
\begin{equation*}\label{eq:ours}
\text{MSE}(\hat \Theta, \trueT) \asymp d^{-(K-1)}, \quad \text{as}\ d\to \infty.
\end{equation*}
Hence, the estimator achieves consistency with polynomial convergence rate. Comparing to max norm based estimator $\tilde{\Theta}$ proposed by~\citet{ghadermarzy2018learning}, $\text{MSE}(\tilde{\Theta}, \trueT) \asymp  d^{-(K-1)/2}$, we see that our bound has a faster convergence rate. We believe the improvement comes from utilizing the exact low-rank structure of $\Theta$, whereas an scale-sensitive surrogate rank measure was employed in~\citet{ghadermarzy2018learning}.

Our bound also generalizes the previous results on ordinal matrices. The convergence rate for rank-constrained matrix denoising was $O(1/\sqrt{d})$~\citep{bhaskar2016probabilistic}, which fits into our special case when $K=2$. Furthermore, our results~\eqref{eq:rate} and~\eqref{eq:KLrate} reveal that the convergence becomes favorably as the order of tensor data increases. 
Intuitively, for tensor data analysis, we can view the number of entries, $\prod_k d_k$, as the sample size, and the number of free parameters in $\tP$ is roughly on the order of $\sum_{k}d_k$, assuming $r_{\max}=O(1)$. Compared to the matrix case, a higher tensor order implies more effective samples per parameter and thus exhibits a faster convergence rate in high dimensions.

We next establish the statistical optimality of our estimator $\hat \Theta$. The result is based on the information theory, and therefore applies to all estimators in $\tP$, including but not limited to $\hat \Theta$ in~\eqref{eq:estimator}. We show that this lower bound nearly matches the upper bound in~\eqref{eq:rate}, as the tensor dimension increases.

\begin{thm}[Minimax lower bound]\label{thm:minimax} 
Assume the same set-up as in Theorem~\ref{thm:rate}, and $d_{\max}=\max_k d_k \geq 8$. Let $\inf_{\hat \Theta}$ denote the infimum over all estimators $\hat \tA\in\tP$ based on the ordinal tensor observation $\tY\in[L]^{d_1\times \cdots \times d_K}$. Then, under the model~\eqref{eq:model},
\begin{align}\label{eq:lower}
\inf_{\hat \Theta }\sup_{\trueT\in\tP} \mathbb{P}\Big\{& \text{MSE}(\hat \Theta, \trueT) \\
& \geq {1\over 256} \min\left( \alpha^2, \ { Cr_{\max}d_{\max} \over \prod_k d_k} \right) \Big\} \geq {1\over 8},
\end{align}
where $C=C(\alpha, L, f,\mb)>0$ is a constant independent of the tensor dimension and rank. 
\end{thm}

\subsection{Sample complexity for tensor completion}

Now we consider the problem of tensor completion, when only a subset of entries $\Omega$ are observed. We consider a general sampling procedure induced by $\Pi$, and assess the accuracy using the weighted squared deviation:
\begin{align}\label{eq:weighted}
\PiFnormSize{}{\Theta-\hat \Theta}^2&\stackrel{\text{def}}{=}
{1\over |\Omega|}\mathbb{E}_{\Omega\sim \Pi}\FnormSize{}{\Theta-\hat \Theta}^2\notag \\
&=\sum_{\omega\in[d_1]\times \cdots \times [d_K]} \pi_{\omega}(\Theta_{\omega}-\hat \Theta_{\omega})^2.
\end{align}

Note that the recovery error depends on the distribution. In particular, the entries with higher sampling probabilities are equipped with better accuracy guarantee, compared to the ones with lower sampling probabilities.

\begin{rmk} If we assume each tensor entry is sampled with some positive probability, i.e., there exits a constant $\mu> 0$ such that 
\[
\pi_\omega\geq {1\over \mu \prod_k d_k},\quad \text{for all}\quad \omega\in[d_1]\times\cdots \times [d_K],
\]
then the error measure in~\eqref{eq:weighted} provides an upper bound for MSE:
\[
\PiFnormSize{}{\Theta-\hat \Theta}^2 \geq {\FnormSize{}{\Theta-\hat \Theta}^2\over \mu \prod_kd_k}={1\over \mu}\text{MSE}(\hat \Theta, \trueT).
\]
The equality is attained in the case of uniform sampling for which $\mu=1$. 
\end{rmk}

\begin{thm}~\label{thm:completion}
Assume the same set-up as in Theorem~\ref{thm:rate}. Suppose that we observe a subset of tensor entries $\{y_\omega\}_{\omega\in\Omega}$, where $\Omega$ is chosen at random with replacement according to a probability distribution $\Pi$. Let $\hat \Theta$ be the solution to~\eqref{eq:estimator}, and assume $r_{\max}=O(1)$. Then, with very high probability over the sampled set $\Omega$, 
\[
\PiFnormSize{}{\Theta-\hat \Theta}^2 \to 0,\quad \text{ as }\quad {|\Omega|\over \sum_k d_k} \ \to \infty. 
\]
\end{thm}
Theorem~\ref{thm:completion} shows that our estimator achieves consistency when sample size is greater than $O(\sum_k d_k)$. {\color{red} add remarks comparing with prior work, and with the degree of freedom}

\section{Numerical Implementation}\label{sec:algorithm}
\subsection{Alternating optimization}
\begin{figure*}[http]
\begin{center}
\includegraphics[width=16cm]{algorithm.pdf}
 \end{center}
\end{figure*}
In this section, we describe the algorithm to seek the optimizer of~\eqref{eq:objective}. The objective function $\logl(\Theta)$ is concave in $\Theta$ whenever $\tP$ is log-concave~\cite{mccullagh1980regression}. However, the feasible set $\tP$ is not a convex set, which makes the optimization~\eqref{eq:objective} a non-convex problem. One approach to handle this problem is utilizing the Tucker decomposition and converting optimization into a block-wise convex problem. Let us denote the low Tucker-rank tensor as
\begin{equation}\label{eq:Tucker}
\Theta=\tC\times_1\mM_1\times_1 \cdots \times_K \mM_K,
\end{equation}
where $\tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is a core tensor and $\mM_k\in\mathbb{R}^{d_k\times r_k}$ are factor matrices with orthogonal columns. From~\eqref{eq:objective} and \eqref{eq:Tucker}, we have $K+2$ blocks of variables in the objective function, one for the cut-points vector $\mb$, one for the core tensor $\tC$, and $K$ for the factor matrices $\mM_k$'s. We can change the optimization problem to simple convex problem if any $N+1$ out of the $N+2$ blocks being fixed. Therefore, we can update one block at a time while other blocks being fixed, and alternate the optimization via iteration. The Algorithm 1 gives the full description.

{\color{red} (highlight that $\mb$ and $\Theta$ are jointly estimated...)}

\subsection{Tuning parameter selection}\label{sec:BIC}
Algorithm~1 takes the rank $\mr$ as an input. In practice, the rank $\mr$ is hardly known and needs to be estimated from the data. We suggest to use Bayesian information criterion (BIC) and choose the rank that minimizes BIC; i.e.
\begin{align}\label{eq:BIC}
\hat \mr&=\argmin_{\mr\in\mathbb{N}^K_{+}} \text{BIC}(\mr)\\
&=\argmin_{\mr\in\mathbb{N}^K_{+}}\left[-2\tL_{\tY}(\hat \Theta(\mr))+p_e(\mr)\log \left(\prod_k d_k\right) \right],
\end{align}
where $\hat \Theta(\mr)$ is the estimator given the rank $\mr$, and $p_e(\mr)\stackrel{\text{def}}{=}\sum_k (d_k-r_k)r_k+\prod_k r_k$ is the effective number of parameters in the model. We select the rank $\hat \mr$ that minimizes BIC through a grid search. Our choice of BIC aims to balance between the goodness-of-fit for the data and the degree of freedom in the population model. We test its empirical performance in Section~\ref{sec:simulation}.

\section{Experiments}
In this section, we evaluate the empirical performance of our rank-constrained M-estimatation. We consider ordinal tensors with various number of categories, and compare the recovery accuracy with other tensor-based methods. Unless otherwise stated, we generate ordinal-valued tensor under the latent-variable model~\eqref{eq:latent} with i.i.d.\ $N(0,\sigma^2)$ noise. The signal tensor is simulated based on~\eqref{eq:Tucker}, with core tensor generated from i.i.d. N(0,1), and the factor matrices uniformly from orthogonal ensembles. The cut-off points are generated i.i.d.from $N(0,1)$. In each simulation study, we report the summary statistics across $n_{\text{sim}}$ = 30 replications.
\subsection{Finite-sample performance}\label{sec:simulation}

\subsection{Comparison with alternative methods}

\section{Real data application}
In this section, we apply our ordinal tensor decomposition method to two real-world datasets of ordinal tensors. In the first application, we use our model to analyze an ordinal tensor consisting of structural connectivity patterns among 68 brain regions for 136 individuals from Human Connec- tome Project (HCP) (Geddes 2016). In the second application, we perform tensor completion from the data with missing values. The data tensor records the ratings from scale 1 to 5 of 42 users to 139 songs on 26 contexts (Baltrunas et al. 2011).

\subsection{Human Connection Project (HCP)}
The human connectome project (HCP) is a 68 × 68 × 136 tensor where the first two modes have 68 indices representing brain regions and the last mode has 136 indices meaning individuals. All the individual images were preprocessed following a standard pipeline (Zhang et al. 2018), and the brain was parcellated to 68 regions of interest following the Desikan atlas (Desikan et al. 2006). The tensor entries consist of $\{1, 2, 3\}$, the strength of fiber connections between 68 brain regions for each of the 136 individuals. First, we apply our ordinal tensor decomposition method with a logistic link function to the HCP data and identify similar brain nodes based on latent parameters. The BIC result suggests $\mr = (23, 23, 8)$ with $\logl(\hat \Theta,\hat \mb) = - 216645.8$.

Also, we compare our ordinal tensor decomposition method with a logistic link function with the classical continuous Tucker decomposition method. We use 5 folded cross validation method for the comparison. Specifically, we randomly split the tensor entries into 5 similar sized pieces and alternatively use each piece of entries as a test data using other 80\% entries as a training data. The entries in the test data are encoded as missing and then predicted based on the tensor decomposition from the training data. We set the rank of the tensor as $\mr = (23,23,8)$ which minimizes BIC value. Table 1 shows RMSE, MAE and error (false prediction rate), averaged over 5 test set results. We can check that our ordinal tensor decomposition model outperforms the classic continuous decomposition in all criteria.

\bibliography{tensor_wang}
\bibliographystyle{icml2020}

\end{document}

% This document was modified from the file originally made available by
% Pat Langley and Andrea Danyluk for ICML-2K. This version was created
% by Iain Murray in 2018, and modified by Alexandre Bouchard in
% 2019 and 2020. Previous contributors include Dan Roy, Lise Getoor and Tobias
% Scheffer, which was slightly modified from the 2010 version by
% Thorsten Joachims & Johannes Fuernkranz, slightly modified from the
% 2009 version by Kiri Wagstaff and Sam Roweis's 2008 version, which is
% slightly modified from Prasad Tadepalli's 2007 version which is a
% lightly changed version of the previous year's version by Andrew
% Moore, which was in turn edited from those of Kristian Kersting and
% Codrina Lauth. Alex Smola contributed to the algorithmic style files.

### 2020-01-21 ICLM/miaoyan/ordinalT_supp.tex
upload draft

\documentclass[11pt]{article}

\usepackage{fancybox}

\usepackage{color}
\usepackage{url}
\usepackage[margin=1in]{geometry}

\renewcommand{\textfraction}{0.0}
\renewcommand{\topfraction}{1.0}
%\renewcommand{\textfloatsep}{5mm}

\usepackage{comment}
% Definitions of handy macros can go here
\usepackage{amsmath,amssymb,amsthm,bm,mathtools}
\usepackage{multirow}
\usepackage{natbib}
%\usepackage{dsfont,multirow,hyperref,setspace,natbib,enumerate}
\usepackage{dsfont,multirow,hyperref,setspace,enumerate}
\hypersetup{colorlinks,linkcolor={blue},citecolor={blue},urlcolor={red}} 
\usepackage{algpseudocode,algorithm}
\algnewcommand\algorithmicinput{\textbf{Input:}}
\algnewcommand\algorithmicoutput{\textbf{Output:}}
\algnewcommand\INPUT{\item[\algorithmicinput]}
\algnewcommand\OUTPUT{\item[\algorithmicoutput]}

\mathtoolsset{showonlyrefs=true}

\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{pro}{Property}
\newtheorem{assumption}{Assumption}

\theoremstyle{definition}
\newtheorem{defn}{Definition}
\newtheorem{cor}{Corollary}
\newtheorem{example}{Example}
\newtheorem{rmk}{Remark}

\renewcommand{\thefigure}{{S\arabic{figure}}}%
\renewcommand{\thetable}{{S\arabic{table}}}%
\renewcommand{\figurename}{{Supplementary Figure}}    
\renewcommand{\tablename}{{Supplementary Table}}    
\setcounter{figure}{0}   
\setcounter{table}{0}

\def\MLET{\hat \Theta_{\text{MLE}}}
\newcommand{\cmt}[1]{{\leavevmode\color{red}{#1}}}

\usepackage{dsfont}

\usepackage{multirow}

\DeclareMathOperator*{\minimize}{minimize}

\usepackage{mathtools}
\mathtoolsset{showonlyrefs}
\newcommand*{\KeepStyleUnderBrace}[1]{%f
  \mathop{%
    \mathchoice
    {\underbrace{\displaystyle#1}}%
    {\underbrace{\textstyle#1}}%
    {\underbrace{\scriptstyle#1}}%
    {\underbrace{\scriptscriptstyle#1}}%
  }\limits
}
\usepackage{xr}
\externaldocument{ordinalT}
\input macros.tex

\title{Supplements for ``Tensor denoising and completion based on ordinal observations''}

%\author{%
%Chanwoo Lee \\
%University of Wisconsin -- Madison\\
 %\texttt{chanwoo.lee@wisc.edu} \\
%\And
%Miaoyan Wang \\
%University of Wisconsin -- Madison\\
%\texttt{miaoyan.wang@wisc.edu} \\
%}

\begin{document}

\begin{center}
\begin{spacing}{1.5}
\textbf{\Large Supplements for ``Tensor denoising and completion based on ordinal observations''}
\end{spacing}
\end{center}

\section{Proofs}
\subsection{Estimation error for tensor denoising}
\begin{proof}[Proof of Theorem~\ref{thm:rate}]
We suppress the subscript $\Omega$ in the proof, because the tensor denoising assumes complete observation $\Omega=[d_1]\times \cdots \times [d_K]$. It follows from the expression of $\flogl(\Theta)$ that
\begin{align}\label{eq:property}
{\partial \flogl\over \partial \theta_\omega}&=\sum_{\ell\in[L]}\mathds{1}_{\{y_{\omega}=\ell\}}
{\dot{g}_\ell(\theta_\omega)\over g_\ell(\theta_\omega)},\notag\\
{\partial^2 \flogl\over \partial \theta_\omega^2}&=\sum_{\ell\in[L]}\mathds{1}_{\{y_\omega=\ell\}}{\ddot{g}_\ell(\theta_\omega)g_\ell(\theta_\omega)-\dot{g}^2_\ell(\theta_\omega)\over g^2_\ell(\theta_\omega)}\ \text{and}\quad
{\partial^2 \flogl\over \partial \theta_\omega \theta_\omega'}=0\ \text{if}\ \omega\neq \omega',
\end{align}
for all $\omega\in[d_1]\times \cdots \times [d_K]$. 
Define $d_{\text{total}}=\prod_k d_k$. Let $\fplogl\in\mathbb{R}^{d_\text{total}}$ denote the vector of gradient with respect to $\Vec(\Theta)\in\mathbb{R}^{d_{\text{total}}}$, and $\fpplogl$ the corresponding Hession matrix of size $d_\text{total}$-by-$d_{\text{total}}$. Here, $\Vec(\cdot)$ denotes the operation that turns a tensor into a vector. By~\eqref{eq:property}, $\fpplogl$ is a diagonal matrix. Recall that
\[
U_\alpha=\max_{\ell\in[L],|\alpha|\leq \alpha}{\dot{g}_\ell(\theta)\over g_\ell(\theta)}>0 \quad \text{and}\quad
L_\alpha=\max_{\ell\in[L],|\alpha|\leq \alpha} {\dot{g}^2_\ell(\theta)-\ddot{g}_\ell(\theta)g_\ell(\theta)\over g^2_\ell(\theta)}>0. 
\]
Therefore, all entries in $\fplogl$ are upper bounded $U_\alpha>0$, and all diagonal entries in $\fpplogl$ are upper bounded by $-L_{\alpha}<0$.

By the second-order Taylor's expansion of $\flogl(\Theta)$ around $\trueT$, we obtain
\begin{equation}\label{eq:taylor}
\flogl(\Theta)=\flogl(\trueT)+\langle\fplogl,\ \Vec(\Theta-\trueT)\rangle+{1\over 2}\Vec(\Theta-\trueT)^T\fpplogl(\check\Theta)\Vec(\Theta-\trueT),
\end{equation}
$\check\Theta=\gamma\trueT+(1-\gamma)\Theta$ for some $\gamma\in[0,1]$, and $\fpplogl(\check\Theta)$ denotes the $\prod_kd_k$-by-$\prod_k d_k$ Hession matrix evaluated at $\check\Theta$.

We first bound the linear term in~\eqref{eq:taylor}. Note that, by Lemma~\ref{lem:inq}, 
\begin{equation}\label{eq:linear}
|\fplogl(\trueT), \Vec(\Theta-\trueT)  \rangle|\leq \snormSize{}{\fplogl(\trueT)} \nnormSize{}{\Theta-\trueT},
\end{equation}
where $\snormSize{}{\cdot}$ denotes the tensor spectral norm and $\nnormSize{}{\cdot}$ denotes the tensor nuclear norm. Define 
\[
s_\omega={\partial \tL_\tY\over \partial \theta_\omega}\Big|_{\Theta=\trueT} \;\; \textrm{ for all } \; \omega\in[d_1]\times\cdots\times [d_K].
\]
Based on~\eqref{eq:property} and the definition of $U_\alpha$, $\fplogl(\trueT)=\entry{s_{\omega}}$ is a random tensor whose entries are independently distributed satisfying
\begin{equation}\label{eq:norm}
\mathbb{E}(s_\omega)=0,\quad |s_\omega|\leq U_\alpha, \quad \text{for all }\omega\in[d_1]\times \cdots \times [d_K].
\end{equation}
By lemma~\ref{lem:noisytensor}, with probability at least $1-\exp(-C_1 \sum_kd_k)$, we have
\begin{equation}\label{eq:normrandom} 
\snormSize{}{\fplogl(\trueT)} \leq C_2 U_\alpha\sqrt{\sum_k d_k},
\end{equation}
where $C_1, C_2$ are two positive constants that depend only on $K$. Furthermore, note that $\text{rank}(\Theta)\leq \mr$, $\text{rank}(\trueT)\leq \mr$, so $\text{rank}(\Theta-\trueT)\leq 2\mr$. By lemma~\ref{lem:nuclear}, $\nnormSize{}{\Theta-\trueT}\leq (2r_{\max})^{K-1\over 2}\FnormSize{}{\Theta-\trueT}$. Combining~\eqref{eq:linear}, \eqref{eq:norm} and \eqref{eq:normrandom}, we have that, with probability at least $1-\exp(-C_1 \sum_kd_k)$,
\begin{equation}\label{eq:linearconclusion}
|\langle \fplogl(\trueT), \Vec(\Theta-\trueT)  \rangle | \leq C_2 U_\alpha  \sqrt{r_{\max}^{K-1} \sum_k d_k}  \FnormSize{}{\Theta-\trueT}.
\end{equation}

We next bound the quadratic term in \eqref{eq:taylor}. Note that
\begin{align}\label{eq:quadratic}
 \Vec(\Theta-\trueT)^T \fpplogl(\check{\Theta})\Vec(\Theta-\trueT)&=\sum_\omega \left( {\partial^2\tL_{\tY}\over \partial \theta^2_\omega} \Big|_{\Theta=\check\Theta} \right)(\theta_\omega-\theta_{{\text{true}},\omega})^2 \nonumber \\
&\leq - L_\alpha\sum_{\omega}(\Theta_{\omega}-\Theta_{\text{true},\omega})^2 \nonumber \\
&=-L_\alpha\FnormSize{}{\Theta-\trueT}^2,
\end{align}
where the second line comes from the fact that  $\mnormSize{}{\check\Theta}\leq \alpha$ and the definition of $L_\alpha$.

Combining~\eqref{eq:taylor}, \eqref{eq:linearconclusion} and~\eqref{eq:quadratic}, we have that, for all $\Theta\in\tP$, with probability at least $1-\exp(-C_1 \sum_kd_k)$, 
\[
\tL_\tY(\Theta)\leq \tL_{\tY}(\trueT)+C_2U_\alpha  \left(r_{\max}^{K-1}\sum_k d_k\right)^{1/2}  \FnormSize{}{\Theta-\trueT}-{L_\alpha\over 2}\FnormSize{}{\Theta-\trueT}^2.
\]
In particular, the above inequality also holds for $\hat \Theta\in\tP$. Therefore, 
\[
\tL_\tY(\hat \Theta)\leq \tL_{\tY}(\trueT)+C_2U_\alpha \left(r_{\max}^{K-1}\sum_k d_k\right)^{1/2}  \FnormSize{}{\hat \Theta-\trueT}-{L_\alpha\over 2} \FnormSize{}{\hat \Theta-\trueT}^2.
\]
Since $\hat \Theta=\arg\max_{\Theta\in\tP}\tL_\tY(\Theta)$, $\tL_\tY(\hat \Theta)-\tL_{\tY}(\trueT)\geq 0$, which gives
\[
C_2U_\alpha \left(r_{\max}^{K-1}\sum_k d_k\right)^{1/2}  \FnormSize{}{\hat \Theta-\trueT}-{L_\alpha\over 2}\FnormSize{}{\hat \Theta-\trueT}^2\geq 0.
\]
Henceforth, 
\[
{1\over \sqrt{\prod_k d_k}} \FnormSize{}{\hat \Theta-\trueT}\leq {2C_2U_\alpha \sqrt{r_{\max}^{K-1}\sum_k d_k}\over L_\alpha \sqrt{\prod_k d_k}}={2C_2U_\alpha r_{\max}^{(K-1)/2}\over L_\alpha} \sqrt{ \sum_k d_k \over \prod_k d_k}.
\]
This completes the proof. 
\end{proof}

\begin{proof}[Proof of Corollary~\ref{cor:prediction}]
The result follows immediately from Theorem~\ref{thm:rate} and Lemma~\ref{lem:KL}.
\end{proof}

\subsection{Sample complexity for tensor completion}
\begin{proof}[Proof of Theorem~\ref{thm:minimax}]

Let $d_{\text{total}}=\prod_{k\in[K]}d_k$, and $\gamma\in[0,1]$ be a constant to be specified later.  Our strategy is to construct a finite set of tensors $\tX=\{\Theta_i\colon i=1,\ldots \}\subset \tP$ satisfying the properties of (i)-(iv) in Lemma~\ref{lem:construction}. By Lemma~\ref{lem:construction}, such a subset of tensors exist. For any tensor  $\Theta\in\tX$, let $\mathbb{P}_{\Theta}$ denote the distribution of $\tY|\Theta$, where $\tY$ is the ordinal tensor. In particular, $\mathbb{P}_{\mathbf{0}}$ is the distribution of $\tY$ induced by the zero parameter tensor $\mathbf{0}$, i.e., the distribution of $\tY$ conditional on the parameter tensor $\Theta=\mathbf{0}$. Based on the Remark for Lemma~\ref{lem:KL}, we have
\begin{equation}\label{eq:KLbound1}
KL(\mathbb{P}_{\Theta}|| \mathbb{P}_{\mathbf{0}})\leq C \FnormSize{}{\Theta}^2,
\end{equation}
where $C={(4L-6) \dot{f}^2(0)\over  A_\alpha}>0$ is a constant independent of the tensor dimension and rank. 
Combining the inequality~\eqref{eq:KLbound1} with property (iii) of $\tX$, we have
\begin{equation}\label{eq:KLbound}
\text{KL}(\mathbb{P}_{\Theta}||\mathbb{P}_{\mathbf{0}})\leq \gamma^2 R_{\max} d_{\max}.
\end{equation}
From~\eqref{eq:KLbound} and the property (i), we deduce that the condition 
\begin{equation}\label{eq:totalKL}
{1\over \text{Card}(\tX)-1}\sum_{\Theta \in\tX}\text{KL}(\mathbb{P}_{\Theta}, \mathbb{P}_{\mathbf{0}})\leq \varepsilon \log\left\{\text{Card}(\tX)-1 \right\}
\end{equation}
holds for any $ \varepsilon \geq 0$ when $\gamma\in[0,1]$ is chosen to be sufficiently small depending on $\varepsilon$, e.g., $\gamma \leq \sqrt{3\varepsilon}$. By applying Theorem~\ref{thm:Tsybakov} to~\eqref{eq:totalKL}, and in view of the property (iv), we obtain that 
\begin{equation}\label{eq:final}
\inf_{\hat \Theta}\sup_{\trueT\in \tX}\mathbb{P}\left(\FnormSize{}{\hat \Theta- \trueT}\geq  {\gamma\over 8} \min\left\{ \alpha\sqrt{d_{\text{total}}}, C^{-1/2}\sqrt{ R_{\max}d_{\max}}\right\} \right)\geq {1\over 2}\left(1-2\varepsilon-\sqrt{16 \varepsilon \over R_{\max}d_{\max}}\right).
\end{equation}
Note that $\text{Loss}(\hat \Theta, \trueT)=\FnormSize{}{\hat \Theta- \trueT}^2/d_{\text{total}}$ and $\tX\subset \tP$. By taking $\varepsilon=1/8$ and $\gamma=1/2$, we conclude from~\eqref{eq:final} that
\begin{equation*}\label{eq:prob}
\inf_{\hat \Theta}\sup_{\trueT\in \tP}\mathbb{P}\left(\text{Loss}(\hat \Theta, \trueT)\geq {1 \over 256}\min\left \{ \alpha^2,  {C^{-1}R_{\max}d_{\max}\over d_{\text{total}}}\right \}\right)\geq {1\over 2}\left({3\over 4}- {2\over R_{\max}d_{\max}} \right)\geq {1\over 8}.
\end{equation*}
This completes the proof. 
\end{proof}

\begin{proof}[Proof of Theorem~\ref{thm:completion}]

For notational convenience, we use $\MFnormSize{}{\Theta}=\sum_{\omega\in\Omega}\Theta^2_\omega$ to denote the sum of squared entries over the observed set $\Omega$, for a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$.

Following a similar argument as in the proof of Theorem~\ref{thm:rate}, we have
\begin{equation}\label{eq:Taylor2}
\logl(\Theta)=\logl(\trueT)+\langle\plogl,\ \Vec(\Theta-\trueT)\rangle+{1\over 2}\Vec(\Theta-\trueT)^T\pplogl(\check\Theta)\Vec(\Theta-\trueT),
\end{equation}
where
\begin{enumerate}
\item $\plogl$ is a length-$d_{\text{total}}$ vector with $|\Omega|$ nonzero entries, and each entry is upper bounded by $L_\alpha>0$. 
\item $\pplogl$ is a diagonal matrix of size $d_{\text{total}}$-by-$d_{\text{total}}$ with $|\Omega|$ nonzero entries, and each entry is upper bounded by $-U_{\alpha}<0$. 
\end{enumerate}

Similar to~\eqref{eq:linear} and~\eqref{eq:quadratic}, we have
\begin{equation}\label{eq:linear2}
|\langle\plogl,\ \Theta-\trueT\rangle|\leq C_2U_\alpha \sqrt{r_{\max}^{K-1}\sum_k d_k}\MFnormSize{}{\Theta-\trueT}
\end{equation}
and
\begin{equation}\label{eq:quadratic2}
\Vec(\Theta-\trueT)^T\fpplogl(\check\Theta)\Vec(\Theta-\trueT)\leq -L_\alpha \MFnormSize{}{\Theta-\trueT}^2.
\end{equation}

Combining~\eqref{eq:Taylor2}-\eqref{eq:quadratic2} with the fact that $\plogl(\hat \Theta)\geq \plogl(\trueT)$, we have
\begin{equation}\label{eq:sample}
\MFnormSize{}{\hat \Theta-\trueT}\leq {2C_2U_\alpha  r_{\max}^{(K-1)/2} \over L_\alpha} \sqrt{\sum_kd_k}.
\end{equation}
Lastly, we invoke the result regarding the closeness of $\Theta$ to its sampled version $\Theta_{\Omega}$, under the entrywise bound condition. Note that $\mnormSize{}{\hat\Theta-\trueT}\leq 2\alpha$ and $\text{rank}(\hat \Theta-\trueT)\leq 2\mr$. By Lemma~\ref{lem:Mnormbound}, $\anormSize{}{\hat \Theta-\trueT}\leq 2^{3(K-1)/2} \left({\prod r_k \over r_{\max}}\right)^{3/2}$. Therefore, the condition in Lemma~\ref{lem:convexity} holds with $\beta=2^{3(K-1)/2} \left({\prod r_k \over r_{\max}}\right)^{3/2}$. Applying Lemma~\ref{lem:convexity} to~\eqref{eq:sample} gives
\begin{align}
 \PiFnormSize{}{\hat \Theta-\trueT}^2&\leq {1\over m}\MFnormSize{}{\hat \Theta-\trueT}^2+c\beta\sqrt{\sum_k d_k\over |\Omega|}\\
 &\leq {C_2  r^{K-1}_{\max}} {\sum_k d_k \over |\Omega|}+C_1 r_{\max}^{3(K-1)/2}\sqrt{\sum_kd_k\over |\Omega|},
\end{align}
with probability at least $1-\exp(-{\sum_kd_k\over \sum_k \log d_k})$ over the sampled set $\Omega$. Here $C_1, C_2>0$ are two constants independent of the tensor dimension and rank. Therefore, 
\[
 \PiFnormSize{}{\hat \Theta-\trueT}^2\to 0,\quad \text{as}\quad {|\Omega|\over \sum_kd_k}\to \infty,
\]
provided that $r_{\max}=O(1)$.
\end{proof}

\subsection{Auxiliary lemmas}
\label{sec:lemma}

We begin with a set of technical lemmas that are useful for the proofs of the main theorems.

\begin{lem}[\cite{tomioka2014spectral}]\label{lem:tensor}
Suppose that $\tS=\entry{s_{\omega}}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is an order-$K$ tensor whose entries are independent random variables that satisfy
\[
\mathbb{E}(s_{\omega})=0,\quad \text{and} \quad\mathbb{E}(e^{ts_{\omega}})\leq e^{t^2L^2/2}.
\]
Then the spectral norm $\snormSize{}{\tS}$ satisfies that, 
\[
\snormSize{}{\tS}\leq \sqrt{{8L^2} \log (12K) \sum_k d_k +\log (2/\delta)},
\]
with probability at least $1-\delta$.
\end{lem}

\begin{rmk}
The above lemma provides the bound on the spectral norm of random tensors. The result was firstly presented in~\cite{nguyen2015tensor}, and we adopt the version from~\cite{tomioka2014spectral}.  
\end{rmk}

\begin{lem} \label{lem:noisytensor}
Suppose that $\tS=\entry{s_{\omega}}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is an order-$K$ tensor whose entries are independent random variables that satisfy
\[
\mathbb{E}(s_{\omega})=0,\quad \text{and}\quad |s_{\omega}|\leq U.
\]
Then we have
\[
\mathbb{P}\left(\snormSize{}{\tS}\geq C_2 U\sqrt{\sum_k d_k} \right)\leq \exp\left(-C_1  \log K \sum_k d_k\right)
\]
where $C_1>0$ is an absolute constant, and $C_2>0$ is a constant that depends only on $K$. 
\end{lem}

\begin{proof}  Note that the random variable $U^{-1}s_{\omega}$ is zero-mean and supported on $[-1,1]$. Therefore, $U^{-1}s_{\omega}$ is sub-Gaussian with parameter ${1-(-1)\over 2}=1$, i.e.
\[
\mathbb{E}(U^{-1}s_{\omega})=0,\quad \text{and}\quad \mathbb{E}(e^{tU^{-1}s_{\omega}})\leq e^{t^2/2}.
\]
It follows from Lemma~\ref{lem:tensor} that, with probability at least $1-\delta$, 
\[
\snormSize{}{U^{-1}\tS}\leq \sqrt{\left(c_0\log K+c_1\right) \sum_k d_k +\log (2/\delta)},
\]
where $c_0, c_1>0$ are two absolute constants. Taking $\delta=\exp (-C_1\log K \sum_k d_k)$ yields the final claim, where $C_2=c_0\log K+c_1+1>0$ is another constant.  
\end{proof}

\begin{lem} \label{lem:nuclear}
Let $\tA\in\mathbb{R}^{d_1\times\cdots\times d_K}$ be an order-$K$ tensor with Tucker $\text{rank}(\tA)=(r_1,\ldots,r_K)$. Then
\[
\nnormSize{}{\tA} \leq \sqrt{\sum_k r_k\over \max_k r_k} \FnormSize{}{\tA},
\]
where $\nnormSize{}{\cdot}$ denotes the nuclear norm of the tensor. 
\end{lem}

\begin{proof}
Without loss of generality, suppose $r_1=\min_k r_k$. Let $\tA_{(k)}$ denote the mode-$k$ matricization of $\tA$ for all $k\in[K]$. By \citet[Corollary 4.11]{wang2017operator}, and the invariance relationship between a tensor and its Tucker core~\citep[Section 6]{jiang2017tensor}, we have
\begin{equation}\label{eq:norminequality}
\nnormSize{}{\tA} \leq \sqrt{\prod_{k\geq 2} r_k \over \max_{k\geq 2} r_k} \nnormSize{}{\tA_{(1)}},
\end{equation}
where $\tA_{(1)}$ is a $d_1$-by-$\prod_{k\geq 2}d_k$ matrix with matrix rank $r_1$. Furthermore, the relationship between the matrix norms implies that $\nnormSize{}{\tA_{(1)}}\leq \sqrt{r_1}\FnormSize{}{\tA_{(1)}}=\sqrt{r_1}\FnormSize{}{\tA}$. Combining this fact with the inequality~\eqref{eq:norminequality} yields the final claim. 
\end{proof}

\begin{lem} \label{lem:inq}
Let $\tA, \; \tB$ be two order-$K$ tensors of the same dimension. Then
\[ 
|\langle \tA,\tB\rangle| \leq \snormSize{}{\tA}   \nnormSize{}{\tB}.
\]
\end{lem}

\begin{proof}
By~\citet[Proposition 3.1]{friedland2018nuclear}, there exists a nuclear norm decomposition of $\tB$, such that
\[
\tB=\sum_{r} \lambda_r \ma^{(1)}_r\otimes \cdots\otimes \ma^{(K)}_r,\quad \ma_r^{(k)}\in\mathbf{S}^{d_k-1}(\mathbb{R}),\quad \text{for all }k\in[K],
\]
and $\nnormSize{}{\tB}=\sum_{r}|\lambda_r|$. Henceforth we have
\begin{align*}
|\langle \tA,\tB\rangle|&=| \langle \tA, \sum_{r} \lambda_r \ma^{(1)}_r\otimes \cdots\otimes \ma^{(K)}_r \rangle|\leq \sum_r |\lambda_r| |\langle \tA, \ma^{(1)}_r \otimes \cdots\otimes \ma^{(K)}_r \rangle|\\
&\leq \sum_{r}|\lambda_r| \snormSize{}{\tA}= \snormSize{}{\tA}\nnormSize{}{\tB},
\end{align*}
which completes the proof. 
\end{proof}

\begin{lem}\label{lem:KLentry} Let $X,\; Y$ be two discrete random variables taking values on $L$ possible categories, with category probabilities $\{p_\ell\}_{\ell\in[L]}$ and $\{q_\ell\}_{\ell\in[L]}$, respectively.  Suppose $p_\ell$, $q_\ell>0$ for all $i\in[L]$. Then, the Kullback-Leibler (KL) divergence satisfies that
\[
\text{KL}(X||Y)\stackrel{\text{def}}{=}-\sum_{\ell\in[L]}\mathbb{P}_X(\ell)\log\left\{{\mathbb{P}_Y(\ell)\over \mathbb{P}_X(\ell)}\right\} \leq \sum_{\ell \in [L]}{(p_\ell-q_\ell)^2 \over q_\ell}.
\]
\end{lem}
\begin{proof} Using the fact $\log x\leq x-1$ for $x>0$, we have that
\begin{align}\label{eq:KL}
\text{KL}(X||Y)&=\sum_{\ell\in[L]}p_\ell\log{p_\ell\over q_\ell}\\
&\leq \sum_{\ell\in[L]} {p_\ell\over q_\ell}(p_\ell-q_\ell)\\
&=\sum_{\ell\in[L]} \left({p_\ell\over q_\ell }- 1\right)(p_\ell-q_\ell)+ \sum_{\ell\in[L]} (p_\ell-q_\ell).
\end{align}
Note that $\sum_{\ell\in[L]}(p_\ell-q_\ell)=0$. Therefore,
\[
\text{KL}(X||Y)\leq \sum_{\ell\in[L]}\left( {p_\ell\over q_\ell}-1\right)\left(p_\ell-q_\ell\right)=\sum_{\ell\in[L]}{(p_\ell-q_\ell)^2\over q_\ell}.
\]
\end{proof}

\begin{lem}~\label{lem:KL}
Let $\tY\in[L]^{d_1\times \cdots \times d_K}$ be an ordinal tensor generated from the model~\eqref{eq:model} with the link function $f$ and parameter tensor $\Theta$. Let $\mathbb{P}_{\Theta}$ denote the joint categorical distribution of $\tY|\Theta$ induced by the parameter tensor $\Theta$, where $\mnormSize{}{\Theta}\leq \alpha$. Define
\begin{equation}\label{eq:ass}
A_\alpha=\min_{\ell\in[L], |\theta|\leq \alpha}\left[f(\theta+b_\ell)-f(\theta+b_{\ell-1})\right].
\end{equation}
Then, for any two tensors $\Theta,\; \Theta^*$ in the parameter spaces, we have
\[
KL(\mathbb{P}_{\Theta}|| \mathbb{P}_{\Theta^*})\leq {2(2L-3)\over A_\alpha} \dot{f}^2(0)\FnormSize{}{\Theta-\Theta^*}^2.
\]
\end{lem}
\begin{proof} Suppose that the distribution over the ordinal tensor $\tY=\entry{y_\omega}$ is induced by $\Theta=\entry{\theta_\omega}$. Then, based on the generative model~\eqref{eq:model},
\[
\mathbb{P}(y_\omega=\ell | \theta_\omega)=f(\theta_\omega+b_{\ell})-f(\theta_\omega+b_{\ell-1}),
\]
for all $\ell\in[L]$ and $\omega\in[d_1]\times \cdots \times [d_K]$. For notational convenience, we suppress the subscribe in $\theta_\omega$ and simply write $\theta$ (and respectively, $\theta^*$). Based on Lemma~\ref{lem:KLentry} and Taylor expansion, 
\begin{align}
\text{KL}(\theta|| \theta^*) & \leq \sum_{\ell\in[L]} {\left[f(\theta+b_{\ell})-f(\theta+b_{\ell-1})-f(\theta^*+b_{\ell})+f(\theta^*+b_{\ell-1})\right]^2\over f(\theta^*+b_{\ell})-f(\theta^*+b_{\ell-1})}\\
 &\leq \sum_{\ell=2}^{L-1} {\left[\dot{f}(\eta_\ell+b_\ell)-\dot{f}(\eta_{\ell-1}+b_{\ell-1})\right]^2 \over f(\theta^*+b_\ell)-f(\theta^*+b_{\ell-1})} \left(\theta-\theta^*\right)^2+{\dot{f}^2(\eta_1+b_1) \over f(\theta^*+b_1)} (\theta-\theta^*)^2\\
 &\quad \quad\quad \quad  +{\dot{f}^2(\eta_{L-1}+b_{L-1})\over 1-f(\theta^*+b_{L-1})}(\theta-\theta^*)^2,
\end{align}
where $\eta_\ell, \eta_{\ell-1}$ fall between $\theta$ and $\theta^*$, and $\check b_{\ell}$ falls between $b_{\ell-1}$ and $b_\ell$. Therefore,
\begin{equation}\label{eq:entrywise}
\text{KL}(\theta|| \theta^*) \leq \left({4(L-2)\over A_\alpha}+ {2\over A_\alpha}\right)\dot{f}^2(0)(\theta-\theta^*)^2={2(2L-3)\over A_\alpha} \dot{f}^2(0)(\theta-\theta^*)^2,
\end{equation}
where we have used the bound~\eqref{eq:ass} and the fact that $\dot{f}(\cdot)$ peaks at zero for an unimodal and symmetric function. Now summing~\eqref{eq:entrywise} over the index set $\omega\in[d_1]\times \cdots \times [d_K]$ gives
\[
\text{KL}(\mathbb{P}_{\Theta}|| \mathbb{P}_{\Theta^*}) =\sum_{\omega\in[d_1]\times \cdots \times[d_K]} \text{KL}(\theta_\omega || \theta^*_\omega) \leq {2(2L-3)\over A_\alpha} \dot{f}^2(0)\FnormSize{}{\Theta-\Theta^*}^2.
\]
\end{proof}

\begin{rmk} In particular, let $\mathbb{P}_{\bf{0}}$ denote the distribution of $\tY|\bf{0}$ induced by the zero parameter tensor. Then we have
\[
\text{KL}(\mathbb{P}_{\Theta}||\mathbb{P}_{\bf{0}})\leq {2(2L-3)\over A_\alpha}  \dot{f}^2(0)\FnormSize{}{\Theta}^2.
\]
\end{rmk}

\begin{lem}\label{lem:construction}
Assume the same setup as in Theorem~\ref{thm:minimax}. Without loss of generality, suppose $d_1=\max_kd_k$. Define $R=\max_k r_k$ and $d_{\text{total}}=\prod_{k\in[K]} d_k$. For any constant $0\leq \gamma \leq 1$, there exist a finite set of tensors $\tX=\{\Theta_i: i=1,\ldots\}\subset \tP$ satisfying the following four properties:
\begin{enumerate}[(i)]
\item $\text{Card}(\tX)\geq 2^{Rd_1/8}+1$, where $\text{Card}$ denotes the cardinality;
\item $\tX$ contains the zero tensor $\mathbf{0}\in\mathbb{R}^{d_1\times \cdots\times d_K}$;
\item $\mnormSize{}{\Theta}\leq \gamma \min\left\{ \alpha ,\ C^{-1/2}\sqrt{Rd_1\over d_{\text{total}}} \right\} $ for any element $\Theta\in\tX$;
\item $\FnormSize{}{\Theta_i-\Theta_j}\geq {\gamma\over 4} \min\left\{ \alpha\sqrt{d_{\text{total}}},\ C^{-1/2}\sqrt{Rd_1}\right\}$ for any two distinct elements $\Theta_i,\; \Theta_j\in\tX$,
\end{enumerate}
Here $C=C(\alpha,L,f,\mb)={(4L-6)\dot{f}^2(0)\over A_\alpha }>0$ is a constant independent of the tensor dimension and rank. 
\end{lem}

\begin{proof}
Given a constant $0\leq \gamma \leq 1$, we define a set of matrices:
\[
\tC=\left\{\mM=(m_{ij})\in\mathbb{R}^{d_1\times R}: a_{ij}\in \left\{ 0,\gamma \min\left\{ \alpha , C^{-1/2}\sqrt{Rd_1\over d_{\text{total}}} \right\}\right\} ,\  \forall (i,j)\in[d_1]\times[R]\right\}.
\]
We then consider the associated set of block tensors:
\begin{align}
\tB=\tB(\tC)=\{\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}\colon& \Theta=\mA\otimes \mathbf{1}_{d_3}\otimes \cdots \otimes \mathbf{1}_{d_K}, \\
&\ \text{where}\ \mA=(\mM|\cdots|\mM|\mO) \in\mathbb{R}^{d_1\times d_2},\ \mM\in\tC\},
\end{align}
where $\mathbf{1}_d$ denotes a length-$d$ vector with all entries 1, $\mO$ denotes the $d_1\times (d_2-R\lfloor d_2/R \rfloor)$ zero matrix, and $\lfloor d_2/ R \rfloor$ is the integer part of $d_2/R$. In other words, the subtensor $\Theta(\mI, \mI,i_3, \ldots,i_K)\in\mathbb{R}^{d_1\times d_2}$ are the same for all fixed $(i_3,\ldots,i_K)\in[d_3]\times \cdots \times [d_K]$, and furthermore, each subtensor $\Theta(\mI,\mI, i_3,\ldots,i_K)$ itself is filled by copying the matrix $\mM\in\mathbb{R}^{d_1\times R}$ as many times as would fit.

By construction, any element of $\tB$, as well as the difference of any two elements of $\tB$, has Tucker rank at most $\max_k r_k\leq R$, and the entries of any tensor in $\tB$ take values in $[0,\alpha]$. Thus, $\tB\subset\tP$. By Lemma~\ref{lem:VGbound}, there exists a subset $\tX\subset \tB$ with cardinality $\text{Card}(\tX)\geq 2^{Rd_1/8}+1$ containing the zero $d_1\times \cdots \times d_K$ tensor, such that, for any two distinct elements $\Theta_i$ and $\Theta_j$ in $\tX$, 
\[
\FnormSize{}{\Theta_i-\Theta_j}^2 \geq {Rd_1\over 8} \gamma^2\min\left\{ \alpha, {C^{-1}Rd_1 \over d_{\text{total}}}\right\} \lfloor {d_2\over R} \rfloor \prod_{k\geq 3}d_k\geq {\gamma^2\min\left\{ \alpha^2 d_{\text{total}}, C^{-1}Rd_1\right\}  \over 16}.
\]
In addition, each entry of $\Theta\in\tX$ is bounded by $\gamma \min\left\{ \alpha , C^{-1/2}\sqrt{Rd_1\over d_{\text{total}}}\right\} $. Therefore the Properties (i) to (iv) are satisfied. 
\end{proof}

\begin{lem}[Varshamov-Gilbert bound]\label{lem:VGbound}
Let $\Omega=\{(w_1,\ldots,w_m)\colon w_i\in\{0,1\}\}$. Suppose $m>8$. Then there exists a subset $\{w^{(0)},\ldots,w^{(M)}\}$ of $\Omega$ such that $w^{(0)}=(0,\ldots,0)$ and
\[
\zeronormSize{}{w^{(j)}-w^{(k)}}\geq {m\over 8},\quad \text{for} \ 0\leq j<k\leq M,
\]
where $\zeronormSize{}{\cdot}$ denotes the Hamming distance, and $M\geq 2^{m/8}$. 
\end{lem}

\begin{thm}[\cite{tsybakov2009introduction}]\label{thm:Tsybakov} 
Assume that a set $\tX$ contains element $\Theta_0, \Theta_1, \ldots,\Theta_M$ ($M\geq 2$) such that
\begin{itemize}
\item $d(\Theta_j,\ \Theta_j)\geq 2s>0$, $\forall 0\leq j\leq k\leq M$;
\item $\mathbb{P}_j\ll\mathbb{P}_0$, $\forall j=1,\ldots,M$, and 
\[
{1\over M}\sum_{j=1}^M KL(\mathbb{P}_j||\mathbb{P}_0)\leq \alpha \log M
\]
where $d\colon \tX\times \tX\mapsto [0,+\infty]$ is a semi-distance function, $0<\alpha<{1/8}$ and $P_j=P_{\Theta_j}$, $j=0,1\ldots,M$. Then
\[
\inf_{\hat \Theta}\sup_{\Theta\in\tX} \mathbb{P}_{\Theta}(d(\hat \Theta, \Theta)\geq s)\geq {\sqrt{M}\over 1+\sqrt{M}}\left(1-2\alpha-\sqrt{2\alpha\over \log M} \right)>0.
\]
\end{itemize}

\end{thm}
\begin{defn}[\cite{ghadermarzy2019near}] 
Define $T_{\pm}=\{\tT\in\{\pm\}^{d_1\times \cdots \times d_K}\colon \text{rank}(\tT)=1\}$. The atomic M-norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
\begin{align}
\anormSize{}{\Theta}&=\inf\{t>0: \Theta\in t\text{conv}(T_{\pm})\}\\
&=\inf\left\{\sum_{\tX\in T_{\pm}}c_x\colon \Theta=\sum_{\tX\in T_{\pm}} c_x\tX, \ c_x>0\right\}.
\end{align}
\end{defn}

\begin{lem}[\cite{ghadermarzy2019near}]\label{lem:Mnormbound}
Let $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ be an order-$K$, rank-$(r_1,\ldots,r_K)$ tensor. Then
\[
\mnormSize{}{\Theta}\leq \anormSize{}{\Theta}\leq \left(\prod_k r_k \over r_{\max}\right)^{3\over 2} \mnormSize{}{\Theta}.
\]
\end{lem}

\begin{lem}[\cite{ghadermarzy2019near}]\label{lem:convexity}
Define $\mathbb{B}_{M}(\beta)=\{\Theta\in \mathbb{R}^{d_1\times \cdots \times d_K}\colon \anormSize{}{\Theta}\leq \beta \}$.  Let $\Omega\subset[d_1]\times\cdots \times [d_K]$ be a random set with $m=|\Omega|$, and assume that each entry in $\Omega$ is drawn with replacement from $[d_1]\times\cdots\times[d_K]$ using probability $\Pi$. Define
\[
\PiFnormSize{}{\Theta}^2={1\over m}\mathbb{E}_{\Omega\in\Pi}\MFnormSize{}{\Theta}^2.
\]
Then, there exists a universal constant $c>0$, such that, with probability at least $1-\exp\left(-{\sum_kd_k \over \sum_k \log d_k} \right)$ over the sampled set $\Omega$, 
\[
{1\over m}\MFnormSize{}{\Theta}^2 \geq \PiFnormSize{}{\Theta}^2-c\beta\sqrt{\sum_k d_k\over m}
\]
holds uniformly for all $\Theta\in\mathbb{B}_M(\beta)$. 
\end{lem}

\bibliographystyle{plainnat}
\bibliography{tensor_wang}

\end{document}

### 2020-01-21 ICLM/miaoyan/ordinalT_supp.tex
update draft

@@
 \end{equation}
-where we have used the bound~\eqref{eq:ass} and the fact that $\dot{f}(\cdot)$ peaks at zero for an unimodal and symmetric function.
+where we have used Taylor expansion, the bound~\eqref{eq:ass}, and the fact that $\dot{f}(\cdot)$ peaks at zero for an unimodal and symmetric function.
 Now summing~\eqref{eq:entrywise} over the index set $\omega\in[d_1]\times \cdots \times [d_K]$ gives

### 2020-01-21 ICLM/miaoyan/ordinalT.tex
update draft

@@
 
+\usepackage{amssymb}
+\usepackage{pifont}
+\newcommand{\cmark}{\ding{51}}%
+\newcommand{\xmark}{\ding{55}}%
+
 \begin{document}
@@
 Our error bound enjoys a faster convergence rate than existing results, and we show that the proposed estimator is minimax optimal under the considered model. 
-An efficient completion method is further developed, which guarantees the consistent recovery of an order-$K$ $(d,\ldots,d)$-dimensional tensor using only $O(Kd)$ noisy, quantized observations.
+An efficient completion method is further developed, which guarantees consistent recovery of an order-$K$ $(d,\ldots,d)$-dimensional tensor using only $O(Kd)$ noisy, quantized observations.
 We demonstrate the outperformance of our approach over previous methods through applications to the tasks of clustering and collaborative filtering.
@@
 The present work focuses on ordinal tensors; namely, the tensor entries are categorical variables with a natural ordering among the categories.
-{\color{red}(Prior work and our contribution. Make a table of comparison.)}
+{\color{red}(Prior work and our contribution. See Table~\ref{tab:compare} for comparison.)}
+
+\begin{table*}[http]
+\begin{center}
+\includegraphics[width=17cm]{compare.pdf}
+\end{center}
+\vspace{-.5cm}
+\caption{Comparison of various low-rank estimation methods based on categorical observations.
+For an order-$K$ dimensional-$(d,\ldots,d)$ ordinal tensor with $L$-level observations, we report the error bound in the recovered signal (for tensor denoising) and required sample complexity (for tensor completion) as functions of tensor dimension.
+{\color{red}(font in the table seems different)}}~\label{tab:compare}
+\end{table*}
 
@@
 \begin{equation}\label{eq:space}
-\tP=\left\{\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K} \colon \text{rank}(\Theta)\leq \mr,\ \mnormSize{}{\Theta}\leq \alpha\right\}.
+\small \tP=\left\{\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K} \colon \text{rank}(\Theta)\leq \mr,\ \mnormSize{}{\Theta}\leq \alpha\right\}.
 \end{equation}
@@
 This suggests that the local optimality itself is not necessarily a severe concern in our context, as long as the objective value at the estimator is large enough.
-{\color{red}In Section~\ref{algorithm}, we will provide an algorithm-dependent rate quantifying the interplay between the statistical and computational errors.}
+In Section~\ref{sec:simulation}, we will perform empirical studies to assess the algorithmic stability and its impact to statistical efficiency.
 
@@
 The result is based on the information theory, and therefore applies to all estimators in $\tP$, including but not limited to $\hat \Theta$ in~\eqref{eq:estimator}.
-We show that this lower bound nearly matches the upper bound in~\eqref{eq:rate}, as the tensor dimension increases.
+We show that this lower bound matches the upper bound in~\eqref{eq:rate} as a function of tensor dimensions.
 
@@
 Theorem~\ref{thm:completion} shows that our estimator achieves consistency when sample size is greater than $O(\sum_k d_k)$.
-{\color{red} add remarks comparing with prior work, and with the degree of freedom}
+{\color{red} add remarks comparing with prior work, and with the degree of freedom.}
 
@@
 In each simulation study, we report the summary statistics across $n_{\text{sim}}$ = 30 replications.
+
 \subsection{Finite-sample performance}\label{sec:simulation}

### 2020-01-22 ICML/miaoyan/compare.tex
upload code

\documentclass[11pt]{article}
\usepackage[paperheight=1.3in,paperwidth=7.4in,margin=0.01in]{geometry}

\usepackage{setspace}
\usepackage{amsmath,amssymb}
\usepackage{amsthm}
\usepackage{fancybox}
\usepackage{url}
\usepackage{enumitem}
\usepackage{multirow}
\usepackage{color}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{comment}
\usepackage{bm}
\usepackage{mathtools}
\mathtoolsset{showonlyrefs=true}

\usepackage{natbib}
\usepackage{xr}
\externaldocument{ordinalT}
%\usepackage{icml2020}

\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

%\input macros.tex

\def\@normalsize{\@setsize\normalsize{11pt}\xpt\@xpt}

\begin{document}
\begin{table}
\normalsize
\begin{tabular}{l|ccc}
&Bhaskar [2016]&Ghadermarzy et al.\ [2018] &This paper\\
\hline 
Higher-order tensors ($K\geq 3$) & \xmark&\cmark& \cmark\\
\hline
Multi-level categories ($L\geq 3$)& \cmark& \xmark&\cmark\\
 \hline
 Error rate for tensor denoising&$d^{-1}$ for $K=2$& $d^{-(K-1)/2}$&$d^{-(K-1)}$\\
\hline
Optimal guarantee under exact low-rankness& unknown & \xmark & \cmark\\
\hline
Sample complexity for tensor completion&$d^2$ for $K=2$& $Kd$&$Kd$\\
\hline
\end{tabular}
\end{table}

%\bibliography{tensor_wang}
\bibliographystyle{icml2020}

\end{document}

### 2020-01-22 code/experiment.R
update code

@@ -1,12 +1,12 @@
-source("function.R")
+source("functions.R")
 ########## brain tensor analysis ##########
-source("B_functions.R")
 load("../data/dti_brain.RData")
 ##ttnsr=tensor[,,attr[,2]=="22-25"] ## subset of individuals
 ttnsr=tensor
 d=dim(ttnsr)
 
-r=c(24,24,7)
+r=c(25,25,7)
+### random initilization
 A_1 = randortho(d[1])[,1:r[1]]
 A_2 = randortho(d[2])[,1:r[2]]
 A_3 = randortho(d[3])[,1:r[3]]
@@ -16,10 +16,70 @@ tic()
 result = fit_ordinal_MM(ttnsr,C,A_1,A_2,A_3,omega=TRUE,alph = TRUE)
 toc()
 
+########## Music analysis #######################
+source("functions.R")###
+load("../data/InCar_Music.RData")
+ttnsr=tensor
+ttnsr[ttnsr==-1]=NA
+missing=which(is.na(ttnsr)==F)
+d=dim(ttnsr)
+set.seed(1)
+r=c(2,2,2) ## need to select via BIC ..
+alpha=100
+A_1 = randortho(d[1])[,1:r[1]]
+A_2 = randortho(d[2])[,1:r[2]]
+A_3 = randortho(d[3])[,1:r[3]]
+C = rand_tensor(modes = r)
+C=C/max(abs(ttl(C,list(A_1,A_2,A_3),ms=1:3)@data))*alpha/10
+
+############# experiement ##########################
+mask=sample(1:2884,2884*0.1,replace=T)
+index=which(is.na(ttnsr)==F)[mask]
+true=ttnsr[index]
+ttnsr[index]=NA
+##### Method 1: continuous tucker #########
+tic()
+result = tucker_missing(ttnsr,r,TRUE,alpha=alpha)
+toc()
+fitted=ttl(result$Z,list(result$U[[1]],result$U[[2]],result$U[[3]]),ms=1:3)
+mean((true-fitted@data[index])^2) #MSE= 18.96803
+mean(abs(true-fitted@data[index])) #MAD= 2.589925
+
+###### Method 2: ordinal tucker #########
+tic()
+result = fit_ordinal(ttnsr,C,A_1,A_2,A_3,omega=TRUE,alph = alpha)
+toc()
[…]

### 2020-01-22 code/functions.R

@@ -2,87 +2,75 @@ library(MASS)
 library(rTensor)
 library(pracma)
 library(doParallel)
-
-
+epsilon=10^-4
 
 ########### simulation ordinal tensors based on logist model with arbitrary k
 realization = function(theta,omega){
-  k = length(omega)
   theta=as.tensor(theta)
   thet <- c(theta@data)
-  p = matrix(nrow = length(thet),ncol = k)
-  for (i in 1:k) {
-    p[,i] = logistic(thet + omega[i])
-  }
-  p =  cbind(p,rep(1,length(thet)))-cbind(rep(0,length(thet)),p)
+  p = theta_to_p(thet,omega)
   for (j in 1:length(thet)) {
     thet[j] <-  sample(1:(k+1),1,prob = p[j,])
   }
   return(as.tensor(array(thet,dim =theta@modes)))
 }      
 
-# estimate the variable which has the largest probability
-estimation = function(theta,omega){
-  k = length(omega)
+
+estimation = function(theta,omega,type){
+  if(is.matrix(theta)){
+  thet=c(theta)
+  }else{
   theta=as.tensor(theta)
   thet <- c(theta@data)
-  p = matrix(nrow = length(thet),ncol = k)
-  for (i in 1:k) {
-    p[,i] = logistic(thet + omega[i])
   }
-  p =  cbind(p,rep(1,length(thet)))-cbind(rep(0,length(thet)),p)
-  for (j in 1:length(thet)) {
-    thet[j] <-  which.max(p[j,])
+  p = theta_to_p(thet,omega)
+  if(type=="max"){  # score prediction based on the mode 
+  for (j in 1:length(thet)) thet[j] <-  which.max(p[j,])
+  }else if(type=="mean"){ # score prediction based on the mean 
+  for (j in 1:length(thet)) thet[j] <-  sum(p[j,]*(1:(k+1))) ## why rounding to integer?
+  }else if(type=="median"){# score prediction based on the median 
+  for (j in 1:length(thet)) thet[j] <-  which(c(cumsum(p[i,]),1)>=0.5)[1] ## median
   }
-  return(as.tensor(array(thet,dim =theta@modes)))
[…]

### 2020-01-22 code/functions_miaoyan.R

@@ -1,0 +1,649 @@
+library(MASS)
+library(rTensor)
+library(pracma)
+library(doParallel)
+epsilon=10^-4
+
+
+########### simulation ordinal tensors based on logist model with arbitrary k
+realization = function(theta,omega){
+  k = length(omega)
+  theta=as.tensor(theta)
+  thet <- c(theta@data)
+  p = matrix(nrow = length(thet),ncol = k)
+  for (i in 1:k) {
+    p[,i] = logistic(thet + omega[i])
+  }
+  p =  cbind(p,rep(1,length(thet)))-cbind(rep(0,length(thet)),p)
+  for (j in 1:length(thet)) {
+    thet[j] <-  sample(1:(k+1),1,prob = p[j,])
+  }
+  return(as.tensor(array(thet,dim =theta@modes)))
+}      
+
+
+
+
+########### cost function  with arbitrary k ###########
+h1 = function(A_1,W1,ttnsr,omega,type="ordinal"){
+  k = length(omega)
+  theta =W1%*%c(A_1)
+  if(type=="Gaussian"){
+      l=sqrt(sum((theta[is.na(ttnsr)==F]-ttnsr[is.na(ttnsr)==F])^2))
+      return(l)  
+  }
+  p = matrix(nrow = length(theta),ncol = k)
+  for (i in 1:k) {
+    p[,i] = as.numeric(logistic(theta + omega[i]))
+  }
+  p =  cbind(p,rep(1,length(theta)))-cbind(rep(0,length(theta)),p)
+  l = lapply(1:(k+1),function(i) -log(p[which(c(ttnsr)==i),i]))
+  return(sum(unlist(l)))
+}
+
+hc = function(A_1,A_2,A_3,C,ttnsr,omega,type="ordinal"){
+  k = length(omega)
+  theta = c(ttl(C,list(A_1,A_2,A_3),ms=1:3)@data)
+  if(type=="Gaussian"){
+      l=sqrt(sum((theta[is.na(ttnsr)==F]-ttnsr[is.na(ttnsr)==F])^2))
+      return(l)
+  }
+  p = matrix(nrow = length(theta),ncol = k)
+  for (i in 1:k) {
+    p[,i] = as.numeric(logistic(theta + omega[i]))
+  }
+  p =  cbind(p,rep(1,length(theta)))-cbind(rep(0,length(theta)),p)
[…]

### 2020-01-22 code/ouputdata/functions_old.R

@@ -1,0 +1,467 @@
+library(MASS)
+library(rTensor)
+library(pracma)
+library(doParallel)
+
+
+
+########### simulation ordinal tensors based on logist model with arbitrary k
+realization = function(theta,omega){
+  k = length(omega)
+  theta=as.tensor(theta)
+  thet <- c(theta@data)
+  p = matrix(nrow = length(thet),ncol = k)
+  for (i in 1:k) {
+    p[,i] = logistic(thet + omega[i])
+  }
+  p =  cbind(p,rep(1,length(thet)))-cbind(rep(0,length(thet)),p)
+  for (j in 1:length(thet)) {
+    thet[j] <-  sample(1:(k+1),1,prob = p[j,])
+  }
+  return(as.tensor(array(thet,dim =theta@modes)))
+}      
+
+
+
+
+########### cost function  with arbitrary k ###########
+h1 = function(A_1,W1,ttnsr,omega){
+  k = length(omega)
+  thet =W1%*%c(A_1)
+  p = matrix(nrow = length(thet),ncol = k)
+  for (i in 1:k) {
+    p[,i] = as.numeric(logistic(thet + omega[i]))
+  }
+  p =  cbind(p,rep(1,length(thet)))-cbind(rep(0,length(thet)),p)
+  l = lapply(1:(k+1),function(i) -log(p[which(c(ttnsr)==i),i]))
+  return(sum(unlist(l)))
+}
+
+hc = function(A_1,A_2,A_3,C,ttnsr,omega){
+  k = length(omega)
+  thet = c(ttl(C,list(A_1,A_2,A_3),ms=1:3)@data)
+  p = matrix(nrow = length(thet),ncol = k)
+  for (i in 1:k) {
+    p[,i] = as.numeric(logistic(thet + omega[i]))
+  }
+  p =  cbind(p,rep(1,length(thet)))-cbind(rep(0,length(thet)),p)
+  l = lapply(1:(k+1),function(i) -log(p[which(c(ttnsr)==i),i]))
+  return(sum(unlist(l)))
+}
+
+# hc(A_1,A_2,A_3,C,ttnsr,omega)
+########### gradient with arbitrary k ###########
+g1 = function(A_1,W,ttnsr,omega){
+  k = length(omega)
+  thet =W%*%c(A_1)
[…]

### 2020-01-22 ICML/miaoyan/ordinalT.tex
update

--- 2020-01-21 ICLM/miaoyan/ordinalT.tex
@@
 \twocolumn[
-\icmltitle{Tensor denoising and completion based on ordinal observations}
+\icmltitle{Optimal tensor denoising and completion based on ordinal observations}
 
@@
 An efficient completion method is further developed, which guarantees consistent recovery of an order-$K$ $(d,\ldots,d)$-dimensional tensor using only $O(Kd)$ noisy, quantized observations.
-We demonstrate the outperformance of our approach over previous methods through applications to the tasks of clustering and collaborative filtering.
+We demonstrate the outperformance of our approach over previous methods on the tasks of clustering and collaborative filtering.
 
@@
 \begin{table*}[http]
-\begin{center}
-\includegraphics[width=17cm]{compare.pdf}
-\end{center}
-\vspace{-.5cm}
+\centerline{\includegraphics[width=16cm]{compare.pdf}}
+\vspace{-.4cm}
 \caption{Comparison of various low-rank estimation methods based on categorical observations.
 For an order-$K$ dimensional-$(d,\ldots,d)$ ordinal tensor with $L$-level observations, we report the error bound in the recovered signal (for tensor denoising) and required sample complexity (for tensor completion) as functions of tensor dimension.
-{\color{red}(font in the table seems different)}}~\label{tab:compare}
+{\color{red}(font in the table seems odd...)}}~\label{tab:compare}
 \end{table*}
@@
 Assume the same set-up as in Theorem~\ref{thm:rate}, and $d_{\max}=\max_k d_k \geq 8$.
-Let $\inf_{\hat \Theta}$ denote the infimum over all estimators $\hat \tA\in\tP$ based on the ordinal tensor observation $\tY\in[L]^{d_1\times \cdots \times d_K}$.
+Let $\inf_{\hat \Theta}$ denote the infimum over all estimators $\hat \Theta\in\tP$ based on the ordinal tensor observation $\tY\in[L]^{d_1\times \cdots \times d_K}$.
 Then, under the model~\eqref{eq:model},
@@
 Let $\hat \Theta$ be the solution to~\eqref{eq:estimator}, and assume $r_{\max}=O(1)$.
-Then, with very high probability over the sampled set $\Omega$,
+Then, with very high probability over the sampled set $\Omega$ ({\color{red} not over $\tY$?}),
 \[

### 2020-01-22 ICML/chanwoo/ICMLdraft.tex
Add files via upload

%%%%%%%% ICML 2020 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%

\documentclass{article}

% Recommended, but optional, packages for figures and better typesetting:
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{booktabs} % for professional tables

% hyperref makes hyperlinks in the resulting PDF.
% If your build breaks (sometimes temporarily if a hyperlink spans a page)
% please comment out the following usepackage line and replace
% \usepackage{icml2020} with \usepackage[nohyperref]{icml2020} above.
\usepackage{hyperref}

% Attempt to make hyperref and algorithmic work together better:
\newcommand{\theHalgorithm}{\arabic{algorithm}}

% Use the following line for the initial blind version submitted for review:
\usepackage{icml2020}

% If accepted, instead use the following line for the camera-ready submission:
%\usepackage[accepted]{icml2020}

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Ordinal tensor denoising and completion}

\usepackage{multirow}
\usepackage{graphicx}
%\usepackage[utf8]{inputenc} % allow utf-8 input
%\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
%\usepackage{booktabs}       % professional-quality tables
\usepackage{amsmath,amssymb}
\usepackage{amsthm}    % blackboard math symbols
%\usepackage{nicefrac}       % compact symbols for 1/2, etc.
%\usepackage{microtype}      % microtypography
\usepackage{bm}
%\usepackage{subfig}
%\usepackage[english]{babel}
%\usepackage{algorithm}
%\usepackage{appendix}
\usepackage{mathtools}
\mathtoolsset{showonlyrefs}

\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{pro}{Property}
\newtheorem{assumption}{Assumption}

\theoremstyle{definition}
\newtheorem{defn}{Definition}
\newtheorem{cor}{Corollary}
\newtheorem{example}{Example}
\newtheorem{rmk}{Remark}

\usepackage{dsfont}
%\usepackage{algpseudocode,algorithm}
%\algnewcommand\algorithmicinput{\textbf{Input:}}
%\algnewcommand\algorithmicoutput{\textbf{Output:}}
%\algnewcommand\INPUT{\item[\algorithmicinput]}
%\algnewcommand\OUTPUT{\item[\algorithmicoutput]}
%\DeclareMathOperator*{\minimize}{minimize}

\newcommand*{\KeepStyleUnderBrace}[1]{%f
  \mathop{%
    \mathchoice
    {\underbrace{\displaystyle#1}}%
    {\underbrace{\textstyle#1}}%
    {\underbrace{\scriptstyle#1}}%
    {\underbrace{\scriptscriptstyle#1}}%
  }\limits
}

\input macros.tex

\begin{document}

\twocolumn[
\icmltitle{Tensor denoising and completion based on ordinal observations}

% It is OKAY to include author information, even for blind
% submissions: the style file will automatically remove it for you
% unless you've provided the [accepted] option to the icml2020
% package.

% List of affiliations: The first argument should be a (short)
% identifier you will use later to specify author affiliations
% Academic affiliations should list Department, University, City, Region, Country
% Industry affiliations should list Company, City, Region, Country

% You can specify symbols, otherwise they are numbered in order.
% Ideally, you should not use this facility. Affiliations will be numbered
% in order of appearance and this is the preferred way.
\icmlsetsymbol{equal}{*}

\begin{icmlauthorlist}
\icmlauthor{Chanwoo Lee}{equal,to}
\icmlauthor{Miaoyan Wang}{equal,to}
\end{icmlauthorlist}

\icmlaffiliation{to}{Department of Statistics, University of Wisconsin at Madison, USA}

\icmlcorrespondingauthor{Miaoyan Wang}{miaoyan.wang@wisc.edu.com}

% You may provide any keywords that you
% find helpful for describing your paper; these are used to populate
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{Higher-order tensors, ordinal data, tensor denoising, tensor completion}

\vskip 0.3in
]

% this must go after the closing bracket ] following \twocolumn[ ...

% This command actually creates the footnote in the first column
% listing the affiliations and the copyright notice.
% The command takes one argument, which is text to display at the start of the footnote.
% The \icmlEqualContribution command is standard text for equal contribution.
% Remove it (just {}) if you do not need this facility.

%\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.

\begin{abstract}

\end{abstract}

\section{Introduction}
Multidimensional array data, a.k.a. a tensor, appears in a huge variety of applications including recommendation systems \citep{Kutty2012APR,Adomavicius2008ContextawareRS,Sun2015ProvableST},
social networks \citep{Sun2009MultiVisCS,Nickel2011ATM},
genomics \citep{Wang2017ThreewayCO},
neuroimaging (EEG, fMRI) \citep{Miwakeichi_decomposingeeg}
and signalprocessing \citep{Sidiropoulos2000ParallelFA,Cichocki2015TensorDF}.
Instead of unfolding those data tensors into matrices where many analysis methods have been proposed, we preserve its inherent multi-modal structure. Studying tensor data while respecting the structure allows us to examine complex interactions among tensor entries. Thereby we can provide extra more interpretation that cannot be addressed by traditional matrix analysis. Also, It has been shown that the tensor preserving analysis improves performance \citep{Zare2018ExtensionOP,Wang2018LearningFB}.
With those reasons, there is a growing need to develop dimension reduction method without losing multi modal structure. In the line of the attempts, a number of tensor decomposition methods have been proposed in many applications. CANDECOMP/PARAFAC (CP) decomposition was first introduced \citep{doi:10.1002/sapm19287139}
and revitalized in psychometrics \citep{Harshman1970FoundationsOT} and in linguistics \citep{Smilde2004MultiwayAW}. The Tucker decomposition was proposed in psychometrics \citep{tucker64extension,RePEc:spr:psycho:v:31:y:1966:i:3:p:279-311}.

Classical tensor completion with those decompositions has treated the entries of data as real-valued.
In many cases, However, we encounter data of which the entries are not real-valued but discrete or quantized i.e. binary-valued or multiple-valued. For example, many survey data takes the integer values. To be specific, the data in the Netflix problem has the 3 modes of `user',`movie' and `date of grade'. The entries of the data are the gradings from the users which take integer value from 1 to 5. Also, there are many cases that the data are quantized in real application. In signal processing, the data are frequently rounded or truncated so that only integer values are available. Also, in graph theory, adjacency matrix can be labeled as from 1 to 3 taking 3 when pairs of vertexes have strong connection and giving 1 when two vertexes have the weak connection according to a given threshold. If we add one more mode on adjacency matrix such as `context' or `individual', the data turns into tensor data with 3 integer values.

Therefore, performance improvement can be achieved when the observations are treated as discrete value not as continuous value.
In matrix case, there has been many achievements to complete matrix for discrete cases: Models for the case of binary or 1-bit were introduced and studied \citep{Davenport20121BitMC,Bhaskar20151bitMC}. Furthermore,  \citet{Bhaskar2016ProbabilisticLM} suggested matrix completion method for general ordinal observations.
In tensor case, however, only binary tensor has gotten an attention and achieved performance improvement using binary tensor decomposition methods \citep{Hore2016TensorDF,Wang2018LearningFB,Hong2018GeneralizedCP,Hu2018TrainingBW}. Accordingly, a general method for the data which has more than 2 ordered label is needed.

We organized this paper as follows. In section 2, we discuss detailed assumptions and descriptions about our probabilistic model in \eqref{multinomial}. Also, we suggest the estimation method for the latent parameters and related algorithm. In section 3, we provide the statistical properties of the upper, lower bounds and the phase-transition. We then provide the numerical experiments. Our model is applied to  real-world data to check validity and performance in section 5. Finally, we wrap up the paper with a discussion.

\section{Preliminaries}
Let $\tY\in\mathbb{R}^{d_1\times \cdots \times d_K}$ denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional tensor. We use $y_\omega$ to denote the tensor entry indexed by $\omega$, where $\omega\in[d_1]\times\cdots\times[d_K]$.  The Frobenius norm of $\tY$ is defined as $\FnormSize{}{\tY}=\sum_\omega y^2_\omega$ and the infinity norm of $\tY$ is defined as $\mnormSize{}{\tY}=\max_{\omega}|y_\omega|$. We use $\tY_{(k)}$ to denote the unfolded matrix of size $d_k$-by-$\prod_{i\neq k}d_k$, obtained by reshaping the tensor along the mode-$k$. The Tucker rank of $\tY$ is defined as a length-$K$ vector $\mr = (r_1,\ldots,r_K)$, where $r_k$ is the rank of matrix $\tY_{(k)}$ for all $k \in[L]$. We say that an event $A$ occurs ``with very high probability'' if $\mathbb{P}(A)$ tends to 1 faster than any polynomial of tensor dimension $d_{\min}=\min\{d_1,\ldots,d_K\}$ as $d_{\min}\to\infty$.

We use lower-case letters (e.g., $a$, $b$, $c$) for scalars/vectors, upper-case boldface letters (e.g., $\mA$, $\mB$, $\mC$) for matrices, and calligraphy letters (e.g., $\tA$, $\tB$, $\tC$) for tensors of order three or greater. For ease of notation, we allow the basic arithmetic operators (e.g., $\leq, +, -$) to be applied to pairs of vectors in an element-wise manner. We use the shorthand $[n]$ to denote the $n$-set $\{1,\ldots,n\}$ for $n \in N_{+}$.

\section{Model}
\subsection{Low-rank cumulative model}
For the K-mode ordinal tensor $\mathcal{Y} = [\![ y_\omega]\!]\in[L]^{d_1\times\cdots\times d_K}$, we assume that its entries are realizations of independent multinomial random variables, such that
\begin{equation}
    \begin{aligned}
        \label{mainmodel}
        \mathbb{P}(y_\omega=l) = f_l(\theta_\omega), \quad \omega\in [d_1]\times\cdots\times[d_K].
    \end{aligned}
\end{equation}
In this model, a twice differentiable function $f_l : \mathbb{R}\rightarrow [0,1]$ with $l\in [L]$ is strictly increasing and satisfies $\sum_l f_l(\theta) = 1$ for a fixed $\theta$.
The tensor $\Theta = [\![\theta_\omega]\!]\in \mathbb{R}^{d_1\times\cdots\times d_K}$
is a hidden parameter which we are interested in. We assume the parameter tensor $\Theta$ is continuous value and admits a rank $\mr = (r_1,\cdots,r_K)$ Tucker decomposition,
\begin{equation}
    \begin{aligned}
        \label{tucker}
        \Theta = \mathcal{C}\times_1\mM_1\cdots\times_N\mM_N,
    \end{aligned}
\end{equation}
where $\mathcal{C}\in \mathbb{R}^{r_1\times\cdots\times r_K}$ is a core tensor and $\mM_k\in \mathbb{R}^{d_k\times r_k}$, for $n \in [L]$ is a factor matrix. We can get information about the influence of each mode by checking factor matrices.

The tensor model \eqref{mainmodel} has an equivalent representation as a latent model with K-level quantization. \citep{Davenport20121BitMC,Lan2014MatrixRF,Bhaskar20151bitMC,Cai2013AMC} where
$\mathcal{Y} = [\![ y_\omega]\!]$ is a quantized value such that,
\begin{equation}
    \label{latentmd}
\begin{aligned}
    y_\omega = \mathcal{Q}(\theta_\omega+\epsilon_\omega),\quad \omega\in [d_1]\times\cdots[d_K],
\end{aligned}
\end{equation}
where $\mathcal{E} = [\![\epsilon_\omega]\!]$ is a noise tensor of $i.i.d.$ from cumulative distribution function $\Phi(\cdot)$ and the function $\mathcal{Q}:\mathbb{R}\rightarrow [L]$ is a quantizer having the following rule.
\begin{equation}
    \label{qfunction}
    \begin{aligned}
        \mathcal{Q}(x) = l,\quad \text{ if } b_{l-1}< x \leq b_{l},\quad l\in [L],
    \end{aligned}
\end{equation}
where $\mb = (b_1,\cdots,b_{L-1})$ is a cutpoints vector such that $ -\infty = b_0<b_1<\cdots<b_L = \infty$. That is, the entries of observed tensor $\mathcal{Y}$ fall in category $l$ when the associated entries of the latent tensor $\Theta+\mathcal{E}$ fall in the $l$th interval of values. Then we have
\begin{equation}
    \label{cumulative}
    \begin{aligned}
        f_l(\theta_\omega) &= \mathbb{P}(y_\omega = l)\\
        &= \Phi(b_l-\theta_\omega)-\Phi(b_{l-1}-\theta_\omega).
    \end{aligned}
\end{equation}
We can diversify our model by the choices of $\Phi(\cdot)$, or equivalently the distribution of $\mathcal{E}$ with given $\mb$.
The followings are 2 common choices of $\Phi(\cdot)$.
\begin{example}[Logistic link/Logistic noise] The logistic model is represented by \eqref{mainmodel} with $f_l(\theta) = \Phi_{log}(\frac{b_l-\theta}{\sigma})-\Phi_{log}(\frac{b_{l-1}-\theta}{\sigma})$  where $\Phi_{log}(x/\sigma) = (1+e^{-x/\sigma})$. Equivalently, the noise $\epsilon_\omega$ in \eqref{latentmd} follows i.i.d. logistic distribution with the scale parameter $\sigma.$
\end{example}

\begin{example}[Probit link/Gaussian noise] The probit model is represented by \eqref{mainmodel} with $f_l(\theta) = \Phi_{norm}(\frac{b_l-\theta}{\sigma})-\Phi_{norm}(\frac{b_{l-1}-\theta}{\sigma})$ where $\Phi_{norm}$ is the cumulative distribution function of the standard Gaussian. Equivalently, the noise $\epsilon_\omega$ in \eqref{latentmd} follows i.i.d.$N(0,\sigma^2)$.
\end{example}

\subsection{Rank-constrained likelihood-based estimation}
Our goal is to estimate unknown parameter tensor $\Theta$ and a cutpoints vector $\mb$ from observed tensor $\mathcal{Y}$ using a constrained likelihood approach. With a little abuse of notation, we use $\Omega$ to denote either the full index set $\Omega = [d_1]\times\cdots\times[d_K]$ or a random subset indeuced from the subsampling distribution. The log-likelihood function for \eqref{mainmodel} is
\begin{equation}
    \label{likelihood}
    \begin{aligned}
    \mathcal{L}_{\mathcal{Y},\Omega}(\Theta,\mb) = \sum_{\omega\in\Omega}\Big[\sum_{l\in[L]}\log(f_l(\theta_\omega))\mathds{1}_{\{y_\omega = l\}}\Big],
    \end{aligned}
\end{equation}
where the cutpoints vector $\mb$ is implicitly contained in the function $f_l$.
Considering the Tucker structure in \eqref{tucker}, we have the following constrained optimization problem.
\begin{equation}
    \label{eq:optimization}
    \begin{aligned}
        &\max_{(\Theta, \mb)\in \mathcal{D}\times\mathcal{B}}\mathcal{L}_{\mathcal{Y},\Omega}(\Theta,\mb),\text{ where }\\
        &\mathcal{D} = \{\Theta\in \mathbb{R}^{d_1\times\cdots\times d_K}: \rank(\Theta) = \bm{r} \text{ and } \mnormSize{}{\Theta}\leq \alpha \}\\
        &\mathcal{B} = \{\mb\in\mathbb{R}^{L-1}:  b_1<\cdots < b_{L-1}\},
    \end{aligned}
\end{equation}
for a given rank $\bm r\in \mathbb{N_+}^K$ and a bound $\alpha \in \mathbb{R}_+$. The search space $\mathcal{D}$ has two constraints on unknown parameter $\Theta$. The first constraint ensures that the unknown parameter $\Theta$ admits the Tucker decomposition with rank $\bm r$. The second constraint makes the entries of $\Theta$ bounded by a constant $\alpha$. This bound condition is a technical assumption to help to recover $\Theta$ in the noiseless case. Similar conditions has been imposed in many literatures for the matrix case \citep{Davenport20121BitMC,Bhaskar20151bitMC,Cai2013AMC,Bhaskar2016ProbabilisticLM}.
The search space $\mathcal{B}$ makes sure that the probability function $f_l$ in \eqref{cumulative} is positive.

\subsection{Optimization}
In this section, we describe the algorithm to seek the optimizer of \eqref{eq:optimization}. The objective function $\mathcal{L}_{\mathcal{Y},\Omega}(\Theta,\mb)$ is concave in $(\Theta,\mb)$ whenever $\Phi(x)$ is log-concave \citep{McCullagh1980RegressionMF,Burridge1981ANO}.
However, the feasible set $\mathcal{D}$ is not a convex set, which makes the optimization \eqref{eq:optimization} a non-convex problem. One approach to handle this problem is utilizing the Tucker decomposition and converting optimization into a block-wise convex problem.
Let us denote the objective function as
\begin{equation}
    \label{blockwise}
    \begin{aligned}
         \mathcal{L}(\mathcal{C},\mM_1,\cdots,\mM_K,\mb) = \mathcal{L}_{\mathcal{Y},\Omega}(\mathcal{C}\times_1\mM_1\cdots\times_K\mM_K,\mb).
    \end{aligned}
\end{equation}
From \eqref{blockwise}, we have $K+2$ blocks of variables in the objective function, one for the cutpoints vector $\mb$, one for the core tensor $\mathcal{C}$ and  K for the factor matrices $\mM_K$'s.
We can change the optimization problem to simple convex problem if any $K+1$ out of the $K+2$ blocks being fixed. Therefore, we can alternatively update one block at a time while other blocks being fixed.
The algorithm \ref{alg} gives the full description.

\begin{algorithm}[tb]
        \caption{Ordinal tensor decomposition }\label{alg}
        \begin{algorithmic}[]
            \STATE{\bfseries Input:}  \text{ Ordinal tensor }
            $\mathcal{Y}\in [L]^{d_1\times\cdots\times d_K}
            $, rank $\mr\in \mathbb{N_+}^{K-1}$, and entry-wise bound $\alpha\in \mathbb{R_+}$.
            \STATE{\bfseries Output:} $(\hat\Theta,\hat{\mb}) =  \argmax_{(\Theta,\mb)\in \mathcal{D}\times\mathcal{B}}  \mathcal{L}_{\mathcal{Y},\Omega}(\Theta,\mb).$
            \STATE Initialize $\mathcal{C}^{(0)}\in\mathbb{R}^{r_1\times\cdots\times r_k}, \{\mM_1^{(0)},\cdots,\mM_K^{(0)}\}$ and $\mb^{(0)} = ( b_1^{(0)},\cdots,b^{(0)}_{L-1})$ s.t. $b_1^{(0)}<\cdots<b_{L-1}^{(0)}$.
            \FOR{t = 1,2,$\cdots$,}
            \STATE{Update $\mM_k$}
            \FOR{k = 1,2,$\cdots$,K}
             \STATE $\mM_k^{(t+1)}\gets\argmax_{\mM_k\in\mathbb{R}^{d_k\times r_k}}\mathcal{L}_{\mathcal{Y},\Omega}(\mM_k)$ given $ \mathcal{C}^{(t)},\{\mM_i^{(t)}: i\neq k\} \text{ and } \mb^{(t)}$ s.t.\\
             $\|\mathcal{C}^{(t)}\times_1\mM_1^{(t+1)}\cdots\times_k\mM_k^{(t+1)}\cdots\times_K\mM_K^{(t)}\|_\infty \leq \alpha$
            \ENDFOR
            \STATE {Update $\mathcal{C}$}
            \STATE $\mathcal{C}^{(t+1)} \gets \argmax_{\mathcal{C}\in \mathbb{R}^{r_1\times\cdots\times r_k}}\mathcal{L}_{\mathcal{Y},\Omega}(\mathcal{C})$
            s.t.\newline $\|\mathcal{C}\times_1\mM_1^{(t+1)}\cdots\times_N\mM_K^{(t+1)}\|_\infty \leq \alpha$

            \STATE $\Theta^{(t+1)} \gets \mathcal{C}^{(t+1)}\times_1\mM_1^{(t+1)}\cdots\times_N\mM_K^{(t+1)}$
            \STATE \textbf{Update $\mb$}
            \STATE $\mb^{(t+1)} \gets \argmax_{\mb\in \mathbb{R}^{L-1}}\mathcal{L}_{\mathcal{Y},\Omega}\big(\Theta^{(t+1)},\mb\big)$  s.t. $b_1^{(0)}<\cdots<b_{L-1}^{(0)}$.
            \ENDFOR
            \STATE \textbf{return}
            $\Theta,\mb$
        \end{algorithmic}
    \end{algorithm}

\subsection{Rank selection}
Algorithm \ref{alg} takes  the rank $\bm r$ as an input variable. In practice, the rank $\bm r$ is hardly known. Estimating an appropriate rank $\bm r$ from a given tensor is an important issue. We suggest to use Bayesian Information Criterion(BIC) to choose the rank.
\begin{equation}
    \label{bic}
    \begin{aligned}
        \hat{\bm r} &= \argmin_{\bm r\in \mathbb{N}_+^K} BIC(\bm r)\\
        &=\argmin_{\bm r\in \mathbb{N}_+^K}\bigg[-2\mathcal{L}_\mathcal{Y}(\hat\Theta(\bm r)) \quad+
        \\&\qquad\big(\prod_{k\in[K]} r_k+\sum_{k\in[K]}(d_kr_k-r_k^2)\big)\log(\prod_{k\in[K]} d_k)\bigg],
    \end{aligned}
\end{equation}
where $\hat\Theta(\bm r)$ is a maximum likelihood estimater given the rank $\bm r$. The second part of the summation in \eqref{bic} is the number of  independent parameters in the model. We select the rank $\hat{\bm r}$ that minimizes BIC value through the grid search method.
\section{Real-world Data Applications}
In this section, we apply our ordinal tensor decomposition method to two real-world datasets of ordinal tensors. In the first application, we use our model to analyze an ordinal tensor consisting of structural connectivity patterns among 68 brain regions for 136 individuals from Human Connectome Project (HCP) \citep{Geddes2016HumanBM}. In the second application, we perform tensor completion from the data with missing values. The data tensor records the ratings from scale 1 to 5 of 42 users to 139 songs on 26 contexts \citep{Baltrunas2011InCarMusicCM}.
\subsection{Human Connectome Project (HCP)}
The human connectome project (HCP) is a $68\times 68 \times 136$ tensor where the first two modes have 68 indices representing brain regions and the last mode has 136 indices meaning individuals. All the individual images were preprocessed following a standard pipeline \citep{Zhang2018MappingPS}, and the brain was parcellated to 68 regions of interest following the Desikan atlas \citep{Desikan2006AnAL}. The tensor entries consist of $\{1,2,3\}$, the strength of fiber connections between 68 brain regions for each of the 136 individuals. First, we apply our ordinal tensor decomposition method with a logistic link function to the HCP data and identify similar brain nodes based on latent parameters. The BIC result suggests $\bm r = (23,23,8)$  with $\mathcal{L}_\mathcal{Y}(\hat{\Theta},\hat{\mb}) = -216645.8$.

Also, we compare our ordinal tensor decomposition method with a logistic link function with the classical continuous Tucker decomposition method. We use 5 folded cross validation method for the comparison. Specifically, we randomly split the tensor entries into 5 similar sized pieces and alternatively use each piece of entries as a test data using other $80\%$ entries as a training data. The entries in the test data are encoded as missing and then predicted based on the tensor decomposition from the training data. We set the rank of the tensor as $\bm r = (23,23,8)$ which minimizes BIC value. Table \ref{CV} shows RMSE, MAE and error(false prediction rate), averaged over 5 test set results. We can check that our ordinal tensor decomposition model outperforms the classic continuous decomposition in all criteria.

\begin{table}[t]
\caption{Tensor completion for the HCP data. Two methods are compared: the proposed ordinal tensor decomposition and the classical continuous Tucker decomposition.}
\label{CV}
\vskip 0.15in
\begin{center}
\begin{small}
\begin{sc}
\begin{tabular}{lcccr}
\toprule
Method & RMSE & MAE & ERROR \\
\midrule
ORDINAL   & 0.1503711 & 0.1502662 & 0.1502151 \\
CONTI & 0.1603685& 0.1600219& 0.1598499\\
\bottomrule
\end{tabular}
\end{sc}
\end{small}
\end{center}
\vskip -0.1in
\end{table}

\subsection{Music data with missing values}

InCarMusic is a mobile application that offers music recommendation to passengers of cars based on contexts \citep{Baltrunas2011InCarMusicCM}.
Our goal is to get the tensor completion from the $42\times139\times 26$ tensor using the ordinal tensor decomposition and thereby we can offer context-specific music recommendation to users.
The tensor entries consist of $\{1,2,3,4,5\}$, the ratings of 42 users to 139 songs on 26 contexts and are encoded as $-1$ for missing values. The number of missing values is $148904$ and the number of available values is $2884$.

\bibliography{ICML.bib}
\bibliographystyle{icml2020}

\end{document}

% This document was modified from the file originally made available by
% Pat Langley and Andrea Danyluk for ICML-2K. This version was created
% by Iain Murray in 2018, and modified by Alexandre Bouchard in
% 2019 and 2020. Previous contributors include Dan Roy, Lise Getoor and Tobias
% Scheffer, which was slightly modified from the 2010 version by
% Thorsten Joachims & Johannes Fuernkranz, slightly modified from the
% 2009 version by Kiri Wagstaff and Sam Roweis's 2008 version, which is
% slightly modified from Prasad Tadepalli's 2007 version which is a
% lightly changed version of the previous year's version by Andrew
% Moore, which was in turn edited from those of Kristian Kersting and
% Codrina Lauth. Alex Smola contributed to the algorithmic style files.

### 2020-01-23 ICML/chanwoo/ICMLdraft.tex
Add files via upload

@@
         \label{tucker}
-        \Theta = \mathcal{C}\times_1\mM_1\cdots\times_N\mM_N,
+        \Theta = \mathcal{C}\times_1\mM_1\cdots\times_K\mM_N,
     \end{aligned}
@@
             $\mathcal{Y}\in [L]^{d_1\times\cdots\times d_K}
-            $, rank $\mr\in \mathbb{N_+}^{K-1}$, and entry-wise bound $\alpha\in \mathbb{R_+}$.
+            $,\\ \hspace{.43in} Rank $\mr\in \mathbb{N_+}^{K-1}$,\\
+            \hspace{.43in} Entry-wise bound $\alpha\in \mathbb{R_+}$.
             \STATE{\bfseries Output:} $(\hat\Theta,\hat{\mb}) =  \argmax_{(\Theta,\mb)\in \mathcal{D}\times\mathcal{B}}  \mathcal{L}_{\mathcal{Y},\Omega}(\Theta,\mb).$
-            \STATE Initialize $\mathcal{C}^{(0)}\in\mathbb{R}^{r_1\times\cdots\times r_k}, \{\mM_1^{(0)},\cdots,\mM_K^{(0)}\}$ and $\mb^{(0)} = ( b_1^{(0)},\cdots,b^{(0)}_{L-1})$ s.t.
-$b_1^{(0)}<\cdots<b_{L-1}^{(0)}$.
+            \STATE Initialize  Core tensor $\mathcal{C}^{(0)}$,\\
+            \hspace{.5in} Factor matrices $ \{\mM_1^{(0)},\cdots,\mM_K^{(0)}\}$,\\
+            \hspace{.5in} Cut-off points $\mb^{(0)}$.
             \FOR{t = 1,2,$\cdots$,}
-            \STATE{Update $\mM_k$}
             \FOR{k = 1,2,$\cdots$,K}
-             \STATE $\mM_k^{(t+1)}\gets\argmax_{\mM_k\in\mathbb{R}^{d_k\times r_k}}\mathcal{L}_{\mathcal{Y},\Omega}(\mM_k)$ given $ \mathcal{C}^{(t)},\{\mM_i^{(t)}: i\neq k\} \text{ and } \mb^{(t)}$ s.t.\\
-             $\|\mathcal{C}^{(t)}\times_1\mM_1^{(t+1)}\cdots\times_k\mM_k^{(t+1)}\cdots\times_K\mM_K^{(t)}\|_\infty \leq \alpha$
+            \STATE{Update $\mM_k$ fixing other blocks.}
+            \STATE $\mM_k^{(t+1)}\gets\argmax_{\mM_k\in\mathbb{R}^{d_k\times r_k}}\mathcal{L}_{\mathcal{Y},\Omega}(\mM_k)$ s.t.
+             $\|\mathcal{C}^{(t)}\times_1\mM_1^{(t+1)}\cdots\times_k\mM_k\cdots\times_K\mM_K^{(t)}\|_\infty \leq \alpha$
             \ENDFOR
-            \STATE {Update $\mathcal{C}$}
+            \STATE {Update $\mathcal{C}$ fixing other blocks.}
             \STATE $\mathcal{C}^{(t+1)} \gets \argmax_{\mathcal{C}\in \mathbb{R}^{r_1\times\cdots\times r_k}}\mathcal{L}_{\mathcal{Y},\Omega}(\mathcal{C})$
-            s.t.\newline $\|\mathcal{C}\times_1\mM_1^{(t+1)}\cdots\times_N\mM_K^{(t+1)}\|_\infty \leq \alpha$
+            s.t.\newline $\|\mathcal{C}\times_1\mM_1^{(t+1)}\cdots\times_K\mM_K^{(t+1)}\|_\infty \leq \alpha$
 
-            \STATE $\Theta^{(t+1)} \gets \mathcal{C}^{(t+1)}\times_1\mM_1^{(t+1)}\cdots\times_N\mM_K^{(t+1)}$
-            \STATE \textbf{Update $\mb$}
-            \STATE $\mb^{(t+1)} \gets \argmax_{\mb\in \mathbb{R}^{L-1}}\mathcal{L}_{\mathcal{Y},\Omega}\big(\Theta^{(t+1)},\mb\big)$  s.t. $b_1^{(0)}<\cdots<b_{L-1}^{(0)}$.
+            \STATE $\Theta^{(t+1)} \gets \mathcal{C}^{(t+1)}\times_1\mM_1^{(t+1)}\cdots\times_K\mM_K^{(t+1)}$
+            \STATE {Update $\mb$ fixing $\Theta^{(t+1)}$.}
+            \STATE $\mb^{(t+1)} \gets \argmax_{\mb\in \mathbb{R}^{L-1}}\mathcal{L}_{\mathcal{Y},\Omega}\big(\Theta^{(t+1)},\mb\big)$
             \ENDFOR
@@
 
-Also, we compare our ordinal tensor decomposition method with a logistic link function with the classical continuous Tucker decomposition method.
+Also, we compare our ordinal tensor decomposition method with a logistic link function with other methods: the continuous tensor decomposition method, the 1 bit-tensor completion method with probit link function and the 1 bit-tensor completion method with logistic link function.
 We use 5 folded cross validation method for the comparison.
 Specifically, we randomly split the tensor entries into 5 similar sized pieces and alternatively use each piece of entries as a test data using other $80\%$ entries as a training data.
-The entries in the test data are encoded as missing and then predicted based on the tensor decomposition from the training data.
-We set the rank of the tensor as $\bm r = (23,23,8)$ which minimizes BIC value.
+The entries in the test data are encoded as missing and then predicted based on each method from the training data.
+We set the rank of the tensor as $\bm r = (23,23,8)$ for the tensor decomposition methods which minimizes BIC value.
 Table \ref{CV} shows RMSE, MAE and error(false prediction rate), averaged over 5 test set results.
-We can check that our ordinal tensor decomposition model outperforms the classic continuous decomposition in all criteria.
+We can check that our ordinal tensor decomposition model outperforms the other methods in all criteria.
 
 \begin{table}[t]
-\caption{Tensor completion for the HCP data. Two methods are compared: the proposed ordinal tensor decomposition and the classical continuous Tucker decomposition.}
 \label{CV}
@@
 \midrule
-ORDINAL   & 0.1503711 & 0.1502662 & 0.1502151 \\
-CONTI & 0.1603685& 0.1600219& 0.1598499\\
+\begin{tabular}[c]{@{}c@{}}Oridnal tensor\\ Decomposition\end{tabular}  & 0.1503711 & 0.1502662 & 0.1502151 \\ \\
+\begin{tabular}[c]{@{}c@{}}Conti tensor\\ Decomposition\end{tabular} & 0.1603685& 0.1600219& 0.1598499\\ \\
+\begin{tabular}[c]{@{}c@{}}1bit-completion\\ (probit)\end{tabular}  & 0.3658390 & 0.3632757 & 0.3619940\\
+\begin{tabular}[c]{@{}c@{}}1bit-completion\\ (logistic)\end{tabular}  &  0.4519253 & 0.4218513 & 0.4068143 
+\\
 \bottomrule
@@
 \vskip -0.1in
+\caption{Results of comparisons among 4 methods on the HCP data predicting the test data.
+Four methods are the ordinal tensor decomposition algorithm, the continuous tensor decomposition algorithm and the 1 bit tensor completion method with probit link function and logistic link function.
+Each method is evaluated by RMSE, MAE, error(false prediction rate).}
 \end{table}

### 2020-01-24 ICML/miaoyan/ordinalT.tex
update

@@
 \caption{Comparison of various low-rank estimation methods based on categorical observations.
-For an order-$K$ dimensional-$(d,\ldots,d)$ ordinal tensor with $L$-level observations, we report the error bound in the recovered signal (for tensor denoising) and required sample complexity (for tensor completion) as functions of tensor dimension.
+For an order-$K$ dimensional-$(d,\ldots,d)$ ordinal tensor with $L$-level observations, we report the error bound in the recovered signal (for tensor denoising) and required sample complexity (for tensor completion) as functions of tensor dimensions.
 {\color{red}(font in the table seems odd...)}}~\label{tab:compare}
@@
 \vspace{-.2cm}
-\item $f'(\theta)$ is unimodal and symmetric with respect to $\theta=0$. 
+\item $f'(\theta)$ is unimodal, and symmetric with respect to $\theta=0$. 
 \end{enumerate}
@@
 In view of equation~\eqref{eq:latent}, these quantities characterize the geometry (such as flatness and convexity) of the latent noise distribution.
-Under the assumption~\ref{ass:link}, all these quantities are strictly positive and independent of the tensor dimension.
+Under the Assumption~\ref{ass:link}, all these quantities are strictly positive and independent of tensor dimensions.
 
@@
 
-A similar conclusion can be obtained for the prediction error, measured in KL divergence, between the two generative distributions. 
+A similar conclusion can be obtained for the prediction error, measured in Kullback-Leibler (KL) divergence, between the two categorical distributions. 
 \begin{cor}[Prediction error]~\label{cor:prediction}
 Assume the same set-up as in Theorem~\ref{thm:rate}.
-Let $\mathbb{P}_{\tY}$ and $\hat{\mathbb{P}}_{\tY}$ denote the distributions generating the $L$-level ($L\geq 2$) ordinal tensor $\tY$, given the true parameter $\Theta$ and its estimator $\hat \Theta$, respectively.
+Let $\mathbb{P}_{\tY}$ and $\hat{\mathbb{P}}_{\tY}$ denote the distributions generating the $L$-level ordinal tensor $\tY$, given the true parameter $\Theta$ and its estimator $\hat \Theta$, respectively.
+Assume $L\geq 2$.
 Then, with very high probability,
@@
 Furthermore, our results~\eqref{eq:rate} and~\eqref{eq:KLrate} reveal that the convergence becomes favorably as the order of tensor data increases.
-Intuitively, for tensor data analysis, we can view the number of entries, $\prod_k d_k$, as the sample size, and the number of free parameters in $\tP$ is roughly on the order of $\sum_{k}d_k$, assuming $r_{\max}=O(1)$.
-Compared to the matrix case, a higher tensor order implies more effective samples per parameter and thus exhibits a faster convergence rate in high dimensions.
+Intuitively, for tensor data analysis, the sample size is the number of tensor entries, $\prod_k d_k$, and the number of free parameters is roughly on the order of $\sum_{k}d_k$, assuming $r_{\max}=O(1)$.
+A higher tensor order implies higher effective sample size per parameter, and thus exhibits a faster convergence rate in high dimensions.
 
@@
 \end{align}
-where $C=C(\alpha, L, f,\mb)>0$ is a constant independent of the tensor dimension and rank. 
+where $C=C(\alpha, L, f,\mb)>0$ is a constant independent of tensor dimensions and the rank. 
 \end{thm}
@@
 Now we consider the problem of tensor completion, when only a subset of entries $\Omega$ are observed.
-We consider a general sampling procedure induced by $\Pi$, and assess the accuracy using the weighted squared deviation:
+We consider a general sampling procedure induced by $\Pi$.
+The recovery accuracy is assessed by the weighted squared deviation:
 \begin{align}\label{eq:weighted}
@@
 
-Note that the recovery error depends on the distribution.
+Note that the recovery error depends on the distribution $\Pi$.
 In particular, the entries with higher sampling probabilities are equipped with better accuracy guarantee, compared to the ones with lower sampling probabilities.
@@
 \subsection{Alternating optimization}
-\begin{figure*}[http]
-\begin{center}
-\includegraphics[width=16cm]{algorithm.pdf}
- \end{center}
-\end{figure*}
+
+\begin{algorithm}[tb]
+        \caption{Ordinal tensor decomposition }\label{alg}
+        \begin{algorithmic}[]
+            \STATE{\bfseries Input:}  \text{ Ordinal tensor }
+            $\mathcal{Y}\in [L]^{d_1\times\cdots\times d_K}
+            $,\\ \hspace{.43in} Rank $\mr\in \mathbb{N_+}^{K}$,\\
+            \hspace{.43in} Entry-wise bound $\alpha\in \mathbb{R_+}$.
+            \STATE{\bfseries Output:} $(\hat\Theta,\hat{\mb}) =  \argmax_{(\Theta,\mb)\in \mathcal{D}\times\mathcal{B}}  \mathcal{L}_{\mathcal{Y},\Omega}(\Theta,\mb).$
+            \STATE Initialize  Core tensor $\mathcal{C}^{(0)}$,\\
+            \hspace{.5in} Factor matrices $ \{\mM_1^{(0)},\cdots,\mM_K^{(0)}\}$,\\
+            \hspace{.5in} Cut-off points $\mb^{(0)}$.
+            \FOR{$t = 1,2,\cdots,$}
+            \FOR{$k = 1,2,\cdots,K$}
+            \STATE{Update $\mM_k$ while fixing other blocks.}
+            \STATE $\mM_k^{(t+1)}\gets\argmax_{\mM_k\in\mathbb{R}^{d_k\times r_K}}\mathcal{L}_{\mathcal{Y},\Omega}(\mM_k)$
+            \newline
+             s.t. $\mnormSize{}{\Theta^{(t+1)}}\leq \alpha$, where $\Theta^{(t+1)}$ is the parameter tensor based on the current block estimates. 
+            \ENDFOR
+            \STATE {Update $\mathcal{C}$ while fixing other blocks: $\mathcal{C}^{(t+1)} \gets \argmax_{\mathcal{C}\in \mathbb{R}^{r_1\times\cdots\times r_K}}\mathcal{L}_{\mathcal{Y},\Omega}(\mathcal{C})$, s.t.
+$\mnormSize{}{\Theta^{(t+1)}}\leq \alpha$.
+               \STATE {Update $\Theta$ based on the current block estimates: $\Theta^{(t+1)} \gets \mathcal{C}^{(t+1)}\times_1\mM_1^{(t+1)}\cdots\times_K\mM_K^{(t+1)}$. 
+            \STATE {Update $\mb$ while fixing $\Theta^{(t+1)}$: $\mb^{(t+1)} \gets \argmax_{\mb\in \mathbb{R}^{L-1}}\mathcal{L}_{\mathcal{Y},\Omega}\big(\Theta^{(t+1)},\mb\big)$
+            \ENDFOR
+            \STATE \textbf{return}
+            $\Theta,\mb$
+        \end{algorithmic}
+    \end{algorithm}
+
 In this section, we describe the algorithm to seek the optimizer of~\eqref{eq:objective}.
@@
 In this section, we apply our ordinal tensor decomposition method to two real-world datasets of ordinal tensors.
-In the first application, we use our model to analyze an ordinal tensor consisting of structural connectivity patterns among 68 brain regions for 136 individuals from Human Connec- tome Project (HCP) (Geddes 2016).
+In the first application, we use our model to analyze an ordinal tensor consisting of structural connectivity patterns among 68 brain regions for 136 individuals from Human Connectome Project (HCP) (Geddes 2016).
 In the second application, we perform tensor completion from the data with missing values.
@@
 
+\begin{table}[H]
+\resizebox{\linewidth}{!}{%
+\begin{tabular}{c|ccc|c}
+\hline
+\multirow{2}{*}{Critera}&\multicolumn{3}{c|}{Our method}&Continuous method\\
+& $\hat y^{(\text{Mean})}_\omega$ & $\hat y^{(\text{Median})}_\omega$ & $\hat y^{(\text{Mode})}_\omega$ &~\cite{filipovic2015tucker}\\
+\hline
+MSE& &  & 0.1504&0.1604\\
+MAD&  &  &0.1504&0.1600\\
+MCR& &&0.1502&0.1598\\
+\hline
+\end{tabular}
+}
+\caption{Comparison of tensor denoising on HCP dataset.}
+\end{table}
+
+\subsection{InCarMusic recommendation system}
+InCarMusic is a mobile application that offers music recommendation to passengers of cars based on contexts (Baltrunas et al.
+2011).
+Our goal is to perform the tensor completion to the $42\times 139\times 26$ ordinal tensor and thereby we can offer context-specific music recommendation to users.
+The tensor entries consist of ordinal observations on the scale 1 to 5, the ratings of 42 users to 139 songs on 26 contexts and are encoded as NA for missing values.
+The number of missing values is 148,904 and the number of available values is 2,884.
+We suggest three estimators $\hat y_\omega$ based on $(\hat \theta_\omega,\hat \mb)$:
+\begin{itemize}
+\item (Mean) $\hat y^{(\text{Mean})}_\omega=\sum_\ell \ell g_\ell(\hat \theta_\omega,\hat \mb) $;
+\item (Median) $\hat y^{(\text{Median})}_\omega=\min\{\ell\in[L]\colon f_\ell(\hat \theta_\omega,\hat \mb)\geq 0.5\}$;
+\item (Mode) $\hat y^{(\text{Mode})}_\omega=\argmax_\ell g_\ell(\hat \theta_\omega,\hat \mb) $
+\end{itemize}
+Note that, under the ordinal tensor model~\eqref{eq:model}, the estimator $\hat y^{(\text{Mean})}_\omega$ (uniformally?) minimizes $\mathbb{E}(y_\omega-\hat y_\omega)^2$, and the estimator $\hat y^{(\text{Median})}_\omega$ minimizes $\mathbb{E}|y_\omega-\hat y_\omega|$ (?).
+In contrast, these three estimators degenerate to a single estimator under the continuous-valued tensor decomposition model.
+We assess the accuracy using three metrics: Mean squared error (MSE), Mean absolute deviation (MAD), and Misclassification rate (MCR).
+
+Results are averaged from 5 fold cross validation. {\color{red} add s.e. to the above table.}
+
+\begin{table}[H]
+\resizebox{\linewidth}{!}{%
+\begin{tabular}{c|ccc|c}
+\hline
+\multirow{2}{*}{Critera}&\multicolumn{3}{c|}{Our method}&Continuous method\\
+& $\hat y^{(\text{Mean})}_\omega$ & $\hat y^{(\text{Median})}_\omega$ & $\hat y^{(\text{Mode})}_\omega$ &~\cite{filipovic2015tucker}\\
+\hline
+MSE& \color{red}2.01&2.14 & 3.64&8.18\\
+MAD& 1.23  &  {\color{red}1.13}&1.26&2.36\\
+MCR& 0.80&0.75&{\color{red}0.54}&0.96\\
+\hline
+\end{tabular}
+}
+\caption{Comparison of tensor completion performance on InCarMusic dataset.}
+\end{table}
+
+We also attempted to run matrix ordinal mehod~\citep{bhaskar2016probabilistic} to the InCarMusic dataset. Unfortunately, the $\mb$ cannot be estimated from their method.
+
 \bibliography{tensor_wang}

### 2020-01-24 code/experiment.R

@@ -24,13 +24,14 @@ ttnsr[ttnsr==-1]=NA
 missing=which(is.na(ttnsr)==F)
 d=dim(ttnsr)
 set.seed(1)
-r=c(2,2,2) ## need to select via BIC ..
+r=c(1,1,1) ## need to select via BIC ..
 alpha=100
-A_1 = randortho(d[1])[,1:r[1]]
-A_2 = randortho(d[2])[,1:r[2]]
-A_3 = randortho(d[3])[,1:r[3]]
+A_1 = as.matrix(randortho(d[1])[,1:r[1]])
+A_2 = as.matrix(randortho(d[2])[,1:r[2]])
+A_3 = as.matrix(randortho(d[3])[,1:r[3]])
 C = rand_tensor(modes = r)
 C=C/max(abs(ttl(C,list(A_1,A_2,A_3),ms=1:3)@data))*alpha/10
+
 
 ############# experiement ##########################
 mask=sample(1:2884,2884*0.1,replace=T)
@@ -44,22 +45,25 @@ toc()
 fitted=ttl(result$Z,list(result$U[[1]],result$U[[2]],result$U[[3]]),ms=1:3)
 mean((true-fitted@data[index])^2) #MSE= 18.96803
 mean(abs(true-fitted@data[index])) #MAD= 2.589925
+mean(true!=round(fitted@data[index])) ## MCR
 
 ###### Method 2: ordinal tucker #########
 tic()
 result = fit_ordinal(ttnsr,C,A_1,A_2,A_3,omega=TRUE,alph = alpha)
 toc()
 fitted=ttl(result$C,list(result$A_1,result$A_2,result$A_3),ms=1:3)@data
-est=estimation(fitted,result$omega,"median")
+bic(ttnsr,fitted,result$omega,d,r)## 10860.46
+est=estimation(fitted,result$omega,"mean")
 mean((true-est@data[index])^2) ## MSE = 2.381944
 mean(abs(true-est@data[index])) ## MAD = 1.256944
+mean(true!=round(est@data[index])) ## MCR
 
 ######### Method 3: ordinal matrix##################
 data=k_unfold(as.tensor(ttnsr),1)@data
 d=dim(data)
-r=2
+r=1
 set.seed(18)
-A_1=  randortho(d[1])[,1:r]
+A_1=  as.matrix(randortho(d[1])[,1:r])
 A_2 = matrix(rnorm(d[2]*r,0,1),nrow=d[2])
[…]

### 2020-01-24 code/functions.R

@@ -4,7 +4,7 @@ library(pracma)
 library(doParallel)
 epsilon=10^-4
 
-########### simulation ordinal tensors based on logist model with arbitrary k
+########### simulate an ordinal tensor based on logist model with arbitrary k
 realization = function(theta,omega){
   theta=as.tensor(theta)
   thet <- c(theta@data)
@@ -15,8 +15,9 @@ realization = function(theta,omega){
   return(as.tensor(array(thet,dim =theta@modes)))
 }      
 
-
-estimation = function(theta,omega,type){
+########### predict the response based on the estimated parameter
+estimation = function(theta,omega,type){ 
+   k=length(omega) 
   if(is.matrix(theta)){
   thet=c(theta)
   }else{
@@ -24,12 +25,12 @@ estimation = function(theta,omega,type){
   thet <- c(theta@data)
   }
   p = theta_to_p(thet,omega)
-  if(type=="max"){  # score prediction based on the mode 
+  if(type=="mode"){  # score prediction based on the mode 
   for (j in 1:length(thet)) thet[j] <-  which.max(p[j,])
   }else if(type=="mean"){ # score prediction based on the mean 
   for (j in 1:length(thet)) thet[j] <-  sum(p[j,]*(1:(k+1))) ## why rounding to integer?
   }else if(type=="median"){# score prediction based on the median 
-  for (j in 1:length(thet)) thet[j] <-  which(c(cumsum(p[i,]),1)>=0.5)[1] ## median
+  for (j in 1:length(thet)) thet[j] <-  which(c(cumsum(p[j,]),1)>=0.5)[1] ## median
   }
   if(is.matrix(theta)) return(as.matrix(thet,dim=dim(theta)))
   else return(as.tensor(array(thet,dim =theta@modes)))
@@ -116,7 +117,7 @@ gradient_tensor=function(A_1,A_2,A_3,C,ttnsr,omega,type="ordinal"){
     return(output)
 }
 
-# gradient with respect to core (gradient not using kronecker product at all)
+# gradient with respect to the core tensor (gradient not using kronecker product at all)
 gc = function(A_1,A_2,A_3,C,ttnsr,omega){
[…]

### 2020-01-28 ICML/miaoyan/ordinalT.tex
update draft
update

@@
             $\mathcal{Y}\in [L]^{d_1\times\cdots\times d_K}
-            $,\\ \hspace{.43in} Rank $\mr\in \mathbb{N_+}^{K}$,\\
+            $,\\ \hspace{.43in} Rank $\mr\in \mathbb{N}_+^{K}$,\\
             \hspace{.43in} Entry-wise bound $\alpha\in \mathbb{R_+}$.
@@
             \newline
-             s.t. $\mnormSize{}{\Theta^{(t+1)}}\leq \alpha$, where $\Theta^{(t+1)}$ is the parameter tensor based on the current block estimates. 
+             s.t.\ $\mnormSize{}{\Theta^{(t+1)}}\leq \alpha$, where $\Theta^{(t+1)}$ is the parameter tensor based on the current block estimates. 
             \ENDFOR
-            \STATE {Update $\mathcal{C}$ while fixing other blocks: $\mathcal{C}^{(t+1)} \gets \argmax_{\mathcal{C}\in \mathbb{R}^{r_1\times\cdots\times r_K}}\mathcal{L}_{\mathcal{Y},\Omega}(\mathcal{C})$, s.t.
-$\mnormSize{}{\Theta^{(t+1)}}\leq \alpha$.
+            \STATE {Update $\mathcal{C}$ while fixing other blocks: $\mathcal{C}^{(t+1)} \gets \argmax_{\mathcal{C}\in \mathbb{R}^{r_1\times\cdots\times r_K}}\mathcal{L}_{\mathcal{Y},\Omega}(\mathcal{C})$, s.t.\ $\mnormSize{}{\Theta^{(t+1)}}\leq \alpha$.
                \STATE {Update $\Theta$ based on the current block estimates: $\Theta^{(t+1)} \gets \mathcal{C}^{(t+1)}\times_1\mM_1^{(t+1)}\cdots\times_K\mM_K^{(t+1)}$. 
@@
 \section{Experiments}
-In this section, we evaluate the empirical performance of our rank-constrained M-estimatation.
+In this section, we evaluate the empirical performance of our rank-constrained estimation.
 We consider ordinal tensors with various number of categories, and compare the recovery accuracy with other tensor-based methods.
@@
 
-\begin{table}[H]
-\resizebox{\linewidth}{!}{%
-\begin{tabular}{c|ccc|c}
-\hline
-\multirow{2}{*}{Critera}&\multicolumn{3}{c|}{Our method}&Continuous method\\
-& $\hat y^{(\text{Mean})}_\omega$ & $\hat y^{(\text{Median})}_\omega$ & $\hat y^{(\text{Mode})}_\omega$ &~\cite{filipovic2015tucker}\\
-\hline
-MSE& &  & 0.1504&0.1604\\
-MAD&  &  &0.1504&0.1600\\
-MCR& &&0.1502&0.1598\\
-\hline
-\end{tabular}
-}
-\caption{Comparison of tensor denoising on HCP dataset.}
-\end{table}
+%\begin{table}[H]
+%\resizebox{\linewidth}{!}{%
+%\begin{tabular}{c|ccc|c}
+%\hline
+%\multirow{2}{*}{Critera}&\multicolumn{3}{c|}{Our method}&Continuous method\\
+%& $\hat y^{(\text{Mean})}_\omega$ & $\hat y^{(\text{Median})}_\omega$ & $\hat y^{(\text{Mode})}_\omega$ &~\cite{filipovic2015tucker}\\
+%\hline
+%MSE& &  & 0.1504&0.1604\\
+%MAD&  &  &0.1504&0.1600\\
+%MCR& &&0.1502&0.1598\\
+%\hline
+%\end{tabular}
+%}
+%\caption{Comparison of tensor denoising on HCP dataset.}
+%\end{table}
 

### 2020-01-28 code/HPC_analysis.R

@@ -1,0 +1,76 @@
+### Cross validation ######
+source("functions.R")
+load("dti_brain.RData")
+set.seed(18)
+ind = sample(1:length(tensor),length(tensor))
+
+### 5 fold ################ Please make changes to the sampling #####
+l = length(tensor)%/%5+1
+i1 = ind[1:l]
+i2 = ind[(l+1):(2*l)]
+i3 = ind[(2*l+1):(3*l)]
+i4 = ind[(3*l+1):(4*l)]
+i5 = ind[(4*l+1):length(tensor)]
+cindex = list(i1,i2,i3,i4,i5)
+
+# `index' is for 1:5  
+# I used following lines when I put codes in the server in sh file 
+# args index ==$SLURM_ARRAY_TASK_ID" filename.R
+test_index = cindex[[index]]
+train_index = setdiff(ind,test_index)
+train_tensor = tensor
+train_tensor[test_index] = NA
+
+#####################################################################################
+################# Continuous Tucker decomposition CV ################################
+d = dim(tensor)
+r = c(23,23,8)
+A_1 = randortho(d[1])[,1:r[1]]
+A_2 = randortho(d[2])[,1:r[2]]
+A_3 = randortho(d[3])[,1:r[3]]
+C = rand_tensor(modes = r)
+result = fit_continuous(train_tensor,C,A_1,A_2,A_3,alpha=TRUE)
+save(result,file = paste("CV23_23_8_",index,".RData",sep = ""))
+
+########## Analysis after getting the above output files ################
+CV = as.data.frame(matrix(nrow = 5, ncol = 3))
+names(CV) = c("MSE","MAE","Error_rate")
+for (i in 1:5) {
+  test_index = cindex[[i]]
+  load(paste("CV23_23_8_",i,".RData",sep = ""))
+  theta =ttl(result$C,list(result$A_1,result$A_2,result$A_3),ms=1:3)@data
+  ##theta = result$theta # alternatively, direct output the theta from result object
+  CV[i,1] = mean((round(theta)[test_index]-tensor[test_index])^2)
+  CV[i,2] = mean(abs(round(theta)[test_index]-tensor[test_index]))
+  CV[i,3] = error_rate = mean(round(theta)[test_index]!=tensor[test_index])
+}
+
+##################### ordinal glm tucker decomposition CV#################
+d = dim(tensor)
+set.seed(18)
+r = c(23,23,8)
[…]

### 2020-01-28 code/InCarMusic.R

@@ -1,71 +1,64 @@
-
-
-make_theta = function(result){
-  return(ttl(result$C,list(result$A_1,result$A_2,result$A_3),ms=1:3))
-}
-
-
-
 ### Cross validation ######
-load("dti_brain.RData")
-set.seed(18)
-ind = sample(1:length(tensor),length(tensor))
+source("functions.R")
+load("../data/InCar_Music.RData")
+tensor[tensor==-1]=NA
+ind=which(is.na(tensor)==F)
+d=dim(tensor)
+set.seed(1)
 
 ### 5 fold ################
-l = length(tensor)%/%5+1
+l = length(ind)%/%5+1
 i1 = ind[1:l]
 i2 = ind[(l+1):(2*l)]
 i3 = ind[(2*l+1):(3*l)]
 i4 = ind[(3*l+1):(4*l)]
-i5 = ind[(4*l+1):length(tensor)]
+i5 = ind[(4*l+1):length(ind)]
 cindex = list(i1,i2,i3,i4,i5)
 
 # `index' is for 1:5  
 # I used following lines when I put codes in the server in sh file 
 # args index ==$SLURM_ARRAY_TASK_ID" filename.R
-
 test_index = cindex[[index]]
 train_index = setdiff(ind,test_index)
 train_tensor = tensor
-train_tensor[test_index] = -1
+train_tensor[test_index] = NA
+
+
 
 #####################################################################################
 ################# Continuous Tucker decomposition CV ################################
-W = array(0, dim(tensor))
-W[train_index] = 1
-
-d = dim(tensor)
-r = c(25,25,7)
-A_1 = randortho(d[1])[,1:r[1]]
-A_2 = randortho(d[2])[,1:r[2]]
-A_3 = randortho(d[3])[,1:r[3]]
+r = rep(1,3)
+A_1 = as.matrix(randortho(d[1])[,1:r[1]])
+A_2 = as.matrix(randortho(d[2])[,1:r[2]])
+A_3 = as.matrix(randortho(d[3])[,1:r[3]])
 C = rand_tensor(modes = r)
-result = tucker_missing(A_1,A_2,A_3,C,train_tensor,W)
-save(result,file = paste("CV25_25_7_",index,".RData",sep = ""))
+#C=C/max(abs(ttl(C,list(A_1,A_2,A_3),ms=1:3)@data))*alpha/10
+
[…]

### 2020-01-28 code/braindataclustering.R

@@ -1,6 +1,6 @@
 load("dti_brain.RData")
 load("rank_23_23_8.RData")
-theta = make_theta(result)
+theta = result$theta
 thetan = tucker(theta,rank = result$C@modes)
 
 # clust1 is clustering from mode 1, cluster 2 is clustering from mode 2

### 2020-01-28 code/experiment.R

@@ -1,366 +1,0 @@
-source("functions.R")
-########## brain tensor analysis ##########
-load("../data/dti_brain.RData")
-##ttnsr=tensor[,,attr[,2]=="22-25"] ## subset of individuals
-ttnsr=tensor
-d=dim(ttnsr)
-
-r=c(25,25,7)
-### random initilization
-A_1 = randortho(d[1])[,1:r[1]]
-A_2 = randortho(d[2])[,1:r[2]]
-A_3 = randortho(d[3])[,1:r[3]]
-C = rand_tensor(modes = r)
-##result = fit_ordinal(ttnsr,C,A_1,A_2,A_3,omega=TRUE,alph = TRUE)
-tic()
-result = fit_ordinal_MM(ttnsr,C,A_1,A_2,A_3,omega=TRUE,alph = TRUE)
-toc()
-
-########## Music analysis #######################
-source("functions.R")###
-load("../data/InCar_Music.RData")
-ttnsr=tensor
-ttnsr[ttnsr==-1]=NA
-missing=which(is.na(ttnsr)==F)
-d=dim(ttnsr)
-set.seed(1)
-r=c(1,1,1) ## need to select via BIC ..
-alpha=100
-A_1 = as.matrix(randortho(d[1])[,1:r[1]])
-A_2 = as.matrix(randortho(d[2])[,1:r[2]])
-A_3 = as.matrix(randortho(d[3])[,1:r[3]])
-C = rand_tensor(modes = r)
-C=C/max(abs(ttl(C,list(A_1,A_2,A_3),ms=1:3)@data))*alpha/10
-
-
-############# experiement ##########################
-mask=sample(1:2884,2884*0.1,replace=T)
-index=which(is.na(ttnsr)==F)[mask]
-true=ttnsr[index]
-ttnsr[index]=NA
-##### Method 1: continuous tucker #########
-tic()
-result = tucker_missing(ttnsr,r,TRUE,alpha=alpha)
-toc()
-fitted=ttl(result$Z,list(result$U[[1]],result$U[[2]],result$U[[3]]),ms=1:3)
-mean((true-fitted@data[index])^2) #MSE= 18.96803
-mean(abs(true-fitted@data[index])) #MAD= 2.589925
-mean(true!=round(fitted@data[index])) ## MCR
-
-###### Method 2: ordinal tucker #########
-tic()
[…]

### 2020-01-28 code/functions.R

@@ -1,50 +1,50 @@
 library(MASS)
 library(rTensor)
 library(pracma)
-library(doParallel)
 epsilon=10^-4
 
-########### simulate an ordinal tensor based on logist model with arbitrary k
+########### simulate ordinal tensors based on logist models with arbitrary k
 realization = function(theta,omega){
+  k=length(omega) 
   theta=as.tensor(theta)
-  thet <- c(theta@data)
-  p = theta_to_p(thet,omega)
-  for (j in 1:length(thet)) {
-    thet[j] <-  sample(1:(k+1),1,prob = p[j,])
+  theta_output <- c(theta@data)
+  p = theta_to_p(theta_output,omega)
+  for (j in 1:length(theta_output)) {
+    theta_output[j] <-  sample(1:(k+1),1,prob = p[j,])
   }
-  return(as.tensor(array(thet,dim =theta@modes)))
+  return(as.tensor(array(theta_output,dim =theta@modes)))
 }      
 
-########### predict the response based on the estimated parameter
+########### predict ordinal observations based on the estimated parameter
 estimation = function(theta,omega,type){ 
-   k=length(omega) 
+  k=length(omega) 
   if(is.matrix(theta)){
-  thet=c(theta)
+   theta_output=c(theta)
   }else{
   theta=as.tensor(theta)
-  thet <- c(theta@data)
+  theta_output <- c(theta@data)
   }
-  p = theta_to_p(thet,omega)
+  p = theta_to_p(theta_output,omega)
   if(type=="mode"){  # score prediction based on the mode 
-  for (j in 1:length(thet)) thet[j] <-  which.max(p[j,])
+  for (j in 1:length(theta_output)) theta_output[j] <-  which.max(p[j,])
   }else if(type=="mean"){ # score prediction based on the mean 
-  for (j in 1:length(thet)) thet[j] <-  sum(p[j,]*(1:(k+1))) ## why rounding to integer?
+  for (j in 1:length(theta_output)) theta_output[j] <-  sum(p[j,]*(1:(k+1))) ## why rounding to integer?
[…]

### 2020-01-28 code/old_code/functions_miaoyan.R

@@ -1,649 +1,0 @@
-library(MASS)
-library(rTensor)
-library(pracma)
-library(doParallel)
-epsilon=10^-4
-
-
-########### simulation ordinal tensors based on logist model with arbitrary k
-realization = function(theta,omega){
-  k = length(omega)
-  theta=as.tensor(theta)
-  thet <- c(theta@data)
-  p = matrix(nrow = length(thet),ncol = k)
-  for (i in 1:k) {
-    p[,i] = logistic(thet + omega[i])
-  }
-  p =  cbind(p,rep(1,length(thet)))-cbind(rep(0,length(thet)),p)
-  for (j in 1:length(thet)) {
-    thet[j] <-  sample(1:(k+1),1,prob = p[j,])
-  }
-  return(as.tensor(array(thet,dim =theta@modes)))
-}      
-
-
-
-
-########### cost function  with arbitrary k ###########
-h1 = function(A_1,W1,ttnsr,omega,type="ordinal"){
-  k = length(omega)
-  theta =W1%*%c(A_1)
-  if(type=="Gaussian"){
-      l=sqrt(sum((theta[is.na(ttnsr)==F]-ttnsr[is.na(ttnsr)==F])^2))
-      return(l)  
-  }
-  p = matrix(nrow = length(theta),ncol = k)
-  for (i in 1:k) {
-    p[,i] = as.numeric(logistic(theta + omega[i]))
-  }
-  p =  cbind(p,rep(1,length(theta)))-cbind(rep(0,length(theta)),p)
-  l = lapply(1:(k+1),function(i) -log(p[which(c(ttnsr)==i),i]))
-  return(sum(unlist(l)))
-}
-
-hc = function(A_1,A_2,A_3,C,ttnsr,omega,type="ordinal"){
-  k = length(omega)
-  theta = c(ttl(C,list(A_1,A_2,A_3),ms=1:3)@data)
-  if(type=="Gaussian"){
-      l=sqrt(sum((theta[is.na(ttnsr)==F]-ttnsr[is.na(ttnsr)==F])^2))
-      return(l)
-  }
-  p = matrix(nrow = length(theta),ncol = k)
-  for (i in 1:k) {
-    p[,i] = as.numeric(logistic(theta + omega[i]))
-  }
-  p =  cbind(p,rep(1,length(theta)))-cbind(rep(0,length(theta)),p)
[…]

### 2020-01-28 code/old_code/unused_functions.R

@@ -1,0 +1,228 @@
+library(MASS)
+library(rTensor)
+library(pracma)
+library(doParallel)
+epsilon=10^-4
+
+
+
+## A_1, A_2, A_3 
+g1_factor=function(A_1,A_2,A_3,C,ttnsr,k,omega,type="ordinal"){
+    g = gradient_tensor(A_1,A_2,A_3,C,ttnsr,omega,type) ## first, take entrywise gradient w.r.t. theta
+    
+    if(k==1){
+        g = ttl(as.tensor(g),list(matrix(1),t(A_2),t(A_3)),ms = 1:3) ## take gradient w.r.t. row in factor 1
+        g = k_unfold(g,1)@data%*%t(k_unfold(C,1)@data)
+    }
+    else if(k==2){
+        g = ttl(as.tensor(g),list(t(A_1),matrix(1),t(A_3)),ms = 1:3) ## take gradient w.r.t. row in factor 2
+        g = k_unfold(g,2)@data%*%t(k_unfold(C,2)@data)
+    }
+    else if(k==3){
+        g = ttl(as.tensor(g),list(t(A_1),t(A_2),matrix(1)),ms = 1:3) ## take gradient w.r.t. row in factor 3
+        g = k_unfold(g,3)@data%*%t(k_unfold(C,3)@data)
+    }
+    return(g)
+}
+
+####### update one factor matrix at a time while holding others fixed ###########
+factorcomb = function(A_1,A_2,A_3,C,ttnsr,k,omega,alpha=TRUE,type="ordinal"){
+    if(k==1){
+        if (alpha==TRUE) {
+            l <- lapply(1:nrow(A_1),function(i){optim(A_1[i,],
+                function(x) hc(t(as.matrix(x)),A_2,A_3,C,ttnsr[i,,],omega,type),
+                function(x) g1_factor(t(as.matrix(x)),A_2,A_3,C,ttnsr[i,,],1,omega,type),method = "BFGS")$par})
+            nA <- matrix(unlist(l),nrow = nrow(A_1),byrow = T)
+        }else{
+            W <- kronecker(A_3,A_2)%*%t(k_unfold(C,1)@data)
+            l <- lapply(1:nrow(A_1),function(i){constrOptim(A_1[i,],
+                function(x) hc(t(as.matrix(x)),A_2,A_3,C,ttnsr[i,,],omega,type),
[…]

### 2020-01-28 code/simulation_figure/Figure1/Figure1.R

@@ -1,0 +1,58 @@
+### Figure: cost function vs. iteration ##
+source("../../functions.R")
+library(ggplot2)
+set.seed(1)
+Output=NULL
+dlist=c(25,30)
+rlist=c(5,10)
+
+for(i in dlist){
+    for(j in rlist){
+        d=rep(i,3); r=rep(j,3);
+        B_1 = randortho(d[1])[,1:r[1]]
+        B_2 = randortho(d[2])[,1:r[2]]
+        B_3 = randortho(d[3])[,1:r[3]]
+        D = rand_tensor(modes = r)
+        theta = ttl(D,list(B_1,B_2,B_3),ms=1:3)@data
+        alpha=10
+        theta = theta/max(abs(theta))*alpha ## specify signal = 10
+        omega = c(-0.3,-0.1,0.1,0.3) ## equal spaced
+        
+        ttnsr <- realization(theta,omega)@data
+        
+        #initial point
+        A_1 = randortho(d[1])[,1:r[1]]
+        A_2 = randortho(d[2])[,1:r[2]]
+        A_3 = randortho(d[3])[,1:r[3]]
+        C = rand_tensor(modes = r)
+        C=C/max(abs(ttl(C,list(A_1,A_2,A_3),ms=1:3)@data))*alpha/10 ## initial theta is in the interior 
+        thetainit=ttl(C,list(A_1,A_2,A_3),ms=1:3)@data
+        
+        time=system.time(result <- fit_ordinal(ttnsr,C,A_1,A_2,A_3,omega=TRUE))
+        time=time[2]/length(result$iteration)
+        
+        Output=rbind(Output,c(time,d[1],r[1],-likelihood(ttnsr,theta,omega),-result$cost))
+    }
+}
+
+colnames(Output)=c("time","dim","rank","true_parameter",paste(rep("iter",15),1:15,sep=""))
+write.table(Output,"Figure2.txt",quote=F)
+
+
+################ plot ############
+Output=read.table("Figure2.txt",header=T)
+cost=c(unlist(table[,5:19]))
+iteration=rep(1:15,rep(4,15))
[…]

### 2020-01-28 code/simulation_figure/Figure2/Figure2.R

@@ -1,0 +1,64 @@
+### Figure: MSE vs. dimension, under various ranks##
+source("../../functions.R")
+set.seed(1)
+Output=NULL
+dlist=seq(from=15,to=40,by=2)
+rlist=c(3,5,8)
+ntotal=30
+MSE_raw=array(0,dim=c(length(rlist),length(dlist),ntotal))
+
+for(j in 1:length(rlist)){
+    for(i in 1:length(dlist)){
+        for(nsim in 1:ntotal){
+        d=rep(dlist[i],3); r=rep(rlist[j],3);
+        B_1 = randortho(d[1])[,1:r[1]]
+        B_2 = randortho(d[2])[,1:r[2]]
+        B_3 = randortho(d[3])[,1:r[3]]
+        D = rand_tensor(modes = r)
+        theta = ttl(D,list(B_1,B_2,B_3),ms=1:3)@data
+        alpha=10
+        theta = theta/max(abs(theta))*alpha ## specify signal = 10
+        omega = logit((1:4)/5) ## equal spaced on the probit scale
+        
+        ttnsr <- realization(theta,omega)@data
+        
+        #initial point
+        A_1 = randortho(d[1])[,1:r[1]]
+        A_2 = randortho(d[2])[,1:r[2]]
+        A_3 = randortho(d[3])[,1:r[3]]
+        C = rand_tensor(modes = r)
+        C=C/max(abs(ttl(C,list(A_1,A_2,A_3),ms=1:3)@data))*alpha/10 ## initial theta is in the interior 
+        thetainit=ttl(C,list(A_1,A_2,A_3),ms=1:3)@data
+        
+        result <- fit_ordinal(ttnsr,C,A_1,A_2,A_3,omega=TRUE,alpha=2*alpha) ## relaxed bound
+        
+        MSE_raw[j,i,nsim]=mean((theta-result$theta)^2)
+    }
+}
+}
+
+Output=data.frame(MSE=c(t(apply(MSE_raw,c(1:2),mean))),rank=as.factor(rep(rlist,rep(length(dlist),length(rlist)))),dim=rep(dlist,length(rlist)),sd=c(t(apply(MSE_raw,c(1:2),sd))))
+save(Output,MSE_raw,file="Figure2.RData")
+
+################ plot ############
+library(ggplot2)
+load("Figure2.RData")
+Output$rank=as.factor(Output$rank)
[…]

### 2020-01-28 code/simulation_figure/Figure2/run_Figure2.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --mail-user=miaoyan@stat.wisc.edu
+#SBATCH -p long
+#SBATCH -t 5-00:00:00
+#SBATCH -n 4
+#SBATCH --mem-per-cpu=10000M
+module load R/R-3.5.3
+export R_LIBS=/workspace/miaoyan/x86_64-pc-linux-gnu-library/3.4/
+R CMD BATCH Figure2.R output_new.Rout

### 2020-01-28 code/simulation_figure/Figure3/Figure3.R

@@ -1,0 +1,72 @@
+### Figure 3: MSE vs. signal level (alpha), under various ranks##
+source("../../functions.R")
+set.seed(1)
+Output=NULL
+rlist=c(3,5,8)
+##alist=seq(from=0.7,to=3.5,by=0.5)
+##alist=seq(from=0.5,to=2.5,by=0.3)
+alist=c(0.6,0.7,seq(from=0.8,to=2.5,by=0.3))
+d=rep(20,3)
+ntotal=30
+MSE_raw=array(0,dim=c(length(rlist),length(alist),ntotal))
+
+for(j in 1:length(rlist)){
+    for(i in 1:length(alist)){
+        for(nsim in 1:ntotal){
+        alpha=10^(alist[i]); r=rep(rlist[j],3);
+        B_1 = randortho(d[1])[,1:r[1]]
+        B_2 = randortho(d[2])[,1:r[2]]
+        B_3 = randortho(d[3])[,1:r[3]]
+        D = rand_tensor(modes = r)
+        theta = ttl(D,list(B_1,B_2,B_3),ms=1:3)@data
+        theta = theta/max(abs(theta))*alpha ## specify signal = alpha
+        omega = logit((1:4)/5) ## equal spaced on the probit scale
+        
+        ttnsr <- realization(theta,omega)@data
+        
+        #initial point
+        A_1 = randortho(d[1])[,1:r[1]]
+        A_2 = randortho(d[2])[,1:r[2]]
+        A_3 = randortho(d[3])[,1:r[3]]
+        C = rand_tensor(modes = r)
+        C=C/max(abs(ttl(C,list(A_1,A_2,A_3),ms=1:3)@data))*alpha/10 ## initial theta is in the interior 
+        thetainit=ttl(C,list(A_1,A_2,A_3),ms=1:3)@data
+        
+        result <- fit_ordinal(ttnsr,C,A_1,A_2,A_3,omega=TRUE,alpha=2*alpha) ## relaxed bound
+        #MSE=c(MSE,mean((theta-result$theta)^2)/mean(theta^2))
+        MSE_raw[j,i,nsim]=mean((theta-result$theta)^2)/mean(theta^2)
+        ##sprintf("%d\n%d\n%d",i,j,nsim)
+    }
+}
+}
+
[…]

### 2020-01-28 code/simulation_figure/Figure4/Figure4.R

@@ -1,0 +1,71 @@
+### Figure: MSE vs. sample size (|Omega|), under various rank##
+source("../../functions.R")
+set.seed(1)
+Output=NULL
+d=rep(20,3)
+rlist=c(3,5,8)
+ntotal=30
+
+rholist=seq(from=0.3,to=1,by=0.1)
+MSE_raw=array(0,dim=c(length(rlist),length(rholist),ntotal))
+for(j in 1:length(rlist)){
+        r=rep(rlist[j],3)
+        B_1 = randortho(d[1])[,1:r[1]]
+        B_2 = randortho(d[2])[,1:r[2]]
+        B_3 = randortho(d[3])[,1:r[3]]
+        D = rand_tensor(modes = r)
+        theta = ttl(D,list(B_1,B_2,B_3),ms=1:3)@data
+        alpha=10
+        theta = theta/max(abs(theta))*alpha ## specify signal = 10
+        omega = logit((1:4)/5) ## equal spaced on the probit scale
+        
+        ttnsr <- realization(theta,omega)@data
+        
+    for(i in 1:length(rholist)){
+        for(nsim in 1:ntotal){
+        rho=rholist[i]
+        data=array(NA,dim=dim(ttnsr))
+        Obs=sample(1:prod(d),prod(d)*rho,replace=F)
+        data[Obs]=ttnsr[Obs]
+        #initial point
+        A_1 = randortho(d[1])[,1:r[1]]
+        A_2 = randortho(d[2])[,1:r[2]]
+        A_3 = randortho(d[3])[,1:r[3]]
+        C = rand_tensor(modes = r)
+        C=C/max(abs(ttl(C,list(A_1,A_2,A_3),ms=1:3)@data))*alpha/10 ## initial theta is in the interior 
+        thetainit=ttl(C,list(A_1,A_2,A_3),ms=1:3)@data
+        
+        result <- fit_ordinal(data,C,A_1,A_2,A_3,omega=TRUE,alpha=2*alpha) #relaxed bound
+        MSE_raw[j,i,nsim]=mean((theta-result$theta)^2)/mean(theta^2)
+        }
+    }
+}
+
+Output=data.frame(MSE=c(t(apply(MSE_raw,c(1:2),mean))),rank=as.factor(rep(rlist,rep(length(rholist),length(rlist)))),rho=rep(rholist,length(rlist)),sd=c(t(apply(MSE_raw,c(1:2),sd))))
[…]

### 2020-01-28 code/simulation_figure/Figure5/Figure5.R

@@ -1,0 +1,80 @@
+### Figure 5: MSE vs. number of ordinal levels; half observations ##
+source("../../functions.R")
+set.seed(1)
+Output=NULL
+d=rep(20,3)
+rlist=c(3,5,8)
+rho=0.5
+Klist=2:7
+ntotal=30
+MSE_raw=array(0,dim=c(length(rlist),length(Klist),ntotal))
+for(j in 1:length(rlist)){      
+        r=rep(rlist[j],3)
+        B_1 = randortho(d[1])[,1:r[1]]
+        B_2 = randortho(d[2])[,1:r[2]]
+        B_3 = randortho(d[3])[,1:r[3]]
+        D = rand_tensor(modes = r)
+        theta = ttl(D,list(B_1,B_2,B_3),ms=1:3)@data
+        alpha=10 #
+        theta = theta/max(abs(theta))*alpha ## specify signal = 10
+        
+        for(i in 1:length(Klist)){
+        K=Klist[i]
+        omega = logit((1:(K-1))/K) ## equal space on the logit scale
+        
+        
+        for(nsim in 1:ntotal){ 
+        ttnsr <- realization(theta,omega)@data
+        data=array(NA,dim=dim(ttnsr))
+        Obs=sample(1:prod(d),prod(d)*rho,replace=F)
+        data[Obs]=ttnsr[Obs]
+        #initial point
+        A_1 = randortho(d[1])[,1:r[1]]
+        A_2 = randortho(d[2])[,1:r[2]]
+        A_3 = randortho(d[3])[,1:r[3]]
+        C = rand_tensor(modes = r)
+        C=C/max(abs(ttl(C,list(A_1,A_2,A_3),ms=1:3)@data))*alpha/10 ## initial theta is in the interior 
+        thetainit=ttl(C,list(A_1,A_2,A_3),ms=1:3)@data
+        
+        result <- fit_ordinal(data,C,A_1,A_2,A_3,omega=TRUE,alpha=2*alpha) ## a slightly relaxed boundary
+        
+        MSE_raw[j,i,nsim]=mean((theta-result$theta)^2)/mean(theta^2) ## more stable using rescaled MSE
+            }
+    }
+}
+
+Output=data.frame(MSE=c(t(apply(MSE_raw,c(1:2),mean))),rank=as.factor(rep(rlist,rep(length(Klist),3))),K=rep(Klist,length(rlist)),sd=c(t(apply(MSE_raw,c(1:2),sd))))
[…]

### 2020-01-28 code/simulation_figure/experiment.R

@@ -1,0 +1,223 @@
+source("functions.R")
+########## brain tensor analysis ##########
+load("../data/dti_brain.RData")
+##ttnsr=tensor[,,attr[,2]=="22-25"] ## subset of individuals
+ttnsr=tensor
+d=dim(ttnsr)
+
+r=c(25,25,7)
+### random initilization
+A_1 = randortho(d[1])[,1:r[1]]
+A_2 = randortho(d[2])[,1:r[2]]
+A_3 = randortho(d[3])[,1:r[3]]
+C = rand_tensor(modes = r)
+########## Music analysis #######################
+source("functions.R")###
+load("../data/InCar_Music.RData")
+ttnsr=tensor
+ttnsr[ttnsr==-1]=NA
+missing=which(is.na(ttnsr)==F)
+d=dim(ttnsr)
+set.seed(1)
+r=c(1,1,1) ## need to select via BIC ..
+alpha=100
+A_1 = as.matrix(randortho(d[1])[,1:r[1]])
+A_2 = as.matrix(randortho(d[2])[,1:r[2]])
+A_3 = as.matrix(randortho(d[3])[,1:r[3]])
+C = rand_tensor(modes = r)
+C=C/max(abs(ttl(C,list(A_1,A_2,A_3),ms=1:3)@data))*alpha/10
+
+
+############# experiement ##########################
+mask=sample(1:2884,2884*0.2,replace=T)
+index=which(is.na(ttnsr)==F)[mask]
+true=ttnsr[index]
+ttnsr[index]=NA
+##### Method 1: continuous tucker #########
+tic()
+result = fit_continuous(ttnsr,C,A_1,A_2,A_3,alpha=alpha)
+toc()
+thetahat=result$theta
+mean((true-thetahat[index])^2) #MSE= 18.96803
+mean(abs(true-thetahat[index])) #MAD= 2.589925
+mean(true!=round(thetahat[index])) ## MCR
+
+###### Method 2: ordinal tucker #########
+tic()
+result2 = fit_ordinal(ttnsr,C,A_1,A_2,A_3,omega=TRUE,alpha =100)
+toc()
+thetahat2=result2$theta
+bic(ttnsr,thetahat2,result2$omega,d,r)## 10860.46
+est=estimation(thetahat2,result2$omega,"mean")
[…]

### 2020-01-28 ICML/miaoyan/ordinalT.tex
update

--- 2020-01-24 ICML/miaoyan/ordinalT.tex
@@
 \icmltitlerunning{Ordinal tensor denoising and completion}
-
+\usepackage{wrapfig}
 \usepackage{multirow}
@@
 \twocolumn[
-\icmltitle{Optimal tensor denoising and completion based on ordinal observations}
+\icmltitle{Tensor denoising and completion based on ordinal observations}
 
@@
 % the "keywords" metadata in the PDF but will not be shown in the document
-\icmlkeywords{Higher-order tensors, ordinal data, tensor denoising, tensor completion}
+\icmlkeywords{Higher-order tensors, ordinal data, tensor decomposition, tensor completion}
 
@@
 Two related problems are studied, one on tensor denoising and another on tensor completion.
-%{\color{red}In contrast to the continuous-valued problems, problems with ordinal tensors exhibit an intrinsic phase transition depending on the signal-to-noise ratio.
+%{\color{red}In contrast to the continuous-valued problems, parameter estimation with ordinal observation exhibit a phase transition depending on the signal-to-noise ratio.
 (Not sure whether this should be mentioned)}
 We propose a multi-linear cumulative link model, develop a rank-constrained M-estimator, and obtain theoretical accuracy guarantees. 
-Our error bound enjoys a faster convergence rate than existing results, and we show that the proposed estimator is minimax optimal under the considered model. 
-An efficient completion method is further developed, which guarantees consistent recovery of an order-$K$ $(d,\ldots,d)$-dimensional tensor using only $O(Kd)$ noisy, quantized observations.
+Our performance bound enjoys a faster convergence rate than previous results, and we show that the proposed estimator is minimax optimal under the class of low-rank models.
+An efficient completion method is further developed, which guarantees consistent recovery of an order-$K$ $(d,\ldots,d)$-dimensional low-rank tensor using only $O(Kd)$ noisy, quantized observations.
 We demonstrate the outperformance of our approach over previous methods on the tasks of clustering and collaborative filtering.
@@
 For an order-$K$ dimensional-$(d,\ldots,d)$ ordinal tensor with $L$-level observations, we report the error bound in the recovered signal (for tensor denoising) and required sample complexity (for tensor completion) as functions of tensor dimensions.
-{\color{red}(font in the table seems odd...)}}~\label{tab:compare}
+{\color{red}(font in the table seems odd...) Add a row ``algorithm stability''}}~\label{tab:compare}
 \end{table*}
@@
 
-\subsection{Latent-variable interpretation.}
+\subsection{Latent-variable interpretation}\label{sec:latent}
 
@@
 
-\subsection{Problem 1: Tensor denoising.}~\label{sec:denoising}
+\subsection{Problem 1: Tensor denoising}~\label{sec:denoising}
 The first question we aim to address is tensor denoising:
 
-(P1) Given the quantization process induced by $f$ and the cut-off points $\mb$, how accurate can we recover the latent signal tensor $\Theta$ from the ordinal observation $\tY$?
+(P1) Given the quantization process induced by $f$ and the cut-off points $\mb$, how accurate can we estimate the latent signal tensor $\Theta$ from the ordinal observation $\tY$?
 
@@
 The first is that $\Theta$ is a low-rank tensor, with $r_k=O(1)$ for all $k\in[K]$.
+Equivalently, $\Theta$ admits the Tucker decomposition:
+\begin{equation}\label{eq:Tucker}
+\Theta=\tC\times_1\mM_1\times_1 \cdots \times_K \mM_K,
+\end{equation}
+where $\tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is a core tensor, $\mM_k\in\mathbb{R}^{d_k\times r_k}$ are factor matrices with orthogonal columns, and $\times_k$ denotes the tensor-by-matrix multiplication~\cite{kolda2009tensor}.
 The Tucker low-rankness is popularly used in tensor data analysis, and is shown to provide a reasonable tradeoff between model complexity and model flexibility.
@@
 
-(P2) Given an incomplete set of ordinal observation $\{y_{\omega}\}_{\omega\in\Omega}$ based on the model~\eqref{eq:model}, how many sampled entries do we need to consistently estimate $\Theta$?
+(P2) Given an incomplete set of ordinal observation $\{y_{\omega}\}_{\omega\in\Omega}$ based on the model~\eqref{eq:model}, how many sampled entries do we need to consistently recover $\Theta$?
 
@@
 
-\subsection{Estimation error for tensor denoising}
+\subsection{Estimation error for tensor denoising}\label{sec:denosing}
 For the tensor denosing problem, we assume the full set of tensor entries are observed. In such a case, we assess the estimation accuracy using the mean squared error (MSE):
@@
 Our bound also generalizes the previous results on ordinal matrices.
-The convergence rate for rank-constrained matrix denoising was $O(1/\sqrt{d})$~\citep{bhaskar2016probabilistic}, which fits into our special case when $K=2$.
+The convergence rate for rank-constrained matrix estimation was $O(1/\sqrt{d})$~\citep{bhaskar2016probabilistic}, which fits into our special case when $K=2$.
 Furthermore, our results~\eqref{eq:rate} and~\eqref{eq:KLrate} reveal that the convergence becomes favorably as the order of tensor data increases.
@@
 \section{Numerical Implementation}\label{sec:algorithm}
-\subsection{Alternating optimization}
+%\subsection{Alternating optimization}
 
 \begin{algorithm}[tb]
-        \caption{Ordinal tensor decomposition }\label{alg}
+        \caption{Ordinal tensor decomposition ({\color{red} allowing missing data})}\label{alg}
         \begin{algorithmic}[]
@@
             \FOR{$k = 1,2,\cdots,K$}
-            \STATE{Update $\mM_k$ while fixing other blocks.}
-            \STATE $\mM_k^{(t+1)}\gets\argmax_{\mM_k\in\mathbb{R}^{d_k\times r_K}}\mathcal{L}_{\mathcal{Y},\Omega}(\mM_k)$
+            \STATE{Update $\mM_k$ while fixing other blocks:}
+            \STATE $\mM_k^{(t+1)}\gets\argmax_{\mM_k\in\mathbb{R}^{d_k\times r_K}}\mathcal{L}_{\mathcal{Y},\Omega}(\mM_k)$,
             \newline
@@
             \ENDFOR
-            \STATE {Update $\mathcal{C}$ while fixing other blocks: $\mathcal{C}^{(t+1)} \gets \argmax_{\mathcal{C}\in \mathbb{R}^{r_1\times\cdots\times r_K}}\mathcal{L}_{\mathcal{Y},\Omega}(\mathcal{C})$, s.t.
-$\mnormSize{}{\Theta^{(t+1)}}\leq \alpha$.
-               \STATE {Update $\Theta$ based on the current block estimates: $\Theta^{(t+1)} \gets \mathcal{C}^{(t+1)}\times_1\mM_1^{(t+1)}\cdots\times_K\mM_K^{(t+1)}$. 
-            \STATE {Update $\mb$ while fixing $\Theta^{(t+1)}$: $\mb^{(t+1)} \gets \argmax_{\mb\in \mathbb{R}^{L-1}}\mathcal{L}_{\mathcal{Y},\Omega}\big(\Theta^{(t+1)},\mb\big)$
+            \STATE {Update $\mathcal{C}$ while fixing other blocks:}
+            \STATE{$\mathcal{C}^{(t+1)} \gets \argmax_{\mathcal{C}\in \mathbb{R}^{r_1\times\cdots\times r_K}}\mathcal{L}_{\mathcal{Y},\Omega}(\mathcal{C})$, s.t.
+$\mnormSize{}{\Theta^{(t+1)}}\leq \alpha$}
+               \STATE {Update $\Theta$ based on the current block estimates: 
+               \STATE{$\Theta^{(t+1)} \gets \mathcal{C}^{(t+1)}\times_1\mM_1^{(t+1)}\cdots\times_K\mM_K^{(t+1)}$}
+            \STATE {Update $\mb$ while fixing $\Theta^{(t+1)}$: \\
+          \STATE{$\mb^{(t+1)} \gets \argmax_{\mb\in \mathbb{R}^{L-1}}\mathcal{L}_{\mathcal{Y},\Omega}\big(\Theta^{(t+1)},\mb\big)$}
             \ENDFOR
@@
 One approach to handle this problem is utilizing the Tucker decomposition and converting optimization into a block-wise convex problem.
-Let us denote the low Tucker-rank tensor as
-\begin{equation}\label{eq:Tucker}
-\Theta=\tC\times_1\mM_1\times_1 \cdots \times_K \mM_K,
-\end{equation}
-where $\tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is a core tensor and $\mM_k\in\mathbb{R}^{d_k\times r_k}$ are factor matrices with orthogonal columns.
 From~\eqref{eq:objective} and \eqref{eq:Tucker}, we have $K+2$ blocks of variables in the objective function, one for the cut-points vector $\mb$, one for the core tensor $\tC$, and $K$ for the factor matrices $\mM_k$'s.
@@
 
-\subsection{Tuning parameter selection}\label{sec:BIC}
-Algorithm~1 takes the rank $\mr$ as an input.
+Before concluding this section, we comment on two implementation details.
+First, the problem~\eqref{eq:estimator} is non-convex, so Algorithm 1 usually has no theoretical guarantee on global optimality.
+Nevertheless, as shown in Section~\ref{sec:denosing}, the desired rate holds not only for the global optimizer, but also for the local optimizer with $\logl(\hat \Theta)\geq \logl(\trueT)$.
+We find that, in practice, the convergence point $\hat \Theta$ is often satisfactory, in that the corresponding objective $\logl(\hat \Theta)$ is close to and actually slightly larger than the objective evaluated at the true parameter $\logl(\trueT)$.
+Figure~\ref{fig:stability} shows the trajectory of the objective function that is output in the default setting of Algorithm 1, with the input tensor generated from probit model~\eqref{eq:model} using $d_1=d_2=d_3=d$ and $r_1=r_2=r_3=r$.
+The dashed line is the objective value at the true parameter $\logl(\trueT)$.
+It is observed that the algorithm generally converges quickly to a desirable value in reasonable number of steps.
+The actual running time per iteration is shown in the plot legend.
+In the case of complete observation, each sub-routine in the algorithm scales linear with the sample size $\tO(\prod_kd_k)$, and this complexity matches the classical tensor methods~\cite{kolda2008scalable,anandkumar2014tensor}.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=5cm]{algorithm.pdf}\label{fig:stability}
+\vspace{-.4cm}
+\caption{Trajectory of the objective function with varing $d$ and $r$.}
+\end{figure}
+%\subsection{Tuning parameter selection}\label{sec:BIC}
+\vspace{-.4cm}
+Second, the algorithm takes the rank $\mr$ as an input.
 In practice, the rank $\mr$ is hardly known and needs to be estimated from the data.
@@
 \hat \mr&=\argmin_{\mr\in\mathbb{N}^K_{+}} \text{BIC}(\mr)\\
-&=\argmin_{\mr\in\mathbb{N}^K_{+}}\left[-2\tL_{\tY}(\hat \Theta(\mr))+p_e(\mr)\log \left(\prod_k d_k\right) \right],
+&=\argmin_{\mr\in\mathbb{N}^K_{+}}\left[-2\tL_{\tY}(\hat \Theta(\mr),\hat \mb)+p_e(\mr)\log (\prod_k d_k) \right],
 \end{align}
@@
 Our choice of BIC aims to balance between the goodness-of-fit for the data and the degree of freedom in the population model.
-We test its empirical performance in Section~\ref{sec:simulation}.
 
 \section{Experiments}
-In this section, we evaluate the empirical performance of our rank-constrained M-estimatation.
-We consider ordinal tensors with various number of categories, and compare the recovery accuracy with other tensor-based methods.
-Unless otherwise stated, we generate ordinal-valued tensor under the latent-variable model~\eqref{eq:latent} with i.i.d.\ $N(0,\sigma^2)$ noise.
-The signal tensor is simulated based on~\eqref{eq:Tucker}, with core tensor generated from i.i.d.
-N(0,1), and the factor matrices uniformly from orthogonal ensembles.
-The cut-off points are generated i.i.d.from $N(0,1)$.
+In this section, we evaluate the empirical performance of our estimation method.
+We consider both the complete and the incomplete settings, and compare the recovery accuracy of our estimator with other tensor-based methods.
+Unless otherwise stated, we generated $L$-level ordinal data tensors from model~\eqref{eq:model} using standard probit link $f$.
+We consider order-3 parameter tensors with $d_1=d_2=d_3=d$ and $r_1=r_2=r_3=r$.
+The parameter tensor is simulated based on~\eqref{eq:Tucker}, where the entries in the core tensor $\tC$ are i.i.d.\ drawn from $N(0,1)$, and the factors $\mM_k$ are uniformly sampled (with respect to Haar measure) from matrices with orthonormal columns.
+We set the cut-off points $b_\ell=f^{-1}(\ell/L)$ for $\ell=0,\ldots,L$, such that $f(b_\ell)$ are evenly spaced from 0 to 1.
 In each simulation study, we report the summary statistics across $n_{\text{sim}}$ = 30 replications.
@@
 \subsection{Finite-sample performance}\label{sec:simulation}
+We first consider the case of complete observation, and assess the empirical relationship between the MSE and model complexity.
+In particular, we consider various settings of dimension $d$, rank $r$ and signal level $\alpha=\mnormSize{}{\Theta}$.
+Figure~\ref{fig:finite}a plots the estimation error versus the tensor dimension $d$ for three different ranks $r\in\{3,5,8\}$.
+The decay in the error appears to behave on the order of $d^{-2}$, which is consistent with our theoretical results~\eqref{eq:rate}.
+It is seen that a higher rank leads to a larger error, as reflected by the upward shift of the curve as $r$ increases.
+Indeed, a higher rank implies a higher intrinsic dimension, thus increasing the difficulty of the estimation.
+
+Figure~\ref{fig:finite}b shows the estimation error versus the signal level.
+A larger estimation error is observed when the signal is either too small or too large.
+This non-monotonic phenomenon may seem surprising, but we believe this is an intrinsic feature with ordinal-valued date problem.
+In light of the latent-variable interpretation (see Section~\ref{sec:latent}), estimation from ordinal observation can be interpreted as an inverse problem of quantization.
+The estimation error will diverge when the noise $\tE$ is negligible, because there is no way to distinguish two different signal tensors, say, $\Theta_1=\ma_1\otimes \ma_2\otimes \ma_3$ and $\Theta_2=\text{sign}(\ma_1)\otimes \text{sign}(\ma_2)\otimes \text{sign}(\ma_3)$, from the quantized observations.
+This behavior is clearly contrary to that of classical continuous-valued tensor problem, and the phenomenon has attracted great interest recently~\cite{davenport2014,sur2018modern}.
+
+We then investigate the case of incomplete observation.
+We sample tensor entries via uniform sampling and assess the recovery error with sample complexity.
+Figure~\ref{fig:finite}c shows the completion error versus the sample size.
+The error reduces roughly at a rate of $\tO(d^{-2})$(?) as the sample size increases.
+This suggests....
+
+%\begin{figure}[http]
+%\includegraphics[width=8.3cm]{panel.pdf}
+%\end{figure}
+
+\begin{figure}[http]
+\includegraphics[width=8.3cm]{panel.pdf}
+\vspace{-.4cm}
+\caption{Finite sample performance}\label{fig:finite}
+\end{figure}
+\vspace{-.5cm}
+
+%\begin{figure}[http]
+%\includegraphics[width=4.2cm]{error_level.pdf}
+%\end{figure}
 
 \subsection{Comparison with alternative methods}
+
+Compare with 
+\begin{enumerate}
+\item continuous tensor method~\cite{filipovic2015tucker}; 
+\item 1-bit tensor method~\cite{ghadermarzy2018learning}; 
+\item ordinal matrix method~\cite{bhaskar2016probabilistic}. 
+\end{enumerate}
 
@@
 \hline
-MSE& &  & 0.1504&0.1604\\
-MAD&  &  &0.1504&0.1600\\
-MCR& &&0.1502&0.1598\\
+MSE& 0.1504&  0.1504 & 0.1504&0.1604\\
+MAD& 0.1503&0.1503 &0.1503&0.1600\\
+MCR& 0.1503& 0.1502&0.1502& 0.1598\\
 \hline
@@
 }
-\caption{Comparison of tensor denoising on HCP dataset.}
+\caption{Comparison of tensor decomposition on HCP dataset.}
 \end{table}
@@
 In contrast, these three estimators degenerate to a single estimator under the continuous-valued tensor decomposition model.
-We assess the accuracy using three metrics: Mean squared error (MSE), Mean absolute deviation (MAD), and Misclassification rate (MCR).
+We then round $\hat y_\omega$ to the nearest integer and assess the accuracy using three metrics: Mean squared error (MSE), Mean absolute deviation (MAD), and Misclassification rate (MCR).
 

### 2020-01-28 code/functions.R
update

@@ -185,7 +185,7 @@ fit_ordinal = function(ttnsr,C,A_1,A_2,A_3,omega=TRUE,alpha = TRUE){
   cost=NULL
   omg = omega
   k=length(unique(as.factor(c(ttnsr))))
-    while ((error > 10^-12)&(iter<15) ) {
+    while ((error > 10^-4)&(iter<50) ) {
       iter = iter +1
       #update omega
       prevtheta <- ttl(C,list(A_1,A_2,A_3),ms=1:3)@data

### 2020-01-28 code/HPC_analysis.R
update

@@ -1,10 +1,11 @@
 ### Cross validation ######
 source("functions.R")
-load("dti_brain.RData")
+load("../data/dti_brain.RData")
 set.seed(18)
-ind = sample(1:length(tensor),length(tensor))
 
-### 5 fold ################ Please make changes to the sampling #####
+
+### 5 fold ################ To Chanwoo: Please changes the sampling per instruction #####
+ind = 1:length(tensor)
 l = length(tensor)%/%5+1
 i1 = ind[1:l]
 i2 = ind[(l+1):(2*l)]

### 2020-01-28 ICML/miaoyan/ordinalT.tex
udate

@@
             \STATE {Update $\mathcal{C}$ while fixing other blocks:}
-            \STATE{$\mathcal{C}^{(t+1)} \gets \argmax_{\mathcal{C}\in \mathbb{R}^{r_1\times\cdots\times r_K}}\mathcal{L}_{\mathcal{Y},\Omega}(\mathcal{C})$, s.t.
-$\mnormSize{}{\Theta^{(t+1)}}\leq \alpha$}
-               \STATE {Update $\Theta$ based on the current block estimates: 
-               \STATE{$\Theta^{(t+1)} \gets \mathcal{C}^{(t+1)}\times_1\mM_1^{(t+1)}\cdots\times_K\mM_K^{(t+1)}$}
-            \STATE {Update $\mb$ while fixing $\Theta^{(t+1)}$: \\
-          \STATE{$\mb^{(t+1)} \gets \argmax_{\mb\in \mathbb{R}^{L-1}}\mathcal{L}_{\mathcal{Y},\Omega}\big(\Theta^{(t+1)},\mb\big)$}
+              \newline$\mathcal{C}^{(t+1)} \gets \argmax_{\mathcal{C}\in \mathbb{R}^{r_1\times\cdots\times r_K}}\mathcal{L}_{\mathcal{Y},\Omega}(\mathcal{C})$, s.t.
+$\mnormSize{}{\Theta^{(t+1)}}\leq \alpha$
+               \STATE {Update $\Theta$ based on the current block estimates:} 
+                   \newline
+                   $\Theta^{(t+1)} \gets \mathcal{C}^{(t+1)}\times_1\mM_1^{(t+1)}\cdots\times_K\mM_K^{(t+1)}$
+            \STATE {Update $\mb$ while fixing $\Theta^{(t+1)}$:}
+              \newline
+              $\mb^{(t+1)} \gets \argmax_{\mb\in \mathbb{R}^{L-1}}\mathcal{L}_{\mathcal{Y},\Omega}\big(\Theta^{(t+1)},\mb\big)$
             \ENDFOR
@@
 
-\begin{table}[H]
-\resizebox{\linewidth}{!}{%
-\begin{tabular}{c|ccc|c}
-\hline
-\multirow{2}{*}{Critera}&\multicolumn{3}{c|}{Our method}&Continuous method\\
-& $\hat y^{(\text{Mean})}_\omega$ & $\hat y^{(\text{Median})}_\omega$ & $\hat y^{(\text{Mode})}_\omega$ &~\cite{filipovic2015tucker}\\
-\hline
-MSE& 0.1504&  0.1504 & 0.1504&0.1604\\
-MAD& 0.1503&0.1503 &0.1503&0.1600\\
-MCR& 0.1503& 0.1502&0.1502& 0.1598\\
-\hline
+\begin{table}[t]
+\label{CV}
+\vskip 0.15in
+\begin{center}
+\begin{small}
+\begin{sc}
+\begin{tabular}{lcccr}
+\toprule
+ Method & RMSE & MAE & ERROR \\
+\midrule
+\begin{tabular}[c]{@{}c@{}}Oridnal tensor\\ Decomposition\end{tabular}  & 0.1504 & 0.1503& 0.1502 \\ \\
+\begin{tabular}[c]{@{}c@{}}Conti tensor\\ Decomposition\end{tabular} & 0.1604& 0.1600& 0.1598\\ \\
+\begin{tabular}[c]{@{}c@{}}1bit-completion\\ (probit)\end{tabular}  & 0.3658 & 0.3633 & 0.3620\\
+\begin{tabular}[c]{@{}c@{}}1bit-completion\\ (logistic)\end{tabular}  &  0.4519& 0.4219 & 0.4068
+\\
+\bottomrule
 \end{tabular}
-}
-\caption{Comparison of tensor decomposition on HCP dataset.}
+\end{sc}
+\end{small}
+\end{center}
+\vskip -0.1in
+\caption{Results of comparisons among 4 methods on the HCP data predicting the test data.
+Four methods are the ordinal tensor decomposition algorithm, the continuous tensor decomposition algorithm and the 1 bit tensor completion method with probit link function and logistic link function.
+Each method is evaluated by RMSE, MAE, error(false prediction rate).}
 \end{table}
+
+%\begin{table}[H]
+%\resizebox{\linewidth}{!}{%
+%\begin{tabular}{c|ccc|c}
+%\hline
+%\multirow{2}{*}{Critera}&\multicolumn{3}{c|}{Our method}&Continuous method\\
+%& $\hat y^{(\text{Mean})}_\omega$ & $\hat y^{(\text{Median})}_\omega$ & $\hat y^{(\text{Mode})}_\omega$ &~\cite{filipovic2015tucker}\\
+%\hline
+%MSE& 0.1504&  0.1504 & 0.1504&0.1604\\
+%MAD& 0.1503&0.1503 &0.1503&0.1600\\
+%MCR& 0.1503& 0.1502&0.1502& 0.1598\\
+%\hline
+%\end{tabular}
+%}
+%\caption{Comparison of tensor decomposition on HCP dataset.}
+%\end{table}
 

### 2020-01-28 code/HPC_analysis.R

@@ -3,7 +3,6 @@ source("functions.R")
 load("../data/dti_brain.RData")
 set.seed(18)
 
-
 ### 5 fold ################ To Chanwoo: Please changes the sampling per instruction #####
 ind = 1:length(tensor)
 l = length(tensor)%/%5+1

### 2020-01-28 code/simulation_figure/Figure2/Figure2.R

@@ -48,7 +48,7 @@ levels(Output$rank)=c("r=3","r=5","r=8")
 
 p=ggplot(data=Output,aes(x=dim,y=MSE))+geom_line(aes(color=rank),size=1.5)+ labs(x='tensor dimension', y='MSE',size=1.5)+theme(text = element_text(size=rel(4)),legend.text = element_text(size = 15),axis.text.x = element_text(size=15),axis.text.y = element_text(size=15))
 
-p=p+stat_function(fun=function(d) 10^2.63/(d^2),lty=2,size=1.5)
+p=p+stat_function(fun=function(d) 10^2.96/(d^2),lty=2,size=1.5)
 g <- ggplot_build(p)
 col=unlist(unique(g$data[[1]]["colour"]))[c(1,2,3)]
 

### 2020-01-28 code/simulation_figure/Figure4/Figure4.R

@@ -45,6 +45,7 @@ Output=data.frame(MSE=c(t(apply(MSE_raw,c(1:2),mean))),rank=as.factor(rep(rlist,
 save(MSE_raw,Output,file="Figure4.RData")
 
 ################ plot ############
+load("Figure4.RData")
 library(ggplot2)
 Output$rank=as.factor(Output$rank)
 levels(Output$rank)=c("r=3","r=5","r=8")

### 2020-01-28 code/simulation_figure/Figure2/Figure2.R
update figures

@@ -48,7 +48,7 @@ levels(Output$rank)=c("r=3","r=5","r=8")
 
 p=ggplot(data=Output,aes(x=dim,y=MSE))+geom_line(aes(color=rank),size=1.5)+ labs(x='tensor dimension', y='MSE',size=1.5)+theme(text = element_text(size=rel(4)),legend.text = element_text(size = 15),axis.text.x = element_text(size=15),axis.text.y = element_text(size=15))
 
-p=p+stat_function(fun=function(d) 10^2.96/(d^2),lty=2,size=1.5)
+p=p+stat_function(fun=function(d) 10^2.96/(d^2),lty=2,size=0.9)
 g <- ggplot_build(p)
 col=unlist(unique(g$data[[1]]["colour"]))[c(1,2,3)]
 

### 2020-01-28 ICML/chanwoo/ICMLdraft.tex
Add files via upload

@@
 \section{Introduction}
-Multidimensional array data, a.k.a. a tensor, appears in a huge variety of applications including recommendation systems \citep{Kutty2012APR,Adomavicius2008ContextawareRS,Sun2015ProvableST},
-social networks \citep{Sun2009MultiVisCS,Nickel2011ATM},
-genomics \citep{Wang2017ThreewayCO},
-neuroimaging (EEG, fMRI) \citep{Miwakeichi_decomposingeeg}
-and signalprocessing \citep{Sidiropoulos2000ParallelFA,Cichocki2015TensorDF}.
-Instead of unfolding those data tensors into matrices where many analysis methods have been proposed, we preserve its inherent multi-modal structure.
+Multidimensional array datasets, a.k.a. a tensor, appear in a huge variety of applications including recommendation systems \citep{kutty2012people,adomavicius2011context,sun2017provable},
+social networks \citep{sun2009multivis,nickel2011three},
+genomics \citep{wang2018learning},
+neuroimaging (EEG, fMRI) \citep{miwakeichi2004decomposing}
+and signalprocessing \citep{sidiropoulos2000parallel,cichocki2015tensor}.
+Instead of unfolding those data tensors into matrices where many analysis methods have been proposed, we suggest preserving multi-modal tensor structure in this paper.
 Studying tensor data while respecting the structure allows us to examine complex interactions among tensor entries.
-Thereby we can provide extra more interpretation that cannot be addressed by traditional matrix analysis.
-Also, It has been shown that the tensor preserving analysis improves performance \citep{Zare2018ExtensionOP,Wang2018LearningFB}.
-With those reasons, there is a growing need to develop dimension reduction method without losing multi modal structure.
+Thereby we can expect to provide extra more interpretation that cannot be addressed by traditional matrix analysis.
+Furthermore, It has been shown that the tensor preserving analysis improves performance \citep{zare2018extension,wang2018learning}.
+With those reasons, there is a growing need to develop dimension reduction method without losing tensor structure.
 In the line of the attempts, a number of tensor decomposition methods have been proposed in many applications.
-CANDECOMP/PARAFAC (CP) decomposition was first introduced \citep{doi:10.1002/sapm19287139}
-and revitalized in psychometrics \citep{Harshman1970FoundationsOT} and in linguistics \citep{Smilde2004MultiwayAW}.
-The Tucker decomposition was proposed in psychometrics \citep{tucker64extension,RePEc:spr:psycho:v:31:y:1966:i:3:p:279-311}.
+CANDECOMP/PARAFAC (CP) decomposition was first introduced \citep{hitchcock1928multiple}
+and made use of in psychometrics \citep{harshman1970foundations} and in linguistics \citep{smilde2005multi}.
+The Tucker decomposition was proposed in psychometrics \citep{tucker1964extension,tucker1966some}.
 
 Classical tensor completion with those decompositions has treated the entries of data as real-valued.
-In many cases, However, we encounter data of which the entries are not real-valued but discrete or quantized i.e.
-binary-valued or multiple-valued.
-For example, many survey data takes the integer values.
-To be specific, the data in the Netflix problem has the 3 modes of `user',`movie' and `date of grade'.
-The entries of the data are the gradings from the users which take integer value from 1 to 5.
-Also, there are many cases that the data are quantized in real application.
-In signal processing, the data are frequently rounded or truncated so that only integer values are available.
-Also, in graph theory, adjacency matrix can be labeled as from 1 to 3 taking 3 when pairs of vertexes have strong connection and giving 1 when two vertexes have the weak connection according to a given threshold.
+In many cases, However, we encounter data sets of which the entries are not real-valued but discrete or quantized i.e.
+binary-valued or ordinal-valued.
+For example, many survey data sets take the integer values.
+To be specific, the data set in the Netflix problem has the three modes of `user',`movie' and `date of grade'.
+The entries of the data are the ratings from the users which take integer value from 1 to 5.
+Another example is the case where the data sets are quantized in real application.
+In signal processing, the data sets are frequently rounded or truncated so that only integer values are available.
+In network analysis, entries of adjacency matrix can be labeled as from 1 to 3 taking 3 when pairs of vertexes have strong connection and giving 1 when two vertexes have the weak connection according to a given threshold.
 If we add one more mode on adjacency matrix such as `context' or `individual', the data turns into tensor data with 3 integer values.
 
-Therefore, performance improvement can be achieved when the observations are treated as discrete value not as continuous value.
-In matrix case, there has been many achievements to complete matrix for discrete cases: Models for the case of binary or 1-bit were introduced and studied \citep{Davenport20121BitMC,Bhaskar20151bitMC}.
-Furthermore,  \citet{Bhaskar2016ProbabilisticLM} suggested matrix completion method for general ordinal observations.
-In tensor case, however, only binary tensor has gotten an attention and achieved performance improvement using binary tensor decomposition methods \citep{Hore2016TensorDF,Wang2018LearningFB,Hong2018GeneralizedCP,Hu2018TrainingBW}.
+We expect that performance improvement can be achieved when the observations are treated as discrete ordinal value not a continuous value.
+In matrix case, there has been many achievements to complete matrix for discrete cases: Models for the case of binary or 1-bit were introduced and studied \citep{davenport20141,bhaskar20151}.
+Furthermore,  \citet{bhaskar2016probabilistic} suggested matrix completion method for general ordinal observations.
+In tensor case, however, only binary tensor has gained an attention and achieved performance improvement using binary tensor decomposition methods \citep{hore2016tensor,wang2018learning,hong2018generalized,hu2018training}.
 Accordingly, a general method for the data which has more than 2 ordered label is needed.
 
-We organized this paper as follows.
-In section 2, we discuss detailed assumptions and descriptions about our probabilistic model in \eqref{multinomial}.
-Also, we suggest the estimation method for the latent parameters and related algorithm.
-In section 3, we provide the statistical properties of the upper, lower bounds and the phase-transition.
+We organized this paper as follows. We specify our notations that we will use over this paper in the next section.
+In section 3, we discuss detailed assumptions and descriptions about our probabilistic model. Also, we suggest the estimation method for the latent parameters and related algorithm.
+In section 4, we provide the statistical properties of the upper, lower bounds and the phase-transition.
 We then provide the numerical experiments.
-Our model is applied to  real-world data to check validity and performance in section 5.
+Our model is applied to  real-world data to check validity and performance in section 6.
 Finally, we wrap up the paper with a discussion.
@@
 \section{Model}
-\subsection{Low-rank cumulative model}
-For the K-mode ordinal tensor $\mathcal{Y} = [\![ y_\omega]\!]\in[L]^{d_1\times\cdots\times d_K}$, we assume that its entries are realizations of independent multinomial random variables, such that
+\subsection{Low-rank ordinal tensor model}
+For the $K$-mode ordinal tensor $\mathcal{Y} = [\![ y_\omega]\!]\in[L]^{d_1\times\cdots\times d_K}$, we assume that its entries are realizations of independent multinomial random variables, such that
 \begin{equation}
@@
         \label{mainmodel}
-        \mathbb{P}(y_\omega=l) = f_l(\theta_\omega), \quad \omega\in [d_1]\times\cdots\times[d_K].
+        \mathbb{P}(y_\omega\leq \ell) = f(\theta_\omega+b_\ell), \quad \omega\in [d_1]\times\cdots\times[d_K].
     \end{aligned}
 \end{equation}
-In this model, a twice differentiable function $f_l : \mathbb{R}\rightarrow [0,1]$ with $l\in [L]$ is strictly increasing and satisfies $\sum_l f_l(\theta) = 1$ for a fixed $\theta$.
+In this model, a twice differentiable function $f: \mathbb{R}\rightarrow [0,1]$  is known and a cumulative distribution function and $\mb = (b_1,\cdots,b_{L-1})$ is a set of unknown scalars satisfying $b_1<\cdots<b_{L-1}$.
+We refer to $\mb$ as the cut-off points and $f$ the link function.
 The tensor $\Theta = [\![\theta_\omega]\!]\in \mathbb{R}^{d_1\times\cdots\times d_K}$
-is a hidden parameter which we are interested in. We assume the parameter tensor $\Theta$ is continuous value and admits a rank $\mr = (r_1,\cdots,r_K)$ Tucker decomposition,
+is a hidden parameter which we are interested in.
+We assume that the entries of the parameter tensor $\Theta$ are continuous valued and $\Theta$ admits a rank $\mr = (r_1,\cdots,r_K)$ Tucker decomposition,
 \begin{equation}
@@
 \end{equation}
-where $\mathcal{C}\in \mathbb{R}^{r_1\times\cdots\times r_K}$ is a core tensor and $\mM_k\in \mathbb{R}^{d_k\times r_k}$, for $n \in [L]$ is a factor matrix.
-We can get information about the influence of each mode by checking factor matrices.
+where $\mathcal{C}\in \mathbb{R}^{r_1\times\cdots\times r_K}$ is a full rank core tensor and $\mM_k\in \mathbb{R}^{d_k\times r_k}$, for $n \in [L]$ is a factor matrix with orthogonal columns.
+This low rank tensor structure allows us to reduce dimension and complete tensors from available entries.
 
-The tensor model \eqref{mainmodel} has an equivalent representation as a latent model with K-level quantization. \citep{Davenport20121BitMC,Lan2014MatrixRF,Bhaskar20151bitMC,Cai2013AMC} where
+The ordinal tensor model \eqref{mainmodel} has the equivalent representation as a latent model with $L$-level quantization.
+\citep{davenport20141,lan2014matrix,bhaskar20151,cai2013max} where the entry of
 $\mathcal{Y} = [\![ y_\omega]\!]$ is a quantized value such that,
@@
 \end{equation}
-where $\mathcal{E} = [\![\epsilon_\omega]\!]$ is a noise tensor of $i.i.d.$ from cumulative distribution function $\Phi(\cdot)$ and the function $\mathcal{Q}:\mathbb{R}\rightarrow [L]$ is a quantizer having the following rule.
+where $\mathcal{E} = [\![\epsilon_\omega]\!]$ is a noise tensor i.i.d.
+entries from cumulative distribution function $\mathcal{F}_\epsilon$ such that $f(\theta) = \mathbb{P}(\epsilon \geq -\theta)$ and  $\mathcal{Q}:\mathbb{R}\rightarrow [L]$ is a quantizer having the following rule.
 \begin{equation}
@@
     \begin{aligned}
-        \mathcal{Q}(x) = l,\quad \text{ if } b_{l-1}< x \leq b_{l},\quad l\in [L],
+        \mathcal{Q}(x) = \ell,\quad \text{ if } b_{\ell-1}< x \leq b_{\ell},\quad \text{ for all } \ell\in [L],
     \end{aligned}
 \end{equation}
-where $\mb = (b_1,\cdots,b_{L-1})$ is a cutpoints vector such that $ -\infty = b_0<b_1<\cdots<b_L = \infty$.
-That is, the entries of observed tensor $\mathcal{Y}$ fall in category $l$ when the associated entries of the latent tensor $\Theta+\mathcal{E}$ fall in the $l$th interval of values.
-Then we have
-\begin{equation}
-    \label{cumulative}
-    \begin{aligned}
-        f_l(\theta_\omega) &= \mathbb{P}(y_\omega = l)\\
-        &= \Phi(b_l-\theta_\omega)-\Phi(b_{l-1}-\theta_\omega).
-    \end{aligned}
-\end{equation}
-We can diversify our model by the choices of $\Phi(\cdot)$, or equivalently the distribution of $\mathcal{E}$ with given $\mb$.
-The followings are 2 common choices of $\Phi(\cdot)$.
-\begin{example}[Logistic link/Logistic noise] The logistic model is represented by \eqref{mainmodel} with $f_l(\theta) = \Phi_{log}(\frac{b_l-\theta}{\sigma})-\Phi_{log}(\frac{b_{l-1}-\theta}{\sigma})$  where $\Phi_{log}(x/\sigma) = (1+e^{-x/\sigma})$.
+That is, the entries of observed tensor $\mathcal{Y}$ fall into category $\ell$ when the associated entries of the latent tensor $\Theta+\mathcal{E}$ fall into the $\ell$-th interval of values.
+From this point of view, we can see the latent parameter $\Theta$ as the signal prior to contamination and quantization.
+We can diversify our model by the choices of $f$, or equivalently the distribution of $\mathcal{E}$.
+The followings are two common choices of $f$.
+
+\begin{example}[Logistic link/Logistic noise] The logistic model is represented by \eqref{mainmodel} with $f(\theta) = \Phi_{log}(\theta/\sigma)$  where $\Phi_{log}(x/\sigma) = (1+e^{-x/\sigma})$.
 Equivalently, the noise $\epsilon_\omega$ in \eqref{latentmd} follows i.i.d.
@@
 
-\begin{example}[Probit link/Gaussian noise] The probit model is represented by \eqref{mainmodel} with $f_l(\theta) = \Phi_{norm}(\frac{b_l-\theta}{\sigma})-\Phi_{norm}(\frac{b_{l-1}-\theta}{\sigma})$ where $\Phi_{norm}$ is the cumulative distribution function of the standard Gaussian.
-Equivalently, the noise $\epsilon_\omega$ in \eqref{latentmd} follows i.i.d.$N(0,\sigma^2)$.
+\begin{example}[Probit link/Gaussian noise] The probit model is represented by \eqref{mainmodel} with $f(\theta) = \Phi_{norm}(\theta/\sigma)
+$ where $\Phi_{norm}$ is the cumulative distribution function of the standard Gaussian. Equivalently, the noise $\epsilon_\omega$ in \eqref{latentmd} follows i.i.d.$N(0,\sigma^2)$.
 \end{example}
@@
 \subsection{Rank-constrained likelihood-based estimation}
-Our goal is to estimate unknown parameter tensor $\Theta$ and a cutpoints vector $\mb$ from observed tensor $\mathcal{Y}$ using a constrained likelihood approach.
-With a little abuse of notation, we use $\Omega$ to denote either the full index set $\Omega = [d_1]\times\cdots\times[d_K]$ or a random subset indeuced from the subsampling distribution.
+Our goal is to estimate unknown parameter tensor $\Theta$ and cut-off points $\mb$ from observed tensor $\mathcal{Y}$ using a constrained likelihood approach.
+With a little abuse of notation, we use $\Omega$ to denote either the full index set $\Omega = [d_1]\times\cdots\times[d_K]$ or a random subset induced from the subsampling distribution.
 The log-likelihood function for \eqref{mainmodel} is
@@
     \begin{aligned}
-    \mathcal{L}_{\mathcal{Y},\Omega}(\Theta,\mb) = \sum_{\omega\in\Omega}\Big[\sum_{l\in[L]}\log(f_l(\theta_\omega))\mathds{1}_{\{y_\omega = l\}}\Big],
+    \mathcal{L}_{\mathcal{Y},\Omega}(\Theta,\mb) = \sum_{\omega\in\Omega}\Big[\sum_{\ell\in[L]}\log(g_\ell(\theta_\omega))\mathds{1}_{\{y_\omega = \ell\}}\Big],
     \end{aligned}
 \end{equation}
-where the cutpoints vector $\mb$ is implicitly contained in the function $f_l$.
+where $g_\ell(x) = \mathbb{P}(y_\omega = \ell) = f(\theta_\omega+b_\ell)-f(\theta_\omega+b_{\ell-1})$ from \eqref{mainmodel}.
+We define $b_0 = -\infty,b_L = \infty$ so that $f(b_0) = 0, f(b_L) = 1$.
+The cut-off points $\mb$ is implicitly contained in the function $g_\ell$.
 Considering the Tucker structure in \eqref{tucker}, we have the following constrained optimization problem.
@@
         &\max_{(\Theta, \mb)\in \mathcal{D}\times\mathcal{B}}\mathcal{L}_{\mathcal{Y},\Omega}(\Theta,\mb),\text{ where }\\
-        &\mathcal{D} = \{\Theta\in \mathbb{R}^{d_1\times\cdots\times d_K}: \rank(\Theta) = \bm{r} \text{ and } \mnormSize{}{\Theta}\leq \alpha \}\\
+        &\mathcal{D} = \{\Theta\in \mathbb{R}^{d_1\times\cdots\times d_K}: \rank(\Theta) \leq \bm{r} \text{ and } \mnormSize{}{\Theta}\leq \alpha \}\\
         &\mathcal{B} = \{\mb\in\mathbb{R}^{L-1}:  b_1<\cdots < b_{L-1}\},
@@
 This bound condition is a technical assumption to help to recover $\Theta$ in the noiseless case.
-Similar conditions has been imposed in many literatures for the matrix case \citep{Davenport20121BitMC,Bhaskar20151bitMC,Cai2013AMC,Bhaskar2016ProbabilisticLM}.
-The search space $\mathcal{B}$ makes sure that the probability function $f_l$ in \eqref{cumulative} is positive.
+Similar conditions has been imposed in many literatures for the matrix case \citep{davenport20141,bhaskar20151,cai2013max,bhaskar2016probabilistic}.
+The search space $\mathcal{B}$ makes sure that the probability function $g_\ell$ in \eqref{likelihood} is strictly positive.
 
@@
 In this section, we describe the algorithm to seek the optimizer of \eqref{eq:optimization}.
-The objective function $\mathcal{L}_{\mathcal{Y},\Omega}(\Theta,\mb)$ is concave in $(\Theta,\mb)$ whenever $\Phi(x)$ is log-concave \citep{McCullagh1980RegressionMF,Burridge1981ANO}.
+The objective function $\mathcal{L}_{\mathcal{Y},\Omega}(\Theta,\mb)$ is concave in $(\Theta,\mb)$ whenever $f(x)$ is log-concave \citep{mccullagh1980regression,burridge1981note}.
 However, the feasible set $\mathcal{D}$ is not a convex set, which makes the optimization \eqref{eq:optimization} a non-convex problem.
 One approach to handle this problem is utilizing the Tucker decomposition and converting optimization into a block-wise convex problem.
-Let us denote the objective function as
-\begin{equation}
-    \label{blockwise}
-    \begin{aligned}
-         \mathcal{L}(\mathcal{C},\mM_1,\cdots,\mM_K,\mb) = \mathcal{L}_{\mathcal{Y},\Omega}(\mathcal{C}\times_1\mM_1\cdots\times_K\mM_K,\mb).
-    \end{aligned}
-\end{equation}
-From \eqref{blockwise}, we have $K+2$ blocks of variables in the objective function, one for the cutpoints vector $\mb$, one for the core tensor $\mathcal{C}$ and  K for the factor matrices $\mM_K$'s.
+From \eqref{likelihood} and \eqref{tucker}, we have $K+2$ blocks of variables in the objective function, one for the cut-points vector $\mb$, one for the core tensor $\mathcal{C}$ and  $K$ for the factor matrices $\mM_K$'s.
 We can change the optimization problem to simple convex problem if any $K+1$ out of the $K+2$ blocks being fixed.
-Therefore, we can alternatively update one block at a time while other blocks being fixed.
+Therefore, we can alternately update one block at a time while other blocks being fixed.
 The algorithm \ref{alg} gives the full description.
 
-\begin{algorithm}[tb]
+\begin{algorithm}[]
         \caption{Ordinal tensor decomposition }\label{alg}
@@
             \FOR{k = 1,2,$\cdots$,K}
-            \STATE{Update $\mM_k$ fixing other blocks.}
-            \STATE $\mM_k^{(t+1)}\gets\argmax_{\mM_k\in\mathbb{R}^{d_k\times r_k}}\mathcal{L}_{\mathcal{Y},\Omega}(\mM_k)$ s.t.
-             $\|\mathcal{C}^{(t)}\times_1\mM_1^{(t+1)}\cdots\times_k\mM_k\cdots\times_K\mM_K^{(t)}\|_\infty \leq \alpha$
+            \STATE{Update $\mM_k$ while fixing other blocks:}
+            \STATE $\mM_k^{(t+1)}\gets\argmax_{\mM_k\in\mathbb{R}^{d_k\times r_k}}\mathcal{L}_{\mathcal{Y},\Omega}(\mM_k)$
+            \newline
+             s.t. $\mnormSize{}{\Theta^{(t+1)}}\leq \alpha$, where $\Theta^{(t+1)}$ is the parameter tensor based on the current block estimates.
             \ENDFOR
-            \STATE {Update $\mathcal{C}$ fixing other blocks.}
+            \STATE {Update $\mathcal{C}$ while fixing other blocks:}
             \STATE $\mathcal{C}^{(t+1)} \gets \argmax_{\mathcal{C}\in \mathbb{R}^{r_1\times\cdots\times r_k}}\mathcal{L}_{\mathcal{Y},\Omega}(\mathcal{C})$
-            s.t.\newline $\|\mathcal{C}\times_1\mM_1^{(t+1)}\cdots\times_K\mM_K^{(t+1)}\|_\infty \leq \alpha$
-
+            s.t.  $\mnormSize{}{\Theta^{(t+1)}}\leq \alpha$
+            \STATE Update $\Theta$ based on the current block estimates:
             \STATE $\Theta^{(t+1)} \gets \mathcal{C}^{(t+1)}\times_1\mM_1^{(t+1)}\cdots\times_K\mM_K^{(t+1)}$
-            \STATE {Update $\mb$ fixing $\Theta^{(t+1)}$.}
-            \STATE $\mb^{(t+1)} \gets \argmax_{\mb\in \mathbb{R}^{L-1}}\mathcal{L}_{\mathcal{Y},\Omega}\big(\Theta^{(t+1)},\mb\big)$
+            \STATE {Update $\mb$ while fixing $\Theta^{(t+1)}$:}
+            \STATE $\mb^{(t+1)} \gets \argmax_{\mb\in \mathbb{R}^{L-1}}\mathcal{L}_{\mathcal{Y},\Omega}\big(\mb\big)$
             \ENDFOR
@@
         \hat{\bm r} &= \argmin_{\bm r\in \mathbb{N}_+^K} BIC(\bm r)\\
-        &=\argmin_{\bm r\in \mathbb{N}_+^K}\bigg[-2\mathcal{L}_\mathcal{Y}(\hat\Theta(\bm r)) \quad+
-        \\&\qquad\big(\prod_{k\in[K]} r_k+\sum_{k\in[K]}(d_kr_k-r_k^2)\big)\log(\prod_{k\in[K]} d_k)\bigg],
+        &=\argmin_{\bm r\in \mathbb{N}_+^K}\bigg[-2\mathcal{L}_{\mathcal{Y},\Omega}(\hat\Theta(\bm r),\hat\mb(\mr)) +
+        p_e(\mr)\log\big(\prod_{k\in[K]} d_k\big)\bigg],
     \end{aligned}
 \end{equation}
-where $\hat\Theta(\bm r)$ is a maximum likelihood estimater given the rank $\bm r$.
-The second part of the summation in \eqref{bic} is the number of  independent parameters in the model.
+where $\hat\Theta(\bm r),\hat\mb(\mr)$ is a maximum likelihood estimate given the rank $\bm r$, and $p_e(\rm)\eqdef \sum_k(d_k-r_k)r_k = \prod_k r_k$ is the effective number of free parameters in the model.
 We select the rank $\hat{\bm r}$ that minimizes BIC value through the grid search method.
@@
 In this section, we apply our ordinal tensor decomposition method to two real-world datasets of ordinal tensors.
-In the first application, we use our model to analyze an ordinal tensor consisting of structural connectivity patterns among 68 brain regions for 136 individuals from Human Connectome Project (HCP) \citep{Geddes2016HumanBM}.
+In the first application, we use our model to analyze an ordinal tensor consisting of structural connectivity patterns among 68 brain regions for 136 individuals from Human Connectome Project (HCP) \citep{geddes2016human}.
 In the second application, we perform tensor completion from the data with missing values.
-The data tensor records the ratings from scale 1 to 5 of 42 users to 139 songs on 26 contexts \citep{Baltrunas2011InCarMusicCM}.
+The data tensor records the ratings from scale 1 to 5 of 42 users to 139 songs on 26 contexts \citep{baltrunas2011incarmusic}.
 \subsection{Human Connectome Project (HCP)}
 The human connectome project (HCP) is a $68\times 68 \times 136$ tensor where the first two modes have 68 indices representing brain regions and the last mode has 136 indices meaning individuals.
-All the individual images were preprocessed following a standard pipeline \citep{Zhang2018MappingPS}, and the brain was parcellated to 68 regions of interest following the Desikan atlas \citep{Desikan2006AnAL}.
+All the individual images were preprocessed following a standard pipeline \citep{zhang2018mapping}, and the brain was parcellated to 68 regions of interest following the Desikan atlas \citep{desikan2006automated}.
 The tensor entries consist of $\{1,2,3\}$, the strength of fiber connections between 68 brain regions for each of the 136 individuals.
-First, we apply our ordinal tensor decomposition method with a logistic link function to the HCP data and identify similar brain nodes based on latent parameters.
+We apply our ordinal tensor decomposition method with a logistic link function to the HCP data.
 The BIC result suggests $\bm r = (23,23,8)$  with $\mathcal{L}_\mathcal{Y}(\hat{\Theta},\hat{\mb}) = -216645.8$.
 
-Also, we compare our ordinal tensor decomposition method with a logistic link function with other methods: the continuous tensor decomposition method, the 1 bit-tensor completion method with probit link function and the 1 bit-tensor completion method with logistic link function.
+We check the performance of our model and compare with other methods: the continuous tensor decomposition method, the 1 bit-tensor completion method with probit link function and the 1 bit-tensor completion method with logistic link function.
 We use 5 folded cross validation method for the comparison.
-Specifically, we randomly split the tensor entries into 5 similar sized pieces and alternatively use each piece of entries as a test data using other $80\%$ entries as a training data.
+Specifically, we randomly split the tensor entries into 5 similar sized pieces and alternately use each piece of entries as a test data using other $80\%$ entries as a training data.
 The entries in the test data are encoded as missing and then predicted based on each method from the training data.
-We set the rank of the tensor as $\bm r = (23,23,8)$ for the tensor decomposition methods which minimizes BIC value.
-Table \ref{CV} shows RMSE, MAE and error(false prediction rate), averaged over 5 test set results.
+In our model,  predict $y_\omega$ as $\hat{y}_\omega^{\text{(mode)}} = \argmax_\ell(g_\ell(
+\hat\theta_\omega,\hat\mb)).$ Table \ref{CV} shows RMSE, MAE and error(false prediction rate), averaged over 5 test set results.
 We can check that our ordinal tensor decomposition model outperforms the other methods in all criteria.
@@
 \toprule
- Method & RMSE & MAE & ERROR \\
+ Method & MSE & MAD & MCR \\
 \midrule
-\begin{tabular}[c]{@{}c@{}}Oridnal tensor\\ Decomposition\end{tabular}  & 0.1503711 & 0.1502662 & 0.1502151 \\ \\
-\begin{tabular}[c]{@{}c@{}}Conti tensor\\ Decomposition\end{tabular} & 0.1603685& 0.1600219& 0.1598499\\ \\
-\begin{tabular}[c]{@{}c@{}}1bit-completion\\ (probit)\end{tabular}  & 0.3658390 & 0.3632757 & 0.3619940\\
-\begin{tabular}[c]{@{}c@{}}1bit-completion\\ (logistic)\end{tabular}  &  0.4519253 & 0.4218513 & 0.4068143 
+\begin{tabular}[c]{@{}c@{}}Oridnal tensor\\ Decomposition\end{tabular}  & 0.1503 & 0.1502& 0.1502\\ \\
+\begin{tabular}[c]{@{}c@{}}Conti tensor\\ Decomposition\end{tabular} & 0.1603& 0.1600& 0.1598\\ \\
+\begin{tabular}[c]{@{}c@{}}1bit-completion\\ (probit)\end{tabular}  & 0.3658 & 0.3632 & 0.3619\\
+\begin{tabular}[c]{@{}c@{}}1bit-completion\\ (logistic)\end{tabular}  &  0.4514 & 0.4215 & 0.4066
 \\
@@
 
-\subsection{Music data with missing values}
+\subsection{InCarMusic recommendation system}
+InCarMusic is a mobile application that offers music recommendation to passengers of cars based on contexts (Baltrunas et al.
+2011).
+To be specific, the music recommendations can be different according to the mood of the driver or the traffic condition in this system.
+Our goal is to perform the tensor completion to the $42\times 139\times 26$ ordinal tensor and thereby we can offer context-specific music recommendation to users.
+The tensor entries consist of ordinal observations on the scale 1 to 5, the ratings of 42 users to 139 songs on 26 contexts and are encoded as NA for missing values.
+The number of missing values is 148,904 and the number of available values is 2,884.
+We suggest three estimators $\hat y_\omega$ based on $(\hat \theta_\omega,\hat \mb)$:
+\begin{itemize}
+\item (Mean) $\hat y^{(\text{Mean})}_\omega=\sum_\ell \ell g_\ell(\hat \theta_\omega,\hat \mb) $;
+\item (Median) $\hat y^{(\text{Median})}_\omega=\min\{\ell\in[L]\colon f_\ell(\hat \theta_\omega,\hat \mb)\geq 0.5\}$;
+\item (Mode) $\hat y^{(\text{Mode})}_\omega=\argmax_\ell g_\ell(\hat \theta_\omega,\hat \mb) $
+\end{itemize}
+Note that, under the ordinal tensor model~\eqref{mainmodel}, the estimator $\hat y^{(\text{Mean})}_\omega$ minimizes $\mathbb{E}_{\hat{\theta}_\omega,\hat{\mb}}(y_\omega-y)^2$, the estimator $\hat y^{(\text{Median})}_\omega$ minimizes $\mathbb{E}_{\hat{\theta}_\omega,\hat{\mb}}|y_\omega-y|$
+and the estimator $\hat y^{(\text{Mode})}_\omega $ minimizes $\mathbb{E}_{\hat{\theta}_\omega,\hat{\mb}}\mathds{1}_{\{y_\omega=y\}}$.
+In contrast, these three estimators degenerate to a single estimator under the continuous-valued tensor decomposition model.
+We assess the accuracy using three metrics: Mean squared error (MSE), Mean absolute deviation (MAD), and Misclassification rate (MCR).
 
-InCarMusic is a mobile application that offers music recommendation to passengers of cars based on contexts \citep{Baltrunas2011InCarMusicCM}.
-Our goal is to get the tensor completion from the $42\times139\times 26$ tensor using the ordinal tensor decomposition and thereby we can offer context-specific music recommendation to users.
-The tensor entries consist of $\{1,2,3,4,5\}$, the ratings of 42 users to 139 songs on 26 contexts and are encoded as $-1$ for missing values.
-The number of missing values is $148904$ and the number of available values is $2884$.
+Results are averaged from 5 fold cross validation. {\color{red} add s.e. to the above table.}
+
+\begin{table}[H]
+\resizebox{\linewidth}{!}{%
+\begin{tabular}{c|ccc|c}
+\hline
+\multirow{2}{*}{Critera}&\multicolumn{3}{c|}{Our method}&Continuous method\\
+& $\hat y^{(\text{Mean})}_\omega$ & $\hat y^{(\text{Median})}_\omega$ & $\hat y^{(\text{Mode})}_\omega$ &~\cite{filipovic2015tucker}\\
+\hline
+MSE& \color{red}2.01&2.14 & 3.64&8.18\\
+MAD& 1.23  &  {\color{red}1.13}&1.26&2.36\\
+MCR& 0.80&0.75&{\color{red}0.54}&0.96\\
+\hline
+\end{tabular}
+}
+\caption{Comparison of tensor completion performance on InCarMusic dataset.}
+\end{table}
+
+We also attempted to run matrix ordinal mehod~\citep{bhaskar2016probabilistic} to the InCarMusic dataset. Unfortunately, the $\mb$ cannot be estimated from their method.
 

### 2020-01-28 ICML/chanwoo/Supplements.tex
Add files via upload

\documentclass[11pt]{article}

\usepackage{fancybox}

\usepackage{color}
\usepackage{url}
\usepackage[margin=1in]{geometry}

\renewcommand{\textfraction}{0.0}
\renewcommand{\topfraction}{1.0}
%\renewcommand{\textfloatsep}{5mm}

\usepackage{comment}
% Definitions of handy macros can go here
\usepackage{amsmath,amssymb,amsthm,bm,mathtools}
\usepackage{multirow}
\usepackage{natbib}
%\usepackage{dsfont,multirow,hyperref,setspace,natbib,enumerate}
\usepackage{dsfont,multirow,hyperref,setspace,enumerate}
\hypersetup{colorlinks,linkcolor={blue},citecolor={blue},urlcolor={red}}
\usepackage{algpseudocode,algorithm}
\algnewcommand\algorithmicinput{\textbf{Input:}}
\algnewcommand\algorithmicoutput{\textbf{Output:}}
\algnewcommand\INPUT{\item[\algorithmicinput]}
\algnewcommand\OUTPUT{\item[\algorithmicoutput]}

\mathtoolsset{showonlyrefs=true}

\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{pro}{Property}
\newtheorem{assumption}{Assumption}

\theoremstyle{definition}
\newtheorem{defn}{Definition}
\newtheorem{cor}{Corollary}
\newtheorem{example}{Example}
\newtheorem{rmk}{Remark}

\renewcommand{\thefigure}{{S\arabic{figure}}}%
\renewcommand{\thetable}{{S\arabic{table}}}%
\renewcommand{\figurename}{{Supplementary Figure}}
\renewcommand{\tablename}{{Supplementary Table}}
\setcounter{figure}{0}
\setcounter{table}{0}

\def\MLET{\hat \Theta_{\text{MLE}}}
\newcommand{\cmt}[1]{{\leavevmode\color{red}{#1}}}

\usepackage{dsfont}

\usepackage{multirow}

\DeclareMathOperator*{\minimize}{minimize}

\usepackage{mathtools}
\mathtoolsset{showonlyrefs}
\newcommand*{\KeepStyleUnderBrace}[1]{%f
  \mathop{%
    \mathchoice
    {\underbrace{\displaystyle#1}}%
    {\underbrace{\textstyle#1}}%
    {\underbrace{\scriptstyle#1}}%
    {\underbrace{\scriptscriptstyle#1}}%
  }\limits
}
\usepackage{xr}
\externaldocument{ordinalT}
\input macros.tex

\title{Some supplements}

%\author{%
%Chanwoo Lee \\
%University of Wisconsin -- Madison\\
 %\texttt{chanwoo.lee@wisc.edu} \\
%\And
%Miaoyan Wang \\
%University of Wisconsin -- Madison\\

%\texttt{miaoyan.wang@wisc.edu} \\
%}

\begin{document}

\begin{center}
\begin{spacing}{1.5}
\textbf{\Large Some supplements}
\end{spacing}
\end{center}
\section{Convexity}
\begin{thm}
\begin{align}
 \logl(\Theta, \mb)&=\sum_{\omega\in\Omega}\sum_{\ell\in[L]} \Big\{\mathds{1}_{\{y_\omega=\ell\}} \log \big[f(\theta_\omega+b_\ell)-  f(\theta_\omega+b_{\ell-1})\big]\Big\}, \text{ where } f(x) = \frac{e^x}{1+e^x}.
 \end{align}
is a concave function to $(\Theta,\mb)$
\end{thm}
\begin{proof}
It is enough to show that $\lambda(u,v) = \log \big[f(u)-  f(v)\big]$
is a concave function to $(u,v)$ where $u>v$. Because if $\lambda(u,v)$ is a concave function and $u,v$ are both linear functions of $(\Theta,\mb) $ such that $u = \ma_1^T(\Theta,\mb), v = \ma_2^T(\Theta,\mb)$, then $\lambda(u,v) = \lambda(\ma_1^T(\Theta,\mb),\ma_2^T(\Theta,\mb)))$ is a concave function to $(\Theta,\mb)$ by the definition of the convexity. With the fact that  $\logl(\Theta, \mb)$
can be written as the form of summations of $\lambda(u,v)$, we can conclude that $\logl(\Theta, \mb) $ is a concave function to $(\Theta,\mb)$ if we prove $\lambda(u,v)$ is concave. Write $\lambda(u,v) = \log\big[f(u)-f(v)\big]$ as $\log\big[\int\mathds{1}_{(u,v)}(x)f'(x)dx\big]$. Notice that  $\log\mathds{1}_{(u,v)}(x)$ is concave to $(u,v,x)$ and $\log f'(x)$ is concave to $x$ because $f'(x) = \frac{e^x}{(1+e^x)^2}$. Thus, $\log\mathds{1}_{(u,v)}(x)f'(x)$
is a concave function to $(u,v,x)$. Therefore, the $\lambda(u,v)$ is a concave function because the following lemma says that the integral of a log concave function with respect to some of its arguments is a log concave function of its remaining argument.
\end{proof}

\begin{lem}[Corollary 3.5 in~\cite{brascamp2002extensions}]
Let $F(x,y):\mathbb{R}^{m+n}\rightarrow \mathbb{R}$ be an integrable function where $x\in \mathbb{R}^{m},y\in \mathbb{R}^n$. Let $$G(x) = \int_{\mathbb{R}^n}F(x,y)dy.$$ If $F(x,y)$ is a log concave function to (x,y), then $G(x)$ is a log concave function.
\end{lem}

\section{Property of an estimator}

\begin{pro}
$y^{(\text{Median})}_\omega$ minimizes $R(y) = \mathbb{E}_{\hat{\theta}_\omega,\hat{\mb}}|y_\omega-y|$.
\end{pro}
\begin{proof}
We can write $R(y) = \sum_{\ell\in[L]}|\ell-y|\hat f_\ell$ where $\hat f_\ell = f_\ell(\hat{\theta}_\omega,\hat\mb)$. Let us denote $\mu = y^{(\text{Median})}_\omega$. Then,
\begin{align*}
  R(y)&=
  \begin{cases}
  \big[- (\hat f_1 \cdots +\hat f_L)\big]y + \big[(L\hat f_L+\cdots +1\hat f_1)\big],& \text{if $y\in(-\infty,1]$},\\
  \big[\hat f_1 - (\hat f_2 \cdots +\hat f_L)\big]y + \big[(L\hat f_L+\cdots +2\hat f_2) -1\hat f_1\big],& \text{if $y\in(1,2]$},\\
  \vdots & \vdots\\
  \big[(\hat f_1 +\cdots + \hat f_{\mu-1}) - (\hat f_{\mu} \cdots +\hat f_L)\big]y + \big[(L\hat f_L+\cdots +\mu\hat f_\mu) -((\mu-1)\hat f_{\mu-1}+\cdots+1\hat f_1)\big],& \text{if $y\in(\mu-1, \mu]$},\\
  \big[(\hat f_1 +\cdots + \hat f_{\mu}) - (\hat f_{\mu+1} \cdots +\hat f_L)\big]y + \big[(L\hat f_L+\cdots +(\mu+1)\hat f_{\mu+1}) -(\mu\hat f_\mu+\cdots+1\hat f_1)\big],& \text{if $y\in(\mu, \mu+1]$} \\
  \vdots  &\vdots\\
  \big[(\hat f_1 +\cdots + \hat f_{L-1}) - \hat f_L)\big]y + \big[(L\hat f_L -((L-1)\hat f_{L-1}+\cdots+1\hat f_1)\big],& \text{if $y\in(L-1, L]$},\\
  \big[(\hat f_1 +\cdots + \hat f_{L})\big]y + \big[(L\hat f_L+\cdots+1\hat f_1)\big],& \text{if $y\in(L, \infty]$}.\\
  \end{cases}
\end{align*}
Based on the above formula, we can find out that $R(y)$ is minimized when $y = \mu$ because $ (\hat f_1 +\cdots + \hat f_{\mu-1}) - (\hat f_{\mu} \cdots +\hat f_L)<0 $ and $ (\hat f_1 +\cdots + \hat f_{\mu}) - (\hat f_{\mu+1} \cdots +\hat f_L)>0$ by the definition of $\mu$.
\end{proof}

\section{Theorems in the case when the cut points is unknown}
When the cut points $\mb$ is unknown, we estimate $(\hat\Theta,\hat\mb)$ by alternately fixing each factor and proceed until each factor does not change from each update. We have the following relationship between $\hat\Theta$ and $\hat\mb$
\begin{equation}
  \begin{aligned}
    \hat\Theta = \argmax_{\Theta\in \mathcal{P}} \logl (\Theta|\hat\mb)\quad \text{ and }\quad \hat\mb = \argmax_{\mb\in \mathcal{B}}\logl(\mb|\hat\Theta).
  \end{aligned}
\end{equation}
We assess the estimation accuracy using the mean squared error (MSE):
$$\text{MSE}(\hat\Theta,\hat\mb) = \frac{1}{\prod_k d_k}\FnormSize{}{(\hat\Theta,\hat\mb)-(\Theta^{\text{true}},\mb^{\text{true}})}. $$
We show that $\frac{1}{\prod_k d_k}\FnormSize{}{(\hat\Theta,\hat\mb)-(\hat\Theta,\mb^{\text{true}})} $ is bounded by $o(\frac{1}{\prod_k d_k})$ with high probability so that $$ \frac{1}{\prod_k d_k}\FnormSize{}{(\hat\Theta,\hat\mb)-(\Theta^{\text{true}},\mb^{\text{true}})} \leq \frac{1}{\prod_k d_k}\FnormSize{}{\hat\Theta-\Theta^{\text{true}}} + o(\frac{1}{\prod_k d_k}).$$
Therefore, we can utilize Theorem 4.1 to establish the upper bound for the proposed $(\hat\Theta,\hat\mb).$
We define a few key quantities that will be used in the proof.
\begin{align}
 U_\alpha &= \max_{\ell\in[L-1],|\theta|\leq \alpha} \max(\frac{\dot{f}(\theta+b_\ell)}{f(\theta+b_\ell)-f(
\theta+b_{\ell-1})},\frac{\dot{f}(\theta+b_\ell)}{f(\theta+b_{\ell+1})-f(
\theta+b_{\ell})})\\
L_\alpha & = \min_{\ell\in[L-1],|\theta|\leq \alpha} \min(-\frac{\ddot{f}(\theta+b_\ell)[f(\theta+b_\ell)-f(\theta+b_{\ell-1})]-\dot{f}(\theta+b_\ell)[\dot{f}(\theta+b_\ell)-\dot{f}(\theta+b_{\ell-1})]}{[f(\theta+b_\ell)-f(\theta+b_{\ell-1})]^2},
\\ &\hspace{1in}\qquad\frac{\ddot{f}(\theta+b_\ell)[f(\theta+b_{\ell+1})-f(\theta+b_{\ell})]-\dot{f}(\theta+b_\ell)[\dot{f}(\theta+b_{\ell+1})-\dot{f}(\theta+b_{\ell})]}{[f(\theta+b_{\ell+1})-f(\theta+b_{\ell})]^2})
\end{align}
We can see $U_\alpha>0$ and $L_\alpha>0$ from the assumption. Especially when $f(x) = \frac{1}{1+e^{-x}}$ , we have $$L_\alpha = \min_{\ell\in[L-1],|\Theta|\leq\alpha}\min\bigg(\frac{e^{\theta+b_\ell} (e^{\theta+b_\ell} - e^{\theta+b_{\ell-1}})^2}{(e^{\theta+b_\ell} + 1)^4 (e^{\theta+b_{\ell-1}} + 1)^2},\frac{e^{\theta+b_\ell} (e^{\theta+b_{\ell+1}} - e^{\theta+b_\ell})^2}{(e^{\theta+b_{\ell+1}} + 1)^2 (e^{\theta+b_\ell} + 1)^4}\bigg).$$
\begin{thm}
  From the same setting in Theorem 4.1 except the knowledge of $\mb$, with very high probability, the estimator satisfies
  \begin{equation}
    \frac{1}{\prod_k d_k}\FnormSize{}{(\hat\Theta,\hat\mb)-(\hat\Theta,\mb^{\text{true}})} = o(\frac{1}{\prod_k d_k}).
  \end{equation}
\end{thm}

\begin{proof}
It follows from the expression of $\logl(\Theta,\mb)$ that
\begin{align}\label{eq:property}
{\partial \flogl\over \partial \beta_\ell}&=\sum_{\omega\in\Omega}\bigg[\mathds{1}_{\{y_{\omega}=\ell\}}
\frac{\dot{f}(\theta_\omega+b_\ell)}{f(\theta_\omega+b_\ell)-f(
\theta_\omega+b_{\ell-1})}-\mathds{1}_{\{y_{\omega}=\ell+1\}}
\frac{\dot{f}(\theta_\omega+b_\ell)}{f(\theta_\omega+b_{\ell+1})-f(\theta_\omega+b_{\ell})}\bigg],\notag\\
{\partial^2 \flogl\over \partial b_\ell^2}&=\sum_{\omega\in\Omega}\bigg[\mathds{1}_{\{y_\omega=\ell\}}\frac{\ddot{f}(\theta_\omega+b_\ell)[f(\theta_\omega+b_\ell)-f(\theta_\omega+b_{\ell-1})]-\dot{f}(\theta_\omega+b_\ell)^2}{[f(\theta_\omega+b_\ell)-f(\theta_\omega+b_{\ell-1})]^2}\\
&\hspace{.6in}-\mathds{1}_{\{y_\omega=\ell+1\}}\frac{\ddot{f}(\theta_\omega+b_\ell)[f(\theta_\omega+b_{\ell+1})-f(\theta_\omega+b_{\ell})]+\dot{f}(\theta_\omega+b_\ell)^2}{[f(\theta_\omega+b_{\ell+1})-f(\theta_\omega+b_{\ell})]^2}\bigg] \text{ for } \ell \in [L-1],\\
{\partial^2 \flogl\over \partial b_\ell \partial b_{\ell+1}}&=\sum_{\omega\in\Omega}\mathds{1}_{\{y_\omega=\ell+1\}}\frac{\dot{f}(\theta_\omega+b_\ell)\dot{f}(\theta_\omega+b_{\ell+1})}{[f(\theta_\omega+b_{\ell+1})-f(\theta_\omega+b_\ell)]^2} \text{ for } \ell \in [L-2] \text{ and } {\partial^2 \flogl\over \partial b_i \partial b_j} = 0 \text{ if } |i-j|>1.
\end{align}
Therefore, all entries in $\frac{1}{\prod_k d_k}\fploglb$ are upper bounded $U>0$, and  $\frac{1}{\prod_k d_k}\fpploglb$ is a tridiagonal matrix.

By the second-order Taylor's expansion of $\flogl(\mb|\hat\Theta)$ around $\mb^{\text{true}}$, we obtain
\begin{equation}\label{eq:taylor}
\flogl(\hat\mb|\hat\Theta)=\flogl(\mb^{\text{true}}|\hat\Theta)+(\trueb-\hat\mb)^T\fploglb(\trueb)+ (\trueb-\hat\mb)^T\fpploglb(\check\mb)(\trueb-\hat\mb),
\end{equation}
$\check\mb=\gamma\trueb+(1-\gamma)\hat\mb$ for some $\gamma\in[0,1]$, and $\fpploglb(\check\mb)$ denotes the $(L-1)$-by-$(L-1)$ Hession matrix evaluated at $\check\mb$.

We first bound the linear term in~\eqref{eq:taylor}. Note that, by Cauchy-Schwartz inequality,
\begin{equation}\label{eq:linear}
(\trueb-\hat\mb)^T\fploglb(\trueb)\leq \|\trueb-\hat\mb\|\|\fploglb(\trueb)\|
\leq \|\trueb-\hat\mb\|U_\alpha\sqrt{L} \prod_k d_k .\end{equation}
The last inequality is followed by
\[
{\partial \tL_\tY\over \partial b_\ell}\Big|_{\mb=\trueb} \leq U_\alpha \;\; \textrm{ for all } \; l\in [L-1].
\]

We next bound the quadratic term in \eqref{eq:taylor}. Note that
\begin{align}\label{eq:quadratic}
 &(\trueb-\hat\mb))^T \fpploglb(\check{\mb})(\trueb-\hat\mb)\\
 &=\sum_{\ell\in [L-1]} \left( {\partial^2\tL_{\tY}\over \partial b^2_\ell} \Big|_{\mb=\check\mb} \right)(\hat b_\ell-b_{{\text{true}},\ell})^2 + 2\sum_{\ell\in[L-1]-\{1\}}\left( {\partial^2\tL_{\tY}\over \partial b_\ell \partial b_{\ell-1}} \Big|_{\mb=\check\mb} \right)(\hat b_\ell-b_{{\text{true}},\ell})(\hat b_{\ell-1}-b_{{\text{true}},\ell-1}) \nonumber \\
&\leq \sum_{\ell\in [L-1]} \left( {\partial^2\tL_{\tY}\over \partial b^2_\ell} \Big|_{\mb=\check\mb} \right)(\hat b_\ell-b_{{\text{true}},\ell})^2 + \sum_{\ell\in[L-1]-\{1\}}\left( {\partial^2\tL_{\tY}\over \partial b_\ell \partial b_{\ell-1}} \Big|_{\mb=\check\mb} \right)\big[(\hat b_\ell-b_{{\text{true}},\ell})^2 +(\hat b_{\ell-1}-b_{{\text{true}},\ell-1})^2\big] \nonumber \\
&=\left( {\partial^2\tL_{\tY}\over \partial b^2_1}+{\partial^2\tL_{\tY}\over \partial b_1 \partial b_2 }\Big|_{\mb=\check\mb} \right)(\hat b_1-b_{{\text{true}},1})^2+\left( {\partial^2\tL_{\tY}\over \partial b^2_{L-1}}+{\partial^2\tL_{\tY}\over \partial b_{L-2} \partial b_{L-1} }\Big|_{\mb=\check\mb} \right)(\hat b_{L-1}-b_{{\text{true}},L-1})^2  \\
&+\sum_{\ell\in[L-2]-\{1\}}\left( {\partial^2\tL_{\tY}\over \partial b^2_\ell}+{\partial^2\tL_{\tY}\over \partial b_\ell \partial b_{\ell-1}}+{\partial^2\tL_{\tY}\over \partial b_{\ell+1} \partial b_{\ell}} \Big|_{\mb=\check\mb} \right)(\hat b_\ell-b_{{\text{true}},\ell})^2     \\
&\leq - L_\alpha\prod_k d_k \sum_{\ell \in [L-1]}(\hat b_\ell-b_{\text{true},\ell})^2 \nonumber \\
&=-L_\alpha\prod_k d_k\FnormSize{}{\hat\mb-\trueb}^2,
\end{align}
Therefore, combining~\eqref{eq:taylor} and the above linear term and quadratic term results, we have that,
\[
\tL_\tY(\hat\mb|\hat\Theta)\leq \tL_{\tY}(\trueb|\hat\Theta)+U_\alpha\sqrt{L}\prod_k d_k\FnormSize{}{\hat\mb-\trueb} -\frac{L_\alpha}{2}\prod_k d_k \FnormSize{}{\hat\mb-\trueb}^2 .
\]

Since $\hat \mb=\arg\max_{\mb\in\tB}\tL_\tY(\mb|\hat\Theta)$, $\tL_\tY(\hat\mb|\hat\Theta))-\tL_\tY(\trueb|\hat\Theta)\geq 0$, which gives
\[
U_\alpha\sqrt{L}\FnormSize{}{\hat\mb-\trueb} -\frac{L_\alpha}{2} \FnormSize{}{\hat\mb-\trueb}^2\geq 0.
\]
Henceforth,
\[
\FnormSize{}{\hat\mb-\trueb}\leq \frac{2U_\alpha \sqrt{L}}{L_\alpha}.
\]
Finally, this complete the proof.
\begin{equation}
  \frac{1}{\prod_k d_k}\FnormSize{}{(\hat\Theta,\hat\mb)-(\hat\Theta,\mb^{\text{true}})} =\frac{1}{\prod_k d_k}\FnormSize{}{\hat\mb-\trueb}
  = o(\frac{1}{\prod_k d_k}).
\end{equation}.

\end{proof}

\bibliography{ICML.bib}
\bibliographystyle{icml2020}
\end{document}

### 2020-02-02 ICML/chanwoo/ICMLdraft.tex
Add files via upload

@@
 The BIC result suggests $\bm r = (23,23,8)$  with $\mathcal{L}_\mathcal{Y}(\hat{\Theta},\hat{\mb}) = -216645.8$.
+First, we estimate latent parameters $\Theta$ from the full dataset and cluster the brain nodes based on $K$-means method from the estimated latent parameters.
+We decide the number of clusters as eight from the elbow method.
+The cluster result shows that most of the brain nodes fall into two biggest groups which can be represented as right-side brain and left-side brain.
+The brain nodes, in each group except the above two biggest clusters, have the similar names and belong to the same location(right or left).
+For example, the fourth cluster group has the nodes having names such as \textit{l.postcentral, l.posteriorcingulate, l.pericalcarine and l.temporalpole,} all of which are encoded as ``SupraM$\_$L'' in the dataset.
+The brain clustering based on the latent estimator successfully catches regional traits of the brain nodes without external knowledge.
+More detailed procedure and result are provided in the supplements.
 
-We check the performance of our model and compare with other methods: the continuous tensor decomposition method, the 1 bit-tensor completion method with probit link function and the 1 bit-tensor completion method with logistic link function.
-We use 5 folded cross validation method for the comparison.
-Specifically, we randomly split the tensor entries into 5 similar sized pieces and alternately use each piece of entries as a test data using other $80\%$ entries as a training data.
-The entries in the test data are encoded as missing and then predicted based on each method from the training data.
-In our model,  predict $y_\omega$ as $\hat{y}_\omega^{\text{(mode)}} = \argmax_\ell(g_\ell(
-\hat\theta_\omega,\hat\mb)).$ Table \ref{CV} shows RMSE, MAE and error(false prediction rate), averaged over 5 test set results.
-We can check that our ordinal tensor decomposition model outperforms the other methods in all criteria.
+Secondly, we check the performance of our model and compare with other methods: the continuous tensor decomposition method\cite{filipovic2015tucker}, the 1 bit-tensor completion method\cite{ghadermarzy2018learning}.
+In 1 bit-tensor completion method, we use the probit link instead of the logistic link because the probit link has the better performance.
+We use stratified 5-folded cross validation method to assess the performances.
+Specifically, we randomly split the tensor entries into five pieces with the same size making sure that every piece has the similar proportion of ordinal observations.
+We alternately use each piece of entries as a test dataset and use the other $80\%$ entries as a training dataset.
+The entries of the test dataset are encoded as missing and then predicted based on each method from the training dataset.
+In our model, We suggest three estimators $\hat y_\omega$ based on $(\hat \theta_\omega,\hat \mb)$:
+\begin{itemize}
+\item (Mean) $\hat y^{(\text{Mean})}_\omega=\sum_\ell \ell g_\ell(\hat \theta_\omega,\hat \mb) $;
+\item (Median) $\hat y^{(\text{Median})}_\omega=\min\{\ell\in[L]\colon f_\ell(\hat \theta_\omega,\hat \mb)\geq 0.5\}$;
+\item (Mode) $\hat y^{(\text{Mode})}_\omega=\argmax_\ell g_\ell(\hat \theta_\omega,\hat \mb) $
+\end{itemize}
+Notice that, under the ordinal tensor model~\eqref{mainmodel}, the estimator $\hat y^{(\text{Mean})}_\omega$ minimizes $\mathbb{E}_{\hat{\theta}_\omega,\hat{\mb}}(y_\omega-y)^2$, the estimator $\hat y^{(\text{Median})}_\omega$ minimizes $\mathbb{E}_{\hat{\theta}_\omega,\hat{\mb}}|y_\omega-y|$
+and the estimator $\hat y^{(\text{Mode})}_\omega $ minimizes $\mathbb{E}_{\hat{\theta}_\omega,\hat{\mb}}\mathds{1}_{\{y_\omega=y\}}$.
+We assess the accuracy using two metrics: Mean absolute deviation (MAD), and Misclassification rate (MCR).
 
-\begin{table}[t]
-\label{CV}
-\vskip 0.15in
-\begin{center}
-\begin{small}
+We repeat stratified cross validation 10 times, average results and calculate standard deviations in Table \ref{table:CV}.
+\begin{table}[H]
+    \label{table:CV}
+\scriptsize
+\center
 \begin{sc}
-\begin{tabular}{lcccr}
-\toprule
- Method & MSE & MAD & MCR \\
-\midrule
-\begin{tabular}[c]{@{}c@{}}Oridnal tensor\\ Decomposition\end{tabular}  & 0.1503 & 0.1502& 0.1502\\ \\
-\begin{tabular}[c]{@{}c@{}}Conti tensor\\ Decomposition\end{tabular} & 0.1603& 0.1600& 0.1598\\ \\
-\begin{tabular}[c]{@{}c@{}}1bit-completion\\ (probit)\end{tabular}  & 0.3658 & 0.3632 & 0.3619\\
-\begin{tabular}[c]{@{}c@{}}1bit-completion\\ (logistic)\end{tabular}  &  0.4514 & 0.4215 & 0.4066
-\\
-\bottomrule
+\begin{tabular}{ll|c|c}
+\hline
+\multicolumn{2}{c|}{Method}                       & MAD & MCR\\
+\hline
+\multirow{2}{*}{Our method} & $\hat y^{(\text{Median})}_\omega$ & \begin{tabular}[c]{@{}c@{}} 0.1607 (0.00046)\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.1606 (0.00047)
+\end{tabular} \\
+&$\hat y^{(\text{Mode})}_\omega$         & \begin{tabular}[c]{@{}c@{}}0.1607  (0.00046)\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.1606 (0.00047)\end{tabular} \\ \cline{1-2}
+\multicolumn{2}{l|}{\begin{tabular}[c]{@{}l@{}}Cotinuous method\\~\cite{filipovic2015tucker}\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.2530  (0.00015)\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.1599 (0.00023)\end{tabular} \\ \cline{1-2}
+\multicolumn{2}{l|}{\begin{tabular}[c]{@{}l@{}}1bit Tensor Completion\\~\cite{ghadermarzy2018learning}\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.3566 (0.00102)\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.3563 (0.00102)\end{tabular} \\ \hline
 \end{tabular}
 \end{sc}
-\end{small}
-\end{center}
-\vskip -0.1in
-\caption{Results of comparisons among 4 methods on the HCP data predicting the test data.
-Four methods are the ordinal tensor decomposition algorithm, the continuous tensor decomposition algorithm and the 1 bit tensor completion method with probit link function and logistic link function.
-Each method is evaluated by RMSE, MAE, error(false prediction rate).}
+\caption{Comparison of tensor completion performance on HCP data: values in the parentheses are standard deviation from 10 repetitions.}
 \end{table}
@@
 To be specific, the music recommendations can be different according to the mood of the driver or the traffic condition in this system.
+Those kinds of external events that have influence on ratings on the musics are regarded as the contexts.
 Our goal is to perform the tensor completion to the $42\times 139\times 26$ ordinal tensor and thereby we can offer context-specific music recommendation to users.
@@
 The number of missing values is 148,904 and the number of available values is 2,884.
-We suggest three estimators $\hat y_\omega$ based on $(\hat \theta_\omega,\hat \mb)$:
-\begin{itemize}
-\item (Mean) $\hat y^{(\text{Mean})}_\omega=\sum_\ell \ell g_\ell(\hat \theta_\omega,\hat \mb) $;
-\item (Median) $\hat y^{(\text{Median})}_\omega=\min\{\ell\in[L]\colon f_\ell(\hat \theta_\omega,\hat \mb)\geq 0.5\}$;
-\item (Mode) $\hat y^{(\text{Mode})}_\omega=\argmax_\ell g_\ell(\hat \theta_\omega,\hat \mb) $
-\end{itemize}
-Note that, under the ordinal tensor model~\eqref{mainmodel}, the estimator $\hat y^{(\text{Mean})}_\omega$ minimizes $\mathbb{E}_{\hat{\theta}_\omega,\hat{\mb}}(y_\omega-y)^2$, the estimator $\hat y^{(\text{Median})}_\omega$ minimizes $\mathbb{E}_{\hat{\theta}_\omega,\hat{\mb}}|y_\omega-y|$
-and the estimator $\hat y^{(\text{Mode})}_\omega $ minimizes $\mathbb{E}_{\hat{\theta}_\omega,\hat{\mb}}\mathds{1}_{\{y_\omega=y\}}$.
-In contrast, these three estimators degenerate to a single estimator under the continuous-valued tensor decomposition model.
-We assess the accuracy using three metrics: Mean squared error (MSE), Mean absolute deviation (MAD), and Misclassification rate (MCR).
-
+We perform 5 fold cross validation to assess the performance our model on InCarMusic data. The estimation methods and the procedure are the same as in HCP data analysis.
 Results are averaged from 5 fold cross validation. {\color{red} add s.e. to the above table.}

### 2020-02-02 ICML/chanwoo/Supplements.tex
Add files via upload

@@
 
+\section{Brain data clustering}
+
+\begin{table}[]
+\begin{tabular}{|c|c|c|c|c|c|c|}
+\hline
+Group       & \multicolumn{3}{c|}{\#1}                                                  & \multicolumn{3}{c|}{\#2}                                      \\ \hline
+Brain nodes & \multicolumn{3}{l|}{\begin{tabular}[c]{@{}l@{}}"RMF\_L","Fpole\_L","Insula\_L",\\ "SupF\_L","CaudMF\_L",\\ "parsTriangularis\_L","parsOpercularis\_L",\\ "preCentral\_L","Tpole\_L","SupT\_L",\\ "SupT\_L","postCentral\_L", "SupP\_L",\\ "IP\_L","LO\_L","MOF\_L","SupF\_L",\\ "IsthmusC\_L","Precuneus\_L",\\ "Cuneus\_L","paraHippo\_L",\\ "Lingual\_L","SupT\_L","LO\_L"\end{tabular}} & \multicolumn{3}{l|}{\begin{tabular}[c]{@{}l@{}}"RMF\_R","Fpole\_R","Insula\_R",\\ "SupF\_R", "CaudMF\_R",\\ "parsTriangularis\_R" ,"parsOpercularis\_R",\\ "preCentral\_R", "Tpole\_R","SupT\_R",\\ "SupT\_R","postCentral\_R","SupP\_R",\\ "IP\_R","LO\_R","MOF\_R", "SupF\_R",\\ "IsthmusC\_R","Precuneus\_R",\\ "Cuneus\_R","paraHippo\_R",\\ "Lingual\_R","SupT\_R","LO\_R"\end{tabular}} \\ \hline
+Group       & \#3    & \#4    & \#5   & \#6  & \#7 & \#8
+\\ \hline
+Brain nodes & \begin{tabular}[c]{@{}c@{}}"SupraM\_L" \\ "SupraM\_L" \\ "SupraM\_L" \\ "SupraM\_L"\end{tabular}                                      & \begin{tabular}[c]{@{}c@{}}"SupraM\_R" \\ "SupraM\_R" \\ "SupraM\_R" \\ "SupraM\_R"\end{tabular}                                      & \begin{tabular}[c]{@{}c@{}}"IT\_L" \\ "IT\_L" \\ "IT\_L"\end{tabular}                                      & \begin{tabular}[c]{@{}c@{}}"IT\_R" \\ "IT\_R" \\ "IT\_R"\end{tabular}                                                          & \begin{tabular}[c]{@{}c@{}}"MT\_L" \\ "MT\_L" \\ "MT\_L"\end{tabular}                                                         & \begin{tabular}[c]{@{}c@{}}"MT\_R" \\ "MT\_R"\\ "MT\_R"\end{tabular}                                                         \\ \hline
+\end{tabular}
+\end{table}
+\begin{figure}[H]
+  \centering
+ \includegraphics[width = 0.45\textwidth]{elbowmethod.png}
+
+  \label{figure:cluster graph}
+  \caption{Brain cluster graph}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+ \includegraphics[width = 0.45\textwidth]{brain1.jpg}
+ \includegraphics[width = 0.45\textwidth]{brain2.jpg}
+ \includegraphics[width = 0.45\textwidth]{brain3.jpg}
+ \includegraphics[width = 0.45\textwidth]{brain4.jpg}
+  \label{figure:clustering}
+  \caption{Brain cluster groups}
+\end{figure}
+
 \end{proof}

### 2020-02-04 ICML/miaoyan/ordinalT_supp_v2.tex
Add files via upload

--- 2020-01-21 ICLM/miaoyan/ordinalT_supp.tex
@@
 
+\usepackage{wrapfig}
+
 \usepackage{mathtools}
@@
 Define $d_{\text{total}}=\prod_k d_k$.
-Let $\fplogl\in\mathbb{R}^{d_\text{total}}$ denote the vector of gradient with respect to $\Vec(\Theta)\in\mathbb{R}^{d_{\text{total}}}$, and $\fpplogl$ the corresponding Hession matrix of size $d_\text{total}$-by-$d_{\text{total}}$.
+Let $\fplogl\in\mathbb{R}^{d_1\times\cdots\times d_K}$ denote the tensor of gradient with respect to $\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}$, and $\fpplogl$ the corresponding Hession matrix of size $d_\text{total}$-by-$d_{\text{total}}$.
 Here, $\Vec(\cdot)$ denotes the operation that turns a tensor into a vector.
@@
 \[
-U_\alpha=\max_{\ell\in[L],|\alpha|\leq \alpha}{\dot{g}_\ell(\theta)\over g_\ell(\theta)}>0 \quad \text{and}\quad
-L_\alpha=\max_{\ell\in[L],|\alpha|\leq \alpha} {\dot{g}^2_\ell(\theta)-\ddot{g}_\ell(\theta)g_\ell(\theta)\over g^2_\ell(\theta)}>0. 
+U_\alpha=\max_{\ell\in[L],|\theta|\leq \alpha}{\dot{g}_\ell(\theta)\over g_\ell(\theta)}>0 \quad \text{and}\quad
+L_\alpha=\min_{\ell\in[L],|\theta|\leq \alpha} {\dot{g}^2_\ell(\theta)-\ddot{g}_\ell(\theta)g_\ell(\theta)\over g^2_\ell(\theta)}>0.
 \]
@@
 \begin{equation}\label{eq:taylor}
-\flogl(\Theta)=\flogl(\trueT)+\langle\fplogl,\ \Vec(\Theta-\trueT)\rangle+{1\over 2}\Vec(\Theta-\trueT)^T\fpplogl(\check\Theta)\Vec(\Theta-\trueT),
+\flogl(\Theta)=\flogl(\trueT)+\langle\Vec(\fplogl),\ \Vec(\Theta-\trueT)\rangle+{1\over 2}\Vec(\Theta-\trueT)^T\fpplogl(\check\Theta)\Vec(\Theta-\trueT),
 \end{equation}
@@
 \begin{equation}\label{eq:linear}
-|\fplogl(\trueT), \Vec(\Theta-\trueT)  \rangle|\leq \snormSize{}{\fplogl(\trueT)} \nnormSize{}{\Theta-\trueT},
+|\langle\Vec(\fplogl(\trueT), \Vec(\Theta-\trueT)  \rangle|\leq \snormSize{}{\fplogl(\trueT)} \nnormSize{}{\Theta-\trueT},
 \end{equation}
@@
 \begin{equation}\label{eq:linearconclusion}
-|\langle \fplogl(\trueT), \Vec(\Theta-\trueT)  \rangle | \leq C_2 U_\alpha  \sqrt{r_{\max}^{K-1} \sum_k d_k}  \FnormSize{}{\Theta-\trueT}.
+|\langle \Vec(\fplogl(\trueT)), \Vec(\Theta-\trueT)  \rangle | \leq C_2 U_\alpha  \sqrt{r_{\max}^{K-1} \sum_k d_k}  \FnormSize{}{\Theta-\trueT}.
 \end{equation}
@@
 \begin{equation}\label{eq:totalKL}
-{1\over \text{Card}(\tX)-1}\sum_{\Theta \in\tX}\text{KL}(\mathbb{P}_{\Theta}, \mathbb{P}_{\mathbf{0}})\leq \varepsilon \log\left\{\text{Card}(\tX)-1 \right\}
+{1\over \text{Card}(\tX)-1}\sum_{\Theta \in\tX}\text{KL}(\mathbb{P}_{\Theta}, \mathbb{P}_{\mathbf{0}})\leq \varepsilon \log_2\left\{\text{Card}(\tX)-1 \right\}
 \end{equation}
 holds for any $ \varepsilon \geq 0$ when $\gamma\in[0,1]$ is chosen to be sufficiently small depending on $\varepsilon$, e.g., $\gamma \leq \sqrt{3\varepsilon}$.
-By applying Theorem~\ref{thm:Tsybakov} to~\eqref{eq:totalKL}, and in view of the property (iv), we obtain that
+By applying Lemma~\ref{lem:Tsybakov} to~\eqref{eq:totalKL}, and in view of the property (iv), we obtain that
 \begin{equation}\label{eq:final}
@@
 \begin{equation}\label{eq:Taylor2}
-\logl(\Theta)=\logl(\trueT)+\langle\plogl,\ \Vec(\Theta-\trueT)\rangle+{1\over 2}\Vec(\Theta-\trueT)^T\pplogl(\check\Theta)\Vec(\Theta-\trueT),
+\logl(\Theta)=\logl(\trueT)+\langle\Vec(\plogl),\ \Vec(\Theta-\trueT)\rangle+{1\over 2}\Vec(\Theta-\trueT)^T\pplogl(\check\Theta)\Vec(\Theta-\trueT),
 \end{equation}
@@
 \begin{enumerate}
-\item $\plogl$ is a length-$d_{\text{total}}$ vector with $|\Omega|$ nonzero entries, and each entry is upper bounded by $L_\alpha>0$. 
-\item $\pplogl$ is a diagonal matrix of size $d_{\text{total}}$-by-$d_{\text{total}}$ with $|\Omega|$ nonzero entries, and each entry is upper bounded by $-U_{\alpha}<0$. 
+\item $\plogl$ is a $d_1\times\cdots\times d_K$ tensor with $|\Omega|$ nonzero entries, and each entry is upper bounded by $U_\alpha>0$.
+\item $\pplogl$ is a diagonal matrix of size $d_{\text{total}}$-by-$d_{\text{total}}$ with $|\Omega|$ nonzero entries, and each entry is upper bounded by $-L_{\alpha}<0$.
 \end{enumerate}
@@
 \begin{equation}\label{eq:linear2}
-|\langle\plogl,\ \Theta-\trueT\rangle|\leq C_2U_\alpha \sqrt{r_{\max}^{K-1}\sum_k d_k}\MFnormSize{}{\Theta-\trueT}
+|\langle\Vec(\plogl),\ \Vec(\Theta-\trueT)\rangle|\leq C_2U_\alpha \sqrt{r_{\max}^{K-1}\sum_k d_k}\MFnormSize{}{\Theta-\trueT}
 \end{equation}
@@
 
-Combining~\eqref{eq:Taylor2}-\eqref{eq:quadratic2} with the fact that $\plogl(\hat \Theta)\geq \plogl(\trueT)$, we have
+Combining~\eqref{eq:Taylor2}-\eqref{eq:quadratic2} with the fact that $\logl(\hat \Theta)\geq \logl(\trueT)$, we have
 \begin{equation}\label{eq:sample}
@@
 Note that $\mnormSize{}{\hat\Theta-\trueT}\leq 2\alpha$ and $\text{rank}(\hat \Theta-\trueT)\leq 2\mr$.
-By Lemma~\ref{lem:Mnormbound}, $\anormSize{}{\hat \Theta-\trueT}\leq 2^{3(K-1)/2} \left({\prod r_k \over r_{\max}}\right)^{3/2}$.
-Therefore, the condition in Lemma~\ref{lem:convexity} holds with $\beta=2^{3(K-1)/2} \left({\prod r_k \over r_{\max}}\right)^{3/2}$.
+By Lemma~\ref{lem:Mnormbound}, $\anormSize{}{\hat \Theta-\trueT}\leq 2^{(3K-1)/2} \alpha \left({\prod r_k \over r_{\max}}\right)^{3/2}$.
+Therefore, the condition in Lemma~\ref{lem:convexity} holds with $\beta=2^{(3K-1)/2}\alpha \left({\prod r_k \over r_{\max}}\right)^{3/2}$.
 Applying Lemma~\ref{lem:convexity} to~\eqref{eq:sample} gives
@@
  \PiFnormSize{}{\hat \Theta-\trueT}^2&\leq {1\over m}\MFnormSize{}{\hat \Theta-\trueT}^2+c\beta\sqrt{\sum_k d_k\over |\Omega|}\\
- &\leq {C_2  r^{K-1}_{\max}} {\sum_k d_k \over |\Omega|}+C_1 r_{\max}^{3(K-1)/2}\sqrt{\sum_kd_k\over |\Omega|},
+ &\leq {C_2  r^{K-1}_{\max}} {\sum_k d_k \over |\Omega|}+C_1 \alpha r_{\max}^{3(K-1)/2}\sqrt{\sum_kd_k\over |\Omega|},
 \end{align}
@@
 
-We begin with a set of technical lemmas that are useful for the proofs of the main theorems.
+We begin with various notion of tensor norms that are useful for the proofs of the main theorems.
 
-\begin{lem}[\cite{tomioka2014spectral}]\label{lem:tensor}
-Suppose that $\tS=\entry{s_{\omega}}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is an order-$K$ tensor whose entries are independent random variables that satisfy
+\begin{defn}[Atomic M-norm~\citep{ghadermarzy2019near}]
+Define $T_{\pm}=\{\tT\in\{\pm 1 \}^{d_1\times \cdots \times d_K}\colon \text{rank}(\tT)=1\}$. The atomic M-norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
+\begin{align}
+\anormSize{}{\Theta}&=\inf\{t>0: \Theta\in t\text{conv}(T_{\pm})\}\\
+&=\inf\left\{\sum_{\tX\in T_{\pm}}c_x\colon \Theta=\sum_{\tX\in T_{\pm}} c_x\tX, \ c_x>0\right\}.
+\end{align}
+\end{defn}
+
+\begin{defn}[Spectral norm~\citep{lim2005singular}]
+The spectral norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
 \[
-\mathbb{E}(s_{\omega})=0,\quad \text{and} \quad\mathbb{E}(e^{ts_{\omega}})\leq e^{t^2L^2/2}.
+\snormSize{}{\Theta}=\sup\left\{\langle \Theta, \mx_1\otimes \cdots \otimes \mx_K\rangle \colon \vnormSize{}{\mx_k}=1,\ \mx_k\in\mathbb{R}^{d_k},\ \text{for all}\ k\in[K]\right\}.
 \]
-Then the spectral norm $\snormSize{}{\tS}$ satisfies that, 
+\end{defn}
+
+\begin{defn}[Nuclear norm~\citep{friedland2018nuclear}]
+The nuclear norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
 \[
-\snormSize{}{\tS}\leq \sqrt{{8L^2} \log (12K) \sum_k d_k +\log (2/\delta)},
+\nnormSize{}{\Theta}=\inf
+\left\{
+\sum_{i\in[r]}|\lambda_i|\colon \Theta=\sum_{i=1}^r \lambda_i\mx^{(i)}_1\otimes \cdots \otimes \mx^{(i)}_K,\ \vnormSize{}{\mx^{(i)}_k}=1,\ \mx^{(i)}_k\in\mathbb{R}^{d_k},\ \text{for all} k\in[K],\ i\in[r]
+\right\},
 \]
-with probability at least $1-\delta$.
+where the infimum is taken over all $r\in\mathbb{N}$ and $\norm{\mx^{(i)}_k}=1$ for all $i\in[r]$ and $k\in[K]$.
+\end{defn}
+
+\begin{lem}[M-norm and infinity norm~\citep{ghadermarzy2019near}]\label{lem:Mnormbound}
+Let $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ be an order-$K$, rank-$(r_1,\ldots,r_K)$ tensor. Then
+\[
+\mnormSize{}{\Theta}\leq \anormSize{}{\Theta}\leq \left(\prod_k r_k \over r_{\max}\right)^{3\over 2} \mnormSize{}{\Theta}.
+\]
 \end{lem}
 
-\begin{rmk}
-The above lemma provides the bound on the spectral norm of random tensors. The result was firstly presented in~\cite{nguyen2015tensor}, and we adopt the version from~\cite{tomioka2014spectral}.  
-\end{rmk}
-
-\begin{lem} \label{lem:noisytensor}
-Suppose that $\tS=\entry{s_{\omega}}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is an order-$K$ tensor whose entries are independent random variables that satisfy
-\[
-\mathbb{E}(s_{\omega})=0,\quad \text{and}\quad |s_{\omega}|\leq U.
-\]
-Then we have
-\[
-\mathbb{P}\left(\snormSize{}{\tS}\geq C_2 U\sqrt{\sum_k d_k} \right)\leq \exp\left(-C_1  \log K \sum_k d_k\right)
-\]
-where $C_1>0$ is an absolute constant, and $C_2>0$ is a constant that depends only on $K$. 
-\end{lem}
-
-\begin{proof}  Note that the random variable $U^{-1}s_{\omega}$ is zero-mean and supported on $[-1,1]$. Therefore, $U^{-1}s_{\omega}$ is sub-Gaussian with parameter ${1-(-1)\over 2}=1$, i.e.
-\[
-\mathbb{E}(U^{-1}s_{\omega})=0,\quad \text{and}\quad \mathbb{E}(e^{tU^{-1}s_{\omega}})\leq e^{t^2/2}.
-\]
-It follows from Lemma~\ref{lem:tensor} that, with probability at least $1-\delta$, 
-\[
-\snormSize{}{U^{-1}\tS}\leq \sqrt{\left(c_0\log K+c_1\right) \sum_k d_k +\log (2/\delta)},
-\]
-where $c_0, c_1>0$ are two absolute constants. Taking $\delta=\exp (-C_1\log K \sum_k d_k)$ yields the final claim, where $C_2=c_0\log K+c_1+1>0$ is another constant.  
-\end{proof}
-
-\begin{lem} \label{lem:nuclear}
+\begin{lem}[Nuclear norm and F-norm] \label{lem:nuclear}
 Let $\tA\in\mathbb{R}^{d_1\times\cdots\times d_K}$ be an order-$K$ tensor with Tucker $\text{rank}(\tA)=(r_1,\ldots,r_K)$. Then
 \[
-\nnormSize{}{\tA} \leq \sqrt{\sum_k r_k\over \max_k r_k} \FnormSize{}{\tA},
+\nnormSize{}{\tA} \leq \sqrt{\prod_k r_k\over \max_k r_k} \FnormSize{}{\tA},
 \]
@@
 
+The following lemma provides the bound on the spectral norm of random tensors. The result was firstly presented in~\cite{nguyen2015tensor}, and we adopt the version from~\cite{tomioka2014spectral}.
+\begin{lem}[\cite{tomioka2014spectral}]\label{lem:tensor}
+Suppose that $\tS=\entry{s_{\omega}}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is an order-$K$ tensor whose entries are independent random variables that satisfy
+\[
+\mathbb{E}(s_{\omega})=0,\quad \text{and} \quad\mathbb{E}(e^{ts_{\omega}})\leq e^{t^2L^2/2}.
+\]
+Then the spectral norm $\snormSize{}{\tS}$ satisfies that,
+\[
+\snormSize{}{\tS}\leq \sqrt{{8L^2} \log (12K) \sum_k d_k +\log (2/\delta)},
+\]
+with probability at least $1-\delta$.
+\end{lem}
+
+\begin{lem} \label{lem:noisytensor}
+Suppose that $\tS=\entry{s_{\omega}}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is an order-$K$ tensor whose entries are independent random variables that satisfy
+\[
+\mathbb{E}(s_{\omega})=0,\quad \text{and}\quad |s_{\omega}|\leq U.
+\]
+Then we have
+\[
+\mathbb{P}\left(\snormSize{}{\tS}\geq C_2 U\sqrt{\sum_k d_k} \right)\leq \exp\left(-C_1  \log K \sum_k d_k\right)
+\]
+where $C_1>0$ is an absolute constant, and $C_2>0$ is a constant that depends only on $K$.
+\end{lem}
+
+\begin{proof}  Note that the random variable $U^{-1}s_{\omega}$ is zero-mean and supported on $[-1,1]$. Therefore, $U^{-1}s_{\omega}$ is sub-Gaussian with parameter ${1-(-1)\over 2}=1$, i.e.
+\[
+\mathbb{E}(U^{-1}s_{\omega})=0,\quad \text{and}\quad \mathbb{E}(e^{tU^{-1}s_{\omega}})\leq e^{t^2/2}.
+\]
+It follows from Lemma~\ref{lem:tensor} that, with probability at least $1-\delta$,
+\[
+\snormSize{}{U^{-1}\tS}\leq \sqrt{\left(c_0\log K+c_1\right) \sum_k d_k +\log (2/\delta)},
+\]
+where $c_0, c_1>0$ are two absolute constants. Taking $\delta=\exp (-C_1\log K \sum_k d_k)$ yields the final claim, where $C_2=c_0\log K+c_1+1>0$ is another constant.
+\end{proof}
+
 \begin{lem}\label{lem:KLentry} Let $X,\; Y$ be two discrete random variables taking values on $L$ possible categories, with category probabilities $\{p_\ell\}_{\ell\in[L]}$ and $\{q_\ell\}_{\ell\in[L]}$, respectively.
@@
 
-\begin{lem}~\label{lem:KL}
+\begin{lem}[KL divergence and F-norm]~\label{lem:KL}
 Let $\tY\in[L]^{d_1\times \cdots \times d_K}$ be an ordinal tensor generated from the model~\eqref{eq:model} with the link function $f$ and parameter tensor $\Theta$.
@@
 \end{align}
-where $\eta_\ell, \eta_{\ell-1}$ fall between $\theta$ and $\theta^*$, and $\check b_{\ell}$ falls between $b_{\ell-1}$ and $b_\ell$. Therefore,
+where $\eta_\ell$ and $\eta_{\ell-1}$ fall between $\theta$ and $\theta^*$. Therefore,
 \begin{equation}\label{eq:entrywise}
@@
 \[
-\FnormSize{}{\Theta_i-\Theta_j}^2 \geq {Rd_1\over 8} \gamma^2\min\left\{ \alpha, {C^{-1}Rd_1 \over d_{\text{total}}}\right\} \lfloor {d_2\over R} \rfloor \prod_{k\geq 3}d_k\geq {\gamma^2\min\left\{ \alpha^2 d_{\text{total}}, C^{-1}Rd_1\right\}  \over 16}.
+\FnormSize{}{\Theta_i-\Theta_j}^2 \geq {Rd_1\over 8} \gamma^2\min\left\{ \alpha^2, {C^{-1}Rd_1 \over d_{\text{total}}}\right\} \lfloor {d_2\over R} \rfloor \prod_{k\geq 3}d_k\geq {\gamma^2\min\left\{ \alpha^2 d_{\text{total}}, C^{-1}Rd_1\right\}  \over 16}.
 \]
@@
 
-\begin{thm}[\cite{tsybakov2009introduction}]\label{thm:Tsybakov} 
+\begin{lem}[Theorem 2.5 in~\cite{tsybakov2009introduction}]\label{lem:Tsybakov}
 Assume that a set $\tX$ contains element $\Theta_0, \Theta_1, \ldots,\Theta_M$ ($M\geq 2$) such that
 \begin{itemize}
-\item $d(\Theta_j,\ \Theta_j)\geq 2s>0$, $\forall 0\leq j\leq k\leq M$;
+\item $d(\Theta_j,\ \Theta_k)\geq 2s>0$, $\forall 0\leq j\leq k\leq M$;
 \item $\mathbb{P}_j\ll\mathbb{P}_0$, $\forall j=1,\ldots,M$, and
@@
 \]
-where $d\colon \tX\times \tX\mapsto [0,+\infty]$ is a semi-distance function, $0<\alpha<{1/8}$ and $P_j=P_{\Theta_j}$, $j=0,1\ldots,M$. Then
+where $d\colon \tX\times \tX\mapsto [0,+\infty]$ is a semi-distance function, $0<\alpha<{1/8}$ and $P_j=P_{\Theta_j}$, $j=0,1\ldots,M$.
+\end{itemize}
+Then
 \[
@@
 \]
-\end{itemize}
-
-\end{thm}
-\begin{defn}[\cite{ghadermarzy2019near}] 
-Define $T_{\pm}=\{\tT\in\{\pm\}^{d_1\times \cdots \times d_K}\colon \text{rank}(\tT)=1\}$. The atomic M-norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
-\begin{align}
-\anormSize{}{\Theta}&=\inf\{t>0: \Theta\in t\text{conv}(T_{\pm})\}\\
-&=\inf\left\{\sum_{\tX\in T_{\pm}}c_x\colon \Theta=\sum_{\tX\in T_{\pm}} c_x\tX, \ c_x>0\right\}.
-\end{align}
-\end{defn}
-
-\begin{lem}[\cite{ghadermarzy2019near}]\label{lem:Mnormbound}
-Let $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ be an order-$K$, rank-$(r_1,\ldots,r_K)$ tensor. Then
-\[
-\mnormSize{}{\Theta}\leq \anormSize{}{\Theta}\leq \left(\prod_k r_k \over r_{\max}\right)^{3\over 2} \mnormSize{}{\Theta}.
-\]
 \end{lem}
 
-\begin{lem}[\cite{ghadermarzy2019near}]\label{lem:convexity}
+\begin{lem}[Lemma 28 in~\cite{ghadermarzy2019near}]\label{lem:convexity}
 Define $\mathbb{B}_{M}(\beta)=\{\Theta\in \mathbb{R}^{d_1\times \cdots \times d_K}\colon \anormSize{}{\Theta}\leq \beta \}$.
@@
 
+\section{Convexity of the likelihood function}
+\begin{thm}\label{thm:convexity}
+\begin{align}
+ \logl(\Theta, \mb)&=\sum_{\omega\in\Omega}\sum_{\ell\in[L]} \Big\{\mathds{1}_{\{y_\omega=\ell\}} \log \big[f(\theta_\omega+b_\ell)-  f(\theta_\omega+b_{\ell-1})\big]\Big\}, \text{ where } f(x) = \frac{e^x}{1+e^x}.
+ \end{align}
+is a concave function to $(\Theta,\mb)$
+\end{thm}
+\begin{proof}
+We use $(\Theta,\mb)$ abusively as $(Vec(\Theta)^T,\mb^T)^T$ in this proof.
+Let us denote $\lambda(u,v) = \log \big[f(u)-  f(v)\big]$.
+It is enough to show that $\lambda(u,v)$
+is a concave function to $(u,v)$ where $u>v$. Because if $\lambda(u,v)$ is a concave function and $u,v$ are both linear functions of $(\Theta,\mb) $ such that
+$$u = \ma_1^T(\Theta,\mb), v = \ma_2^T(\Theta,\mb) \text{ for some }\ma_1,\ma_2\in \mathbb{R}^{d_{\text{total}}+L-1},$$ then $\lambda(u,v) = \lambda(\ma_1^T(\Theta,\mb),\ma_2^T(\Theta,\mb)))$ is a concave function to $(\Theta,\mb)$
+by the definition of the convexity. Furthermore, we can conclude that $\logl(\Theta, \mb) $ is a concave function to $(\Theta,\mb)$ because $\logl(\Theta, \mb)$
+can be written as the form of summations of $\lambda(u,v)$. To prove the convexity of $\lambda(u,v)$, we have the following.
+\begin{equation*}
+\lambda(u,v) = \log\big[f(u)-f(v)\big]=\log\big[\int\mathds{1}_{(u,v)}(x)f'(x)dx\big].
+\end{equation*}
+Notice that  $\mathds{1}_{(u,v)}(x)$ is log-concave to $(u,v,x)$ and $f'(x)$ is log-concave to $x$ because $f'(x) = \frac{e^x}{(1+e^x)^2}$. Those convexities imply that $\mathds{1}_{(u,v)}(x)f'(x)$
+is a log-concave function to $(u,v,x)$. Therefore, we can conclude that the $\lambda(u,v)$ is a concave function from the lemma~\ref{lem:convexity}.
+\end{proof}
+
+\begin{lem}[Corollary 3.5 in~\cite{brascamp2002extensions}]\label{lem:convexity}
+Let $F(x,y):\mathbb{R}^{m+n}\rightarrow \mathbb{R}$ be an integrable function where $x\in \mathbb{R}^{m},y\in \mathbb{R}^n$.
+Let $$G(x) = \int_{\mathbb{R}^n}F(x,y)dy.$$ If $F(x,y)$ is a log concave function to (x,y), then $G(x)$ is a log concave function.
+\end{lem}
+
+\section{Brain clustering method and result}
+\subsection{Clustering method}
+We perform clustering based on the estimated latent parameter $\hat\Theta$.
+Our clustering method is based on multidimensional version of Principal Component Analysis (PCA) and Singular Value Decomposition (SVD).
+In matrices case, we perform clustering on a $m\times n$ data matrix $X$ based on the following procedure.
+First, we factorize $X$ into
+\begin{equation*}
+X = U\Sigma V^T,
+\end{equation*}
+where $\Sigma$ is a diagonal matrix and $ U,V$ are factor matrices with orthogonal columns.
+We interpret each column of $V$ as a principal direction or axis and $U\Sigma$ as principal components.
+From this interpretation, we perform $K$-means method to $m$ rows of $U\Sigma$.
+We suggest to apply this procedure to the estimated latent parameter tensor $\hat\Theta$.
+For the notational convenience, we use $\Theta$ as $\hat\Theta$ in this section.
+From the Tucker decomposition which is multidimensional version of SVD, we have
+\begin{equation*}
+\Theta = \tC\times_1\mM_1\times_2\cdots\times_K\mM_K,
+\end{equation*}
+where $\tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is a core tensor, $\mM_k\in\mathbb{R}^{d_k\times r_k}$ are factor matrices with orthogonal columns, and $\times_k$ denotes the tensor-by-matrix multiplication~\cite{kolda2009tensor}.
+We perform mode-$k$ matricization of the tensor $\Theta$ to cluster data from mode-$k$ variable as follows.
+\begin{equation*}
+\Theta_{(k)} = \mM_k\tC_{(k)}\big(\mM_K\otimes\cdots\otimes\mM_1\big)
+\end{equation*}
+We interpret $\big(\mM_K\otimes\cdots\otimes\mM_1\big)$ as principal axes and $\mM_k\tC_{(k)}$ as principal components.
+By putting other mode variables as principal axes, we can exclude effects from those variables and can see $k$-th mode effect from checking principal components, $\mM_k\tC_{(k)}$.
+Therefore, we perform $K$-means method to $d_k$ rows of the principal components like in the matrices case.
+
+\subsection{Clustering result on HCP}
+We perform $K$-means method to cluster brain nodes based on the suggested above method.
+We have the estimated latent parameter $\hat\Theta$ with the rank $(23,23,8)$ which minimizes the BIC value.
+The number of clusters is chosen to be eight from the elbow method in figure ~\ref{figure:elbow}.
+The table in \ref{table:clustering} shows eight cluster groups and their entries.
+Most of brain nodes fall into the cluster $\#1$ and the cluster $\#2$ which can be represented as the left side brain and the right side brain.
+From the group $\#3$ to the group $\#8$ , entries of each cluster share the same name encoded in the data and play a similar role in the brain.
+For example, the group $\#3$ and the group $\#4$ are the left and the right side of the regions called Supramarginal gyrus.
+Those regions are known to be in charge of interpreting tactile sensory data and involved in perception of space and limbs location \cite{carlson2012physiology,reed1994nature}.
+The figure \ref{figure:brain image} shows the nodes in the dataset on the brain image.
+Each node is connected by edges within the same cluster and the groups can be paired as the right part and the left part of the brain.
+The results show that the clustering based on our model successfully group the brain nodes without knowledge of the brain region.
+\begin{figure}[H]
+  \label{figure:elbow}
+  \begin{center}
+    \includegraphics[width=0.48\textwidth]{elbowmethod.png}
+  \end{center}
+  \caption{The elbow plot for the number of clusters}
+\end{figure}
+
+\begin{table}[H]
+  \label{table:clustering}
+  \center
+  \begin{small}
+    \footnotesize
+  \begin{sc}
+\begin{tabular}{|c|c|c|c|c|c|c|}
+\hline
+Group       & \multicolumn{3}{c|}{\#1}                                                  & \multicolumn{3}{c|}{\#2}                                      \\ \hline
+Brain nodes & \multicolumn{3}{l|}{\begin{tabular}[c]{@{}l@{}}"RMF\_L","Fpole\_L","Insula\_L",\\ "SupF\_L","CaudMF\_L",\\ "parsTriangularis\_L","parsOpercularis\_L",\\ "preCentral\_L","Tpole\_L","SupT\_L",\\ "SupT\_L","postCentral\_L", "SupP\_L",\\ "IP\_L","LO\_L","MOF\_L","SupF\_L",\\ "IsthmusC\_L","Precuneus\_L",\\ "Cuneus\_L","paraHippo\_L",\\ "Lingual\_L","SupT\_L","LO\_L"\end{tabular}} & \multicolumn{3}{l|}{\begin{tabular}[c]{@{}l@{}}"RMF\_R","Fpole\_R","Insula\_R",\\ "SupF\_R", "CaudMF\_R",\\ "parsTriangularis\_R" ,"parsOpercularis\_R",\\ "preCentral\_R", "Tpole\_R","SupT\_R",\\ "SupT\_R","postCentral\_R","SupP\_R",\\ "IP\_R","LO\_R","MOF\_R", "SupF\_R",\\ "IsthmusC\_R","Precuneus\_R",\\ "Cuneus\_R","paraHippo\_R",\\ "Lingual\_R","SupT\_R","LO\_R"\end{tabular}} \\ \hline
+Group       & \#3    & \#4    & \#5   & \#6  & \#7 & \#8
+\\ \hline
+Brain nodes & \begin{tabular}[c]{@{}c@{}}"SupraM\_L" \\ "SupraM\_L" \\ "SupraM\_L" \\ "SupraM\_L"\end{tabular}                                      & \begin{tabular}[c]{@{}c@{}}"SupraM\_R" \\ "SupraM\_R" \\ "SupraM\_R" \\ "SupraM\_R"\end{tabular}                                      & \begin{tabular}[c]{@{}c@{}}"IT\_L" \\ "IT\_L" \\ "IT\_L"\end{tabular}                                      & \begin{tabular}[c]{@{}c@{}}"IT\_R" \\ "IT\_R" \\ "IT\_R"\end{tabular}                                                          & \begin{tabular}[c]{@{}c@{}}"MT\_L" \\ "MT\_L" \\ "MT\_L"\end{tabular}                                                         & \begin{tabular}[c]{@{}c@{}}"MT\_R" \\ "MT\_R"\\ "MT\_R"\end{tabular}                                                         \\ \hline
+\end{tabular}
+\end{sc}
+\end{small}
+\caption{Result of the brain nodes clustering: the last alphabet in each node name means the left side for `L' and the right side for `R' in the brain}
+\end{table}
+
+\begin{figure}[H]
+  \centering
+ \includegraphics[width = 0.45\textwidth]{brain1.jpg}
+ \includegraphics[width = 0.45\textwidth]{brain2.jpg}
+ \includegraphics[width = 0.45\textwidth]{brain3.jpg}
+ \includegraphics[width = 0.45\textwidth]{brain4.jpg}
+  \label{figure:brain image}
+  \caption{Brain image of eight clusters: entries of each cluster are connected by edges}
+\end{figure}
+
 \bibliographystyle{plainnat}

### 2020-02-04 code/InCarMusic.R
upload figure/draft

@@ -5,76 +5,95 @@ tensor[tensor==-1]=NA
 ind=which(is.na(tensor)==F)
 d=dim(tensor)
 set.seed(1)
+library("R.matlab")
+Err1=Err2=Err3=Err4=Err5=Err6=matrix(0,nrow=5,ncol=10)
 
 ### 5 fold ################
-l = length(ind)%/%5+1
-i1 = ind[1:l]
-i2 = ind[(l+1):(2*l)]
-i3 = ind[(2*l+1):(3*l)]
-i4 = ind[(3*l+1):(4*l)]
-i5 = ind[(4*l+1):length(ind)]
+for(nsim in 1:10){
+i1=i2=i3=i4=i5=NULL
+
+for(i in 1:5){ ### 5-fold sampling while preserving the lable proportion
+    label=sample(which(tensor==i),length(which(tensor==i))) 
+    l = length(label)%/%5+1
+    i1 = c(i1,label[1:l])
+    i2 = c(i2,label[(l+1):(2*l)])
+    i3 = c(i3,label[(2*l+1):(3*l)])
+    i4 = c(i4,label[(3*l+1):(4*l)])
+    i5 = c(i5,label[(4*l+1):length(label)])
+}
 cindex = list(i1,i2,i3,i4,i5)
+
 
 # `index' is for 1:5  
 # I used following lines when I put codes in the server in sh file 
 # args index ==$SLURM_ARRAY_TASK_ID" filename.R
+for(index in 1:5){
 test_index = cindex[[index]]
 train_index = setdiff(ind,test_index)
 train_tensor = tensor
 train_tensor[test_index] = NA
 
+save(test_index,train_index,file=sprintf("InCarMusic/input_%d_sim%d.RData",index,nsim))
 
+### Method 1. 1-bit tensor completion
+input=M_to_one(train_tensor,5,type="sign")
+writeMat(sprintf("InCarMusic/input_CV%d_sim%d.mat",index,nsim),data=input$binary,E=input$E,ave=input$ave,scale=input$scale)
+}
+}
 
 #####################################################################################
-################# Continuous Tucker decomposition CV ################################
-r = rep(1,3)
+################# Methods 2 and 3: Tucker decomposition CV ################################
+for(nsim in 1:10){
+    for(index in 1:5){
+   load(sprintf("InCarMusic/input_%d_sim%d.RData",index,nsim))     
+   train_tensor = tensor
+   train_tensor[test_index] = NA
+   
[…]

### 2020-02-04 code/functions.R

@@ -32,7 +32,7 @@ estimation = function(theta,omega,type){
   }else if(type=="median"){# score prediction based on the median 
   for (j in 1:length(theta_output)) theta_output[j] <-  which(c(cumsum(p[j,]),1)>=0.5)[1] ## median
   }
-  if(is.matrix(theta)) return(as.matrix(theta_output,dim=dim(theta)))
+  if(is.matrix(theta)) return(matrix(theta_output,nrow=dim(theta)[1],ncol=dim(theta)[2]))
   else return(as.tensor(array(theta_output,dim =theta@modes)))
 }   
 
@@ -186,7 +186,7 @@ fit_ordinal = function(ttnsr,C,A_1,A_2,A_3,omega=TRUE,alpha = TRUE){
   cost=NULL
   omg = omega
   k=length(unique(as.factor(c(ttnsr))))-is.element(NA,ttnsr) ## for NA not being included in length
-    while ((error > 10^-4)&(iter<50) ) {
+    while ((error > 10^-4)&(iter<10) ) {
       iter = iter +1
       #update omega
       prevtheta <- ttl(C,list(A_1,A_2,A_3),ms=1:3)@data
@@ -205,7 +205,7 @@ fit_ordinal = function(ttnsr,C,A_1,A_2,A_3,omega=TRUE,alpha = TRUE){
 
       prev <- likelihood(ttnsr,prevtheta,omega)
       
-    
+      
       #update A_1
       W <-kronecker(A_3,A_2)%*%t(k_unfold(C,1)@data)
       A_1<- comb(A_1,W,ttnsr,1,omega,alpha_minus2)
@@ -229,22 +229,24 @@ fit_ordinal = function(ttnsr,C,A_1,A_2,A_3,omega=TRUE,alpha = TRUE){
       qr_res=qr(A_3)
       A_3=qr.Q(qr_res)
       C=ttm(C,qr.R(qr_res),3)
-      
-      # update C
-      C_update <- corecomb(A_1,A_2,A_3,C,ttnsr,omega)
-      
-      if(!is.logical(alpha) & max(abs(ttl(C_update,list(A_1,A_2,A_3),ms=1:3)@data))>=alpha_minus2){
-          theta <- ttl(C,list(A_1,A_2,A_3),ms=1:3)@data
-          new <- likelihood(ttnsr,theta,omega)
-          cost = c(cost,new); break 
-      }else{
-          C=C_update
[…]

### 2020-02-04 code/simulation_figure 2/Figure2/plot_Figure2.R

@@ -1,0 +1,25 @@
+### Figure: MSE vs. dimension, under various ranks##
+library(ggplot2)
+dlist=seq(from=15,to=40,by=2)
+rlist=c(3,5,8)
+################ plot ############
+load("Figure2.RData")
+Output$rank=as.factor(Output$rank)
+levels(Output$rank)=c("r=3","r=5","r=8")
+
+p=ggplot(data=Output,aes(x=dim,y=MSE))+geom_line(aes(color=rank),size=1.5)+ labs(x=expression("tensor dimension"~(d)), y='MSE',size=1.5)+theme(text = element_text(size=rel(4)),legend.text = element_text(size = 15),axis.text.x = element_text(size=15),axis.text.y = element_text(size=15))
+
+p=p+stat_function(fun=function(d) 10^2.96/(d^2),lty=2,size=0.9)
+g <- ggplot_build(p)
+col=unlist(unique(g$data[[1]]["colour"]))[c(1,2,3)]
+
+p=p+geom_errorbar(aes(ymin=MSE-sd, ymax=MSE+sd), width=0.3,
+position=position_dodge(0.05),color=col[rep(rep(1:3),length(dlist))],size=0.5)+geom_point(aes(shape=rank),size=1.5)
+
+pdf("error_dimension.pdf",width=5,height=4)
+p
+dev.off()
+
+
+
+

### 2020-02-04 code/simulation_figure 2/Figure3/Figure3.R

@@ -1,0 +1,44 @@
+### Figure 3: MSE vs. signal level (alpha), under various ranks##
+source("../../functions.R")
+set.seed(1)
+Output=NULL
+rlist=c(3,5,8)
+##alist=seq(from=0.7,to=3.5,by=0.5)
+##alist=seq(from=0.5,to=2.5,by=0.3)
+alist=c(0.6,0.7,seq(from=0.8,to=2.5,by=0.3))
+d=rep(20,3)
+ntotal=30
+MSE_raw=array(0,dim=c(length(rlist),length(alist),ntotal))
+
+for(j in 1:length(rlist)){
+    for(i in 1:length(alist)){
+        for(nsim in 1:ntotal){
+        alpha=10^(alist[i]); r=rep(rlist[j],3);
+        B_1 = randortho(d[1])[,1:r[1]]
+        B_2 = randortho(d[2])[,1:r[2]]
+        B_3 = randortho(d[3])[,1:r[3]]
+        D = rand_tensor(modes = r)
+        theta = ttl(D,list(B_1,B_2,B_3),ms=1:3)@data
+        theta = theta/max(abs(theta))*alpha ## specify signal = alpha
+        omega = logit((1:4)/5) ## equal spaced on the probit scale
+        
+        ttnsr <- realization(theta,omega)@data
+        
+        #initial point
+        A_1 = randortho(d[1])[,1:r[1]]
+        A_2 = randortho(d[2])[,1:r[2]]
+        A_3 = randortho(d[3])[,1:r[3]]
+        C = rand_tensor(modes = r)
+        C=C/max(abs(ttl(C,list(A_1,A_2,A_3),ms=1:3)@data))*alpha/10 ## initial theta is in the interior 
+        thetainit=ttl(C,list(A_1,A_2,A_3),ms=1:3)@data
+        
+        result <- fit_ordinal(ttnsr,C,A_1,A_2,A_3,omega=TRUE,alpha=2*alpha) ## relaxed bound
+        #MSE=c(MSE,mean((theta-result$theta)^2)/mean(theta^2))
+        MSE_raw[j,i,nsim]=mean((theta-result$theta)^2)/mean(theta^2)
+        ##sprintf("%d\n%d\n%d",i,j,nsim)
+    }
+}
+}
+
[…]

### 2020-02-04 code/simulation_figure 2/Figure4/Figure4.R

@@ -1,0 +1,45 @@
+### Figure: MSE vs. sample size (|Omega|), under various rank##
+source("../../functions.R")
+set.seed(1)
+Output=NULL
+d=rep(20,3)
+rlist=c(3,5,8)
+ntotal=30
+
+rholist=seq(from=0.3,to=1,by=0.1)
+MSE_raw=array(0,dim=c(length(rlist),length(rholist),ntotal))
+for(j in 1:length(rlist)){
+        r=rep(rlist[j],3)
+        B_1 = randortho(d[1])[,1:r[1]]
+        B_2 = randortho(d[2])[,1:r[2]]
+        B_3 = randortho(d[3])[,1:r[3]]
+        D = rand_tensor(modes = r)
+        theta = ttl(D,list(B_1,B_2,B_3),ms=1:3)@data
+        alpha=10
+        theta = theta/max(abs(theta))*alpha ## specify signal = 10
+        omega = logit((1:4)/5) ## equal spaced on the probit scale
+        
+        ttnsr <- realization(theta,omega)@data
+        
+    for(i in 1:length(rholist)){
+        for(nsim in 1:ntotal){
+        rho=rholist[i]
+        data=array(NA,dim=dim(ttnsr))
+        Obs=sample(1:prod(d),prod(d)*rho,replace=F)
+        data[Obs]=ttnsr[Obs]
+        #initial point
+        A_1 = randortho(d[1])[,1:r[1]]
+        A_2 = randortho(d[2])[,1:r[2]]
+        A_3 = randortho(d[3])[,1:r[3]]
+        C = rand_tensor(modes = r)
+        C=C/max(abs(ttl(C,list(A_1,A_2,A_3),ms=1:3)@data))*alpha/10 ## initial theta is in the interior 
+        thetainit=ttl(C,list(A_1,A_2,A_3),ms=1:3)@data
+        
+        result <- fit_ordinal(data,C,A_1,A_2,A_3,omega=TRUE,alpha=alpha) #relaxed bound
+        MSE_raw[j,i,nsim]=mean((theta-result$theta)^2)/mean(theta^2)
+        }
+    }
+}
+
+Output=data.frame(MSE=c(t(apply(MSE_raw,c(1:2),mean))),rank=as.factor(rep(rlist,rep(length(rholist),length(rlist)))),rho=rep(rholist,length(rlist)),sd=c(t(apply(MSE_raw,c(1:2),sd))))
+save(MSE_raw,Output,file="Figure4.RData")

### 2020-02-04 code/simulation_figure 2/Figure5/Figure5.R

@@ -1,0 +1,48 @@
+### Figure 5: MSE vs. number of ordinal levels; half observations ##
+source("../../functions.R")
+set.seed(1)
+Output=NULL
+d=rep(20,3)
+rlist=c(3,5,8)
+rho=0.5
+Klist=2:7
+ntotal=30
+MSE_raw=array(0,dim=c(length(rlist),length(Klist),ntotal))
+for(j in 1:length(rlist)){      
+        r=rep(rlist[j],3)
+        B_1 = randortho(d[1])[,1:r[1]]
+        B_2 = randortho(d[2])[,1:r[2]]
+        B_3 = randortho(d[3])[,1:r[3]]
+        D = rand_tensor(modes = r)
+        theta = ttl(D,list(B_1,B_2,B_3),ms=1:3)@data
+        alpha=10 #
+        theta = theta/max(abs(theta))*alpha ## specify signal = 10
+        
+        for(i in 1:length(Klist)){
+        K=Klist[i]
+        omega = logit((1:(K-1))/K) ## equal space on the logit scale
+        
+        
+        for(nsim in 1:ntotal){ 
+        ttnsr <- realization(theta,omega)@data
+        data=array(NA,dim=dim(ttnsr))
+        Obs=sample(1:prod(d),prod(d)*rho,replace=F)
+        data[Obs]=ttnsr[Obs]
+        #initial point
+        A_1 = randortho(d[1])[,1:r[1]]
+        A_2 = randortho(d[2])[,1:r[2]]
+        A_3 = randortho(d[3])[,1:r[3]]
+        C = rand_tensor(modes = r)
+        C=C/max(abs(ttl(C,list(A_1,A_2,A_3),ms=1:3)@data))*alpha/10 ## initial theta is in the interior 
+        thetainit=ttl(C,list(A_1,A_2,A_3),ms=1:3)@data
+        
+        result <- fit_ordinal(data,C,A_1,A_2,A_3,omega=TRUE,alpha=2*alpha) ## a slightly relaxed boundary
+        
+        MSE_raw[j,i,nsim]=mean((theta-result$theta)^2)/mean(theta^2) ## more stable using rescaled MSE
+            }
+    }
+}
+
[…]

### 2020-02-04 code/simulation_figure 2/Figure7/matlab_Figure7.sh

@@ -1,0 +1,13 @@
+#!/bin/bash
+#SBATCH --mail-user=miaoyan@stat.wisc.edu
+#SBATCH -p long
+#SBATCH -t 5-00:00:00
+#SBATCH -n 4
+#SBATCH --mem-per-cpu=10000M
+export MATLABPATH=/workspace/miaoyan/TensorCompletion_1bit_noisy-master/
+export MATLABPATH=/workspace/miaoyan/TensorCompletion_1bit_noisy-master/tensorlab/
+matlab -nodisplay -nosplash < One_bit_simulation_Fig8.m 
+
+
+
+

### 2020-02-04 code/simulation_figure 2/Figure7/plot_Figure7.R

@@ -1,0 +1,49 @@
+### Figure: MSE vs. number of levels (L), under different methods ##
+source("../../functions.R")
+library(ggplot2)
+library("R.matlab")
+d=rep(20,3)
+r=rep(5,3)
+rho=0.8
+Klist=2:7
+load("Figure8.RData")
+
+Err1_mean=apply(Err1,c(1,2),mean)
+Err1_sd=apply(Err1,c(1,2),sd)
+Err2_mean=apply(Err2,c(1,2),mean)
+Err2_sd=apply(Err2,c(1,2),sd)
+Err3_mean=apply(Err3,c(1,2),mean)
+Err3_sd=apply(Err3,c(1,2),sd)
+Err4_mean=apply(Err4,c(1,2),mean)
+Err4_sd=apply(Err4,c(1,2),sd)
+Err5_mean=apply(Err5,c(1,2),mean)
+Err5_sd=apply(Err5,c(1,2),sd)
+
+index=3
+MAD_for_mode=c(Err1_mean[,index],Err2_mean[,index],Err3_mean[,index],Err4_mean[,index],Err5_mean[,index])
+MAD_for_mode_sd=c(Err1_sd[,index],Err2_sd[,index],Err3_sd[,index],Err4_sd[,index],Err5_sd[,index])
+method=c("Ordinal-T (ours)","Continuous-T","Ordinal-M","1bit-caregory-T","1bit-sign-T")
+method=rep(method,rep(length(Klist),5))
+K=rep(Klist,5)
+Output=data.frame(MAD=MAD_for_mode,sd=MAD_for_mode_sd,method=method,K=K)
+
+################ plot ############
+p=ggplot(data=Output,aes(x=K,y=MAD))+geom_line(aes(color=method),size=1.2)+geom_point(aes(shape=method))+ labs(x=expression('Number of Ordinal Levels'~(L)), y='MAD')+theme(text = element_text(size=rel(4)),legend.text = element_text(size = 15),axis.text.x = element_text(size=15),axis.text.y = element_text(size=15))
+
+g <- ggplot_build(p)
+col=unlist(unique(g$data[[1]]["colour"]))[c(5,3,4,1,2)]
+
+p=p+geom_errorbar(aes(ymin=MAD-sd, ymax=MAD+sd), width=0.01,
[…]

### 2020-02-04 ICML/miaoyan/ordinalT_supp_v2.tex
Update ordinalT_supp_v2.tex

@@
 Notice that  $\mathds{1}_{(u,v)}(x)$ is log-concave to $(u,v,x)$ and $f'(x)$ is log-concave to $x$ because $f'(x) = \frac{e^x}{(1+e^x)^2}$. Those convexities imply that $\mathds{1}_{(u,v)}(x)f'(x)$
-is a log-concave function to $(u,v,x)$. Therefore, we can conclude that the $\lambda(u,v)$ is a concave function from the lemma~\ref{lem:convexity}.
+is a log-concave function to $(u,v,x)$. Therefore, we can conclude that the $\lambda(u,v)$ is a concave function from the lemma~\ref{lem:lossconvexity}.
 \end{proof}
 
-\begin{lem}[Corollary 3.5 in~\cite{brascamp2002extensions}]\label{lem:convexity}
+\begin{lem}[Corollary 3.5 in~\cite{brascamp2002extensions}]\label{lem:lossconvexity}
 Let $F(x,y):\mathbb{R}^{m+n}\rightarrow \mathbb{R}$ be an integrable function where $x\in \mathbb{R}^{m},y\in \mathbb{R}^n$.

### 2020-02-05 ICML/submit/ordinalT.tex
paper for submission

%%%%%%%% ICML 2020 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%

\documentclass{article}

% Recommended, but optional, packages for figures and better typesetting:
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{booktabs} % for professional tables

% hyperref makes hyperlinks in the resulting PDF.
% If your build breaks (sometimes temporarily if a hyperlink spans a page)
% please comment out the following usepackage line and replace
% \usepackage{icml2020} with \usepackage[nohyperref]{icml2020} above.
\usepackage{hyperref}

% Attempt to make hyperref and algorithmic work together better:
\newcommand{\theHalgorithm}{\arabic{algorithm}}

% Use the following line for the initial blind version submitted for review:
\usepackage{icml2020}

% If accepted, instead use the following line for the camera-ready submission:
%\usepackage[accepted]{icml2020}

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Ordinal tensor denoising and completion}
\usepackage{wrapfig}
\usepackage{multirow}
\usepackage{graphicx}
%\usepackage[utf8]{inputenc} % allow utf-8 input
%\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
%\usepackage{booktabs}       % professional-quality tables
\usepackage{amsmath,amssymb} 
\usepackage{amsthm}    % blackboard math symbols
%\usepackage{nicefrac}       % compact symbols for 1/2, etc.
%\usepackage{microtype}      % microtypography
\usepackage{bm}
%\usepackage{subfig}
%\usepackage[english]{babel}
%\usepackage{algorithm}
%\usepackage{appendix}
\usepackage{mathtools}
\mathtoolsset{showonlyrefs}
\usepackage{enumitem}
\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{pro}{Property}
\newtheorem{assumption}{Assumption}

\theoremstyle{definition}
\newtheorem{defn}{Definition}
\newtheorem{cor}{Corollary}
\newtheorem{example}{Example}
\newtheorem{rmk}{Remark}

\usepackage{dsfont}
%\usepackage{algpseudocode,algorithm}
%\algnewcommand\algorithmicinput{\textbf{Input:}}
%\algnewcommand\algorithmicoutput{\textbf{Output:}}
%\algnewcommand\INPUT{\item[\algorithmicinput]}
%\algnewcommand\OUTPUT{\item[\algorithmicoutput]}
%\DeclareMathOperator*{\minimize}{minimize}

\newcommand*{\KeepStyleUnderBrace}[1]{%f
  \mathop{%
    \mathchoice
    {\underbrace{\displaystyle#1}}%
    {\underbrace{\textstyle#1}}%
    {\underbrace{\scriptstyle#1}}%
    {\underbrace{\scriptscriptstyle#1}}%
  }\limits
}
\usepackage{makecell}
\input macros.tex

\usepackage{amssymb}
\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

\begin{document}

\twocolumn[
\icmltitle{Tensor denoising and completion based on ordinal observations}

% It is OKAY to include author information, even for blind
% submissions: the style file will automatically remove it for you
% unless you've provided the [accepted] option to the icml2020
% package.

% List of affiliations: The first argument should be a (short)
% identifier you will use later to specify author affiliations
% Academic affiliations should list Department, University, City, Region, Country
% Industry affiliations should list Company, City, Region, Country

% You can specify symbols, otherwise they are numbered in order.
% Ideally, you should not use this facility. Affiliations will be numbered
% in order of appearance and this is the preferred way.
\icmlsetsymbol{equal}{*}

\begin{icmlauthorlist}
\icmlauthor{Chanwoo Lee}{equal,to}
\icmlauthor{Miaoyan Wang}{equal,to}
\end{icmlauthorlist}

\icmlaffiliation{to}{Department of Statistics, University of Wisconsin at Madison, USA}

\icmlcorrespondingauthor{Miaoyan Wang}{miaoyan.wang@wisc.edu.com}

% You may provide any keywords that you
% find helpful for describing your paper; these are used to populate
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{Higher-order tensors, ordinal observation, tensor decomposition, tensor completion}

\vskip 0.3in
]

% this must go after the closing bracket ] following \twocolumn[ ...

% This command actually creates the footnote in the first column
% listing the affiliations and the copyright notice.
% The command takes one argument, which is text to display at the start of the footnote.
% The \icmlEqualContribution command is standard text for equal contribution.
% Remove it (just {}) if you do not need this facility.

%\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.

\begin{abstract}
Higher-order tensors arise frequently in applications such as neuroimaging, recommendation system, social network analysis, and psychological studies. We consider the problem of low-rank tensor estimation from possibly incomplete, ordinal-valued observations. Two related problems are studied, one on tensor denoising and another on tensor completion. We propose a multi-linear cumulative link model, develop a rank-constrained M-estimator, and obtain theoretical accuracy guarantees. Our mean squared error bound enjoys a faster convergence rate than previous results, and we show that the proposed estimator is minimax optimal under the class of low-rank models. Furthermore, the procedure developed serves as an efficient completion method which guarantees consistent recovery of an order-$K$ $(d,\ldots,d)$-dimensional low-rank tensor using only $\tO(Kd)$ noisy, quantized observations. We demonstrate the outperformance of our approach over previous methods on the tasks of clustering and collaborative filtering.

\end{abstract}

\section{Introduction}
Multidimensional arrays, a.k.a.\ tensors, arise in a variety of applications including recommendation systems~\citep{adomavicius2011context}, social networks~\citep{nickel2011three}, genomics~\citep{wang2017three}, and neuroimaging~\citep{zhou2013tensor}. There is a growing need to develop dimension reduction method for analyzing these noisy, high-dimensional datasets. In one direction, \emph{tensor denoising} -- which aims to recover a signal tensor from its noisy entries -- has gained increased attention in both theory and applications~\citep{xia2019sup,xie2016multispectral}. In a related direction, \emph{tensor completion} examines the minimum number of entries needed for a consistent recovery~\citep{ghadermarzy2018learning,ghadermarzy2019near}. A low-rank structure is often imposed to the signal tensor, which efficiently reduces the intrinsic dimension in both problems.

In the line of these attempts, a number of low-rank tensor estimation methods have been proposed, revitalizing classical methods such as CANDECOMP/PARAFAC (CP) decomposition~\citep{hitchcock1927expression} and Tucker decomposition~\citep{tucker1966some}. These tensor methods have treated the entries as continuous-valued. In many cases, however, we encounter datasets of which the entries are qualitative. For example, the Netflix problem records the ratings of users to movies over weeks. Each data is a rating on a nominal scale \{{\it very like, like, neutral, dislike, very dislike}\}. Another example is in the signal processing, where the digits are frequently rounded or truncated so that only integer values are available. The qualitative observations take values in a limited set of categories, making the learning problem harder compared to the finer, continuous observations.

The analyses of ordinal tensors are complicated by several factors. Ordinal entries are categorical variables with an ordering among the categories; for example, {\it very like} $\prec$ {\it like} $\prec$ {\it neutral} $\prec \cdots$. A reasonable ordinal tensor model has two key properties. First, the model should be invariant under a reversal of categories, say, {\it very like} $\succ$ {\it like} $\succ$ {\it neutral} $\succ \cdots$, but not under arbitrary label permutations. Second, the parameter interpretations should be consistent under merging or splitting of contiguous categories. The classical continuous tensor model~\cite{kolda2009tensor, ghadermarzy2019near} fails in the first aspect, whereas the binary tensor model~\cite{ghadermarzy2018learning} lacks the second property. An appropriate model for ordinal tensors has yet to be studied.

\begin{table*}[ht]
\centerline{\includegraphics[width=16cm]{compare.pdf}}
\vspace{-.3cm}
\caption{Comparison with previous work. For ease of presentation, we summarize the error rate and sample complexity assuming equal tensor dimension in all modes. $K$: tensor order; $L$: number of ordinal levels; $d$: dimension at each mode. }~\label{tab:compare}
\vspace{-.5cm}
\end{table*}

{\bf Our contribution.} This paper presents an efficient low-rank estimation method and theory for tensors with ordinal-valued entries. Our main contributions are summarized in Table~\ref{tab:compare}. We propose a cumulative link model for higher-order tensors, develop a rank-constrained M-estimator, and obtain theoretical accuracy guarantees. The mean squared error is established, and we show that the obtained rate is minimax optimal in high dimensions under the low-rank model. Our estimator enjoys a faster convergence rate $\tO(d^{-(K-1)/2})$ than $\tO(d^{-K})$ in~\citet{ghadermarzy2018learning}, which is a substantial improvement as the order $K$ increases. Furthermore, our proposal serves as an efficient completion algorithm that guarantees consistent recovery of an order-$K$ $(d,\ldots,d)$-dimensional low-rank tensor using only $\tO(Kd)$ noisy, quantized observations.

{\bf Related work.} Our work is related to, but clearly distinctive from, several lines of existing literature. Matrix completion from quantized samples was firstly introduced for binary observations~\cite{cai2013max,davenport2014,bhaskar20151} and then extended to ordinal observations~\cite{bhaskar2016probabilistic}. As we show in Section~\ref{sec:theory}, applying existing matrix methods to an ordinal tensor will result in a suboptimal estimator with a slower convergence rate. Therefore, a full exploitation of the tensor structure is necessary; this is the focus of the current paper.

Our work is also connected to non-Gaussian tensor decomposition. Existing work focuses exclusively on univariate observations such as binary- or continuous-valued entries~\citep{wang2018learning,hong2018generalized}. As we mentioned earlier, the ordinal observations adds considerable challenges to the model formulation. From statistical perspective, our proposed model generalizes the usual binary tensor model while preserving palindromic invariance~\cite{mccullagh1980regression} for ordinal observations. From algorithm perspective, our alternating optimization compares favorably to the approximate (non-convex) algorithm developed in the context of binary tensors~\cite{ghadermarzy2018learning}. We numerically compare the two approaches in Section~\ref{sec:experiment}.

\section{Preliminaries}
Let $\tY\in\mathbb{R}^{d_1\times \cdots \times d_K}$ denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional tensor. We use $y_\omega$ to denote the tensor entry indexed by $\omega$, where $\omega\in[d_1]\times\cdots\times[d_K]$.  The Frobenius norm of $\tY$ is defined as $\FnormSize{}{\tY}=\sum_\omega y^2_\omega$ and the infinity norm of $\tY$ is defined as $\mnormSize{}{\tY}=\max_{\omega}|y_\omega|$. We use $\tY_{(k)}$ to denote the unfolded matrix of size $d_k$-by-$\prod_{i\neq k}d_k$, obtained by reshaping the tensor along the mode-$k$. The Tucker rank of $\tY$ is defined as a length-$K$ vector $\mr = (r_1,\ldots,r_K)$, where $r_k$ is the rank of matrix $\tY_{(k)}$ for all $k \in[K]$. We say that an event $A$ occurs ``with very high probability'' if $\mathbb{P}(A)$ tends to 1 faster than any polynomial of tensor dimension $d_{\min}=\min\{d_1,\ldots,d_K\} \to\infty$.

We use lower-case letters ($a, b, c, \ldots$) for scalars/vectors, upper-case boldface letters ($\mA, \mB, \mC, \ldots$) for matrices, and calligraphy letters ($\tA,\tB,\tC,\ldots$) for tensors of order three or greater. For ease of notation, we allow the basic arithmetic operators (e.g., $\leq, +, -$) to be applied to pairs of tensors in an element-wise manner. We use the shorthand $[n]$ to denote the $n$-set $\{1,\ldots,n\}$ for $n \in N_{+}$.

\section{Model formulation and motivation}
\subsection{Observation model}
Let $\tY$ denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional data tensor. Suppose the entries of $\tY$ are ordinal-valued, and the observation space consists of $L$ ordered levels, denoted by $[L]:=\{1,\ldots,L\}$. We propose a cumulative link model for the ordinal tensor $\tY=\entry{y_\omega}\in[L]^{d_1\times \cdots\times d_K}$. Specifically, assume the entries $y_\omega$ are (conditionally) independently distributed with cumulative probabilities:
\begin{equation}\label{eq:model}
\mathbb{P}(y_\omega\leq \ell)=f(b_\ell-\theta_\omega),\ \text{for all}\ \ell\in[L-1],
\end{equation}
where $\mb=(b_1,\ldots,b_{L-1})$ is a set of unknown scalars satisfying $b_1<\cdots <b_{L-1}$, $\Theta=\entry{\theta_\omega}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is a continuous-valued parameter tensor satisfying certain low-dimensional structure (to be specified later), and $f(\cdot):\mathbb{R}\mapsto[0,1]$ is a known, strictly increasing function. We refer to $\mb$ as the cut-off points and $f$ the link function.

The formulation~\eqref{eq:model} imposes an additive model to the transformed probability of cumulative categories. This modeling choice is to respect the ordering structure among the categories. For example, if we choose the inverse link $f^{-1}(x)=\log {x\over 1-x}$ to be the log odds, then the model~\eqref{eq:model} implies linear spacing between the proportional odds:
\begin{equation}\label{eq:logodd}
\log {\mathbb{P}(y_{\omega}\leq \ell) \over \mathbb{P}(y_\omega >  \ell) } - \log {\mathbb{P}(y_{\omega}\leq {\ell-1}) \over \mathbb{P}(y_\omega >  {\ell-1}) } = b_\ell-b_{\ell-1},
\end{equation}
for all tensor entries $y_\omega$. When there are only two categories in the observation space (e.g. binary tensors), the cumulative model~\eqref{eq:model} is equivalent to the usual multinomial link model. In general, however, when the number of categories $L\geq 3$, the proportional odds assumption~\eqref{eq:logodd} is more parsimonious, in that, the ordered categories can be envisaged as contiguous intervals on the continuous scale, where the points of division are exactly $b_1<\cdots <b_{L-1}$. This interpretation will be made more explicit in the next section.

\subsection{Latent-variable interpretation}\label{sec:latent} 
We show that the ordinal tensor model~\eqref{eq:model} is equivalent to an $L$-level quantization model on $\tY=\entry{y_\omega}$:
\begin{align}\label{eq:quantization}
y_\omega&=
\begin{cases}
1,& \text{if $y^*_\omega\in(-\infty, b_1]$},\\
2,& \text{if $y^*_\omega\in(b_1, b_2]$},\\
\vdots  &\vdots\\
L,& \text{if $y^*_\omega\in(b_{L-1}, \infty)$},\\
\end{cases}
\end{align}
for all $\omega\in[d_1]\times \cdots \times [d_k]$. Here, $\tY^*=\entry{y^*_\omega}$ is a latent continuous-valued tensor following an additive noise model:
\begin{align}\label{eq:latent}
\KeepStyleUnderBrace{\tY^*}_{\text{latent continuous-valued tensor}}=\KeepStyleUnderBrace{\Theta}_{\text{signal tensor}}+\KeepStyleUnderBrace{\tE}_{\text{i.i.d.\ noise}},
\end{align}
where $\tE=\entry{\varepsilon_\omega}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is a noise tensor with i.i.d.\ entries according to distribution $\mathbb{P}(\varepsilon)$. From the viewpoint of~\eqref{eq:latent}, the parameter tensor $\Theta$ can be interpreted as the latent signal tensor prior to contamination and quantization.

The equivalence between the latent-variable model~\eqref{eq:quantization} and the cumulative link model~\eqref{eq:model} is established if the link $f$ is chosen to be athe cumulative distribution function of noise $\varepsilon$, i.e., $f(\theta)=\mathbb{P}(\varepsilon\leq \theta)$. We describe two common choices of link $f$, or equivalently, the distribution of $\varepsilon$.

\begin{example}[Logistic model] The logistic model is characterized by~\eqref{eq:model} with $f(\theta)=(1+e^{-\theta/\sigma})^{-1}$, where $\sigma>0$ is the scale parameter. Equivalently, the noise $\varepsilon_\omega$ in~\eqref{eq:quantization} follows i.i.d.\ logistic distribution with scale parameter $\sigma$. 
\end{example}
\begin{example}[Probit model] The probit model is characterized by~\eqref{eq:model} with
$f(\theta)=\mathbb{P}(z\leq \theta/\sigma)$, where $z\sim N(0,1)$. Equivalently, the noise $\varepsilon_\omega$ in~\eqref{eq:quantization} follows i.i.d.\ $N(0,\sigma^2)$.
\end{example}
Other link functions are also possible, such as Laplace, Cauchy, etc~\cite{mccullagh1980regression}. All the models share the property that the ordered categories can be thought of as contiguous interval on some continuous scale. We should point out that, although the latent-variable interpretation is incisive, our estimation procedure does not refer to the existence of $\tY^*$. Therefore, our model~\eqref{eq:model} is general and still valid in the absence of quantization process. More generally, we make the following assumptions about the link $f$.

\begin{assumption}\label{ass:link}
The link function is assumed to satisfy:
\vspace{-.4cm}
\begin{enumerate}
\item $f(\theta)$ is strictly increasing and twice-differentiable in $\theta\in\mathbb{R}/\{0\}$.
\vspace{-.2cm}
\item $f'(\theta)$ is strictly log-concave and symmetric with respect to $\theta=0$. 
\end{enumerate}
\end{assumption}

\subsection{Problem 1: Tensor denoising}~\label{sec:denoising}
The first question we aim to address is tensor denoising:

(P1) Given the quantization process induced by $f$ and the cut-off points $\mb$, how accurate can we estimate the latent signal tensor $\Theta$ from the ordinal observation $\tY$?

Clearly, the problem (P1) cannot be solved uniformly for all possible $\Theta$. We focus on a class of ``low-rank'' and ``flat'' signal tensors, which is a plausible assumption in practical applications~\cite{zhou2013tensor,bhaskar20151}. Specifically, we consider the parameter space:
\begin{equation}\label{eq:space}
\small \tP=\left\{\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K} \colon \text{rank}(\Theta)\leq \mr,\ \mnormSize{}{\Theta}\leq \alpha\right\}.
\end{equation}
where $\mr=(r_1,\ldots,r_K)$ denotes the Tucker rank of $\Theta$.

The parameter tensor of our interest satisfies two constraints. The first is that $\Theta$ is a low-rank tensor, with $r_k=\tO(1)$ for all $k\in[K]$. Equivalently, $\Theta$ admits the Tucker decomposition:
\begin{equation}\label{eq:Tucker}
\Theta=\tC\times_1\mM_1\times_1 \cdots \times_K \mM_K,
\end{equation}
where $\tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is a core tensor, $\mM_k\in\mathbb{R}^{d_k\times r_k}$ are factor matrices with orthogonal columns, and $\times_k$ denotes the tensor-by-matrix multiplication~\cite{kolda2009tensor}. The Tucker low-rankness is popularly imposed in tensor data analysis, and is shown to provide a reasonable tradeoff between model complexity and model flexibility. Note that, unlike matrices, there are various notations of tensor low-rankness, such as CP rank~\cite{hitchcock1927expression} and train rank~\cite{oseledets2011tensor}. Some notation of low-rankness may lead to mathematically ill-defined optimization; for example, the best low CP-rank tensor approximation may not exist~\cite{de2008tensor}. We choose Tucker representation for parsimony and easy interpretation.

The second constraint is that the entries of $\Theta$ are uniformly bounded in magnitude by a constant $\alpha \in \mathbb{R}_{+}$. In view of~\eqref{eq:latent}, we refer to $\alpha$ as the signal level. The entry-wise bound assumption is a technical condition that avoids the degeneracy in probability estimation with ordinal observations.

\subsection{Problem 2: Tensor completion}
Motivated by applications in collaborative filtering, we also consider a more general setup when only a subset of tensor entries $y_\omega$ are observed. Let $\Omega\subset[d_1]\times \cdots\times[d_K]$ denote the set of observed indices. The second question we aim to address is stated as follows:

(P2) Given an incomplete set of ordinal observations $\{y_{\omega}\}_{\omega\in\Omega}$, how many sampled entries do we need to consistently recover $\Theta$ based on the model~\eqref{eq:model}?

The answer to (P2) depends on the choice of $\Omega$. We consider a general model on $\Omega$ that allows both uniform and non-uniform sampling. Specifically, let $\Pi=\{\pi_{i_1,\ldots,i_K}\}$ denote a predefine probability distribution over the index set such that $\sum_{\omega\in[d_1]\times \cdots \times [d_K]} \pi_\omega =1$. We assume that each index in $\Omega$ is drawn with replacement using distribution $\Pi$. This sampling model relaxes the uniform sampling in literature and is arguably a better fit in applications.

We consider the same parameter space~\eqref{eq:space} for the completion problem. In addition to the reasons mentioned in Section~\ref{sec:denoising}, the entrywise bound assumption also serves as the incoherence requirement for completion. In classical matrix completion, the incoherence is often imposed on the singular vectors. This assumption is recently relaxed for ``flat'' matrices with bounded magnitude~\cite{negahban2011estimation,cai2013max,bhaskar20151}. We adopt the same assumption for higher-order tensors.

\section{Rank-constrained M-estimator}\label{sec:theory}
We present a general treatment to both problems mentioned above. With a little abuse of notation, we use $\Omega$ to denote either the full index set $\Omega=[d_1]\times \cdots \times [d_K]$ (for the tensor denoising) or a random subset induced from the sampling distribution $\Pi$ (for the tensor completion). Define $b_0=-\infty$, $b_L=\infty$, $f(-\infty)=0$ and $f(\infty)=1$. The log-likelihood associated with the observed entries is
\begin{align}\label{eq:objective}
\logl(\Theta, \mb)&=\sum_{\omega\in\Omega}\sum_{\ell\in[L]} \Big\{\mathds{1}_{\{y_\omega=\ell\}} \log \big[f(b_\ell-\theta_\omega)- \notag \\
&\quad \quad \quad \quad \quad \quad \quad  f(b_{\ell-1}-\theta_\omega)\big]\Big\}.
\end{align}
We propose a rank-constrained maximum likelihood estimator (a.k.a.\ M-estimator) for $\Theta$:
\begin{align}\label{eq:estimator}
\hat \Theta&=\argmax_{\Theta\in \tP}\logl(\Theta, \mb),\ \text{where}\notag \\
\tP&=\left\{\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K} \colon \text{rank}(\Theta)\leq \mr,\ \mnormSize{}{\Theta}\leq \alpha\right\}.
 \end{align}
In practice, the cut-off points $\mb$ are unknown and should be jointly estimated with $\Theta$. For technical convenience, we assume in this section that the cut-off points $\mb$ are known. The adaptation of unknown $\mb$ is addressed in Section~\ref{sec:algorithm} and the Supplement. Similar treatment applies to $\mr$.

We define a few key quantities that will be used in our theory. Let $g_\ell=f(\theta+b_\ell)-f(\theta+b_{\ell-1})$ for all $\ell\in[L]$, and
\begin{align}\label{eq:regular}
A_\alpha &= \min_{\ell\in[L], |\theta|\leq \alpha} g_\ell(\theta),\quad  U_\alpha=\max_{\ell\in[L], |\theta|\leq \alpha } {\dot{g}_\ell(\theta)\over g_\ell(\theta)},\\
 L_\alpha &= \min_{\ell\in[L], |\theta|\leq \alpha }\left[{\dot{g}_\ell^2 (\theta)\over g_\ell^2(\theta)} -{\ddot{g}_\ell(\theta)\over g_\ell(\theta)}\right],
\end{align}
where $\dot{g}(\theta)=dg(\theta)/d\theta$, and $\alpha$ is the entrywise bound of $\Theta$. In view of equation~\eqref{eq:latent}, these quantities characterize the geometry including flatness and convexity of the latent noise distribution. Under the Assumption~\ref{ass:link}, all these quantities are strictly positive and independent of tensor dimension.

\subsection{Estimation error for tensor denoising}\label{sec:denosing}
For the tensor denoising problem, we assume the full set of tensor entries are observed. We assess the estimation accuracy using the mean squared error (MSE):
\[
\text{MSE}(\hat \Theta, \trueT)={1\over \prod_k d_k}\FnormSize{}{\Theta-\trueT}^2.
\]
The next theorem establishes the upper bound for the proposed $\hat \Theta$ in~\eqref{eq:estimator}.

\begin{thm}[Statistical convergence] \label{thm:rate}
Consider an ordinal tensor $\tY\in[L]^{d_1\times\dots\times d_K}$ generated from model~\eqref{eq:model}, with the link function $f$ and the true coefficient tensor $\trueT\in\tP$. Define $r_{\max}=\max_k r_k$. Then, with very high probability, the estimator in~\eqref{eq:estimator} satisfies
\begin{equation}\label{eq:rate}
\text{MSE}(\hat \Theta, \trueT) \leq \min\left( 4\alpha^2,\ {c_2  U^2_\alpha r_{\max}^{K-1}  \over  L^2_\alpha } {\sum_kd_k\over  \prod_k d_k} \right),
\end{equation}
where $c_1, c_2>0$ are two constants that depend only on $K$. 
\end{thm}
Theorem~\ref{thm:rate} establishes the statistical convergence for the estimator~\eqref{eq:estimator}. In fact, the proof of this theorem (see the Supplement) shows that the same statistical rate holds, not only for the global optimizer~\eqref{eq:estimator}, but also for any local optimizer $\check \Theta$ in the level set $\{\check\Theta\in\tP: \logl(\check\Theta)\geq \logl(\trueT)\}$. This suggests that the local optimality itself is not necessarily a severe concern in our context, as long as the objective value upon convergence is large enough. In Sections~\ref{sec:algorithm} and~\ref{sec:experiment}, we perform empirical studies to assess the algorithmic stability and its impact to statistical efficiency.

A similar conclusion is obtained for the prediction error, measured in Kullback-Leibler (KL) divergence, between the categorical distributions in the observation space. 
\begin{cor}[Prediction error]~\label{cor:prediction}
Assume the same set-up as in Theorem~\ref{thm:rate}. Let $\mathbb{P}_{\tY}$ and $\hat{\mathbb{P}}_{\tY}$ denote the distributions generating the $L$-level ordinal tensor $\tY$, given the true parameter $\Theta$ and its estimator $\hat \Theta$, respectively. Assume $L\geq 2$. Then, with very high probability, 
\begin{equation}\label{eq:KLrate}
\text{KL}(\mathbb{P}_{\tY} || \hat{\mathbb{P}}_{\tY}) \leq  {c_2 U^2_\alpha r_{\max}^{K-1}  \over L^2_\alpha } {(4L-6)\dot{f}^2(0)  \over A_\alpha} {\sum_kd_k\over  \prod_k d_k},
\end{equation}
where $c_1, c_2>0$ are the same constants as in Theorem~\ref{thm:rate}.  
\end{cor}
To gain insight into these bounds, we consider a special setting with equal dimension in all modes, i.e., $d_1=\cdots=d_K=d$. In such a case, our bound \eqref{eq:rate} reduces to
\begin{equation*}\label{eq:ours}
\text{MSE}(\hat \Theta, \trueT) \asymp d^{-(K-1)}, \quad \text{as}\ d\to \infty.
\end{equation*}
Hence, our estimator achieves consistency with polynomial convergence rate. We compare the bound with existing literature. In the special case $L=2$, \citet{ghadermarzy2018learning} proposed a max-norm constrained estimator $\tilde{\Theta}$ with $\text{MSE}(\tilde{\Theta}, \trueT) \asymp  d^{-(K-1)/2}$. In contrast, our estimator converges at a rate of $d^{-(K-1)}$, which is substantially faster than theirs. This provides a positive answer to the open question posed in~\citet{ghadermarzy2018learning} whether the square root in the bound is removable. The improvement stems from utilizing the exact low-rankness of $\Theta$, whereas the surrogate rank measure employed in~\citet{ghadermarzy2018learning} is scale-sensitive.

Our bound also generalizes the previous results on ordinal matrices. The convergence rate for rank-constrained matrix estimation was $\tO(1/\sqrt{d})$~\citep{bhaskar2016probabilistic}, which fits into our special case when $K=2$. Furthermore, our results~\eqref{eq:rate} and~\eqref{eq:KLrate} reveal that the convergence becomes favorably as the order of data tensor increases. Intuitively, the sample size for tensor data analysis is the number of entries, $\prod_k d_k$, and the number of free parameters is roughly on the order of $\sum_{k}d_k$, assuming $r_{\max}=\tO(1)$. A higher tensor order implies higher effective sample size per parameter, and thus exhibits a faster convergence rate in high dimensions.

We next show the statistical optimality of our estimator $\hat \Theta$. The result is based on the information theory, and applies to all estimators in $\tP$, including but not limited to $\hat \Theta$ in~\eqref{eq:estimator}.

\begin{thm}[Minimax lower bound]\label{thm:minimax} 
Assume the same set-up as in Theorem~\ref{thm:rate}, and $d_{\max}=\max_k d_k \geq 8$. Let $\inf_{\hat \Theta}$ denote the infimum over all estimators $\hat \Theta\in\tP$ based on the ordinal tensor observation $\tY\in[L]^{d_1\times \cdots \times d_K}$. Then, under the model~\eqref{eq:model},
\begin{align}\label{eq:lower}
\inf_{\hat \Theta }\sup_{\trueT\in\tP} \mathbb{P}\Big\{& \text{MSE}(\hat \Theta, \trueT) \\
& \geq {1\over 256} \min\left( \alpha^2, \ { Cr_{\max}d_{\max} \over \prod_k d_k} \right) \Big\} \geq {1\over 8},
\end{align}
where $C=C(\alpha, L, f,\mb)>0$ is a constant independent of tensor dimension and the rank. 
\end{thm}
We see that the lower bound matches the upper bound in~\eqref{eq:rate} on the polynomial order of tensor dimension. Therefore, our estimator~\eqref{eq:estimator} is order-optimal.

\subsection{Sample complexity for tensor completion}
We now consider the tensor completion problem, when only a subset of entries $\Omega$ are observed. We consider a general sampling procedure induced by $\Pi$. The recovery accuracy is assessed by the weighted squared error:
\begin{align}\label{eq:weighted}
\PiFnormSize{}{\Theta-\hat \Theta}^2&\stackrel{\text{def}}{=}
{1\over |\Omega|}\mathbb{E}_{\Omega\sim \Pi}\FnormSize{}{\Theta-\hat \Theta}^2\notag \\
&=\sum_{\omega\in[d_1]\times \cdots \times [d_K]} \pi_{\omega}(\Theta_{\omega}-\hat \Theta_{\omega})^2.
\end{align}

Note that the recovery error depends on the distribution $\Pi$. In particular, tensor entries with higher sampling probabilities are equipped with better accuracy guarantee, compared to the ones with lower sampling probabilities.

\begin{rmk} If we assume each entry is sampled with strictly positive probability; i.e.\ there exits a constant $\mu> 0$ s.t.
\[
\ \pi_\omega\geq {1\over \mu \prod_k d_k},\quad \text{for all}\ \omega\in[d_1]\times\cdots \times [d_K],
\]
then the error in~\eqref{eq:weighted} provides an upper bound for MSE:
\[
\PiFnormSize{}{\Theta-\hat \Theta}^2 \geq {\FnormSize{}{\Theta-\hat \Theta}^2\over \mu \prod_kd_k}={1\over \mu}\text{MSE}(\hat \Theta, \trueT).
\]
The equality is attained under uniform sampling with $\mu=1$. 
\end{rmk}

\begin{thm}~\label{thm:completion}
Assume the same set-up as in Theorem~\ref{thm:rate}. Suppose that we observe a subset of tensor entries $\{y_\omega\}_{\omega\in\Omega}$, where $\Omega$ is chosen at random with replacement according to a probability distribution $\Pi$. Let $\hat \Theta$ be the solution to~\eqref{eq:estimator}, and assume $r_{\max}=\tO(1)$. Then, with very high probability, 
\[
\PiFnormSize{}{\Theta-\hat \Theta}^2 \to 0,\quad \text{ as }\quad {|\Omega|\over \sum_k d_k} \ \to \infty. 
\]
\end{thm}
Theorem~\ref{thm:completion} shows that our estimator achieves consistent recovery using as few as $\tO(Kd)$ noisy, quantized observations from an order-$K$ $(d,\ldots,d)$-dimensional tensor. Note that $\tO(Kd)$ matches the degree of freedom for an order-$K$ tensor of fixed rank $\mr$, suggesting the optimality of our sample requirement. This sample complexity substantially improves over earlier result $\tO(d^{\lceil K/2\rceil})$ based on square matrixization~\cite{mu2014square}, or $\tO(d^{N/2})$ based on tensor nuclear-norm regularization~\cite{yuan2016tensor}. Existing methods that achieve $\tO(Kd)$ sample complexity require either a deterministic cross sampling design~\cite{zhang2019cross} or univariate measurements~\cite{ghadermarzy2018learning}. Our method extends the conclusion to multi-level measurements under a broader class of sampling schemes.

\section{Numerical Implementation}\label{sec:algorithm}
We describe the algorithm to seek the optimizer of~\eqref{eq:objective}. In practice, the cut-off points $\mb$ are often unknown, so we choose to maximize $\logl$ jointly over $(\Theta,\mb)\in\tP\times \tB$. The objective $\logl$ is concave in $(\Theta,\mb)$ whenever $f'$ is log-concave (See the Supplement). However, the feasible set $\tP$ is non-convex, which makes the optimization~\eqref{eq:objective} a non-convex problem. We employ the alternating optimization approach by utilizing the Tucker representation of $\Theta$. Specifically, based on~\eqref{eq:Tucker} and~\eqref{eq:objective}, the objective function consists of $K+2$ blocks of variables, one for the cut-off points $\mb$, one for the core tensor $\tC$, and $K$ for the factor matrices $\mM_k$'s. The optimization is a simple convex problem if any $K+1$ out of the $K+2$ blocks are fixed. We update one block at a time while holding others fixed, and alternate the optimization throughout the iteration. The convergence is guaranteed whenever $\logl$ is bounded from above, since the alternating procedure monotonically increases the objective. The Algorithm~\ref{alg} gives the full description.

\begin{algorithm}[t]
        \caption{Ordinal tensor decomposition and completion}\label{alg}
        \begin{algorithmic}[]
            \STATE{\bfseries Input:}  \text{ Ordinal data tensor }
            $\mathcal{Y}\in [L]^{d_1\times\cdots\times d_K}$, rank $\mr\in \mathbb{N}_{+}^{K}$, entry-wise bound $\alpha\in \mathbb{R_+}$.
            \STATE{\bfseries Output:} $(\hat\Theta,\hat{\mb}) =  \argmax_{(\Theta,\mb)\in \mathcal{P}\times\mathcal{B}}  \mathcal{L}_{\mathcal{Y},\Omega}(\Theta,\mb).$
            \STATE Initialize core tensor $\mathcal{C}^{(0)}$, factor matrices $ \{\mM_k^{(0)}\}$, and cut-off points $\mb^{(0)}$.
            \FOR{$t = 1,2,\cdots,$}
            \FOR{$k = 1,2,\cdots,K$}
            \STATE{Update $\mM^{(t+1)}_k$ while fixing other blocks:}
            \STATE $\mM_k^{(t+1)}\gets\argmax_{\mM_k\in\mathbb{R}^{d_k\times r_K}}\mathcal{L}_{\mathcal{Y},\Omega}(\mM_k)$,
            \newline
             s.t. $\mnormSize{}{\Theta^{(t+1)}}\leq \alpha$, where $\Theta^{(t+1)}$ is the parameter tensor based on the current block estimates. 
            \ENDFOR
            \STATE {Update $\mathcal{C}^{(t+1)}$ while fixing other blocks:}
              \newline$\mathcal{C}^{(t+1)} \gets \argmax_{\mathcal{C}\in \mathbb{R}^{r_1\times\cdots\times r_K}}\mathcal{L}_{\mathcal{Y},\Omega}(\mathcal{C})$, s.t. $\mnormSize{}{\Theta^{(t+1)}}\leq \alpha.$
               \STATE {Update $\Theta^{(t+1)}$ based on the current block estimates:} 
                   \newline
                   $\Theta^{(t+1)} \gets \mathcal{C}^{(t+1)}\times_1\mM_1^{(t+1)}\cdots\times_K\mM_K^{(t+1)}.$
            \STATE {Update $\mb^{(t+1)}$ while fixing $\Theta^{(t+1)}$:}
              \newline
              $\mb^{(t+1)} \gets \argmax_{\mb\in \mathbb{R}^{L-1}}\mathcal{L}_{\mathcal{Y},\Omega}\big(\Theta^{(t+1)},\mb\big).$
            \ENDFOR
            \STATE \textbf{return}
            $(\hat \Theta,\hat \mb)$
        \end{algorithmic}
 \end{algorithm}

Before concluding this section, we comment on two implementation details. First, the problem~\eqref{eq:estimator} is non-convex, so Algorithm 1 usually has no theoretical guarantee on global optimality. Nevertheless, as shown in Section~\ref{sec:denosing}, the desired rate holds not only for the global optimizer, but also for the local optimizer with $\logl(\hat \Theta)\geq \logl(\trueT)$. In practice, we find the convergence point $\hat \Theta$ is often satisfactory, in that the corresponding objective $\logl(\hat \Theta)$ is close to and actually slightly larger than the objective evaluated at the true parameter $\logl(\trueT)$. Figure~\ref{fig:stability} shows the trajectory of the objective function that is output in the default setting of Algorithm 1, with the input tensor generated from probit model~\eqref{eq:model} with $d_1=d_2=d_3=d$ and $r_1=r_2=r_3=r$. The dashed line is the objective value at the true parameter $\logl(\trueT)$. It is observed that the algorithm generally converges quickly to a desirable value in reasonable number of steps. The actual running time per iteration is shown in the plot legend.

\begin{figure}[H]
\centering
\vspace{-.5cm}
\includegraphics[width=8cm]{algorithm.pdf}\label{fig:stability}
\vspace{-.3cm}
\caption{Trajectory of objective function with various $d$ and $r$.}
\vspace{-.7cm}
\end{figure}

Second, the algorithm takes the rank $\mr$ as an input. In practice, the rank $\mr$ is hardly known and needs to be estimated from the data. We suggest to use Bayesian information criterion (BIC) and choose the rank that minimizes BIC; i.e.
\begin{align}\label{eq:BIC}
\hat \mr&=\argmin_{\mr\in\mathbb{N}^K_{+}} \text{BIC}(\mr)\\
&=\argmin_{\mr\in\mathbb{N}^K_{+}}\{-2\tL_{\tY}(\hat \Theta(\mr),\hat \mb(\mr))+p_e(\mr)\log (\prod_k d_k)\},
\end{align}
where $\hat \Theta(\mr), \hat\mb(\mr)$ are the estimates given the rank $\mr$, and $p_e(\mr)\stackrel{\text{def}}{=}\sum_k (d_k-r_k)r_k+\prod_k r_k$ is the effective number of parameters in the model. We select $\hat \mr$ that minimizes BIC through a grid search. The choice of BIC is intended to balance between the goodness-of-fit for the data and the degrees of freedom in the population model.

\section{Experiments}\label{sec:experiment}
In this section, we evaluate the empirical performance of our method. We investigate both the complete and the incomplete settings, and compare the recovery accuracy with other tensor-based methods. Unless otherwise stated, the ordinal data tensors are generated from model~\eqref{eq:model} using standard probit link $f$. We consider the setting with $K=3$, $d_1=d_2=d_3=d$, and $r_1=r_2=r_3=r$. The parameter tensors are simulated based on~\eqref{eq:Tucker}, where the core tensor entries are i.i.d.\ drawn from $N(0,1)$, and the factors $\mM_k$ are uniformly sampled (with respect to Haar measure) from matrices with orthonormal columns. We set the cut-off points $b_\ell=f^{-1}(\ell/L)$ for $\ell\in[L]$, such that $f(b_\ell)$ are evenly spaced from 0 to 1. In each simulation study, we report the summary statistics across $n_{\text{sim}}$ = 30 replications.

\subsection{Finite-sample performance}\label{sec:simulation}

The first experiment examines the performance under complete observations. We assess the empirical relationship between the MSE and various aspects of model complexity, such as dimension $d$, rank $r$, and signal level $\alpha=\mnormSize{}{\Theta}$. Figure~\ref{fig:finite}a plots the estimation error versus the tensor dimension $d$ for three different ranks $r\in\{3,5,8\}$. The decay in the error appears to behave on the order of $d^{-2}$, which is consistent with our theoretical results~\eqref{eq:rate}. It is seen that a higher rank leads to a larger error, as reflected by the upward shift of the curve as $r$ increases. Indeed, a higher rank implies a higher intrinsic dimension, thus increasing the difficulty of the estimation. Figure~\ref{fig:finite}b shows the estimation error versus the signal level under $d=20$. Interestingly, a larger estimation error is observed when the signal is either too small or too large. The non-monotonic behavior may seem surprising, but this is in fact an intrinsic feature in the estimation with ordinal data. In view of the latent-variable interpretation (see Section~\ref{sec:latent}), estimation from ordinal observation can be interpreted as an inverse problem of quantization. Therefore, the estimation error will diverge when the noise $\tE$ is negligible, because there will be no way to distinguish two different signal tensors, e.g., $\Theta_1=\ma_1\otimes \ma_2\otimes \ma_3$ and $\Theta_2=\text{sign}(\ma_1)\otimes \text{sign}(\ma_2)\otimes \text{sign}(\ma_3)$, from the quantized observations. This phenomenon is clearly contrary to the classical continuous-valued tensor problem~\cite{davenport2014,sur2018modern}.

The second experiment investigates the incomplete observations. We consider $L$-level tensors with $d=20$, $\alpha=10$ and choose a subset of tensor entries via uniform sampling. Figure~\ref{fig:finite}c shows the estimation error of $\hat \Theta$ versus the fraction of observation $\rho=|\Omega|/d^K$. As expected, the error reduces with increased $\rho$ or decreased $r$. Figure~\ref{fig:finite}d evaluates the impact of ordinal levels $L$ to estimation accuracy, under the setting $\rho=0.5$. It is seen that the performance improves as $L$ grows, especially from binary observations ($L=2$) to multi-level ordinal observations ($L\geq 3$). The result showcases the benefit of multi-level observations provide compared to binary observations.

\begin{figure}[http]
\includegraphics[width=8.3cm]{panel.pdf}
\vspace{-.4cm}
\caption{Empirical relationship between (relative) MSE versus (a) dimension $d$, (b) signal level $\alpha$, (c) observation fraction $\rho$, and (d) number of ordinal levels $L$. In panels (b)-(d), we plot the relative MSE $=\FnormSize{}{\hat \Theta-\trueT}/\FnormSize{}{\trueT}$ for better visualization.}
\label{fig:finite}
\vspace{-.2cm}
\end{figure}

\subsection{Comparison with alternative methods}
Next, we compare our ordinal tensor method ({\bf Ordinal-T}) with three popular low-rank methods:

\begin{itemize}[itemsep=0pt,topsep=-2pt,leftmargin=*]
\item Continuous tensor decomposition ({\bf Continuous-T})~\cite{filipovic2015tucker,oh2018scalable} is the classical Tucker approximation method. 
\item One-bit tensor completion ({\bf 1bit-T})~\cite{ghadermarzy2018learning} is a max-norm penalized tensor learning method based on partial binary observations. 
\item Ordinal matrix completion ({\bf Ordinal-M})~\cite{bhaskar2016probabilistic} is a rank-constrained matrix estimation method based on noisy, quantized observations. 
\vspace{-.4cm}
\end{itemize}

We apply each of the above methods to $L$-level ordinal tensors $\tY$ generated from model~\eqref{eq:model}. The {\bf Continuous-T} is applied directly to $\tY$ by treating the $L$ levels as continuous observations. The {\bf Ordinal-M} is applied to the matrix $\tY_{(1)}$ obtained via 1-mode unfolding. The {\bf 1bit-T} is applied to $\tY$ in two ways. The first approach ({\bf 1bit-sign-T}) follows from~\citet{ghadermarzy2018learning} that transforms $\tY$ to a binary tensor of equal size, by taking the entrywise sign of the mean-adjusted tensor, $\tY- |\Omega|^{-1}\sum_\omega y_\omega$. The second approach ({\bf 1bit-category-T}) transforms the order-3 ordinal tensor $\tY$ to an order-4 binary tensor $\tY^{\sharp}=\entry{y^{\sharp}_{ijkl}}$ via dummy variable encoding; i.e., $y^{\sharp}_{ijk\ell}=\mathds{1}_{\{y_{ijk}=\ell\}}$ for all $\ell\in[L-1]$. We evaluate the methods by their capabilities in predicting the most frequent label for each entry, i.e.,\ $y_\omega^{\text{mode}}=\arg\max_\ell\mathbb{P}(y_\omega=\ell)$. Two performance metrics are considered: mean absolute deviation, $\text{MAD}=d^{-K}\sum_\omega |y_\omega^{\text{mode}}-\hat y_\omega^{\text{mode}}|$, and misclassification rate, $\text{MCR}=d^{-K}\sum_\omega|y_\omega^{\text{mode}}-\text{round}(\hat y_\omega^{\text{mode}})|$, where $\text{round}(\cdot)$ denotes the nearest integer of the (possibly continuous-valued) prediction, e.g. returned by {\bf Continuous-T}. Note that MAD penalizes the large deviation more heavily than does MCR.

Figure~\ref{fig:compare} compares the prediction accuracy under the setting $\alpha=10$, $d=20$, and $r=5$. The problem size we considered is comparable to~\citet{ghadermarzy2018learning}. It is seen that our method outperforms the others in both MAD and MCR. In particular, methods built on multi-level observations ({\bf Ordinal-T}, {\bf Ordinal-M}, {\bf 1bit-category-T}) exhibit stable MCR over $\rho$ and $L$, whereas the others two methods ({\bf Continuous-T}, {\bf 1bit-sign-T}) generally fail except for $L=2$ (Figures~\ref{fig:compare}a-b). This observation highlights the necessity of modeling multi-level probabilities in classification task. Interestingly, although both {\bf 1bit-category-T} and our method {\bf Ordinal-T} behave similarly for binary tensors ($L=2$), the improvement of our method is substantial as $L$ increases (Figures~\ref{fig:compare}a and~\ref{fig:compare}c). One possible reason is that our method incorporates the intrinsic ordering among the $L$ levels via proportional odds assumption~\eqref{eq:logodd}, whereas {\bf 1bit-category-T} ignores the ordinal structure and dependence among the induced binary entries. Figures~\ref{fig:compare}c-d assess the prediction accuracy with sample size. We see a clear advantage of our method ({\bf Ordinal-T}) over the matricization ({\bf Ordinal-M}) in both complete and non-complete observations. When the observation fraction is small, e.g., ${|\Omega|/ d^K}=0.4$, the tensor-based completion shows $\sim$ 30\% reduction in error compared to the matricization.

\begin{figure}[t]
\includegraphics[width=8.3cm]{panel2.pdf}
\vspace{-.4cm}
\caption{Performance comparison in MSE (a-b) and MAD (c-d). (a, b) Prediction errors versus the number of ordinal levels $L$ when $\rho=0.8$. (c, d) Prediction errors versus sample complexity $\rho=|\Omega|/d^K$ when $L=5$. }\label{fig:compare}
\vspace{-.4cm}
\end{figure}

We also compare the methods by their abilities in predicting the median score, $y_\omega^{\text{median}}=\min\{\ell\colon\mathbb{P}(y_\omega=\ell)\geq 0.5\}$. Note that, under the latent variable model~\eqref{eq:latent} and Assumption~\ref{ass:link}, the median score can be directly interpreted as the quantized $\Theta$ without noise; i.e.\ $y_\omega^{\text{median}}=\sum_\ell \mathds{1}_{\theta_\omega\in(b_{\ell-1},b_\ell]}$. Similar results demonstrate the outperformance of our method (see the Supplement and Figure S1).

\section{Data Applications}
We apply our ordinal tensor method to two real-world datasets. In the first application, we use our model to analyze an ordinal tensor consisting of structural connectivities among 68 brain regions for 136 individuals from Human Connectome Project (HCP)~\citep{geddes2016human}. In the second application, we perform tensor completion to an ordinal dataset with missing values. The data tensor records the ratings on a scale of 1 to 5 from 42 users to 139 songs on 26 contexts~\citep{baltrunas2011incarmusic}.

\subsection{Human Connectome Project (HCP)}
Each entry in the HCP dataset takes value on a nominal scale, \{{\it high, moderate, low}\}, indicating the strength level of fiber connection. We convert the dataset to a 3-level ordinal tensor $\tY\in[3]^{68\times 68\times 136}$ and apply the ordinal tensor method with a logistic link function. The BIC suggests $\mr = (23,23,8)$ with $\logl(\hat{\Theta},\hat{\mb}) = -216,646$. Based on the estimated Tucker factors $\{\hat \mM_k\}$, we perform a clustering analysis via K-mean on the brain nodes (see detailed procedure in the Supplement). We find that the 68 brain nodes are grouped into eight clusters, and the clustering capture the spatial separation of brain nodes very well (see the Supplement). In particular, the top two clusters represent the left and right hemispheres; the smaller clusters represent local regions driving by similar nodes. For example, the cluster IV consists of brain nodes {\it postcentral}, {\it posteriorcingulate}, {\it pericalcarine} and {\it temporalpole}, all of which are located at {\it SupraM}. The identified similarity among brain nodes suggests the applicability of our method.

We compare the goodness-of-fit of various tensor methods on the HPC data. Table~\ref{table:CV} summarizes the prediction error via 5-fold stratified cross-validation averaged over 10 runs. It is clearly seen that our method outperform the others especially in MAD.

\begin{table}[http]
\resizebox{.48\textwidth}{!}{
\begin{tabular}{c|c|c}
\Xhline{2\arrayrulewidth}
\multicolumn{3}{c}{Human Connectome Project (HCP) dataset}\\
\Xhline{2\arrayrulewidth}
Method                    & MAD & MCR\\
\hline
Ordinal-T (ours) & 0.1607 (0.0005)& 0.1606 (0.0005)\\
 Continuous-T&0.2530 (0.0002)&0.1599 (0.0002)\\
 1bit-sign-T&0.3566 (0.0010)&0.1563 (0.0010)\\
 \Xhline{2\arrayrulewidth}
\multicolumn{3}{c}{InCarMusic dataset}\\
\Xhline{2\arrayrulewidth}
Method                    & MAD & MCR\\
\hline
Ordinal-T (ours) & 1.37 (0.039) &  0.59 (0.009)\\
Continuous-T & 2.39 (0.152)&0.94 (0.027)\\
1bit-sign-T&1.39 (0.003)& 0.81 (0.005)\\
\hline
\end{tabular}
}
\caption{Comparison of prediction error in the HPC and InCarMusic analyses. Standard errors are reported in parentheses.}\label{table:CV} 
\vspace{-.4cm}
\end{table}

\subsection{InCarMusic recommendation system}
In the second application, we apply ordinal tensor completion to a recommendation system {\it InCarMusic}. {\it InCarMusic} is a mobile application that offers music recommendation to passengers of cars based on contexts~\cite{baltrunas2011incarmusic}. The data tensor consists of partially observed ratings from scale 1 to 5 of 42 users to 139 songs on 26 contexts. Our goal is to perform tensor completion to impute the unobserved entries in the $42\times 139\times 26$ ordinal tensor and thereby we can offer context-specific music recommendation to users. Table~\ref{table:CV} shows the averaged prediction error via 5 fold cross validation. Again, our method achieves the best prediction error among the three. The data tensor has 98\% missing rate which makes the accurate classification challenging. Nevertheless, our method correctly classifies half of the unseen entries.

\section{Conclusion}
We have developed a low-rank tensor estimation method based on possibly incomplete, ordinal-valued observations. A sharp error bound is established, and we demonstrate the outperformance of our approach compared to other methods. The work unlocked several directions of future research. One interesting question would be the inference problem, i.e.. to assess the uncertainty of the obtained estimates and the imputation. Other directions include the trade-off between (non)convex optimization and statistical/computational efficiency. While convex relaxation is a popular approach for matrix-based problem, computing convex regularizer could be NP hard for tensors of order $K\geq 3$~\cite{hillar2013most, friedland2018nuclear}. The interplay between computational efficiency and statistical accuracy in general tensor problems warrants future research.

\newpage
\bibliography{tensor_wang}
\bibliographystyle{icml2020}

\end{document}

% This document was modified from the file originally made available by
% Pat Langley and Andrea Danyluk for ICML-2K. This version was created
% by Iain Murray in 2018, and modified by Alexandre Bouchard in
% 2019 and 2020. Previous contributors include Dan Roy, Lise Getoor and Tobias
% Scheffer, which was slightly modified from the 2010 version by
% Thorsten Joachims & Johannes Fuernkranz, slightly modified from the
% 2009 version by Kiri Wagstaff and Sam Roweis's 2008 version, which is
% slightly modified from Prasad Tadepalli's 2007 version which is a
% lightly changed version of the previous year's version by Andrew
% Moore, which was in turn edited from those of Kristian Kersting and
% Codrina Lauth. Alex Smola contributed to the algorithmic style files.

### 2020-02-05 ICML/submit/ordinalT.tex
update paper

@@
 There is a growing need to develop dimension reduction method for analyzing these noisy, high-dimensional datasets.
-In one direction, \emph{tensor denoising} -- which aims to recover a signal tensor from its noisy entries -- has gained increased attention in both theory and applications~\citep{xia2019sup,xie2016multispectral}.
-In a related direction, \emph{tensor completion} examines the minimum number of entries needed for a consistent recovery~\citep{ghadermarzy2018learning,ghadermarzy2019near}.
-A low-rank structure is often imposed to the signal tensor, which efficiently reduces the intrinsic dimension in both problems.
+The problem of tensor denoising -- which aims to recover a signal tensor from its noisy entries -- has gained increased attention in both theory and applications~\citep{xia2019sup,xie2016multispectral}.
+A related problem, tensor completion, examines the minimum number of entries needed for a consistent recovery~\citep{ghadermarzy2018learning,ghadermarzy2019near}.
+Low-rankness is often imposed to the signal tensor, which efficiently reduces the intrinsic dimension in both problems.
 
-In the line of these attempts, a number of low-rank tensor estimation methods have been proposed, revitalizing classical methods such as CANDECOMP/PARAFAC (CP) decomposition~\citep{hitchcock1927expression} and Tucker decomposition~\citep{tucker1966some}.
-These tensor methods have treated the entries as continuous-valued.
+A number of low-rank tensor estimation methods have been proposed~\citep{kolda2009tensor,acar2010scalable}, revitalizing classical methods such as CANDECOMP/PARAFAC (CP) decomposition~\citep{hitchcock1927expression} and Tucker decomposition~\citep{tucker1966some}.
+These tensor methods treat the entries as continuous-valued.
 In many cases, however, we encounter datasets of which the entries are qualitative.
@@
 \begin{algorithm}[t]
-        \caption{Ordinal tensor decomposition and completion}\label{alg}
+        \caption{Ordinal tensor decomposition}\label{alg}
         \begin{algorithmic}[]
@@
 
-Before concluding this section, we comment on two implementation details.
+We comment on two implementation details before concluding this section.
 First, the problem~\eqref{eq:estimator} is non-convex, so Algorithm 1 usually has no theoretical guarantee on global optimality.
@@
 Interestingly, a larger estimation error is observed when the signal is either too small or too large.
-The non-monotonic behavior may seem surprising, but this is in fact an intrinsic feature in the estimation with ordinal data.
+The non-monotonic behavior may seem surprising, but this is an intrinsic feature in the estimation with ordinal data.
 In view of the latent-variable interpretation (see Section~\ref{sec:latent}), estimation from ordinal observation can be interpreted as an inverse problem of quantization.
-Therefore, the estimation error will diverge when the noise $\tE$ is negligible, because there will be no way to distinguish two different signal tensors, e.g., $\Theta_1=\ma_1\otimes \ma_2\otimes \ma_3$ and $\Theta_2=\text{sign}(\ma_1)\otimes \text{sign}(\ma_2)\otimes \text{sign}(\ma_3)$, from the quantized observations.
+Therefore, the estimation error diverges in the absence of noise $\tE$, because it is impossible to distinguish two different signal tensors, e.g., $\Theta_1=\ma_1\otimes \ma_2\otimes \ma_3$ and $\Theta_2=\text{sign}(\ma_1)\otimes \text{sign}(\ma_2)\otimes \text{sign}(\ma_3)$, from the quantized observations.
 This phenomenon is clearly contrary to the classical continuous-valued tensor problem~\cite{davenport2014,sur2018modern}.
@@
 Figure~\ref{fig:finite}d evaluates the impact of ordinal levels $L$ to estimation accuracy, under the setting $\rho=0.5$.
-It is seen that the performance improves as $L$ grows, especially from binary observations ($L=2$) to multi-level ordinal observations ($L\geq 3$).
+An improved performance is observed as $L$ grows, especially from binary observations ($L=2$) to multi-level ordinal observations ($L\geq 3$).
 The result showcases the benefit of multi-level observations provide compared to binary observations.
@@
 
-\begin{itemize}[itemsep=0pt,topsep=-2pt,leftmargin=*]
+\begin{itemize}[itemsep=-.5pt,topsep=-.5pt,leftmargin=*,partopsep=-.5pt]
 \item Continuous tensor decomposition ({\bf Continuous-T})~\cite{filipovic2015tucker,oh2018scalable} is the classical Tucker approximation method. 
@@
 \item Ordinal matrix completion ({\bf Ordinal-M})~\cite{bhaskar2016probabilistic} is a rank-constrained matrix estimation method based on noisy, quantized observations. 
-\vspace{-.4cm}
 \end{itemize}
@@
 The problem size we considered is comparable to~\citet{ghadermarzy2018learning}.
-It is seen that our method outperforms the others in both MAD and MCR.
+We find that our method outperforms the others in both MAD and MCR.
 In particular, methods built on multi-level observations ({\bf Ordinal-T}, {\bf Ordinal-M}, {\bf 1bit-category-T}) exhibit stable MCR over $\rho$ and $L$, whereas the others two methods ({\bf Continuous-T}, {\bf 1bit-sign-T}) generally fail except for $L=2$ (Figures~\ref{fig:compare}a-b).
@@
 }\label{fig:compare}
-\vspace{-.4cm}
+\vspace{-.5cm}
 \end{figure}
 
-We also compare the methods by their abilities in predicting the median score, $y_\omega^{\text{median}}=\min\{\ell\colon\mathbb{P}(y_\omega=\ell)\geq 0.5\}$.
-Note that, under the latent variable model~\eqref{eq:latent} and Assumption~\ref{ass:link}, the median score can be directly interpreted as the quantized $\Theta$ without noise; i.e.\ $y_\omega^{\text{median}}=\sum_\ell \mathds{1}_{\theta_\omega\in(b_{\ell-1},b_\ell]}$.
-Similar results demonstrate the outperformance of our method (see the Supplement and Figure S1).
+We also compare the methods by their abilities in predicting the median level, $y_\omega^{\text{median}}=\min\{\ell\colon\mathbb{P}(y_\omega=\ell)\geq 0.5\}$.
+Under the latent variable model~\eqref{eq:latent} and Assumption~\ref{ass:link}, the median level is the quantized $\Theta$ without noise; i.e.\ $y_\omega^{\text{median}}=\sum_\ell \mathds{1}_{\theta_\omega\in(b_{\ell-1},b_\ell]}$.
+Similar results demonstrate the outperformance of our method (see the Supplement).
 
@@
 Table~\ref{table:CV} summarizes the prediction error via 5-fold stratified cross-validation averaged over 10 runs.
-It is clearly seen that our method outperform the others especially in MAD.
+Our method outperform the others, especially in MAD.
 
@@
 }
+\vspace{-.2cm}
 \caption{Comparison of prediction error in the HPC and InCarMusic analyses. Standard errors are reported in parentheses.}\label{table:CV} 
-\vspace{-.4cm}
+\vspace{-.3cm}
 \end{table}
@@
 \subsection{InCarMusic recommendation system}
-In the second application, we apply ordinal tensor completion to a recommendation system {\it InCarMusic}.
+We apply ordinal tensor completion to a recommendation system {\it InCarMusic}.
 {\it InCarMusic} is a mobile application that offers music recommendation to passengers of cars based on contexts~\cite{baltrunas2011incarmusic}.
-The data tensor consists of partially observed ratings from scale 1 to 5 of 42 users to 139 songs on 26 contexts.
 Our goal is to perform tensor completion to impute the unobserved entries in the $42\times 139\times 26$ ordinal tensor and thereby we can offer context-specific music recommendation to users.
-Table~\ref{table:CV} shows the averaged prediction error via 5 fold cross validation.
-Again, our method achieves the best prediction error among the three.
-The data tensor has 98\% missing rate which makes the accurate classification challenging.
-Nevertheless, our method correctly classifies half of the unseen entries.
+The data tensor consists of 2,884 observed entries.
+Table~\ref{table:CV} shows the averaged prediction error via 5-fold cross validation.
+The high missing rate makes the accurate classification challenging.
+Nevertheless, our method achieves the best performance among the three.
 
@@
 Other directions include the trade-off between (non)convex optimization and statistical/computational efficiency.
-While convex relaxation is a popular approach for matrix-based problem, computing convex regularizer could be NP hard for tensors of order $K\geq 3$~\cite{hillar2013most, friedland2018nuclear}.
+While convex relaxations are popular approach for matrix-based problem, they are often slow in practice~\cite{chen2019non}.
 The interplay between computational efficiency and statistical accuracy in general tensor problems warrants future research.

### 2020-02-05 ICML/submit/ordinalT.tex
update

@@
 \section{Introduction}
-Multidimensional arrays, a.k.a.\ tensors, arise in a variety of applications including recommendation systems~\citep{adomavicius2011context}, social networks~\citep{nickel2011three}, genomics~\citep{wang2017three}, and neuroimaging~\citep{zhou2013tensor}.
-There is a growing need to develop dimension reduction method for analyzing these noisy, high-dimensional datasets.
-The problem of tensor denoising -- which aims to recover a signal tensor from its noisy entries -- has gained increased attention in both theory and applications~\citep{xia2019sup,xie2016multispectral}.
+Multidimensional arrays, a.k.a.\ tensors, arise in a variety of applications including recommendation systems~\citep{adomavicius2011context}, social networks~\citep{nickel2011three}, genomics~\citep{hore2016tensor}, and neuroimaging~\citep{zhou2013tensor}.
+There is a growing need to develop dimension reduction methods for analyzing these noisy, high-dimensional datasets.
+The problem of tensor denoising -- which aims to recover a signal tensor from its noisy entries -- has gained increased attention in theory and applications~\citep{xia2019sup,zeng2019multiway}.
 A related problem, tensor completion, examines the minimum number of entries needed for a consistent recovery~\citep{ghadermarzy2018learning,ghadermarzy2019near}.
@@
 Therefore, the estimation error diverges in the absence of noise $\tE$, because it is impossible to distinguish two different signal tensors, e.g., $\Theta_1=\ma_1\otimes \ma_2\otimes \ma_3$ and $\Theta_2=\text{sign}(\ma_1)\otimes \text{sign}(\ma_2)\otimes \text{sign}(\ma_3)$, from the quantized observations.
-This phenomenon is clearly contrary to the classical continuous-valued tensor problem~\cite{davenport2014,sur2018modern}.
+This phenomenon~\cite{davenport2014,sur2019modern} is clearly contrary to the classical continuous-valued tensor problem.
 
@@
 \begin{itemize}[itemsep=-.5pt,topsep=-.5pt,leftmargin=*,partopsep=-.5pt]
-\item Continuous tensor decomposition ({\bf Continuous-T})~\cite{filipovic2015tucker,oh2018scalable} is the classical Tucker approximation method. 
+\item Continuous tensor decomposition ({\bf Continuous-T})~\cite{filipovic2015tucker,acar2010scalable} is the classical Tucker approximation method. 
 \item One-bit tensor completion ({\bf 1bit-T})~\cite{ghadermarzy2018learning} is a max-norm penalized tensor learning method based on partial binary observations. 

### 2020-02-06 ICML/submit/ordinalTv2.tex
Add files via upload

--- 2020-02-05 ICML/submit/ordinalT.tex
@@
 Multidimensional arrays, a.k.a.\ tensors, arise in a variety of applications including recommendation systems~\citep{adomavicius2011context}, social networks~\citep{nickel2011three}, genomics~\citep{hore2016tensor}, and neuroimaging~\citep{zhou2013tensor}.
-There is a growing need to develop dimension reduction methods for analyzing these noisy, high-dimensional datasets.
+There is a growing need to develop general methods for analyzing these noisy, high-dimensional datasets that can handle two main problems.
 The problem of tensor denoising -- which aims to recover a signal tensor from its noisy entries -- has gained increased attention in theory and applications~\citep{xia2019sup,zeng2019multiway}.
@@
 Another example is in the signal processing, where the digits are frequently rounded or truncated so that only integer values are available.
-The qualitative observations take values in a limited set of categories, making the learning problem harder compared to the finer, continuous observations.
+The qualitative observations take values in a limited set of categories, making the learning problem harder compared to  continuous observations.
 
-The analyses of ordinal tensors are complicated by several factors.
  Ordinal entries are categorical variables with an ordering among the categories; for example, {\it very like} $\prec$ {\it like} $\prec$ {\it neutral} $\prec \cdots$.
-A reasonable ordinal tensor model has two key properties.
-First, the model should be invariant under a reversal of categories, say, {\it very like} $\succ$ {\it like} $\succ$ {\it neutral} $\succ \cdots$, but not under arbitrary label permutations.
+The analyses of tensors with the ordinal entries are mainly complicated by two key properties needed for a reasonable model.
+First, the model should be invariant under a reversal of categories, say from the example, {\it very like} $\succ$ {\it like} $\succ$ {\it neutral} $\succ \cdots$, but not under arbitrary label permutations.
 Second, the parameter interpretations should be consistent under merging or splitting of contiguous categories.
 The classical continuous tensor model~\cite{kolda2009tensor, ghadermarzy2019near} fails in the first aspect, whereas the binary tensor model~\cite{ghadermarzy2018learning} lacks the second property.
-An appropriate model for ordinal tensors has yet to be studied.
+An appropriate model for ordinal tensors has not yet to be studied.
 
@@
 We propose a cumulative link model for higher-order tensors, develop a rank-constrained M-estimator, and obtain theoretical accuracy guarantees.
-The mean squared error is established, and we show that the obtained rate is minimax optimal in high dimensions under the low-rank model.
+The upper bound for the mean squared error is established, and we show that the obtained bound has minimax optimal rate in high dimensions under the low-rank model.
 Our estimator enjoys a faster convergence rate $\tO(d^{-(K-1)/2})$ than $\tO(d^{-K})$ in~\citet{ghadermarzy2018learning}, which is a substantial improvement as the order $K$ increases.
@@
 Matrix completion from quantized samples was firstly introduced for binary observations~\cite{cai2013max,davenport2014,bhaskar20151} and then extended to ordinal observations~\cite{bhaskar2016probabilistic}.
-As we show in Section~\ref{sec:theory}, applying existing matrix methods to an ordinal tensor will result in a suboptimal estimator with a slower convergence rate.
+As we show in Section~\ref{sec:theory}, applying existing matrix methods to an ordinal tensor results in a suboptimal estimator with a slower convergence rate.
 Therefore, a full exploitation of the tensor structure is necessary; this is the focus of the current paper.
@@
 Existing work focuses exclusively on univariate observations such as binary- or continuous-valued entries~\citep{wang2018learning,hong2018generalized}.
-As we mentioned earlier, the ordinal observations adds considerable challenges to the model formulation.
+As we mentioned earlier, the ordinal observations add considerable challenges to the model formulation.
+We handle the problems from two perspectives.
 From statistical perspective, our proposed model generalizes the usual binary tensor model while preserving palindromic invariance~\cite{mccullagh1980regression} for ordinal observations.
@@
 \subsection{Latent-variable interpretation}\label{sec:latent}
-We show that the ordinal tensor model~\eqref{eq:model} is equivalent to an $L$-level quantization model on $\tY=\entry{y_\omega}$:
+The ordinal tensor model~\eqref{eq:model} with certain types of link $f$ has the equivalent representation as an $L$-level quantization model on $\tY=\entry{y_\omega}$:
 \begin{align}\label{eq:quantization}
@@
 
-The equivalence between the latent-variable model~\eqref{eq:quantization} and the cumulative link model~\eqref{eq:model} is established if the link $f$ is chosen to be athe cumulative distribution function of noise $\varepsilon$, i.e., $f(\theta)=\mathbb{P}(\varepsilon\leq \theta)$.
+The equivalence between the latent-variable model~\eqref{eq:quantization} and the cumulative link model~\eqref{eq:model} is established if the link $f$ is chosen to be the cumulative distribution function of noise $\varepsilon$, i.e., $f(\theta)=\mathbb{P}(\varepsilon\leq \theta)$.
 We describe two common choices of link $f$, or equivalently, the distribution of $\varepsilon$.
@@
 
-(P1) Given the quantization process induced by $f$ and the cut-off points $\mb$, how accurate can we estimate the latent signal tensor $\Theta$ from the ordinal observation $\tY$?
+(P1) Given the quantization process induced by $f$ and the cut-off points $\mb$, how accurately can we estimate the latent signal tensor $\Theta$ from the ordinal observation $\tY$?
 
@@
 Some notation of low-rankness may lead to mathematically ill-defined optimization; for example, the best low CP-rank tensor approximation may not exist~\cite{de2008tensor}.
-We choose Tucker representation for parsimony and easy interpretation.
+We choose Tucker representation for well definedness of optimization and easy interpretation.
 
@@
 The adaptation of unknown $\mb$ is addressed in Section~\ref{sec:algorithm} and the Supplement.
-Similar treatment applies to $\mr$.
 
@@
 \subsection{Estimation error for tensor denoising}\label{sec:denosing}
-For the tensor denoising problem, we assume the full set of tensor entries are observed. We assess the estimation accuracy using the mean squared error (MSE):
+For the tensor denoising problem, we assume that the full set of tensor entries are observed. We assess the estimation accuracy using the mean squared error (MSE):
 \[
@@
 \]
-The next theorem establishes the upper bound for the proposed $\hat \Theta$ in~\eqref{eq:estimator}.
+The next theorem establishes the upper bound for the MSE of the proposed $\hat \Theta$ in~\eqref{eq:estimator}.
 
@@
 In fact, the proof of this theorem (see the Supplement) shows that the same statistical rate holds, not only for the global optimizer~\eqref{eq:estimator}, but also for any local optimizer $\check \Theta$ in the level set $\{\check\Theta\in\tP: \logl(\check\Theta)\geq \logl(\trueT)\}$.
-This suggests that the local optimality itself is not necessarily a severe concern in our context, as long as the objective value upon convergence is large enough.
-In Sections~\ref{sec:algorithm} and~\ref{sec:experiment}, we perform empirical studies to assess the algorithmic stability and its impact to statistical efficiency.
+This suggests that the local optimality itself is not necessarily a severe concern in our context, as long as the convergent objective is large enough.
+In Section ~\ref{sec:algorithm}, we perform empirical studies to assess the algorithmic stability.
 
@@
 The convergence rate for rank-constrained matrix estimation was $\tO(1/\sqrt{d})$~\citep{bhaskar2016probabilistic}, which fits into our special case when $K=2$.
-Furthermore, our results~\eqref{eq:rate} and~\eqref{eq:KLrate} reveal that the convergence becomes favorably as the order of data tensor increases.
+Furthermore, our results~\eqref{eq:rate} and~\eqref{eq:KLrate} reveal that the convergence becomes favorable as the order of data tensor increases.
 Intuitively, the sample size for tensor data analysis is the number of entries, $\prod_k d_k$, and the number of free parameters is roughly on the order of $\sum_{k}d_k$, assuming $r_{\max}=\tO(1)$.
@@
 Note that the recovery error depends on the distribution $\Pi$.
-In particular, tensor entries with higher sampling probabilities are equipped with better accuracy guarantee, compared to the ones with lower sampling probabilities.
+In particular, tensor entries with higher sampling probabilities have more influence on the recovery accuracy, compared to the ones with lower sampling probabilities.
 
@@
 \end{thm}
-Theorem~\ref{thm:completion} shows that our estimator achieves consistent recovery using as few as $\tO(Kd)$ noisy, quantized observations from an order-$K$ $(d,\ldots,d)$-dimensional tensor.
+Theorem~\ref{thm:completion} shows that our estimator achieves consistent recovery using as few as $\tO(Kd)$ {\color{red}($\tO(kd^{1+\epsilon})$?)} noisy, quantized observations from an order-$K$ $(d,\ldots,d)$-dimensional tensor.
 Note that $\tO(Kd)$ matches the degree of freedom for an order-$K$ tensor of fixed rank $\mr$, suggesting the optimality of our sample requirement.
@@
 Existing methods that achieve $\tO(Kd)$ sample complexity require either a deterministic cross sampling design~\cite{zhang2019cross} or univariate measurements~\cite{ghadermarzy2018learning}.
-Our method extends the conclusion to multi-level measurements under a broader class of sampling schemes.
+Our method extends the conclusions to multi-level measurements under a broader class of sampling schemes.
 
@@
 \centering
-\vspace{-.5cm}
+\vspace{-.4cm}
 \includegraphics[width=8cm]{algorithm.pdf}\label{fig:stability}
@@
 It is seen that a higher rank leads to a larger error, as reflected by the upward shift of the curve as $r$ increases.
-Indeed, a higher rank implies a higher intrinsic dimension, thus increasing the difficulty of the estimation.
+Indeed, a higher rank implies the higher number of parameters to estimate, thus increasing the difficulty of the estimation.
 Figure~\ref{fig:finite}b shows the estimation error versus the signal level under $d=20$.
@@
 An improved performance is observed as $L$ grows, especially from binary observations ($L=2$) to multi-level ordinal observations ($L\geq 3$).
-The result showcases the benefit of multi-level observations provide compared to binary observations.
+The result showcases the benefit of multi-level observations compared to binary observations.
 
@@
 \begin{itemize}[itemsep=-.5pt,topsep=-.5pt,leftmargin=*,partopsep=-.5pt]
-\item Continuous tensor decomposition ({\bf Continuous-T})~\cite{filipovic2015tucker,acar2010scalable} is the classical Tucker approximation method. 
+\item Continuous tensor decomposition ({\bf Continuous-T})~\cite{acar2010scalable} is a low-rank approximation method based on classical Tucker model.
 \item One-bit tensor completion ({\bf 1bit-T})~\cite{ghadermarzy2018learning} is a max-norm penalized tensor learning method based on partial binary observations.
@@
 The {\bf 1bit-T} is applied to $\tY$ in two ways.
-The first approach ({\bf 1bit-sign-T}) follows from~\citet{ghadermarzy2018learning} that transforms $\tY$ to a binary tensor of equal size, by taking the entrywise sign of the mean-adjusted tensor, $\tY- |\Omega|^{-1}\sum_\omega y_\omega$.
+The first approach ({\bf 1bit-sign-T}) follows from~\citet{ghadermarzy2018learning} that transforms
+$\tY$ to a binary tensor, by taking the entrywise sign of the mean-adjusted tensor, $\tY- |\Omega|^{-1}\sum_\omega y_\omega$.
 The second approach ({\bf 1bit-category-T}) transforms the order-3 ordinal tensor $\tY$ to an order-4 binary tensor $\tY^{\sharp}=\entry{y^{\sharp}_{ijkl}}$ via dummy variable encoding; i.e., $y^{\sharp}_{ijk\ell}=\mathds{1}_{\{y_{ijk}=\ell\}}$ for all $\ell\in[L-1]$.
-We evaluate the methods by their capabilities in predicting the most frequent label for each entry, i.e.,\ $y_\omega^{\text{mode}}=\arg\max_\ell\mathbb{P}(y_\omega=\ell)$.
-Two performance metrics are considered: mean absolute deviation, $\text{MAD}=d^{-K}\sum_\omega |y_\omega^{\text{mode}}-\hat y_\omega^{\text{mode}}|$, and misclassification rate, $\text{MCR}=d^{-K}\sum_\omega|y_\omega^{\text{mode}}-\text{round}(\hat y_\omega^{\text{mode}})|$, where $\text{round}(\cdot)$ denotes the nearest integer of the (possibly continuous-valued) prediction, e.g.
-returned by {\bf Continuous-T}.
-Note that MAD penalizes the large deviation more heavily than does MCR.
+We evaluate the methods by their capabilities in predicting the highest probability label for each entry, i.e.,\ $y_\omega^{\text{mode}}=\arg\max_\ell\mathbb{P}(y_\omega=\ell)$.
+Two performance metrics are considered: mean absolute deviation, $\text{MAD}=d^{-K}\sum_\omega |y_\omega^{\text{mode}}-\hat y_\omega^{\text{mode}}|$, and misclassification rate, $\text{MCR}=d^{-K}\sum_\omega\mathds{1}_{y_\omega^{\text{mode}}\neq\text{round}(\hat y_\omega^{\text{mode}})}$,
+where $\text{round}(\cdot)$ denotes the nearest integer of the  prediction (possibly continuous-valued returned by {\bf Continuous-T}).
+Note that MAD penalizes the large deviation more heavily than MCR does.
 
@@
 \vspace{-.4cm}
-\caption{Performance comparison in MSE (a-b) and MAD (c-d).
-(a, b) Prediction errors versus the number of ordinal levels $L$ when $\rho=0.8$.
-(c, d) Prediction errors versus sample complexity $\rho=|\Omega|/d^K$ when $L=5$.
+\caption{Performance comparison in MCR (a, b) and MAD (c, d).
+(b, d)   Prediction errors versus the number of ordinal levels $L$ when $\rho=0.8$.
+(a, c) Prediction errors versus sample complexity $\rho=|\Omega|/d^K$ when $L=5$.
 }\label{fig:compare}
@@
 
-We also compare the methods by their abilities in predicting the median level, $y_\omega^{\text{median}}=\min\{\ell\colon\mathbb{P}(y_\omega=\ell)\geq 0.5\}$.
-Under the latent variable model~\eqref{eq:latent} and Assumption~\ref{ass:link}, the median level is the quantized $\Theta$ without noise; i.e.\ $y_\omega^{\text{median}}=\sum_\ell \mathds{1}_{\theta_\omega\in(b_{\ell-1},b_\ell]}$.
+We also compare the methods by their performance in predicting the median lebel, $y_\omega^{\text{median}}=\min\{\ell\colon\mathbb{P}(y_\omega=\ell)\geq 0.5\}$.
+Under the latent variable model~\eqref{eq:latent} and Assumption~\ref{ass:link}, the median lebel is the quantized $\Theta$ without noise; i.e.\ $y_\omega^{\text{median}}=\sum_\ell \mathds{1}_{\theta_\omega\in(b_{\ell-1},b_\ell]}$.
 Similar results demonstrate the outperformance of our method (see the Supplement).
@@
 Table~\ref{table:CV} summarizes the prediction error via 5-fold stratified cross-validation averaged over 10 runs.
-Our method outperform the others, especially in MAD.
+Our method outperforms the others, especially in MAD.
 

### 2020-02-07 ICML/submit/ordinalTv3.tex
final version

--- 2020-02-06 ICML/submit/ordinalTv2.tex
@@
 Our mean squared error bound enjoys a faster convergence rate than previous results, and we show that the proposed estimator is minimax optimal under the class of low-rank models.
-Furthermore, the procedure developed serves as an efficient completion method which guarantees consistent recovery of an order-$K$ $(d,\ldots,d)$-dimensional low-rank tensor using only $\tO(Kd)$ noisy, quantized observations.
+Furthermore, the procedure developed serves as an efficient completion method which guarantees consistent recovery of an order-$K$ $(d,\ldots,d)$-dimensional low-rank tensor using only $\tilde \tO(Kd)$ noisy, quantized observations.
 We demonstrate the outperformance of our approach over previous methods on the tasks of clustering and collaborative filtering.
@@
 \section{Introduction}
-Multidimensional arrays, a.k.a.\ tensors, arise in a variety of applications including recommendation systems~\citep{adomavicius2011context}, social networks~\citep{nickel2011three}, genomics~\citep{hore2016tensor}, and neuroimaging~\citep{zhou2013tensor}.
+Multidimensional arrays, a.k.a.\ tensors, arise in a variety of applications including recommendation systems~\citep{baltrunas2011incarmusic}, social networks~\citep{nickel2011three}, genomics~\citep{hore2016tensor}, and neuroimaging~\citep{zhou2013tensor}.
 There is a growing need to develop general methods for analyzing these noisy, high-dimensional datasets that can handle two main problems.
@@
 In many cases, however, we encounter datasets of which the entries are qualitative.
-For example, the Netflix problem records the ratings of users to movies over weeks.
+For example, the Netflix problem records the ratings of users on movies over time.
 Each data is a rating on a nominal scale \{{\it very like, like, neutral, dislike, very dislike}\}.
@@
 The analyses of tensors with the ordinal entries are mainly complicated by two key properties needed for a reasonable model.
-First, the model should be invariant under a reversal of categories, say from the example, {\it very like} $\succ$ {\it like} $\succ$ {\it neutral} $\succ \cdots$, but not under arbitrary label permutations.
+First, the model should be invariant under a reversal of categories, say, from the Netflix example, {\it very like} $\succ$ {\it like} $\succ$ {\it neutral} $\succ \cdots$, but not under arbitrary label permutations.
 Second, the parameter interpretations should be consistent under merging or splitting of contiguous categories.
 The classical continuous tensor model~\cite{kolda2009tensor, ghadermarzy2019near} fails in the first aspect, whereas the binary tensor model~\cite{ghadermarzy2018learning} lacks the second property.
-An appropriate model for ordinal tensors has not yet to be studied.
+An appropriate model for ordinal tensors has yet to be studied.
 
@@
 We propose a cumulative link model for higher-order tensors, develop a rank-constrained M-estimator, and obtain theoretical accuracy guarantees.
-The upper bound for the mean squared error is established, and we show that the obtained bound has minimax optimal rate in high dimensions under the low-rank model.
+The mean squared error bound is established, and we show that the obtained bound has minimax optimal rate in high dimensions under the low-rank model.
 Our estimator enjoys a faster convergence rate $\tO(d^{-(K-1)/2})$ than $\tO(d^{-K})$ in~\citet{ghadermarzy2018learning}, which is a substantial improvement as the order $K$ increases.
-Furthermore, our proposal serves as an efficient completion algorithm that guarantees consistent recovery of an order-$K$ $(d,\ldots,d)$-dimensional low-rank tensor using only $\tO(Kd)$ noisy, quantized observations.
+Furthermore, our proposal serves as an efficient completion algorithm that guarantees consistent recovery of an order-$K$ $(d,\ldots,d)$-dimensional low-rank tensor using only $\tilde \tO(Kd)$ noisy, quantized observations.
 
@@
 Our work is also connected to non-Gaussian tensor decomposition.
-Existing work focuses exclusively on univariate observations such as binary- or continuous-valued entries~\citep{wang2018learning,hong2018generalized}.
+Existing work focuses exclusively on univariate observations such as binary- or continuous-valued entries~\citep{wang2018learning,hong2018generalized,ghadermarzy2018learning}.
 As we mentioned earlier, the ordinal observations add considerable challenges to the model formulation.
-We handle the problems from two perspectives.
+We address the problems from two perspectives.
 From statistical perspective, our proposed model generalizes the usual binary tensor model while preserving palindromic invariance~\cite{mccullagh1980regression} for ordinal observations.
@@
 
+\vspace{-.2cm}
 \section{Model formulation and motivation}
@@
 Note that, unlike matrices, there are various notations of tensor low-rankness, such as CP rank~\cite{hitchcock1927expression} and train rank~\cite{oseledets2011tensor}.
-Some notation of low-rankness may lead to mathematically ill-defined optimization; for example, the best low CP-rank tensor approximation may not exist~\cite{de2008tensor}.
-We choose Tucker representation for well definedness of optimization and easy interpretation.
+Some notation of low-rankness may lead to mathematically ill-posed optimization; for example, the best low CP-rank tensor approximation may not exist~\cite{de2008tensor}.
+We choose Tucker representation for well-posedness of optimization and easy interpretation.
 
@@
 \end{thm}
-Theorem~\ref{thm:completion} shows that our estimator achieves consistent recovery using as few as $\tO(Kd)$ {\color{red}($\tO(kd^{1+\epsilon})$?)} noisy, quantized observations from an order-$K$ $(d,\ldots,d)$-dimensional tensor.
-Note that $\tO(Kd)$ matches the degree of freedom for an order-$K$ tensor of fixed rank $\mr$, suggesting the optimality of our sample requirement.
-This sample complexity substantially improves over earlier result $\tO(d^{\lceil K/2\rceil})$ based on square matrixization~\cite{mu2014square}, or $\tO(d^{N/2})$ based on tensor nuclear-norm regularization~\cite{yuan2016tensor}.
-Existing methods that achieve $\tO(Kd)$ sample complexity require either a deterministic cross sampling design~\cite{zhang2019cross} or univariate measurements~\cite{ghadermarzy2018learning}.
+Theorem~\ref{thm:completion} shows that our estimator achieves consistent recovery using as few as $\tilde{\tO}(Kd)$ noisy, quantized observations from an order-$K$ $(d,\ldots,d)$-dimensional tensor.
+Note that $\tilde \tO(Kd)$ roughly matches the degree of freedom for an order-$K$ tensor of fixed rank $\mr$, suggesting the optimality of our sample requirement.
+This sample complexity substantially improves over earlier result $\tO(d^{\lceil K/2\rceil})$ based on square matricization~\cite{mu2014square}, or $\tO(d^{N/2})$ based on tensor nuclear-norm regularization~\cite{yuan2016tensor}.
+Existing methods that achieve $\tilde \tO(Kd)$ sample complexity require either a deterministic cross sampling design~\cite{zhang2019cross} or univariate measurements~\cite{ghadermarzy2018learning}.
 Our method extends the conclusions to multi-level measurements under a broader class of sampling schemes.
 
+\vspace{-.2cm}
 \section{Numerical Implementation}\label{sec:algorithm}
@@
             \STATE{\bfseries Output:} $(\hat\Theta,\hat{\mb}) =  \argmax_{(\Theta,\mb)\in \mathcal{P}\times\mathcal{B}}  \mathcal{L}_{\mathcal{Y},\Omega}(\Theta,\mb).$
-            \STATE Initialize core tensor $\mathcal{C}^{(0)}$, factor matrices $ \{\mM_k^{(0)}\}$, and cut-off points $\mb^{(0)}$.
+            \STATE Random initialization of core tensor $\mathcal{C}^{(0)}$, factor matrices $ \{\mM_k^{(0)}\}$, and cut-off points $\mb^{(0)}$.
             \FOR{$t = 1,2,\cdots,$}
@@
               \newline
-              $\mb^{(t+1)} \gets \argmax_{\mb\in \mathbb{R}^{L-1}}\mathcal{L}_{\mathcal{Y},\Omega}\big(\Theta^{(t+1)},\mb\big).$
+              $\mb^{(t+1)} \gets \argmax_{\mb\in \tB}\mathcal{L}_{\mathcal{Y},\Omega}\big(\Theta^{(t+1)},\mb\big).$
             \ENDFOR
@@
  \end{algorithm}
-
 We comment on two implementation details before concluding this section.
@@
 Nevertheless, as shown in Section~\ref{sec:denosing}, the desired rate holds not only for the global optimizer, but also for the local optimizer with $\logl(\hat \Theta)\geq \logl(\trueT)$.
-In practice, we find the convergence point $\hat \Theta$ is often satisfactory, in that the corresponding objective $\logl(\hat \Theta)$ is close to and actually slightly larger than the objective evaluated at the true parameter $\logl(\trueT)$.
+In practice, we find the convergence point $\hat \Theta$ upon random initialization is often satisfactory, in that the corresponding objective $\logl(\hat \Theta)$ is close to and actually slightly larger than the objective evaluated at the true parameter $\logl(\trueT)$.
 Figure~\ref{fig:stability} shows the trajectory of the objective function that is output in the default setting of Algorithm 1, with the input tensor generated from probit model~\eqref{eq:model} with $d_1=d_2=d_3=d$ and $r_1=r_2=r_3=r$.
@@
 \caption{Trajectory of objective function with various $d$ and $r$.}
-\vspace{-.7cm}
+\vspace{-.6cm}
 \end{figure}
@@
 
+\vspace{-.2cm}
 \section{Experiments}\label{sec:experiment}
@@
 
+\vspace{-.1cm}
 \subsection{Finite-sample performance}\label{sec:simulation}
-
+\vspace{-.1cm}
 The first experiment examines the performance under complete observations.
@@
 
+\vspace{-.1cm}
 \subsection{Comparison with alternative methods}
+\vspace{-.1cm}
 Next, we compare our ordinal tensor method ({\bf Ordinal-T}) with three popular low-rank methods:
@@
 The second approach ({\bf 1bit-category-T}) transforms the order-3 ordinal tensor $\tY$ to an order-4 binary tensor $\tY^{\sharp}=\entry{y^{\sharp}_{ijkl}}$ via dummy variable encoding; i.e., $y^{\sharp}_{ijk\ell}=\mathds{1}_{\{y_{ijk}=\ell\}}$ for all $\ell\in[L-1]$.
-We evaluate the methods by their capabilities in predicting the highest probability label for each entry, i.e.,\ $y_\omega^{\text{mode}}=\arg\max_\ell\mathbb{P}(y_\omega=\ell)$.
-Two performance metrics are considered: mean absolute deviation, $\text{MAD}=d^{-K}\sum_\omega |y_\omega^{\text{mode}}-\hat y_\omega^{\text{mode}}|$, and misclassification rate, $\text{MCR}=d^{-K}\sum_\omega\mathds{1}_{y_\omega^{\text{mode}}\neq\text{round}(\hat y_\omega^{\text{mode}})}$,
+We evaluate the methods by their capabilities in predicting the most likely label for each entry, i.e.,\ $y_\omega^{\text{mode}}=\arg\max_\ell\mathbb{P}(y_\omega=\ell)$.
+Two performance metrics are considered: mean absolute deviation, $\text{MAD}=d^{-K}\sum_\omega |y_\omega^{\text{mode}}-\hat y_\omega^{\text{mode}}|$, and misclassification rate, $\text{MCR}=d^{-K}\sum_\omega\mathds{1}_{\{y_\omega^{\text{mode}}\neq\text{round}(\hat y_\omega^{\text{mode}})\}}$,
 where $\text{round}(\cdot)$ denotes the nearest integer of the prediction (possibly continuous-valued returned by {\bf Continuous-T}).
-Note that MAD penalizes the large deviation more heavily than MCR does.
+Note that MAD penalizes the large deviation more heavily than MCR.
 
@@
 }\label{fig:compare}
-\vspace{-.5cm}
+\vspace{-.6cm}
 \end{figure}
 
-We also compare the methods by their performance in predicting the median lebel, $y_\omega^{\text{median}}=\min\{\ell\colon\mathbb{P}(y_\omega=\ell)\geq 0.5\}$.
-Under the latent variable model~\eqref{eq:latent} and Assumption~\ref{ass:link}, the median lebel is the quantized $\Theta$ without noise; i.e.\ $y_\omega^{\text{median}}=\sum_\ell \mathds{1}_{\theta_\omega\in(b_{\ell-1},b_\ell]}$.
+We also compare the methods by their performance in predicting the median label, $y_\omega^{\text{median}}=\min\{\ell\colon\mathbb{P}(y_\omega=\ell)\geq 0.5\}$.
+Under the latent variable model~\eqref{eq:latent} and Assumption~\ref{ass:link}, the median label is the quantized $\Theta$ without noise; i.e.\ $y_\omega^{\text{median}}=\sum_\ell \mathds{1}_{\theta_\omega\in(b_{\ell-1},b_\ell]}$.
 Similar results demonstrate the outperformance of our method (see the Supplement).
 
+\vspace{-.2cm}
 \section{Data Applications}
+\vspace{-.1cm}
 We apply our ordinal tensor method to two real-world datasets.
@@
 
+\vspace{-.1cm}
 \subsection{Human Connectome Project (HCP)}
+\vspace{-.1cm}
 Each entry in the HCP dataset takes value on a nominal scale, \{{\it high, moderate, low}\}, indicating the strength level of fiber connection.
@@
 
+\vspace{-.1cm}
 \subsection{InCarMusic recommendation system}
+\vspace{-.1cm}
 We apply ordinal tensor completion to a recommendation system {\it InCarMusic}.
@@
 
+\vspace{-.2cm}
 \section{Conclusion}
+\vspace{-.1cm}
 We have developed a low-rank tensor estimation method based on possibly incomplete, ordinal-valued observations.

### 2020-02-11 ICML/arxiv/ordinalT_supp_v3.tex
upload

--- 2020-02-04 ICML/miaoyan/ordinalT_supp_v2.tex
@@
 \usepackage{xr}
-\externaldocument{ordinalT}
+\externaldocument{ordinalTv3}
 \input macros.tex
+\setlength\parindent{0pt}
+\allowdisplaybreaks
 
@@
 Recall that
-\[
-U_\alpha=\max_{\ell\in[L],|\theta|\leq \alpha}{\dot{g}_\ell(\theta)\over g_\ell(\theta)}>0 \quad \text{and}\quad
+\begin{equation}\label{eq:bound}
+U_\alpha=\max_{\ell\in[L],|\theta|\leq \alpha}{|\dot{g}_\ell(\theta)|\over g_\ell(\theta)}>0 \quad \text{and}\quad
 L_\alpha=\min_{\ell\in[L],|\theta|\leq \alpha} {\dot{g}^2_\ell(\theta)-\ddot{g}_\ell(\theta)g_\ell(\theta)\over g^2_\ell(\theta)}>0.
-\]
-Therefore, all entries in $\fplogl$ are upper bounded $U_\alpha>0$, and all diagonal entries in $\fpplogl$ are upper bounded by $-L_{\alpha}<0$.
+\end{equation}
+Therefore, the entries in $\fplogl$ are upper bounded in magnitude by $U_\alpha>0$, and all diagonal entries in $\fpplogl$ are upper bounded by $-L_{\alpha}<0$.
 
@@
 
-\subsection{Sample complexity for tensor completion}
 \begin{proof}[Proof of Theorem~\ref{thm:minimax}]
@@
 \end{proof}
+
+\subsection{Sample complexity for tensor completion}
 
@@
 \end{proof}
+
+\subsection{Convexity of the log-likelihood function}
+\begin{thm}\label{thm:convexity}
+Define the function
+\begin{align}\label{eq:function}
+ \logl(\Theta, \mb)&=\sum_{\omega\in\Omega}\sum_{\ell\in[L]} \left\{\mathds{1}_{\{y_\omega=\ell\}} \log \big[f(\theta_\omega+b_\ell)-  f(\theta_\omega+b_{\ell-1})\big]\right\},
+ \end{align}
+where $f(\cdot)$ satisfies Assumption~\eqref{eq:ass}. Then, $\logl(\Theta, \mb)$ is concave in $(\Theta,\mb)$. 
+\end{thm}
+
+\begin{proof}
+Define $d_{\text{total}}=\prod_k d_k$.
+With a little abuse of notation, we use $(\Theta,\mb)$ to denote a length-$(d_{\text{total}}+L-1)$-vector collecting all parameters together.
+Let us denote a bivariate function
+\begin{align}
+\lambda: &\mathbb{R}^2\mapsto \mathbb{R}\\
+(u,v)&\mapsto\lambda(u,v) = \log \big[f(u)-  f(v)\big].
+\end{align}
+It suffices to show that $\lambda(u,v)$ is concave in $(u,v)$ where $u>v$.
+
+Suppose that the claim holds (which we will prove in the next paragraph). Based on~\eqref{eq:function}, $u,v$ are both linear functions of $(\Theta,\mb)$:
+\[
+u = \ma_1^T(\Theta,\mb), \quad v = \ma_2^T(\Theta,\mb), \quad \text{ for some}\ \ma_1,\ma_2\in \mathbb{R}^{d_{\text{total}}+L-1}.
+\]
+Then, $\lambda(u,v) = \lambda(\ma_1^T(\Theta,\mb),\ \ma_2^T(\Theta,\mb))$ is concave in $(\Theta,\mb)$ by the definition of concavity.
+Furthermore, we can conclude that $\logl(\Theta, \mb) $ is concave in $(\Theta,\mb)$ because $\logl(\Theta, \mb)$ is the sum of $\lambda(u,v)$.
+
+Now, we prove the concavity of $\lambda(u,v)$. Note that
+\begin{equation*}
+\lambda(u,v) = \log\big[f(u)-f(v)\big]=\log\big[\int\mathds{1}_{[u,v]}(x)f'(x)dx\big],
+\end{equation*}
+where $\mathds{1}_{[u,v]}$ is an indicator function that equals 1 in the interval $[u,v]$, and 0 elsewhere.
+Furthermore, $\mathds{1}_{[u,v]}(x)$ is log-concave in $(u,v,x)$ and $f'(x) = \frac{e^x}{(1+e^x)^2}$ is log-concave in $x$.
+The convexities imply that $\mathds{1}_{(u,v)}(x)f'(x)$
+is a log-concave in $(u,v,x)$. By Lemma~\ref{lem:lossconvexity}, we conclude that $\lambda(u,v)$ is concave in $(u,v)$ where $u>v$. 
+\end{proof}
+
+\begin{lem}[Corollary 3.5 in~\cite{brascamp2002extensions}]\label{lem:lossconvexity}
+Let $F(x,y):\mathbb{R}^{m+n}\rightarrow \mathbb{R}$ be an integrable function where $x\in \mathbb{R}^{m},y\in \mathbb{R}^n$. Let 
+\[
+G(x) = \int_{\mathbb{R}^n}F(x,y)dy.
+\]
+If $F(x,y)$ is log concave in $(x,y)$, then $G(x)$ is log concave in $x$.
+\end{lem}
 
@@
 
-\section{Convexity of the likelihood function}
-\begin{thm}\label{thm:convexity}
+\section{Extension of Theorem~\ref{thm:rate} to unknown cut-off points}
+
+When the cut-off points $\mb$ is unknown, we estimate $(\hat \Theta,\hat \mb)$ by
+\begin{equation}\label{eq:joint}
+(\hat \Theta,\hat \mb)=\argmax_{(\Theta,\mb)\in\tP\times \tB}\logl(\Theta,\mb),
+\end{equation}
+where 
+\[
+\tP=\{\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}\colon \text{rank}(\tP)\leq \mr,\ \mnormSize{}{\Theta}\leq \alpha\}, \quad \tB=\{\mb\in\mathbb{R}^{L-1}\colon \mnormSize{}{\mb}\leq \beta,\ \min_{\ell}(b_\ell-b_{\ell-1}) \geq \Delta\}.
+\]
+The estimation accuracy is assessed using the mean squared error (MSE):
+\[
+\text{MSE}(\hat \Theta,\Theta)={1\over \prod_k d_k}\FnormSize{}{\hat \Theta-\Theta},\quad \text{MSE}(\hat \mb,\mb)={1\over L-1}\FnormSize{}{\hat \mb-\mb}.
+\]
+
+We introduce several quantities that will used in our theory:
+\begin{enumerate}
+\item We make the convention that $b_0=-\infty$, $b_L=\infty$, $f(-\infty)=0$, $f(\infty)=1$, and $\dot{f}(-\infty)=\ddot{f}(-\infty)=\dot{f}(\infty)=\ddot{f}(\infty)=0$.
+
+\item The difference function $g_\ell(\theta)$ is defined as $g_\ell(\theta)=f(\theta+b_\ell)-f(\theta+b_{\ell-1})$ for all $\theta\in\mathbb{R}$ and $\ell\in[L]$.
+
+\item Define $n_\ell=\sum_{\omega\in\Omega}\mathds{1}_{\{y_\omega=\ell\}}$, i.e., the number of tensor entries taking value on $\ell\in[L]$.
+
+\item With a little abuse of notation, we re-define the constants in~\eqref{eq:bound} as
+\begin{equation}\label{eq:UL}
+U_{\alpha,\beta,\Delta}=\max_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}}\max_{\ell\in[L-1]}{|\dot{g}_\ell(\theta)|\over g_\ell(\theta)}, \quad \text{and}\quad
+L_{\alpha,\beta,\Delta}=\min_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}}\min_{\ell\in[L-1]} {\dot{g}^2_\ell(\theta)-\ddot{g}_\ell(\theta)g_\ell(\theta)\over g^2_\ell(\theta)}.
+\end{equation}
+\item We define two additional constants:
+\begin{align}\label{eq:CD}
+C_{\alpha,\beta,\Delta} &= \max_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}}  \max_{\ell\in[L-1]}\max\left\{\frac{\dot{f}(\theta+b_\ell)}{g_\ell(\theta)},\ \frac{\dot{f}(\theta+b_\ell)}{g_{\ell+1}(\theta)}\right\},\notag \\
+D_{\alpha,\beta,\Delta} &=\min_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}} \min_{\ell\in[L-1]}\min\left\{ - {\partial \over \partial \theta}\left({\dot{f}(\theta+b_{\ell})\over g_\ell(\theta)}\right),\ {\partial \over \partial \theta}\left({\dot{f}(\theta+b_{\ell})\over g_{\ell+1}(\theta)}\right)\right\}\notag \\
+&=  \min_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}}\min_{\ell\in[L-1]}\min \left\{-\frac{\ddot{f}(\theta+b_\ell)g_\ell(\theta)-\dot{f}(\theta+b_\ell)\dot{g}_\ell(\theta)}{g^2_\ell(\theta)},\ \right.
+\notag \\
+&\hspace{1.5in} \left. \frac{\ddot{f}(\theta+b_\ell)g_{\ell+1}(\theta)-\dot{f}(\theta+b_\ell)\dot{g}_{\ell+1}(\theta)}{g^2_{\ell+1}(\theta)}\right\}.
+\end{align} 
+\end{enumerate}
+
+We make the following assumptions about the link function.  
+\begin{assumption}\label{ass:joint}
+The link function $f\colon \mathbb{R}\mapsto [0,1]$ satisfies the following properties:
+\begin{enumerate}
+\item $f(\theta)$ is twice-differentiable and strictly increasing in $\theta$.
+\item $\dot{f}(\theta)$ is strictly log-concave and symmetric with respect to $\theta=0$. 
+\item The function ${\dot{f}(\theta+b_\ell)\over g_\ell(\theta)}$ is strictly decreasing with respect to $\theta$ for all $\mb\in\tB$. 
+\item The function ${\dot{f}(\theta+b_\ell)\over g_{\ell+1}(\theta)}$ is strictly increasing with respect to $\theta$ for all $\mb\in\tB$. 
+\end{enumerate}
+\end{assumption}
+
+\begin{rmk}
+The condition $\Delta=\min_{\ell}(b_{\ell}-b_{\ell-1})>0$ on the feasible set $\tB$ guarantees the strict positiveness of $g_{\ell}(\theta)=f(\theta+b_{\ell})-f(\theta+b_{\ell-1})$.
+Therefore, the denominators in the above quantities $U_{\alpha,\beta,\Delta}, L_{\alpha,\beta,\Delta}, C_{\alpha,\beta,\Delta}, D_{\alpha,\beta,\Delta}$ are well-defined.
+Furthermore, by Theorem~\ref{thm:convexity}, $g_\ell(\theta)$ is strictly log-concave in $\theta$ for all $\mb\in\tB$.
+Based on Assumption~\ref{ass:joint} and closeness of the feasible set, we have $U_{\alpha,\beta,\Delta}>0$, $L_{\alpha,\beta,\Delta}>0$, $C_{\alpha,\beta,\Delta}>0$, $D_{\alpha,\beta,\Delta}>0$.
+\end{rmk}
+\begin{rmk}
+In particular, for logistic link $f(x) = \frac{1}{1+e^{-x}}$ , we have 
 \begin{align}
- \logl(\Theta, \mb)&=\sum_{\omega\in\Omega}\sum_{\ell\in[L]} \Big\{\mathds{1}_{\{y_\omega=\ell\}} \log \big[f(\theta_\omega+b_\ell)-  f(\theta_\omega+b_{\ell-1})\big]\Big\}, \text{ where } f(x) = \frac{e^x}{1+e^x}.
+C_{\alpha,\beta,\Delta}&=\max_{\substack{|\theta|\leq \alpha,\\ \mb\in\tB}}\max_{\ell\in[L-1]}\max
+\left\{
+{ 1\over 1-e^{-(b_{\ell}-b_{\ell-1})}} { 1+e^{\theta+b_{\ell-1}} \over 1+e^{\theta+b_{\ell}}},\
+{ 1\over e^{(b_{\ell+1}-b_{\ell})}-1} { 1+e^{\theta+b_{\ell+1}} \over 1+e^{\theta+b_{\ell}}}
+\right\}>0,
+\\
+D_{\alpha,\beta,\Delta}&=\min_{\substack{|\theta|\leq \alpha,\\ \mb\in\tB}}\min_{\ell\in[L-1]} {1\over (1+e^{\theta+b_\ell})(1+e^{-(\theta+b_\ell)})}>0.
 \end{align}
-is a concave function to $(\Theta,\mb)$
+\end{rmk}
+
+\begin{thm}[Statistical convergence with unknown $\mb$]\label{thm:ratejoint}
+Consider an ordinal tensor $\tY\in[L]^{d_1\times \cdots \times d_K}$ generated from model~\eqref{eq:model} with the link function $f$ and parameters $(\trueT, \trueb)\in\tP\times \tB$.
+Suppose the link function $f$ satisfies Assumption~\ref{ass:joint}.
+Define $r_{\max}=\max_k r_k$.
+Then with very high probability, the estimator in~\eqref{eq:joint} satisfies
+\[
+\text{MSE}(\hat \Theta, \trueT)\leq {c_1U_{\alpha,\beta,\Delta}\over L_{\alpha,\beta,\Delta}} r_{\max}^{(K-1)/2}  \sqrt{ \sum_k d_k \over \prod_k d_k},
+\]
+and
+\[
+\text{MSE}(\hat\mb, \trueb) \leq  {c_2 C_{\alpha,\beta,\Delta} \over D_{\alpha,\beta,\Delta}}{\max_{\ell}(n_\ell+n_{\ell+1})\over \min_{\ell}(n_{\ell}+n_{\ell+1})}  \sqrt{1\over L-1},
+\]
+where $c_1,c_2>0$ are two constants independent of the tensor dimension and rank, and $U_{\alpha,\beta,\Delta}$, $L_{\alpha,\beta,\Delta}$, $C_{\alpha,\beta,\Delta}$, $D_{\alpha,\beta,\Delta}>0$ are constants defined in~\eqref{eq:UL} and~\eqref{eq:CD}.
 \end{thm}
-\begin{proof}
-We use $(\Theta,\mb)$ abusively as $(Vec(\Theta)^T,\mb^T)^T$ in this proof.
-Let us denote $\lambda(u,v) = \log \big[f(u)-  f(v)\big]$.
-It is enough to show that $\lambda(u,v)$
-is a concave function to $(u,v)$ where $u>v$. Because if $\lambda(u,v)$ is a concave function and $u,v$ are both linear functions of $(\Theta,\mb) $ such that
-$$u = \ma_1^T(\Theta,\mb), v = \ma_2^T(\Theta,\mb) \text{ for some }\ma_1,\ma_2\in \mathbb{R}^{d_{\text{total}}+L-1},$$ then $\lambda(u,v) = \lambda(\ma_1^T(\Theta,\mb),\ma_2^T(\Theta,\mb)))$ is a concave function to $(\Theta,\mb)$
-by the definition of the convexity. Furthermore, we can conclude that $\logl(\Theta, \mb) $ is a concave function to $(\Theta,\mb)$ because $\logl(\Theta, \mb)$
-can be written as the form of summations of $\lambda(u,v)$. To prove the convexity of $\lambda(u,v)$, we have the following.
-\begin{equation*}
-\lambda(u,v) = \log\big[f(u)-f(v)\big]=\log\big[\int\mathds{1}_{(u,v)}(x)f'(x)dx\big].
-\end{equation*}
-Notice that  $\mathds{1}_{(u,v)}(x)$ is log-concave to $(u,v,x)$ and $f'(x)$ is log-concave to $x$ because $f'(x) = \frac{e^x}{(1+e^x)^2}$. Those convexities imply that $\mathds{1}_{(u,v)}(x)f'(x)$
-is a log-concave function to $(u,v,x)$. Therefore, we can conclude that the $\lambda(u,v)$ is a concave function from the lemma~\ref{lem:lossconvexity}.
+
+\begin{rmk} It may seem counter-intuitive that the MSE of $\hat \Theta$ shows no dependence of $\hat \mb$ (and vice versa).
+In fact, from the proof of Theorem~\ref{thm:ratejoint}, the uniform constants $U_{\alpha,\beta,\Delta}, L_{\alpha,\beta,\Delta}$ can be replaced by
+\[
+U_{\alpha,\hat \mb}=\max_{|\theta|\leq \alpha}\max_{\ell\in[L-1]}{|\dot{g}_\ell(\theta)\over g_\ell(\theta)|},\quad \text{and}\quad L_{\alpha,\hat \mb}=\min_{|\theta|\leq \alpha}\min_{\ell\in[L-1]} {\dot{g}^2_\ell(\theta)-\ddot{g}_\ell(\theta)g_\ell(\theta)\over g^2_\ell(\theta)}.
+\]
+Therefore, the accuracy of $\hat \Theta$ depends on the accuracy of $\hat \mb$ implicitly through $U_{\alpha, \hat \mb}$ and $ L_{\alpha,\hat \mb}$.
+
+\end{rmk}
+
+\begin{proof}[Proof of Theorem~\ref{thm:ratejoint}]
+Similar to the proof of Theorem~\ref{thm:rate}, we suppress $\Omega$ in the subscript. Based on the definition of $(\hat \Theta, \hat \mb)$, 
+we have the following inequalities:
+\begin{equation}\label{eq:level}
+\flogl(\hat \Theta,\hat \mb)\geq \flogl(\trueT,\hat \mb)\quad \text{and}\quad \flogl(\hat \Theta,\hat \mb)\geq \flogl(\hat \Theta, \trueb).
+\end{equation}
+
+Following the same argument as in Theorem~\ref{thm:rate} and the first inequality in~\eqref{eq:level}, we obtain that
+\begin{equation}\label{eq:ratenew}
+\text{MSE}(\hat \Theta, \trueT)\leq {CU_{\alpha,\beta,\Delta}\over L_{\alpha,\beta,\Delta}} r_{\max}^{(K-1)/2}  \sqrt{ \sum_k d_k \over \prod_k d_k},
+\end{equation}
+where $U_{\alpha,\beta,\Delta}, L_{\alpha,\beta,\Delta}>0$ are two universal constants independent of the tensor dimension and rank.
+Next we bound $\text{MSE}(\hat \mb,\trueb)$. It follows from the expression of $\flogl(\Theta,\mb)$ that
+\begin{align}\label{eq:property}
+{\partial \flogl\over \partial b_\ell}&=\sum_{\omega\in\Omega}\left[\mathds{1}_{\{y_{\omega}=\ell\}}
+\frac{\dot{f}(\theta_\omega+b_\ell)}{g_\ell(\theta_\omega)}-\mathds{1}_{\{y_{\omega}=\ell+1\}}
+\frac{\dot{f}(\theta_\omega+b_\ell)}{g_{\ell+1}(\theta_\omega)}\right],\notag\\
+{\partial^2 \flogl\over \partial b_\ell^2}&=\sum_{\omega\in\Omega}\left[\mathds{1}_{\{y_\omega=\ell\}}\frac{\ddot{f}(\theta_\omega+b_\ell)g_\ell(\theta_\omega)-\dot{f}(\theta_\omega+b_\ell)^2}{g^2_\ell(\theta_\omega)}-\mathds{1}_{\{y_\omega=\ell+1\}}\frac{\ddot{f}(\theta_\omega+b_\ell)g_{\ell+1}(\theta_\omega)+\dot{f}(\theta_\omega+b_\ell)^2}{g^2_{\ell+1}(\theta_\omega)}\right],\notag\\
+&\quad  \text{ for all } \ell \in [L-1],\notag \\
+{\partial^2 \flogl\over \partial b_\ell \partial b_{\ell+1}}&=\sum_{\omega\in\Omega}\mathds{1}_{\{y_\omega=\ell+1\}}\frac{\dot{f}(\theta_\omega+b_\ell)\dot{f}(\theta_\omega+b_{\ell+1})}{g^2_{\ell+1}(\theta_\omega)}\quad \text{and}\quad
+\ {\partial^2 \flogl\over \partial b_\ell \partial b_\ell'} = 0 \text{ if } |\ell-\ell'|>1.
+\end{align}
+
+Therefore, all entries in $\fploglb$ are upper bounded by $\left\{C_{\alpha,\beta,\Delta} \max_\ell (n_\ell+n_{\ell+1})\right\}>0$, and $\fpploglb$ is a tridiagonal matrix.
+
+We consider the profile log-likelihood $\flogl(\mb, \hat \Theta)$ as a function of $\mb\in\tB$.
+For notational convenience, we drop $\hat \Theta$ from $\flogl(\mb, \hat \Theta)$ and simply write $\flogl(\mb)$.
+By the second-order Taylor's expansion of $\flogl(\mb)$ around $\trueb$, we obtain
+\begin{equation}\label{eq:taylorb}
+\flogl(\hat\mb)=\flogl(\trueb)+(\trueb-\hat\mb)^T\fploglb(\trueb)+ {1\over 2}(\trueb-\hat\mb)^T\fpploglb(\check\mb)(\trueb-\hat\mb),
+\end{equation}
+where $\check\mb=\gamma\trueb+(1-\gamma)\hat\mb$ for some $\gamma\in[0,1]$, and $\fpploglb(\check\mb)$ denotes the $(L-1)$-by-$(L-1)$ Hession matrix evaluated at $\check\mb$.
+
+The linear term in~\eqref{eq:taylorb} can be bounded by Cauchy-Schwartz inequality,
+\begin{equation}\label{eq:linearb}
+(\trueb-\hat\mb)^T\fploglb(\trueb)\leq \FnormSize{}{\trueb-\hat\mb}\FnormSize{}{\fploglb(\trueb)}
+\leq \FnormSize{}{\trueb-\hat\mb}\sqrt{L-1}C_{\alpha,\beta,\Delta}\max_{\ell\in[L-1]}(n_\ell+n_{\ell+1}),
+\end{equation}
+where the last inequality is followed from
+\[
+\left|{\partial \tL_\tY\over \partial b_\ell}\Big|_{\mb=\trueb}\right| \leq C_{\alpha,\beta,\Delta}\max_{\ell\in[L-1]}(n_{\ell}+n_{\ell+1}), \quad \textrm{ for all } \; \ell\in [L-1].
+\]
+
+We next bound the quadratic term in \eqref{eq:taylorb}. Note that
+\begin{align}\label{eq:quadraticb}
+ &(\trueb-\hat\mb)^T \fpploglb(\check{\mb})(\trueb-\hat\mb)\\
+ &=\sum_{\ell\in [L-1]} \left( {\partial^2\tL_{\tY}\over \partial b^2_\ell} \Big|_{\mb=\check\mb} \right)(\hat b_\ell-b^{\text{true}}_{\ell})^2 + 2\sum_{\ell\in[L-1]/\{1\}}\left( {\partial^2\tL_{\tY}\over \partial b_\ell \partial b_{\ell-1}} \Big|_{\mb=\check\mb} \right)(\hat b_\ell-b^{\text{true}}_{\ell})(\hat b_{\ell-1}-b^{\text{true}}_{\ell-1}) \nonumber \\
+&\leq \sum_{\ell\in [L-1]} \left( {\partial^2\tL_{\tY}\over \partial b^2_\ell} \Big|_{\mb=\check\mb} \right)(\hat b_\ell-b^{\text{true}}_{\ell})^2 + \sum_{\ell\in[L-1]/\{1\}}\left( {\partial^2\tL_{\tY}\over \partial b_\ell \partial b_{\ell-1}} \Big|_{\mb=\check\mb} \right)\big[(\hat b_\ell-b^{\text{true}}_{\ell})^2 +(\hat b_{\ell-1}-b^{\text{true}}_{\ell-1})^2\big] \nonumber \\
+&=\left( {\partial^2\tL_{\tY}\over \partial b^2_1}+{\partial^2\tL_{\tY}\over \partial b_1 \partial b_2 }\right)\Big|_{\mb=\check\mb}(\hat b_1-b^{\text{true}}_{1})^2+\left( {\partial^2\tL_{\tY}\over \partial b^2_{L-1}}+{\partial^2\tL_{\tY}\over \partial b_{L-2} \partial b_{L-1} }\right)\Big|_{\mb=\check\mb}(\hat b_{L-1}-b^{\text{true}}_{L-1})^2  \\
+&\quad\quad +\sum_{\ell\in[L-2]/\{1\}}\left( {\partial^2\tL_{\tY}\over \partial b^2_\ell}+{\partial^2\tL_{\tY}\over \partial b_\ell \partial b_{\ell-1}}+{\partial^2\tL_{\tY}\over \partial b_{\ell+1} \partial b_{\ell}} \right)\Big|_{\mb=\check\mb} (\hat b_\ell-b^{\text{true}}_{\ell})^2     \\
+&\leq - \tilde{D}_{\alpha,\beta,\Delta} \sum_{\ell \in [L-1]}(\hat b_\ell-b_{\text{true},\ell})^2 \nonumber \\
+&=-\tilde{D}_{\alpha,\beta,\Delta}\FnormSize{}{\hat\mb-\trueb}^2,
+\end{align}
+where
+\begin{align}
+\tilde{D}_{\alpha,\beta,\Delta}&=\min_{\substack{|\theta|\leq \alpha,\\ \mb\in\tB}}\min_{\ell\in[L-1]} -\left( {\partial^2\tL_{\tY}\over \partial b^2_\ell}+{\partial^2\tL_{\tY}\over \partial b_\ell \partial b_{\ell-1}}+{\partial^2\tL_{\tY}\over \partial b_{\ell+1} \partial b_{\ell}} \right)\\
+&=\min_{\substack{|\theta|\leq \alpha,\\ \mb\in\tB}} \min_{\ell\in[L-1]}\left\{\sum_{\omega\in\Omega}-\mathds{1}_{\{y_\omega=\ell\}}\left( {\ddot{f}(\theta_\omega+b_{\ell}) g_\ell(\theta_\omega)- \dot{f}(\theta_\omega+b_{\ell})\dot{g}_\ell(\theta_\omega) \over g^2_\ell(\theta_\omega) }\right)\right.
+\\
+&\left.
+\hspace{1in}+\sum_{\omega\in\Omega}\mathds{1}_{\{y_\omega=\ell+1\}} \left( {\ddot{f}(\theta_\omega+b_\ell) g_{\ell+1}(\theta_\omega) - \dot{f}_{\theta_\omega+b_\ell}\dot{g}_{\ell+1} (\theta_\omega)\over g^2_{\ell+1}(\theta_\omega)} \right)\right\}\\
+&=\min_{\substack{|\theta|\leq \alpha,\\ \mb\in\tB}} \min_{\ell\in[L-1]}\left\{\KeepStyleUnderBrace{-\sum_{\omega\in\Omega} \mathds{1}_{\{y_\omega=\ell\}}{\partial \over \partial \theta_\omega}\left( \dot{f}(\theta_\omega+b_\ell) \over g_\ell(\theta_\omega) \right)}_{>0}+\KeepStyleUnderBrace{\sum_{\omega\in\Omega}\mathds{1}_{\{y_\omega=\ell+1\}} {\partial \over \partial \theta_\omega}\left( \dot{f}(\theta_\omega+b_\ell) \over g_{\ell+1}(\theta_\omega) \right) }_{>0}\right\}\\
+&\geq D_{\alpha,\beta,\Delta}  \min_{\ell\in[L-1]}\left(n_{\ell}+n_{\ell+1}\right) .
+\end{align}
+
+Combining inequalities~\eqref{eq:taylorb}, \eqref{eq:linearb} and~\eqref{eq:quadraticb} yields
+\[
+\flogl(\hat \mb)\leq \flogl(\trueb)+C_{\alpha,\beta,\Delta}\sqrt{L-1}\max_{\ell}(n_\ell+n_{\ell+1}) \FnormSize{}{\hat\mb-\trueb} -\frac{D_{\alpha,\beta,\Delta}}{2}\min_{\ell}(n_\ell+n_{\ell+1}) \FnormSize{}{\hat\mb-\trueb}^2 .
+\]
+Since $\hat \mb$ satisfies $\flogl(\hat \mb)-\flogl(\trueb)\geq 0$, we have that
+\[
+C_{\alpha,\beta,\Delta}\sqrt{L-1}\max_{\ell}(n_\ell+n_{\ell+1}) \FnormSize{}{\hat\mb-\trueb} -\frac{D_{\alpha,\beta,\Delta}}{2}\min_{\ell}(n_\ell+n_{\ell+1}) \FnormSize{}{\hat\mb-\trueb}^2\geq 0
+\]
+Finally,
+\[
+\text{MSE}(\hat \mb,\trueb)={1\over L-1}\FnormSize{}{\hat\mb-\trueb}\leq {2C_{\alpha,\beta,\Delta} \over D_{\alpha,\beta,\Delta}}{\max_{\ell}(n_\ell+n_{\ell+1})\over \min_{\ell}(n_{\ell}+n_{\ell+1})}  \sqrt{1\over L-1},
+\]
+which completes the proof.
 \end{proof}
 
-\begin{lem}[Corollary 3.5 in~\cite{brascamp2002extensions}]\label{lem:lossconvexity}
-Let $F(x,y):\mathbb{R}^{m+n}\rightarrow \mathbb{R}$ be an integrable function where $x\in \mathbb{R}^{m},y\in \mathbb{R}^n$.
-Let $$G(x) = \int_{\mathbb{R}^n}F(x,y)dy.$$ If $F(x,y)$ is a log concave function to (x,y), then $G(x)$ is a log concave function.
-\end{lem}
+\section{Additional results on simulations}
+In the main paper, we compare our method ({\bf Ordinal-T}) with other four methods ({\bf Ordinal-M}, {\bf 1bit-category-T},{\bf 1bit-sign-T}, and {\bf Continuous-T}).
+The methods are evaluated by their capabilities in predicting the most likely label for each entry, i.e.,\ $y_\omega^{\text{mode}}=\arg\max_\ell\mathbb{P}(y_\omega=\ell)$.
+Here we perform additional simulations to compare the methods by their performance in predicting the median label, $y_\omega^{\text{median}}=\min\{\ell\colon\mathbb{P}(y_\omega=\ell)\geq 0.5\}$.
+Under the latent variable model~\eqref{eq:latent} and Assumption~\ref{ass:link}, the median label is the quantized $\Theta$ without noise; i.e.\ $y_\omega^{\text{median}}=\sum_\ell \mathds{1}_{\theta_\omega\in(b_{\ell-1},b_\ell]}$.
 
-\section{Brain clustering method and result}
-\subsection{Clustering method}
-We perform clustering based on the estimated latent parameter $\hat\Theta$.
-Our clustering method is based on multidimensional version of Principal Component Analysis (PCA) and Singular Value Decomposition (SVD).
-In matrices case, we perform clustering on a $m\times n$ data matrix $X$ based on the following procedure.
+\fixme{miaoyan}{Please add a few sentence to describe Figure S}
+\begin{figure}[H]
+  \begin{center}
+    \includegraphics[width=.8\textwidth]{panel3.pdf}
+  \end{center}
+  \caption{Performance comparison in predicting the median label.
+(a,c) Prediction errors versus sample complexity $\rho=|\Omega|/d^k$ when $L=5$.
+(b,d) Prediction errors versus the number of ordinal levels $L$, when $\rho=0.8.$}
+\end{figure}
+
+\section{Additional results on HCP analysis}
+\subsection{Clustering based on Tucker representation}\label{sec:clustering}
+We perform clustering analyses based on the Tucker representation of the estimated tensor parameter $\hat\Theta$.
+The procedure is motivated from the higher-order extension of Principal Component Analysis (PCA) or Singular Value Decomposition (SVD).
+Recall that, in the matrix case, we usually perform clustering on an $m\times n$ (normalized) matrix $X$ based on the following procedure.
 First, we factorize $X$ into
-\begin{equation*}
+\begin{equation}
 X = U\Sigma V^T,
-\end{equation*}
+\end{equation}
 where $\Sigma$ is a diagonal matrix and $ U,V$ are factor matrices with orthogonal columns.
-We interpret each column of $V$ as a principal direction or axis and $U\Sigma$ as principal components.
-From this interpretation, we perform $K$-means method to $m$ rows of $U\Sigma$.
-We suggest to apply this procedure to the estimated latent parameter tensor $\hat\Theta$.
-For the notational convenience, we use $\Theta$ as $\hat\Theta$ in this section.
-From the Tucker decomposition which is multidimensional version of SVD, we have
-\begin{equation*}
-\Theta = \tC\times_1\mM_1\times_2\cdots\times_K\mM_K,
-\end{equation*}
-where $\tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is a core tensor, $\mM_k\in\mathbb{R}^{d_k\times r_k}$ are factor matrices with orthogonal columns, and $\times_k$ denotes the tensor-by-matrix multiplication~\cite{kolda2009tensor}.
-We perform mode-$k$ matricization of the tensor $\Theta$ to cluster data from mode-$k$ variable as follows.
-\begin{equation*}
-\Theta_{(k)} = \mM_k\tC_{(k)}\big(\mM_K\otimes\cdots\otimes\mM_1\big)
-\end{equation*}
-We interpret $\big(\mM_K\otimes\cdots\otimes\mM_1\big)$ as principal axes and $\mM_k\tC_{(k)}$ as principal components.
-By putting other mode variables as principal axes, we can exclude effects from those variables and can see $k$-th mode effect from checking principal components, $\mM_k\tC_{(k)}$.
-Therefore, we perform $K$-means method to $d_k$ rows of the principal components like in the matrices case.
+Second, we take each column of $V$ as a principal axis and $U\Sigma$ as principal components.
+A subsequent multivariate clustering method (such as $K$-means) is then applied to the $m$ rows of $U\Sigma$.
 
-\subsection{Clustering result on HCP}
-We perform $K$-means method to cluster brain nodes based on the suggested above method.
-We have the estimated latent parameter $\hat\Theta$ with the rank $(23,23,8)$ which minimizes the BIC value.
-The number of clusters is chosen to be eight from the elbow method in figure ~\ref{figure:elbow}.
-The table in \ref{table:clustering} shows eight cluster groups and their entries.
-Most of brain nodes fall into the cluster $\#1$ and the cluster $\#2$ which can be represented as the left side brain and the right side brain.
-From the group $\#3$ to the group $\#8$ , entries of each cluster share the same name encoded in the data and play a similar role in the brain.
-For example, the group $\#3$ and the group $\#4$ are the left and the right side of the regions called Supramarginal gyrus.
-Those regions are known to be in charge of interpreting tactile sensory data and involved in perception of space and limbs location \cite{carlson2012physiology,reed1994nature}.
-The figure \ref{figure:brain image} shows the nodes in the dataset on the brain image.
-Each node is connected by edges within the same cluster and the groups can be paired as the right part and the left part of the brain.
-The results show that the clustering based on our model successfully group the brain nodes without knowledge of the brain region.
+We apply a similar clustering procedure to the estimated parameter tensor $\hat\Theta$. Based on Tucker decomposition, we have
+\begin{equation}\label{eq:Tuckerest}
+\hat \Theta = \hat \tC\times_1\hat \mM_1\times_2\cdots\times_K\hat \mM_K,
+\end{equation}
+where $\hat \tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is the estimated core tensor, $\hat \mM_k\in\mathbb{R}^{d_k\times r_k}$ are estimated factor matrices with orthogonal columns, and $\times_k$ denotes the tensor-by-matrix multiplication~\citep{kolda2009tensor}.
+The mode-$k$ matricization of~\eqref{eq:Tuckerest} gives
+\begin{equation}
+\hat \Theta_{(k)} = \hat \mM_k\hat \tC_{(k)}\left(\hat \mM_K\otimes\cdots\otimes\hat \mM_1\right),
+\end{equation}
+where $\hat \Theta_{(k)} , \hat \tC_{(k)}$ denote the mode-$k$ unfolding of $\hat \Theta$ and $\hat \tC$, respectively.
+Then, the mode-$k$ clustering can be performed as follows.
+First, we take $\left(\hat \mM_K\otimes\cdots\otimes\hat \mM_1\right)$ as principal axes and $\hat \mM_k\hat \tC_{(k)}$ as principal components.
+Then, we perform $K$-means clustering method to the $d_k$ rows of the principal components $\hat \mM_k\hat \tC_{(k)}$.
+
+\subsection{Clustering results on HCP}
+We perform clustering analysis to the 68 brain nodes using the the produce described in Section~\ref{sec:clustering}.
+Recall that our ordinal tensor method outputs the estimated parameter tensor $\hat\Theta\in\mathbb{R}^{68\times 68\times136}$ with rank $(23,23,8)$.
+The number of clusters in the $K$-means is chosen based on the elbow method (see figure~\ref{figure:elbow}). Table~\ref{table:clustering} shows the eight clusters of brain nodes. 
 \begin{figure}[H]
@@
 \end{figure}
+
+We find that most brain nodes fall into cluster I and the cluster II, which can be represented as the left and right hemispheres, respectively (Figure~\ref{figure:brain image}a).
+From cluster III to the cluster VIII, entries within each cluster share the same name encoded in the data.
+For example, cluster III and cluster IV represent the {\it Supramarginal gyrus} region on the left brain and right brain.
+These regions are known to be in charge of interpreting tactile sensory data and involved in perception of space and limbs location~\cite{carlson2012physiology,reed1994nature}.
+\fixme{Miaoyan}{(I do not understand the figure and text.
+Why are there four panels in the Figure?
+Did you just plot all pairwise connections in Cluster I?
+Please change the edge weights using averaged connection strengths.
+Step 1.
+take average of connections over individuals.
+2.
+plot the averaged strength between nodes within a same cluster.
+You also need to tune the sparsity level to make the figure interpretable...) (Texts need to be rewritten...) Figure~\ref{figure:brain image} displays the averaged connection within each cluster brain nodes.
+Each node is connected by edges within the same cluster and the groups can be paired as the right part and the left part of the brain.}
+The results demonstrate that our clustering successfully groups similar brain nodes together without knowledge of external annotation.
 
@@
 \hline
-Group       & \multicolumn{3}{c|}{\#1}                                                  & \multicolumn{3}{c|}{\#2}                                      \\ \hline
-Brain nodes & \multicolumn{3}{l|}{\begin{tabular}[c]{@{}l@{}}"RMF\_L","Fpole\_L","Insula\_L",\\ "SupF\_L","CaudMF\_L",\\ "parsTriangularis\_L","parsOpercularis\_L",\\ "preCentral\_L","Tpole\_L","SupT\_L",\\ "SupT\_L","postCentral\_L", "SupP\_L",\\ "IP\_L","LO\_L","MOF\_L","SupF\_L",\\ "IsthmusC\_L","Precuneus\_L",\\ "Cuneus\_L","paraHippo\_L",\\ "Lingual\_L","SupT\_L","LO\_L"\end{tabular}} & \multicolumn{3}{l|}{\begin{tabular}[c]{@{}l@{}}"RMF\_R","Fpole\_R","Insula\_R",\\ "SupF\_R", "CaudMF\_R",\\ "parsTriangularis\_R" ,"parsOpercularis\_R",\\ "preCentral\_R", "Tpole\_R","SupT\_R",\\ "SupT\_R","postCentral\_R","SupP\_R",\\ "IP\_R","LO\_R","MOF\_R", "SupF\_R",\\ "IsthmusC\_R","Precuneus\_R",\\ "Cuneus\_R","paraHippo\_R",\\ "Lingual\_R","SupT\_R","LO\_R"\end{tabular}} \\ \hline
-Group       & \#3    & \#4    & \#5   & \#6  & \#7 & \#8
+Cluster     & \multicolumn{3}{c|}{I}                                                  & \multicolumn{3}{c|}{II}                                      \\ \hline
+Brain nodes & \multicolumn{3}{l|}{\begin{tabular}[c]{@{}l@{}} ``RMF\_L'',``'Fpole\_L",``Insula\_L",\\ ``SupF\_L",``CaudMF\_L",\\ ``parsTriangularis\_L",``parsOpercularis\_L",\\ ``preCentral\_L",``Tpole\_L",``SupT\_L",\\ ``SupT\_L",``postCentral\_L", ``SupP\_L",\\ ``IP\_L",``LO\_L",``MOF\_L",``SupF\_L",\\ ``IsthmusC\_L",``Precuneus\_L",\\ ``Cuneus\_L",``paraHippo\_L",\\ ``Lingual\_L",``SupT\_L",``LO\_L"\end{tabular}} & \multicolumn{3}{l|}{\begin{tabular}[c]{@{}l@{}}``RMF\_R",``Fpole\_R",``Insula\_R",\\ ``SupF\_R", ``CaudMF\_R",\\ ``parsTriangularis\_R" ,``parsOpercularis\_R",\\ ``preCentral\_R", ``Tpole\_R",``SupT\_R",\\ ``SupT\_R",``postCentral\_R",``SupP\_R",\\ ``IP\_R",``LO\_R",``MOF\_R", ``SupF\_R",\\ ``IsthmusC\_R",``Precuneus\_R",\\ ``Cuneus\_R",``paraHippo\_R",\\ ``Lingual\_R",``SupT\_R",``LO\_R"\end{tabular}} \\ \hline
+Cluster      & III    & IV    & V   & VI  & VII & VIII
 \\ \hline
-Brain nodes & \begin{tabular}[c]{@{}c@{}}"SupraM\_L" \\ "SupraM\_L" \\ "SupraM\_L" \\ "SupraM\_L"\end{tabular}                                      & \begin{tabular}[c]{@{}c@{}}"SupraM\_R" \\ "SupraM\_R" \\ "SupraM\_R" \\ "SupraM\_R"\end{tabular}                                      & \begin{tabular}[c]{@{}c@{}}"IT\_L" \\ "IT\_L" \\ "IT\_L"\end{tabular}                                      & \begin{tabular}[c]{@{}c@{}}"IT\_R" \\ "IT\_R" \\ "IT\_R"\end{tabular}                                                          & \begin{tabular}[c]{@{}c@{}}"MT\_L" \\ "MT\_L" \\ "MT\_L"\end{tabular}                                                         & \begin{tabular}[c]{@{}c@{}}"MT\_R" \\ "MT\_R"\\ "MT\_R"\end{tabular}                                                         \\ \hline
+Brain nodes & \begin{tabular}[c]{@{}c@{}}``SupraM\_L" \\ ``SupraM\_L" \\ ``SupraM\_L" \\ ``SupraM\_L"\end{tabular}                                      & \begin{tabular}[c]{@{}c@{}}``SupraM\_R" \\ ``SupraM\_R" \\ ``SupraM\_R" \\ ``SupraM\_R"\end{tabular}                                      & \begin{tabular}[c]{@{}c@{}}``IT\_L" \\ ``IT\_L" \\ ``IT\_L"\end{tabular}                                      & \begin{tabular}[c]{@{}c@{}}``IT\_R" \\ ``IT\_R" \\ ``IT\_R"\end{tabular}                                                          & \begin{tabular}[c]{@{}c@{}}``MT\_L" \\ ``MT\_L" \\ ``MT\_L"\end{tabular}                                                         & \begin{tabular}[c]{@{}c@{}}``MT\_R" \\ ``MT\_R"\\ ``MT\_R"\end{tabular}                                                         \\ \hline
 \end{tabular}
@@
 \end{small}
-\caption{Result of the brain nodes clustering: the last alphabet in each node name means the left side for `L' and the right side for `R' in the brain}
+\caption{Clustering result of brain nodes.
+The last alphabet in the node name indicates the left brain (`L' ) and the right brain (`R') {\color{red} Please use node names from Github.../../BrainNetViewer\_20171031/Data/miaoyan/Desikan\_miaoyan.node.txt.
+Please remove the quote.} }
 \end{table}
@@
   \label{figure:brain image}
-  \caption{Brain image of eight clusters: entries of each cluster are connected by edges}
+  \caption{Averaged connection within each of the eight clusters. \fixme{Miaoyan}{Nodes within each cluster are connected by edges?}}
 \end{figure}

### 2020-02-11 ICML/arxiv/ordinalT_supp_v3.tex
update

@@
  \end{align}
-where $f(\cdot)$ satisfies Assumption~\eqref{eq:ass}. Then, $\logl(\Theta, \mb)$ is concave in $(\Theta,\mb)$. 
+where $f(\cdot)$ satisfies Assumption~\ref{ass:link}. Then, $\logl(\Theta, \mb)$ is concave in $(\Theta,\mb)$. 
 \end{thm}
@@
 Define $d_{\text{total}}=\prod_k d_k$.
-With a little abuse of notation, we use $(\Theta,\mb)$ to denote a length-$(d_{\text{total}}+L-1)$-vector collecting all parameters together.
+With a little abuse of notation, we use $(\Theta,\mb)$ to denote the length-$(d_{\text{total}}+L-1)$-vector collecting all parameters together.
 Let us denote a bivariate function
@@
 Then, $\lambda(u,v) = \lambda(\ma_1^T(\Theta,\mb),\ \ma_2^T(\Theta,\mb))$ is concave in $(\Theta,\mb)$ by the definition of concavity.
-Furthermore, we can conclude that $\logl(\Theta, \mb) $ is concave in $(\Theta,\mb)$ because $\logl(\Theta, \mb)$ is the sum of $\lambda(u,v)$.
+Therefore, we can conclude that $\logl(\Theta, \mb) $ is concave in $(\Theta,\mb)$ because $\logl(\Theta, \mb)$ is the sum of $\lambda(u,v)$.
 
@@
 where $\mathds{1}_{[u,v]}$ is an indicator function that equals 1 in the interval $[u,v]$, and 0 elsewhere.
-Furthermore, $\mathds{1}_{[u,v]}(x)$ is log-concave in $(u,v,x)$ and $f'(x) = \frac{e^x}{(1+e^x)^2}$ is log-concave in $x$.
-The convexities imply that $\mathds{1}_{(u,v)}(x)f'(x)$
+Furthermore, $\mathds{1}_{[u,v]}(x)$ is log-concave in $(u,v,x)$, and by Assumption~\ref{ass:link}, $f'(x)$ is log-concave in $x$.
+It follows that $\mathds{1}_{(u,v)}(x)f'(x)$
 is a log-concave in $(u,v,x)$. By Lemma~\ref{lem:lossconvexity}, we conclude that $\lambda(u,v)$ is concave in $(u,v)$ where $u>v$. 
@@
 \label{sec:lemma}
-
-We begin with various notion of tensor norms that are useful for the proofs of the main theorems.
+This section collects lemmas that are useful for the proofs of the main theorems.
 
@@
 \[
-U_{\alpha,\hat \mb}=\max_{|\theta|\leq \alpha}\max_{\ell\in[L-1]}{|\dot{g}_\ell(\theta)\over g_\ell(\theta)|},\quad \text{and}\quad L_{\alpha,\hat \mb}=\min_{|\theta|\leq \alpha}\min_{\ell\in[L-1]} {\dot{g}^2_\ell(\theta)-\ddot{g}_\ell(\theta)g_\ell(\theta)\over g^2_\ell(\theta)}.
+U_{\alpha,\hat \mb}=\max_{|\theta|\leq \alpha}\max_{\ell\in[L-1]}{|\dot{g}_\ell(\theta)|\over g_\ell(\theta)},\quad \text{and}\quad L_{\alpha,\hat \mb}=\min_{|\theta|\leq \alpha}\min_{\ell\in[L-1]} {\dot{g}^2_\ell(\theta)-\ddot{g}_\ell(\theta)g_\ell(\theta)\over g^2_\ell(\theta)}.
 \]
@@
   \end{center}
-  \caption{Performance comparison in predicting the median label.
+  \caption{Performance comparison for predicting the median label.
 (a,c) Prediction errors versus sample complexity $\rho=|\Omega|/d^k$ when $L=5$.
@@
 where $\Sigma$ is a diagonal matrix and $ U,V$ are factor matrices with orthogonal columns.
-Second, we take each column of $V$ as a principal axis and $U\Sigma$ as principal components.
+Second, we take each column of $V$ as a principal axis and each row in $U\Sigma$ as principal component.
 A subsequent multivariate clustering method (such as $K$-means) is then applied to the $m$ rows of $U\Sigma$.
 
-We apply a similar clustering procedure to the estimated parameter tensor $\hat\Theta$. Based on Tucker decomposition, we have
+We apply a similar clustering procedure to the estimated parameter tensor $\hat\Theta$. Based on Tucker representation of $\hat \Theta$, we have
 \begin{equation}\label{eq:Tuckerest}
@@
 Then, the mode-$k$ clustering can be performed as follows.
-First, we take $\left(\hat \mM_K\otimes\cdots\otimes\hat \mM_1\right)$ as principal axes and $\hat \mM_k\hat \tC_{(k)}$ as principal components.
-Then, we perform $K$-means clustering method to the $d_k$ rows of the principal components $\hat \mM_k\hat \tC_{(k)}$.
+First, we take colums in $\left(\hat \mM_K\otimes\cdots\otimes\hat \mM_1\right)$ as principal axes and rows in $\hat \mM_k\hat \tC_{(k)}$ as principal components.
+Then, we perform $K$-means clustering method to the $d_k$ rows of the matrix $\hat \mM_k\hat \tC_{(k)}$.
 
 \subsection{Clustering results on HCP}
-We perform clustering analysis to the 68 brain nodes using the the produce described in Section~\ref{sec:clustering}.
+We perform clustering analysis on the 68 brain nodes using the the produce described in Section~\ref{sec:clustering}.
 Recall that our ordinal tensor method outputs the estimated parameter tensor $\hat\Theta\in\mathbb{R}^{68\times 68\times136}$ with rank $(23,23,8)$.
-The number of clusters in the $K$-means is chosen based on the elbow method (see figure~\ref{figure:elbow}). Table~\ref{table:clustering} shows the eight clusters of brain nodes. 
+We apply $K$-means to the mode-1 principal component matrix of size $68\times 184$ ($184=23\times8$).
+The elbow method suggests eight clusters among the 68 nodes (see Figure~\ref{figure:elbow}).
+Table~\ref{table:clustering} shows the eight clusters and their compositions.
 \begin{figure}[H]

### 2020-02-13 ICML/arxiv/icml_arxiv.tex
Add files via upload

--- 2020-02-07 ICML/submit/ordinalTv3.tex
@@
-%%%%%%%% ICML 2020 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
+\documentclass[11pt]{article}
 
-\documentclass{article}
+\usepackage{fancybox}
 
-% Recommended, but optional, packages for figures and better typesetting:
-\usepackage{microtype}
-\usepackage{graphicx}
-\usepackage{subfigure}
-\usepackage{booktabs} % for professional tables
+\usepackage{color}
+\usepackage{url}
+\usepackage[margin=1in]{geometry}
 
-% hyperref makes hyperlinks in the resulting PDF.
-% If your build breaks (sometimes temporarily if a hyperlink spans a page)
-% please comment out the following usepackage line and replace
-% \usepackage{icml2020} with \usepackage[nohyperref]{icml2020} above.
-\usepackage{hyperref}
+\renewcommand{\textfraction}{0.0}
+\renewcommand{\topfraction}{1.0}
+%\renewcommand{\textfloatsep}{5mm}
 
-% Attempt to make hyperref and algorithmic work together better:
 \newcommand{\theHalgorithm}{\arabic{algorithm}}
 
-% Use the following line for the initial blind version submitted for review:
-\usepackage{icml2020}
+\usepackage{comment}
+% Definitions of handy macros can go here
+\usepackage{amsmath,amssymb,amsthm,bm,mathtools}
+\usepackage{natbib}
+%\usepackage{dsfont,multirow,hyperref,setspace,natbib,enumerate}
+\usepackage{dsfont,multirow,hyperref,setspace,enumerate}
+\hypersetup{colorlinks,linkcolor={blue},citecolor={blue},urlcolor={red}}
+\usepackage{algorithm}
+\usepackage{algorithmic}
+\usepackage{makecell}
 
-% If accepted, instead use the following line for the camera-ready submission:
-%\usepackage[accepted]{icml2020}
+\mathtoolsset{showonlyrefs=true}
 
-% The \icmltitle you define below is probably too long as a header.
-% Therefore, a short form for the running title is supplied here:
-\icmltitlerunning{Ordinal tensor denoising and completion}
-\usepackage{wrapfig}
-\usepackage{multirow}
-\usepackage{graphicx}
-%\usepackage[utf8]{inputenc} % allow utf-8 input
-%\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-\usepackage{hyperref}       % hyperlinks
-\usepackage{url}            % simple URL typesetting
-%\usepackage{booktabs}       % professional-quality tables
-\usepackage{amsmath,amssymb}
-\usepackage{amsthm}    % blackboard math symbols
-%\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-%\usepackage{microtype}      % microtypography
-\usepackage{bm}
-%\usepackage{subfig}
-%\usepackage[english]{babel}
-%\usepackage{algorithm}
-%\usepackage{appendix}
-\usepackage{mathtools}
-\mathtoolsset{showonlyrefs}
-\usepackage{enumitem}
 \theoremstyle{plain}
@@
 
+\setcounter{figure}{0}
+\setcounter{table}{0}
+
+\def\MLET{\hat \Theta_{\text{MLE}}}
+\newcommand{\cmt}[1]{{\leavevmode\color{red}{#1}}}
+
 \usepackage{dsfont}
-%\usepackage{algpseudocode,algorithm}
-%\algnewcommand\algorithmicinput{\textbf{Input:}}
-%\algnewcommand\algorithmicoutput{\textbf{Output:}}
-%\algnewcommand\INPUT{\item[\algorithmicinput]}
-%\algnewcommand\OUTPUT{\item[\algorithmicoutput]}
-%\DeclareMathOperator*{\minimize}{minimize}
 
+\DeclareMathOperator*{\minimize}{minimize}
+
+\usepackage{wrapfig}
+
+\usepackage{mathtools}
+\mathtoolsset{showonlyrefs}
 \newcommand*{\KeepStyleUnderBrace}[1]{%f
@@
 }
-\usepackage{makecell}
+\usepackage{xr}
+\externaldocument{ordinalTv3}
 \input macros.tex
+\allowdisplaybreaks
 
-\usepackage{amssymb}
-\usepackage{pifont}
-\newcommand{\cmark}{\ding{51}}%
-\newcommand{\xmark}{\ding{55}}%
+\title{Tensor denoising and completion based on ordinal observations}
+\date{}
+\author{%
+Chanwoo Lee \\
+University of Wisconsin -- Madison\\
+\texttt{chanwoo.lee@wisc.edu} \\
+\and
+Miaoyan Wang \\
+University of Wisconsin -- Madison\\
+\texttt{miaoyan.wang@wisc.edu} \\
+}
 
 \begin{document}
-
-\twocolumn[
-\icmltitle{Tensor denoising and completion based on ordinal observations}
-
-% It is OKAY to include author information, even for blind
-% submissions: the style file will automatically remove it for you
-% unless you've provided the [accepted] option to the icml2020
-% package.
-
-% List of affiliations: The first argument should be a (short)
-% identifier you will use later to specify author affiliations
-% Academic affiliations should list Department, University, City, Region, Country
-% Industry affiliations should list Company, City, Region, Country
-
-% You can specify symbols, otherwise they are numbered in order.
-% Ideally, you should not use this facility. Affiliations will be numbered
-% in order of appearance and this is the preferred way.
-\icmlsetsymbol{equal}{*}
-
-\begin{icmlauthorlist}
-\icmlauthor{Chanwoo Lee}{equal,to}
-\icmlauthor{Miaoyan Wang}{equal,to}
-\end{icmlauthorlist}
-
-\icmlaffiliation{to}{Department of Statistics, University of Wisconsin at Madison, USA}
-
-\icmlcorrespondingauthor{Miaoyan Wang}{miaoyan.wang@wisc.edu.com}
-
-% You may provide any keywords that you
-% find helpful for describing your paper; these are used to populate
-% the "keywords" metadata in the PDF but will not be shown in the document
-\icmlkeywords{Higher-order tensors, ordinal observation, tensor decomposition, tensor completion}
-
-\vskip 0.3in
-]
-
-% this must go after the closing bracket ] following \twocolumn[ ...
-
-% This command actually creates the footnote in the first column
-% listing the affiliations and the copyright notice.
-% The command takes one argument, which is text to display at the start of the footnote.
-% The \icmlEqualContribution command is standard text for equal contribution.
-% Remove it (just {}) if you do not need this facility.
-
-%\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
-\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.
-
+\maketitle
 \begin{abstract}
@@
 We demonstrate the outperformance of our approach over previous methods on the tasks of clustering and collaborative filtering.
-
+\\[0.5cm]
+\textbf{Keywords:} Higher-order tensors, ordinal observation, tensor decomposition, tensor completion
 \end{abstract}
@@
 \centerline{\includegraphics[width=16cm]{compare.pdf}}
-\vspace{-.3cm}
 \caption{Comparison with previous work.
@@
 }~\label{tab:compare}
-\vspace{-.5cm}
 \end{table*}
@@
 
-\vspace{-.2cm}
 \section{Model formulation and motivation}
@@
 The link function is assumed to satisfy:
-\vspace{-.4cm}
 \begin{enumerate}
 \item $f(\theta)$ is strictly increasing and twice-differentiable in $\theta\in\mathbb{R}/\{0\}$.
-\vspace{-.2cm}
 \item $f'(\theta)$ is strictly log-concave and symmetric with respect to $\theta=0$.
@@
 
-We define a few key quantities that will be used in our theory. Let $g_\ell=f(\theta+b_\ell)-f(\theta+b_{\ell-1})$ for all $\ell\in[L]$, and
+We define a few key quantities that will be used in our theory. Let $g_\ell=f(b_\ell-\theta)-f(b_{\ell-1}-\theta)$ for all $\ell\in[L]$, and
 \begin{align}\label{eq:regular}
-A_\alpha &= \min_{\ell\in[L], |\theta|\leq \alpha} g_\ell(\theta),\quad  U_\alpha=\max_{\ell\in[L], |\theta|\leq \alpha } {\dot{g}_\ell(\theta)\over g_\ell(\theta)},\\
+A_\alpha &= \min_{\ell\in[L], |\theta|\leq \alpha} g_\ell(\theta),\quad  U_\alpha=\max_{\ell\in[L], |\theta|\leq \alpha } {|\dot{g}_\ell(\theta)|\over g_\ell(\theta)},\\
  L_\alpha &= \min_{\ell\in[L], |\theta|\leq \alpha }\left[{\dot{g}_\ell^2 (\theta)\over g_\ell^2(\theta)} -{\ddot{g}_\ell(\theta)\over g_\ell(\theta)}\right],
@@
 Theorem~\ref{thm:rate} establishes the statistical convergence for the estimator~\eqref{eq:estimator}.
-In fact, the proof of this theorem (see the Supplement) shows that the same statistical rate holds, not only for the global optimizer~\eqref{eq:estimator}, but also for any local optimizer $\check \Theta$ in the level set $\{\check\Theta\in\tP: \logl(\check\Theta)\geq \logl(\trueT)\}$.
+In fact, the proof of this theorem (see Section~\ref{sec:proof}) shows that the same statistical rate holds, not only for the global optimizer~\eqref{eq:estimator}, but also for any local optimizer $\check \Theta$ in the level set $\{\check\Theta\in\tP: \logl(\check\Theta)\geq \logl(\trueT)\}$.
 This suggests that the local optimality itself is not necessarily a severe concern in our context, as long as the convergent objective is large enough.
@@
 \begin{align}\label{eq:lower}
-\inf_{\hat \Theta }\sup_{\trueT\in\tP} \mathbb{P}\Big\{& \text{MSE}(\hat \Theta, \trueT) \\
-& \geq {1\over 256} \min\left( \alpha^2, \ { Cr_{\max}d_{\max} \over \prod_k d_k} \right) \Big\} \geq {1\over 8},
+\inf_{\hat \Theta }\sup_{\trueT\in\tP} \mathbb{P}\Big\{& \text{MSE}(\hat \Theta, \trueT)  \geq c \min\left( \alpha^2, \ { Cr_{\max}d_{\max} \over \prod_k d_k} \right) \Big\} \geq {1\over 8},
 \end{align}
-where $C=C(\alpha, L, f,\mb)>0$ is a constant independent of tensor dimension and the rank.
+where $C=C(\alpha, L, f,\mb)>0$ and $c>0$ are constants independent of tensor dimension and the rank.
 \end{thm}
@@
 \end{thm}
+
 Theorem~\ref{thm:completion} shows that our estimator achieves consistent recovery using as few as $\tilde{\tO}(Kd)$ noisy, quantized observations from an order-$K$ $(d,\ldots,d)$-dimensional tensor.
@@
 
-\vspace{-.2cm}
 \section{Numerical Implementation}\label{sec:algorithm}
+
 We describe the algorithm to seek the optimizer of~\eqref{eq:objective}.
 In practice, the cut-off points $\mb$ are often unknown, so we choose to maximize $\logl$ jointly over $(\Theta,\mb)\in\tP\times \tB$.
-The objective $\logl$ is concave in $(\Theta,\mb)$ whenever $f'$ is log-concave (See the Supplement).
+The objective $\logl$ is concave in $(\Theta,\mb)$ whenever $f'$ is log-concave (see Section~\ref{sec:proof}).
 However, the feasible set $\tP$ is non-convex, which makes the optimization~\eqref{eq:objective} a non-convex problem.
@@
  \end{algorithm}
+
 We comment on two implementation details before concluding this section.
@@
 The dashed line is the objective value at the true parameter $\logl(\trueT)$.
-It is observed that the algorithm generally converges quickly to a desirable value in reasonable number of steps.
+We find that the algorithm generally converges quickly to a desirable value in reasonable number of steps.
 The actual running time per iteration is shown in the plot legend.
@@
 \centering
-\vspace{-.4cm}
-\includegraphics[width=8cm]{algorithm.pdf}\label{fig:stability}
-\vspace{-.3cm}
+\includegraphics[width=0.7\textwidth]{algorithm.pdf}\label{fig:stability}
 \caption{Trajectory of objective function with various $d$ and $r$.}
-\vspace{-.6cm}
 \end{figure}
@@
 
-\vspace{-.2cm}
 \section{Experiments}\label{sec:experiment}
@@
 
-\vspace{-.1cm}
 \subsection{Finite-sample performance}\label{sec:simulation}
-\vspace{-.1cm}
 The first experiment examines the performance under complete observations.
@@
 The decay in the error appears to behave on the order of $d^{-2}$, which is consistent with our theoretical results~\eqref{eq:rate}.
-It is seen that a higher rank leads to a larger error, as reflected by the upward shift of the curve as $r$ increases.
+We find that a higher rank leads to a larger error, as reflected by the upward shift of the curve as $r$ increases.
 Indeed, a higher rank implies the higher number of parameters to estimate, thus increasing the difficulty of the estimation.
@@
 
-\begin{figure}[http]
-\includegraphics[width=8.3cm]{panel.pdf}
-\vspace{-.4cm}
+\begin{figure}[H]
+  \centering
+\includegraphics[width=0.6\textwidth]{panel.pdf}
 \caption{Empirical relationship between (relative) MSE versus (a) dimension $d$, (b) signal level $\alpha$, (c) observation fraction $\rho$, and (d) number of ordinal levels $L$.
@@
 \label{fig:finite}
-\vspace{-.2cm}
 \end{figure}
 
-\vspace{-.1cm}
-\subsection{Comparison with alternative methods}
-\vspace{-.1cm}
+\subsection{Comparison with alternative methods}\label{sec:compare}
 Next, we compare our ordinal tensor method ({\bf Ordinal-T}) with three popular low-rank methods:
 
-\begin{itemize}[itemsep=-.5pt,topsep=-.5pt,leftmargin=*,partopsep=-.5pt]
+\begin{itemize}
 \item Continuous tensor decomposition ({\bf Continuous-T})~\cite{acar2010scalable} is a low-rank approximation method based on classical Tucker model.
@@
 \begin{figure}[t]
-\includegraphics[width=8.3cm]{panel2.pdf}
-\vspace{-.4cm}
+  \centering
+\includegraphics[width=0.6\textwidth]{panel2.pdf}
 \caption{Performance comparison in MCR (a, b) and MAD (c, d).
@@
 }\label{fig:compare}
-\vspace{-.6cm}
 \end{figure}
@@
 
-\vspace{-.2cm}
 \section{Data Applications}
-\vspace{-.1cm}
 We apply our ordinal tensor method to two real-world datasets.
@@
 
-\vspace{-.1cm}
 \subsection{Human Connectome Project (HCP)}
-\vspace{-.1cm}
 Each entry in the HCP dataset takes value on a nominal scale, \{{\it high, moderate, low}\}, indicating the strength level of fiber connection.
@@
 In particular, the top two clusters represent the left and right hemispheres; the smaller clusters represent local regions driving by similar nodes.
-For example, the cluster IV consists of brain nodes {\it postcentral}, {\it posteriorcingulate}, {\it pericalcarine} and {\it temporalpole}, all of which are located at {\it SupraM}.
+For example, the cluster IV consists of brain nodes all of which are called {\it R.SUPRAMARGINAL} and located at the same region.
 The identified similarity among brain nodes suggests the applicability of our method.
@@
 \begin{table}[http]
-\resizebox{.48\textwidth}{!}{
-\begin{tabular}{c|c|c}
+\begin{tabular}{c|c|c||c|c|c}
 \Xhline{2\arrayrulewidth}
-\multicolumn{3}{c}{Human Connectome Project (HCP) dataset}\\
+\multicolumn{3}{c||}{Human Connectome Project (HCP) dataset} & \multicolumn{3}{c}{InCarMusic dataset}\\
 \Xhline{2\arrayrulewidth}
-Method                    & MAD & MCR\\
+Method                    & MAD & MCR & Method                    & MAD & MCR\\
 \hline
-Ordinal-T (ours) & 0.1607 (0.0005)& 0.1606 (0.0005)\\
- Continuous-T&0.2530 (0.0002)&0.1599 (0.0002)\\
- 1bit-sign-T&0.3566 (0.0010)&0.1563 (0.0010)\\
- \Xhline{2\arrayrulewidth}
-\multicolumn{3}{c}{InCarMusic dataset}\\
-\Xhline{2\arrayrulewidth}
-Method                    & MAD & MCR\\
-\hline
-Ordinal-T (ours) & 1.37 (0.039) &  0.59 (0.009)\\
-Continuous-T & 2.39 (0.152)&0.94 (0.027)\\
-1bit-sign-T&1.39 (0.003)& 0.81 (0.005)\\
+Ordinal-T (ours) & 0.1607 (0.0005)& 0.1606 (0.0005)& Ordinal-T (ours) & 1.37 (0.039) &  0.59 (0.009)\\
+ Continuous-T&0.2530 (0.0002)&0.1599 (0.0002)& Continuous-T & 2.39 (0.152)&0.94 (0.027)\\
+
+ 1bit-sign-T&0.3566 (0.0010)&0.1563 (0.0010)& 1bit-sign-T&1.39 (0.003)& 0.81 (0.005)\\
 \hline
 \end{tabular}
-}
-\vspace{-.2cm}
+
 \caption{Comparison of prediction error in the HPC and InCarMusic analyses. Standard errors are reported in parentheses.}\label{table:CV}
-\vspace{-.3cm}
 \end{table}
 
-\vspace{-.1cm}
 \subsection{InCarMusic recommendation system}
-\vspace{-.1cm}
 We apply ordinal tensor completion to a recommendation system {\it InCarMusic}.
@@
 
-\vspace{-.2cm}
-\section{Conclusion}
-\vspace{-.1cm}
+\section{Conclusions}
 We have developed a low-rank tensor estimation method based on possibly incomplete, ordinal-valued observations.
 A sharp error bound is established, and we demonstrate the outperformance of our approach compared to other methods.
-The work unlocked several directions of future research.
+The work unlocks several directions of future research.
 One interesting question would be the inference problem, i.e..
@@
 
-\newpage
+\section{Proofs}\label{sec:proof}
+\subsection{Estimation error for tensor denoising}
+\begin{proof}[Proof of Theorem~\ref{thm:rate}]
+We suppress the subscript $\Omega$ in the proof, because the tensor denoising assumes complete observation $\Omega=[d_1]\times \cdots \times [d_K]$.
+It follows from the expression of $\flogl(\Theta)$ that
+\begin{align}\label{eq:property}
+{\partial \flogl\over \partial \theta_\omega}&=\sum_{\ell\in[L]}\mathds{1}_{\{y_{\omega}=\ell\}}
+{\dot{g}_\ell(\theta_\omega)\over g_\ell(\theta_\omega)},\notag\\
+{\partial^2 \flogl\over \partial \theta_\omega^2}&=\sum_{\ell\in[L]}\mathds{1}_{\{y_\omega=\ell\}}{\ddot{g}_\ell(\theta_\omega)g_\ell(\theta_\omega)-\dot{g}^2_\ell(\theta_\omega)\over g^2_\ell(\theta_\omega)}\ \text{and}\quad
+{\partial^2 \flogl\over \partial \theta_\omega \theta_\omega'}=0\ \text{if}\ \omega\neq \omega',
+\end{align}
+for all $\omega\in[d_1]\times \cdots \times [d_K]$.
+Define $d_{\text{total}}=\prod_k d_k$.
+Let $\fplogl\in\mathbb{R}^{d_1\times\cdots\times d_K}$ denote the tensor of gradient with respect to $\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}$, and $\fpplogl$ the corresponding Hession matrix of size $d_\text{total}$-by-$d_{\text{total}}$.
+Here, $\Vec(\cdot)$ denotes the operation that turns a tensor into a vector.
+By~\eqref{eq:property}, $\fpplogl$ is a diagonal matrix.
+Recall that
+\begin{equation}\label{eq:bound}
+U_\alpha=\max_{\ell\in[L],|\theta|\leq \alpha}{|\dot{g}_\ell(\theta)|\over g_\ell(\theta)}>0 \quad \text{and}\quad
+L_\alpha=\min_{\ell\in[L],|\theta|\leq \alpha} {\dot{g}^2_\ell(\theta)-\ddot{g}_\ell(\theta)g_\ell(\theta)\over g^2_\ell(\theta)}>0.
+\end{equation}
+Therefore, the entries in $\fplogl$ are upper bounded in magnitude by $U_\alpha>0$, and all diagonal entries in $\fpplogl$ are upper bounded by $-L_{\alpha}<0$.
+
+By the second-order Taylor's expansion of $\flogl(\Theta)$ around $\trueT$, we obtain
+\begin{equation}\label{eq:taylor}
+\flogl(\Theta)=\flogl(\trueT)+\langle\Vec(\fplogl),\ \Vec(\Theta-\trueT)\rangle+{1\over 2}\Vec(\Theta-\trueT)^T\fpplogl(\check\Theta)\Vec(\Theta-\trueT),
+\end{equation}
+$\check\Theta=\gamma\trueT+(1-\gamma)\Theta$ for some $\gamma\in[0,1]$, and $\fpplogl(\check\Theta)$ denotes the $\prod_kd_k$-by-$\prod_k d_k$ Hession matrix evaluated at $\check\Theta$.
+
+We first bound the linear term in~\eqref{eq:taylor}. Note that, by Lemma~\ref{lem:inq},
+\begin{equation}\label{eq:linear}
+|\langle\Vec(\fplogl(\trueT), \Vec(\Theta-\trueT)  \rangle|\leq \snormSize{}{\fplogl(\trueT)} \nnormSize{}{\Theta-\trueT},
+\end{equation}
+where $\snormSize{}{\cdot}$ denotes the tensor spectral norm and $\nnormSize{}{\cdot}$ denotes the tensor nuclear norm. Define
+\[
+s_\omega={\partial \tL_\tY\over \partial \theta_\omega}\Big|_{\Theta=\trueT} \;\; \textrm{ for all } \; \omega\in[d_1]\times\cdots\times [d_K].
+\]
+Based on~\eqref{eq:property} and the definition of $U_\alpha$, $\fplogl(\trueT)=\entry{s_{\omega}}$ is a random tensor whose entries are independently distributed satisfying
+\begin{equation}\label{eq:norm}
+\mathbb{E}(s_\omega)=0,\quad |s_\omega|\leq U_\alpha, \quad \text{for all }\omega\in[d_1]\times \cdots \times [d_K].
+\end{equation}
+By lemma~\ref{lem:noisytensor}, with probability at least $1-\exp(-C_1 \sum_kd_k)$, we have
+\begin{equation}\label{eq:normrandom}
+\snormSize{}{\fplogl(\trueT)} \leq C_2 U_\alpha\sqrt{\sum_k d_k},
+\end{equation}
+where $C_1, C_2$ are two positive constants that depend only on $K$.
+Furthermore, note that $\text{rank}(\Theta)\leq \mr$, $\text{rank}(\trueT)\leq \mr$, so $\text{rank}(\Theta-\trueT)\leq 2\mr$.
+By lemma~\ref{lem:nuclear}, $\nnormSize{}{\Theta-\trueT}\leq (2r_{\max})^{K-1\over 2}\FnormSize{}{\Theta-\trueT}$.
+Combining~\eqref{eq:linear}, \eqref{eq:norm} and \eqref{eq:normrandom}, we have that, with probability at least $1-\exp(-C_1 \sum_kd_k)$,
+\begin{equation}\label{eq:linearconclusion}
+|\langle \Vec(\fplogl(\trueT)), \Vec(\Theta-\trueT)  \rangle | \leq C_2 U_\alpha  \sqrt{r_{\max}^{K-1} \sum_k d_k}  \FnormSize{}{\Theta-\trueT}.
+\end{equation}
+
+We next bound the quadratic term in \eqref{eq:taylor}. Note that
+\begin{align}\label{eq:quadratic}
+ \Vec(\Theta-\trueT)^T \fpplogl(\check{\Theta})\Vec(\Theta-\trueT)&=\sum_\omega \left( {\partial^2\tL_{\tY}\over \partial \theta^2_\omega} \Big|_{\Theta=\check\Theta} \right)(\theta_\omega-\theta_{{\text{true}},\omega})^2 \nonumber \\
+&\leq - L_\alpha\sum_{\omega}(\Theta_{\omega}-\Theta_{\text{true},\omega})^2 \nonumber \\
+&=-L_\alpha\FnormSize{}{\Theta-\trueT}^2,
+\end{align}
+where the second line comes from the fact that  $\mnormSize{}{\check\Theta}\leq \alpha$ and the definition of $L_\alpha$.
+
+Combining~\eqref{eq:taylor}, \eqref{eq:linearconclusion} and~\eqref{eq:quadratic}, we have that, for all $\Theta\in\tP$, with probability at least $1-\exp(-C_1 \sum_kd_k)$,
+\[
+\tL_\tY(\Theta)\leq \tL_{\tY}(\trueT)+C_2U_\alpha  \left(r_{\max}^{K-1}\sum_k d_k\right)^{1/2}  \FnormSize{}{\Theta-\trueT}-{L_\alpha\over 2}\FnormSize{}{\Theta-\trueT}^2.
+\]
+In particular, the above inequality also holds for $\hat \Theta\in\tP$. Therefore,
+\[
+\tL_\tY(\hat \Theta)\leq \tL_{\tY}(\trueT)+C_2U_\alpha \left(r_{\max}^{K-1}\sum_k d_k\right)^{1/2}  \FnormSize{}{\hat \Theta-\trueT}-{L_\alpha\over 2} \FnormSize{}{\hat \Theta-\trueT}^2.
+\]
+Since $\hat \Theta=\arg\max_{\Theta\in\tP}\tL_\tY(\Theta)$, $\tL_\tY(\hat \Theta)-\tL_{\tY}(\trueT)\geq 0$, which gives
+\[
+C_2U_\alpha \left(r_{\max}^{K-1}\sum_k d_k\right)^{1/2}  \FnormSize{}{\hat \Theta-\trueT}-{L_\alpha\over 2}\FnormSize{}{\hat \Theta-\trueT}^2\geq 0.
+\]
+Henceforth,
+\[
+{1\over \sqrt{\prod_k d_k}} \FnormSize{}{\hat \Theta-\trueT}\leq {2C_2U_\alpha \sqrt{r_{\max}^{K-1}\sum_k d_k}\over L_\alpha \sqrt{\prod_k d_k}}={2C_2U_\alpha r_{\max}^{(K-1)/2}\over L_\alpha} \sqrt{ \sum_k d_k \over \prod_k d_k}.
+\]
+This completes the proof.
+\end{proof}
+
+\begin{proof}[Proof of Corollary~\ref{cor:prediction}]
+The result follows immediately from Theorem~\ref{thm:rate} and Lemma~\ref{lem:KL}.
+\end{proof}
+
+\begin{proof}[Proof of Theorem~\ref{thm:minimax}]
+
+Let $d_{\text{total}}=\prod_{k\in[K]}d_k$, and $\gamma\in[0,1]$ be a constant to be specified later.
+Our strategy is to construct a finite set of tensors $\tX=\{\Theta_i\colon i=1,\ldots \}\subset \tP$ satisfying the properties of (i)-(iv) in Lemma~\ref{lem:construction}.
+By Lemma~\ref{lem:construction}, such a subset of tensors exist.
+For any tensor  $\Theta\in\tX$, let $\mathbb{P}_{\Theta}$ denote the distribution of $\tY|\Theta$, where $\tY$ is the ordinal tensor.
+In particular, $\mathbb{P}_{\mathbf{0}}$ is the distribution of $\tY$ induced by the zero parameter tensor $\mathbf{0}$, i.e., the distribution of $\tY$ conditional on the parameter tensor $\Theta=\mathbf{0}$.
+Based on the Remark for Lemma~\ref{lem:KL}, we have
+\begin{equation}\label{eq:KLbound1}
+\mathrm{KL}(\mathbb{P}_{\Theta}|| \mathbb{P}_{\mathbf{0}})\leq C \FnormSize{}{\Theta}^2,
+\end{equation}
+where $C={(4L-6) \dot{f}^2(0)\over  A_\alpha}>0$ is a constant independent of the tensor dimension and rank.
+Combining the inequality~\eqref{eq:KLbound1} with property (iii) of $\tX$, we have
+\begin{equation}\label{eq:KLbound}
+\text{KL}(\mathbb{P}_{\Theta}||\mathbb{P}_{\mathbf{0}})\leq \gamma^2 R_{\max} d_{\max}.
+\end{equation}
+From~\eqref{eq:KLbound} and the property (i), we deduce that the condition
+\begin{equation}\label{eq:totalKL}
+{1\over \text{Card}(\tX)-1}\sum_{\Theta \in\tX}\text{KL}(\mathbb{P}_{\Theta}, \mathbb{P}_{\mathbf{0}})\leq \varepsilon \log_2\left\{\text{Card}(\tX)-1 \right\}
+\end{equation}
+holds for any $ \varepsilon \geq 0$ when $\gamma\in[0,1]$ is chosen to be sufficiently small depending on $\varepsilon$, e.g., $\gamma \leq \sqrt{\varepsilon\log2\over8}$.
+By applying Lemma~\ref{lem:Tsybakov} to~\eqref{eq:totalKL}, and in view of the property (iv), we obtain that
+\begin{equation}\label{eq:final}
+\inf_{\hat \Theta}\sup_{\trueT\in \tX}\mathbb{P}\left(\FnormSize{}{\hat \Theta- \trueT}\geq  {\gamma\over 8} \min\left\{ \alpha\sqrt{d_{\text{total}}}, C^{-1/2}\sqrt{ R_{\max}d_{\max}}\right\} \right)\geq {1\over 2}\left(1-2\varepsilon-\sqrt{16 \varepsilon \over R_{\max}d_{\max}\log2}\right).
+\end{equation}
+Note that $\text{Loss}(\hat \Theta, \trueT)=\FnormSize{}{\hat \Theta- \trueT}^2/d_{\text{total}}$ and $\tX\subset \tP$.
+By taking $\varepsilon=1/10$ and $\gamma=1/11$, we conclude from~\eqref{eq:final} that
+\begin{equation*}\label{eq:prob}
+\inf_{\hat \Theta}\sup_{\trueT\in \tP}\mathbb{P}\left(\text{Loss}(\hat \Theta, \trueT)\geq c\min\left \{ \alpha^2,  {C^{-1}R_{\max}d_{\max}\over d_{\text{total}}}\right \}\right)\geq {1\over 2}\left({4\over 5}- \sqrt{1.6\over R_{\max}d_{\max}\log2} \right)\geq {1\over 8},
+\end{equation*}
+where $c = {1 \over 88^2}$ and the last inequality comes from the condition for $d_{\text{max}}$.
+This completes the proof.
+\end{proof}
+
+\subsection{Sample complexity for tensor completion}
+
+\begin{proof}[Proof of Theorem~\ref{thm:completion}]
+
+For notational convenience, we use $\MFnormSize{}{\Theta}=\sum_{\omega\in\Omega}\Theta^2_\omega$ to denote the sum of squared entries over the observed set $\Omega$, for a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$.
+
+Following a similar argument as in the proof of Theorem~\ref{thm:rate}, we have
+\begin{equation}\label{eq:Taylor2}
+\logl(\Theta)=\logl(\trueT)+\langle\Vec(\plogl),\ \Vec(\Theta-\trueT)\rangle+{1\over 2}\Vec(\Theta-\trueT)^T\pplogl(\check\Theta)\Vec(\Theta-\trueT),
+\end{equation}
+where
+\begin{enumerate}
+\item $\plogl$ is a $d_1\times\cdots\times d_K$ tensor with $|\Omega|$ nonzero entries, and each entry is upper bounded by $U_\alpha>0$.
+\item $\pplogl$ is a diagonal matrix of size $d_{\text{total}}$-by-$d_{\text{total}}$ with $|\Omega|$ nonzero entries, and each entry is upper bounded by $-L_{\alpha}<0$.
+\end{enumerate}
+
+Similar to~\eqref{eq:linear} and~\eqref{eq:quadratic}, we have
+\begin{equation}\label{eq:linear2}
+|\langle\Vec(\plogl),\ \Vec(\Theta-\trueT)\rangle|\leq C_2U_\alpha \sqrt{r_{\max}^{K-1}\sum_k d_k}\MFnormSize{}{\Theta-\trueT}
+\end{equation}
+and
+\begin{equation}\label{eq:quadratic2}
+\Vec(\Theta-\trueT)^T\fpplogl(\check\Theta)\Vec(\Theta-\trueT)\leq -L_\alpha \MFnormSize{}{\Theta-\trueT}^2.
+\end{equation}
+
+Combining~\eqref{eq:Taylor2}-\eqref{eq:quadratic2} with the fact that $\logl(\hat \Theta)\geq \logl(\trueT)$, we have
+\begin{equation}\label{eq:sample}
+\MFnormSize{}{\hat \Theta-\trueT}\leq {2C_2U_\alpha  r_{\max}^{(K-1)/2} \over L_\alpha} \sqrt{\sum_kd_k}.
+\end{equation}
+Lastly, we invoke the result regarding the closeness of $\Theta$ to its sampled version $\Theta_{\Omega}$, under the entrywise bound condition.
+Note that $\mnormSize{}{\hat\Theta-\trueT}\leq 2\alpha$ and $\text{rank}(\hat \Theta-\trueT)\leq 2\mr$.
+By Lemma~\ref{lem:Mnormbound}, $\anormSize{}{\hat \Theta-\trueT}\leq 2^{(3K-1)/2} \alpha \left({\prod r_k \over r_{\max}}\right)^{3/2}$.
+Therefore, the condition in Lemma~\ref{lem:convexity} holds with $\beta=2^{(3K-1)/2}\alpha \left({\prod r_k \over r_{\max}}\right)^{3/2}$.
+Applying Lemma~\ref{lem:convexity} to~\eqref{eq:sample} gives
+\begin{align}
+ \PiFnormSize{}{\hat \Theta-\trueT}^2&\leq {1\over m}\MFnormSize{}{\hat \Theta-\trueT}^2+c\beta\sqrt{\sum_k d_k\over |\Omega|}\\
+ &\leq {C_2  r^{K-1}_{\max}} {\sum_k d_k \over |\Omega|}+C_1 \alpha r_{\max}^{3(K-1)/2}\sqrt{\sum_kd_k\over |\Omega|},
+\end{align}
+with probability at least $1-\exp(-{\sum_kd_k\over \sum_k \log d_k})$ over the sampled set $\Omega$. Here $C_1, C_2>0$ are two constants independent of the tensor dimension and rank. Therefore,
+\[
+ \PiFnormSize{}{\hat \Theta-\trueT}^2\to 0,\quad \text{as}\quad {|\Omega|\over \sum_kd_k}\to \infty,
+\]
+provided that $r_{\max}=O(1)$.
+\end{proof}
+
+\subsection{Convexity of the log-likelihood function}
+\begin{thm}\label{thm:convexity}
+Define the function
+\begin{align}\label{eq:function}
+ \logl(\Theta, \mb)&=\sum_{\omega\in\Omega}\sum_{\ell\in[L]} \left\{\mathds{1}_{\{y_\omega=\ell\}} \log \big[f(b_\ell-\theta_\omega)-  f(b_{\ell-1}-\theta_\omega)\big]\right\},
+ \end{align}
+where $f(\cdot)$ satisfies Assumption~\eqref{eq:ass}. Then, $\logl(\Theta, \mb)$ is concave in $(\Theta,\mb)$.
+\end{thm}
+
+\begin{proof}
+Define $d_{\text{total}}=\prod_k d_k$.
+With a little abuse of notation, we use $(\Theta,\mb)$ to denote a length-$(d_{\text{total}}+L-1)$-vector collecting all parameters together.
+Let us denote a bivariate function
+\begin{align}
+\lambda: &\mathbb{R}^2\mapsto \mathbb{R}\\
+(u,v)&\mapsto\lambda(u,v) = \log \big[f(u)-  f(v)\big].
+\end{align}
+It suffices to show that $\lambda(u,v)$ is concave in $(u,v)$ where $u>v$.
+
+Suppose that the claim holds (which we will prove in the next paragraph). Based on~\eqref{eq:function}, $u,v$ are both linear functions of $(\Theta,\mb)$:
+\[
+u = \ma_1^T(\Theta,\mb), \quad v = \ma_2^T(\Theta,\mb), \quad \text{ for some}\ \ma_1,\ma_2\in \mathbb{R}^{d_{\text{total}}+L-1}.
+\]
+Then, $\lambda(u,v) = \lambda(\ma_1^T(\Theta,\mb),\ \ma_2^T(\Theta,\mb))$ is concave in $(\Theta,\mb)$ by the definition of concavity.
+Furthermore, we can conclude that $\logl(\Theta, \mb) $ is concave in $(\Theta,\mb)$ because $\logl(\Theta, \mb)$ is the sum of $\lambda(u,v)$.
+
+Now, we prove the concavity of $\lambda(u,v)$. Note that
+\begin{equation*}
+\lambda(u,v) = \log\big[f(u)-f(v)\big]=\log\big[\int\mathds{1}_{[u,v]}(x)f'(x)dx\big],
+\end{equation*}
+where $\mathds{1}_{[u,v]}$ is an indicator function that equals 1 in the interval $[u,v]$, and 0 elsewhere.
+Furthermore, $\mathds{1}_{[u,v]}(x)$ is log-concave in $(u,v,x)$ and $f'(x) = \frac{e^x}{(1+e^x)^2}$ is log-concave in $x$.
+The convexities imply that $\mathds{1}_{(u,v)}(x)f'(x)$
+is a log-concave in $(u,v,x)$. By Lemma~\ref{lem:lossconvexity}, we conclude that $\lambda(u,v)$ is concave in $(u,v)$ where $u>v$.
+\end{proof}
+
+\begin{lem}[Corollary 3.5 in~\cite{brascamp2002extensions}]\label{lem:lossconvexity}
+Let $F(x,y):\mathbb{R}^{m+n}\rightarrow \mathbb{R}$ be an integrable function where $x\in \mathbb{R}^{m},y\in \mathbb{R}^n$. Let
+\[
+G(x) = \int_{\mathbb{R}^n}F(x,y)dy.
+\]
+If $F(x,y)$ is log concave in $(x,y)$, then $G(x)$ is log concave in $x$.
+\end{lem}
+
+\subsection{Auxiliary lemmas}
+\label{sec:lemma}
+
+We begin with various notion of tensor norms that are useful for the proofs of the main theorems.
+
+\begin{defn}[Atomic M-norm~\citep{ghadermarzy2019near}]
+Define $T_{\pm}=\{\tT\in\{\pm 1 \}^{d_1\times \cdots \times d_K}\colon \text{rank}(\tT)=1\}$. The atomic M-norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
+\begin{align}
+\anormSize{}{\Theta}&=\inf\{t>0: \Theta\in t\text{conv}(T_{\pm})\}\\
+&=\inf\left\{\sum_{\tX\in T_{\pm}}c_x\colon \Theta=\sum_{\tX\in T_{\pm}} c_x\tX, \ c_x>0\right\}.
+\end{align}
+\end{defn}
+
+\begin{defn}[Spectral norm~\citep{lim2005singular}]
+The spectral norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
+\[
+\snormSize{}{\Theta}=\sup\left\{\langle \Theta, \mx_1\otimes \cdots \otimes \mx_K\rangle \colon \vnormSize{}{\mx_k}=1,\ \mx_k\in\mathbb{R}^{d_k},\ \text{for all}\ k\in[K]\right\}.
+\]
+\end{defn}
+
+\begin{defn}[Nuclear norm~\citep{friedland2018nuclear}]
+The nuclear norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
+\[
+\nnormSize{}{\Theta}=\inf
+\left\{
+\sum_{i\in[r]}|\lambda_i|\colon \Theta=\sum_{i=1}^r \lambda_i\mx^{(i)}_1\otimes \cdots \otimes \mx^{(i)}_K,\ \vnormSize{}{\mx^{(i)}_k}=1,\ \mx^{(i)}_k\in\mathbb{R}^{d_k},\ \text{for all} k\in[K],\ i\in[r]
+\right\},
+\]
+where the infimum is taken over all $r\in\mathbb{N}$ and $\vnormSize{}{\mx^{(i)}_k}=1$ for all $i\in[r]$ and $k\in[K]$.
+\end{defn}
+
+\begin{lem}[M-norm and infinity norm~\citep{ghadermarzy2019near}]\label{lem:Mnormbound}
+Let $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ be an order-$K$, rank-$(r_1,\ldots,r_K)$ tensor. Then
+\[
+\mnormSize{}{\Theta}\leq \anormSize{}{\Theta}\leq \left(\prod_k r_k \over r_{\max}\right)^{3\over 2} \mnormSize{}{\Theta}.
+\]
+\end{lem}
+
+\begin{lem}[Nuclear norm and F-norm] \label{lem:nuclear}
+Let $\tA\in\mathbb{R}^{d_1\times\cdots\times d_K}$ be an order-$K$ tensor with Tucker $\text{rank}(\tA)=(r_1,\ldots,r_K)$. Then
+\[
+\nnormSize{}{\tA} \leq \sqrt{\prod_k r_k\over \max_k r_k} \FnormSize{}{\tA},
+\]
+where $\nnormSize{}{\cdot}$ denotes the nuclear norm of the tensor.
+\end{lem}
+
+\begin{proof}
+Without loss of generality, suppose $r_1=\min_k r_k$.
+Let $\tA_{(k)}$ denote the mode-$k$ matricization of $\tA$ for all $k\in[K]$.
+By \citet[Corollary 4.11]{wang2017operator}, and the invariance relationship between a tensor and its Tucker core~\citep[Section 6]{jiang2017tensor}, we have
+\begin{equation}\label{eq:norminequality}
+\nnormSize{}{\tA} \leq \sqrt{\prod_{k\geq 2} r_k \over \max_{k\geq 2} r_k} \nnormSize{}{\tA_{(1)}},
+\end{equation}
+where $\tA_{(1)}$ is a $d_1$-by-$\prod_{k\geq 2}d_k$ matrix with matrix rank $r_1$.
+Furthermore, the relationship between the matrix norms implies that $\nnormSize{}{\tA_{(1)}}\leq \sqrt{r_1}\FnormSize{}{\tA_{(1)}}=\sqrt{r_1}\FnormSize{}{\tA}$.
+Combining this fact with the inequality~\eqref{eq:norminequality} yields the final claim.
+\end{proof}
+
+\begin{lem} \label{lem:inq}
+Let $\tA, \; \tB$ be two order-$K$ tensors of the same dimension. Then
+\[
+|\langle \tA,\tB\rangle| \leq \snormSize{}{\tA}   \nnormSize{}{\tB}.
+\]
+\end{lem}
+
+\begin{proof}
+By~\citet[Proposition 3.1]{friedland2018nuclear}, there exists a nuclear norm decomposition of $\tB$, such that
+\[
+\tB=\sum_{r} \lambda_r \ma^{(1)}_r\otimes \cdots\otimes \ma^{(K)}_r,\quad \ma_r^{(k)}\in\mathbf{S}^{d_k-1}(\mathbb{R}),\quad \text{for all }k\in[K],
+\]
+and $\nnormSize{}{\tB}=\sum_{r}|\lambda_r|$. Henceforth we have
+\begin{align*}
+|\langle \tA,\tB\rangle|&=| \langle \tA, \sum_{r} \lambda_r \ma^{(1)}_r\otimes \cdots\otimes \ma^{(K)}_r \rangle|\leq \sum_r |\lambda_r| |\langle \tA, \ma^{(1)}_r \otimes \cdots\otimes \ma^{(K)}_r \rangle|\\
+&\leq \sum_{r}|\lambda_r| \snormSize{}{\tA}= \snormSize{}{\tA}\nnormSize{}{\tB},
+\end{align*}
+which completes the proof.
+\end{proof}
+
+The following lemma provides the bound on the spectral norm of random tensors. The result was firstly presented in~\cite{nguyen2015tensor}, and we adopt the version from~\cite{tomioka2014spectral}.
+\begin{lem}[\cite{tomioka2014spectral}]\label{lem:tensor}
+Suppose that $\tS=\entry{s_{\omega}}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is an order-$K$ tensor whose entries are independent random variables that satisfy
+\[
+\mathbb{E}(s_{\omega})=0,\quad \text{and} \quad\mathbb{E}(e^{ts_{\omega}})\leq e^{t^2L^2/2}.
+\]
+Then the spectral norm $\snormSize{}{\tS}$ satisfies that,
+\[
+\snormSize{}{\tS}\leq \sqrt{{8L^2} \log (12K) \sum_k d_k +\log (2/\delta)},
+\]
+with probability at least $1-\delta$.
+\end{lem}
+
+\begin{lem} \label{lem:noisytensor}
+Suppose that $\tS=\entry{s_{\omega}}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is an order-$K$ tensor whose entries are independent random variables that satisfy
+\[
+\mathbb{E}(s_{\omega})=0,\quad \text{and}\quad |s_{\omega}|\leq U.
+\]
+Then we have
+\[
+\mathbb{P}\left(\snormSize{}{\tS}\geq C_2 U\sqrt{\sum_k d_k} \right)\leq \exp\left(-C_1  \log K \sum_k d_k\right)
+\]
+where $C_1>0$ is an absolute constant, and $C_2>0$ is a constant that depends only on $K$.
+\end{lem}
+
+\begin{proof}  Note that the random variable $U^{-1}s_{\omega}$ is zero-mean and supported on $[-1,1]$. Therefore, $U^{-1}s_{\omega}$ is sub-Gaussian with parameter ${1-(-1)\over 2}=1$, i.e.
+\[
+\mathbb{E}(U^{-1}s_{\omega})=0,\quad \text{and}\quad \mathbb{E}(e^{tU^{-1}s_{\omega}})\leq e^{t^2/2}.
+\]
+It follows from Lemma~\ref{lem:tensor} that, with probability at least $1-\delta$,
+\[
+\snormSize{}{U^{-1}\tS}\leq \sqrt{\left(c_0\log K+c_1\right) \sum_k d_k +\log (2/\delta)},
+\]
+where $c_0, c_1>0$ are two absolute constants. Taking $\delta=\exp (-C_1\log K \sum_k d_k)$ yields the final claim, where $C_2=c_0\log K+c_1+1>0$ is another constant.
+\end{proof}
+
+\begin{lem}\label{lem:KLentry} Let $X,\; Y$ be two discrete random variables taking values on $L$ possible categories, with category probabilities $\{p_\ell\}_{\ell\in[L]}$ and $\{q_\ell\}_{\ell\in[L]}$, respectively.
+Suppose $p_\ell$, $q_\ell>0$ for all $i\in[L]$.
+Then, the Kullback-Leibler (KL) divergence satisfies that
+\[
+\mathrm{KL}(X||Y)\stackrel{\text{def}}{=}-\sum_{\ell\in[L]}\mathbb{P}_X(\ell)\log\left\{{\mathbb{P}_Y(\ell)\over \mathbb{P}_X(\ell)}\right\} \leq \sum_{\ell \in [L]}{(p_\ell-q_\ell)^2 \over q_\ell}.
+\]
+\end{lem}
+\begin{proof} Using the fact $\log x\leq x-1$ for $x>0$, we have that
+\begin{align}\label{eq:KL}
+\text{KL}(X||Y)&=\sum_{\ell\in[L]}p_\ell\log{p_\ell\over q_\ell}\\
+&\leq \sum_{\ell\in[L]} {p_\ell\over q_\ell}(p_\ell-q_\ell)\\
+&=\sum_{\ell\in[L]} \left({p_\ell\over q_\ell }- 1\right)(p_\ell-q_\ell)+ \sum_{\ell\in[L]} (p_\ell-q_\ell).
+\end{align}
+Note that $\sum_{\ell\in[L]}(p_\ell-q_\ell)=0$. Therefore,
+\[
+\text{KL}(X||Y)\leq \sum_{\ell\in[L]}\left( {p_\ell\over q_\ell}-1\right)\left(p_\ell-q_\ell\right)=\sum_{\ell\in[L]}{(p_\ell-q_\ell)^2\over q_\ell}.
+\]
+\end{proof}
+
+\begin{lem}[KL divergence and F-norm]~\label{lem:KL}
+Let $\tY\in[L]^{d_1\times \cdots \times d_K}$ be an ordinal tensor generated from the model~\eqref{eq:model} with the link function $f$ and parameter tensor $\Theta$.
+Let $\mathbb{P}_{\Theta}$ denote the joint categorical distribution of $\tY|\Theta$ induced by the parameter tensor $\Theta$, where $\mnormSize{}{\Theta}\leq \alpha$.
+Define
+\begin{equation}\label{eq:ass}
+A_\alpha=\min_{\ell\in[L], |\theta|\leq \alpha}\left[f(b_\ell-\theta)-f(b_{\ell-1}-\theta)\right].
+\end{equation}
+Then, for any two tensors $\Theta,\; \Theta^*$ in the parameter spaces, we have
+\[
+\mathrm{KL}(\mathbb{P}_{\Theta}|| \mathbb{P}_{\Theta^*})\leq {2(2L-3)\over A_\alpha} \dot{f}^2(0)\FnormSize{}{\Theta-\Theta^*}^2.
+\]
+\end{lem}
+\begin{proof} Suppose that the distribution over the ordinal tensor $\tY=\entry{y_\omega}$ is induced by $\Theta=\entry{\theta_\omega}$. Then, based on the generative model~\eqref{eq:model},
+\[
+\mathbb{P}(y_\omega=\ell | \theta_\omega)=f(b_{\ell}-\theta_\omega)-f(b_{\ell-1}-\theta_\omega),
+\]
+for all $\ell\in[L]$ and $\omega\in[d_1]\times \cdots \times [d_K]$.
+For notational convenience, we suppress the subscribe in $\theta_\omega$ and simply write $\theta$ (and respectively, $\theta^*$).
+Based on Lemma~\ref{lem:KLentry} and Taylor expansion,
+\begin{align}
+\text{KL}(\theta|| \theta^*) & \leq \sum_{\ell\in[L]} {\left[f(b_{\ell}-\theta)-f(b_{\ell-1}-\theta)-f(b_{\ell}-\theta^*)+f(b_{\ell-1}-\theta^*)\right]^2\over f(b_{\ell}-\theta^*)-f(b_{\ell-1}-\theta^*)}\\
+ &\leq \sum_{\ell=2}^{L-1} {\left[\dot{f}(b_\ell-\eta_\ell)-\dot{f}(b_{\ell-1}-\eta_{\ell-1})\right]^2 \over f(b_\ell-\theta^*)-f(b_{\ell-1}-\theta^*)} \left(\theta-\theta^*\right)^2+{\dot{f}^2(b_1-\eta_1) \over f(b_1-\theta^*)} (\theta-\theta^*)^2\\
+ &\quad \quad\quad \quad  +{\dot{f}^2(b_{L-1}-\eta_{L-1})\over 1-f(b_{L-1}-\theta^*)}(\theta-\theta^*)^2,
+\end{align}
+where $\eta_\ell$ and $\eta_{\ell-1}$ fall between $\theta$ and $\theta^*$. Therefore,
+\begin{equation}\label{eq:entrywise}
+\text{KL}(\theta|| \theta^*) \leq \left({4(L-2)\over A_\alpha}+ {2\over A_\alpha}\right)\dot{f}^2(0)(\theta-\theta^*)^2={2(2L-3)\over A_\alpha} \dot{f}^2(0)(\theta-\theta^*)^2,
+\end{equation}
+where we have used Taylor expansion, the bound~\eqref{eq:ass}, and the fact that $\dot{f}(\cdot)$ peaks at zero for an unimodal and symmetric function.
+Now summing~\eqref{eq:entrywise} over the index set $\omega\in[d_1]\times \cdots \times [d_K]$ gives
+\[
+\text{KL}(\mathbb{P}_{\Theta}|| \mathbb{P}_{\Theta^*}) =\sum_{\omega\in[d_1]\times \cdots \times[d_K]} \text{KL}(\theta_\omega || \theta^*_\omega) \leq {2(2L-3)\over A_\alpha} \dot{f}^2(0)\FnormSize{}{\Theta-\Theta^*}^2.
+\]
+\end{proof}
+
+\begin{rmk} In particular, let $\mathbb{P}_{\bf{0}}$ denote the distribution of $\tY|\bf{0}$ induced by the zero parameter tensor. Then we have
+\[
+\text{KL}(\mathbb{P}_{\Theta}||\mathbb{P}_{\bf{0}})\leq {2(2L-3)\over A_\alpha}  \dot{f}^2(0)\FnormSize{}{\Theta}^2.
+\]
+\end{rmk}
+
+\begin{lem}\label{lem:construction}
+Assume the same setup as in Theorem~\ref{thm:minimax}.
+Without loss of generality, suppose $d_1=\max_kd_k$.
+Define $R=\max_k r_k$ and $d_{\text{total}}=\prod_{k\in[K]} d_k$.
+For any constant $0\leq \gamma \leq 1$, there exist a finite set of tensors $\tX=\{\Theta_i: i=1,\ldots\}\subset \tP$ satisfying the following four properties:
+\begin{enumerate}[(i)]
+\item $\text{Card}(\tX)\geq 2^{Rd_1/8}+1$, where $\text{Card}$ denotes the cardinality;
+\item $\tX$ contains the zero tensor $\mathbf{0}\in\mathbb{R}^{d_1\times \cdots\times d_K}$;
+\item $\mnormSize{}{\Theta}\leq \gamma \min\left\{ \alpha ,\ C^{-1/2}\sqrt{Rd_1\over d_{\text{total}}} \right\} $ for any element $\Theta\in\tX$;
+\item $\FnormSize{}{\Theta_i-\Theta_j}\geq {\gamma\over 4} \min\left\{ \alpha\sqrt{d_{\text{total}}},\ C^{-1/2}\sqrt{Rd_1}\right\}$ for any two distinct elements $\Theta_i,\; \Theta_j\in\tX$,
+\end{enumerate}
+Here $C=C(\alpha,L,f,\mb)={(4L-6)\dot{f}^2(0)\over A_\alpha }>0$ is a constant independent of the tensor dimension and rank.
+\end{lem}
+
+\begin{proof}
+Given a constant $0\leq \gamma \leq 1$, we define a set of matrices:
+\[
+\tC=\left\{\mM=(m_{ij})\in\mathbb{R}^{d_1\times R}: a_{ij}\in \left\{ 0,\gamma \min\left\{ \alpha , C^{-1/2}\sqrt{Rd_1\over d_{\text{total}}} \right\}\right\} ,\  \forall (i,j)\in[d_1]\times[R]\right\}.
+\]
+We then consider the associated set of block tensors:
+\begin{align}
+\tB=\tB(\tC)=\{\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}\colon& \Theta=\mA\otimes \mathbf{1}_{d_3}\otimes \cdots \otimes \mathbf{1}_{d_K}, \\
+&\ \text{where}\ \mA=(\mM|\cdots|\mM|\mO) \in\mathbb{R}^{d_1\times d_2},\ \mM\in\tC\},
+\end{align}
+where $\mathbf{1}_d$ denotes a length-$d$ vector with all entries 1, $\mO$ denotes the $d_1\times (d_2-R\lfloor d_2/R \rfloor)$ zero matrix, and $\lfloor d_2/ R \rfloor$ is the integer part of $d_2/R$.
+In other words, the subtensor $\Theta(\mI, \mI,i_3, \ldots,i_K)\in\mathbb{R}^{d_1\times d_2}$ are the same for all fixed $(i_3,\ldots,i_K)\in[d_3]\times \cdots \times [d_K]$, and furthermore, each subtensor $\Theta(\mI,\mI, i_3,\ldots,i_K)$ itself is filled by copying the matrix $\mM\in\mathbb{R}^{d_1\times R}$ as many times as would fit.
+
+By construction, any element of $\tB$, as well as the difference of any two elements of $\tB$, has Tucker rank at most $\max_k r_k\leq R$, and the entries of any tensor in $\tB$ take values in $[0,\alpha]$.
+Thus, $\tB\subset\tP$.
+By Lemma~\ref{lem:VGbound}, there exists a subset $\tX\subset \tB$ with cardinality $\text{Card}(\tX)\geq 2^{Rd_1/8}+1$ containing the zero $d_1\times \cdots \times d_K$ tensor, such that, for any two distinct elements $\Theta_i$ and $\Theta_j$ in $\tX$,
+\[
+\FnormSize{}{\Theta_i-\Theta_j}^2 \geq {Rd_1\over 8} \gamma^2\min\left\{ \alpha^2, {C^{-1}Rd_1 \over d_{\text{total}}}\right\} \lfloor {d_2\over R} \rfloor \prod_{k\geq 3}d_k\geq {\gamma^2\min\left\{ \alpha^2 d_{\text{total}}, C^{-1}Rd_1\right\}  \over 16}.
+\]
+In addition, each entry of $\Theta\in\tX$ is bounded by $\gamma \min\left\{ \alpha , C^{-1/2}\sqrt{Rd_1\over d_{\text{total}}}\right\} $. Therefore the Properties (i) to (iv) are satisfied.
+\end{proof}
+
+\begin{lem}[Varshamov-Gilbert bound]\label{lem:VGbound}
+Let $\Omega=\{(w_1,\ldots,w_m)\colon w_i\in\{0,1\}\}$. Suppose $m>8$. Then there exists a subset $\{w^{(0)},\ldots,w^{(M)}\}$ of $\Omega$ such that $w^{(0)}=(0,\ldots,0)$ and
+\[
+\zeronormSize{}{w^{(j)}-w^{(k)}}\geq {m\over 8},\quad \text{for} \ 0\leq j<k\leq M,
+\]
+where $\zeronormSize{}{\cdot}$ denotes the Hamming distance, and $M\geq 2^{m/8}$.
+\end{lem}
+
+\begin{defn}[Absolutely continuous]
+A measure $\lambda$ is absolutely continuous with respect to another measure $\mu$ if $\lambda(E)=0$ for every set with $\mu(E)=0$ and is denoted as $\lambda\ll\mu$.
+\end{defn}
+
+\begin{lem}[Theorem 2.5 in~\cite{tsybakov2009introduction}]\label{lem:Tsybakov}
+Assume that a set $\tX$ contains element $\Theta_0, \Theta_1, \ldots,\Theta_M$ ($M\geq 2$) such that
+\begin{itemize}
+\item $d(\Theta_j,\ \Theta_k)\geq 2s>0$, $\forall 0\leq j\leq k\leq M$;
+\item $\mathbb{P}_j\ll\mathbb{P}_0$, $\forall j=1,\ldots,M$, and
+\[
+{1\over M}\sum_{j=1}^M \mathrm{KL}(\mathbb{P}_j||\mathbb{P}_0)\leq \alpha \log M
+\]
+where $d\colon \tX\times \tX\mapsto [0,+\infty]$ is a semi-distance function, $0<\alpha<{1/8}$ and $P_j=P_{\Theta_j}$, $j=0,1\ldots,M$.
+\end{itemize}
+Then
+\[
+\inf_{\hat \Theta}\sup_{\Theta\in\tX} \mathbb{P}_{\Theta}(d(\hat \Theta, \Theta)\geq s)\geq {\sqrt{M}\over 1+\sqrt{M}}\left(1-2\alpha-\sqrt{2\alpha\over \log M} \right)>0.
+\]
+\end{lem}
+
+\begin{lem}[Lemma 28 in~\cite{ghadermarzy2019near}]\label{lem:convexity}
+Define $\mathbb{B}_{M}(\beta)=\{\Theta\in \mathbb{R}^{d_1\times \cdots \times d_K}\colon \anormSize{}{\Theta}\leq \beta \}$.
+Let $\Omega\subset[d_1]\times\cdots \times [d_K]$ be a random set with $m=|\Omega|$, and assume that each entry in $\Omega$ is drawn with replacement from $[d_1]\times\cdots\times[d_K]$ using probability $\Pi$.
+Define
+\[
+\PiFnormSize{}{\Theta}^2={1\over m}\mathbb{E}_{\Omega\in\Pi}\MFnormSize{}{\Theta}^2.
+\]
+Then, there exists a universal constant $c>0$, such that, with probability at least $1-\exp\left(-{\sum_kd_k \over \sum_k \log d_k} \right)$ over the sampled set $\Omega$,
+\[
+{1\over m}\MFnormSize{}{\Theta}^2 \geq \PiFnormSize{}{\Theta}^2-c\beta\sqrt{\sum_k d_k\over m}
+\]
+holds uniformly for all $\Theta\in\mathbb{B}_M(\beta)$.
+\end{lem}
+
+\appendix
+\section{Extension of Theorem~\ref{thm:rate} to unknown cut-off points}
+
+When the cut-off points $\mb$ is unknown, we estimate $(\hat \Theta,\hat \mb)$ by
+\begin{equation}\label{eq:joint}
+(\hat \Theta,\hat \mb)=\argmax_{(\Theta,\mb)\in\tP\times \tB}\logl(\Theta,\mb),
+\end{equation}
+where
+\[
+\tP=\{\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}\colon \text{rank}(\tP)\leq \mr,\ \mnormSize{}{\Theta}\leq \alpha\}, \quad \tB=\{\mb\in\mathbb{R}^{L-1}\colon \mnormSize{}{\mb}\leq \beta,\ \min_{\ell}(b_\ell-b_{\ell-1}) \geq \Delta\}.
+\]
+The estimation accuracy is assessed using the mean squared error (MSE):
+\[
+\text{MSE}(\hat \Theta,\Theta)={1\over \prod_k d_k}\FnormSize{}{\hat \Theta-\Theta},\quad \text{MSE}(\hat \mb,\mb)={1\over L-1}\FnormSize{}{\hat \mb-\mb}.
+\]
+
+We introduce several quantities that will used in our theory:
+\begin{enumerate}
+\item We make the convention that $b_0=-\infty$, $b_L=\infty$, $f(-\infty)=0$, $f(\infty)=1$, and $\dot{f}(-\infty)=\ddot{f}(-\infty)=\dot{f}(\infty)=\ddot{f}(\infty)=0$.
+
+\item The difference function $g_\ell(\theta)$ is defined as $g_\ell(\theta)=f(b_\ell-\theta)-f(b_{\ell-1}-\theta)$ for all $\theta\in\mathbb{R}$ and $\ell\in[L]$.
+
+\item Define $n_\ell=\sum_{\omega\in\Omega}\mathds{1}_{\{y_\omega=\ell\}}$, i.e., the number of tensor entries taking value on $\ell\in[L]$.
+
+\item With a little abuse of notation, we re-define the constants in~\eqref{eq:bound} as
+\begin{equation}\label{eq:UL}
+U_{\alpha,\beta,\Delta}=\max_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}}\max_{\ell\in[L-1]}{|\dot{g}_\ell(\theta)|\over g_\ell(\theta)}, \quad \text{and}\quad
+L_{\alpha,\beta,\Delta}=\min_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}}\min_{\ell\in[L-1]} {\dot{g}^2_\ell(\theta)-\ddot{g}_\ell(\theta)g_\ell(\theta)\over g^2_\ell(\theta)}.
+\end{equation}
+\item We define two additional constants:
+\begin{align}\label{eq:CD}
+C_{\alpha,\beta,\Delta} &= \max_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}}  \max_{\ell\in[L-1]}\max\left\{\frac{\dot{f}(b_\ell-\theta)}{g_\ell(\theta)},\ \frac{\dot{f}(b_\ell-\theta)}{g_{\ell+1}(\theta)}\right\},\notag \\
+D_{\alpha,\beta,\Delta} &=\min_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}} \min_{\ell\in[L-1]}\min\left\{  {\partial \over \partial \theta}\left({\dot{f}(b_{\ell}-\theta)\over g_\ell(\theta)}\right),\ -{\partial \over \partial \theta}\left({\dot{f}(b_{\ell}-\theta)\over g_{\ell+1}(\theta)}\right)\right\}\notag \\
+&=  \min_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}}\min_{\ell\in[L-1]}\min \left\{-\frac{\ddot{f}(b_\ell-\theta)g_\ell(\theta)-\dot{f}(b_\ell-\theta)\dot{g}_\ell(\theta)}{g^2_\ell(\theta)},\ \right.
+\notag \\
+&\hspace{1.5in} \left. \frac{\ddot{f}(b_\ell-\theta)g_{\ell+1}(\theta)-\dot{f}(b_\ell-\theta)\dot{g}_{\ell+1}(\theta)}{g^2_{\ell+1}(\theta)}\right\}.
+\end{align}
+\end{enumerate}
+
+We make the following assumptions about the link function.
+\begin{assumption}\label{ass:joint}
+The link function $f\colon \mathbb{R}\mapsto [0,1]$ satisfies the following properties:
+\begin{enumerate}
+\item $f(\theta)$ is twice-differentiable and strictly increasing in $\theta$.
+\item $\dot{f}(\theta)$ is strictly log-concave and symmetric with respect to $\theta=0$.
+\item The function ${\dot{f}(b_\ell-\theta)\over g_\ell(\theta)}$ is strictly increasing with respect to $\theta$ for all $\mb\in\tB$.
+\item The function ${\dot{f}(b_\ell-\theta)\over g_{\ell+1}(\theta)}$ is strictly decreasing with respect to $\theta$ for all $\mb\in\tB$.
+\end{enumerate}
+\end{assumption}
+
+\begin{rmk}
+The condition $\Delta=\min_{\ell}(b_{\ell}-b_{\ell-1})>0$ on the feasible set $\tB$ guarantees the strict positiveness of $g_{\ell}(\theta)=f(b_{\ell}-\theta)-f(b_{\ell-1}-\theta)$.
+Therefore, the denominators in the above quantities $U_{\alpha,\beta,\Delta}, L_{\alpha,\beta,\Delta}, C_{\alpha,\beta,\Delta}, D_{\alpha,\beta,\Delta}$ are well-defined.
+Furthermore, by Theorem~\ref{thm:convexity}, $g_\ell(\theta)$ is strictly log-concave in $\theta$ for all $\mb\in\tB$.
+Based on Assumption~\ref{ass:joint} and closeness of the feasible set, we have $U_{\alpha,\beta,\Delta}>0$, $L_{\alpha,\beta,\Delta}>0$, $C_{\alpha,\beta,\Delta}>0$, $D_{\alpha,\beta,\Delta}>0$.
+\end{rmk}
+\begin{rmk}
+In particular, for logistic link $f(x) = \frac{1}{1+e^{-x}}$ , we have
+\begin{align}
+C_{\alpha,\beta,\Delta}&=\max_{\substack{|\theta|\leq \alpha,\\ \mb\in\tB}}\max_{\ell\in[L-1]}\max
+\left\{
+{ 1\over e^{b_{\ell}-b_{\ell-1}}-1} { 1+e^{-(b_{\ell-1}-\theta)} \over 1+e^{-(b_{\ell}-\theta)}},\
+{ 1\over 1-e^{-(b_{\ell+1}-b_{\ell})}} { 1+e^{-(b_{\ell+1}-\theta)} \over 1+e^{-(b_{\ell}-\theta)}}
+\right\}>0,
+\\
+D_{\alpha,\beta,\Delta}&=\min_{\substack{|\theta|\leq \alpha,\\ \mb\in\tB}}\min_{\ell\in[L-1]} {e^{-(b_\ell-\theta)}\over (1+e^{-(b_\ell-\theta)})^2}>0.
+\end{align}
+\end{rmk}
+
+\begin{thm}[Statistical convergence with unknown $\mb$]\label{thm:ratejoint}
+Consider an ordinal tensor $\tY\in[L]^{d_1\times \cdots \times d_K}$ generated from model~\eqref{eq:model} with the link function $f$ and parameters $(\trueT, \trueb)\in\tP\times \tB$.
+Suppose the link function $f$ satisfies Assumption~\ref{ass:joint}.
+Define $r_{\max}=\max_k r_k$.
+Then with very high probability, the estimator in~\eqref{eq:joint} satisfies
+\[
+\text{MSE}(\hat \Theta, \trueT)\leq {c_1U_{\alpha,\beta,\Delta}\over L_{\alpha,\beta,\Delta}} r_{\max}^{(K-1)/2}  \sqrt{ \sum_k d_k \over \prod_k d_k},
+\]
+and
+\[
+\text{MSE}(\hat\mb, \trueb) \leq  {c_2 C_{\alpha,\beta,\Delta} \over D_{\alpha,\beta,\Delta}}{\max_{\ell}(n_\ell+n_{\ell+1})\over \min_{\ell}(n_{\ell}+n_{\ell+1})}  \sqrt{1\over L-1},
+\]
+where $c_1,c_2>0$ are two constants independent of the tensor dimension and rank, and $U_{\alpha,\beta,\Delta}$, $L_{\alpha,\beta,\Delta}$, $C_{\alpha,\beta,\Delta}$, $D_{\alpha,\beta,\Delta}>0$ are constants defined in~\eqref{eq:UL} and~\eqref{eq:CD}.
+\end{thm}
+
+\begin{rmk} It may seem counter-intuitive that the MSE of $\hat \Theta$ shows no dependence of $\hat \mb$ (and vice versa).
+In fact, from the proof of Theorem~\ref{thm:ratejoint}, the uniform constants $U_{\alpha,\beta,\Delta}, L_{\alpha,\beta,\Delta}$ can be replaced by
+\[
+U_{\alpha,\hat \mb}=\max_{|\theta|\leq \alpha}\max_{\ell\in[L-1]}{|\dot{g}_\ell(\theta)|\over g_\ell(\theta)},\quad \text{and}\quad L_{\alpha,\hat \mb}=\min_{|\theta|\leq \alpha}\min_{\ell\in[L-1]} {\dot{g}^2_\ell(\theta)-\ddot{g}_\ell(\theta)g_\ell(\theta)\over g^2_\ell(\theta)}.
+\]
+Therefore, the accuracy of $\hat \Theta$ depends on the accuracy of $\hat \mb$ implicitly through $U_{\alpha, \hat \mb}$ and $ L_{\alpha,\hat \mb}$.
+
+\end{rmk}
+
+\begin{proof}[Proof of Theorem~\ref{thm:ratejoint}]
+Similar to the proof of Theorem~\ref{thm:rate}, we suppress $\Omega$ in the subscript. Based on the definition of $(\hat \Theta, \hat \mb)$,
+we have the following inequalities:
+\begin{equation}\label{eq:level}
+\flogl(\hat \Theta,\hat \mb)\geq \flogl(\trueT,\hat \mb)\quad \text{and}\quad \flogl(\hat \Theta,\hat \mb)\geq \flogl(\hat \Theta, \trueb).
+\end{equation}
+
+Following the same argument as in Theorem~\ref{thm:rate} and the first inequality in~\eqref{eq:level}, we obtain that
+\begin{equation}\label{eq:ratenew}
+\text{MSE}(\hat \Theta, \trueT)\leq {CU_{\alpha,\beta,\Delta}\over L_{\alpha,\beta,\Delta}} r_{\max}^{(K-1)/2}  \sqrt{ \sum_k d_k \over \prod_k d_k},
+\end{equation}
+where $U_{\alpha,\beta,\Delta}, L_{\alpha,\beta,\Delta}>0$ are two universal constants independent of the tensor dimension and rank.
+Next we bound $\text{MSE}(\hat \mb,\trueb)$. It follows from the expression of $\flogl(\Theta,\mb)$ that
+\begin{align}\label{eq:property}
+{\partial \flogl\over \partial b_\ell}&=\sum_{\omega\in\Omega}\left[\mathds{1}_{\{y_{\omega}=\ell\}}
+\frac{\dot{f}(b_\ell-\theta_\omega)}{g_\ell(\theta_\omega)}-\mathds{1}_{\{y_{\omega}=\ell+1\}}
+\frac{\dot{f}(b_\ell-\theta_\omega)}{g_{\ell+1}(\theta_\omega)}\right],\notag\\
+{\partial^2 \flogl\over \partial b_\ell^2}&=\sum_{\omega\in\Omega}\left[\mathds{1}_{\{y_\omega=\ell\}}\frac{\ddot{f}(b_\ell-\theta_\omega)g_\ell(\theta_\omega)-\dot{f}(b_\ell-\theta_\omega)^2}{g^2_\ell(\theta_\omega)}-\mathds{1}_{\{y_\omega=\ell+1\}}\frac{\ddot{f}(b_\ell-\theta_\omega)g_{\ell+1}(\theta_\omega)+\dot{f}(b_\ell-\theta_\omega)^2}{g^2_{\ell+1}(\theta_\omega)}\right],\notag\\
+&\quad  \text{ for all } \ell \in [L-1],\notag \\
+{\partial^2 \flogl\over \partial b_\ell \partial b_{\ell+1}}&=\sum_{\omega\in\Omega}\mathds{1}_{\{y_\omega=\ell+1\}}\frac{\dot{f}(b_\ell-\theta_\omega)\dot{f}(b_{\ell+1}-\theta_\omega)}{g^2_{\ell+1}(\theta_\omega)}\quad \text{and}\quad
+\ {\partial^2 \flogl\over \partial b_\ell \partial b_\ell'} = 0 \text{ if } |\ell-\ell'|>1.
+\end{align}
+
+Therefore, all entries in $\fploglb$ are upper bounded by $\left\{C_{\alpha,\beta,\Delta} \max_\ell (n_\ell+n_{\ell+1})\right\}>0$, and $\fpploglb$ is a tridiagonal matrix.
+
+We consider the profile log-likelihood $\flogl(\mb, \hat \Theta)$ as a function of $\mb\in\tB$.
+For notational convenience, we drop $\hat \Theta$ from $\flogl(\mb, \hat \Theta)$ and simply write $\flogl(\mb)$.
+By the second-order Taylor's expansion of $\flogl(\mb)$ around $\trueb$, we obtain
+\begin{equation}\label{eq:taylorb}
+\flogl(\hat\mb)=\flogl(\trueb)+(\trueb-\hat\mb)^T\fploglb(\trueb)+ {1\over 2}(\trueb-\hat\mb)^T\fpploglb(\check\mb)(\trueb-\hat\mb),
+\end{equation}
+where $\check\mb=\gamma\trueb+(1-\gamma)\hat\mb$ for some $\gamma\in[0,1]$, and $\fpploglb(\check\mb)$ denotes the $(L-1)$-by-$(L-1)$ Hession matrix evaluated at $\check\mb$.
+
+The linear term in~\eqref{eq:taylorb} can be bounded by Cauchy-Schwartz inequality,
+\begin{equation}\label{eq:linearb}
+(\trueb-\hat\mb)^T\fploglb(\trueb)\leq \FnormSize{}{\trueb-\hat\mb}\FnormSize{}{\fploglb(\trueb)}
+\leq \FnormSize{}{\trueb-\hat\mb}\sqrt{L-1}C_{\alpha,\beta,\Delta}\max_{\ell\in[L-1]}(n_\ell+n_{\ell+1}),
+\end{equation}
+where the last inequality is followed from
+\[
+\left|{\partial \tL_\tY\over \partial b_\ell}\Big|_{\mb=\trueb}\right| \leq C_{\alpha,\beta,\Delta}\max_{\ell\in[L-1]}(n_{\ell}+n_{\ell+1}), \quad \textrm{ for all } \; \ell\in [L-1].
+\]
+
+We next bound the quadratic term in \eqref{eq:taylorb}. Note that
+\begin{align}\label{eq:quadraticb}
+ &(\trueb-\hat\mb)^T \fpploglb(\check{\mb})(\trueb-\hat\mb)\\
+ &=\sum_{\ell\in [L-1]} \left( {\partial^2\tL_{\tY}\over \partial b^2_\ell} \Big|_{\mb=\check\mb} \right)(\hat b_\ell-b^{\text{true}}_{\ell})^2 + 2\sum_{\ell\in[L-1]/\{1\}}\left( {\partial^2\tL_{\tY}\over \partial b_\ell \partial b_{\ell-1}} \Big|_{\mb=\check\mb} \right)(\hat b_\ell-b^{\text{true}}_{\ell})(\hat b_{\ell-1}-b^{\text{true}}_{\ell-1}) \nonumber \\
+&\leq \sum_{\ell\in [L-1]} \left( {\partial^2\tL_{\tY}\over \partial b^2_\ell} \Big|_{\mb=\check\mb} \right)(\hat b_\ell-b^{\text{true}}_{\ell})^2 + \sum_{\ell\in[L-1]/\{1\}}\left( {\partial^2\tL_{\tY}\over \partial b_\ell \partial b_{\ell-1}} \Big|_{\mb=\check\mb} \right)\big[(\hat b_\ell-b^{\text{true}}_{\ell})^2 +(\hat b_{\ell-1}-b^{\text{true}}_{\ell-1})^2\big] \nonumber \\
+&=\left( {\partial^2\tL_{\tY}\over \partial b^2_1}+{\partial^2\tL_{\tY}\over \partial b_1 \partial b_2 }\right)\Big|_{\mb=\check\mb}(\hat b_1-b^{\text{true}}_{1})^2+\left( {\partial^2\tL_{\tY}\over \partial b^2_{L-1}}+{\partial^2\tL_{\tY}\over \partial b_{L-2} \partial b_{L-1} }\right)\Big|_{\mb=\check\mb}(\hat b_{L-1}-b^{\text{true}}_{L-1})^2  \\
+&\quad\quad +\sum_{\ell\in[L-2]/\{1\}}\left( {\partial^2\tL_{\tY}\over \partial b^2_\ell}+{\partial^2\tL_{\tY}\over \partial b_\ell \partial b_{\ell-1}}+{\partial^2\tL_{\tY}\over \partial b_{\ell+1} \partial b_{\ell}} \right)\Big|_{\mb=\check\mb} (\hat b_\ell-b^{\text{true}}_{\ell})^2     \\
+&\leq - \tilde{D}_{\alpha,\beta,\Delta} \sum_{\ell \in [L-1]}(\hat b_\ell-b_{\text{true},\ell})^2 \nonumber \\
+&=-\tilde{D}_{\alpha,\beta,\Delta}\FnormSize{}{\hat\mb-\trueb}^2,
+\end{align}
+where
+\begin{align}
+\tilde{D}_{\alpha,\beta,\Delta}&=\min_{\substack{|\theta|\leq \alpha,\\ \mb\in\tB}}\min_{\ell\in[L-1]} -\left( {\partial^2\tL_{\tY}\over \partial b^2_\ell}+{\partial^2\tL_{\tY}\over \partial b_\ell \partial b_{\ell-1}}+{\partial^2\tL_{\tY}\over \partial b_{\ell+1} \partial b_{\ell}} \right)\\
+&=\min_{\substack{|\theta|\leq \alpha,\\ \mb\in\tB}} \min_{\ell\in[L-1]}\left\{\sum_{\omega\in\Omega}-\mathds{1}_{\{y_\omega=\ell\}}\left( {\ddot{f}(b_{\ell}-\theta_\omega) g_\ell(\theta_\omega)- \dot{f}(b_{\ell}-\theta_\omega)\dot{g}_\ell(\theta_\omega) \over g^2_\ell(\theta_\omega) }\right)\right.
+\\
+&\left.
+\hspace{1in}+\sum_{\omega\in\Omega}\mathds{1}_{\{y_\omega=\ell+1\}} \left( {\ddot{f}(b_\ell-\theta_\omega) g_{\ell+1}(\theta_\omega) - \dot{f}(b_\ell-\theta_\omega)\dot{g}_{\ell+1} (\theta_\omega)\over g^2_{\ell+1}(\theta_\omega)} \right)\right\}\\
+&=\min_{\substack{|\theta|\leq \alpha,\\ \mb\in\tB}} \min_{\ell\in[L-1]}\left\{\KeepStyleUnderBrace{-\sum_{\omega\in\Omega} \mathds{1}_{\{y_\omega=\ell\}}{\partial \over \partial \theta_\omega}\left( \dot{f}(b_\ell-\theta_\omega) \over g_\ell(\theta_\omega) \right)}_{>0}+\KeepStyleUnderBrace{\sum_{\omega\in\Omega}\mathds{1}_{\{y_\omega=\ell+1\}} {\partial \over \partial \theta_\omega}\left( \dot{f}(b_\ell-\theta_\omega) \over g_{\ell+1}(\theta_\omega) \right) }_{>0}\right\}\\
+&\geq D_{\alpha,\beta,\Delta}  \min_{\ell\in[L-1]}\left(n_{\ell}+n_{\ell+1}\right) .
+\end{align}
+
+Combining inequalities~\eqref{eq:taylorb}, \eqref{eq:linearb} and~\eqref{eq:quadraticb} yields
+\[
+\flogl(\hat \mb)\leq \flogl(\trueb)+C_{\alpha,\beta,\Delta}\sqrt{L-1}\max_{\ell}(n_\ell+n_{\ell+1}) \FnormSize{}{\hat\mb-\trueb} -\frac{D_{\alpha,\beta,\Delta}}{2}\min_{\ell}(n_\ell+n_{\ell+1}) \FnormSize{}{\hat\mb-\trueb}^2 .
+\]
+Since $\hat \mb$ satisfies $\flogl(\hat \mb)-\flogl(\trueb)\geq 0$, we have that
+\[
+C_{\alpha,\beta,\Delta}\sqrt{L-1}\max_{\ell}(n_\ell+n_{\ell+1}) \FnormSize{}{\hat\mb-\trueb} -\frac{D_{\alpha,\beta,\Delta}}{2}\min_{\ell}(n_\ell+n_{\ell+1}) \FnormSize{}{\hat\mb-\trueb}^2\geq 0
+\]
+Finally,
+\[
+\text{MSE}(\hat \mb,\trueb)={1\over L-1}\FnormSize{}{\hat\mb-\trueb}\leq {2C_{\alpha,\beta,\Delta} \over D_{\alpha,\beta,\Delta}}{\max_{\ell}(n_\ell+n_{\ell+1})\over \min_{\ell}(n_{\ell}+n_{\ell+1})}  \sqrt{1\over L-1},
+\]
+which completes the proof.
+\end{proof}
+
+\section{Additional results on simulations}
+In the main paper, we compare our method ({\bf Ordinal-T}) with other four methods ({\bf Ordinal-M}, {\bf 1bit-category-T},{\bf 1bit-sign-T}, and {\bf Continuous-T}).
+The methods are evaluated by their capabilities in predicting the most likely label for each entry, i.e.,\ $y_\omega^{\text{mode}}=\arg\max_\ell\mathbb{P}(y_\omega=\ell)$.
+Here we perform additional simulations to compare the methods by their performance in predicting the median label, $y_\omega^{\text{median}}=\min\{\ell\colon\mathbb{P}(y_\omega=\ell)\geq 0.5\}$.
+Under the latent variable model~\eqref{eq:latent} and Assumption~\ref{ass:link}, the median label is the quantized $\Theta$ without noise; i.e.\ $y_\omega^{\text{median}}=\sum_\ell \mathds{1}_{\theta_\omega\in(b_{\ell-1},b_\ell]}$.
+
+\begin{figure}[H]
+  \begin{center}
+    \includegraphics[width=.6\textwidth]{panel3.pdf}
+  \end{center}
+  \caption{Performance comparison in predicting the median label.
+(a,c) Prediction errors versus sample complexity $\rho=|\Omega|/d^k$ when $L=5$.
+(b,d) Prediction errors versus the number of ordinal levels $L$, when $\rho=0.8.$}
+  \label{fig:compare2}
+\end{figure}
+
+Figure~\ref{fig:compare2} compares the prediction accuracy under the setting $\alpha = 10, d = 20$ and $r = 5$.
+We see that our method outperforms the others in both MAD and MCR.
+The methods with probability model ({\bf Ordinal-T, Ordinal-M, 1bit-category-T}) have the tendency to be improved in MAD and be worsened in MCR when  $y_\omega^{\text{median}}$ is used (Figure~\ref{fig:compare2}) compared to the results when $y_\omega^{\text{mode}}$ is used (Figure~\ref{fig:compare}).
+This tendency is from the property that the estimator $\hat y^{(\text{Median})}_\omega$ minimizes $\mathbb{E}_{\hat{\theta}_\omega,\hat{\mb}}|y_\omega-y|$
+and the estimator $\hat y^{(\text{Mode})}_\omega $ minimizes $\mathbb{E}_{\hat{\theta}_\omega,\hat{\mb}}\mathds{1}_{\{y_\omega=y\}}$.
+By the similar reasons in Section~\ref{sec:compare} such as making use of tensor format of the datasets, accepting multi-level observation and preserving the ordinal structure, our model is benefited.
+
+\section{Additional results on HCP analysis}
+\subsection{Clustering based on Tucker representation}\label{sec:clustering}
+We perform clustering analyses based on the Tucker representation of the estimated tensor parameter $\hat\Theta$.
+The procedure is motivated from the higher-order extension of Principal Component Analysis (PCA) or Singular Value Decomposition (SVD).
+Recall that, in the matrix case, we usually perform clustering on an $m\times n$ (normalized) matrix $X$ based on the following procedure.
+First, we factorize $X$ into
+\begin{equation}
+X = U\Sigma V^T,
+\end{equation}
+where $\Sigma$ is a diagonal matrix and $ U,V$ are factor matrices with orthogonal columns.
+Second, we take each column of $V$ as a principal axis and $U\Sigma$ as principal components.
+A subsequent multivariate clustering method (such as $K$-means) is then applied to the $m$ rows of $U\Sigma$.
+
+We apply a similar clustering procedure to the estimated parameter tensor $\hat\Theta$. Based on Tucker decomposition, we have
+\begin{equation}\label{eq:Tuckerest}
+\hat \Theta = \hat \tC\times_1\hat \mM_1\times_2\cdots\times_K\hat \mM_K,
+\end{equation}
+where $\hat \tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is the estimated core tensor, $\hat \mM_k\in\mathbb{R}^{d_k\times r_k}$ are estimated factor matrices with orthogonal columns, and $\times_k$ denotes the tensor-by-matrix multiplication~\citep{kolda2009tensor}.
+The mode-$k$ matricization of~\eqref{eq:Tuckerest} gives
+\begin{equation}
+\hat \Theta_{(k)} = \hat \mM_k\hat \tC_{(k)}\left(\hat \mM_K\otimes\cdots\otimes\hat \mM_1\right),
+\end{equation}
+where $\hat \Theta_{(k)} , \hat \tC_{(k)}$ denote the mode-$k$ unfolding of $\hat \Theta$ and $\hat \tC$, respectively.
+Then, the mode-$k$ clustering can be performed as follows.
+First, we take $\left(\hat \mM_K\otimes\cdots\otimes\hat \mM_1\right)$ as principal axes and $\hat \mM_k\hat \tC_{(k)}$ as principal components.
+Then, we perform $K$-means clustering method to the $d_k$ rows of the principal components $\hat \mM_k\hat \tC_{(k)}$.
+
+\subsection{Clustering results on HCP}
+We perform clustering analysis to the 68 brain nodes using the the produce described in Section~\ref{sec:clustering}.
+Recall that our ordinal tensor method outputs the estimated parameter tensor $\hat\Theta\in\mathbb{R}^{68\times 68\times136}$ with rank $(23,23,8)$.
+The number of clusters in the $K$-means is chosen based on the elbow method (see figure~\ref{figure:elbow}). Table~\ref{table:clustering} shows the eight clusters of brain nodes.
+\begin{figure}[H]
+  \label{figure:elbow}
+  \begin{center}
+    \includegraphics[width=0.49\textwidth]{elbowm.pdf}
+  \end{center}
+  \caption{The elbow plot for the number of clusters}
+\end{figure}
+
+We find that most brain nodes fall into cluster I and the cluster II, which can be represented as the left and right hemispheres, respectively (Figure~\ref{figure:brain image}a).
+From cluster III to the cluster VIII, entries within each cluster share the same name encoded in the data.
+For example, cluster III and cluster IV represent the {\it Supramarginal gyrus} region on the left brain and right brain.
+These regions are known to be in charge of interpreting tactile sensory data and involved in perception of space and limbs location~\cite{carlson2012physiology,reed1994nature}.
+The results demonstrate that our clustering successfully groups similar brain nodes together without knowledge of external annotation.
+
+\begin{table}[H]
+  \label{table:clustering}
+  \center
+  \begin{small}
+    \footnotesize
+  \begin{sc}
+\begin{tabular}{|c|c|c|c|}
+\hline
+Cluster     & \multicolumn{3}{c|}{I}\\ \hline
+Brain nodes & \multicolumn{3}{l|}{\begin{tabular}[c]{@{}l@{}} l.superiorfrontal(2),  l.superiortemporal(2),    l.insula ,             l.frontalpole,         \\
+l.caudalmiddlefrontal, l.parstriangularis,     l.parsopercularis,     l.precentral,          \\
+l.temporalpole,        l.postcentral,
+l.superiorparietal,    l.inferiorparietal,    \\ l.lateraloccipital,    l.medialorbitofrontal,
+l.superiorfrontal,       \\ l.precuneus,           l.cuneus,
+l.parahippocampal,     l.lingual,              l.superiortemporal,\\  l.isthmuscingulate,   l.lateraloccipital
+\end{tabular}} \\
+\hline
+Cluster & \multicolumn{3}{c|}{II}\\ \hline
+Brain nodes & \multicolumn{3}{l|}{\begin{tabular}[c]{@{}l@{}}
+r.superiorfrontal(2),  r.superiortemporal(2),  r.insula,              r.frontalpole,         \\
+r.caudalmiddlefrontal, r.parstriangularis,     r.parsopercularis,     r.precentral,          \\
+r.temporalpole,        r.postcentral,
+r.superiorparietal,    r.inferiorparietal,    \\ r.lateraloccipital,   r.medialorbitofrontal,
+r.superiorfrontal,       \\ r.precuneus,
+r.cuneus,
+r.parahippocampal,     r.lingual,              r.superiortemporal,    \\r.isthmuscingulate,  r.lateraloccipital\end{tabular}} \\ \hline
+Cluster      & III    & IV    & V
+\\ \hline
+Brain nodes & \begin{tabular}[c]{@{}c@{}}l.supramarginal(4)\end{tabular}&
+\begin{tabular}[c]{@{}c@{}}r.supramarginal(4)\end{tabular}&
+\begin{tabular}[c]{@{}c@{}}l.inferiortemporal(3)\end{tabular}\\
+\hline
+Cluster & VI  & VII & VIII
+\\ \hline
+Brain nodes &
+\begin{tabular}[c]{@{}c@{}}r.inferiortemporal(3)\end{tabular}&
+\begin{tabular}[c]{@{}c@{}}l.middletemporal(3)\end{tabular}& \begin{tabular}[c]{@{}c@{}}r.middletemporal(3)\end{tabular}
+\\
+\hline
+\end{tabular}
+\end{sc}
+\end{small}
+\caption{Clustering result of brain nodes.
+The first alphabet in the node name indicates the left brain (L) and the right brain (R) and the values in the parentheses are the number of nodes that have the same name in each cluster.
+}
+\end{table}
+
+\begin{figure}[H]
+  \centering
+ \includegraphics[width = 0.45\textwidth]{braincluster.pdf}
+ \includegraphics[width = 0.45\textwidth]{braincluster2.pdf}
+  \label{figure:brain image}
+  \caption{In the left brain image, yellow nodes are from  cluster I and Green nodes are from cluster II.
+Nodes from each cluster from III to VIII have almost same locations.
+The right brain image shows the extent of connectivity among nodes in each cluster I and cluster II.
+We take average of connections among nodes over 136 individuals.
+Yellow color means strong connectivity and green color weak connectivity.
+}
+\end{figure}
+
+\bibliographystyle{plainnat}
 \bibliography{tensor_wang}
-\bibliographystyle{icml2020}
 
 \end{document}
-
-% This document was modified from the file originally made available by
-% Pat Langley and Andrea Danyluk for ICML-2K. This version was created
-% by Iain Murray in 2018, and modified by Alexandre Bouchard in
-% 2019 and 2020. Previous contributors include Dan Roy, Lise Getoor and Tobias
-% Scheffer, which was slightly modified from the 2010 version by
-% Thorsten Joachims & Johannes Fuernkranz, slightly modified from the
-% 2009 version by Kiri Wagstaff and Sam Roweis's 2008 version, which is
-% slightly modified from Prasad Tadepalli's 2007 version which is a
-% lightly changed version of the previous year's version by Andrew
-% Moore, which was in turn edited from those of Kristian Kersting and
-% Codrina Lauth. Alex Smola contributed to the algorithmic style files.

### 2020-02-13 ICML/arxiv/ordinalT_supp_v4.tex
Add files via upload

--- 2020-02-13 ICML/arxiv/icml_arxiv.tex
@@
 
-\newcommand{\theHalgorithm}{\arabic{algorithm}}
-
 \usepackage{comment}
@@
 \usepackage{amsmath,amssymb,amsthm,bm,mathtools}
+\usepackage{multirow}
 \usepackage{natbib}
@@
 \hypersetup{colorlinks,linkcolor={blue},citecolor={blue},urlcolor={red}}
-\usepackage{algorithm}
-\usepackage{algorithmic}
-\usepackage{makecell}
+\usepackage{algpseudocode,algorithm}
+\algnewcommand\algorithmicinput{\textbf{Input:}}
+\algnewcommand\algorithmicoutput{\textbf{Output:}}
+\algnewcommand\INPUT{\item[\algorithmicinput]}
+\algnewcommand\OUTPUT{\item[\algorithmicoutput]}
 
@@
 
+\renewcommand{\thefigure}{{S\arabic{figure}}}%
+\renewcommand{\thetable}{{S\arabic{table}}}%
+\renewcommand{\figurename}{{Supplementary Figure}}
+\renewcommand{\tablename}{{Supplementary Table}}
 \setcounter{figure}{0}
@@
 \usepackage{dsfont}
+
+\usepackage{multirow}
 
@@
 \input macros.tex
+\setlength\parindent{0pt}
 \allowdisplaybreaks
 
-\title{Tensor denoising and completion based on ordinal observations}
-\date{}
-\author{%
-Chanwoo Lee \\
-University of Wisconsin -- Madison\\
-\texttt{chanwoo.lee@wisc.edu} \\
-\and
-Miaoyan Wang \\
-University of Wisconsin -- Madison\\
-\texttt{miaoyan.wang@wisc.edu} \\
-}
+\title{Supplements for ``Tensor denoising and completion based on ordinal observations''}
+
+%\author{%
+%Chanwoo Lee \\
+%University of Wisconsin -- Madison\\
+ %\texttt{chanwoo.lee@wisc.edu} \\
+%\And
+%Miaoyan Wang \\
+%University of Wisconsin -- Madison\\
+%\texttt{miaoyan.wang@wisc.edu} \\
+%}
 
 \begin{document}
-\maketitle
-\begin{abstract}
-Higher-order tensors arise frequently in applications such as neuroimaging, recommendation system, social network analysis, and psychological studies.
-We consider the problem of low-rank tensor estimation from possibly incomplete, ordinal-valued observations.
-Two related problems are studied, one on tensor denoising and another on tensor completion.
-We propose a multi-linear cumulative link model, develop a rank-constrained M-estimator, and obtain theoretical accuracy guarantees.
-Our mean squared error bound enjoys a faster convergence rate than previous results, and we show that the proposed estimator is minimax optimal under the class of low-rank models.
-Furthermore, the procedure developed serves as an efficient completion method which guarantees consistent recovery of an order-$K$ $(d,\ldots,d)$-dimensional low-rank tensor using only $\tilde \tO(Kd)$ noisy, quantized observations.
-We demonstrate the outperformance of our approach over previous methods on the tasks of clustering and collaborative filtering.
-\\[0.5cm]
-\textbf{Keywords:} Higher-order tensors, ordinal observation, tensor decomposition, tensor completion
-\end{abstract}
-
-\section{Introduction}
-Multidimensional arrays, a.k.a.\ tensors, arise in a variety of applications including recommendation systems~\citep{baltrunas2011incarmusic}, social networks~\citep{nickel2011three}, genomics~\citep{hore2016tensor}, and neuroimaging~\citep{zhou2013tensor}.
-There is a growing need to develop general methods for analyzing these noisy, high-dimensional datasets that can handle two main problems.
-The problem of tensor denoising -- which aims to recover a signal tensor from its noisy entries -- has gained increased attention in theory and applications~\citep{xia2019sup,zeng2019multiway}.
-A related problem, tensor completion, examines the minimum number of entries needed for a consistent recovery~\citep{ghadermarzy2018learning,ghadermarzy2019near}.
-Low-rankness is often imposed to the signal tensor, which efficiently reduces the intrinsic dimension in both problems.
-
-A number of low-rank tensor estimation methods have been proposed~\citep{kolda2009tensor,acar2010scalable}, revitalizing classical methods such as CANDECOMP/PARAFAC (CP) decomposition~\citep{hitchcock1927expression} and Tucker decomposition~\citep{tucker1966some}.
-These tensor methods treat the entries as continuous-valued.
-In many cases, however, we encounter datasets of which the entries are qualitative.
-For example, the Netflix problem records the ratings of users on movies over time.
-Each data is a rating on a nominal scale \{{\it very like, like, neutral, dislike, very dislike}\}.
-Another example is in the signal processing, where the digits are frequently rounded or truncated so that only integer values are available.
-The qualitative observations take values in a limited set of categories, making the learning problem harder compared to  continuous observations.
-
- Ordinal entries are categorical variables with an ordering among the categories; for example, {\it very like} $\prec$ {\it like} $\prec$ {\it neutral} $\prec \cdots$.
-The analyses of tensors with the ordinal entries are mainly complicated by two key properties needed for a reasonable model.
-First, the model should be invariant under a reversal of categories, say, from the Netflix example, {\it very like} $\succ$ {\it like} $\succ$ {\it neutral} $\succ \cdots$, but not under arbitrary label permutations.
-Second, the parameter interpretations should be consistent under merging or splitting of contiguous categories.
-The classical continuous tensor model~\cite{kolda2009tensor, ghadermarzy2019near} fails in the first aspect, whereas the binary tensor model~\cite{ghadermarzy2018learning} lacks the second property.
-An appropriate model for ordinal tensors has yet to be studied.
-
-\begin{table*}[ht]
-\centerline{\includegraphics[width=16cm]{compare.pdf}}
-\caption{Comparison with previous work.
-For ease of presentation, we summarize the error rate and sample complexity assuming equal tensor dimension in all modes.
-$K$: tensor order; $L$: number of ordinal levels; $d$: dimension at each mode.
-}~\label{tab:compare}
-\end{table*}
-
-{\bf Our contribution.} This paper presents an efficient low-rank estimation method and theory for tensors with ordinal-valued entries.
-Our main contributions are summarized in Table~\ref{tab:compare}.
-We propose a cumulative link model for higher-order tensors, develop a rank-constrained M-estimator, and obtain theoretical accuracy guarantees.
-The mean squared error bound is established, and we show that the obtained bound has minimax optimal rate in high dimensions under the low-rank model.
-Our estimator enjoys a faster convergence rate $\tO(d^{-(K-1)/2})$ than $\tO(d^{-K})$ in~\citet{ghadermarzy2018learning}, which is a substantial improvement as the order $K$ increases.
-Furthermore, our proposal serves as an efficient completion algorithm that guarantees consistent recovery of an order-$K$ $(d,\ldots,d)$-dimensional low-rank tensor using only $\tilde \tO(Kd)$ noisy, quantized observations.
-
-{\bf Related work.} Our work is related to, but clearly distinctive from, several lines of existing literature.
-Matrix completion from quantized samples was firstly introduced for binary observations~\cite{cai2013max,davenport2014,bhaskar20151} and then extended to ordinal observations~\cite{bhaskar2016probabilistic}.
-As we show in Section~\ref{sec:theory}, applying existing matrix methods to an ordinal tensor results in a suboptimal estimator with a slower convergence rate.
-Therefore, a full exploitation of the tensor structure is necessary; this is the focus of the current paper.
-
-Our work is also connected to non-Gaussian tensor decomposition.
-Existing work focuses exclusively on univariate observations such as binary- or continuous-valued entries~\citep{wang2018learning,hong2018generalized,ghadermarzy2018learning}.
-As we mentioned earlier, the ordinal observations add considerable challenges to the model formulation.
-We address the problems from two perspectives.
-From statistical perspective, our proposed model generalizes the usual binary tensor model while preserving palindromic invariance~\cite{mccullagh1980regression} for ordinal observations.
-From algorithm perspective, our alternating optimization compares favorably to the approximate (non-convex) algorithm developed in the context of binary tensors~\cite{ghadermarzy2018learning}.
-We numerically compare the two approaches in Section~\ref{sec:experiment}.
-
-\section{Preliminaries}
-Let $\tY\in\mathbb{R}^{d_1\times \cdots \times d_K}$ denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional tensor.
-We use $y_\omega$ to denote the tensor entry indexed by $\omega$, where $\omega\in[d_1]\times\cdots\times[d_K]$.
-The Frobenius norm of $\tY$ is defined as $\FnormSize{}{\tY}=\sum_\omega y^2_\omega$ and the infinity norm of $\tY$ is defined as $\mnormSize{}{\tY}=\max_{\omega}|y_\omega|$.
-We use $\tY_{(k)}$ to denote the unfolded matrix of size $d_k$-by-$\prod_{i\neq k}d_k$, obtained by reshaping the tensor along the mode-$k$.
-The Tucker rank of $\tY$ is defined as a length-$K$ vector $\mr = (r_1,\ldots,r_K)$, where $r_k$ is the rank of matrix $\tY_{(k)}$ for all $k \in[K]$.
-We say that an event $A$ occurs ``with very high probability'' if $\mathbb{P}(A)$ tends to 1 faster than any polynomial of tensor dimension $d_{\min}=\min\{d_1,\ldots,d_K\} \to\infty$.
-
-We use lower-case letters ($a, b, c, \ldots$) for scalars/vectors, upper-case boldface letters ($\mA, \mB, \mC, \ldots$) for matrices, and calligraphy letters ($\tA,\tB,\tC,\ldots$) for tensors of order three or greater.
-For ease of notation, we allow the basic arithmetic operators (e.g., $\leq, +, -$) to be applied to pairs of tensors in an element-wise manner.
-We use the shorthand $[n]$ to denote the $n$-set $\{1,\ldots,n\}$ for $n \in N_{+}$.
-
-\section{Model formulation and motivation}
-\subsection{Observation model}
-Let $\tY$ denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional data tensor.
-Suppose the entries of $\tY$ are ordinal-valued, and the observation space consists of $L$ ordered levels, denoted by $[L]:=\{1,\ldots,L\}$.
-We propose a cumulative link model for the ordinal tensor $\tY=\entry{y_\omega}\in[L]^{d_1\times \cdots\times d_K}$.
-Specifically, assume the entries $y_\omega$ are (conditionally) independently distributed with cumulative probabilities:
-\begin{equation}\label{eq:model}
-\mathbb{P}(y_\omega\leq \ell)=f(b_\ell-\theta_\omega),\ \text{for all}\ \ell\in[L-1],
-\end{equation}
-where $\mb=(b_1,\ldots,b_{L-1})$ is a set of unknown scalars satisfying $b_1<\cdots <b_{L-1}$, $\Theta=\entry{\theta_\omega}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is a continuous-valued parameter tensor satisfying certain low-dimensional structure (to be specified later), and $f(\cdot):\mathbb{R}\mapsto[0,1]$ is a known, strictly increasing function.
-We refer to $\mb$ as the cut-off points and $f$ the link function.
-
-The formulation~\eqref{eq:model} imposes an additive model to the transformed probability of cumulative categories.
-This modeling choice is to respect the ordering structure among the categories.
-For example, if we choose the inverse link $f^{-1}(x)=\log {x\over 1-x}$ to be the log odds, then the model~\eqref{eq:model} implies linear spacing between the proportional odds:
-\begin{equation}\label{eq:logodd}
-\log {\mathbb{P}(y_{\omega}\leq \ell) \over \mathbb{P}(y_\omega >  \ell) } - \log {\mathbb{P}(y_{\omega}\leq {\ell-1}) \over \mathbb{P}(y_\omega >  {\ell-1}) } = b_\ell-b_{\ell-1},
-\end{equation}
-for all tensor entries $y_\omega$.
-When there are only two categories in the observation space (e.g.
-binary tensors), the cumulative model~\eqref{eq:model} is equivalent to the usual multinomial link model.
-In general, however, when the number of categories $L\geq 3$, the proportional odds assumption~\eqref{eq:logodd} is more parsimonious, in that, the ordered categories can be envisaged as contiguous intervals on the continuous scale, where the points of division are exactly $b_1<\cdots <b_{L-1}$.
-This interpretation will be made more explicit in the next section.
-
-\subsection{Latent-variable interpretation}\label{sec:latent}
-The ordinal tensor model~\eqref{eq:model} with certain types of link $f$ has the equivalent representation as an $L$-level quantization model on $\tY=\entry{y_\omega}$:
-\begin{align}\label{eq:quantization}
-y_\omega&=
-\begin{cases}
-1,& \text{if $y^*_\omega\in(-\infty, b_1]$},\\
-2,& \text{if $y^*_\omega\in(b_1, b_2]$},\\
-\vdots  &\vdots\\
-L,& \text{if $y^*_\omega\in(b_{L-1}, \infty)$},\\
-\end{cases}
-\end{align}
-for all $\omega\in[d_1]\times \cdots \times [d_k]$. Here, $\tY^*=\entry{y^*_\omega}$ is a latent continuous-valued tensor following an additive noise model:
-\begin{align}\label{eq:latent}
-\KeepStyleUnderBrace{\tY^*}_{\text{latent continuous-valued tensor}}=\KeepStyleUnderBrace{\Theta}_{\text{signal tensor}}+\KeepStyleUnderBrace{\tE}_{\text{i.i.d.\ noise}},
-\end{align}
-where $\tE=\entry{\varepsilon_\omega}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is a noise tensor with i.i.d.\ entries according to distribution $\mathbb{P}(\varepsilon)$.
-From the viewpoint of~\eqref{eq:latent}, the parameter tensor $\Theta$ can be interpreted as the latent signal tensor prior to contamination and quantization.
-
-The equivalence between the latent-variable model~\eqref{eq:quantization} and the cumulative link model~\eqref{eq:model} is established if the link $f$ is chosen to be the cumulative distribution function of noise $\varepsilon$, i.e., $f(\theta)=\mathbb{P}(\varepsilon\leq \theta)$.
-We describe two common choices of link $f$, or equivalently, the distribution of $\varepsilon$.
-
-\begin{example}[Logistic model] The logistic model is characterized by~\eqref{eq:model} with $f(\theta)=(1+e^{-\theta/\sigma})^{-1}$, where $\sigma>0$ is the scale parameter.
-Equivalently, the noise $\varepsilon_\omega$ in~\eqref{eq:quantization} follows i.i.d.\ logistic distribution with scale parameter $\sigma$.
-\end{example}
-\begin{example}[Probit model] The probit model is characterized by~\eqref{eq:model} with
-$f(\theta)=\mathbb{P}(z\leq \theta/\sigma)$, where $z\sim N(0,1)$. Equivalently, the noise $\varepsilon_\omega$ in~\eqref{eq:quantization} follows i.i.d.\ $N(0,\sigma^2)$.
-\end{example}
-Other link functions are also possible, such as Laplace, Cauchy, etc~\cite{mccullagh1980regression}.
-All the models share the property that the ordered categories can be thought of as contiguous interval on some continuous scale.
-We should point out that, although the latent-variable interpretation is incisive, our estimation procedure does not refer to the existence of $\tY^*$.
-Therefore, our model~\eqref{eq:model} is general and still valid in the absence of quantization process.
-More generally, we make the following assumptions about the link $f$.
-
-\begin{assumption}\label{ass:link}
-The link function is assumed to satisfy:
-\begin{enumerate}
-\item $f(\theta)$ is strictly increasing and twice-differentiable in $\theta\in\mathbb{R}/\{0\}$.
-\item $f'(\theta)$ is strictly log-concave and symmetric with respect to $\theta=0$.
-\end{enumerate}
-\end{assumption}
-
-\subsection{Problem 1: Tensor denoising}~\label{sec:denoising}
-The first question we aim to address is tensor denoising:
-
-(P1) Given the quantization process induced by $f$ and the cut-off points $\mb$, how accurately can we estimate the latent signal tensor $\Theta$ from the ordinal observation $\tY$?
-
-Clearly, the problem (P1) cannot be solved uniformly for all possible $\Theta$.
-We focus on a class of ``low-rank'' and ``flat'' signal tensors, which is a plausible assumption in practical applications~\cite{zhou2013tensor,bhaskar20151}.
-Specifically, we consider the parameter space:
-\begin{equation}\label{eq:space}
-\small \tP=\left\{\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K} \colon \text{rank}(\Theta)\leq \mr,\ \mnormSize{}{\Theta}\leq \alpha\right\}.
-\end{equation}
-where $\mr=(r_1,\ldots,r_K)$ denotes the Tucker rank of $\Theta$.
-
-The parameter tensor of our interest satisfies two constraints.
-The first is that $\Theta$ is a low-rank tensor, with $r_k=\tO(1)$ for all $k\in[K]$.
-Equivalently, $\Theta$ admits the Tucker decomposition:
-\begin{equation}\label{eq:Tucker}
-\Theta=\tC\times_1\mM_1\times_1 \cdots \times_K \mM_K,
-\end{equation}
-where $\tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is a core tensor, $\mM_k\in\mathbb{R}^{d_k\times r_k}$ are factor matrices with orthogonal columns, and $\times_k$ denotes the tensor-by-matrix multiplication~\cite{kolda2009tensor}.
-The Tucker low-rankness is popularly imposed in tensor data analysis, and is shown to provide a reasonable tradeoff between model complexity and model flexibility.
-Note that, unlike matrices, there are various notations of tensor low-rankness, such as CP rank~\cite{hitchcock1927expression} and train rank~\cite{oseledets2011tensor}.
-Some notation of low-rankness may lead to mathematically ill-posed optimization; for example, the best low CP-rank tensor approximation may not exist~\cite{de2008tensor}.
-We choose Tucker representation for well-posedness of optimization and easy interpretation.
-
-The second constraint is that the entries of $\Theta$ are uniformly bounded in magnitude by a constant $\alpha \in \mathbb{R}_{+}$.
-In view of~\eqref{eq:latent}, we refer to $\alpha$ as the signal level.
-The entry-wise bound assumption is a technical condition that avoids the degeneracy in probability estimation with ordinal observations.
-
-\subsection{Problem 2: Tensor completion}
-Motivated by applications in collaborative filtering, we also consider a more general setup when only a subset of tensor entries $y_\omega$ are observed.
-Let $\Omega\subset[d_1]\times \cdots\times[d_K]$ denote the set of observed indices.
-The second question we aim to address is stated as follows:
-
-(P2) Given an incomplete set of ordinal observations $\{y_{\omega}\}_{\omega\in\Omega}$, how many sampled entries do we need to consistently recover $\Theta$ based on the model~\eqref{eq:model}?
-
-The answer to (P2) depends on the choice of $\Omega$.
-We consider a general model on $\Omega$ that allows both uniform and non-uniform sampling.
-Specifically, let $\Pi=\{\pi_{i_1,\ldots,i_K}\}$ denote a predefine probability distribution over the index set such that $\sum_{\omega\in[d_1]\times \cdots \times [d_K]} \pi_\omega =1$.
-We assume that each index in $\Omega$ is drawn with replacement using distribution $\Pi$.
-This sampling model relaxes the uniform sampling in literature and is arguably a better fit in applications.
-
-We consider the same parameter space~\eqref{eq:space} for the completion problem.
-In addition to the reasons mentioned in Section~\ref{sec:denoising}, the entrywise bound assumption also serves as the incoherence requirement for completion.
-In classical matrix completion, the incoherence is often imposed on the singular vectors.
-This assumption is recently relaxed for ``flat'' matrices with bounded magnitude~\cite{negahban2011estimation,cai2013max,bhaskar20151}.
-We adopt the same assumption for higher-order tensors.
-
-\section{Rank-constrained M-estimator}\label{sec:theory}
-We present a general treatment to both problems mentioned above.
-With a little abuse of notation, we use $\Omega$ to denote either the full index set $\Omega=[d_1]\times \cdots \times [d_K]$ (for the tensor denoising) or a random subset induced from the sampling distribution $\Pi$ (for the tensor completion).
-Define $b_0=-\infty$, $b_L=\infty$, $f(-\infty)=0$ and $f(\infty)=1$.
-The log-likelihood associated with the observed entries is
-\begin{align}\label{eq:objective}
-\logl(\Theta, \mb)&=\sum_{\omega\in\Omega}\sum_{\ell\in[L]} \Big\{\mathds{1}_{\{y_\omega=\ell\}} \log \big[f(b_\ell-\theta_\omega)- \notag \\
-&\quad \quad \quad \quad \quad \quad \quad  f(b_{\ell-1}-\theta_\omega)\big]\Big\}.
-\end{align}
-We propose a rank-constrained maximum likelihood estimator (a.k.a.\ M-estimator) for $\Theta$:
-\begin{align}\label{eq:estimator}
-\hat \Theta&=\argmax_{\Theta\in \tP}\logl(\Theta, \mb),\ \text{where}\notag \\
-\tP&=\left\{\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K} \colon \text{rank}(\Theta)\leq \mr,\ \mnormSize{}{\Theta}\leq \alpha\right\}.
- \end{align}
-In practice, the cut-off points $\mb$ are unknown and should be jointly estimated with $\Theta$.
-For technical convenience, we assume in this section that the cut-off points $\mb$ are known.
-The adaptation of unknown $\mb$ is addressed in Section~\ref{sec:algorithm} and the Supplement.
-
-We define a few key quantities that will be used in our theory. Let $g_\ell=f(b_\ell-\theta)-f(b_{\ell-1}-\theta)$ for all $\ell\in[L]$, and
-\begin{align}\label{eq:regular}
-A_\alpha &= \min_{\ell\in[L], |\theta|\leq \alpha} g_\ell(\theta),\quad  U_\alpha=\max_{\ell\in[L], |\theta|\leq \alpha } {|\dot{g}_\ell(\theta)|\over g_\ell(\theta)},\\
- L_\alpha &= \min_{\ell\in[L], |\theta|\leq \alpha }\left[{\dot{g}_\ell^2 (\theta)\over g_\ell^2(\theta)} -{\ddot{g}_\ell(\theta)\over g_\ell(\theta)}\right],
-\end{align}
-where $\dot{g}(\theta)=dg(\theta)/d\theta$, and $\alpha$ is the entrywise bound of $\Theta$.
-In view of equation~\eqref{eq:latent}, these quantities characterize the geometry including flatness and convexity of the latent noise distribution.
-Under the Assumption~\ref{ass:link}, all these quantities are strictly positive and independent of tensor dimension.
-
-\subsection{Estimation error for tensor denoising}\label{sec:denosing}
-For the tensor denoising problem, we assume that the full set of tensor entries are observed. We assess the estimation accuracy using the mean squared error (MSE):
-\[
-\text{MSE}(\hat \Theta, \trueT)={1\over \prod_k d_k}\FnormSize{}{\Theta-\trueT}^2.
-\]
-The next theorem establishes the upper bound for the MSE of the proposed $\hat \Theta$ in~\eqref{eq:estimator}.
-
-\begin{thm}[Statistical convergence] \label{thm:rate}
-Consider an ordinal tensor $\tY\in[L]^{d_1\times\dots\times d_K}$ generated from model~\eqref{eq:model}, with the link function $f$ and the true coefficient tensor $\trueT\in\tP$.
-Define $r_{\max}=\max_k r_k$.
-Then, with very high probability, the estimator in~\eqref{eq:estimator} satisfies
-\begin{equation}\label{eq:rate}
-\text{MSE}(\hat \Theta, \trueT) \leq \min\left( 4\alpha^2,\ {c_2  U^2_\alpha r_{\max}^{K-1}  \over  L^2_\alpha } {\sum_kd_k\over  \prod_k d_k} \right),
-\end{equation}
-where $c_1, c_2>0$ are two constants that depend only on $K$.
-\end{thm}
-Theorem~\ref{thm:rate} establishes the statistical convergence for the estimator~\eqref{eq:estimator}.
-In fact, the proof of this theorem (see Section~\ref{sec:proof}) shows that the same statistical rate holds, not only for the global optimizer~\eqref{eq:estimator}, but also for any local optimizer $\check \Theta$ in the level set $\{\check\Theta\in\tP: \logl(\check\Theta)\geq \logl(\trueT)\}$.
-This suggests that the local optimality itself is not necessarily a severe concern in our context, as long as the convergent objective is large enough.
-In Section ~\ref{sec:algorithm}, we perform empirical studies to assess the algorithmic stability.
-
-A similar conclusion is obtained for the prediction error, measured in Kullback-Leibler (KL) divergence, between the categorical distributions in the observation space.
-\begin{cor}[Prediction error]~\label{cor:prediction}
-Assume the same set-up as in Theorem~\ref{thm:rate}.
-Let $\mathbb{P}_{\tY}$ and $\hat{\mathbb{P}}_{\tY}$ denote the distributions generating the $L$-level ordinal tensor $\tY$, given the true parameter $\Theta$ and its estimator $\hat \Theta$, respectively.
-Assume $L\geq 2$.
-Then, with very high probability,
-\begin{equation}\label{eq:KLrate}
-\text{KL}(\mathbb{P}_{\tY} || \hat{\mathbb{P}}_{\tY}) \leq  {c_2 U^2_\alpha r_{\max}^{K-1}  \over L^2_\alpha } {(4L-6)\dot{f}^2(0)  \over A_\alpha} {\sum_kd_k\over  \prod_k d_k},
-\end{equation}
-where $c_1, c_2>0$ are the same constants as in Theorem~\ref{thm:rate}.
-\end{cor}
-To gain insight into these bounds, we consider a special setting with equal dimension in all modes, i.e., $d_1=\cdots=d_K=d$. In such a case, our bound \eqref{eq:rate} reduces to
-\begin{equation*}\label{eq:ours}
-\text{MSE}(\hat \Theta, \trueT) \asymp d^{-(K-1)}, \quad \text{as}\ d\to \infty.
-\end{equation*}
-Hence, our estimator achieves consistency with polynomial convergence rate.
-We compare the bound with existing literature.
-In the special case $L=2$, \citet{ghadermarzy2018learning} proposed a max-norm constrained estimator $\tilde{\Theta}$ with $\text{MSE}(\tilde{\Theta}, \trueT) \asymp  d^{-(K-1)/2}$.
-In contrast, our estimator converges at a rate of $d^{-(K-1)}$, which is substantially faster than theirs.
-This provides a positive answer to the open question posed in~\citet{ghadermarzy2018learning} whether the square root in the bound is removable.
-The improvement stems from utilizing the exact low-rankness of $\Theta$, whereas the surrogate rank measure employed in~\citet{ghadermarzy2018learning} is scale-sensitive.
-
-Our bound also generalizes the previous results on ordinal matrices.
-The convergence rate for rank-constrained matrix estimation was $\tO(1/\sqrt{d})$~\citep{bhaskar2016probabilistic}, which fits into our special case when $K=2$.
-Furthermore, our results~\eqref{eq:rate} and~\eqref{eq:KLrate} reveal that the convergence becomes favorable as the order of data tensor increases.
-Intuitively, the sample size for tensor data analysis is the number of entries, $\prod_k d_k$, and the number of free parameters is roughly on the order of $\sum_{k}d_k$, assuming $r_{\max}=\tO(1)$.
-A higher tensor order implies higher effective sample size per parameter, and thus exhibits a faster convergence rate in high dimensions.
-
-We next show the statistical optimality of our estimator $\hat \Theta$.
-The result is based on the information theory, and applies to all estimators in $\tP$, including but not limited to $\hat \Theta$ in~\eqref{eq:estimator}.
-
-\begin{thm}[Minimax lower bound]\label{thm:minimax}
-Assume the same set-up as in Theorem~\ref{thm:rate}, and $d_{\max}=\max_k d_k \geq 8$.
-Let $\inf_{\hat \Theta}$ denote the infimum over all estimators $\hat \Theta\in\tP$ based on the ordinal tensor observation $\tY\in[L]^{d_1\times \cdots \times d_K}$.
-Then, under the model~\eqref{eq:model},
-\begin{align}\label{eq:lower}
-\inf_{\hat \Theta }\sup_{\trueT\in\tP} \mathbb{P}\Big\{& \text{MSE}(\hat \Theta, \trueT)  \geq c \min\left( \alpha^2, \ { Cr_{\max}d_{\max} \over \prod_k d_k} \right) \Big\} \geq {1\over 8},
-\end{align}
-where $C=C(\alpha, L, f,\mb)>0$ and $c>0$ are constants independent of tensor dimension and the rank.
-\end{thm}
-We see that the lower bound matches the upper bound in~\eqref{eq:rate} on the polynomial order of tensor dimension. Therefore, our estimator~\eqref{eq:estimator} is order-optimal.
-
-\subsection{Sample complexity for tensor completion}
-We now consider the tensor completion problem, when only a subset of entries $\Omega$ are observed.
-We consider a general sampling procedure induced by $\Pi$.
-The recovery accuracy is assessed by the weighted squared error:
-\begin{align}\label{eq:weighted}
-\PiFnormSize{}{\Theta-\hat \Theta}^2&\stackrel{\text{def}}{=}
-{1\over |\Omega|}\mathbb{E}_{\Omega\sim \Pi}\FnormSize{}{\Theta-\hat \Theta}^2\notag \\
-&=\sum_{\omega\in[d_1]\times \cdots \times [d_K]} \pi_{\omega}(\Theta_{\omega}-\hat \Theta_{\omega})^2.
-\end{align}
-
-Note that the recovery error depends on the distribution $\Pi$.
-In particular, tensor entries with higher sampling probabilities have more influence on the recovery accuracy, compared to the ones with lower sampling probabilities.
-
-\begin{rmk} If we assume each entry is sampled with strictly positive probability; i.e.\ there exits a constant $\mu> 0$ s.t.
-\[
-\ \pi_\omega\geq {1\over \mu \prod_k d_k},\quad \text{for all}\ \omega\in[d_1]\times\cdots \times [d_K],
-\]
-then the error in~\eqref{eq:weighted} provides an upper bound for MSE:
-\[
-\PiFnormSize{}{\Theta-\hat \Theta}^2 \geq {\FnormSize{}{\Theta-\hat \Theta}^2\over \mu \prod_kd_k}={1\over \mu}\text{MSE}(\hat \Theta, \trueT).
-\]
-The equality is attained under uniform sampling with $\mu=1$.
-\end{rmk}
-
-\begin{thm}~\label{thm:completion}
-Assume the same set-up as in Theorem~\ref{thm:rate}.
-Suppose that we observe a subset of tensor entries $\{y_\omega\}_{\omega\in\Omega}$, where $\Omega$ is chosen at random with replacement according to a probability distribution $\Pi$.
-Let $\hat \Theta$ be the solution to~\eqref{eq:estimator}, and assume $r_{\max}=\tO(1)$.
-Then, with very high probability,
-\[
-\PiFnormSize{}{\Theta-\hat \Theta}^2 \to 0,\quad \text{ as }\quad {|\Omega|\over \sum_k d_k} \ \to \infty.
-\]
-\end{thm}
-
-Theorem~\ref{thm:completion} shows that our estimator achieves consistent recovery using as few as $\tilde{\tO}(Kd)$ noisy, quantized observations from an order-$K$ $(d,\ldots,d)$-dimensional tensor.
-Note that $\tilde \tO(Kd)$ roughly matches the degree of freedom for an order-$K$ tensor of fixed rank $\mr$, suggesting the optimality of our sample requirement.
-This sample complexity substantially improves over earlier result $\tO(d^{\lceil K/2\rceil})$ based on square matricization~\cite{mu2014square}, or $\tO(d^{N/2})$ based on tensor nuclear-norm regularization~\cite{yuan2016tensor}.
-Existing methods that achieve $\tilde \tO(Kd)$ sample complexity require either a deterministic cross sampling design~\cite{zhang2019cross} or univariate measurements~\cite{ghadermarzy2018learning}.
-Our method extends the conclusions to multi-level measurements under a broader class of sampling schemes.
-
-\section{Numerical Implementation}\label{sec:algorithm}
-
-We describe the algorithm to seek the optimizer of~\eqref{eq:objective}.
-In practice, the cut-off points $\mb$ are often unknown, so we choose to maximize $\logl$ jointly over $(\Theta,\mb)\in\tP\times \tB$.
-The objective $\logl$ is concave in $(\Theta,\mb)$ whenever $f'$ is log-concave (see Section~\ref{sec:proof}).
-However, the feasible set $\tP$ is non-convex, which makes the optimization~\eqref{eq:objective} a non-convex problem.
-We employ the alternating optimization approach by utilizing the Tucker representation of $\Theta$.
-Specifically, based on~\eqref{eq:Tucker} and~\eqref{eq:objective}, the objective function consists of $K+2$ blocks of variables, one for the cut-off points $\mb$, one for the core tensor $\tC$, and $K$ for the factor matrices $\mM_k$'s.
-The optimization is a simple convex problem if any $K+1$ out of the $K+2$ blocks are fixed.
-We update one block at a time while holding others fixed, and alternate the optimization throughout the iteration.
-The convergence is guaranteed whenever $\logl$ is bounded from above, since the alternating procedure monotonically increases the objective.
-The Algorithm~\ref{alg} gives the full description.
-
-\begin{algorithm}[t]
-        \caption{Ordinal tensor decomposition}\label{alg}
-        \begin{algorithmic}[]
-            \STATE{\bfseries Input:}  \text{ Ordinal data tensor }
-            $\mathcal{Y}\in [L]^{d_1\times\cdots\times d_K}$, rank $\mr\in \mathbb{N}_{+}^{K}$, entry-wise bound $\alpha\in \mathbb{R_+}$.
-            \STATE{\bfseries Output:} $(\hat\Theta,\hat{\mb}) =  \argmax_{(\Theta,\mb)\in \mathcal{P}\times\mathcal{B}}  \mathcal{L}_{\mathcal{Y},\Omega}(\Theta,\mb).$
-            \STATE Random initialization of core tensor $\mathcal{C}^{(0)}$, factor matrices $ \{\mM_k^{(0)}\}$, and cut-off points $\mb^{(0)}$.
-            \FOR{$t = 1,2,\cdots,$}
-            \FOR{$k = 1,2,\cdots,K$}
-            \STATE{Update $\mM^{(t+1)}_k$ while fixing other blocks:}
-            \STATE $\mM_k^{(t+1)}\gets\argmax_{\mM_k\in\mathbb{R}^{d_k\times r_K}}\mathcal{L}_{\mathcal{Y},\Omega}(\mM_k)$,
-            \newline
-             s.t. $\mnormSize{}{\Theta^{(t+1)}}\leq \alpha$, where $\Theta^{(t+1)}$ is the parameter tensor based on the current block estimates.
-            \ENDFOR
-            \STATE {Update $\mathcal{C}^{(t+1)}$ while fixing other blocks:}
-              \newline$\mathcal{C}^{(t+1)} \gets \argmax_{\mathcal{C}\in \mathbb{R}^{r_1\times\cdots\times r_K}}\mathcal{L}_{\mathcal{Y},\Omega}(\mathcal{C})$, s.t.
-$\mnormSize{}{\Theta^{(t+1)}}\leq \alpha.$
-               \STATE {Update $\Theta^{(t+1)}$ based on the current block estimates:}
-                   \newline
-                   $\Theta^{(t+1)} \gets \mathcal{C}^{(t+1)}\times_1\mM_1^{(t+1)}\cdots\times_K\mM_K^{(t+1)}.$
-            \STATE {Update $\mb^{(t+1)}$ while fixing $\Theta^{(t+1)}$:}
-              \newline
-              $\mb^{(t+1)} \gets \argmax_{\mb\in \tB}\mathcal{L}_{\mathcal{Y},\Omega}\big(\Theta^{(t+1)},\mb\big).$
-            \ENDFOR
-            \STATE \textbf{return}
-            $(\hat \Theta,\hat \mb)$
-        \end{algorithmic}
- \end{algorithm}
-
-We comment on two implementation details before concluding this section.
-First, the problem~\eqref{eq:estimator} is non-convex, so Algorithm 1 usually has no theoretical guarantee on global optimality.
-Nevertheless, as shown in Section~\ref{sec:denosing}, the desired rate holds not only for the global optimizer, but also for the local optimizer with $\logl(\hat \Theta)\geq \logl(\trueT)$.
-In practice, we find the convergence point $\hat \Theta$ upon random initialization is often satisfactory, in that the corresponding objective $\logl(\hat \Theta)$ is close to and actually slightly larger than the objective evaluated at the true parameter $\logl(\trueT)$.
-Figure~\ref{fig:stability} shows the trajectory of the objective function that is output in the default setting of Algorithm 1, with the input tensor generated from probit model~\eqref{eq:model} with $d_1=d_2=d_3=d$ and $r_1=r_2=r_3=r$.
-The dashed line is the objective value at the true parameter $\logl(\trueT)$.
-We find that the algorithm generally converges quickly to a desirable value in reasonable number of steps.
-The actual running time per iteration is shown in the plot legend.
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=0.7\textwidth]{algorithm.pdf}\label{fig:stability}
-\caption{Trajectory of objective function with various $d$ and $r$.}
-\end{figure}
-
-Second, the algorithm takes the rank $\mr$ as an input.
-In practice, the rank $\mr$ is hardly known and needs to be estimated from the data.
-We suggest to use Bayesian information criterion (BIC) and choose the rank that minimizes BIC; i.e.
-\begin{align}\label{eq:BIC}
-\hat \mr&=\argmin_{\mr\in\mathbb{N}^K_{+}} \text{BIC}(\mr)\\
-&=\argmin_{\mr\in\mathbb{N}^K_{+}}\{-2\tL_{\tY}(\hat \Theta(\mr),\hat \mb(\mr))+p_e(\mr)\log (\prod_k d_k)\},
-\end{align}
-where $\hat \Theta(\mr), \hat\mb(\mr)$ are the estimates given the rank $\mr$, and $p_e(\mr)\stackrel{\text{def}}{=}\sum_k (d_k-r_k)r_k+\prod_k r_k$ is the effective number of parameters in the model.
-We select $\hat \mr$ that minimizes BIC through a grid search.
-The choice of BIC is intended to balance between the goodness-of-fit for the data and the degrees of freedom in the population model.
-
-\section{Experiments}\label{sec:experiment}
-In this section, we evaluate the empirical performance of our method.
-We investigate both the complete and the incomplete settings, and compare the recovery accuracy with other tensor-based methods.
-Unless otherwise stated, the ordinal data tensors are generated from model~\eqref{eq:model} using standard probit link $f$.
-We consider the setting with $K=3$, $d_1=d_2=d_3=d$, and $r_1=r_2=r_3=r$.
-The parameter tensors are simulated based on~\eqref{eq:Tucker}, where the core tensor entries are i.i.d.\ drawn from $N(0,1)$, and the factors $\mM_k$ are uniformly sampled (with respect to Haar measure) from matrices with orthonormal columns.
-We set the cut-off points $b_\ell=f^{-1}(\ell/L)$ for $\ell\in[L]$, such that $f(b_\ell)$ are evenly spaced from 0 to 1.
-In each simulation study, we report the summary statistics across $n_{\text{sim}}$ = 30 replications.
-
-\subsection{Finite-sample performance}\label{sec:simulation}
-The first experiment examines the performance under complete observations.
-We assess the empirical relationship between the MSE and various aspects of model complexity, such as dimension $d$, rank $r$, and signal level $\alpha=\mnormSize{}{\Theta}$.
-Figure~\ref{fig:finite}a plots the estimation error versus the tensor dimension $d$ for three different ranks $r\in\{3,5,8\}$.
-The decay in the error appears to behave on the order of $d^{-2}$, which is consistent with our theoretical results~\eqref{eq:rate}.
-We find that a higher rank leads to a larger error, as reflected by the upward shift of the curve as $r$ increases.
-Indeed, a higher rank implies the higher number of parameters to estimate, thus increasing the difficulty of the estimation.
-Figure~\ref{fig:finite}b shows the estimation error versus the signal level under $d=20$.
-Interestingly, a larger estimation error is observed when the signal is either too small or too large.
-The non-monotonic behavior may seem surprising, but this is an intrinsic feature in the estimation with ordinal data.
-In view of the latent-variable interpretation (see Section~\ref{sec:latent}), estimation from ordinal observation can be interpreted as an inverse problem of quantization.
-Therefore, the estimation error diverges in the absence of noise $\tE$, because it is impossible to distinguish two different signal tensors, e.g., $\Theta_1=\ma_1\otimes \ma_2\otimes \ma_3$ and $\Theta_2=\text{sign}(\ma_1)\otimes \text{sign}(\ma_2)\otimes \text{sign}(\ma_3)$, from the quantized observations.
-This phenomenon~\cite{davenport2014,sur2019modern} is clearly contrary to the classical continuous-valued tensor problem.
-
-The second experiment investigates the incomplete observations.
-We consider $L$-level tensors with $d=20$, $\alpha=10$ and choose a subset of tensor entries via uniform sampling.
-Figure~\ref{fig:finite}c shows the estimation error of $\hat \Theta$ versus the fraction of observation $\rho=|\Omega|/d^K$.
-As expected, the error reduces with increased $\rho$ or decreased $r$.
-Figure~\ref{fig:finite}d evaluates the impact of ordinal levels $L$ to estimation accuracy, under the setting $\rho=0.5$.
-An improved performance is observed as $L$ grows, especially from binary observations ($L=2$) to multi-level ordinal observations ($L\geq 3$).
-The result showcases the benefit of multi-level observations compared to binary observations.
-
-\begin{figure}[H]
-  \centering
-\includegraphics[width=0.6\textwidth]{panel.pdf}
-\caption{Empirical relationship between (relative) MSE versus (a) dimension $d$, (b) signal level $\alpha$, (c) observation fraction $\rho$, and (d) number of ordinal levels $L$.
-In panels (b)-(d), we plot the relative MSE $=\FnormSize{}{\hat \Theta-\trueT}/\FnormSize{}{\trueT}$ for better visualization.}
-\label{fig:finite}
-\end{figure}
-
-\subsection{Comparison with alternative methods}\label{sec:compare}
-Next, we compare our ordinal tensor method ({\bf Ordinal-T}) with three popular low-rank methods:
-
-\begin{itemize}
-\item Continuous tensor decomposition ({\bf Continuous-T})~\cite{acar2010scalable} is a low-rank approximation method based on classical Tucker model.
-\item One-bit tensor completion ({\bf 1bit-T})~\cite{ghadermarzy2018learning} is a max-norm penalized tensor learning method based on partial binary observations.
-\item Ordinal matrix completion ({\bf Ordinal-M})~\cite{bhaskar2016probabilistic} is a rank-constrained matrix estimation method based on noisy, quantized observations.
-\end{itemize}
-
-We apply each of the above methods to $L$-level ordinal tensors $\tY$ generated from model~\eqref{eq:model}.
-The {\bf Continuous-T} is applied directly to $\tY$ by treating the $L$ levels as continuous observations.
-The {\bf Ordinal-M} is applied to the matrix $\tY_{(1)}$ obtained via 1-mode unfolding.
-The {\bf 1bit-T} is applied to $\tY$ in two ways.
-The first approach ({\bf 1bit-sign-T}) follows from~\citet{ghadermarzy2018learning} that transforms
-$\tY$ to a binary tensor, by taking the entrywise sign of the mean-adjusted tensor, $\tY- |\Omega|^{-1}\sum_\omega y_\omega$.
-The second approach ({\bf 1bit-category-T}) transforms the order-3 ordinal tensor $\tY$ to an order-4 binary tensor $\tY^{\sharp}=\entry{y^{\sharp}_{ijkl}}$ via dummy variable encoding; i.e., $y^{\sharp}_{ijk\ell}=\mathds{1}_{\{y_{ijk}=\ell\}}$ for all $\ell\in[L-1]$.
-We evaluate the methods by their capabilities in predicting the most likely label for each entry, i.e.,\ $y_\omega^{\text{mode}}=\arg\max_\ell\mathbb{P}(y_\omega=\ell)$.
-Two performance metrics are considered: mean absolute deviation, $\text{MAD}=d^{-K}\sum_\omega |y_\omega^{\text{mode}}-\hat y_\omega^{\text{mode}}|$, and misclassification rate, $\text{MCR}=d^{-K}\sum_\omega\mathds{1}_{\{y_\omega^{\text{mode}}\neq\text{round}(\hat y_\omega^{\text{mode}})\}}$,
-where $\text{round}(\cdot)$ denotes the nearest integer of the prediction (possibly continuous-valued returned by {\bf Continuous-T}).
-Note that MAD penalizes the large deviation more heavily than MCR.
-
-Figure~\ref{fig:compare} compares the prediction accuracy under the setting $\alpha=10$, $d=20$, and $r=5$.
-The problem size we considered is comparable to~\citet{ghadermarzy2018learning}.
-We find that our method outperforms the others in both MAD and MCR.
-In particular, methods built on multi-level observations ({\bf Ordinal-T}, {\bf Ordinal-M}, {\bf 1bit-category-T}) exhibit stable MCR over $\rho$ and $L$, whereas the others two methods ({\bf Continuous-T}, {\bf 1bit-sign-T}) generally fail except for $L=2$ (Figures~\ref{fig:compare}a-b).
-This observation highlights the necessity of modeling multi-level probabilities in classification task.
-Interestingly, although both {\bf 1bit-category-T} and our method {\bf Ordinal-T} behave similarly for binary tensors ($L=2$), the improvement of our method is substantial as $L$ increases (Figures~\ref{fig:compare}a and~\ref{fig:compare}c).
-One possible reason is that our method incorporates the intrinsic ordering among the $L$ levels via proportional odds assumption~\eqref{eq:logodd}, whereas {\bf 1bit-category-T} ignores the ordinal structure and dependence among the induced binary entries.
-Figures~\ref{fig:compare}c-d assess the prediction accuracy with sample size.
-We see a clear advantage of our method ({\bf Ordinal-T}) over the matricization ({\bf Ordinal-M}) in both complete and non-complete observations.
-When the observation fraction is small, e.g., ${|\Omega|/ d^K}=0.4$, the tensor-based completion shows $\sim$ 30\% reduction in error compared to the matricization.
-
-\begin{figure}[t]
-  \centering
-\includegraphics[width=0.6\textwidth]{panel2.pdf}
-\caption{Performance comparison in MCR (a, b) and MAD (c, d).
-(b, d) Prediction errors versus the number of ordinal levels $L$ when $\rho=0.8$.
-(a, c) Prediction errors versus sample complexity $\rho=|\Omega|/d^K$ when $L=5$.
-}\label{fig:compare}
-\end{figure}
-
-We also compare the methods by their performance in predicting the median label, $y_\omega^{\text{median}}=\min\{\ell\colon\mathbb{P}(y_\omega=\ell)\geq 0.5\}$.
-Under the latent variable model~\eqref{eq:latent} and Assumption~\ref{ass:link}, the median label is the quantized $\Theta$ without noise; i.e.\ $y_\omega^{\text{median}}=\sum_\ell \mathds{1}_{\theta_\omega\in(b_{\ell-1},b_\ell]}$.
-Similar results demonstrate the outperformance of our method (see the Supplement).
-
-\section{Data Applications}
-We apply our ordinal tensor method to two real-world datasets.
-In the first application, we use our model to analyze an ordinal tensor consisting of structural connectivities among 68 brain regions for 136 individuals from Human Connectome Project (HCP)~\citep{geddes2016human}.
-In the second application, we perform tensor completion to an ordinal dataset with missing values.
-The data tensor records the ratings on a scale of 1 to 5 from 42 users to 139 songs on 26 contexts~\citep{baltrunas2011incarmusic}.
-
-\subsection{Human Connectome Project (HCP)}
-Each entry in the HCP dataset takes value on a nominal scale, \{{\it high, moderate, low}\}, indicating the strength level of fiber connection.
-We convert the dataset to a 3-level ordinal tensor $\tY\in[3]^{68\times 68\times 136}$ and apply the ordinal tensor method with a logistic link function.
-The BIC suggests $\mr = (23,23,8)$ with $\logl(\hat{\Theta},\hat{\mb}) = -216,646$.
-Based on the estimated Tucker factors $\{\hat \mM_k\}$, we perform a clustering analysis via K-mean on the brain nodes (see detailed procedure in the Supplement).
-We find that the 68 brain nodes are grouped into eight clusters, and the clustering capture the spatial separation of brain nodes very well (see the Supplement).
-In particular, the top two clusters represent the left and right hemispheres; the smaller clusters represent local regions driving by similar nodes.
-For example, the cluster IV consists of brain nodes all of which are called {\it R.SUPRAMARGINAL} and located at the same region.
-The identified similarity among brain nodes suggests the applicability of our method.
-
-We compare the goodness-of-fit of various tensor methods on the HPC data.
-Table~\ref{table:CV} summarizes the prediction error via 5-fold stratified cross-validation averaged over 10 runs.
-Our method outperforms the others, especially in MAD.
-
-\begin{table}[http]
-\begin{tabular}{c|c|c||c|c|c}
-\Xhline{2\arrayrulewidth}
-\multicolumn{3}{c||}{Human Connectome Project (HCP) dataset} & \multicolumn{3}{c}{InCarMusic dataset}\\
-\Xhline{2\arrayrulewidth}
-Method                    & MAD & MCR & Method                    & MAD & MCR\\
-\hline
-Ordinal-T (ours) & 0.1607 (0.0005)& 0.1606 (0.0005)& Ordinal-T (ours) & 1.37 (0.039) &  0.59 (0.009)\\
- Continuous-T&0.2530 (0.0002)&0.1599 (0.0002)& Continuous-T & 2.39 (0.152)&0.94 (0.027)\\
-
- 1bit-sign-T&0.3566 (0.0010)&0.1563 (0.0010)& 1bit-sign-T&1.39 (0.003)& 0.81 (0.005)\\
-\hline
-\end{tabular}
-
-\caption{Comparison of prediction error in the HPC and InCarMusic analyses. Standard errors are reported in parentheses.}\label{table:CV}
-\end{table}
-
-\subsection{InCarMusic recommendation system}
-We apply ordinal tensor completion to a recommendation system {\it InCarMusic}.
-{\it InCarMusic} is a mobile application that offers music recommendation to passengers of cars based on contexts~\cite{baltrunas2011incarmusic}.
-Our goal is to perform tensor completion to impute the unobserved entries in the $42\times 139\times 26$ ordinal tensor and thereby we can offer context-specific music recommendation to users.
-The data tensor consists of 2,884 observed entries.
-Table~\ref{table:CV} shows the averaged prediction error via 5-fold cross validation.
-The high missing rate makes the accurate classification challenging.
-Nevertheless, our method achieves the best performance among the three.
-
-\section{Conclusions}
-We have developed a low-rank tensor estimation method based on possibly incomplete, ordinal-valued observations.
-A sharp error bound is established, and we demonstrate the outperformance of our approach compared to other methods.
-The work unlocks several directions of future research.
-One interesting question would be the inference problem, i.e..
-to assess the uncertainty of the obtained estimates and the imputation.
-Other directions include the trade-off between (non)convex optimization and statistical/computational efficiency.
-While convex relaxations are popular approach for matrix-based problem, they are often slow in practice~\cite{chen2019non}.
-The interplay between computational efficiency and statistical accuracy in general tensor problems warrants future research.
-
-\section{Proofs}\label{sec:proof}
+
+\begin{center}
+\begin{spacing}{1.5}
+\textbf{\Large Supplements for ``Tensor denoising and completion based on ordinal observations''}
+\end{spacing}
+\end{center}
+
+\section{Proofs}
 \subsection{Estimation error for tensor denoising}
@@
 
-\appendix
 \section{Extension of Theorem~\ref{thm:rate} to unknown cut-off points}
@@
 
+Figure~\ref{fig:compare2} compares the prediction accuracy under the setting $\alpha = 10, d = 20$ and $r = 5$.
+We see that our method outperforms the others in both MAD and MCR.
+The methods with probability model ({\bf Ordinal-T, Ordinal-M, 1bit-category-T}) have the tendency to be improved in MAD and be worsened in MCR when  $y_\omega^{\text{median}}$ is used (Figure~\ref{fig:compare2}) compared to the results when $y_\omega^{\text{mode}}$ is used (Figure~\ref{fig:compare}).
+This tendency is from the property that the estimator $\hat y^{(\text{Median})}_\omega$ minimizes $\mathbb{E}_{\hat{\theta}_\omega,\hat{\mb}}|y_\omega-y|$
+and the estimator $\hat y^{(\text{Mode})}_\omega $ minimizes $\mathbb{E}_{\hat{\theta}_\omega,\hat{\mb}}\mathds{1}_{\{y_\omega=y\}}$.
+By the similar reasons in Section~\ref{sec:compare} such as making use of tensor format of the datasets, accepting multi-level observation and preserving the ordinal structure, our model is benefited.
+
 \begin{figure}[H]
   \begin{center}
-    \includegraphics[width=.6\textwidth]{panel3.pdf}
+    \includegraphics[width=.8\textwidth]{panel3.pdf}
   \end{center}
@@
 \end{figure}
-
-Figure~\ref{fig:compare2} compares the prediction accuracy under the setting $\alpha = 10, d = 20$ and $r = 5$.
-We see that our method outperforms the others in both MAD and MCR.
-The methods with probability model ({\bf Ordinal-T, Ordinal-M, 1bit-category-T}) have the tendency to be improved in MAD and be worsened in MCR when  $y_\omega^{\text{median}}$ is used (Figure~\ref{fig:compare2}) compared to the results when $y_\omega^{\text{mode}}$ is used (Figure~\ref{fig:compare}).
-This tendency is from the property that the estimator $\hat y^{(\text{Median})}_\omega$ minimizes $\mathbb{E}_{\hat{\theta}_\omega,\hat{\mb}}|y_\omega-y|$
-and the estimator $\hat y^{(\text{Mode})}_\omega $ minimizes $\mathbb{E}_{\hat{\theta}_\omega,\hat{\mb}}\mathds{1}_{\{y_\omega=y\}}$.
-By the similar reasons in Section~\ref{sec:compare} such as making use of tensor format of the datasets, accepting multi-level observation and preserving the ordinal structure, our model is benefited.
 

### 2020-02-13 ICML/arxiv/ordinalT_arxiv_v2.tex
arxiv preprint

--- 2020-02-13 ICML/arxiv/icml_arxiv.tex
@@
 \usepackage{fancybox}
-
 \usepackage{color}
@@
 \usepackage[margin=1in]{geometry}
-
-\renewcommand{\textfraction}{0.0}
-\renewcommand{\topfraction}{1.0}
-%\renewcommand{\textfloatsep}{5mm}
-
 \newcommand{\theHalgorithm}{\arabic{algorithm}}
-
 \usepackage{comment}
-% Definitions of handy macros can go here
 \usepackage{amsmath,amssymb,amsthm,bm,mathtools}
 \usepackage{natbib}
-%\usepackage{dsfont,multirow,hyperref,setspace,natbib,enumerate}
 \usepackage{dsfont,multirow,hyperref,setspace,enumerate}
@@
 
-\mathtoolsset{showonlyrefs=true}
+\setlength\parindent{0pt}
+\usepackage[parfill]{parskip}
 
@@
 \newtheorem{assumption}{Assumption}
+\newtheorem{cor}{Corollary}[section]
 
@@
 \newtheorem{defn}{Definition}
-\newtheorem{cor}{Corollary}
 \newtheorem{example}{Example}
@@
 
-\setcounter{figure}{0}
-\setcounter{table}{0}
-
-\def\MLET{\hat \Theta_{\text{MLE}}}
-\newcommand{\cmt}[1]{{\leavevmode\color{red}{#1}}}
+\usepackage{pifont}
+\newcommand{\cmark}{\ding{51}}%
+\newcommand{\xmark}{\ding{55}}%
 
 \usepackage{dsfont}
-
-\DeclareMathOperator*{\minimize}{minimize}
-
 \usepackage{wrapfig}
-
 \usepackage{mathtools}
 \mathtoolsset{showonlyrefs}
+
 \newcommand*{\KeepStyleUnderBrace}[1]{%f
@@
 }
-\usepackage{xr}
-\externaldocument{ordinalTv3}
+\begingroup
+    \makeatletter
+    \@for\theoremstyle:=definition,remark,plain\do{%
+        \expandafter\g@addto@macro\csname th@\theoremstyle\endcsname{%
+            \addtolength\thm@preskip\parskip
+            }%
+        }
+\endgroup
+
 \input macros.tex
@@
 Multidimensional arrays, a.k.a.\ tensors, arise in a variety of applications including recommendation systems~\citep{baltrunas2011incarmusic}, social networks~\citep{nickel2011three}, genomics~\citep{hore2016tensor}, and neuroimaging~\citep{zhou2013tensor}.
-There is a growing need to develop general methods for analyzing these noisy, high-dimensional datasets that can handle two main problems.
-The problem of tensor denoising -- which aims to recover a signal tensor from its noisy entries -- has gained increased attention in theory and applications~\citep{xia2019sup,zeng2019multiway}.
-A related problem, tensor completion, examines the minimum number of entries needed for a consistent recovery~\citep{ghadermarzy2018learning,ghadermarzy2019near}.
-Low-rankness is often imposed to the signal tensor, which efficiently reduces the intrinsic dimension in both problems.
-
-A number of low-rank tensor estimation methods have been proposed~\citep{kolda2009tensor,acar2010scalable}, revitalizing classical methods such as CANDECOMP/PARAFAC (CP) decomposition~\citep{hitchcock1927expression} and Tucker decomposition~\citep{tucker1966some}.
+There is a growing need to develop general methods that can handle two main problems for analyzing these noisy, high-dimensional datasets.
+The first problem is tensor denoising which aims to recover a signal tensor from its noisy entries~\citep{hong2018generalized,zeng2019multiway}.
+A second problem is tensor completion which examines the minimum number of entries needed for a consistent recovery~\citep{ghadermarzy2018learning,ghadermarzy2019near}.
+Low-rankness is often imposed to the signal tensor, thereby efficiently reducing the intrinsic dimension in both problems.
+
+A number of low-rank tensor estimation methods have been proposed~\citep{anandkumar2014tensor,wang2017tensor}, revitalizing classical methods such as CANDECOMP/PARAFAC (CP) decomposition~\citep{hitchcock1927expression} and Tucker decomposition~\citep{tucker1966some}.
 These tensor methods treat the entries as continuous-valued.
@@
 For example, the Netflix problem records the ratings of users on movies over time.
-Each data is a rating on a nominal scale \{{\it very like, like, neutral, dislike, very dislike}\}.
+Each entry is a rating on a nominal scale \{{\it very like, like, neutral, dislike, very dislike}\}.
 Another example is in the signal processing, where the digits are frequently rounded or truncated so that only integer values are available.
@@
 Second, the parameter interpretations should be consistent under merging or splitting of contiguous categories.
-The classical continuous tensor model~\cite{kolda2009tensor, ghadermarzy2019near} fails in the first aspect, whereas the binary tensor model~\cite{ghadermarzy2018learning} lacks the second property.
+The classical continuous tensor model~\citep{kolda2009tensor, ghadermarzy2019near} fails in the first aspect, whereas the binary tensor model~\citep{ghadermarzy2018learning} lacks the second property.
 An appropriate model for ordinal tensors has yet to be studied.
-
-\begin{table*}[ht]
-\centerline{\includegraphics[width=16cm]{compare.pdf}}
-\caption{Comparison with previous work.
-For ease of presentation, we summarize the error rate and sample complexity assuming equal tensor dimension in all modes.
-$K$: tensor order; $L$: number of ordinal levels; $d$: dimension at each mode.
-}~\label{tab:compare}
-\end{table*}
 
@@
 The mean squared error bound is established, and we show that the obtained bound has minimax optimal rate in high dimensions under the low-rank model.
-Our estimator enjoys a faster convergence rate $\tO(d^{-(K-1)/2})$ than $\tO(d^{-K})$ in~\citet{ghadermarzy2018learning}, which is a substantial improvement as the order $K$ increases.
+Our estimator enjoys a faster convergence rate $\tO(d^{-(K-1)/2})$ than $\tO(d^{-K})$ in~\citet{ghadermarzy2018learning}, which is a substantial improvement for higher-order tensors.
 Furthermore, our proposal serves as an efficient completion algorithm that guarantees consistent recovery of an order-$K$ $(d,\ldots,d)$-dimensional low-rank tensor using only $\tilde \tO(Kd)$ noisy, quantized observations.
 
+\begin{table}[ht]
+\resizebox{\columnwidth}{!}{
+\begin{tabular}{l|ccc}
+&\cite{bhaskar2016probabilistic}&\cite{ghadermarzy2018learning} &This paper\\
+\hline 
+Higher-order tensors ($K\geq 3$) & \xmark&\cmark& \cmark\\
+\hline
+Multi-level categories ($L\geq 3$)& \cmark& \xmark&\cmark\\
+ \hline
+ Error rate for tensor denoising&$d^{-1}$ for $K=2$& $d^{-(K-1)/2}$&$d^{-(K-1)}$\\
+\hline
+Optimality guarantee under low-rank models& unknown & \xmark & \cmark\\
+\hline
+Sample complexity for tensor completion&$d^K$& $Kd$&$Kd$\\
+\hline
+\end{tabular}
+}
+\caption{Comparison with previous work when tensor rank $r=O(1)$.
+For ease of presentation, we summarize the error rate and sample complexity (neglecting log factors) when the tensor dimensions are equal in all modes.
+$K$: tensor order; $L$: number of ordinal levels; $d$: dimension at each mode.
+}~\label{tab:compare}
+\end{table}
+
 {\bf Related work.} Our work is related to, but clearly distinctive from, several lines of existing literature.
-Matrix completion from quantized samples was firstly introduced for binary observations~\cite{cai2013max,davenport2014,bhaskar20151} and then extended to ordinal observations~\cite{bhaskar2016probabilistic}.
+Matrix completion from quantized samples was firstly introduced for binary observations~\citep{cai2013max,davenport2014,bhaskar20151} and then extended to ordinal observations~\citep{bhaskar2016probabilistic}.
 As we show in Section~\ref{sec:theory}, applying existing matrix methods to an ordinal tensor results in a suboptimal estimator with a slower convergence rate.
@@
 We address the problems from two perspectives.
-From statistical perspective, our proposed model generalizes the usual binary tensor model while preserving palindromic invariance~\cite{mccullagh1980regression} for ordinal observations.
-From algorithm perspective, our alternating optimization compares favorably to the approximate (non-convex) algorithm developed in the context of binary tensors~\cite{ghadermarzy2018learning}.
+From statistical perspective, our proposed model generalizes the usual binary tensor model while preserving palindromic invariance~\citep{mccullagh1980regression} for ordinal observations.
+From algorithm perspective, our alternating optimization compares favorably to the approximate (non-convex) algorithm developed in the context of binary tensors~\citep{ghadermarzy2018learning}.
 We numerically compare the two approaches in Section~\ref{sec:experiment}.
@@
 We say that an event $A$ occurs ``with very high probability'' if $\mathbb{P}(A)$ tends to 1 faster than any polynomial of tensor dimension $d_{\min}=\min\{d_1,\ldots,d_K\} \to\infty$.
+For any two functions $f, g$ depending on $(d_1,\ldots,d_K)$, we write $f=\tO(g)$ to indicate that $f\leq Cg$, where $C>0$ is a constant independent of tensor dimension.
+We write $f=\tilde\tO(g)$ to indicate that $f\leq C(\log d_{\min})^\beta g$ for some $\beta>0$.
 
@@
 \end{example}
+
 \begin{example}[Probit model] The probit model is characterized by~\eqref{eq:model} with
@@
 \end{example}
-Other link functions are also possible, such as Laplace, Cauchy, etc~\cite{mccullagh1980regression}.
+
+Other link functions are also possible, such as Laplace, Cauchy, inverse log-log, etc~\citep{mccullagh1980regression}.
 All the models share the property that the ordered categories can be thought of as contiguous interval on some continuous scale.
@@
 Therefore, our model~\eqref{eq:model} is general and still valid in the absence of quantization process.
-More generally, we make the following assumptions about the link $f$.
+More generally, we make the following assumptions about the link $f(\cdot): \mathbb{R}\mapsto[0,1]$.
 
@@
 \begin{enumerate}
-\item $f(\theta)$ is strictly increasing and twice-differentiable in $\theta\in\mathbb{R}/\{0\}$.
+\item $f(\theta)$ is strictly increasing and twice-differentiable in $\theta$.
 \item $f'(\theta)$ is strictly log-concave and symmetric with respect to $\theta=0$.
@@
 
-(P1) Given the quantization process induced by $f$ and the cut-off points $\mb$, how accurately can we estimate the latent signal tensor $\Theta$ from the ordinal observation $\tY$?
+{\bf (P1)} Given the quantization process induced by $f$ and the cut-off points $\mb$, how accurately can we estimate the latent signal tensor $\Theta$ from the ordinal observation $\tY$?
 
 Clearly, the problem (P1) cannot be solved uniformly for all possible $\Theta$.
-We focus on a class of ``low-rank'' and ``flat'' signal tensors, which is a plausible assumption in practical applications~\cite{zhou2013tensor,bhaskar20151}.
+We focus on a class of ``low-rank'' and ``flat'' signal tensors, which is a plausible assumption in practical applications~\citep{zhou2013tensor,bhaskar20151}.
 Specifically, we consider the parameter space:
 \begin{equation}\label{eq:space}
-\small \tP=\left\{\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K} \colon \text{rank}(\Theta)\leq \mr,\ \mnormSize{}{\Theta}\leq \alpha\right\}.
+ \tP=\left\{\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K} \colon \text{rank}(\Theta)\leq \mr,\ \mnormSize{}{\Theta}\leq \alpha\right\}.
 \end{equation}
@@
 \end{equation}
-where $\tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is a core tensor, $\mM_k\in\mathbb{R}^{d_k\times r_k}$ are factor matrices with orthogonal columns, and $\times_k$ denotes the tensor-by-matrix multiplication~\cite{kolda2009tensor}.
+where $\tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is a core tensor, $\mM_k\in\mathbb{R}^{d_k\times r_k}$ are factor matrices with orthogonal columns, and $\times_k$ denotes the tensor-by-matrix multiplication~\citep{kolda2009tensor}.
 The Tucker low-rankness is popularly imposed in tensor data analysis, and is shown to provide a reasonable tradeoff between model complexity and model flexibility.
-Note that, unlike matrices, there are various notations of tensor low-rankness, such as CP rank~\cite{hitchcock1927expression} and train rank~\cite{oseledets2011tensor}.
-Some notation of low-rankness may lead to mathematically ill-posed optimization; for example, the best low CP-rank tensor approximation may not exist~\cite{de2008tensor}.
+Note that, unlike matrices, there are various notations of tensor low-rankness, such as CP rank~\citep{hitchcock1927expression} and train rank~\citep{oseledets2011tensor}.
+Some notation of low-rankness may lead to mathematically ill-posed optimization; for example, the best low CP-rank tensor approximation may not exist~\citep{de2008tensor}.
 We choose Tucker representation for well-posedness of optimization and easy interpretation.
@@
 
-(P2) Given an incomplete set of ordinal observations $\{y_{\omega}\}_{\omega\in\Omega}$, how many sampled entries do we need to consistently recover $\Theta$ based on the model~\eqref{eq:model}?
+{\bf (P2)} Given an incomplete set of ordinal observations $\{y_{\omega}\}_{\omega\in\Omega}$, how many sampled entries do we need to consistently recover $\Theta$ based on the model~\eqref{eq:model}?
 
@@
 In classical matrix completion, the incoherence is often imposed on the singular vectors.
-This assumption is recently relaxed for ``flat'' matrices with bounded magnitude~\cite{negahban2011estimation,cai2013max,bhaskar20151}.
+This assumption is recently relaxed for ``flat'' matrices with bounded magnitude~\citep{negahban2011estimation,cai2013max,bhaskar20151}.
 We adopt the same assumption for higher-order tensors.
@@
 \begin{align}\label{eq:objective}
-\logl(\Theta, \mb)&=\sum_{\omega\in\Omega}\sum_{\ell\in[L]} \Big\{\mathds{1}_{\{y_\omega=\ell\}} \log \big[f(b_\ell-\theta_\omega)- \notag \\
-&\quad \quad \quad \quad \quad \quad \quad  f(b_{\ell-1}-\theta_\omega)\big]\Big\}.
+\logl(\Theta, \mb)&=\sum_{\omega\in\Omega}\sum_{\ell\in[L]} \Big\{\mathds{1}_{\{y_\omega=\ell\}} \log \big[f(b_\ell-\theta_\omega)-  f(b_{\ell-1}-\theta_\omega)\big]\Big\}.
 \end{align}
@@
 For technical convenience, we assume in this section that the cut-off points $\mb$ are known.
-The adaptation of unknown $\mb$ is addressed in Section~\ref{sec:algorithm} and the Supplement.
-
-We define a few key quantities that will be used in our theory. Let $g_\ell=f(b_\ell-\theta)-f(b_{\ell-1}-\theta)$ for all $\ell\in[L]$, and
-\begin{align}\label{eq:regular}
-A_\alpha &= \min_{\ell\in[L], |\theta|\leq \alpha} g_\ell(\theta),\quad  U_\alpha=\max_{\ell\in[L], |\theta|\leq \alpha } {|\dot{g}_\ell(\theta)|\over g_\ell(\theta)},\\
- L_\alpha &= \min_{\ell\in[L], |\theta|\leq \alpha }\left[{\dot{g}_\ell^2 (\theta)\over g_\ell^2(\theta)} -{\ddot{g}_\ell(\theta)\over g_\ell(\theta)}\right],
-\end{align}
+The adaptation of unknown $\mb$ is addressed in Section~\ref{sec:algorithm} and Appendix~\ref{sec:extention}.
+
+We define a few key quantities that will be used in our theory. Let $g_\ell(\theta)=f(b_\ell-\theta)-f(b_{\ell-1}-\theta)$ for all $\ell\in[L]$, and
+\begin{equation}\label{eq:regular}
+A_\alpha = \min_{\ell\in[L], |\theta|\leq \alpha} g_\ell(\theta),\quad  U_\alpha=\max_{\ell\in[L], |\theta|\leq \alpha } {|\dot{g}_\ell(\theta)|\over g_\ell(\theta)},\quad \text{and}\quad
+ L_\alpha = \min_{\ell\in[L], |\theta|\leq \alpha }\left[{\dot{g}_\ell^2 (\theta)\over g_\ell^2(\theta)} -{\ddot{g}_\ell(\theta)\over g_\ell(\theta)}\right],
+\end{equation}
 where $\dot{g}(\theta)=dg(\theta)/d\theta$, and $\alpha$ is the entrywise bound of $\Theta$.
@@
 \[
-\text{MSE}(\hat \Theta, \trueT)={1\over \prod_k d_k}\FnormSize{}{\Theta-\trueT}^2.
+\mathrm{MSE}(\hat \Theta, \trueT)={1\over \prod_k d_k}\FnormSize{}{\Theta-\trueT}^2.
 \]
@@
 \begin{equation}\label{eq:rate}
-\text{MSE}(\hat \Theta, \trueT) \leq \min\left( 4\alpha^2,\ {c_2  U^2_\alpha r_{\max}^{K-1}  \over  L^2_\alpha } {\sum_kd_k\over  \prod_k d_k} \right),
-\end{equation}
-where $c_1, c_2>0$ are two constants that depend only on $K$.
+\mathrm{MSE}(\hat \Theta, \trueT) \leq \min\left( 4\alpha^2,\ {c_2  U^2_\alpha r_{\max}^{K-1}  \over  L^2_\alpha } {\sum_kd_k\over  \prod_k d_k} \right),
+\end{equation}
+where $c_1, c_2>0$ are two constants that depend only on $K$.\\
 \end{thm}
+
 Theorem~\ref{thm:rate} establishes the statistical convergence for the estimator~\eqref{eq:estimator}.
-In fact, the proof of this theorem (see Section~\ref{sec:proof}) shows that the same statistical rate holds, not only for the global optimizer~\eqref{eq:estimator}, but also for any local optimizer $\check \Theta$ in the level set $\{\check\Theta\in\tP: \logl(\check\Theta)\geq \logl(\trueT)\}$.
+In fact, the proof of this theorem (see Section~\ref{sec:proofMSE}) shows that the same statistical rate holds, not only for the global optimizer~\eqref{eq:estimator}, but also for any local optimizer $\check \Theta$ in the level set $\{\check\Theta\in\tP: \logl(\check\Theta)\geq \logl(\trueT)\}$.
 This suggests that the local optimality itself is not necessarily a severe concern in our context, as long as the convergent objective is large enough.
-In Section ~\ref{sec:algorithm}, we perform empirical studies to assess the algorithmic stability.
+In Section~\ref{sec:algorithm}, we perform empirical studies to assess the algorithmic stability.
 
 A similar conclusion is obtained for the prediction error, measured in Kullback-Leibler (KL) divergence, between the categorical distributions in the observation space.
+
 \begin{cor}[Prediction error]~\label{cor:prediction}
@@
 \begin{equation}\label{eq:KLrate}
-\text{KL}(\mathbb{P}_{\tY} || \hat{\mathbb{P}}_{\tY}) \leq  {c_2 U^2_\alpha r_{\max}^{K-1}  \over L^2_\alpha } {(4L-6)\dot{f}^2(0)  \over A_\alpha} {\sum_kd_k\over  \prod_k d_k},
+\mathrm{KL}(\mathbb{P}_{\tY} || \hat{\mathbb{P}}_{\tY}) \leq  {c_2 U^2_\alpha r_{\max}^{K-1}  \over L^2_\alpha } {(4L-6)\dot{f}^2(0)  \over A_\alpha} {\sum_kd_k\over  \prod_k d_k},
 \end{equation}
@@
 \end{cor}
+
 To gain insight into these bounds, we consider a special setting with equal dimension in all modes, i.e., $d_1=\cdots=d_K=d$. In such a case, our bound \eqref{eq:rate} reduces to
-\begin{equation*}\label{eq:ours}
+\begin{equation}\label{eq:ours}
 \text{MSE}(\hat \Theta, \trueT) \asymp d^{-(K-1)}, \quad \text{as}\ d\to \infty.
-\end{equation*}
+\end{equation}
 Hence, our estimator achieves consistency with polynomial convergence rate.
@@
 \end{align}
-where $C=C(\alpha, L, f,\mb)>0$ and $c>0$ are constants independent of tensor dimension and the rank.
+where $C=C(\alpha, L, f,\mb)>0$ and $c>0$ are constants independent of tensor dimension and the rank.\\
 \end{thm}
+
 We see that the lower bound matches the upper bound in~\eqref{eq:rate} on the polynomial order of tensor dimension. Therefore, our estimator~\eqref{eq:estimator} is order-optimal.
@@
 Note that $\tilde \tO(Kd)$ roughly matches the degree of freedom for an order-$K$ tensor of fixed rank $\mr$, suggesting the optimality of our sample requirement.
-This sample complexity substantially improves over earlier result $\tO(d^{\lceil K/2\rceil})$ based on square matricization~\cite{mu2014square}, or $\tO(d^{N/2})$ based on tensor nuclear-norm regularization~\cite{yuan2016tensor}.
-Existing methods that achieve $\tilde \tO(Kd)$ sample complexity require either a deterministic cross sampling design~\cite{zhang2019cross} or univariate measurements~\cite{ghadermarzy2018learning}.
+This sample complexity substantially improves over earlier result $\tO(d^{\lceil K/2\rceil})$ based on square matricization~\citep{mu2014square}, or $\tO(d^{N/2})$ based on tensor nuclear-norm regularization~\citep{yuan2016tensor}.
+Existing methods that achieve $\tilde \tO(Kd)$ sample complexity require either a deterministic cross sampling design~\citep{zhang2019cross} or univariate measurements~\citep{ghadermarzy2018learning}.
 Our method extends the conclusions to multi-level measurements under a broader class of sampling schemes.
@@
 In practice, the cut-off points $\mb$ are often unknown, so we choose to maximize $\logl$ jointly over $(\Theta,\mb)\in\tP\times \tB$.
-The objective $\logl$ is concave in $(\Theta,\mb)$ whenever $f'$ is log-concave (see Section~\ref{sec:proof}).
+The objective $\logl$ is concave in $(\Theta,\mb)$ whenever $f'$ is log-concave (see Section~\ref{sec:proofconvexity}).
 However, the feasible set $\tP$ is non-convex, which makes the optimization~\eqref{eq:objective} a non-convex problem.
@@
             \STATE $\mM_k^{(t+1)}\gets\argmax_{\mM_k\in\mathbb{R}^{d_k\times r_K}}\mathcal{L}_{\mathcal{Y},\Omega}(\mM_k)$,
-            \newline
            s.t. $\mnormSize{}{\Theta^{(t+1)}}\leq \alpha$, where $\Theta^{(t+1)}$ is the parameter tensor based on the current block estimates.
@@
             \STATE {Update $\mb^{(t+1)}$ while fixing $\Theta^{(t+1)}$:} 
-              \newline
               $\mb^{(t+1)} \gets \argmax_{\mb\in \tB}\mathcal{L}_{\mathcal{Y},\Omega}\big(\Theta^{(t+1)},\mb\big).$
@@
 
-\begin{figure}[H]
+\begin{figure}[ht]
 \centering
@@
 \hat \mr&=\argmin_{\mr\in\mathbb{N}^K_{+}} \text{BIC}(\mr)\\
-&=\argmin_{\mr\in\mathbb{N}^K_{+}}\{-2\tL_{\tY}(\hat \Theta(\mr),\hat \mb(\mr))+p_e(\mr)\log (\prod_k d_k)\},
+&=\argmin_{\mr\in\mathbb{N}^K_{+}}\left\{-2\tL_{\tY}(\hat \Theta(\mr),\hat \mb(\mr))+p_e(\mr)\log \left(\prod_k d_k\right)\right\},
 \end{align}
@@
 We consider the setting with $K=3$, $d_1=d_2=d_3=d$, and $r_1=r_2=r_3=r$.
-The parameter tensors are simulated based on~\eqref{eq:Tucker}, where the core tensor entries are i.i.d.\ drawn from $N(0,1)$, and the factors $\mM_k$ are uniformly sampled (with respect to Haar measure) from matrices with orthonormal columns.
+The parameter tensors are simulated based on~\eqref{eq:Tucker}, where the core tensor entries are i.i.d.\ drawn from $\tN(0,1)$, and the factors $\mM_k$ are uniformly sampled (with respect to Haar measure) from matrices with orthonormal columns.
 We set the cut-off points $b_\ell=f^{-1}(\ell/L)$ for $\ell\in[L]$, such that $f(b_\ell)$ are evenly spaced from 0 to 1.
@@
 Therefore, the estimation error diverges in the absence of noise $\tE$, because it is impossible to distinguish two different signal tensors, e.g., $\Theta_1=\ma_1\otimes \ma_2\otimes \ma_3$ and $\Theta_2=\text{sign}(\ma_1)\otimes \text{sign}(\ma_2)\otimes \text{sign}(\ma_3)$, from the quantized observations.
-This phenomenon~\cite{davenport2014,sur2019modern} is clearly contrary to the classical continuous-valued tensor problem.
+This phenomenon~\citep{davenport2014,sur2019modern} is clearly contrary to the classical continuous-valued tensor problem.
 
@@
 
-\begin{figure}[H]
+\begin{figure}[ht]
   \centering
@@
 \begin{itemize}
-\item Continuous tensor decomposition ({\bf Continuous-T})~\cite{acar2010scalable} is a low-rank approximation method based on classical Tucker model.
-\item One-bit tensor completion ({\bf 1bit-T})~\cite{ghadermarzy2018learning} is a max-norm penalized tensor learning method based on partial binary observations.
-\item Ordinal matrix completion ({\bf Ordinal-M})~\cite{bhaskar2016probabilistic} is a rank-constrained matrix estimation method based on noisy, quantized observations.
+\item Continuous tensor decomposition ({\bf Continuous-T})~\citep{acar2010scalable} is a low-rank approximation method based on classical Tucker model.
+\item One-bit tensor completion ({\bf 1bit-T})~\citep{ghadermarzy2018learning} is a max-norm penalized tensor learning method based on partial binary observations.
+\item Ordinal matrix completion ({\bf Ordinal-M})~\citep{bhaskar2016probabilistic} is a rank-constrained matrix estimation method based on noisy, quantized observations.
 \end{itemize}
@@
 The {\bf 1bit-T} is applied to $\tY$ in two ways.
-The first approach ({\bf 1bit-sign-T}) follows from~\citet{ghadermarzy2018learning} that transforms
-$\tY$ to a binary tensor, by taking the entrywise sign of the mean-adjusted tensor, $\tY- |\Omega|^{-1}\sum_\omega y_\omega$.
+The first approach ({\bf 1bit-sign-T}) follows from~\citet{ghadermarzy2018learning} that transforms $\tY$ to a binary tensor, by taking the entrywise sign of the mean-adjusted tensor, $\tY- |\Omega|^{-1}\sum_\omega y_\omega$.
 The second approach ({\bf 1bit-category-T}) transforms the order-3 ordinal tensor $\tY$ to an order-4 binary tensor $\tY^{\sharp}=\entry{y^{\sharp}_{ijkl}}$ via dummy variable encoding; i.e., $y^{\sharp}_{ijk\ell}=\mathds{1}_{\{y_{ijk}=\ell\}}$ for all $\ell\in[L-1]$.
+
 We evaluate the methods by their capabilities in predicting the most likely label for each entry, i.e.,\ $y_\omega^{\text{mode}}=\arg\max_\ell\mathbb{P}(y_\omega=\ell)$.
-Two performance metrics are considered: mean absolute deviation, $\text{MAD}=d^{-K}\sum_\omega |y_\omega^{\text{mode}}-\hat y_\omega^{\text{mode}}|$, and misclassification rate, $\text{MCR}=d^{-K}\sum_\omega\mathds{1}_{\{y_\omega^{\text{mode}}\neq\text{round}(\hat y_\omega^{\text{mode}})\}}$,
+Two performance metrics are considered: mean absolute deviation, $\text{MAD}(\tY^{\text{mode}}, \hat \tY^{\text{mode}})=d^{-K}\sum_\omega |y_\omega^{\text{mode}}-\hat y_\omega^{\text{mode}}|$, and misclassification rate, $\text{MCR}(\tY^{\text{mode}}, \hat \tY^{\text{mode}})=d^{-K}\sum_\omega\mathds{1}_{\{y_\omega^{\text{mode}}\neq\text{round}(\hat y_\omega^{\text{mode}})\}}$,
 where $\text{round}(\cdot)$ denotes the nearest integer of the prediction (possibly continuous-valued returned by {\bf Continuous-T}).
+Both metrics are widely used for evaluation of prediction accuracy.
 Note that MAD penalizes the large deviation more heavily than MCR.
@@
 
-\begin{figure}[t]
+\begin{figure}[ht]
   \centering
 \includegraphics[width=0.6\textwidth]{panel2.pdf}
-\caption{Performance comparison in MCR (a, b) and MAD (c, d).
-(b, d) Prediction errors versus the number of ordinal levels $L$ when $\rho=0.8$.
-(a, c) Prediction errors versus sample complexity $\rho=|\Omega|/d^K$ when $L=5$.
+\caption{Performance comparison for predicting most likely labels.
+(a, c) Prediction error versus sample complexity $\rho=|\Omega|/d^K$ when $L=5$.
+(b, d) Prediction error versus the number of ordinal levels $L$ when $\rho=0.8$.
 }\label{fig:compare}
@@
 
-We also compare the methods by their performance in predicting the median label, $y_\omega^{\text{median}}=\min\{\ell\colon\mathbb{P}(y_\omega=\ell)\geq 0.5\}$.
-Under the latent variable model~\eqref{eq:latent} and Assumption~\ref{ass:link}, the median label is the quantized $\Theta$ without noise; i.e.\ $y_\omega^{\text{median}}=\sum_\ell \mathds{1}_{\theta_\omega\in(b_{\ell-1},b_\ell]}$.
-Similar results demonstrate the outperformance of our method (see the Supplement).
+We also compare the methods by their performance in predicting the median labels, $y_\omega^{\text{median}}=\min\{\ell\colon\mathbb{P}(y_\omega=\ell)\geq 0.5\}$.
+Under the latent variable model~\eqref{eq:latent} and Assumption~\ref{ass:link}, the median label is the quantized $\theta_\omega$ without noise; i.e.\ $y_\omega^{\text{median}}=\sum_\ell \mathds{1}_{\theta_\omega\in(b_{\ell-1},b_\ell]}$.
+We utilize the same simulation setting as in the earlier experiment.
+Figure~\ref{fig:compare2} shows that our method outperforms the others in both MCR and MAD.
+The improved accuracy comes from the incorporation of multilinear low-rank structure, multi-level observations, and the ordinal structure.
+Interestingly, for the three multilevel methods ({\bf 1bit-sign-T}, {\bf Ordinal-M}, and {\bf Ordinal-T}), the median estimator tends to yield smaller MAD than the mode estimator, $\text{MAD}(\tY^{\text{median}}, \hat \tY^{\text{median}})\leq \text{MAD}(\tY^{\text{mode}}, \hat \tY^{\text{mode}})$  (Figures~\ref{fig:compare}a-b vs.\ Figures~\ref{fig:compare2}a-b).
+On the other hand, the mode estimator tends to yield smaller MCR than the median estimator, $\text{MCR}(\tY^{\text{mode}}, \hat \tY^{\text{mode}})\leq\text{MCR}(\tY^{\text{median}}, \hat \tY^{\text{median}})$ (Figures~\ref{fig:compare}c-d vs.\ Figures~\ref{fig:compare2}c-d).
+This tendency is from the property that the median estimator $\hat y^{(\text{median})}_\omega$ minimizes $g_1(z)=\mathbb{E}_{y_\omega}|y_\omega-z|$, whereas the mode estimator $\hat y^{(\text{mode})}_\omega $ minimizes $g_2(z)=\mathbb{E}_{y_\omega}\mathds{1}_{\{y_\omega=z\}}$.
+Here the expectation is over the categorical distribution of $y_\omega$ given parameters $\hat \Theta$ and $\hat \mb$.
+
+\begin{figure}[ht]
+\begin{center}
+\includegraphics[width=.6\textwidth]{panel3.pdf}
+\end{center}
+\caption{Performance comparison for predicting median labels.
+(a, c) Prediction error versus sample complexity $\rho=|\Omega|/d^K$ when $L=5$.
+(b, d) Prediction error versus the number of ordinal levels $L$, when $\rho=0.8.$}\label{fig:compare2}
+\end{figure}
 
@@
 We apply our ordinal tensor method to two real-world datasets.
-In the first application, we use our model to analyze an ordinal tensor consisting of structural connectivities among 68 brain regions for 136 individuals from Human Connectome Project (HCP)~\citep{geddes2016human}.
+In the first application, we use our model to analyze an ordinal tensor consisting of structural connectivities among 68 brain regions for 136 individuals from Human Connectome Project (HCP)~\citep{van2013wu}.
 In the second application, we perform tensor completion to an ordinal dataset with missing values.
@@
 
-\subsection{Human Connectome Project (HCP)}
-Each entry in the HCP dataset takes value on a nominal scale, \{{\it high, moderate, low}\}, indicating the strength level of fiber connection.
+{\bf Human Connectome Project (HCP).} Each entry in the HCP dataset takes value on a nominal scale, \{{\it high, moderate, low}\}, indicating the strength level of fiber connection.
 We convert the dataset to a 3-level ordinal tensor $\tY\in[3]^{68\times 68\times 136}$ and apply the ordinal tensor method with a logistic link function.
 The BIC suggests $\mr = (23,23,8)$ with $\logl(\hat{\Theta},\hat{\mb}) = -216,646$.
-Based on the estimated Tucker factors $\{\hat \mM_k\}$, we perform a clustering analysis via K-mean on the brain nodes (see detailed procedure in the Supplement).
-We find that the 68 brain nodes are grouped into eight clusters, and the clustering capture the spatial separation of brain nodes very well (see the Supplement).
-In particular, the top two clusters represent the left and right hemispheres; the smaller clusters represent local regions driving by similar nodes.
-For example, the cluster IV consists of brain nodes all of which are called {\it R.SUPRAMARGINAL} and located at the same region.
-The identified similarity among brain nodes suggests the applicability of our method.
-
-We compare the goodness-of-fit of various tensor methods on the HPC data.
-Table~\ref{table:CV} summarizes the prediction error via 5-fold stratified cross-validation averaged over 10 runs.
+Based on the estimated Tucker factors $\{\hat \mM_k\}$, we perform a clustering analysis via K-mean on the brain nodes (see Appendix~\ref{sec:clustering} for detailed procedure).
+The 68 brain nodes are grouped into eight clusters.
+We find that the clustering capture the spatial separation between brain regions (Supplementary Figure~\ref{figure:brain image}).
+In particular, the top two clusters represent the left and right hemispheres; the smaller clusters represent local regions driving by similar nodes (Supplementary Table~\ref{table:clustering}).
+For example, the cluster III/IV consists of nodes in the supramarginal gyrus region in the left/right hemisphere.
+This region is known to be involved in visual word recognition and reading~\citep{stoeckel2009supramarginal}.
+The spatial separation between clusters suggests the applicability of our clustering method even without knowledge of external annotations.
+
+We also compare the goodness-of-fit of various tensor methods on the HPC data.
+We perform 5-fold cross-validation while preserving the label proportions in train/test sets.
+Table~\ref{table:CV} summarizes the prediction error averaged over 10 runs.
 Our method outperforms the others, especially in MAD.
 
-\begin{table}[http]
+{\bf InCarMusic recommendation system.} We apply ordinal tensor completion to a recommendation system InCarMusic.
+InCarMusic is a mobile application that offers music recommendation to passengers of cars based on contexts~\citep{baltrunas2011incarmusic}.
+Our goal is to perform tensor completion to impute the unobserved entries in the $42\times 139\times 26$ ordinal tensor and thereby we can offer context-specific music recommendation to users.
+The data tensor consists of 2,884 observed entries.
+Table~\ref{table:CV} shows the averaged prediction error via 5-fold cross validation.
+The high missing rate makes the accurate classification challenging.
+Nevertheless, our method achieves the best performance among the three.
+
+\begin{table}[ht]
+\resizebox{\columnwidth}{!}{
 \begin{tabular}{c|c|c||c|c|c}
@@
 Continuous-T&0.2530 (0.0002)&0.1599 (0.0002)& Continuous-T & 2.39 (0.152)&0.94 (0.027)\\
-
 1bit-sign-T&0.3566 (0.0010)&0.1563 (0.0010)& 1bit-sign-T&1.39 (0.003)& 0.81 (0.005)\\
@@
 \end{tabular}
-
+}
 \caption{Comparison of prediction error in the HPC and InCarMusic analyses. Standard errors are reported in parentheses.}\label{table:CV}
@@
 
-\subsection{InCarMusic recommendation system}
-We apply ordinal tensor completion to a recommendation system {\it InCarMusic}.
-{\it InCarMusic} is a mobile application that offers music recommendation to passengers of cars based on contexts~\cite{baltrunas2011incarmusic}.
-Our goal is to perform tensor completion to impute the unobserved entries in the $42\times 139\times 26$ ordinal tensor and thereby we can offer context-specific music recommendation to users.
-The data tensor consists of 2,884 observed entries.
-Table~\ref{table:CV} shows the averaged prediction error via 5-fold cross validation.
-The high missing rate makes the accurate classification challenging.
-Nevertheless, our method achieves the best performance among the three.
-
-\section{Conclusions}
-We have developed a low-rank tensor estimation method based on possibly incomplete, ordinal-valued observations.
-A sharp error bound is established, and we demonstrate the outperformance of our approach compared to other methods.
-The work unlocks several directions of future research.
-One interesting question would be the inference problem, i.e..
-to assess the uncertainty of the obtained estimates and the imputation.
-Other directions include the trade-off between (non)convex optimization and statistical/computational efficiency.
-While convex relaxations are popular approach for matrix-based problem, they are often slow in practice~\cite{chen2019non}.
-The interplay between computational efficiency and statistical accuracy in general tensor problems warrants future research.
-
 \section{Proofs}\label{sec:proof}
-\subsection{Estimation error for tensor denoising}
+Here, we provide proofs of the theoretical results presented in Sections~\ref{sec:theory}.
+
+\subsection{Estimation error for tensor denoising}\label{sec:proofMSE}
 \begin{proof}[Proof of Theorem~\ref{thm:rate}]
@@
 
-\subsection{Convexity of the log-likelihood function}
+\subsection{Convexity of the log-likelihood function}\label{sec:proofconvexity}
 \begin{thm}\label{thm:convexity}
@@
  \end{align}
-where $f(\cdot)$ satisfies Assumption~\eqref{eq:ass}. Then, $\logl(\Theta, \mb)$ is concave in $(\Theta,\mb)$.
+where $f(\cdot)$ satisfies Assumption~\ref{ass:link}. Then, $\logl(\Theta, \mb)$ is concave in $(\Theta,\mb)$.
 \end{thm}
@@
 Define $d_{\text{total}}=\prod_k d_k$.
-With a little abuse of notation, we use $(\Theta,\mb)$ to denote a length-$(d_{\text{total}}+L-1)$-vector collecting all parameters together.
+By abuse of notation, we use $(\Theta,\mb)$ to denote the length-$(d_{\text{total}}+L-1)$-vector collecting all parameters together.
 Let us denote a bivariate function
@@
 Then, $\lambda(u,v) = \lambda(\ma_1^T(\Theta,\mb),\ \ma_2^T(\Theta,\mb))$ is concave in $(\Theta,\mb)$ by the definition of concavity.
-Furthermore, we can conclude that $\logl(\Theta, \mb) $ is concave in $(\Theta,\mb)$ because $\logl(\Theta, \mb)$ is the sum of $\lambda(u,v)$.
+Therefore, we can conclude that $\logl(\Theta, \mb) $ is concave in $(\Theta,\mb)$ because $\logl(\Theta, \mb)$ is the sum of $\lambda(u,v)$.
 
 Now, we prove the concavity of $\lambda(u,v)$. Note that
-\begin{equation*}
+\begin{equation}
 \lambda(u,v) = \log\big[f(u)-f(v)\big]=\log\big[\int\mathds{1}_{[u,v]}(x)f'(x)dx\big],
-\end{equation*}
+\end{equation}
 where $\mathds{1}_{[u,v]}$ is an indicator function that equals 1 in the interval $[u,v]$, and 0 elsewhere.
-Furthermore, $\mathds{1}_{[u,v]}(x)$ is log-concave in $(u,v,x)$ and $f'(x) = \frac{e^x}{(1+e^x)^2}$ is log-concave in $x$.
-The convexities imply that $\mathds{1}_{(u,v)}(x)f'(x)$
+Furthermore, $\mathds{1}_{[u,v]}(x)$ is log-concave in $(u,v,x)$, and by Assumption~\ref{ass:link}, $f'(x)$ is log-concave in $x$.
+It follows that $\mathds{1}_{[u,v]}(x)f'(x)$
 is a log-concave in $(u,v,x)$. By Lemma~\ref{lem:lossconvexity}, we conclude that $\lambda(u,v)$ is concave in $(u,v)$ where $u>v$. 
@@
 
-We begin with various notion of tensor norms that are useful for the proofs of the main theorems.
+This section collects lemmas that are useful for the proofs of the main theorems.
 
 \begin{defn}[Atomic M-norm~\citep{ghadermarzy2019near}]
-Define $T_{\pm}=\{\tT\in\{\pm 1 \}^{d_1\times \cdots \times d_K}\colon \text{rank}(\tT)=1\}$. The atomic M-norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
+Define $T_{\pm}=\{\tT\in\{\pm 1 \}^{d_1\times \cdots \times d_K}\colon$ $\text{rank}(\tT)=1\}$. The atomic M-norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
 \begin{align}
@@
 The following lemma provides the bound on the spectral norm of random tensors. The result was firstly presented in~\cite{nguyen2015tensor}, and we adopt the version from~\cite{tomioka2014spectral}.
-\begin{lem}[\cite{tomioka2014spectral}]\label{lem:tensor}
+\begin{lem}[Spectral norm of random tensors~\citep{tomioka2014spectral}]\label{lem:tensor}
 Suppose that $\tS=\entry{s_{\omega}}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is an order-$K$ tensor whose entries are independent random variables that satisfy
@@
 
-\begin{defn}[Absolutely continuous]
-A measure $\lambda$ is absolutely continuous with respect to another measure $\mu$ if $\lambda(E)=0$ for every set with $\mu(E)=0$ and is denoted as $\lambda\ll\mu$.
+\begin{defn}[Absolute continuity]
+We use $\lambda\ll\mu$ to denote absolutely continuity of a measure $\lambda$ with respect to another measure $\mu$; that is, $\lambda(E)=0$ whenever $\mu(E)=0$. 
 \end{defn}
 
-\begin{lem}[Theorem 2.5 in~\cite{tsybakov2009introduction}]\label{lem:Tsybakov}
+\begin{lem}[Theorem 2.5 in~\cite{tsybakov2008introduction}]\label{lem:Tsybakov}
 Assume that a set $\tX$ contains element $\Theta_0, \Theta_1, \ldots,\Theta_M$ ($M\geq 2$) such that
@@
 
+\section{Conclusions}
+We have developed a low-rank tensor estimation method based on possibly incomplete, ordinal-valued observations.
+A sharp error bound is established, and we demonstrate the outperformance of our approach compared to other methods.
+The work unlocks several directions of future research.
+One interesting question would be the inference problem, i.e..
+to assess the uncertainty of the obtained estimates and the imputation.
+Other directions include the trade-off between (non)convex optimization and statistical/computational efficiency.
+While convex relaxations are popular approach for matrix/tensor problem, they are often slow in practice~\citep{ge2017optimization,chen2019non}.
+The interplay between computational efficiency and statistical accuracy in general tensor problems warrants future research.
+
+\section*{Acknowledgments}
+This research is supported in part by NSF grant DMS-1915978 and the University of Wisconsin-Madison, Office of the Vice Chancellor for Research and Graduate Education with funding from the Wisconsin Alumni Research Foundation.
+\newpage
+
 \appendix
-\section{Extension of Theorem~\ref{thm:rate} to unknown cut-off points}
+\setcounter{figure}{0}  
+\setcounter{table}{0}  
+\renewcommand{\thetable}{S\arabic{table}}
+\renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\figurename}{Supplemental Figure}
+\renewcommand{\tablename}{Supplemental Table}
+
+\section*{Appendix}
+\section{Extension of Theorem~\ref{thm:rate} to unknown cut-off points}\label{sec:extention}
 
@@
 
-We introduce several quantities that will used in our theory:
+We introduce several quantities that will be used in our theory:
 \begin{enumerate}
@@
 \[
-\text{MSE}(\hat \Theta, \trueT)\leq {c_1U_{\alpha,\beta,\Delta}\over L_{\alpha,\beta,\Delta}} r_{\max}^{(K-1)/2}  \sqrt{ \sum_k d_k \over \prod_k d_k},
+\mathrm{MSE}(\hat \Theta, \trueT)\leq \min\left\{ 4\alpha^2,\ {c_1U_{\alpha,\beta,\Delta}\over L_{\alpha,\beta,\Delta}} r_{\max}^{(K-1)/2}  \sqrt{ \sum_k d_k \over \prod_k d_k}\right\},
 \]
@@
 \[
-\text{MSE}(\hat\mb, \trueb) \leq  {c_2 C_{\alpha,\beta,\Delta} \over D_{\alpha,\beta,\Delta}}{\max_{\ell}(n_\ell+n_{\ell+1})\over \min_{\ell}(n_{\ell}+n_{\ell+1})}  \sqrt{1\over L-1},
+\text{MSE}(\hat\mb, \trueb) \leq  \min\left\{ 4\beta^2,\ {c_2 C_{\alpha,\beta,\Delta} \over D_{\alpha,\beta,\Delta}}{\max_{\ell}(n_\ell+n_{\ell+1})\over \min_{\ell}(n_{\ell}+n_{\ell+1})}  \sqrt{1\over L-1}\right\},
 \]
@@
 \begin{proof}[Proof of Theorem~\ref{thm:ratejoint}]
-Similar to the proof of Theorem~\ref{thm:rate}, we suppress $\Omega$ in the subscript. Based on the definition of $(\hat \Theta, \hat \mb)$,
+The constant bounds $4\alpha^2$ and $4\beta^2$ follow trivially from the definition of the feasible sets.
+We now focus on proving the non-constant bounds.
+Similar to the proof of Theorem~\ref{thm:rate}, we suppress $\Omega$ in the subscript.
+Based on the definition of $(\hat \Theta, \hat \mb)$,
 we have the following inequalities:
@@
 Next we bound $\text{MSE}(\hat \mb,\trueb)$. It follows from the expression of $\flogl(\Theta,\mb)$ that
-\begin{align}\label{eq:property}
+\begin{align}\label{eq:propertyb}
 {\partial \flogl\over \partial b_\ell}&=\sum_{\omega\in\Omega}\left[\mathds{1}_{\{y_{\omega}=\ell\}}
@@
 
-We consider the profile log-likelihood $\flogl(\mb, \hat \Theta)$ as a function of $\mb\in\tB$.
-For notational convenience, we drop $\hat \Theta$ from $\flogl(\mb, \hat \Theta)$ and simply write $\flogl(\mb)$.
+We consider the profile log-likelihood $\flogl(\hat \Theta,\mb)$ as a function of $\mb\in\tB$.
+For notational convenience, we drop $\hat \Theta$ from $\flogl(\hat \Theta,\mb)$ and simply write $\flogl(\mb)$.
 By the second-order Taylor's expansion of $\flogl(\mb)$ around $\trueb$, we obtain
@@
 
-\section{Additional results on simulations}
-In the main paper, we compare our method ({\bf Ordinal-T}) with other four methods ({\bf Ordinal-M}, {\bf 1bit-category-T},{\bf 1bit-sign-T}, and {\bf Continuous-T}).
-The methods are evaluated by their capabilities in predicting the most likely label for each entry, i.e.,\ $y_\omega^{\text{mode}}=\arg\max_\ell\mathbb{P}(y_\omega=\ell)$.
-Here we perform additional simulations to compare the methods by their performance in predicting the median label, $y_\omega^{\text{median}}=\min\{\ell\colon\mathbb{P}(y_\omega=\ell)\geq 0.5\}$.
-Under the latent variable model~\eqref{eq:latent} and Assumption~\ref{ass:link}, the median label is the quantized $\Theta$ without noise; i.e.\ $y_\omega^{\text{median}}=\sum_\ell \mathds{1}_{\theta_\omega\in(b_{\ell-1},b_\ell]}$.
-
-\begin{figure}[H]
-  \begin{center}
-    \includegraphics[width=.6\textwidth]{panel3.pdf}
-  \end{center}
-  \caption{Performance comparison in predicting the median label.
-(a,c) Prediction errors versus sample complexity $\rho=|\Omega|/d^k$ when $L=5$.
-(b,d) Prediction errors versus the number of ordinal levels $L$, when $\rho=0.8.$}
-  \label{fig:compare2}
-\end{figure}
-
-Figure~\ref{fig:compare2} compares the prediction accuracy under the setting $\alpha = 10, d = 20$ and $r = 5$.
-We see that our method outperforms the others in both MAD and MCR.
-The methods with probability model ({\bf Ordinal-T, Ordinal-M, 1bit-category-T}) have the tendency to be improved in MAD and be worsened in MCR when  $y_\omega^{\text{median}}$ is used (Figure~\ref{fig:compare2}) compared to the results when $y_\omega^{\text{mode}}$ is used (Figure~\ref{fig:compare}).
-This tendency is from the property that the estimator $\hat y^{(\text{Median})}_\omega$ minimizes $\mathbb{E}_{\hat{\theta}_\omega,\hat{\mb}}|y_\omega-y|$
-and the estimator $\hat y^{(\text{Mode})}_\omega $ minimizes $\mathbb{E}_{\hat{\theta}_\omega,\hat{\mb}}\mathds{1}_{\{y_\omega=y\}}$.
-By the similar reasons in Section~\ref{sec:compare} such as making use of tensor format of the datasets, accepting multi-level observation and preserving the ordinal structure, our model is benefited.
-
-\section{Additional results on HCP analysis}
+\section{Additional results of HCP analysis}\label{sec:additionalHCP}
 \subsection{Clustering based on Tucker representation}\label{sec:clustering}
@@
 Then, the mode-$k$ clustering can be performed as follows.
-First, we take $\left(\hat \mM_K\otimes\cdots\otimes\hat \mM_1\right)$ as principal axes and $\hat \mM_k\hat \tC_{(k)}$ as principal components.
-Then, we perform $K$-means clustering method to the $d_k$ rows of the principal components $\hat \mM_k\hat \tC_{(k)}$.
-
-\subsection{Clustering results on HCP}
-We perform clustering analysis to the 68 brain nodes using the the produce described in Section~\ref{sec:clustering}.
+First, we take columns in $\left(\hat \mM_K\otimes\cdots\otimes\hat \mM_1\right)$ as principal axes and rows in $\hat \mM_k\hat \tC_{(k)}$ as principal components.
+Then, we perform $K$-means clustering method to the $d_k$ rows of the matrix $\hat \mM_k\hat \tC_{(k)}$.
+
+\subsection{Clustering results of HCP analysis}
+We perform a clustering analysis on the 68 brain nodes using the produce described above.
 Recall that our ordinal tensor method outputs the estimated parameter tensor $\hat\Theta\in\mathbb{R}^{68\times 68\times136}$ with rank $(23,23,8)$.
-The number of clusters in the $K$-means is chosen based on the elbow method (see figure~\ref{figure:elbow}). Table~\ref{table:clustering} shows the eight clusters of brain nodes.
+We apply $K$-means to the mode-1 principal component matrix of size $68\times 184$ ($184=23\times8$).
+The elbow method suggests eight clusters among the 68 nodes (see Figure~\ref{figure:elbow}).
+
+Table~\ref{table:clustering} shows the eight clusters and their compositions.
+We find that most brain nodes fall into cluster I and the cluster II, which can be represented as the left and right hemispheres, respectively (Figure~\ref{figure:brain image}).
+Clusters III-VIII captures brain regions with same annotations in the dataset.
+For example, clusters III/IV represents the supramarginal gyrus region on the left/right hemisphere, respectively.
+This region is known to be involved in visual word recognition and reading~\citep{stoeckel2009supramarginal}.
+The results demonstrate that our clustering successfully groups similar brain regions together without knowledge of external annotation.
+
 \begin{figure}[H]
-  \label{figure:elbow}
 \begin{center}
-    \includegraphics[width=0.49\textwidth]{elbowm.pdf}
+ \includegraphics[width=0.4\textwidth]{elbowm.pdf}
 \end{center}
-  \caption{The elbow plot for the number of clusters}
+ \caption{Elbow plot for determining the number of clusters in $K$-means.}\label{figure:elbow}
 \end{figure}
 
-We find that most brain nodes fall into cluster I and the cluster II, which can be represented as the left and right hemispheres, respectively (Figure~\ref{figure:brain image}a).
-From cluster III to the cluster VIII, entries within each cluster share the same name encoded in the data.
-For example, cluster III and cluster IV represent the {\it Supramarginal gyrus} region on the left brain and right brain.
-These regions are known to be in charge of interpreting tactile sensory data and involved in perception of space and limbs location~\cite{carlson2012physiology,reed1994nature}.
-The results demonstrate that our clustering successfully groups similar brain nodes together without knowledge of external annotation.
+\begin{figure}[H]
+  \centering
+ \includegraphics[width = 0.45\textwidth]{braincluster.pdf}
+ \includegraphics[width = 0.45\textwidth]{braincluster3.pdf}
+  \caption{Top two clusters of brain nodes.
+(a) Nodes in yellow are from cluster I (left hemisphere), and nodes in green are from cluster II (right hemisphere).
+(b) Predicted connectivity within cluster I and cluster II.
+Edges are colored based on predicted strength level averaged across individuals.}  \label{figure:brain image}
+\end{figure}
 
 \begin{table}[H]
-  \label{table:clustering}
  \center
@@
 \caption{Clustering result of brain nodes.
-The first alphabet in the node name indicates the left brain (L) and the right brain (R) and the values in the parentheses are the number of nodes that have the same name in each cluster.
-}
+The first alphabet in the node name indicates the left (L) or right (R) hemisphere.
+The number in the parentheses indicates the node count in each cluster.
+}  \label{table:clustering}
 \end{table}
 
-\begin{figure}[H]
-  \centering
- \includegraphics[width = 0.45\textwidth]{braincluster.pdf}
- \includegraphics[width = 0.45\textwidth]{braincluster2.pdf}
-  \label{figure:brain image}
-  \caption{In the left brain image, yellow nodes are from  cluster I and Green nodes are from cluster II.
-Nodes from each cluster from III to VIII have almost same locations.
-The right brain image shows the extent of connectivity among nodes in each cluster I and cluster II.
-We take average of connections among nodes over 136 individuals.
-Yellow color means strong connectivity and green color weak connectivity.
-}
-\end{figure}
-
-\bibliographystyle{plainnat}
 \bibliography{tensor_wang}
-
+\bibliographystyle{apalike}
 \end{document}

### 2020-02-13 ICML/arxiv/old_file/ordinalTv3.tex
update arxiv

--- 2020-02-07 ICML/submit/ordinalTv3.tex
@@
 In practice, the cut-off points $\mb$ are often unknown, so we choose to maximize $\logl$ jointly over $(\Theta,\mb)\in\tP\times \tB$.
-The objective $\logl$ is concave in $(\Theta,\mb)$ whenever $f'$ is log-concave (See the Supplement).
+The objective $\logl$ is concave in $(\Theta,\mb)$ whenever $f'$ is log-concave (see the Supplement).
 However, the feasible set $\tP$ is non-convex, which makes the optimization~\eqref{eq:objective} a non-convex problem.
@@
 \centering
-\vspace{-.4cm}
 \includegraphics[width=8cm]{algorithm.pdf}\label{fig:stability}
-\vspace{-.3cm}
+\vspace{-.2cm}
 \caption{Trajectory of objective function with various $d$ and $r$.}
-\vspace{-.6cm}
+\vspace{-.5cm}
 \end{figure}
@@
 
-\vspace{-.2cm}
+\vspace{-.1cm}
 \section{Data Applications}
@@
 \vspace{-.2cm}
-\section{Conclusion}
+\section{Conclusions}
 \vspace{-.1cm}
@@
 A sharp error bound is established, and we demonstrate the outperformance of our approach compared to other methods.
-The work unlocked several directions of future research.
+The work unlocks several directions of future research.
 One interesting question would be the inference problem, i.e..

### 2020-02-13 code/functions.R
update

@@ -169,6 +169,7 @@ corecomb = function(A_1,A_2,A_3,C,ttnsr,omega,alpha=TRUE,type="ordinal"){
 
 ### main function ##
 fit_ordinal = function(ttnsr,C,A_1,A_2,A_3,omega=TRUE,alpha = TRUE){
+    
   if(is.logical(alpha)) alpha_minus=alpha_minus2=TRUE
   else{
         alpha_minus=alpha-epsilon
@@ -186,6 +187,9 @@ fit_ordinal = function(ttnsr,C,A_1,A_2,A_3,omega=TRUE,alpha = TRUE){
   cost=NULL
   omg = omega
   k=length(unique(as.factor(c(ttnsr))))-is.element(NA,ttnsr) ## for NA not being included in length
+  d=dim(ttnsr)
+  ttnsr=array(as.numeric(as.factor(ttnsr)),dim=d) ## code labels from 1 to K 
+  
     while ((error > 10^-4)&(iter<10) ) {
       iter = iter +1
       #update omega
@@ -278,7 +282,7 @@ theta_to_p=function(theta,omega){
         p[,i] = as.numeric(logistic(theta + omega[i]))
     }
     p =  cbind(p,rep(1,length(theta)))-cbind(rep(0,length(theta)),p)
-    p[p<epsilon]=epsilon
+    p[p<epsilon]=epsilon ## regularize
     return(p)
 }
 
@@ -383,6 +387,8 @@ fit_ordinal_matrix = function(ttnsr,A_1,A_2,omega=TRUE,alpha = TRUE){
     cost=NULL
     omg = omega
     k=length(unique(as.factor(c(ttnsr))))-is.element(NA,ttnsr) 
+    d=dim(ttnsr)
+    ttnsr=array(as.numeric(as.factor(ttnsr)),dim=d) ## code labels from 1 to K 
       
     while ((error > 10^-4)&(iter<10) ) {
         iter = iter +1

### 2020-02-13 ICML/arxiv/ordinalT_arxiv_v2.tex
upload

@@
 By taking $\varepsilon=1/10$ and $\gamma=1/11$, we conclude from~\eqref{eq:final} that
-\begin{equation*}\label{eq:prob}
+\begin{equation}\label{eq:prob}
 \inf_{\hat \Theta}\sup_{\trueT\in \tP}\mathbb{P}\left(\text{Loss}(\hat \Theta, \trueT)\geq c\min\left \{ \alpha^2,  {C^{-1}R_{\max}d_{\max}\over d_{\text{total}}}\right \}\right)\geq {1\over 2}\left({4\over 5}- \sqrt{1.6\over R_{\max}d_{\max}\log2} \right)\geq {1\over 8},
-\end{equation*}
+\end{equation}
 where $c = {1 \over 88^2}$ and the last inequality comes from the condition for $d_{\text{max}}$.
@@
 We find that most brain nodes fall into cluster I and the cluster II, which can be represented as the left and right hemispheres, respectively (Figure~\ref{figure:brain image}).
-Clusters III-VIII captures brain regions with same annotations in the dataset.
+Clusters III-VIII captures brain regions with same annotations.
 For example, clusters III/IV represents the supramarginal gyrus region on the left/right hemisphere, respectively.
@@
 (a) Nodes in yellow are from cluster I (left hemisphere), and nodes in green are from cluster II (right hemisphere).
-(b) Predicted connectivity within cluster I and cluster II.
+(b) Predicted connectivity within cluster I/II.
 Edges are colored based on predicted strength level averaged across individuals.}  \label{figure:brain image}

### 2020-02-13 ICML/arxiv/ordinalT_arxiv_v2.tex
final_version_minus_appendix_a

@@
 We also compare the goodness-of-fit of various tensor methods on the HPC data.
-We perform 5-fold cross-validation while preserving the label proportions in train/test sets.
+We perform 5-fold cross-validation while preserving the same label proportions in train/test sets.
 Table~\ref{table:CV} summarizes the prediction error averaged over 10 runs.
@@
 
-\section{Conclusions}
+\section{Conclusion}
 We have developed a low-rank tensor estimation method based on possibly incomplete, ordinal-valued observations.
@@
 \[
-\text{MSE}(\hat\mb, \trueb) \leq  \min\left\{ 4\beta^2,\ {c_2 C_{\alpha,\beta,\Delta} \over D_{\alpha,\beta,\Delta}}{\max_{\ell}(n_\ell+n_{\ell+1})\over \min_{\ell}(n_{\ell}+n_{\ell+1})}  \sqrt{1\over L-1}\right\},
+\mathrm{MSE}(\hat\mb, \trueb) \leq  \min\left\{ 4\beta^2,\ {c_2 C_{\alpha,\beta,\Delta} \over D_{\alpha,\beta,\Delta}}{\max_{\ell}(n_\ell+n_{\ell+1})\over \min_{\ell}(n_{\ell}+n_{\ell+1})}  \sqrt{1\over L-1}\right\},
 \]
@@
 \frac{\dot{f}(b_\ell-\theta_\omega)}{g_{\ell+1}(\theta_\omega)}\right],\notag\\
-{\partial^2 \flogl\over \partial b_\ell^2}&=\sum_{\omega\in\Omega}\left[\mathds{1}_{\{y_\omega=\ell\}}\frac{\ddot{f}(b_\ell-\theta_\omega)g_\ell(\theta_\omega)-\dot{f}(b_\ell-\theta_\omega)^2}{g^2_\ell(\theta_\omega)}-\mathds{1}_{\{y_\omega=\ell+1\}}\frac{\ddot{f}(b_\ell-\theta_\omega)g_{\ell+1}(\theta_\omega)+\dot{f}(b_\ell-\theta_\omega)^2}{g^2_{\ell+1}(\theta_\omega)}\right],\notag\\
+{\partial^2 \flogl\over \partial b^2_\ell}&=\sum_{\omega\in\Omega}\left[\mathds{1}_{\{y_\omega=\ell\}}\frac{\ddot{f}(b_\ell-\theta_\omega)g_\ell(\theta_\omega)-\dot{f}^2(b_\ell-\theta_\omega)}{g^2_\ell(\theta_\omega)}-\mathds{1}_{\{y_\omega=\ell+1\}}\frac{\ddot{f}(b_\ell-\theta_\omega)g_{\ell+1}(\theta_\omega)+\dot{f}^2(b_\ell-\theta_\omega)}{g^2_{\ell+1}(\theta_\omega)}\right],\notag\\
 &\quad  \text{ for all } \ell \in [L-1],\notag \\

### 2020-02-15 ICML/arxiv/ordinalT_arxiv_v3.tex
Add files via upload

--- 2020-02-13 ICML/arxiv/ordinalT_arxiv_v2.tex
@@
 The first problem is tensor denoising which aims to recover a signal tensor from its noisy entries~\citep{hong2018generalized,zeng2019multiway}.
-A second problem is tensor completion which examines the minimum number of entries needed for a consistent recovery~\citep{ghadermarzy2018learning,ghadermarzy2019near}.
+The second problem is tensor completion which examines the minimum number of entries needed for a consistent recovery~\citep{ghadermarzy2018learning,ghadermarzy2019near}.
 Low-rankness is often imposed to the signal tensor, thereby efficiently reducing the intrinsic dimension in both problems.
@@
 \begin{align}\label{eq:lower}
-\inf_{\hat \Theta }\sup_{\trueT\in\tP} \mathbb{P}\Big\{& \text{MSE}(\hat \Theta, \trueT)  \geq c \min\left( \alpha^2, \ { Cr_{\max}d_{\max} \over \prod_k d_k} \right) \Big\} \geq {1\over 8},
+\inf_{\hat \Theta }\sup_{\trueT\in\tP} \mathbb{P}\Big\{&\mathrm{MSE}(\hat \Theta, \trueT)  \geq c \min\left( \alpha^2, \ { Cr_{\max}d_{\max} \over \prod_k d_k} \right) \Big\} \geq {1\over 8},
 \end{align}
@@
 \[
-\text{MSE}(\hat \Theta,\Theta)={1\over \prod_k d_k}\FnormSize{}{\hat \Theta-\Theta},\quad \text{MSE}(\hat \mb,\mb)={1\over L-1}\FnormSize{}{\hat \mb-\mb}.
-\]
-
+\text{MSE}\left((\hat \Theta,\hat \mb),(\Theta,\mb)\right)={1\over \prod_k d_k+L-1}\FnormSize{}{(\hat \Theta,\hat \mb)-(\Theta,\mb)}^2
+\]
 We introduce several quantities that will be used in our theory:
@@
 \end{equation}
-\item We define two additional constants:
+\item We define three additional constants:
 \begin{align}\label{eq:CD}
@@
 \notag \\
-&\hspace{1.5in} \left. \frac{\ddot{f}(b_\ell-\theta)g_{\ell+1}(\theta)-\dot{f}(b_\ell-\theta)\dot{g}_{\ell+1}(\theta)}{g^2_{\ell+1}(\theta)}\right\}.
+&\hspace{1.5in} \left. \frac{\ddot{f}(b_\ell-\theta)g_{\ell+1}(\theta)-\dot{f}(b_\ell-\theta)\dot{g}_{\ell+1}(\theta)}{g^2_{\ell+1}(\theta)}\right\}\\
+\tilde{D}_{\alpha,\beta,\Delta} &=\max_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}} \max_{\ell\in[L-1]}\max\left\{  {\partial \over \partial \theta}\left({\dot{f}(b_{\ell}-\theta)\over g_\ell(\theta)}\right),\ -{\partial \over \partial \theta}\left({\dot{f}(b_{\ell}-\theta)\over g_{\ell+1}(\theta)}\right)\right\}\notag.
 \end{align}
 \end{enumerate}
-
 We make the following assumptions about the link function.
@@
 The condition $\Delta=\min_{\ell}(b_{\ell}-b_{\ell-1})>0$ on the feasible set $\tB$ guarantees the strict positiveness of $g_{\ell}(\theta)=f(b_{\ell}-\theta)-f(b_{\ell-1}-\theta)$.
-Therefore, the denominators in the above quantities $U_{\alpha,\beta,\Delta}, L_{\alpha,\beta,\Delta}, C_{\alpha,\beta,\Delta}, D_{\alpha,\beta,\Delta}$ are well-defined.
+Therefore, the denominators in the above quantities $U_{\alpha,\beta,\Delta}, L_{\alpha,\beta,\Delta}, C_{\alpha,\beta,\Delta}, D_{\alpha,\beta,\Delta},\tilde{D}_{\alpha,\beta,\Delta}$ are well-defined.
 Furthermore, by Theorem~\ref{thm:convexity}, $g_\ell(\theta)$ is strictly log-concave in $\theta$ for all $\mb\in\tB$.
-Based on Assumption~\ref{ass:joint} and closeness of the feasible set, we have $U_{\alpha,\beta,\Delta}>0$, $L_{\alpha,\beta,\Delta}>0$, $C_{\alpha,\beta,\Delta}>0$, $D_{\alpha,\beta,\Delta}>0$.
+Based on Assumption~\ref{ass:joint} and closeness of the feasible set, we have $U_{\alpha,\beta,\Delta}>0$, $L_{\alpha,\beta,\Delta}>0$, $C_{\alpha,\beta,\Delta}>0$, $D_{\alpha,\beta,\Delta}>0$, $\tilde{D}_{\alpha,\beta,\Delta}>0$.
 \end{rmk}
@@
 Then with very high probability, the estimator in~\eqref{eq:joint} satisfies
-\[
-\mathrm{MSE}(\hat \Theta, \trueT)\leq \min\left\{ 4\alpha^2,\ {c_1U_{\alpha,\beta,\Delta}\over L_{\alpha,\beta,\Delta}} r_{\max}^{(K-1)/2}  \sqrt{ \sum_k d_k \over \prod_k d_k}\right\},
-\]
-and
-\[
-\mathrm{MSE}(\hat\mb, \trueb) \leq  \min\left\{ 4\beta^2,\ {c_2 C_{\alpha,\beta,\Delta} \over D_{\alpha,\beta,\Delta}}{\max_{\ell}(n_\ell+n_{\ell+1})\over \min_{\ell}(n_{\ell}+n_{\ell+1})}  \sqrt{1\over L-1}\right\},
-\]
-where $c_1,c_2>0$ are two constants independent of the tensor dimension and rank, and $U_{\alpha,\beta,\Delta}$, $L_{\alpha,\beta,\Delta}$, $C_{\alpha,\beta,\Delta}$, $D_{\alpha,\beta,\Delta}>0$ are constants defined in~\eqref{eq:UL} and~\eqref{eq:CD}.
+\begin{align}
+\mathrm{MSE}\left((\hat \Theta,\hat \mb),(\trueT,\trueb)\right) & \leq {c_1U_{\alpha,\beta,\Delta}^2 \over L_{\alpha,\beta,\Delta}^2} r_{\max}^{(K-1)} { \sum_k d_k \over \prod_k d_k}+{C_2 \over D_{\alpha,\beta,\Delta}^2}{(\sum_k d_k)^2\over \min_{\ell}(n_{\ell}+n_{\ell+1})^2}.
+\end{align}
+Briefly,
+\begin{align}
+\mathrm{MSE}\left((\hat \Theta,\hat \mb),(\trueT,\trueb)\right) &= \tO\left({\sum_k d_k\over\prod_k d_k}\right)+\tO\left({(\sum_k d_k)^2\over \min_\ell(n_\ell+n_{\ell+1})^2}\right)\\& = \tO\left({(\sum_k d_k)^2\over \min\left\{\sum_k d_k\prod_k d_k ,\min_\ell(n_\ell+n_{\ell+1})^2\right\}}\right).
+\end{align}
+where $c_1>0$ is a constant independent of the tensor dimension and rank, and $U_{\alpha,\beta,\Delta}$, $L_{\alpha,\beta,\Delta}$,  $D_{\alpha,\beta,\Delta}>0$ are constants defined in~\eqref{eq:UL} and~\eqref{eq:CD} and $C_2>0$ is a constant that depends on $ C_{\alpha,\beta,\Delta},\tilde{D}_{\alpha,\beta,\Delta},L$.
 \end{thm}
 
-\begin{rmk} It may seem counter-intuitive that the MSE of $\hat \Theta$ shows no dependence of $\hat \mb$ (and vice versa).
-In fact, from the proof of Theorem~\ref{thm:ratejoint}, the uniform constants $U_{\alpha,\beta,\Delta}, L_{\alpha,\beta,\Delta}$ can be replaced by
-\[
-U_{\alpha,\hat \mb}=\max_{|\theta|\leq \alpha}\max_{\ell\in[L-1]}{|\dot{g}_\ell(\theta)|\over g_\ell(\theta)},\quad \text{and}\quad L_{\alpha,\hat \mb}=\min_{|\theta|\leq \alpha}\min_{\ell\in[L-1]} {\dot{g}^2_\ell(\theta)-\ddot{g}_\ell(\theta)g_\ell(\theta)\over g^2_\ell(\theta)}.
-\]
-Therefore, the accuracy of $\hat \Theta$ depends on the accuracy of $\hat \mb$ implicitly through $U_{\alpha, \hat \mb}$ and $ L_{\alpha,\hat \mb}$.
-
+\begin{rmk}
+The total MSE has two components where $\tO\left({\sum_k d_k\over\prod_k d_k}\right)$ is from the error in $\hat\Theta$ and $\tO\left({(\sum_k d_k)^2\over \min_\ell(n_\ell+n_{\ell+1})^2}\right)$ is from the error in $\hat \mb.$ When $\min_\ell(n_\ell+n_{\ell+1})\gg \sum_k d_k\prod_k d_k$, the error in $\hat\Theta$ dominates the total MSE.
+When $\min_\ell(n_\ell+n_{\ell+1})\ll \sum_k d_k\prod_k d_k $, the error in $\hat\mb$ dominates the total MSE.
 \end{rmk}
 
+\begin{rmk}
+  The constant bounds with $\alpha$ and $\beta$ can be obtained trivially from the definition of the feasible sets. For notational convenience, we ignore this constant bound.
+\end{rmk}
+
 \begin{proof}[Proof of Theorem~\ref{thm:ratejoint}]
-The constant bounds $4\alpha^2$ and $4\beta^2$ follow trivially from the definition of the feasible sets.
-We now focus on proving the non-constant bounds.
-Similar to the proof of Theorem~\ref{thm:rate}, we suppress $\Omega$ in the subscript.
-Based on the definition of $(\hat \Theta, \hat \mb)$,
+Similar to the proof of Theorem~\ref{thm:rate}, we suppress $\Omega$ in the subscript. Based on the definition of $(\hat \Theta, \hat \mb)$,
 we have the following inequalities:
@@
 \end{equation}
-
 Following the same argument as in Theorem~\ref{thm:rate} and the first inequality in~\eqref{eq:level}, we obtain that
 \begin{equation}\label{eq:ratenew}
-\text{MSE}(\hat \Theta, \trueT)\leq {CU_{\alpha,\beta,\Delta}\over L_{\alpha,\beta,\Delta}} r_{\max}^{(K-1)/2}  \sqrt{ \sum_k d_k \over \prod_k d_k},
+\FnormSize{}{\hat \Theta-\trueT}^2\leq {c_1^2U_{\alpha,\beta,\Delta}^2\over L_{\alpha,\beta,\Delta}^2} r_{\max}^{(K-1)}  \sum_k d_k ,
 \end{equation}
 where $U_{\alpha,\beta,\Delta}, L_{\alpha,\beta,\Delta}>0$ are two universal constants independent of the tensor dimension and rank.
-Next we bound $\text{MSE}(\hat \mb,\trueb)$. It follows from the expression of $\flogl(\Theta,\mb)$ that
+Next we bound $\FnormSize{}{\hat \mb-\trueb}^2$ given $\hat\Theta$. It follows from the expression of $\flogl(\Theta,\mb)$ that
 \begin{align}\label{eq:propertyb}
@@
 \end{align}
-
 Therefore, all entries in $\fploglb$ are upper bounded by $\left\{C_{\alpha,\beta,\Delta} \max_\ell (n_\ell+n_{\ell+1})\right\}>0$, and $\fpploglb$ is a tridiagonal matrix.
@@
 (\trueb-\hat\mb)^T\fploglb(\trueb)\leq \FnormSize{}{\trueb-\hat\mb}\FnormSize{}{\fploglb(\trueb)}
-\leq \FnormSize{}{\trueb-\hat\mb}\sqrt{L-1}C_{\alpha,\beta,\Delta}\max_{\ell\in[L-1]}(n_\ell+n_{\ell+1}),
-\end{equation}
-where the last inequality is followed from
-\[
-\left|{\partial \tL_\tY\over \partial b_\ell}\Big|_{\mb=\trueb}\right| \leq C_{\alpha,\beta,\Delta}\max_{\ell\in[L-1]}(n_{\ell}+n_{\ell+1}), \quad \textrm{ for all } \; \ell\in [L-1].
-\]
-
+\leq \FnormSize{}{\trueb-\hat\mb}C\sqrt{L-1}\sum_k d_k\sqrt{\prod_k d_k},
+\end{equation}
+where the last inequality is followed from Lemma \ref{lem:sharpergd} stated as,
+\[
+\left|{\partial \flogl\over \partial b_\ell}\bigg|_{(\hat \Theta,\trueb)}\right|\leq C\sum_k d_k\sqrt{\prod_k d_k},\quad \text{ for all } \ell \in [L-1] \text{ with high probability}.
+\]
 We next bound the quadratic term in \eqref{eq:taylorb}. Note that
@@
 &\quad\quad +\sum_{\ell\in[L-2]/\{1\}}\left( {\partial^2\tL_{\tY}\over \partial b^2_\ell}+{\partial^2\tL_{\tY}\over \partial b_\ell \partial b_{\ell-1}}+{\partial^2\tL_{\tY}\over \partial b_{\ell+1} \partial b_{\ell}} \right)\Big|_{\mb=\check\mb} (\hat b_\ell-b^{\text{true}}_{\ell})^2     \\
-&\leq - \tilde{D}_{\alpha,\beta,\Delta} \sum_{\ell \in [L-1]}(\hat b_\ell-b_{\text{true},\ell})^2 \nonumber \\
-&=-\tilde{D}_{\alpha,\beta,\Delta}\FnormSize{}{\hat\mb-\trueb}^2,
+&\leq - D_{\alpha,\beta,\Delta}' \sum_{\ell \in [L-1]}(\hat b_\ell-b_{\text{true},\ell})^2 \nonumber \\
+&=-D_{\alpha,\beta,\Delta}'\FnormSize{}{\hat\mb-\trueb}^2,
 \end{align}
@@
 \begin{align}
-\tilde{D}_{\alpha,\beta,\Delta}&=\min_{\substack{|\theta|\leq \alpha,\\ \mb\in\tB}}\min_{\ell\in[L-1]} -\left( {\partial^2\tL_{\tY}\over \partial b^2_\ell}+{\partial^2\tL_{\tY}\over \partial b_\ell \partial b_{\ell-1}}+{\partial^2\tL_{\tY}\over \partial b_{\ell+1} \partial b_{\ell}} \right)\\
+D_{\alpha,\beta,\Delta}'&=\min_{\substack{|\theta|\leq \alpha,\\ \mb\in\tB}}\min_{\ell\in[L-1]} -\left( {\partial^2\tL_{\tY}\over \partial b^2_\ell}+{\partial^2\tL_{\tY}\over \partial b_\ell \partial b_{\ell-1}}+{\partial^2\tL_{\tY}\over \partial b_{\ell+1} \partial b_{\ell}} \right)\\
 &=\min_{\substack{|\theta|\leq \alpha,\\ \mb\in\tB}} \min_{\ell\in[L-1]}\left\{\sum_{\omega\in\Omega}-\mathds{1}_{\{y_\omega=\ell\}}\left( {\ddot{f}(b_{\ell}-\theta_\omega) g_\ell(\theta_\omega)- \dot{f}(b_{\ell}-\theta_\omega)\dot{g}_\ell(\theta_\omega) \over g^2_\ell(\theta_\omega) }\right)\right.
@@
 \end{align}
-
 Combining inequalities~\eqref{eq:taylorb}, \eqref{eq:linearb} and~\eqref{eq:quadraticb} yields
 \[
-\flogl(\hat \mb)\leq \flogl(\trueb)+C_{\alpha,\beta,\Delta}\sqrt{L-1}\max_{\ell}(n_\ell+n_{\ell+1}) \FnormSize{}{\hat\mb-\trueb} -\frac{D_{\alpha,\beta,\Delta}}{2}\min_{\ell}(n_\ell+n_{\ell+1}) \FnormSize{}{\hat\mb-\trueb}^2 .
+\flogl(\hat \mb)\leq \flogl(\trueb)+C\sqrt{L-1} \sum_k d_k\sqrt{\prod_k d_k}) \FnormSize{}{\hat\mb-\trueb} -\frac{D_{\alpha,\beta,\Delta}}{2}\min_{\ell}(n_\ell+n_{\ell+1}) \FnormSize{}{\hat\mb-\trueb}^2 .
 \]
@@
 \[
-C_{\alpha,\beta,\Delta}\sqrt{L-1}\max_{\ell}(n_\ell+n_{\ell+1}) \FnormSize{}{\hat\mb-\trueb} -\frac{D_{\alpha,\beta,\Delta}}{2}\min_{\ell}(n_\ell+n_{\ell+1}) \FnormSize{}{\hat\mb-\trueb}^2\geq 0
+C\sqrt{L-1} \sum_k d_k\sqrt{\prod_k d_k}) \FnormSize{}{\hat\mb-\trueb} -\frac{D_{\alpha,\beta,\Delta}}{2}\min_{\ell}(n_\ell+n_{\ell+1}) \FnormSize{}{\hat\mb-\trueb}^2\geq 0
 \]
@@
 \[
-\text{MSE}(\hat \mb,\trueb)={1\over L-1}\FnormSize{}{\hat\mb-\trueb}\leq {2C_{\alpha,\beta,\Delta} \over D_{\alpha,\beta,\Delta}}{\max_{\ell}(n_\ell+n_{\ell+1})\over \min_{\ell}(n_{\ell}+n_{\ell+1})}  \sqrt{1\over L-1},
-\]
-which completes the proof.
+\FnormSize{}{\hat\mb-\trueb}^2\leq {C_2 \over D_{\alpha,\beta,\Delta}^2}{(\sum_k d_k)^2\prod_k d_k\over \min_{\ell}(n_{\ell}+n_{\ell+1})^2},
+\]
+where $C_2(C_{\alpha,\beta,\Delta},\tilde{D}_{\alpha,\beta,\Delta},L)$ is a positive constant.
+The following completes the proof.
+\begin{align}
+\text{MSE}\left((\hat \Theta,\hat \mb),(\trueT,\trueb)\right)&\leq{1\over \prod_k d_k}\FnormSize{}{(\hat \Theta,\hat \mb)-(\trueT,\trueb)}^2\\
+&={1\over \prod_k d_k}\FnormSize{}{(\hat \Theta,\hat \mb)-(\hat \Theta,\trueb)+(\hat \Theta,\trueb)-(\trueT,\trueb)}^2\\
+&\leq {1\over\prod_kd_k}\FnormSize{}{(\hat \Theta- \trueT)}^2+ {1\over\prod_kd_k}\FnormSize{}{\hat \mb -\trueb}^2.
+\end{align}
+\end{proof}
+
+\begin{rmk}[MSE for $\hat\mb$]
+In the proof, we have shown that
+\[
+\FnormSize{}{\hat\mb-\trueb}\leq {\sqrt{C_2} \over D_{\alpha,\beta,\Delta}}{(\sum_k d_k)\sqrt{\prod_k d_k}\over \min_{\ell}(n_{\ell}+n_{\ell+1})}.
+\]
+This bound is sharper than the trivial bound $\FnormSize{}{\trueb-\hat\mb}\leq 2\beta\sqrt{L-1}$ which is from $|b_\ell^{\text{true}}-\hat b_\ell|\leq 2\beta.$ In particular, $\FnormSize{}{\hat\mb-\trueb}\rightarrow 0$ as $\min_\ell(n_\ell+n_{\ell+1})\asymp \tilde \tO(\sum_k d_k\sqrt{\prod_k d_k})\rightarrow \infty.$
+\end{rmk}
+
+\begin{lem}[CLT for i.d. Bernoulli]\label{lem:clt}
+Suppose $\{X_n\}$ is a series of independent Bernoulli trials with possibly different success probabilities $p_n$. Let $s_n = \sum_{m=1}^n p_m(1-p_m)$ and $Y_n = X_n-p_n$. Then,
+\[{1\over s_n}\sum_{m=1}^n Y_n \stackrel{\mathcal{D}}{\rightarrow} N(0,1), \quad \text{ as } n \rightarrow \infty,\]
+whenever $s_n = \sum_{m=1}^n p_m(1-p_m)\rightarrow \infty.$
+\end{lem}
+
+\begin{proof}
+Let us look for a Lyapunov condition for $\delta = 1$.
+First, we have
+\[
+\mathbb{E}[|Y_n|^3] = p_n(1-p_n)^3+ (1-p_n)p_n^3\leq p_n(1-p_n)[(1-p_n)^2+p_n^2]\leq p_n(1-p_n).
+\]
+Hence, summation of the above inequality shows,
+\[
+\sum_{m=1}^n\mathbb{E}[|Y_n|^3]\leq \sum_{m=1}^n p_m(1-p_m) = s_n^2.
+\]
+Thus, the Lyapunov condition is satisfied whenever,
+\[{1\over s_n^3}\sum_{m=1}^n \mathbb{E}[|Y_n|^3]\leq {s_n^2\over s_n^3}\rightarrow 0.\]
+or simply $s_n = \sum_{m=1}^n p_m(1-p_m)\rightarrow \infty.$ Central limit theorem on $Y_n$ completes the proof.
+
+\end{proof}
+
+\begin{lem}[Bound on Gradients]\label{lem:sharpergd}
+Consider the same set-up as in Theorem \ref{thm:ratejoint}. Then, with very high probability,
+\[
+\left|{\partial \flogl\over \partial b_\ell}\bigg|_{(\hat \Theta,\trueb)}\right|\leq \left(C_{\alpha,\beta,\Delta}+\tilde{D}_{\alpha,\beta,\Delta}\FnormSize{}{\trueT-\hat\Theta}^2\right)\sqrt{\prod_k d_k},\quad \text{ for all } \ell \in [L-1],
+\]
+In particular, there exists a constant $d_0\in \mathbb{N}_+$ such that for all $d_k\geq d_0$ for $k \in [K]$,
+\[
+\left|{\partial \flogl\over \partial b_\ell}\bigg|_{(\hat \Theta,\trueb)}\right|\leq C\sum_k d_k\sqrt{\prod_k d_k},\quad \text{ for all } \ell \in [L-1],
+\]
+where $C(C_{\alpha,\beta,\Delta},\tilde{D}_{\alpha,\beta,\Delta})>0$ is a constant.
+\end{lem}
+\begin{proof}
+We only prove the case for $\ell = 1$. Other cases can be proved similarly.
+Note that
+\begin{equation}\label{eq:sharpergd}
+  {\partial \flogl\over \partial b_1}\bigg|_{(\hat \Theta,\trueb)} = \KeepStyleUnderBrace{{\partial \flogl\over \partial b_1}\bigg|_{(\hat \Theta,\trueb)}-\mathbb{E}_\mathcal{Y}\left[{\partial \flogl\over \partial b_1}\bigg|_{(\hat \Theta,\trueb)}\right]}_{\coloneqq A}+\KeepStyleUnderBrace{\mathbb{E}_\mathcal{Y}\left[{\partial \flogl\over \partial b_1}\bigg|_{(\hat \Theta,\trueb)}\right]-\mathbb{E}_\mathcal{Y}\left[{\partial \flogl\over \partial b_1}\bigg|_{(\hat \trueT,\trueb)}\right] }_{\coloneqq B}.
+\end{equation}
+We have used the fact that the score function has mean zero, $\mathbb{E}_\mathcal{Y}\left[{\partial \flogl\over \partial b_1}\bigg|_{(\hat \trueT,\trueb)}\right]=0 $.
+Here all expectations are taken with respect to $\mathcal{Y}\sim \mathbb{P}(\trueT,\trueb).$\\
+We now bound the two deviation terms in \eqref{eq:sharpergd} separately. The term $A$ in \eqref{eq:sharpergd} is the stochastic deviation of log-likelihood to its expectation:
+\begin{align}
+A &= \sum_{\omega\in\Omega}\left\{[\mathds{1}_{y_\omega = 1}-g_1(\theta^{\text{true}}_\omega)]{\dot{f}(b_1-\hat\theta_\omega)\over g_1(\hat\theta_\omega)}-[\mathds{1}_{y_\omega = 2}-g_2(\theta^{\text{true}}_\omega)]{\dot{f}(b_1-\hat\theta_\omega)\over g_2(\hat\theta_\omega)}\right\}\\
+&\leq C_{\alpha,\beta,\Delta}\sum_{\omega\in\Omega}\KeepStyleUnderBrace{[\mathds{1}_{y_\omega = 1}+\mathds{1}_{y_\omega = 2}- g_1(\theta^{\text{true}}_\omega)-g_2(\theta^{\text{true}}_\omega)]}_{\coloneqq W_\omega}.
+\end{align}
+Note that $\{W_\omega\}$ are centered independent Bernoulli random variables with success probabilities $g_1(\theta^{\text{true}}_\omega)+g_2(\theta^{\text{true}}_\omega)$.
+By Lemma \ref{lem:clt}, we have
+\[
+\sum_{\omega\in\Omega}W_\omega \stackrel{\mathcal{D}}{\rightarrow} N\left(0, \sum_{\omega\in\Omega}\left(g_1(\theta^{\text{true}}_\omega)+g_2(\theta^{\text{true}}_\omega)\right)\left(1-g_1(\theta^{\text{true}}_\omega)+g_2(\theta^{\text{true}}_\omega)\right)\right).
+\]
+Hence, with the fact that $\sum_{\omega\in\Omega}\left(g_1(\theta^{\text{true}}_\omega)+g_2(\theta^{\text{true}}_\omega)\right)\left(1-g_1(\theta^{\text{true}}_\omega)+g_2(\theta^{\text{true}}_\omega)\right)\leq {1\over 4}\prod_k d_k$,
+\begin{equation}\label{eq:A}
+|A| \leq C_{\alpha,\beta,\Delta}\left|\sum_{\omega\in\Omega}W_\omega\right|\leq C_{\alpha,\beta,\Delta}\sqrt{\prod_k d_k}, \quad \text{ with high probability. }
+\end{equation}
+The second term $B$ in \eqref{eq:sharpergd} is the bias induced by the inaccuracy of $\hat\Theta$:
+\begin{align}
+|B| &= \left|\sum_{\omega\in\Omega} g_1(\theta^{\text{true}}_\omega)\left({\dot{f}(b_1-\hat\theta_\omega)\over g_1(\hat\theta_\omega)}-{\dot{f}(b_1-\theta^{\text{true}}_\omega)\over g_1(\theta^{\text{true}}_\omega)}\right)-\sum_{\omega\in\Omega} g_2(\theta^{\text{true}}_\omega)\left({\dot{f}(b_1-\hat\theta_\omega)\over g_2(\hat\theta_\omega)}-{\dot{f}(b_1-\theta^{\text{true}}_\omega)\over g_2(\theta^{\text{true}}_\omega)}\right)\right|\\
+&= \Bigg|\sum_{\omega\in\Omega} g_1(\theta^{\text{true}}_\omega)(\theta^{\text{true}}_\omega-\hat\theta_\omega)\left\{{\partial\over\partial}\left({\dot{f}(b_1-\theta)\over g_1(\theta)}\right)\bigg|_{\rho_\omega\hat\theta_\omega+(1-\rho_\omega)\theta^{\text{true}}_\omega}\right\}\\
+&\hspace{0.5in}-\sum_{\omega\in\Omega} g_2(\theta^{\text{true}}_\omega)(\theta^{\text{true}}_\omega-\hat\theta_\omega)\left\{{\partial\over\partial}\left({\dot{f}(b_1-\theta)\over g_2(\theta)}\right)\bigg|_{\rho_\omega'\hat\theta_\omega+(1-\rho_\omega')\theta^{\text{true}}_\omega}\right\}\Bigg|\\
+&\leq \tilde{D}_{\alpha,\beta,\Delta}\sum_{\omega\in\Omega}|g_1(\theta^{\text{true}}_\omega)+g_2(\theta^{\text{true}}_\omega)||\theta^{\text{true}}_\omega-\hat\theta_\omega|.
+\end{align}
+By Cauchy-Schwartz inequality with the fact that $g_1(\theta)+g_2(\theta)<1$,
+\begin{equation}\label{eq:B}
+|B| \leq \tilde{D}_{\alpha,\beta,\Delta}\sqrt{\prod_k d_k} \FnormSize{}{\trueT-\hat\Theta}.
+\end{equation}
+Plugging \eqref{eq:A} and \eqref{eq:B} back to \eqref{eq:sharpergd} yields that
+\[
+{\partial \flogl\over \partial b_1}\bigg|_{(\hat \Theta,\trueb)}\leq \left(C_{\alpha,\beta,\Delta}+\tilde{D}_{\alpha,\beta,\Delta}\FnormSize{}{\trueT-\hat\Theta}\right)\sqrt{\prod_k d_k}.
+\]
+holds with very high probability. The second inequality in the lemma comes from the fact that $\FnormSize{}{\trueT-\hat\Theta}\leq \tO(\sum_k d_k).$
 \end{proof}

### 2020-02-15 ICML/arxiv/ordinalT_arxiv_v3.tex
Add files via upload

@@
 Table~\ref{table:clustering} shows the eight clusters and their compositions.
-We find that most brain nodes fall into cluster I and the cluster II, which can be represented as the left and right hemispheres, respectively (Figure~\ref{figure:brain image}).
-Clusters III-VIII captures brain regions with same annotations.
+We find that most brain nodes fall into cluster I and the cluster II, which can be represented as the left and right hemispheres, respectively (Figure~\ref{figure:brain image} a,b).
+Clusters III-VIII capture brain regions with same annotations.
 For example, clusters III/IV represents the supramarginal gyrus region on the left/right hemisphere, respectively.
 This region is known to be involved in visual word recognition and reading~\citep{stoeckel2009supramarginal}.
+We increase the number of clusters to 11 and 13.
+When the number of clusters is 11, there are three main clusters which most of brain nodes fall into (Figure~\ref{figure:brain image} c,d).
+Two of them are represented as the left and right hemispheres like the previous cluster result.
+The other main cluster is newly captured and located in the bottom-middle part of the brain.
+Other small clusters are more or less the same as the small clusters in the result when the number of cluster is eight.
+When the number of clusters is 13, we find the four biggest clusters.
+We can characterize those clusters as top-left, bottom-left, top-right and bottom-right hemispheres (Figure~\ref{figure:brain image} e,f).
+Other clusters remain almost the same as the small clusters in the previous results, which share same annotations in each cluster.
 The results demonstrate that our clustering successfully groups similar brain regions together without knowledge of external annotation.
@@
   \centering
- \includegraphics[width = 0.45\textwidth]{braincluster.pdf}
- \includegraphics[width = 0.45\textwidth]{braincluster3.pdf}
-  \caption{Top two clusters of brain nodes.
-(a) Nodes in yellow are from cluster I (left hemisphere), and nodes in green are from cluster II (right hemisphere).
-(b) Predicted connectivity within cluster I/II.
+  \includegraphics[width = 0.45\textwidth]{clustnode8.pdf}
+  \includegraphics[width = 0.45\textwidth]{clustedge8.pdf}
+  \includegraphics[width = 0.45\textwidth]{clustnode11.pdf}
+  \includegraphics[width = 0.45\textwidth]{clustedge11.pdf}
+  \includegraphics[width = 0.45\textwidth]{clustnode13.pdf}
+  \includegraphics[width = 0.45\textwidth]{clustedge13.pdf}
+  \caption{Cluster image of brain nodes when the number of clusters is 8 (a,b), 11 (c,d) and 13 (e,f). (a,c,e)
+  Nodes from the same cluster are colored with the same color.
+(b,d,f) Predicted connectivity within clusters.
 Edges are colored based on predicted strength level averaged across individuals.}  \label{figure:brain image}

### 2020-02-16 ICML/arxiv/old_file/sharper_MSE.tex
upload v4

\documentclass[11pt]{article}

\usepackage{fancybox}
\usepackage{color}
\usepackage{url}
\usepackage[margin=1in]{geometry}
\newcommand{\theHalgorithm}{\arabic{algorithm}}
\usepackage{comment}
\usepackage{amsmath,amssymb,amsthm,bm,mathtools}
\usepackage{natbib}
\usepackage{dsfont,multirow,hyperref,setspace,enumerate}
\hypersetup{colorlinks,linkcolor={blue},citecolor={blue},urlcolor={red}}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{makecell}

\setlength\parindent{0pt}
\usepackage[parfill]{parskip}

\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{pro}{Property}
\newtheorem{assumption}{Assumption}
\newtheorem{cor}{Corollary}

\theoremstyle{definition}
\newtheorem{defn}{Definition}
\newtheorem{example}{Example}
\newtheorem{rmk}{Remark}

\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

\usepackage{dsfont}
\usepackage{wrapfig}
\usepackage{mathtools}
\mathtoolsset{showonlyrefs}

\newcommand*{\KeepStyleUnderBrace}[1]{%f
 \mathop{%
    \mathchoice
    {\underbrace{\displaystyle#1}}%
    {\underbrace{\textstyle#1}}%
    {\underbrace{\scriptstyle#1}}%
    {\underbrace{\scriptscriptstyle#1}}%
  }\limits
}
\begingroup
    \makeatletter
    \@for\theoremstyle:=definition,remark,plain\do{%
        \expandafter\g@addto@macro\csname th@\theoremstyle\endcsname{%
            \addtolength\thm@preskip\parskip
            }%
        }
\endgroup

\input macros.tex
\allowdisplaybreaks

\begin{document}
\begin{center}
{\large Joint estimation of $\hat \Theta$ and $\hat \mb$ in the tensor ordinal model}\\
\vspace{.1cm}
Miaoyan Wang, Feb 14, 2020. 
\end{center}
\begin{rmk}[Notation]
For notational convenience, we consider the special case $d_1=\cdots=d_K=d$. We focus on the asymptotic region $d\to \infty$ and treat all other quantities as constants, i.e., $\mr, L, U_{\alpha,\beta,\Delta}, L_{\alpha,\beta,\Delta},...=\tO(1)$. In particular, 
\begin{enumerate}
\item $\prod_k d_k \asymp d^K$, $\sum_k d_k\asymp d$. 
\item Denote $n_{\max}=\max_\ell n_{\ell}$ and $n_{\min}=\min_\ell(n_\ell+n_{\ell+1})$. Then, $n_{\max}\asymp d^K$ (important!) but $\tO(1)\leq n_{\min}\leq \tO(d^K)$. 
\end{enumerate}
Note: we use $C>$ to denote a positive constant whose value may change from line to line. 
\end{rmk}

\section{Results}
\subsection{Current bound}
Total MSE:
\begin{align}\label{eq:bound1}
\text{MSE}((\hat \Theta,\hat \mb), (\trueT,\trueb))&\stackrel{\text{def}}{=}{\prod_k d_k \mathrm{MSE}(\hat \Theta, \trueT)+(L-1)\mathrm{MSE}(\hat \Theta, \trueT) \over \prod_k d_k+L-1} \notag \\
&\leq {c_1\sum_k d_k + c_2  {n^2_{\max} \over n^2_{\min}} \over \prod_k d_k+L-1} \notag \\
&=\tO\left({\sum_k d_k\over \prod_k d_k}\right)+\tO\left({n^2_{\max} \over n^2_{\min}} {1\over \prod_k d_k}\right)
\end{align}

\begin{rmk}
Unfortunately, this total MSE bound does not converge to zero. In the worse case, the bound can be $\asymp d^K$;  for example, the bound diverges when $n_{\min}\asymp 1$ (and $n_{\max}\asymp d^K$). 
\end{rmk}
\begin{rmk}
The total MSE bound converges to zero only when 
\begin{equation}\label{eq:1}
{n_{\max}\over n_{\min}}\ll d^{K/2},\quad \text{or equivalently,}\quad n_{\min} \gg d^{K/2}.
\end{equation}
In other words, the current bound~\eqref{eq:bound1} tolerates only certain imbalanced classes for which $\tO(d^{K/2})\leq n_{\min}\leq \tO(d^K)$. 
\end{rmk}

\subsection{Sharper bound}
We will prove a sharper bound on the linear term
\begin{align}\label{eq:linearb}
(\trueb-\hat\mb)^T\fploglb(\trueb)&\leq \FnormSize{}{\trueb-\hat\mb}\FnormSize{}{\fploglb(\trueb)}\\
&\leq C\FnormSize{}{\trueb-\hat\mb} \sqrt{d^{K+2}},
\end{align}
where the last inequality is followed from a sharper bound on the gradient:
\begin{equation}\label{eq:bound}
\left|{\partial \tL_\tY\over \partial b_\ell}\Big|_{(\hat \Theta, \trueb)}\right|   \leq C \sqrt{d^{K+2}}, \quad \textrm{ for all } \; \ell\in [L-1].
\end{equation}
Suppose~\eqref{eq:bound} holds. Following the same line in the current proof (i.e.,\ Taylor expansion, quadratic bound, etc.), we have
\[
\FnormSize{}{\trueb-\hat \mb}\leq C{d^{(K+2)/2}  \over n_{\min}}.
\]
{\color{red} (Final results.) Therefore, the total MSE:
\begin{equation}\label{eq:boundnew}
\text{Total MSE}\leq \tO\left({1\over d^{K-1}}\right)+\tO\left(d^{K+2}\over n^2_{\min} d^K\right)=\tO\left({d^2\over \min\{d^{K+1}, n^2_{\min}\}}\right)
\end{equation}
which convergences to zero whenever
\begin{equation}\label{eq:requirement}
n_{\min} \gg d.
\end{equation}
}
In other words, the new bound~\eqref{eq:boundnew} tolerates highly imbalanced classes for which $\tO(\sqrt{d})\leq n_{\min}\leq \tO(d^K)$.

\begin{rmk}
The consistency condition~\eqref{eq:requirement} is more relaxed than~\eqref{eq:1} for all $K\geq 3$. Both bounds agree in the matrix case ($K=2$). 
\end{rmk}

\begin{rmk}
When $n_{\min}\gg d^{(K+1)/2}$, the error in $\hat \Theta$ dominates the total MSE. Then $n_{\min}\ll d^{(K+1)/2}$, the error in $\hat \mb$ dominates the total MSE
\end{rmk}

\section{Proofs}
Now we prove the inequality~\eqref{eq:bound}. 
\begin{lem}[Sharper Bound on Gradients]\label{lem:gradient}
Consider the same set-up as in Theorem A.1. Then, with very high probability,
\[
\left|{\partial \tL_\tY\over \partial b_\ell}\Big|_{(\hat \Theta, \trueb)}\right| \leq C_1d^{K/2}+C_2d^{K/2}\FnormSize{}{\trueT-\hat \Theta},\quad \textrm{ for all } \; \ell\in [L-1].
\]
where $C_1,C_2>0$ are two constants. In particular, there exists a constant $d_0\in\mathbb{N}_{+}$, such that for all $d\geq d_0$,
\begin{equation}
\left|{\partial \tL_\tY\over \partial b_\ell}\Big|_{(\hat \Theta, \trueb)}\right|   \leq Cd^{(K+2)/2},\quad \textrm{ for all } \; \ell\in [L-1].
\end{equation}
\end{lem}

\begin{cor}[MSE for $\hat \mb$] Under the same set-up as in Theorem A.1, we have
\begin{equation}\label{eq:boundb}
\FnormSize{}{\hat \mb-\trueb}\leq { C_1d^{K/2}+C_2d^{K/2}\FnormSize{}{\hat \Theta-\trueT} \over n_{\min}}\leq {Cd^{(K+2)/2}\over n_{\min}}.
\end{equation}
\end{cor}
\begin{rmk} The bound~\eqref{eq:bound} is sharper than the trivial bound $|b^{\text{true}}_\ell-\hat b_\ell|\leq 2\beta$. In particular, $\FnormSize{}{\hat \mb-\trueb}\to 0$ as $n_{\min}\asymp (d^{(K+2)/2})\to \infty$. 
\end{rmk}
\begin{proof}[Proof of Lemma~\ref{lem:gradient}]
We only prove the case for $\ell=1$. Other cases can be proved similarly.

Note that 
\begin{align}\label{eq:diff}
{\partial \tL_\tY\over \partial b_1}\Big|_{(\hat \Theta, \trueb)}
 &=\KeepStyleUnderBrace{ {\partial \tL_\tY\over \partial b_1}\Big|_{(\hat \Theta, \trueb)}-\mathbb{E}_{\tY}\left[{\partial \tL_\tY\over \partial b_1}\Big|_{(\hat \Theta, \trueb)}\right]}_{\text{:=A}}+\KeepStyleUnderBrace{\mathbb{E}_{\tY}\left[{\partial \tL_\tY\over \partial b_1}\Big|_{(\hat \Theta, \trueb)}\right]-\mathbb{E}_{\tY}\left[{\partial \tL_\tY\over \partial b_1}\Big|_{(\trueT, \trueb)}\right]}_{:=B},
\end{align}
where we have used the fact that the score function has mean zero, $\mathbb{E}_{\tY}\left[{\partial \tL_\tY\over \partial b_1}\Big|_{(\trueT, \trueb)}\right]=0$. Here all expectations are taken with respect to $\tY\sim \mathbb{P}(\trueT, \trueb)$.

We now bound the two deviation terms in~\eqref{eq:diff} separately.  The term $A$ in~\eqref{eq:diff} is the stochastic deviation of log-likelihood to its expectation:
\begin{align}
A=\sum_{\omega\in\Omega}\left\{\KeepStyleUnderBrace{\left[\mathds{1}_{\{y_{\omega}=1\}}-g_1(\theta^{\text{true}}_\omega)\right]
\frac{\dot{f}(b_1-\hat \theta_\omega)}{g_1(\hat \theta_\omega)}-\left[\mathds{1}_{\{y_{\omega}=2\}}-g_2(\theta^{\text{true}}_\omega)\right]
\frac{\dot{f}(b_1-\hat \theta_\omega)}{g_{2}(\hat \theta_\omega)}}_{:=W_\omega}\right\}.
\end{align}
Note that $\{W_\omega\}$ are i.i.d.\ random variables, and each $W_\omega$ has zero mean and bounded variance: 
\[
\text{Var}(W_\omega)\leq C\left(g_1(\theta_\omega^{\text{true}}) (1-g_1(\theta_\omega^{\text{true}}))+g_2(\theta_\omega^{\text{true}}) (1-g_2(\theta_\omega^{\text{true}}))\right)\leq {C/2}.
\]
By central limit theorem (verify..)
\[
\sum_{\omega\in\Omega}W_\omega \stackrel{\tD}{\rightarrow}N(0, {Cd^K}),\quad \text{as}\quad d^K\to\infty.
\]
Hence, with very high probability,
\begin{equation}\label{eq:A}
|A|=\left|\sum_{\omega\in\Omega}W_\omega \right|\leq Cd^{K/2} .
\end{equation}

The second term $B$ in~\eqref{eq:diff} is the bias induced by the inaccuracy of $\hat \Theta$:
\begin{align}
B&=\sum_{\omega \in \Omega}g_1(\theta^{\text{true}}_\omega) \left(  \frac{\dot{f}(b_1-\hat \theta_\omega)}{g_1(\hat \theta_\omega)}-\frac{\dot{f}(b_1-\theta^{\text{true}}_\omega)}{g_1(\theta^{\text{true}}_\omega)}\right)-\sum_{\omega \in \Omega}g_2(\theta^{\text{true}}_\omega) \left(  \frac{\dot{f}(b_2-\hat \theta_\omega)}{g_2(\hat \theta_\omega)}-\frac{\dot{f}(b_2-\theta^{\text{true}}_\omega)}{g_2(\theta^{\text{true}}_\omega)}\right) \notag\\
&\leq \sum_{\omega \in \Omega} g_1(\theta^{\text{true}}_\omega)  (\theta^{\text{true}}_\omega-\hat \theta_\omega )\left\{{\partial \over \partial \theta}\left({\dot{f}(b_1-\theta)\over g_1(\theta)}\right)\bigg|_{\rho \hat \theta_\omega+(1-\rho)\theta^{\text{true}}_\omega}\right\}\notag\\
&\quad - \sum_{\omega \in \Omega} g_2(\theta^{\text{true}}_\omega) (\theta^{\text{true}}_\omega-\hat \theta_\omega )  \left\{{\partial \over \partial \theta}\left({\dot{f}(b_2-\theta)\over g_2(\theta)}\right)\bigg|_{\rho' \hat \theta_\omega+(1-\rho')\theta^{\text{true}}_\omega}\right\} \notag\\
&\leq C\sum_{\omega\in\Omega} \left[g_1(\theta^{\text{true}}_\omega)- g_2(\theta^{\text{true}}_\omega)\right]\left(\theta^{\text{true}}_\omega-\hat \theta_\omega \right).
\end{align}
By Cauchy-Schwartz inequality,
\begin{equation}\label{eq:B}
|B|\leq Cd^{K/2}\FnormSize{}{\trueT-\hat \Theta}.
\end{equation}
Plugging \eqref{eq:A} and \eqref{eq:B} back to~\eqref{eq:diff} yields that
\[
\left|{\partial \tL_\tY\over \partial b_\ell}\Big|_{(\hat \Theta, \trueb)}\right|\leq C_1d^{K/2}+C_2d^{K/2}\FnormSize{}{\trueT-\hat \Theta}
\]
holds with very high probability. The second inequality in the conclusion comes from the fact that $\FnormSize{}{\trueT-\hat \Theta}\leq \tO(d)$ as $d\to \infty$. 
\end{proof}

\bibliography{tensor_wang}
\bibliographystyle{apalike}
\end{document}

### 2020-02-16 ICML/arxiv/ordinalT_arxiv_v4.tex
upload v4

--- 2020-02-15 ICML/arxiv/ordinalT_arxiv_v3.tex
@@
 }
+
 \begingroup
@@
 \begin{equation}\label{eq:rate}
-\mathrm{MSE}(\hat \Theta, \trueT) \leq \min\left( 4\alpha^2,\ {c_2  U^2_\alpha r_{\max}^{K-1}  \over  L^2_\alpha } {\sum_kd_k\over  \prod_k d_k} \right),
-\end{equation}
-where $c_1, c_2>0$ are two constants that depend only on $K$.\\
+\mathrm{MSE}(\hat \Theta, \trueT) \leq \min\left( 4\alpha^2,\ {c_1  U^2_\alpha r_{\max}^{K-1}  \over  L^2_\alpha } {\sum_kd_k\over  \prod_k d_k} \right),
+\end{equation}
+where $c_1>0$ is a constant that depends only on $K$.\\
 \end{thm}
@@
 \begin{equation}\label{eq:KLrate}
-\mathrm{KL}(\mathbb{P}_{\tY} || \hat{\mathbb{P}}_{\tY}) \leq  {c_2 U^2_\alpha r_{\max}^{K-1}  \over L^2_\alpha } {(4L-6)\dot{f}^2(0)  \over A_\alpha} {\sum_kd_k\over  \prod_k d_k},
-\end{equation}
-where $c_1, c_2>0$ are the same constants as in Theorem~\ref{thm:rate}.
+\mathrm{KL}(\mathbb{P}_{\tY} || \hat{\mathbb{P}}_{\tY}) \leq  {c_1 U^2_\alpha r_{\max}^{K-1}  \over L^2_\alpha } {(4L-6)\dot{f}^2(0)  \over A_\alpha} {\sum_kd_k\over  \prod_k d_k},
+\end{equation}
+where $c_1>0$ is the same constant as in Theorem~\ref{thm:rate}.
 \end{cor}
@@
 \section{Numerical Implementation}\label{sec:algorithm}
-
 We describe the algorithm to seek the optimizer of~\eqref{eq:objective}.
-In practice, the cut-off points $\mb$ are often unknown, so we choose to maximize $\logl$ jointly over $(\Theta,\mb)\in\tP\times \tB$.
+In practice, the cut-off points $\mb$ are often unknown, so we choose to maximize $\logl$ jointly over $(\Theta,\mb)\in\tP\times \tB$ (see Appendix~\ref{sec:extention} for details).
 The objective $\logl$ is concave in $(\Theta,\mb)$ whenever $f'$ is log-concave (see Section~\ref{sec:proofconvexity}).
@@
 \STATE {Update $\mathcal{C}^{(t+1)}$ while fixing other blocks:}
-              \newline$\mathcal{C}^{(t+1)} \gets \argmax_{\mathcal{C}\in \mathbb{R}^{r_1\times\cdots\times r_K}}\mathcal{L}_{\mathcal{Y},\Omega}(\mathcal{C})$, s.t.
-$\mnormSize{}{\Theta^{(t+1)}}\leq \alpha.$
+\newline$\mathcal{C}^{(t+1)} \gets \argmax_{\mathcal{C}\in \mathbb{R}^{r_1\times\cdots\times r_K}}\mathcal{L}_{\mathcal{Y},\Omega}(\mathcal{C})$, s.t. $\mnormSize{}{\Theta^{(t+1)}}\leq \alpha.$
 \STATE {Update $\Theta^{(t+1)}$ based on the current block estimates:}
@@
 \centering
-\includegraphics[width=0.6\textwidth]{panel.pdf}
+\includegraphics[width=0.7\textwidth]{panel.pdf}
 \caption{Empirical relationship between (relative) MSE versus (a) dimension $d$, (b) signal level $\alpha$, (c) observation fraction $\rho$, and (d) number of ordinal levels $L$.
@@
 \centering
-\includegraphics[width=0.6\textwidth]{panel2.pdf}
+\includegraphics[width=0.7\textwidth]{panel2.pdf}
 \caption{Performance comparison for predicting most likely labels.
@@
 On the other hand, the mode estimator tends to yield smaller MCR than the median estimator, $\text{MCR}(\tY^{\text{mode}}, \hat \tY^{\text{mode}})\leq\text{MCR}(\tY^{\text{median}}, \hat \tY^{\text{median}})$ (Figures~\ref{fig:compare}c-d vs.\ Figures~\ref{fig:compare2}c-d).
-This tendency is from the property that the median estimator $\hat y^{(\text{median})}_\omega$ minimizes $g_1(z)=\mathbb{E}_{y_\omega}|y_\omega-z|$, whereas the mode estimator $\hat y^{(\text{mode})}_\omega $ minimizes $g_2(z)=\mathbb{E}_{y_\omega}\mathds{1}_{\{y_\omega=z\}}$.
+This tendency is from the property that the median estimator $\hat y^{(\text{median})}_\omega$ minimizes $R_1(z)=\mathbb{E}_{y_\omega}|y_\omega-z|$, whereas the mode estimator $\hat y^{(\text{mode})}_\omega $ minimizes $R_2(z)=\mathbb{E}_{y_\omega}\mathds{1}_{\{y_\omega=z\}}$.
 Here the expectation is over the categorical distribution of $y_\omega$ given parameters $\hat \Theta$ and $\hat \mb$.
@@
 \begin{center}
-\includegraphics[width=.6\textwidth]{panel3.pdf}
+\includegraphics[width=.7\textwidth]{panel3.pdf}
 \end{center}
@@
 We apply our ordinal tensor method to two real-world datasets.
-In the first application, we use our model to analyze an ordinal tensor consisting of structural connectivities among 68 brain regions for 136 individuals from Human Connectome Project (HCP)~\citep{van2013wu}.
+In the first application, we use our model to analyze an ordinal tensor consisting of structural connectivities among 68 brain nodes for 136 individuals from Human Connectome Project (HCP)~\citep{van2013wu}.
 In the second application, we perform tensor completion to an ordinal dataset with missing values.
@@
 
-{\bf Human Connectome Project (HCP).} Each entry in the HCP dataset takes value on a nominal scale, \{{\it high, moderate, low}\}, indicating the strength level of fiber connection.
+\subsection{Human Connectome Project (HCP)}
+
+Each entry in the HCP dataset takes value on a nominal scale, \{{\it high, moderate, low}\}, indicating the strength level of fiber connection.
 We convert the dataset to a 3-level ordinal tensor $\tY\in[3]^{68\times 68\times 136}$ and apply the ordinal tensor method with a logistic link function.
 The BIC suggests $\mr = (23,23,8)$ with $\logl(\hat{\Theta},\hat{\mb}) = -216,646$.
-Based on the estimated Tucker factors $\{\hat \mM_k\}$, we perform a clustering analysis via K-mean on the brain nodes (see Appendix~\ref{sec:clustering} for detailed procedure).
-The 68 brain nodes are grouped into eight clusters.
-We find that the clustering capture the spatial separation between brain regions (Supplementary Figure~\ref{figure:brain image}).
-In particular, the top two clusters represent the left and right hemispheres; the smaller clusters represent local regions driving by similar nodes (Supplementary Table~\ref{table:clustering}).
-For example, the cluster III/IV consists of nodes in the supramarginal gyrus region in the left/right hemisphere.
+Based on the estimated Tucker factors $\{\hat \mM_k\}$, we perform a clustering analysis via K-mean on the brain nodes (see Appendix~\ref{sec:additionalHCP} for detailed procedure).
+The 68 brain nodes are grouped into 11 clusters.
+We find that the clustering captures the spatial separation between brain regions very well (Table~\ref{table:clustering}).
+In particular, cluster I represents the connection between the left and right hemispheres, whereas clusters II-III represent the connection within each of the half brains (Figure~\ref{figure:brain image}).
+Other smaller clusters represent local regions driving by similar nodes (Table~\ref{table:clustering}).
+For example, the cluster IV/VII consists of nodes in the supramarginal gyrus region in the left/right hemisphere.
 This region is known to be involved in visual word recognition and reading~\citep{stoeckel2009supramarginal}.
-The spatial separation between clusters suggests the applicability of our clustering method even without knowledge of external annotations.
+The identified similarities among nodes without external annotations illustrate the applicability of our method to clustering analysis.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width = 1\textwidth]{HCP.pdf}
+\caption{Top three clusters in the HCP analysis.
+(a) Cluster I reflects the connections between two brain hemispheres.
+(b)-(c) Cluster II/III consists of nodes within left/right hemisphere only.
+Node name are shown in abbreviation.
+Edges are colored based on predicted connection level averaged across individuals.
+}  \label{figure:brain image}
+\end{figure}
+
+\begin{table}[H]
+\center
+\begin{sc}
+\resizebox{\columnwidth}{!}{
+\begin{tabular}{|c|c|c|c|}
+\hline
+Cluster     & \multicolumn{3}{c|}{I}\\ \hline
+Brain nodes & \multicolumn{3}{l|}{\begin{tabular}[c]{@{}l@{}} l.frontalpole, l.temporalpole, l.medialorbitofrontal,  l.cuneus, l.parahippocampal, l.lingual,            \\
+r.frontalpole, r.temporalpole,  r.medialorbitofrontal,  r.cuneus,  r.parahippocampal \\
+\end{tabular}} \\
+\hline
+Cluster & \multicolumn{3}{c|}{II}\\ \hline
+Brain nodes & \multicolumn{3}{l|}{\begin{tabular}[c]{@{}l@{}}
+l.caudalmiddlefrontal,  l.inferiorparietal,  l.insula,  l.isthmuscingulate, l.lateraloccipital(2), \\
+l.parsopercularis, l.parstriangularis, l.postcentral, l.precuneus, l.superiorfrontal, l.superiortemporal(3)
+\end{tabular}} \\ 
+\hline
+Cluster & \multicolumn{3}{c|}{III}\\ \hline
+Brain nodes & \multicolumn{3}{l|}{\begin{tabular}[c]{@{}l@{}}
+r.caudalmiddlefrontal, r.inferiorparietal, r.insula, r.isthmuscingulate, r.lateraloccipital(2), r.lingual, \\
+r.parsopercularis, r.parstriangularis, r.postcentral, r.precentral, r.precuneus, r.superiorfrontal(3), \\
+r.superiorparietal,  r.superiortemporal(3)
+\end{tabular}} \\ 
+\hline
+Cluster      & IV    & V  & VI\\
+\hline
+Brain nodes & \begin{tabular}[c]{@{}c@{}}l.supramarginal(4)\end{tabular}&
+\begin{tabular}[c]{@{}c@{}}l.inferiortemporal(3)\end{tabular}&
+\begin{tabular}[c]{@{}c@{}}l.middletemporal(3)\end{tabular}\\
+\hline
+Cluster & VII  & VIII & VIIII\\
+\hline
+Brain nodes &
+\begin{tabular}[c]{@{}c@{}}r.supramarginal(4)\end{tabular}&
+\begin{tabular}[c]{@{}c@{}}r.inferiortemporal(3)\end{tabular}&
+\begin{tabular}[c]{@{}c@{}}r.middletemporal(3)\end{tabular}\\
+\hline
+Cluster &X & \multicolumn{2}{c|}{XI}  \\
+\hline
+Brain nodes &
+\begin{tabular}[c]{@{}c@{}}l.superiorfrontal(2)\end{tabular}&
+ \multicolumn{2}{c|}{l.precentral,  l.superiorparietal}\\
+ \hline
+\end{tabular}
+}
+\end{sc}
+\caption{Node clusters in the HCP analysis.
+The first alphabet in the node name indicates the left (L) or right (R) hemisphere.
+The number in the parentheses indicates the node count in each cluster.
+}  \label{table:clustering}
+\end{table}
 
@@
 
-{\bf InCarMusic recommendation system.} We apply ordinal tensor completion to a recommendation system InCarMusic.
+\subsection{InCarMusic recommendation system}
+
+We apply ordinal tensor completion to a recommendation system InCarMusic.
 InCarMusic is a mobile application that offers music recommendation to passengers of cars based on contexts~\citep{baltrunas2011incarmusic}.
@@
 \begin{equation}\label{eq:sample}
-\MFnormSize{}{\hat \Theta-\trueT}\leq {2C_2U_\alpha  r_{\max}^{(K-1)/2} \over L_\alpha} \sqrt{\sum_kd_k}.
-\end{equation}
+\MFnormSize{}{\hat \Theta-\trueT}\leq {2C_2U_\alpha  r_{\max}^{(K-1)/2} \over L_\alpha} \sqrt{\sum_kd_k},
+\end{equation}
+with probability at least $1-\exp(-C_1 \sum_k d_k)$.
 Lastly, we invoke the result regarding the closeness of $\Theta$ to its sampled version $\Theta_{\Omega}$, under the entrywise bound condition.
@@
 
-\subsection{Auxiliary lemmas}
-\label{sec:lemma}
-
+\subsection{Auxiliary lemmas}\label{sec:lemma}
 This section collects lemmas that are useful for the proofs of the main theorems.
@@
 
-\begin{defn}[Absolute continuity]
-We use $\lambda\ll\mu$ to denote absolutely continuity of a measure $\lambda$ with respect to another measure $\mu$; that is, $\lambda(E)=0$ whenever $\mu(E)=0$.
-\end{defn}
-
 \begin{lem}[Theorem 2.5 in~\cite{tsybakov2008introduction}]\label{lem:Tsybakov}
@@
 \item $d(\Theta_j,\ \Theta_k)\geq 2s>0$, $\forall 0\leq j\leq k\leq M$;
-\item $\mathbb{P}_j\ll\mathbb{P}_0$, $\forall j=1,\ldots,M$, and
+\item $\mathbb{P}_0$ is absolutely continuous with respect to $\mathbb{P}_j$, $\forall j=1,\ldots,M$, and
 \[
@@
 \]
-where $d\colon \tX\times \tX\mapsto [0,+\infty]$ is a semi-distance function, $0<\alpha<{1/8}$ and $P_j=P_{\Theta_j}$, $j=0,1\ldots,M$.
+where $d\colon \tX\times \tX\mapsto [0,+\infty]$ is a semi-distance function, $0<\alpha<{1/8}$ and $\mathbb{P}_j=\mathbb{P}_{\Theta_j}$, $j=0,1\ldots,M$.
 \end{itemize}
@@
 \section{Extension of Theorem~\ref{thm:rate} to unknown cut-off points}\label{sec:extention}
-
 When the cut-off points $\mb$ is unknown, we estimate $(\hat \Theta,\hat \mb)$ by
@@
 \[
-\text{MSE}\left((\hat \Theta,\hat \mb),(\Theta,\mb)\right)={1\over \prod_k d_k+L-1}\FnormSize{}{(\hat \Theta,\hat \mb)-(\Theta,\mb)}^2
+\text{MSE}\left( \hat \Theta, \trueT \right)={1\over \prod_k d_k}\FnormSize{}{\hat \Theta-\trueT},\quad \text{MSE}\left( \hat \mb, \trueb \right)={1\over L-1}\FnormSize{}{\hat \mb-\trueb}.
+\]
+By abuse of notation, we use $(\Theta,\mb)$ to denote the length-$(d_{\text{total}}+L-1)$-vector collecting all parameters together. The total MSE is defined as
+\[
+\text{MSE}\left((\hat \Theta,\hat \mb),\ (\trueT,\trueb)\right)={1\over \prod_k d_k+L-1}\FnormSize{}{(\hat \Theta,\hat \mb)-(\trueT,\trueb)}^2.
 \]
@@
 \notag \\
-&\hspace{1.5in} \left. \frac{\ddot{f}(b_\ell-\theta)g_{\ell+1}(\theta)-\dot{f}(b_\ell-\theta)\dot{g}_{\ell+1}(\theta)}{g^2_{\ell+1}(\theta)}\right\}\\
-\tilde{D}_{\alpha,\beta,\Delta} &=\max_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}} \max_{\ell\in[L-1]}\max\left\{  {\partial \over \partial \theta}\left({\dot{f}(b_{\ell}-\theta)\over g_\ell(\theta)}\right),\ -{\partial \over \partial \theta}\left({\dot{f}(b_{\ell}-\theta)\over g_{\ell+1}(\theta)}\right)\right\}\notag.
+&\hspace{1.5in} \left. \frac{\ddot{f}(b_\ell-\theta)g_{\ell+1}(\theta)-\dot{f}(b_\ell-\theta)\dot{g}_{\ell+1}(\theta)}{g^2_{\ell+1}(\theta)}\right\},\notag \\
+\tilde{D}_{\alpha,\beta,\Delta} &=\max_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}} \max_{\ell\in[L-1]}\max\left\{  {\partial \over \partial \theta}\left({\dot{f}(b_{\ell}-\theta)\over g_\ell(\theta)}\right),\ -{\partial \over \partial \theta}\left({\dot{f}(b_{\ell}-\theta)\over g_{\ell+1}(\theta)}\right)\right\}.
 \end{align}
@@
 \end{rmk}
+
 \begin{rmk}
@@
 Suppose the link function $f$ satisfies Assumption~\ref{ass:joint}.
-Define $r_{\max}=\max_k r_k$.
+Define $r_{\max}=\max_k r_k$, and assume $r_{\max}=\tO(1)$.
 Then with very high probability, the estimator in~\eqref{eq:joint} satisfies
+\begin{equation}\label{eq:jointTheta}
+\mathrm{MSE}\left(\hat \Theta, \trueT \right)\leq\min\left( 4\alpha^2,\  {c_1U_{\alpha,\beta,\Delta}^2  r_{\max}^{K-1}  \over L_{\alpha,\beta,\Delta}^2}{ \sum_k d_k \over \prod_k d_k}\right),
+\end{equation}
+and
+\begin{equation}\label{eq:jointb}
+\mathrm{MSE}\left(\hat \mb, \trueb \right)\leq \min\left(4\beta^2,\ {C'_{\alpha,\beta,\Delta}r^{K-1}_{\max} \over D_{\alpha,\beta,\Delta}^2}{\left(\sum_kd_k\right)\left(\prod_k d_k\right) \over \min_{\ell}(n_{\ell}+n_{\ell+1})^2}\right),
+\end{equation}
+where $c_1, U_{\alpha,\beta,\Delta}, L_{\alpha,\beta,\Delta}, C'_{\alpha,\beta,\Delta}, \tilde D_{\alpha,\beta,\Delta}$ are constants independent of the tensor dimension and rank.
+
+In particular, we have
 \begin{align}
-\mathrm{MSE}\left((\hat \Theta,\hat \mb),(\trueT,\trueb)\right) & \leq {c_1U_{\alpha,\beta,\Delta}^2 \over L_{\alpha,\beta,\Delta}^2} r_{\max}^{(K-1)} { \sum_k d_k \over \prod_k d_k}+{C_2 \over D_{\alpha,\beta,\Delta}^2}{(\sum_k d_k)^2\over \min_{\ell}(n_{\ell}+n_{\ell+1})^2}.
-\end{align}
-Briefly,
-\begin{align}
-\mathrm{MSE}\left((\hat \Theta,\hat \mb),(\trueT,\trueb)\right) &= \tO\left({\sum_k d_k\over\prod_k d_k}\right)+\tO\left({(\sum_k d_k)^2\over \min_\ell(n_\ell+n_{\ell+1})^2}\right)\\& = \tO\left({(\sum_k d_k)^2\over \min\left\{\sum_k d_k\prod_k d_k ,\min_\ell(n_\ell+n_{\ell+1})^2\right\}}\right).
-\end{align}
-where $c_1>0$ is a constant independent of the tensor dimension and rank, and $U_{\alpha,\beta,\Delta}$, $L_{\alpha,\beta,\Delta}$,  $D_{\alpha,\beta,\Delta}>0$ are constants defined in~\eqref{eq:UL} and~\eqref{eq:CD} and $C_2>0$ is a constant that depends on $ C_{\alpha,\beta,\Delta},\tilde{D}_{\alpha,\beta,\Delta},L$.
+\mathrm{MSE}\left((\hat \Theta,\hat \mb),\ (\trueT,\trueb)\right) = \tO\left({\sum_k d_k \over \min\left\{\prod_k d_k ,\ \min_\ell(n_\ell+n_{\ell+1})^2\right\}}\right),\quad \text{as}\quad d_{\min}\to\infty.
+\end{align}
 \end{thm}
 
-\begin{rmk}
-The total MSE has two components where $\tO\left({\sum_k d_k\over\prod_k d_k}\right)$ is from the error in $\hat\Theta$ and $\tO\left({(\sum_k d_k)^2\over \min_\ell(n_\ell+n_{\ell+1})^2}\right)$ is from the error in $\hat \mb.$ When $\min_\ell(n_\ell+n_{\ell+1})\gg \sum_k d_k\prod_k d_k$, the error in $\hat\Theta$ dominates the total MSE.
-When $\min_\ell(n_\ell+n_{\ell+1})\ll \sum_k d_k\prod_k d_k $, the error in $\hat\mb$ dominates the total MSE.
+\begin{rmk} The total MSE has two components, where the first component $\tO\left({\sum_k d_k \over\prod_k d_k}\right)$ is from the error in $\hat\Theta$, and the second component $\tO\left({\sum_k d_k \over \min_\ell(n_\ell+n_{\ell+1})^2}\right)$ is from the error in $\hat \mb.$ When the $L$ labels are moderately balanced in that $\min_\ell(n_\ell+n_{\ell+1})\gg \sqrt{\prod_kd_k}$, the error in $\hat\Theta$ dominates the total MSE.
+When the $L$ labels are severely imbalanced in that $\min_\ell(n_\ell+n_{\ell+1})\ll \sqrt{\prod_kd_k}$, the error in $\hat\mb$ dominates the total MSE.
 \end{rmk}
 
-\begin{rmk}
-  The constant bounds with $\alpha$ and $\beta$ can be obtained trivially from the definition of the feasible sets. For notational convenience, we ignore this constant bound.
+\begin{rmk} We discuss the impact of imbalance between labels to the estimation accuracy.
+Consider the special case $d_1=\cdots =d_K=d$.
+The bounds~\eqref{eq:jointTheta} and~\eqref{eq:jointb} demonstrate that, both $\hat \Theta$ and $\hat \mb$ achieve consistency as long as $\min_\ell(n_\ell+n_{\ell+1})$ is slightly larger than $\tO(d^{(K+1)/2})$.
+Note that in this case ${\min_\ell(n_\ell+n_{\ell+1}) \over \max_\ell n_{\ell}}\asymp {\min_\ell(n_\ell+n_{\ell+1})\over d^K} \to 0$, indicating a long-run imbalance between labels.
+The obtained consistency highlights the robustness of our proposed estimator~\eqref{eq:joint} to the label imbalance.
 \end{rmk}
@@
 \begin{proof}[Proof of Theorem~\ref{thm:ratejoint}]
+The constant bounds with $\alpha$ and $\beta$ can be obtained trivially from the definition of the feasible sets. We focus on proving the non-constant bounds.
+
 Similar to the proof of Theorem~\ref{thm:rate}, we suppress $\Omega$ in the subscript. Based on the definition of $(\hat \Theta, \hat \mb)$,
@@
 \begin{equation}\label{eq:ratenew}
-\FnormSize{}{\hat \Theta-\trueT}^2\leq {c_1^2U_{\alpha,\beta,\Delta}^2\over L_{\alpha,\beta,\Delta}^2} r_{\max}^{(K-1)}  \sum_k d_k ,
-\end{equation}
-where $U_{\alpha,\beta,\Delta}, L_{\alpha,\beta,\Delta}>0$ are two universal constants independent of the tensor dimension and rank.
-Next we bound $\FnormSize{}{\hat \mb-\trueb}^2$ given $\hat\Theta$. It follows from the expression of $\flogl(\Theta,\mb)$ that
+\FnormSize{}{\hat \Theta-\trueT}^2\leq {c_1^2U_{\alpha,\beta,\Delta}^2 r_{\max}^{K-1} \over L_{\alpha,\beta,\Delta}^2}   \sum_k d_k ,
+\end{equation}
+where $U_{\alpha,\beta,\Delta}, L_{\alpha,\beta,\Delta}>0$ are two constants defined in~\eqref{eq:UL}.
+
+Next we bound $\FnormSize{}{\hat \mb-\trueb}^2$ given $\hat\Theta$.
+We consider the profile log-likelihood $\flogl(\hat \Theta,\mb)$ as a function of $\mb\in\tB$.
+For notational convenience, we drop $\hat \Theta$ from $\flogl(\hat \Theta,\mb)$ and simply write $\flogl(\mb)$.
+It follows from the expression of $\flogl(\mb)$ that
 \begin{align}\label{eq:propertyb}
@@
 
-We consider the profile log-likelihood $\flogl(\hat \Theta,\mb)$ as a function of $\mb\in\tB$.
-For notational convenience, we drop $\hat \Theta$ from $\flogl(\hat \Theta,\mb)$ and simply write $\flogl(\mb)$.
 By the second-order Taylor's expansion of $\flogl(\mb)$ around $\trueb$, we obtain
@@
 The linear term in~\eqref{eq:taylorb} can be bounded by Cauchy-Schwartz inequality,
-\begin{equation}\label{eq:linearb}
-(\trueb-\hat\mb)^T\fploglb(\trueb)\leq \FnormSize{}{\trueb-\hat\mb}\FnormSize{}{\fploglb(\trueb)}
-\leq \FnormSize{}{\trueb-\hat\mb}C\sqrt{L-1}\sum_k d_k\sqrt{\prod_k d_k},
-\end{equation}
-where the last inequality is followed from Lemma \ref{lem:sharpergd} stated as,
-\[
-\left|{\partial \flogl\over \partial b_\ell}\bigg|_{(\hat \Theta,\trueb)}\right|\leq C\sum_k d_k\sqrt{\prod_k d_k},\quad \text{ for all } \ell \in [L-1] \text{ with high probability}.
-\]
+\begin{align}\label{eq:linearb}
+(\trueb-\hat\mb)^T\fploglb(\trueb)&\leq \FnormSize{}{\trueb-\hat\mb}\FnormSize{}{\fploglb(\trueb)}\\
+&\leq C'_{\alpha,\beta,\Delta}\FnormSize{}{\trueb-\hat\mb}\sqrt{L-1}\sqrt{\left(\sum_k d_k\right)\left( \prod_k d_k\right)},
+\end{align}
+for some constant $C'_{\alpha,\beta,\Delta}>0$, where the last inequality follows from Lemma \ref{lem:sharpergd},
+\[
+\left|{\partial \flogl\over \partial b_\ell}\bigg|_{(\hat \Theta,\trueb)}\right|\leq C'_{\alpha,\beta,\Delta}\sqrt{\left(\sum_k d_k\right)\left(\prod_k d_k\right)},\quad \text{ for all } \ell \in [L-1] \text{ with very high probability}.
+\]
+
 We next bound the quadratic term in \eqref{eq:taylorb}. Note that
@@
 \[
-\flogl(\hat \mb)\leq \flogl(\trueb)+C\sqrt{L-1} \sum_k d_k\sqrt{\prod_k d_k}) \FnormSize{}{\hat\mb-\trueb} -\frac{D_{\alpha,\beta,\Delta}}{2}\min_{\ell}(n_\ell+n_{\ell+1}) \FnormSize{}{\hat\mb-\trueb}^2 .
+\flogl(\hat \mb)\leq \flogl(\trueb)+C'_{\alpha,\beta,\Delta} \sqrt{(L-1)\left(\sum_k d_k\right)\left(\prod_k d_k\right)} \FnormSize{}{\hat\mb-\trueb} -\frac{D_{\alpha,\beta,\Delta}}{2}\min_{\ell}(n_\ell+n_{\ell+1}) \FnormSize{}{\hat\mb-\trueb}^2 .
 \]
@@
 \[
-C\sqrt{L-1} \sum_k d_k\sqrt{\prod_k d_k}) \FnormSize{}{\hat\mb-\trueb} -\frac{D_{\alpha,\beta,\Delta}}{2}\min_{\ell}(n_\ell+n_{\ell+1}) \FnormSize{}{\hat\mb-\trueb}^2\geq 0
+C'_{\alpha,\beta,\Delta} \sqrt{(L-1)\left(\sum_k d_k\right)\left(\prod_k d_k\right)} \FnormSize{}{\hat\mb-\trueb} -\frac{D_{\alpha,\beta,\Delta}}{2}\min_{\ell}(n_\ell+n_{\ell+1}) \FnormSize{}{\hat\mb-\trueb}^2\geq 0.
 \]
 Finally,
-\[
-\FnormSize{}{\hat\mb-\trueb}^2\leq {C_2 \over D_{\alpha,\beta,\Delta}^2}{(\sum_k d_k)^2\prod_k d_k\over \min_{\ell}(n_{\ell}+n_{\ell+1})^2},
-\]
-where $C_2(C_{\alpha,\beta,\Delta},\tilde{D}_{\alpha,\beta,\Delta},L)$ is a positive constant.
-The following completes the proof.
+\begin{equation}\label{eq:rateb}
+\FnormSize{}{\hat\mb-\trueb}^2\leq {C'_{\alpha,\beta,\Delta} (L-1) \over D_{\alpha,\beta,\Delta}^2}{\left(\sum_k d_k\right)\left(\prod_k d_k\right)\over \min_{\ell}(n_{\ell}+n_{\ell+1})^2}.
+\end{equation}
+Combing~\eqref{eq:ratenew} and~\eqref{eq:rateb} with the following inequality completes the proof.
 \begin{align}
@@
 
-\begin{rmk}[MSE for $\hat\mb$]
-In the proof, we have shown that
-\[
-\FnormSize{}{\hat\mb-\trueb}\leq {\sqrt{C_2} \over D_{\alpha,\beta,\Delta}}{(\sum_k d_k)\sqrt{\prod_k d_k}\over \min_{\ell}(n_{\ell}+n_{\ell+1})}.
-\]
-This bound is sharper than the trivial bound $\FnormSize{}{\trueb-\hat\mb}\leq 2\beta\sqrt{L-1}$ which is from $|b_\ell^{\text{true}}-\hat b_\ell|\leq 2\beta.$ In particular, $\FnormSize{}{\hat\mb-\trueb}\rightarrow 0$ as $\min_\ell(n_\ell+n_{\ell+1})\asymp \tilde \tO(\sum_k d_k\sqrt{\prod_k d_k})\rightarrow \infty.$
-\end{rmk}
-
-\begin{lem}[CLT for i.d. Bernoulli]\label{lem:clt}
-Suppose $\{X_n\}$ is a series of independent Bernoulli trials with possibly different success probabilities $p_n$. Let $s_n = \sum_{m=1}^n p_m(1-p_m)$ and $Y_n = X_n-p_n$. Then,
-\[{1\over s_n}\sum_{m=1}^n Y_n \stackrel{\mathcal{D}}{\rightarrow} N(0,1), \quad \text{ as } n \rightarrow \infty,\]
-whenever $s_n = \sum_{m=1}^n p_m(1-p_m)\rightarrow \infty.$
+\begin{lem}[CLT for independent Bernoulli r.v.'s]\label{lem:clt}
+Let $\{X_n\}$ be a series of independent Bernoulli random variables with possibly different success probabilities $\{p_n\}$.
+Suppose the success probabilities $\{p_n\}$ are bounded, in that, $c_1\leq p_n\leq 1-c_2$ for some positive constants $c_1,c_2>0$ and all $n\in[m]$.
+Define $s^2_m = \sum_{n=1}^m p_n(1-p_n)$ and $Y_n = X_n-p_n$.
+Then,
+\[
+{1\over s_m}\sum_{n=1}^m Y_n \stackrel{\mathcal{D}}{\rightarrow} N(0,1), \quad \text{ as } m \rightarrow \infty.
+\]
 \end{lem}
@@
 \begin{proof}
-Let us look for a Lyapunov condition for $\delta = 1$.
-First, we have
+We apply Lyapunov central limit theorem (CLT) to $\{Y_n\}$. Let us verify Lyapunov's condition for $\delta = 1$:
 \[
@@
 \]
-Hence, summation of the above inequality shows,
-\[
-\sum_{m=1}^n\mathbb{E}[|Y_n|^3]\leq \sum_{m=1}^n p_m(1-p_m) = s_n^2.
-\]
-Thus, the Lyapunov condition is satisfied whenever,
-\[{1\over s_n^3}\sum_{m=1}^n \mathbb{E}[|Y_n|^3]\leq {s_n^2\over s_n^3}\rightarrow 0.\]
-or simply $s_n = \sum_{m=1}^n p_m(1-p_m)\rightarrow \infty.$ Central limit theorem on $Y_n$ completes the proof.
-
+Summation of the above inequality shows
+\[
+\sum_{n=1}^m\mathbb{E}[|Y_n|^3]\leq \sum_{n=1}^m p_n(1-p_n) = s_m^2.
+\]
+Thus, the Lyapunov condition is satisfied whenever
+\[
+{1\over s_m^3}\sum_{n=1}^m \mathbb{E}[|Y_n|^3]\leq {s_m^2\over s_m^3}\rightarrow 0,
+\]
+or simply $s_m = \sum_{n=1}^m p_n(1-p_n) \geq m\min\{c_1(1-c_1), c_2(1-c_2)\} \rightarrow \infty.$ Applying Lyapunov CLT to $\{Y_n\}$ completes the proof.
 \end{proof}
 
-\begin{lem}[Bound on Gradients]\label{lem:sharpergd}
-Consider the same set-up as in Theorem \ref{thm:ratejoint}. Then, with very high probability,
-\[
-\left|{\partial \flogl\over \partial b_\ell}\bigg|_{(\hat \Theta,\trueb)}\right|\leq \left(C_{\alpha,\beta,\Delta}+\tilde{D}_{\alpha,\beta,\Delta}\FnormSize{}{\trueT-\hat\Theta}^2\right)\sqrt{\prod_k d_k},\quad \text{ for all } \ell \in [L-1],
-\]
-In particular, there exists a constant $d_0\in \mathbb{N}_+$ such that for all $d_k\geq d_0$ for $k \in [K]$,
-\[
-\left|{\partial \flogl\over \partial b_\ell}\bigg|_{(\hat \Theta,\trueb)}\right|\leq C\sum_k d_k\sqrt{\prod_k d_k},\quad \text{ for all } \ell \in [L-1],
-\]
-where $C(C_{\alpha,\beta,\Delta},\tilde{D}_{\alpha,\beta,\Delta})>0$ is a constant.
+\begin{lem}[Bound on score function]\label{lem:sharpergd}
+Consider the same set-up as in Theorem~\ref{thm:ratejoint}. Let $\tL_\tY(\Theta,\mb)$ denote the log-likelihood function of $(\Theta, \mb)$ given data tensor $\tY$. Then, with very high probability,
+\[
+\left|{\partial \flogl\over \partial b_\ell}\bigg|_{(\Theta,\trueb)}\right|\leq \left(C_{\alpha,\beta,\Delta}+\tilde{D}_{\alpha,\beta,\Delta}\FnormSize{}{\trueT-\Theta}^2\right)\sqrt{\prod_k d_k},\quad \text{ for all } \ell \in [L-1],
+\]
+In particular, with very high probability,
+\[
+\left|{\partial \flogl\over \partial b_\ell}\bigg|_{(\hat \Theta,\trueb)}\right|\leq C'_{\alpha,\beta,\Delta}\sqrt{\left(\sum_k d_k\right)\left( \prod_k d_k\right)},\quad \text{ for all } \ell \in [L-1],
+\]
+where $C'_{\alpha,\beta,\Delta}>0$ is a constant that depends on $U_{\alpha,\beta,\Delta}, L_{\alpha,\beta,\Delta}, C_{\alpha,\beta,\Delta},  \tilde D_{\alpha,\beta,\Delta}$ defined in~\eqref{eq:UL} and~\eqref{eq:CD}.
 \end{lem}
+
 \begin{proof}
 We only prove the case for $\ell = 1$. Other cases can be proved similarly.
-Note that
+
 \begin{equation}\label{eq:sharpergd}
-  {\partial \flogl\over \partial b_1}\bigg|_{(\hat \Theta,\trueb)} = \KeepStyleUnderBrace{{\partial \flogl\over \partial b_1}\bigg|_{(\hat \Theta,\trueb)}-\mathbb{E}_\mathcal{Y}\left[{\partial \flogl\over \partial b_1}\bigg|_{(\hat \Theta,\trueb)}\right]}_{\coloneqq A}+\KeepStyleUnderBrace{\mathbb{E}_\mathcal{Y}\left[{\partial \flogl\over \partial b_1}\bigg|_{(\hat \Theta,\trueb)}\right]-\mathbb{E}_\mathcal{Y}\left[{\partial \flogl\over \partial b_1}\bigg|_{(\hat \trueT,\trueb)}\right] }_{\coloneqq B}.
-\end{equation}
-We have used the fact that the score function has mean zero, $\mathbb{E}_\mathcal{Y}\left[{\partial \flogl\over \partial b_1}\bigg|_{(\hat \trueT,\trueb)}\right]=0 $.
+  {\partial \flogl\over \partial b_1}\bigg|_{(\Theta,\trueb)} = \KeepStyleUnderBrace{{\partial \flogl\over \partial b_1}\bigg|_{( \Theta,\trueb)}-\mathbb{E}_\mathcal{Y}\left[{\partial \flogl\over \partial b_1}\bigg|_{(\Theta,\trueb)}\right]}_{\coloneqq A}+\KeepStyleUnderBrace{\mathbb{E}_\mathcal{Y}\left[{\partial \flogl\over \partial b_1}\bigg|_{( \Theta,\trueb)}\right]-\mathbb{E}_\mathcal{Y}\left[{\partial \flogl\over \partial b_1}\bigg|_{(\trueT,\trueb)}\right] }_{\coloneqq B}.
+\end{equation}
+We have used the fact that the score function has mean zero, $\mathbb{E}_\mathcal{Y}\left[{\partial \flogl\over \partial b_1}\big|_{(\trueT,\trueb)}\right]=0 $.
 Here all expectations are taken with respect to $\mathcal{Y}\sim \mathbb{P}(\trueT,\trueb).$\\
@@
 \begin{align}
-A &= \sum_{\omega\in\Omega}\left\{[\mathds{1}_{y_\omega = 1}-g_1(\theta^{\text{true}}_\omega)]{\dot{f}(b_1-\hat\theta_\omega)\over g_1(\hat\theta_\omega)}-[\mathds{1}_{y_\omega = 2}-g_2(\theta^{\text{true}}_\omega)]{\dot{f}(b_1-\hat\theta_\omega)\over g_2(\hat\theta_\omega)}\right\}\\
+A &= \sum_{\omega\in\Omega}\left\{[\mathds{1}_{y_\omega = 1}-g_1(\theta^{\text{true}}_\omega)]{\dot{f}(b_1-\theta_\omega)\over g_1(\theta_\omega)}-[\mathds{1}_{y_\omega = 2}-g_2(\theta^{\text{true}}_\omega)]{\dot{f}(b_1-\theta_\omega)\over g_2(\theta_\omega)}\right\}\\
 &\leq C_{\alpha,\beta,\Delta}\sum_{\omega\in\Omega}\KeepStyleUnderBrace{[\mathds{1}_{y_\omega = 1}+\mathds{1}_{y_\omega = 2}- g_1(\theta^{\text{true}}_\omega)-g_2(\theta^{\text{true}}_\omega)]}_{\coloneqq W_\omega}.
 \end{align}
-Note that $\{W_\omega\}$ are centered independent Bernoulli random variables with success probabilities $g_1(\theta^{\text{true}}_\omega)+g_2(\theta^{\text{true}}_\omega)$.
+Note that $\{W_\omega\}$ are independent, centered Bernoulli random variables with bounded success probabilities $g_1(\theta^{\text{true}}_\omega)+g_2(\theta^{\text{true}}_\omega)$.
 By Lemma \ref{lem:clt}, we have
 \[
-\sum_{\omega\in\Omega}W_\omega \stackrel{\mathcal{D}}{\rightarrow} N\left(0, \sum_{\omega\in\Omega}\left(g_1(\theta^{\text{true}}_\omega)+g_2(\theta^{\text{true}}_\omega)\right)\left(1-g_1(\theta^{\text{true}}_\omega)+g_2(\theta^{\text{true}}_\omega)\right)\right).
-\]
-Hence, with the fact that $\sum_{\omega\in\Omega}\left(g_1(\theta^{\text{true}}_\omega)+g_2(\theta^{\text{true}}_\omega)\right)\left(1-g_1(\theta^{\text{true}}_\omega)+g_2(\theta^{\text{true}}_\omega)\right)\leq {1\over 4}\prod_k d_k$,
+\sum_{\omega\in\Omega}W_\omega \stackrel{\mathcal{D}}{\rightarrow} N\left(0,\ \sum_{\omega\in\Omega}\left(g_1(\theta^{\text{true}}_\omega)+g_2(\theta^{\text{true}}_\omega)\right)\left(1-g_1(\theta^{\text{true}}_\omega)-g_2(\theta^{\text{true}}_\omega)\right)\right).
+\]
+Hence, with the fact that $\sum_{\omega\in\Omega}\left(g_1(\theta^{\text{true}}_\omega)+g_2(\theta^{\text{true}}_\omega)\right)\left(1-g_1(\theta^{\text{true}}_\omega)-g_2(\theta^{\text{true}}_\omega)\right)\leq {1\over 4}\prod_k d_k$,
 \begin{equation}\label{eq:A}
-|A| \leq C_{\alpha,\beta,\Delta}\left|\sum_{\omega\in\Omega}W_\omega\right|\leq C_{\alpha,\beta,\Delta}\sqrt{\prod_k d_k}, \quad \text{ with high probability. }
-\end{equation}
-The second term $B$ in \eqref{eq:sharpergd} is the bias induced by the inaccuracy of $\hat\Theta$:
+|A| \leq C_{\alpha,\beta,\Delta}\left|\sum_{\omega\in\Omega}W_\omega\right|\leq C_{\alpha,\beta,\Delta}\sqrt{\prod_k d_k}, \quad \text{ with very high probability. }
+\end{equation}
+The second term $B$ in \eqref{eq:sharpergd} is the bias induced by $\Theta$:
 \begin{align}
-|B| &= \left|\sum_{\omega\in\Omega} g_1(\theta^{\text{true}}_\omega)\left({\dot{f}(b_1-\hat\theta_\omega)\over g_1(\hat\theta_\omega)}-{\dot{f}(b_1-\theta^{\text{true}}_\omega)\over g_1(\theta^{\text{true}}_\omega)}\right)-\sum_{\omega\in\Omega} g_2(\theta^{\text{true}}_\omega)\left({\dot{f}(b_1-\hat\theta_\omega)\over g_2(\hat\theta_\omega)}-{\dot{f}(b_1-\theta^{\text{true}}_\omega)\over g_2(\theta^{\text{true}}_\omega)}\right)\right|\\
-&= \Bigg|\sum_{\omega\in\Omega} g_1(\theta^{\text{true}}_\omega)(\theta^{\text{true}}_\omega-\hat\theta_\omega)\left\{{\partial\over\partial}\left({\dot{f}(b_1-\theta)\over g_1(\theta)}\right)\bigg|_{\rho_\omega\hat\theta_\omega+(1-\rho_\omega)\theta^{\text{true}}_\omega}\right\}\\
-&\hspace{0.5in}-\sum_{\omega\in\Omega} g_2(\theta^{\text{true}}_\omega)(\theta^{\text{true}}_\omega-\hat\theta_\omega)\left\{{\partial\over\partial}\left({\dot{f}(b_1-\theta)\over g_2(\theta)}\right)\bigg|_{\rho_\omega'\hat\theta_\omega+(1-\rho_\omega')\theta^{\text{true}}_\omega}\right\}\Bigg|\\
-&\leq \tilde{D}_{\alpha,\beta,\Delta}\sum_{\omega\in\Omega}|g_1(\theta^{\text{true}}_\omega)+g_2(\theta^{\text{true}}_\omega)||\theta^{\text{true}}_\omega-\hat\theta_\omega|.
-\end{align}
-By Cauchy-Schwartz inequality with the fact that $g_1(\theta)+g_2(\theta)<1$,
+|B| &= \left|\sum_{\omega\in\Omega} g_1(\theta^{\text{true}}_\omega)\left({\dot{f}(b_1-\theta_\omega)\over g_1(\theta_\omega)}-{\dot{f}(b_1-\theta^{\text{true}}_\omega)\over g_1(\theta^{\text{true}}_\omega)}\right)-\sum_{\omega\in\Omega} g_2(\theta^{\text{true}}_\omega)\left({\dot{f}(b_1-\theta_\omega)\over g_2(\theta_\omega)}-{\dot{f}(b_1-\theta^{\text{true}}_\omega)\over g_2(\theta^{\text{true}}_\omega)}\right)\right|\\
+&= \left|\sum_{\omega\in\Omega} g_1(\theta^{\text{true}}_\omega)(\theta^{\text{true}}_\omega-\theta_\omega)\left\{{\partial\over\partial \theta}\left({\dot{f}(b_1-\theta)\over g_1(\theta)}\right)\bigg|_{\rho_\omega\theta_\omega+(1-\rho_\omega)\theta^{\text{true}}_\omega}\right\}\right.\\
+&\hspace{0.5in}\left.-\sum_{\omega\in\Omega} g_2(\theta^{\text{true}}_\omega)(\theta^{\text{true}}_\omega-\theta_\omega)\left\{{\partial\over\partial \theta}\left({\dot{f}(b_1-\theta)\over g_2(\theta)}\right)\bigg|_{\rho_\omega'\theta_\omega+(1-\rho_\omega')\theta^{\text{true}}_\omega}\right\}\right|\\
+&\leq \tilde{D}_{\alpha,\beta,\Delta}\sum_{\omega\in\Omega}|g_1(\theta^{\text{true}}_\omega)+g_2(\theta^{\text{true}}_\omega)||\theta^{\text{true}}_\omega-\theta_\omega|.
+\end{align}
+By Cauchy-Schwartz inequality with the fact that $g_1(\theta^{\text{true}}_\omega)+g_2(\theta^{\text{true}}_\omega)\leq 1$,
 \begin{equation}\label{eq:B}
-|B| \leq \tilde{D}_{\alpha,\beta,\Delta}\sqrt{\prod_k d_k} \FnormSize{}{\trueT-\hat\Theta}.
+|B| \leq \tilde{D}_{\alpha,\beta,\Delta}\sqrt{\prod_k d_k} \FnormSize{}{\trueT-\Theta}.
 \end{equation}
@@
 \[
-{\partial \flogl\over \partial b_1}\bigg|_{(\hat \Theta,\trueb)}\leq \left(C_{\alpha,\beta,\Delta}+\tilde{D}_{\alpha,\beta,\Delta}\FnormSize{}{\trueT-\hat\Theta}\right)\sqrt{\prod_k d_k}.
-\]
-holds with very high probability. The second inequality in the lemma comes from the fact that $\FnormSize{}{\trueT-\hat\Theta}\leq \tO(\sum_k d_k).$
+{\partial \flogl\over \partial b_1}\bigg|_{(\Theta,\trueb)}\leq \left(C_{\alpha,\beta,\Delta}+\tilde{D}_{\alpha,\beta,\Delta}\FnormSize{}{\trueT-\Theta}\right)\sqrt{\prod_k d_k}.
+\]
+holds with very high probability. The second inequality in the lemma comes from~\eqref{eq:ratenew} that $\FnormSize{}{\trueT-\hat\Theta}\leq \tO(\sqrt{\sum_k d_k}).$
 \end{proof}
@@
 \section{Additional results of HCP analysis}\label{sec:additionalHCP}
-\subsection{Clustering based on Tucker representation}\label{sec:clustering}
 We perform clustering analyses based on the Tucker representation of the estimated tensor parameter $\hat\Theta$.
@@
 
-\subsection{Clustering results of HCP analysis}
 We perform a clustering analysis on the 68 brain nodes using the produce described above.
-Recall that our ordinal tensor method outputs the estimated parameter tensor $\hat\Theta\in\mathbb{R}^{68\times 68\times136}$ with rank $(23,23,8)$.
+Our ordinal tensor method outputs the estimated parameter tensor $\hat\Theta\in\mathbb{R}^{68\times 68\times136}$ with rank $(23,23,8)$.
 We apply $K$-means to the mode-1 principal component matrix of size $68\times 184$ ($184=23\times8$).
-The elbow method suggests eight clusters among the 68 nodes (see Figure~\ref{figure:elbow}).
-
-Table~\ref{table:clustering} shows the eight clusters and their compositions.
-We find that most brain nodes fall into cluster I and the cluster II, which can be represented as the left and right hemispheres, respectively (Figure~\ref{figure:brain image}).
-Clusters III-VIII captures brain regions with same annotations.
-For example, clusters III/IV represents the supramarginal gyrus region on the left/right hemisphere, respectively.
-This region is known to be involved in visual word recognition and reading~\citep{stoeckel2009supramarginal}.
-The results demonstrate that our clustering successfully groups similar brain regions together without knowledge of external annotation.
+The elbow method suggests 11 clusters among the 68 nodes (see Figure~\ref{figure:elbow}).
 
@@
 
-\begin{figure}[H]
-  \centering
- \includegraphics[width = 0.45\textwidth]{braincluster.pdf}
- \includegraphics[width = 0.45\textwidth]{braincluster3.pdf}
-  \caption{Top two clusters of brain nodes.
-(a) Nodes in yellow are from cluster I (left hemisphere), and nodes in green are from cluster II (right hemisphere).
-(b) Predicted connectivity within cluster I/II.
-Edges are colored based on predicted strength level averaged across individuals.}  \label{figure:brain image}
-\end{figure}
-
-\begin{table}[H]
- \center
-  \begin{small}
-    \footnotesize
-  \begin{sc}
-\begin{tabular}{|c|c|c|c|}
-\hline
-Cluster     & \multicolumn{3}{c|}{I}\\ \hline
-Brain nodes & \multicolumn{3}{l|}{\begin{tabular}[c]{@{}l@{}} l.superiorfrontal(2),  l.superiortemporal(2),    l.insula ,             l.frontalpole,         \\
-l.caudalmiddlefrontal, l.parstriangularis,     l.parsopercularis,     l.precentral,          \\
-l.temporalpole,        l.postcentral,
-l.superiorparietal,    l.inferiorparietal,    \\ l.lateraloccipital,    l.medialorbitofrontal,
-l.superiorfrontal,       \\ l.precuneus,           l.cuneus,
-l.parahippocampal,     l.lingual,              l.superiortemporal,\\  l.isthmuscingulate,   l.lateraloccipital
-\end{tabular}} \\
-\hline
-Cluster & \multicolumn{3}{c|}{II}\\ \hline
-Brain nodes & \multicolumn{3}{l|}{\begin{tabular}[c]{@{}l@{}}
-r.superiorfrontal(2),  r.superiortemporal(2),  r.insula,              r.frontalpole,         \\
-r.caudalmiddlefrontal, r.parstriangularis,     r.parsopercularis,     r.precentral,          \\
-r.temporalpole,        r.postcentral,
-r.superiorparietal,    r.inferiorparietal,    \\ r.lateraloccipital,   r.medialorbitofrontal,
-r.superiorfrontal,       \\ r.precuneus,
-r.cuneus,
-r.parahippocampal,     r.lingual,              r.superiortemporal,    \\r.isthmuscingulate,  r.lateraloccipital\end{tabular}} \\ \hline
-Cluster      & III    & IV    & V
-\\ \hline
-Brain nodes & \begin{tabular}[c]{@{}c@{}}l.supramarginal(4)\end{tabular}&
-\begin{tabular}[c]{@{}c@{}}r.supramarginal(4)\end{tabular}&
-\begin{tabular}[c]{@{}c@{}}l.inferiortemporal(3)\end{tabular}\\
-\hline
-Cluster & VI  & VII & VIII
-\\ \hline
-Brain nodes &
-\begin{tabular}[c]{@{}c@{}}r.inferiortemporal(3)\end{tabular}&
-\begin{tabular}[c]{@{}c@{}}l.middletemporal(3)\end{tabular}& \begin{tabular}[c]{@{}c@{}}r.middletemporal(3)\end{tabular}
-\\
-\hline
-\end{tabular}
-\end{sc}
-\end{small}
-\caption{Clustering result of brain nodes.
-The first alphabet in the node name indicates the left (L) or right (R) hemisphere.
-The number in the parentheses indicates the node count in each cluster.
-}  \label{table:clustering}
-\end{table}
-
 \bibliography{tensor_wang}

### 2020-02-18 ICML/submit/supplements/latexsource/ordinalTv3.tex
Add files via upload

--- 2020-02-13 ICML/arxiv/old_file/ordinalTv3.tex
@@
 \begin{icmlauthorlist}
-\icmlauthor{Chanwoo Lee}{equal,to}
-\icmlauthor{Miaoyan Wang}{equal,to}
+\icmlauthor{}{}
 \end{icmlauthorlist}
 
-\icmlaffiliation{to}{Department of Statistics, University of Wisconsin at Madison, USA}
-
-\icmlcorrespondingauthor{Miaoyan Wang}{miaoyan.wang@wisc.edu.com}
-
+\icmlaffiliation{}{}
+\icmlcorrespondingauthor{}{}
 % You may provide any keywords that you
@@
 The problem of tensor denoising -- which aims to recover a signal tensor from its noisy entries -- has gained increased attention in theory and applications~\citep{xia2019sup,zeng2019multiway}.
-A related problem, tensor completion, examines the minimum number of entries needed for a consistent recovery~\citep{ghadermarzy2018learning,ghadermarzy2019near}.
-Low-rankness is often imposed to the signal tensor, which efficiently reduces the intrinsic dimension in both problems.
+The second problem, tensor completion, examines the minimum number of entries needed for a consistent recovery~\citep{ghadermarzy2018learning,ghadermarzy2019near}.
+Low-rankness is often imposed to the signal tensor, Thereby efficiently reducing the intrinsic dimension in both problems.
 
@@
 
+\vspace{-.2cm}
 \section{Preliminaries}
+\vspace{-.1cm}
 Let $\tY\in\mathbb{R}^{d_1\times \cdots \times d_K}$ denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional tensor.
@@
 \section{Model formulation and motivation}
+\vspace{-.1cm}
 \subsection{Observation model}
+\vspace{-.1cm}
 Let $\tY$ denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional data tensor.
@@
 This interpretation will be made more explicit in the next section.
+\vspace{-.1cm}
+\subsection{Latent-variable interpretation}\label{sec:latent}
+\vspace{-.1cm}
 
-\subsection{Latent-variable interpretation}\label{sec:latent}
 The ordinal tensor model~\eqref{eq:model} with certain types of link $f$ has the equivalent representation as an $L$-level quantization model on $\tY=\entry{y_\omega}$:
@@
 \end{assumption}
-
+\vspace{-.1cm}
 \subsection{Problem 1: Tensor denoising}~\label{sec:denoising}
+\vspace{-.1cm}
 The first question we aim to address is tensor denoising:
@@
 The entry-wise bound assumption is a technical condition that avoids the degeneracy in probability estimation with ordinal observations.
-
+\vspace{-.1cm}
 \subsection{Problem 2: Tensor completion}
+\vspace{-.1cm}
 Motivated by applications in collaborative filtering, we also consider a more general setup when only a subset of tensor entries $y_\omega$ are observed.
@@
 We adopt the same assumption for higher-order tensors.
-
+\vspace{-.2cm}
 \section{Rank-constrained M-estimator}\label{sec:theory}
+\vspace{-.1cm}
 We present a general treatment to both problems mentioned above.
@@
 \begin{align}\label{eq:regular}
-A_\alpha &= \min_{\ell\in[L], |\theta|\leq \alpha} g_\ell(\theta),\quad  U_\alpha=\max_{\ell\in[L], |\theta|\leq \alpha } {\dot{g}_\ell(\theta)\over g_\ell(\theta)},\\
+A_\alpha &= \min_{\ell\in[L], |\theta|\leq \alpha} g_\ell(\theta),\quad  U_\alpha=\max_{\ell\in[L], |\theta|\leq \alpha } {|\dot{g}_\ell(\theta)|\over g_\ell(\theta)},\\
  L_\alpha &= \min_{\ell\in[L], |\theta|\leq \alpha }\left[{\dot{g}_\ell^2 (\theta)\over g_\ell^2(\theta)} -{\ddot{g}_\ell(\theta)\over g_\ell(\theta)}\right],
@@
 Under the Assumption~\ref{ass:link}, all these quantities are strictly positive and independent of tensor dimension.
-
+\vspace{-.1cm}
 \subsection{Estimation error for tensor denoising}\label{sec:denosing}
+\vspace{-.1cm}
 For the tensor denoising problem, we assume that the full set of tensor entries are observed. We assess the estimation accuracy using the mean squared error (MSE):
@@
 \begin{equation}\label{eq:rate}
-\text{MSE}(\hat \Theta, \trueT) \leq \min\left( 4\alpha^2,\ {c_2  U^2_\alpha r_{\max}^{K-1}  \over  L^2_\alpha } {\sum_kd_k\over  \prod_k d_k} \right),
+\mathrm{MSE}(\hat \Theta, \trueT) \leq \min\left( 4\alpha^2,\ {c_1  U^2_\alpha r_{\max}^{K-1}  \over  L^2_\alpha } {\sum_kd_k\over  \prod_k d_k} \right),
 \end{equation}
-where $c_1, c_2>0$ are two constants that depend only on $K$.
+where $c_1 >0$ is a constant that depends only on $K$.
 \end{thm}
@@
 \begin{equation}\label{eq:KLrate}
-\text{KL}(\mathbb{P}_{\tY} || \hat{\mathbb{P}}_{\tY}) \leq  {c_2 U^2_\alpha r_{\max}^{K-1}  \over L^2_\alpha } {(4L-6)\dot{f}^2(0)  \over A_\alpha} {\sum_kd_k\over  \prod_k d_k},
+\text{KL}(\mathbb{P}_{\tY} || \hat{\mathbb{P}}_{\tY}) \leq  {c_1 U^2_\alpha r_{\max}^{K-1}  \over L^2_\alpha } {(4L-6)\dot{f}^2(0)  \over A_\alpha} {\sum_kd_k\over  \prod_k d_k},
 \end{equation}
-where $c_1, c_2>0$ are the same constants as in Theorem~\ref{thm:rate}.
+where $c_1 >0$ is the same constants as in Theorem~\ref{thm:rate}.
 \end{cor}
 To gain insight into these bounds, we consider a special setting with equal dimension in all modes, i.e., $d_1=\cdots=d_K=d$. In such a case, our bound \eqref{eq:rate} reduces to
-\begin{equation*}\label{eq:ours}
+\begin{equation}\label{eq:ours}
 \text{MSE}(\hat \Theta, \trueT) \asymp d^{-(K-1)}, \quad \text{as}\ d\to \infty.
-\end{equation*}
+\end{equation}
 Hence, our estimator achieves consistency with polynomial convergence rate.
@@
 \inf_{\hat \Theta }\sup_{\trueT\in\tP} \mathbb{P}\Big\{& \text{MSE}(\hat \Theta, \trueT) \\
-& \geq {1\over 256} \min\left( \alpha^2, \ { Cr_{\max}d_{\max} \over \prod_k d_k} \right) \Big\} \geq {1\over 8},
+& \geq c\min\left( \alpha^2, \ { Cr_{\max}d_{\max} \over \prod_k d_k} \right) \Big\} \geq {1\over 8},
 \end{align}
-where $C=C(\alpha, L, f,\mb)>0$ is a constant independent of tensor dimension and the rank.
+where $C=C(\alpha, L, f,\mb)>0$ and $c>0$ are constants independent of tensor dimension and the rank.
 \end{thm}
@@
 \section{Numerical Implementation}\label{sec:algorithm}
+\vspace{-.1cm}
 We describe the algorithm to seek the optimizer of~\eqref{eq:objective}.
@@
 Based on the estimated Tucker factors $\{\hat \mM_k\}$, we perform a clustering analysis via K-mean on the brain nodes (see detailed procedure in the Supplement).
-We find that the 68 brain nodes are grouped into eight clusters, and the clustering capture the spatial separation of brain nodes very well (see the Supplement).
-In particular, the top two clusters represent the left and right hemispheres; the smaller clusters represent local regions driving by similar nodes.
-For example, the cluster IV consists of brain nodes {\it postcentral}, {\it posteriorcingulate}, {\it pericalcarine} and {\it temporalpole}, all of which are located at {\it SupraM}.
+We find that the 68 brain nodes are grouped into 11 clusters, and the clustering captures the spatial separation of brain nodes very well (see the Supplement).
+In particular, the top three clusters represent the left, right and connection between two hemispheres; the smaller clusters represent local regions driving by similar nodes.
+For example, the cluster IV consists of nodes in the left supramarginal gyrus region.
 The identified similarity among brain nodes suggests the applicability of our method.
 
-We compare the goodness-of-fit of various tensor methods on the HPC data.
+We also compare the goodness-of-fit of various tensor methods on the HPC data.
 Table~\ref{table:CV} summarizes the prediction error via 5-fold stratified cross-validation averaged over 10 runs.
@@
 \newpage
-\bibliography{tensor_wang}
+\bibliography{ordinalTv3.bbl}
 \bibliographystyle{icml2020}

### 2020-02-18 ICML/submit/supplements/latexsource_supp/ordinalT_supp.tex
Add files via upload

--- 2020-02-16 ICML/arxiv/ordinalT_arxiv_v4.tex
@@
 \usepackage{fancybox}
+
 \usepackage{color}
@@
 \usepackage[margin=1in]{geometry}
-\newcommand{\theHalgorithm}{\arabic{algorithm}}
+
+\renewcommand{\textfraction}{0.0}
+\renewcommand{\topfraction}{1.0}
+%\renewcommand{\textfloatsep}{5mm}
+
 \usepackage{comment}
+% Definitions of handy macros can go here
 \usepackage{amsmath,amssymb,amsthm,bm,mathtools}
+\usepackage{multirow}
 \usepackage{natbib}
+%\usepackage{dsfont,multirow,hyperref,setspace,natbib,enumerate}
 \usepackage{dsfont,multirow,hyperref,setspace,enumerate}
 \hypersetup{colorlinks,linkcolor={blue},citecolor={blue},urlcolor={red}}
-\usepackage{algorithm}
-\usepackage{algorithmic}
-\usepackage{makecell}
-
-\setlength\parindent{0pt}
-\usepackage[parfill]{parskip}
+\usepackage{algpseudocode,algorithm}
+\algnewcommand\algorithmicinput{\textbf{Input:}}
+\algnewcommand\algorithmicoutput{\textbf{Output:}}
+\algnewcommand\INPUT{\item[\algorithmicinput]}
+\algnewcommand\OUTPUT{\item[\algorithmicoutput]}
+
+\mathtoolsset{showonlyrefs=true}
 
@@
 \newtheorem{assumption}{Assumption}
-\newtheorem{cor}{Corollary}[section]
 
@@
 \newtheorem{defn}{Definition}
+\newtheorem{cor}{Corollary}
 \newtheorem{example}{Example}
@@
 
-\usepackage{pifont}
-\newcommand{\cmark}{\ding{51}}%
-\newcommand{\xmark}{\ding{55}}%
+\renewcommand{\thefigure}{{S\arabic{figure}}}%
+\renewcommand{\thetable}{{S\arabic{table}}}%
+\renewcommand{\figurename}{{Supplementary Figure}}
+\renewcommand{\tablename}{{Supplementary Table}}
+\setcounter{figure}{0}
+\setcounter{table}{0}
+
+\def\MLET{\hat \Theta_{\text{MLE}}}
+\newcommand{\cmt}[1]{{\leavevmode\color{red}{#1}}}
 
 \usepackage{dsfont}
+
+\usepackage{multirow}
+
+\DeclareMathOperator*{\minimize}{minimize}
+
 \usepackage{wrapfig}
+
 \usepackage{mathtools}
 \mathtoolsset{showonlyrefs}
-
 \newcommand*{\KeepStyleUnderBrace}[1]{%f
@@
 }
-
-\begingroup
-\makeatletter
-\@for\theoremstyle:=definition,remark,plain\do{%
-\expandafter\g@addto@macro\csname th@\theoremstyle\endcsname{%
-\addtolength\thm@preskip\parskip
-}%
-}
-\endgroup
-
+\usepackage{xr}
+\externaldocument{ordinalTv3}
 \input macros.tex
-\allowdisplaybreaks
-
-\title{Tensor denoising and completion based on ordinal observations}
-\date{}
-\author{%
-Chanwoo Lee \\
-University of Wisconsin -- Madison\\
-\texttt{chanwoo.lee@wisc.edu} \\
-\and
-Miaoyan Wang \\
-University of Wisconsin -- Madison\\
-\texttt{miaoyan.wang@wisc.edu} \\
-}
+
+\title{Supplements for ``Tensor denoising and completion based on ordinal observations''}
 
 \begin{document}
-\maketitle
-\begin{abstract}
-Higher-order tensors arise frequently in applications such as neuroimaging, recommendation system, social network analysis, and psychological studies.
-We consider the problem of low-rank tensor estimation from possibly incomplete, ordinal-valued observations.
-Two related problems are studied, one on tensor denoising and another on tensor completion.
-We propose a multi-linear cumulative link model, develop a rank-constrained M-estimator, and obtain theoretical accuracy guarantees.
-Our mean squared error bound enjoys a faster convergence rate than previous results, and we show that the proposed estimator is minimax optimal under the class of low-rank models.
-Furthermore, the procedure developed serves as an efficient completion method which guarantees consistent recovery of an order-$K$ $(d,\ldots,d)$-dimensional low-rank tensor using only $\tilde \tO(Kd)$ noisy, quantized observations.
-We demonstrate the outperformance of our approach over previous methods on the tasks of clustering and collaborative filtering.
-\\[0.5cm]
-\textbf{Keywords:} Higher-order tensors, ordinal observation, tensor decomposition, tensor completion
-\end{abstract}
-
-\section{Introduction}
-Multidimensional arrays, a.k.a.\ tensors, arise in a variety of applications including recommendation systems~\citep{baltrunas2011incarmusic}, social networks~\citep{nickel2011three}, genomics~\citep{hore2016tensor}, and neuroimaging~\citep{zhou2013tensor}.
-There is a growing need to develop general methods that can handle two main problems for analyzing these noisy, high-dimensional datasets.
-The first problem is tensor denoising which aims to recover a signal tensor from its noisy entries~\citep{hong2018generalized,zeng2019multiway}.
-The second problem is tensor completion which examines the minimum number of entries needed for a consistent recovery~\citep{ghadermarzy2018learning,ghadermarzy2019near}.
-Low-rankness is often imposed to the signal tensor, thereby efficiently reducing the intrinsic dimension in both problems.
-
-A number of low-rank tensor estimation methods have been proposed~\citep{anandkumar2014tensor,wang2017tensor}, revitalizing classical methods such as CANDECOMP/PARAFAC (CP) decomposition~\citep{hitchcock1927expression} and Tucker decomposition~\citep{tucker1966some}.
-These tensor methods treat the entries as continuous-valued.
-In many cases, however, we encounter datasets of which the entries are qualitative.
-For example, the Netflix problem records the ratings of users on movies over time.
-Each entry is a rating on a nominal scale \{{\it very like, like, neutral, dislike, very dislike}\}.
-Another example is in the signal processing, where the digits are frequently rounded or truncated so that only integer values are available.
-The qualitative observations take values in a limited set of categories, making the learning problem harder compared to  continuous observations.
-
- Ordinal entries are categorical variables with an ordering among the categories; for example, {\it very like} $\prec$ {\it like} $\prec$ {\it neutral} $\prec \cdots$.
-The analyses of tensors with the ordinal entries are mainly complicated by two key properties needed for a reasonable model.
-First, the model should be invariant under a reversal of categories, say, from the Netflix example, {\it very like} $\succ$ {\it like} $\succ$ {\it neutral} $\succ \cdots$, but not under arbitrary label permutations.
-Second, the parameter interpretations should be consistent under merging or splitting of contiguous categories.
-The classical continuous tensor model~\citep{kolda2009tensor, ghadermarzy2019near} fails in the first aspect, whereas the binary tensor model~\citep{ghadermarzy2018learning} lacks the second property.
-An appropriate model for ordinal tensors has yet to be studied.
-
-{\bf Our contribution.} This paper presents an efficient low-rank estimation method and theory for tensors with ordinal-valued entries.
-Our main contributions are summarized in Table~\ref{tab:compare}.
-We propose a cumulative link model for higher-order tensors, develop a rank-constrained M-estimator, and obtain theoretical accuracy guarantees.
-The mean squared error bound is established, and we show that the obtained bound has minimax optimal rate in high dimensions under the low-rank model.
-Our estimator enjoys a faster convergence rate $\tO(d^{-(K-1)/2})$ than $\tO(d^{-K})$ in~\citet{ghadermarzy2018learning}, which is a substantial improvement for higher-order tensors.
-Furthermore, our proposal serves as an efficient completion algorithm that guarantees consistent recovery of an order-$K$ $(d,\ldots,d)$-dimensional low-rank tensor using only $\tilde \tO(Kd)$ noisy, quantized observations.
-
-\begin{table}[ht]
-\resizebox{\columnwidth}{!}{
-\begin{tabular}{l|ccc}
-&\cite{bhaskar2016probabilistic}&\cite{ghadermarzy2018learning} &This paper\\
-\hline
-Higher-order tensors ($K\geq 3$) & \xmark&\cmark& \cmark\\
-\hline
-Multi-level categories ($L\geq 3$)& \cmark& \xmark&\cmark\\
- \hline
- Error rate for tensor denoising&$d^{-1}$ for $K=2$& $d^{-(K-1)/2}$&$d^{-(K-1)}$\\
-\hline
-Optimality guarantee under low-rank models& unknown & \xmark & \cmark\\
-\hline
-Sample complexity for tensor completion&$d^K$& $Kd$&$Kd$\\
-\hline
-\end{tabular}
-}
-\caption{Comparison with previous work when tensor rank $r=O(1)$.
-For ease of presentation, we summarize the error rate and sample complexity (neglecting log factors) when the tensor dimensions are equal in all modes.
-$K$: tensor order; $L$: number of ordinal levels; $d$: dimension at each mode.
-}~\label{tab:compare}
-\end{table}
-
-{\bf Related work.} Our work is related to, but clearly distinctive from, several lines of existing literature.
-Matrix completion from quantized samples was firstly introduced for binary observations~\citep{cai2013max,davenport2014,bhaskar20151} and then extended to ordinal observations~\citep{bhaskar2016probabilistic}.
-As we show in Section~\ref{sec:theory}, applying existing matrix methods to an ordinal tensor results in a suboptimal estimator with a slower convergence rate.
-Therefore, a full exploitation of the tensor structure is necessary; this is the focus of the current paper.
-
-Our work is also connected to non-Gaussian tensor decomposition.
-Existing work focuses exclusively on univariate observations such as binary- or continuous-valued entries~\citep{wang2018learning,hong2018generalized,ghadermarzy2018learning}.
-As we mentioned earlier, the ordinal observations add considerable challenges to the model formulation.
-We address the problems from two perspectives.
-From statistical perspective, our proposed model generalizes the usual binary tensor model while preserving palindromic invariance~\citep{mccullagh1980regression} for ordinal observations.
-From algorithm perspective, our alternating optimization compares favorably to the approximate (non-convex) algorithm developed in the context of binary tensors~\citep{ghadermarzy2018learning}.
-We numerically compare the two approaches in Section~\ref{sec:experiment}.
-
-\section{Preliminaries}
-Let $\tY\in\mathbb{R}^{d_1\times \cdots \times d_K}$ denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional tensor.
-We use $y_\omega$ to denote the tensor entry indexed by $\omega$, where $\omega\in[d_1]\times\cdots\times[d_K]$.
-The Frobenius norm of $\tY$ is defined as $\FnormSize{}{\tY}=\sum_\omega y^2_\omega$ and the infinity norm of $\tY$ is defined as $\mnormSize{}{\tY}=\max_{\omega}|y_\omega|$.
-We use $\tY_{(k)}$ to denote the unfolded matrix of size $d_k$-by-$\prod_{i\neq k}d_k$, obtained by reshaping the tensor along the mode-$k$.
-The Tucker rank of $\tY$ is defined as a length-$K$ vector $\mr = (r_1,\ldots,r_K)$, where $r_k$ is the rank of matrix $\tY_{(k)}$ for all $k \in[K]$.
-We say that an event $A$ occurs ``with very high probability'' if $\mathbb{P}(A)$ tends to 1 faster than any polynomial of tensor dimension $d_{\min}=\min\{d_1,\ldots,d_K\} \to\infty$.
-For any two functions $f, g$ depending on $(d_1,\ldots,d_K)$, we write $f=\tO(g)$ to indicate that $f\leq Cg$, where $C>0$ is a constant independent of tensor dimension.
-We write $f=\tilde\tO(g)$ to indicate that $f\leq C(\log d_{\min})^\beta g$ for some $\beta>0$.
-
-We use lower-case letters ($a, b, c, \ldots$) for scalars/vectors, upper-case boldface letters ($\mA, \mB, \mC, \ldots$) for matrices, and calligraphy letters ($\tA,\tB,\tC,\ldots$) for tensors of order three or greater.
-For ease of notation, we allow the basic arithmetic operators (e.g., $\leq, +, -$) to be applied to pairs of tensors in an element-wise manner.
-We use the shorthand $[n]$ to denote the $n$-set $\{1,\ldots,n\}$ for $n \in N_{+}$.
-
-\section{Model formulation and motivation}
-\subsection{Observation model}
-Let $\tY$ denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional data tensor.
-Suppose the entries of $\tY$ are ordinal-valued, and the observation space consists of $L$ ordered levels, denoted by $[L]:=\{1,\ldots,L\}$.
-We propose a cumulative link model for the ordinal tensor $\tY=\entry{y_\omega}\in[L]^{d_1\times \cdots\times d_K}$.
-Specifically, assume the entries $y_\omega$ are (conditionally) independently distributed with cumulative probabilities:
-\begin{equation}\label{eq:model}
-\mathbb{P}(y_\omega\leq \ell)=f(b_\ell-\theta_\omega),\ \text{for all}\ \ell\in[L-1],
-\end{equation}
-where $\mb=(b_1,\ldots,b_{L-1})$ is a set of unknown scalars satisfying $b_1<\cdots <b_{L-1}$, $\Theta=\entry{\theta_\omega}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is a continuous-valued parameter tensor satisfying certain low-dimensional structure (to be specified later), and $f(\cdot):\mathbb{R}\mapsto[0,1]$ is a known, strictly increasing function.
-We refer to $\mb$ as the cut-off points and $f$ the link function.
-
-The formulation~\eqref{eq:model} imposes an additive model to the transformed probability of cumulative categories.
-This modeling choice is to respect the ordering structure among the categories.
-For example, if we choose the inverse link $f^{-1}(x)=\log {x\over 1-x}$ to be the log odds, then the model~\eqref{eq:model} implies linear spacing between the proportional odds:
-\begin{equation}\label{eq:logodd}
-\log {\mathbb{P}(y_{\omega}\leq \ell) \over \mathbb{P}(y_\omega >  \ell) } - \log {\mathbb{P}(y_{\omega}\leq {\ell-1}) \over \mathbb{P}(y_\omega >  {\ell-1}) } = b_\ell-b_{\ell-1},
-\end{equation}
-for all tensor entries $y_\omega$.
-When there are only two categories in the observation space (e.g.
-binary tensors), the cumulative model~\eqref{eq:model} is equivalent to the usual multinomial link model.
-In general, however, when the number of categories $L\geq 3$, the proportional odds assumption~\eqref{eq:logodd} is more parsimonious, in that, the ordered categories can be envisaged as contiguous intervals on the continuous scale, where the points of division are exactly $b_1<\cdots <b_{L-1}$.
-This interpretation will be made more explicit in the next section.
-
-\subsection{Latent-variable interpretation}\label{sec:latent}
-The ordinal tensor model~\eqref{eq:model} with certain types of link $f$ has the equivalent representation as an $L$-level quantization model on $\tY=\entry{y_\omega}$:
-\begin{align}\label{eq:quantization}
-y_\omega&=
-\begin{cases}
-1,& \text{if $y^*_\omega\in(-\infty, b_1]$},\\
-2,& \text{if $y^*_\omega\in(b_1, b_2]$},\\
-\vdots  &\vdots\\
-L,& \text{if $y^*_\omega\in(b_{L-1}, \infty)$},\\
-\end{cases}
-\end{align}
-for all $\omega\in[d_1]\times \cdots \times [d_k]$. Here, $\tY^*=\entry{y^*_\omega}$ is a latent continuous-valued tensor following an additive noise model:
-\begin{align}\label{eq:latent}
-\KeepStyleUnderBrace{\tY^*}_{\text{latent continuous-valued tensor}}=\KeepStyleUnderBrace{\Theta}_{\text{signal tensor}}+\KeepStyleUnderBrace{\tE}_{\text{i.i.d.\ noise}},
-\end{align}
-where $\tE=\entry{\varepsilon_\omega}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is a noise tensor with i.i.d.\ entries according to distribution $\mathbb{P}(\varepsilon)$.
-From the viewpoint of~\eqref{eq:latent}, the parameter tensor $\Theta$ can be interpreted as the latent signal tensor prior to contamination and quantization.
-
-The equivalence between the latent-variable model~\eqref{eq:quantization} and the cumulative link model~\eqref{eq:model} is established if the link $f$ is chosen to be the cumulative distribution function of noise $\varepsilon$, i.e., $f(\theta)=\mathbb{P}(\varepsilon\leq \theta)$.
-We describe two common choices of link $f$, or equivalently, the distribution of $\varepsilon$.
-
-\begin{example}[Logistic model] The logistic model is characterized by~\eqref{eq:model} with $f(\theta)=(1+e^{-\theta/\sigma})^{-1}$, where $\sigma>0$ is the scale parameter.
-Equivalently, the noise $\varepsilon_\omega$ in~\eqref{eq:quantization} follows i.i.d.\ logistic distribution with scale parameter $\sigma$.
-\end{example}
-
-\begin{example}[Probit model] The probit model is characterized by~\eqref{eq:model} with
-$f(\theta)=\mathbb{P}(z\leq \theta/\sigma)$, where $z\sim N(0,1)$. Equivalently, the noise $\varepsilon_\omega$ in~\eqref{eq:quantization} follows i.i.d.\ $N(0,\sigma^2)$.
-\end{example}
-
-Other link functions are also possible, such as Laplace, Cauchy, inverse log-log, etc~\citep{mccullagh1980regression}.
-All the models share the property that the ordered categories can be thought of as contiguous interval on some continuous scale.
-We should point out that, although the latent-variable interpretation is incisive, our estimation procedure does not refer to the existence of $\tY^*$.
-Therefore, our model~\eqref{eq:model} is general and still valid in the absence of quantization process.
-More generally, we make the following assumptions about the link $f(\cdot): \mathbb{R}\mapsto[0,1]$.
-
-\begin{assumption}\label{ass:link}
-The link function is assumed to satisfy:
+
+\begin{center}
+\begin{spacing}{1.5}
+\textbf{\Large Supplements for ``Tensor denoising and completion based on ordinal observations''}
+\end{spacing}
+\end{center}
+
+\section{Proofs}\label{sec:proof}
+Here, we provide proofs of the theoretical results presented in Sections~\ref{sec:theory}.
+
+\subsection{Estimation error for tensor denoising}\label{sec:proofMSE}
+\begin{proof}[Proof of Theorem~\ref{thm:rate}]
+We suppress the subscript $\Omega$ in the proof, because the tensor denoising assumes complete observation $\Omega=[d_1]\times \cdots \times [d_K]$.
+It follows from the expression of $\flogl(\Theta)$ that
+\begin{align}\label{eq:property}
+{\partial \flogl\over \partial \theta_\omega}&=\sum_{\ell\in[L]}\mathds{1}_{\{y_{\omega}=\ell\}}
+{\dot{g}_\ell(\theta_\omega)\over g_\ell(\theta_\omega)},\notag\\
+{\partial^2 \flogl\over \partial \theta_\omega^2}&=\sum_{\ell\in[L]}\mathds{1}_{\{y_\omega=\ell\}}{\ddot{g}_\ell(\theta_\omega)g_\ell(\theta_\omega)-\dot{g}^2_\ell(\theta_\omega)\over g^2_\ell(\theta_\omega)}\ \text{and}\quad
+{\partial^2 \flogl\over \partial \theta_\omega \theta_\omega'}=0\ \text{if}\ \omega\neq \omega',
+\end{align}
+for all $\omega\in[d_1]\times \cdots \times [d_K]$.
+Define $d_{\text{total}}=\prod_k d_k$.
+Let $\fplogl\in\mathbb{R}^{d_1\times\cdots\times d_K}$ denote the tensor of gradient with respect to $\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}$, and $\fpplogl$ the corresponding Hession matrix of size $d_\text{total}$-by-$d_{\text{total}}$.
+Here, $\Vec(\cdot)$ denotes the operation that turns a tensor into a vector.
+By~\eqref{eq:property}, $\fpplogl$ is a diagonal matrix.
+Recall that
+\begin{equation}\label{eq:bound}
+U_\alpha=\max_{\ell\in[L],|\theta|\leq \alpha}{|\dot{g}_\ell(\theta)|\over g_\ell(\theta)}>0 \quad \text{and}\quad
+L_\alpha=\min_{\ell\in[L],|\theta|\leq \alpha} {\dot{g}^2_\ell(\theta)-\ddot{g}_\ell(\theta)g_\ell(\theta)\over g^2_\ell(\theta)}>0.
+\end{equation}
+Therefore, the entries in $\fplogl$ are upper bounded in magnitude by $U_\alpha>0$, and all diagonal entries in $\fpplogl$ are upper bounded by $-L_{\alpha}<0$.
+
+By the second-order Taylor's expansion of $\flogl(\Theta)$ around $\trueT$, we obtain
+\begin{equation}\label{eq:taylor}
+\flogl(\Theta)=\flogl(\trueT)+\langle\Vec(\fplogl),\ \Vec(\Theta-\trueT)\rangle+{1\over 2}\Vec(\Theta-\trueT)^T\fpplogl(\check\Theta)\Vec(\Theta-\trueT),
+\end{equation}
+$\check\Theta=\gamma\trueT+(1-\gamma)\Theta$ for some $\gamma\in[0,1]$, and $\fpplogl(\check\Theta)$ denotes the $\prod_kd_k$-by-$\prod_k d_k$ Hession matrix evaluated at $\check\Theta$.
+
+We first bound the linear term in~\eqref{eq:taylor}. Note that, by Lemma~\ref{lem:inq},
+\begin{equation}\label{eq:linear}
+|\langle\Vec(\fplogl(\trueT), \Vec(\Theta-\trueT)  \rangle|\leq \snormSize{}{\fplogl(\trueT)} \nnormSize{}{\Theta-\trueT},
+\end{equation}
+where $\snormSize{}{\cdot}$ denotes the tensor spectral norm and $\nnormSize{}{\cdot}$ denotes the tensor nuclear norm. Define
+\[
+s_\omega={\partial \tL_\tY\over \partial \theta_\omega}\Big|_{\Theta=\trueT} \;\; \textrm{ for all } \; \omega\in[d_1]\times\cdots\times [d_K].
+\]
+Based on~\eqref{eq:property} and the definition of $U_\alpha$, $\fplogl(\trueT)=\entry{s_{\omega}}$ is a random tensor whose entries are independently distributed satisfying
+\begin{equation}\label{eq:norm}
+\mathbb{E}(s_\omega)=0,\quad |s_\omega|\leq U_\alpha, \quad \text{for all }\omega\in[d_1]\times \cdots \times [d_K].
+\end{equation}
+By lemma~\ref{lem:noisytensor}, with probability at least $1-\exp(-C_1 \sum_kd_k)$, we have
+\begin{equation}\label{eq:normrandom}
+\snormSize{}{\fplogl(\trueT)} \leq C_2 U_\alpha\sqrt{\sum_k d_k},
+\end{equation}
+where $C_1, C_2$ are two positive constants that depend only on $K$.
+Furthermore, note that $\text{rank}(\Theta)\leq \mr$, $\text{rank}(\trueT)\leq \mr$, so $\text{rank}(\Theta-\trueT)\leq 2\mr$.
+By lemma~\ref{lem:nuclear}, $\nnormSize{}{\Theta-\trueT}\leq (2r_{\max})^{K-1\over 2}\FnormSize{}{\Theta-\trueT}$.
+Combining~\eqref{eq:linear}, \eqref{eq:norm} and \eqref{eq:normrandom}, we have that, with probability at least $1-\exp(-C_1 \sum_kd_k)$,
+\begin{equation}\label{eq:linearconclusion}
+|\langle \Vec(\fplogl(\trueT)), \Vec(\Theta-\trueT)  \rangle | \leq C_2 U_\alpha  \sqrt{r_{\max}^{K-1} \sum_k d_k}  \FnormSize{}{\Theta-\trueT}.
+\end{equation}
+
+We next bound the quadratic term in \eqref{eq:taylor}. Note that
+\begin{align}\label{eq:quadratic}
+ \Vec(\Theta-\trueT)^T \fpplogl(\check{\Theta})\Vec(\Theta-\trueT)&=\sum_\omega \left( {\partial^2\tL_{\tY}\over \partial \theta^2_\omega} \Big|_{\Theta=\check\Theta} \right)(\theta_\omega-\theta_{{\text{true}},\omega})^2 \nonumber \\
+&\leq - L_\alpha\sum_{\omega}(\Theta_{\omega}-\Theta_{\text{true},\omega})^2 \nonumber \\
+&=-L_\alpha\FnormSize{}{\Theta-\trueT}^2,
+\end{align}
+where the second line comes from the fact that  $\mnormSize{}{\check\Theta}\leq \alpha$ and the definition of $L_\alpha$.
+
+Combining~\eqref{eq:taylor}, \eqref{eq:linearconclusion} and~\eqref{eq:quadratic}, we have that, for all $\Theta\in\tP$, with probability at least $1-\exp(-C_1 \sum_kd_k)$,
+\[
+\tL_\tY(\Theta)\leq \tL_{\tY}(\trueT)+C_2U_\alpha  \left(r_{\max}^{K-1}\sum_k d_k\right)^{1/2}  \FnormSize{}{\Theta-\trueT}-{L_\alpha\over 2}\FnormSize{}{\Theta-\trueT}^2.
+\]
+In particular, the above inequality also holds for $\hat \Theta\in\tP$. Therefore,
+\[
+\tL_\tY(\hat \Theta)\leq \tL_{\tY}(\trueT)+C_2U_\alpha \left(r_{\max}^{K-1}\sum_k d_k\right)^{1/2}  \FnormSize{}{\hat \Theta-\trueT}-{L_\alpha\over 2} \FnormSize{}{\hat \Theta-\trueT}^2.
+\]
+Since $\hat \Theta=\arg\max_{\Theta\in\tP}\tL_\tY(\Theta)$, $\tL_\tY(\hat \Theta)-\tL_{\tY}(\trueT)\geq 0$, which gives
+\[
+C_2U_\alpha \left(r_{\max}^{K-1}\sum_k d_k\right)^{1/2}  \FnormSize{}{\hat \Theta-\trueT}-{L_\alpha\over 2}\FnormSize{}{\hat \Theta-\trueT}^2\geq 0.
+\]
+Henceforth,
+\[
+{1\over \sqrt{\prod_k d_k}} \FnormSize{}{\hat \Theta-\trueT}\leq {2C_2U_\alpha \sqrt{r_{\max}^{K-1}\sum_k d_k}\over L_\alpha \sqrt{\prod_k d_k}}={2C_2U_\alpha r_{\max}^{(K-1)/2}\over L_\alpha} \sqrt{ \sum_k d_k \over \prod_k d_k}.
+\]
+This completes the proof.
+\end{proof}
+
+\begin{proof}[Proof of Corollary~\ref{cor:prediction}]
+The result follows immediately from Theorem~\ref{thm:rate} and Lemma~\ref{lem:KL}.
+\end{proof}
+
+\begin{proof}[Proof of Theorem~\ref{thm:minimax}]
+
+Let $d_{\text{total}}=\prod_{k\in[K]}d_k$, and $\gamma\in[0,1]$ be a constant to be specified later.
+Our strategy is to construct a finite set of tensors $\tX=\{\Theta_i\colon i=1,\ldots \}\subset \tP$ satisfying the properties of (i)-(iv) in Lemma~\ref{lem:construction}.
+By Lemma~\ref{lem:construction}, such a subset of tensors exist.
+For any tensor  $\Theta\in\tX$, let $\mathbb{P}_{\Theta}$ denote the distribution of $\tY|\Theta$, where $\tY$ is the ordinal tensor.
+In particular, $\mathbb{P}_{\mathbf{0}}$ is the distribution of $\tY$ induced by the zero parameter tensor $\mathbf{0}$, i.e., the distribution of $\tY$ conditional on the parameter tensor $\Theta=\mathbf{0}$.
+Based on the Remark for Lemma~\ref{lem:KL}, we have
+\begin{equation}\label{eq:KLbound1}
+\mathrm{KL}(\mathbb{P}_{\Theta}|| \mathbb{P}_{\mathbf{0}})\leq C \FnormSize{}{\Theta}^2,
+\end{equation}
+where $C={(4L-6) \dot{f}^2(0)\over  A_\alpha}>0$ is a constant independent of the tensor dimension and rank.
+Combining the inequality~\eqref{eq:KLbound1} with property (iii) of $\tX$, we have
+\begin{equation}\label{eq:KLbound}
+\text{KL}(\mathbb{P}_{\Theta}||\mathbb{P}_{\mathbf{0}})\leq \gamma^2 R_{\max} d_{\max}.
+\end{equation}
+From~\eqref{eq:KLbound} and the property (i), we deduce that the condition
+\begin{equation}\label{eq:totalKL}
+{1\over \text{Card}(\tX)-1}\sum_{\Theta \in\tX}\text{KL}(\mathbb{P}_{\Theta}, \mathbb{P}_{\mathbf{0}})\leq \varepsilon \log_2\left\{\text{Card}(\tX)-1 \right\}
+\end{equation}
+holds for any $ \varepsilon \geq 0$ when $\gamma\in[0,1]$ is chosen to be sufficiently small depending on $\varepsilon$, e.g., $\gamma \leq \sqrt{\varepsilon\log2\over8}$.
+By applying Lemma~\ref{lem:Tsybakov} to~\eqref{eq:totalKL}, and in view of the property (iv), we obtain that
+\begin{equation}\label{eq:final}
+\inf_{\hat \Theta}\sup_{\trueT\in \tX}\mathbb{P}\left(\FnormSize{}{\hat \Theta- \trueT}\geq  {\gamma\over 8} \min\left\{ \alpha\sqrt{d_{\text{total}}}, C^{-1/2}\sqrt{ R_{\max}d_{\max}}\right\} \right)\geq {1\over 2}\left(1-2\varepsilon-\sqrt{16 \varepsilon \over R_{\max}d_{\max}\log2}\right).
+\end{equation}
+Note that $\text{Loss}(\hat \Theta, \trueT)=\FnormSize{}{\hat \Theta- \trueT}^2/d_{\text{total}}$ and $\tX\subset \tP$.
+By taking $\varepsilon=1/10$ and $\gamma=1/11$, we conclude from~\eqref{eq:final} that
+\begin{equation}\label{eq:prob}
+\inf_{\hat \Theta}\sup_{\trueT\in \tP}\mathbb{P}\left(\text{Loss}(\hat \Theta, \trueT)\geq c\min\left \{ \alpha^2,  {C^{-1}R_{\max}d_{\max}\over d_{\text{total}}}\right \}\right)\geq {1\over 2}\left({4\over 5}- \sqrt{1.6\over R_{\max}d_{\max}\log2} \right)\geq {1\over 8},
+\end{equation}
+where $c = {1 \over 88^2}$ and the last inequality comes from the condition for $d_{\text{max}}$.
+This completes the proof.
+\end{proof}
+
+\subsection{Sample complexity for tensor completion}
+
+\begin{proof}[Proof of Theorem~\ref{thm:completion}]
+
+For notational convenience, we use $\MFnormSize{}{\Theta}=\sum_{\omega\in\Omega}\Theta^2_\omega$ to denote the sum of squared entries over the observed set $\Omega$, for a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$.
+
+Following a similar argument as in the proof of Theorem~\ref{thm:rate}, we have
+\begin{equation}\label{eq:Taylor2}
+\logl(\Theta)=\logl(\trueT)+\langle\Vec(\plogl),\ \Vec(\Theta-\trueT)\rangle+{1\over 2}\Vec(\Theta-\trueT)^T\pplogl(\check\Theta)\Vec(\Theta-\trueT),
+\end{equation}
+where
 \begin{enumerate}
-\item $f(\theta)$ is strictly increasing and twice-differentiable in $\theta$.
-\item $f'(\theta)$ is strictly log-concave and symmetric with respect to $\theta=0$.
+\item $\plogl$ is a $d_1\times\cdots\times d_K$ tensor with $|\Omega|$ nonzero entries, and each entry is upper bounded by $U_\alpha>0$.
+\item $\pplogl$ is a diagonal matrix of size $d_{\text{total}}$-by-$d_{\text{total}}$ with $|\Omega|$ nonzero entries, and each entry is upper bounded by $-L_{\alpha}<0$.
+\end{enumerate}
+
+Similar to~\eqref{eq:linear} and~\eqref{eq:quadratic}, we have
+\begin{equation}\label{eq:linear2}
+|\langle\Vec(\plogl),\ \Vec(\Theta-\trueT)\rangle|\leq C_2U_\alpha \sqrt{r_{\max}^{K-1}\sum_k d_k}\MFnormSize{}{\Theta-\trueT}
+\end{equation}
+and
+\begin{equation}\label{eq:quadratic2}
+\Vec(\Theta-\trueT)^T\fpplogl(\check\Theta)\Vec(\Theta-\trueT)\leq -L_\alpha \MFnormSize{}{\Theta-\trueT}^2.
+\end{equation}
+
+Combining~\eqref{eq:Taylor2}-\eqref{eq:quadratic2} with the fact that $\logl(\hat \Theta)\geq \logl(\trueT)$, we have
+\begin{equation}\label{eq:sample}
+\MFnormSize{}{\hat \Theta-\trueT}\leq {2C_2U_\alpha  r_{\max}^{(K-1)/2} \over L_\alpha} \sqrt{\sum_kd_k},
+\end{equation}
+with probability at least $1-\exp(-C_1 \sum_k d_k)$.
+Lastly, we invoke the result regarding the closeness of $\Theta$ to its sampled version $\Theta_{\Omega}$, under the entrywise bound condition.
+Note that $\mnormSize{}{\hat\Theta-\trueT}\leq 2\alpha$ and $\text{rank}(\hat \Theta-\trueT)\leq 2\mr$.
+By Lemma~\ref{lem:Mnormbound}, $\anormSize{}{\hat \Theta-\trueT}\leq 2^{(3K-1)/2} \alpha \left({\prod r_k \over r_{\max}}\right)^{3/2}$.
+Therefore, the condition in Lemma~\ref{lem:convexity} holds with $\beta=2^{(3K-1)/2}\alpha \left({\prod r_k \over r_{\max}}\right)^{3/2}$.
+Applying Lemma~\ref{lem:convexity} to~\eqref{eq:sample} gives
+\begin{align}
+ \PiFnormSize{}{\hat \Theta-\trueT}^2&\leq {1\over m}\MFnormSize{}{\hat \Theta-\trueT}^2+c\beta\sqrt{\sum_k d_k\over |\Omega|}\\
+ &\leq {C_2  r^{K-1}_{\max}} {\sum_k d_k \over |\Omega|}+C_1 \alpha r_{\max}^{3(K-1)/2}\sqrt{\sum_kd_k\over |\Omega|},
+\end{align}
+with probability at least $1-\exp(-{\sum_kd_k\over \sum_k \log d_k})$ over the sampled set $\Omega$. Here $C_1, C_2>0$ are two constants independent of the tensor dimension and rank. Therefore,
+\[
+ \PiFnormSize{}{\hat \Theta-\trueT}^2\to 0,\quad \text{as}\quad {|\Omega|\over \sum_kd_k}\to \infty,
+\]
+provided that $r_{\max}=O(1)$.
+\end{proof}
+
+\subsection{Convexity of the log-likelihood function}\label{sec:proofconvexity}
+\begin{thm}\label{thm:convexity}
+Define the function
+\begin{align}\label{eq:function}
+ \logl(\Theta, \mb)&=\sum_{\omega\in\Omega}\sum_{\ell\in[L]} \left\{\mathds{1}_{\{y_\omega=\ell\}} \log \big[f(b_\ell-\theta_\omega)-  f(b_{\ell-1}-\theta_\omega)\big]\right\},
+ \end{align}
+where $f(\cdot)$ satisfies Assumption~\ref{ass:link}. Then, $\logl(\Theta, \mb)$ is concave in $(\Theta,\mb)$.
+\end{thm}
+
+\begin{proof}
+Define $d_{\text{total}}=\prod_k d_k$.
+By abuse of notation, we use $(\Theta,\mb)$ to denote the length-$(d_{\text{total}}+L-1)$-vector collecting all parameters together.
+Let us denote a bivariate function
+\begin{align}
+\lambda: &\mathbb{R}^2\mapsto \mathbb{R}\\
+(u,v)&\mapsto\lambda(u,v) = \log \big[f(u)-  f(v)\big].
+\end{align}
+It suffices to show that $\lambda(u,v)$ is concave in $(u,v)$ where $u>v$.
+
+Suppose that the claim holds (which we will prove in the next paragraph). Based on~\eqref{eq:function}, $u,v$ are both linear functions of $(\Theta,\mb)$:
+\[
+u = \ma_1^T(\Theta,\mb), \quad v = \ma_2^T(\Theta,\mb), \quad \text{ for some}\ \ma_1,\ma_2\in \mathbb{R}^{d_{\text{total}}+L-1}.
+\]
+Then, $\lambda(u,v) = \lambda(\ma_1^T(\Theta,\mb),\ \ma_2^T(\Theta,\mb))$ is concave in $(\Theta,\mb)$ by the definition of concavity.
+Therefore, we can conclude that $\logl(\Theta, \mb) $ is concave in $(\Theta,\mb)$ because $\logl(\Theta, \mb)$ is the sum of $\lambda(u,v)$.
+
+Now, we prove the concavity of $\lambda(u,v)$. Note that
+\begin{equation}
+\lambda(u,v) = \log\big[f(u)-f(v)\big]=\log\big[\int\mathds{1}_{[u,v]}(x)f'(x)dx\big],
+\end{equation}
+where $\mathds{1}_{[u,v]}$ is an indicator function that equals 1 in the interval $[u,v]$, and 0 elsewhere.
+Furthermore, $\mathds{1}_{[u,v]}(x)$ is log-concave in $(u,v,x)$, and by Assumption~\ref{ass:link}, $f'(x)$ is log-concave in $x$.
+It follows that $\mathds{1}_{[u,v]}(x)f'(x)$
+is a log-concave in $(u,v,x)$. By Lemma~\ref{lem:lossconvexity}, we conclude that $\lambda(u,v)$ is concave in $(u,v)$ where $u>v$.
+\end{proof}
+
+\begin{lem}[Corollary 3.5 in~\cite{brascamp2002extensions}]\label{lem:lossconvexity}
+Let $F(x,y):\mathbb{R}^{m+n}\rightarrow \mathbb{R}$ be an integrable function where $x\in \mathbb{R}^{m},y\in \mathbb{R}^n$. Let
+\[
+G(x) = \int_{\mathbb{R}^n}F(x,y)dy.
+\]
+If $F(x,y)$ is log concave in $(x,y)$, then $G(x)$ is log concave in $x$.
+\end{lem}
+
+\subsection{Auxiliary lemmas}\label{sec:lemma}
+This section collects lemmas that are useful for the proofs of the main theorems.
+
+\begin{defn}[Atomic M-norm~\citep{ghadermarzy2019near}]
+Define $T_{\pm}=\{\tT\in\{\pm 1 \}^{d_1\times \cdots \times d_K}\colon$ $\text{rank}(\tT)=1\}$. The atomic M-norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
+\begin{align}
+\anormSize{}{\Theta}&=\inf\{t>0: \Theta\in t\text{conv}(T_{\pm})\}\\
+&=\inf\left\{\sum_{\tX\in T_{\pm}}c_x\colon \Theta=\sum_{\tX\in T_{\pm}} c_x\tX, \ c_x>0\right\}.
+\end{align}
+\end{defn}
+
+\begin{defn}[Spectral norm~\citep{lim2005singular}]
+The spectral norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
+\[
+\snormSize{}{\Theta}=\sup\left\{\langle \Theta, \mx_1\otimes \cdots \otimes \mx_K\rangle \colon \vnormSize{}{\mx_k}=1,\ \mx_k\in\mathbb{R}^{d_k},\ \text{for all}\ k\in[K]\right\}.
+\]
+\end{defn}
+
+\begin{defn}[Nuclear norm~\citep{friedland2018nuclear}]
+The nuclear norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
+\[
+\nnormSize{}{\Theta}=\inf
+\left\{
+\sum_{i\in[r]}|\lambda_i|\colon \Theta=\sum_{i=1}^r \lambda_i\mx^{(i)}_1\otimes \cdots \otimes \mx^{(i)}_K,\ \vnormSize{}{\mx^{(i)}_k}=1,\ \mx^{(i)}_k\in\mathbb{R}^{d_k},\ \text{for all} k\in[K],\ i\in[r]
+\right\},
+\]
+where the infimum is taken over all $r\in\mathbb{N}$ and $\vnormSize{}{\mx^{(i)}_k}=1$ for all $i\in[r]$ and $k\in[K]$.
+\end{defn}
+
+\begin{lem}[M-norm and infinity norm~\citep{ghadermarzy2019near}]\label{lem:Mnormbound}
+Let $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ be an order-$K$, rank-$(r_1,\ldots,r_K)$ tensor. Then
+\[
+\mnormSize{}{\Theta}\leq \anormSize{}{\Theta}\leq \left(\prod_k r_k \over r_{\max}\right)^{3\over 2} \mnormSize{}{\Theta}.
+\]
+\end{lem}
+
+\begin{lem}[Nuclear norm and F-norm] \label{lem:nuclear}
+Let $\tA\in\mathbb{R}^{d_1\times\cdots\times d_K}$ be an order-$K$ tensor with Tucker $\text{rank}(\tA)=(r_1,\ldots,r_K)$. Then
+\[
+\nnormSize{}{\tA} \leq \sqrt{\prod_k r_k\over \max_k r_k} \FnormSize{}{\tA},
+\]
+where $\nnormSize{}{\cdot}$ denotes the nuclear norm of the tensor.
+\end{lem}
+
+\begin{proof}
+Without loss of generality, suppose $r_1=\min_k r_k$.
+Let $\tA_{(k)}$ denote the mode-$k$ matricization of $\tA$ for all $k\in[K]$.
+By \citet[Corollary 4.11]{wang2017operator}, and the invariance relationship between a tensor and its Tucker core~\citep[Section 6]{jiang2017tensor}, we have
+\begin{equation}\label{eq:norminequality}
+\nnormSize{}{\tA} \leq \sqrt{\prod_{k\geq 2} r_k \over \max_{k\geq 2} r_k} \nnormSize{}{\tA_{(1)}},
+\end{equation}
+where $\tA_{(1)}$ is a $d_1$-by-$\prod_{k\geq 2}d_k$ matrix with matrix rank $r_1$.
+Furthermore, the relationship between the matrix norms implies that $\nnormSize{}{\tA_{(1)}}\leq \sqrt{r_1}\FnormSize{}{\tA_{(1)}}=\sqrt{r_1}\FnormSize{}{\tA}$.
+Combining this fact with the inequality~\eqref{eq:norminequality} yields the final claim.
+\end{proof}
+
+\begin{lem} \label{lem:inq}
+Let $\tA, \; \tB$ be two order-$K$ tensors of the same dimension. Then
+\[
+|\langle \tA,\tB\rangle| \leq \snormSize{}{\tA}   \nnormSize{}{\tB}.
+\]
+\end{lem}
+
+\begin{proof}
+By~\citet[Proposition 3.1]{friedland2018nuclear}, there exists a nuclear norm decomposition of $\tB$, such that
+\[
+\tB=\sum_{r} \lambda_r \ma^{(1)}_r\otimes \cdots\otimes \ma^{(K)}_r,\quad \ma_r^{(k)}\in\mathbf{S}^{d_k-1}(\mathbb{R}),\quad \text{for all }k\in[K],
+\]
+and $\nnormSize{}{\tB}=\sum_{r}|\lambda_r|$. Henceforth we have
+\begin{align*}
+|\langle \tA,\tB\rangle|&=| \langle \tA, \sum_{r} \lambda_r \ma^{(1)}_r\otimes \cdots\otimes \ma^{(K)}_r \rangle|\leq \sum_r |\lambda_r| |\langle \tA, \ma^{(1)}_r \otimes \cdots\otimes \ma^{(K)}_r \rangle|\\
+&\leq \sum_{r}|\lambda_r| \snormSize{}{\tA}= \snormSize{}{\tA}\nnormSize{}{\tB},
+\end{align*}
+which completes the proof.
+\end{proof}
+
+The following lemma provides the bound on the spectral norm of random tensors. The result was firstly presented in~\cite{nguyen2015tensor}, and we adopt the version from~\cite{tomioka2014spectral}.
+\begin{lem}[Spectral norm of random tensors~\citep{tomioka2014spectral}]\label{lem:tensor}
+Suppose that $\tS=\entry{s_{\omega}}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is an order-$K$ tensor whose entries are independent random variables that satisfy
+\[
+\mathbb{E}(s_{\omega})=0,\quad \text{and} \quad\mathbb{E}(e^{ts_{\omega}})\leq e^{t^2L^2/2}.
+\]
+Then the spectral norm $\snormSize{}{\tS}$ satisfies that,
+\[
+\snormSize{}{\tS}\leq \sqrt{{8L^2} \log (12K) \sum_k d_k +\log (2/\delta)},
+\]
+with probability at least $1-\delta$.
+\end{lem}
+
+\begin{lem} \label{lem:noisytensor}
+Suppose that $\tS=\entry{s_{\omega}}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is an order-$K$ tensor whose entries are independent random variables that satisfy
+\[
+\mathbb{E}(s_{\omega})=0,\quad \text{and}\quad |s_{\omega}|\leq U.
+\]
+Then we have
+\[
+\mathbb{P}\left(\snormSize{}{\tS}\geq C_2 U\sqrt{\sum_k d_k} \right)\leq \exp\left(-C_1  \log K \sum_k d_k\right)
+\]
+where $C_1>0$ is an absolute constant, and $C_2>0$ is a constant that depends only on $K$.
+\end{lem}
+
+\begin{proof}  Note that the random variable $U^{-1}s_{\omega}$ is zero-mean and supported on $[-1,1]$. Therefore, $U^{-1}s_{\omega}$ is sub-Gaussian with parameter ${1-(-1)\over 2}=1$, i.e.
+\[
+\mathbb{E}(U^{-1}s_{\omega})=0,\quad \text{and}\quad \mathbb{E}(e^{tU^{-1}s_{\omega}})\leq e^{t^2/2}.
+\]
+It follows from Lemma~\ref{lem:tensor} that, with probability at least $1-\delta$,
+\[
+\snormSize{}{U^{-1}\tS}\leq \sqrt{\left(c_0\log K+c_1\right) \sum_k d_k +\log (2/\delta)},
+\]
+where $c_0, c_1>0$ are two absolute constants. Taking $\delta=\exp (-C_1\log K \sum_k d_k)$ yields the final claim, where $C_2=c_0\log K+c_1+1>0$ is another constant.
+\end{proof}
+
+\begin{lem}\label{lem:KLentry} Let $X,\; Y$ be two discrete random variables taking values on $L$ possible categories, with category probabilities $\{p_\ell\}_{\ell\in[L]}$ and $\{q_\ell\}_{\ell\in[L]}$, respectively.
+Suppose $p_\ell$, $q_\ell>0$ for all $i\in[L]$.
+Then, the Kullback-Leibler (KL) divergence satisfies that
+\[
+\mathrm{KL}(X||Y)\stackrel{\text{def}}{=}-\sum_{\ell\in[L]}\mathbb{P}_X(\ell)\log\left\{{\mathbb{P}_Y(\ell)\over \mathbb{P}_X(\ell)}\right\} \leq \sum_{\ell \in [L]}{(p_\ell-q_\ell)^2 \over q_\ell}.
+\]
+\end{lem}
+\begin{proof} Using the fact $\log x\leq x-1$ for $x>0$, we have that
+\begin{align}\label{eq:KL}
+\text{KL}(X||Y)&=\sum_{\ell\in[L]}p_\ell\log{p_\ell\over q_\ell}\\
+&\leq \sum_{\ell\in[L]} {p_\ell\over q_\ell}(p_\ell-q_\ell)\\
+&=\sum_{\ell\in[L]} \left({p_\ell\over q_\ell }- 1\right)(p_\ell-q_\ell)+ \sum_{\ell\in[L]} (p_\ell-q_\ell).
+\end{align}
+Note that $\sum_{\ell\in[L]}(p_\ell-q_\ell)=0$. Therefore,
+\[
+\text{KL}(X||Y)\leq \sum_{\ell\in[L]}\left( {p_\ell\over q_\ell}-1\right)\left(p_\ell-q_\ell\right)=\sum_{\ell\in[L]}{(p_\ell-q_\ell)^2\over q_\ell}.
+\]
+\end{proof}
+
+\begin{lem}[KL divergence and F-norm]~\label{lem:KL}
+Let $\tY\in[L]^{d_1\times \cdots \times d_K}$ be an ordinal tensor generated from the model~\eqref{eq:model} with the link function $f$ and parameter tensor $\Theta$.
+Let $\mathbb{P}_{\Theta}$ denote the joint categorical distribution of $\tY|\Theta$ induced by the parameter tensor $\Theta$, where $\mnormSize{}{\Theta}\leq \alpha$.
+Define
+\begin{equation}\label{eq:ass}
+A_\alpha=\min_{\ell\in[L], |\theta|\leq \alpha}\left[f(b_\ell-\theta)-f(b_{\ell-1}-\theta)\right].
+\end{equation}
+Then, for any two tensors $\Theta,\; \Theta^*$ in the parameter spaces, we have
+\[
+\mathrm{KL}(\mathbb{P}_{\Theta}|| \mathbb{P}_{\Theta^*})\leq {2(2L-3)\over A_\alpha} \dot{f}^2(0)\FnormSize{}{\Theta-\Theta^*}^2.
+\]
+\end{lem}
+\begin{proof} Suppose that the distribution over the ordinal tensor $\tY=\entry{y_\omega}$ is induced by $\Theta=\entry{\theta_\omega}$. Then, based on the generative model~\eqref{eq:model},
+\[
+\mathbb{P}(y_\omega=\ell | \theta_\omega)=f(b_{\ell}-\theta_\omega)-f(b_{\ell-1}-\theta_\omega),
+\]
+for all $\ell\in[L]$ and $\omega\in[d_1]\times \cdots \times [d_K]$.
+For notational convenience, we suppress the subscribe in $\theta_\omega$ and simply write $\theta$ (and respectively, $\theta^*$).
+Based on Lemma~\ref{lem:KLentry} and Taylor expansion,
+\begin{align}
+\text{KL}(\theta|| \theta^*) & \leq \sum_{\ell\in[L]} {\left[f(b_{\ell}-\theta)-f(b_{\ell-1}-\theta)-f(b_{\ell}-\theta^*)+f(b_{\ell-1}-\theta^*)\right]^2\over f(b_{\ell}-\theta^*)-f(b_{\ell-1}-\theta^*)}\\
+ &\leq \sum_{\ell=2}^{L-1} {\left[\dot{f}(b_\ell-\eta_\ell)-\dot{f}(b_{\ell-1}-\eta_{\ell-1})\right]^2 \over f(b_\ell-\theta^*)-f(b_{\ell-1}-\theta^*)} \left(\theta-\theta^*\right)^2+{\dot{f}^2(b_1-\eta_1) \over f(b_1-\theta^*)} (\theta-\theta^*)^2\\
+ &\quad \quad\quad \quad  +{\dot{f}^2(b_{L-1}-\eta_{L-1})\over 1-f(b_{L-1}-\theta^*)}(\theta-\theta^*)^2,
+\end{align}
+where $\eta_\ell$ and $\eta_{\ell-1}$ fall between $\theta$ and $\theta^*$. Therefore,
+\begin{equation}\label{eq:entrywise}
+\text{KL}(\theta|| \theta^*) \leq \left({4(L-2)\over A_\alpha}+ {2\over A_\alpha}\right)\dot{f}^2(0)(\theta-\theta^*)^2={2(2L-3)\over A_\alpha} \dot{f}^2(0)(\theta-\theta^*)^2,
+\end{equation}
+where we have used Taylor expansion, the bound~\eqref{eq:ass}, and the fact that $\dot{f}(\cdot)$ peaks at zero for an unimodal and symmetric function.
+Now summing~\eqref{eq:entrywise} over the index set $\omega\in[d_1]\times \cdots \times [d_K]$ gives
+\[
+\text{KL}(\mathbb{P}_{\Theta}|| \mathbb{P}_{\Theta^*}) =\sum_{\omega\in[d_1]\times \cdots \times[d_K]} \text{KL}(\theta_\omega || \theta^*_\omega) \leq {2(2L-3)\over A_\alpha} \dot{f}^2(0)\FnormSize{}{\Theta-\Theta^*}^2.
+\]
+\end{proof}
+
+\begin{rmk} In particular, let $\mathbb{P}_{\bf{0}}$ denote the distribution of $\tY|\bf{0}$ induced by the zero parameter tensor. Then we have
+\[
+\text{KL}(\mathbb{P}_{\Theta}||\mathbb{P}_{\bf{0}})\leq {2(2L-3)\over A_\alpha}  \dot{f}^2(0)\FnormSize{}{\Theta}^2.
+\]
+\end{rmk}
+
+\begin{lem}\label{lem:construction}
+Assume the same setup as in Theorem~\ref{thm:minimax}.
+Without loss of generality, suppose $d_1=\max_kd_k$.
+Define $R=\max_k r_k$ and $d_{\text{total}}=\prod_{k\in[K]} d_k$.
+For any constant $0\leq \gamma \leq 1$, there exist a finite set of tensors $\tX=\{\Theta_i: i=1,\ldots\}\subset \tP$ satisfying the following four properties:
+\begin{enumerate}[(i)]
+\item $\text{Card}(\tX)\geq 2^{Rd_1/8}+1$, where $\text{Card}$ denotes the cardinality;
+\item $\tX$ contains the zero tensor $\mathbf{0}\in\mathbb{R}^{d_1\times \cdots\times d_K}$;
+\item $\mnormSize{}{\Theta}\leq \gamma \min\left\{ \alpha ,\ C^{-1/2}\sqrt{Rd_1\over d_{\text{total}}} \right\} $ for any element $\Theta\in\tX$;
+\item $\FnormSize{}{\Theta_i-\Theta_j}\geq {\gamma\over 4} \min\left\{ \alpha\sqrt{d_{\text{total}}},\ C^{-1/2}\sqrt{Rd_1}\right\}$ for any two distinct elements $\Theta_i,\; \Theta_j\in\tX$,
+\end{enumerate}
+Here $C=C(\alpha,L,f,\mb)={(4L-6)\dot{f}^2(0)\over A_\alpha }>0$ is a constant independent of the tensor dimension and rank.
+\end{lem}
+
+\begin{proof}
+Given a constant $0\leq \gamma \leq 1$, we define a set of matrices:
+\[
+\tC=\left\{\mM=(m_{ij})\in\mathbb{R}^{d_1\times R}: a_{ij}\in \left\{ 0,\gamma \min\left\{ \alpha , C^{-1/2}\sqrt{Rd_1\over d_{\text{total}}} \right\}\right\} ,\  \forall (i,j)\in[d_1]\times[R]\right\}.
+\]
+We then consider the associated set of block tensors:
+\begin{align}
+\tB=\tB(\tC)=\{\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}\colon& \Theta=\mA\otimes \mathbf{1}_{d_3}\otimes \cdots \otimes \mathbf{1}_{d_K}, \\
+&\ \text{where}\ \mA=(\mM|\cdots|\mM|\mO) \in\mathbb{R}^{d_1\times d_2},\ \mM\in\tC\},
+\end{align}
+where $\mathbf{1}_d$ denotes a length-$d$ vector with all entries 1, $\mO$ denotes the $d_1\times (d_2-R\lfloor d_2/R \rfloor)$ zero matrix, and $\lfloor d_2/ R \rfloor$ is the integer part of $d_2/R$.
+In other words, the subtensor $\Theta(\mI, \mI,i_3, \ldots,i_K)\in\mathbb{R}^{d_1\times d_2}$ are the same for all fixed $(i_3,\ldots,i_K)\in[d_3]\times \cdots \times [d_K]$, and furthermore, each subtensor $\Theta(\mI,\mI, i_3,\ldots,i_K)$ itself is filled by copying the matrix $\mM\in\mathbb{R}^{d_1\times R}$ as many times as would fit.
+
+By construction, any element of $\tB$, as well as the difference of any two elements of $\tB$, has Tucker rank at most $\max_k r_k\leq R$, and the entries of any tensor in $\tB$ take values in $[0,\alpha]$.
+Thus, $\tB\subset\tP$.
+By Lemma~\ref{lem:VGbound}, there exists a subset $\tX\subset \tB$ with cardinality $\text{Card}(\tX)\geq 2^{Rd_1/8}+1$ containing the zero $d_1\times \cdots \times d_K$ tensor, such that, for any two distinct elements $\Theta_i$ and $\Theta_j$ in $\tX$,
+\[
+\FnormSize{}{\Theta_i-\Theta_j}^2 \geq {Rd_1\over 8} \gamma^2\min\left\{ \alpha^2, {C^{-1}Rd_1 \over d_{\text{total}}}\right\} \lfloor {d_2\over R} \rfloor \prod_{k\geq 3}d_k\geq {\gamma^2\min\left\{ \alpha^2 d_{\text{total}}, C^{-1}Rd_1\right\}  \over 16}.
+\]
+In addition, each entry of $\Theta\in\tX$ is bounded by $\gamma \min\left\{ \alpha , C^{-1/2}\sqrt{Rd_1\over d_{\text{total}}}\right\} $. Therefore the Properties (i) to (iv) are satisfied.
+\end{proof}
+
+\begin{lem}[Varshamov-Gilbert bound]\label{lem:VGbound}
+Let $\Omega=\{(w_1,\ldots,w_m)\colon w_i\in\{0,1\}\}$. Suppose $m>8$. Then there exists a subset $\{w^{(0)},\ldots,w^{(M)}\}$ of $\Omega$ such that $w^{(0)}=(0,\ldots,0)$ and
+\[
+\zeronormSize{}{w^{(j)}-w^{(k)}}\geq {m\over 8},\quad \text{for} \ 0\leq j<k\leq M,
+\]
+where $\zeronormSize{}{\cdot}$ denotes the Hamming distance, and $M\geq 2^{m/8}$.
+\end{lem}
+
+\begin{lem}[Theorem 2.5 in~\cite{tsybakov2008introduction}]\label{lem:Tsybakov}
+Assume that a set $\tX$ contains element $\Theta_0, \Theta_1, \ldots,\Theta_M$ ($M\geq 2$) such that
+\begin{itemize}
+\item $d(\Theta_j,\ \Theta_k)\geq 2s>0$, $\forall 0\leq j\leq k\leq M$;
+\item $\mathbb{P}_0$ is absolutely continuous with respect to $\mathbb{P}_j$, $\forall j=1,\ldots,M$, and
+\[
+{1\over M}\sum_{j=1}^M \mathrm{KL}(\mathbb{P}_j||\mathbb{P}_0)\leq \alpha \log M
+\]
+where $d\colon \tX\times \tX\mapsto [0,+\infty]$ is a semi-distance function, $0<\alpha<{1/8}$ and $\mathbb{P}_j=\mathbb{P}_{\Theta_j}$, $j=0,1\ldots,M$.
+\end{itemize}
+Then
+\[
+\inf_{\hat \Theta}\sup_{\Theta\in\tX} \mathbb{P}_{\Theta}(d(\hat \Theta, \Theta)\geq s)\geq {\sqrt{M}\over 1+\sqrt{M}}\left(1-2\alpha-\sqrt{2\alpha\over \log M} \right)>0.
+\]
+\end{lem}
+
+\begin{lem}[Lemma 28 in~\cite{ghadermarzy2019near}]\label{lem:convexity}
+Define $\mathbb{B}_{M}(\beta)=\{\Theta\in \mathbb{R}^{d_1\times \cdots \times d_K}\colon \anormSize{}{\Theta}\leq \beta \}$.
+Let $\Omega\subset[d_1]\times\cdots \times [d_K]$ be a random set with $m=|\Omega|$, and assume that each entry in $\Omega$ is drawn with replacement from $[d_1]\times\cdots\times[d_K]$ using probability $\Pi$.
+Define
+\[
+\PiFnormSize{}{\Theta}^2={1\over m}\mathbb{E}_{\Omega\in\Pi}\MFnormSize{}{\Theta}^2.
+\]
+Then, there exists a universal constant $c>0$, such that, with probability at least $1-\exp\left(-{\sum_kd_k \over \sum_k \log d_k} \right)$ over the sampled set $\Omega$,
+\[
+{1\over m}\MFnormSize{}{\Theta}^2 \geq \PiFnormSize{}{\Theta}^2-c\beta\sqrt{\sum_k d_k\over m}
+\]
+holds uniformly for all $\Theta\in\mathbb{B}_M(\beta)$.
+\end{lem}
+
+\section{Extension of Theorem~\ref{thm:rate} to unknown cut-off points}\label{sec:extention}
+When the cut-off points $\mb$ is unknown, we estimate $(\hat \Theta,\hat \mb)$ by
+\begin{equation}\label{eq:joint}
+(\hat \Theta,\hat \mb)=\argmax_{(\Theta,\mb)\in\tP\times \tB}\logl(\Theta,\mb),
+\end{equation}
+where
+\[
+\tP=\{\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}\colon \text{rank}(\tP)\leq \mr,\ \mnormSize{}{\Theta}\leq \alpha\}, \quad \tB=\{\mb\in\mathbb{R}^{L-1}\colon \mnormSize{}{\mb}\leq \beta,\ \min_{\ell}(b_\ell-b_{\ell-1}) \geq \Delta\}.
+\]
+The estimation accuracy is assessed using the mean squared error (MSE):
+\[
+\text{MSE}\left( \hat \Theta, \trueT \right)={1\over \prod_k d_k}\FnormSize{}{\hat \Theta-\trueT},\quad \text{MSE}\left( \hat \mb, \trueb \right)={1\over L-1}\FnormSize{}{\hat \mb-\trueb}.
+\]
+By abuse of notation, we use $(\Theta,\mb)$ to denote the length-$(d_{\text{total}}+L-1)$-vector collecting all parameters together. The total MSE is defined as
+\[
+\text{MSE}\left((\hat \Theta,\hat \mb),\ (\trueT,\trueb)\right)={1\over \prod_k d_k+L-1}\FnormSize{}{(\hat \Theta,\hat \mb)-(\trueT,\trueb)}^2.
+\]
+We introduce several quantities that will be used in our theory:
+\begin{enumerate}
+\item We make the convention that $b_0=-\infty$, $b_L=\infty$, $f(-\infty)=0$, $f(\infty)=1$, and $\dot{f}(-\infty)=\ddot{f}(-\infty)=\dot{f}(\infty)=\ddot{f}(\infty)=0$.
+
+\item The difference function $g_\ell(\theta)$ is defined as $g_\ell(\theta)=f(b_\ell-\theta)-f(b_{\ell-1}-\theta)$ for all $\theta\in\mathbb{R}$ and $\ell\in[L]$.
+
+\item Define $n_\ell=\sum_{\omega\in\Omega}\mathds{1}_{\{y_\omega=\ell\}}$, i.e., the number of tensor entries taking value on $\ell\in[L]$.
+
+\item With a little abuse of notation, we re-define the constants in~\eqref{eq:bound} as
+\begin{equation}\label{eq:UL}
+U_{\alpha,\beta,\Delta}=\max_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}}\max_{\ell\in[L-1]}{|\dot{g}_\ell(\theta)|\over g_\ell(\theta)}, \quad \text{and}\quad
+L_{\alpha,\beta,\Delta}=\min_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}}\min_{\ell\in[L-1]} {\dot{g}^2_\ell(\theta)-\ddot{g}_\ell(\theta)g_\ell(\theta)\over g^2_\ell(\theta)}.
+\end{equation}
+\item We define three additional constants:
+\begin{align}\label{eq:CD}
+C_{\alpha,\beta,\Delta} &= \max_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}}  \max_{\ell\in[L-1]}\max\left\{\frac{\dot{f}(b_\ell-\theta)}{g_\ell(\theta)},\ \frac{\dot{f}(b_\ell-\theta)}{g_{\ell+1}(\theta)}\right\},\notag \\
+D_{\alpha,\beta,\Delta} &=\min_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}} \min_{\ell\in[L-1]}\min\left\{  {\partial \over \partial \theta}\left({\dot{f}(b_{\ell}-\theta)\over g_\ell(\theta)}\right),\ -{\partial \over \partial \theta}\left({\dot{f}(b_{\ell}-\theta)\over g_{\ell+1}(\theta)}\right)\right\}\notag \\
+&=  \min_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}}\min_{\ell\in[L-1]}\min \left\{-\frac{\ddot{f}(b_\ell-\theta)g_\ell(\theta)-\dot{f}(b_\ell-\theta)\dot{g}_\ell(\theta)}{g^2_\ell(\theta)},\ \right.
+\notag \\
+&\hspace{1.5in} \left. \frac{\ddot{f}(b_\ell-\theta)g_{\ell+1}(\theta)-\dot{f}(b_\ell-\theta)\dot{g}_{\ell+1}(\theta)}{g^2_{\ell+1}(\theta)}\right\},\notag \\
+\tilde{D}_{\alpha,\beta,\Delta} &=\max_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}} \max_{\ell\in[L-1]}\max\left\{  {\partial \over \partial \theta}\left({\dot{f}(b_{\ell}-\theta)\over g_\ell(\theta)}\right),\ -{\partial \over \partial \theta}\left({\dot{f}(b_{\ell}-\theta)\over g_{\ell+1}(\theta)}\right)\right\}.
+\end{align}
+\end{enumerate}
+We make the following assumptions about the link function.
+\begin{assumption}\label{ass:joint}
+The link function $f\colon \mathbb{R}\mapsto [0,1]$ satisfies the following properties:
+\begin{enumerate}
+\item $f(\theta)$ is twice-differentiable and strictly increasing in $\theta$.
+\item $\dot{f}(\theta)$ is strictly log-concave and symmetric with respect to $\theta=0$.
+\item The function ${\dot{f}(b_\ell-\theta)\over g_\ell(\theta)}$ is strictly increasing with respect to $\theta$ for all $\mb\in\tB$.
+\item The function ${\dot{f}(b_\ell-\theta)\over g_{\ell+1}(\theta)}$ is strictly decreasing with respect to $\theta$ for all $\mb\in\tB$.
 \end{enumerate}
@@
 
-\subsection{Problem 1: Tensor denoising}~\label{sec:denoising}
-The first question we aim to address is tensor denoising:
-
-{\bf (P1)} Given the quantization process induced by $f$ and the cut-off points $\mb$, how accurately can we estimate the latent signal tensor $\Theta$ from the ordinal observation $\tY$?
-
-Clearly, the problem (P1) cannot be solved uniformly for all possible $\Theta$.
-We focus on a class of ``low-rank'' and ``flat'' signal tensors, which is a plausible assumption in practical applications~\citep{zhou2013tensor,bhaskar20151}.
-Specifically, we consider the parameter space:
-\begin{equation}\label{eq:space}
- \tP=\left\{\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K} \colon \text{rank}(\Theta)\leq \mr,\ \mnormSize{}{\Theta}\leq \alpha\right\}.
-\end{equation}
-where $\mr=(r_1,\ldots,r_K)$ denotes the Tucker rank of $\Theta$.
-
-The parameter tensor of our interest satisfies two constraints.
-The first is that $\Theta$ is a low-rank tensor, with $r_k=\tO(1)$ for all $k\in[K]$.
-Equivalently, $\Theta$ admits the Tucker decomposition:
-\begin{equation}\label{eq:Tucker}
-\Theta=\tC\times_1\mM_1\times_1 \cdots \times_K \mM_K,
-\end{equation}
-where $\tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is a core tensor, $\mM_k\in\mathbb{R}^{d_k\times r_k}$ are factor matrices with orthogonal columns, and $\times_k$ denotes the tensor-by-matrix multiplication~\citep{kolda2009tensor}.
-The Tucker low-rankness is popularly imposed in tensor data analysis, and is shown to provide a reasonable tradeoff between model complexity and model flexibility.
-Note that, unlike matrices, there are various notations of tensor low-rankness, such as CP rank~\citep{hitchcock1927expression} and train rank~\citep{oseledets2011tensor}.
-Some notation of low-rankness may lead to mathematically ill-posed optimization; for example, the best low CP-rank tensor approximation may not exist~\citep{de2008tensor}.
-We choose Tucker representation for well-posedness of optimization and easy interpretation.
-
-The second constraint is that the entries of $\Theta$ are uniformly bounded in magnitude by a constant $\alpha \in \mathbb{R}_{+}$.
-In view of~\eqref{eq:latent}, we refer to $\alpha$ as the signal level.
-The entry-wise bound assumption is a technical condition that avoids the degeneracy in probability estimation with ordinal observations.
-
-\subsection{Problem 2: Tensor completion}
-Motivated by applications in collaborative filtering, we also consider a more general setup when only a subset of tensor entries $y_\omega$ are observed.
-Let $\Omega\subset[d_1]\times \cdots\times[d_K]$ denote the set of observed indices.
-The second question we aim to address is stated as follows:
-
-{\bf (P2)} Given an incomplete set of ordinal observations $\{y_{\omega}\}_{\omega\in\Omega}$, how many sampled entries do we need to consistently recover $\Theta$ based on the model~\eqref{eq:model}?
-
-The answer to (P2) depends on the choice of $\Omega$.
-We consider a general model on $\Omega$ that allows both uniform and non-uniform sampling.
-Specifically, let $\Pi=\{\pi_{i_1,\ldots,i_K}\}$ denote a predefine probability distribution over the index set such that $\sum_{\omega\in[d_1]\times \cdots \times [d_K]} \pi_\omega =1$.
-We assume that each index in $\Omega$ is drawn with replacement using distribution $\Pi$.
-This sampling model relaxes the uniform sampling in literature and is arguably a better fit in applications.
-
-We consider the same parameter space~\eqref{eq:space} for the completion problem.
-In addition to the reasons mentioned in Section~\ref{sec:denoising}, the entrywise bound assumption also serves as the incoherence requirement for completion.
-In classical matrix completion, the incoherence is often imposed on the singular vectors.
-This assumption is recently relaxed for ``flat'' matrices with bounded magnitude~\citep{negahban2011estimation,cai2013max,bhaskar20151}.
-We adopt the same assumption for higher-order tensors.
-
-\section{Rank-constrained M-estimator}\label{sec:theory}
-We present a general treatment to both problems mentioned above.
-With a little abuse of notation, we use $\Omega$ to denote either the full index set $\Omega=[d_1]\times \cdots \times [d_K]$ (for the tensor denoising) or a random subset induced from the sampling distribution $\Pi$ (for the tensor completion).
-Define $b_0=-\infty$, $b_L=\infty$, $f(-\infty)=0$ and $f(\infty)=1$.
-The log-likelihood associated with the observed entries is
-\begin{align}\label{eq:objective}
-\logl(\Theta, \mb)&=\sum_{\omega\in\Omega}\sum_{\ell\in[L]} \Big\{\mathds{1}_{\{y_\omega=\ell\}} \log \big[f(b_\ell-\theta_\omega)-  f(b_{\ell-1}-\theta_\omega)\big]\Big\}.
-\end{align}
-We propose a rank-constrained maximum likelihood estimator (a.k.a.\ M-estimator) for $\Theta$:
-\begin{align}\label{eq:estimator}
-\hat \Theta&=\argmax_{\Theta\in \tP}\logl(\Theta, \mb),\ \text{where}\notag \\
-\tP&=\left\{\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K} \colon \text{rank}(\Theta)\leq \mr,\ \mnormSize{}{\Theta}\leq \alpha\right\}.
- \end{align}
-In practice, the cut-off points $\mb$ are unknown and should be jointly estimated with $\Theta$.
-For technical convenience, we assume in this section that the cut-off points $\mb$ are known.
-The adaptation of unknown $\mb$ is addressed in Section~\ref{sec:algorithm} and Appendix~\ref{sec:extention}.
-
-We define a few key quantities that will be used in our theory. Let $g_\ell(\theta)=f(b_\ell-\theta)-f(b_{\ell-1}-\theta)$ for all $\ell\in[L]$, and
-\begin{equation}\label{eq:regular}
-A_\alpha = \min_{\ell\in[L], |\theta|\leq \alpha} g_\ell(\theta),\quad  U_\alpha=\max_{\ell\in[L], |\theta|\leq \alpha } {|\dot{g}_\ell(\theta)|\over g_\ell(\theta)},\quad \text{and}\quad
- L_\alpha = \min_{\ell\in[L], |\theta|\leq \alpha }\left[{\dot{g}_\ell^2 (\theta)\over g_\ell^2(\theta)} -{\ddot{g}_\ell(\theta)\over g_\ell(\theta)}\right],
-\end{equation}
-where $\dot{g}(\theta)=dg(\theta)/d\theta$, and $\alpha$ is the entrywise bound of $\Theta$.
-In view of equation~\eqref{eq:latent}, these quantities characterize the geometry including flatness and convexity of the latent noise distribution.
-Under the Assumption~\ref{ass:link}, all these quantities are strictly positive and independent of tensor dimension.
-
-\subsection{Estimation error for tensor denoising}\label{sec:denosing}
-For the tensor denoising problem, we assume that the full set of tensor entries are observed. We assess the estimation accuracy using the mean squared error (MSE):
-\[
-\mathrm{MSE}(\hat \Theta, \trueT)={1\over \prod_k d_k}\FnormSize{}{\Theta-\trueT}^2.
-\]
-The next theorem establishes the upper bound for the MSE of the proposed $\hat \Theta$ in~\eqref{eq:estimator}.
-
-\begin{thm}[Statistical convergence] \label{thm:rate}
-Consider an ordinal tensor $\tY\in[L]^{d_1\times\dots\times d_K}$ generated from model~\eqref{eq:model}, with the link function $f$ and the true coefficient tensor $\trueT\in\tP$.
-Define $r_{\max}=\max_k r_k$.
-Then, with very high probability, the estimator in~\eqref{eq:estimator} satisfies
-\begin{equation}\label{eq:rate}
-\mathrm{MSE}(\hat \Theta, \trueT) \leq \min\left( 4\alpha^2,\ {c_1  U^2_\alpha r_{\max}^{K-1}  \over  L^2_\alpha } {\sum_kd_k\over  \prod_k d_k} \right),
-\end{equation}
-where $c_1>0$ is a constant that depends only on $K$.\\
+\begin{rmk}
+The condition $\Delta=\min_{\ell}(b_{\ell}-b_{\ell-1})>0$ on the feasible set $\tB$ guarantees the strict positiveness of $g_{\ell}(\theta)=f(b_{\ell}-\theta)-f(b_{\ell-1}-\theta)$.
+Therefore, the denominators in the above quantities $U_{\alpha,\beta,\Delta}, L_{\alpha,\beta,\Delta}, C_{\alpha,\beta,\Delta}, D_{\alpha,\beta,\Delta},\tilde{D}_{\alpha,\beta,\Delta}$ are well-defined.
+Furthermore, by Theorem~\ref{thm:convexity}, $g_\ell(\theta)$ is strictly log-concave in $\theta$ for all $\mb\in\tB$.
+Based on Assumption~\ref{ass:joint} and closeness of the feasible set, we have $U_{\alpha,\beta,\Delta}>0$, $L_{\alpha,\beta,\Delta}>0$, $C_{\alpha,\beta,\Delta}>0$, $D_{\alpha,\beta,\Delta}>0$, $\tilde{D}_{\alpha,\beta,\Delta}>0$.
+\end{rmk}
+
+\begin{rmk}
+In particular, for logistic link $f(x) = \frac{1}{1+e^{-x}}$ , we have
+\begin{align}
+C_{\alpha,\beta,\Delta}&=\max_{\substack{|\theta|\leq \alpha,\\ \mb\in\tB}}\max_{\ell\in[L-1]}\max
+\left\{
+{ 1\over e^{b_{\ell}-b_{\ell-1}}-1} { 1+e^{-(b_{\ell-1}-\theta)} \over 1+e^{-(b_{\ell}-\theta)}},\
+{ 1\over 1-e^{-(b_{\ell+1}-b_{\ell})}} { 1+e^{-(b_{\ell+1}-\theta)} \over 1+e^{-(b_{\ell}-\theta)}}
+\right\}>0,
+\\
+D_{\alpha,\beta,\Delta}&=\min_{\substack{|\theta|\leq \alpha,\\ \mb\in\tB}}\min_{\ell\in[L-1]} {e^{-(b_\ell-\theta)}\over (1+e^{-(b_\ell-\theta)})^2}>0.
+\end{align}
+\end{rmk}
+
+\begin{thm}[Statistical convergence with unknown $\mb$]\label{thm:ratejoint}
+Consider an ordinal tensor $\tY\in[L]^{d_1\times \cdots \times d_K}$ generated from model~\eqref{eq:model} with the link function $f$ and parameters $(\trueT, \trueb)\in\tP\times \tB$.
+Suppose the link function $f$ satisfies Assumption~\ref{ass:joint}.
+Define $r_{\max}=\max_k r_k$, and assume $r_{\max}=\tO(1)$.
+Then with very high probability, the estimator in~\eqref{eq:joint} satisfies
+\begin{equation}\label{eq:jointTheta}
+\mathrm{MSE}\left(\hat \Theta, \trueT \right)\leq\min\left( 4\alpha^2,\  {c_1U_{\alpha,\beta,\Delta}^2  r_{\max}^{K-1}  \over L_{\alpha,\beta,\Delta}^2}{ \sum_k d_k \over \prod_k d_k}\right),
+\end{equation}
+and
+\begin{equation}\label{eq:jointb}
+\mathrm{MSE}\left(\hat \mb, \trueb \right)\leq \min\left(4\beta^2,\ {C'_{\alpha,\beta,\Delta}r^{K-1}_{\max} \over D_{\alpha,\beta,\Delta}^2}{\left(\sum_kd_k\right)\left(\prod_k d_k\right) \over \min_{\ell}(n_{\ell}+n_{\ell+1})^2}\right),
+\end{equation}
+where $c_1, U_{\alpha,\beta,\Delta}, L_{\alpha,\beta,\Delta}, C'_{\alpha,\beta,\Delta}, \tilde D_{\alpha,\beta,\Delta}$ are constants independent of the tensor dimension and rank.
+
+In particular, we have
+\begin{align}
+\mathrm{MSE}\left((\hat \Theta,\hat \mb),\ (\trueT,\trueb)\right) = \tO\left({\sum_k d_k \over \min\left\{\prod_k d_k ,\ \min_\ell(n_\ell+n_{\ell+1})^2\right\}}\right),\quad \text{as}\quad d_{\min}\to\infty.
+\end{align}
 \end{thm}
 
-Theorem~\ref{thm:rate} establishes the statistical convergence for the estimator~\eqref{eq:estimator}.
-In fact, the proof of this theorem (see Section~\ref{sec:proofMSE}) shows that the same statistical rate holds, not only for the global optimizer~\eqref{eq:estimator}, but also for any local optimizer $\check \Theta$ in the level set $\{\check\Theta\in\tP: \logl(\check\Theta)\geq \logl(\trueT)\}$.
-This suggests that the local optimality itself is not necessarily a severe concern in our context, as long as the convergent objective is large enough.
-In Section~\ref{sec:algorithm}, we perform empirical studies to assess the algorithmic stability.
-
-A similar conclusion is obtained for the prediction error, measured in Kullback-Leibler (KL) divergence, between the categorical distributions in the observation space.
-
-\begin{cor}[Prediction error]~\label{cor:prediction}
-Assume the same set-up as in Theorem~\ref{thm:rate}.
-Let $\mathbb{P}_{\tY}$ and $\hat{\mathbb{P}}_{\tY}$ denote the distributions generating the $L$-level ordinal tensor $\tY$, given the true parameter $\Theta$ and its estimator $\hat \Theta$, respectively.
-Assume $L\geq 2$.
-Then, with very high probability,
-\begin{equation}\label{eq:KLrate}
-\mathrm{KL}(\mathbb{P}_{\tY} || \hat{\mathbb{P}}_{\tY}) \leq  {c_1 U^2_\alpha r_{\max}^{K-1}  \over L^2_\alpha } {(4L-6)\dot{f}^2(0)  \over A_\alpha} {\sum_kd_k\over  \prod_k d_k},
-\end{equation}
-where $c_1>0$ is the same constant as in Theorem~\ref{thm:rate}.
-\end{cor}
-
-To gain insight into these bounds, we consider a special setting with equal dimension in all modes, i.e., $d_1=\cdots=d_K=d$. In such a case, our bound \eqref{eq:rate} reduces to
-\begin{equation}\label{eq:ours}
-\text{MSE}(\hat \Theta, \trueT) \asymp d^{-(K-1)}, \quad \text{as}\ d\to \infty.
-\end{equation}
-Hence, our estimator achieves consistency with polynomial convergence rate.
-We compare the bound with existing literature.
-In the special case $L=2$, \citet{ghadermarzy2018learning} proposed a max-norm constrained estimator $\tilde{\Theta}$ with $\text{MSE}(\tilde{\Theta}, \trueT) \asymp  d^{-(K-1)/2}$.
-In contrast, our estimator converges at a rate of $d^{-(K-1)}$, which is substantially faster than theirs.
-This provides a positive answer to the open question posed in~\citet{ghadermarzy2018learning} whether the square root in the bound is removable.
-The improvement stems from utilizing the exact low-rankness of $\Theta$, whereas the surrogate rank measure employed in~\citet{ghadermarzy2018learning} is scale-sensitive.
-
-Our bound also generalizes the previous results on ordinal matrices.
-The convergence rate for rank-constrained matrix estimation was $\tO(1/\sqrt{d})$~\citep{bhaskar2016probabilistic}, which fits into our special case when $K=2$.
-Furthermore, our results~\eqref{eq:rate} and~\eqref{eq:KLrate} reveal that the convergence becomes favorable as the order of data tensor increases.
-Intuitively, the sample size for tensor data analysis is the number of entries, $\prod_k d_k$, and the number of free parameters is roughly on the order of $\sum_{k}d_k$, assuming $r_{\max}=\tO(1)$.
-A higher tensor order implies higher effective sample size per parameter, and thus exhibits a faster convergence rate in high dimensions.
-
-We next show the statistical optimality of our estimator $\hat \Theta$.
-The result is based on the information theory, and applies to all estimators in $\tP$, including but not limited to $\hat \Theta$ in~\eqref{eq:estimator}.
-
-\begin{thm}[Minimax lower bound]\label{thm:minimax}
-Assume the same set-up as in Theorem~\ref{thm:rate}, and $d_{\max}=\max_k d_k \geq 8$.
-Let $\inf_{\hat \Theta}$ denote the infimum over all estimators $\hat \Theta\in\tP$ based on the ordinal tensor observation $\tY\in[L]^{d_1\times \cdots \times d_K}$.
-Then, under the model~\eqref{eq:model},
-\begin{align}\label{eq:lower}
-\inf_{\hat \Theta }\sup_{\trueT\in\tP} \mathbb{P}\Big\{&\mathrm{MSE}(\hat \Theta, \trueT)  \geq c \min\left( \alpha^2, \ { Cr_{\max}d_{\max} \over \prod_k d_k} \right) \Big\} \geq {1\over 8},
-\end{align}
-where $C=C(\alpha, L, f,\mb)>0$ and $c>0$ are constants independent of tensor dimension and the rank.\\
-\end{thm}
-
-We see that the lower bound matches the upper bound in~\eqref{eq:rate} on the polynomial order of tensor dimension. Therefore, our estimator~\eqref{eq:estimator} is order-optimal.
-
-\subsection{Sample complexity for tensor completion}
-We now consider the tensor completion problem, when only a subset of entries $\Omega$ are observed.
-We consider a general sampling procedure induced by $\Pi$.
-The recovery accuracy is assessed by the weighted squared error:
-\begin{align}\label{eq:weighted}
-\PiFnormSize{}{\Theta-\hat \Theta}^2&\stackrel{\text{def}}{=}
-{1\over |\Omega|}\mathbb{E}_{\Omega\sim \Pi}\FnormSize{}{\Theta-\hat \Theta}^2\notag \\
-&=\sum_{\omega\in[d_1]\times \cdots \times [d_K]} \pi_{\omega}(\Theta_{\omega}-\hat \Theta_{\omega})^2.
-\end{align}
-
-Note that the recovery error depends on the distribution $\Pi$.
-In particular, tensor entries with higher sampling probabilities have more influence on the recovery accuracy, compared to the ones with lower sampling probabilities.
-
-\begin{rmk} If we assume each entry is sampled with strictly positive probability; i.e.\ there exits a constant $\mu> 0$ s.t.
-\[
-\ \pi_\omega\geq {1\over \mu \prod_k d_k},\quad \text{for all}\ \omega\in[d_1]\times\cdots \times [d_K],
-\]
-then the error in~\eqref{eq:weighted} provides an upper bound for MSE:
-\[
-\PiFnormSize{}{\Theta-\hat \Theta}^2 \geq {\FnormSize{}{\Theta-\hat \Theta}^2\over \mu \prod_kd_k}={1\over \mu}\text{MSE}(\hat \Theta, \trueT).
-\]
-The equality is attained under uniform sampling with $\mu=1$.
+\begin{rmk} The total MSE has two components, where the first component $\tO\left({\sum_k d_k \over\prod_k d_k}\right)$ is from the error in $\hat\Theta$, and the second component $\tO\left({\sum_k d_k \over \min_\ell(n_\ell+n_{\ell+1})^2}\right)$ is from the error in $\hat \mb.$ When the $L$ labels are moderately balanced in that $\min_\ell(n_\ell+n_{\ell+1})\gg \sqrt{\prod_kd_k}$, the error in $\hat\Theta$ dominates the total MSE.
+When the $L$ labels are severely imbalanced in that $\min_\ell(n_\ell+n_{\ell+1})\ll \sqrt{\prod_kd_k}$, the error in $\hat\mb$ dominates the total MSE.
 \end{rmk}
 
-\begin{thm}~\label{thm:completion}
-Assume the same set-up as in Theorem~\ref{thm:rate}.
-Suppose that we observe a subset of tensor entries $\{y_\omega\}_{\omega\in\Omega}$, where $\Omega$ is chosen at random with replacement according to a probability distribution $\Pi$.
-Let $\hat \Theta$ be the solution to~\eqref{eq:estimator}, and assume $r_{\max}=\tO(1)$.
-Then, with very high probability,
-\[
-\PiFnormSize{}{\Theta-\hat \Theta}^2 \to 0,\quad \text{ as }\quad {|\Omega|\over \sum_k d_k} \ \to \infty.
-\]
-\end{thm}
-
-Theorem~\ref{thm:completion} shows that our estimator achieves consistent recovery using as few as $\tilde{\tO}(Kd)$ noisy, quantized observations from an order-$K$ $(d,\ldots,d)$-dimensional tensor.
-Note that $\tilde \tO(Kd)$ roughly matches the degree of freedom for an order-$K$ tensor of fixed rank $\mr$, suggesting the optimality of our sample requirement.
-This sample complexity substantially improves over earlier result $\tO(d^{\lceil K/2\rceil})$ based on square matricization~\citep{mu2014square}, or $\tO(d^{N/2})$ based on tensor nuclear-norm regularization~\citep{yuan2016tensor}.
-Existing methods that achieve $\tilde \tO(Kd)$ sample complexity require either a deterministic cross sampling design~\citep{zhang2019cross} or univariate measurements~\citep{ghadermarzy2018learning}.
-Our method extends the conclusions to multi-level measurements under a broader class of sampling schemes.
-
-\section{Numerical Implementation}\label{sec:algorithm}
-We describe the algorithm to seek the optimizer of~\eqref{eq:objective}.
-In practice, the cut-off points $\mb$ are often unknown, so we choose to maximize $\logl$ jointly over $(\Theta,\mb)\in\tP\times \tB$ (see Appendix~\ref{sec:extention} for details).
-The objective $\logl$ is concave in $(\Theta,\mb)$ whenever $f'$ is log-concave (see Section~\ref{sec:proofconvexity}).
-However, the feasible set $\tP$ is non-convex, which makes the optimization~\eqref{eq:objective} a non-convex problem.
-We employ the alternating optimization approach by utilizing the Tucker representation of $\Theta$.
-Specifically, based on~\eqref{eq:Tucker} and~\eqref{eq:objective}, the objective function consists of $K+2$ blocks of variables, one for the cut-off points $\mb$, one for the core tensor $\tC$, and $K$ for the factor matrices $\mM_k$'s.
-The optimization is a simple convex problem if any $K+1$ out of the $K+2$ blocks are fixed.
-We update one block at a time while holding others fixed, and alternate the optimization throughout the iteration.
-The convergence is guaranteed whenever $\logl$ is bounded from above, since the alternating procedure monotonically increases the objective.
-The Algorithm~\ref{alg} gives the full description.
-
-\begin{algorithm}[t]
-\caption{Ordinal tensor decomposition}\label{alg}
-\begin{algorithmic}[]
-\STATE{\bfseries Input:}  \text{ Ordinal data tensor }
-$\mathcal{Y}\in [L]^{d_1\times\cdots\times d_K}$, rank $\mr\in \mathbb{N}_{+}^{K}$, entry-wise bound $\alpha\in \mathbb{R_+}$.
-\STATE{\bfseries Output:} $(\hat\Theta,\hat{\mb}) =  \argmax_{(\Theta,\mb)\in \mathcal{P}\times\mathcal{B}}  \mathcal{L}_{\mathcal{Y},\Omega}(\Theta,\mb).$
-\STATE Random initialization of core tensor $\mathcal{C}^{(0)}$, factor matrices $ \{\mM_k^{(0)}\}$, and cut-off points $\mb^{(0)}$.
-\FOR{$t = 1,2,\cdots,$}
-\FOR{$k = 1,2,\cdots,K$}
-\STATE{Update $\mM^{(t+1)}_k$ while fixing other blocks:}
-\STATE $\mM_k^{(t+1)}\gets\argmax_{\mM_k\in\mathbb{R}^{d_k\times r_K}}\mathcal{L}_{\mathcal{Y},\Omega}(\mM_k)$,
-s.t. $\mnormSize{}{\Theta^{(t+1)}}\leq \alpha$, where $\Theta^{(t+1)}$ is the parameter tensor based on the current block estimates.
-\ENDFOR
-\STATE {Update $\mathcal{C}^{(t+1)}$ while fixing other blocks:}
-\newline$\mathcal{C}^{(t+1)} \gets \argmax_{\mathcal{C}\in \mathbb{R}^{r_1\times\cdots\times r_K}}\mathcal{L}_{\mathcal{Y},\Omega}(\mathcal{C})$, s.t. $\mnormSize{}{\Theta^{(t+1)}}\leq \alpha.$
-\STATE {Update $\Theta^{(t+1)}$ based on the current block estimates:}
-\newline
-$\Theta^{(t+1)} \gets \mathcal{C}^{(t+1)}\times_1\mM_1^{(t+1)}\cdots\times_K\mM_K^{(t+1)}.$
-\STATE {Update $\mb^{(t+1)}$ while fixing $\Theta^{(t+1)}$:}
-$\mb^{(t+1)} \gets \argmax_{\mb\in \tB}\mathcal{L}_{\mathcal{Y},\Omega}\big(\Theta^{(t+1)},\mb\big).$
-\ENDFOR
-\STATE \textbf{return}
-$(\hat \Theta,\hat \mb)$
-\end{algorithmic}
-\end{algorithm}
-
-We comment on two implementation details before concluding this section.
-First, the problem~\eqref{eq:estimator} is non-convex, so Algorithm 1 usually has no theoretical guarantee on global optimality.
-Nevertheless, as shown in Section~\ref{sec:denosing}, the desired rate holds not only for the global optimizer, but also for the local optimizer with $\logl(\hat \Theta)\geq \logl(\trueT)$.
-In practice, we find the convergence point $\hat \Theta$ upon random initialization is often satisfactory, in that the corresponding objective $\logl(\hat \Theta)$ is close to and actually slightly larger than the objective evaluated at the true parameter $\logl(\trueT)$.
-Figure~\ref{fig:stability} shows the trajectory of the objective function that is output in the default setting of Algorithm 1, with the input tensor generated from probit model~\eqref{eq:model} with $d_1=d_2=d_3=d$ and $r_1=r_2=r_3=r$.
-The dashed line is the objective value at the true parameter $\logl(\trueT)$.
-We find that the algorithm generally converges quickly to a desirable value in reasonable number of steps.
-The actual running time per iteration is shown in the plot legend.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=0.7\textwidth]{algorithm.pdf}\label{fig:stability}
-\caption{Trajectory of objective function with various $d$ and $r$.}
-\end{figure}
-
-Second, the algorithm takes the rank $\mr$ as an input.
-In practice, the rank $\mr$ is hardly known and needs to be estimated from the data.
-We suggest to use Bayesian information criterion (BIC) and choose the rank that minimizes BIC; i.e.
-\begin{align}\label{eq:BIC}
-\hat \mr&=\argmin_{\mr\in\mathbb{N}^K_{+}} \text{BIC}(\mr)\\
-&=\argmin_{\mr\in\mathbb{N}^K_{+}}\left\{-2\tL_{\tY}(\hat \Theta(\mr),\hat \mb(\mr))+p_e(\mr)\log \left(\prod_k d_k\right)\right\},
-\end{align}
-where $\hat \Theta(\mr), \hat\mb(\mr)$ are the estimates given the rank $\mr$, and $p_e(\mr)\stackrel{\text{def}}{=}\sum_k (d_k-r_k)r_k+\prod_k r_k$ is the effective number of parameters in the model.
-We select $\hat \mr$ that minimizes BIC through a grid search.
-The choice of BIC is intended to balance between the goodness-of-fit for the data and the degrees of freedom in the population model.
-
-\section{Experiments}\label{sec:experiment}
-In this section, we evaluate the empirical performance of our method.
-We investigate both the complete and the incomplete settings, and compare the recovery accuracy with other tensor-based methods.
-Unless otherwise stated, the ordinal data tensors are generated from model~\eqref{eq:model} using standard probit link $f$.
-We consider the setting with $K=3$, $d_1=d_2=d_3=d$, and $r_1=r_2=r_3=r$.
-The parameter tensors are simulated based on~\eqref{eq:Tucker}, where the core tensor entries are i.i.d.\ drawn from $\tN(0,1)$, and the factors $\mM_k$ are uniformly sampled (with respect to Haar measure) from matrices with orthonormal columns.
-We set the cut-off points $b_\ell=f^{-1}(\ell/L)$ for $\ell\in[L]$, such that $f(b_\ell)$ are evenly spaced from 0 to 1.
-In each simulation study, we report the summary statistics across $n_{\text{sim}}$ = 30 replications.
-
-\subsection{Finite-sample performance}\label{sec:simulation}
-The first experiment examines the performance under complete observations.
-We assess the empirical relationship between the MSE and various aspects of model complexity, such as dimension $d$, rank $r$, and signal level $\alpha=\mnormSize{}{\Theta}$.
-Figure~\ref{fig:finite}a plots the estimation error versus the tensor dimension $d$ for three different ranks $r\in\{3,5,8\}$.
-The decay in the error appears to behave on the order of $d^{-2}$, which is consistent with our theoretical results~\eqref{eq:rate}.
-We find that a higher rank leads to a larger error, as reflected by the upward shift of the curve as $r$ increases.
-Indeed, a higher rank implies the higher number of parameters to estimate, thus increasing the difficulty of the estimation.
-Figure~\ref{fig:finite}b shows the estimation error versus the signal level under $d=20$.
-Interestingly, a larger estimation error is observed when the signal is either too small or too large.
-The non-monotonic behavior may seem surprising, but this is an intrinsic feature in the estimation with ordinal data.
-In view of the latent-variable interpretation (see Section~\ref{sec:latent}), estimation from ordinal observation can be interpreted as an inverse problem of quantization.
-Therefore, the estimation error diverges in the absence of noise $\tE$, because it is impossible to distinguish two different signal tensors, e.g., $\Theta_1=\ma_1\otimes \ma_2\otimes \ma_3$ and $\Theta_2=\text{sign}(\ma_1)\otimes \text{sign}(\ma_2)\otimes \text{sign}(\ma_3)$, from the quantized observations.
-This phenomenon~\citep{davenport2014,sur2019modern} is clearly contrary to the classical continuous-valued tensor problem.
-
-The second experiment investigates the incomplete observations.
-We consider $L$-level tensors with $d=20$, $\alpha=10$ and choose a subset of tensor entries via uniform sampling.
-Figure~\ref{fig:finite}c shows the estimation error of $\hat \Theta$ versus the fraction of observation $\rho=|\Omega|/d^K$.
-As expected, the error reduces with increased $\rho$ or decreased $r$.
-Figure~\ref{fig:finite}d evaluates the impact of ordinal levels $L$ to estimation accuracy, under the setting $\rho=0.5$.
-An improved performance is observed as $L$ grows, especially from binary observations ($L=2$) to multi-level ordinal observations ($L\geq 3$).
-The result showcases the benefit of multi-level observations compared to binary observations.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=0.7\textwidth]{panel.pdf}
-\caption{Empirical relationship between (relative) MSE versus (a) dimension $d$, (b) signal level $\alpha$, (c) observation fraction $\rho$, and (d) number of ordinal levels $L$.
-In panels (b)-(d), we plot the relative MSE $=\FnormSize{}{\hat \Theta-\trueT}/\FnormSize{}{\trueT}$ for better visualization.}
-\label{fig:finite}
-\end{figure}
-
-\subsection{Comparison with alternative methods}\label{sec:compare}
-Next, we compare our ordinal tensor method ({\bf Ordinal-T}) with three popular low-rank methods:
-
-\begin{itemize}
-\item Continuous tensor decomposition ({\bf Continuous-T})~\citep{acar2010scalable} is a low-rank approximation method based on classical Tucker model.
-\item One-bit tensor completion ({\bf 1bit-T})~\citep{ghadermarzy2018learning} is a max-norm penalized tensor learning method based on partial binary observations.
-\item Ordinal matrix completion ({\bf Ordinal-M})~\citep{bhaskar2016probabilistic} is a rank-constrained matrix estimation method based on noisy, quantized observations.
-\end{itemize}
-
-We apply each of the above methods to $L$-level ordinal tensors $\tY$ generated from model~\eqref{eq:model}.
-The {\bf Continuous-T} is applied directly to $\tY$ by treating the $L$ levels as continuous observations.
-The {\bf Ordinal-M} is applied to the matrix $\tY_{(1)}$ obtained via 1-mode unfolding.
-The {\bf 1bit-T} is applied to $\tY$ in two ways.
-The first approach ({\bf 1bit-sign-T}) follows from~\citet{ghadermarzy2018learning} that transforms $\tY$ to a binary tensor, by taking the entrywise sign of the mean-adjusted tensor, $\tY- |\Omega|^{-1}\sum_\omega y_\omega$.
-The second approach ({\bf 1bit-category-T}) transforms the order-3 ordinal tensor $\tY$ to an order-4 binary tensor $\tY^{\sharp}=\entry{y^{\sharp}_{ijkl}}$ via dummy variable encoding; i.e., $y^{\sharp}_{ijk\ell}=\mathds{1}_{\{y_{ijk}=\ell\}}$ for all $\ell\in[L-1]$.
-
-We evaluate the methods by their capabilities in predicting the most likely label for each entry, i.e.,\ $y_\omega^{\text{mode}}=\arg\max_\ell\mathbb{P}(y_\omega=\ell)$.
-Two performance metrics are considered: mean absolute deviation, $\text{MAD}(\tY^{\text{mode}}, \hat \tY^{\text{mode}})=d^{-K}\sum_\omega |y_\omega^{\text{mode}}-\hat y_\omega^{\text{mode}}|$, and misclassification rate, $\text{MCR}(\tY^{\text{mode}}, \hat \tY^{\text{mode}})=d^{-K}\sum_\omega\mathds{1}_{\{y_\omega^{\text{mode}}\neq\text{round}(\hat y_\omega^{\text{mode}})\}}$,
-where $\text{round}(\cdot)$ denotes the nearest integer of the prediction (possibly continuous-valued returned by {\bf Continuous-T}).
-Both metrics are widely used for evaluation of prediction accuracy.
-Note that MAD penalizes the large deviation more heavily than MCR.
-
-Figure~\ref{fig:compare} compares the prediction accuracy under the setting $\alpha=10$, $d=20$, and $r=5$.
-The problem size we considered is comparable to~\citet{ghadermarzy2018learning}.
-We find that our method outperforms the others in both MAD and MCR.
-In particular, methods built on multi-level observations ({\bf Ordinal-T}, {\bf Ordinal-M}, {\bf 1bit-category-T}) exhibit stable MCR over $\rho$ and $L$, whereas the others two methods ({\bf Continuous-T}, {\bf 1bit-sign-T}) generally fail except for $L=2$ (Figures~\ref{fig:compare}a-b).
-This observation highlights the necessity of modeling multi-level probabilities in classification task.
-Interestingly, although both {\bf 1bit-category-T} and our method {\bf Ordinal-T} behave similarly for binary tensors ($L=2$), the improvement of our method is substantial as $L$ increases (Figures~\ref{fig:compare}a and~\ref{fig:compare}c).
-One possible reason is that our method incorporates the intrinsic ordering among the $L$ levels via proportional odds assumption~\eqref{eq:logodd}, whereas {\bf 1bit-category-T} ignores the ordinal structure and dependence among the induced binary entries.
-Figures~\ref{fig:compare}c-d assess the prediction accuracy with sample size.
-We see a clear advantage of our method ({\bf Ordinal-T}) over the matricization ({\bf Ordinal-M}) in both complete and non-complete observations.
-When the observation fraction is small, e.g., ${|\Omega|/ d^K}=0.4$, the tensor-based completion shows $\sim$ 30\% reduction in error compared to the matricization.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=0.7\textwidth]{panel2.pdf}
-\caption{Performance comparison for predicting most likely labels.
-(a, c) Prediction error versus sample complexity $\rho=|\Omega|/d^K$ when $L=5$.
-(b, d) Prediction error versus the number of ordinal levels $L$ when $\rho=0.8$.
-}\label{fig:compare}
-\end{figure}
-
+\begin{rmk} We discuss the impact of imbalance between labels to the estimation accuracy.
+Consider the special case $d_1=\cdots =d_K=d$.
+The bounds~\eqref{eq:jointTheta} and~\eqref{eq:jointb} demonstrate that, both $\hat \Theta$ and $\hat \mb$ achieve consistency as long as $\min_\ell(n_\ell+n_{\ell+1})$ is slightly larger than $\tO(d^{(K+1)/2})$.
+Note that in this case ${\min_\ell(n_\ell+n_{\ell+1}) \over \max_\ell n_{\ell}}\asymp {\min_\ell(n_\ell+n_{\ell+1})\over d^K} \to 0$, indicating a long-run imbalance between labels.
+The obtained consistency highlights the robustness of our proposed estimator~\eqref{eq:joint} to the label imbalance.
+\end{rmk}
+
+\begin{proof}[Proof of Theorem~\ref{thm:ratejoint}]
+The constant bounds with $\alpha$ and $\beta$ can be obtained trivially from the definition of the feasible sets. We focus on proving the non-constant bounds.
+
+Similar to the proof of Theorem~\ref{thm:rate}, we suppress $\Omega$ in the subscript. Based on the definition of $(\hat \Theta, \hat \mb)$,
+we have the following inequalities:
+\begin{equation}\label{eq:level}
+\flogl(\hat \Theta,\hat \mb)\geq \flogl(\trueT,\hat \mb)\quad \text{and}\quad \flogl(\hat \Theta,\hat \mb)\geq \flogl(\hat \Theta, \trueb).
+\end{equation}
+Following the same argument as in Theorem~\ref{thm:rate} and the first inequality in~\eqref{eq:level}, we obtain that
+\begin{equation}\label{eq:ratenew}
+\FnormSize{}{\hat \Theta-\trueT}^2\leq {c_1^2U_{\alpha,\beta,\Delta}^2 r_{\max}^{K-1} \over L_{\alpha,\beta,\Delta}^2}   \sum_k d_k ,
+\end{equation}
+where $U_{\alpha,\beta,\Delta}, L_{\alpha,\beta,\Delta}>0$ are two constants defined in~\eqref{eq:UL}.
+
+Next we bound $\FnormSize{}{\hat \mb-\trueb}^2$ given $\hat\Theta$.
+We consider the profile log-likelihood $\flogl(\hat \Theta,\mb)$ as a function of $\mb\in\tB$.
+For notational convenience, we drop $\hat \Theta$ from $\flogl(\hat \Theta,\mb)$ and simply write $\flogl(\mb)$.
+It follows from the expression of $\flogl(\mb)$ that
+\begin{align}\label{eq:propertyb}
+{\partial \flogl\over \partial b_\ell}&=\sum_{\omega\in\Omega}\left[\mathds{1}_{\{y_{\omega}=\ell\}}
+\frac{\dot{f}(b_\ell-\theta_\omega)}{g_\ell(\theta_\omega)}-\mathds{1}_{\{y_{\omega}=\ell+1\}}
+\frac{\dot{f}(b_\ell-\theta_\omega)}{g_{\ell+1}(\theta_\omega)}\right],\notag\\
+{\partial^2 \flogl\over \partial b^2_\ell}&=\sum_{\omega\in\Omega}\left[\mathds{1}_{\{y_\omega=\ell\}}\frac{\ddot{f}(b_\ell-\theta_\omega)g_\ell(\theta_\omega)-\dot{f}^2(b_\ell-\theta_\omega)}{g^2_\ell(\theta_\omega)}-\mathds{1}_{\{y_\omega=\ell+1\}}\frac{\ddot{f}(b_\ell-\theta_\omega)g_{\ell+1}(\theta_\omega)+\dot{f}^2(b_\ell-\theta_\omega)}{g^2_{\ell+1}(\theta_\omega)}\right],\notag\\
+&\quad  \text{ for all } \ell \in [L-1],\notag \\
+{\partial^2 \flogl\over \partial b_\ell \partial b_{\ell+1}}&=\sum_{\omega\in\Omega}\mathds{1}_{\{y_\omega=\ell+1\}}\frac{\dot{f}(b_\ell-\theta_\omega)\dot{f}(b_{\ell+1}-\theta_\omega)}{g^2_{\ell+1}(\theta_\omega)}\quad \text{and}\quad
+\ {\partial^2 \flogl\over \partial b_\ell \partial b_\ell'} = 0 \text{ if } |\ell-\ell'|>1.
+\end{align}
+Therefore, all entries in $\fploglb$ are upper bounded by $\left\{C_{\alpha,\beta,\Delta} \max_\ell (n_\ell+n_{\ell+1})\right\}>0$, and $\fpploglb$ is a tridiagonal matrix.
+
+By the second-order Taylor's expansion of $\flogl(\mb)$ around $\trueb$, we obtain
+\begin{equation}\label{eq:taylorb}
+\flogl(\hat\mb)=\flogl(\trueb)+(\trueb-\hat\mb)^T\fploglb(\trueb)+ {1\over 2}(\trueb-\hat\mb)^T\fpploglb(\check\mb)(\trueb-\hat\mb),
+\end{equation}
+where $\check\mb=\gamma\trueb+(1-\gamma)\hat\mb$ for some $\gamma\in[0,1]$, and $\fpploglb(\check\mb)$ denotes the $(L-1)$-by-$(L-1)$ Hession matrix evaluated at $\check\mb$.
+
+The linear term in~\eqref{eq:taylorb} can be bounded by Cauchy-Schwartz inequality,
+\begin{align}\label{eq:linearb}
+(\trueb-\hat\mb)^T\fploglb(\trueb)&\leq \FnormSize{}{\trueb-\hat\mb}\FnormSize{}{\fploglb(\trueb)}\\
+&\leq C'_{\alpha,\beta,\Delta}\FnormSize{}{\trueb-\hat\mb}\sqrt{L-1}\sqrt{\left(\sum_k d_k\right)\left( \prod_k d_k\right)},
+\end{align}
+for some constant $C'_{\alpha,\beta,\Delta}>0$, where the last inequality follows from Lemma \ref{lem:sharpergd},
+\[
+\left|{\partial \flogl\over \partial b_\ell}\bigg|_{(\hat \Theta,\trueb)}\right|\leq C'_{\alpha,\beta,\Delta}\sqrt{\left(\sum_k d_k\right)\left(\prod_k d_k\right)},\quad \text{ for all } \ell \in [L-1] \text{ with very high probability}.
+\]
+
+We next bound the quadratic term in \eqref{eq:taylorb}. Note that
+\begin{align}\label{eq:quadraticb}
+ &(\trueb-\hat\mb)^T \fpploglb(\check{\mb})(\trueb-\hat\mb)\\
+ &=\sum_{\ell\in [L-1]} \left( {\partial^2\tL_{\tY}\over \partial b^2_\ell} \Big|_{\mb=\check\mb} \right)(\hat b_\ell-b^{\text{true}}_{\ell})^2 + 2\sum_{\ell\in[L-1]/\{1\}}\left( {\partial^2\tL_{\tY}\over \partial b_\ell \partial b_{\ell-1}} \Big|_{\mb=\check\mb} \right)(\hat b_\ell-b^{\text{true}}_{\ell})(\hat b_{\ell-1}-b^{\text{true}}_{\ell-1}) \nonumber \\
+&\leq \sum_{\ell\in [L-1]} \left( {\partial^2\tL_{\tY}\over \partial b^2_\ell} \Big|_{\mb=\check\mb} \right)(\hat b_\ell-b^{\text{true}}_{\ell})^2 + \sum_{\ell\in[L-1]/\{1\}}\left( {\partial^2\tL_{\tY}\over \partial b_\ell \partial b_{\ell-1}} \Big|_{\mb=\check\mb} \right)\big[(\hat b_\ell-b^{\text{true}}_{\ell})^2 +(\hat b_{\ell-1}-b^{\text{true}}_{\ell-1})^2\big] \nonumber \\
+&=\left( {\partial^2\tL_{\tY}\over \partial b^2_1}+{\partial^2\tL_{\tY}\over \partial b_1 \partial b_2 }\right)\Big|_{\mb=\check\mb}(\hat b_1-b^{\text{true}}_{1})^2+\left( {\partial^2\tL_{\tY}\over \partial b^2_{L-1}}+{\partial^2\tL_{\tY}\over \partial b_{L-2} \partial b_{L-1} }\right)\Big|_{\mb=\check\mb}(\hat b_{L-1}-b^{\text{true}}_{L-1})^2  \\
+&\quad\quad +\sum_{\ell\in[L-2]/\{1\}}\left( {\partial^2\tL_{\tY}\over \partial b^2_\ell}+{\partial^2\tL_{\tY}\over \partial b_\ell \partial b_{\ell-1}}+{\partial^2\tL_{\tY}\over \partial b_{\ell+1} \partial b_{\ell}} \right)\Big|_{\mb=\check\mb} (\hat b_\ell-b^{\text{true}}_{\ell})^2     \\
+&\leq - D_{\alpha,\beta,\Delta}' \sum_{\ell \in [L-1]}(\hat b_\ell-b_{\text{true},\ell})^2 \nonumber \\
+&=-D_{\alpha,\beta,\Delta}'\FnormSize{}{\hat\mb-\trueb}^2,
+\end{align}
+where
+\begin{align}
+D_{\alpha,\beta,\Delta}'&=\min_{\substack{|\theta|\leq \alpha,\\ \mb\in\tB}}\min_{\ell\in[L-1]} -\left( {\partial^2\tL_{\tY}\over \partial b^2_\ell}+{\partial^2\tL_{\tY}\over \partial b_\ell \partial b_{\ell-1}}+{\partial^2\tL_{\tY}\over \partial b_{\ell+1} \partial b_{\ell}} \right)\\
+&=\min_{\substack{|\theta|\leq \alpha,\\ \mb\in\tB}} \min_{\ell\in[L-1]}\left\{\sum_{\omega\in\Omega}-\mathds{1}_{\{y_\omega=\ell\}}\left( {\ddot{f}(b_{\ell}-\theta_\omega) g_\ell(\theta_\omega)- \dot{f}(b_{\ell}-\theta_\omega)\dot{g}_\ell(\theta_\omega) \over g^2_\ell(\theta_\omega) }\right)\right.
+\\
+&\left.
+\hspace{1in}+\sum_{\omega\in\Omega}\mathds{1}_{\{y_\omega=\ell+1\}} \left( {\ddot{f}(b_\ell-\theta_\omega) g_{\ell+1}(\theta_\omega) - \dot{f}(b_\ell-\theta_\omega)\dot{g}_{\ell+1} (\theta_\omega)\over g^2_{\ell+1}(\theta_\omega)} \right)\right\}\\
+&=\min_{\substack{|\theta|\leq \alpha,\\ \mb\in\tB}} \min_{\ell\in[L-1]}\left\{\KeepStyleUnderBrace{-\sum_{\omega\in\Omega} \mathds{1}_{\{y_\omega=\ell\}}{\partial \over \partial \theta_\omega}\left( \dot{f}(b_\ell-\theta_\omega) \over g_\ell(\theta_\omega) \right)}_{>0}+\KeepStyleUnderBrace{\sum_{\omega\in\Omega}\mathds{1}_{\{y_\omega=\ell+1\}} {\partial \over \partial \theta_\omega}\left( \dot{f}(b_\ell-\theta_\omega) \over g_{\ell+1}(\theta_\omega) \right) }_{>0}\right\}\\
+&\geq D_{\alpha,\beta,\Delta}  \min_{\ell\in[L-1]}\left(n_{\ell}+n_{\ell+1}\right) .
+\end{align}
+Combining inequalities~\eqref{eq:taylorb}, \eqref{eq:linearb} and~\eqref{eq:quadraticb} yields
+\[
+\flogl(\hat \mb)\leq \flogl(\trueb)+C'_{\alpha,\beta,\Delta} \sqrt{(L-1)\left(\sum_k d_k\right)\left(\prod_k d_k\right)} \FnormSize{}{\hat\mb-\trueb} -\frac{D_{\alpha,\beta,\Delta}}{2}\min_{\ell}(n_\ell+n_{\ell+1}) \FnormSize{}{\hat\mb-\trueb}^2 .
+\]
+Since $\hat \mb$ satisfies $\flogl(\hat \mb)-\flogl(\trueb)\geq 0$, we have that
+\[
+C'_{\alpha,\beta,\Delta} \sqrt{(L-1)\left(\sum_k d_k\right)\left(\prod_k d_k\right)} \FnormSize{}{\hat\mb-\trueb} -\frac{D_{\alpha,\beta,\Delta}}{2}\min_{\ell}(n_\ell+n_{\ell+1}) \FnormSize{}{\hat\mb-\trueb}^2\geq 0.
+\]
+Finally,
+\begin{equation}\label{eq:rateb}
+\FnormSize{}{\hat\mb-\trueb}^2\leq {C'_{\alpha,\beta,\Delta} (L-1) \over D_{\alpha,\beta,\Delta}^2}{\left(\sum_k d_k\right)\left(\prod_k d_k\right)\over \min_{\ell}(n_{\ell}+n_{\ell+1})^2}.
+\end{equation}
+Combing~\eqref{eq:ratenew} and~\eqref{eq:rateb} with the following inequality completes the proof.
+\begin{align}
+\text{MSE}\left((\hat \Theta,\hat \mb),(\trueT,\trueb)\right)&\leq{1\over \prod_k d_k}\FnormSize{}{(\hat \Theta,\hat \mb)-(\trueT,\trueb)}^2\\
+&={1\over \prod_k d_k}\FnormSize{}{(\hat \Theta,\hat \mb)-(\hat \Theta,\trueb)+(\hat \Theta,\trueb)-(\trueT,\trueb)}^2\\
+&\leq {1\over\prod_kd_k}\FnormSize{}{(\hat \Theta- \trueT)}^2+ {1\over\prod_kd_k}\FnormSize{}{\hat \mb -\trueb}^2.
+\end{align}
+\end{proof}
+
+\begin{lem}[CLT for independent Bernoulli r.v.'s]\label{lem:clt}
+Let $\{X_n\}$ be a series of independent Bernoulli random variables with possibly different success probabilities $\{p_n\}$.
+Suppose the success probabilities $\{p_n\}$ are bounded, in that, $c_1\leq p_n\leq 1-c_2$ for some positive constants $c_1,c_2>0$ and all $n\in[m]$.
+Define $s^2_m = \sum_{n=1}^m p_n(1-p_n)$ and $Y_n = X_n-p_n$.
+Then,
+\[
+{1\over s_m}\sum_{n=1}^m Y_n \stackrel{\mathcal{D}}{\rightarrow} N(0,1), \quad \text{ as } m \rightarrow \infty.
+\]
+\end{lem}
+
+\begin{proof}
+We apply Lyapunov central limit theorem (CLT) to $\{Y_n\}$. Let us verify Lyapunov's condition for $\delta = 1$:
+\[
+\mathbb{E}[|Y_n|^3] = p_n(1-p_n)^3+ (1-p_n)p_n^3\leq p_n(1-p_n)[(1-p_n)^2+p_n^2]\leq p_n(1-p_n).
+\]
+Summation of the above inequality shows
+\[
+\sum_{n=1}^m\mathbb{E}[|Y_n|^3]\leq \sum_{n=1}^m p_n(1-p_n) = s_m^2.
+\]
+Thus, the Lyapunov condition is satisfied whenever
+\[
+{1\over s_m^3}\sum_{n=1}^m \mathbb{E}[|Y_n|^3]\leq {s_m^2\over s_m^3}\rightarrow 0,
+\]
+or simply $s_m = \sum_{n=1}^m p_n(1-p_n) \geq m\min\{c_1(1-c_1), c_2(1-c_2)\} \rightarrow \infty.$ Applying Lyapunov CLT to $\{Y_n\}$ completes the proof.
+\end{proof}
+
+\begin{lem}[Bound on score function]\label{lem:sharpergd}
+Consider the same set-up as in Theorem~\ref{thm:ratejoint}. Let $\tL_\tY(\Theta,\mb)$ denote the log-likelihood function of $(\Theta, \mb)$ given data tensor $\tY$. Then, with very high probability,
+\[
+\left|{\partial \flogl\over \partial b_\ell}\bigg|_{(\Theta,\trueb)}\right|\leq \left(C_{\alpha,\beta,\Delta}+\tilde{D}_{\alpha,\beta,\Delta}\FnormSize{}{\trueT-\Theta}^2\right)\sqrt{\prod_k d_k},\quad \text{ for all } \ell \in [L-1],
+\]
+In particular, with very high probability,
+\[
+\left|{\partial \flogl\over \partial b_\ell}\bigg|_{(\hat \Theta,\trueb)}\right|\leq C'_{\alpha,\beta,\Delta}\sqrt{\left(\sum_k d_k\right)\left( \prod_k d_k\right)},\quad \text{ for all } \ell \in [L-1],
+\]
+where $C'_{\alpha,\beta,\Delta}>0$ is a constant that depends on $U_{\alpha,\beta,\Delta}, L_{\alpha,\beta,\Delta}, C_{\alpha,\beta,\Delta},  \tilde D_{\alpha,\beta,\Delta}$ defined in~\eqref{eq:UL} and~\eqref{eq:CD}.
+\end{lem}
+
+\begin{proof}
+We only prove the case for $\ell = 1$. Other cases can be proved similarly.
+
+\begin{equation}\label{eq:sharpergd}
+  {\partial \flogl\over \partial b_1}\bigg|_{(\Theta,\trueb)} = \KeepStyleUnderBrace{{\partial \flogl\over \partial b_1}\bigg|_{( \Theta,\trueb)}-\mathbb{E}_\mathcal{Y}\left[{\partial \flogl\over \partial b_1}\bigg|_{(\Theta,\trueb)}\right]}_{\coloneqq A}+\KeepStyleUnderBrace{\mathbb{E}_\mathcal{Y}\left[{\partial \flogl\over \partial b_1}\bigg|_{( \Theta,\trueb)}\right]-\mathbb{E}_\mathcal{Y}\left[{\partial \flogl\over \partial b_1}\bigg|_{(\trueT,\trueb)}\right] }_{\coloneqq B}.
+\end{equation}
+We have used the fact that the score function has mean zero, $\mathbb{E}_\mathcal{Y}\left[{\partial \flogl\over \partial b_1}\big|_{(\trueT,\trueb)}\right]=0 $.
+Here all expectations are taken with respect to $\mathcal{Y}\sim \mathbb{P}(\trueT,\trueb).$\\
+We now bound the two deviation terms in \eqref{eq:sharpergd} separately. The term $A$ in \eqref{eq:sharpergd} is the stochastic deviation of log-likelihood to its expectation:
+\begin{align}
+A &= \sum_{\omega\in\Omega}\left\{[\mathds{1}_{y_\omega = 1}-g_1(\theta^{\text{true}}_\omega)]{\dot{f}(b_1-\theta_\omega)\over g_1(\theta_\omega)}-[\mathds{1}_{y_\omega = 2}-g_2(\theta^{\text{true}}_\omega)]{\dot{f}(b_1-\theta_\omega)\over g_2(\theta_\omega)}\right\}\\
+&\leq C_{\alpha,\beta,\Delta}\sum_{\omega\in\Omega}\KeepStyleUnderBrace{[\mathds{1}_{y_\omega = 1}+\mathds{1}_{y_\omega = 2}- g_1(\theta^{\text{true}}_\omega)-g_2(\theta^{\text{true}}_\omega)]}_{\coloneqq W_\omega}.
+\end{align}
+Note that $\{W_\omega\}$ are independent, centered Bernoulli random variables with bounded success probabilities $g_1(\theta^{\text{true}}_\omega)+g_2(\theta^{\text{true}}_\omega)$.
+By Lemma \ref{lem:clt}, we have
+\[
+\sum_{\omega\in\Omega}W_\omega \stackrel{\mathcal{D}}{\rightarrow} N\left(0,\ \sum_{\omega\in\Omega}\left(g_1(\theta^{\text{true}}_\omega)+g_2(\theta^{\text{true}}_\omega)\right)\left(1-g_1(\theta^{\text{true}}_\omega)-g_2(\theta^{\text{true}}_\omega)\right)\right).
+\]
+Hence, with the fact that $\sum_{\omega\in\Omega}\left(g_1(\theta^{\text{true}}_\omega)+g_2(\theta^{\text{true}}_\omega)\right)\left(1-g_1(\theta^{\text{true}}_\omega)-g_2(\theta^{\text{true}}_\omega)\right)\leq {1\over 4}\prod_k d_k$,
+\begin{equation}\label{eq:A}
+|A| \leq C_{\alpha,\beta,\Delta}\left|\sum_{\omega\in\Omega}W_\omega\right|\leq C_{\alpha,\beta,\Delta}\sqrt{\prod_k d_k}, \quad \text{ with very high probability. }
+\end{equation}
+The second term $B$ in \eqref{eq:sharpergd} is the bias induced by $\Theta$:
+\begin{align}
+|B| &= \left|\sum_{\omega\in\Omega} g_1(\theta^{\text{true}}_\omega)\left({\dot{f}(b_1-\theta_\omega)\over g_1(\theta_\omega)}-{\dot{f}(b_1-\theta^{\text{true}}_\omega)\over g_1(\theta^{\text{true}}_\omega)}\right)-\sum_{\omega\in\Omega} g_2(\theta^{\text{true}}_\omega)\left({\dot{f}(b_1-\theta_\omega)\over g_2(\theta_\omega)}-{\dot{f}(b_1-\theta^{\text{true}}_\omega)\over g_2(\theta^{\text{true}}_\omega)}\right)\right|\\
+&= \left|\sum_{\omega\in\Omega} g_1(\theta^{\text{true}}_\omega)(\theta^{\text{true}}_\omega-\theta_\omega)\left\{{\partial\over\partial \theta}\left({\dot{f}(b_1-\theta)\over g_1(\theta)}\right)\bigg|_{\rho_\omega\theta_\omega+(1-\rho_\omega)\theta^{\text{true}}_\omega}\right\}\right.\\
+&\hspace{0.5in}\left.-\sum_{\omega\in\Omega} g_2(\theta^{\text{true}}_\omega)(\theta^{\text{true}}_\omega-\theta_\omega)\left\{{\partial\over\partial \theta}\left({\dot{f}(b_1-\theta)\over g_2(\theta)}\right)\bigg|_{\rho_\omega'\theta_\omega+(1-\rho_\omega')\theta^{\text{true}}_\omega}\right\}\right|\\
+&\leq \tilde{D}_{\alpha,\beta,\Delta}\sum_{\omega\in\Omega}|g_1(\theta^{\text{true}}_\omega)+g_2(\theta^{\text{true}}_\omega)||\theta^{\text{true}}_\omega-\theta_\omega|.
+\end{align}
+By Cauchy-Schwartz inequality with the fact that $g_1(\theta^{\text{true}}_\omega)+g_2(\theta^{\text{true}}_\omega)\leq 1$,
+\begin{equation}\label{eq:B}
+|B| \leq \tilde{D}_{\alpha,\beta,\Delta}\sqrt{\prod_k d_k} \FnormSize{}{\trueT-\Theta}.
+\end{equation}
+Plugging \eqref{eq:A} and \eqref{eq:B} back to \eqref{eq:sharpergd} yields that
+\[
+{\partial \flogl\over \partial b_1}\bigg|_{(\Theta,\trueb)}\leq \left(C_{\alpha,\beta,\Delta}+\tilde{D}_{\alpha,\beta,\Delta}\FnormSize{}{\trueT-\Theta}\right)\sqrt{\prod_k d_k}.
+\]
+holds with very high probability. The second inequality in the lemma comes from~\eqref{eq:ratenew} that $\FnormSize{}{\trueT-\hat\Theta}\leq \tO(\sqrt{\sum_k d_k}).$
+\end{proof}
+
+\section{Additional results on simulations}
 We also compare the methods by their performance in predicting the median labels, $y_\omega^{\text{median}}=\min\{\ell\colon\mathbb{P}(y_\omega=\ell)\geq 0.5\}$.
@@
 \begin{center}
-\includegraphics[width=.7\textwidth]{panel3.pdf}
+\includegraphics[width=.65\textwidth]{panel3.pdf}
 \end{center}
@@
 
-\section{Data Applications}
-We apply our ordinal tensor method to two real-world datasets.
-In the first application, we use our model to analyze an ordinal tensor consisting of structural connectivities among 68 brain nodes for 136 individuals from Human Connectome Project (HCP)~\citep{van2013wu}.
-In the second application, we perform tensor completion to an ordinal dataset with missing values.
-The data tensor records the ratings on a scale of 1 to 5 from 42 users to 139 songs on 26 contexts~\citep{baltrunas2011incarmusic}.
-
-\subsection{Human Connectome Project (HCP)}
-
-Each entry in the HCP dataset takes value on a nominal scale, \{{\it high, moderate, low}\}, indicating the strength level of fiber connection.
-We convert the dataset to a 3-level ordinal tensor $\tY\in[3]^{68\times 68\times 136}$ and apply the ordinal tensor method with a logistic link function.
-The BIC suggests $\mr = (23,23,8)$ with $\logl(\hat{\Theta},\hat{\mb}) = -216,646$.
-Based on the estimated Tucker factors $\{\hat \mM_k\}$, we perform a clustering analysis via K-mean on the brain nodes (see Appendix~\ref{sec:additionalHCP} for detailed procedure).
-The 68 brain nodes are grouped into 11 clusters.
+\section{Additional explanations of HCP analysis}\label{sec:additionalHCP}
+We perform clustering analyses based on the Tucker representation of the estimated tensor parameter $\hat\Theta$.
+The procedure is motivated from the higher-order extension of Principal Component Analysis (PCA) or Singular Value Decomposition (SVD).
+Recall that, in the matrix case, we usually perform clustering on an $m\times n$ (normalized) matrix $X$ based on the following procedure.
+First, we factorize $X$ into
+\begin{equation}
+X = U\Sigma V^T,
+\end{equation}
+where $\Sigma$ is a diagonal matrix and $ U,V$ are factor matrices with orthogonal columns.
+Second, we take each column of $V$ as a principal axis and each row in $U\Sigma$ as principal component.
+A subsequent multivariate clustering method (such as $K$-means) is then applied to the $m$ rows of $U\Sigma$.
+
+We apply a similar clustering procedure to the estimated parameter tensor $\hat\Theta$. Based on Tucker representation of $\hat \Theta$, we have
+\begin{equation}\label{eq:Tuckerest}
+\hat \Theta = \hat \tC\times_1\hat \mM_1\times_2\cdots\times_K\hat \mM_K,
+\end{equation}
+where $\hat \tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is the estimated core tensor, $\hat \mM_k\in\mathbb{R}^{d_k\times r_k}$ are estimated factor matrices with orthogonal columns, and $\times_k$ denotes the tensor-by-matrix multiplication~\citep{kolda2009tensor}.
+The mode-$k$ matricization of~\eqref{eq:Tuckerest} gives
+\begin{equation}
+\hat \Theta_{(k)} = \hat \mM_k\hat \tC_{(k)}\left(\hat \mM_K\otimes\cdots\otimes\hat \mM_1\right),
+\end{equation}
+where $\hat \Theta_{(k)} , \hat \tC_{(k)}$ denote the mode-$k$ unfolding of $\hat \Theta$ and $\hat \tC$, respectively.
+Then, the mode-$k$ clustering can be performed as follows.
+First, we take columns in $\left(\hat \mM_K\otimes\cdots\otimes\hat \mM_1\right)$ as principal axes and rows in $\hat \mM_k\hat \tC_{(k)}$ as principal components.
+Then, we perform $K$-means clustering method to the $d_k$ rows of the matrix $\hat \mM_k\hat \tC_{(k)}$.
+
+We perform a clustering analysis on the 68 brain nodes using the produce described above.
+Our ordinal tensor method outputs the estimated parameter tensor $\hat\Theta\in\mathbb{R}^{68\times 68\times136}$ with rank $(23,23,8)$.
+We apply $K$-means to the mode-1 principal component matrix of size $68\times 184$ ($184=23\times8$).
+The elbow method suggests 11 clusters among the 68 nodes (see Figure~\ref{figure:elbow}).
 We find that the clustering captures the spatial separation between brain regions very well (Table~\ref{table:clustering}).
@@
 
+\begin{figure}[ht]
+\begin{center}
+\includegraphics[width=0.4\textwidth]{elbowm.pdf}
+\end{center}
+\caption{Elbow plot for determining the number of clusters in $K$-means.}\label{figure:elbow}
+\end{figure}
+
 \begin{figure}[H]
 \centering
-\includegraphics[width = 1\textwidth]{HCP.pdf}
+\includegraphics[width = .7\textwidth]{HCP.pdf}
 \caption{Top three clusters in the HCP analysis.
@@
 
-\begin{table}[H]
+\begin{table}[ht]
 \center
@@
 
-We also compare the goodness-of-fit of various tensor methods on the HPC data.
-We perform 5-fold cross-validation while preserving the same label proportions in train/test sets.
-Table~\ref{table:CV} summarizes the prediction error averaged over 10 runs.
-Our method outperforms the others, especially in MAD.
-
-\subsection{InCarMusic recommendation system}
-
-We apply ordinal tensor completion to a recommendation system InCarMusic.
-InCarMusic is a mobile application that offers music recommendation to passengers of cars based on contexts~\citep{baltrunas2011incarmusic}.
-Our goal is to perform tensor completion to impute the unobserved entries in the $42\times 139\times 26$ ordinal tensor and thereby we can offer context-specific music recommendation to users.
-The data tensor consists of 2,884 observed entries.
-Table~\ref{table:CV} shows the averaged prediction error via 5-fold cross validation.
-The high missing rate makes the accurate classification challenging.
-Nevertheless, our method achieves the best performance among the three.
-
-\begin{table}[ht]
-\resizebox{\columnwidth}{!}{
-\begin{tabular}{c|c|c||c|c|c}
-\Xhline{2\arrayrulewidth}
-\multicolumn{3}{c||}{Human Connectome Project (HCP) dataset} & \multicolumn{3}{c}{InCarMusic dataset}\\
-\Xhline{2\arrayrulewidth}
-Method                    & MAD & MCR & Method                    & MAD & MCR\\
-\hline
-Ordinal-T (ours) & 0.1607 (0.0005)& 0.1606 (0.0005)& Ordinal-T (ours) & 1.37 (0.039) &  0.59 (0.009)\\
-Continuous-T&0.2530 (0.0002)&0.1599 (0.0002)& Continuous-T & 2.39 (0.152)&0.94 (0.027)\\
-1bit-sign-T&0.3566 (0.0010)&0.1563 (0.0010)& 1bit-sign-T&1.39 (0.003)& 0.81 (0.005)\\
-\hline
-\end{tabular}
-}
-\caption{Comparison of prediction error in the HPC and InCarMusic analyses. Standard errors are reported in parentheses.}\label{table:CV}
-\end{table}
-
-\section{Proofs}\label{sec:proof}
-Here, we provide proofs of the theoretical results presented in Sections~\ref{sec:theory}.
-
-\subsection{Estimation error for tensor denoising}\label{sec:proofMSE}
-\begin{proof}[Proof of Theorem~\ref{thm:rate}]
-We suppress the subscript $\Omega$ in the proof, because the tensor denoising assumes complete observation $\Omega=[d_1]\times \cdots \times [d_K]$.
-It follows from the expression of $\flogl(\Theta)$ that
-\begin{align}\label{eq:property}
-{\partial \flogl\over \partial \theta_\omega}&=\sum_{\ell\in[L]}\mathds{1}_{\{y_{\omega}=\ell\}}
-{\dot{g}_\ell(\theta_\omega)\over g_\ell(\theta_\omega)},\notag\\
-{\partial^2 \flogl\over \partial \theta_\omega^2}&=\sum_{\ell\in[L]}\mathds{1}_{\{y_\omega=\ell\}}{\ddot{g}_\ell(\theta_\omega)g_\ell(\theta_\omega)-\dot{g}^2_\ell(\theta_\omega)\over g^2_\ell(\theta_\omega)}\ \text{and}\quad
-{\partial^2 \flogl\over \partial \theta_\omega \theta_\omega'}=0\ \text{if}\ \omega\neq \omega',
-\end{align}
-for all $\omega\in[d_1]\times \cdots \times [d_K]$.
-Define $d_{\text{total}}=\prod_k d_k$.
-Let $\fplogl\in\mathbb{R}^{d_1\times\cdots\times d_K}$ denote the tensor of gradient with respect to $\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}$, and $\fpplogl$ the corresponding Hession matrix of size $d_\text{total}$-by-$d_{\text{total}}$.
-Here, $\Vec(\cdot)$ denotes the operation that turns a tensor into a vector.
-By~\eqref{eq:property}, $\fpplogl$ is a diagonal matrix.
-Recall that
-\begin{equation}\label{eq:bound}
-U_\alpha=\max_{\ell\in[L],|\theta|\leq \alpha}{|\dot{g}_\ell(\theta)|\over g_\ell(\theta)}>0 \quad \text{and}\quad
-L_\alpha=\min_{\ell\in[L],|\theta|\leq \alpha} {\dot{g}^2_\ell(\theta)-\ddot{g}_\ell(\theta)g_\ell(\theta)\over g^2_\ell(\theta)}>0.
-\end{equation}
-Therefore, the entries in $\fplogl$ are upper bounded in magnitude by $U_\alpha>0$, and all diagonal entries in $\fpplogl$ are upper bounded by $-L_{\alpha}<0$.
-
-By the second-order Taylor's expansion of $\flogl(\Theta)$ around $\trueT$, we obtain
-\begin{equation}\label{eq:taylor}
-\flogl(\Theta)=\flogl(\trueT)+\langle\Vec(\fplogl),\ \Vec(\Theta-\trueT)\rangle+{1\over 2}\Vec(\Theta-\trueT)^T\fpplogl(\check\Theta)\Vec(\Theta-\trueT),
-\end{equation}
-$\check\Theta=\gamma\trueT+(1-\gamma)\Theta$ for some $\gamma\in[0,1]$, and $\fpplogl(\check\Theta)$ denotes the $\prod_kd_k$-by-$\prod_k d_k$ Hession matrix evaluated at $\check\Theta$.
-
-We first bound the linear term in~\eqref{eq:taylor}. Note that, by Lemma~\ref{lem:inq},
-\begin{equation}\label{eq:linear}
-|\langle\Vec(\fplogl(\trueT), \Vec(\Theta-\trueT)  \rangle|\leq \snormSize{}{\fplogl(\trueT)} \nnormSize{}{\Theta-\trueT},
-\end{equation}
-where $\snormSize{}{\cdot}$ denotes the tensor spectral norm and $\nnormSize{}{\cdot}$ denotes the tensor nuclear norm. Define
-\[
-s_\omega={\partial \tL_\tY\over \partial \theta_\omega}\Big|_{\Theta=\trueT} \;\; \textrm{ for all } \; \omega\in[d_1]\times\cdots\times [d_K].
-\]
-Based on~\eqref{eq:property} and the definition of $U_\alpha$, $\fplogl(\trueT)=\entry{s_{\omega}}$ is a random tensor whose entries are independently distributed satisfying
-\begin{equation}\label{eq:norm}
-\mathbb{E}(s_\omega)=0,\quad |s_\omega|\leq U_\alpha, \quad \text{for all }\omega\in[d_1]\times \cdots \times [d_K].
-\end{equation}
-By lemma~\ref{lem:noisytensor}, with probability at least $1-\exp(-C_1 \sum_kd_k)$, we have
-\begin{equation}\label{eq:normrandom}
-\snormSize{}{\fplogl(\trueT)} \leq C_2 U_\alpha\sqrt{\sum_k d_k},
-\end{equation}
-where $C_1, C_2$ are two positive constants that depend only on $K$.
-Furthermore, note that $\text{rank}(\Theta)\leq \mr$, $\text{rank}(\trueT)\leq \mr$, so $\text{rank}(\Theta-\trueT)\leq 2\mr$.
-By lemma~\ref{lem:nuclear}, $\nnormSize{}{\Theta-\trueT}\leq (2r_{\max})^{K-1\over 2}\FnormSize{}{\Theta-\trueT}$.
-Combining~\eqref{eq:linear}, \eqref{eq:norm} and \eqref{eq:normrandom}, we have that, with probability at least $1-\exp(-C_1 \sum_kd_k)$,
-\begin{equation}\label{eq:linearconclusion}
-|\langle \Vec(\fplogl(\trueT)), \Vec(\Theta-\trueT)  \rangle | \leq C_2 U_\alpha  \sqrt{r_{\max}^{K-1} \sum_k d_k}  \FnormSize{}{\Theta-\trueT}.
-\end{equation}
-
-We next bound the quadratic term in \eqref{eq:taylor}. Note that
-\begin{align}\label{eq:quadratic}
- \Vec(\Theta-\trueT)^T \fpplogl(\check{\Theta})\Vec(\Theta-\trueT)&=\sum_\omega \left( {\partial^2\tL_{\tY}\over \partial \theta^2_\omega} \Big|_{\Theta=\check\Theta} \right)(\theta_\omega-\theta_{{\text{true}},\omega})^2 \nonumber \\
-&\leq - L_\alpha\sum_{\omega}(\Theta_{\omega}-\Theta_{\text{true},\omega})^2 \nonumber \\
-&=-L_\alpha\FnormSize{}{\Theta-\trueT}^2,
-\end{align}
-where the second line comes from the fact that  $\mnormSize{}{\check\Theta}\leq \alpha$ and the definition of $L_\alpha$.
-
-Combining~\eqref{eq:taylor}, \eqref{eq:linearconclusion} and~\eqref{eq:quadratic}, we have that, for all $\Theta\in\tP$, with probability at least $1-\exp(-C_1 \sum_kd_k)$,
-\[
-\tL_\tY(\Theta)\leq \tL_{\tY}(\trueT)+C_2U_\alpha  \left(r_{\max}^{K-1}\sum_k d_k\right)^{1/2}  \FnormSize{}{\Theta-\trueT}-{L_\alpha\over 2}\FnormSize{}{\Theta-\trueT}^2.
-\]
-In particular, the above inequality also holds for $\hat \Theta\in\tP$. Therefore,
-\[
-\tL_\tY(\hat \Theta)\leq \tL_{\tY}(\trueT)+C_2U_\alpha \left(r_{\max}^{K-1}\sum_k d_k\right)^{1/2}  \FnormSize{}{\hat \Theta-\trueT}-{L_\alpha\over 2} \FnormSize{}{\hat \Theta-\trueT}^2.
-\]
-Since $\hat \Theta=\arg\max_{\Theta\in\tP}\tL_\tY(\Theta)$, $\tL_\tY(\hat \Theta)-\tL_{\tY}(\trueT)\geq 0$, which gives
-\[
-C_2U_\alpha \left(r_{\max}^{K-1}\sum_k d_k\right)^{1/2}  \FnormSize{}{\hat \Theta-\trueT}-{L_\alpha\over 2}\FnormSize{}{\hat \Theta-\trueT}^2\geq 0.
-\]
-Henceforth,
-\[
-{1\over \sqrt{\prod_k d_k}} \FnormSize{}{\hat \Theta-\trueT}\leq {2C_2U_\alpha \sqrt{r_{\max}^{K-1}\sum_k d_k}\over L_\alpha \sqrt{\prod_k d_k}}={2C_2U_\alpha r_{\max}^{(K-1)/2}\over L_\alpha} \sqrt{ \sum_k d_k \over \prod_k d_k}.
-\]
-This completes the proof.
-\end{proof}
-
-\begin{proof}[Proof of Corollary~\ref{cor:prediction}]
-The result follows immediately from Theorem~\ref{thm:rate} and Lemma~\ref{lem:KL}.
-\end{proof}
-
-\begin{proof}[Proof of Theorem~\ref{thm:minimax}]
-
-Let $d_{\text{total}}=\prod_{k\in[K]}d_k$, and $\gamma\in[0,1]$ be a constant to be specified later.
-Our strategy is to construct a finite set of tensors $\tX=\{\Theta_i\colon i=1,\ldots \}\subset \tP$ satisfying the properties of (i)-(iv) in Lemma~\ref{lem:construction}.
-By Lemma~\ref{lem:construction}, such a subset of tensors exist.
-For any tensor  $\Theta\in\tX$, let $\mathbb{P}_{\Theta}$ denote the distribution of $\tY|\Theta$, where $\tY$ is the ordinal tensor.
-In particular, $\mathbb{P}_{\mathbf{0}}$ is the distribution of $\tY$ induced by the zero parameter tensor $\mathbf{0}$, i.e., the distribution of $\tY$ conditional on the parameter tensor $\Theta=\mathbf{0}$.
-Based on the Remark for Lemma~\ref{lem:KL}, we have
-\begin{equation}\label{eq:KLbound1}
-\mathrm{KL}(\mathbb{P}_{\Theta}|| \mathbb{P}_{\mathbf{0}})\leq C \FnormSize{}{\Theta}^2,
-\end{equation}
-where $C={(4L-6) \dot{f}^2(0)\over  A_\alpha}>0$ is a constant independent of the tensor dimension and rank.
-Combining the inequality~\eqref{eq:KLbound1} with property (iii) of $\tX$, we have
-\begin{equation}\label{eq:KLbound}
-\text{KL}(\mathbb{P}_{\Theta}||\mathbb{P}_{\mathbf{0}})\leq \gamma^2 R_{\max} d_{\max}.
-\end{equation}
-From~\eqref{eq:KLbound} and the property (i), we deduce that the condition
-\begin{equation}\label{eq:totalKL}
-{1\over \text{Card}(\tX)-1}\sum_{\Theta \in\tX}\text{KL}(\mathbb{P}_{\Theta}, \mathbb{P}_{\mathbf{0}})\leq \varepsilon \log_2\left\{\text{Card}(\tX)-1 \right\}
-\end{equation}
-holds for any $ \varepsilon \geq 0$ when $\gamma\in[0,1]$ is chosen to be sufficiently small depending on $\varepsilon$, e.g., $\gamma \leq \sqrt{\varepsilon\log2\over8}$.
-By applying Lemma~\ref{lem:Tsybakov} to~\eqref{eq:totalKL}, and in view of the property (iv), we obtain that
-\begin{equation}\label{eq:final}
-\inf_{\hat \Theta}\sup_{\trueT\in \tX}\mathbb{P}\left(\FnormSize{}{\hat \Theta- \trueT}\geq  {\gamma\over 8} \min\left\{ \alpha\sqrt{d_{\text{total}}}, C^{-1/2}\sqrt{ R_{\max}d_{\max}}\right\} \right)\geq {1\over 2}\left(1-2\varepsilon-\sqrt{16 \varepsilon \over R_{\max}d_{\max}\log2}\right).
-\end{equation}
-Note that $\text{Loss}(\hat \Theta, \trueT)=\FnormSize{}{\hat \Theta- \trueT}^2/d_{\text{total}}$ and $\tX\subset \tP$.
-By taking $\varepsilon=1/10$ and $\gamma=1/11$, we conclude from~\eqref{eq:final} that
-\begin{equation}\label{eq:prob}
-\inf_{\hat \Theta}\sup_{\trueT\in \tP}\mathbb{P}\left(\text{Loss}(\hat \Theta, \trueT)\geq c\min\left \{ \alpha^2,  {C^{-1}R_{\max}d_{\max}\over d_{\text{total}}}\right \}\right)\geq {1\over 2}\left({4\over 5}- \sqrt{1.6\over R_{\max}d_{\max}\log2} \right)\geq {1\over 8},
-\end{equation}
-where $c = {1 \over 88^2}$ and the last inequality comes from the condition for $d_{\text{max}}$.
-This completes the proof.
-\end{proof}
-
-\subsection{Sample complexity for tensor completion}
-
-\begin{proof}[Proof of Theorem~\ref{thm:completion}]
-
-For notational convenience, we use $\MFnormSize{}{\Theta}=\sum_{\omega\in\Omega}\Theta^2_\omega$ to denote the sum of squared entries over the observed set $\Omega$, for a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$.
-
-Following a similar argument as in the proof of Theorem~\ref{thm:rate}, we have
-\begin{equation}\label{eq:Taylor2}
-\logl(\Theta)=\logl(\trueT)+\langle\Vec(\plogl),\ \Vec(\Theta-\trueT)\rangle+{1\over 2}\Vec(\Theta-\trueT)^T\pplogl(\check\Theta)\Vec(\Theta-\trueT),
-\end{equation}
-where
-\begin{enumerate}
-\item $\plogl$ is a $d_1\times\cdots\times d_K$ tensor with $|\Omega|$ nonzero entries, and each entry is upper bounded by $U_\alpha>0$.
-\item $\pplogl$ is a diagonal matrix of size $d_{\text{total}}$-by-$d_{\text{total}}$ with $|\Omega|$ nonzero entries, and each entry is upper bounded by $-L_{\alpha}<0$.
-\end{enumerate}
-
-Similar to~\eqref{eq:linear} and~\eqref{eq:quadratic}, we have
-\begin{equation}\label{eq:linear2}
-|\langle\Vec(\plogl),\ \Vec(\Theta-\trueT)\rangle|\leq C_2U_\alpha \sqrt{r_{\max}^{K-1}\sum_k d_k}\MFnormSize{}{\Theta-\trueT}
-\end{equation}
-and
-\begin{equation}\label{eq:quadratic2}
-\Vec(\Theta-\trueT)^T\fpplogl(\check\Theta)\Vec(\Theta-\trueT)\leq -L_\alpha \MFnormSize{}{\Theta-\trueT}^2.
-\end{equation}
-
-Combining~\eqref{eq:Taylor2}-\eqref{eq:quadratic2} with the fact that $\logl(\hat \Theta)\geq \logl(\trueT)$, we have
-\begin{equation}\label{eq:sample}
-\MFnormSize{}{\hat \Theta-\trueT}\leq {2C_2U_\alpha  r_{\max}^{(K-1)/2} \over L_\alpha} \sqrt{\sum_kd_k},
-\end{equation}
-with probability at least $1-\exp(-C_1 \sum_k d_k)$.
-Lastly, we invoke the result regarding the closeness of $\Theta$ to its sampled version $\Theta_{\Omega}$, under the entrywise bound condition.
-Note that $\mnormSize{}{\hat\Theta-\trueT}\leq 2\alpha$ and $\text{rank}(\hat \Theta-\trueT)\leq 2\mr$.
-By Lemma~\ref{lem:Mnormbound}, $\anormSize{}{\hat \Theta-\trueT}\leq 2^{(3K-1)/2} \alpha \left({\prod r_k \over r_{\max}}\right)^{3/2}$.
-Therefore, the condition in Lemma~\ref{lem:convexity} holds with $\beta=2^{(3K-1)/2}\alpha \left({\prod r_k \over r_{\max}}\right)^{3/2}$.
-Applying Lemma~\ref{lem:convexity} to~\eqref{eq:sample} gives
-\begin{align}
- \PiFnormSize{}{\hat \Theta-\trueT}^2&\leq {1\over m}\MFnormSize{}{\hat \Theta-\trueT}^2+c\beta\sqrt{\sum_k d_k\over |\Omega|}\\
- &\leq {C_2  r^{K-1}_{\max}} {\sum_k d_k \over |\Omega|}+C_1 \alpha r_{\max}^{3(K-1)/2}\sqrt{\sum_kd_k\over |\Omega|},
-\end{align}
-with probability at least $1-\exp(-{\sum_kd_k\over \sum_k \log d_k})$ over the sampled set $\Omega$. Here $C_1, C_2>0$ are two constants independent of the tensor dimension and rank. Therefore,
-\[
- \PiFnormSize{}{\hat \Theta-\trueT}^2\to 0,\quad \text{as}\quad {|\Omega|\over \sum_kd_k}\to \infty,
-\]
-provided that $r_{\max}=O(1)$.
-\end{proof}
-
-\subsection{Convexity of the log-likelihood function}\label{sec:proofconvexity}
-\begin{thm}\label{thm:convexity}
-Define the function
-\begin{align}\label{eq:function}
- \logl(\Theta, \mb)&=\sum_{\omega\in\Omega}\sum_{\ell\in[L]} \left\{\mathds{1}_{\{y_\omega=\ell\}} \log \big[f(b_\ell-\theta_\omega)-  f(b_{\ell-1}-\theta_\omega)\big]\right\},
- \end{align}
-where $f(\cdot)$ satisfies Assumption~\ref{ass:link}. Then, $\logl(\Theta, \mb)$ is concave in $(\Theta,\mb)$.
-\end{thm}
-
-\begin{proof}
-Define $d_{\text{total}}=\prod_k d_k$.
-By abuse of notation, we use $(\Theta,\mb)$ to denote the length-$(d_{\text{total}}+L-1)$-vector collecting all parameters together.
-Let us denote a bivariate function
-\begin{align}
-\lambda: &\mathbb{R}^2\mapsto \mathbb{R}\\
-(u,v)&\mapsto\lambda(u,v) = \log \big[f(u)-  f(v)\big].
-\end{align}
-It suffices to show that $\lambda(u,v)$ is concave in $(u,v)$ where $u>v$.
-
-Suppose that the claim holds (which we will prove in the next paragraph). Based on~\eqref{eq:function}, $u,v$ are both linear functions of $(\Theta,\mb)$:
-\[
-u = \ma_1^T(\Theta,\mb), \quad v = \ma_2^T(\Theta,\mb), \quad \text{ for some}\ \ma_1,\ma_2\in \mathbb{R}^{d_{\text{total}}+L-1}.
-\]
-Then, $\lambda(u,v) = \lambda(\ma_1^T(\Theta,\mb),\ \ma_2^T(\Theta,\mb))$ is concave in $(\Theta,\mb)$ by the definition of concavity.
-Therefore, we can conclude that $\logl(\Theta, \mb) $ is concave in $(\Theta,\mb)$ because $\logl(\Theta, \mb)$ is the sum of $\lambda(u,v)$.
-
-Now, we prove the concavity of $\lambda(u,v)$. Note that
-\begin{equation}
-\lambda(u,v) = \log\big[f(u)-f(v)\big]=\log\big[\int\mathds{1}_{[u,v]}(x)f'(x)dx\big],
-\end{equation}
-where $\mathds{1}_{[u,v]}$ is an indicator function that equals 1 in the interval $[u,v]$, and 0 elsewhere.
-Furthermore, $\mathds{1}_{[u,v]}(x)$ is log-concave in $(u,v,x)$, and by Assumption~\ref{ass:link}, $f'(x)$ is log-concave in $x$.
-It follows that $\mathds{1}_{[u,v]}(x)f'(x)$
-is a log-concave in $(u,v,x)$. By Lemma~\ref{lem:lossconvexity}, we conclude that $\lambda(u,v)$ is concave in $(u,v)$ where $u>v$.
-\end{proof}
-
-\begin{lem}[Corollary 3.5 in~\cite{brascamp2002extensions}]\label{lem:lossconvexity}
-Let $F(x,y):\mathbb{R}^{m+n}\rightarrow \mathbb{R}$ be an integrable function where $x\in \mathbb{R}^{m},y\in \mathbb{R}^n$. Let
-\[
-G(x) = \int_{\mathbb{R}^n}F(x,y)dy.
-\]
-If $F(x,y)$ is log concave in $(x,y)$, then $G(x)$ is log concave in $x$.
-\end{lem}
-
-\subsection{Auxiliary lemmas}\label{sec:lemma}
-This section collects lemmas that are useful for the proofs of the main theorems.
-
-\begin{defn}[Atomic M-norm~\citep{ghadermarzy2019near}]
-Define $T_{\pm}=\{\tT\in\{\pm 1 \}^{d_1\times \cdots \times d_K}\colon$ $\text{rank}(\tT)=1\}$. The atomic M-norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
-\begin{align}
-\anormSize{}{\Theta}&=\inf\{t>0: \Theta\in t\text{conv}(T_{\pm})\}\\
-&=\inf\left\{\sum_{\tX\in T_{\pm}}c_x\colon \Theta=\sum_{\tX\in T_{\pm}} c_x\tX, \ c_x>0\right\}.
-\end{align}
-\end{defn}
-
-\begin{defn}[Spectral norm~\citep{lim2005singular}]
-The spectral norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
-\[
-\snormSize{}{\Theta}=\sup\left\{\langle \Theta, \mx_1\otimes \cdots \otimes \mx_K\rangle \colon \vnormSize{}{\mx_k}=1,\ \mx_k\in\mathbb{R}^{d_k},\ \text{for all}\ k\in[K]\right\}.
-\]
-\end{defn}
-
-\begin{defn}[Nuclear norm~\citep{friedland2018nuclear}]
-The nuclear norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
-\[
-\nnormSize{}{\Theta}=\inf
-\left\{
-\sum_{i\in[r]}|\lambda_i|\colon \Theta=\sum_{i=1}^r \lambda_i\mx^{(i)}_1\otimes \cdots \otimes \mx^{(i)}_K,\ \vnormSize{}{\mx^{(i)}_k}=1,\ \mx^{(i)}_k\in\mathbb{R}^{d_k},\ \text{for all} k\in[K],\ i\in[r]
-\right\},
-\]
-where the infimum is taken over all $r\in\mathbb{N}$ and $\vnormSize{}{\mx^{(i)}_k}=1$ for all $i\in[r]$ and $k\in[K]$.
-\end{defn}
-
-\begin{lem}[M-norm and infinity norm~\citep{ghadermarzy2019near}]\label{lem:Mnormbound}
-Let $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ be an order-$K$, rank-$(r_1,\ldots,r_K)$ tensor. Then
-\[
-\mnormSize{}{\Theta}\leq \anormSize{}{\Theta}\leq \left(\prod_k r_k \over r_{\max}\right)^{3\over 2} \mnormSize{}{\Theta}.
-\]
-\end{lem}
-
-\begin{lem}[Nuclear norm and F-norm] \label{lem:nuclear}
-Let $\tA\in\mathbb{R}^{d_1\times\cdots\times d_K}$ be an order-$K$ tensor with Tucker $\text{rank}(\tA)=(r_1,\ldots,r_K)$. Then
-\[
-\nnormSize{}{\tA} \leq \sqrt{\prod_k r_k\over \max_k r_k} \FnormSize{}{\tA},
-\]
-where $\nnormSize{}{\cdot}$ denotes the nuclear norm of the tensor.
-\end{lem}
-
-\begin{proof}
-Without loss of generality, suppose $r_1=\min_k r_k$.
-Let $\tA_{(k)}$ denote the mode-$k$ matricization of $\tA$ for all $k\in[K]$.
-By \citet[Corollary 4.11]{wang2017operator}, and the invariance relationship between a tensor and its Tucker core~\citep[Section 6]{jiang2017tensor}, we have
-\begin{equation}\label{eq:norminequality}
-\nnormSize{}{\tA} \leq \sqrt{\prod_{k\geq 2} r_k \over \max_{k\geq 2} r_k} \nnormSize{}{\tA_{(1)}},
-\end{equation}
-where $\tA_{(1)}$ is a $d_1$-by-$\prod_{k\geq 2}d_k$ matrix with matrix rank $r_1$.
-Furthermore, the relationship between the matrix norms implies that $\nnormSize{}{\tA_{(1)}}\leq \sqrt{r_1}\FnormSize{}{\tA_{(1)}}=\sqrt{r_1}\FnormSize{}{\tA}$.
-Combining this fact with the inequality~\eqref{eq:norminequality} yields the final claim.
-\end{proof}
-
-\begin{lem} \label{lem:inq}
-Let $\tA, \; \tB$ be two order-$K$ tensors of the same dimension. Then
-\[
-|\langle \tA,\tB\rangle| \leq \snormSize{}{\tA}   \nnormSize{}{\tB}.
-\]
-\end{lem}
-
-\begin{proof}
-By~\citet[Proposition 3.1]{friedland2018nuclear}, there exists a nuclear norm decomposition of $\tB$, such that
-\[
-\tB=\sum_{r} \lambda_r \ma^{(1)}_r\otimes \cdots\otimes \ma^{(K)}_r,\quad \ma_r^{(k)}\in\mathbf{S}^{d_k-1}(\mathbb{R}),\quad \text{for all }k\in[K],
-\]
-and $\nnormSize{}{\tB}=\sum_{r}|\lambda_r|$. Henceforth we have
-\begin{align*}
-|\langle \tA,\tB\rangle|&=| \langle \tA, \sum_{r} \lambda_r \ma^{(1)}_r\otimes \cdots\otimes \ma^{(K)}_r \rangle|\leq \sum_r |\lambda_r| |\langle \tA, \ma^{(1)}_r \otimes \cdots\otimes \ma^{(K)}_r \rangle|\\
-&\leq \sum_{r}|\lambda_r| \snormSize{}{\tA}= \snormSize{}{\tA}\nnormSize{}{\tB},
-\end{align*}
-which completes the proof.
-\end{proof}
-
-The following lemma provides the bound on the spectral norm of random tensors. The result was firstly presented in~\cite{nguyen2015tensor}, and we adopt the version from~\cite{tomioka2014spectral}.
-\begin{lem}[Spectral norm of random tensors~\citep{tomioka2014spectral}]\label{lem:tensor}
-Suppose that $\tS=\entry{s_{\omega}}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is an order-$K$ tensor whose entries are independent random variables that satisfy
-\[
-\mathbb{E}(s_{\omega})=0,\quad \text{and} \quad\mathbb{E}(e^{ts_{\omega}})\leq e^{t^2L^2/2}.
-\]
-Then the spectral norm $\snormSize{}{\tS}$ satisfies that,
-\[
-\snormSize{}{\tS}\leq \sqrt{{8L^2} \log (12K) \sum_k d_k +\log (2/\delta)},
-\]
-with probability at least $1-\delta$.
-\end{lem}
-
-\begin{lem} \label{lem:noisytensor}
-Suppose that $\tS=\entry{s_{\omega}}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is an order-$K$ tensor whose entries are independent random variables that satisfy
-\[
-\mathbb{E}(s_{\omega})=0,\quad \text{and}\quad |s_{\omega}|\leq U.
-\]
-Then we have
-\[
-\mathbb{P}\left(\snormSize{}{\tS}\geq C_2 U\sqrt{\sum_k d_k} \right)\leq \exp\left(-C_1  \log K \sum_k d_k\right)
-\]
-where $C_1>0$ is an absolute constant, and $C_2>0$ is a constant that depends only on $K$.
-\end{lem}
-
-\begin{proof}  Note that the random variable $U^{-1}s_{\omega}$ is zero-mean and supported on $[-1,1]$. Therefore, $U^{-1}s_{\omega}$ is sub-Gaussian with parameter ${1-(-1)\over 2}=1$, i.e.
-\[
-\mathbb{E}(U^{-1}s_{\omega})=0,\quad \text{and}\quad \mathbb{E}(e^{tU^{-1}s_{\omega}})\leq e^{t^2/2}.
-\]
-It follows from Lemma~\ref{lem:tensor} that, with probability at least $1-\delta$,
-\[
-\snormSize{}{U^{-1}\tS}\leq \sqrt{\left(c_0\log K+c_1\right) \sum_k d_k +\log (2/\delta)},
-\]
-where $c_0, c_1>0$ are two absolute constants. Taking $\delta=\exp (-C_1\log K \sum_k d_k)$ yields the final claim, where $C_2=c_0\log K+c_1+1>0$ is another constant.
-\end{proof}
-
-\begin{lem}\label{lem:KLentry} Let $X,\; Y$ be two discrete random variables taking values on $L$ possible categories, with category probabilities $\{p_\ell\}_{\ell\in[L]}$ and $\{q_\ell\}_{\ell\in[L]}$, respectively.
-Suppose $p_\ell$, $q_\ell>0$ for all $i\in[L]$.
-Then, the Kullback-Leibler (KL) divergence satisfies that
-\[
-\mathrm{KL}(X||Y)\stackrel{\text{def}}{=}-\sum_{\ell\in[L]}\mathbb{P}_X(\ell)\log\left\{{\mathbb{P}_Y(\ell)\over \mathbb{P}_X(\ell)}\right\} \leq \sum_{\ell \in [L]}{(p_\ell-q_\ell)^2 \over q_\ell}.
-\]
-\end{lem}
-\begin{proof} Using the fact $\log x\leq x-1$ for $x>0$, we have that
-\begin{align}\label{eq:KL}
-\text{KL}(X||Y)&=\sum_{\ell\in[L]}p_\ell\log{p_\ell\over q_\ell}\\
-&\leq \sum_{\ell\in[L]} {p_\ell\over q_\ell}(p_\ell-q_\ell)\\
-&=\sum_{\ell\in[L]} \left({p_\ell\over q_\ell }- 1\right)(p_\ell-q_\ell)+ \sum_{\ell\in[L]} (p_\ell-q_\ell).
-\end{align}
-Note that $\sum_{\ell\in[L]}(p_\ell-q_\ell)=0$. Therefore,
-\[
-\text{KL}(X||Y)\leq \sum_{\ell\in[L]}\left( {p_\ell\over q_\ell}-1\right)\left(p_\ell-q_\ell\right)=\sum_{\ell\in[L]}{(p_\ell-q_\ell)^2\over q_\ell}.
-\]
-\end{proof}
-
-\begin{lem}[KL divergence and F-norm]~\label{lem:KL}
-Let $\tY\in[L]^{d_1\times \cdots \times d_K}$ be an ordinal tensor generated from the model~\eqref{eq:model} with the link function $f$ and parameter tensor $\Theta$.
-Let $\mathbb{P}_{\Theta}$ denote the joint categorical distribution of $\tY|\Theta$ induced by the parameter tensor $\Theta$, where $\mnormSize{}{\Theta}\leq \alpha$.
-Define
-\begin{equation}\label{eq:ass}
-A_\alpha=\min_{\ell\in[L], |\theta|\leq \alpha}\left[f(b_\ell-\theta)-f(b_{\ell-1}-\theta)\right].
-\end{equation}
-Then, for any two tensors $\Theta,\; \Theta^*$ in the parameter spaces, we have
-\[
-\mathrm{KL}(\mathbb{P}_{\Theta}|| \mathbb{P}_{\Theta^*})\leq {2(2L-3)\over A_\alpha} \dot{f}^2(0)\FnormSize{}{\Theta-\Theta^*}^2.
-\]
-\end{lem}
-\begin{proof} Suppose that the distribution over the ordinal tensor $\tY=\entry{y_\omega}$ is induced by $\Theta=\entry{\theta_\omega}$. Then, based on the generative model~\eqref{eq:model},
-\[
-\mathbb{P}(y_\omega=\ell | \theta_\omega)=f(b_{\ell}-\theta_\omega)-f(b_{\ell-1}-\theta_\omega),
-\]
-for all $\ell\in[L]$ and $\omega\in[d_1]\times \cdots \times [d_K]$.
-For notational convenience, we suppress the subscribe in $\theta_\omega$ and simply write $\theta$ (and respectively, $\theta^*$).
-Based on Lemma~\ref{lem:KLentry} and Taylor expansion,
-\begin{align}
-\text{KL}(\theta|| \theta^*) & \leq \sum_{\ell\in[L]} {\left[f(b_{\ell}-\theta)-f(b_{\ell-1}-\theta)-f(b_{\ell}-\theta^*)+f(b_{\ell-1}-\theta^*)\right]^2\over f(b_{\ell}-\theta^*)-f(b_{\ell-1}-\theta^*)}\\
- &\leq \sum_{\ell=2}^{L-1} {\left[\dot{f}(b_\ell-\eta_\ell)-\dot{f}(b_{\ell-1}-\eta_{\ell-1})\right]^2 \over f(b_\ell-\theta^*)-f(b_{\ell-1}-\theta^*)} \left(\theta-\theta^*\right)^2+{\dot{f}^2(b_1-\eta_1) \over f(b_1-\theta^*)} (\theta-\theta^*)^2\\
- &\quad \quad\quad \quad  +{\dot{f}^2(b_{L-1}-\eta_{L-1})\over 1-f(b_{L-1}-\theta^*)}(\theta-\theta^*)^2,
-\end{align}
-where $\eta_\ell$ and $\eta_{\ell-1}$ fall between $\theta$ and $\theta^*$. Therefore,
-\begin{equation}\label{eq:entrywise}
-\text{KL}(\theta|| \theta^*) \leq \left({4(L-2)\over A_\alpha}+ {2\over A_\alpha}\right)\dot{f}^2(0)(\theta-\theta^*)^2={2(2L-3)\over A_\alpha} \dot{f}^2(0)(\theta-\theta^*)^2,
-\end{equation}
-where we have used Taylor expansion, the bound~\eqref{eq:ass}, and the fact that $\dot{f}(\cdot)$ peaks at zero for an unimodal and symmetric function.
-Now summing~\eqref{eq:entrywise} over the index set $\omega\in[d_1]\times \cdots \times [d_K]$ gives
-\[
-\text{KL}(\mathbb{P}_{\Theta}|| \mathbb{P}_{\Theta^*}) =\sum_{\omega\in[d_1]\times \cdots \times[d_K]} \text{KL}(\theta_\omega || \theta^*_\omega) \leq {2(2L-3)\over A_\alpha} \dot{f}^2(0)\FnormSize{}{\Theta-\Theta^*}^2.
-\]
-\end{proof}
-
-\begin{rmk} In particular, let $\mathbb{P}_{\bf{0}}$ denote the distribution of $\tY|\bf{0}$ induced by the zero parameter tensor. Then we have
-\[
-\text{KL}(\mathbb{P}_{\Theta}||\mathbb{P}_{\bf{0}})\leq {2(2L-3)\over A_\alpha}  \dot{f}^2(0)\FnormSize{}{\Theta}^2.
-\]
-\end{rmk}
-
-\begin{lem}\label{lem:construction}
-Assume the same setup as in Theorem~\ref{thm:minimax}.
-Without loss of generality, suppose $d_1=\max_kd_k$.
-Define $R=\max_k r_k$ and $d_{\text{total}}=\prod_{k\in[K]} d_k$.
-For any constant $0\leq \gamma \leq 1$, there exist a finite set of tensors $\tX=\{\Theta_i: i=1,\ldots\}\subset \tP$ satisfying the following four properties:
-\begin{enumerate}[(i)]
-\item $\text{Card}(\tX)\geq 2^{Rd_1/8}+1$, where $\text{Card}$ denotes the cardinality;
-\item $\tX$ contains the zero tensor $\mathbf{0}\in\mathbb{R}^{d_1\times \cdots\times d_K}$;
-\item $\mnormSize{}{\Theta}\leq \gamma \min\left\{ \alpha ,\ C^{-1/2}\sqrt{Rd_1\over d_{\text{total}}} \right\} $ for any element $\Theta\in\tX$;
-\item $\FnormSize{}{\Theta_i-\Theta_j}\geq {\gamma\over 4} \min\left\{ \alpha\sqrt{d_{\text{total}}},\ C^{-1/2}\sqrt{Rd_1}\right\}$ for any two distinct elements $\Theta_i,\; \Theta_j\in\tX$,
-\end{enumerate}
-Here $C=C(\alpha,L,f,\mb)={(4L-6)\dot{f}^2(0)\over A_\alpha }>0$ is a constant independent of the tensor dimension and rank.
-\end{lem}
-
-\begin{proof}
-Given a constant $0\leq \gamma \leq 1$, we define a set of matrices:
-\[
-\tC=\left\{\mM=(m_{ij})\in\mathbb{R}^{d_1\times R}: a_{ij}\in \left\{ 0,\gamma \min\left\{ \alpha , C^{-1/2}\sqrt{Rd_1\over d_{\text{total}}} \right\}\right\} ,\  \forall (i,j)\in[d_1]\times[R]\right\}.
-\]
-We then consider the associated set of block tensors:
-\begin{align}
-\tB=\tB(\tC)=\{\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}\colon& \Theta=\mA\otimes \mathbf{1}_{d_3}\otimes \cdots \otimes \mathbf{1}_{d_K}, \\
-&\ \text{where}\ \mA=(\mM|\cdots|\mM|\mO) \in\mathbb{R}^{d_1\times d_2},\ \mM\in\tC\},
-\end{align}
-where $\mathbf{1}_d$ denotes a length-$d$ vector with all entries 1, $\mO$ denotes the $d_1\times (d_2-R\lfloor d_2/R \rfloor)$ zero matrix, and $\lfloor d_2/ R \rfloor$ is the integer part of $d_2/R$.
-In other words, the subtensor $\Theta(\mI, \mI,i_3, \ldots,i_K)\in\mathbb{R}^{d_1\times d_2}$ are the same for all fixed $(i_3,\ldots,i_K)\in[d_3]\times \cdots \times [d_K]$, and furthermore, each subtensor $\Theta(\mI,\mI, i_3,\ldots,i_K)$ itself is filled by copying the matrix $\mM\in\mathbb{R}^{d_1\times R}$ as many times as would fit.
-
-By construction, any element of $\tB$, as well as the difference of any two elements of $\tB$, has Tucker rank at most $\max_k r_k\leq R$, and the entries of any tensor in $\tB$ take values in $[0,\alpha]$.
-Thus, $\tB\subset\tP$.
-By Lemma~\ref{lem:VGbound}, there exists a subset $\tX\subset \tB$ with cardinality $\text{Card}(\tX)\geq 2^{Rd_1/8}+1$ containing the zero $d_1\times \cdots \times d_K$ tensor, such that, for any two distinct elements $\Theta_i$ and $\Theta_j$ in $\tX$,
-\[
-\FnormSize{}{\Theta_i-\Theta_j}^2 \geq {Rd_1\over 8} \gamma^2\min\left\{ \alpha^2, {C^{-1}Rd_1 \over d_{\text{total}}}\right\} \lfloor {d_2\over R} \rfloor \prod_{k\geq 3}d_k\geq {\gamma^2\min\left\{ \alpha^2 d_{\text{total}}, C^{-1}Rd_1\right\}  \over 16}.
-\]
-In addition, each entry of $\Theta\in\tX$ is bounded by $\gamma \min\left\{ \alpha , C^{-1/2}\sqrt{Rd_1\over d_{\text{total}}}\right\} $. Therefore the Properties (i) to (iv) are satisfied.
-\end{proof}
-
-\begin{lem}[Varshamov-Gilbert bound]\label{lem:VGbound}
-Let $\Omega=\{(w_1,\ldots,w_m)\colon w_i\in\{0,1\}\}$. Suppose $m>8$. Then there exists a subset $\{w^{(0)},\ldots,w^{(M)}\}$ of $\Omega$ such that $w^{(0)}=(0,\ldots,0)$ and
-\[
-\zeronormSize{}{w^{(j)}-w^{(k)}}\geq {m\over 8},\quad \text{for} \ 0\leq j<k\leq M,
-\]
-where $\zeronormSize{}{\cdot}$ denotes the Hamming distance, and $M\geq 2^{m/8}$.
-\end{lem}
-
-\begin{lem}[Theorem 2.5 in~\cite{tsybakov2008introduction}]\label{lem:Tsybakov}
-Assume that a set $\tX$ contains element $\Theta_0, \Theta_1, \ldots,\Theta_M$ ($M\geq 2$) such that
-\begin{itemize}
-\item $d(\Theta_j,\ \Theta_k)\geq 2s>0$, $\forall 0\leq j\leq k\leq M$;
-\item $\mathbb{P}_0$ is absolutely continuous with respect to $\mathbb{P}_j$, $\forall j=1,\ldots,M$, and
-\[
-{1\over M}\sum_{j=1}^M \mathrm{KL}(\mathbb{P}_j||\mathbb{P}_0)\leq \alpha \log M
-\]
-where $d\colon \tX\times \tX\mapsto [0,+\infty]$ is a semi-distance function, $0<\alpha<{1/8}$ and $\mathbb{P}_j=\mathbb{P}_{\Theta_j}$, $j=0,1\ldots,M$.
-\end{itemize}
-Then
-\[
-\inf_{\hat \Theta}\sup_{\Theta\in\tX} \mathbb{P}_{\Theta}(d(\hat \Theta, \Theta)\geq s)\geq {\sqrt{M}\over 1+\sqrt{M}}\left(1-2\alpha-\sqrt{2\alpha\over \log M} \right)>0.
-\]
-\end{lem}
-
-\begin{lem}[Lemma 28 in~\cite{ghadermarzy2019near}]\label{lem:convexity}
-Define $\mathbb{B}_{M}(\beta)=\{\Theta\in \mathbb{R}^{d_1\times \cdots \times d_K}\colon \anormSize{}{\Theta}\leq \beta \}$.
-Let $\Omega\subset[d_1]\times\cdots \times [d_K]$ be a random set with $m=|\Omega|$, and assume that each entry in $\Omega$ is drawn with replacement from $[d_1]\times\cdots\times[d_K]$ using probability $\Pi$.
-Define
-\[
-\PiFnormSize{}{\Theta}^2={1\over m}\mathbb{E}_{\Omega\in\Pi}\MFnormSize{}{\Theta}^2.
-\]
-Then, there exists a universal constant $c>0$, such that, with probability at least $1-\exp\left(-{\sum_kd_k \over \sum_k \log d_k} \right)$ over the sampled set $\Omega$,
-\[
-{1\over m}\MFnormSize{}{\Theta}^2 \geq \PiFnormSize{}{\Theta}^2-c\beta\sqrt{\sum_k d_k\over m}
-\]
-holds uniformly for all $\Theta\in\mathbb{B}_M(\beta)$.
-\end{lem}
-
-\section{Conclusion}
-We have developed a low-rank tensor estimation method based on possibly incomplete, ordinal-valued observations.
-A sharp error bound is established, and we demonstrate the outperformance of our approach compared to other methods.
-The work unlocks several directions of future research.
-One interesting question would be the inference problem, i.e..
-to assess the uncertainty of the obtained estimates and the imputation.
-Other directions include the trade-off between (non)convex optimization and statistical/computational efficiency.
-While convex relaxations are popular approach for matrix/tensor problem, they are often slow in practice~\citep{ge2017optimization,chen2019non}.
-The interplay between computational efficiency and statistical accuracy in general tensor problems warrants future research.
-
-\section*{Acknowledgments}
-This research is supported in part by NSF grant DMS-1915978 and the University of Wisconsin-Madison, Office of the Vice Chancellor for Research and Graduate Education with funding from the Wisconsin Alumni Research Foundation.
-\newpage
-
-\appendix
-\setcounter{figure}{0}
-\setcounter{table}{0}
-\renewcommand{\thetable}{S\arabic{table}}
-\renewcommand{\thefigure}{S\arabic{figure}}
-\renewcommand{\figurename}{Supplemental Figure}
-\renewcommand{\tablename}{Supplemental Table}
-
-\section*{Appendix}
-\section{Extension of Theorem~\ref{thm:rate} to unknown cut-off points}\label{sec:extention}
-When the cut-off points $\mb$ is unknown, we estimate $(\hat \Theta,\hat \mb)$ by
-\begin{equation}\label{eq:joint}
-(\hat \Theta,\hat \mb)=\argmax_{(\Theta,\mb)\in\tP\times \tB}\logl(\Theta,\mb),
-\end{equation}
-where
-\[
-\tP=\{\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}\colon \text{rank}(\tP)\leq \mr,\ \mnormSize{}{\Theta}\leq \alpha\}, \quad \tB=\{\mb\in\mathbb{R}^{L-1}\colon \mnormSize{}{\mb}\leq \beta,\ \min_{\ell}(b_\ell-b_{\ell-1}) \geq \Delta\}.
-\]
-The estimation accuracy is assessed using the mean squared error (MSE):
-\[
-\text{MSE}\left( \hat \Theta, \trueT \right)={1\over \prod_k d_k}\FnormSize{}{\hat \Theta-\trueT},\quad \text{MSE}\left( \hat \mb, \trueb \right)={1\over L-1}\FnormSize{}{\hat \mb-\trueb}.
-\]
-By abuse of notation, we use $(\Theta,\mb)$ to denote the length-$(d_{\text{total}}+L-1)$-vector collecting all parameters together. The total MSE is defined as
-\[
-\text{MSE}\left((\hat \Theta,\hat \mb),\ (\trueT,\trueb)\right)={1\over \prod_k d_k+L-1}\FnormSize{}{(\hat \Theta,\hat \mb)-(\trueT,\trueb)}^2.
-\]
-We introduce several quantities that will be used in our theory:
-\begin{enumerate}
-\item We make the convention that $b_0=-\infty$, $b_L=\infty$, $f(-\infty)=0$, $f(\infty)=1$, and $\dot{f}(-\infty)=\ddot{f}(-\infty)=\dot{f}(\infty)=\ddot{f}(\infty)=0$.
-
-\item The difference function $g_\ell(\theta)$ is defined as $g_\ell(\theta)=f(b_\ell-\theta)-f(b_{\ell-1}-\theta)$ for all $\theta\in\mathbb{R}$ and $\ell\in[L]$.
-
-\item Define $n_\ell=\sum_{\omega\in\Omega}\mathds{1}_{\{y_\omega=\ell\}}$, i.e., the number of tensor entries taking value on $\ell\in[L]$.
-
-\item With a little abuse of notation, we re-define the constants in~\eqref{eq:bound} as
-\begin{equation}\label{eq:UL}
-U_{\alpha,\beta,\Delta}=\max_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}}\max_{\ell\in[L-1]}{|\dot{g}_\ell(\theta)|\over g_\ell(\theta)}, \quad \text{and}\quad
-L_{\alpha,\beta,\Delta}=\min_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}}\min_{\ell\in[L-1]} {\dot{g}^2_\ell(\theta)-\ddot{g}_\ell(\theta)g_\ell(\theta)\over g^2_\ell(\theta)}.
-\end{equation}
-\item We define three additional constants:
-\begin{align}\label{eq:CD}
-C_{\alpha,\beta,\Delta} &= \max_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}}  \max_{\ell\in[L-1]}\max\left\{\frac{\dot{f}(b_\ell-\theta)}{g_\ell(\theta)},\ \frac{\dot{f}(b_\ell-\theta)}{g_{\ell+1}(\theta)}\right\},\notag \\
-D_{\alpha,\beta,\Delta} &=\min_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}} \min_{\ell\in[L-1]}\min\left\{  {\partial \over \partial \theta}\left({\dot{f}(b_{\ell}-\theta)\over g_\ell(\theta)}\right),\ -{\partial \over \partial \theta}\left({\dot{f}(b_{\ell}-\theta)\over g_{\ell+1}(\theta)}\right)\right\}\notag \\
-&=  \min_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}}\min_{\ell\in[L-1]}\min \left\{-\frac{\ddot{f}(b_\ell-\theta)g_\ell(\theta)-\dot{f}(b_\ell-\theta)\dot{g}_\ell(\theta)}{g^2_\ell(\theta)},\ \right.
-\notag \\
-&\hspace{1.5in} \left. \frac{\ddot{f}(b_\ell-\theta)g_{\ell+1}(\theta)-\dot{f}(b_\ell-\theta)\dot{g}_{\ell+1}(\theta)}{g^2_{\ell+1}(\theta)}\right\},\notag \\
-\tilde{D}_{\alpha,\beta,\Delta} &=\max_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}} \max_{\ell\in[L-1]}\max\left\{  {\partial \over \partial \theta}\left({\dot{f}(b_{\ell}-\theta)\over g_\ell(\theta)}\right),\ -{\partial \over \partial \theta}\left({\dot{f}(b_{\ell}-\theta)\over g_{\ell+1}(\theta)}\right)\right\}.
-\end{align}
-\end{enumerate}
-We make the following assumptions about the link function.
-\begin{assumption}\label{ass:joint}
-The link function $f\colon \mathbb{R}\mapsto [0,1]$ satisfies the following properties:
-\begin{enumerate}
-\item $f(\theta)$ is twice-differentiable and strictly increasing in $\theta$.
-\item $\dot{f}(\theta)$ is strictly log-concave and symmetric with respect to $\theta=0$.
-\item The function ${\dot{f}(b_\ell-\theta)\over g_\ell(\theta)}$ is strictly increasing with respect to $\theta$ for all $\mb\in\tB$.
-\item The function ${\dot{f}(b_\ell-\theta)\over g_{\ell+1}(\theta)}$ is strictly decreasing with respect to $\theta$ for all $\mb\in\tB$.
-\end{enumerate}
-\end{assumption}
-
-\begin{rmk}
-The condition $\Delta=\min_{\ell}(b_{\ell}-b_{\ell-1})>0$ on the feasible set $\tB$ guarantees the strict positiveness of $g_{\ell}(\theta)=f(b_{\ell}-\theta)-f(b_{\ell-1}-\theta)$.
-Therefore, the denominators in the above quantities $U_{\alpha,\beta,\Delta}, L_{\alpha,\beta,\Delta}, C_{\alpha,\beta,\Delta}, D_{\alpha,\beta,\Delta},\tilde{D}_{\alpha,\beta,\Delta}$ are well-defined.
-Furthermore, by Theorem~\ref{thm:convexity}, $g_\ell(\theta)$ is strictly log-concave in $\theta$ for all $\mb\in\tB$.
-Based on Assumption~\ref{ass:joint} and closeness of the feasible set, we have $U_{\alpha,\beta,\Delta}>0$, $L_{\alpha,\beta,\Delta}>0$, $C_{\alpha,\beta,\Delta}>0$, $D_{\alpha,\beta,\Delta}>0$, $\tilde{D}_{\alpha,\beta,\Delta}>0$.
-\end{rmk}
-
-\begin{rmk}
-In particular, for logistic link $f(x) = \frac{1}{1+e^{-x}}$ , we have
-\begin{align}
-C_{\alpha,\beta,\Delta}&=\max_{\substack{|\theta|\leq \alpha,\\ \mb\in\tB}}\max_{\ell\in[L-1]}\max
-\left\{
-{ 1\over e^{b_{\ell}-b_{\ell-1}}-1} { 1+e^{-(b_{\ell-1}-\theta)} \over 1+e^{-(b_{\ell}-\theta)}},\
-{ 1\over 1-e^{-(b_{\ell+1}-b_{\ell})}} { 1+e^{-(b_{\ell+1}-\theta)} \over 1+e^{-(b_{\ell}-\theta)}}
-\right\}>0,
-\\
-D_{\alpha,\beta,\Delta}&=\min_{\substack{|\theta|\leq \alpha,\\ \mb\in\tB}}\min_{\ell\in[L-1]} {e^{-(b_\ell-\theta)}\over (1+e^{-(b_\ell-\theta)})^2}>0.
-\end{align}
-\end{rmk}
-
-\begin{thm}[Statistical convergence with unknown $\mb$]\label{thm:ratejoint}
-Consider an ordinal tensor $\tY\in[L]^{d_1\times \cdots \times d_K}$ generated from model~\eqref{eq:model} with the link function $f$ and parameters $(\trueT, \trueb)\in\tP\times \tB$.
-Suppose the link function $f$ satisfies Assumption~\ref{ass:joint}.
-Define $r_{\max}=\max_k r_k$, and assume $r_{\max}=\tO(1)$.
-Then with very high probability, the estimator in~\eqref{eq:joint} satisfies
-\begin{equation}\label{eq:jointTheta}
-\mathrm{MSE}\left(\hat \Theta, \trueT \right)\leq\min\left( 4\alpha^2,\  {c_1U_{\alpha,\beta,\Delta}^2  r_{\max}^{K-1}  \over L_{\alpha,\beta,\Delta}^2}{ \sum_k d_k \over \prod_k d_k}\right),
-\end{equation}
-and
-\begin{equation}\label{eq:jointb}
-\mathrm{MSE}\left(\hat \mb, \trueb \right)\leq \min\left(4\beta^2,\ {C'_{\alpha,\beta,\Delta}r^{K-1}_{\max} \over D_{\alpha,\beta,\Delta}^2}{\left(\sum_kd_k\right)\left(\prod_k d_k\right) \over \min_{\ell}(n_{\ell}+n_{\ell+1})^2}\right),
-\end{equation}
-where $c_1, U_{\alpha,\beta,\Delta}, L_{\alpha,\beta,\Delta}, C'_{\alpha,\beta,\Delta}, \tilde D_{\alpha,\beta,\Delta}$ are constants independent of the tensor dimension and rank.
-
-In particular, we have
-\begin{align}
-\mathrm{MSE}\left((\hat \Theta,\hat \mb),\ (\trueT,\trueb)\right) = \tO\left({\sum_k d_k \over \min\left\{\prod_k d_k ,\ \min_\ell(n_\ell+n_{\ell+1})^2\right\}}\right),\quad \text{as}\quad d_{\min}\to\infty.
-\end{align}
-\end{thm}
-
-\begin{rmk} The total MSE has two components, where the first component $\tO\left({\sum_k d_k \over\prod_k d_k}\right)$ is from the error in $\hat\Theta$, and the second component $\tO\left({\sum_k d_k \over \min_\ell(n_\ell+n_{\ell+1})^2}\right)$ is from the error in $\hat \mb.$ When the $L$ labels are moderately balanced in that $\min_\ell(n_\ell+n_{\ell+1})\gg \sqrt{\prod_kd_k}$, the error in $\hat\Theta$ dominates the total MSE.
-When the $L$ labels are severely imbalanced in that $\min_\ell(n_\ell+n_{\ell+1})\ll \sqrt{\prod_kd_k}$, the error in $\hat\mb$ dominates the total MSE.
-\end{rmk}
-
-\begin{rmk} We discuss the impact of imbalance between labels to the estimation accuracy.
-Consider the special case $d_1=\cdots =d_K=d$.
-The bounds~\eqref{eq:jointTheta} and~\eqref{eq:jointb} demonstrate that, both $\hat \Theta$ and $\hat \mb$ achieve consistency as long as $\min_\ell(n_\ell+n_{\ell+1})$ is slightly larger than $\tO(d^{(K+1)/2})$.
-Note that in this case ${\min_\ell(n_\ell+n_{\ell+1}) \over \max_\ell n_{\ell}}\asymp {\min_\ell(n_\ell+n_{\ell+1})\over d^K} \to 0$, indicating a long-run imbalance between labels.
-The obtained consistency highlights the robustness of our proposed estimator~\eqref{eq:joint} to the label imbalance.
-\end{rmk}
-
-\begin{proof}[Proof of Theorem~\ref{thm:ratejoint}]
-The constant bounds with $\alpha$ and $\beta$ can be obtained trivially from the definition of the feasible sets. We focus on proving the non-constant bounds.
-
-Similar to the proof of Theorem~\ref{thm:rate}, we suppress $\Omega$ in the subscript. Based on the definition of $(\hat \Theta, \hat \mb)$,
-we have the following inequalities:
-\begin{equation}\label{eq:level}
-\flogl(\hat \Theta,\hat \mb)\geq \flogl(\trueT,\hat \mb)\quad \text{and}\quad \flogl(\hat \Theta,\hat \mb)\geq \flogl(\hat \Theta, \trueb).
-\end{equation}
-Following the same argument as in Theorem~\ref{thm:rate} and the first inequality in~\eqref{eq:level}, we obtain that
-\begin{equation}\label{eq:ratenew}
-\FnormSize{}{\hat \Theta-\trueT}^2\leq {c_1^2U_{\alpha,\beta,\Delta}^2 r_{\max}^{K-1} \over L_{\alpha,\beta,\Delta}^2}   \sum_k d_k ,
-\end{equation}
-where $U_{\alpha,\beta,\Delta}, L_{\alpha,\beta,\Delta}>0$ are two constants defined in~\eqref{eq:UL}.
-
-Next we bound $\FnormSize{}{\hat \mb-\trueb}^2$ given $\hat\Theta$.
-We consider the profile log-likelihood $\flogl(\hat \Theta,\mb)$ as a function of $\mb\in\tB$.
-For notational convenience, we drop $\hat \Theta$ from $\flogl(\hat \Theta,\mb)$ and simply write $\flogl(\mb)$.
-It follows from the expression of $\flogl(\mb)$ that
-\begin{align}\label{eq:propertyb}
-{\partial \flogl\over \partial b_\ell}&=\sum_{\omega\in\Omega}\left[\mathds{1}_{\{y_{\omega}=\ell\}}
-\frac{\dot{f}(b_\ell-\theta_\omega)}{g_\ell(\theta_\omega)}-\mathds{1}_{\{y_{\omega}=\ell+1\}}
-\frac{\dot{f}(b_\ell-\theta_\omega)}{g_{\ell+1}(\theta_\omega)}\right],\notag\\
-{\partial^2 \flogl\over \partial b^2_\ell}&=\sum_{\omega\in\Omega}\left[\mathds{1}_{\{y_\omega=\ell\}}\frac{\ddot{f}(b_\ell-\theta_\omega)g_\ell(\theta_\omega)-\dot{f}^2(b_\ell-\theta_\omega)}{g^2_\ell(\theta_\omega)}-\mathds{1}_{\{y_\omega=\ell+1\}}\frac{\ddot{f}(b_\ell-\theta_\omega)g_{\ell+1}(\theta_\omega)+\dot{f}^2(b_\ell-\theta_\omega)}{g^2_{\ell+1}(\theta_\omega)}\right],\notag\\
-&\quad  \text{ for all } \ell \in [L-1],\notag \\
-{\partial^2 \flogl\over \partial b_\ell \partial b_{\ell+1}}&=\sum_{\omega\in\Omega}\mathds{1}_{\{y_\omega=\ell+1\}}\frac{\dot{f}(b_\ell-\theta_\omega)\dot{f}(b_{\ell+1}-\theta_\omega)}{g^2_{\ell+1}(\theta_\omega)}\quad \text{and}\quad
-\ {\partial^2 \flogl\over \partial b_\ell \partial b_\ell'} = 0 \text{ if } |\ell-\ell'|>1.
-\end{align}
-Therefore, all entries in $\fploglb$ are upper bounded by $\left\{C_{\alpha,\beta,\Delta} \max_\ell (n_\ell+n_{\ell+1})\right\}>0$, and $\fpploglb$ is a tridiagonal matrix.
-
-By the second-order Taylor's expansion of $\flogl(\mb)$ around $\trueb$, we obtain
-\begin{equation}\label{eq:taylorb}
-\flogl(\hat\mb)=\flogl(\trueb)+(\trueb-\hat\mb)^T\fploglb(\trueb)+ {1\over 2}(\trueb-\hat\mb)^T\fpploglb(\check\mb)(\trueb-\hat\mb),
-\end{equation}
-where $\check\mb=\gamma\trueb+(1-\gamma)\hat\mb$ for some $\gamma\in[0,1]$, and $\fpploglb(\check\mb)$ denotes the $(L-1)$-by-$(L-1)$ Hession matrix evaluated at $\check\mb$.
-
-The linear term in~\eqref{eq:taylorb} can be bounded by Cauchy-Schwartz inequality,
-\begin{align}\label{eq:linearb}
-(\trueb-\hat\mb)^T\fploglb(\trueb)&\leq \FnormSize{}{\trueb-\hat\mb}\FnormSize{}{\fploglb(\trueb)}\\
-&\leq C'_{\alpha,\beta,\Delta}\FnormSize{}{\trueb-\hat\mb}\sqrt{L-1}\sqrt{\left(\sum_k d_k\right)\left( \prod_k d_k\right)},
-\end{align}
-for some constant $C'_{\alpha,\beta,\Delta}>0$, where the last inequality follows from Lemma \ref{lem:sharpergd},
-\[
-\left|{\partial \flogl\over \partial b_\ell}\bigg|_{(\hat \Theta,\trueb)}\right|\leq C'_{\alpha,\beta,\Delta}\sqrt{\left(\sum_k d_k\right)\left(\prod_k d_k\right)},\quad \text{ for all } \ell \in [L-1] \text{ with very high probability}.
-\]
-
-We next bound the quadratic term in \eqref{eq:taylorb}. Note that
-\begin{align}\label{eq:quadraticb}
- &(\trueb-\hat\mb)^T \fpploglb(\check{\mb})(\trueb-\hat\mb)\\
- &=\sum_{\ell\in [L-1]} \left( {\partial^2\tL_{\tY}\over \partial b^2_\ell} \Big|_{\mb=\check\mb} \right)(\hat b_\ell-b^{\text{true}}_{\ell})^2 + 2\sum_{\ell\in[L-1]/\{1\}}\left( {\partial^2\tL_{\tY}\over \partial b_\ell \partial b_{\ell-1}} \Big|_{\mb=\check\mb} \right)(\hat b_\ell-b^{\text{true}}_{\ell})(\hat b_{\ell-1}-b^{\text{true}}_{\ell-1}) \nonumber \\
-&\leq \sum_{\ell\in [L-1]} \left( {\partial^2\tL_{\tY}\over \partial b^2_\ell} \Big|_{\mb=\check\mb} \right)(\hat b_\ell-b^{\text{true}}_{\ell})^2 + \sum_{\ell\in[L-1]/\{1\}}\left( {\partial^2\tL_{\tY}\over \partial b_\ell \partial b_{\ell-1}} \Big|_{\mb=\check\mb} \right)\big[(\hat b_\ell-b^{\text{true}}_{\ell})^2 +(\hat b_{\ell-1}-b^{\text{true}}_{\ell-1})^2\big] \nonumber \\
-&=\left( {\partial^2\tL_{\tY}\over \partial b^2_1}+{\partial^2\tL_{\tY}\over \partial b_1 \partial b_2 }\right)\Big|_{\mb=\check\mb}(\hat b_1-b^{\text{true}}_{1})^2+\left( {\partial^2\tL_{\tY}\over \partial b^2_{L-1}}+{\partial^2\tL_{\tY}\over \partial b_{L-2} \partial b_{L-1} }\right)\Big|_{\mb=\check\mb}(\hat b_{L-1}-b^{\text{true}}_{L-1})^2  \\
-&\quad\quad +\sum_{\ell\in[L-2]/\{1\}}\left( {\partial^2\tL_{\tY}\over \partial b^2_\ell}+{\partial^2\tL_{\tY}\over \partial b_\ell \partial b_{\ell-1}}+{\partial^2\tL_{\tY}\over \partial b_{\ell+1} \partial b_{\ell}} \right)\Big|_{\mb=\check\mb} (\hat b_\ell-b^{\text{true}}_{\ell})^2     \\
-&\leq - D_{\alpha,\beta,\Delta}' \sum_{\ell \in [L-1]}(\hat b_\ell-b_{\text{true},\ell})^2 \nonumber \\
-&=-D_{\alpha,\beta,\Delta}'\FnormSize{}{\hat\mb-\trueb}^2,
-\end{align}
-where
-\begin{align}
-D_{\alpha,\beta,\Delta}'&=\min_{\substack{|\theta|\leq \alpha,\\ \mb\in\tB}}\min_{\ell\in[L-1]} -\left( {\partial^2\tL_{\tY}\over \partial b^2_\ell}+{\partial^2\tL_{\tY}\over \partial b_\ell \partial b_{\ell-1}}+{\partial^2\tL_{\tY}\over \partial b_{\ell+1} \partial b_{\ell}} \right)\\
-&=\min_{\substack{|\theta|\leq \alpha,\\ \mb\in\tB}} \min_{\ell\in[L-1]}\left\{\sum_{\omega\in\Omega}-\mathds{1}_{\{y_\omega=\ell\}}\left( {\ddot{f}(b_{\ell}-\theta_\omega) g_\ell(\theta_\omega)- \dot{f}(b_{\ell}-\theta_\omega)\dot{g}_\ell(\theta_\omega) \over g^2_\ell(\theta_\omega) }\right)\right.
-\\
-&\left.
-\hspace{1in}+\sum_{\omega\in\Omega}\mathds{1}_{\{y_\omega=\ell+1\}} \left( {\ddot{f}(b_\ell-\theta_\omega) g_{\ell+1}(\theta_\omega) - \dot{f}(b_\ell-\theta_\omega)\dot{g}_{\ell+1} (\theta_\omega)\over g^2_{\ell+1}(\theta_\omega)} \right)\right\}\\
-&=\min_{\substack{|\theta|\leq \alpha,\\ \mb\in\tB}} \min_{\ell\in[L-1]}\left\{\KeepStyleUnderBrace{-\sum_{\omega\in\Omega} \mathds{1}_{\{y_\omega=\ell\}}{\partial \over \partial \theta_\omega}\left( \dot{f}(b_\ell-\theta_\omega) \over g_\ell(\theta_\omega) \right)}_{>0}+\KeepStyleUnderBrace{\sum_{\omega\in\Omega}\mathds{1}_{\{y_\omega=\ell+1\}} {\partial \over \partial \theta_\omega}\left( \dot{f}(b_\ell-\theta_\omega) \over g_{\ell+1}(\theta_\omega) \right) }_{>0}\right\}\\
-&\geq D_{\alpha,\beta,\Delta}  \min_{\ell\in[L-1]}\left(n_{\ell}+n_{\ell+1}\right) .
-\end{align}
-Combining inequalities~\eqref{eq:taylorb}, \eqref{eq:linearb} and~\eqref{eq:quadraticb} yields
-\[
-\flogl(\hat \mb)\leq \flogl(\trueb)+C'_{\alpha,\beta,\Delta} \sqrt{(L-1)\left(\sum_k d_k\right)\left(\prod_k d_k\right)} \FnormSize{}{\hat\mb-\trueb} -\frac{D_{\alpha,\beta,\Delta}}{2}\min_{\ell}(n_\ell+n_{\ell+1}) \FnormSize{}{\hat\mb-\trueb}^2 .
-\]
-Since $\hat \mb$ satisfies $\flogl(\hat \mb)-\flogl(\trueb)\geq 0$, we have that
-\[
-C'_{\alpha,\beta,\Delta} \sqrt{(L-1)\left(\sum_k d_k\right)\left(\prod_k d_k\right)} \FnormSize{}{\hat\mb-\trueb} -\frac{D_{\alpha,\beta,\Delta}}{2}\min_{\ell}(n_\ell+n_{\ell+1}) \FnormSize{}{\hat\mb-\trueb}^2\geq 0.
-\]
-Finally,
-\begin{equation}\label{eq:rateb}
-\FnormSize{}{\hat\mb-\trueb}^2\leq {C'_{\alpha,\beta,\Delta} (L-1) \over D_{\alpha,\beta,\Delta}^2}{\left(\sum_k d_k\right)\left(\prod_k d_k\right)\over \min_{\ell}(n_{\ell}+n_{\ell+1})^2}.
-\end{equation}
-Combing~\eqref{eq:ratenew} and~\eqref{eq:rateb} with the following inequality completes the proof.
-\begin{align}
-\text{MSE}\left((\hat \Theta,\hat \mb),(\trueT,\trueb)\right)&\leq{1\over \prod_k d_k}\FnormSize{}{(\hat \Theta,\hat \mb)-(\trueT,\trueb)}^2\\
-&={1\over \prod_k d_k}\FnormSize{}{(\hat \Theta,\hat \mb)-(\hat \Theta,\trueb)+(\hat \Theta,\trueb)-(\trueT,\trueb)}^2\\
-&\leq {1\over\prod_kd_k}\FnormSize{}{(\hat \Theta- \trueT)}^2+ {1\over\prod_kd_k}\FnormSize{}{\hat \mb -\trueb}^2.
-\end{align}
-\end{proof}
-
-\begin{lem}[CLT for independent Bernoulli r.v.'s]\label{lem:clt}
-Let $\{X_n\}$ be a series of independent Bernoulli random variables with possibly different success probabilities $\{p_n\}$.
-Suppose the success probabilities $\{p_n\}$ are bounded, in that, $c_1\leq p_n\leq 1-c_2$ for some positive constants $c_1,c_2>0$ and all $n\in[m]$.
-Define $s^2_m = \sum_{n=1}^m p_n(1-p_n)$ and $Y_n = X_n-p_n$.
-Then,
-\[
-{1\over s_m}\sum_{n=1}^m Y_n \stackrel{\mathcal{D}}{\rightarrow} N(0,1), \quad \text{ as } m \rightarrow \infty.
-\]
-\end{lem}
-
-\begin{proof}
-We apply Lyapunov central limit theorem (CLT) to $\{Y_n\}$. Let us verify Lyapunov's condition for $\delta = 1$:
-\[
-\mathbb{E}[|Y_n|^3] = p_n(1-p_n)^3+ (1-p_n)p_n^3\leq p_n(1-p_n)[(1-p_n)^2+p_n^2]\leq p_n(1-p_n).
-\]
-Summation of the above inequality shows
-\[
-\sum_{n=1}^m\mathbb{E}[|Y_n|^3]\leq \sum_{n=1}^m p_n(1-p_n) = s_m^2.
-\]
-Thus, the Lyapunov condition is satisfied whenever
-\[
-{1\over s_m^3}\sum_{n=1}^m \mathbb{E}[|Y_n|^3]\leq {s_m^2\over s_m^3}\rightarrow 0,
-\]
-or simply $s_m = \sum_{n=1}^m p_n(1-p_n) \geq m\min\{c_1(1-c_1), c_2(1-c_2)\} \rightarrow \infty.$ Applying Lyapunov CLT to $\{Y_n\}$ completes the proof.
-\end{proof}
-
-\begin{lem}[Bound on score function]\label{lem:sharpergd}
-Consider the same set-up as in Theorem~\ref{thm:ratejoint}. Let $\tL_\tY(\Theta,\mb)$ denote the log-likelihood function of $(\Theta, \mb)$ given data tensor $\tY$. Then, with very high probability,
-\[
-\left|{\partial \flogl\over \partial b_\ell}\bigg|_{(\Theta,\trueb)}\right|\leq \left(C_{\alpha,\beta,\Delta}+\tilde{D}_{\alpha,\beta,\Delta}\FnormSize{}{\trueT-\Theta}^2\right)\sqrt{\prod_k d_k},\quad \text{ for all } \ell \in [L-1],
-\]
-In particular, with very high probability,
-\[
-\left|{\partial \flogl\over \partial b_\ell}\bigg|_{(\hat \Theta,\trueb)}\right|\leq C'_{\alpha,\beta,\Delta}\sqrt{\left(\sum_k d_k\right)\left( \prod_k d_k\right)},\quad \text{ for all } \ell \in [L-1],
-\]
-where $C'_{\alpha,\beta,\Delta}>0$ is a constant that depends on $U_{\alpha,\beta,\Delta}, L_{\alpha,\beta,\Delta}, C_{\alpha,\beta,\Delta},  \tilde D_{\alpha,\beta,\Delta}$ defined in~\eqref{eq:UL} and~\eqref{eq:CD}.
-\end{lem}
-
-\begin{proof}
-We only prove the case for $\ell = 1$. Other cases can be proved similarly.
-
-\begin{equation}\label{eq:sharpergd}
-  {\partial \flogl\over \partial b_1}\bigg|_{(\Theta,\trueb)} = \KeepStyleUnderBrace{{\partial \flogl\over \partial b_1}\bigg|_{( \Theta,\trueb)}-\mathbb{E}_\mathcal{Y}\left[{\partial \flogl\over \partial b_1}\bigg|_{(\Theta,\trueb)}\right]}_{\coloneqq A}+\KeepStyleUnderBrace{\mathbb{E}_\mathcal{Y}\left[{\partial \flogl\over \partial b_1}\bigg|_{( \Theta,\trueb)}\right]-\mathbb{E}_\mathcal{Y}\left[{\partial \flogl\over \partial b_1}\bigg|_{(\trueT,\trueb)}\right] }_{\coloneqq B}.
-\end{equation}
-We have used the fact that the score function has mean zero, $\mathbb{E}_\mathcal{Y}\left[{\partial \flogl\over \partial b_1}\big|_{(\trueT,\trueb)}\right]=0 $.
-Here all expectations are taken with respect to $\mathcal{Y}\sim \mathbb{P}(\trueT,\trueb).$\\
-We now bound the two deviation terms in \eqref{eq:sharpergd} separately. The term $A$ in \eqref{eq:sharpergd} is the stochastic deviation of log-likelihood to its expectation:
-\begin{align}
-A &= \sum_{\omega\in\Omega}\left\{[\mathds{1}_{y_\omega = 1}-g_1(\theta^{\text{true}}_\omega)]{\dot{f}(b_1-\theta_\omega)\over g_1(\theta_\omega)}-[\mathds{1}_{y_\omega = 2}-g_2(\theta^{\text{true}}_\omega)]{\dot{f}(b_1-\theta_\omega)\over g_2(\theta_\omega)}\right\}\\
-&\leq C_{\alpha,\beta,\Delta}\sum_{\omega\in\Omega}\KeepStyleUnderBrace{[\mathds{1}_{y_\omega = 1}+\mathds{1}_{y_\omega = 2}- g_1(\theta^{\text{true}}_\omega)-g_2(\theta^{\text{true}}_\omega)]}_{\coloneqq W_\omega}.
-\end{align}
-Note that $\{W_\omega\}$ are independent, centered Bernoulli random variables with bounded success probabilities $g_1(\theta^{\text{true}}_\omega)+g_2(\theta^{\text{true}}_\omega)$.
-By Lemma \ref{lem:clt}, we have
-\[
-\sum_{\omega\in\Omega}W_\omega \stackrel{\mathcal{D}}{\rightarrow} N\left(0,\ \sum_{\omega\in\Omega}\left(g_1(\theta^{\text{true}}_\omega)+g_2(\theta^{\text{true}}_\omega)\right)\left(1-g_1(\theta^{\text{true}}_\omega)-g_2(\theta^{\text{true}}_\omega)\right)\right).
-\]
-Hence, with the fact that $\sum_{\omega\in\Omega}\left(g_1(\theta^{\text{true}}_\omega)+g_2(\theta^{\text{true}}_\omega)\right)\left(1-g_1(\theta^{\text{true}}_\omega)-g_2(\theta^{\text{true}}_\omega)\right)\leq {1\over 4}\prod_k d_k$,
-\begin{equation}\label{eq:A}
-|A| \leq C_{\alpha,\beta,\Delta}\left|\sum_{\omega\in\Omega}W_\omega\right|\leq C_{\alpha,\beta,\Delta}\sqrt{\prod_k d_k}, \quad \text{ with very high probability. }
-\end{equation}
-The second term $B$ in \eqref{eq:sharpergd} is the bias induced by $\Theta$:
-\begin{align}
-|B| &= \left|\sum_{\omega\in\Omega} g_1(\theta^{\text{true}}_\omega)\left({\dot{f}(b_1-\theta_\omega)\over g_1(\theta_\omega)}-{\dot{f}(b_1-\theta^{\text{true}}_\omega)\over g_1(\theta^{\text{true}}_\omega)}\right)-\sum_{\omega\in\Omega} g_2(\theta^{\text{true}}_\omega)\left({\dot{f}(b_1-\theta_\omega)\over g_2(\theta_\omega)}-{\dot{f}(b_1-\theta^{\text{true}}_\omega)\over g_2(\theta^{\text{true}}_\omega)}\right)\right|\\
-&= \left|\sum_{\omega\in\Omega} g_1(\theta^{\text{true}}_\omega)(\theta^{\text{true}}_\omega-\theta_\omega)\left\{{\partial\over\partial \theta}\left({\dot{f}(b_1-\theta)\over g_1(\theta)}\right)\bigg|_{\rho_\omega\theta_\omega+(1-\rho_\omega)\theta^{\text{true}}_\omega}\right\}\right.\\
-&\hspace{0.5in}\left.-\sum_{\omega\in\Omega} g_2(\theta^{\text{true}}_\omega)(\theta^{\text{true}}_\omega-\theta_\omega)\left\{{\partial\over\partial \theta}\left({\dot{f}(b_1-\theta)\over g_2(\theta)}\right)\bigg|_{\rho_\omega'\theta_\omega+(1-\rho_\omega')\theta^{\text{true}}_\omega}\right\}\right|\\
-&\leq \tilde{D}_{\alpha,\beta,\Delta}\sum_{\omega\in\Omega}|g_1(\theta^{\text{true}}_\omega)+g_2(\theta^{\text{true}}_\omega)||\theta^{\text{true}}_\omega-\theta_\omega|.
-\end{align}
-By Cauchy-Schwartz inequality with the fact that $g_1(\theta^{\text{true}}_\omega)+g_2(\theta^{\text{true}}_\omega)\leq 1$,
-\begin{equation}\label{eq:B}
-|B| \leq \tilde{D}_{\alpha,\beta,\Delta}\sqrt{\prod_k d_k} \FnormSize{}{\trueT-\Theta}.
-\end{equation}
-Plugging \eqref{eq:A} and \eqref{eq:B} back to \eqref{eq:sharpergd} yields that
-\[
-{\partial \flogl\over \partial b_1}\bigg|_{(\Theta,\trueb)}\leq \left(C_{\alpha,\beta,\Delta}+\tilde{D}_{\alpha,\beta,\Delta}\FnormSize{}{\trueT-\Theta}\right)\sqrt{\prod_k d_k}.
-\]
-holds with very high probability. The second inequality in the lemma comes from~\eqref{eq:ratenew} that $\FnormSize{}{\trueT-\hat\Theta}\leq \tO(\sqrt{\sum_k d_k}).$
-\end{proof}
-
-\section{Additional results of HCP analysis}\label{sec:additionalHCP}
-We perform clustering analyses based on the Tucker representation of the estimated tensor parameter $\hat\Theta$.
-The procedure is motivated from the higher-order extension of Principal Component Analysis (PCA) or Singular Value Decomposition (SVD).
-Recall that, in the matrix case, we usually perform clustering on an $m\times n$ (normalized) matrix $X$ based on the following procedure.
-First, we factorize $X$ into
-\begin{equation}
-X = U\Sigma V^T,
-\end{equation}
-where $\Sigma$ is a diagonal matrix and $ U,V$ are factor matrices with orthogonal columns.
-Second, we take each column of $V$ as a principal axis and each row in $U\Sigma$ as principal component.
-A subsequent multivariate clustering method (such as $K$-means) is then applied to the $m$ rows of $U\Sigma$.
-
-We apply a similar clustering procedure to the estimated parameter tensor $\hat\Theta$. Based on Tucker representation of $\hat \Theta$, we have
-\begin{equation}\label{eq:Tuckerest}
-\hat \Theta = \hat \tC\times_1\hat \mM_1\times_2\cdots\times_K\hat \mM_K,
-\end{equation}
-where $\hat \tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is the estimated core tensor, $\hat \mM_k\in\mathbb{R}^{d_k\times r_k}$ are estimated factor matrices with orthogonal columns, and $\times_k$ denotes the tensor-by-matrix multiplication~\citep{kolda2009tensor}.
-The mode-$k$ matricization of~\eqref{eq:Tuckerest} gives
-\begin{equation}
-\hat \Theta_{(k)} = \hat \mM_k\hat \tC_{(k)}\left(\hat \mM_K\otimes\cdots\otimes\hat \mM_1\right),
-\end{equation}
-where $\hat \Theta_{(k)} , \hat \tC_{(k)}$ denote the mode-$k$ unfolding of $\hat \Theta$ and $\hat \tC$, respectively.
-Then, the mode-$k$ clustering can be performed as follows.
-First, we take columns in $\left(\hat \mM_K\otimes\cdots\otimes\hat \mM_1\right)$ as principal axes and rows in $\hat \mM_k\hat \tC_{(k)}$ as principal components.
-Then, we perform $K$-means clustering method to the $d_k$ rows of the matrix $\hat \mM_k\hat \tC_{(k)}$.
-
-We perform a clustering analysis on the 68 brain nodes using the produce described above.
-Our ordinal tensor method outputs the estimated parameter tensor $\hat\Theta\in\mathbb{R}^{68\times 68\times136}$ with rank $(23,23,8)$.
-We apply $K$-means to the mode-1 principal component matrix of size $68\times 184$ ($184=23\times8$).
-The elbow method suggests 11 clusters among the 68 nodes (see Figure~\ref{figure:elbow}).
-
-\begin{figure}[H]
-\begin{center}
-\includegraphics[width=0.4\textwidth]{elbowm.pdf}
-\end{center}
-\caption{Elbow plot for determining the number of clusters in $K$-means.}\label{figure:elbow}
-\end{figure}
-
-\bibliography{tensor_wang}
-\bibliographystyle{apalike}
+\bibliographystyle{plainnat}
+\bibliography{ordinalT_supp.bbl}
+
 \end{document}

### 2020-02-18 ICML/submit/supplements/functions.R

@@ -1,0 +1,522 @@
+library(MASS)
+library(rTensor)
+library(pracma)
+epsilon=10^-4
+
+################### some functions used for analysis #######################################
+#### Output: randomly sampled ordinal tensor based on logistic models given input parameters
+#### Input : 
+#### theta: continuous parameter tensor (latent parameter)
+#### omega: the cut-off points (k-levels)
+realization = function(theta,omega){
+  k=length(omega) 
+  theta=as.tensor(theta)
+  theta_output <- c(theta@data)
+  p = theta_to_p(theta_output,omega)
+  for (j in 1:length(theta_output)) {
+    theta_output[j] <-  sample(1:(k+1),1,prob = p[j,])
+  }
+  return(as.tensor(array(theta_output,dim =theta@modes)))
+}      
+
+#### Output: predicted ordinal tensor from the given input parameters and types of predction
+#### Input : 
+#### theta: continuous parameter tensor (latent parameter)
+#### omega: the cut-off points (k-levels)
+#### type:`mode' the mostly likely label,`mean' the mean label, `median' the median label
+estimation = function(theta,omega,type){ 
+  k=length(omega) 
+  if(is.matrix(theta)){
+    theta_output=c(theta)
+  }else{
+    theta=as.tensor(theta)
+    theta_output <- c(theta@data)
+  }
+  p = theta_to_p(theta_output,omega)
+  if(type=="mode"){  # score prediction based on the mode 
+    for (j in 1:length(theta_output)) theta_output[j] <-  which.max(p[j,])
+  }else if(type=="mean"){ # score prediction based on the mean 
+    for (j in 1:length(theta_output)) theta_output[j] <-  sum(p[j,]*(1:(k+1))) 
+  }else if(type=="median"){# score prediction based on the median 
+    for (j in 1:length(theta_output)) theta_output[j] <-  which(c(cumsum(p[j,]),1)>=0.5)[1] ## median
+  }
+  if(is.matrix(theta)) return(matrix(theta_output,nrow=dim(theta)[1],ncol=dim(theta)[2]))
[…]

### 2020-02-19 ICML/submit/supplements/functions.R
Delete functions.R

@@ -1,522 +1,0 @@
-library(MASS)
-library(rTensor)
-library(pracma)
-epsilon=10^-4
-
-################### some functions used for analysis #######################################
-#### Output: randomly sampled ordinal tensor based on logistic models given input parameters
-#### Input : 
-#### theta: continuous parameter tensor (latent parameter)
-#### omega: the cut-off points (k-levels)
-realization = function(theta,omega){
-  k=length(omega) 
-  theta=as.tensor(theta)
-  theta_output <- c(theta@data)
-  p = theta_to_p(theta_output,omega)
-  for (j in 1:length(theta_output)) {
-    theta_output[j] <-  sample(1:(k+1),1,prob = p[j,])
-  }
-  return(as.tensor(array(theta_output,dim =theta@modes)))
-}      
-
-#### Output: predicted ordinal tensor from the given input parameters and types of predction
-#### Input : 
-#### theta: continuous parameter tensor (latent parameter)
-#### omega: the cut-off points (k-levels)
-#### type:`mode' the mostly likely label,`mean' the mean label, `median' the median label
-estimation = function(theta,omega,type){ 
-  k=length(omega) 
-  if(is.matrix(theta)){
-    theta_output=c(theta)
-  }else{
-    theta=as.tensor(theta)
-    theta_output <- c(theta@data)
-  }
-  p = theta_to_p(theta_output,omega)
-  if(type=="mode"){  # score prediction based on the mode 
-    for (j in 1:length(theta_output)) theta_output[j] <-  which.max(p[j,])
-  }else if(type=="mean"){ # score prediction based on the mean 
-    for (j in 1:length(theta_output)) theta_output[j] <-  sum(p[j,]*(1:(k+1))) 
-  }else if(type=="median"){# score prediction based on the median 
-    for (j in 1:length(theta_output)) theta_output[j] <-  which(c(cumsum(p[j,]),1)>=0.5)[1] ## median
-  }
-  if(is.matrix(theta)) return(matrix(theta_output,nrow=dim(theta)[1],ncol=dim(theta)[2]))
[…]

### 2020-02-19 ICML/submit/supplements/demo.R
Add files via upload

@@ -1,0 +1,44 @@
+source("functions.R")
+
+### This file contains the code of a simple demo for the tensor estimations and predictions.
+
+### set tensor dimension and rank of the tensor
+d=c(30,30,30) 
+r=c(10,10,10)
+
+
+### set true latent parameter
+A_1 = randortho(d[1])[,1:r[1]]
+A_2 = randortho(d[2])[,1:r[2]]
+A_3 = randortho(d[3])[,1:r[3]]
+C = rand_tensor(modes = r)
+theta = ttl(C,list(A_1,A_2,A_3),ms=1:3)@data*10
+
+### set true cut-off points
+omega = c(-1,1)
+
+### Draw a randomly drawn tensor from above parameters with logistic model.
+ttnsr = realization(theta,omega)
+
+
+##########################################################################################
+### Estimation of the latent parameter theta and cut-off points omega from the observed tensor.
+
+### Initial points for `ordinal_fit' function ( we can use BIC to estimate rank )
+A_1 = randortho(d[1])[,1:r[1]]
+A_2 = randortho(d[2])[,1:r[2]]
+A_3 = randortho(d[3])[,1:r[3]]
+C = rand_tensor(modes = r)
+
+### Estimation for parameters 
+parameter_est = fit_ordinal(ttnsr@data,C,A_1,A_2,A_3,omega = TRUE,alpha = 100)
+
+print(paste("MSE of theta: ",sum((parameter_est$theta-theta)^2)/prod(d),
+            ",   MSE of omega: ",sum(parameter_est$omega-omega)^2/length(omega)))
+
+### Predction from estimated parameters.
+mode_prediction = estimation(parameter_est$theta,parameter_est$omega,"mode")
+
+print(paste("MAD : ",mean(abs((mode_prediction-ttnsr)@data)),
+            ",    MCR : ",mean(mode_prediction@data!=ttnsr@data)))
+

### 2020-02-19 ICML/submit/supplements/functions.R

@@ -1,0 +1,543 @@
+library(MASS)
+library(rTensor)
+library(pracma)
+epsilon=10^-4
+
+################### some functions used for analysis #######################################
+#### Output: randomly sampled ordinal tensor based on logistic models given input parameters
+#### Input : 
+#### theta: continuous parameter tensor (latent parameter)
+#### omega: the cut-off points (k-levels)
+realization = function(theta,omega){
+  k=length(omega) 
+  theta=as.tensor(theta)
+  theta_output <- c(theta@data)
+  p = theta_to_p(theta_output,omega)
+  for (j in 1:length(theta_output)) {
+    theta_output[j] <-  sample(1:(k+1),1,prob = p[j,])
+  }
+  return(as.tensor(array(theta_output,dim =theta@modes)))
+}      
+
+#### Output: predicted ordinal tensor from the given input parameters and types of predction
+#### Input : 
+#### theta: continuous parameter tensor (latent parameter)
+#### omega: the cut-off points (k-levels)
+#### type:`mode' the mostly likely label,`mean' the mean label, `median' the median label
+estimation = function(theta,omega,type){ 
+  k=length(omega) 
+  if(is.matrix(theta)){
+    theta_output=c(theta)
+  }else{
+    theta=as.tensor(theta)
+    theta_output <- c(theta@data)
+  }
+  p = theta_to_p(theta_output,omega)
+  if(type=="mode"){  # score prediction based on the mode 
+    for (j in 1:length(theta_output)) theta_output[j] <-  which.max(p[j,])
+  }else if(type=="mean"){ # score prediction based on the mean 
+    for (j in 1:length(theta_output)) theta_output[j] <-  sum(p[j,]*(1:(k+1))) 
+  }else if(type=="median"){# score prediction based on the median 
+    for (j in 1:length(theta_output)) theta_output[j] <-  which(c(cumsum(p[j,]),1)>=0.5)[1] ## median
+  }
+  if(is.matrix(theta)) return(matrix(theta_output,nrow=dim(theta)[1],ncol=dim(theta)[2]))
+  else return(as.tensor(array(theta_output,dim =theta@modes)))
[…]

### 2020-02-19 old_files/old_preprint/miaoyan/ordinalT_supp.tex
clean up

--- 2020-01-21 ICLM/miaoyan/ordinalT_supp.tex
@@
 holds for any $ \varepsilon \geq 0$ when $\gamma\in[0,1]$ is chosen to be sufficiently small depending on $\varepsilon$, e.g., $\gamma \leq \sqrt{3\varepsilon}$.
-By applying Theorem~\ref{thm:Tsybakov} to~\eqref{eq:totalKL}, and in view of the property (iv), we obtain that
+By applying Lemma~\ref{lem:Tsybakov} to~\eqref{eq:totalKL}, and in view of the property (iv), we obtain that
 \begin{equation}\label{eq:final}
@@
 
-We begin with a set of technical lemmas that are useful for the proofs of the main theorems.
+We begin with various notion of tensor norms that are useful for the proofs of the main theorems.
 
-\begin{lem}[\cite{tomioka2014spectral}]\label{lem:tensor}
-Suppose that $\tS=\entry{s_{\omega}}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is an order-$K$ tensor whose entries are independent random variables that satisfy
+\begin{defn}[Atomic M-norm~\citep{ghadermarzy2019near}]
+Define $T_{\pm}=\{\tT\in\{\pm 1 \}^{d_1\times \cdots \times d_K}\colon \text{rank}(\tT)=1\}$. The atomic M-norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
+\begin{align}
+\anormSize{}{\Theta}&=\inf\{t>0: \Theta\in t\text{conv}(T_{\pm})\}\\
+&=\inf\left\{\sum_{\tX\in T_{\pm}}c_x\colon \Theta=\sum_{\tX\in T_{\pm}} c_x\tX, \ c_x>0\right\}.
+\end{align}
+\end{defn}
+
+\begin{defn}[Spectral norm~\citep{lim2005singular}]
+The spectral norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
 \[
-\mathbb{E}(s_{\omega})=0,\quad \text{and} \quad\mathbb{E}(e^{ts_{\omega}})\leq e^{t^2L^2/2}.
+\snormSize{}{\Theta}=\sup\left\{\langle \Theta, \mx_1\otimes \cdots \otimes \mx_K\rangle \colon \vnormSize{}{\mx_k}=1,\ \mx_k\in\mathbb{R}^{d_k},\ \text{for all}\ k\in[K]\right\}.
 \]
-Then the spectral norm $\snormSize{}{\tS}$ satisfies that, 
+\end{defn}
+
+\begin{defn}[Nuclear norm~\citep{friedland2018nuclear}]
+The nuclear norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
 \[
-\snormSize{}{\tS}\leq \sqrt{{8L^2} \log (12K) \sum_k d_k +\log (2/\delta)},
+\nnormSize{}{\Theta}=\inf
+\left\{
+\sum_{i\in[r]}|\lambda_i|\colon \Theta=\sum_{i=1}^r \lambda_i\mx^{(i)}_1\otimes \cdots \otimes \mx^{(i)}_K,\ \vnormSize{}{\mx^{(i)}_k}=1,\ \mx^{(i)}_k\in\mathbb{R}^{d_k},\ \text{for all} k\in[K],\ i\in[r]
+\right\},
 \]
-with probability at least $1-\delta$.
+where the infimum is taken over all $r\in\mathbb{N}$ and $\norm{\mx^{(i)}_k}=1$ for all $i\in[r]$ and $k\in[K]$.
+\end{defn}
+
+\begin{lem}[M-norm and infinity norm~\citep{ghadermarzy2019near}]\label{lem:Mnormbound}
+Let $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ be an order-$K$, rank-$(r_1,\ldots,r_K)$ tensor. Then
+\[
+\mnormSize{}{\Theta}\leq \anormSize{}{\Theta}\leq \left(\prod_k r_k \over r_{\max}\right)^{3\over 2} \mnormSize{}{\Theta}.
+\]
 \end{lem}
 
-\begin{rmk}
-The above lemma provides the bound on the spectral norm of random tensors. The result was firstly presented in~\cite{nguyen2015tensor}, and we adopt the version from~\cite{tomioka2014spectral}.  
-\end{rmk}
-
-\begin{lem} \label{lem:noisytensor}
-Suppose that $\tS=\entry{s_{\omega}}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is an order-$K$ tensor whose entries are independent random variables that satisfy
-\[
-\mathbb{E}(s_{\omega})=0,\quad \text{and}\quad |s_{\omega}|\leq U.
-\]
-Then we have
-\[
-\mathbb{P}\left(\snormSize{}{\tS}\geq C_2 U\sqrt{\sum_k d_k} \right)\leq \exp\left(-C_1  \log K \sum_k d_k\right)
-\]
-where $C_1>0$ is an absolute constant, and $C_2>0$ is a constant that depends only on $K$. 
-\end{lem}
-
-\begin{proof}  Note that the random variable $U^{-1}s_{\omega}$ is zero-mean and supported on $[-1,1]$. Therefore, $U^{-1}s_{\omega}$ is sub-Gaussian with parameter ${1-(-1)\over 2}=1$, i.e.
-\[
-\mathbb{E}(U^{-1}s_{\omega})=0,\quad \text{and}\quad \mathbb{E}(e^{tU^{-1}s_{\omega}})\leq e^{t^2/2}.
-\]
-It follows from Lemma~\ref{lem:tensor} that, with probability at least $1-\delta$, 
-\[
-\snormSize{}{U^{-1}\tS}\leq \sqrt{\left(c_0\log K+c_1\right) \sum_k d_k +\log (2/\delta)},
-\]
-where $c_0, c_1>0$ are two absolute constants. Taking $\delta=\exp (-C_1\log K \sum_k d_k)$ yields the final claim, where $C_2=c_0\log K+c_1+1>0$ is another constant.  
-\end{proof}
-
-\begin{lem} \label{lem:nuclear}
+\begin{lem}[Nuclear norm and F-norm] \label{lem:nuclear}
 Let $\tA\in\mathbb{R}^{d_1\times\cdots\times d_K}$ be an order-$K$ tensor with Tucker $\text{rank}(\tA)=(r_1,\ldots,r_K)$. Then
@@
 
+The following lemma provides the bound on the spectral norm of random tensors. The result was firstly presented in~\cite{nguyen2015tensor}, and we adopt the version from~\cite{tomioka2014spectral}.  
+\begin{lem}[\cite{tomioka2014spectral}]\label{lem:tensor}
+Suppose that $\tS=\entry{s_{\omega}}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is an order-$K$ tensor whose entries are independent random variables that satisfy
+\[
+\mathbb{E}(s_{\omega})=0,\quad \text{and} \quad\mathbb{E}(e^{ts_{\omega}})\leq e^{t^2L^2/2}.
+\]
+Then the spectral norm $\snormSize{}{\tS}$ satisfies that, 
+\[
+\snormSize{}{\tS}\leq \sqrt{{8L^2} \log (12K) \sum_k d_k +\log (2/\delta)},
+\]
+with probability at least $1-\delta$.
+\end{lem}
+
+\begin{lem} \label{lem:noisytensor}
+Suppose that $\tS=\entry{s_{\omega}}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is an order-$K$ tensor whose entries are independent random variables that satisfy
+\[
+\mathbb{E}(s_{\omega})=0,\quad \text{and}\quad |s_{\omega}|\leq U.
+\]
+Then we have
+\[
+\mathbb{P}\left(\snormSize{}{\tS}\geq C_2 U\sqrt{\sum_k d_k} \right)\leq \exp\left(-C_1  \log K \sum_k d_k\right)
+\]
+where $C_1>0$ is an absolute constant, and $C_2>0$ is a constant that depends only on $K$. 
+\end{lem}
+
+\begin{proof}  Note that the random variable $U^{-1}s_{\omega}$ is zero-mean and supported on $[-1,1]$. Therefore, $U^{-1}s_{\omega}$ is sub-Gaussian with parameter ${1-(-1)\over 2}=1$, i.e.
+\[
+\mathbb{E}(U^{-1}s_{\omega})=0,\quad \text{and}\quad \mathbb{E}(e^{tU^{-1}s_{\omega}})\leq e^{t^2/2}.
+\]
+It follows from Lemma~\ref{lem:tensor} that, with probability at least $1-\delta$, 
+\[
+\snormSize{}{U^{-1}\tS}\leq \sqrt{\left(c_0\log K+c_1\right) \sum_k d_k +\log (2/\delta)},
+\]
+where $c_0, c_1>0$ are two absolute constants. Taking $\delta=\exp (-C_1\log K \sum_k d_k)$ yields the final claim, where $C_2=c_0\log K+c_1+1>0$ is another constant.  
+\end{proof}
+
 \begin{lem}\label{lem:KLentry} Let $X,\; Y$ be two discrete random variables taking values on $L$ possible categories, with category probabilities $\{p_\ell\}_{\ell\in[L]}$ and $\{q_\ell\}_{\ell\in[L]}$, respectively.
@@
 
-\begin{thm}[\cite{tsybakov2009introduction}]\label{thm:Tsybakov} 
+\begin{lem}[Theorem 2.5 in~\cite{tsybakov2009introduction}]\label{lem:Tsybakov} 
 Assume that a set $\tX$ contains element $\Theta_0, \Theta_1, \ldots,\Theta_M$ ($M\geq 2$) such that
@@
 \]
-where $d\colon \tX\times \tX\mapsto [0,+\infty]$ is a semi-distance function, $0<\alpha<{1/8}$ and $P_j=P_{\Theta_j}$, $j=0,1\ldots,M$. Then
+where $d\colon \tX\times \tX\mapsto [0,+\infty]$ is a semi-distance function, $0<\alpha<{1/8}$ and $P_j=P_{\Theta_j}$, $j=0,1\ldots,M$. 
+\end{itemize} 
+Then
 \[

### 2020-02-19 old_files/old_preprint/miaoyan/ordinalT_supp_v2.tex
clean up

--- 2020-02-04 ICML/miaoyan/ordinalT_supp_v2.tex
@@
 Most of brain nodes fall into the cluster $\#1$ and the cluster $\#2$ which can be represented as the left side brain and the right side brain.
-From the group $\#3$ to the group $\#8$ , entries of each cluster share the same name encoded in the data and play a similar role in the brain.
-For example, the group $\#3$ and the group $\#4$ are the left and the right side of the regions called Supramarginal gyrus.
+From the cluster $\#3$ to the cluster $\#8$ , entries of each cluster share the same name encoded in the data and play a similar role in the brain.
+For example, the cluster $\#3$ and the cluster $\#4$ are the left and the right side of the regions called Supramarginal gyrus.
 Those regions are known to be in charge of interpreting tactile sensory data and involved in perception of space and limbs location \cite{carlson2012physiology,reed1994nature}.
@@
 Group       & \multicolumn{3}{c|}{\#1}                                                  & \multicolumn{3}{c|}{\#2}                                      \\ \hline
-Brain nodes & \multicolumn{3}{l|}{\begin{tabular}[c]{@{}l@{}}"RMF\_L","Fpole\_L","Insula\_L",\\ "SupF\_L","CaudMF\_L",\\ "parsTriangularis\_L","parsOpercularis\_L",\\ "preCentral\_L","Tpole\_L","SupT\_L",\\ "SupT\_L","postCentral\_L", "SupP\_L",\\ "IP\_L","LO\_L","MOF\_L","SupF\_L",\\ "IsthmusC\_L","Precuneus\_L",\\ "Cuneus\_L","paraHippo\_L",\\ "Lingual\_L","SupT\_L","LO\_L"\end{tabular}} & \multicolumn{3}{l|}{\begin{tabular}[c]{@{}l@{}}"RMF\_R","Fpole\_R","Insula\_R",\\ "SupF\_R", "CaudMF\_R",\\ "parsTriangularis\_R" ,"parsOpercularis\_R",\\ "preCentral\_R", "Tpole\_R","SupT\_R",\\ "SupT\_R","postCentral\_R","SupP\_R",\\ "IP\_R","LO\_R","MOF\_R", "SupF\_R",\\ "IsthmusC\_R","Precuneus\_R",\\ "Cuneus\_R","paraHippo\_R",\\ "Lingual\_R","SupT\_R","LO\_R"\end{tabular}} \\ \hline
+Brain nodes & \multicolumn{3}{l|}{\begin{tabular}[c]{@{}l@{}} ``RMF\_L'',``'Fpole\_L",``Insula\_L",\\ ``SupF\_L",``CaudMF\_L",\\ ``parsTriangularis\_L",``parsOpercularis\_L",\\ ``preCentral\_L",``Tpole\_L",``SupT\_L",\\ ``SupT\_L",``postCentral\_L", ``SupP\_L",\\ ``IP\_L",``LO\_L",``MOF\_L",``SupF\_L",\\ ``IsthmusC\_L",``Precuneus\_L",\\ ``Cuneus\_L",``paraHippo\_L",\\ ``Lingual\_L",``SupT\_L",``LO\_L"\end{tabular}} & \multicolumn{3}{l|}{\begin{tabular}[c]{@{}l@{}}``RMF\_R",``Fpole\_R",``Insula\_R",\\ ``SupF\_R", ``CaudMF\_R",\\ ``parsTriangularis\_R" ,``parsOpercularis\_R",\\ ``preCentral\_R", ``Tpole\_R",``SupT\_R",\\ ``SupT\_R",``postCentral\_R",``SupP\_R",\\ ``IP\_R",``LO\_R",``MOF\_R", ``SupF\_R",\\ ``IsthmusC\_R",``Precuneus\_R",\\ ``Cuneus\_R",``paraHippo\_R",\\ ``Lingual\_R",``SupT\_R",``LO\_R"\end{tabular}} \\ \hline
 Group       & \#3    & \#4    & \#5   & \#6  & \#7 & \#8
 \\ \hline
-Brain nodes & \begin{tabular}[c]{@{}c@{}}"SupraM\_L" \\ "SupraM\_L" \\ "SupraM\_L" \\ "SupraM\_L"\end{tabular}                                      & \begin{tabular}[c]{@{}c@{}}"SupraM\_R" \\ "SupraM\_R" \\ "SupraM\_R" \\ "SupraM\_R"\end{tabular}                                      & \begin{tabular}[c]{@{}c@{}}"IT\_L" \\ "IT\_L" \\ "IT\_L"\end{tabular}                                      & \begin{tabular}[c]{@{}c@{}}"IT\_R" \\ "IT\_R" \\ "IT\_R"\end{tabular}                                                          & \begin{tabular}[c]{@{}c@{}}"MT\_L" \\ "MT\_L" \\ "MT\_L"\end{tabular}                                                         & \begin{tabular}[c]{@{}c@{}}"MT\_R" \\ "MT\_R"\\ "MT\_R"\end{tabular}                                                         \\ \hline
+Brain nodes & \begin{tabular}[c]{@{}c@{}}``SupraM\_L" \\ ``SupraM\_L" \\ ``SupraM\_L" \\ ``SupraM\_L"\end{tabular}                                      & \begin{tabular}[c]{@{}c@{}}``SupraM\_R" \\ ``SupraM\_R" \\ ``SupraM\_R" \\ ``SupraM\_R"\end{tabular}                                      & \begin{tabular}[c]{@{}c@{}}``IT\_L" \\ ``IT\_L" \\ ``IT\_L"\end{tabular}                                      & \begin{tabular}[c]{@{}c@{}}``IT\_R" \\ ``IT\_R" \\ ``IT\_R"\end{tabular}                                                          & \begin{tabular}[c]{@{}c@{}}``MT\_L" \\ ``MT\_L" \\ ``MT\_L"\end{tabular}                                                         & \begin{tabular}[c]{@{}c@{}}``MT\_R" \\ ``MT\_R"\\ ``MT\_R"\end{tabular}                                                         \\ \hline
 \end{tabular}

### 2020-02-19 old_files/old_preprint/old_arXiv/ordinalT_supp_v3.tex
clean up

--- 2020-02-11 ICML/arxiv/ordinalT_supp_v3.tex
@@
 By taking $\varepsilon=1/8$ and $\gamma=1/2$, we conclude from~\eqref{eq:final} that
-\begin{equation*}\label{eq:prob}
+\begin{equation}\label{eq:prob}
 \inf_{\hat \Theta}\sup_{\trueT\in \tP}\mathbb{P}\left(\text{Loss}(\hat \Theta, \trueT)\geq {1 \over 256}\min\left \{ \alpha^2,  {C^{-1}R_{\max}d_{\max}\over d_{\text{total}}}\right \}\right)\geq {1\over 2}\left({3\over 4}- {2\over R_{\max}d_{\max}} \right)\geq {1\over 8}.
-\end{equation*}
+\end{equation}
 This completes the proof.
@@
 Now, we prove the concavity of $\lambda(u,v)$. Note that
-\begin{equation*}
+\begin{equation}
 \lambda(u,v) = \log\big[f(u)-f(v)\big]=\log\big[\int\mathds{1}_{[u,v]}(x)f'(x)dx\big],
-\end{equation*}
+\end{equation}
 where $\mathds{1}_{[u,v]}$ is an indicator function that equals 1 in the interval $[u,v]$, and 0 elsewhere.
@@
 \]
-where the infimum is taken over all $r\in\mathbb{N}$ and $\norm{\mx^{(i)}_k}=1$ for all $i\in[r]$ and $k\in[K]$.
+where the infimum is taken over all $r\in\mathbb{N}$ and $\vnormSize{}{\mx^{(i)}_k}=1$ for all $i\in[r]$ and $k\in[K]$.
 \end{defn}

### 2020-02-19 old_files/old_preprint/old_arXiv/ordinalTv4.tex
clean up

--- 2020-02-13 ICML/arxiv/old_file/ordinalTv3.tex
@@
 \inf_{\hat \Theta }\sup_{\trueT\in\tP} \mathbb{P}\Big\{& \text{MSE}(\hat \Theta, \trueT) \\
-& \geq {1\over 256} \min\left( \alpha^2, \ { Cr_{\max}d_{\max} \over \prod_k d_k} \right) \Big\} \geq {1\over 8},
+& \geq c \min\left( \alpha^2, \ { Cr_{\max}d_{\max} \over \prod_k d_k} \right) \Big\} \geq {1\over 8},
 \end{align}
-where $C=C(\alpha, L, f,\mb)>0$ is a constant independent of tensor dimension and the rank.
+where $C=C(\alpha, L, f,\mb)>0$ and $c>0$ are constants independent of tensor dimension and the rank.
 \end{thm}
@@
 \vspace{-.1cm}
-\subsection{Comparison with alternative methods}
+\subsection{Comparison with alternative methods}\label{sec:compare}
 \vspace{-.1cm}

### 2020-02-19 old_files/old_preprint/ordinalT.tex
clean up

--- 2020-02-05 ICML/submit/ordinalT.tex
@@
 In many cases, however, we encounter datasets of which the entries are qualitative.
-For example, the Netflix problem records the ratings of users to movies over weeks.
+For example, the Netflix problem records the ratings of users to movies over time.
 Each data is a rating on a nominal scale \{{\it very like, like, neutral, dislike, very dislike}\}.
@@
 \begin{itemize}[itemsep=-.5pt,topsep=-.5pt,leftmargin=*,partopsep=-.5pt]
-\item Continuous tensor decomposition ({\bf Continuous-T})~\cite{filipovic2015tucker,acar2010scalable} is the classical Tucker approximation method. 
+\item Continuous tensor decomposition ({\bf Continuous-T})~\cite{acar2010scalable} is a low-rank approximation method based on classical Tucker model. 
 \item One-bit tensor completion ({\bf 1bit-T})~\cite{ghadermarzy2018learning} is a max-norm penalized tensor learning method based on partial binary observations. 

### 2020-02-19 ICML/submit/supplements/demo.R

@@ -1,44 +1,0 @@
-source("functions.R")
-
-### This file contains the code of a simple demo for the tensor estimations and predictions.
-
-### set tensor dimension and rank of the tensor
-d=c(30,30,30) 
-r=c(10,10,10)
-
-
-### set true latent parameter
-A_1 = randortho(d[1])[,1:r[1]]
-A_2 = randortho(d[2])[,1:r[2]]
-A_3 = randortho(d[3])[,1:r[3]]
-C = rand_tensor(modes = r)
-theta = ttl(C,list(A_1,A_2,A_3),ms=1:3)@data*10
-
-### set true cut-off points
-omega = c(-1,1)
-
-### Draw a randomly drawn tensor from above parameters with logistic model.
-ttnsr = realization(theta,omega)
-
-
-##########################################################################################
-### Estimation of the latent parameter theta and cut-off points omega from the observed tensor.
-
-### Initial points for `ordinal_fit' function ( we can use BIC to estimate rank )
-A_1 = randortho(d[1])[,1:r[1]]
-A_2 = randortho(d[2])[,1:r[2]]
-A_3 = randortho(d[3])[,1:r[3]]
-C = rand_tensor(modes = r)
-
-### Estimation for parameters 
-parameter_est = fit_ordinal(ttnsr@data,C,A_1,A_2,A_3,omega = TRUE,alpha = 100)
-
-print(paste("MSE of theta: ",sum((parameter_est$theta-theta)^2)/prod(d),
-            ",   MSE of omega: ",sum(parameter_est$omega-omega)^2/length(omega)))
-
-### Predction from estimated parameters.
-mode_prediction = estimation(parameter_est$theta,parameter_est$omega,"mode")
-
-print(paste("MAD : ",mean(abs((mode_prediction-ttnsr)@data)),
-            ",    MCR : ",mean(mode_prediction@data!=ttnsr@data)))
-

### 2020-02-19 ICML/submit/supplements/functions.R

@@ -1,543 +1,0 @@
-library(MASS)
-library(rTensor)
-library(pracma)
-epsilon=10^-4
-
-################### some functions used for analysis #######################################
-#### Output: randomly sampled ordinal tensor based on logistic models given input parameters
-#### Input : 
-#### theta: continuous parameter tensor (latent parameter)
-#### omega: the cut-off points (k-levels)
-realization = function(theta,omega){
-  k=length(omega) 
-  theta=as.tensor(theta)
-  theta_output <- c(theta@data)
-  p = theta_to_p(theta_output,omega)
-  for (j in 1:length(theta_output)) {
-    theta_output[j] <-  sample(1:(k+1),1,prob = p[j,])
-  }
-  return(as.tensor(array(theta_output,dim =theta@modes)))
-}      
-
-#### Output: predicted ordinal tensor from the given input parameters and types of predction
-#### Input : 
-#### theta: continuous parameter tensor (latent parameter)
-#### omega: the cut-off points (k-levels)
-#### type:`mode' the mostly likely label,`mean' the mean label, `median' the median label
-estimation = function(theta,omega,type){ 
-  k=length(omega) 
-  if(is.matrix(theta)){
-    theta_output=c(theta)
-  }else{
-    theta=as.tensor(theta)
-    theta_output <- c(theta@data)
-  }
-  p = theta_to_p(theta_output,omega)
-  if(type=="mode"){  # score prediction based on the mode 
-    for (j in 1:length(theta_output)) theta_output[j] <-  which.max(p[j,])
-  }else if(type=="mean"){ # score prediction based on the mean 
-    for (j in 1:length(theta_output)) theta_output[j] <-  sum(p[j,]*(1:(k+1))) 
-  }else if(type=="median"){# score prediction based on the median 
-    for (j in 1:length(theta_output)) theta_output[j] <-  which(c(cumsum(p[j,]),1)>=0.5)[1] ## median
-  }
-  if(is.matrix(theta)) return(matrix(theta_output,nrow=dim(theta)[1],ncol=dim(theta)[2]))
-  else return(as.tensor(array(theta_output,dim =theta@modes)))
[…]

### 2020-02-19 code/InCarMusic.R

@@ -2,6 +2,17 @@
 source("functions.R")
 load("../data/InCar_Music.RData")
 tensor[tensor==-1]=NA
+###
+#m=tensor
+#m[is.na(m)]=0
+#m[m!=0]=1
+#count1=apply(m,1,mean)
+#count2=apply(m,2,mean)
+#count3=apply(m,3,mean)
+#tensor=tensor[count1>=0.015,count2>=0.015,count3>=0.015]
+###
+
+tensor[tensor==-1]=NA
 ind=which(is.na(tensor)==F)
 d=dim(tensor)
 set.seed(1)
@@ -33,11 +44,11 @@ train_index = setdiff(ind,test_index)
 train_tensor = tensor
 train_tensor[test_index] = NA
 
-save(test_index,train_index,file=sprintf("InCarMusic/input_%d_sim%d.RData",index,nsim))
+save(test_index,train_index,file=sprintf("InCarMusic_subset/input_%d_sim%d.RData",index,nsim))
 
 ### Method 1. 1-bit tensor completion
 input=M_to_one(train_tensor,5,type="sign")
-writeMat(sprintf("InCarMusic/input_CV%d_sim%d.mat",index,nsim),data=input$binary,E=input$E,ave=input$ave,scale=input$scale)
+writeMat(sprintf("InCarMusic_subset/input_CV%d_sim%d.mat",index,nsim),data=input$binary,E=input$E,ave=input$ave,scale=input$scale)
 }
 }
 
@@ -45,21 +56,21 @@ writeMat(sprintf("InCarMusic/input_CV%d_sim%d.mat",index,nsim),data=input$binary
 ################# Methods 2 and 3: Tucker decomposition CV ################################
 for(nsim in 1:10){
     for(index in 1:5){
-   load(sprintf("InCarMusic/input_%d_sim%d.RData",index,nsim))     
+   load(sprintf("InCarMusic_subset/input_%d_sim%d.RData",index,nsim))     
    train_tensor = tensor
    train_tensor[test_index] = NA
    
 d=dim(train_tensor)
-r=c(1,1,1)
+r=c(2,2,2)
 A_1 = as.matrix(randortho(d[1])[,1:r[1]])
 A_2 = as.matrix(randortho(d[2])[,1:r[2]])
 A_3 = as.matrix(randortho(d[3])[,1:r[3]])
 C = rand_tensor(modes = r)
[…]

### 2020-02-19 code/One_bit_InCar.m

@@ -1,0 +1,22 @@
+p_cc=0.1; 
+doingmax=1; 
+doing1bit=1; 
+f      = @(x) 1./(1+exp(-x));
+fprime = @(x) exp(-x)./((1+exp(-x)).*(1+exp(-x)));
+iter=10; 
+alternations=10;
+infbound=10; 
+
+%%% start simulation
+for nsim = 1:10
+for i = 1:5
+%% categorical encoding
+input=importdata(sprintf("InCarMusic_subset/input_CV%d_sim%d.mat",i,nsim));
+T_recovered=main_noisyor1bit_TC(input.data,input.E,p_cc,doingmax,doing1bit,f,fprime,iter,alternations,infbound);
+T_recovered=T_recovered/norm(T_recovered(:))*input.scale;
+T_recovered=T_recovered+input.ave;
+save(sprintf("InCarMusic_subset/output_CV%d_sim%d.mat",i,nsim),'T_recovered');
+end
+end
+%%% end simulation
+

### 2020-02-19 code/old_code/Musicdata.R

@@ -1,154 +1,0 @@
-source("functions.R")
-load("InCar_Music.RData")
-
-## Choosing argmax index
-expectation = function(theta,omega){
-  k = length(omega)
-  theta=as.tensor(theta)
-  thet <- c(theta@data)
-  p = matrix(nrow = length(thet),ncol = k)
-  for (i in 1:k) {
-    p[,i] = logistic(thet + omega[i])
-  }
-  p =  cbind(p,rep(1,length(thet)))-cbind(rep(0,length(thet)),p)
-  for (j in 1:length(thet)) {
-    thet[j] <-  which.max(p[j,])
-  }
-  return(as.tensor(array(thet,dim =theta@modes)))
-}      
-
-
-fit_missing = function(tensor,ttnsr,C,A_1,A_2,A_3,omega=TRUE,alph){
-  alphbound <- alph+10^-4
-  result = list()
-  error<- 3
-  iter = 0
-  cost=NULL
-  omg = omega
-  
-  while ((error > 10^-4)&(iter<50) ) {
-    iter = iter +1
-    
-    #update omega
-    theta <- ttl(C,list(A_1,A_2,A_3),ms=1:3)@data
-    if(sum(omg)==TRUE) {
-      omega <- polr(as.factor(c(ttnsr[tensor>0]))~offset(-c(theta[tensor>0])))$zeta
-    }
-    (prev <- likelihood(ttnsr[tensor>0],theta[tensor>0],omega))
-    
-    
-    #update A_1
-    W =kronecker(A_3,A_2)%*%t(k_unfold(C,1)@data)
-    A_1 <- comb(A_1,W,ttnsr,1,omega,alphbound)
-    #orthognalize A_1
-    qr_res=qr(A_1)
-    A_1=qr.Q(qr_res)
-    C=ttm(C,qr.R(qr_res),1)
-    theta <- ttl(C,list(A_1,A_2,A_3),ms=1:3)@data
-    (new <- likelihood(ttnsr[tensor>0],theta[tensor>0],omega))
-    
-    if(max(abs(theta))>=alph) break
-    
-    # update A_2
-    W <- kronecker(A_3,A_1)%*%t(k_unfold(C,2)@data)
-    A_2 <- comb(A_2,W,ttnsr,2,omega,alphbound)
-    #orthognalize A_2
-    qr_res=qr(A_2)
-    A_2=qr.Q(qr_res)
[…]

### 2020-02-19 code/old_code/missingTucker.R

@@ -1,175 +1,0 @@
-library(MASS)
-library(rTensor)
-library(pracma)
-
-
-# initial point
-A_1 = randortho(d[1])[,1:r[1]]
-A_2 = randortho(d[2])[,1:r[2]]
-A_3 = randortho(d[3])[,1:r[3]]
-C = rand_tensor(modes = r)
-prevtheta <- ttl(C,list(A_1,A_2,A_3),ms=1:3)@data
-theta=prevtheta-2*gradient_tensor(A_1,A_2,A_3,C,ttnsr,omega)
-
-# constant term for the missing values
-W = array(0, dim(tensor))
-W[tensor>0] = 1
-
-costf = function(C,A_1,A_2,A_3,theta,W){
-  return(sum((W*(ttl(C,list(A_1,A_2,A_3),ms = 1:3)@data-theta))^2)/2)
-}
-
-cgm = function(C,A_1,A_2,A_3,theta,W,i){
-  val = as.tensor(W*(ttl(C,list(A_1,A_2,A_3),ms = 1:3)@data-theta))
-  if(i == 1){
-    g = k_unfold(val,1)@data%*%t(k_unfold(ttl(C,list(A_2,A_3),ms = c(2,3)),1)@data)
-  }else if(i == 2){
-    g = k_unfold(val,2)@data%*%t(k_unfold(ttl(C,list(A_1,A_3),ms = c(1,3)),2)@data)
-  }else if(i == 3){
-    g = k_unfold(val,3)@data%*%t(k_unfold(ttl(C,list(A_1,A_2),ms = c(1,2)),3)@data)
-  }else{
-    g = ttl(val,list(t(A_1),t(A_2),t(A_3)),ms = 1:3)
-  }
-  return(g)
-}
-
-cgma = function(A,G,theta,W,i){
-  g = (k_unfold(as.tensor(W),i)@data*(A%*%G-k_unfold(as.tensor(theta),i)@data))%*%t(G)
-  return(g)
-}
-
-
-tucker_missing = function(A_1,A_2,A_3,C,theta,W,alph = TRUE){
-  alphbound <- alph+10^-4
-  result = list()
-  error<- 3
-  iter = 0
-  cost=NULL
-  if (alph == TRUE) {
-    while ((error > 10^-4)&(iter<50) ) {
-      iter = iter +1
-      (prev = costf(C,A_1,A_2,A_3,theta,W))
-    
-      # update A_1
-      # tic()
-      # l <- lapply(1:nrow(A_1),
[…]

### 2020-02-19 code/old_code/unused_functions.R

@@ -1,228 +1,0 @@
-library(MASS)
-library(rTensor)
-library(pracma)
-library(doParallel)
-epsilon=10^-4
-
-
-
-## A_1, A_2, A_3 
-g1_factor=function(A_1,A_2,A_3,C,ttnsr,k,omega,type="ordinal"){
-    g = gradient_tensor(A_1,A_2,A_3,C,ttnsr,omega,type) ## first, take entrywise gradient w.r.t. theta
-    
-    if(k==1){
-        g = ttl(as.tensor(g),list(matrix(1),t(A_2),t(A_3)),ms = 1:3) ## take gradient w.r.t. row in factor 1
-        g = k_unfold(g,1)@data%*%t(k_unfold(C,1)@data)
-    }
-    else if(k==2){
-        g = ttl(as.tensor(g),list(t(A_1),matrix(1),t(A_3)),ms = 1:3) ## take gradient w.r.t. row in factor 2
-        g = k_unfold(g,2)@data%*%t(k_unfold(C,2)@data)
-    }
-    else if(k==3){
-        g = ttl(as.tensor(g),list(t(A_1),t(A_2),matrix(1)),ms = 1:3) ## take gradient w.r.t. row in factor 3
-        g = k_unfold(g,3)@data%*%t(k_unfold(C,3)@data)
-    }
-    return(g)
-}
-
-####### update one factor matrix at a time while holding others fixed ###########
-factorcomb = function(A_1,A_2,A_3,C,ttnsr,k,omega,alpha=TRUE,type="ordinal"){
-    if(k==1){
-        if (alpha==TRUE) {
-            l <- lapply(1:nrow(A_1),function(i){optim(A_1[i,],
-                function(x) hc(t(as.matrix(x)),A_2,A_3,C,ttnsr[i,,],omega,type),
-                function(x) g1_factor(t(as.matrix(x)),A_2,A_3,C,ttnsr[i,,],1,omega,type),method = "BFGS")$par})
-            nA <- matrix(unlist(l),nrow = nrow(A_1),byrow = T)
-        }else{
-            W <- kronecker(A_3,A_2)%*%t(k_unfold(C,1)@data)
-            l <- lapply(1:nrow(A_1),function(i){constrOptim(A_1[i,],
-                function(x) hc(t(as.matrix(x)),A_2,A_3,C,ttnsr[i,,],omega,type),
[…]

### 2020-02-19 code/ouputdata_HPC/functions_old.R

@@ -1,467 +1,0 @@
-library(MASS)
-library(rTensor)
-library(pracma)
-library(doParallel)
-
-
-
-########### simulation ordinal tensors based on logist model with arbitrary k
-realization = function(theta,omega){
-  k = length(omega)
-  theta=as.tensor(theta)
-  thet <- c(theta@data)
-  p = matrix(nrow = length(thet),ncol = k)
-  for (i in 1:k) {
-    p[,i] = logistic(thet + omega[i])
-  }
-  p =  cbind(p,rep(1,length(thet)))-cbind(rep(0,length(thet)),p)
-  for (j in 1:length(thet)) {
-    thet[j] <-  sample(1:(k+1),1,prob = p[j,])
-  }
-  return(as.tensor(array(thet,dim =theta@modes)))
-}      
-
-
-
-
-########### cost function  with arbitrary k ###########
-h1 = function(A_1,W1,ttnsr,omega){
-  k = length(omega)
-  thet =W1%*%c(A_1)
-  p = matrix(nrow = length(thet),ncol = k)
-  for (i in 1:k) {
-    p[,i] = as.numeric(logistic(thet + omega[i]))
-  }
-  p =  cbind(p,rep(1,length(thet)))-cbind(rep(0,length(thet)),p)
-  l = lapply(1:(k+1),function(i) -log(p[which(c(ttnsr)==i),i]))
-  return(sum(unlist(l)))
-}
-
-hc = function(A_1,A_2,A_3,C,ttnsr,omega){
-  k = length(omega)
-  thet = c(ttl(C,list(A_1,A_2,A_3),ms=1:3)@data)
-  p = matrix(nrow = length(thet),ncol = k)
-  for (i in 1:k) {
-    p[,i] = as.numeric(logistic(thet + omega[i]))
-  }
-  p =  cbind(p,rep(1,length(thet)))-cbind(rep(0,length(thet)),p)
-  l = lapply(1:(k+1),function(i) -log(p[which(c(ttnsr)==i),i]))
-  return(sum(unlist(l)))
-}
-
-# hc(A_1,A_2,A_3,C,ttnsr,omega)
-########### gradient with arbitrary k ###########
-g1 = function(A_1,W,ttnsr,omega){
-  k = length(omega)
-  thet =W%*%c(A_1)
[…]

### 2020-02-19 code/ouputdata_HPC/speedup2.R

@@ -1,518 +1,0 @@
-library(MASS)
-library(rTensor)
-library(pracma)
-library(ggplot2)
-library(ggthemes)
-library(gridExtra)
-
-realization = function(tnsr,alpha){
-  thet <- k_unfold(tnsr,1)@data
-  theta1 <- thet + alpha[1]
-  theta2 <- thet + alpha[2]
-  result <- k_unfold(tnsr,1)@data
-  p1 <- logistic(theta1)
-  p2 <- logistic(theta2)-logistic(theta1)
-  p3 <- matrix(1,nrow = nrow(thet),ncol = ncol(thet))-logistic(theta2)
-  for (i in 1:nrow(thet)) {
-    for(j in 1:ncol(thet)){
-      result[i,j] <- sample(c(1,2,3),1,prob= c(p1[i,j],p2[i,j],p3[i,j]))
-    }
-  }
-  return(k_fold(result,1,modes = tnsr@modes))
-}      
-
-
-
-Hessi = function(A_1,W4,ttnsr,omega){
-  thet =W4%*%c(A_1)
-  p1 = logistic(thet + omega[1])
-  p2 = logistic(thet + omega[2])
-  q1 = p1*(1-p1)
-  q2 = p2*(1-p2)+p1*(1-p1)
-  q3 = p2*(1-p2)
-  H = t(W4[which(c(ttnsr)==1),])%*%diag(q1[which(c(ttnsr)==1)])%*%W4[which(c(ttnsr)==1),]+
-    t(W4[which(c(ttnsr)==2),])%*%diag(q2[which(c(ttnsr)==2)])%*%W4[which(c(ttnsr)==2),]+
-    t(W4[which(c(ttnsr)==3),])%*%diag(q3[which(c(ttnsr)==3)])%*%W4[which(c(ttnsr)==3),]
-  return(H)
-}
-
-
-
-h1 = function(A_1,W1,ttnsr,omega){
-  thet =W1%*%c(A_1)
-  p1 = logistic(thet + omega[1])
-  p2 = logistic(thet + omega[2])
-  p = cbind(p1,p2-p1,1-p2)
-  return(-sum(log(c(p[which(c(ttnsr)==1),1],p[which(c(ttnsr)==2),2],p[which(c(ttnsr)==3),3]))))
-}
-
-
-g1 = function(A_1,W1,ttnsr,omega){
-  thet =W1%*%c(A_1)
[…]

### 2020-02-19 code/simulation_code/Figure1/plot_Figure1.R

@@ -1,0 +1,22 @@
+### Figure: cost function vs. iteration ##
+library(ggplot2)
+dlist=c(25,30)
+rlist=c(5,10)
+
+################ plot ############
+Output=read.table("Figure1.txt",header=T)
+cost=c(unlist(Output[,5:19]))
+iteration=rep(1:15,rep(4,15))
+Setting=rep(c(sprintf("d=25,r=5 (%.2f sec/iter)",Output[1,1]),sprintf("d=25,r=10 (%.2f sec/iter)",Output[2,1]),sprintf("d=30,r=5 (%.2f sec/iter)",Output[3,1]),sprintf("d=30,r=10 (%.2f sec/iter)",Output[4,1])),15)
+data=data.frame(cost=cost,Setting=Setting,iteration=iteration)
+####
+p=ggplot(data=data,aes(x=iteration,y=cost/(10^4)))+geom_line(aes(color=Setting),size=1.2)+geom_point(aes(shape=Setting))+ labs(x='iteration', y=expression("log-likelihood"%*%10^4))
+
+g <- ggplot_build(p)
+col=unique(g$data[[1]]$colour)[c(2,1,4,3)] ## match the line color to curve
+for(r in 1:4){
+    p=p+geom_hline(yintercept=Output[r,4]/(10^4),lty=2,col=col[r])
+}
+pdf("algorithm.pdf",width=6,height=2.8)
+p
+dev.off()

### 2020-02-19 code/simulation_code/Figure2/Figure2.R

@@ -1,0 +1,41 @@
+### Figure: MSE vs. dimension, under various ranks##
+source("../../functions.R")
+set.seed(1)
+Output=NULL
+dlist=seq(from=15,to=40,by=2)
+rlist=c(3,5,8)
+ntotal=30
+MSE_raw=array(0,dim=c(length(rlist),length(dlist),ntotal))
+
+for(j in 1:length(rlist)){
+    for(i in 1:length(dlist)){
+        for(nsim in 1:ntotal){
+        d=rep(dlist[i],3); r=rep(rlist[j],3);
+        B_1 = randortho(d[1])[,1:r[1]]
+        B_2 = randortho(d[2])[,1:r[2]]
+        B_3 = randortho(d[3])[,1:r[3]]
+        D = rand_tensor(modes = r)
+        theta = ttl(D,list(B_1,B_2,B_3),ms=1:3)@data
+        alpha=10
+        theta = theta/max(abs(theta))*alpha ## specify signal = 10
+        omega = logit((1:4)/5) ## equal spaced on the probit scale
+        
+        ttnsr <- realization(theta,omega)@data
+        
+        #initial point
+        A_1 = randortho(d[1])[,1:r[1]]
+        A_2 = randortho(d[2])[,1:r[2]]
+        A_3 = randortho(d[3])[,1:r[3]]
+        C = rand_tensor(modes = r)
+        C=C/max(abs(ttl(C,list(A_1,A_2,A_3),ms=1:3)@data))*alpha/10 ## initial theta is in the interior 
+        thetainit=ttl(C,list(A_1,A_2,A_3),ms=1:3)@data
+        
+        result <- fit_ordinal(ttnsr,C,A_1,A_2,A_3,omega=TRUE,alpha=2*alpha) ## relaxed bound
+        
+        MSE_raw[j,i,nsim]=mean((theta-result$theta)^2)
+    }
+}
+}
+
+Output=data.frame(MSE=c(t(apply(MSE_raw,c(1:2),mean))),rank=as.factor(rep(rlist,rep(length(dlist),length(rlist)))),dim=rep(dlist,length(rlist)),sd=c(t(apply(MSE_raw,c(1:2),sd))))
+save(Output,MSE_raw,file="Figure2.RData")

### 2020-02-19 code/simulation_code/Figure2/plot_Figure2.R

@@ -1,0 +1,25 @@
+### Figure: MSE vs. dimension, under various ranks##
+library(ggplot2)
+dlist=seq(from=15,to=40,by=2)
+rlist=c(3,5,8)
+################ plot ############
+load("Figure2.RData")
+Output$rank=as.factor(Output$rank)
+levels(Output$rank)=c("r=3","r=5","r=8")
+
+p=ggplot(data=Output,aes(x=dim,y=MSE))+geom_line(aes(color=rank),size=1.5)+ labs(x=expression("tensor dimension"~(d)), y='MSE',size=1.5)+theme(text = element_text(size=rel(4)),legend.text = element_text(size = 15),axis.text.x = element_text(size=15),axis.text.y = element_text(size=15))
+
+p=p+stat_function(fun=function(d) 10^2.96/(d^2),lty=2,size=0.9)
+g <- ggplot_build(p)
+col=unlist(unique(g$data[[1]]["colour"]))[c(1,2,3)]
+
+p=p+geom_errorbar(aes(ymin=MSE-sd, ymax=MSE+sd), width=0.3,
+position=position_dodge(0.05),color=col[rep(rep(1:3),length(dlist))],size=0.5)+geom_point(aes(shape=rank),size=2)+scale_shape_manual(values = c(0,2,16))
+
+pdf("error_dimension.pdf",width=5,height=4)
+p
+dev.off()
+
+
+
+

### 2020-02-19 code/simulation_code/Figure3/plot_Figure3.R

@@ -1,0 +1,33 @@
+### Figure 3: MSE vs. signal level (alpha), under various ranks##
+library(ggplot2)
+rlist=c(3,5,8)
+alist=c(0.6,0.7,seq(from=0.8,to=2.5,by=0.3))
+d=rep(20,3)
+################ plot ############
+load("Figure3.RData")
+Output$rank=as.factor(Output$rank)
+levels(Output$rank)=c("r=3","r=5","r=8")
+alist=alist[alist>0.6]
+Output=Output[Output[,3]>0.6,]
+
+#p=ggplot(data=Output,aes(x=alpha,y=MSE))+geom_line(aes(color=rank),size=1.2)+ labs(x=expression("signal level (log10 scale)"), y='relative MSE',size=1.5)+theme(text = element_text(size=rel(4)),legend.text = element_text(size = 15),axis.text.x = element_text(size=15),axis.text.y = element_text(size=15))
+
+p=ggplot(data=Output,aes(x=alpha,y=MSE))+geom_line(aes(color=rank),size=1.5)+ labs(x=expression("signal level"~(log[10]~alpha)), y='relative MSE',size=1.5)+theme(text = element_text(size=rel(4)),legend.text = element_text(size = 15),axis.text.x = element_text(size=15),axis.text.y = element_text(size=15))
+
+g <- ggplot_build(p)
+col=unlist(unique(g$data[[1]]["colour"]))[c(1,2,3)]
+
+
+p=p+geom_errorbar(aes(ymin=MSE-sd, ymax=MSE+sd), width=0.01,
+position=position_dodge(0.05),color=col[rep(rep(1:length(rlist)),length(alist))],size=0.5)+geom_point(aes(shape=rank),size=2)+scale_shape_manual(values = c(0,2,16))
+
+pdf("error_alpha.pdf",width=5,height=4)
+p
+dev.off()
+
+
+
+
+
+
+

### 2020-02-19 code/simulation_code/Figure4/Figure4.R

@@ -1,0 +1,48 @@
+### Figure: MSE vs. sample size (|Omega|), under various rank##
+source("../../functions.R")
+set.seed(1)
+Output=NULL
+d=rep(20,3)
+rlist=c(3,5,8)
+ntotal=30
+rholist=seq(from=0.3,to=1,by=0.1)
+
+#ntotal=1
+#rholist=seq(from=0.9,to=1,by=0.02)
+
+MSE_raw=array(0,dim=c(length(rlist),length(rholist),ntotal))
+for(j in 1:length(rlist)){
+        r=rep(rlist[j],3)
+        B_1 = randortho(d[1])[,1:r[1]]
+        B_2 = randortho(d[2])[,1:r[2]]
+        B_3 = randortho(d[3])[,1:r[3]]
+        D = rand_tensor(modes = r)
+        theta = ttl(D,list(B_1,B_2,B_3),ms=1:3)@data
+        alpha=10
+        theta = theta/max(abs(theta))*alpha ## specify signal = 10
+        omega = logit((1:4)/5) ## equal spaced on the probit scale
+        
+        ttnsr <- realization(theta,omega)@data
+        
+    for(i in 1:length(rholist)){
+        for(nsim in 1:ntotal){
+        rho=rholist[i]
+        data=array(NA,dim=dim(ttnsr))
+        Obs=sample(1:prod(d),prod(d)*rho,replace=F)
+        data[Obs]=ttnsr[Obs]
+        #initial point
+        A_1 = randortho(d[1])[,1:r[1]]
+        A_2 = randortho(d[2])[,1:r[2]]
+        A_3 = randortho(d[3])[,1:r[3]]
+        C = rand_tensor(modes = r)
+        C=C/max(abs(ttl(C,list(A_1,A_2,A_3),ms=1:3)@data))*alpha/10 ## initial theta is in the interior 
+        thetainit=ttl(C,list(A_1,A_2,A_3),ms=1:3)@data
+        
+        result <- fit_ordinal(data,C,A_1,A_2,A_3,omega=TRUE,alpha=alpha) #relaxed bound
+        MSE_raw[j,i,nsim]=mean((theta-result$theta)^2)/mean(theta^2)
+        }
+    }
+}
+
[…]

### 2020-02-19 code/simulation_code/Figure4/plot_Figure4.R

@@ -1,0 +1,30 @@
+### Figure: MSE vs. sample size (|Omega|), under various rank##
+library(ggplot2)
+d=rep(20,3)
+rlist=c(3,5,8)
+rholist=seq(from=0.3,to=1,by=0.1)
+load("Figure4.RData")
+
+Output$rank=as.factor(Output$rank)
+levels(Output$rank)=c("r=3","r=5","r=8")
+
+p=ggplot(data=Output,aes(x=rho,y=MSE))+geom_line(aes(color=rank),size=1.5)+geom_point(aes(shape=rank))+ labs(x=expression('observation fraction'~(abs(Omega)/d^K)), y='relative MSE',size=1.2)+theme(text = element_text(size=rel(4)),legend.text = element_text(size = 15),axis.text.x = element_text(size=15),axis.text.y = element_text(size=15))
+
+#p=p+stat_function(fun=function(d)10^5/(d^2),lty=2)
+g <- ggplot_build(p)
+col=unlist(unique(g$data[[1]]["colour"]))[c(1,2,3)]
+
+p=p+geom_errorbar(aes(ymin=MSE-sd, ymax=MSE+sd), width=0.01,
+position=position_dodge(0.05),color=col[rep(rep(1:length(rlist)),length(rholist))],size=0.5)+geom_point(aes(shape=rank),size=2)+scale_shape_manual(values = c(0,2,16))
+
+pdf("error_sample.pdf",width=5,height=4)
+p
+dev.off()
+
+
+
+
+
+
+
+

### 2020-02-19 code/simulation_code/Figure5/plot_Figure5.R

@@ -1,0 +1,24 @@
+### Figure 5: MSE vs. number of ordinal levels; half observations ##
+library(ggplot2)
+d=rep(20,3)
+rlist=c(3,5,8)
+Klist=2:7
+################ plot ############
+load("Figure5.RData")
+Output$rank=as.factor(Output$rank)
+levels(Output$rank)=c("r=3","r=5","r=8")
+
+p=ggplot(data=Output,aes(x=K,y=MSE))+geom_line(aes(color=rank),size=1.2)+ labs(x=expression('number of ordinal levels'~(L)), y='relative MSE')+xlim(2,7)+theme(text = element_text(size=rel(4)),legend.text = element_text(size = 15),axis.text.x = element_text(size=15),axis.text.y = element_text(size=15))+ylim(0,2)
+
+g <- ggplot_build(p)
+col=unlist(unique(g$data[[1]]["colour"]))[c(1,2,3)]
+
+p=p+geom_errorbar(aes(ymin=MSE-sd, ymax=MSE+sd), width=0.1,
+position=position_dodge(0),color=col[rep(rep(1:3),length(Klist))],size=0.5)+geom_point(aes(shape=rank),size=2)+scale_shape_manual(values = c(0,2,16))
+
+
+pdf("error_level.pdf",width=5,height=4)
+p
+dev.off()
+
+

### 2020-02-19 code/simulation_code/Figure5/run_Figure5.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --mail-user=miaoyan@stat.wisc.edu
+#SBATCH -p long
+#SBATCH -t 5-00:00:00
+#SBATCH -n 4
+#SBATCH --mem-per-cpu=10000M
+module load R/R-3.5.3
+export R_LIBS=/workspace/miaoyan/x86_64-pc-linux-gnu-library/3.4/
+R CMD BATCH Figure5.R output_new.Rout

### 2020-02-19 code/simulation_code/Figure6/1-bit_simulation_Fig6.m

@@ -1,0 +1,27 @@
+p_cc=0.1; 
+doingmax=1; 
+doing1bit=1; 
+f      = @(x) 1./(1+exp(-x));
+fprime = @(x) exp(-x)./((1+exp(-x)).*(1+exp(-x)));
+iter=10; 
+alternations=10;
+infbound=10; 
+
+rholist=[0.4,0.5,0.6,0.7,0.8,0.9,1.0];
+%%% start simulation
+for nsim = 1:30
+for i = 1:7
+%% categorical encoding
+input=importdata(sprintf("simulation_figure/Figure6/alpha_matlab/input_level%.1f_sim%d_categorical.mat",rholist(i),nsim));
+T_recovered=main_noisyor1bit_TC(input.data,input.E,p_cc,doingmax,doing1bit,f,fprime,iter,alternations,infbound);
+save(sprintf("simulation_figure/Figure6/alpha_matlab/output_level%.1f_sim%d_categorical.mat",rholist(i),nsim),'T_recovered');
+
+%% sign encoding
+input=importdata(sprintf("simulation_figure/Figure6/alpha_matlab/input_level%.1f_sim%d_sign.mat",rholist(i),nsim));
+T_recovered=main_noisyor1bit_TC(input.data,input.E,p_cc,doingmax,doing1bit,f,fprime,iter,alternations,infbound);
+T_recovered=T_recovered/norm(T_recovered(:))*input.scale;
+T_recovered=T_recovered+input.ave;
+save(sprintf("simulation_figure/Figure6/alpha_matlab/output_level%.1f_sim%d_sign.mat",rholist(i),nsim),'T_recovered');
+end
+end
+%%% end simulation

### 2020-02-19 code/simulation_code/Figure6/Figure6.R

@@ -1,0 +1,136 @@
+### Figure: MSE vs. sample size (|Omega|) under different methods ##
+source("../../functions.R")
+library("R.matlab")
+set.seed(1)
+d=rep(20,3)
+r=rep(5,3)
+ntotal=30
+K=5
+rholist=seq(from=0.4,to=1,by=0.1)
+Err1=Err2=Err3=Err4=Err5=array(0,dim=c(length(rholist),4,ntotal))
+
+for(nsim in 1:ntotal){
+B_1 = as.matrix(randortho(d[1])[,1:r[1]])
+B_2 = as.matrix(randortho(d[2])[,1:r[2]])
+B_3 = as.matrix(randortho(d[3])[,1:r[3]])
+D = rand_tensor(modes = r)
+theta = ttl(D,list(B_1,B_2,B_3),ms=1:3)@data
+alpha=10
+theta = theta/max(abs(theta))*alpha ## specify signal = 10
+omega = logit((1:4)/5) ## equal spaced on the probit scale
+#omega=0   
+K=length(omega)+1
+##omega=logit((1:2)/3)
+ttnsr <- realization(theta,omega)@data
+ 
+######## prepare input files ############################
+for(i in 1:length(rholist)){
+rho=rholist[i]
+## generate observation for fit_ordinal
+data=array(NA,dim=dim(ttnsr))
+Obs=sample(1:prod(d),prod(d)*rho,replace=F)
+data[Obs]=ttnsr[Obs] 
+save(data,theta,omega,file=sprintf("../Figure6/alpha_R/input_level%.1f_sim%d.RData",rholist[i],nsim))
+#### generate observation for 1-bit tensor
+input=M_to_one(data,K,"category")
+writeMat(sprintf("../Figure6/alpha_matlab/input_level%.1f_sim%d_categorical.mat",rholist[i],nsim),data=input$binary,E=input$E)
+input=M_to_one(data,K,"sign")
+writeMat(sprintf("../Figure6/alpha_matlab/input_level%.1f_sim%d_sign.mat",rholist[i],nsim),data=input$binary,E=input$E,ave=input$ave,scale=input$scale)
+}
+}
+
+######################## decomposition ################################
+for(nsim in 1:ntotal){
+for(i in 1:length(rholist)){
+    load(sprintf("../Figure6/alpha_R/input_level%.1f_sim%d.RData",rholist[i],nsim))
[…]

### 2020-02-19 code/simulation_code/Figure6/plot_Figure6.R

@@ -1,0 +1,50 @@
+### Figure: MSE vs. sample size (|Omega|) under different methods ##
+source("../../functions.R")
+library(ggplot2)
+library("R.matlab")
+d=rep(20,3)
+r=rep(5,3)
+rholist=seq(from=0.4,to=1,by=0.1)
+K=5
+load("Figure6.RData")
+
+Err1_mean=apply(Err1,c(1,2),mean)
+Err1_sd=apply(Err1,c(1,2),sd)
+Err2_mean=apply(Err2,c(1,2),mean)
+Err2_sd=apply(Err2,c(1,2),sd)
+Err3_mean=apply(Err3,c(1,2),mean)
+Err3_sd=apply(Err3,c(1,2),sd)
+Err4_mean=apply(Err4,c(1,2),mean)
+Err4_sd=apply(Err4,c(1,2),sd)
+Err5_mean=apply(Err5,c(1,2),mean)
+Err5_sd=apply(Err5,c(1,2),sd)
+
+index=4
+MAD_for_mode=c(Err1_mean[,index],Err2_mean[,index],Err3_mean[,index],Err4_mean[,index],Err5_mean[,index])
+MAD_for_mode_sd=c(Err1_sd[,index],Err2_sd[,index],Err3_sd[,index],Err4_sd[,index],Err5_sd[,index])
+method=c("Ordinal-T (ours)","Continuous-T","Ordinal-M","1bit-category-T","1bit-sign-T")
+method=rep(method,rep(length(rholist),5))
+rho=rep(rholist,5)
+Output=data.frame(MAD=MAD_for_mode,sd=MAD_for_mode_sd,method=method,rho=rho)
+
+################ plot ############
+p=ggplot(data=Output,aes(x=rho,y=MAD))+geom_line(aes(color=method),size=1.2)+ labs(x=expression('Observation fraction'~(abs(Omega)/d^K)), y='MCR')+theme(text = element_text(size=rel(4)),legend.text = element_text(size = 15),axis.text.x = element_text(size=15),axis.text.y = element_text(size=15))+ylim(0,1)
+
+g <- ggplot_build(p)
+col=unlist(unique(g$data[[1]]["colour"]))[c(5,3,4,1,2)]
+
+p=p+geom_errorbar(aes(ymin=MAD-sd, ymax=MAD+sd), width=0.008,
[…]

### 2020-02-19 code/simulation_code/Figure6/run_Figure6.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --mail-user=miaoyan@stat.wisc.edu
+#SBATCH -p long
+#SBATCH -t 5-00:00:00
+#SBATCH -n 4
+#SBATCH --mem-per-cpu=10000M
+module load R/R-3.5.3
+export R_LIBS=/workspace/miaoyan/x86_64-pc-linux-gnu-library/3.4/
+R CMD BATCH Figure6.R output_new.Rout

### 2020-02-19 code/simulation_code/Figure7/Figure7.R

@@ -1,0 +1,138 @@
+### Figure: MSE vs. number of levels (K) under different methods ##
+source("../../functions.R")
+library("R.matlab")
+set.seed(1)
+d=rep(20,3)
+r=rep(5,3)
+ntotal=30
+Klist=2:7
+rho=0.8
+Err1=Err2=Err3=Err4=Err5=array(0,dim=c(length(Klist),4,ntotal))
+
+for(nsim in 1:ntotal){
+B_1 = as.matrix(randortho(d[1])[,1:r[1]])
+B_2 = as.matrix(randortho(d[2])[,1:r[2]])
+B_3 = as.matrix(randortho(d[3])[,1:r[3]])
+D = rand_tensor(modes = r)
+theta = ttl(D,list(B_1,B_2,B_3),ms=1:3)@data
+alpha=10
+theta = theta/max(abs(theta))*alpha ## specify signal = 10
+for(i in 1:length(Klist)){
+K=Klist[i]
+omega = logit((1:(K-1))/K) ## equal spaced on the probit scale
+#omega=0   
+#K=length(omega)+1
+##omega=logit((1:2)/3)
+ttnsr <- realization(theta,omega)@data
+ 
+######## prepare input files ############################
+## generate observation for fit_ordinal
+data=array(NA,dim=dim(ttnsr))
+Obs=sample(1:prod(d),prod(d)*rho,replace=F)
+data[Obs]=ttnsr[Obs] 
+save(data,theta,omega,file=sprintf("../Figure8/K_R/input_level%.1f_sim%d.RData",Klist[i],nsim))
+#### generate observation for 1-bit tensor
+input=M_to_one(data,K,"category")
+writeMat(sprintf("../Figure8/K_matlab/input_level%.1f_sim%d_categorical.mat",Klist[i],nsim),data=input$binary,E=input$E)
+input=M_to_one(data,K,"sign")
+writeMat(sprintf("../Figure8/K_matlab/input_level%.1f_sim%d_sign.mat",Klist[i],nsim),data=input$binary,E=input$E,ave=input$ave,scale=input$scale)
+}
+}
+
+######################## decomposition ################################
+for(nsim in 1:ntotal){
+for(i in 1:length(Klist)){
+    K=Klist[i]
+    load(sprintf("../Figure8/K_R/input_level%.1f_sim%d.RData",Klist[i],nsim))
+    ######### Method 1: ordinal tensor ##################
[…]

### 2020-02-19 code/simulation_code/Figure7/One_bit_simulation_Fig7.m

@@ -1,0 +1,30 @@
+addpath('/workspace/miaoyan/TensorCompletion_1bit_noisy-master/tensorlab/') 
+addpath('/workspace/miaoyan/TensorCompletion_1bit_noisy-master') 
+
+p_cc=0.1; 
+doingmax=1; 
+doing1bit=1; 
+f      = @(x) 1./(1+exp(-x));
+fprime = @(x) exp(-x)./((1+exp(-x)).*(1+exp(-x)));
+iter=10; 
+alternations=10;
+infbound=10; 
+
+Klist=[2,3,4,5,6,7];
+%%% start simulation
+for nsim = 1:30
+for i = 1:6
+%% categorical encoding
+input=importdata(sprintf("K_matlab/input_level%.1f_sim%d_categorical.mat",Klist(i),nsim));
+T_recovered=main_noisyor1bit_TC(input.data,input.E,p_cc,doingmax,doing1bit,f,fprime,iter,alternations,infbound);
+save(sprintf("K_matlab/output_level%.1f_sim%d_categorical.mat",Klist(i),nsim),'T_recovered');
+
+%% sign encoding
+input=importdata(sprintf("K_matlab/input_level%.1f_sim%d_sign.mat",Klist(i),nsim));
+T_recovered=main_noisyor1bit_TC(input.data,input.E,p_cc,doingmax,doing1bit,f,fprime,iter,alternations,infbound);
+T_recovered=T_recovered/norm(T_recovered(:))*input.scale;
+T_recovered=T_recovered+input.ave;
+save(sprintf("K_matlab/output_level%.1f_sim%d_sign.mat",Klist(i),nsim),'T_recovered');
+end
+end
+%%% end simulation

### 2020-02-19 code/simulation_code/Figure7/plot_Figure7.R

@@ -1,0 +1,49 @@
+### Figure: MSE vs. number of levels (L), under different methods ##
+source("../../functions.R")
+library(ggplot2)
+library("R.matlab")
+d=rep(20,3)
+r=rep(5,3)
+rho=0.8
+Klist=2:7
+load("Figure7.RData")
+
+Err1_mean=apply(Err1,c(1,2),mean)
+Err1_sd=apply(Err1,c(1,2),sd)
+Err2_mean=apply(Err2,c(1,2),mean)
+Err2_sd=apply(Err2,c(1,2),sd)
+Err3_mean=apply(Err3,c(1,2),mean)
+Err3_sd=apply(Err3,c(1,2),sd)
+Err4_mean=apply(Err4,c(1,2),mean)
+Err4_sd=apply(Err4,c(1,2),sd)
+Err5_mean=apply(Err5,c(1,2),mean)
+Err5_sd=apply(Err5,c(1,2),sd)
+
+index=2
+MAD_for_mode=c(Err1_mean[,index],Err2_mean[,index],Err3_mean[,index],Err4_mean[,index],Err5_mean[,index])
+MAD_for_mode_sd=c(Err1_sd[,index],Err2_sd[,index],Err3_sd[,index],Err4_sd[,index],Err5_sd[,index])
+method=c("Ordinal-T (ours)","Continuous-T","Ordinal-M","1bit-category-T","1bit-sign-T")
+method=rep(method,rep(length(Klist),5))
+K=rep(Klist,5)
+Output=data.frame(MAD=MAD_for_mode,sd=MAD_for_mode_sd,method=method,K=K)
+
+################ plot ############
+p=ggplot(data=Output,aes(x=K,y=MAD))+geom_line(aes(color=method),size=1.2)+ labs(x=expression('Number of ordinal levels'~(L)), y='MCR')+theme(text = element_text(size=rel(4)),legend.text = element_text(size = 15),axis.text.x = element_text(size=15),axis.text.y = element_text(size=15))+ylim(0,1)
+
+g <- ggplot_build(p)
+col=unlist(unique(g$data[[1]]["colour"]))[c(5,3,4,1,2)]
+
+p=p+geom_errorbar(aes(ymin=MAD-sd, ymax=MAD+sd), width=0.01,
[…]

### 2020-02-19 code/simulation_figure 2/Figure1/Figure1.R

@@ -1,40 +1,0 @@
-### Figure: cost function vs. iteration ##
-source("../../functions.R")
-set.seed(1)
-Output=NULL
-dlist=c(25,30)
-rlist=c(5,10)
-
-for(i in dlist){
-    for(j in rlist){
-        d=rep(i,3); r=rep(j,3);
-        B_1 = randortho(d[1])[,1:r[1]]
-        B_2 = randortho(d[2])[,1:r[2]]
-        B_3 = randortho(d[3])[,1:r[3]]
-        D = rand_tensor(modes = r)
-        theta = ttl(D,list(B_1,B_2,B_3),ms=1:3)@data
-        alpha=10
-        theta = theta/max(abs(theta))*alpha ## specify signal = 10
-        omega = c(-0.3,-0.1,0.1,0.3) ## equal spaced
-        
-        ttnsr <- realization(theta,omega)@data
-        
-        #initial point
-        A_1 = randortho(d[1])[,1:r[1]]
-        A_2 = randortho(d[2])[,1:r[2]]
-        A_3 = randortho(d[3])[,1:r[3]]
-        C = rand_tensor(modes = r)
-        C=C/max(abs(ttl(C,list(A_1,A_2,A_3),ms=1:3)@data))*alpha/10 ## initial theta is in the interior 
-        thetainit=ttl(C,list(A_1,A_2,A_3),ms=1:3)@data
-        
-        time=system.time(result <- fit_ordinal(ttnsr,C,A_1,A_2,A_3,omega=TRUE))
-        time=time[2]/length(result$iteration)
-        
-        Output=rbind(Output,c(time,d[1],r[1],-likelihood(ttnsr,theta,omega),-result$cost))
-    }
-}
-
-colnames(Output)=c("time","dim","rank","true_parameter",paste(rep("iter",15),1:15,sep=""))
-write.table(Output,"Figure1.txt",quote=F)
-
-

### 2020-02-19 code/simulation_figure 2/Figure1/plot_Figure1.R

@@ -1,22 +1,0 @@
-### Figure: cost function vs. iteration ##
-library(ggplot2)
-dlist=c(25,30)
-rlist=c(5,10)
-
-################ plot ############
-Output=read.table("Figure1.txt",header=T)
-cost=c(unlist(Output[,5:19]))
-iteration=rep(1:15,rep(4,15))
-Setting=rep(c(sprintf("d=25,r=5 (%.2f sec/iter)",Output[1,1]),sprintf("d=25,r=10 (%.2f sec/iter)",Output[2,1]),sprintf("d=30,r=5 (%.2f sec/iter)",Output[3,1]),sprintf("d=30,r=10 (%.2f sec/iter)",Output[4,1])),15)
-data=data.frame(cost=cost,Setting=Setting,iteration=iteration)
-####
-p=ggplot(data=data,aes(x=iteration,y=cost/(10^4)))+geom_line(aes(color=Setting),size=1.2)+geom_point(aes(shape=Setting))+ labs(x='iteration', y=expression("log-likelihood"%*%10^4))
-
-g <- ggplot_build(p)
-col=unique(g$data[[1]]$colour)[c(2,1,4,3)] ## match the line color to curve
-for(r in 1:4){
-    p=p+geom_hline(yintercept=Output[r,4]/(10^4),lty=2,col=col[r])
-}
-pdf("algorithm.pdf",width=6,height=2.8)
-p
-dev.off()

### 2020-02-19 code/simulation_figure 2/Figure2/Figure2.R

@@ -1,41 +1,0 @@
-### Figure: MSE vs. dimension, under various ranks##
-source("../../functions.R")
-set.seed(1)
-Output=NULL
-dlist=seq(from=15,to=40,by=2)
-rlist=c(3,5,8)
-ntotal=30
-MSE_raw=array(0,dim=c(length(rlist),length(dlist),ntotal))
-
-for(j in 1:length(rlist)){
-    for(i in 1:length(dlist)){
-        for(nsim in 1:ntotal){
-        d=rep(dlist[i],3); r=rep(rlist[j],3);
-        B_1 = randortho(d[1])[,1:r[1]]
-        B_2 = randortho(d[2])[,1:r[2]]
-        B_3 = randortho(d[3])[,1:r[3]]
-        D = rand_tensor(modes = r)
-        theta = ttl(D,list(B_1,B_2,B_3),ms=1:3)@data
-        alpha=10
-        theta = theta/max(abs(theta))*alpha ## specify signal = 10
-        omega = logit((1:4)/5) ## equal spaced on the probit scale
-        
-        ttnsr <- realization(theta,omega)@data
-        
-        #initial point
-        A_1 = randortho(d[1])[,1:r[1]]
-        A_2 = randortho(d[2])[,1:r[2]]
-        A_3 = randortho(d[3])[,1:r[3]]
-        C = rand_tensor(modes = r)
-        C=C/max(abs(ttl(C,list(A_1,A_2,A_3),ms=1:3)@data))*alpha/10 ## initial theta is in the interior 
-        thetainit=ttl(C,list(A_1,A_2,A_3),ms=1:3)@data
-        
-        result <- fit_ordinal(ttnsr,C,A_1,A_2,A_3,omega=TRUE,alpha=2*alpha) ## relaxed bound
-        
-        MSE_raw[j,i,nsim]=mean((theta-result$theta)^2)
-    }
-}
-}
-
-Output=data.frame(MSE=c(t(apply(MSE_raw,c(1:2),mean))),rank=as.factor(rep(rlist,rep(length(dlist),length(rlist)))),dim=rep(dlist,length(rlist)),sd=c(t(apply(MSE_raw,c(1:2),sd))))
-save(Output,MSE_raw,file="Figure2.RData")

### 2020-02-19 code/simulation_figure 2/Figure2/plot_Figure2.R

@@ -1,25 +1,0 @@
-### Figure: MSE vs. dimension, under various ranks##
-library(ggplot2)
-dlist=seq(from=15,to=40,by=2)
-rlist=c(3,5,8)
-################ plot ############
-load("Figure2.RData")
-Output$rank=as.factor(Output$rank)
-levels(Output$rank)=c("r=3","r=5","r=8")
-
-p=ggplot(data=Output,aes(x=dim,y=MSE))+geom_line(aes(color=rank),size=1.5)+ labs(x=expression("tensor dimension"~(d)), y='MSE',size=1.5)+theme(text = element_text(size=rel(4)),legend.text = element_text(size = 15),axis.text.x = element_text(size=15),axis.text.y = element_text(size=15))
-
-p=p+stat_function(fun=function(d) 10^2.96/(d^2),lty=2,size=0.9)
-g <- ggplot_build(p)
-col=unlist(unique(g$data[[1]]["colour"]))[c(1,2,3)]
-
-p=p+geom_errorbar(aes(ymin=MSE-sd, ymax=MSE+sd), width=0.3,
-position=position_dodge(0.05),color=col[rep(rep(1:3),length(dlist))],size=0.5)+geom_point(aes(shape=rank),size=1.5)
-
-pdf("error_dimension.pdf",width=5,height=4)
-p
-dev.off()
-
-
-
-

### 2020-02-19 code/simulation_figure 2/Figure2/run_Figure2.sh

@@ -1,9 +1,0 @@
-#!/bin/bash
-#SBATCH --mail-user=miaoyan@stat.wisc.edu
-#SBATCH -p long
-#SBATCH -t 5-00:00:00
-#SBATCH -n 4
-#SBATCH --mem-per-cpu=10000M
-module load R/R-3.5.3
-export R_LIBS=/workspace/miaoyan/x86_64-pc-linux-gnu-library/3.4/
-R CMD BATCH Figure2.R output_new.Rout

### 2020-02-19 code/simulation_figure 2/Figure3/Figure3.R

@@ -1,44 +1,0 @@
-### Figure 3: MSE vs. signal level (alpha), under various ranks##
-source("../../functions.R")
-set.seed(1)
-Output=NULL
-rlist=c(3,5,8)
-##alist=seq(from=0.7,to=3.5,by=0.5)
-##alist=seq(from=0.5,to=2.5,by=0.3)
-alist=c(0.6,0.7,seq(from=0.8,to=2.5,by=0.3))
-d=rep(20,3)
-ntotal=30
-MSE_raw=array(0,dim=c(length(rlist),length(alist),ntotal))
-
-for(j in 1:length(rlist)){
-    for(i in 1:length(alist)){
-        for(nsim in 1:ntotal){
-        alpha=10^(alist[i]); r=rep(rlist[j],3);
-        B_1 = randortho(d[1])[,1:r[1]]
-        B_2 = randortho(d[2])[,1:r[2]]
-        B_3 = randortho(d[3])[,1:r[3]]
-        D = rand_tensor(modes = r)
-        theta = ttl(D,list(B_1,B_2,B_3),ms=1:3)@data
-        theta = theta/max(abs(theta))*alpha ## specify signal = alpha
-        omega = logit((1:4)/5) ## equal spaced on the probit scale
-        
-        ttnsr <- realization(theta,omega)@data
-        
-        #initial point
-        A_1 = randortho(d[1])[,1:r[1]]
-        A_2 = randortho(d[2])[,1:r[2]]
-        A_3 = randortho(d[3])[,1:r[3]]
-        C = rand_tensor(modes = r)
-        C=C/max(abs(ttl(C,list(A_1,A_2,A_3),ms=1:3)@data))*alpha/10 ## initial theta is in the interior 
-        thetainit=ttl(C,list(A_1,A_2,A_3),ms=1:3)@data
-        
-        result <- fit_ordinal(ttnsr,C,A_1,A_2,A_3,omega=TRUE,alpha=2*alpha) ## relaxed bound
-        #MSE=c(MSE,mean((theta-result$theta)^2)/mean(theta^2))
-        MSE_raw[j,i,nsim]=mean((theta-result$theta)^2)/mean(theta^2)
-        ##sprintf("%d\n%d\n%d",i,j,nsim)
-    }
-}
-}
-
[…]

### 2020-02-19 code/simulation_figure 2/Figure3/plot_Figure3.R

@@ -1,31 +1,0 @@
-### Figure 3: MSE vs. signal level (alpha), under various ranks##
-library(ggplot2)
-rlist=c(3,5,8)
-alist=c(0.6,0.7,seq(from=0.8,to=2.5,by=0.3))
-d=rep(20,3)
-################ plot ############
-load("Figure3.RData")
-Output$rank=as.factor(Output$rank)
-levels(Output$rank)=c("r=3","r=5","r=8")
-
-#p=ggplot(data=Output,aes(x=alpha,y=MSE))+geom_line(aes(color=rank),size=1.2)+ labs(x=expression("signal level (log10 scale)"), y='relative MSE',size=1.5)+theme(text = element_text(size=rel(4)),legend.text = element_text(size = 15),axis.text.x = element_text(size=15),axis.text.y = element_text(size=15))
-
-p=ggplot(data=Output,aes(x=alpha,y=MSE))+geom_line(aes(color=rank),size=1.5)+ labs(x=expression("signal level"~(log[10]~alpha)), y='relative MSE',size=1.5)+theme(text = element_text(size=rel(4)),legend.text = element_text(size = 15),axis.text.x = element_text(size=15),axis.text.y = element_text(size=15))
-
-g <- ggplot_build(p)
-col=unlist(unique(g$data[[1]]["colour"]))[c(1,2,3)]
-
-
-p=p+geom_errorbar(aes(ymin=MSE-sd, ymax=MSE+sd), width=0.01,
-position=position_dodge(0.05),color=col[rep(rep(1:length(rlist)),length(alist))],size=0.5)+geom_point(aes(shape=rank),size=1.5)
-
-pdf("error_alpha.pdf",width=5,height=4)
-p
-dev.off()
-
-
-
-
-
-
-

### 2020-02-19 code/simulation_figure 2/Figure4/Figure4.R

@@ -1,45 +1,0 @@
-### Figure: MSE vs. sample size (|Omega|), under various rank##
-source("../../functions.R")
-set.seed(1)
-Output=NULL
-d=rep(20,3)
-rlist=c(3,5,8)
-ntotal=30
-
-rholist=seq(from=0.3,to=1,by=0.1)
-MSE_raw=array(0,dim=c(length(rlist),length(rholist),ntotal))
-for(j in 1:length(rlist)){
-        r=rep(rlist[j],3)
-        B_1 = randortho(d[1])[,1:r[1]]
-        B_2 = randortho(d[2])[,1:r[2]]
-        B_3 = randortho(d[3])[,1:r[3]]
-        D = rand_tensor(modes = r)
-        theta = ttl(D,list(B_1,B_2,B_3),ms=1:3)@data
-        alpha=10
-        theta = theta/max(abs(theta))*alpha ## specify signal = 10
-        omega = logit((1:4)/5) ## equal spaced on the probit scale
-        
-        ttnsr <- realization(theta,omega)@data
-        
-    for(i in 1:length(rholist)){
-        for(nsim in 1:ntotal){
-        rho=rholist[i]
-        data=array(NA,dim=dim(ttnsr))
-        Obs=sample(1:prod(d),prod(d)*rho,replace=F)
-        data[Obs]=ttnsr[Obs]
-        #initial point
-        A_1 = randortho(d[1])[,1:r[1]]
-        A_2 = randortho(d[2])[,1:r[2]]
-        A_3 = randortho(d[3])[,1:r[3]]
-        C = rand_tensor(modes = r)
-        C=C/max(abs(ttl(C,list(A_1,A_2,A_3),ms=1:3)@data))*alpha/10 ## initial theta is in the interior 
-        thetainit=ttl(C,list(A_1,A_2,A_3),ms=1:3)@data
-        
-        result <- fit_ordinal(data,C,A_1,A_2,A_3,omega=TRUE,alpha=alpha) #relaxed bound
-        MSE_raw[j,i,nsim]=mean((theta-result$theta)^2)/mean(theta^2)
-        }
-    }
-}
-
-Output=data.frame(MSE=c(t(apply(MSE_raw,c(1:2),mean))),rank=as.factor(rep(rlist,rep(length(rholist),length(rlist)))),rho=rep(rholist,length(rlist)),sd=c(t(apply(MSE_raw,c(1:2),sd))))
-save(MSE_raw,Output,file="Figure4.RData")

### 2020-02-19 code/simulation_figure 2/Figure4/plot_Figure4.R

@@ -1,30 +1,0 @@
-### Figure: MSE vs. sample size (|Omega|), under various rank##
-library(ggplot2)
-d=rep(20,3)
-rlist=c(3,5,8)
-rholist=seq(from=0.3,to=1,by=0.1)
-load("Figure4.RData")
-
-Output$rank=as.factor(Output$rank)
-levels(Output$rank)=c("r=3","r=5","r=8")
-
-p=ggplot(data=Output,aes(x=rho,y=MSE))+geom_line(aes(color=rank),size=1.5)+geom_point(aes(shape=rank))+ labs(x=expression('Observation fraction'~(abs(Omega)/d^K)), y='relative MSE',size=1.2)+theme(text = element_text(size=rel(4)),legend.text = element_text(size = 15),axis.text.x = element_text(size=15),axis.text.y = element_text(size=15))
-
-#p=p+stat_function(fun=function(d)10^5/(d^2),lty=2)
-g <- ggplot_build(p)
-col=unlist(unique(g$data[[1]]["colour"]))[c(1,2,3)]
-
-p=p+geom_errorbar(aes(ymin=MSE-sd, ymax=MSE+sd), width=0.01,
-position=position_dodge(0.05),color=col[rep(rep(1:length(rlist)),length(rholist))],size=0.5)+geom_point(aes(shape=rank),size=1.5)
-
-pdf("error_sample.pdf",width=5,height=4)
-p
-dev.off()
-
-
-
-
-
-
-
-

### 2020-02-19 code/simulation_figure 2/Figure5/Figure5.R

@@ -1,48 +1,0 @@
-### Figure 5: MSE vs. number of ordinal levels; half observations ##
-source("../../functions.R")
-set.seed(1)
-Output=NULL
-d=rep(20,3)
-rlist=c(3,5,8)
-rho=0.5
-Klist=2:7
-ntotal=30
-MSE_raw=array(0,dim=c(length(rlist),length(Klist),ntotal))
-for(j in 1:length(rlist)){      
-        r=rep(rlist[j],3)
-        B_1 = randortho(d[1])[,1:r[1]]
-        B_2 = randortho(d[2])[,1:r[2]]
-        B_3 = randortho(d[3])[,1:r[3]]
-        D = rand_tensor(modes = r)
-        theta = ttl(D,list(B_1,B_2,B_3),ms=1:3)@data
-        alpha=10 #
-        theta = theta/max(abs(theta))*alpha ## specify signal = 10
-        
-        for(i in 1:length(Klist)){
-        K=Klist[i]
-        omega = logit((1:(K-1))/K) ## equal space on the logit scale
-        
-        
-        for(nsim in 1:ntotal){ 
-        ttnsr <- realization(theta,omega)@data
-        data=array(NA,dim=dim(ttnsr))
-        Obs=sample(1:prod(d),prod(d)*rho,replace=F)
-        data[Obs]=ttnsr[Obs]
-        #initial point
-        A_1 = randortho(d[1])[,1:r[1]]
-        A_2 = randortho(d[2])[,1:r[2]]
-        A_3 = randortho(d[3])[,1:r[3]]
-        C = rand_tensor(modes = r)
-        C=C/max(abs(ttl(C,list(A_1,A_2,A_3),ms=1:3)@data))*alpha/10 ## initial theta is in the interior 
-        thetainit=ttl(C,list(A_1,A_2,A_3),ms=1:3)@data
-        
-        result <- fit_ordinal(data,C,A_1,A_2,A_3,omega=TRUE,alpha=2*alpha) ## a slightly relaxed boundary
-        
-        MSE_raw[j,i,nsim]=mean((theta-result$theta)^2)/mean(theta^2) ## more stable using rescaled MSE
-            }
-    }
-}
-
[…]

### 2020-02-19 code/simulation_figure 2/Figure5/plot_Figure5.R

@@ -1,23 +1,0 @@
-### Figure 5: MSE vs. number of ordinal levels; half observations ##
-library(ggplot2)
-d=rep(20,3)
-rlist=c(3,5,8)
-Klist=2:7
-################ plot ############
-load("Figure5.RData")
-Output$rank=as.factor(Output$rank)
-levels(Output$rank)=c("r=3","r=5","r=8")
-
-p=ggplot(data=Output,aes(x=K,y=MSE))+geom_line(aes(color=rank),size=1.2)+ labs(x=expression('Number of Ordinal Levels'~(L)), y='relative MSE')+xlim(2,7)+theme(text = element_text(size=rel(4)),legend.text = element_text(size = 15),axis.text.x = element_text(size=15),axis.text.y = element_text(size=15))+ylim(0,2)
-
-g <- ggplot_build(p)
-col=unlist(unique(g$data[[1]]["colour"]))[c(1,2,3)]
-
-p=p+geom_errorbar(aes(ymin=MSE-sd, ymax=MSE+sd), width=0.1,
-position=position_dodge(0),color=col[rep(rep(1:3),length(Klist))],size=0.5)+geom_point(aes(shape=rank),size=1.5)
-
-pdf("error_level.pdf",width=5,height=4)
-p
-dev.off()
-
-

### 2020-02-19 code/simulation_figure 2/Figure6/1-bit_simulation_Fig6.m

@@ -1,27 +1,0 @@
-p_cc=0.1; 
-doingmax=1; 
-doing1bit=1; 
-f      = @(x) 1./(1+exp(-x));
-fprime = @(x) exp(-x)./((1+exp(-x)).*(1+exp(-x)));
-iter=10; 
-alternations=10;
-infbound=10; 
-
-rholist=[0.4,0.5,0.6,0.7,0.8,0.9,1.0];
-%%% start simulation
-for nsim = 1:30
-for i = 1:7
-%% categorical encoding
-input=importdata(sprintf("simulation_figure/Figure6/alpha_matlab/input_level%.1f_sim%d_categorical.mat",rholist(i),nsim));
-T_recovered=main_noisyor1bit_TC(input.data,input.E,p_cc,doingmax,doing1bit,f,fprime,iter,alternations,infbound);
-save(sprintf("simulation_figure/Figure6/alpha_matlab/output_level%.1f_sim%d_categorical.mat",rholist(i),nsim),'T_recovered');
-
-%% sign encoding
-input=importdata(sprintf("simulation_figure/Figure6/alpha_matlab/input_level%.1f_sim%d_sign.mat",rholist(i),nsim));
-T_recovered=main_noisyor1bit_TC(input.data,input.E,p_cc,doingmax,doing1bit,f,fprime,iter,alternations,infbound);
-T_recovered=T_recovered/norm(T_recovered(:))*input.scale;
-T_recovered=T_recovered+input.ave;
-save(sprintf("simulation_figure/Figure6/alpha_matlab/output_level%.1f_sim%d_sign.mat",rholist(i),nsim),'T_recovered');
-end
-end
-%%% end simulation

### 2020-02-19 code/simulation_figure 2/Figure6/Figure6.R

@@ -1,136 +1,0 @@
-### Figure: MSE vs. sample size (|Omega|) under different methods ##
-source("../../functions.R")
-library("R.matlab")
-set.seed(1)
-d=rep(20,3)
-r=rep(5,3)
-ntotal=30
-K=5
-rholist=seq(from=0.4,to=1,by=0.1)
-Err1=Err2=Err3=Err4=Err5=array(0,dim=c(length(rholist),4,ntotal))
-
-for(nsim in 1:ntotal){
-B_1 = as.matrix(randortho(d[1])[,1:r[1]])
-B_2 = as.matrix(randortho(d[2])[,1:r[2]])
-B_3 = as.matrix(randortho(d[3])[,1:r[3]])
-D = rand_tensor(modes = r)
-theta = ttl(D,list(B_1,B_2,B_3),ms=1:3)@data
-alpha=10
-theta = theta/max(abs(theta))*alpha ## specify signal = 10
-omega = logit((1:4)/5) ## equal spaced on the probit scale
-#omega=0   
-K=length(omega)+1
-##omega=logit((1:2)/3)
-ttnsr <- realization(theta,omega)@data
- 
-######## prepare input files ############################
-for(i in 1:length(rholist)){
-rho=rholist[i]
-## generate observation for fit_ordinal
-data=array(NA,dim=dim(ttnsr))
-Obs=sample(1:prod(d),prod(d)*rho,replace=F)
-data[Obs]=ttnsr[Obs] 
-save(data,theta,omega,file=sprintf("../Figure6/alpha_R/input_level%.1f_sim%d.RData",rholist[i],nsim))
-#### generate observation for 1-bit tensor
-input=M_to_one(data,K,"category")
-writeMat(sprintf("../Figure6/alpha_matlab/input_level%.1f_sim%d_categorical.mat",rholist[i],nsim),data=input$binary,E=input$E)
-input=M_to_one(data,K,"sign")
-writeMat(sprintf("../Figure6/alpha_matlab/input_level%.1f_sim%d_sign.mat",rholist[i],nsim),data=input$binary,E=input$E,ave=input$ave,scale=input$scale)
-}
-}
-
-######################## decomposition ################################
-for(nsim in 1:ntotal){
-for(i in 1:length(rholist)){
-    load(sprintf("../Figure6/alpha_R/input_level%.1f_sim%d.RData",rholist[i],nsim))
[…]

### 2020-02-19 code/simulation_figure 2/Figure6/plot_Figure6.R

@@ -1,49 +1,0 @@
-### Figure: MSE vs. sample size (|Omega|) under different methods ##
-source("../../functions.R")
-library(ggplot2)
-library("R.matlab")
-d=rep(20,3)
-r=rep(5,3)
-rholist=seq(from=0.4,to=1,by=0.1)
-K=5
-load("Figure6.RData")
-
-Err1_mean=apply(Err1,c(1,2),mean)
-Err1_sd=apply(Err1,c(1,2),sd)
-Err2_mean=apply(Err2,c(1,2),mean)
-Err2_sd=apply(Err2,c(1,2),sd)
-Err3_mean=apply(Err3,c(1,2),mean)
-Err3_sd=apply(Err3,c(1,2),sd)
-Err4_mean=apply(Err4,c(1,2),mean)
-Err4_sd=apply(Err4,c(1,2),sd)
-Err5_mean=apply(Err5,c(1,2),mean)
-Err5_sd=apply(Err5,c(1,2),sd)
-
-index=1
-MAD_for_mode=c(Err1_mean[,index],Err2_mean[,index],Err3_mean[,index],Err4_mean[,index],Err5_mean[,index])
-MAD_for_mode_sd=c(Err1_sd[,index],Err2_sd[,index],Err3_sd[,index],Err4_sd[,index],Err5_sd[,index])
-method=c("Ordinal-T (ours)","Continuous-T","Ordinal-M","1bit-caregory-T","1bit-sign-T")
-method=rep(method,rep(length(rholist),5))
-rho=rep(rholist,5)
-Output=data.frame(MAD=MAD_for_mode,sd=MAD_for_mode_sd,method=method,rho=rho)
-
-################ plot ############
-p=ggplot(data=Output,aes(x=rho,y=MAD))+geom_line(aes(color=method),size=1.2)+geom_point(aes(shape=method))+ labs(x=expression('Observation fraction'~(abs(Omega)/d^K)), y='MAD')+theme(text = element_text(size=rel(4)),legend.text = element_text(size = 15),axis.text.x = element_text(size=15),axis.text.y = element_text(size=15))+ylim(0,2.1)
-
-g <- ggplot_build(p)
-col=unlist(unique(g$data[[1]]["colour"]))[c(5,3,4,1,2)]
-
-p=p+geom_errorbar(aes(ymin=MAD-sd, ymax=MAD+sd), width=0.008,
[…]

### 2020-02-19 code/simulation_figure 2/Figure6/run_Figure6.sh

@@ -1,9 +1,0 @@
-#!/bin/bash
-#SBATCH --mail-user=miaoyan@stat.wisc.edu
-#SBATCH -p long
-#SBATCH -t 5-00:00:00
-#SBATCH -n 4
-#SBATCH --mem-per-cpu=10000M
-module load R/R-3.5.3
-export R_LIBS=/workspace/miaoyan/x86_64-pc-linux-gnu-library/3.4/
-R CMD BATCH Figure6.R output_new.Rout

### 2020-02-19 code/simulation_figure 2/Figure7/Figure7.R

@@ -1,138 +1,0 @@
-### Figure: MSE vs. number of levels (K) under different methods ##
-source("../../functions.R")
-library("R.matlab")
-set.seed(1)
-d=rep(20,3)
-r=rep(5,3)
-ntotal=30
-Klist=2:7
-rho=0.8
-Err1=Err2=Err3=Err4=Err5=array(0,dim=c(length(Klist),4,ntotal))
-
-for(nsim in 1:ntotal){
-B_1 = as.matrix(randortho(d[1])[,1:r[1]])
-B_2 = as.matrix(randortho(d[2])[,1:r[2]])
-B_3 = as.matrix(randortho(d[3])[,1:r[3]])
-D = rand_tensor(modes = r)
-theta = ttl(D,list(B_1,B_2,B_3),ms=1:3)@data
-alpha=10
-theta = theta/max(abs(theta))*alpha ## specify signal = 10
-for(i in 1:length(Klist)){
-K=Klist[i]
-omega = logit((1:(K-1))/K) ## equal spaced on the probit scale
-#omega=0   
-#K=length(omega)+1
-##omega=logit((1:2)/3)
-ttnsr <- realization(theta,omega)@data
- 
-######## prepare input files ############################
-## generate observation for fit_ordinal
-data=array(NA,dim=dim(ttnsr))
-Obs=sample(1:prod(d),prod(d)*rho,replace=F)
-data[Obs]=ttnsr[Obs] 
-save(data,theta,omega,file=sprintf("../Figure8/K_R/input_level%.1f_sim%d.RData",Klist[i],nsim))
-#### generate observation for 1-bit tensor
-input=M_to_one(data,K,"category")
-writeMat(sprintf("../Figure8/K_matlab/input_level%.1f_sim%d_categorical.mat",Klist[i],nsim),data=input$binary,E=input$E)
-input=M_to_one(data,K,"sign")
-writeMat(sprintf("../Figure8/K_matlab/input_level%.1f_sim%d_sign.mat",Klist[i],nsim),data=input$binary,E=input$E,ave=input$ave,scale=input$scale)
-}
-}
-
-######################## decomposition ################################
-for(nsim in 1:ntotal){
-for(i in 1:length(Klist)){
-    K=Klist[i]
-    load(sprintf("../Figure8/K_R/input_level%.1f_sim%d.RData",Klist[i],nsim))
-    ######### Method 1: ordinal tensor ##################
[…]

### 2020-02-19 code/simulation_figure 2/Figure7/One_bit_simulation_Fig7.m

@@ -1,30 +1,0 @@
-addpath('/workspace/miaoyan/TensorCompletion_1bit_noisy-master/tensorlab/') 
-addpath('/workspace/miaoyan/TensorCompletion_1bit_noisy-master') 
-
-p_cc=0.1; 
-doingmax=1; 
-doing1bit=1; 
-f      = @(x) 1./(1+exp(-x));
-fprime = @(x) exp(-x)./((1+exp(-x)).*(1+exp(-x)));
-iter=10; 
-alternations=10;
-infbound=10; 
-
-Klist=[2,3,4,5,6,7];
-%%% start simulation
-for nsim = 1:30
-for i = 1:6
-%% categorical encoding
-input=importdata(sprintf("K_matlab/input_level%.1f_sim%d_categorical.mat",Klist(i),nsim));
-T_recovered=main_noisyor1bit_TC(input.data,input.E,p_cc,doingmax,doing1bit,f,fprime,iter,alternations,infbound);
-save(sprintf("K_matlab/output_level%.1f_sim%d_categorical.mat",Klist(i),nsim),'T_recovered');
-
-%% sign encoding
-input=importdata(sprintf("K_matlab/input_level%.1f_sim%d_sign.mat",Klist(i),nsim));
-T_recovered=main_noisyor1bit_TC(input.data,input.E,p_cc,doingmax,doing1bit,f,fprime,iter,alternations,infbound);
-T_recovered=T_recovered/norm(T_recovered(:))*input.scale;
-T_recovered=T_recovered+input.ave;
-save(sprintf("K_matlab/output_level%.1f_sim%d_sign.mat",Klist(i),nsim),'T_recovered');
-end
-end
-%%% end simulation

### 2020-02-19 code/simulation_figure 2/Figure7/matlab_Figure7.sh

@@ -1,13 +1,0 @@
-#!/bin/bash
-#SBATCH --mail-user=miaoyan@stat.wisc.edu
-#SBATCH -p long
-#SBATCH -t 5-00:00:00
-#SBATCH -n 4
-#SBATCH --mem-per-cpu=10000M
-export MATLABPATH=/workspace/miaoyan/TensorCompletion_1bit_noisy-master/
-export MATLABPATH=/workspace/miaoyan/TensorCompletion_1bit_noisy-master/tensorlab/
-matlab -nodisplay -nosplash < One_bit_simulation_Fig8.m 
-
-
-
-

### 2020-02-19 code/simulation_figure 2/Figure7/plot_Figure7.R

@@ -1,49 +1,0 @@
-### Figure: MSE vs. number of levels (L), under different methods ##
-source("../../functions.R")
-library(ggplot2)
-library("R.matlab")
-d=rep(20,3)
-r=rep(5,3)
-rho=0.8
-Klist=2:7
-load("Figure8.RData")
-
-Err1_mean=apply(Err1,c(1,2),mean)
-Err1_sd=apply(Err1,c(1,2),sd)
-Err2_mean=apply(Err2,c(1,2),mean)
-Err2_sd=apply(Err2,c(1,2),sd)
-Err3_mean=apply(Err3,c(1,2),mean)
-Err3_sd=apply(Err3,c(1,2),sd)
-Err4_mean=apply(Err4,c(1,2),mean)
-Err4_sd=apply(Err4,c(1,2),sd)
-Err5_mean=apply(Err5,c(1,2),mean)
-Err5_sd=apply(Err5,c(1,2),sd)
-
-index=3
-MAD_for_mode=c(Err1_mean[,index],Err2_mean[,index],Err3_mean[,index],Err4_mean[,index],Err5_mean[,index])
-MAD_for_mode_sd=c(Err1_sd[,index],Err2_sd[,index],Err3_sd[,index],Err4_sd[,index],Err5_sd[,index])
-method=c("Ordinal-T (ours)","Continuous-T","Ordinal-M","1bit-caregory-T","1bit-sign-T")
-method=rep(method,rep(length(Klist),5))
-K=rep(Klist,5)
-Output=data.frame(MAD=MAD_for_mode,sd=MAD_for_mode_sd,method=method,K=K)
-
-################ plot ############
-p=ggplot(data=Output,aes(x=K,y=MAD))+geom_line(aes(color=method),size=1.2)+geom_point(aes(shape=method))+ labs(x=expression('Number of Ordinal Levels'~(L)), y='MAD')+theme(text = element_text(size=rel(4)),legend.text = element_text(size = 15),axis.text.x = element_text(size=15),axis.text.y = element_text(size=15))
-
-g <- ggplot_build(p)
-col=unlist(unique(g$data[[1]]["colour"]))[c(5,3,4,1,2)]
-
-p=p+geom_errorbar(aes(ymin=MAD-sd, ymax=MAD+sd), width=0.01,
[…]

### 2020-02-19 code/simulation_figure 2/Figure7/run_Figure7.sh

@@ -1,9 +1,0 @@
-#!/bin/bash
-#SBATCH --mail-user=miaoyan@stat.wisc.edu
-#SBATCH -p long
-#SBATCH -t 5-00:00:00
-#SBATCH -n 4
-#SBATCH --mem-per-cpu=10000M
-module load R/R-3.5.3
-export R_LIBS=/workspace/miaoyan/x86_64-pc-linux-gnu-library/3.4/
-R CMD BATCH Figure8.R output_new.Rout

### 2020-02-19 code/simulation_figure 2/experiment.R

@@ -1,223 +1,0 @@
-source("functions.R")
-########## brain tensor analysis ##########
-load("../data/dti_brain.RData")
-##ttnsr=tensor[,,attr[,2]=="22-25"] ## subset of individuals
-ttnsr=tensor
-d=dim(ttnsr)
-
-r=c(25,25,7)
-### random initilization
-A_1 = randortho(d[1])[,1:r[1]]
-A_2 = randortho(d[2])[,1:r[2]]
-A_3 = randortho(d[3])[,1:r[3]]
-C = rand_tensor(modes = r)
-########## Music analysis #######################
-source("functions.R")###
-load("../data/InCar_Music.RData")
-ttnsr=tensor
-ttnsr[ttnsr==-1]=NA
-missing=which(is.na(ttnsr)==F)
-d=dim(ttnsr)
-set.seed(1)
-r=c(1,1,1) ## need to select via BIC ..
-alpha=100
-A_1 = as.matrix(randortho(d[1])[,1:r[1]])
-A_2 = as.matrix(randortho(d[2])[,1:r[2]])
-A_3 = as.matrix(randortho(d[3])[,1:r[3]])
-C = rand_tensor(modes = r)
-C=C/max(abs(ttl(C,list(A_1,A_2,A_3),ms=1:3)@data))*alpha/10
-
-
-############# experiement ##########################
-mask=sample(1:2884,2884*0.2,replace=T)
-index=which(is.na(ttnsr)==F)[mask]
-true=ttnsr[index]
-ttnsr[index]=NA
-##### Method 1: continuous tucker #########
-tic()
-result = fit_continuous(ttnsr,C,A_1,A_2,A_3,alpha=alpha)
-toc()
-thetahat=result$theta
-mean((true-thetahat[index])^2) #MSE= 18.96803
-mean(abs(true-thetahat[index])) #MAD= 2.589925
-mean(true!=round(thetahat[index])) ## MCR
-
-###### Method 2: ordinal tucker #########
-tic()
-result2 = fit_ordinal(ttnsr,C,A_1,A_2,A_3,omega=TRUE,alpha =100)
-toc()
-thetahat2=result2$theta
-bic(ttnsr,thetahat2,result2$omega,d,r)## 10860.46
-est=estimation(thetahat2,result2$omega,"mean")
[…]

### 2020-02-19 code/simulation_figure/Figure1/Figure1.R

@@ -1,58 +1,0 @@
-### Figure: cost function vs. iteration ##
-source("../../functions.R")
-library(ggplot2)
-set.seed(1)
-Output=NULL
-dlist=c(25,30)
-rlist=c(5,10)
-
-for(i in dlist){
-    for(j in rlist){
-        d=rep(i,3); r=rep(j,3);
-        B_1 = randortho(d[1])[,1:r[1]]
-        B_2 = randortho(d[2])[,1:r[2]]
-        B_3 = randortho(d[3])[,1:r[3]]
-        D = rand_tensor(modes = r)
-        theta = ttl(D,list(B_1,B_2,B_3),ms=1:3)@data
-        alpha=10
-        theta = theta/max(abs(theta))*alpha ## specify signal = 10
-        omega = c(-0.3,-0.1,0.1,0.3) ## equal spaced
-        
-        ttnsr <- realization(theta,omega)@data
-        
-        #initial point
-        A_1 = randortho(d[1])[,1:r[1]]
-        A_2 = randortho(d[2])[,1:r[2]]
-        A_3 = randortho(d[3])[,1:r[3]]
-        C = rand_tensor(modes = r)
-        C=C/max(abs(ttl(C,list(A_1,A_2,A_3),ms=1:3)@data))*alpha/10 ## initial theta is in the interior 
-        thetainit=ttl(C,list(A_1,A_2,A_3),ms=1:3)@data
-        
-        time=system.time(result <- fit_ordinal(ttnsr,C,A_1,A_2,A_3,omega=TRUE))
-        time=time[2]/length(result$iteration)
-        
-        Output=rbind(Output,c(time,d[1],r[1],-likelihood(ttnsr,theta,omega),-result$cost))
-    }
-}
-
-colnames(Output)=c("time","dim","rank","true_parameter",paste(rep("iter",15),1:15,sep=""))
-write.table(Output,"Figure2.txt",quote=F)
-
-
-################ plot ############
-Output=read.table("Figure2.txt",header=T)
-cost=c(unlist(table[,5:19]))
-iteration=rep(1:15,rep(4,15))
[…]

### 2020-02-19 code/simulation_figure/Figure2/Figure2.R

@@ -1,64 +1,0 @@
-### Figure: MSE vs. dimension, under various ranks##
-source("../../functions.R")
-set.seed(1)
-Output=NULL
-dlist=seq(from=15,to=40,by=2)
-rlist=c(3,5,8)
-ntotal=30
-MSE_raw=array(0,dim=c(length(rlist),length(dlist),ntotal))
-
-for(j in 1:length(rlist)){
-    for(i in 1:length(dlist)){
-        for(nsim in 1:ntotal){
-        d=rep(dlist[i],3); r=rep(rlist[j],3);
-        B_1 = randortho(d[1])[,1:r[1]]
-        B_2 = randortho(d[2])[,1:r[2]]
-        B_3 = randortho(d[3])[,1:r[3]]
-        D = rand_tensor(modes = r)
-        theta = ttl(D,list(B_1,B_2,B_3),ms=1:3)@data
-        alpha=10
-        theta = theta/max(abs(theta))*alpha ## specify signal = 10
-        omega = logit((1:4)/5) ## equal spaced on the probit scale
-        
-        ttnsr <- realization(theta,omega)@data
-        
-        #initial point
-        A_1 = randortho(d[1])[,1:r[1]]
-        A_2 = randortho(d[2])[,1:r[2]]
-        A_3 = randortho(d[3])[,1:r[3]]
-        C = rand_tensor(modes = r)
-        C=C/max(abs(ttl(C,list(A_1,A_2,A_3),ms=1:3)@data))*alpha/10 ## initial theta is in the interior 
-        thetainit=ttl(C,list(A_1,A_2,A_3),ms=1:3)@data
-        
-        result <- fit_ordinal(ttnsr,C,A_1,A_2,A_3,omega=TRUE,alpha=2*alpha) ## relaxed bound
-        
-        MSE_raw[j,i,nsim]=mean((theta-result$theta)^2)
-    }
-}
-}
-
-Output=data.frame(MSE=c(t(apply(MSE_raw,c(1:2),mean))),rank=as.factor(rep(rlist,rep(length(dlist),length(rlist)))),dim=rep(dlist,length(rlist)),sd=c(t(apply(MSE_raw,c(1:2),sd))))
-save(Output,MSE_raw,file="Figure2.RData")
-
-################ plot ############
-library(ggplot2)
-load("Figure2.RData")
-Output$rank=as.factor(Output$rank)
[…]

### 2020-02-19 code/simulation_figure/Figure3/Figure3.R

@@ -1,72 +1,0 @@
-### Figure 3: MSE vs. signal level (alpha), under various ranks##
-source("../../functions.R")
-set.seed(1)
-Output=NULL
-rlist=c(3,5,8)
-##alist=seq(from=0.7,to=3.5,by=0.5)
-##alist=seq(from=0.5,to=2.5,by=0.3)
-alist=c(0.6,0.7,seq(from=0.8,to=2.5,by=0.3))
-d=rep(20,3)
-ntotal=30
-MSE_raw=array(0,dim=c(length(rlist),length(alist),ntotal))
-
-for(j in 1:length(rlist)){
-    for(i in 1:length(alist)){
-        for(nsim in 1:ntotal){
-        alpha=10^(alist[i]); r=rep(rlist[j],3);
-        B_1 = randortho(d[1])[,1:r[1]]
-        B_2 = randortho(d[2])[,1:r[2]]
-        B_3 = randortho(d[3])[,1:r[3]]
-        D = rand_tensor(modes = r)
-        theta = ttl(D,list(B_1,B_2,B_3),ms=1:3)@data
-        theta = theta/max(abs(theta))*alpha ## specify signal = alpha
-        omega = logit((1:4)/5) ## equal spaced on the probit scale
-        
-        ttnsr <- realization(theta,omega)@data
-        
-        #initial point
-        A_1 = randortho(d[1])[,1:r[1]]
-        A_2 = randortho(d[2])[,1:r[2]]
-        A_3 = randortho(d[3])[,1:r[3]]
-        C = rand_tensor(modes = r)
-        C=C/max(abs(ttl(C,list(A_1,A_2,A_3),ms=1:3)@data))*alpha/10 ## initial theta is in the interior 
-        thetainit=ttl(C,list(A_1,A_2,A_3),ms=1:3)@data
-        
-        result <- fit_ordinal(ttnsr,C,A_1,A_2,A_3,omega=TRUE,alpha=2*alpha) ## relaxed bound
-        #MSE=c(MSE,mean((theta-result$theta)^2)/mean(theta^2))
-        MSE_raw[j,i,nsim]=mean((theta-result$theta)^2)/mean(theta^2)
-        ##sprintf("%d\n%d\n%d",i,j,nsim)
-    }
-}
-}
-
[…]

### 2020-02-19 code/simulation_figure/Figure3/run_Figure3.sh

@@ -1,9 +1,0 @@
-#!/bin/bash
-#SBATCH --mail-user=miaoyan@stat.wisc.edu
-#SBATCH -p long
-#SBATCH -t 5-00:00:00
-#SBATCH -n 4
-#SBATCH --mem-per-cpu=10000M
-module load R/R-3.5.3
-export R_LIBS=/workspace/miaoyan/x86_64-pc-linux-gnu-library/3.4/
-R CMD BATCH Figure3.R output_new.Rout

### 2020-02-19 code/simulation_figure/Figure4/Figure4.R

@@ -1,72 +1,0 @@
-### Figure: MSE vs. sample size (|Omega|), under various rank##
-source("../../functions.R")
-set.seed(1)
-Output=NULL
-d=rep(20,3)
-rlist=c(3,5,8)
-ntotal=30
-
-rholist=seq(from=0.3,to=1,by=0.1)
-MSE_raw=array(0,dim=c(length(rlist),length(rholist),ntotal))
-for(j in 1:length(rlist)){
-        r=rep(rlist[j],3)
-        B_1 = randortho(d[1])[,1:r[1]]
-        B_2 = randortho(d[2])[,1:r[2]]
-        B_3 = randortho(d[3])[,1:r[3]]
-        D = rand_tensor(modes = r)
-        theta = ttl(D,list(B_1,B_2,B_3),ms=1:3)@data
-        alpha=10
-        theta = theta/max(abs(theta))*alpha ## specify signal = 10
-        omega = logit((1:4)/5) ## equal spaced on the probit scale
-        
-        ttnsr <- realization(theta,omega)@data
-        
-    for(i in 1:length(rholist)){
-        for(nsim in 1:ntotal){
-        rho=rholist[i]
-        data=array(NA,dim=dim(ttnsr))
-        Obs=sample(1:prod(d),prod(d)*rho,replace=F)
-        data[Obs]=ttnsr[Obs]
-        #initial point
-        A_1 = randortho(d[1])[,1:r[1]]
-        A_2 = randortho(d[2])[,1:r[2]]
-        A_3 = randortho(d[3])[,1:r[3]]
-        C = rand_tensor(modes = r)
-        C=C/max(abs(ttl(C,list(A_1,A_2,A_3),ms=1:3)@data))*alpha/10 ## initial theta is in the interior 
-        thetainit=ttl(C,list(A_1,A_2,A_3),ms=1:3)@data
-        
-        result <- fit_ordinal(data,C,A_1,A_2,A_3,omega=TRUE,alpha=2*alpha) #relaxed bound
-        MSE_raw[j,i,nsim]=mean((theta-result$theta)^2)/mean(theta^2)
-        }
-    }
-}
-
-Output=data.frame(MSE=c(t(apply(MSE_raw,c(1:2),mean))),rank=as.factor(rep(rlist,rep(length(rholist),length(rlist)))),rho=rep(rholist,length(rlist)),sd=c(t(apply(MSE_raw,c(1:2),sd))))
[…]

### 2020-02-19 code/simulation_figure/Figure4/run_Figure4.sh

@@ -1,9 +1,0 @@
-#!/bin/bash
-#SBATCH --mail-user=miaoyan@stat.wisc.edu
-#SBATCH -p long
-#SBATCH -t 5-00:00:00
-#SBATCH -n 4
-#SBATCH --mem-per-cpu=10000M
-module load R/R-3.5.3
-export R_LIBS=/workspace/miaoyan/x86_64-pc-linux-gnu-library/3.4/
-R CMD BATCH Figure4.R output_new.Rout

### 2020-02-19 code/simulation_figure/Figure5/Figure5.R

@@ -1,80 +1,0 @@
-### Figure 5: MSE vs. number of ordinal levels; half observations ##
-source("../../functions.R")
-set.seed(1)
-Output=NULL
-d=rep(20,3)
-rlist=c(3,5,8)
-rho=0.5
-Klist=2:7
-ntotal=30
-MSE_raw=array(0,dim=c(length(rlist),length(Klist),ntotal))
-for(j in 1:length(rlist)){      
-        r=rep(rlist[j],3)
-        B_1 = randortho(d[1])[,1:r[1]]
-        B_2 = randortho(d[2])[,1:r[2]]
-        B_3 = randortho(d[3])[,1:r[3]]
-        D = rand_tensor(modes = r)
-        theta = ttl(D,list(B_1,B_2,B_3),ms=1:3)@data
-        alpha=10 #
-        theta = theta/max(abs(theta))*alpha ## specify signal = 10
-        
-        for(i in 1:length(Klist)){
-        K=Klist[i]
-        omega = logit((1:(K-1))/K) ## equal space on the logit scale
-        
-        
-        for(nsim in 1:ntotal){ 
-        ttnsr <- realization(theta,omega)@data
-        data=array(NA,dim=dim(ttnsr))
-        Obs=sample(1:prod(d),prod(d)*rho,replace=F)
-        data[Obs]=ttnsr[Obs]
-        #initial point
-        A_1 = randortho(d[1])[,1:r[1]]
-        A_2 = randortho(d[2])[,1:r[2]]
-        A_3 = randortho(d[3])[,1:r[3]]
-        C = rand_tensor(modes = r)
-        C=C/max(abs(ttl(C,list(A_1,A_2,A_3),ms=1:3)@data))*alpha/10 ## initial theta is in the interior 
-        thetainit=ttl(C,list(A_1,A_2,A_3),ms=1:3)@data
-        
-        result <- fit_ordinal(data,C,A_1,A_2,A_3,omega=TRUE,alpha=2*alpha) ## a slightly relaxed boundary
-        
-        MSE_raw[j,i,nsim]=mean((theta-result$theta)^2)/mean(theta^2) ## more stable using rescaled MSE
-            }
-    }
-}
-
-Output=data.frame(MSE=c(t(apply(MSE_raw,c(1:2),mean))),rank=as.factor(rep(rlist,rep(length(Klist),3))),K=rep(Klist,length(rlist)),sd=c(t(apply(MSE_raw,c(1:2),sd))))
[…]

### 2020-02-19 code/simulation_figure/Figure5/run_Figure5.sh

@@ -1,9 +1,0 @@
-#!/bin/bash
-#SBATCH --mail-user=miaoyan@stat.wisc.edu
-#SBATCH -p long
-#SBATCH -t 5-00:00:00
-#SBATCH -n 4
-#SBATCH --mem-per-cpu=10000M
-module load R/R-3.5.3
-export R_LIBS=/workspace/miaoyan/x86_64-pc-linux-gnu-library/3.4/
-R CMD BATCH Figure5.R output_new.Rout

### 2020-02-19 code/simulation_figure/simulation.R

@@ -1,176 +1,0 @@
-source("functions.R")
-########## simulate tucker tensor with d, r ##############################
-set.seed(18)
-dlist=c(20,30,40,50,60)
-r=c(3,3,3)
-RMSE=matrix(0,nrow=5,ncol=5)
-set.seed(18)
-
-for(i in 1:5){
-for(nsim in 1:5){
-d=rep(dlist[i],3)
-B_1 = randortho(d[1])[,1:r[1]]
-B_2 = randortho(d[2])[,1:r[2]]
-B_3 = randortho(d[3])[,1:r[3]]
-D = rand_tensor(modes = r)
-theta = ttl(D,list(B_1,B_2,B_3),ms=1:3)@data
-alpha=10
-theta = theta/max(abs(theta))*alpha ## specify signal = 10
-omega = sort(rnorm(2,0,1))
-
-ttnsr <- realization(theta,omega)@data
-#missing = 1*(rand_tensor(modes = d)>0)@data ## uniform missingness with p=0.5
-#ttnsr[missing==1]=NA
-
-
-#initial point
-A_1 = randortho(d[1])[,1:r[1]]
-A_2 = randortho(d[2])[,1:r[2]]
-A_3 = randortho(d[3])[,1:r[3]]
-C = rand_tensor(modes = r)
-C=C/max(abs(ttl(C,list(A_1,A_2,A_3),ms=1:3)@data))*alpha/10 ## make sure theta is in the interior 
-thetainit=ttl(C,list(A_1,A_2,A_3),ms=1:3)@data
-
-tic()
-result <- fit_ordinal(ttnsr,C,A_1,A_2,A_3,omega=TRUE,alpha=10)
-toc()
-
-RMSE[i,nsim]=sqrt(mean((result$theta-theta)^2))
-}
-}
-
-###
-
-ggplot(data=figure1, aes(x=rescalednpq,y=sqrtmse))+geom_line(aes(color=d1d2d3))+
-scale_shape_manual(values=seq(0,15))+
-geom_point(aes(shape=d1d2d3))+
-labs(x='Rescaled sample size N', y="Root mean squared error (RMSE)", 
-color="Number of clusters", shape="Number of clusters")+
-geom_point(data=rs.order4,aes(x=rescalednpqs,y=sqrtmse,shape="(4,4,4,4)"))+
-geom_line(data=rs.order4,aes(x=rescalednpqs,y=sqrtmse,color="(4,4,4,4)"))
-
-
-
[…]

### 2020-02-19 old_files/1-bit_simulation.m

@@ -1,0 +1,30 @@
+p_cc=0.1; 
+doingmax=1; 
+doing1bit=1; 
+f      = @(x) 1./(1+exp(-x));
+fprime = @(x) exp(-x)./((1+exp(-x)).*(1+exp(-x)));
+iter=10; 
+alternations=10;
+infbound=10; 
+
+rholist=[0.4,0.5,0.6,0.7,0.8,0.9,1.0];
+%%% start simulation
+for i = 1:7
+%% categorical encoding
+input=importdata(sprintf("simulation_figure/Figure6/alpha_matlab/input_level%.2f_categorical.mat",rholist(i)));
+T_recovered=main_noisyor1bit_TC(input.data,input.E,p_cc,doingmax,doing1bit,f,fprime,iter,alternations,infbound);
+save(sprintf("simulation_figure/Figure6/alpha_matlab/output_level%.2f_categorical.mat",rholist(i)),'T_recovered');
+
+%% cumulative encoding
+input=importdata(sprintf("simulation_figure/Figure6/alpha_matlab/input_level%.2f_cumulative.mat",rholist(i)));
+T_recovered=main_noisyor1bit_TC(input.data,input.E,p_cc,doingmax,doing1bit,f,fprime,iter,alternations,infbound);
+save(sprintf("simulation_figure/Figure6/alpha_matlab/output_level%.2f_cumulative.mat",rholist(i)),'T_recovered');
+
+%% sign encoding
+input=importdata(sprintf("simulation_figure/Figure6/alpha_matlab/input_level%.2f_sign.mat",rholist(i)));
+T_recovered=main_noisyor1bit_TC(input.data,input.E,p_cc,doingmax,doing1bit,f,fprime,iter,alternations,infbound);
+T_recovered=T_recovered/norm(T_recovered(:))*input.scale;
+T_recovered=T_recovered+input.ave;
+save(sprintf("simulation_figure/Figure6/alpha_matlab/output_level%.2f_sign.mat",rholist(i)),'T_recovered');
+end
+%%% end simulation

### 2020-02-19 old_files/Figure6.R

@@ -1,0 +1,177 @@
+### Figure: MSE vs. sample size (|Omega|), under various rank##
+source("../../functions.R")
+library("R.matlab")
+set.seed(1)
+Output=NULL
+d=rep(20,3)
+r=rep(5,3) ## r cannot 1, otherwise worse than other methods...
+ntotal=30
+rholist=seq(from=0.5,to=1,by=0.05)
+Err=Err2=Err3=Err4=Err5=Err6=matrix(0,nrow=length(rholist),ncol=4)
+
+#### start
+B_1 = as.matrix(randortho(d[1])[,1:r[1]])
+B_2 = as.matrix(randortho(d[2])[,1:r[2]])
+B_3 = as.matrix(randortho(d[3])[,1:r[3]])
+D = rand_tensor(modes = r)
+theta = ttl(D,list(B_1,B_2,B_3),ms=1:3)@data
+alpha=10
+theta = theta/max(abs(theta))*alpha ## specify signal = 10
+omega = logit((1:4)/5) ## equal spaced on the probit scale
+#omega=0   
+K=length(omega)+1
+##omega=logit((1:2)/3)
+ttnsr <- realization(theta,omega)@data
+ 
+######## start############################
+for(i in 1:length(rholist)){
+rho=rholist[i]
+## generate observation for fit_ordinal
+data=array(NA,dim=dim(ttnsr))
+Obs=sample(1:prod(d),prod(d)*rho,replace=F)
+data[Obs]=ttnsr[Obs] 
+save(data,theta,omega,file=sprintf("../Figure6/alpha_R/input_level%.2f.RData",rholist[i]))
+#### generate observation for 1-bit tensor
+input=M_to_one(data,K,"category")
+writeMat(sprintf("../Figure6/alpha_matlab/input_level%.2f_categorical.mat",rholist[i]),data=input$binary,E=input$E)
+input=M_to_one(data,K,"cumulative")
+writeMat(sprintf("../Figure6/alpha_matlab/input_level%.2f_cumulative.mat",rholist[i]),data=input$binary,E=input$E)
+input=M_to_one(data,K,"sign")
+writeMat(sprintf("../Figure6/alpha_matlab/input_level%.2f_sign.mat",rholist[i]),data=input$binary,E=input$E,ave=input$ave,scale=input$scale)
+}
+
+######################## decomposition ################################
+for(i in 1:length(rholist)){
+    load(sprintf("../Figure6/alpha_R/input_level%.2f.RData",rholist[i]))
[…]

### 2020-02-19 old_files/HPC_1bit.m

@@ -1,0 +1,23 @@
+p_cc=0.1; 
+doingmax=1; 
+doing1bit=1; 
+f      = @(x) 1./(1+exp(-x));
+fprime = @(x) exp(-x)./((1+exp(-x)).*(1+exp(-x)));
+iter=10; 
+alternations=10;
+infbound=10; 
+
+index=[1,2,3,4,5];
+
+%%% begin HPC analysis
+for i = 1:5
+input=importdata(sprintf("HPC/input_CV23_23_8_%.0f.mat",index(i)));
+T_recovered=main_noisyor1bit_TC(input.data,input.E,p_cc,doingmax,doing1bit,f,fprime,iter,alternations,infbound);
+save(sprintf("HPC/output_CV23_23_8_%.0f.mat",index(i),'T_recovered');
+end
+%%% end HPC analysis
+
+     
+     
+
+

### 2020-02-19 old_files/HPC_1bit_pre:post_processing.R

@@ -1,0 +1,44 @@
+############ Method 3: 1-bit tensor completion CV ########
+##The code below is for a single instance of train_tensor. Please modify the code when multiple instances (either from cross-validation or from replicates) are needed. 
+
+## j-th testset. Need to prepare cindex beforhand. lines 5-8 are copied from lines 38-41 in HPC.analysis.R
+test_index = cindex[[j]]
+train_index = setdiff(1:length(tensor),test_index)
+train_tensor = tensor
+train_tensor[test_index] = NA
+
+###### Step 1. prepare input for matlab
+input=array(0,dim=c(dim(train_tensor),2)) ## suppose the data tensor is a 3-level ordinal tensor. 
+for(k in 1:2){
+    prepare=train_tensor  
+    prepare[train_tensor!=k]=-1
+    prepare[train_tensor==k]=1 ## dummy variable encoding; convert multi-level encoding to binary encodding. 
+    input[,,,k]=prepare
+}
+E=array(1,dim=dim(input))
+E[is.na(input)]=0
+input[E==0]=0 ## missingness encoded as zero in 1-bit tensor completion algorithm. 
+writeMat(paste("HPC/input.mat",data=input,E=E)
+
+######Step 2. open matlab, run the script HPC_1bit.m. 
+
+######Step 3. analysis after getting the above output files
+BCV = as.data.frame(matrix(nrow = 5, ncol = 4))
+names(BCV) = c("mode_MAD","mode_MCR","median_MAD","median_MCR")
+
+for (i in 1:5) { ##5-fold CV
+    test_index = cindex[[i]]
+    output=readMat(paste("HPC/output.mat"))
+    p_sum=apply(logistic(output$T.recovered),c(1:3),sum) ## estimated probability for labels 1 to K-1
+    prob_total=array(0,dim=c(dim(data),3)) ## estimated probability for labels 1 to K
+    prob_total[,,,1:2]=logistic(output$T.recovered)
+    prob_total[,,,3]=1-p_sum
+    est1=apply(prob_total,1:3,function(x)which.max(x)) ## prediction via mode
+    BCV[i,1] = mean(est1[test_index]-tensor[test_index])) ##MAD
+    BCV[i,2] = error_rate = mean(est1[test_index]!=tensor[test_index]) ##MCR
+    est2=apply(prob_total,1:3,function(x)which(cumsum(x)>=0.5)[1])## prediction via median
[…]

### 2020-02-19 old_files/HPC_analysis_miaoyan.R

@@ -1,0 +1,116 @@
+### Cross validation ######
+source("functions.R")
+load("../data/dti_brain.RData")
+## I changed the number of iteration from 50 to 10 in fit_continous/fit_ordinal
+set.seed(18)
+i1=i2=i3=i4=i5=NULL
+
+for(i in 1:3){ ### 5-fold sampling while preserving the lable proportion
+    label=sample(which(tensor==i),length(which(tensor==i))) 
+    l = length(label)%/%5+1
+    i1 = c(i1,label[1:l])
+    i2 = c(i2,label[(l+1):(2*l)])
+    i3 = c(i3,label[(2*l+1):(3*l)])
+    i4 = c(i4,label[(3*l+1):(4*l)])
+    i5 = c(i5,label[(4*l+1):length(ind)])
+}
+cindex = list(i1,i2,i3,i4,i5)
+
+# `index' is for 1:5  
+# I used following lines when I put codes in the server in sh file 
+# args index ==$SLURM_ARRAY_TASK_ID" filename.R
+test_index = cindex[[index]]
+train_index = setdiff(1:length(tensor),test_index)
+train_tensor = tensor
+train_tensor[test_index] = NA
+
+
+#####################################################################################
+################# Method 1: 1-bit tensor completion CV ################################
+###### Step 1. prepare input for matlab
+input=array(0,dim=c(dim(train_tensor),2))
+for(k in 1:2){
+    prepare=train_tensor  
+    prepare[train_tensor!=k]=-1
+    prepare[train_tensor==k]=1
+    input[,,,k]=prepare
+}
+E=array(1,dim=dim(input))
+E[is.na(input)]=0
+input[E==0]=0 ## missingness encoded as zero in 1-bit tensor completion algorithm. 
+writeMat(paste("HPC/input_CV23_23_8_",index,".mat",sep = ""),data=input,E=E)
+
+######Step 2. open matlab, run the script HPC_1bit.m. 
+######Step 3. analysis after getting the above output files
+BCV = as.data.frame(matrix(nrow = 5, ncol = 4))
+names(BCV) = c("mode_MAD","mode_MCR","median_MAD","median_MCR")
+for (i in 1:5) {
+test_index = cindex[[i]]
+output=readMat(paste("HPC/output_CV23_23_8_",index,".mat",sep = ""))
[…]

### 2020-02-19 old_files/Musicdata.R

@@ -1,0 +1,154 @@
+source("functions.R")
+load("InCar_Music.RData")
+
+## Choosing argmax index
+expectation = function(theta,omega){
+  k = length(omega)
+  theta=as.tensor(theta)
+  thet <- c(theta@data)
+  p = matrix(nrow = length(thet),ncol = k)
+  for (i in 1:k) {
+    p[,i] = logistic(thet + omega[i])
+  }
+  p =  cbind(p,rep(1,length(thet)))-cbind(rep(0,length(thet)),p)
+  for (j in 1:length(thet)) {
+    thet[j] <-  which.max(p[j,])
+  }
+  return(as.tensor(array(thet,dim =theta@modes)))
+}      
+
+
+fit_missing = function(tensor,ttnsr,C,A_1,A_2,A_3,omega=TRUE,alph){
+  alphbound <- alph+10^-4
+  result = list()
+  error<- 3
+  iter = 0
+  cost=NULL
+  omg = omega
+  
+  while ((error > 10^-4)&(iter<50) ) {
+    iter = iter +1
+    
+    #update omega
+    theta <- ttl(C,list(A_1,A_2,A_3),ms=1:3)@data
+    if(sum(omg)==TRUE) {
+      omega <- polr(as.factor(c(ttnsr[tensor>0]))~offset(-c(theta[tensor>0])))$zeta
+    }
+    (prev <- likelihood(ttnsr[tensor>0],theta[tensor>0],omega))
+    
+    
+    #update A_1
+    W =kronecker(A_3,A_2)%*%t(k_unfold(C,1)@data)
+    A_1 <- comb(A_1,W,ttnsr,1,omega,alphbound)
+    #orthognalize A_1
+    qr_res=qr(A_1)
+    A_1=qr.Q(qr_res)
+    C=ttm(C,qr.R(qr_res),1)
+    theta <- ttl(C,list(A_1,A_2,A_3),ms=1:3)@data
+    (new <- likelihood(ttnsr[tensor>0],theta[tensor>0],omega))
+    
+    if(max(abs(theta))>=alph) break
+    
+    # update A_2
+    W <- kronecker(A_3,A_1)%*%t(k_unfold(C,2)@data)
+    A_2 <- comb(A_2,W,ttnsr,2,omega,alphbound)
+    #orthognalize A_2
+    qr_res=qr(A_2)
+    A_2=qr.Q(qr_res)
[…]

### 2020-02-19 old_files/missingTucker.R

@@ -1,0 +1,175 @@
+library(MASS)
+library(rTensor)
+library(pracma)
+
+
+# initial point
+A_1 = randortho(d[1])[,1:r[1]]
+A_2 = randortho(d[2])[,1:r[2]]
+A_3 = randortho(d[3])[,1:r[3]]
+C = rand_tensor(modes = r)
+prevtheta <- ttl(C,list(A_1,A_2,A_3),ms=1:3)@data
+theta=prevtheta-2*gradient_tensor(A_1,A_2,A_3,C,ttnsr,omega)
+
+# constant term for the missing values
+W = array(0, dim(tensor))
+W[tensor>0] = 1
+
+costf = function(C,A_1,A_2,A_3,theta,W){
+  return(sum((W*(ttl(C,list(A_1,A_2,A_3),ms = 1:3)@data-theta))^2)/2)
+}
+
+cgm = function(C,A_1,A_2,A_3,theta,W,i){
+  val = as.tensor(W*(ttl(C,list(A_1,A_2,A_3),ms = 1:3)@data-theta))
+  if(i == 1){
+    g = k_unfold(val,1)@data%*%t(k_unfold(ttl(C,list(A_2,A_3),ms = c(2,3)),1)@data)
+  }else if(i == 2){
+    g = k_unfold(val,2)@data%*%t(k_unfold(ttl(C,list(A_1,A_3),ms = c(1,3)),2)@data)
+  }else if(i == 3){
+    g = k_unfold(val,3)@data%*%t(k_unfold(ttl(C,list(A_1,A_2),ms = c(1,2)),3)@data)
+  }else{
+    g = ttl(val,list(t(A_1),t(A_2),t(A_3)),ms = 1:3)
+  }
+  return(g)
+}
+
+cgma = function(A,G,theta,W,i){
+  g = (k_unfold(as.tensor(W),i)@data*(A%*%G-k_unfold(as.tensor(theta),i)@data))%*%t(G)
+  return(g)
+}
+
+
+tucker_missing = function(A_1,A_2,A_3,C,theta,W,alph = TRUE){
+  alphbound <- alph+10^-4
+  result = list()
+  error<- 3
+  iter = 0
+  cost=NULL
+  if (alph == TRUE) {
+    while ((error > 10^-4)&(iter<50) ) {
+      iter = iter +1
+      (prev = costf(C,A_1,A_2,A_3,theta,W))
+    
+      # update A_1
+      # tic()
+      # l <- lapply(1:nrow(A_1),
[…]

### 2020-03-02 paper/arxiv/correction.tex
update
update

\documentclass[11pt]{article}

\usepackage{fancybox}
\usepackage{color}
\usepackage{url}
\usepackage[margin=1in]{geometry}
\newcommand{\theHalgorithm}{\arabic{algorithm}}
\usepackage{comment}
\usepackage{amsmath,amssymb,amsthm,bm,mathtools}
\usepackage{natbib}
\usepackage{dsfont,multirow,hyperref,setspace,enumerate}
\hypersetup{colorlinks,linkcolor={blue},citecolor={blue},urlcolor={red}}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{makecell}

\setlength\parindent{0pt}
\usepackage[parfill]{parskip}

\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{pro}{Property}
\newtheorem{assumption}{Assumption}
\newtheorem{cor}{Corollary}[section]

\theoremstyle{definition}
\newtheorem{defn}{Definition}
\newtheorem{example}{Example}
\newtheorem{rmk}{Remark}

\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

\usepackage{dsfont}
\usepackage{wrapfig}
\usepackage{mathtools}
\mathtoolsset{showonlyrefs}

\newcommand*{\KeepStyleUnderBrace}[1]{%f
\mathop{%
\mathchoice
{\underbrace{\displaystyle#1}}%
{\underbrace{\textstyle#1}}%
{\underbrace{\scriptstyle#1}}%
{\underbrace{\scriptscriptstyle#1}}%
}\limits
}

\begingroup
\makeatletter
\@for\theoremstyle:=definition,remark,plain\do{%
\expandafter\g@addto@macro\csname th@\theoremstyle\endcsname{%
\addtolength\thm@preskip\parskip
}%
}
\endgroup

\input macros.tex
\allowdisplaybreaks

\title{Tensor denoising and completion based on ordinal observations}
\date{}
\author{%
Chanwoo Lee \\
University of Wisconsin -- Madison\\
\texttt{chanwoo.lee@wisc.edu} \\
\and
Miaoyan Wang \\
University of Wisconsin -- Madison\\
\texttt{miaoyan.wang@wisc.edu} \\
}
\usepackage{xr}
\externaldocument{ordinalT_arxiv_v4}

\begin{document}
\begin{center}
{\bf \large Correction of Theorem~\ref{thm:ratejoint}}\\
Miaoyan Wang, March 1, 2020\\
\end{center}

We now extend Theorem~\ref{thm:rate} to the case of unknown cut-off points $\mb$. Assume that the true parameters $(\trueT, \trueb)\in \tP\times \tB$, where the feasible sets are defined as
\begin{align}
\tP&=\{\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}\colon \text{rank}(\tP)\leq \mr,\ \langle \Theta, \tJ \rangle=0,\ \mnormSize{}{\Theta}\leq \alpha\},\\
 \tB&=\{\mb\in\mathbb{R}^{L-1}\colon \mnormSize{}{\mb}\leq \beta,\ \min_{\ell}(b_\ell-b_{\ell-1}) \geq \Delta\}.
\end{align}
Here, $\tJ=\entry{1}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ denotes a tensor of all ones. The constraint $\langle \Theta, \tJ \rangle=0$ is imposed to ensure the identifiability of $\Theta$ and $\mb$. We propose the constrained M-estimator
\begin{equation}\label{eq:joint}
(\hat \Theta,\hat \mb)=\argmax_{(\Theta,\mb)\in\tP\times \tB}\flogl(\Theta,\mb).
\end{equation}
The estimation accuracy is assessed using the mean squared error (MSE):
\[
\text{MSE}\left( \hat \Theta, \trueT \right)={1\over \prod_k d_k}\FnormSize{}{\hat \Theta-\trueT},\quad \text{MSE}\left( \hat \mb, \trueb \right)={1\over L-1}\FnormSize{}{\hat \mb-\trueb}.
\]

To facilitate the examination of MSE, we define an order-$(K+1)$ tensor, $\tZ=\entry{z_{\omega,\ell}}\in\mathbb{R}^{d_1\times \cdots \times d_K\times (L-1)}$, by stacking the parameters $\Theta=\entry{\theta_\omega}$ and $\mb=\entry{b_\ell}$ together. Specifically, let $z_{\omega,\ell}=-\theta_\omega+b_\ell$ for all $\omega\in[d_1]\times \cdots \times [d_K]$ and $\ell\in[L-1]$; that is,
\[
\tZ=-\Theta\otimes \mathbf{1}+\tJ\otimes \mb,
\]
where $\mathbf{1}$ denotes a length-$(L-1)$ vector of all ones. Under the identifiability constraint $\langle \Theta, \tJ \rangle=0$, there is an one-to-one mapping between $\tZ$ and $(\Theta, \mb)$, with $\text{rank}(\tZ)=(\text{rank}(\Theta),2)^T$. Furthermore, 
\begin{equation}\label{eq:decomp}
\FnormSize{}{\hat \tZ-\trueZ}^2=\FnormSize{}{\hat \Theta-\trueT}^2 (L-1)+\FnormSize{}{\hat \mb-\trueb}^2 \left(\prod_kd_k\right),
\end{equation}
where $\trueZ=\trueT\otimes \mathbf{1}+\tJ\otimes \trueb$ and $\hat \tZ=\hat \Theta\otimes \mathbf{1}+\tJ\otimes \hat \mb$.

We make the following assumptions about the link function.
\begin{assumption}\label{ass:joint}
The link function $f\colon \mathbb{R}\mapsto [0,1]$ satisfies the following properties:
\begin{enumerate}
\item $f(z)$ is twice-differentiable and strictly increasing in $z$.
\item $\dot{f}(z)$ is strictly log-concave and symmetric with respect to $z=0$.
\end{enumerate}
\end{assumption}

We define the following constants:
\begin{align}\label{eq:constantZ}
C_{\alpha,\beta,\Delta}&=\max_{|z|\leq \alpha+\beta}\max_{\substack{z'\leq z-\Delta\\z''\geq z+\Delta}}\max\left\{ {\dot{f}(z) \over f(z)-f(z')},\ {\dot{f}(z) \over f(z'')-f(z)}\right\},\\
D_{\alpha,\beta,\Delta}&=\max_{|z|\leq \alpha+\beta}\max_{\substack{z'\leq z-\Delta\\z''\geq z+\Delta}}\max\left\{- {\partial\over \partial z}\left({\dot{f}(z) \over f(z)-f(z')}\right),\  {\partial\over \partial z}\left({\dot{f}(z) \over f(z'')-f(z)}\right) \right\},\\
A_{\alpha,\beta,\Delta}&=\min_{|z|\leq \alpha+\beta}\min_{z'\leq z-\Delta} \left(f(z)-f(z')\right).
\end{align}

\begin{rmk}
The condition $\Delta=\min_{\ell}(b_{\ell}-b_{\ell-1})>0$ on the feasible set $\tB$ guarantees the strict positiveness of $f(z)-f(z')$ and $f(z'')-f(z)$. Therefore, the denominators in the above quantities $C_{\alpha,\beta,\Delta}, D_{\alpha,\beta,\Delta}$ are well-defined. Furthermore, by Theorem~\ref{thm:convexity}, $f(z)-f(z')$ is strictly log-concave in $(z,z')$ for $z\leq z'-\Delta, z,z'\in[-\alpha-\beta,\ \alpha+\beta]$. Based on Assumption~\ref{ass:joint} and closeness of the feasible set, we have $C_{\alpha,\beta,\Delta}>0$, $D_{\alpha,\beta,\Delta}>0$, $A_{\alpha,\beta,\Delta}>0$.
\end{rmk}

\begin{thm}[Statistical convergence with unknown $\mb$]\label{thm:ratejoint}
Consider an ordinal tensor $\tY\in[L]^{d_1\times \cdots \times d_K}$ generated from model~\eqref{eq:model} with the link function $f$ and parameters $(\trueT, \trueb)\in\tP\times \tB$. Suppose the link function $f$ satisfies Assumption~\ref{ass:joint}. Define $r_{\max}=\max_k r_k$, and assume $r_{\max}=\tO(1)$.

Then with very high probability, the estimator in~\eqref{eq:joint} satisfies
\begin{equation}\label{eq:scoreZ}
\FnormSize{}{\hat \tZ-\trueZ}^2 \leq {c_1r^{K-1}_{\max}C^2_{\alpha,\beta,\Delta}\over A^2_{\alpha,\beta,\Delta}D^2_{\alpha,\beta,\Delta}}\left(L-1+\sum_k d_k\right),
\end{equation}
In particular,
\begin{equation}\label{eq:jointTheta}
\mathrm{MSE}\left(\hat \Theta, \trueT \right)\leq\min\left( 4\alpha^2,\  {c_1 r_{\max}^{K-1}  C_{\alpha,\beta,\Delta}^2  \over A^2_{\alpha,\beta,\Delta}D_{\alpha,\beta,\Delta}^2}{ L-1+\sum_k d_k \over (L-1)\prod_k d_k}\right),
\end{equation}
and
\begin{equation}\label{eq:jointb}
\mathrm{MSE}\left(\hat \mb, \trueb \right)\leq \min\left(4\beta^2,\ {c_1 r^{K-1}_{\max} C^2_{\alpha,\beta,\Delta} \over A^2_{\alpha,\beta,\Delta}D_{\alpha,\beta,\Delta}^2}{L-1+\sum_kd_k \over (L-1)\prod_k d_K}\right),
\end{equation}
where $c_1, C_{\alpha,\beta,\Delta}, D_{\alpha,\beta,\Delta}$ are positive constants independent of the tensor dimension, rank, and number of ordinal levels. 
\end{thm}
\begin{proof} (sketch)

Let $\plogZ=\entry{{\partial \tL_\tY\over \partial z_{\omega,\ell}}}\in\mathbb{R}^{d_1\times \cdots \times d_K\times[L-1]}$ denote the score function, and $\mH=\pplogZ$ the Hession matrix. Following the same argument in the previous version (Taylor expansion, $r_{\max}(\tZ)=r_{\max}(\Theta)$, etc), we have
\begin{equation}\label{eq:boundZ}
\FnormSize{}{\hat \tZ-\trueZ}^2 \leq c_1r^{K-1}_{\max}{\snormSize{}{\plogZ(\trueZ)}^2\over \lambda^2_{1} \left(\mH(\check\tZ)\right)},
\end{equation}
where $\plogZ(\trueZ)$ is the score evaluated at $\trueZ$, $\mH(\check \tZ)$ is the Hession evaluated at $\check\tZ$, for some $\check\tZ$ between $\hat \tZ$ and $\trueZ$, and $\lambda_{1}(\cdot)$ is the largest matrix eigenvalue.

Hence, it suffices to bound the score and the Hession.

\begin{enumerate}
\item (Score.) The $(\omega,\ell)$-th entry in $\plogZ$ is 
\[
{\partial \tL_\tY\over \partial z_{\omega,\ell}}=\mathds{1}_{\{y_\omega=\ell\}}{\dot{f}(z)\over f(z)-f(z')}\Bigg|_{(z,\ z')=(z_{\omega,\ell},\ z_{\omega,\ell-1})} - \mathds{1}_{\{y_\omega=\ell+1\}}{\dot{f}(z)\over f(z'')-f(z)}\Bigg|_{(z'',\ z)=(z_{\omega,\ell+1},\ z_{\omega,\ell})},
\]
which is upper bounded in magnitude by $C_{\alpha,\beta,\Delta}>0$. Therefore, with very high probability,
\[
\snormSize{}{\plogZ(\trueZ)}\leq C_{\alpha,\beta,\Delta}\sqrt{L-1+\sum_k d_k}.
\]
\item (Hession.) The entries in the Hession matrix are
\begin{align}
\text{Diagonal: }&{\partial^2 \tL_\tY\over \partial z^2_{\omega,\ell}}=\mathds{1}_{\{y_\omega=\ell\}}{\ddot{f}(z)\left(f(z)-f(z')\right)-\dot{f}^2(z)\over \left(f(z)-f(z')\right)^2}\Bigg|_{(z,\ z')=(z_{\omega,\ell},\ z_{\omega,\ell-1})}-\\
&\hspace{.65in}\mathds{1}_{\{y_\omega=\ell+1\}}{\ddot{f}(z)\left(f(z'')-f(z)\right)+\dot{f}^2(z)\over \left(f(z'')-f(z)\right)^2}\Bigg|_{(z'',\ z)=(z_{\omega,\ell+1},\ z_{\omega,\ell})},\\
\text{Off-diagonal: }&
{\partial^2 \tL_\tY\over \partial z_{\omega,\ell}z_{\omega,\ell+1}}=\mathds{1}_{\{y_\omega=\ell+1\}}{\dot{f}(z_{\omega,\ell})\dot{f}(z_{\omega,\ell+1})\over \left(f(z_{\omega,\ell+1})-f(z_{\omega,\ell})\right)^2}\quad \text{and}\quad {\partial^2 \tL_\tY\over \partial z_{\omega,\ell}z_{\omega',\ell'}}=0 \text{ otherwise}. 
\end{align}

Based on Assumption~\ref{ass:joint}, the Hession matrix $\mH$ has the following three properties:
\begin{enumerate}
\item The Hession matrix is a block matrix, $\mH=\text{diag}\{\mH_\omega: \omega\in[d_1]\times \cdots \times [d_K]\}$, and each block $\mH_{\omega}\in\mathbb{R}^{(L-1)\times (L-1)}$ is a tridiagonal matrix. 
\item The off-diagonal entries are either zero or strictly positive.
\item The diagonal entries are either zero or strictly negative. Furthermore, 
\begin{align}
&\mH_\omega(\ell,\ell)+ \mH_\omega(\ell,\ell-1)+\mH_\omega(\ell,\ell+1)\\
=&{\partial^2 \tL_\tY\over \partial z^2_{\omega,\ell}}+{\partial^2 \tL_\tY\over \partial z_{\omega,\ell}z_{\omega,\ell+1}}+{\partial^2 \tL_\tY\over \partial z_{\omega,\ell-1}z_{\omega,\ell}}\\
=&\mathds{1}_{\{y_\omega=\ell\}}{\partial\over \partial z}\left({\dot{f}(z) \over f(z)-f(z')}\right)\Bigg|_{(z,\ z')=(z_{\omega,\ell},\ z_{\omega,\ell-1})}  -\mathds{1}_{\{y_\omega=\ell+1\}}{\partial\over \partial z}\left({\dot{f}(z) \over f(z)-f(z')}\right)\Bigg|_{(z'',\ z)=(z_{\omega,\ell+1},\ z_{\omega,\ell})}\\
\leq &-D_{\alpha,\beta,\Delta}<0.
\end{align}
\end{enumerate}
We will show that, with very high probability over $\tY$, $\mH$ is negative definite in that
\begin{equation}\label{eq:HessionZ}
\lambda_{1}(\mH)=\max_{\mz}{\mz^T\mH\mz\over\FnormSize{}{\mz}^2} \leq -c_2A_{\alpha,\beta,\Delta}D_{\alpha,\beta,\Delta},
\end{equation}
where $A_{\alpha,\beta,\Delta}, D_{\alpha,\beta,\Delta}>0$ are constants defined in~\eqref{eq:constantZ}, and $c_1>0$ is a constant.

Let $\mz_\omega=(z_{\omega,1},\ldots,z_{\omega,L-1})^T\in\mathbb{R}^{L-1}$ and $\mz=(\mz_{1,\ldots,1},\ldots,\mz_{d_1,\ldots,d_K})^T\in\mathbb{R}^{(L-1)\prod_k d_k}$. It follows from property (a) that
\[
\mz^T\mH\mz =\sum_{\omega}\mz_\omega^T \mH_{\omega} \mz_\omega.
\]
Furthermore, properties (b) and (c) (or similar arguments as in page 29, arXiv preprint) implies that
\begin{equation}
\mz_\omega^T \mH_{\omega} \mz_\omega \leq -D_{\alpha,\beta,\Delta} \sum_{\ell}z^2_{\omega,\ell}\KeepStyleUnderBrace{\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}}_{\text{Bernoulli r.v.\ with probability bounded by $A_{\alpha,\beta,\Delta}$}}.
\end{equation}
Therefore,
\begin{equation}\label{eq:sum}
\mz^T\mH\mz =\sum_{\omega}\mz_\omega^T \mH_{\omega} \mz_\omega \leq -D_{\alpha,\beta,\Delta}\sum_{\omega}\sum_{\ell} z^2_{\omega,\ell}\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}
\end{equation}
Based on central limit theorem, as the tensor dimension goes to infinity, 
\begin{equation}\label{eq:clt}
\sum_{\omega}\sum_{\ell} z^2_{\omega,\ell}\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}} \rightarrow\sum_{\omega}\sum_{\ell} z^2_{\omega,\ell}\mathbb{P}(y_\omega=\ell \text{ or }\ell+1)\geq c_2A_{\alpha,\beta,\Delta}\FnormSize{}{\mz}^2. 
\end{equation}
By~\eqref{eq:sum} and~\eqref{eq:clt}, we have
\begin{equation}\label{eq:boundH}
\mz^T\mH\mz \leq-c_2A_{\alpha,\beta,\Delta}D_{\alpha,\beta,\Delta}\FnormSize{}{\mz}^2,
\end{equation}
and therefore~\eqref{eq:HessionZ} is proved. Plugging~\eqref{eq:scoreZ} and~\eqref{eq:HessionZ} into~\eqref{eq:boundZ} yields
\[
\FnormSize{}{\hat \tZ-\trueZ}^2 \leq {c_1r^{K-1}_{\max}C^2_{\alpha,\beta,\Delta}\over A^2_{\alpha,\beta,\Delta}D^2_{\alpha,\beta,\Delta}}\left(L-1+\sum_k d_k\right).
\]
The MSEs for $\hat \Theta$ and $\hat \mb$ readily follow from~\eqref{eq:decomp}. 
\end{proof}

\bibliography{tensor_wang}
\bibliographystyle{apalike}
\end{document}

### 2020-03-02 paper/arxiv/correction.tex
upload correction

@@
 where $\mathbf{1}$ denotes a length-$(L-1)$ vector of all ones.
-Under the identifiability constraint $\langle \Theta, \tJ \rangle=0$, there is an one-to-one mapping between $\tZ$ and $(\Theta, \mb)$, with $\text{rank}(\tZ)=(\text{rank}(\Theta),2)^T$.
+Under the identifiability constraint $\langle \Theta, \tJ \rangle=0$, there is an one-to-one mapping between $\tZ$ and $(\Theta, \mb)$, with $\text{rank}(\tZ)\leq (\text{rank}(\Theta),2)^T$.
 Furthermore,
@@
 
-We define the following constants:
+We define the following constants that will be used in the theory:
 \begin{align}\label{eq:constantZ}
@@
 
+\begin{rmk} Add the specific bound for logistic link. 
+\end{rmk}
 \begin{thm}[Statistical convergence with unknown $\mb$]\label{thm:ratejoint}
@@
 \]
-Furthermore, properties (b) and (c) (or similar arguments as in page 29, arXiv preprint) implies that
+Furthermore, properties (b) and (c) (or similar arguments as in page 29, arXiv preprint) imply that
 \begin{equation}
@@
 \begin{equation}\label{eq:sum}
-\mz^T\mH\mz =\sum_{\omega}\mz_\omega^T \mH_{\omega} \mz_\omega \leq -D_{\alpha,\beta,\Delta}\sum_{\omega}\sum_{\ell} z^2_{\omega,\ell}\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}
+\mz^T\mH\mz =\sum_{\omega}\mz_\omega^T \mH_{\omega} \mz_\omega \leq -D_{\alpha,\beta,\Delta}\sum_{\omega}\sum_{\ell} z^2_{\omega,\ell}\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}.
 \end{equation}
@@
 \begin{equation}\label{eq:clt}
-\sum_{\omega}\sum_{\ell} z^2_{\omega,\ell}\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}} \rightarrow\sum_{\omega}\sum_{\ell} z^2_{\omega,\ell}\mathbb{P}(y_\omega=\ell \text{ or }\ell+1)\geq c_2A_{\alpha,\beta,\Delta}\FnormSize{}{\mz}^2.
+\sum_{\omega}\sum_{\ell} z^2_{\omega,\ell}\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}} \rightarrow\sum_{\omega}\sum_{\ell} z^2_{\omega,\ell}\mathbb{P}(y_\omega=\ell \text{ or }\ell+1)\geq c_2A_{\alpha,\beta,\Delta}\FnormSize{}{\mz}^2
 \end{equation}
+holds with very high probability. 
 By~\eqref{eq:sum} and~\eqref{eq:clt}, we have

### 2020-03-05 paper/arxiv/correction.tex
correction
update correction.pdf

@@
 \end{equation}
-Based on central limit theorem, as the tensor dimension goes to infinity, 
+
+{\color{red}
+Based on central limit theorem (and concentration properties of Bernoulli r.v.'s), as the tensor dimension goes to infinity, 
 \begin{equation}\label{eq:clt}
@@
 holds with very high probability. 
+}
+
 By~\eqref{eq:sum} and~\eqref{eq:clt}, we have
@@
 The MSEs for $\hat \Theta$ and $\hat \mb$ readily follow from~\eqref{eq:decomp}.
+
+{ \color{red}Correction of~\eqref{eq:clt}. Define the subspace:
+\[
+\tS=\{\Vec(\tZ): \tZ=-\Theta\otimes \mathbf{1}+\tJ\otimes \mb,\ (\Theta,\mb)\in(\tP,\tB)\}.
+\]
+We show that Hession is definite negative restricted in the subspace $\tS$. Specifically, for any vector $\mz=\entry{z_{\omega,\ell}}\in\tS$,
+\begin{align}
+\sum_{\omega,\ell} z^2_{\omega,\ell}\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}&=\sum_{\omega,\ell}(-\theta_\omega+b_\ell)^2\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}\\
+&=\sum_{\omega,\ell}(\theta^2_\omega-2\theta_\omega b_\ell + b^2_\ell)\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}\\
+&=\sum_{\omega,\ell}\theta^2_\omega\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}-2\sum_{\omega,\ell} \theta_\omega b_\ell \mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}+\sum_{\omega,\ell}b^2_\ell\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}\\
+&\geq \sum_{\omega}\theta^2_\omega -2\sum_{\omega,\ell}\theta_\omega b_\ell+\sum_{\ell}b_\ell^2\left(n_{\ell}+n_{\ell+1}\right)\\
+&=\sum_{\omega}\theta^2_\omega+\min_\ell\left(n_{\ell}+n_{\ell+1}\right)\sum_{\ell}b_\ell^2
+\end{align}
+
+On the other hand,
+\[
+\FnormSize{}{\mz}^2=\sum_{\omega,\ell}z^2_{\omega,\ell}=\sum_{\omega,\ell}(-\theta_\omega+b_\ell)^2=(L-1)\sum_{\omega}\theta^2_\omega+\left(\prod_k d_k\right)\sum_\ell b^2_\ell.
+\]
+
+Therefore, there exists a positive constant $c_1>0$ (? perhaps depending on $\alpha, \beta,\Delta$ etc...needs verification..) such that
+\[
+{{\mz^T\mH\mz}\over \FnormSize{}{\mz}^2}\geq c_1 {\min_{\ell}(n_\ell+n_{\ell_1})\over (L-1)\prod_k d_k}.
+\]
+}
+\end{enumerate}
 \end{proof}
-
 \bibliography{tensor_wang}

### 2020-03-05 paper/arxiv/correction.tex
Hession bound

--- 2020-03-02 paper/arxiv/correction.tex
@@
 The MSEs for $\hat \Theta$ and $\hat \mb$ readily follow from~\eqref{eq:decomp}.
+
+{ \color{red}Correction of~\eqref{eq:clt}. Define the subspace:
+\[
+\tS=\{\Vec(\tZ): \tZ=-\Theta\otimes \mathbf{1}+\tJ\otimes \mb,\ (\Theta,\mb)\in(\tP,\tB)\}.
+\]
+We show that Hession is definite negative restricted in the subspace $\tS$. Specifically, for any vector $\mz=\entry{z_{\omega,\ell}}\in\tS$,
+\begin{align}
+\sum_{\omega,\ell} z^2_{\omega,\ell}\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}&=\sum_{\omega,\ell}(-\theta_\omega+b_\ell)^2\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}\\
+&=\sum_{\omega,\ell}(\theta^2_\omega-2\theta_\omega b_\ell + b^2_\ell)\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}\\
+&=\sum_{\omega,\ell}\theta^2_\omega\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}-2\sum_{\omega,\ell} \theta_\omega b_\ell \mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}+\sum_{\omega,\ell}b^2_\ell\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}\\
+&\geq \sum_{\omega}\theta^2_\omega -2\sum_{\omega,\ell}\theta_\omega b_\ell+\sum_{\ell}b_\ell^2\left(n_{\ell}+n_{\ell+1}\right)\\
+&\geq \sum_{\omega}\theta^2_\omega+\min_\ell\left(n_{\ell}+n_{\ell+1}\right)\sum_{\ell}b_\ell^2
+\end{align}
+
+On the other hand,
+\[
+\FnormSize{}{\mz}^2=\sum_{\omega,\ell}z^2_{\omega,\ell}=\sum_{\omega,\ell}(-\theta_\omega+b_\ell)^2=(L-1)\sum_{\omega}\theta^2_\omega+\left(\prod_k d_k\right)\sum_\ell b^2_\ell.
+\]
+
+Therefore, there exists a positive constant $c_1>0$ (? perhaps depending on $\alpha, \beta,\Delta$ etc...needs some caculation..) such that
+\[
+\max_{\mz\in\tS,\mz\neq \mathbf{0}}{\sum_{\omega,\ell} z^2_{\omega,\ell}\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}} \over \FnormSize{}{\mz}^2}\geq c_1 {\min_{\ell}(n_\ell+n_{\ell+1})\over (L-1)\prod_k d_k}.
+\]
+The conclusion follows by noting that the ratio ${\min_{\ell}(n_\ell+n_{\ell+1})\over \prod_k d_k} \geq c'A_{\alpha,\beta,\Delta}$ in high probability. 
+}
+
+\end{enumerate}1
 \end{proof}
-
 \bibliography{tensor_wang}

### 2020-03-05 paper/arxiv/correction.tex
update correction

@@
 \]
-We show that Hession is definite negative restricted in the subspace $\tS$. Specifically, for any vector $\mz=\entry{z_{\omega,\ell}}\in\tS$,
+It suffices to prove the negative definiteness of Hession when restricted in the subspace $\tS$. Specifically, for any vector $\mz=\entry{z_{\omega,\ell}}\in\tS$,
 \begin{align}
@@
 
-Therefore, there exists a positive constant $c_1>0$ (? perhaps depending on $\alpha, \beta,\Delta$ etc...needs some caculation..) such that
+Therefore, there exists a positive constant $c_1>0$ (? perhaps depending on $\alpha, \beta,\Delta$ etc...needs some calculation..) such that
 \[
-\max_{\mz\in\tS,\mz\neq \mathbf{0}}{\sum_{\omega,\ell} z^2_{\omega,\ell}\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}} \over \FnormSize{}{\mz}^2}\geq c_1 {\min_{\ell}(n_\ell+n_{\ell+1})\over (L-1)\prod_k d_k}.
+\max_{\mz\in\tS,\mz\neq \mathbf{0}}{\sum_{\omega,\ell} z^2_{\omega,\ell}\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}} \over \FnormSize{}{\mz}^2}\geq c_1 {\min_{\ell}(n_\ell+n_{\ell+1})\over (L-1)\prod_k d_k}\geq c_1 {\min_{\ell}(n_\ell+n_{\ell+1})\over (L-1)\prod_k d_k}.
 \]
-The conclusion follows by noting that the ratio ${\min_{\ell}(n_\ell+n_{\ell+1})\over \prod_k d_k} \geq c'A_{\alpha,\beta,\Delta}$ in high probability. 
+The conclusion~\eqref{eq:clt} then follows by applying the law of large number and the uniform bound $\min_{z_{\omega,\ell}}\mathbb{P}(y_{\omega}=\ell \text{ or }\ell+1|z_{\omega,\ell})\geq 2A_{\alpha,\beta,\Delta}$ to the ratio:
+\[
+{\min_{\ell}(n_\ell+n_{\ell+1})\over \prod_k d_k}\geq 2A_{\alpha,\beta,\Delta}, \quad \text{ in high probability as $d_{\min}\to \infty$}.
+\] 
 }
-
 \end{enumerate}1

### 2020-03-05 paper/arxiv/correction.tex
upload correction.tex

@@
 \]
-The conclusion~\eqref{eq:clt} then follows by applying the law of large number and the uniform bound $\min_{z_{\omega,\ell}}\mathbb{P}(y_{\omega}=\ell \text{ or }\ell+1|z_{\omega,\ell})\geq 2A_{\alpha,\beta,\Delta}$ to the ratio:
+The conclusion~\eqref{eq:clt} then follows by applying the law of large numbers and the uniform bound $\min_{z_{\omega,\ell}}\mathbb{P}(y_{\omega}=\ell \text{ or }\ell+1|z_{\omega,\ell})\geq 2A_{\alpha,\beta,\Delta}$ to the ratio:
 \[

### 2020-04-11 paper/ICML/rebuttal.txt
Create rebuttal.txt

Reviewer #5
Q1. Distinction between classical tensors and ordinal tensors 
( because we are using Tucker decomposition?, He or she wants clear distinction and challenges with regards to ordinal tensor, is it special case of classical analysis)

A1. Our ordinal tensor analysis is conceptually different from continuous tensor analysis method. We make use of quantization to guarantee categorical responses. This quantization model ensure that the estimation is invariant under a reversal of categories and the parameter interpretations are consistent, which is not the case in classical tensor analysis. Our method is not a special case of classical analysis, but it is one step more from classical analysis. In addition, this latent model helps us to interpret the model with signal and noise concept. With this concept, the model has the phenomenon that estimation becomes harder in high signal level in Figure 2-b which is distinctive from classical analysis.
(Also, it is not exponential family)

Q2. The tools adopted is from classical tensor analysis (rank constrained M estimation) -This also can be related to reviewer #3’s question,

A2. I agree that rank constrained M estimation is common in previous works. In our paper, however, we established the theoretical results guaranteeing optimal convergence and complexity of M estimation. Those theories make our paper distinctive from others and justify using constrained M estimation.

Reviewer #4
Q1. The initialization of “b” is quite important to obtain good local optima. Do you have any good strategy to initialize b?(Initialization problem, may be the reviewer concerns identifiability issue too.)

A1.First, we perform multiple initialization methods to increase the probability of getting good initialization. Actual value of intercept b cannot be decided by itself but with consideration of signal tensor and response variables, which means if signal tensor shifted left then b will follow. Identifiability issue can arise from this point. So we set signal tensor sum to zero for each update and update intercept b to solve identifiability issue. In this way, we initialize intercept b to be fitted on the initialization of signal tensor which sum to zero.  This initialization strategy for b combining with multiple initialization makes algorithm stable and start in good position to obtain good local optima.

Q2. Computational cost for the optimization.
MovieLens: dataset represents rating tuples of the form: <user, item, rating, timestamp>. There are diverse datasets such as 
100k:100,000 ratings from 1000 users on 1700 movies
1M:1 million ratings from 6000 users on 4000 movies
10M:10 million ratings and 100,000 tag applications applied to 10,000 movies by 72,000 users
20M:20 million ratings and 465,000 tag applications applied to 27,000 movies by 138,000 users
https://papers.nips.cc/paper/8860-singleshot-a-scalable-tucker-tensor-decomposition.pdf (Traore, A., Bérar, M., & Rakotomamonjy, A. (2019). Singleshot : a scalable Tucker tensor decomposition. NeurIPS.)
https://arxiv.org/pdf/1710.02261v1.pdf(Oh, S., Park, N., Sael, L., & Kang, U. (2018). Scalable Tucker Factorization for Sparse Tensors - Algorithms and Discoveries. 2018 IEEE 34th International Conference on Data Engineering (ICDE), 1120-1131.)
: uses data where 6040 users, 100209 ratings from 1 to 6 on 3076 movies , between April,2000 and February,2003
(It assumes sparse tensor which means rank is really low)

A2. For bigger dataset, like Movielens dataset, we can use small rank to reduce computations like other papers have done for the application on Movielens such as
(Traore, A., Bérar, M., & Rakotomamonjy, A. (2019). Singleshot : a scalable Tucker tensor decomposition. NeurIPS,
Oh, S., Park, N., Sael, L., & Kang, U. (2018). Scalable Tucker Factorization for Sparse Tensors - Algorithms and Discoveries. 2018 IEEE 34th International Conference on Data Engineering (ICDE), 1120-1131.) Here they used Tucker rank (5,5,5) and (10,10,10). This small rank makes implementation much faster (our brain dataset rank was (23,23,8) which is not quite small)
There are several rooms for us to improve our current algorithm for big data applications. Movielens datasets have more than 100000 ratings which makes computation of gradient challenging. However our loss function has good structure to apply stochastic gradient descent method. SGD is one way to improve our computation time. In addition, we can select a rowdies update method for factor matrices updates. The key benefit of this approach is that all rows of factor matrices are independent of each other in the sense of minimizing our loss function. This property enables us to apply parallel computing on updating factor matrices. Those approaches enable us to apply big dataset like MovieLens.

Reviewer #3
Q1. Can not find tractable algorithms for MLE ( He or she concerns about the gap between theoretical result and real application)

A1.
It is fair argument. However, no other approaches succeed to have tractable algorithm  to estimate global optimizer. The other methods include the nuclear norm approaches sacrificing rank constraints but still having computation hard algorithm. Also, our paper is meaningful in the sense that we show that this straightforward method works better than other approaches even with relaxed local minimum condition. Our paper shows theoretical reason for this outpeformance: MLE has optimal convergence rate and optimal complexity which makes even local minimum estimation outperforms the other methods with suboptimal rate. This paper provides one step closer to ultimate goal of tensor completion method for ordinal dataset. Another fact is multiple initialization can handle local minimization problem too. In addition, It is known that with multiple random restarts and deflation is capable of finding all components ui with high probability.

(I am taking a look at articles that study about spurious local minima but can not find good articles to apply in our method yet. I will do work more on that)

### 2020-04-14 paper/ICML/rebuttal.txt
Update rebuttal.txt

@@
 A1.First, we perform multiple initialization methods to increase the probability of getting good initialization.
-Actual value of intercept b cannot be decided by itself but with consideration of signal tensor and response variables, which means if signal tensor shifted left then b will follow.
+Actual value of threshold parameter b cannot be decided by itself but with consideration of signal tensor and response variables, which means if signal tensor shifted left then b will follow (consider P(y\geq l) = f(b_l-theta)).
 Identifiability issue can arise from this point.
-So we set signal tensor sum to zero for each update and update intercept b to solve identifiability issue.
-In this way, we initialize intercept b to be fitted on the initialization of signal tensor which sum to zero.
+So we set signal tensor sum to zero to solve the identifiability issue.
+Since we have mean 0 signal tensor, it is natiural for us to initialize the threshold parameter b is to set an increasing sequence such that the cumulative density of logisic distribution between consecutative threshold is the same.
 This initialization strategy for b combining with multiple initialization makes algorithm stable and start in good position to obtain good local optima.

### 2020-04-14 paper/ICML/rebuttal.txt
Update rebuttal.txt

@@
 So we set signal tensor sum to zero to solve the identifiability issue.
-Since we have mean 0 signal tensor, it is natiural for us to initialize the threshold parameter b is to set an increasing sequence such that the cumulative density of logisic distribution between consecutative threshold is the same.
+Since we have mean 0 signal tensor, it is natiural for us to initialize the threshold parameter b as an increasing sequence such that the cumulative density of the logistic distribution between consecutative threshold is the same (or an increasing seqeunece such that the cumulative density of the logistic distribution between consecutative threshold reflects the proportion of ordinal values in the data).
 This initialization strategy for b combining with multiple initialization makes algorithm stable and start in good position to obtain good local optima.

### 2020-04-14 paper/ICML/rebuttal.txt
Update rebuttal.txt

@@
 SGD is one way to improve our computation time.
-In addition, we can select a rowdies update method for factor matrices updates.
-The key benefit of this approach is that all rows of factor matrices are independent of each other in the sense of minimizing our loss function.
-This property enables us to apply parallel computing on updating factor matrices.
-Those approaches enable us to apply big dataset like MovieLens.
+In addition, we can select a row-wise update method for factor matrices.
+In our alternating gradient based algorithm, we can show that all rows of factor matrices are independent of each other in the sense of minimizing our loss function.
+This property enables us to use parallel computing which significantly reduces memory requirement for updating factor matrices.
+I think those approaches enable us to apply big dataset like MovieLens.
 

### 2020-04-14 paper/ICML/rebuttal.txt
Update rebuttal.txt

@@
 This paper provides one step closer to ultimate goal of tensor completion method for ordinal dataset.
-Another fact is multiple initialization can handle local minimization problem too.
-In addition, It is known that with multiple random restarts and deflation is capable of finding all components ui with high probability.
 
-(I am taking a look at articles that study about spurious local minima but can not find good articles to apply in our method yet. I will do work more on that)
+In the matrix case of our optimization problem, many papers have shown that one would expect to achieve global convergence for the rank constrained optimization problem provided algorithms converge to a local minimum.
+(R. Salakhutdinov and A. Mnih. Probabilistic matrix factorization. In Advances in Neural Information Processing Systems, 2008.,
+T. Cai and W.-X. Zhou. A Max-Norm Constrained Minimization Approach to 1-Bit Matrix Completion. Journal of Machine Learning Research, 14:3619–3647, 2013.,
+Ge, R., Jin, C., & Zheng, Y.
+(2017, August).
+No spurious local minima in nonconvex low rank problems: A unified geometric analysis.
+In Proceedings of the 34th International Conference on Machine Learning-Volume 70 (pp.
+1233-1242).
+JMLR.
+org.)
+
+In the line of those work, many studies try to show there is no spurious local minima in our tensor optimization problem.
+(Haeffele, B. D., & Vidal, R. (2015). Global optimality in tensor factorization, deep learning, and beyond. arXiv preprint arXiv:1506.07540.,
+Ge, R., & Ma, T. (2017). On the optimization landscape of tensor decompositions. In Advances in Neural Information Processing Systems (pp. 3653-3663).,
+Sanjabi, M., Baharlouei, S., Razaviyayn, M., & Lee, J. D. (2019). When Does Non-Orthogonal Tensor Decomposition Have No Spurious Local Minima?. arXiv preprint arXiv:1911.09815.
+). 
+Even though the optimization problems in those papers are the special case of our problem (for example assuming that factor matrices are identical or adding extra regularization term), they are giving us positive conjecture that our algorithm may succeed finding global minima.

### 2020-04-14 paper/ICML/rebuttal.txt
Update rebuttal.txt

@@
 It is fair argument.
-However, no other approaches succeed to have tractable algorithm  to estimate global optimizer.
-The other methods include the nuclear norm approaches sacrificing rank constraints but still having computation hard algorithm.
-Also, our paper is meaningful in the sense that we show that this straightforward method works better than other approaches even with relaxed local minimum condition.
-Our paper shows theoretical reason for this outpeformance: MLE has optimal convergence rate and optimal complexity which makes even local minimum estimation outperforms the other methods with suboptimal rate.
-This paper provides one step closer to ultimate goal of tensor completion method for ordinal dataset.
+However, I want to emphasize that our theorems are not only applicable to the global optimal MLE but any local optimums suffice as long as it has smaller loss function than at true parameter.
+This implies that the local optimums to the MLE have the optimal rate in the condition.
+Therefore, I would say that the optimality of the MLE is still useful.
+In the same line,  we argue that the finding a local minimum is suffice not only with experiments but also with theoretical analysis in our proof.
+The experiments are coincide with the theorems in the sense that our approach with possible local minimas outperforms other approaches (optimal vs suboptimal).
 
-In the matrix case of our optimization problem, many papers have shown that one would expect to achieve global convergence for the rank constrained optimization problem provided algorithms converge to a local minimum.
+As you mentioned, the sufficiency of the finding local optima of nonconvex problems is very active area.
+Most of works are focusing on showing that local optima is global optima.
+Those approaches are different from our approach but wil provide another good point of our estimation method.
+
+In the matrix case, many papers have shown that one would expect to achieve global convergence for the rank constrained optimization problem provided algorithms converge to a local minimum.
 (R. Salakhutdinov and A. Mnih. Probabilistic matrix factorization. In Advances in Neural Information Processing Systems, 2008.,
@@
 
-In the line of those work, many studies try to show there is no spurious local minima in our tensor optimization problem.
+In the tensor case, many studies try to show there is no spurious local minima for the rank constrained optimization problem.
 (Haeffele, B. D., & Vidal, R. (2015). Global optimality in tensor factorization, deep learning, and beyond. arXiv preprint arXiv:1506.07540.,

### 2020-04-19 paper/ICML/rebuttal0419.txt
update rebuttle

We thank the reviewers for helpful comments and feedback. Our responses are detailed below.

To Reviewer 3:
* Regarding theoretical contributions and MLE approach

We highlight two main contributions that set our work apart from current literature. 
1) [New challenges with non-classical observations] The problem of ordinal/quantized tensor is fundamentally more challenging than previously well-studied tensors for two reasons: (a) the entries do not belong to exponential family distribution, and (b) the observation contains much less information, because neither the underlying signal nor the quantization operator is known. The limited information makes the statistical recovery notably hard, not to mention the associated computation.

We are among the first to establish the recovery theory for both the signal tensors and the quantization operators from a limited number of highly discrete entries. Our model achieves invariance under a reversal of categories and consistency under merging/splitting of contiguous categories. A distinctive non-monotonic phase-transition pattern is demonstrated (Figure 2b). None of these features are ever considered in classical tensor analysis. Furthermore, we prove that, surprisingly, recovery from quantized tensors achieves equally good information-theoretical convergence as the continuous tensors. These results fill the gap between classical and non-classical (ordinal) observations, thereby greatly enriching the tensor-model literature.

2) [New challenges with regularized MLE] The properties of regularized MLE are challenging in our tensor model. Current approaches for solving similar problems fall into two categories: convex (e.g., nuclear-norm regularized) MLE and non-convex MLE (our approach). Most work focuses on the former. However, in the tensor case, convex MLE suffers from both computational in-traceability and statistical sub-optimality. This is because computing nuclear norm is NP-hard for tensors.

We advocate the non-convex approach and provide strong evidence for its success in our setting. We improve the statistical error bound from d^{-(K-1)/2} to d^{-(K-1)}, which is substantial as the tensor order K increases. This solves the earlier open question whether the square root can be removed in an unlimited computational setting [Ghadermarzy et al 2019]. As a by-effect, we demonstrate the statistical optimality of the non-convex approach over the (computationally intractably) convex MLE approach. While the computational limit remains open (as in many modern ML problems), we believe our improved statistical bound provides a useful benchmark for future algorithm development.

* Regarding the optimization

We agree with the reviewer that MLE is computationally hard. While our main contributions lie in the aforementioned two aspects, we do provide useful results on optimization. We show that the alternating algorithm convergences to local optimums, and we prove (in Thm 4.1 and Supplements) that the local optimums with reasonably high objective suffice for accurate recovery. In practice, we find that the estimation is robust under randomized initialization, which resembles the `benign non-convexity’ phenomenon in literature. In special cases such as rank-1 or orthogonally decomposable tensors, we are able to leverage recent advances on optimization to obtain the asymptotical number of initializations for global guarantee, however, at the cost of more stringent assumptions on the underlying signal tensor. For this reason, we choose not to pursue this direction but leave framework general for the current work.

Tensor learning is a clear challenge for further research. We believe our work provides a (somewhat surprising) new idea for quantized tensor problem, and we obtain results that were previously difficult to establish. We hope the work open up new inquiry that allows more machine learning researchers to contribute to this field.

To Reviewer 4:
* Initialization of b
We initialize b such that the each of the estimated categories has roughly equal size. Furthermore, multiple initializations are utilized and we choose the one that maximizes the objective. We find that such initialization strategy gives stable results in the simulations and the two applications considered. 
* Computational costs
We choose these two datasets for fair comparison because they were used in the previous work. We agree that applying to large datasets such as MovieLens would be interesting. In fact, our alternating approach lends itself well to this context. The sub-step for solving tensor factors could be paralleled and be boosted by fast algorithms such as stochastic gradient descent. We would be glad to include this application, space permitting.

To Reviewer 5:
* Difference to classical tensors
See our response to reviewer 3 regarding `New challenges with non-classical observations.'
* rank-constrained MLE
See our response to reviewer 3 regarding `New challenges with regularizeMLE.'

### 2020-04-19 paper/ICML/rebuttal0419.txt
update

@@
 We choose these two datasets for fair comparison because they were used in the previous work.
-We agree that applying to large datasets such as MovieLens would be interesting.
+We agree that application to large datasets such as MovieLens would be interesting.
 In fact, our alternating approach lends itself well to this context.
@@
 * Difference to classical tensors
-See our response to reviewer 3 regarding `New challenges with non-classical observations.'
+See our response to reviewer 3 regarding `New challenges with non-classical observations'
 * rank-constrained MLE
-See our response to reviewer 3 regarding `New challenges with regularizeMLE.'
+See our response to reviewer 3 regarding `New challenges with regularizeMLE'

### 2020-04-19 paper/ICML/rebuttal0419.txt
update

@@
 We show that the alternating algorithm convergences to local optimums, and we prove (in Thm 4.1 and Supplements) that the local optimums with reasonably high objective suffice for accurate recovery.
-In practice, we find that the estimation is robust under randomized initialization, which resembles the `benign non-convexity’ phenomenon in literature.
+In practice, we find that the estimation is robust under randomized initialization, which resembles the `benign non-convex’ phenomenon in literature.
 In special cases such as rank-1 or orthogonally decomposable tensors, we are able to leverage recent advances on optimization to obtain the asymptotical number of initializations for global guarantee, however, at the cost of more stringent assumptions on the underlying signal tensor.
@@
 * rank-constrained MLE
-See our response to reviewer 3 regarding `New challenges with regularizeMLE'
+See our response to reviewer 3 regarding `New challenges with regularized MLE'

### 2020-04-19 paper/ICML/rebuttal0419.txt
update

@@
 We find that such initialization strategy gives stable results in the simulations and the two applications considered.
+
 * Computational costs
@@
 See our response to reviewer 3 regarding `New challenges with non-classical observations'
+
 * rank-constrained MLE

### 2020-04-19 paper/ICML/rebuttal0419.txt
Update
Update rebuttal0419.txt

@@
 Most work focuses on the former.
-However, in the tensor case, convex MLE suffers from both computational in-traceability and statistical sub-optimality.
+However, in the tensor case, convex MLE suffers from both computational in-tractability and statistical sub-optimality.
 This is because computing nuclear norm is NP-hard for tensors.
@@
 * Initialization of b
-We initialize b such that the each of the estimated categories has roughly equal size.
+We initialize b such that each of the estimated categories has roughly equal size.
 Furthermore, multiple initializations are utilized and we choose the one that maximizes the objective.

### 2020-04-19 paper/ICML/rebuttal0419.txt
Update rebuttal0419.txt
update

@@
 We improve the statistical error bound from d^{-(K-1)/2} to d^{-(K-1)}, which is substantial as the tensor order K increases.
-This solves the earlier open question whether the square root can be removed in an unlimited computational setting [Ghadermarzy et al 2019].
-As a by-effect, we demonstrate the statistical optimality of the non-convex approach over the (computationally intractably) convex MLE approach.
+This solves the earlier open question whether the square root can be removed in an unlimited computational setting [Ghadermarzy et al 2018].
+As a by-effect, we demonstrate the statistical optimality of the non-convex approach over the (computationally intractable) convex MLE approach.
 While the computational limit remains open (as in many modern ML problems), we believe our improved statistical bound provides a useful benchmark for future algorithm development.
@@
 While our main contributions lie in the aforementioned two aspects, we do provide useful results on optimization.
-We show that the alternating algorithm convergences to local optimums, and we prove (in Thm 4.1 and Supplements) that the local optimums with reasonably high objective suffice for accurate recovery.
+We show that the alternating algorithm converges to stationary points, and we prove (in Thm 4.1 and Supplements) that the local optimums with reasonably high objective suffice for accurate recovery.
 In practice, we find that the estimation is robust under randomized initialization, which resembles the `benign non-convex’ phenomenon in literature.
@@
 We believe our work provides a (somewhat surprising) new idea for quantized tensor problem, and we obtain results that were previously difficult to establish.
-We hope the work open up new inquiry that allows more machine learning researchers to contribute to this field.
+We hope the work opens up new inquiry that allows more machine learning researchers to contribute to this field.
 
+—
 To Reviewer 4:
@@
 
+—
 To Reviewer 5:

### 2020-05-30 paper/NeurIPS/ordinalT.tex
Add files via upload

--- 2020-02-18 ICML/submit/supplements/latexsource/ordinalTv3.tex
@@
-%%%%%%%% ICML 2020 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
-
 \documentclass{article}
 
-% Recommended, but optional, packages for figures and better typesetting:
-\usepackage{microtype}
+% if you need to pass options to natbib, use, e.g.:
+  %  \PassOptionsToPackage{numbers, compress}{natbib}
+% before loading neurips_2020
+
+% ready for submission
+\usepackage{neurips_2020}
  \usepackage{graphicx}
-\usepackage{subfigure}
-\usepackage{booktabs} % for professional tables
+ \usepackage{bm}
 
-% hyperref makes hyperlinks in the resulting PDF.
-% If your build breaks (sometimes temporarily if a hyperlink spans a page)
-% please comment out the following usepackage line and replace
-% \usepackage{icml2020} with \usepackage[nohyperref]{icml2020} above.
-\usepackage{hyperref}
+% to compile a preprint version, e.g., for submission to arXiv, add add the
+% [preprint] option:
+%     \usepackage[preprint]{neurips_2020}
 
-% Attempt to make hyperref and algorithmic work together better:
-\newcommand{\theHalgorithm}{\arabic{algorithm}}
+% to compile a camera-ready version, add the [final] option, e.g.:
+%     \usepackage[final]{neurips_2020}
 
-% Use the following line for the initial blind version submitted for review:
-\usepackage{icml2020}
-
-% If accepted, instead use the following line for the camera-ready submission:
-%\usepackage[accepted]{icml2020}
-
-% The \icmltitle you define below is probably too long as a header.
-% Therefore, a short form for the running title is supplied here:
-\icmltitlerunning{Ordinal tensor denoising and completion}
-\usepackage{wrapfig}
-\usepackage{multirow}
-\usepackage{graphicx}
-%\usepackage[utf8]{inputenc} % allow utf-8 input
-%\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
+% to avoid loading the natbib package, add option nonatbib:
+     %\usepackage[nonatbib]{neurips_2020}
+\usepackage{amsmath,amssymb}
+\usepackage{amsthm}
+\usepackage[utf8]{inputenc} % allow utf-8 input
+\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
 \usepackage{hyperref}       % hyperlinks
 \usepackage{url}            % simple URL typesetting
-%\usepackage{booktabs}       % professional-quality tables
-\usepackage{amsmath,amssymb}
-\usepackage{amsthm}    % blackboard math symbols
-%\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-%\usepackage{microtype}      % microtypography
-\usepackage{bm}
-%\usepackage{subfig}
-%\usepackage[english]{babel}
-%\usepackage{algorithm}
-%\usepackage{appendix}
+\usepackage{booktabs}       % professional-quality tables
+\usepackage{amsfonts}       % blackboard math symbols
+\usepackage{nicefrac}       % compact symbols for 1/2, etc.
+\usepackage{microtype}      % microtypography
+
 \usepackage{mathtools}
@@
 \newtheorem{assumption}{Assumption}
-
 \theoremstyle{definition}
@@
 \newtheorem{rmk}{Remark}
-
-\usepackage{dsfont}
-%\usepackage{algpseudocode,algorithm}
-%\algnewcommand\algorithmicinput{\textbf{Input:}}
-%\algnewcommand\algorithmicoutput{\textbf{Output:}}
-%\algnewcommand\INPUT{\item[\algorithmicinput]}
-%\algnewcommand\OUTPUT{\item[\algorithmicoutput]}
-%\DeclareMathOperator*{\minimize}{minimize}
-
 \newcommand*{\KeepStyleUnderBrace}[1]{%f
@@
 }
+\usepackage{dsfont}
+\newcommand{\theHalgorithm}{\arabic{algorithm}}
+\usepackage{algorithm}
+\usepackage{algorithmic}
 \usepackage{makecell}
+
+\title{Tensor denoising and completion based on ordinal observations}
+
+% The \author macro works with any number of authors. There are two commands
+% used to separate the names and addresses of multiple authors: \And and \AND.
+%
+% Using \And between authors leaves it to LaTeX to determine where to break the
+% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
+% authors names on the first line, and the last on the second line, try using
+% \AND instead of \And before the third author name.
+
+\author{
+ Chanwoo Lee \\
+  Department of Statistics\\
+  University of Wisconsin - Madison\\
+  \texttt{chanwoo.lee@wisc.edu} \\
+  % examples of more authors
+   \And
+  Miaoyan Wang \\
+  Department of Statistics\\
+  University of Wisconsin - Madison\\
+  \texttt{miaoyan.wang@wisc.edu} \\
+  % \AND
+  % Coauthor \\
+  % Affiliation \\
+  % Address \\
+  % \texttt{email} \\
+  % \And
+  % Coauthor \\
+  % Affiliation \\
+  % Address \\
+  % \texttt{email} \\
+  % \And
+  % Coauthor \\
+  % Affiliation \\
+  % Address \\
+  % \texttt{email} \\
+}
 \input macros.tex
 
-\usepackage{amssymb}
-\usepackage{pifont}
-\newcommand{\cmark}{\ding{51}}%
-\newcommand{\xmark}{\ding{55}}%
-
 \begin{document}
-
-\twocolumn[
-\icmltitle{Tensor denoising and completion based on ordinal observations}
-
-% It is OKAY to include author information, even for blind
-% submissions: the style file will automatically remove it for you
-% unless you've provided the [accepted] option to the icml2020
-% package.
-
-% List of affiliations: The first argument should be a (short)
-% identifier you will use later to specify author affiliations
-% Academic affiliations should list Department, University, City, Region, Country
-% Industry affiliations should list Company, City, Region, Country
-
-% You can specify symbols, otherwise they are numbered in order.
-% Ideally, you should not use this facility. Affiliations will be numbered
-% in order of appearance and this is the preferred way.
-\icmlsetsymbol{equal}{*}
-
-\begin{icmlauthorlist}
-\icmlauthor{}{}
-\end{icmlauthorlist}
-
-\icmlaffiliation{}{}
-\icmlcorrespondingauthor{}{}
-% You may provide any keywords that you
-% find helpful for describing your paper; these are used to populate
-% the "keywords" metadata in the PDF but will not be shown in the document
-\icmlkeywords{Higher-order tensors, ordinal observation, tensor decomposition, tensor completion}
-
-\vskip 0.3in
-]
-
-% this must go after the closing bracket ] following \twocolumn[ ...
-
-% This command actually creates the footnote in the first column
-% listing the affiliations and the copyright notice.
-% The command takes one argument, which is text to display at the start of the footnote.
-% The \icmlEqualContribution command is standard text for equal contribution.
-% Remove it (just {}) if you do not need this facility.
-
-%\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
-\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.
-
+\maketitle
 \begin{abstract}
@@
 We demonstrate the outperformance of our approach over previous methods on the tasks of clustering and collaborative filtering.
-
 \end{abstract}
-
+\vspace{-.3cm}
 \section{Introduction}
+\vspace{-.2cm}
 Multidimensional arrays, a.k.a.\ tensors, arise in a variety of applications including recommendation systems~\citep{baltrunas2011incarmusic}, social networks~\citep{nickel2011three}, genomics~\citep{hore2016tensor}, and neuroimaging~\citep{zhou2013tensor}.
-There is a growing need to develop general methods for analyzing these noisy, high-dimensional datasets that can handle two main problems.
-The problem of tensor denoising -- which aims to recover a signal tensor from its noisy entries -- has gained increased attention in theory and applications~\citep{xia2019sup,zeng2019multiway}.
-The second problem, tensor completion, examines the minimum number of entries needed for a consistent recovery~\citep{ghadermarzy2018learning,ghadermarzy2019near}.
-Low-rankness is often imposed to the signal tensor, Thereby efficiently reducing the intrinsic dimension in both problems.
+Two main problems have gained increased attention for analyzing those noisy, high-dimensional datasets: tensor denoising, tensor completion.
+Tensor denoising aims to recover a signal tensor from its noisy entrie~\citep{xia2019sup,zeng2019multiway}.
+Tensor completion examines the minimum number of entries needed for a consistent recovery~\citep{ghadermarzy2018learning,ghadermarzy2019near}.
+Low-rankness is often imposed to the signal tensor,
+thereby efficiently reducing the intrinsic dimension in both problems.
 
@@
 Another example is in the signal processing, where the digits are frequently rounded or truncated so that only integer values are available.
-The qualitative observations take values in a limited set of categories, making the learning problem harder compared to  continuous observations.
+Those qualitative observations take values in a limited set of categories, making the learning problem harder compared to  continuous observations.
 
@@
 The analyses of tensors with the ordinal entries are mainly complicated by two key properties needed for a reasonable model.
-First, the model should be invariant under a reversal of categories, say, from the Netflix example, {\it very like} $\succ$ {\it like} $\succ$ {\it neutral} $\succ \cdots$, but not under arbitrary label permutations.
+First, the model should be invariant under a reversal of categories, say, {\it very like} $\succ$ {\it like} $\succ$ {\it neutral} $\succ \cdots$, but not under arbitrary label permutations.
 Second, the parameter interpretations should be consistent under merging or splitting of contiguous categories.
@@
 \begin{table*}[ht]
-\centerline{\includegraphics[width=16cm]{compare.pdf}}
+\centerline{\includegraphics[width=13cm]{compare.pdf}}
 \vspace{-.3cm}
@@
 We propose a cumulative link model for higher-order tensors, develop a rank-constrained M-estimator, and obtain theoretical accuracy guarantees.
-The mean squared error bound is established, and we show that the obtained bound has minimax optimal rate in high dimensions under the low-rank model.
+We successfully recover true signal where neither the underlying signal nor the quantization operator is known.
+We show that our mean squared error bound has minimax optimal rate under the low-rank model.
 Our estimator enjoys a faster convergence rate $\tO(d^{-(K-1)/2})$ than $\tO(d^{-K})$ in~\citet{ghadermarzy2018learning}, which is a substantial improvement as the order $K$ increases.
-Furthermore, our proposal serves as an efficient completion algorithm that guarantees consistent recovery of an order-$K$ $(d,\ldots,d)$-dimensional low-rank tensor using only $\tilde \tO(Kd)$ noisy, quantized observations.
+Proposed tensor completion algorithm guarantees consistent recovery of an order-$K$ $(d,\ldots,d)$-dimensional low-rank tensor using only $\tilde \tO(Kd)$ noisy, quantized observations.
 
@@
 As we mentioned earlier, the ordinal observations add considerable challenges to the model formulation.
-We address the problems from two perspectives.
+We address the challenges from two perspectives.
 From statistical perspective, our proposed model generalizes the usual binary tensor model while preserving palindromic invariance~\cite{mccullagh1980regression} for ordinal observations.
-From algorithm perspective, our alternating optimization compares favorably to the approximate (non-convex) algorithm developed in the context of binary tensors~\cite{ghadermarzy2018learning}.
+From algorithm perspective, our alternating optimization compares favorably to the approximated (convex) algorithm developed in the context of binary tensors~\cite{ghadermarzy2018learning}.
 We numerically compare the two approaches in Section~\ref{sec:experiment}.
 
-\vspace{-.2cm}
+\vspace{-.3cm}
 \section{Preliminaries}
-\vspace{-.1cm}
+\vspace{-.3cm}
 Let $\tY\in\mathbb{R}^{d_1\times \cdots \times d_K}$ denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional tensor.
@@
 We say that an event $A$ occurs ``with very high probability'' if $\mathbb{P}(A)$ tends to 1 faster than any polynomial of tensor dimension $d_{\min}=\min\{d_1,\ldots,d_K\} \to\infty$.
-
 We use lower-case letters ($a, b, c, \ldots$) for scalars/vectors, upper-case boldface letters ($\mA, \mB, \mC, \ldots$) for matrices, and calligraphy letters ($\tA,\tB,\tC,\ldots$) for tensors of order three or greater.
@@
 
+\vspace{-.3cm}
+\section{Model formulation and motivation}
 \vspace{-.2cm}
-\section{Model formulation and motivation}
-\vspace{-.1cm}
 \subsection{Observation model}
-\vspace{-.1cm}
+\vspace{-.2cm}
 Let $\tY$ denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional data tensor.
-Suppose the entries of $\tY$ are ordinal-valued, and the observation space consists of $L$ ordered levels, denoted by $[L]:=\{1,\ldots,L\}$.
+Suppose the entries of $\tY$ are ordinal-valued, and the observation space is denoted by $[L]:=\{1,\ldots,L\}$.
 We propose a cumulative link model for the ordinal tensor $\tY=\entry{y_\omega}\in[L]^{d_1\times \cdots\times d_K}$.
@@
 \end{equation}
-where $\mb=(b_1,\ldots,b_{L-1})$ is a set of unknown scalars satisfying $b_1<\cdots <b_{L-1}$, $\Theta=\entry{\theta_\omega}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is a continuous-valued parameter tensor satisfying certain low-dimensional structure (to be specified later), and $f(\cdot):\mathbb{R}\mapsto[0,1]$ is a known, strictly increasing function.
+where $\mb=(b_1,\ldots,b_{L-1})$ is a set of unknown scalars satisfying $b_1<\cdots <b_{L-1}$, $\Theta=\entry{\theta_\omega}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is a continuous-valued parameter tensor satisfying certain low-rank structure (to be specified later), and $f(\cdot):\mathbb{R}\mapsto[0,1]$ is a known, strictly increasing function.
 We refer to $\mb$ as the cut-off points and $f$ the link function.
@@
 \end{equation}
-for all tensor entries $y_\omega$.
+for all $y_\omega$.
 When there are only two categories in the observation space (e.g.
-binary tensors), the cumulative model~\eqref{eq:model} is equivalent to the usual multinomial link model.
-In general, however, when the number of categories $L\geq 3$, the proportional odds assumption~\eqref{eq:logodd} is more parsimonious, in that, the ordered categories can be envisaged as contiguous intervals on the continuous scale, where the points of division are exactly $b_1<\cdots <b_{L-1}$.
-This interpretation will be made more explicit in the next section.
-\vspace{-.1cm}
+binary tensors), the cumulative model~\eqref{eq:model} is equivalent to the usual binomial link model.
+In general, however, when the number of categories $L\geq 3$, the proportional odds assumption~\eqref{eq:logodd} is more parsimonious.
+The ordered categories can be envisaged as contiguous intervals on the continuous scale, where the points of division are exactly $b_1<\cdots <b_{L-1}$.
+This interpretation will be made explicit in the next section.
+\vspace{-.2cm}
 \subsection{Latent-variable interpretation}\label{sec:latent}
-\vspace{-.1cm}
+\vspace{-.2cm}
 
@@
 \begin{align}\label{eq:quantization}
-y_\omega&=
-\begin{cases}
-1,& \text{if $y^*_\omega\in(-\infty, b_1]$},\\
-2,& \text{if $y^*_\omega\in(b_1, b_2]$},\\
-\vdots  &\vdots\\
-L,& \text{if $y^*_\omega\in(b_{L-1}, \infty)$},\\
-\end{cases}
+y_\omega&=\sum_{\ell = 1}^L \ell\mathds{1}_{y^*_\omega \in (b_{\ell-1},b_\ell]},
 \end{align}
-for all $\omega\in[d_1]\times \cdots \times [d_k]$. Here, $\tY^*=\entry{y^*_\omega}$ is a latent continuous-valued tensor following an additive noise model:
+for all $\omega\in[d_1]\times \cdots \times [d_k]$.
+We define $b_0 = -\infty$ and $b_L = \infty$.
+Here, $\tY^*=\entry{y^*_\omega}$ is a latent continuous-valued tensor following an additive noise model:
 \begin{align}\label{eq:latent}
@@
 \end{example}
+\vspace{-.2cm}
 Other link functions are also possible, such as Laplace, Cauchy, etc~\cite{mccullagh1980regression}.
 All the models share the property that the ordered categories can be thought of as contiguous interval on some continuous scale.
-We should point out that, although the latent-variable interpretation is incisive, our estimation procedure does not refer to the existence of $\tY^*$.
+We point out that, although the latent-variable interpretation is incisive, our estimation procedure does not refer to the existence of $\tY^*$.
 Therefore, our model~\eqref{eq:model} is general and still valid in the absence of quantization process.
@@
 \end{assumption}
-\vspace{-.1cm}
+\vspace{-.2cm}
 \subsection{Problem 1: Tensor denoising}~\label{sec:denoising}
-\vspace{-.1cm}
+\vspace{-.2cm}
 The first question we aim to address is tensor denoising:
@@
 The Tucker low-rankness is popularly imposed in tensor data analysis, and is shown to provide a reasonable tradeoff between model complexity and model flexibility.
-Note that, unlike matrices, there are various notations of tensor low-rankness, such as CP rank~\cite{hitchcock1927expression} and train rank~\cite{oseledets2011tensor}.
-Some notation of low-rankness may lead to mathematically ill-posed optimization; for example, the best low CP-rank tensor approximation may not exist~\cite{de2008tensor}.
 We choose Tucker representation for well-posedness of optimization and easy interpretation.
@@
 The entry-wise bound assumption is a technical condition that avoids the degeneracy in probability estimation with ordinal observations.
-\vspace{-.1cm}
+\vspace{-.2cm}
 \subsection{Problem 2: Tensor completion}
-\vspace{-.1cm}
+\vspace{-.2cm}
 Motivated by applications in collaborative filtering, we also consider a more general setup when only a subset of tensor entries $y_\omega$ are observed.
 Let $\Omega\subset[d_1]\times \cdots\times[d_K]$ denote the set of observed indices.
-The second question we aim to address is stated as follows:
+The second question is stated as follows:
 
@@
 We adopt the same assumption for higher-order tensors.
+\vspace{-.3cm}
+\section{Rank-constrained M-estimator}\label{sec:theory}
 \vspace{-.2cm}
-\section{Rank-constrained M-estimator}\label{sec:theory}
-\vspace{-.1cm}
 We present a general treatment to both problems mentioned above.
 With a little abuse of notation, we use $\Omega$ to denote either the full index set $\Omega=[d_1]\times \cdots \times [d_K]$ (for the tensor denoising) or a random subset induced from the sampling distribution $\Pi$ (for the tensor completion).
-Define $b_0=-\infty$, $b_L=\infty$, $f(-\infty)=0$ and $f(\infty)=1$.
+Define $f(-\infty)=0$ and $f(\infty)=1$.
 The log-likelihood associated with the observed entries is
 \begin{align}\label{eq:objective}
-\logl(\Theta, \mb)&=\sum_{\omega\in\Omega}\sum_{\ell\in[L]} \Big\{\mathds{1}_{\{y_\omega=\ell\}} \log \big[f(b_\ell-\theta_\omega)- \notag \\
-&\quad \quad \quad \quad \quad \quad \quad  f(b_{\ell-1}-\theta_\omega)\big]\Big\}.
+\logl(\Theta, \mb)=\sum_{\omega\in\Omega}\sum_{\ell\in[L]} \Big\{\mathds{1}_{\{y_\omega=\ell\}} \log \big[f(b_\ell-\theta_\omega)-  f(b_{\ell-1}-\theta_\omega)\big]\Big\}.
 \end{align}
@@
 \begin{align}\label{eq:estimator}
-\hat \Theta&=\argmax_{\Theta\in \tP}\logl(\Theta, \mb),\ \text{where}\notag \\
-\tP&=\left\{\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K} \colon \text{rank}(\Theta)\leq \mr,\ \mnormSize{}{\Theta}\leq \alpha\right\}.
+\hat \Theta=\argmax_{\Theta\in \tP}\logl(\Theta, \mb),\ \text{ where } 
+\tP=\left\{\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K} \colon \text{rank}(\Theta)\leq \mr,\ \mnormSize{}{\Theta}\leq \alpha\right\}.
  \end{align}
@@
 \begin{align}\label{eq:regular}
-A_\alpha &= \min_{\ell\in[L], |\theta|\leq \alpha} g_\ell(\theta),\quad  U_\alpha=\max_{\ell\in[L], |\theta|\leq \alpha } {|\dot{g}_\ell(\theta)|\over g_\ell(\theta)},\\
- L_\alpha &= \min_{\ell\in[L], |\theta|\leq \alpha }\left[{\dot{g}_\ell^2 (\theta)\over g_\ell^2(\theta)} -{\ddot{g}_\ell(\theta)\over g_\ell(\theta)}\right],
+A_\alpha = \min_{\ell\in[L], |\theta|\leq \alpha} g_\ell(\theta),\quad  U_\alpha=\max_{\ell\in[L], |\theta|\leq \alpha } {|\dot{g}_\ell(\theta)|\over g_\ell(\theta)},
+ L_\alpha = \min_{\ell\in[L], |\theta|\leq \alpha }\left[{\dot{g}_\ell^2 (\theta)\over g_\ell^2(\theta)} -{\ddot{g}_\ell(\theta)\over g_\ell(\theta)}\right],
 \end{align}
@@
 Under the Assumption~\ref{ass:link}, all these quantities are strictly positive and independent of tensor dimension.
-\vspace{-.1cm}
+\vspace{-.2cm}
 \subsection{Estimation error for tensor denoising}\label{sec:denosing}
-\vspace{-.1cm}
+\vspace{-.2cm}
 For the tensor denoising problem, we assume that the full set of tensor entries are observed. We assess the estimation accuracy using the mean squared error (MSE):
@@
 \begin{align}\label{eq:lower}
-\inf_{\hat \Theta }\sup_{\trueT\in\tP} \mathbb{P}\Big\{& \text{MSE}(\hat \Theta, \trueT) \\
-& \geq c\min\left( \alpha^2, \ { Cr_{\max}d_{\max} \over \prod_k d_k} \right) \Big\} \geq {1\over 8},
+\inf_{\hat \Theta }\sup_{\trueT\in\tP} \mathbb{P}\Big\{ \text{MSE}(\hat \Theta, \trueT) 
+\geq c\min\left( \alpha^2, \ { Cr_{\max}d_{\max} \over \prod_k d_k} \right) \Big\} \geq {1\over 8},
 \end{align}
@@
 We see that the lower bound matches the upper bound in~\eqref{eq:rate} on the polynomial order of tensor dimension. Therefore, our estimator~\eqref{eq:estimator} is order-optimal.
-
+\vspace{-.2cm}
 \subsection{Sample complexity for tensor completion}
+\vspace{-.2cm}
 We now consider the tensor completion problem, when only a subset of entries $\Omega$ are observed.
@@
 \begin{align}\label{eq:weighted}
-\PiFnormSize{}{\Theta-\hat \Theta}^2&\stackrel{\text{def}}{=}
-{1\over |\Omega|}\mathbb{E}_{\Omega\sim \Pi}\FnormSize{}{\Theta-\hat \Theta}^2\notag \\
-&=\sum_{\omega\in[d_1]\times \cdots \times [d_K]} \pi_{\omega}(\Theta_{\omega}-\hat \Theta_{\omega})^2.
+\PiFnormSize{}{\Theta-\hat \Theta}^2\stackrel{\text{def}}{=}
+{1\over |\Omega|}\mathbb{E}_{\Omega\sim \Pi}\FnormSize{}{\Theta-\hat \Theta}^2
+=\sum_{\omega\in[d_1]\times \cdots \times [d_K]} \pi_{\omega}(\Theta_{\omega}-\hat \Theta_{\omega})^2.
 \end{align}
@@
 
+\vspace{-.3cm}
+\section{Numerical Implementation}\label{sec:algorithm}
 \vspace{-.2cm}
-\section{Numerical Implementation}\label{sec:algorithm}
-\vspace{-.1cm}
 We describe the algorithm to seek the optimizer of~\eqref{eq:objective}.
 In practice, the cut-off points $\mb$ are often unknown, so we choose to maximize $\logl$ jointly over $(\Theta,\mb)\in\tP\times \tB$.
-The objective $\logl$ is concave in $(\Theta,\mb)$ whenever $f'$ is log-concave (see the Supplement).
-However, the feasible set $\tP$ is non-convex, which makes the optimization~\eqref{eq:objective} a non-convex problem.
+The non convexity of feasible set $\tP$  makes the optimization~\eqref{eq:objective} a non-convex problem.
 We employ the alternating optimization approach by utilizing the Tucker representation of $\Theta$.
@@
 The convergence is guaranteed whenever $\logl$ is bounded from above, since the alternating procedure monotonically increases the objective.
-The Algorithm~\ref{alg} gives the full description.
 
-\begin{algorithm}[t]
-        \caption{Ordinal tensor decomposition}\label{alg}
-        \begin{algorithmic}[]
-            \STATE{\bfseries Input:}  \text{ Ordinal data tensor }
-            $\mathcal{Y}\in [L]^{d_1\times\cdots\times d_K}$, rank $\mr\in \mathbb{N}_{+}^{K}$, entry-wise bound $\alpha\in \mathbb{R_+}$.
-            \STATE{\bfseries Output:} $(\hat\Theta,\hat{\mb}) =  \argmax_{(\Theta,\mb)\in \mathcal{P}\times\mathcal{B}}  \mathcal{L}_{\mathcal{Y},\Omega}(\Theta,\mb).$
-            \STATE Random initialization of core tensor $\mathcal{C}^{(0)}$, factor matrices $ \{\mM_k^{(0)}\}$, and cut-off points $\mb^{(0)}$.
-            \FOR{$t = 1,2,\cdots,$}
-            \FOR{$k = 1,2,\cdots,K$}
-            \STATE{Update $\mM^{(t+1)}_k$ while fixing other blocks:}
-            \STATE $\mM_k^{(t+1)}\gets\argmax_{\mM_k\in\mathbb{R}^{d_k\times r_K}}\mathcal{L}_{\mathcal{Y},\Omega}(\mM_k)$,
-            \newline
-             s.t. $\mnormSize{}{\Theta^{(t+1)}}\leq \alpha$, where $\Theta^{(t+1)}$ is the parameter tensor based on the current block estimates.
-            \ENDFOR
-            \STATE {Update $\mathcal{C}^{(t+1)}$ while fixing other blocks:}
-              \newline$\mathcal{C}^{(t+1)} \gets \argmax_{\mathcal{C}\in \mathbb{R}^{r_1\times\cdots\times r_K}}\mathcal{L}_{\mathcal{Y},\Omega}(\mathcal{C})$, s.t.
-$\mnormSize{}{\Theta^{(t+1)}}\leq \alpha.$
-               \STATE {Update $\Theta^{(t+1)}$ based on the current block estimates:}
-                   \newline
-                   $\Theta^{(t+1)} \gets \mathcal{C}^{(t+1)}\times_1\mM_1^{(t+1)}\cdots\times_K\mM_K^{(t+1)}.$
-            \STATE {Update $\mb^{(t+1)}$ while fixing $\Theta^{(t+1)}$:}
-              \newline
-              $\mb^{(t+1)} \gets \argmax_{\mb\in \tB}\mathcal{L}_{\mathcal{Y},\Omega}\big(\Theta^{(t+1)},\mb\big).$
-            \ENDFOR
-            \STATE \textbf{return}
-            $(\hat \Theta,\hat \mb)$
-        \end{algorithmic}
- \end{algorithm}
 We comment on two implementation details before concluding this section.
-First, the problem~\eqref{eq:estimator} is non-convex, so Algorithm 1 usually has no theoretical guarantee on global optimality.
+First, the problem~\eqref{eq:estimator} is non-convex, so the algorithm usually has no theoretical guarantee on global optimality.
 Nevertheless, as shown in Section~\ref{sec:denosing}, the desired rate holds not only for the global optimizer, but also for the local optimizer with $\logl(\hat \Theta)\geq \logl(\trueT)$.
 In practice, we find the convergence point $\hat \Theta$ upon random initialization is often satisfactory, in that the corresponding objective $\logl(\hat \Theta)$ is close to and actually slightly larger than the objective evaluated at the true parameter $\logl(\trueT)$.
-Figure~\ref{fig:stability} shows the trajectory of the objective function that is output in the default setting of Algorithm 1, with the input tensor generated from probit model~\eqref{eq:model} with $d_1=d_2=d_3=d$ and $r_1=r_2=r_3=r$.
+Figure~\ref{fig:stability} shows the trajectory of the objective function from the algorithm with the input tensor generated from probit model~\eqref{eq:model} with $d_1=d_2=d_3=d$ and $r_1=r_2=r_3=r$.
 The dashed line is the objective value at the true parameter $\logl(\trueT)$.
-We find that the algorithm generally converges quickly to a desirable value in reasonable number of steps.
+The algorithm generally converges quickly to a desirable value in reasonable number of steps.
 The actual running time per iteration is shown in the plot legend.
@@
 \centering
-\includegraphics[width=8cm]{algorithm.pdf}\label{fig:stability}
+\includegraphics[width=7cm]{algorithm.pdf}\label{fig:stability}
 \vspace{-.2cm}
-\caption{Trajectory of objective function with various $d$ and $r$.}
+\caption{\footnotesize Trajectory of objective function with various $d$ and $r$.}
 \vspace{-.5cm}
@@
 \begin{align}\label{eq:BIC}
-\hat \mr&=\argmin_{\mr\in\mathbb{N}^K_{+}} \text{BIC}(\mr)\\
-&=\argmin_{\mr\in\mathbb{N}^K_{+}}\{-2\tL_{\tY}(\hat \Theta(\mr),\hat \mb(\mr))+p_e(\mr)\log (\prod_k d_k)\},
+\hat \mr=\argmin_{\mr\in\mathbb{N}^K_{+}} \text{BIC}(\mr)
+=\argmin_{\mr\in\mathbb{N}^K_{+}}\{-2\tL_{\tY}(\hat \Theta(\mr),\hat \mb(\mr))+p_e(\mr)\log (\prod_k d_k)\},
 \end{align}
@@
 
+\vspace{-.3cm}
+\section{Experiments}\label{sec:experiment}
 \vspace{-.2cm}
-\section{Experiments}\label{sec:experiment}
 In this section, we evaluate the empirical performance of our method.
@@
 We consider the setting with $K=3$, $d_1=d_2=d_3=d$, and $r_1=r_2=r_3=r$.
-The parameter tensors are simulated based on~\eqref{eq:Tucker}, where the core tensor entries are i.i.d.\ drawn from $N(0,1)$, and the factors $\mM_k$ are uniformly sampled (with respect to Haar measure) from matrices with orthonormal columns.
+The parameter tensors are simulated based on~\eqref{eq:Tucker}, where the core tensor entries are i.i.d.\ drawn from $N(0,1)$, and the factors $\mM_k$ are uniformly sampled from matrices with orthonormal columns.
 We set the cut-off points $b_\ell=f^{-1}(\ell/L)$ for $\ell\in[L]$, such that $f(b_\ell)$ are evenly spaced from 0 to 1.
@@
 
-\vspace{-.1cm}
+\vspace{-.2cm}
 \subsection{Finite-sample performance}\label{sec:simulation}
-\vspace{-.1cm}
+\vspace{-.2cm}
 The first experiment examines the performance under complete observations.
 We assess the empirical relationship between the MSE and various aspects of model complexity, such as dimension $d$, rank $r$, and signal level $\alpha=\mnormSize{}{\Theta}$.
-Figure~\ref{fig:finite}a plots the estimation error versus the tensor dimension $d$ for three different ranks $r\in\{3,5,8\}$.
+Figure~\ref{fig:finite}a plots the estimation error versus the tensor dimension $d$ for ranks $r\in\{3,5,8\}$.
 The decay in the error appears to behave on the order of $d^{-2}$, which is consistent with our theoretical results~\eqref{eq:rate}.
 We find that a higher rank leads to a larger error, as reflected by the upward shift of the curve as $r$ increases.
-Indeed, a higher rank implies the higher number of parameters to estimate, thus increasing the difficulty of the estimation.
+Indeed, a higher rank implies the more parameters to estimate, thus increasing the difficulty of the estimation.
 Figure~\ref{fig:finite}b shows the estimation error versus the signal level under $d=20$.
@@
 Therefore, the estimation error diverges in the absence of noise $\tE$, because it is impossible to distinguish two different signal tensors, e.g., $\Theta_1=\ma_1\otimes \ma_2\otimes \ma_3$ and $\Theta_2=\text{sign}(\ma_1)\otimes \text{sign}(\ma_2)\otimes \text{sign}(\ma_3)$, from the quantized observations.
-This phenomenon~\cite{davenport2014,sur2019modern} is clearly contrary to the classical continuous-valued tensor problem.
+This phenomenon~\cite{davenport2014,sur2019modern} is  contrary to the classical continuous-valued tensor problem.
 
@@
 \begin{figure}[http]
-\includegraphics[width=8.3cm]{panel.pdf}
+\centering\includegraphics[width=14cm]{panel1.png}
 \vspace{-.4cm}
-\caption{Empirical relationship between (relative) MSE versus (a) dimension $d$, (b) signal level $\alpha$, (c) observation fraction $\rho$, and (d) number of ordinal levels $L$.
+\caption{\footnotesize Empirical relationship between (relative) MSE versus (a) dimension $d$, (b) signal level $\alpha$, (c) observation fraction $\rho$, and (d) number of ordinal levels $L$.
 In panels (b)-(d), we plot the relative MSE $=\FnormSize{}{\hat \Theta-\trueT}/\FnormSize{}{\trueT}$ for better visualization.}
@@
 
-\vspace{-.1cm}
+\vspace{-.2cm}
 \subsection{Comparison with alternative methods}
-\vspace{-.1cm}
+\vspace{-.2cm}
 Next, we compare our ordinal tensor method ({\bf Ordinal-T}) with three popular low-rank methods:
@@
 The first approach ({\bf 1bit-sign-T}) follows from~\citet{ghadermarzy2018learning} that transforms
-$\tY$ to a binary tensor, by taking the entrywise sign of the mean-adjusted tensor, $\tY- |\Omega|^{-1}\sum_\omega y_\omega$.
-The second approach ({\bf 1bit-category-T}) transforms the order-3 ordinal tensor $\tY$ to an order-4 binary tensor $\tY^{\sharp}=\entry{y^{\sharp}_{ijkl}}$ via dummy variable encoding; i.e., $y^{\sharp}_{ijk\ell}=\mathds{1}_{\{y_{ijk}=\ell\}}$ for all $\ell\in[L-1]$.
-We evaluate the methods by their capabilities in predicting the most likely label for each entry, i.e.,\ $y_\omega^{\text{mode}}=\arg\max_\ell\mathbb{P}(y_\omega=\ell)$.
+$\tY$ to a binary tensor, by taking the entrywise sign of the mean-adjusted tensor, %$\tY- |\Omega|^{-1}\sum_\omega y_\omega$. 
+The second approach ({\bf 1bit-category-T}) transforms the order-3 ordinal tensor 
+$\tY$ to an order-4 binary tensor $\tY^{\sharp}=\entry{y^{\sharp}_{ijkl}}$ via dummy variable encoding.
+%; i.e., $y^{\sharp}_{ijk\ell}=\mathds{1}_{\{y_{ijk}=\ell\}}$ for all $\ell\in[L-1]$.
+We evaluate the methods by capabilities in predicting %the most likely label for each entry, i.e.,\ 
+$y_\omega^{\text{mode}}=\arg\max_\ell\mathbb{P}(y_\omega=\ell)$.
 Two performance metrics are considered: mean absolute deviation, $\text{MAD}=d^{-K}\sum_\omega |y_\omega^{\text{mode}}-\hat y_\omega^{\text{mode}}|$, and misclassification rate, $\text{MCR}=d^{-K}\sum_\omega\mathds{1}_{\{y_\omega^{\text{mode}}\neq\text{round}(\hat y_\omega^{\text{mode}})\}}$,
-where $\text{round}(\cdot)$ denotes the nearest integer of the prediction (possibly continuous-valued returned by {\bf Continuous-T}).
-Note that MAD penalizes the large deviation more heavily than MCR.
+where $\text{round}(\cdot)$ gives the nearest integer of the prediction.
+%(possibly continuous-valued from {\bf Continuous-T}). 
+%Note that MAD penalizes the large deviation more heavily than MCR.
 
@@
 In particular, methods built on multi-level observations ({\bf Ordinal-T}, {\bf Ordinal-M}, {\bf 1bit-category-T}) exhibit stable MCR over $\rho$ and $L$, whereas the others two methods ({\bf Continuous-T}, {\bf 1bit-sign-T}) generally fail except for $L=2$ (Figures~\ref{fig:compare}a-b).
-This observation highlights the necessity of modeling multi-level probabilities in classification task.
+This observation highlights the necessity of modeling multi-level probabilities.
+%in classification task. 
 Interestingly, although both {\bf 1bit-category-T} and our method {\bf Ordinal-T} behave similarly for binary tensors ($L=2$), the improvement of our method is substantial as $L$ increases (Figures~\ref{fig:compare}a and~\ref{fig:compare}c).
@@
 We see a clear advantage of our method ({\bf Ordinal-T}) over the matricization ({\bf Ordinal-M}) in both complete and non-complete observations.
-When the observation fraction is small, e.g., ${|\Omega|/ d^K}=0.4$, the tensor-based completion shows $\sim$ 30\% reduction in error compared to the matricization.
+When the observation fraction is small, e.g., ${|\Omega|/ d^K}=0.4$, the tensor-based completion shows $\sim$ 30\% error reduction  compared to the matricization.
 
 \begin{figure}[t]
-\includegraphics[width=8.3cm]{panel2.pdf}
+\centering\includegraphics[width=14cm]{panel2.png}
 \vspace{-.4cm}
-\caption{Performance comparison in MCR (a, b) and MAD (c, d).
+\caption{\footnotesize Performance comparison in MCR (a, b) and MAD (c, d).
 (b, d) Prediction errors versus the number of ordinal levels $L$ when $\rho=0.8$.
@@
 
-We also compare the methods by their performance in predicting the median label, $y_\omega^{\text{median}}=\min\{\ell\colon\mathbb{P}(y_\omega=\ell)\geq 0.5\}$.
+%We also compare the methods by their performance in predicting the median label, $y_\omega^{\text{median}}=\min\{\ell\colon\mathbb{P}(y_\omega=\ell)\geq 0.5\}$.
 Under the latent variable model~\eqref{eq:latent} and Assumption~\ref{ass:link}, the median label is the quantized $\Theta$ without noise; i.e.\ $y_\omega^{\text{median}}=\sum_\ell \mathds{1}_{\theta_\omega\in(b_{\ell-1},b_\ell]}$.
@@
 
-\vspace{-.1cm}
+\vspace{-.3cm}
 \section{Data Applications}
-\vspace{-.1cm}
-We apply our ordinal tensor method to two real-world datasets.
-In the first application, we use our model to analyze an ordinal tensor consisting of structural connectivities among 68 brain regions for 136 individuals from Human Connectome Project (HCP)~\citep{geddes2016human}.
-In the second application, we perform tensor completion to an ordinal dataset with missing values.
-The data tensor records the ratings on a scale of 1 to 5 from 42 users to 139 songs on 26 contexts~\citep{baltrunas2011incarmusic}.
+\vspace{-.2cm}
+We apply our method to two datasets.
+In the first application, we analyze an ordinal tensor consisting of structural connectivities among 68 brain regions from 136 individuals in Human Connectome Project (HCP)~\citep{van2013wu}.
+In the second application, we complete ordinal tensor data which records the ratings on a scale of 1 to 5 from 42 users to 139 songs on 26 contexts~\citep{baltrunas2011incarmusic} with missing values.
 
-\vspace{-.1cm}
+\vspace{-.2cm}
 \subsection{Human Connectome Project (HCP)}
-\vspace{-.1cm}
+\vspace{-.2cm}
 Each entry in the HCP dataset takes value on a nominal scale, \{{\it high, moderate, low}\}, indicating the strength level of fiber connection.
-We convert the dataset to a 3-level ordinal tensor $\tY\in[3]^{68\times 68\times 136}$ and apply the ordinal tensor method with a logistic link function.
-The BIC suggests $\mr = (23,23,8)$ with $\logl(\hat{\Theta},\hat{\mb}) = -216,646$.
-Based on the estimated Tucker factors $\{\hat \mM_k\}$, we perform a clustering analysis via K-mean on the brain nodes (see detailed procedure in the Supplement).
-We find that the 68 brain nodes are grouped into 11 clusters, and the clustering captures the spatial separation of brain nodes very well (see the Supplement).
+We convert the dataset to a 3-level ordinal tensor $\tY\in[3]^{68\times 68\times 136}$ and apply our method with a logistic link function.
+The BIC suggests $\mr = (23,23,8)$.
+Based on the estimated factor matrices $\{\hat \mM_k\}$, we perform a clustering analysis via K-mean (see detailed procedure in the Supplement).
+The 68 brain nodes are grouped into 11 clusters, and the clustering captures the spatial separation of brain nodes (see the Supplement).
 In particular, the top three clusters represent the left, right and connection between two hemispheres; the smaller clusters represent local regions driving by similar nodes.
-For example, the cluster IV consists of nodes in the left supramarginal gyrus region.
-The identified similarity among brain nodes suggests the applicability of our method.
-
-We also compare the goodness-of-fit of various tensor methods on the HPC data.
+We compare the goodness-of-fit of various tensor methods on the HPC data.
 Table~\ref{table:CV} summarizes the prediction error via 5-fold stratified cross-validation averaged over 10 runs.
-Our method outperforms the others, especially in MAD.
+Our method outperforms the others.
 
 \begin{table}[http]
-\resizebox{.48\textwidth}{!}{
-\begin{tabular}{c|c|c}
+\centering
+\resizebox{.9\textwidth}{!}{
+\begin{tabular}{c|c|c|c|c}
 \Xhline{2\arrayrulewidth}
-\multicolumn{3}{c}{Human Connectome Project (HCP) dataset}\\
+&\multicolumn{2}{c|}{Human Connectome Project (HCP) dataset}&\multicolumn{2}{|c}{InCarMusic dataset}\\
 \Xhline{2\arrayrulewidth}
-Method                    & MAD & MCR\\
+Method                    & MAD & MCR& MAD & MCR\\
 \hline
-Ordinal-T (ours) & 0.1607 (0.0005)& 0.1606 (0.0005)\\
- Continuous-T&0.2530 (0.0002)&0.1599 (0.0002)\\
- 1bit-sign-T&0.3566 (0.0010)&0.1563 (0.0010)\\
+Ordinal-T (ours) & 0.1607 (0.0005)& 0.1606 (0.0005) & 1.37 (0.039) &  0.59 (0.009)\\
+ Continuous-T&0.2530 (0.0002)&0.1599 (0.0002) & 2.39 (0.152)&0.94 (0.027)\\
+ 1bit-sign-T&0.3566 (0.0010)&0.1563 (0.0010)&1.39 (0.003)& 0.81 (0.005)\\
  \Xhline{2\arrayrulewidth}
-\multicolumn{3}{c}{InCarMusic dataset}\\
-\Xhline{2\arrayrulewidth}
-Method                    & MAD & MCR\\
-\hline
-Ordinal-T (ours) & 1.37 (0.039) &  0.59 (0.009)\\
-Continuous-T & 2.39 (0.152)&0.94 (0.027)\\
-1bit-sign-T&1.39 (0.003)& 0.81 (0.005)\\
-\hline
 \end{tabular}
 }
-\vspace{-.2cm}
-\caption{Comparison of prediction error in the HPC and InCarMusic analyses. Standard errors are reported in parentheses.}\label{table:CV}
-\vspace{-.3cm}
+
+\caption{\footnotesize Comparison of prediction error in the HPC and InCarMusic analyses. Standard errors are reported in parentheses.}\label{table:CV}
 \end{table}
 
-\vspace{-.1cm}
+\vspace{-.4cm}
 \subsection{InCarMusic recommendation system}
-\vspace{-.1cm}
+\vspace{-.2cm}
 We apply ordinal tensor completion to a recommendation system {\it InCarMusic}.
-{\it InCarMusic} is a mobile application that offers music recommendation to passengers of cars based on contexts~\cite{baltrunas2011incarmusic}.
-Our goal is to perform tensor completion to impute the unobserved entries in the $42\times 139\times 26$ ordinal tensor and thereby we can offer context-specific music recommendation to users.
-The data tensor consists of 2,884 observed entries.
+{\it InCarMusic} is a mobile application that offers music recommendation to passengers based on contexts~\cite{baltrunas2011incarmusic}.
+we conduct tensor completion on the $42\times 139\times 26$ ordinal tensor with 2,844 observed entries.
  Table~\ref{table:CV} shows the averaged prediction error via 5-fold cross validation.
@@
 
+\vspace{-.3cm}
+\section{Conclusions}
 \vspace{-.2cm}
-\section{Conclusions}
-\vspace{-.1cm}
 We have developed a low-rank tensor estimation method based on possibly incomplete, ordinal-valued observations.
@@
 
-\newpage
-\bibliography{ordinalTv3.bbl}
-\bibliographystyle{icml2020}
+\bibliographystyle{plainnat}
+\bibliography{tensor_wang.bib}
 
 \end{document}
-
-% This document was modified from the file originally made available by
-% Pat Langley and Andrea Danyluk for ICML-2K. This version was created
-% by Iain Murray in 2018, and modified by Alexandre Bouchard in
-% 2019 and 2020. Previous contributors include Dan Roy, Lise Getoor and Tobias
-% Scheffer, which was slightly modified from the 2010 version by
-% Thorsten Joachims & Johannes Fuernkranz, slightly modified from the
-% 2009 version by Kiri Wagstaff and Sam Roweis's 2008 version, which is
-% slightly modified from Prasad Tadepalli's 2007 version which is a
-% lightly changed version of the previous year's version by Andrew
-% Moore, which was in turn edited from those of Kristian Kersting and
-% Codrina Lauth. Alex Smola contributed to the algorithmic style files.

### 2020-05-31 paper/NeurIPS/ordinalT.tex
Add files via upload

@@
  \usepackage{bm}
+\usepackage{caption}
 
@@
 \begin{table*}[ht]
-\centerline{\includegraphics[width=13cm]{compare.pdf}}
+\centerline{\includegraphics[width=14.2cm]{compare.pdf}}
 \vspace{-.3cm}
-\caption{Comparison with previous work.
-For ease of presentation, we summarize the error rate and sample complexity assuming equal tensor dimension in all modes.
+\caption{\footnotesize Comparison with previous work.
+We summarize the error rate and sample complexity assuming equal tensor dimension in all modes.
 $K$: tensor order; $L$: number of ordinal levels; $d$: dimension at each mode.
@@
 
-{\bf Our contribution.} This paper presents an efficient low-rank estimation method and theory for tensors with ordinal-valued entries.
-Our main contributions are summarized in Table~\ref{tab:compare}.
-We propose a cumulative link model for higher-order tensors, develop a rank-constrained M-estimator, and obtain theoretical accuracy guarantees.
-We successfully recover true signal where neither the underlying signal nor the quantization operator is known.
-We show that our mean squared error bound has minimax optimal rate under the low-rank model.
-Our estimator enjoys a faster convergence rate $\tO(d^{-(K-1)/2})$ than $\tO(d^{-K})$ in~\citet{ghadermarzy2018learning}, which is a substantial improvement as the order $K$ increases.
+{\bf Our contribution.} This paper presents a low-rank estimation method and theory for tensors with ordinal-valued entries. 
+Our main contributions are summarized in Table~\ref{tab:compare}. %We propose a cumulative link model, develop a rank-constrained M-estimator, and obtain theoretical accuracy guarantees. 
+We are among the first to establish the recovery theory for both signal tensor and the quantization operators from a limited number of highly discrete entries.
+We show that our mean squared error bound is minimax optimal under the low-rank model.
 Proposed tensor completion algorithm guarantees consistent recovery of an order-$K$ $(d,\ldots,d)$-dimensional low-rank tensor using only $\tilde \tO(Kd)$ noisy, quantized observations.
@@
 Existing work focuses exclusively on univariate observations such as binary- or continuous-valued entries~\citep{wang2018learning,hong2018generalized,ghadermarzy2018learning}.
-As we mentioned earlier, the ordinal observations add considerable challenges to the model formulation.
-We address the challenges from two perspectives.
-From statistical perspective, our proposed model generalizes the usual binary tensor model while preserving palindromic invariance~\cite{mccullagh1980regression} for ordinal observations.
-From algorithm perspective, our alternating optimization compares favorably to the approximated (convex) algorithm developed in the context of binary tensors~\cite{ghadermarzy2018learning}.
-We numerically compare the two approaches in Section~\ref{sec:experiment}.
+We address ordinal tensor observations from two perspective.
+From statistical perspective, our proposed model generalizes the binary tensor model while preserving palindromic invariance~\cite{mccullagh1980regression} for ordinal observations. 
+From algorithm perspective,  we address regularized MLE with  alternating optimization (Non-convex).
+The approximated (convex) algorithm is developed in the context of binary tensors~\cite{ghadermarzy2018learning}.
+However, we advocate our approach and provide strong evidence.
+We show the improvement of error bound from $\tO(d^{-(K-1)/2})$ to $\tO(d^{-K})$ in~\cite{ghadermarzy2018learning} and numerically compare the two approaches.
 
@@
 We use lower-case letters ($a, b, c, \ldots$) for scalars/vectors, upper-case boldface letters ($\mA, \mB, \mC, \ldots$) for matrices, and calligraphy letters ($\tA,\tB,\tC,\ldots$) for tensors of order three or greater.
-For ease of notation, we allow the basic arithmetic operators (e.g., $\leq, +, -$) to be applied to pairs of tensors in an element-wise manner.
 We use the shorthand $[n]$ to denote the $n$-set $\{1,\ldots,n\}$ for $n \in N_{+}$.
@@
 \end{equation}
-where $\mb=(b_1,\ldots,b_{L-1})$ is a set of unknown scalars satisfying $b_1<\cdots <b_{L-1}$, $\Theta=\entry{\theta_\omega}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is a continuous-valued parameter tensor satisfying certain low-rank structure (to be specified later), and $f(\cdot):\mathbb{R}\mapsto[0,1]$ is a known, strictly increasing function.
+where $\mb=(b_1,\ldots,b_{L-1})$ is a set of  scalars satisfying $b_1<\cdots <b_{L-1}$, $\Theta=\entry{\theta_\omega}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is a continuous-valued parameter tensor satisfying  low-rank structure, and $f(\cdot):\mathbb{R}\mapsto[0,1]$ is a known, strictly increasing function.
 We refer to $\mb$ as the cut-off points and $f$ the link function.
@@
 \begin{example}[Logistic model] The logistic model is characterized by~\eqref{eq:model} with $f(\theta)=(1+e^{-\theta/\sigma})^{-1}$, where $\sigma>0$ is the scale parameter.
-Equivalently, the noise $\varepsilon_\omega$ in~\eqref{eq:quantization} follows i.i.d.\ logistic distribution with scale parameter $\sigma$.
+Equivalently, the noise $\varepsilon_\omega$ in~\eqref{eq:quantization} follows i.i.d.\ $\text{Logistic}(0,\sigma)$.
 \end{example}
@@
 The link function is assumed to satisfy:
-\vspace{-.4cm}
+\vspace{-.3cm}
 \begin{enumerate}
@@
 \end{assumption}
+
 \vspace{-.2cm}
 \subsection{Problem 1: Tensor denoising}~\label{sec:denoising}
-\vspace{-.2cm}
 The first question we aim to address is tensor denoising:
@@
 
-Clearly, the problem (P1) cannot be solved uniformly for all possible $\Theta$.
 We focus on a class of ``low-rank'' and ``flat'' signal tensors, which is a plausible assumption in practical applications~\cite{zhou2013tensor,bhaskar20151}.
@@
 \begin{equation}\label{eq:space}
-\small \tP=\left\{\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K} \colon \text{rank}(\Theta)\leq \mr,\ \mnormSize{}{\Theta}\leq \alpha\right\}.
+\small \tP=\left\{\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K} \colon \text{rank}(\Theta)\leq \mr,\ \mnormSize{}{\Theta}\leq \alpha\right\},\text{ where }\mr=(r_1,\ldots,r_K) \text{ denotes the Tucker rank of } \Theta.
 \end{equation}
-where $\mr=(r_1,\ldots,r_K)$ denotes the Tucker rank of $\Theta$.
 
@@
 We choose Tucker representation for well-posedness of optimization and easy interpretation.
-
 The second constraint is that the entries of $\Theta$ are uniformly bounded in magnitude by a constant $\alpha \in \mathbb{R}_{+}$.
@@
 We consider a general model on $\Omega$ that allows both uniform and non-uniform sampling.
-Specifically, let $\Pi=\{\pi_{i_1,\ldots,i_K}\}$ denote a predefine probability distribution over the index set such that $\sum_{\omega\in[d_1]\times \cdots \times [d_K]} \pi_\omega =1$.
+Let $\Pi=\{\pi_{i_1,\ldots,i_K}\}$ denote a predefine probability distribution over the index set such that $\sum_{\omega\in[d_1]\times \cdots \times [d_K]} \pi_\omega =1$.
 We assume that each index in $\Omega$ is drawn with replacement using distribution $\Pi$.
-This sampling model relaxes the uniform sampling in literature and is arguably a better fit in applications.
+This sampling model relaxes the uniform sampling in literature.
 
-We consider the same parameter space~\eqref{eq:space} for the completion problem.
-In addition to the reasons mentioned in Section~\ref{sec:denoising}, the entrywise bound assumption also serves as the incoherence requirement for completion.
-In classical matrix completion, the incoherence is often imposed on the singular vectors.
-This assumption is recently relaxed for ``flat'' matrices with bounded magnitude~\cite{negahban2011estimation,cai2013max,bhaskar20151}.
-We adopt the same assumption for higher-order tensors.
 \vspace{-.3cm}
@@
 This suggests that the local optimality itself is not necessarily a severe concern in our context, as long as the convergent objective is large enough.
-In Section ~\ref{sec:algorithm}, we perform empirical studies to assess the algorithmic stability.
 
-A similar conclusion is obtained for the prediction error, measured in Kullback-Leibler (KL) divergence, between the categorical distributions in the observation space.
-\begin{cor}[Prediction error]~\label{cor:prediction}
-Assume the same set-up as in Theorem~\ref{thm:rate}.
-Let $\mathbb{P}_{\tY}$ and $\hat{\mathbb{P}}_{\tY}$ denote the distributions generating the $L$-level ordinal tensor $\tY$, given the true parameter $\Theta$ and its estimator $\hat \Theta$, respectively.
-Assume $L\geq 2$.
-Then, with very high probability,
-\begin{equation}\label{eq:KLrate}
-\text{KL}(\mathbb{P}_{\tY} || \hat{\mathbb{P}}_{\tY}) \leq  {c_1 U^2_\alpha r_{\max}^{K-1}  \over L^2_\alpha } {(4L-6)\dot{f}^2(0)  \over A_\alpha} {\sum_kd_k\over  \prod_k d_k},
-\end{equation}
-where $c_1 >0$ is the same constants as in Theorem~\ref{thm:rate}.
-\end{cor}
 To gain insight into these bounds, we consider a special setting with equal dimension in all modes, i.e., $d_1=\cdots=d_K=d$. In such a case, our bound \eqref{eq:rate} reduces to
@@
 \end{equation}
-Hence, our estimator achieves consistency with polynomial convergence rate.
+Our estimator achieves consistency with polynomial convergence rate.
 We compare the bound with existing literature.
-In the special case $L=2$, \citet{ghadermarzy2018learning} proposed a max-norm constrained estimator $\tilde{\Theta}$ with $\text{MSE}(\tilde{\Theta}, \trueT) \asymp  d^{-(K-1)/2}$.
-In contrast, our estimator converges at a rate of $d^{-(K-1)}$, which is substantially faster than theirs.
-This provides a positive answer to the open question posed in~\citet{ghadermarzy2018learning} whether the square root in the bound is removable.
-The improvement stems from utilizing the exact low-rankness of $\Theta$, whereas the surrogate rank measure employed in~\citet{ghadermarzy2018learning} is scale-sensitive.
+In the special case $L=2$, \citep{ghadermarzy2018learning} proposed a max-norm constrained estimator $\tilde{\Theta}$ with $\text{MSE}(\tilde{\Theta}, \trueT) \asymp  d^{-(K-1)/2}$.
+In contrast, our estimator converges at a rate of $d^{-(K-1)}$, which is much faster than theirs.
+This provides an answer to the open question posed in~\citep{ghadermarzy2018learning} whether the square root in the bound is removable.
+The improvement stems from utilizing the exact low-rankness of $\Theta$, whereas the surrogate rank measure employed in~\citep{ghadermarzy2018learning} is scale-sensitive.
 
@@
 The convergence rate for rank-constrained matrix estimation was $\tO(1/\sqrt{d})$~\citep{bhaskar2016probabilistic}, which fits into our special case when $K=2$.
-Furthermore, our results~\eqref{eq:rate} and~\eqref{eq:KLrate} reveal that the convergence becomes favorable as the order of data tensor increases.
-Intuitively, the sample size for tensor data analysis is the number of entries, $\prod_k d_k$, and the number of free parameters is roughly on the order of $\sum_{k}d_k$, assuming $r_{\max}=\tO(1)$.
-A higher tensor order implies higher effective sample size per parameter, and thus exhibits a faster convergence rate in high dimensions.
+Furthermore, \eqref{eq:rate} reveals that the convergence becomes favorable as the order of data tensor increases.
+Intuitively, the sample size for tensor analysis is the number of entries, $\prod_k d_k$, and the number of free parameters is roughly on the order of $\sum_{k}d_k$, assuming $r_{\max}=\tO(1)$.
+A higher tensor order implies more effective sample size per parameter, and thus has a faster convergence rate in high dimensions.
 
@@
 \vspace{-.2cm}
-We describe the algorithm to seek the optimizer of~\eqref{eq:objective}.
  In practice, the cut-off points $\mb$ are often unknown, so we choose to maximize $\logl$ jointly over $(\Theta,\mb)\in\tP\times \tB$.
 The non convexity of feasible set $\tP$  makes the optimization~\eqref{eq:objective} a non-convex problem.
-We employ the alternating optimization approach by utilizing the Tucker representation of $\Theta$.
-Specifically, based on~\eqref{eq:Tucker} and~\eqref{eq:objective}, the objective function consists of $K+2$ blocks of variables, one for the cut-off points $\mb$, one for the core tensor $\tC$, and $K$ for the factor matrices $\mM_k$'s.
-The optimization is a simple convex problem if any $K+1$ out of the $K+2$ blocks are fixed.
-We update one block at a time while holding others fixed, and alternate the optimization throughout the iteration.
-The convergence is guaranteed whenever $\logl$ is bounded from above, since the alternating procedure monotonically increases the objective.
+We employ the alternating optimization approach by utilizing the Tucker representation of $\Theta$ and make each update convex problem.
+Multiple initializations are utilized, the sub-step for solving tensor factors is paralleled, and can be boosted by stochastic gradient descent.
+The detailed algorithm is in the Supplements.
 
-We comment on two implementation details before concluding this section.
-First, the problem~\eqref{eq:estimator} is non-convex, so the algorithm usually has no theoretical guarantee on global optimality.
+The problem~\eqref{eq:estimator} is non-convex, so the algorithm usually has no theoretical guarantee on global optimality.
 Nevertheless, as shown in Section~\ref{sec:denosing}, the desired rate holds not only for the global optimizer, but also for the local optimizer with $\logl(\hat \Theta)\geq \logl(\trueT)$.
 In practice, we find the convergence point $\hat \Theta$ upon random initialization is often satisfactory, in that the corresponding objective $\logl(\hat \Theta)$ is close to and actually slightly larger than the objective evaluated at the true parameter $\logl(\trueT)$.
-Figure~\ref{fig:stability} shows the trajectory of the objective function from the algorithm with the input tensor generated from probit model~\eqref{eq:model} with $d_1=d_2=d_3=d$ and $r_1=r_2=r_3=r$.
-The dashed line is the objective value at the true parameter $\logl(\trueT)$.
+Figure~\ref{fig:stability} shows the trajectory of the objective function from the algorithm.
 The algorithm generally converges quickly to a desirable value in reasonable number of steps.
@@
 
-\begin{figure}[H]
+\begin{minipage}{0.58\textwidth}
 \centering
-\includegraphics[width=7cm]{algorithm.pdf}\label{fig:stability}
-\vspace{-.2cm}
-\caption{\footnotesize Trajectory of objective function with various $d$ and $r$.}
-\vspace{-.5cm}
-\end{figure}
+\includegraphics[width=6cm]{algorithm.pdf}\label{fig:stability}
+\end{minipage}
+\begin{minipage}{0.4\textwidth}
+\captionof{figure}{\footnotesize  Trajectory of objective function with the input tensor generated from probit model~\eqref{eq:model} with $d_1=d_2=d_3=d$ and $r_1=r_2=r_3=r$.
+The dashed line is the objective value at the true parameter $\logl(\trueT)$.}
+\end{minipage}
 
-Second, the algorithm takes the rank $\mr$ as an input.
+The algorithm takes the rank $\mr$ as an input.
 In practice, the rank $\mr$ is hardly known and needs to be estimated from the data.
@@
 
-\vspace{-.2cm}
+\vspace{-.3cm}
 \subsection{Finite-sample performance}\label{sec:simulation}
@@
 The first experiment examines the performance under complete observations.
-We assess the empirical relationship between the MSE and various aspects of model complexity, such as dimension $d$, rank $r$, and signal level $\alpha=\mnormSize{}{\Theta}$.
-Figure~\ref{fig:finite}a plots the estimation error versus the tensor dimension $d$ for ranks $r\in\{3,5,8\}$.
+We assess the empirical relationship between the MSE and various  model complexity, such as dimension $d$, rank $r$, and signal level $\alpha=\mnormSize{}{\Theta}$.
+Figure~\ref{fig:finite}a plots the MSE versus the tensor dimension $d$ for ranks $r\in\{3,5,8\}$.
 The decay in the error appears to behave on the order of $d^{-2}$, which is consistent with our theoretical results~\eqref{eq:rate}.
@@
 Indeed, a higher rank implies the more parameters to estimate, thus increasing the difficulty of the estimation.
-Figure~\ref{fig:finite}b shows the estimation error versus the signal level under $d=20$.
+Figure~\ref{fig:finite}b shows the MSE versus the signal level under $d=20$.
 Interestingly, a larger estimation error is observed when the signal is either too small or too large.
@@
 We consider $L$-level tensors with $d=20$, $\alpha=10$ and choose a subset of tensor entries via uniform sampling.
-Figure~\ref{fig:finite}c shows the estimation error of $\hat \Theta$ versus the fraction of observation $\rho=|\Omega|/d^K$.
+Figure~\ref{fig:finite}c shows the  MSE of $\hat \Theta$ versus the fraction of observation $\rho=|\Omega|/d^K$.
 As expected, the error reduces with increased $\rho$ or decreased $r$.
@@
 The result showcases the benefit of multi-level observations compared to binary observations.
-
+\vspace{-.3cm}
 \begin{figure}[http]
-\centering\includegraphics[width=14cm]{panel1.png}
+\centering\includegraphics[width=14cm]{panel.pdf}
 \vspace{-.4cm}
-\caption{\footnotesize Empirical relationship between (relative) MSE versus (a) dimension $d$, (b) signal level $\alpha$, (c) observation fraction $\rho$, and (d) number of ordinal levels $L$.
-In panels (b)-(d), we plot the relative MSE $=\FnormSize{}{\hat \Theta-\trueT}/\FnormSize{}{\trueT}$ for better visualization.}
+\caption{\footnotesize Empirical relationship between (relative) MSE versus (a) dimension $d$, (b) signal level $\alpha$, (c) observation fraction $\rho$, and (d) number of ordinal levels $L$.}
 \label{fig:finite}
-\vspace{-.2cm}
+\vspace{-.4cm}
 \end{figure}
@@
 
-We apply each of the above methods to $L$-level ordinal tensors $\tY$ generated from model~\eqref{eq:model}.
-The {\bf Continuous-T} is applied directly to $\tY$ by treating the $L$ levels as continuous observations.
+We apply the each method to $L$-level ordinal tensors $\tY$ generated from model~\eqref{eq:model}.
+The {\bf Continuous-T} is applied to $\tY$ treating the $L$ levels as continuous observations.
 The {\bf Ordinal-M} is applied to the matrix $\tY_{(1)}$ obtained via 1-mode unfolding.
 The {\bf 1bit-T} is applied to $\tY$ in two ways.
-The first approach ({\bf 1bit-sign-T}) follows from~\citet{ghadermarzy2018learning} that transforms
-$\tY$ to a binary tensor, by taking the entrywise sign of the mean-adjusted tensor, %$\tY- |\Omega|^{-1}\sum_\omega y_\omega$. 
+The first approach ({\bf 1bit-sign-T}) follows from~\citep{ghadermarzy2018learning} that transforms
+$\tY$ to a binary tensor, by taking the entrywise sign of the mean-adjusted tensor. %$\tY- |\Omega|^{-1}\sum_\omega y_\omega$. 
 The second approach ({\bf 1bit-category-T}) transforms the order-3 ordinal tensor 
-$\tY$ to an order-4 binary tensor $\tY^{\sharp}=\entry{y^{\sharp}_{ijkl}}$ via dummy variable encoding.
+$\tY$ to an order-4 binary tensor $\tY^{\sharp}=\entry{y^{\sharp}_{ijkl}}$ via dummy variable.
 %; i.e., $y^{\sharp}_{ijk\ell}=\mathds{1}_{\{y_{ijk}=\ell\}}$ for all $\ell\in[L-1]$.
-We evaluate the methods by capabilities in predicting %the most likely label for each entry, i.e.,\ 
-$y_\omega^{\text{mode}}=\arg\max_\ell\mathbb{P}(y_\omega=\ell)$.
-Two performance metrics are considered: mean absolute deviation, $\text{MAD}=d^{-K}\sum_\omega |y_\omega^{\text{mode}}-\hat y_\omega^{\text{mode}}|$, and misclassification rate, $\text{MCR}=d^{-K}\sum_\omega\mathds{1}_{\{y_\omega^{\text{mode}}\neq\text{round}(\hat y_\omega^{\text{mode}})\}}$,
-where $\text{round}(\cdot)$ gives the nearest integer of the prediction.
+We evaluate the method by mean absolute deviation (MAD) and misclassification rate (MCR).
 %(possibly continuous-valued from {\bf Continuous-T}). 
@@
 Figure~\ref{fig:compare} compares the prediction accuracy under the setting $\alpha=10$, $d=20$, and $r=5$.
-The problem size we considered is comparable to~\citet{ghadermarzy2018learning}.
+The problem size we considered is comparable to~\citep{ghadermarzy2018learning}.
 We find that our method outperforms the others in both MAD and MCR.
@@
 Figures~\ref{fig:compare}c-d assess the prediction accuracy with sample size.
-We see a clear advantage of our method ({\bf Ordinal-T}) over the matricization ({\bf Ordinal-M}) in both complete and non-complete observations.
+We see a clear advantage of our method ({\bf Ordinal-T}) over the matricization ({\bf Ordinal-M}).
 When the observation fraction is small, e.g., ${|\Omega|/ d^K}=0.4$, the tensor-based completion shows $\sim$ 30\% error reduction  compared to the matricization.
@@
 \begin{figure}[t]
-\centering\includegraphics[width=14cm]{panel2.png}
+\centering\includegraphics[width=14cm]{panel2.pdf}
 \vspace{-.4cm}
@@
 In the first application, we analyze an ordinal tensor consisting of structural connectivities among 68 brain regions from 136 individuals in Human Connectome Project (HCP)~\citep{van2013wu}.
-In the second application, we complete ordinal tensor data which records the ratings on a scale of 1 to 5 from 42 users to 139 songs on 26 contexts~\citep{baltrunas2011incarmusic} with missing values.
-
-\vspace{-.2cm}
-\subsection{Human Connectome Project (HCP)}
-\vspace{-.2cm}
 Each entry in the HCP dataset takes value on a nominal scale, \{{\it high, moderate, low}\}, indicating the strength level of fiber connection.
@@
 The BIC suggests $\mr = (23,23,8)$.
-Based on the estimated factor matrices $\{\hat \mM_k\}$, we perform a clustering analysis via K-mean (see detailed procedure in the Supplement).
-The 68 brain nodes are grouped into 11 clusters, and the clustering captures the spatial separation of brain nodes (see the Supplement).
+Based on the estimated factor matrices $\{\hat \mM_k\}$, we perform a clustering analysis via K-mean (see detailed procedure  and results in Supplement).
+The 68 brain nodes are grouped into 11 clusters, and the clustering captures the spatial separation of brain nodes.
 In particular, the top three clusters represent the left, right and connection between two hemispheres; the smaller clusters represent local regions driving by similar nodes.
-We compare the goodness-of-fit of various tensor methods on the HPC data.
+We compare the prediction performance with the alternative methods.
 Table~\ref{table:CV} summarizes the prediction error via 5-fold stratified cross-validation averaged over 10 runs.
@@
 \centering
-\resizebox{.9\textwidth}{!}{
+\resizebox{.86\textwidth}{!}{
 \begin{tabular}{c|c|c|c|c}
@@
 
-\caption{\footnotesize Comparison of prediction error in the HPC and InCarMusic analyses. Standard errors are reported in parentheses.}\label{table:CV}
+\caption{\footnotesize Comparison of prediction error in the HPC and InCarMusic. Standard errors are reported in parentheses.}\label{table:CV}
 \end{table}
-
 \vspace{-.4cm}
-\subsection{InCarMusic recommendation system}
-\vspace{-.2cm}
-We apply ordinal tensor completion to a recommendation system {\it InCarMusic}.
-{\it InCarMusic} is a mobile application that offers music recommendation to passengers based on contexts~\cite{baltrunas2011incarmusic}.
-we conduct tensor completion on the $42\times 139\times 26$ ordinal tensor with 2,844 observed entries.
+In the second application, we analyze ordinal tensor which records the ratings of 139 songs on a scale of 1 to 5 from 42 users  on 26 contexts~\citep{baltrunas2011incarmusic}. 
+we conduct tensor completion on the $\tY\in[5]^{42\times139\times26}$ ordinal tensor with only 2,844 observed entries.
  Table~\ref{table:CV} shows the averaged prediction error via 5-fold cross validation.
 The high missing rate  makes the accurate classification challenging.
-Nevertheless, our method achieves the best performance among the three.
+Nevertheless, our method achieves the best performance.
 
@@
 We have developed a low-rank tensor estimation method based on possibly incomplete, ordinal-valued observations.
-A sharp error bound is established, and we demonstrate the outperformance of our approach compared to other methods.
+A sharp error bound is established, and we demonstrate the outperformance of our approach over other alternative methods.
 The work unlocks several directions of future research.
@@
 
-\bibliographystyle{plainnat}
+\bibliographystyle{plain}
 \bibliography{tensor_wang.bib}

### 2020-06-06 paper/ICML/supplements/latexsource/ordinalTv3.tex
Add files via upload

--- 2020-02-18 ICML/submit/supplements/latexsource/ordinalTv3.tex
@@
 % Use the following line for the initial blind version submitted for review:
-\usepackage{icml2020}
 
 % If accepted, instead use the following line for the camera-ready submission:
-%\usepackage[accepted]{icml2020}
+\usepackage[accepted]{icml2020}
 
@@
 \begin{icmlauthorlist}
-\icmlauthor{}{}
+\icmlauthor{Chanwoo Lee}{equal,wi}
+\icmlauthor{Miaoyan Wang}{equal,wi}
 \end{icmlauthorlist}
 
-\icmlaffiliation{}{}
-\icmlcorrespondingauthor{}{}
+\icmlaffiliation{wi}{Department of Statistics, University of Wisconsin - Madison, Wisconsin, USA}
+\icmlcorrespondingauthor{Chanwoo Lee}{chanwoo.lee@wisc.edu}
+\icmlcorrespondingauthor{Miaoyan Wang}{miaoyan.wang@wisc.edu}
+
 % You may provide any keywords that you
@@
 We propose a cumulative link model for higher-order tensors, develop a rank-constrained M-estimator, and obtain theoretical accuracy guarantees.
+We are among the first to establish the recovery theory for both signal tensor and the quantization operators from limited number of highly discrete entries.
 The mean squared error bound is established, and we show that the obtained bound has minimax optimal rate in high dimensions under the low-rank model.
-Our estimator enjoys a faster convergence rate $\tO(d^{-(K-1)/2})$ than $\tO(d^{-K})$ in~\citet{ghadermarzy2018learning}, which is a substantial improvement as the order $K$ increases.
 Furthermore, our proposal serves as an efficient completion algorithm that guarantees consistent recovery of an order-$K$ $(d,\ldots,d)$-dimensional low-rank tensor using only $\tilde \tO(Kd)$ noisy, quantized observations.
@@
 From statistical perspective, our proposed model generalizes the usual binary tensor model while preserving palindromic invariance~\cite{mccullagh1980regression} for ordinal observations.
-From algorithm perspective, our alternating optimization compares favorably to the approximate (non-convex) algorithm developed in the context of binary tensors~\cite{ghadermarzy2018learning}.
-We numerically compare the two approaches in Section~\ref{sec:experiment}.
+From algorithm perspective, we address regularized MLE with alternating optimization (Non-convex).
+The approximated (convex) algorithm is developed in the context of binary tensors~\cite{ghadermarzy2018learning}.
+However, we advocate our approach and provide strong evidence.
+We show the improvement of error bound from  $\tO(d^{-(K-1)/2})$ in~\citet{ghadermarzy2018learning} to $\tO(d^{-(K-1)})$  and numerically compare two approach.
 
@@
 When there are only two categories in the observation space (e.g.
-binary tensors), the cumulative model~\eqref{eq:model} is equivalent to the usual multinomial link model.
+binary tensors), the cumulative model~\eqref{eq:model} is equivalent to the usual binomial link model.
 In general, however, when the number of categories $L\geq 3$, the proportional odds assumption~\eqref{eq:logodd} is more parsimonious, in that, the ordered categories can be envisaged as contiguous intervals on the continuous scale, where the points of division are exactly $b_1<\cdots <b_{L-1}$.
@@
 In the second application, we perform tensor completion to an ordinal dataset with missing values.
-The data tensor records the ratings on a scale of 1 to 5 from 42 users to 139 songs on 26 contexts~\citep{baltrunas2011incarmusic}.
+The data tensor records the ratings of 139 songs on a scale of 1 to 5 from 42 users on 26 contexts~\citep{baltrunas2011incarmusic}.
 
@@
 Based on the estimated Tucker factors $\{\hat \mM_k\}$, we perform a clustering analysis via K-mean on the brain nodes (see detailed procedure in the Supplement).
-We find that the 68 brain nodes are grouped into 11 clusters, and the clustering captures the spatial separation of brain nodes very well (see the Supplement).
+We find that the 68 brain nodes are grouped into 11 clusters, and the clustering captures the spatial separation of brain nodes (see the Supplement).
 In particular, the top three clusters represent the left, right and connection between two hemispheres; the smaller clusters represent local regions driving by similar nodes.
@@
 {\it InCarMusic} is a mobile application that offers music recommendation to passengers of cars based on contexts~\cite{baltrunas2011incarmusic}.
-Our goal is to perform tensor completion to impute the unobserved entries in the $42\times 139\times 26$ ordinal tensor and thereby we can offer context-specific music recommendation to users.
-The data tensor consists of 2,884 observed entries.
+We conduct tensor completion on the $\tY\in[5]^{42\times139\times26}$ with only 2844 observed entries.
 Table~\ref{table:CV} shows the averaged prediction error via 5-fold cross validation.

### 2020-06-06 paper/ICML/supplements/latexsource/ordinalTv4.tex
upload

--- 2020-06-06 paper/ICML/supplements/latexsource/ordinalTv3.tex
@@
 % in order of appearance and this is the preferred way.
-\icmlsetsymbol{equal}{*}
+%\icmlsetsymbol{equal}{*}
 
 \begin{icmlauthorlist}
-\icmlauthor{Chanwoo Lee}{equal,wi}
-\icmlauthor{Miaoyan Wang}{equal,wi}
+\icmlauthor{Chanwoo Lee}{wi}
+\icmlauthor{Miaoyan Wang}{wi}
 \end{icmlauthorlist}
 
-\icmlaffiliation{wi}{Department of Statistics, University of Wisconsin - Madison, Wisconsin, USA}
-\icmlcorrespondingauthor{Chanwoo Lee}{chanwoo.lee@wisc.edu}
+\icmlaffiliation{wi}{Department of Statistics, University of Wisconsin at Madison, Wisconsin, USA}
 \icmlcorrespondingauthor{Miaoyan Wang}{miaoyan.wang@wisc.edu}
@@
 
-%\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
-\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.
+\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
+%\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.
 
 \begin{abstract}
-Higher-order tensors arise frequently in applications such as neuroimaging, recommendation system, social network analysis, and psychological studies.
+Higher-order tensors arise frequently in applications such as neuroimaging, recommendation system, and social network analysis.
 We consider the problem of low-rank tensor estimation from possibly incomplete, ordinal-valued observations.
@@
 \section{Introduction}
-Multidimensional arrays, a.k.a.\ tensors, arise in a variety of applications including recommendation systems~\citep{baltrunas2011incarmusic}, social networks~\citep{nickel2011three}, genomics~\citep{hore2016tensor}, and neuroimaging~\citep{zhou2013tensor}.
+Multidimensional arrays, a.k.a.\ tensors, arise in a variety of applications including recommendation systems~\citep{baltrunas2011incarmusic}, social networks~\citep{nickel2011three}, genomics~\citep{wang2019three}, and neuroimaging~\citep{zhou2013tensor}.
 There is a growing need to develop general methods for analyzing these noisy, high-dimensional datasets that can handle two main problems.
-The problem of tensor denoising -- which aims to recover a signal tensor from its noisy entries -- has gained increased attention in theory and applications~\citep{xia2019sup,zeng2019multiway}.
-The second problem, tensor completion, examines the minimum number of entries needed for a consistent recovery~\citep{ghadermarzy2018learning,ghadermarzy2019near}.
-Low-rankness is often imposed to the signal tensor, Thereby efficiently reducing the intrinsic dimension in both problems.
+The problem of tensor denoising -- which aims to recover a signal tensor from its noisy entries -- has gained increased attention in theory and applications~\citep{hong2018generalized,wang2019multiway}.
+The second problem, tensor completion, examines the minimum number of entries needed for a consistent recovery~\citep{ghadermarzy2018learning,montanari2018spectral}.
+Low-rankness is imposed to the signal tensor, thereby efficiently reducing the intrinsic dimension in both problems.
 
-A number of low-rank tensor estimation methods have been proposed~\citep{kolda2009tensor,acar2010scalable}, revitalizing classical methods such as CANDECOMP/PARAFAC (CP) decomposition~\citep{hitchcock1927expression} and Tucker decomposition~\citep{tucker1966some}.
+A number of low-rank tensor estimation methods have been proposed~\citep{hong2018generalized,wang2017tensor}, revitalizing classical methods such as CANDECOMP/PARAFAC (CP) decomposition~\citep{hitchcock1927expression} and Tucker decomposition~\citep{tucker1966some}.
 These tensor methods treat the entries as continuous-valued.
@@
 \begin{table*}[ht]
-\centerline{\includegraphics[width=16cm]{compare.pdf}}
+\centerline{\includegraphics[width=.98\textwidth]{compare.pdf}}
 \vspace{-.3cm}
@@
 }~\label{tab:compare}
-\vspace{-.5cm}
+\vspace{-.7cm}
 \end{table*}
 
-{\bf Our contribution.} This paper presents an efficient low-rank estimation method and theory for tensors with ordinal-valued entries.
+{\bf Our contributions.} We are among the first to establish the recovery theory for signal tensors and quantization operators simultaneously from a limited number of highly discrete entries.
 Our main contributions are summarized in Table~\ref{tab:compare}.
 We propose a cumulative link model for higher-order tensors, develop a rank-constrained M-estimator, and obtain theoretical accuracy guarantees.
-We are among the first to establish the recovery theory for both signal tensor and the quantization operators from limited number of highly discrete entries.
 The mean squared error bound is established, and we show that the obtained bound has minimax optimal rate in high dimensions under the low-rank model.
-Furthermore, our proposal serves as an efficient completion algorithm that guarantees consistent recovery of an order-$K$ $(d,\ldots,d)$-dimensional low-rank tensor using only $\tilde \tO(Kd)$ noisy, quantized observations.
+Furthermore, our proposal guarantees consistent recovery of an order-$K$ $(d,\ldots,d)$-dimensional low-rank tensor using only $\tilde \tO(Kd)$ noisy, quantized observations.
 
-{\bf Related work.} Our work is related to, but clearly distinctive from, several lines of existing literature.
-Matrix completion from quantized samples was firstly introduced for binary observations~\cite{cai2013max,davenport2014,bhaskar20151} and then extended to ordinal observations~\cite{bhaskar2016probabilistic}.
-As we show in Section~\ref{sec:theory}, applying existing matrix methods to an ordinal tensor results in a suboptimal estimator with a slower convergence rate.
-Therefore, a full exploitation of the tensor structure is necessary; this is the focus of the current paper.
+{\bf Related work.} Our work is connected to non-Gaussian tensor decomposition.
+Existing work focuses exclusively on univariate observations such as binary- or continuous-valued entries~\citep{hong2018generalized,ghadermarzy2018learning}.
+The problem of ordinal/quantized tensor is fundamentally more challenging than previously well-studied tensors for two reasons: (a) the entries do not belong to exponential family distribution, and (b) the observation contains much less information, because neither the underlying signal nor the quantization operator is known.
+The limited information makes the statistical recovery notably hard.
 
-Our work is also connected to non-Gaussian tensor decomposition.
-Existing work focuses exclusively on univariate observations such as binary- or continuous-valued entries~\citep{wang2018learning,hong2018generalized,ghadermarzy2018learning}.
-As we mentioned earlier, the ordinal observations add considerable challenges to the model formulation.
-We address the problems from two perspectives.
-From statistical perspective, our proposed model generalizes the usual binary tensor model while preserving palindromic invariance~\cite{mccullagh1980regression} for ordinal observations.
-From algorithm perspective, we address regularized MLE with alternating optimization (Non-convex).
-The approximated (convex) algorithm is developed in the context of binary tensors~\cite{ghadermarzy2018learning}.
-However, we advocate our approach and provide strong evidence.
-We show the improvement of error bound from  $\tO(d^{-(K-1)/2})$ in~\citet{ghadermarzy2018learning} to $\tO(d^{-(K-1)})$  and numerically compare two approach.
+We address the challenge by proposing a cumulative link model that enjoys palindromic invariance~\cite{mccullagh1980regression}.
+A distinctive non-monotonic, phase-transition pattern is demonstrated (Fig~\ref{fig:finite}b).
+We prove that recovery from quantized tensors achieves equally good information-theoretical convergence as the continuous tensors.
+These results fills the gap between classical and non-classical (ordinal) observations, thereby greatly enriching the tensor-model literature.
 
-\vspace{-.2cm}
+From algorithm perspective, we address the challenge using the (non-convex) alternating algorithm.
+Earlier work has proposed an approximate (convex) algorithm for binary tensor completion~\cite{ghadermarzy2018learning}.
+Unlike matrix problems, convex-relaxation for low-rank tensors suffers from both computational intractability~\cite{hillar2013most} and statistical suboptimality.
+We improve the error bound from $\tO(d^{-(K-1)/2})$ in~\citet{ghadermarzy2018learning} to $\tO(d^{-(K-1)})$ and numerically compare the two approaches.
+
+We also highlight the challenge associated with higher-order tensors.
+Matrix completion has been proposed for binary observations~\cite{cai2013max,davenport2014,bhaskar20151} and for ordinal observations~\cite{bhaskar2016probabilistic}.
+We show that, applying existing matrix methods to higher-order tensors results in suboptimal estimates.
+A full exploitation of the higher-order structure is needed; this is another challenge we address in this paper.
+
 \section{Preliminaries}
-\vspace{-.1cm}
 Let $\tY\in\mathbb{R}^{d_1\times \cdots \times d_K}$ denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional tensor.
 We use $y_\omega$ to denote the tensor entry indexed by $\omega$, where $\omega\in[d_1]\times\cdots\times[d_K]$.
-The Frobenius norm of $\tY$ is defined as $\FnormSize{}{\tY}=\sum_\omega y^2_\omega$ and the infinity norm of $\tY$ is defined as $\mnormSize{}{\tY}=\max_{\omega}|y_\omega|$.
+The Frobenius norm of $\tY$ is defined as $\FnormSize{}{\tY}=\sum_\omega y^2_\omega$ and the infinity norm is defined as $\mnormSize{}{\tY}=\max_{\omega}|y_\omega|$.
 We use $\tY_{(k)}$ to denote the unfolded matrix of size $d_k$-by-$\prod_{i\neq k}d_k$, obtained by reshaping the tensor along the mode-$k$.
-The Tucker rank of $\tY$ is defined as a length-$K$ vector $\mr = (r_1,\ldots,r_K)$, where $r_k$ is the rank of matrix $\tY_{(k)}$ for all $k \in[K]$.
-We say that an event $A$ occurs ``with very high probability'' if $\mathbb{P}(A)$ tends to 1 faster than any polynomial of tensor dimension $d_{\min}=\min\{d_1,\ldots,d_K\} \to\infty$.
+The Tucker rank of $\tY$ is defined as a length-$K$ vector $\mr = (r_1,\ldots,r_K)$, where $r_k$ is the rank of matrix $\tY_{(k)}$ for $k \in[K]$.
 
-We use lower-case letters ($a, b, c, \ldots$) for scalars/vectors, upper-case boldface letters ($\mA, \mB, \mC, \ldots$) for matrices, and calligraphy letters ($\tA,\tB,\tC,\ldots$) for tensors of order three or greater.
-For ease of notation, we allow the basic arithmetic operators (e.g., $\leq, +, -$) to be applied to pairs of tensors in an element-wise manner.
-We use the shorthand $[n]$ to denote the $n$-set $\{1,\ldots,n\}$ for $n \in N_{+}$.
+We use lower-case letters ($a, b, \ldots$) for scalars/vectors, upper-case boldface letters ($\mA, \mB,  \ldots$) for matrices, and calligraphy letters ($\tA,\tB,\ldots$) for tensors of order three or greater.
+An event $A$ is said to occur ``with very high probability'' if $\mathbb{P}(A)$ tends to 1 faster than any polynomial of tensor dimension $d_{\min}=\min\{d_1,\ldots,d_K\} \to\infty$.
+The indicator function of an event $A$ is denoted as $\mathds{1}\{A\}$.
+For ease of notation, we allow basic arithmetic operators (e.g., $\leq, +, -$) to be applied to pairs of tensors in an element-wise manner.
+We use the shorthand $[n]$ to denote $\{1,\ldots,n\}$ for $n \in N_{+}$.
 
-\vspace{-.2cm}
+\vspace{-.1cm}
 \section{Model formulation and motivation}
@@
 This interpretation will be made more explicit in the next section.
+
 \vspace{-.1cm}
@@
 The ordinal tensor model~\eqref{eq:model} with certain types of link $f$ has the equivalent representation as an $L$-level quantization model on $\tY=\entry{y_\omega}$:
-\begin{align}\label{eq:quantization}
-y_\omega&=
-\begin{cases}
-1,& \text{if $y^*_\omega\in(-\infty, b_1]$},\\
-2,& \text{if $y^*_\omega\in(b_1, b_2]$},\\
-\vdots  &\vdots\\
-L,& \text{if $y^*_\omega\in(b_{L-1}, \infty)$},\\
-\end{cases}
-\end{align}
-for all $\omega\in[d_1]\times \cdots \times [d_k]$. Here, $\tY^*=\entry{y^*_\omega}$ is a latent continuous-valued tensor following an additive noise model:
+\begin{equation}\label{eq:quantization}
+y_\omega=\sum_{\ell\in[L]} \ell \mathds{1}\{y^*_\omega \in(b_{\ell-1},b_\ell]\},\ \forall \ \omega\in[d_1]\times \cdots \times [d_k],
+\end{equation}
+%y_\omega&=
+%\begin{cases}
+%1,& \text{if $y^*_\omega\in(-\infty, b_1]$},\\
+%2,& \text{if $y^*_\omega\in(b_1, b_2]$},\\
+%\vdots  &\vdots\\
+%L,& \text{if $y^*_\omega\in(b_{L-1}, \infty)$},\\
+%\end{cases}
+%\end{equation}
+where we define $b_0=-\infty$ and $b_L=\infty$. Here, $\tY^*=\entry{y^*_\omega}$ is a latent continuous-valued tensor following an additive noise model:
 \begin{align}\label{eq:latent}
@@
 The link function is assumed to satisfy:
-\vspace{-.4cm}
+\vspace{-.2cm}
 \begin{enumerate}
 \item $f(\theta)$ is strictly increasing and twice-differentiable in $\theta\in\mathbb{R}/\{0\}$.
-\vspace{-.2cm}
 \item $f'(\theta)$ is strictly log-concave and symmetric with respect to $\theta=0$.
@@
 \end{assumption}
-\vspace{-.1cm}
+
 \subsection{Problem 1: Tensor denoising}~\label{sec:denoising}
@@
 The entry-wise bound assumption is a technical condition that avoids the degeneracy in probability estimation with ordinal observations.
-\vspace{-.1cm}
+
 \subsection{Problem 2: Tensor completion}
-\vspace{-.1cm}
 Motivated by applications in collaborative filtering, we also consider a more general setup when only a subset of tensor entries $y_\omega$ are observed.
@@
 We adopt the same assumption for higher-order tensors.
-\vspace{-.2cm}
+
 \section{Rank-constrained M-estimator}\label{sec:theory}
@@
 In practice, the cut-off points $\mb$ are unknown and should be jointly estimated with $\Theta$.
-For technical convenience, we assume in this section that the cut-off points $\mb$ are known.
-The adaptation of unknown $\mb$ is addressed in Section~\ref{sec:algorithm} and the Supplement.
+We will focus on the theory and algorithm with known $\mb$ in the main paper.
+The adaptation for unknown $\mb$ is addressed in Section~\ref{sec:algorithm} and the Supplement.
 
@@
 Under the Assumption~\ref{ass:link}, all these quantities are strictly positive and independent of tensor dimension.
-\vspace{-.1cm}
+
 \subsection{Estimation error for tensor denoising}\label{sec:denosing}
@@
 In particular, tensor entries with higher sampling probabilities have more influence on the recovery accuracy, compared to the ones with lower sampling probabilities.
-
 \begin{rmk} If we assume each entry is sampled with strictly positive probability; i.e.\ there exits a constant $\mu> 0$ s.t.
@@
 
-\vspace{-.2cm}
 \section{Numerical Implementation}\label{sec:algorithm}
-\vspace{-.1cm}
 We describe the algorithm to seek the optimizer of~\eqref{eq:objective}.
@@
 \caption{Trajectory of objective function with various $d$ and $r$.}
-\vspace{-.5cm}
+\vspace{-.2cm}
 \end{figure}
@@
 \includegraphics[width=8.3cm]{panel.pdf}
-\vspace{-.4cm}
+\vspace{-.1cm}
 \caption{Empirical relationship between (relative) MSE versus (a) dimension $d$, (b) signal level $\alpha$, (c) observation fraction $\rho$, and (d) number of ordinal levels $L$.
@@
 \label{fig:finite}
-\vspace{-.2cm}
+\vspace{-.1cm}
 \end{figure}
 
-\vspace{-.1cm}
 \subsection{Comparison with alternative methods}
-\vspace{-.1cm}
 Next, we compare our ordinal tensor method ({\bf Ordinal-T}) with three popular low-rank methods:
 
-\begin{itemize}[itemsep=-.5pt,topsep=-.5pt,leftmargin=*,partopsep=-.5pt]
+\begin{itemize}[itemsep=0pt,topsep=0pt,leftmargin=*,partopsep=0pt]
 \item Continuous tensor decomposition ({\bf Continuous-T})~\cite{acar2010scalable} is a low-rank approximation method based on classical Tucker model.
@@
 The {\bf Continuous-T} is applied directly to $\tY$ by treating the $L$ levels as continuous observations.
-The {\bf Ordinal-M} is applied to the matrix $\tY_{(1)}$ obtained via 1-mode unfolding.
+The {\bf Ordinal-M} is applied to the 1-mode unfolding matrix $\tY_{(1)}$.
 The {\bf 1bit-T} is applied to $\tY$ in two ways.
@@
 $\tY$ to a binary tensor, by taking the entrywise sign of the mean-adjusted tensor, $\tY- |\Omega|^{-1}\sum_\omega y_\omega$.
-The second approach ({\bf 1bit-category-T}) transforms the order-3 ordinal tensor $\tY$ to an order-4 binary tensor $\tY^{\sharp}=\entry{y^{\sharp}_{ijkl}}$ via dummy variable encoding; i.e., $y^{\sharp}_{ijk\ell}=\mathds{1}_{\{y_{ijk}=\ell\}}$ for all $\ell\in[L-1]$.
-We evaluate the methods by their capabilities in predicting the most likely label for each entry, i.e.,\ $y_\omega^{\text{mode}}=\arg\max_\ell\mathbb{P}(y_\omega=\ell)$.
-Two performance metrics are considered: mean absolute deviation, $\text{MAD}=d^{-K}\sum_\omega |y_\omega^{\text{mode}}-\hat y_\omega^{\text{mode}}|$, and misclassification rate, $\text{MCR}=d^{-K}\sum_\omega\mathds{1}_{\{y_\omega^{\text{mode}}\neq\text{round}(\hat y_\omega^{\text{mode}})\}}$,
-where $\text{round}(\cdot)$ denotes the nearest integer of the prediction (possibly continuous-valued returned by {\bf Continuous-T}).
-Note that MAD penalizes the large deviation more heavily than MCR.
+The second approach ({\bf 1bit-category-T}) transforms the order-3 ordinal tensor $\tY$ to an order-4 binary tensor, $\tY^{\sharp}=\entry{y^{\sharp}_{ijkl}}$, via dummy variable encoding; i.e., $y^{\sharp}_{ijk\ell}=\mathds{1}\{y_{ijk}=\ell\}$ for $\ell\in[L-1]$.
+We evaluate the methods by their capabilities in predicting the most likely labels, $y_\omega^{\text{mode}}=\arg\max_\ell\mathbb{P}(y_\omega=\ell)$.
+Two performance metrics are considered: mean absolute deviation (MAD) $=d^{-K}\sum_\omega |y_\omega^{\text{mode}}-\hat y_\omega^{\text{mode}}|$, and misclassification rate (MCR) $=d^{-K}\sum_\omega\mathds{1}\{y_\omega^{\text{mode}}\neq\text{round}(\hat y_\omega^{\text{mode}})\}$,
+where $\text{round}(\cdot)$ denotes the nearest integer. Note that MAD penalizes the large deviation more heavily than MCR.
 
 Figure~\ref{fig:compare} compares the prediction accuracy under the setting $\alpha=10$, $d=20$, and $r=5$.
-The problem size we considered is comparable to~\citet{ghadermarzy2018learning}.
 We find that our method outperforms the others in both MAD and MCR.
@@
 One possible reason is that our method incorporates the intrinsic ordering among the $L$ levels via proportional odds assumption~\eqref{eq:logodd}, whereas {\bf 1bit-category-T} ignores the ordinal structure and dependence among the induced binary entries.
+
 Figures~\ref{fig:compare}c-d assess the prediction accuracy with sample size.
@@
 }\label{fig:compare}
-\vspace{-.6cm}
+\vspace{-.4cm}
 \end{figure}
@@
 We also compare the methods by their performance in predicting the median label, $y_\omega^{\text{median}}=\min\{\ell\colon\mathbb{P}(y_\omega=\ell)\geq 0.5\}$.
-Under the latent variable model~\eqref{eq:latent} and Assumption~\ref{ass:link}, the median label is the quantized $\Theta$ without noise; i.e.\ $y_\omega^{\text{median}}=\sum_\ell \mathds{1}_{\theta_\omega\in(b_{\ell-1},b_\ell]}$.
-Similar results demonstrate the outperformance of our method (see the Supplement).
+Under the latent variable model~\eqref{eq:latent} and Assumption~\ref{ass:link}, the median label is the quantized $\Theta$ without noise; i.e.\ $y_\omega^{\text{median}}=\sum_\ell \mathds{1}\{\theta_\omega\in(b_{\ell-1},b_\ell]\}$.
+Similar results show the outperformance of our method (see the Supplement).
 
-\vspace{-.1cm}
 \section{Data Applications}
-\vspace{-.1cm}
 We apply our ordinal tensor method to two real-world datasets.
@@
 
-\vspace{-.1cm}
 \subsection{InCarMusic recommendation system}
-\vspace{-.1cm}
 We apply ordinal tensor completion to a recommendation system {\it InCarMusic}.
 {\it InCarMusic} is a mobile application that offers music recommendation to passengers of cars based on contexts~\cite{baltrunas2011incarmusic}.
-We conduct tensor completion on the $\tY\in[5]^{42\times139\times26}$ with only 2844 observed entries.
+We conduct tensor completion on the $\tY\in[5]^{42\times139\times26}$ with 2,844 observed entries only.
 Table~\ref{table:CV} shows the averaged prediction error via 5-fold cross validation.
@@
 
-\vspace{-.2cm}
 \section{Conclusions}
-\vspace{-.1cm}
 We have developed a low-rank tensor estimation method based on possibly incomplete, ordinal-valued observations.
@@
 Other directions include the trade-off between (non)convex optimization and statistical/computational efficiency.
-While convex relaxations are popular approach for matrix-based problem, they are often slow in practice~\cite{chen2019non}.
+While we have provided numerical evidence for the success of nonconvex approach, the full landscape of the optimization remains open.
 The interplay between computational efficiency and statistical accuracy in general tensor problems warrants future research.
@@
 \newpage
-\bibliography{ordinalTv3.bbl}
+\bibliography{tensor_wang.bib}
 \bibliographystyle{icml2020}

### 2020-06-06 paper/ICML/supplements/latexsource/ordinalTv4.tex
update

@@
 
-\icmlaffiliation{wi}{Department of Statistics, University of Wisconsin at Madison, Wisconsin, USA}
+\icmlaffiliation{wi}{Department of Statistics, University of Wisconsin-Madison, Wisconsin, USA}
 \icmlcorrespondingauthor{Miaoyan Wang}{miaoyan.wang@wisc.edu}
@@
 For example, the Netflix problem records the ratings of users on movies over time.
-Each data is a rating on a nominal scale \{{\it very like, like, neutral, dislike, very dislike}\}.
+Each data entry is a rating on a nominal scale \{{\it very like, like, neutral, dislike, very dislike}\}.
 Another example is in the signal processing, where the digits are frequently rounded or truncated so that only integer values are available.

### 2020-06-06 paper/ICML/supplements/latexsource/ordinalTv4.tex
update

@@
 \section{Introduction}
-Multidimensional arrays, a.k.a.\ tensors, arise in a variety of applications including recommendation systems~\citep{baltrunas2011incarmusic}, social networks~\citep{nickel2011three}, genomics~\citep{wang2019three}, and neuroimaging~\citep{zhou2013tensor}.
-There is a growing need to develop general methods for analyzing these noisy, high-dimensional datasets that can handle two main problems.
+Multidimensional arrays, a.k.a.\ tensors, arise in wide applications including recommendation systems~\citep{baltrunas2011incarmusic}, social networks~\citep{nickel2011three}, genomics~\citep{wang2019three}, and neuroimaging~\citep{zhou2013tensor}.
+There is a growing need to develop general methods that can address two main problems for analyzing these noisy, high-dimensional datasets.
 The problem of tensor denoising -- which aims to recover a signal tensor from its noisy entries -- has gained increased attention in theory and applications~\citep{hong2018generalized,wang2019multiway}.
 The second problem, tensor completion, examines the minimum number of entries needed for a consistent recovery~\citep{ghadermarzy2018learning,montanari2018spectral}.
-Low-rankness is imposed to the signal tensor, thereby efficiently reducing the intrinsic dimension in both problems.
+Low-rankness is often imposed to the signal tensor, thereby efficiently reducing the intrinsic dimension in both problems.
 

### 2020-06-06 paper/ICML/supplements/latexsource/ordinalTv4.tex
update

--- 2020-06-06 paper/ICML/supplements/latexsource/ordinalTv4.tex
@@
 Multidimensional arrays, a.k.a.\ tensors, arise in a variety of applications including recommendation systems~\citep{baltrunas2011incarmusic}, social networks~\citep{nickel2011three}, genomics~\citep{wang2019three}, and neuroimaging~\citep{zhou2013tensor}.
-There is a growing need to develop general methods for analyzing these noisy, high-dimensional datasets that can handle two main problems.
-The problem of tensor denoising -- which aims to recover a signal tensor from its noisy entries -- has gained increased attention in theory and applications~\citep{hong2018generalized,wang2019multiway}.
-The second problem, tensor completion, examines the minimum number of entries needed for a consistent recovery~\citep{ghadermarzy2018learning,montanari2018spectral}.
-Low-rankness is imposed to the signal tensor, thereby efficiently reducing the intrinsic dimension in both problems.
+There is a growing need to develop general methods that can address two main problems for analyzing these noisy, high-dimensional datasets.
+The first problem is tensor denoising which aims to recover a signal tensor from its noisy entries~\citep{hong2018generalized,wang2019multiway}.
+The second problem is tensor completion which examines the minimum number of entries needed for a consistent recovery~\citep{ghadermarzy2018learning,montanari2018spectral}.
+Low-rankness is often imposed to the signal tensor, thereby efficiently reducing the intrinsic dimension in both problems.
 
@@
 \begin{enumerate}
-\item $f(\theta)$ is strictly increasing and twice-differentiable in $\theta\in\mathbb{R}/\{0\}$.
+\item $f(\theta)$ is strictly increasing and twice-differentiable in $\theta\in \mathbb{R}$.
 \item $f'(\theta)$ is strictly log-concave and symmetric with respect to $\theta=0$.
@@
 \end{equation}
-where $c_1 >0$ is the same constants as in Theorem~\ref{thm:rate}.
+where $c_1 >0$ is the same constant as in Theorem~\ref{thm:rate}.
 \end{cor}
@@
 We describe the algorithm to seek the optimizer of~\eqref{eq:objective}.
-In practice, the cut-off points $\mb$ are often unknown, so we choose to maximize $\logl$ jointly over $(\Theta,\mb)\in\tP\times \tB$.
-The objective $\logl$ is concave in $(\Theta,\mb)$ whenever $f'$ is log-concave (see the Supplement).
+In practice, the cut-off points $\mb$ are often unknown, so we choose to maximize $\logl$ jointly over $(\Theta,\mb)\in\tP\times \tB$ (see the Supplement for details).
+The objective $\logl$ is concave in $(\Theta,\mb)$ whenever $f'$ is log-concave.
 However, the feasible set $\tP$ is non-convex, which makes the optimization~\eqref{eq:objective} a non-convex problem.
@@
 
-We also compare the methods by their performance in predicting the median label, $y_\omega^{\text{median}}=\min\{\ell\colon\mathbb{P}(y_\omega=\ell)\geq 0.5\}$.
-Under the latent variable model~\eqref{eq:latent} and Assumption~\ref{ass:link}, the median label is the quantized $\Theta$ without noise; i.e.\ $y_\omega^{\text{median}}=\sum_\ell \mathds{1}\{\theta_\omega\in(b_{\ell-1},b_\ell]\}$.
+We also compare the methods by their performance in predicting the median labels, $y_\omega^{\text{median}}=\min\{\ell\colon\mathbb{P}(y_\omega=\ell)\geq 0.5\}$.
+Under the latent variable model~\eqref{eq:latent} and Assumption~\ref{ass:link}, the median label is the quantized $\theta_\omega$ without noise; i.e.\ $y_\omega^{\text{median}}=\sum_\ell \mathds{1}\{\theta_\omega\in(b_{\ell-1},b_\ell]\}$.
 Similar results show the outperformance of our method (see the Supplement).
@@
 We find that the 68 brain nodes are grouped into 11 clusters, and the clustering captures the spatial separation of brain nodes (see the Supplement).
-In particular, the top three clusters represent the left, right and connection between two hemispheres; the smaller clusters represent local regions driving by similar nodes.
+In particular, the top three clusters represent the connections in the left-, right-, and inter-hemispheres; the smaller clusters represent local regions driving by similar nodes.
 For example, the cluster IV consists of nodes in the left supramarginal gyrus region.
@@
 {\it InCarMusic} is a mobile application that offers music recommendation to passengers of cars based on contexts~\cite{baltrunas2011incarmusic}.
-We conduct tensor completion on the $\tY\in[5]^{42\times139\times26}$ with 2,844 observed entries only.
+We conduct tensor completion on the 42-by-139-by-26 tensor with 2,844 observed entries only.
 Table~\ref{table:CV} shows the averaged prediction error via 5-fold cross validation.

### 2020-06-06 paper/ICML/supplements/latexsource/ordinalTv4.tex
update
typo

@@
 \begin{table*}[ht]
-\centerline{\includegraphics[width=.98\textwidth]{compare.pdf}}
+\centerline{\includegraphics[width=1\textwidth]{compare.pdf}}
 \vspace{-.3cm}
@@
 Note that $\tilde \tO(Kd)$ roughly matches the degree of freedom for an order-$K$ tensor of fixed rank $\mr$, suggesting the optimality of our sample requirement.
-This sample complexity substantially improves over earlier result $\tO(d^{\lceil K/2\rceil})$ based on square matricization~\cite{mu2014square}, or $\tO(d^{N/2})$ based on tensor nuclear-norm regularization~\cite{yuan2016tensor}.
+This sample complexity substantially improves over earlier result $\tO(d^{\lceil K/2\rceil})$ based on square matricization~\cite{mu2014square}, or $\tO(d^{K/2})$ based on tensor nuclear-norm regularization~\cite{yuan2016tensor}.
 Existing methods that achieve $\tilde \tO(Kd)$ sample complexity require either a deterministic cross sampling design~\cite{zhang2019cross} or univariate measurements~\cite{ghadermarzy2018learning}.

### 2020-06-06 paper/ICML/supplements/latexsource/ordinalTv4.tex
typo

@@
 \end{thm}
-We see that the lower bound matches the upper bound in~\eqref{eq:rate} on the polynomial order of tensor dimension. Therefore, our estimator~\eqref{eq:estimator} is order-optimal.
+We see that the lower bound matches the upper bound in~\eqref{eq:rate} on the polynomial order of tensor dimension. Therefore, our estimator~\eqref{eq:estimator} is rate-optimal.
 
@@
 The work unlocks several directions of future research.
-One interesting question would be the inference problem, i.e..
-to assess the uncertainty of the obtained estimates and the imputation.
+One interesting question would be the inference problem, i.e.,\ to assess the uncertainty of the obtained estimates and the imputation.
 Other directions include the trade-off between (non)convex optimization and statistical/computational efficiency.

### 2020-06-08 paper/ICML/supplements/latexsource/ordinalTv4.tex
add funding info

@@
 \section{Experiments}\label{sec:experiment}
-In this section, we evaluate the empirical performance of our method.
+In this section, we evaluate the empirical performance of our methods\footnote{Software:  https://CRAN.R-project.org/package=tensorordinal}.
 We investigate both the complete and the incomplete settings, and compare the recovery accuracy with other tensor-based methods.
@@
 \includegraphics[width=8.3cm]{panel.pdf}
-\vspace{-.1cm}
+\vspace{-.2cm}
 \caption{Empirical relationship between (relative) MSE versus (a) dimension $d$, (b) signal level $\alpha$, (c) observation fraction $\rho$, and (d) number of ordinal levels $L$.
@@
 \label{fig:finite}
-\vspace{-.1cm}
+\vspace{-.3cm}
 \end{figure}
@@
 In particular, methods built on multi-level observations ({\bf Ordinal-T}, {\bf Ordinal-M}, {\bf 1bit-category-T}) exhibit stable MCR over $\rho$ and $L$, whereas the others two methods ({\bf Continuous-T}, {\bf 1bit-sign-T}) generally fail except for $L=2$ (Figures~\ref{fig:compare}a-b).
-This observation highlights the necessity of modeling multi-level probabilities in classification task.
-Interestingly, although both {\bf 1bit-category-T} and our method {\bf Ordinal-T} behave similarly for binary tensors ($L=2$), the improvement of our method is substantial as $L$ increases (Figures~\ref{fig:compare}a and~\ref{fig:compare}c).
+This observation highlights the benefits of multi-level modeling in the classification task.
+Although {\bf 1bit-category-T} and our method {\bf Ordinal-T} behave similarly for binary tensors ($L=2$), the improvement of our method is substantial as $L$ increases (Figures~\ref{fig:compare}a and~\ref{fig:compare}c).
 One possible reason is that our method incorporates the intrinsic ordering among the $L$ levels via proportional odds assumption~\eqref{eq:logodd}, whereas {\bf 1bit-category-T} ignores the ordinal structure and dependence among the induced binary entries.
@@
 \newpage
+\section*{Acknowledgements}
+This research is supported in part by NSF grant DMS-1915978 and the University of Wisconsin-Madison, Office of the Vice Chancellor for Research and Graduate Education with funding from the Wisconsin Alumni Research Foundation.
+
 \bibliography{tensor_wang.bib}

### 2020-06-09 paper/ICML/supplements/latexsource/ordinalTv4.tex
remove space adjustment

@@
 
-\vspace{-.1cm}
 \section{Model formulation and motivation}
-\vspace{-.1cm}
 \subsection{Observation model}
@@
 \end{equation}
-where $\mb=(b_1,\ldots,b_{L-1})$ is a set of unknown scalars satisfying $b_1<\cdots <b_{L-1}$, $\Theta=\entry{\theta_\omega}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is a continuous-valued parameter tensor satisfying certain low-dimensional structure (to be specified later), and $f(\cdot):\mathbb{R}\mapsto[0,1]$ is a known, strictly increasing function.
+where $\mb=(b_1,\ldots,b_{L-1})$ is a set of unknown scalars satisfying $b_1<\cdots <b_{L-1}$, $\Theta=\entry{\theta_\omega}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is a continuous-valued parameter tensor satisfying certain low-dimensional structure (to be specified later), and $f(\cdot)\colon \mathbb{R}\mapsto[0,1]$ is a known, strictly increasing function.
 We refer to $\mb$ as the cut-off points and $f$ the link function.
@@
 
-\vspace{-.1cm}
 \subsection{Latent-variable interpretation}\label{sec:latent}
-\vspace{-.1cm}
 
@@
 \subsection{Problem 1: Tensor denoising}~\label{sec:denoising}
-\vspace{-.1cm}
 The first question we aim to address is tensor denoising:
@@
 One interesting question would be the inference problem, i.e.,\ to assess the uncertainty of the obtained estimates and the imputation.
-Other directions include the trade-off between (non)convex optimization and statistical/computational efficiency.
+Other directions include the trade-off between (non)convex optimization and statistica efficiency.
 While we have provided numerical evidence for the success of nonconvex approach, the full landscape of the optimization remains open.

### 2020-06-14 paper/Camera ready/ordinalTv6.tex
upload

--- 2020-02-16 ICML/arxiv/ordinalT_arxiv_v4.tex
@@
-\documentclass[11pt]{article}
-
-\usepackage{fancybox}
-\usepackage{color}
-\usepackage{url}
-\usepackage[margin=1in]{geometry}
+%%%%%%%% ICML 2020 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
+
+\documentclass{article}
+
+% Recommended, but optional, packages for figures and better typesetting:
+\usepackage{microtype}
+\usepackage{graphicx}
+\usepackage{subfigure}
+\usepackage{booktabs} % for professional tables
+
+% hyperref makes hyperlinks in the resulting PDF.
+% If your build breaks (sometimes temporarily if a hyperlink spans a page)
+% please comment out the following usepackage line and replace
+% \usepackage{icml2020} with \usepackage[nohyperref]{icml2020} above.
+\usepackage{hyperref}
+
+% Attempt to make hyperref and algorithmic work together better:
 \newcommand{\theHalgorithm}{\arabic{algorithm}}
-\usepackage{comment}
-\usepackage{amsmath,amssymb,amsthm,bm,mathtools}
-\usepackage{natbib}
-\usepackage{dsfont,multirow,hyperref,setspace,enumerate}
-\hypersetup{colorlinks,linkcolor={blue},citecolor={blue},urlcolor={red}}
-\usepackage{algorithm}
-\usepackage{algorithmic}
-\usepackage{makecell}
-
-\setlength\parindent{0pt}
-\usepackage[parfill]{parskip}
-
+
+% Use the following line for the initial blind version submitted for review:
+
+% If accepted, instead use the following line for the camera-ready submission:
+\usepackage[accepted]{icml2020}
+
+% The \icmltitle you define below is probably too long as a header.
+% Therefore, a short form for the running title is supplied here:
+\icmltitlerunning{Ordinal tensor denoising and completion}
+\usepackage{wrapfig}
+\usepackage{multirow}
+\usepackage{graphicx}
+%\usepackage[utf8]{inputenc} % allow utf-8 input
+%\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
+\usepackage{hyperref}       % hyperlinks
+\usepackage{url}            % simple URL typesetting
+%\usepackage{booktabs}       % professional-quality tables
+\usepackage{amsmath,amssymb}
+\usepackage{amsthm}    % blackboard math symbols
+%\usepackage{nicefrac}       % compact symbols for 1/2, etc.
+%\usepackage{microtype}      % microtypography
+\usepackage{bm}
+%\usepackage{subfig}
+%\usepackage[english]{babel}
+%\usepackage{algorithm}
+%\usepackage{appendix}
+\usepackage{mathtools}
+\mathtoolsset{showonlyrefs}
+\usepackage{enumitem}
 \theoremstyle{plain}
@@
 
-\usepackage{pifont}
-\newcommand{\cmark}{\ding{51}}%
-\newcommand{\xmark}{\ding{55}}%
-
 \usepackage{dsfont}
-\usepackage{wrapfig}
-\usepackage{mathtools}
-\mathtoolsset{showonlyrefs}
+%\usepackage{algpseudocode,algorithm}
+%\algnewcommand\algorithmicinput{\textbf{Input:}}
+%\algnewcommand\algorithmicoutput{\textbf{Output:}}
+%\algnewcommand\INPUT{\item[\algorithmicinput]}
+%\algnewcommand\OUTPUT{\item[\algorithmicoutput]}
+%\DeclareMathOperator*{\minimize}{minimize}
 
@@
 }
-
-\begingroup
-\makeatletter
-\@for\theoremstyle:=definition,remark,plain\do{%
-\expandafter\g@addto@macro\csname th@\theoremstyle\endcsname{%
-\addtolength\thm@preskip\parskip
-}%
-}
-\endgroup
-
+\usepackage{makecell}
 \input macros.tex
-\allowdisplaybreaks
-
-\title{Tensor denoising and completion based on ordinal observations}
-\date{}
-\author{%
-Chanwoo Lee \\
-University of Wisconsin -- Madison\\
-\texttt{chanwoo.lee@wisc.edu} \\
-\and
-Miaoyan Wang \\
-University of Wisconsin -- Madison\\
-\texttt{miaoyan.wang@wisc.edu} \\
-}
+
+\usepackage{amssymb}
+\usepackage{pifont}
+\newcommand{\cmark}{\ding{51}}%
+\newcommand{\xmark}{\ding{55}}%
 
 \begin{document}
-\maketitle
+
+\twocolumn[
+\icmltitle{Tensor denoising and completion based on ordinal observations}
+
+% It is OKAY to include author information, even for blind
+% submissions: the style file will automatically remove it for you
+% unless you've provided the [accepted] option to the icml2020
+% package.
+
+% List of affiliations: The first argument should be a (short)
+% identifier you will use later to specify author affiliations
+% Academic affiliations should list Department, University, City, Region, Country
+% Industry affiliations should list Company, City, Region, Country
+
+% You can specify symbols, otherwise they are numbered in order.
+% Ideally, you should not use this facility. Affiliations will be numbered
+% in order of appearance and this is the preferred way.
+%\icmlsetsymbol{equal}{*}
+
+\begin{icmlauthorlist}
+\icmlauthor{Chanwoo Lee}{wi}
+\icmlauthor{Miaoyan Wang}{wi}
+\end{icmlauthorlist}
+
+\icmlaffiliation{wi}{Department of Statistics, University of Wisconsin-Madison, Wisconsin, USA}
+\icmlcorrespondingauthor{Miaoyan Wang}{miaoyan.wang@wisc.edu}
+
+% You may provide any keywords that you
+% find helpful for describing your paper; these are used to populate
+% the "keywords" metadata in the PDF but will not be shown in the document
+\icmlkeywords{Higher-order tensors, ordinal observation, tensor decomposition, tensor completion}
+
+\vskip 0.3in
+]
+
+% this must go after the closing bracket ] following \twocolumn[ ...
+
+% This command actually creates the footnote in the first column
+% listing the affiliations and the copyright notice.
+% The command takes one argument, which is text to display at the start of the footnote.
+% The \icmlEqualContribution command is standard text for equal contribution.
+% Remove it (just {}) if you do not need this facility.
+
+\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
+%\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.
+
 \begin{abstract}
-Higher-order tensors arise frequently in applications such as neuroimaging, recommendation system, social network analysis, and psychological studies.
+Higher-order tensors arise frequently in applications such as neuroimaging, recommendation system, and social network analysis.
 We consider the problem of low-rank tensor estimation from possibly incomplete, ordinal-valued observations.
@@
 We demonstrate the outperformance of our approach over previous methods on the tasks of clustering and collaborative filtering.
-\\[0.5cm]
-\textbf{Keywords:} Higher-order tensors, ordinal observation, tensor decomposition, tensor completion
+
 \end{abstract}
@@
 \section{Introduction}
-Multidimensional arrays, a.k.a.\ tensors, arise in a variety of applications including recommendation systems~\citep{baltrunas2011incarmusic}, social networks~\citep{nickel2011three}, genomics~\citep{hore2016tensor}, and neuroimaging~\citep{zhou2013tensor}.
-There is a growing need to develop general methods that can handle two main problems for analyzing these noisy, high-dimensional datasets.
-The first problem is tensor denoising which aims to recover a signal tensor from its noisy entries~\citep{hong2018generalized,zeng2019multiway}.
-The second problem is tensor completion which examines the minimum number of entries needed for a consistent recovery~\citep{ghadermarzy2018learning,ghadermarzy2019near}.
+Multidimensional arrays, a.k.a.\ tensors, arise in a variety of applications including recommendation systems~\citep{baltrunas2011incarmusic}, social networks~\citep{nickel2011three}, genomics~\citep{wang2019three}, and neuroimaging~\citep{zhou2013tensor}.
+There is a growing need to develop general methods that can address two main problems for analyzing these noisy, high-dimensional datasets.
+The first problem is tensor denoising which aims to recover a signal tensor from its noisy entries~\citep{hong2020generalized,wang2019multiway}.
+The second problem is tensor completion which examines the minimum number of entries needed for a consistent recovery~\citep{ghadermarzy2018learning,montanari2018spectral}.
 Low-rankness is often imposed to the signal tensor, thereby efficiently reducing the intrinsic dimension in both problems.
 
-A number of low-rank tensor estimation methods have been proposed~\citep{anandkumar2014tensor,wang2017tensor}, revitalizing classical methods such as CANDECOMP/PARAFAC (CP) decomposition~\citep{hitchcock1927expression} and Tucker decomposition~\citep{tucker1966some}.
+A number of low-rank tensor estimation methods have been proposed~\citep{hong2020generalized,wang2017tensor}, revitalizing classical methods such as CANDECOMP/PARAFAC (CP) decomposition~\citep{hitchcock1927expression} and Tucker decomposition~\citep{tucker1966some}.
 These tensor methods treat the entries as continuous-valued.
@@
 For example, the Netflix problem records the ratings of users on movies over time.
-Each entry is a rating on a nominal scale \{{\it very like, like, neutral, dislike, very dislike}\}.
+Each data entry is a rating on a nominal scale \{{\it very like, like, neutral, dislike, very dislike}\}.
 Another example is in the signal processing, where the digits are frequently rounded or truncated so that only integer values are available.
@@
 Second, the parameter interpretations should be consistent under merging or splitting of contiguous categories.
-The classical continuous tensor model~\citep{kolda2009tensor, ghadermarzy2019near} fails in the first aspect, whereas the binary tensor model~\citep{ghadermarzy2018learning} lacks the second property.
+The classical continuous tensor model~\cite{kolda2009tensor, ghadermarzy2019near} fails in the first aspect, whereas the binary tensor model~\cite{ghadermarzy2018learning} lacks the second property.
 An appropriate model for ordinal tensors has yet to be studied.
 
-{\bf Our contribution.} This paper presents an efficient low-rank estimation method and theory for tensors with ordinal-valued entries.
+\begin{table*}[ht]
+\centerline{\includegraphics[width=1\textwidth]{compare.pdf}}
+\vspace{-.3cm}
+\caption{Comparison with previous work.
+For ease of presentation, we summarize the error rate and sample complexity assuming equal tensor dimension in all modes.
+$K$: tensor order; $L$: number of ordinal levels; $d$: dimension at each mode.
+}~\label{tab:compare}
+\vspace{-.7cm}
+\end{table*}
+
+{\bf Our contributions.} We are among the first to establish the recovery theory for signal tensors and quantization operators simultaneously from a limited number of highly discrete entries.
 Our main contributions are summarized in Table~\ref{tab:compare}.
@@
 The mean squared error bound is established, and we show that the obtained bound has minimax optimal rate in high dimensions under the low-rank model.
-Our estimator enjoys a faster convergence rate $\tO(d^{-(K-1)/2})$ than $\tO(d^{-K})$ in~\citet{ghadermarzy2018learning}, which is a substantial improvement for higher-order tensors.
-Furthermore, our proposal serves as an efficient completion algorithm that guarantees consistent recovery of an order-$K$ $(d,\ldots,d)$-dimensional low-rank tensor using only $\tilde \tO(Kd)$ noisy, quantized observations.
-
-\begin{table}[ht]
-\resizebox{\columnwidth}{!}{
-\begin{tabular}{l|ccc}
-&\cite{bhaskar2016probabilistic}&\cite{ghadermarzy2018learning} &This paper\\
-\hline
-Higher-order tensors ($K\geq 3$) & \xmark&\cmark& \cmark\\
-\hline
-Multi-level categories ($L\geq 3$)& \cmark& \xmark&\cmark\\
- \hline
- Error rate for tensor denoising&$d^{-1}$ for $K=2$& $d^{-(K-1)/2}$&$d^{-(K-1)}$\\
-\hline
-Optimality guarantee under low-rank models& unknown & \xmark & \cmark\\
-\hline
-Sample complexity for tensor completion&$d^K$& $Kd$&$Kd$\\
-\hline
-\end{tabular}
-}
-\caption{Comparison with previous work when tensor rank $r=O(1)$.
-For ease of presentation, we summarize the error rate and sample complexity (neglecting log factors) when the tensor dimensions are equal in all modes.
-$K$: tensor order; $L$: number of ordinal levels; $d$: dimension at each mode.
-}~\label{tab:compare}
-\end{table}
-
-{\bf Related work.} Our work is related to, but clearly distinctive from, several lines of existing literature.
-Matrix completion from quantized samples was firstly introduced for binary observations~\citep{cai2013max,davenport2014,bhaskar20151} and then extended to ordinal observations~\citep{bhaskar2016probabilistic}.
-As we show in Section~\ref{sec:theory}, applying existing matrix methods to an ordinal tensor results in a suboptimal estimator with a slower convergence rate.
-Therefore, a full exploitation of the tensor structure is necessary; this is the focus of the current paper.
-
-Our work is also connected to non-Gaussian tensor decomposition.
-Existing work focuses exclusively on univariate observations such as binary- or continuous-valued entries~\citep{wang2018learning,hong2018generalized,ghadermarzy2018learning}.
-As we mentioned earlier, the ordinal observations add considerable challenges to the model formulation.
-We address the problems from two perspectives.
-From statistical perspective, our proposed model generalizes the usual binary tensor model while preserving palindromic invariance~\citep{mccullagh1980regression} for ordinal observations.
-From algorithm perspective, our alternating optimization compares favorably to the approximate (non-convex) algorithm developed in the context of binary tensors~\citep{ghadermarzy2018learning}.
-We numerically compare the two approaches in Section~\ref{sec:experiment}.
+Furthermore, our proposal guarantees consistent recovery of an order-$K$ $(d,\ldots,d)$-dimensional low-rank tensor using only $\tilde \tO(Kd)$ noisy, quantized observations.
+
+{\bf Related work.} Our work is connected to non-Gaussian tensor decomposition.
+Existing work focuses exclusively on univariate observations such as binary- or continuous-valued entries~\citep{wang2018learning,ghadermarzy2018learning}.
+The problem of ordinal/quantized tensor is fundamentally more challenging than previously well-studied tensors for two reasons: (a) the entries do not belong to exponential family distribution, and (b) the observation contains much less information, because neither the underlying signal nor the quantization operator is known.
+The limited information makes the statistical recovery notably hard.
+
+We address the challenge by proposing a cumulative link model that enjoys palindromic invariance~\cite{mccullagh1980regression}.
+A distinctive non-monotonic, phase-transition pattern is demonstrated (Fig~\ref{fig:finite}b).
+We prove that recovery from quantized tensors achieves equally good information-theoretical convergence as the continuous tensors.
+These results fills the gap between classical and non-classical (ordinal) observations, thereby greatly enriching the tensor-model literature.
+
+From algorithm perspective, we address the challenge using the (non-convex) alternating algorithm.
+Earlier work has proposed an approximate (convex) algorithm for binary tensor completion~\cite{ghadermarzy2018learning}.
+Unlike matrix problems, convex-relaxation for low-rank tensors suffers from both computational intractability~\cite{hillar2013most} and statistical suboptimality.
+We improve the error bound from $\tO(d^{-(K-1)/2})$ in~\citet{ghadermarzy2018learning} to $\tO(d^{-(K-1)})$ and numerically compare the two approaches.
+
+We also highlight the challenge associated with higher-order tensors.
+Matrix completion has been proposed for binary observations~\cite{cai2013max,davenport2014,bhaskar20151} and for ordinal observations~\cite{bhaskar2016probabilistic}.
+We show that, applying existing matrix methods to higher-order tensors results in suboptimal estimates.
+A full exploitation of the higher-order structure is needed; this is another challenge we address in this paper.
 
@@
 We use $y_\omega$ to denote the tensor entry indexed by $\omega$, where $\omega\in[d_1]\times\cdots\times[d_K]$.
-The Frobenius norm of $\tY$ is defined as $\FnormSize{}{\tY}=\sum_\omega y^2_\omega$ and the infinity norm of $\tY$ is defined as $\mnormSize{}{\tY}=\max_{\omega}|y_\omega|$.
+The Frobenius norm of $\tY$ is defined as $\FnormSize{}{\tY}=\sum_\omega y^2_\omega$ and the infinity norm is defined as $\mnormSize{}{\tY}=\max_{\omega}|y_\omega|$.
 We use $\tY_{(k)}$ to denote the unfolded matrix of size $d_k$-by-$\prod_{i\neq k}d_k$, obtained by reshaping the tensor along the mode-$k$.
-The Tucker rank of $\tY$ is defined as a length-$K$ vector $\mr = (r_1,\ldots,r_K)$, where $r_k$ is the rank of matrix $\tY_{(k)}$ for all $k \in[K]$.
-We say that an event $A$ occurs ``with very high probability'' if $\mathbb{P}(A)$ tends to 1 faster than any polynomial of tensor dimension $d_{\min}=\min\{d_1,\ldots,d_K\} \to\infty$.
-For any two functions $f, g$ depending on $(d_1,\ldots,d_K)$, we write $f=\tO(g)$ to indicate that $f\leq Cg$, where $C>0$ is a constant independent of tensor dimension.
-We write $f=\tilde\tO(g)$ to indicate that $f\leq C(\log d_{\min})^\beta g$ for some $\beta>0$.
-
-We use lower-case letters ($a, b, c, \ldots$) for scalars/vectors, upper-case boldface letters ($\mA, \mB, \mC, \ldots$) for matrices, and calligraphy letters ($\tA,\tB,\tC,\ldots$) for tensors of order three or greater.
-For ease of notation, we allow the basic arithmetic operators (e.g., $\leq, +, -$) to be applied to pairs of tensors in an element-wise manner.
-We use the shorthand $[n]$ to denote the $n$-set $\{1,\ldots,n\}$ for $n \in N_{+}$.
+The Tucker rank of $\tY$ is defined as a length-$K$ vector $\mr = (r_1,\ldots,r_K)$, where $r_k$ is the rank of matrix $\tY_{(k)}$ for $k \in[K]$.
+
+We use lower-case letters ($a, b, \ldots$) for scalars/vectors, upper-case boldface letters ($\mA, \mB,  \ldots$) for matrices, and calligraphy letters ($\tA,\tB,\ldots$) for tensors of order three or greater.
+An event $A$ is said to occur ``with very high probability'' if $\mathbb{P}(A)$ tends to 1 faster than any polynomial of tensor dimension $d_{\min}=\min\{d_1,\ldots,d_K\} \to\infty$.
+The indicator function of an event $A$ is denoted as $\mathds{1}\{A\}$.
+For ease of notation, we allow basic arithmetic operators (e.g., $\leq, +, -$) to be applied to pairs of tensors in an element-wise manner.
+We use the shorthand $[n]$ to denote $\{1,\ldots,n\}$ for $n \in N_{+}$.
 
@@
 \subsection{Observation model}
+\vspace{-.1cm}
 Let $\tY$ denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional data tensor.
@@
 \subsection{Latent-variable interpretation}\label{sec:latent}
+
 The ordinal tensor model~\eqref{eq:model} with certain types of link $f$ has the equivalent representation as an $L$-level quantization model on $\tY=\entry{y_\omega}$:
-\begin{align}\label{eq:quantization}
-y_\omega&=
-\begin{cases}
-1,& \text{if $y^*_\omega\in(-\infty, b_1]$},\\
-2,& \text{if $y^*_\omega\in(b_1, b_2]$},\\
-\vdots  &\vdots\\
-L,& \text{if $y^*_\omega\in(b_{L-1}, \infty)$},\\
-\end{cases}
-\end{align}
-for all $\omega\in[d_1]\times \cdots \times [d_k]$. Here, $\tY^*=\entry{y^*_\omega}$ is a latent continuous-valued tensor following an additive noise model:
+\begin{equation}\label{eq:quantization}
+y_\omega=\sum_{\ell\in[L]} \ell \mathds{1}\{y^*_\omega \in(b_{\ell-1},b_\ell]\},\ \forall \ \omega\in[d_1]\times \cdots \times [d_k],
+\end{equation}
+%y_\omega&=
+%\begin{cases}
+%1,& \text{if $y^*_\omega\in(-\infty, b_1]$},\\
+%2,& \text{if $y^*_\omega\in(b_1, b_2]$},\\
+%\vdots  &\vdots\\
+%L,& \text{if $y^*_\omega\in(b_{L-1}, \infty)$},\\
+%\end{cases}
+%\end{equation}
+where we define $b_0=-\infty$ and $b_L=\infty$. Here, $\tY^*=\entry{y^*_\omega}$ is a latent continuous-valued tensor following an additive noise model:
 \begin{align}\label{eq:latent}
@@
 \end{example}
-
 \begin{example}[Probit model] The probit model is characterized by~\eqref{eq:model} with
@@
 \end{example}
-
-Other link functions are also possible, such as Laplace, Cauchy, inverse log-log, etc~\citep{mccullagh1980regression}.
+Other link functions are also possible, such as Laplace, Cauchy, etc~\cite{mccullagh1980regression}.
 All the models share the property that the ordered categories can be thought of as contiguous interval on some continuous scale.
@@
 Therefore, our model~\eqref{eq:model} is general and still valid in the absence of quantization process.
-More generally, we make the following assumptions about the link $f(\cdot): \mathbb{R}\mapsto[0,1]$.
+More generally, we make the following assumptions about the link $f$.
 
@@
 The link function is assumed to satisfy:
+\vspace{-.2cm}
 \begin{enumerate}
-\item $f(\theta)$ is strictly increasing and twice-differentiable in $\theta$.
+\item $f(\theta)$ is strictly increasing and twice-differentiable in $\theta\in \mathbb{R}$.
 \item $f'(\theta)$ is strictly log-concave and symmetric with respect to $\theta=0$.
@@
 
-{\bf (P1)} Given the quantization process induced by $f$ and the cut-off points $\mb$, how accurately can we estimate the latent signal tensor $\Theta$ from the ordinal observation $\tY$?
+(P1) Given the quantization process induced by $f$ and the cut-off points $\mb$, how accurately can we estimate the latent signal tensor $\Theta$ from the ordinal observation $\tY$?
 
 Clearly, the problem (P1) cannot be solved uniformly for all possible $\Theta$.
-We focus on a class of ``low-rank'' and ``flat'' signal tensors, which is a plausible assumption in practical applications~\citep{zhou2013tensor,bhaskar20151}.
+We focus on a class of ``low-rank'' and ``flat'' signal tensors, which is a plausible assumption in practical applications~\cite{zhou2013tensor,bhaskar20151}.
 Specifically, we consider the parameter space:
@@
 \end{equation}
-where $\tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is a core tensor, $\mM_k\in\mathbb{R}^{d_k\times r_k}$ are factor matrices with orthogonal columns, and $\times_k$ denotes the tensor-by-matrix multiplication~\citep{kolda2009tensor}.
+where $\tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is a core tensor, $\mM_k\in\mathbb{R}^{d_k\times r_k}$ are factor matrices with orthogonal columns, and $\times_k$ denotes the tensor-by-matrix multiplication~\cite{kolda2009tensor}.
 The Tucker low-rankness is popularly imposed in tensor data analysis, and is shown to provide a reasonable tradeoff between model complexity and model flexibility.
-Note that, unlike matrices, there are various notations of tensor low-rankness, such as CP rank~\citep{hitchcock1927expression} and train rank~\citep{oseledets2011tensor}.
-Some notation of low-rankness may lead to mathematically ill-posed optimization; for example, the best low CP-rank tensor approximation may not exist~\citep{de2008tensor}.
+Note that, unlike matrices, there are various notations of tensor low-rankness, such as CP rank~\cite{hitchcock1927expression} and train rank~\cite{oseledets2011tensor}.
+Some notation of low-rankness may lead to mathematically ill-posed optimization; for example, the best low CP-rank tensor approximation may not exist~\cite{de2008tensor}.
 We choose Tucker representation for well-posedness of optimization and easy interpretation.
@@
 
-{\bf (P2)} Given an incomplete set of ordinal observations $\{y_{\omega}\}_{\omega\in\Omega}$, how many sampled entries do we need to consistently recover $\Theta$ based on the model~\eqref{eq:model}?
+(P2) Given an incomplete set of ordinal observations $\{y_{\omega}\}_{\omega\in\Omega}$, how many sampled entries do we need to consistently recover $\Theta$ based on the model~\eqref{eq:model}?
 
@@
 In classical matrix completion, the incoherence is often imposed on the singular vectors.
-This assumption is recently relaxed for ``flat'' matrices with bounded magnitude~\citep{negahban2011estimation,cai2013max,bhaskar20151}.
+This assumption is recently relaxed for ``flat'' matrices with bounded magnitude~\cite{negahban2011estimation,cai2013max,bhaskar20151}.
 We adopt the same assumption for higher-order tensors.
@@
 \section{Rank-constrained M-estimator}\label{sec:theory}
+\vspace{-.1cm}
 We present a general treatment to both problems mentioned above.
@@
 \begin{align}\label{eq:objective}
-\logl(\Theta, \mb)&=\sum_{\omega\in\Omega}\sum_{\ell\in[L]} \Big\{\mathds{1}_{\{y_\omega=\ell\}} \log \big[f(b_\ell-\theta_\omega)-  f(b_{\ell-1}-\theta_\omega)\big]\Big\}.
+\logl(\Theta, \mb)&=\sum_{\omega\in\Omega}\sum_{\ell\in[L]} \Big\{\mathds{1}\{y_\omega=\ell\} \log \big[f(b_\ell-\theta_\omega)- \notag \\
+&\quad \quad \quad \quad \quad \quad \quad  f(b_{\ell-1}-\theta_\omega)\big]\Big\}.
 \end{align}
@@
 In practice, the cut-off points $\mb$ are unknown and should be jointly estimated with $\Theta$.
-For technical convenience, we assume in this section that the cut-off points $\mb$ are known.
-The adaptation of unknown $\mb$ is addressed in Section~\ref{sec:algorithm} and Appendix~\ref{sec:extention}.
-
-We define a few key quantities that will be used in our theory. Let $g_\ell(\theta)=f(b_\ell-\theta)-f(b_{\ell-1}-\theta)$ for all $\ell\in[L]$, and
-\begin{equation}\label{eq:regular}
-A_\alpha = \min_{\ell\in[L], |\theta|\leq \alpha} g_\ell(\theta),\quad  U_\alpha=\max_{\ell\in[L], |\theta|\leq \alpha } {|\dot{g}_\ell(\theta)|\over g_\ell(\theta)},\quad \text{and}\quad
- L_\alpha = \min_{\ell\in[L], |\theta|\leq \alpha }\left[{\dot{g}_\ell^2 (\theta)\over g_\ell^2(\theta)} -{\ddot{g}_\ell(\theta)\over g_\ell(\theta)}\right],
-\end{equation}
+We will focus on the theory and algorithm with known $\mb$ in the main paper.
+The adaptation for unknown $\mb$ is addressed in Section~\ref{sec:algorithm} and the Supplement.
+
+We define a few key quantities that will be used in our theory. Let $g_\ell=f(\theta+b_\ell)-f(\theta+b_{\ell-1})$ for all $\ell\in[L]$, and
+\begin{align}\label{eq:regular}
+A_\alpha &= \min_{\ell\in[L], |\theta|\leq \alpha} g_\ell(\theta),\quad  U_\alpha=\max_{\ell\in[L], |\theta|\leq \alpha } {|\dot{g}_\ell(\theta)|\over g_\ell(\theta)},\\
+ L_\alpha &= \min_{\ell\in[L], |\theta|\leq \alpha }\left[{\dot{g}_\ell^2 (\theta)\over g_\ell^2(\theta)} -{\ddot{g}_\ell(\theta)\over g_\ell(\theta)}\right],
+\end{align}
 where $\dot{g}(\theta)=dg(\theta)/d\theta$, and $\alpha$ is the entrywise bound of $\Theta$.
@@
 \subsection{Estimation error for tensor denoising}\label{sec:denosing}
+\vspace{-.1cm}
 For the tensor denoising problem, we assume that the full set of tensor entries are observed. We assess the estimation accuracy using the mean squared error (MSE):
 \[
-\mathrm{MSE}(\hat \Theta, \trueT)={1\over \prod_k d_k}\FnormSize{}{\Theta-\trueT}^2.
+\text{MSE}(\hat \Theta, \trueT)={1\over \prod_k d_k}\FnormSize{}{\Theta-\trueT}^2.
 \]
@@
 \end{equation}
-where $c_1>0$ is a constant that depends only on $K$.\\
+where $c_1 >0$ is a constant that depends only on $K$.
 \end{thm}
-
 Theorem~\ref{thm:rate} establishes the statistical convergence for the estimator~\eqref{eq:estimator}.
-In fact, the proof of this theorem (see Section~\ref{sec:proofMSE}) shows that the same statistical rate holds, not only for the global optimizer~\eqref{eq:estimator}, but also for any local optimizer $\check \Theta$ in the level set $\{\check\Theta\in\tP: \logl(\check\Theta)\geq \logl(\trueT)\}$.
+In fact, the proof of this theorem (see the Supplement) shows that the same statistical rate holds, not only for the global optimizer~\eqref{eq:estimator}, but also for any local optimizer $\check \Theta$ in the level set $\{\check\Theta\in\tP: \logl(\check\Theta)\geq \logl(\trueT)\}$.
 This suggests that the local optimality itself is not necessarily a severe concern in our context, as long as the convergent objective is large enough.
-In Section~\ref{sec:algorithm}, we perform empirical studies to assess the algorithmic stability.
+In Section ~\ref{sec:algorithm}, we perform empirical studies to assess the algorithmic stability.
 
 A similar conclusion is obtained for the prediction error, measured in Kullback-Leibler (KL) divergence, between the categorical distributions in the observation space.
-
 \begin{cor}[Prediction error]~\label{cor:prediction}
@@
 \begin{equation}\label{eq:KLrate}
-\mathrm{KL}(\mathbb{P}_{\tY} || \hat{\mathbb{P}}_{\tY}) \leq  {c_1 U^2_\alpha r_{\max}^{K-1}  \over L^2_\alpha } {(4L-6)\dot{f}^2(0)  \over A_\alpha} {\sum_kd_k\over  \prod_k d_k},
-\end{equation}
-where $c_1>0$ is the same constant as in Theorem~\ref{thm:rate}.
+\text{KL}(\mathbb{P}_{\tY} || \hat{\mathbb{P}}_{\tY}) \leq  {c_1 U^2_\alpha r_{\max}^{K-1}  \over L^2_\alpha } {(4L-6)\dot{f}^2(0)  \over A_\alpha} {\sum_kd_k\over  \prod_k d_k},
+\end{equation}
+where $c_1 >0$ is the same constant as in Theorem~\ref{thm:rate}.
 \end{cor}
-
 To gain insight into these bounds, we consider a special setting with equal dimension in all modes, i.e., $d_1=\cdots=d_K=d$. In such a case, our bound \eqref{eq:rate} reduces to
@@
 \begin{align}\label{eq:lower}
-\inf_{\hat \Theta }\sup_{\trueT\in\tP} \mathbb{P}\Big\{&\mathrm{MSE}(\hat \Theta, \trueT)  \geq c \min\left( \alpha^2, \ { Cr_{\max}d_{\max} \over \prod_k d_k} \right) \Big\} \geq {1\over 8},
-\end{align}
-where $C=C(\alpha, L, f,\mb)>0$ and $c>0$ are constants independent of tensor dimension and the rank.\\
+\inf_{\hat \Theta }\sup_{\trueT\in\tP} \mathbb{P}\Big\{& \text{MSE}(\hat \Theta, \trueT) \\
+& \geq c\min\left( \alpha^2, \ { Cr_{\max}d_{\max} \over \prod_k d_k} \right) \Big\} \geq {1\over 8},
+\end{align}
+where $C=C(\alpha, L, f,\mb)>0$ and $c>0$ are constants independent of tensor dimension and the rank.
 \end{thm}
-
-We see that the lower bound matches the upper bound in~\eqref{eq:rate} on the polynomial order of tensor dimension. Therefore, our estimator~\eqref{eq:estimator} is order-optimal.
+We see that the lower bound matches the upper bound in~\eqref{eq:rate} on the polynomial order of tensor dimension. Therefore, our estimator~\eqref{eq:estimator} is rate-optimal.
 
@@
 \end{thm}
-
 Theorem~\ref{thm:completion} shows that our estimator achieves consistent recovery using as few as $\tilde{\tO}(Kd)$ noisy, quantized observations from an order-$K$ $(d,\ldots,d)$-dimensional tensor.
 Note that $\tilde \tO(Kd)$ roughly matches the degree of freedom for an order-$K$ tensor of fixed rank $\mr$, suggesting the optimality of our sample requirement.
-This sample complexity substantially improves over earlier result $\tO(d^{\lceil K/2\rceil})$ based on square matricization~\citep{mu2014square}, or $\tO(d^{N/2})$ based on tensor nuclear-norm regularization~\citep{yuan2016tensor}.
-Existing methods that achieve $\tilde \tO(Kd)$ sample complexity require either a deterministic cross sampling design~\citep{zhang2019cross} or univariate measurements~\citep{ghadermarzy2018learning}.
+This sample complexity substantially improves over earlier result $\tO(d^{\lceil K/2\rceil})$ based on square matricization~\cite{mu2014square}, or $\tO(d^{K/2})$ based on tensor nuclear-norm regularization~\cite{yuan2016tensor}.
+Existing methods that achieve $\tilde \tO(Kd)$ sample complexity require either a deterministic cross sampling design~\cite{zhang2019cross} or univariate measurements~\cite{ghadermarzy2018learning}.
 Our method extends the conclusions to multi-level measurements under a broader class of sampling schemes.
@@
 We describe the algorithm to seek the optimizer of~\eqref{eq:objective}.
-In practice, the cut-off points $\mb$ are often unknown, so we choose to maximize $\logl$ jointly over $(\Theta,\mb)\in\tP\times \tB$ (see Appendix~\ref{sec:extention} for details).
-The objective $\logl$ is concave in $(\Theta,\mb)$ whenever $f'$ is log-concave (see Section~\ref{sec:proofconvexity}).
+In practice, the cut-off points $\mb$ are often unknown, so we choose to maximize $\logl$ jointly over $(\Theta,\mb)\in\tP\times \tB$ (see the Supplement for details).
+The objective $\logl$ is concave in $(\Theta,\mb)$ whenever $f'$ is log-concave.
 However, the feasible set $\tP$ is non-convex, which makes the optimization~\eqref{eq:objective} a non-convex problem.
@@
             \STATE $\mM_k^{(t+1)}\gets\argmax_{\mM_k\in\mathbb{R}^{d_k\times r_K}}\mathcal{L}_{\mathcal{Y},\Omega}(\mM_k)$,
+            \newline
              s.t. $\mnormSize{}{\Theta^{(t+1)}}\leq \alpha$, where $\Theta^{(t+1)}$ is the parameter tensor based on the current block estimates.
@@
             \STATE {Update $\mathcal{C}^{(t+1)}$ while fixing other blocks:}
-\newline$\mathcal{C}^{(t+1)} \gets \argmax_{\mathcal{C}\in \mathbb{R}^{r_1\times\cdots\times r_K}}\mathcal{L}_{\mathcal{Y},\Omega}(\mathcal{C})$, s.t. $\mnormSize{}{\Theta^{(t+1)}}\leq \alpha.$
+              \newline$\mathcal{C}^{(t+1)} \gets \argmax_{\mathcal{C}\in \mathbb{R}^{r_1\times\cdots\times r_K}}\mathcal{L}_{\mathcal{Y},\Omega}(\mathcal{C})$, s.t.
+$\mnormSize{}{\Theta^{(t+1)}}\leq \alpha.$
                \STATE {Update $\Theta^{(t+1)}$ based on the current block estimates:}
@@
             \STATE {Update $\mb^{(t+1)}$ while fixing $\Theta^{(t+1)}$:}
+              \newline
               $\mb^{(t+1)} \gets \argmax_{\mb\in \tB}\mathcal{L}_{\mathcal{Y},\Omega}\big(\Theta^{(t+1)},\mb\big).$
@@
 
-\begin{figure}[ht]
+\begin{figure}[H]
 \centering
-\includegraphics[width=0.7\textwidth]{algorithm.pdf}\label{fig:stability}
+\includegraphics[width=8cm]{algorithm.pdf}\label{fig:stability}
+\vspace{-.2cm}
 \caption{Trajectory of objective function with various $d$ and $r$.}
+\vspace{-.2cm}
 \end{figure}
@@
 \hat \mr&=\argmin_{\mr\in\mathbb{N}^K_{+}} \text{BIC}(\mr)\\
-&=\argmin_{\mr\in\mathbb{N}^K_{+}}\left\{-2\tL_{\tY}(\hat \Theta(\mr),\hat \mb(\mr))+p_e(\mr)\log \left(\prod_k d_k\right)\right\},
+&=\argmin_{\mr\in\mathbb{N}^K_{+}}\{-2\tL_{\tY}(\hat \Theta(\mr),\hat \mb(\mr))+p_e(\mr)\log (\prod_k d_k)\},
 \end{align}
@@
 Therefore, the estimation error diverges in the absence of noise $\tE$, because it is impossible to distinguish two different signal tensors, e.g., $\Theta_1=\ma_1\otimes \ma_2\otimes \ma_3$ and $\Theta_2=\text{sign}(\ma_1)\otimes \text{sign}(\ma_2)\otimes \text{sign}(\ma_3)$, from the quantized observations.
-This phenomenon~\citep{davenport2014,sur2019modern} is clearly contrary to the classical continuous-valued tensor problem.
+This phenomenon~\cite{davenport2014,sur2019modern} is clearly contrary to the classical continuous-valued tensor problem.
 
@@
 
-\begin{figure}[ht]
-\centering
-\includegraphics[width=0.7\textwidth]{panel.pdf}
+\begin{figure}[http]
+\includegraphics[width=8.3cm]{panel.pdf}
+\vspace{-.2cm}
 \caption{Empirical relationship between (relative) MSE versus (a) dimension $d$, (b) signal level $\alpha$, (c) observation fraction $\rho$, and (d) number of ordinal levels $L$.
@@
 \label{fig:finite}
+\vspace{-.3cm}
 \end{figure}
 
-\subsection{Comparison with alternative methods}\label{sec:compare}
+\subsection{Comparison with alternative methods}
 Next, we compare our ordinal tensor method ({\bf Ordinal-T}) with three popular low-rank methods:
 
-\begin{itemize}
-\item Continuous tensor decomposition ({\bf Continuous-T})~\citep{acar2010scalable} is a low-rank approximation method based on classical Tucker model.
-\item One-bit tensor completion ({\bf 1bit-T})~\citep{ghadermarzy2018learning} is a max-norm penalized tensor learning method based on partial binary observations.
-\item Ordinal matrix completion ({\bf Ordinal-M})~\citep{bhaskar2016probabilistic} is a rank-constrained matrix estimation method based on noisy, quantized observations.
+\begin{itemize}[itemsep=0pt,topsep=0pt,leftmargin=*,partopsep=0pt]
+\item Continuous tensor decomposition ({\bf Continuous-T})~\cite{acar2010scalable} is a low-rank approximation method based on classical Tucker model.
+\item One-bit tensor completion ({\bf 1bit-T})~\cite{ghadermarzy2018learning} is a max-norm penalized tensor learning method based on partial binary observations.
+\item Ordinal matrix completion ({\bf Ordinal-M})~\cite{bhaskar2016probabilistic} is a rank-constrained matrix estimation method based on noisy, quantized observations.
 \end{itemize}
@@
 The {\bf Continuous-T} is applied directly to $\tY$ by treating the $L$ levels as continuous observations.
-The {\bf Ordinal-M} is applied to the matrix $\tY_{(1)}$ obtained via 1-mode unfolding.
+The {\bf Ordinal-M} is applied to the 1-mode unfolding matrix $\tY_{(1)}$.
 The {\bf 1bit-T} is applied to $\tY$ in two ways.
-The first approach ({\bf 1bit-sign-T}) follows from~\citet{ghadermarzy2018learning} that transforms $\tY$ to a binary tensor, by taking the entrywise sign of the mean-adjusted tensor, $\tY- |\Omega|^{-1}\sum_\omega y_\omega$.
-The second approach ({\bf 1bit-category-T}) transforms the order-3 ordinal tensor $\tY$ to an order-4 binary tensor $\tY^{\sharp}=\entry{y^{\sharp}_{ijkl}}$ via dummy variable encoding; i.e., $y^{\sharp}_{ijk\ell}=\mathds{1}_{\{y_{ijk}=\ell\}}$ for all $\ell\in[L-1]$.
-
-We evaluate the methods by their capabilities in predicting the most likely label for each entry, i.e.,\ $y_\omega^{\text{mode}}=\arg\max_\ell\mathbb{P}(y_\omega=\ell)$.
-Two performance metrics are considered: mean absolute deviation, $\text{MAD}(\tY^{\text{mode}}, \hat \tY^{\text{mode}})=d^{-K}\sum_\omega |y_\omega^{\text{mode}}-\hat y_\omega^{\text{mode}}|$, and misclassification rate, $\text{MCR}(\tY^{\text{mode}}, \hat \tY^{\text{mode}})=d^{-K}\sum_\omega\mathds{1}_{\{y_\omega^{\text{mode}}\neq\text{round}(\hat y_\omega^{\text{mode}})\}}$,
-where $\text{round}(\cdot)$ denotes the nearest integer of the prediction (possibly continuous-valued returned by {\bf Continuous-T}).
-Both metrics are widely used for evaluation of prediction accuracy.
-Note that MAD penalizes the large deviation more heavily than MCR.
+The first approach ({\bf 1bit-sign-T}) follows from~\citet{ghadermarzy2018learning} that transforms
+$\tY$ to a binary tensor, by taking the entrywise sign of the mean-adjusted tensor, $\tY- |\Omega|^{-1}\sum_\omega y_\omega$.
+The second approach ({\bf 1bit-category-T}) transforms the order-3 ordinal tensor $\tY$ to an order-4 binary tensor, $\tY^{\sharp}=\entry{y^{\sharp}_{ijkl}}$, via dummy variable encoding; i.e., $y^{\sharp}_{ijk\ell}=\mathds{1}\{y_{ijk}=\ell\}$ for $\ell\in[L-1]$.
+We evaluate the methods by their capabilities in predicting the most likely labels, $y_\omega^{\text{mode}}=\arg\max_\ell\mathbb{P}(y_\omega=\ell)$.
+Two performance metrics are considered: mean absolute deviation (MAD) $=d^{-K}\sum_\omega |y_\omega^{\text{mode}}-\hat y_\omega^{\text{mode}}|$, and misclassification rate (MCR) $=d^{-K}\sum_\omega\mathds{1}\{y_\omega^{\text{mode}}\neq\text{round}(\hat y_\omega^{\text{mode}})\}$,
+where $\text{round}(\cdot)$ denotes the nearest integer. Note that MAD penalizes the large deviation more heavily than MCR.
 
 Figure~\ref{fig:compare} compares the prediction accuracy under the setting $\alpha=10$, $d=20$, and $r=5$.
-The problem size we considered is comparable to~\citet{ghadermarzy2018learning}.
 We find that our method outperforms the others in both MAD and MCR.
 In particular, methods built on multi-level observations ({\bf Ordinal-T}, {\bf Ordinal-M}, {\bf 1bit-category-T}) exhibit stable MCR over $\rho$ and $L$, whereas the others two methods ({\bf Continuous-T}, {\bf 1bit-sign-T}) generally fail except for $L=2$ (Figures~\ref{fig:compare}a-b).
-This observation highlights the necessity of modeling multi-level probabilities in classification task.
-Interestingly, although both {\bf 1bit-category-T} and our method {\bf Ordinal-T} behave similarly for binary tensors ($L=2$), the improvement of our method is substantial as $L$ increases (Figures~\ref{fig:compare}a and~\ref{fig:compare}c).
+This observation highlights the benefits of multi-level modeling in the classification task.
+Although {\bf 1bit-category-T} and our method {\bf Ordinal-T} behave similarly for binary tensors ($L=2$), the improvement of our method is substantial as $L$ increases (Figures~\ref{fig:compare}a and~\ref{fig:compare}c).
 One possible reason is that our method incorporates the intrinsic ordering among the $L$ levels via proportional odds assumption~\eqref{eq:logodd}, whereas {\bf 1bit-category-T} ignores the ordinal structure and dependence among the induced binary entries.
+
 Figures~\ref{fig:compare}c-d assess the prediction accuracy with sample size.
@@
 
-\begin{figure}[ht]
-\centering
-\includegraphics[width=0.7\textwidth]{panel2.pdf}
-\caption{Performance comparison for predicting most likely labels.
-(a, c) Prediction error versus sample complexity $\rho=|\Omega|/d^K$ when $L=5$.
-(b, d) Prediction error versus the number of ordinal levels $L$ when $\rho=0.8$.
+\begin{figure}[t]
+\includegraphics[width=8.3cm]{panel2.pdf}
+\vspace{-.4cm}
+\caption{Performance comparison in MCR (a, b) and MAD (c, d).
+(b, d) Prediction errors versus the number of ordinal levels $L$ when $\rho=0.8$.
+(a, c) Prediction errors versus sample complexity $\rho=|\Omega|/d^K$ when $L=5$.
 }\label{fig:compare}
+\vspace{-.4cm}
 \end{figure}
@@
 We also compare the methods by their performance in predicting the median labels, $y_\omega^{\text{median}}=\min\{\ell\colon\mathbb{P}(y_\omega=\ell)\geq 0.5\}$.
-Under the latent variable model~\eqref{eq:latent} and Assumption~\ref{ass:link}, the median label is the quantized $\theta_\omega$ without noise; i.e.\ $y_\omega^{\text{median}}=\sum_\ell \mathds{1}_{\theta_\omega\in(b_{\ell-1},b_\ell]}$.
+Under the latent variable model~\eqref{eq:latent} and Assumption~\ref{ass:link}, the median label is the quantized $\theta_\omega$ without noise; i.e.\ $y_\omega^{\text{median}}=\sum_\ell \mathds{1}\{\theta_\omega\in(b_{\ell-1},b_\ell]\}$.
+Similar results show the outperformance of our method (see the Supplement).
+
+\section{Data Applications}
+We apply our ordinal tensor method to two real-world datasets.
+In the first application, we use our model to analyze an ordinal tensor consisting of structural connectivities among 68 brain regions for 136 individuals from Human Connectome Project (HCP)~\citep{geddes2016human}.
+In the second application, we perform tensor completion to an ordinal dataset with missing values.
+The data tensor records the ratings of 139 songs on a scale of 1 to 5 from 42 users on 26 contexts~\citep{baltrunas2011incarmusic}.
+
+\subsection{Human Connectome Project (HCP)}
+Each entry in the HCP dataset takes value on a nominal scale, \{{\it high, moderate, low}\}, indicating the strength level of fiber connection.
+We convert the dataset to a 3-level ordinal tensor $\tY\in[3]^{68\times 68\times 136}$ and apply the ordinal tensor method with a logistic link function.
+The BIC suggests $\mr = (23,23,8)$ with $\logl(\hat{\Theta},\hat{\mb}) = -216,646$.
+Based on the estimated Tucker factors $\{\hat \mM_k\}$, we perform a clustering analysis via K-mean on the brain nodes (see detailed procedure in the Supplement).
+We find that the 68 brain nodes are grouped into 11 clusters, and the clustering captures the spatial separation of brain nodes (see the Supplement).
+In particular, the top three clusters represent the connections in the left-, right-, and inter-hemispheres; the smaller clusters represent local regions driving by similar nodes.
+For example, the cluster IV consists of nodes in the left supramarginal gyrus region.
+The identified similarity among brain nodes suggests the applicability of our method.
+
+We also compare the goodness-of-fit of various tensor methods on the HPC data.
+Table~\ref{table:CV} summarizes the prediction error via 5-fold stratified cross-validation averaged over 10 runs.
+Our method outperforms the others, especially in MAD.
+
+\begin{table}[http]
+\resizebox{.48\textwidth}{!}{
+\begin{tabular}{c|c|c}
+\Xhline{2\arrayrulewidth}
+\multicolumn{3}{c}{Human Connectome Project (HCP) dataset}\\
+\Xhline{2\arrayrulewidth}
+Method                    & MAD & MCR\\
+\hline
+Ordinal-T (ours) & 0.1607 (0.0005)& 0.1606 (0.0005)\\
+ Continuous-T&0.2530 (0.0002)&0.1599 (0.0002)\\
+ 1bit-sign-T&0.3566 (0.0010)&0.1563 (0.0010)\\
+ \Xhline{2\arrayrulewidth}
+\multicolumn{3}{c}{InCarMusic dataset}\\
+\Xhline{2\arrayrulewidth}
+Method                    & MAD & MCR\\
+\hline
+Ordinal-T (ours) & 1.37 (0.039) &  0.59 (0.009)\\
+Continuous-T & 2.39 (0.152)&0.94 (0.027)\\
+1bit-sign-T&1.39 (0.003)& 0.81 (0.005)\\
+\hline
+\end{tabular}
+}
+\vspace{-.2cm}
+\caption{Comparison of prediction error in the HPC and InCarMusic analyses. Standard errors are reported in parentheses.}\label{table:CV}
+\vspace{-.3cm}
+\end{table}
+
+\subsection{InCarMusic recommendation system}
+We apply ordinal tensor completion to a recommendation system {\it InCarMusic}.
+{\it InCarMusic} is a mobile application that offers music recommendation to passengers of cars based on contexts~\cite{baltrunas2011incarmusic}.
+We conduct tensor completion on the 42-by-139-by-26 tensor with 2,844 observed entries only.
+Table~\ref{table:CV} shows the averaged prediction error via 5-fold cross validation.
+The high missing rate makes the accurate classification challenging.
+Nevertheless, our method achieves the best performance among the three.
+
+\section{Conclusions}
+We have developed a low-rank tensor estimation method based on possibly incomplete, ordinal-valued observations.
+A sharp error bound is established, and we demonstrate the outperformance of our approach compared to other methods.
+The work unlocks several directions of future research.
+One interesting question would be the inference problem, i.e.,\ to assess the uncertainty of the obtained estimates and the imputation.
+Other directions include the trade-off between (non)convex optimization and statistical efficiency.
+While we have provided numerical evidence for the success of nonconvex approach, the full landscape of the optimization remains open.
+The interplay between computational efficiency and statistical accuracy in general tensor problems warrants future research.
+
+\newpage
+\section*{Acknowledgements}
+This research is supported in part by NSF grant DMS-1915978 and Wisconsin Alumni Research Foundation.
+
+\bibliography{tensor_wang}
+\bibliographystyle{icml2020}
+
+\clearpage
+\appendix
+\onecolumn
+
+\icmltitle{Supplements for ``Tensor denoising and completion based on \\ ordinal observations''}
+
+\section{Proofs}\label{sec:proof}
+Here, we provide proofs of the theoretical results presented in Sections~\ref{sec:theory}.
+
+\subsection{Estimation error for tensor denoising}\label{sec:proofMSE}
+\begin{proof}[Proof of Theorem~\ref{thm:rate}]
+We suppress the subscript $\Omega$ in the proof, because the tensor denoising assumes complete observation $\Omega=[d_1]\times \cdots \times [d_K]$.
+It follows from the expression of $\flogl(\Theta)$ that
+\begin{align}\label{eq:property}
+{\partial \flogl\over \partial \theta_\omega}&=\sum_{\ell\in[L]}\mathds{1}\{y_{\omega}=\ell\}
+{\dot{g}_\ell(\theta_\omega)\over g_\ell(\theta_\omega)},\notag\\
+{\partial^2 \flogl\over \partial \theta_\omega^2}&=\sum_{\ell\in[L]}\mathds{1}\{y_\omega=\ell\}{\ddot{g}_\ell(\theta_\omega)g_\ell(\theta_\omega)-\dot{g}^2_\ell(\theta_\omega)\over g^2_\ell(\theta_\omega)}\ \text{and}\quad
+{\partial^2 \flogl\over \partial \theta_\omega \theta_\omega'}=0\ \text{if}\ \omega\neq \omega',
+\end{align}
+for all $\omega\in[d_1]\times \cdots \times [d_K]$.
+Define $d_{\text{total}}=\prod_k d_k$.
+Let $\fplogl\in\mathbb{R}^{d_1\times\cdots\times d_K}$ denote the tensor of gradient with respect to $\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}$, and $\fpplogl$ the corresponding Hession matrix of size $d_\text{total}$-by-$d_{\text{total}}$.
+Here, $\Vec(\cdot)$ denotes the operation that turns a tensor into a vector.
+By~\eqref{eq:property}, $\fpplogl$ is a diagonal matrix.
+Recall that
+\begin{equation}\label{eq:bound}
+U_\alpha=\max_{\ell\in[L],|\theta|\leq \alpha}{|\dot{g}_\ell(\theta)|\over g_\ell(\theta)}>0 \quad \text{and}\quad
+L_\alpha=\min_{\ell\in[L],|\theta|\leq \alpha} {\dot{g}^2_\ell(\theta)-\ddot{g}_\ell(\theta)g_\ell(\theta)\over g^2_\ell(\theta)}>0.
+\end{equation}
+Therefore, the entries in $\fplogl$ are upper bounded in magnitude by $U_\alpha>0$, and all diagonal entries in $\fpplogl$ are upper bounded by $-L_{\alpha}<0$.
+
+By the second-order Taylor's expansion of $\flogl(\Theta)$ around $\trueT$, we obtain
+\begin{equation}\label{eq:taylor}
+\flogl(\Theta)=\flogl(\trueT)+\langle\Vec(\fplogl(\trueT)),\ \Vec(\Theta-\trueT)\rangle+{1\over 2}\Vec(\Theta-\trueT)^T\fpplogl(\check\Theta)\Vec(\Theta-\trueT),
+\end{equation}
+where $\check\Theta=\gamma\trueT+(1-\gamma)\Theta$ for some $\gamma\in[0,1]$, and $\fpplogl(\check\Theta)$ denotes the $d_{\text{total}}$-by-$d_\text{total}$ Hession matrix evaluated at $\check\Theta$.
+
+We first bound the linear term in~\eqref{eq:taylor}. Note that, by Lemma~\ref{lem:inq},
+\begin{equation}\label{eq:linear}
+|\langle\Vec(\fplogl(\trueT), \Vec(\Theta-\trueT)  \rangle|\leq \snormSize{}{\fplogl(\trueT)} \nnormSize{}{\Theta-\trueT},
+\end{equation}
+where $\snormSize{}{\cdot}$ denotes the tensor spectral norm and $\nnormSize{}{\cdot}$ denotes the tensor nuclear norm. Define
+\[
+s_\omega={\partial \tL_\tY\over \partial \theta_\omega}\Big|_{\Theta=\trueT} \;\; \textrm{ for all } \; \omega\in[d_1]\times\cdots\times [d_K].
+\]
+Based on~\eqref{eq:property} and the definition of $U_\alpha$, $\fplogl(\trueT)=\entry{s_{\omega}}$ is a random tensor whose entries are independently distributed satisfying
+\begin{equation}\label{eq:norm}
+\mathbb{E}(s_\omega)=0,\quad |s_\omega|\leq U_\alpha, \quad \text{for all }\omega\in[d_1]\times \cdots \times [d_K].
+\end{equation}
+By lemma~\ref{lem:noisytensor}, with probability at least $1-\exp(-C_1 \sum_kd_k)$, we have
+\begin{equation}\label{eq:normrandom}
+\snormSize{}{\fplogl(\trueT)} \leq C_2 U_\alpha\sqrt{\sum_k d_k},
+\end{equation}
+where $C_1, C_2$ are two positive constants that depend only on $K$.
+Furthermore, note that $\text{rank}(\Theta)\leq \mr$, $\text{rank}(\trueT)\leq \mr$, so $\text{rank}(\Theta-\trueT)\leq 2\mr$.
+By lemma~\ref{lem:nuclear}, $\nnormSize{}{\Theta-\trueT}\leq (2r_{\max})^{K-1\over 2}\FnormSize{}{\Theta-\trueT}$.
+Combining~\eqref{eq:linear}, \eqref{eq:norm} and \eqref{eq:normrandom}, we have that, with probability at least $1-\exp(-C_1 \sum_kd_k)$,
+\begin{equation}\label{eq:linearconclusion}
+|\langle \Vec(\fplogl(\trueT)), \Vec(\Theta-\trueT)  \rangle | \leq C_2 U_\alpha  \sqrt{r_{\max}^{K-1} \sum_k d_k}  \FnormSize{}{\Theta-\trueT}.
+\end{equation}
+
+We next bound the quadratic term in \eqref{eq:taylor}. Note that
+\begin{align}\label{eq:quadratic}
+ \Vec(\Theta-\trueT)^T \fpplogl(\check{\Theta})\Vec(\Theta-\trueT)&=\sum_\omega \left( {\partial^2\tL_{\tY}\over \partial \theta^2_\omega} \Big|_{\Theta=\check\Theta} \right)(\theta_\omega-\theta_{{\text{true}},\omega})^2 \nonumber \\
+&\leq - L_\alpha\sum_{\omega}(\Theta_{\omega}-\Theta_{\text{true},\omega})^2 \nonumber \\
+&=-L_\alpha\FnormSize{}{\Theta-\trueT}^2,
+\end{align}
+where the second line comes from the fact that  $\mnormSize{}{\check\Theta}\leq \alpha$ and the definition of $L_\alpha$.
+
+Combining~\eqref{eq:taylor}, \eqref{eq:linearconclusion} and~\eqref{eq:quadratic}, we have that, for all $\Theta\in\tP$, with probability at least $1-\exp(-C_1 \sum_kd_k)$,
+\[
+\tL_\tY(\Theta)\leq \tL_{\tY}(\trueT)+C_2U_\alpha  \left(r_{\max}^{K-1}\sum_k d_k\right)^{1/2}  \FnormSize{}{\Theta-\trueT}-{L_\alpha\over 2}\FnormSize{}{\Theta-\trueT}^2.
+\]
+In particular, the above inequality also holds for $\hat \Theta\in\tP$. Therefore,
+\[
+\tL_\tY(\hat \Theta)\leq \tL_{\tY}(\trueT)+C_2U_\alpha \left(r_{\max}^{K-1}\sum_k d_k\right)^{1/2}  \FnormSize{}{\hat \Theta-\trueT}-{L_\alpha\over 2} \FnormSize{}{\hat \Theta-\trueT}^2.
+\]
+Since $\hat \Theta=\arg\max_{\Theta\in\tP}\tL_\tY(\Theta)$, $\tL_\tY(\hat \Theta)-\tL_{\tY}(\trueT)\geq 0$, which gives
+\[
+C_2U_\alpha \left(r_{\max}^{K-1}\sum_k d_k\right)^{1/2}  \FnormSize{}{\hat \Theta-\trueT}-{L_\alpha\over 2}\FnormSize{}{\hat \Theta-\trueT}^2\geq 0.
+\]
+Henceforth,
+\[
+{1\over \sqrt{\prod_k d_k}} \FnormSize{}{\hat \Theta-\trueT}\leq {2C_2U_\alpha \sqrt{r_{\max}^{K-1}\sum_k d_k}\over L_\alpha \sqrt{\prod_k d_k}}={2C_2U_\alpha r_{\max}^{(K-1)/2}\over L_\alpha} \sqrt{ \sum_k d_k \over \prod_k d_k}.
+\]
+This completes the proof.
+\end{proof}
+
+\begin{proof}[Proof of Corollary~\ref{cor:prediction}]
+The result follows immediately from Theorem~\ref{thm:rate} and Lemma~\ref{lem:KL}.
+\end{proof}
+
+\begin{proof}[Proof of Theorem~\ref{thm:minimax}]
+
+Let $d_{\text{total}}=\prod_{k\in[K]}d_k$, and $\gamma\in[0,1]$ be a constant to be specified later.
+Our strategy is to construct a finite set of tensors $\tX=\{\Theta_i\colon i=1,\ldots \}\subset \tP$ satisfying the properties of (i)-(iv) in Lemma~\ref{lem:construction}.
+By Lemma~\ref{lem:construction}, such a subset of tensors exist.
+For any tensor  $\Theta\in\tX$, let $\mathbb{P}_{\Theta}$ denote the distribution of $\tY|\Theta$, where $\tY$ is the ordinal tensor.
+In particular, $\mathbb{P}_{\mathbf{0}}$ is the distribution of $\tY$ induced by the zero parameter tensor $\mathbf{0}$, i.e., the distribution of $\tY$ conditional on the parameter tensor $\Theta=\mathbf{0}$.
+Based on the Remark for Lemma~\ref{lem:KL}, we have
+\begin{equation}\label{eq:KLbound1}
+\mathrm{KL}(\mathbb{P}_{\Theta}|| \mathbb{P}_{\mathbf{0}})\leq C \FnormSize{}{\Theta}^2,
+\end{equation}
+where $C={(4L-6) \dot{f}^2(0)\over  A_\alpha}>0$ is a constant independent of the tensor dimension and rank.
+Combining the inequality~\eqref{eq:KLbound1} with property (iii) of $\tX$, we have
+\begin{equation}\label{eq:KLbound}
+\text{KL}(\mathbb{P}_{\Theta}||\mathbb{P}_{\mathbf{0}})\leq \gamma^2 r_{\max} d_{\max}.
+\end{equation}
+From~\eqref{eq:KLbound} and the property (i), we deduce that the condition
+\begin{equation}\label{eq:totalKL}
+{1\over \text{Card}(\tX)-1}\sum_{\Theta \in\tX}\text{KL}(\mathbb{P}_{\Theta}, \mathbb{P}_{\mathbf{0}})\leq \varepsilon \log_2\left\{\text{Card}(\tX)-1 \right\}
+\end{equation}
+holds for any $ \varepsilon \geq 0$ when $\gamma\in[0,1]$ is chosen to be sufficiently small depending on $\varepsilon$, e.g., $\gamma \leq \sqrt{\varepsilon\log2\over8}$.
+By applying Lemma~\ref{lem:Tsybakov} to~\eqref{eq:totalKL}, and in view of the property (iv), we obtain that
+\begin{equation}\label{eq:final}
+\inf_{\hat \Theta}\sup_{\trueT\in \tX}\mathbb{P}\left(\FnormSize{}{\hat \Theta- \trueT}\geq  {\gamma\over 8} \min\left\{ \alpha\sqrt{d_{\text{total}}}, C^{-1/2}\sqrt{ r_{\max}d_{\max}}\right\} \right)\geq {1\over 2}\left(1-2\varepsilon-\sqrt{16 \varepsilon \over r_{\max}d_{\max}\log2}\right).
+\end{equation}
+Note that $\text{Loss}(\hat \Theta, \trueT)=\FnormSize{}{\hat \Theta- \trueT}^2/d_{\text{total}}$ and $\tX\subset \tP$.
+By taking $\varepsilon=1/10$ and $\gamma=1/11$, we conclude from~\eqref{eq:final} that
+\begin{equation}\label{eq:prob}
+\inf_{\hat \Theta}\sup_{\trueT\in \tP}\mathbb{P}\left(\text{Loss}(\hat \Theta, \trueT)\geq c\min\left \{ \alpha^2,  {C^{-1}r_{\max}d_{\max}\over d_{\text{total}}}\right \}\right)\geq {1\over 2}\left({4\over 5}- \sqrt{1.6\over r_{\max}d_{\max}\log2} \right)\geq {1\over 8},
+\end{equation}
+where $c = {1 \over 88^2}$ and the last inequality comes from the condition for $d_{\text{max}}$.
+This completes the proof.
+\end{proof}
+
+\subsection{Sample complexity for tensor completion}
+
+\begin{proof}[Proof of Theorem~\ref{thm:completion}]
+
+For notational convenience, we use $\MFnormSize{}{\Theta}=\sum_{\omega\in\Omega}\Theta^2_\omega$ to denote the sum of squared entries over the observed set $\Omega$, for a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$.
+
+Following a similar argument as in the proof of Theorem~\ref{thm:rate}, we have
+\begin{equation}\label{eq:Taylor2}
+\logl(\Theta)=\logl(\trueT)+\langle\Vec(\plogl),\ \Vec(\Theta-\trueT)\rangle+{1\over 2}\Vec(\Theta-\trueT)^T\pplogl(\check\Theta)\Vec(\Theta-\trueT),
+\end{equation}
+where
+\begin{enumerate}
+\item $\plogl$ is a $d_1\times\cdots\times d_K$ tensor with $|\Omega|$ nonzero entries, and each entry is upper bounded by $U_\alpha>0$.
+\item $\pplogl$ is a diagonal matrix of size $d_{\text{total}}$-by-$d_{\text{total}}$ with $|\Omega|$ nonzero entries, and each entry is upper bounded by $-L_{\alpha}<0$.
+\end{enumerate}
+
+Similar to~\eqref{eq:linear} and~\eqref{eq:quadratic}, we have
+\begin{equation}\label{eq:linear2}
+|\langle\Vec(\plogl),\ \Vec(\Theta-\trueT)\rangle|\leq C_2U_\alpha \sqrt{r_{\max}^{K-1}\sum_k d_k}\MFnormSize{}{\Theta-\trueT}
+\end{equation}
+and
+\begin{equation}\label{eq:quadratic2}
+\Vec(\Theta-\trueT)^T\fpplogl(\check\Theta)\Vec(\Theta-\trueT)\leq -L_\alpha \MFnormSize{}{\Theta-\trueT}^2.
+\end{equation}
+
+Combining~\eqref{eq:Taylor2}-\eqref{eq:quadratic2} with the fact that $\logl(\hat \Theta)\geq \logl(\trueT)$, we have
+\begin{equation}\label{eq:sample}
+\MFnormSize{}{\hat \Theta-\trueT}\leq {2C_2U_\alpha  r_{\max}^{(K-1)/2} \over L_\alpha} \sqrt{\sum_kd_k},
+\end{equation}
+with probability at least $1-\exp(-C_1 \sum_k d_k)$.
+Lastly, we invoke the result regarding the closeness of $\Theta$ to its sampled version $\Theta_{\Omega}$, under the entrywise bound condition.
+Note that $\mnormSize{}{\hat\Theta-\trueT}\leq 2\alpha$ and $\text{rank}(\hat \Theta-\trueT)\leq 2\mr$.
+By Lemma~\ref{lem:Mnormbound}, $\anormSize{}{\hat \Theta-\trueT}\leq 2^{(3K-1)/2} \alpha \left({\prod r_k \over r_{\max}}\right)^{3/2}$.
+Therefore, the condition in Lemma~\ref{lem:convexity} holds with $\beta=2^{(3K-1)/2}\alpha \left({\prod r_k \over r_{\max}}\right)^{3/2}$.
+Applying Lemma~\ref{lem:convexity} to~\eqref{eq:sample} gives
+\begin{align}
+ \PiFnormSize{}{\hat \Theta-\trueT}^2&\leq {1\over m}\MFnormSize{}{\hat \Theta-\trueT}^2+c\beta\sqrt{\sum_k d_k\over |\Omega|}\\
+ &\leq {C_2  r^{K-1}_{\max}} {\sum_k d_k \over |\Omega|}+C_1 \alpha r_{\max}^{3(K-1)/2}\sqrt{\sum_kd_k\over |\Omega|},
+\end{align}
+with probability at least $1-\exp(-{\sum_kd_k\over \sum_k \log d_k})$ over the sampled set $\Omega$. Here $C_1, C_2>0$ are two constants independent of the tensor dimension and rank. Therefore,
+\[
+ \PiFnormSize{}{\hat \Theta-\trueT}^2\to 0,\quad \text{as}\quad {|\Omega|\over \sum_kd_k}\to \infty,
+\]
+provided that $r_{\max}=O(1)$.
+\end{proof}
+
+\subsection{Convexity of the log-likelihood function}\label{sec:proofconvexity}
+\begin{thm}\label{thm:convexity}
+Define the function
+\begin{align}\label{eq:function}
+ \logl(\Theta, \mb)&=\sum_{\omega\in\Omega}\sum_{\ell\in[L]} \big(\mathds{1}\{y_\omega=\ell\}\log \left[f(b_\ell-\theta_\omega)-  f(b_{\ell-1}-\theta_\omega)\right]\big),
+ \end{align}
+where $f(\cdot)$ satisfies Assumption~\ref{ass:link}. Then, $\logl(\Theta, \mb)$ is concave in $(\Theta,\mb)$.
+\end{thm}
+
+\begin{proof}
+Define $d_{\text{total}}=\prod_k d_k$.
+By abuse of notation, we use $(\Theta,\mb)$ to denote the length-$(d_{\text{total}}+L-1)$-vector collecting all parameters together.
+Let us denote a bivariate function
+\begin{align}
+\lambda\colon &\mathbb{R}^2\mapsto \mathbb{R}\\
+(u,v)&\mapsto\lambda(u,v) = \log \big[f(u)-  f(v)\big].
+\end{align}
+It suffices to show that $\lambda(u,v)$ is concave in $(u,v)$ where $u>v$.
+
+Suppose that the claim holds (which we will prove in the next paragraph). Based on~\eqref{eq:function}, $u,v$ are both linear functions of $(\Theta,\mb)$:
+\[
+u = \ma_1^T(\Theta,\mb), \quad v = \ma_2^T(\Theta,\mb), \quad \text{ for some}\ \ma_1,\ma_2\in \mathbb{R}^{d_{\text{total}}+L-1}.
+\]
+Then, $\lambda(u,v) = \lambda(\ma_1^T(\Theta,\mb),\ \ma_2^T(\Theta,\mb))$ is concave in $(\Theta,\mb)$ by the definition of concavity.
+Therefore, we can conclude that $\logl(\Theta, \mb) $ is concave in $(\Theta,\mb)$ because $\logl(\Theta, \mb)$ is the sum of $\lambda(u,v)$.
+
+Now, we prove the concavity of $\lambda(u,v)$. Note that
+\begin{equation}
+\lambda(u,v) = \log\big[f(u)-f(v)\big]=\log\big[\int\mathds{1}_{[u,v]}(x)f'(x)dx\big],
+\end{equation}
+where $\mathds{1}_{[u,v]}$ is an indicator function that equals 1 in the interval $[u,v]$, and 0 elsewhere.
+Furthermore, $\mathds{1}_{[u,v]}(x)$ is log-concave in $(u,v,x)$, and by Assumption~\ref{ass:link}, $f'(x)$ is log-concave in $x$.
+It follows that $\mathds{1}_{[u,v]}(x)f'(x)$
+is a log-concave in $(u,v,x)$. By Lemma~\ref{lem:lossconvexity}, we conclude that $\lambda(u,v)$ is concave in $(u,v)$ where $u>v$.
+\end{proof}
+
+\begin{lem}[Corollary 3.5 in~\citet{brascamp2002extensions}]\label{lem:lossconvexity}
+Let $F(x,y)\colon \mathbb{R}^{m+n}\rightarrow \mathbb{R}$ be an integrable function where $x\in \mathbb{R}^{m},y\in \mathbb{R}^n$. Let
+\[
+G(x) = \int_{\mathbb{R}^n}F(x,y)dy.
+\]
+If $F(x,y)$ is log concave in $(x,y)$, then $G(x)$ is log concave in $x$.
+\end{lem}
+
+\subsection{Auxiliary lemmas}\label{sec:lemma}
+This section collects lemmas that are useful for the proofs of the main theorems.
+
+\begin{defn}[Atomic M-norm~\citep{ghadermarzy2019near}]
+Define $T_{\pm}=\{\tT\in\{\pm 1 \}^{d_1\times \cdots \times d_K}\colon \text{rank}(\tT)=1\}$. The atomic M-norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
+\begin{align}
+\anormSize{}{\Theta}&=\inf\{t>0: \Theta\in t\text{conv}(T_{\pm})\}\\
+&=\inf\left\{\sum_{\tX\in T_{\pm}}c_x\colon \Theta=\sum_{\tX\in T_{\pm}} c_x\tX, \ c_x>0\right\}.
+\end{align}
+\end{defn}
+
+\begin{defn}[Spectral norm~\citep{lim2005singular}]
+The spectral norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
+\[
+\snormSize{}{\Theta}=\sup\left\{\langle \Theta, \mx_1\otimes \cdots \otimes \mx_K\rangle \colon \vnormSize{}{\mx_k}=1,\ \mx_k\in\mathbb{R}^{d_k},\ \text{for all}\ k\in[K]\right\}.
+\]
+\end{defn}
+
+\begin{defn}[Nuclear norm~\citep{friedland2018nuclear}]
+The nuclear norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
+\[
+\nnormSize{}{\Theta}=\inf
+\left\{
+\sum_{i\in[r]}|\lambda_i|\colon \Theta=\sum_{i=1}^r \lambda_i\mx^{(i)}_1\otimes \cdots \otimes \mx^{(i)}_K,\ \vnormSize{}{\mx^{(i)}_k}=1,\ \mx^{(i)}_k\in\mathbb{R}^{d_k},\ \text{for all}\ k\in[K],\ i\in[r]
+\right\},
+\]
+where the infimum is taken over all $r\in\mathbb{N}$ and $\vnormSize{}{\mx^{(i)}_k}=1$ for all $i\in[r]$ and $k\in[K]$.
+\end{defn}
+
+\begin{lem}[M-norm and infinity norm~\citep{ghadermarzy2019near}]\label{lem:Mnormbound}
+Let $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ be an order-$K$, rank-$(r_1,\ldots,r_K)$ tensor. Then
+\[
+\mnormSize{}{\Theta}\leq \anormSize{}{\Theta}\leq \left(\prod_k r_k \over r_{\max}\right)^{3\over 2} \mnormSize{}{\Theta}.
+\]
+\end{lem}
+
+\begin{lem}[Nuclear norm and F-norm] \label{lem:nuclear}
+Let $\tA\in\mathbb{R}^{d_1\times\cdots\times d_K}$ be an order-$K$ tensor with Tucker $\text{rank}(\tA)=(r_1,\ldots,r_K)$. Then
+\[
+\nnormSize{}{\tA} \leq \sqrt{\prod_k r_k\over \max_k r_k} \FnormSize{}{\tA},
+\]
+where $\nnormSize{}{\cdot}$ denotes the nuclear norm of the tensor.
+\end{lem}
+
+\begin{proof}
+Without loss of generality, suppose $r_1=\min_k r_k$.
+Let $\tA_{(k)}$ denote the mode-$k$ matricization of $\tA$ for all $k\in[K]$.
+By \citet[Corollary 4.11]{wang2017operator}, and the invariance relationship between a tensor and its Tucker core~\citep[Section 6]{jiang2017tensor}, we have
+\begin{equation}\label{eq:norminequality}
+\nnormSize{}{\tA} \leq \sqrt{\prod_{k\geq 2} r_k \over \max_{k\geq 2} r_k} \nnormSize{}{\tA_{(1)}},
+\end{equation}
+where $\tA_{(1)}$ is a $d_1$-by-$\prod_{k\geq 2}d_k$ matrix with matrix rank $r_1$.
+Furthermore, the relationship between the matrix norms implies that $\nnormSize{}{\tA_{(1)}}\leq \sqrt{r_1}\FnormSize{}{\tA_{(1)}}=\sqrt{r_1}\FnormSize{}{\tA}$.
+Combining this fact with the inequality~\eqref{eq:norminequality} yields the final claim.
+\end{proof}
+
+\begin{lem} \label{lem:inq}
+Let $\tA, \; \tB$ be two order-$K$ tensors of the same dimension. Then
+\[
+|\langle \tA,\tB\rangle| \leq \snormSize{}{\tA}   \nnormSize{}{\tB}.
+\]
+\end{lem}
+
+\begin{proof}
+By~\citet[Proposition 3.1]{friedland2018nuclear}, there exists a nuclear norm decomposition of $\tB$, such that
+\[
+\tB=\sum_{r} \lambda_r \ma^{(1)}_r\otimes \cdots\otimes \ma^{(K)}_r,\quad \ma_r^{(k)}\in\mathbf{S}^{d_k-1}(\mathbb{R}),\quad \text{for all }k\in[K],
+\]
+and $\nnormSize{}{\tB}=\sum_{r}|\lambda_r|$. Henceforth we have
+\begin{align*}
+|\langle \tA,\tB\rangle|&=| \langle \tA, \sum_{r} \lambda_r \ma^{(1)}_r\otimes \cdots\otimes \ma^{(K)}_r \rangle|\leq \sum_r |\lambda_r| |\langle \tA, \ma^{(1)}_r \otimes \cdots\otimes \ma^{(K)}_r \rangle|\\
+&\leq \sum_{r}|\lambda_r| \snormSize{}{\tA}= \snormSize{}{\tA}\nnormSize{}{\tB},
+\end{align*}
+which completes the proof.
+\end{proof}
+
+The following lemma provides the bound on the spectral norm of random tensors. The result was firstly presented in~\citet{nguyen2015tensor}, and we adopt the version from~\citet{tomioka2014spectral}.
+\begin{lem}[Spectral norm of random tensors~\citep{tomioka2014spectral}]\label{lem:tensor}
+Suppose that $\tS=\entry{s_{\omega}}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is an order-$K$ tensor whose entries are independent random variables that satisfy
+\[
+\mathbb{E}(s_{\omega})=0,\quad \text{and} \quad\mathbb{E}(e^{ts_{\omega}})\leq e^{t^2L^2/2}.
+\]
+Then the spectral norm $\snormSize{}{\tS}$ satisfies that,
+\[
+\snormSize{}{\tS}\leq \sqrt{{8L^2} \log (12K) \sum_k d_k +\log (2/\delta)},
+\]
+with probability at least $1-\delta$.
+\end{lem}
+
+\begin{lem} \label{lem:noisytensor}
+Suppose that $\tS=\entry{s_{\omega}}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is an order-$K$ tensor whose entries are independent random variables that satisfy
+\[
+\mathbb{E}(s_{\omega})=0,\quad \text{and}\quad |s_{\omega}|\leq U.
+\]
+Then we have
+\[
+\mathbb{P}\left(\snormSize{}{\tS}\geq C_2 U\sqrt{\sum_k d_k} \right)\leq \exp\left(-C_1  \log K \sum_k d_k\right)
+\]
+where $C_1>0$ is an absolute constant, and $C_2>0$ is a constant that depends only on $K$.
+\end{lem}
+
+\begin{proof}  Note that the random variable $U^{-1}s_{\omega}$ is zero-mean and supported on $[-1,1]$. Therefore, $U^{-1}s_{\omega}$ is sub-Gaussian with parameter ${1-(-1)\over 2}=1$, i.e.
+\[
+\mathbb{E}(U^{-1}s_{\omega})=0,\quad \text{and}\quad \mathbb{E}(e^{tU^{-1}s_{\omega}})\leq e^{t^2/2}.
+\]
+It follows from Lemma~\ref{lem:tensor} that, with probability at least $1-\delta$,
+\[
+\snormSize{}{U^{-1}\tS}\leq \sqrt{\left(c_0\log K+c_1\right) \sum_k d_k +\log (2/\delta)},
+\]
+where $c_0, c_1>0$ are two absolute constants. Taking $\delta=\exp (-C_1\log K \sum_k d_k)$ yields the final claim, where $C_2=c_0\log K+c_1+1>0$ is another constant.
+\end{proof}
+
+\begin{lem}\label{lem:KLentry} Let $X,\; Y$ be two discrete random variables taking values on $L$ possible categories, with point mass probabilities $\{p_\ell\}_{\ell\in[L]}$ and $\{q_\ell\}_{\ell\in[L]}$, respectively.
+Suppose $p_\ell$, $q_\ell>0$ for all $\ell\in[L]$.
+Then, the Kullback-Leibler (KL) divergence satisfies that
+\[
+\mathrm{KL}(X||Y)\stackrel{\text{def}}{=}-\sum_{\ell\in[L]}\mathbb{P}_X(\ell)\log\left\{{\mathbb{P}_Y(\ell)\over \mathbb{P}_X(\ell)}\right\} \leq \sum_{\ell \in [L]}{(p_\ell-q_\ell)^2 \over q_\ell}.
+\]
+\end{lem}
+\begin{proof} Using the fact $\log x\leq x-1$ for $x>0$, we have that
+\begin{align}\label{eq:KL}
+\text{KL}(X||Y)&=\sum_{\ell\in[L]}p_\ell\log{p_\ell\over q_\ell}\\
+&\leq \sum_{\ell\in[L]} {p_\ell\over q_\ell}(p_\ell-q_\ell)\\
+&=\sum_{\ell\in[L]} \left({p_\ell\over q_\ell }- 1\right)(p_\ell-q_\ell)+ \sum_{\ell\in[L]} (p_\ell-q_\ell).
+\end{align}
+Note that $\sum_{\ell\in[L]}(p_\ell-q_\ell)=0$. Therefore,
+\[
+\text{KL}(X||Y)\leq \sum_{\ell\in[L]}\left( {p_\ell\over q_\ell}-1\right)\left(p_\ell-q_\ell\right)=\sum_{\ell\in[L]}{(p_\ell-q_\ell)^2\over q_\ell}.
+\]
+\end{proof}
+
+\begin{lem}[KL divergence and F-norm]~\label{lem:KL}
+Let $\tY\in[L]^{d_1\times \cdots \times d_K}$ be an ordinal tensor generated from the model~\eqref{eq:model} with the link function $f$ and parameter tensor $\Theta$.
+Let $\mathbb{P}_{\Theta}$ denote the joint categorical distribution of $\tY|\Theta$ induced by the parameter tensor $\Theta$, where $\mnormSize{}{\Theta}\leq \alpha$.
+Define
+\begin{equation}\label{eq:ass}
+A_\alpha=\min_{\ell\in[L], |\theta|\leq \alpha}\left[f(b_\ell-\theta)-f(b_{\ell-1}-\theta)\right].
+\end{equation}
+Then, for any two tensors $\Theta,\; \Theta^*$ in the parameter spaces, we have
+\[
+\mathrm{KL}(\mathbb{P}_{\Theta}|| \mathbb{P}_{\Theta^*})\leq {2(2L-3)\over A_\alpha} \dot{f}^2(0)\FnormSize{}{\Theta-\Theta^*}^2.
+\]
+\end{lem}
+\begin{proof} Suppose that the distribution over the ordinal tensor $\tY=\entry{y_\omega}$ is induced by $\Theta=\entry{\theta_\omega}$. Then, based on the generative model~\eqref{eq:model},
+\[
+\mathbb{P}(y_\omega=\ell | \theta_\omega)=f(b_{\ell}-\theta_\omega)-f(b_{\ell-1}-\theta_\omega),
+\]
+for all $\ell\in[L]$ and $\omega\in[d_1]\times \cdots \times [d_K]$.
+For notational convenience, we suppress the subscribe in $\theta_\omega$ and simply write $\theta$ (and respectively, $\theta^*$).
+Based on Lemma~\ref{lem:KLentry} and Taylor expansion,
+\begin{align}
+\text{KL}(\theta|| \theta^*) & \leq \sum_{\ell\in[L]} {\left[f(b_{\ell}-\theta)-f(b_{\ell-1}-\theta)-f(b_{\ell}-\theta^*)+f(b_{\ell-1}-\theta^*)\right]^2\over f(b_{\ell}-\theta^*)-f(b_{\ell-1}-\theta^*)}\\
+ &\leq \sum_{\ell=2}^{L-1} {\left[\dot{f}(b_\ell-\eta_\ell)-\dot{f}(b_{\ell-1}-\eta_{\ell-1})\right]^2 \over f(b_\ell-\theta^*)-f(b_{\ell-1}-\theta^*)} \left(\theta-\theta^*\right)^2+{\dot{f}^2(b_1-\eta_1) \over f(b_1-\theta^*)} (\theta-\theta^*)^2\\
+ &\quad \quad\quad \quad  +{\dot{f}^2(b_{L-1}-\eta_{L-1})\over 1-f(b_{L-1}-\theta^*)}(\theta-\theta^*)^2,
+\end{align}
+where $\eta_\ell$ and $\eta_{\ell-1}$ fall between $\theta$ and $\theta^*$. Therefore,
+\begin{equation}\label{eq:entrywise}
+\text{KL}(\theta|| \theta^*) \leq \left({4(L-2)\over A_\alpha}+ {2\over A_\alpha}\right)\dot{f}^2(0)(\theta-\theta^*)^2={2(2L-3)\over A_\alpha} \dot{f}^2(0)(\theta-\theta^*)^2,
+\end{equation}
+where we have used Taylor expansion, the bound~\eqref{eq:ass}, and the fact that $\dot{f}(\cdot)$ peaks at zero for an unimodal and symmetric function.
+Now summing~\eqref{eq:entrywise} over the index set $\omega\in[d_1]\times \cdots \times [d_K]$ gives
+\[
+\text{KL}(\mathbb{P}_{\Theta}|| \mathbb{P}_{\Theta^*}) =\sum_{\omega\in[d_1]\times \cdots \times[d_K]} \text{KL}(\theta_\omega || \theta^*_\omega) \leq {2(2L-3)\over A_\alpha} \dot{f}^2(0)\FnormSize{}{\Theta-\Theta^*}^2.
+\]
+\end{proof}
+
+\begin{rmk} In particular, let $\mathbb{P}_{\bf{0}}$ denote the distribution of $\tY|\bf{0}$ induced by the zero parameter tensor. Then we have
+\[
+\text{KL}(\mathbb{P}_{\Theta}||\mathbb{P}_{\bf{0}})\leq {2(2L-3)\over A_\alpha}  \dot{f}^2(0)\FnormSize{}{\Theta}^2.
+\]
+\end{rmk}
+
+\begin{lem}\label{lem:construction}
+Assume the same setup as in Theorem~\ref{thm:minimax}.
+Without loss of generality, suppose $d_1=\max_kd_k$.
+Define $R=\max_k r_k$ and $d_{\text{total}}=\prod_{k\in[K]} d_k$.
+For any constant $0\leq \gamma \leq 1$, there exist a finite set of tensors $\tX=\{\Theta_i: i=1,\ldots\}\subset \tP$ satisfying the following four properties:
+\begin{enumerate}
+\item $\text{Card}(\tX)\geq 2^{Rd_1/8}+1$, where $\text{Card}$ denotes the cardinality;
+\item $\tX$ contains the zero tensor $\mathbf{0}\in\mathbb{R}^{d_1\times \cdots\times d_K}$;
+\item $\mnormSize{}{\Theta}\leq \gamma \min\left\{ \alpha ,\ C^{-1/2}\sqrt{Rd_1\over d_{\text{total}}} \right\} $ for any element $\Theta\in\tX$;
+\item $\FnormSize{}{\Theta_i-\Theta_j}\geq {\gamma\over 4} \min\left\{ \alpha\sqrt{d_{\text{total}}},\ C^{-1/2}\sqrt{Rd_1}\right\}$ for any two distinct elements $\Theta_i,\; \Theta_j\in\tX$,
+\end{enumerate}
+Here $C=C(\alpha,L,f,\mb)={(4L-6)\dot{f}^2(0)\over A_\alpha }>0$ is a constant independent of the tensor dimension and rank.
+\end{lem}
+
+\begin{proof}
+Given a constant $0\leq \gamma \leq 1$, we define a set of matrices:
+\[
+\tC=\left\{\mM=(m_{ij})\in\mathbb{R}^{d_1\times R}: a_{ij}\in \left\{ 0,\gamma \min\left\{ \alpha , C^{-1/2}\sqrt{Rd_1\over d_{\text{total}}} \right\}\right\} ,\  \forall (i,j)\in[d_1]\times[R]\right\}.
+\]
+We then consider the associated set of block tensors:
+\begin{align}
+\tB=\tB(\tC)=\{\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}\colon& \Theta=\mA\otimes \mathbf{1}_{d_3}\otimes \cdots \otimes \mathbf{1}_{d_K}, \ \text{where}\ \mA=(\mM|\cdots|\mM|\mO) \in\mathbb{R}^{d_1\times d_2},\ \mM\in\tC\},
+\end{align}
+where $\mathbf{1}_d$ denotes a length-$d$ vector with all entries 1, $\mO$ denotes the $d_1\times (d_2-R\lfloor d_2/R \rfloor)$ zero matrix, and $\lfloor d_2/ R \rfloor$ is the integer part of $d_2/R$.
+In other words, the subtensor $\Theta(\mI, \mI,i_3, \ldots,i_K)\in\mathbb{R}^{d_1\times d_2}$ are the same for all fixed $(i_3,\ldots,i_K)\in[d_3]\times \cdots \times [d_K]$, and furthermore, each subtensor $\Theta(\mI,\mI, i_3,\ldots,i_K)$ itself is filled by copying the matrix $\mM\in\mathbb{R}^{d_1\times R}$ as many times as would fit.
+
+By construction, any element of $\tB$, as well as the difference of any two elements of $\tB$, has Tucker rank at most $\max_k r_k\leq R$, and the entries of any tensor in $\tB$ take values in $[0,\alpha]$.
+Thus, $\tB\subset\tP$.
+By Lemma~\ref{lem:VGbound}, there exists a subset $\tX\subset \tB$ with cardinality $\text{Card}(\tX)\geq 2^{Rd_1/8}+1$ containing the zero $d_1\times \cdots \times d_K$ tensor, such that, for any two distinct elements $\Theta_i$ and $\Theta_j$ in $\tX$,
+\[
+\FnormSize{}{\Theta_i-\Theta_j}^2 \geq {Rd_1\over 8} \gamma^2\min\left\{ \alpha^2, {C^{-1}Rd_1 \over d_{\text{total}}}\right\} \Big\lfloor {d_2\over R}\Big \rfloor \prod_{k\geq 3}d_k\geq {\gamma^2\min\left\{ \alpha^2 d_{\text{total}}, C^{-1}Rd_1\right\}  \over 16}.
+\]
+In addition, each entry of $\Theta\in\tX$ is bounded by $\gamma \min\left\{ \alpha , C^{-1/2}\sqrt{Rd_1\over d_{\text{total}}}\right\} $. Therefore the Properties (i) to (iv) are satisfied.
+\end{proof}
+
+\begin{lem}[Varshamov-Gilbert bound]\label{lem:VGbound}
+Let $\Omega=\{(w_1,\ldots,w_m)\colon w_i\in\{0,1\}\}$. Suppose $m>8$. Then there exists a subset $\{w^{(0)},\ldots,w^{(M)}\}$ of $\Omega$ such that $w^{(0)}=(0,\ldots,0)$ and
+\[
+\zeronormSize{}{w^{(j)}-w^{(k)}}\geq {m\over 8},\quad \text{for } \ 0\leq j<k\leq M,
+\]
+where $\zeronormSize{}{\cdot}$ denotes the Hamming distance, and $M\geq 2^{m/8}$.
+\end{lem}
+
+\begin{lem}[Theorem 2.5 in~\citet{tsybakov2008introduction}]\label{lem:Tsybakov}
+Assume that a set $\tX$ contains element $\Theta_0, \Theta_1, \ldots,\Theta_M$ ($M\geq 2$) such that
+\begin{itemize}
+\item $d(\Theta_j,\ \Theta_k)\geq 2s>0$, $\forall 0\leq j\leq k\leq M$;
+\item $\mathbb{P}_0$ is absolutely continuous with respect to $\mathbb{P}_j$, $\forall j=1,\ldots,M$, and
+\[
+{1\over M}\sum_{j=1}^M \mathrm{KL}(\mathbb{P}_j||\mathbb{P}_0)\leq \alpha \log M
+\]
+where $d\colon \tX\times \tX\mapsto [0,+\infty]$ is a semi-distance function, $0<\alpha<{1/8}$ and $\mathbb{P}_j=\mathbb{P}_{\Theta_j}$, $j=0,1\ldots,M$.
+\end{itemize}
+Then
+\[
+\inf_{\hat \Theta}\sup_{\Theta\in\tX} \mathbb{P}_{\Theta}(d(\hat \Theta, \Theta)\geq s)\geq {\sqrt{M}\over 1+\sqrt{M}}\left(1-2\alpha-\sqrt{2\alpha\over \log M} \right)>0.
+\]
+\end{lem}
+
+\begin{lem}[Lemma 28 in~\citet{ghadermarzy2019near}]\label{lem:convexity}
+Define $\mathbb{B}_{M}(\beta)=\{\Theta\in \mathbb{R}^{d_1\times \cdots \times d_K}\colon \anormSize{}{\Theta}\leq \beta \}$.
+Let $\Omega\subset[d_1]\times\cdots \times [d_K]$ be a random set with $m=|\Omega|$, and assume that each entry in $\Omega$ is drawn with replacement from $[d_1]\times\cdots\times[d_K]$ using probability $\Pi$.
+Define
+\[
+\PiFnormSize{}{\Theta}^2={1\over m}\mathbb{E}_{\Omega\in\Pi}\MFnormSize{}{\Theta}^2.
+\]
+Then, there exists a universal constant $c>0$, such that, with probability at least $1-\exp\left(-{\sum_kd_k \over \sum_k \log d_k} \right)$ over the sampled set $\Omega$,
+\[
+{1\over m}\MFnormSize{}{\Theta}^2 \geq \PiFnormSize{}{\Theta}^2-c\beta\sqrt{\sum_k d_k\over m}
+\]
+holds uniformly for all $\Theta\in\mathbb{B}_M(\beta)$.
+\end{lem}
+
+\section{Extension of Theorem~\ref{thm:rate} to unknown cut-off points}\label{sec:extention}
+We now extend Theorem~\ref{thm:rate} to the case of unknown cut-off points $\mb$. Assume that the true parameters $(\trueT, \trueb)\in \tP\times \tB$, where the feasible sets are defined as
+\begin{align}
+\tP&=\{\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}\colon \text{rank}(\tP)\leq \mr,\ \langle \Theta, \tJ \rangle=0,\ \mnormSize{}{\Theta}\leq \alpha\},\\
+ \tB&=\{\mb\in\mathbb{R}^{L-1}\colon \mnormSize{}{\mb}\leq \beta,\ \min_{\ell}(b_\ell-b_{\ell-1}) \geq \Delta\},
+\end{align}
+with positive constants $\alpha,\beta,\Delta>0$ and a given rank $\mr\in\mathbb{N}^K_{+}$.
+Here, $\tJ=\entry{1}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ denotes a tensor of all ones.
+The constraint $\langle \Theta, \tJ \rangle=0$ is imposed to ensure the identifiability of $\Theta$ and $\mb$.
+We propose the constrained M-estimator
+\begin{equation}\label{eq:joint}
+(\hat \Theta,\hat \mb)=\argmax_{(\Theta,\mb)\in\tP\times \tB}\flogl(\Theta,\mb).
+\end{equation}
+The estimation accuracy is assessed using the mean squared error (MSE):
+\[
+\text{MSE}\left( \hat \Theta, \trueT \right)={1\over \prod_k d_k}\FnormSize{}{\hat \Theta-\trueT}^2,\quad \text{MSE}\left( \hat \mb, \trueb \right)={1\over L-1}\FnormSize{}{\hat \mb-\trueb}^2.
+\]
+
+To facilitate the examination of MSE, we define an order-$(K+1)$ tensor, $\tZ=\entry{z_{\omega,\ell}}\in\mathbb{R}^{d_1\times \cdots \times d_K\times (L-1)}$, by stacking the parameters $\Theta=\entry{\theta_\omega}$ and $\mb=\entry{b_\ell}$ together.
+Specifically, let $z_{\omega,\ell}=-\theta_\omega+b_\ell$ for all $\omega\in[d_1]\times \cdots \times [d_K]$ and $\ell\in[L-1]$; that is,
+\[
+\tZ=-\Theta\otimes \mathbf{1}+\tJ\otimes \mb,
+\]
+where $\mathbf{1}$ denotes a length-$(L-1)$ vector of all ones.
+Under the identifiability constraint $\langle \Theta, \tJ \rangle=0$, there is an one-to-one mapping between $\tZ$ and $(\Theta, \mb)$, with $\text{rank}(\tZ)\leq (\text{rank}(\Theta)+1,\ 2)^T$.
+Furthermore, 
+\begin{equation}\label{eq:decomp}
+\FnormSize{}{\hat \tZ-\trueZ}^2=\FnormSize{}{\hat \Theta-\trueT}^2 (L-1)+\FnormSize{}{\hat \mb-\trueb}^2 \left(\prod_kd_k\right),
+\end{equation}
+where $\trueZ=-\trueT\otimes \mathbf{1}+\tJ\otimes \trueb$ and $\hat \tZ=-\hat \Theta\otimes \mathbf{1}+\tJ\otimes \hat \mb$.
+
+We make the following assumptions about the link function.
+\begin{assumption}\label{ass:joint}
+The link function $f\colon \mathbb{R}\mapsto [0,1]$ satisfies the following properties:
+\begin{enumerate}
+\item $f(z)$ is twice-differentiable and strictly increasing in $z$.
+\item $\dot{f}(z)$ is strictly log-concave and symmetric with respect to $z=0$.
+\end{enumerate}
+\end{assumption}
+
+We define the following constants that will be used in the theory:
+\begin{align}\label{eq:constantZ}
+C_{\alpha,\beta,\Delta}&=\max_{|z|\leq \alpha+\beta}\max_{\substack{z'\leq z-\Delta\\z''\geq z+\Delta}}\max\left\{ {\dot{f}(z) \over f(z)-f(z')},\ {\dot{f}(z) \over f(z'')-f(z)}\right\},\\
+D_{\alpha,\beta,\Delta}&=\min_{|z|\leq \alpha+\beta}\min_{\substack{z'\leq z-\Delta\\z''\geq z+\Delta}}\min\left\{- {\partial\over \partial z}\left({\dot{f}(z) \over f(z)-f(z')}\right),\  {\partial\over \partial z}\left({\dot{f}(z) \over f(z'')-f(z)}\right) \right\},\\
+A_{\alpha,\beta,\Delta}&=\min_{|z|\leq \alpha+\beta}\min_{z'\leq z-\Delta} \left(f(z)-f(z')\right).
+\end{align}
+
+\begin{rmk}
+The condition $\Delta=\min_{\ell}(b_{\ell}-b_{\ell-1})>0$ on the feasible set $\tB$ guarantees the strict positiveness of $f(z)-f(z')$ and $f(z'')-f(z)$.
+Therefore, the denominators in the above quantities $C_{\alpha,\beta,\Delta},\ D_{\alpha,\beta,\Delta}$ are well-defined.
+Furthermore, by Theorem~\ref{thm:convexity}, $f(z)-f(z')$ is strictly log-concave in $(z,z')$ for $z\leq z'-\Delta,\ z,z'\in[-\alpha-\beta,\ \alpha+\beta]$.
+Based on Assumption~\ref{ass:joint} and closeness of the feasible set, we have $C_{\alpha,\beta,\Delta}>0$, $D_{\alpha,\beta,\Delta}>0$, $A_{\alpha,\beta,\Delta}>0$.
+\end{rmk}
+\begin{rmk} In particular, for logistic link $f(x)= \frac{1}{1+e^{-x}}$, we have
+\begin{align}
+C_{\alpha,\beta,\Delta}&=\max_{|z|\leq \alpha+\beta}\max_{\substack{z'\leq z-\Delta\\z''\geq z+\Delta}}\max\left\{ \frac{1}{e^\Delta-1}\left(\frac{1+e^{-z'}}{1+e^{-z}}\right),\ \frac{1}{1-e^{-\Delta}}\left(\frac{1+e^{-z''}}{1+e^{-z}}\right)\right\}>0,\\
+D_{\alpha,\beta,\Delta}&=\min_{|z|\leq \alpha+\beta}\frac{e^{-z}}{(1+e^{-z})^2}>0.\end{align}
+\end{rmk}
+\begin{thm}[Statistical convergence with unknown $\mb$]\label{thm:ratejoint}
+Consider an ordinal tensor $\tY\in[L]^{d_1\times \cdots \times d_K}$ generated from model~\eqref{eq:model} with the link function $f$ and parameters $(\trueT, \trueb)\in\tP\times \tB$.
+Suppose the link function $f$ satisfies Assumption~\ref{ass:joint}.
+Define $r_{\max}=\max_k r_k$, and assume $r_{\max}=\tO(1)$.
+
+Then with very high probability, the estimator in~\eqref{eq:joint} satisfies
+\begin{equation}\label{eq:scoreZ}
+\FnormSize{}{\hat \tZ-\trueZ}^2 \leq {c_1r^{K}_{\max}C^2_{\alpha,\beta,\Delta}\over A^2_{\alpha,\beta,\Delta}D^2_{\alpha,\beta,\Delta}}\left(L-1+\sum_k d_k\right),
+\end{equation}
+In particular,
+\begin{equation}\label{eq:jointTheta}
+\mathrm{MSE}\left(\hat \Theta, \trueT \right)\leq\min\left\{ 4\alpha^2,\  {c_1 r_{\max}^{K}  C_{\alpha,\beta,\Delta}^2 \over A^2_{\alpha,\beta,\Delta}D_{\alpha,\beta,\Delta}^2}\left({ L-1+\sum_k d_k \over \prod_k d_k}\right)\right\},
+\end{equation}
+and
+\begin{equation}\label{eq:jointb}
+\mathrm{MSE}\left(\hat \mb, \trueb \right)\leq \min\left\{4\beta^2,\ {c_1 r^{K}_{\max} C^2_{\alpha,\beta,\Delta} \over A^2_{\alpha,\beta,\Delta}D_{\alpha,\beta,\Delta}^2}\left({L-1+\sum_kd_k \over \prod_k d_K}\right)\right\},
+\end{equation}
+where $c_1, C_{\alpha,\beta,\Delta}, D_{\alpha,\beta,\Delta}$ are positive constants independent of the tensor dimension, rank, and number of ordinal levels. 
+\end{thm}
+\begin{proof} 
+The log-likelihood associated with the observed entries in terms of $\tZ$ is
+\begin{equation}\label{eq:loglz}
+\flogl(\tZ) = \sum_{\omega\in\Omega}\sum_{\ell\in[L]}\mathds{1}\{y_\omega=\ell\}\log\left[f(z_{\omega,\ell})-f(z_{\omega,\ell-1})\right].
+\end{equation}
+
+Let $\plogZ=\entry{{\partial \tL_\tY\over \partial z_{\omega,\ell}}}\in\mathbb{R}^{d_1\times \cdots \times d_K\times[L-1]}$ denote the score function, and $\mH=\pplogZ$ the Hession matrix. 
+Based on the definition of $\hat\tZ$, we have the following inequality:
+\begin{equation}\label{eq:ineql}
+\flogl(\hat\tZ) \geq \flogl(\trueZ).
+\end{equation}
+
+Following the similar argument in Theorem \ref{thm:rate} and the inequality \eqref{eq:ineql}, we obtain that
+\begin{equation}\label{eq:boundZ}
+\FnormSize{}{\hat \tZ-\trueZ}^2 \leq c_1r^{K}_{\max}{\snormSize{}{\plogZ(\trueZ)}^2\over \lambda^2_{1} \left(\mH(\check\tZ)\right)},
+\end{equation}
+where $\plogZ(\trueZ)$ is the score evaluated at $\trueZ$, $\mH(\check \tZ)$ is the Hession evaluated at $\check\tZ$, for some $\check\tZ$ between $\hat \tZ$ and $\trueZ$,  and $\lambda_{1}(\cdot)$ is the largest matrix eigenvalue.
+
+We bound the score and the Hessian to obtain \eqref{eq:scoreZ}.
+
+\begin{enumerate}
+\item (Score.) The $(\omega,\ell)$-th entry in $\plogZ$ is 
+\[
+{\partial \tL_\tY\over \partial z_{\omega,\ell}}=\mathds{1}\{y_\omega=\ell\}{\dot{f}(z)\over f(z)-f(z')}\Bigg|_{(z,\ z')=(z_{\omega,\ell},\ z_{\omega,\ell-1})} - \mathds{1}\{y_\omega=\ell+1\}{\dot{f}(z)\over f(z'')-f(z)}\Bigg|_{(z'',\ z)=(z_{\omega,\ell+1},\ z_{\omega,\ell})},
+\]
+which is upper bounded in magnitude by $C_{\alpha,\beta,\Delta}>0$ with zero mean. By Lemma \ref{lem:noisytensor}, with probability at least $1-\exp\left(-c_2'
+\left(\sum_k d_k+ L-1\right)\right)$, we have
+\begin{align}\label{eq:scorebound}
+\snormSize{}{\plogZ(\trueZ)}\leq c_2C_{\alpha,\beta,\Delta}\sqrt{L-1+\sum_k d_k},
+\end{align}
+
+where $c_2,c_2'$ are two positive constants that depend only on $K$.
+\item (Hession.) The entries in the Hession matrix are
+\begin{align}
+\text{Diagonal: }&{\partial^2 \tL_\tY\over \partial z^2_{\omega,\ell}}=\mathds{1}\{y_\omega=\ell\}{\ddot{f}(z)\left(f(z)-f(z')\right)-\dot{f}^2(z)\over \left(f(z)-f(z')\right)^2}\Bigg|_{(z,\ z')=(z_{\omega,\ell},\ z_{\omega,\ell-1})}-\\
+&\hspace{.65in}\mathds{1}\{y_\omega=\ell+1\}{\ddot{f}(z)\left(f(z'')-f(z)\right)+\dot{f}^2(z)\over \left(f(z'')-f(z)\right)^2}\Bigg|_{(z'',\ z)=(z_{\omega,\ell+1},\ z_{\omega,\ell})},\\
+\text{Off-diagonal: }&
+{\partial^2 \tL_\tY\over \partial z_{\omega,\ell}z_{\omega,\ell+1}}=\mathds{1}\{y_\omega=\ell+1\}{\dot{f}(z_{\omega,\ell})\dot{f}(z_{\omega,\ell+1})\over \left(f(z_{\omega,\ell+1})-f(z_{\omega,\ell})\right)^2}\quad \text{and}\quad {\partial^2 \tL_\tY\over \partial z_{\omega,\ell}z_{\omega',\ell'}}=0 \text{ otherwise}.
+\end{align}
+
+Based on Assumption~\ref{ass:joint}, the Hession matrix $\mH$ has the following three properties:
+\begin{enumerate}
+\item The Hession matrix is a block matrix, $\mH=\text{diag}\{\mH_\omega\colon \omega\in[d_1]\times \cdots \times [d_K]\}$, and each block $\mH_{\omega}\in\mathbb{R}^{(L-1)\times (L-1)}$ is a tridiagonal matrix.
+\item The off-diagonal entries are either zero or strictly positive.
+\item The diagonal entries are either zero or strictly negative. Furthermore, 
+\begin{align}
+&\mH_\omega(\ell,\ell)+ \mH_\omega(\ell,\ell-1)+\mH_\omega(\ell,\ell+1)\\
+=&{\partial^2 \tL_\tY\over \partial z^2_{\omega,\ell}}+{\partial^2 \tL_\tY\over \partial z_{\omega,\ell}z_{\omega,\ell+1}}+{\partial^2 \tL_\tY\over \partial z_{\omega,\ell-1}z_{\omega,\ell}}\\
+=&\mathds{1}\{y_\omega=\ell\}{\partial\over \partial z}\left({\dot{f}(z) \over f(z)-f(z')}\right)\Bigg|_{(z,\ z')=(z_{\omega,\ell},\ z_{\omega,\ell-1})}  -\mathds{1}\{y_\omega=\ell+1\}{\partial\over \partial z}\left({\dot{f}(z) \over f(z)-f(z')}\right)\Bigg|_{(z'',\ z)=(z_{\omega,\ell+1},\ z_{\omega,\ell})}\\
+\leq &-D_{\alpha,\beta,\Delta}\mathds{1}\{y_\omega = \ell \text{ or } \ell+1\}.
+\end{align}
+\end{enumerate}
+We will show that, with very high probability over $\tY$, $\mH$ is negative definite in that
+\begin{equation}\label{eq:HessionZ}
+\lambda_{1}(\mH)=\max_{\mz\neq 0}{\mz^T\mH\mz\over\FnormSize{}{\mz}^2} \leq - cA_{\alpha,\beta,\Delta}D_{\alpha,\beta,\Delta},
+\end{equation}
+where $A_{\alpha,\beta,\Delta},\ D_{\alpha,\beta,\Delta}>0$ are constants defined in~\eqref{eq:constantZ}, and $c>0$ is a constant.
+
+Let $\mz_\omega=(z_{\omega,1},\ldots,z_{\omega,L-1})^T\in\mathbb{R}^{L-1}$ and $\mz=(\mz_{1,\ldots,1,1},\ldots,\mz_{d_1,\ldots,d_K,L-1})^T\in\mathbb{R}^{(L-1)\prod_k d_k}$.
+It follows from property (a) that
+\[
+\mz^T\mH\mz =\sum_{\omega}\mz_\omega^T \mH_{\omega} \mz_\omega.
+\]
+Furthermore, from properties (b) and (c) we have
+\begin{align}
+\mz_\omega^T \mH_{\omega} \mz_\omega &=
+\sum_{\ell\in[L-1]}\mH_\omega(\ell,\ell)z_{\omega,\ell}^2+\sum_{\ell\in[L-1]/\{1\}}2\mH_{\omega}(\ell,\ell-1)z_{\omega,\ell}z_{\omega,\ell-1}\\
+&\leq \sum_{\ell\in[L-1]}\mH(\ell,\ell)z^2_{\omega,\ell}+\sum_{\ell\in[L-1]/\{1\}}\mH(\ell,\ell-1)\left[z_{\omega,\ell}^2+z_{\omega,\ell-1}^2\right]\\
+&= \left(\mH(1,1)+\mH(1,2)\right)z_{\omega,1}^2+ \left(\mH(L-1,L-1)+\mH(L-1,L-2)\right)z_{\omega,L-1}^2\\&\quad\quad+\sum_{\ell\in[L-2]/\{1\}}\left(\mH(\ell,\ell)+\mH(\ell,\ell-1)+\mH(\ell,\ell+1)\right)z_{\omega,\ell}^2\\
+& \leq -D_{\alpha,\beta,\Delta} \sum_{\ell}z^2_{\omega,\ell}\mathds{1}\{y_\omega=\ell \text{ or }\ell+1\}.
+\end{align}
+Therefore,
+\begin{equation}\label{eq:sum}
+\mz^T\mH\mz =\sum_{\omega}\mz_\omega^T \mH_{\omega} \mz_\omega \leq -D_{\alpha,\beta,\Delta}\sum_{\omega}\sum_{\ell} z^2_{\omega,\ell}\mathds{1}\{y_\omega=\ell \text{ or }\ell+1\}.
+\end{equation}
+
+Define the subspace:
+\[
+\tS=\{\Vec(\tZ): \tZ=-\Theta\otimes \mathbf{1}+\tJ\otimes \mb,\ (\Theta,\mb)\in(\tP,\tB)\}.
+\]
+It suffices to prove the negative definiteness of Hession when restricted in the subspace $\tS$. Specifically, for any vector $\mz=\entry{z_{\omega,\ell}}\in\tS$,
+\begin{align}
+\sum_{\omega,\ell} z^2_{\omega,\ell}\mathds{1}\{y_\omega=\ell \text{ or }\ell+1\}&=\sum_{\omega,\ell}(-\theta_\omega+b_\ell)^2\mathds{1}\{y_\omega=\ell \text{ or }\ell+1\}\\
+&=\sum_{\omega,\ell}(\theta^2_\omega-2\theta_\omega b_\ell + b^2_\ell)\mathds{1}\{y_\omega=\ell \text{ or }\ell+1\}\\
+&=\sum_{\omega,\ell}\theta^2_\omega\mathds{1}\{y_\omega=\ell \text{ or }\ell+1\}-2\sum_{\omega,\ell} \theta_\omega b_\ell \mathds{1}\{y_\omega=\ell \text{ or }\ell+1\}+\sum_{\omega,\ell}b^2_\ell\mathds{1}\{y_\omega=\ell \text{ or }\ell+1\}\\
+&\geq \sum_{\omega}\theta^2_\omega -2\sum_{\omega,\ell}\theta_\omega b_\ell+\sum_{\ell}b_\ell^2\left(n_{\ell}+n_{\ell+1}\right)\\
+&\geq \sum_{\omega}\theta^2_\omega+\min_\ell\left(n_{\ell}+n_{\ell+1}\right)\sum_{\ell}b_\ell^2
+\end{align}
+
+On the other hand,
+\[
+\FnormSize{}{\mz}^2=\sum_{\omega,\ell}z^2_{\omega,\ell}=\sum_{\omega,\ell}(-\theta_\omega+b_\ell)^2=L_{\text{total}}\sum_{\omega}\theta^2_\omega+d_{\text{total}}\sum_\ell b^2_\ell,
+\]
+where $L_{\text{total}}:=(L-1)$ and $d_{\text{total}}:=\prod_k d_k$.
+
+Therefore, we have
+\begin{align}\label{eq:negativeH}
+\max_{\mz\in\tS,\mz\neq \mathbf{0}}{\sum_{\omega,\ell} z^2_{\omega,\ell}\mathds{1}\{\{y_\omega=\ell \text{ or }\ell+1\}\} \over \FnormSize{}{\mz}^2}&\geq {\sum_{\omega}\theta^2_\omega+\min_\ell\left(n_{\ell}+n_{\ell+1}\right)\sum_{\ell}b_\ell^2\over L_{\text{total}} \sum_{\omega}\theta^2_\omega+d_{\text{total}}\sum_\ell b^2_\ell}
+\geq{\min_{\ell}(n_\ell+n_{\ell+1})\over (1+{\alpha^2\over c\Delta^2})d_{\text{total}}}\notag\\
+&\geq {2A_{\alpha,\beta,\Delta}\over 1+{\alpha^2\over c\Delta^2}}\quad \text{in high probability as $d_{\min}\to \infty$}. 
+\end{align}
+The second inequality in \eqref{eq:negativeH} is from the condition that 
+\[
+\sum_{\omega}\theta_\omega^2\in [0,\ \alpha^2 d_{\text{total}})\quad \text{and}\quad \sum_\ell b_\ell^2 \in[c L_{\text{total}}\Delta^2,\  L_{\text{total}}\beta^2],
+\]
+for some universal constant $c>0$. 
+The last inequality in \eqref{eq:negativeH} follows by applying the law of large numbers and the uniform bound $\min_{z_{\omega,\ell}}\mathbb{P}(y_{\omega}=\ell \text{ or }\ell+1|z_{\omega,\ell})\geq 2A_{\alpha,\beta,\Delta}$ to the empirical ratio:
+\[
+{\min_{\ell}(n_\ell+n_{\ell+1})\over d_{\text{total}}}\overset{p}{\to}\min_{\ell} \mathbb{P}(y_{\omega}=\ell \text{ or }\ell+1|z_{\omega,\ell}) \geq 2A_{\alpha,\beta,\Delta}, \quad \text{ in high probability as $d_{\min}\to \infty$}.
+\]
+
+By~\eqref{eq:sum} and~\eqref{eq:negativeH}, we have
+\begin{equation}\label{eq:boundH}
+\mz^T\mH\mz \leq- c'A_{\alpha,\beta,\Delta}D_{\alpha,\beta,\Delta}\FnormSize{}{\mz}^2,
+\end{equation}
+for some constant $c'>0$, therefore~\eqref{eq:HessionZ} is proved.
+\end{enumerate}
+
+ Plugging~\eqref{eq:scorebound} and~\eqref{eq:HessionZ} into~\eqref{eq:boundZ} yields
+\[
+\FnormSize{}{\hat \tZ-\trueZ}^2 \leq {c_1r^{K}_{\max}C^2_{\alpha,\beta,\Delta}\over A^2_{\alpha,\beta,\Delta}D^2_{\alpha,\beta,\Delta}}\left(L-1+\sum_k d_k\right).
+\]
+The MSEs for $\hat \Theta$ and $\hat \mb$ readily follow from~\eqref{eq:decomp}.
+
+\end{proof}
+
+\section{Additional results on simulations}
+We also compare the methods by their performance in predicting the median labels, $y_\omega^{\text{median}}=\min\{\ell\colon\mathbb{P}(y_\omega=\ell)\geq 0.5\}$.
+Under the latent variable model~\eqref{eq:latent} and Assumption~\ref{ass:link}, the median label is the quantized $\theta_\omega$ without noise; i.e.\ $y_\omega^{\text{median}}=\sum_\ell \mathds{1}\{\theta_\omega\in(b_{\ell-1},b_\ell]\}$.
  We utilize the same simulation setting as in the earlier experiment.
@@
 The improved accuracy comes from the incorporation of multilinear low-rank structure, multi-level observations, and the ordinal structure.
-Interestingly, for the three multilevel methods ({\bf 1bit-sign-T}, {\bf Ordinal-M}, and {\bf Ordinal-T}), the median estimator tends to yield smaller MAD than the mode estimator, $\text{MAD}(\tY^{\text{median}}, \hat \tY^{\text{median}})\leq \text{MAD}(\tY^{\text{mode}}, \hat \tY^{\text{mode}})$  (Figures~\ref{fig:compare}a-b vs.\ Figures~\ref{fig:compare2}a-b).
-On the other hand, the mode estimator tends to yield smaller MCR than the median estimator, $\text{MCR}(\tY^{\text{mode}}, \hat \tY^{\text{mode}})\leq\text{MCR}(\tY^{\text{median}}, \hat \tY^{\text{median}})$ (Figures~\ref{fig:compare}c-d vs.\ Figures~\ref{fig:compare2}c-d).
-This tendency is from the property that the median estimator $\hat y^{(\text{median})}_\omega$ minimizes $R_1(z)=\mathbb{E}_{y_\omega}|y_\omega-z|$, whereas the mode estimator $\hat y^{(\text{mode})}_\omega $ minimizes $R_2(z)=\mathbb{E}_{y_\omega}\mathds{1}_{\{y_\omega=z\}}$.
+Interestingly, the median estimator tends to yield smaller MAD than the mode estimator, $\text{MAD}(\tY^{\text{median}}, \hat \tY^{\text{median}})\leq \text{MAD}(\tY^{\text{mode}}, \hat \tY^{\text{mode}})$  (Figures~\ref{fig:compare}a-b vs.\ Figures~\ref{fig:compare2}a-b), for the three multilevel methods ({\bf 1bit-sign-T}, {\bf Ordinal-M}, and {\bf Ordinal-T}).
+The mode estimator, on the other hand, tends to yield smaller MCR than the median estimator, $\text{MCR}(\tY^{\text{mode}}, \hat \tY^{\text{mode}})\leq\text{MCR}(\tY^{\text{median}}, \hat \tY^{\text{median}})$ (Figures~\ref{fig:compare}c-d vs.\ Figures~\ref{fig:compare2}c-d).
+This tendency is from the property that the median estimator $\hat y^{(\text{median})}_\omega$ minimizes $R_1(z)=\mathbb{E}_{y_\omega}|y_\omega-z|$, whereas the mode estimator $\hat y^{(\text{mode})}_\omega $ minimizes $R_2(z)=\mathbb{E}_{y_\omega}\mathds{1}\{y_\omega=z\}$.
 Here the expectation is over the categorical distribution of $y_\omega$ given parameters $\hat \Theta$ and $\hat \mb$.
@@
 \begin{center}
-\includegraphics[width=.7\textwidth]{panel3.pdf}
+\includegraphics[width=.55\textwidth]{panel3.pdf}
 \end{center}
@@
 
-\section{Data Applications}
-We apply our ordinal tensor method to two real-world datasets.
-In the first application, we use our model to analyze an ordinal tensor consisting of structural connectivities among 68 brain nodes for 136 individuals from Human Connectome Project (HCP)~\citep{van2013wu}.
-In the second application, we perform tensor completion to an ordinal dataset with missing values.
-The data tensor records the ratings on a scale of 1 to 5 from 42 users to 139 songs on 26 contexts~\citep{baltrunas2011incarmusic}.
-
-\subsection{Human Connectome Project (HCP)}
-
-Each entry in the HCP dataset takes value on a nominal scale, \{{\it high, moderate, low}\}, indicating the strength level of fiber connection.
-We convert the dataset to a 3-level ordinal tensor $\tY\in[3]^{68\times 68\times 136}$ and apply the ordinal tensor method with a logistic link function.
-The BIC suggests $\mr = (23,23,8)$ with $\logl(\hat{\Theta},\hat{\mb}) = -216,646$.
-Based on the estimated Tucker factors $\{\hat \mM_k\}$, we perform a clustering analysis via K-mean on the brain nodes (see Appendix~\ref{sec:additionalHCP} for detailed procedure).
-The 68 brain nodes are grouped into 11 clusters.
-We find that the clustering captures the spatial separation between brain regions very well (Table~\ref{table:clustering}).
+\section{Additional explanations of HCP analysis}\label{sec:additionalHCP}
+We perform clustering analyses based on the Tucker representation of the estimated signal tensor $\hat\Theta$.
+The procedure is motivated from the higher-order extension of Principal Component Analysis (PCA) or Singular Value Decomposition (SVD).
+Recall that, in the matrix case, we perform clustering on an $m\times n$ (normalized) matrix $\mX$ based on the following procedure.
+First, we factorize $X$ into
+\begin{equation}
+\mX = \mU\mSigma \mV^T,
+\end{equation}
+where $\mSigma$ is a diagonal matrix and $\mU,\mV$ are factor matrices with orthogonal columns.
+Second, we take each column of $\mV$ as a principal axis and each row in $\mU\mSigma$ as principal component.
+A subsequent multivariate clustering method (such as $K$-means) is then applied to the $m$ rows of $\mU\mSigma$.
+
+We apply a similar clustering procedure to the estimated signal tensor $\hat\Theta$. We factorize $\hat \Theta$ based on Tucker decomposition.
+\begin{equation}\label{eq:Tuckerest}
+\hat \Theta = \hat \tC\times_1\hat \mM_1\times_2\cdots\times_K\hat \mM_K,
+\end{equation}
+where $\hat \tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is the estimated core tensor, $\hat \mM_k\in\mathbb{R}^{d_k\times r_k}$ are estimated factor matrices with orthogonal columns, and $\times_k$ denotes the tensor-by-matrix multiplication~\citep{kolda2009tensor}.
+The mode-$k$ matricization of~\eqref{eq:Tuckerest} gives
+\begin{equation}
+\hat \Theta_{(k)} = \hat \mM_k\hat \tC_{(k)}\left(\hat \mM_K\otimes\cdots\otimes\hat \mM_1\right),
+\end{equation}
+where $\hat \Theta_{(k)} , \hat \tC_{(k)}$ denote the mode-$k$ unfolding of $\hat \Theta$ and $\hat \tC$, respectively.
+We conduct clustering on this  the mode-$k$ unfolded signal tensor.
+We take columns in $\left(\hat \mM_K\otimes\cdots\otimes\hat \mM_1\right)$ as principal axes and rows in $\hat \mM_k\hat \tC_{(k)}$ as principal components.
+Then, we apply $K$-means clustering method to the $d_k$ rows of the matrix $\hat \mM_k\hat \tC_{(k)}$.
+
+We perform a clustering analysis on the 68 brain nodes using the procedure described above.
+Our ordinal tensor method outputs the estimated parameter tensor $\hat\Theta\in\mathbb{R}^{68\times 68\times136}$ with rank $(23,23,8)$.
+We apply $K$-means to the mode-1 principal component matrix of size $68\times 184$ ($184=23\times8$).
+The elbow method suggests 11 clusters among the 68 nodes (see Figure~\ref{figure:elbow}).
+We find that the clustering successfully captures the spatial separation between brain regions (Table~\ref{table:clustering}).
 In particular, cluster I represents the connection between the left and right hemispheres, whereas clusters II-III represent the connection within each of the half brains (Figure~\ref{figure:brain image}).
@@
 
-We also compare the goodness-of-fit of various tensor methods on the HPC data.
-We perform 5-fold cross-validation while preserving the same label proportions in train/test sets.
-Table~\ref{table:CV} summarizes the prediction error averaged over 10 runs.
-Our method outperforms the others, especially in MAD.
-
-\subsection{InCarMusic recommendation system}
-
-We apply ordinal tensor completion to a recommendation system InCarMusic.
-InCarMusic is a mobile application that offers music recommendation to passengers of cars based on contexts~\citep{baltrunas2011incarmusic}.
-Our goal is to perform tensor completion to impute the unobserved entries in the $42\times 139\times 26$ ordinal tensor and thereby we can offer context-specific music recommendation to users.
-The data tensor consists of 2,884 observed entries.
-Table~\ref{table:CV} shows the averaged prediction error via 5-fold cross validation.
-The high missing rate makes the accurate classification challenging.
-Nevertheless, our method achieves the best performance among the three.
-
-\begin{table}[ht]
-\resizebox{\columnwidth}{!}{
-\begin{tabular}{c|c|c||c|c|c}
-\Xhline{2\arrayrulewidth}
-\multicolumn{3}{c||}{Human Connectome Project (HCP) dataset} & \multicolumn{3}{c}{InCarMusic dataset}\\
-\Xhline{2\arrayrulewidth}
-Method                    & MAD & MCR & Method                    & MAD & MCR\\
-\hline
-Ordinal-T (ours) & 0.1607 (0.0005)& 0.1606 (0.0005)& Ordinal-T (ours) & 1.37 (0.039) &  0.59 (0.009)\\
-Continuous-T&0.2530 (0.0002)&0.1599 (0.0002)& Continuous-T & 2.39 (0.152)&0.94 (0.027)\\
-1bit-sign-T&0.3566 (0.0010)&0.1563 (0.0010)& 1bit-sign-T&1.39 (0.003)& 0.81 (0.005)\\
-\hline
-\end{tabular}
-}
-\caption{Comparison of prediction error in the HPC and InCarMusic analyses. Standard errors are reported in parentheses.}\label{table:CV}
-\end{table}
-
-\section{Proofs}\label{sec:proof}
-Here, we provide proofs of the theoretical results presented in Sections~\ref{sec:theory}.
-
-\subsection{Estimation error for tensor denoising}\label{sec:proofMSE}
-\begin{proof}[Proof of Theorem~\ref{thm:rate}]
-We suppress the subscript $\Omega$ in the proof, because the tensor denoising assumes complete observation $\Omega=[d_1]\times \cdots \times [d_K]$.
-It follows from the expression of $\flogl(\Theta)$ that
-\begin{align}\label{eq:property}
-{\partial \flogl\over \partial \theta_\omega}&=\sum_{\ell\in[L]}\mathds{1}_{\{y_{\omega}=\ell\}}
-{\dot{g}_\ell(\theta_\omega)\over g_\ell(\theta_\omega)},\notag\\
-{\partial^2 \flogl\over \partial \theta_\omega^2}&=\sum_{\ell\in[L]}\mathds{1}_{\{y_\omega=\ell\}}{\ddot{g}_\ell(\theta_\omega)g_\ell(\theta_\omega)-\dot{g}^2_\ell(\theta_\omega)\over g^2_\ell(\theta_\omega)}\ \text{and}\quad
-{\partial^2 \flogl\over \partial \theta_\omega \theta_\omega'}=0\ \text{if}\ \omega\neq \omega',
-\end{align}
-for all $\omega\in[d_1]\times \cdots \times [d_K]$.
-Define $d_{\text{total}}=\prod_k d_k$.
-Let $\fplogl\in\mathbb{R}^{d_1\times\cdots\times d_K}$ denote the tensor of gradient with respect to $\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}$, and $\fpplogl$ the corresponding Hession matrix of size $d_\text{total}$-by-$d_{\text{total}}$.
-Here, $\Vec(\cdot)$ denotes the operation that turns a tensor into a vector.
-By~\eqref{eq:property}, $\fpplogl$ is a diagonal matrix.
-Recall that
-\begin{equation}\label{eq:bound}
-U_\alpha=\max_{\ell\in[L],|\theta|\leq \alpha}{|\dot{g}_\ell(\theta)|\over g_\ell(\theta)}>0 \quad \text{and}\quad
-L_\alpha=\min_{\ell\in[L],|\theta|\leq \alpha} {\dot{g}^2_\ell(\theta)-\ddot{g}_\ell(\theta)g_\ell(\theta)\over g^2_\ell(\theta)}>0.
-\end{equation}
-Therefore, the entries in $\fplogl$ are upper bounded in magnitude by $U_\alpha>0$, and all diagonal entries in $\fpplogl$ are upper bounded by $-L_{\alpha}<0$.
-
-By the second-order Taylor's expansion of $\flogl(\Theta)$ around $\trueT$, we obtain
-\begin{equation}\label{eq:taylor}
-\flogl(\Theta)=\flogl(\trueT)+\langle\Vec(\fplogl),\ \Vec(\Theta-\trueT)\rangle+{1\over 2}\Vec(\Theta-\trueT)^T\fpplogl(\check\Theta)\Vec(\Theta-\trueT),
-\end{equation}
-$\check\Theta=\gamma\trueT+(1-\gamma)\Theta$ for some $\gamma\in[0,1]$, and $\fpplogl(\check\Theta)$ denotes the $\prod_kd_k$-by-$\prod_k d_k$ Hession matrix evaluated at $\check\Theta$.
-
-We first bound the linear term in~\eqref{eq:taylor}. Note that, by Lemma~\ref{lem:inq},
-\begin{equation}\label{eq:linear}
-|\langle\Vec(\fplogl(\trueT), \Vec(\Theta-\trueT)  \rangle|\leq \snormSize{}{\fplogl(\trueT)} \nnormSize{}{\Theta-\trueT},
-\end{equation}
-where $\snormSize{}{\cdot}$ denotes the tensor spectral norm and $\nnormSize{}{\cdot}$ denotes the tensor nuclear norm. Define
-\[
-s_\omega={\partial \tL_\tY\over \partial \theta_\omega}\Big|_{\Theta=\trueT} \;\; \textrm{ for all } \; \omega\in[d_1]\times\cdots\times [d_K].
-\]
-Based on~\eqref{eq:property} and the definition of $U_\alpha$, $\fplogl(\trueT)=\entry{s_{\omega}}$ is a random tensor whose entries are independently distributed satisfying
-\begin{equation}\label{eq:norm}
-\mathbb{E}(s_\omega)=0,\quad |s_\omega|\leq U_\alpha, \quad \text{for all }\omega\in[d_1]\times \cdots \times [d_K].
-\end{equation}
-By lemma~\ref{lem:noisytensor}, with probability at least $1-\exp(-C_1 \sum_kd_k)$, we have
-\begin{equation}\label{eq:normrandom}
-\snormSize{}{\fplogl(\trueT)} \leq C_2 U_\alpha\sqrt{\sum_k d_k},
-\end{equation}
-where $C_1, C_2$ are two positive constants that depend only on $K$.
-Furthermore, note that $\text{rank}(\Theta)\leq \mr$, $\text{rank}(\trueT)\leq \mr$, so $\text{rank}(\Theta-\trueT)\leq 2\mr$.
-By lemma~\ref{lem:nuclear}, $\nnormSize{}{\Theta-\trueT}\leq (2r_{\max})^{K-1\over 2}\FnormSize{}{\Theta-\trueT}$.
-Combining~\eqref{eq:linear}, \eqref{eq:norm} and \eqref{eq:normrandom}, we have that, with probability at least $1-\exp(-C_1 \sum_kd_k)$,
-\begin{equation}\label{eq:linearconclusion}
-|\langle \Vec(\fplogl(\trueT)), \Vec(\Theta-\trueT)  \rangle | \leq C_2 U_\alpha  \sqrt{r_{\max}^{K-1} \sum_k d_k}  \FnormSize{}{\Theta-\trueT}.
-\end{equation}
-
-We next bound the quadratic term in \eqref{eq:taylor}. Note that
-\begin{align}\label{eq:quadratic}
- \Vec(\Theta-\trueT)^T \fpplogl(\check{\Theta})\Vec(\Theta-\trueT)&=\sum_\omega \left( {\partial^2\tL_{\tY}\over \partial \theta^2_\omega} \Big|_{\Theta=\check\Theta} \right)(\theta_\omega-\theta_{{\text{true}},\omega})^2 \nonumber \\
-&\leq - L_\alpha\sum_{\omega}(\Theta_{\omega}-\Theta_{\text{true},\omega})^2 \nonumber \\
-&=-L_\alpha\FnormSize{}{\Theta-\trueT}^2,
-\end{align}
-where the second line comes from the fact that  $\mnormSize{}{\check\Theta}\leq \alpha$ and the definition of $L_\alpha$.
-
-Combining~\eqref{eq:taylor}, \eqref{eq:linearconclusion} and~\eqref{eq:quadratic}, we have that, for all $\Theta\in\tP$, with probability at least $1-\exp(-C_1 \sum_kd_k)$,
-\[
-\tL_\tY(\Theta)\leq \tL_{\tY}(\trueT)+C_2U_\alpha  \left(r_{\max}^{K-1}\sum_k d_k\right)^{1/2}  \FnormSize{}{\Theta-\trueT}-{L_\alpha\over 2}\FnormSize{}{\Theta-\trueT}^2.
-\]
-In particular, the above inequality also holds for $\hat \Theta\in\tP$. Therefore,
-\[
-\tL_\tY(\hat \Theta)\leq \tL_{\tY}(\trueT)+C_2U_\alpha \left(r_{\max}^{K-1}\sum_k d_k\right)^{1/2}  \FnormSize{}{\hat \Theta-\trueT}-{L_\alpha\over 2} \FnormSize{}{\hat \Theta-\trueT}^2.
-\]
-Since $\hat \Theta=\arg\max_{\Theta\in\tP}\tL_\tY(\Theta)$, $\tL_\tY(\hat \Theta)-\tL_{\tY}(\trueT)\geq 0$, which gives
-\[
-C_2U_\alpha \left(r_{\max}^{K-1}\sum_k d_k\right)^{1/2}  \FnormSize{}{\hat \Theta-\trueT}-{L_\alpha\over 2}\FnormSize{}{\hat \Theta-\trueT}^2\geq 0.
-\]
-Henceforth,
-\[
-{1\over \sqrt{\prod_k d_k}} \FnormSize{}{\hat \Theta-\trueT}\leq {2C_2U_\alpha \sqrt{r_{\max}^{K-1}\sum_k d_k}\over L_\alpha \sqrt{\prod_k d_k}}={2C_2U_\alpha r_{\max}^{(K-1)/2}\over L_\alpha} \sqrt{ \sum_k d_k \over \prod_k d_k}.
-\]
-This completes the proof.
-\end{proof}
-
-\begin{proof}[Proof of Corollary~\ref{cor:prediction}]
-The result follows immediately from Theorem~\ref{thm:rate} and Lemma~\ref{lem:KL}.
-\end{proof}
-
-\begin{proof}[Proof of Theorem~\ref{thm:minimax}]
-
-Let $d_{\text{total}}=\prod_{k\in[K]}d_k$, and $\gamma\in[0,1]$ be a constant to be specified later.
-Our strategy is to construct a finite set of tensors $\tX=\{\Theta_i\colon i=1,\ldots \}\subset \tP$ satisfying the properties of (i)-(iv) in Lemma~\ref{lem:construction}.
-By Lemma~\ref{lem:construction}, such a subset of tensors exist.
-For any tensor  $\Theta\in\tX$, let $\mathbb{P}_{\Theta}$ denote the distribution of $\tY|\Theta$, where $\tY$ is the ordinal tensor.
-In particular, $\mathbb{P}_{\mathbf{0}}$ is the distribution of $\tY$ induced by the zero parameter tensor $\mathbf{0}$, i.e., the distribution of $\tY$ conditional on the parameter tensor $\Theta=\mathbf{0}$.
-Based on the Remark for Lemma~\ref{lem:KL}, we have
-\begin{equation}\label{eq:KLbound1}
-\mathrm{KL}(\mathbb{P}_{\Theta}|| \mathbb{P}_{\mathbf{0}})\leq C \FnormSize{}{\Theta}^2,
-\end{equation}
-where $C={(4L-6) \dot{f}^2(0)\over  A_\alpha}>0$ is a constant independent of the tensor dimension and rank.
-Combining the inequality~\eqref{eq:KLbound1} with property (iii) of $\tX$, we have
-\begin{equation}\label{eq:KLbound}
-\text{KL}(\mathbb{P}_{\Theta}||\mathbb{P}_{\mathbf{0}})\leq \gamma^2 R_{\max} d_{\max}.
-\end{equation}
-From~\eqref{eq:KLbound} and the property (i), we deduce that the condition
-\begin{equation}\label{eq:totalKL}
-{1\over \text{Card}(\tX)-1}\sum_{\Theta \in\tX}\text{KL}(\mathbb{P}_{\Theta}, \mathbb{P}_{\mathbf{0}})\leq \varepsilon \log_2\left\{\text{Card}(\tX)-1 \right\}
-\end{equation}
-holds for any $ \varepsilon \geq 0$ when $\gamma\in[0,1]$ is chosen to be sufficiently small depending on $\varepsilon$, e.g., $\gamma \leq \sqrt{\varepsilon\log2\over8}$.
-By applying Lemma~\ref{lem:Tsybakov} to~\eqref{eq:totalKL}, and in view of the property (iv), we obtain that
-\begin{equation}\label{eq:final}
-\inf_{\hat \Theta}\sup_{\trueT\in \tX}\mathbb{P}\left(\FnormSize{}{\hat \Theta- \trueT}\geq  {\gamma\over 8} \min\left\{ \alpha\sqrt{d_{\text{total}}}, C^{-1/2}\sqrt{ R_{\max}d_{\max}}\right\} \right)\geq {1\over 2}\left(1-2\varepsilon-\sqrt{16 \varepsilon \over R_{\max}d_{\max}\log2}\right).
-\end{equation}
-Note that $\text{Loss}(\hat \Theta, \trueT)=\FnormSize{}{\hat \Theta- \trueT}^2/d_{\text{total}}$ and $\tX\subset \tP$.
-By taking $\varepsilon=1/10$ and $\gamma=1/11$, we conclude from~\eqref{eq:final} that
-\begin{equation}\label{eq:prob}
-\inf_{\hat \Theta}\sup_{\trueT\in \tP}\mathbb{P}\left(\text{Loss}(\hat \Theta, \trueT)\geq c\min\left \{ \alpha^2,  {C^{-1}R_{\max}d_{\max}\over d_{\text{total}}}\right \}\right)\geq {1\over 2}\left({4\over 5}- \sqrt{1.6\over R_{\max}d_{\max}\log2} \right)\geq {1\over 8},
-\end{equation}
-where $c = {1 \over 88^2}$ and the last inequality comes from the condition for $d_{\text{max}}$.
-This completes the proof.
-\end{proof}
-
-\subsection{Sample complexity for tensor completion}
-
-\begin{proof}[Proof of Theorem~\ref{thm:completion}]
-
-For notational convenience, we use $\MFnormSize{}{\Theta}=\sum_{\omega\in\Omega}\Theta^2_\omega$ to denote the sum of squared entries over the observed set $\Omega$, for a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$.
-
-Following a similar argument as in the proof of Theorem~\ref{thm:rate}, we have
-\begin{equation}\label{eq:Taylor2}
-\logl(\Theta)=\logl(\trueT)+\langle\Vec(\plogl),\ \Vec(\Theta-\trueT)\rangle+{1\over 2}\Vec(\Theta-\trueT)^T\pplogl(\check\Theta)\Vec(\Theta-\trueT),
-\end{equation}
-where
-\begin{enumerate}
-\item $\plogl$ is a $d_1\times\cdots\times d_K$ tensor with $|\Omega|$ nonzero entries, and each entry is upper bounded by $U_\alpha>0$.
-\item $\pplogl$ is a diagonal matrix of size $d_{\text{total}}$-by-$d_{\text{total}}$ with $|\Omega|$ nonzero entries, and each entry is upper bounded by $-L_{\alpha}<0$.
-\end{enumerate}
-
-Similar to~\eqref{eq:linear} and~\eqref{eq:quadratic}, we have
-\begin{equation}\label{eq:linear2}
-|\langle\Vec(\plogl),\ \Vec(\Theta-\trueT)\rangle|\leq C_2U_\alpha \sqrt{r_{\max}^{K-1}\sum_k d_k}\MFnormSize{}{\Theta-\trueT}
-\end{equation}
-and
-\begin{equation}\label{eq:quadratic2}
-\Vec(\Theta-\trueT)^T\fpplogl(\check\Theta)\Vec(\Theta-\trueT)\leq -L_\alpha \MFnormSize{}{\Theta-\trueT}^2.
-\end{equation}
-
-Combining~\eqref{eq:Taylor2}-\eqref{eq:quadratic2} with the fact that $\logl(\hat \Theta)\geq \logl(\trueT)$, we have
-\begin{equation}\label{eq:sample}
-\MFnormSize{}{\hat \Theta-\trueT}\leq {2C_2U_\alpha  r_{\max}^{(K-1)/2} \over L_\alpha} \sqrt{\sum_kd_k},
-\end{equation}
-with probability at least $1-\exp(-C_1 \sum_k d_k)$.
-Lastly, we invoke the result regarding the closeness of $\Theta$ to its sampled version $\Theta_{\Omega}$, under the entrywise bound condition.
-Note that $\mnormSize{}{\hat\Theta-\trueT}\leq 2\alpha$ and $\text{rank}(\hat \Theta-\trueT)\leq 2\mr$.
-By Lemma~\ref{lem:Mnormbound}, $\anormSize{}{\hat \Theta-\trueT}\leq 2^{(3K-1)/2} \alpha \left({\prod r_k \over r_{\max}}\right)^{3/2}$.
-Therefore, the condition in Lemma~\ref{lem:convexity} holds with $\beta=2^{(3K-1)/2}\alpha \left({\prod r_k \over r_{\max}}\right)^{3/2}$.
-Applying Lemma~\ref{lem:convexity} to~\eqref{eq:sample} gives
-\begin{align}
- \PiFnormSize{}{\hat \Theta-\trueT}^2&\leq {1\over m}\MFnormSize{}{\hat \Theta-\trueT}^2+c\beta\sqrt{\sum_k d_k\over |\Omega|}\\
- &\leq {C_2  r^{K-1}_{\max}} {\sum_k d_k \over |\Omega|}+C_1 \alpha r_{\max}^{3(K-1)/2}\sqrt{\sum_kd_k\over |\Omega|},
-\end{align}
-with probability at least $1-\exp(-{\sum_kd_k\over \sum_k \log d_k})$ over the sampled set $\Omega$. Here $C_1, C_2>0$ are two constants independent of the tensor dimension and rank. Therefore,
-\[
- \PiFnormSize{}{\hat \Theta-\trueT}^2\to 0,\quad \text{as}\quad {|\Omega|\over \sum_kd_k}\to \infty,
-\]
-provided that $r_{\max}=O(1)$.
-\end{proof}
-
-\subsection{Convexity of the log-likelihood function}\label{sec:proofconvexity}
-\begin{thm}\label{thm:convexity}
-Define the function
-\begin{align}\label{eq:function}
- \logl(\Theta, \mb)&=\sum_{\omega\in\Omega}\sum_{\ell\in[L]} \left\{\mathds{1}_{\{y_\omega=\ell\}} \log \big[f(b_\ell-\theta_\omega)-  f(b_{\ell-1}-\theta_\omega)\big]\right\},
- \end{align}
-where $f(\cdot)$ satisfies Assumption~\ref{ass:link}. Then, $\logl(\Theta, \mb)$ is concave in $(\Theta,\mb)$.
-\end{thm}
-
-\begin{proof}
-Define $d_{\text{total}}=\prod_k d_k$.
-By abuse of notation, we use $(\Theta,\mb)$ to denote the length-$(d_{\text{total}}+L-1)$-vector collecting all parameters together.
-Let us denote a bivariate function
-\begin{align}
-\lambda: &\mathbb{R}^2\mapsto \mathbb{R}\\
-(u,v)&\mapsto\lambda(u,v) = \log \big[f(u)-  f(v)\big].
-\end{align}
-It suffices to show that $\lambda(u,v)$ is concave in $(u,v)$ where $u>v$.
-
-Suppose that the claim holds (which we will prove in the next paragraph). Based on~\eqref{eq:function}, $u,v$ are both linear functions of $(\Theta,\mb)$:
-\[
-u = \ma_1^T(\Theta,\mb), \quad v = \ma_2^T(\Theta,\mb), \quad \text{ for some}\ \ma_1,\ma_2\in \mathbb{R}^{d_{\text{total}}+L-1}.
-\]
-Then, $\lambda(u,v) = \lambda(\ma_1^T(\Theta,\mb),\ \ma_2^T(\Theta,\mb))$ is concave in $(\Theta,\mb)$ by the definition of concavity.
-Therefore, we can conclude that $\logl(\Theta, \mb) $ is concave in $(\Theta,\mb)$ because $\logl(\Theta, \mb)$ is the sum of $\lambda(u,v)$.
-
-Now, we prove the concavity of $\lambda(u,v)$. Note that
-\begin{equation}
-\lambda(u,v) = \log\big[f(u)-f(v)\big]=\log\big[\int\mathds{1}_{[u,v]}(x)f'(x)dx\big],
-\end{equation}
-where $\mathds{1}_{[u,v]}$ is an indicator function that equals 1 in the interval $[u,v]$, and 0 elsewhere.
-Furthermore, $\mathds{1}_{[u,v]}(x)$ is log-concave in $(u,v,x)$, and by Assumption~\ref{ass:link}, $f'(x)$ is log-concave in $x$.
-It follows that $\mathds{1}_{[u,v]}(x)f'(x)$
-is a log-concave in $(u,v,x)$. By Lemma~\ref{lem:lossconvexity}, we conclude that $\lambda(u,v)$ is concave in $(u,v)$ where $u>v$.
-\end{proof}
-
-\begin{lem}[Corollary 3.5 in~\cite{brascamp2002extensions}]\label{lem:lossconvexity}
-Let $F(x,y):\mathbb{R}^{m+n}\rightarrow \mathbb{R}$ be an integrable function where $x\in \mathbb{R}^{m},y\in \mathbb{R}^n$. Let
-\[
-G(x) = \int_{\mathbb{R}^n}F(x,y)dy.
-\]
-If $F(x,y)$ is log concave in $(x,y)$, then $G(x)$ is log concave in $x$.
-\end{lem}
-
-\subsection{Auxiliary lemmas}\label{sec:lemma}
-This section collects lemmas that are useful for the proofs of the main theorems.
-
-\begin{defn}[Atomic M-norm~\citep{ghadermarzy2019near}]
-Define $T_{\pm}=\{\tT\in\{\pm 1 \}^{d_1\times \cdots \times d_K}\colon$ $\text{rank}(\tT)=1\}$. The atomic M-norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
-\begin{align}
-\anormSize{}{\Theta}&=\inf\{t>0: \Theta\in t\text{conv}(T_{\pm})\}\\
-&=\inf\left\{\sum_{\tX\in T_{\pm}}c_x\colon \Theta=\sum_{\tX\in T_{\pm}} c_x\tX, \ c_x>0\right\}.
-\end{align}
-\end{defn}
-
-\begin{defn}[Spectral norm~\citep{lim2005singular}]
-The spectral norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
-\[
-\snormSize{}{\Theta}=\sup\left\{\langle \Theta, \mx_1\otimes \cdots \otimes \mx_K\rangle \colon \vnormSize{}{\mx_k}=1,\ \mx_k\in\mathbb{R}^{d_k},\ \text{for all}\ k\in[K]\right\}.
-\]
-\end{defn}
-
-\begin{defn}[Nuclear norm~\citep{friedland2018nuclear}]
-The nuclear norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
-\[
-\nnormSize{}{\Theta}=\inf
-\left\{
-\sum_{i\in[r]}|\lambda_i|\colon \Theta=\sum_{i=1}^r \lambda_i\mx^{(i)}_1\otimes \cdots \otimes \mx^{(i)}_K,\ \vnormSize{}{\mx^{(i)}_k}=1,\ \mx^{(i)}_k\in\mathbb{R}^{d_k},\ \text{for all} k\in[K],\ i\in[r]
-\right\},
-\]
-where the infimum is taken over all $r\in\mathbb{N}$ and $\vnormSize{}{\mx^{(i)}_k}=1$ for all $i\in[r]$ and $k\in[K]$.
-\end{defn}
-
-\begin{lem}[M-norm and infinity norm~\citep{ghadermarzy2019near}]\label{lem:Mnormbound}
-Let $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ be an order-$K$, rank-$(r_1,\ldots,r_K)$ tensor. Then
-\[
-\mnormSize{}{\Theta}\leq \anormSize{}{\Theta}\leq \left(\prod_k r_k \over r_{\max}\right)^{3\over 2} \mnormSize{}{\Theta}.
-\]
-\end{lem}
-
-\begin{lem}[Nuclear norm and F-norm] \label{lem:nuclear}
-Let $\tA\in\mathbb{R}^{d_1\times\cdots\times d_K}$ be an order-$K$ tensor with Tucker $\text{rank}(\tA)=(r_1,\ldots,r_K)$. Then
-\[
-\nnormSize{}{\tA} \leq \sqrt{\prod_k r_k\over \max_k r_k} \FnormSize{}{\tA},
-\]
-where $\nnormSize{}{\cdot}$ denotes the nuclear norm of the tensor.
-\end{lem}
-
-\begin{proof}
-Without loss of generality, suppose $r_1=\min_k r_k$.
-Let $\tA_{(k)}$ denote the mode-$k$ matricization of $\tA$ for all $k\in[K]$.
-By \citet[Corollary 4.11]{wang2017operator}, and the invariance relationship between a tensor and its Tucker core~\citep[Section 6]{jiang2017tensor}, we have
-\begin{equation}\label{eq:norminequality}
-\nnormSize{}{\tA} \leq \sqrt{\prod_{k\geq 2} r_k \over \max_{k\geq 2} r_k} \nnormSize{}{\tA_{(1)}},
-\end{equation}
-where $\tA_{(1)}$ is a $d_1$-by-$\prod_{k\geq 2}d_k$ matrix with matrix rank $r_1$.
-Furthermore, the relationship between the matrix norms implies that $\nnormSize{}{\tA_{(1)}}\leq \sqrt{r_1}\FnormSize{}{\tA_{(1)}}=\sqrt{r_1}\FnormSize{}{\tA}$.
-Combining this fact with the inequality~\eqref{eq:norminequality} yields the final claim.
-\end{proof}
-
-\begin{lem} \label{lem:inq}
-Let $\tA, \; \tB$ be two order-$K$ tensors of the same dimension. Then
-\[
-|\langle \tA,\tB\rangle| \leq \snormSize{}{\tA}   \nnormSize{}{\tB}.
-\]
-\end{lem}
-
-\begin{proof}
-By~\citet[Proposition 3.1]{friedland2018nuclear}, there exists a nuclear norm decomposition of $\tB$, such that
-\[
-\tB=\sum_{r} \lambda_r \ma^{(1)}_r\otimes \cdots\otimes \ma^{(K)}_r,\quad \ma_r^{(k)}\in\mathbf{S}^{d_k-1}(\mathbb{R}),\quad \text{for all }k\in[K],
-\]
-and $\nnormSize{}{\tB}=\sum_{r}|\lambda_r|$. Henceforth we have
-\begin{align*}
-|\langle \tA,\tB\rangle|&=| \langle \tA, \sum_{r} \lambda_r \ma^{(1)}_r\otimes \cdots\otimes \ma^{(K)}_r \rangle|\leq \sum_r |\lambda_r| |\langle \tA, \ma^{(1)}_r \otimes \cdots\otimes \ma^{(K)}_r \rangle|\\
-&\leq \sum_{r}|\lambda_r| \snormSize{}{\tA}= \snormSize{}{\tA}\nnormSize{}{\tB},
-\end{align*}
-which completes the proof.
-\end{proof}
-
-The following lemma provides the bound on the spectral norm of random tensors. The result was firstly presented in~\cite{nguyen2015tensor}, and we adopt the version from~\cite{tomioka2014spectral}.
-\begin{lem}[Spectral norm of random tensors~\citep{tomioka2014spectral}]\label{lem:tensor}
-Suppose that $\tS=\entry{s_{\omega}}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is an order-$K$ tensor whose entries are independent random variables that satisfy
-\[
-\mathbb{E}(s_{\omega})=0,\quad \text{and} \quad\mathbb{E}(e^{ts_{\omega}})\leq e^{t^2L^2/2}.
-\]
-Then the spectral norm $\snormSize{}{\tS}$ satisfies that,
-\[
-\snormSize{}{\tS}\leq \sqrt{{8L^2} \log (12K) \sum_k d_k +\log (2/\delta)},
-\]
-with probability at least $1-\delta$.
-\end{lem}
-
-\begin{lem} \label{lem:noisytensor}
-Suppose that $\tS=\entry{s_{\omega}}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is an order-$K$ tensor whose entries are independent random variables that satisfy
-\[
-\mathbb{E}(s_{\omega})=0,\quad \text{and}\quad |s_{\omega}|\leq U.
-\]
-Then we have
-\[
-\mathbb{P}\left(\snormSize{}{\tS}\geq C_2 U\sqrt{\sum_k d_k} \right)\leq \exp\left(-C_1  \log K \sum_k d_k\right)
-\]
-where $C_1>0$ is an absolute constant, and $C_2>0$ is a constant that depends only on $K$.
-\end{lem}
-
-\begin{proof}  Note that the random variable $U^{-1}s_{\omega}$ is zero-mean and supported on $[-1,1]$. Therefore, $U^{-1}s_{\omega}$ is sub-Gaussian with parameter ${1-(-1)\over 2}=1$, i.e.
-\[
-\mathbb{E}(U^{-1}s_{\omega})=0,\quad \text{and}\quad \mathbb{E}(e^{tU^{-1}s_{\omega}})\leq e^{t^2/2}.
-\]
-It follows from Lemma~\ref{lem:tensor} that, with probability at least $1-\delta$,
-\[
-\snormSize{}{U^{-1}\tS}\leq \sqrt{\left(c_0\log K+c_1\right) \sum_k d_k +\log (2/\delta)},
-\]
-where $c_0, c_1>0$ are two absolute constants. Taking $\delta=\exp (-C_1\log K \sum_k d_k)$ yields the final claim, where $C_2=c_0\log K+c_1+1>0$ is another constant.
-\end{proof}
-
-\begin{lem}\label{lem:KLentry} Let $X,\; Y$ be two discrete random variables taking values on $L$ possible categories, with category probabilities $\{p_\ell\}_{\ell\in[L]}$ and $\{q_\ell\}_{\ell\in[L]}$, respectively.
-Suppose $p_\ell$, $q_\ell>0$ for all $i\in[L]$.
-Then, the Kullback-Leibler (KL) divergence satisfies that
-\[
-\mathrm{KL}(X||Y)\stackrel{\text{def}}{=}-\sum_{\ell\in[L]}\mathbb{P}_X(\ell)\log\left\{{\mathbb{P}_Y(\ell)\over \mathbb{P}_X(\ell)}\right\} \leq \sum_{\ell \in [L]}{(p_\ell-q_\ell)^2 \over q_\ell}.
-\]
-\end{lem}
-\begin{proof} Using the fact $\log x\leq x-1$ for $x>0$, we have that
-\begin{align}\label{eq:KL}
-\text{KL}(X||Y)&=\sum_{\ell\in[L]}p_\ell\log{p_\ell\over q_\ell}\\
-&\leq \sum_{\ell\in[L]} {p_\ell\over q_\ell}(p_\ell-q_\ell)\\
-&=\sum_{\ell\in[L]} \left({p_\ell\over q_\ell }- 1\right)(p_\ell-q_\ell)+ \sum_{\ell\in[L]} (p_\ell-q_\ell).
-\end{align}
-Note that $\sum_{\ell\in[L]}(p_\ell-q_\ell)=0$. Therefore,
-\[
-\text{KL}(X||Y)\leq \sum_{\ell\in[L]}\left( {p_\ell\over q_\ell}-1\right)\left(p_\ell-q_\ell\right)=\sum_{\ell\in[L]}{(p_\ell-q_\ell)^2\over q_\ell}.
-\]
-\end{proof}
-
-\begin{lem}[KL divergence and F-norm]~\label{lem:KL}
-Let $\tY\in[L]^{d_1\times \cdots \times d_K}$ be an ordinal tensor generated from the model~\eqref{eq:model} with the link function $f$ and parameter tensor $\Theta$.
-Let $\mathbb{P}_{\Theta}$ denote the joint categorical distribution of $\tY|\Theta$ induced by the parameter tensor $\Theta$, where $\mnormSize{}{\Theta}\leq \alpha$.
-Define
-\begin{equation}\label{eq:ass}
-A_\alpha=\min_{\ell\in[L], |\theta|\leq \alpha}\left[f(b_\ell-\theta)-f(b_{\ell-1}-\theta)\right].
-\end{equation}
-Then, for any two tensors $\Theta,\; \Theta^*$ in the parameter spaces, we have
-\[
-\mathrm{KL}(\mathbb{P}_{\Theta}|| \mathbb{P}_{\Theta^*})\leq {2(2L-3)\over A_\alpha} \dot{f}^2(0)\FnormSize{}{\Theta-\Theta^*}^2.
-\]
-\end{lem}
-\begin{proof} Suppose that the distribution over the ordinal tensor $\tY=\entry{y_\omega}$ is induced by $\Theta=\entry{\theta_\omega}$. Then, based on the generative model~\eqref{eq:model},
-\[
-\mathbb{P}(y_\omega=\ell | \theta_\omega)=f(b_{\ell}-\theta_\omega)-f(b_{\ell-1}-\theta_\omega),
-\]
-for all $\ell\in[L]$ and $\omega\in[d_1]\times \cdots \times [d_K]$.
-For notational convenience, we suppress the subscribe in $\theta_\omega$ and simply write $\theta$ (and respectively, $\theta^*$).
-Based on Lemma~\ref{lem:KLentry} and Taylor expansion,
-\begin{align}
-\text{KL}(\theta|| \theta^*) & \leq \sum_{\ell\in[L]} {\left[f(b_{\ell}-\theta)-f(b_{\ell-1}-\theta)-f(b_{\ell}-\theta^*)+f(b_{\ell-1}-\theta^*)\right]^2\over f(b_{\ell}-\theta^*)-f(b_{\ell-1}-\theta^*)}\\
- &\leq \sum_{\ell=2}^{L-1} {\left[\dot{f}(b_\ell-\eta_\ell)-\dot{f}(b_{\ell-1}-\eta_{\ell-1})\right]^2 \over f(b_\ell-\theta^*)-f(b_{\ell-1}-\theta^*)} \left(\theta-\theta^*\right)^2+{\dot{f}^2(b_1-\eta_1) \over f(b_1-\theta^*)} (\theta-\theta^*)^2\\
- &\quad \quad\quad \quad  +{\dot{f}^2(b_{L-1}-\eta_{L-1})\over 1-f(b_{L-1}-\theta^*)}(\theta-\theta^*)^2,
-\end{align}
-where $\eta_\ell$ and $\eta_{\ell-1}$ fall between $\theta$ and $\theta^*$. Therefore,
-\begin{equation}\label{eq:entrywise}
-\text{KL}(\theta|| \theta^*) \leq \left({4(L-2)\over A_\alpha}+ {2\over A_\alpha}\right)\dot{f}^2(0)(\theta-\theta^*)^2={2(2L-3)\over A_\alpha} \dot{f}^2(0)(\theta-\theta^*)^2,
-\end{equation}
-where we have used Taylor expansion, the bound~\eqref{eq:ass}, and the fact that $\dot{f}(\cdot)$ peaks at zero for an unimodal and symmetric function.
-Now summing~\eqref{eq:entrywise} over the index set $\omega\in[d_1]\times \cdots \times [d_K]$ gives
-\[
-\text{KL}(\mathbb{P}_{\Theta}|| \mathbb{P}_{\Theta^*}) =\sum_{\omega\in[d_1]\times \cdots \times[d_K]} \text{KL}(\theta_\omega || \theta^*_\omega) \leq {2(2L-3)\over A_\alpha} \dot{f}^2(0)\FnormSize{}{\Theta-\Theta^*}^2.
-\]
-\end{proof}
-
-\begin{rmk} In particular, let $\mathbb{P}_{\bf{0}}$ denote the distribution of $\tY|\bf{0}$ induced by the zero parameter tensor. Then we have
-\[
-\text{KL}(\mathbb{P}_{\Theta}||\mathbb{P}_{\bf{0}})\leq {2(2L-3)\over A_\alpha}  \dot{f}^2(0)\FnormSize{}{\Theta}^2.
-\]
-\end{rmk}
-
-\begin{lem}\label{lem:construction}
-Assume the same setup as in Theorem~\ref{thm:minimax}.
-Without loss of generality, suppose $d_1=\max_kd_k$.
-Define $R=\max_k r_k$ and $d_{\text{total}}=\prod_{k\in[K]} d_k$.
-For any constant $0\leq \gamma \leq 1$, there exist a finite set of tensors $\tX=\{\Theta_i: i=1,\ldots\}\subset \tP$ satisfying the following four properties:
-\begin{enumerate}[(i)]
-\item $\text{Card}(\tX)\geq 2^{Rd_1/8}+1$, where $\text{Card}$ denotes the cardinality;
-\item $\tX$ contains the zero tensor $\mathbf{0}\in\mathbb{R}^{d_1\times \cdots\times d_K}$;
-\item $\mnormSize{}{\Theta}\leq \gamma \min\left\{ \alpha ,\ C^{-1/2}\sqrt{Rd_1\over d_{\text{total}}} \right\} $ for any element $\Theta\in\tX$;
-\item $\FnormSize{}{\Theta_i-\Theta_j}\geq {\gamma\over 4} \min\left\{ \alpha\sqrt{d_{\text{total}}},\ C^{-1/2}\sqrt{Rd_1}\right\}$ for any two distinct elements $\Theta_i,\; \Theta_j\in\tX$,
-\end{enumerate}
-Here $C=C(\alpha,L,f,\mb)={(4L-6)\dot{f}^2(0)\over A_\alpha }>0$ is a constant independent of the tensor dimension and rank.
-\end{lem}
-
-\begin{proof}
-Given a constant $0\leq \gamma \leq 1$, we define a set of matrices:
-\[
-\tC=\left\{\mM=(m_{ij})\in\mathbb{R}^{d_1\times R}: a_{ij}\in \left\{ 0,\gamma \min\left\{ \alpha , C^{-1/2}\sqrt{Rd_1\over d_{\text{total}}} \right\}\right\} ,\  \forall (i,j)\in[d_1]\times[R]\right\}.
-\]
-We then consider the associated set of block tensors:
-\begin{align}
-\tB=\tB(\tC)=\{\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}\colon& \Theta=\mA\otimes \mathbf{1}_{d_3}\otimes \cdots \otimes \mathbf{1}_{d_K}, \\
-&\ \text{where}\ \mA=(\mM|\cdots|\mM|\mO) \in\mathbb{R}^{d_1\times d_2},\ \mM\in\tC\},
-\end{align}
-where $\mathbf{1}_d$ denotes a length-$d$ vector with all entries 1, $\mO$ denotes the $d_1\times (d_2-R\lfloor d_2/R \rfloor)$ zero matrix, and $\lfloor d_2/ R \rfloor$ is the integer part of $d_2/R$.
-In other words, the subtensor $\Theta(\mI, \mI,i_3, \ldots,i_K)\in\mathbb{R}^{d_1\times d_2}$ are the same for all fixed $(i_3,\ldots,i_K)\in[d_3]\times \cdots \times [d_K]$, and furthermore, each subtensor $\Theta(\mI,\mI, i_3,\ldots,i_K)$ itself is filled by copying the matrix $\mM\in\mathbb{R}^{d_1\times R}$ as many times as would fit.
-
-By construction, any element of $\tB$, as well as the difference of any two elements of $\tB$, has Tucker rank at most $\max_k r_k\leq R$, and the entries of any tensor in $\tB$ take values in $[0,\alpha]$.
-Thus, $\tB\subset\tP$.
-By Lemma~\ref{lem:VGbound}, there exists a subset $\tX\subset \tB$ with cardinality $\text{Card}(\tX)\geq 2^{Rd_1/8}+1$ containing the zero $d_1\times \cdots \times d_K$ tensor, such that, for any two distinct elements $\Theta_i$ and $\Theta_j$ in $\tX$,
-\[
-\FnormSize{}{\Theta_i-\Theta_j}^2 \geq {Rd_1\over 8} \gamma^2\min\left\{ \alpha^2, {C^{-1}Rd_1 \over d_{\text{total}}}\right\} \lfloor {d_2\over R} \rfloor \prod_{k\geq 3}d_k\geq {\gamma^2\min\left\{ \alpha^2 d_{\text{total}}, C^{-1}Rd_1\right\}  \over 16}.
-\]
-In addition, each entry of $\Theta\in\tX$ is bounded by $\gamma \min\left\{ \alpha , C^{-1/2}\sqrt{Rd_1\over d_{\text{total}}}\right\} $. Therefore the Properties (i) to (iv) are satisfied.
-\end{proof}
-
-\begin{lem}[Varshamov-Gilbert bound]\label{lem:VGbound}
-Let $\Omega=\{(w_1,\ldots,w_m)\colon w_i\in\{0,1\}\}$. Suppose $m>8$. Then there exists a subset $\{w^{(0)},\ldots,w^{(M)}\}$ of $\Omega$ such that $w^{(0)}=(0,\ldots,0)$ and
-\[
-\zeronormSize{}{w^{(j)}-w^{(k)}}\geq {m\over 8},\quad \text{for} \ 0\leq j<k\leq M,
-\]
-where $\zeronormSize{}{\cdot}$ denotes the Hamming distance, and $M\geq 2^{m/8}$.
-\end{lem}
-
-\begin{lem}[Theorem 2.5 in~\cite{tsybakov2008introduction}]\label{lem:Tsybakov}
-Assume that a set $\tX$ contains element $\Theta_0, \Theta_1, \ldots,\Theta_M$ ($M\geq 2$) such that
-\begin{itemize}
-\item $d(\Theta_j,\ \Theta_k)\geq 2s>0$, $\forall 0\leq j\leq k\leq M$;
-\item $\mathbb{P}_0$ is absolutely continuous with respect to $\mathbb{P}_j$, $\forall j=1,\ldots,M$, and
-\[
-{1\over M}\sum_{j=1}^M \mathrm{KL}(\mathbb{P}_j||\mathbb{P}_0)\leq \alpha \log M
-\]
-where $d\colon \tX\times \tX\mapsto [0,+\infty]$ is a semi-distance function, $0<\alpha<{1/8}$ and $\mathbb{P}_j=\mathbb{P}_{\Theta_j}$, $j=0,1\ldots,M$.
-\end{itemize}
-Then
-\[
-\inf_{\hat \Theta}\sup_{\Theta\in\tX} \mathbb{P}_{\Theta}(d(\hat \Theta, \Theta)\geq s)\geq {\sqrt{M}\over 1+\sqrt{M}}\left(1-2\alpha-\sqrt{2\alpha\over \log M} \right)>0.
-\]
-\end{lem}
-
-\begin{lem}[Lemma 28 in~\cite{ghadermarzy2019near}]\label{lem:convexity}
-Define $\mathbb{B}_{M}(\beta)=\{\Theta\in \mathbb{R}^{d_1\times \cdots \times d_K}\colon \anormSize{}{\Theta}\leq \beta \}$.
-Let $\Omega\subset[d_1]\times\cdots \times [d_K]$ be a random set with $m=|\Omega|$, and assume that each entry in $\Omega$ is drawn with replacement from $[d_1]\times\cdots\times[d_K]$ using probability $\Pi$.
-Define
-\[
-\PiFnormSize{}{\Theta}^2={1\over m}\mathbb{E}_{\Omega\in\Pi}\MFnormSize{}{\Theta}^2.
-\]
-Then, there exists a universal constant $c>0$, such that, with probability at least $1-\exp\left(-{\sum_kd_k \over \sum_k \log d_k} \right)$ over the sampled set $\Omega$,
-\[
-{1\over m}\MFnormSize{}{\Theta}^2 \geq \PiFnormSize{}{\Theta}^2-c\beta\sqrt{\sum_k d_k\over m}
-\]
-holds uniformly for all $\Theta\in\mathbb{B}_M(\beta)$.
-\end{lem}
-
-\section{Conclusion}
-We have developed a low-rank tensor estimation method based on possibly incomplete, ordinal-valued observations.
-A sharp error bound is established, and we demonstrate the outperformance of our approach compared to other methods.
-The work unlocks several directions of future research.
-One interesting question would be the inference problem, i.e..
-to assess the uncertainty of the obtained estimates and the imputation.
-Other directions include the trade-off between (non)convex optimization and statistical/computational efficiency.
-While convex relaxations are popular approach for matrix/tensor problem, they are often slow in practice~\citep{ge2017optimization,chen2019non}.
-The interplay between computational efficiency and statistical accuracy in general tensor problems warrants future research.
-
-\section*{Acknowledgments}
-This research is supported in part by NSF grant DMS-1915978 and the University of Wisconsin-Madison, Office of the Vice Chancellor for Research and Graduate Education with funding from the Wisconsin Alumni Research Foundation.
-\newpage
-
-\appendix
-\setcounter{figure}{0}
-\setcounter{table}{0}
-\renewcommand{\thetable}{S\arabic{table}}
-\renewcommand{\thefigure}{S\arabic{figure}}
-\renewcommand{\figurename}{Supplemental Figure}
-\renewcommand{\tablename}{Supplemental Table}
-
-\section*{Appendix}
-\section{Extension of Theorem~\ref{thm:rate} to unknown cut-off points}\label{sec:extention}
-When the cut-off points $\mb$ is unknown, we estimate $(\hat \Theta,\hat \mb)$ by
-\begin{equation}\label{eq:joint}
-(\hat \Theta,\hat \mb)=\argmax_{(\Theta,\mb)\in\tP\times \tB}\logl(\Theta,\mb),
-\end{equation}
-where
-\[
-\tP=\{\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}\colon \text{rank}(\tP)\leq \mr,\ \mnormSize{}{\Theta}\leq \alpha\}, \quad \tB=\{\mb\in\mathbb{R}^{L-1}\colon \mnormSize{}{\mb}\leq \beta,\ \min_{\ell}(b_\ell-b_{\ell-1}) \geq \Delta\}.
-\]
-The estimation accuracy is assessed using the mean squared error (MSE):
-\[
-\text{MSE}\left( \hat \Theta, \trueT \right)={1\over \prod_k d_k}\FnormSize{}{\hat \Theta-\trueT},\quad \text{MSE}\left( \hat \mb, \trueb \right)={1\over L-1}\FnormSize{}{\hat \mb-\trueb}.
-\]
-By abuse of notation, we use $(\Theta,\mb)$ to denote the length-$(d_{\text{total}}+L-1)$-vector collecting all parameters together. The total MSE is defined as
-\[
-\text{MSE}\left((\hat \Theta,\hat \mb),\ (\trueT,\trueb)\right)={1\over \prod_k d_k+L-1}\FnormSize{}{(\hat \Theta,\hat \mb)-(\trueT,\trueb)}^2.
-\]
-We introduce several quantities that will be used in our theory:
-\begin{enumerate}
-\item We make the convention that $b_0=-\infty$, $b_L=\infty$, $f(-\infty)=0$, $f(\infty)=1$, and $\dot{f}(-\infty)=\ddot{f}(-\infty)=\dot{f}(\infty)=\ddot{f}(\infty)=0$.
-
-\item The difference function $g_\ell(\theta)$ is defined as $g_\ell(\theta)=f(b_\ell-\theta)-f(b_{\ell-1}-\theta)$ for all $\theta\in\mathbb{R}$ and $\ell\in[L]$.
-
-\item Define $n_\ell=\sum_{\omega\in\Omega}\mathds{1}_{\{y_\omega=\ell\}}$, i.e., the number of tensor entries taking value on $\ell\in[L]$.
-
-\item With a little abuse of notation, we re-define the constants in~\eqref{eq:bound} as
-\begin{equation}\label{eq:UL}
-U_{\alpha,\beta,\Delta}=\max_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}}\max_{\ell\in[L-1]}{|\dot{g}_\ell(\theta)|\over g_\ell(\theta)}, \quad \text{and}\quad
-L_{\alpha,\beta,\Delta}=\min_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}}\min_{\ell\in[L-1]} {\dot{g}^2_\ell(\theta)-\ddot{g}_\ell(\theta)g_\ell(\theta)\over g^2_\ell(\theta)}.
-\end{equation}
-\item We define three additional constants:
-\begin{align}\label{eq:CD}
-C_{\alpha,\beta,\Delta} &= \max_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}}  \max_{\ell\in[L-1]}\max\left\{\frac{\dot{f}(b_\ell-\theta)}{g_\ell(\theta)},\ \frac{\dot{f}(b_\ell-\theta)}{g_{\ell+1}(\theta)}\right\},\notag \\
-D_{\alpha,\beta,\Delta} &=\min_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}} \min_{\ell\in[L-1]}\min\left\{  {\partial \over \partial \theta}\left({\dot{f}(b_{\ell}-\theta)\over g_\ell(\theta)}\right),\ -{\partial \over \partial \theta}\left({\dot{f}(b_{\ell}-\theta)\over g_{\ell+1}(\theta)}\right)\right\}\notag \\
-&=  \min_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}}\min_{\ell\in[L-1]}\min \left\{-\frac{\ddot{f}(b_\ell-\theta)g_\ell(\theta)-\dot{f}(b_\ell-\theta)\dot{g}_\ell(\theta)}{g^2_\ell(\theta)},\ \right.
-\notag \\
-&\hspace{1.5in} \left. \frac{\ddot{f}(b_\ell-\theta)g_{\ell+1}(\theta)-\dot{f}(b_\ell-\theta)\dot{g}_{\ell+1}(\theta)}{g^2_{\ell+1}(\theta)}\right\},\notag \\
-\tilde{D}_{\alpha,\beta,\Delta} &=\max_{ \substack{|\theta|\leq \alpha,\\ \mb \in\tB}} \max_{\ell\in[L-1]}\max\left\{  {\partial \over \partial \theta}\left({\dot{f}(b_{\ell}-\theta)\over g_\ell(\theta)}\right),\ -{\partial \over \partial \theta}\left({\dot{f}(b_{\ell}-\theta)\over g_{\ell+1}(\theta)}\right)\right\}.
-\end{align}
-\end{enumerate}
-We make the following assumptions about the link function.
-\begin{assumption}\label{ass:joint}
-The link function $f\colon \mathbb{R}\mapsto [0,1]$ satisfies the following properties:
-\begin{enumerate}
-\item $f(\theta)$ is twice-differentiable and strictly increasing in $\theta$.
-\item $\dot{f}(\theta)$ is strictly log-concave and symmetric with respect to $\theta=0$.
-\item The function ${\dot{f}(b_\ell-\theta)\over g_\ell(\theta)}$ is strictly increasing with respect to $\theta$ for all $\mb\in\tB$.
-\item The function ${\dot{f}(b_\ell-\theta)\over g_{\ell+1}(\theta)}$ is strictly decreasing with respect to $\theta$ for all $\mb\in\tB$.
-\end{enumerate}
-\end{assumption}
-
-\begin{rmk}
-The condition $\Delta=\min_{\ell}(b_{\ell}-b_{\ell-1})>0$ on the feasible set $\tB$ guarantees the strict positiveness of $g_{\ell}(\theta)=f(b_{\ell}-\theta)-f(b_{\ell-1}-\theta)$.
-Therefore, the denominators in the above quantities $U_{\alpha,\beta,\Delta}, L_{\alpha,\beta,\Delta}, C_{\alpha,\beta,\Delta}, D_{\alpha,\beta,\Delta},\tilde{D}_{\alpha,\beta,\Delta}$ are well-defined.
-Furthermore, by Theorem~\ref{thm:convexity}, $g_\ell(\theta)$ is strictly log-concave in $\theta$ for all $\mb\in\tB$.
-Based on Assumption~\ref{ass:joint} and closeness of the feasible set, we have $U_{\alpha,\beta,\Delta}>0$, $L_{\alpha,\beta,\Delta}>0$, $C_{\alpha,\beta,\Delta}>0$, $D_{\alpha,\beta,\Delta}>0$, $\tilde{D}_{\alpha,\beta,\Delta}>0$.
-\end{rmk}
-
-\begin{rmk}
-In particular, for logistic link $f(x) = \frac{1}{1+e^{-x}}$ , we have
-\begin{align}
-C_{\alpha,\beta,\Delta}&=\max_{\substack{|\theta|\leq \alpha,\\ \mb\in\tB}}\max_{\ell\in[L-1]}\max
-\left\{
-{ 1\over e^{b_{\ell}-b_{\ell-1}}-1} { 1+e^{-(b_{\ell-1}-\theta)} \over 1+e^{-(b_{\ell}-\theta)}},\
-{ 1\over 1-e^{-(b_{\ell+1}-b_{\ell})}} { 1+e^{-(b_{\ell+1}-\theta)} \over 1+e^{-(b_{\ell}-\theta)}}
-\right\}>0,
-\\
-D_{\alpha,\beta,\Delta}&=\min_{\substack{|\theta|\leq \alpha,\\ \mb\in\tB}}\min_{\ell\in[L-1]} {e^{-(b_\ell-\theta)}\over (1+e^{-(b_\ell-\theta)})^2}>0.
-\end{align}
-\end{rmk}
-
-\begin{thm}[Statistical convergence with unknown $\mb$]\label{thm:ratejoint}
-Consider an ordinal tensor $\tY\in[L]^{d_1\times \cdots \times d_K}$ generated from model~\eqref{eq:model} with the link function $f$ and parameters $(\trueT, \trueb)\in\tP\times \tB$.
-Suppose the link function $f$ satisfies Assumption~\ref{ass:joint}.
-Define $r_{\max}=\max_k r_k$, and assume $r_{\max}=\tO(1)$.
-Then with very high probability, the estimator in~\eqref{eq:joint} satisfies
-\begin{equation}\label{eq:jointTheta}
-\mathrm{MSE}\left(\hat \Theta, \trueT \right)\leq\min\left( 4\alpha^2,\  {c_1U_{\alpha,\beta,\Delta}^2  r_{\max}^{K-1}  \over L_{\alpha,\beta,\Delta}^2}{ \sum_k d_k \over \prod_k d_k}\right),
-\end{equation}
-and
-\begin{equation}\label{eq:jointb}
-\mathrm{MSE}\left(\hat \mb, \trueb \right)\leq \min\left(4\beta^2,\ {C'_{\alpha,\beta,\Delta}r^{K-1}_{\max} \over D_{\alpha,\beta,\Delta}^2}{\left(\sum_kd_k\right)\left(\prod_k d_k\right) \over \min_{\ell}(n_{\ell}+n_{\ell+1})^2}\right),
-\end{equation}
-where $c_1, U_{\alpha,\beta,\Delta}, L_{\alpha,\beta,\Delta}, C'_{\alpha,\beta,\Delta}, \tilde D_{\alpha,\beta,\Delta}$ are constants independent of the tensor dimension and rank.
-
-In particular, we have
-\begin{align}
-\mathrm{MSE}\left((\hat \Theta,\hat \mb),\ (\trueT,\trueb)\right) = \tO\left({\sum_k d_k \over \min\left\{\prod_k d_k ,\ \min_\ell(n_\ell+n_{\ell+1})^2\right\}}\right),\quad \text{as}\quad d_{\min}\to\infty.
-\end{align}
-\end{thm}
-
-\begin{rmk} The total MSE has two components, where the first component $\tO\left({\sum_k d_k \over\prod_k d_k}\right)$ is from the error in $\hat\Theta$, and the second component $\tO\left({\sum_k d_k \over \min_\ell(n_\ell+n_{\ell+1})^2}\right)$ is from the error in $\hat \mb.$ When the $L$ labels are moderately balanced in that $\min_\ell(n_\ell+n_{\ell+1})\gg \sqrt{\prod_kd_k}$, the error in $\hat\Theta$ dominates the total MSE.
-When the $L$ labels are severely imbalanced in that $\min_\ell(n_\ell+n_{\ell+1})\ll \sqrt{\prod_kd_k}$, the error in $\hat\mb$ dominates the total MSE.
-\end{rmk}
-
-\begin{rmk} We discuss the impact of imbalance between labels to the estimation accuracy.
-Consider the special case $d_1=\cdots =d_K=d$.
-The bounds~\eqref{eq:jointTheta} and~\eqref{eq:jointb} demonstrate that, both $\hat \Theta$ and $\hat \mb$ achieve consistency as long as $\min_\ell(n_\ell+n_{\ell+1})$ is slightly larger than $\tO(d^{(K+1)/2})$.
-Note that in this case ${\min_\ell(n_\ell+n_{\ell+1}) \over \max_\ell n_{\ell}}\asymp {\min_\ell(n_\ell+n_{\ell+1})\over d^K} \to 0$, indicating a long-run imbalance between labels.
-The obtained consistency highlights the robustness of our proposed estimator~\eqref{eq:joint} to the label imbalance.
-\end{rmk}
-
-\begin{proof}[Proof of Theorem~\ref{thm:ratejoint}]
-The constant bounds with $\alpha$ and $\beta$ can be obtained trivially from the definition of the feasible sets. We focus on proving the non-constant bounds.
-
-Similar to the proof of Theorem~\ref{thm:rate}, we suppress $\Omega$ in the subscript. Based on the definition of $(\hat \Theta, \hat \mb)$,
-we have the following inequalities:
-\begin{equation}\label{eq:level}
-\flogl(\hat \Theta,\hat \mb)\geq \flogl(\trueT,\hat \mb)\quad \text{and}\quad \flogl(\hat \Theta,\hat \mb)\geq \flogl(\hat \Theta, \trueb).
-\end{equation}
-Following the same argument as in Theorem~\ref{thm:rate} and the first inequality in~\eqref{eq:level}, we obtain that
-\begin{equation}\label{eq:ratenew}
-\FnormSize{}{\hat \Theta-\trueT}^2\leq {c_1^2U_{\alpha,\beta,\Delta}^2 r_{\max}^{K-1} \over L_{\alpha,\beta,\Delta}^2}   \sum_k d_k ,
-\end{equation}
-where $U_{\alpha,\beta,\Delta}, L_{\alpha,\beta,\Delta}>0$ are two constants defined in~\eqref{eq:UL}.
-
-Next we bound $\FnormSize{}{\hat \mb-\trueb}^2$ given $\hat\Theta$.
-We consider the profile log-likelihood $\flogl(\hat \Theta,\mb)$ as a function of $\mb\in\tB$.
-For notational convenience, we drop $\hat \Theta$ from $\flogl(\hat \Theta,\mb)$ and simply write $\flogl(\mb)$.
-It follows from the expression of $\flogl(\mb)$ that
-\begin{align}\label{eq:propertyb}
-{\partial \flogl\over \partial b_\ell}&=\sum_{\omega\in\Omega}\left[\mathds{1}_{\{y_{\omega}=\ell\}}
-\frac{\dot{f}(b_\ell-\theta_\omega)}{g_\ell(\theta_\omega)}-\mathds{1}_{\{y_{\omega}=\ell+1\}}
-\frac{\dot{f}(b_\ell-\theta_\omega)}{g_{\ell+1}(\theta_\omega)}\right],\notag\\
-{\partial^2 \flogl\over \partial b^2_\ell}&=\sum_{\omega\in\Omega}\left[\mathds{1}_{\{y_\omega=\ell\}}\frac{\ddot{f}(b_\ell-\theta_\omega)g_\ell(\theta_\omega)-\dot{f}^2(b_\ell-\theta_\omega)}{g^2_\ell(\theta_\omega)}-\mathds{1}_{\{y_\omega=\ell+1\}}\frac{\ddot{f}(b_\ell-\theta_\omega)g_{\ell+1}(\theta_\omega)+\dot{f}^2(b_\ell-\theta_\omega)}{g^2_{\ell+1}(\theta_\omega)}\right],\notag\\
-&\quad  \text{ for all } \ell \in [L-1],\notag \\
-{\partial^2 \flogl\over \partial b_\ell \partial b_{\ell+1}}&=\sum_{\omega\in\Omega}\mathds{1}_{\{y_\omega=\ell+1\}}\frac{\dot{f}(b_\ell-\theta_\omega)\dot{f}(b_{\ell+1}-\theta_\omega)}{g^2_{\ell+1}(\theta_\omega)}\quad \text{and}\quad
-\ {\partial^2 \flogl\over \partial b_\ell \partial b_\ell'} = 0 \text{ if } |\ell-\ell'|>1.
-\end{align}
-Therefore, all entries in $\fploglb$ are upper bounded by $\left\{C_{\alpha,\beta,\Delta} \max_\ell (n_\ell+n_{\ell+1})\right\}>0$, and $\fpploglb$ is a tridiagonal matrix.
-
-By the second-order Taylor's expansion of $\flogl(\mb)$ around $\trueb$, we obtain
-\begin{equation}\label{eq:taylorb}
-\flogl(\hat\mb)=\flogl(\trueb)+(\trueb-\hat\mb)^T\fploglb(\trueb)+ {1\over 2}(\trueb-\hat\mb)^T\fpploglb(\check\mb)(\trueb-\hat\mb),
-\end{equation}
-where $\check\mb=\gamma\trueb+(1-\gamma)\hat\mb$ for some $\gamma\in[0,1]$, and $\fpploglb(\check\mb)$ denotes the $(L-1)$-by-$(L-1)$ Hession matrix evaluated at $\check\mb$.
-
-The linear term in~\eqref{eq:taylorb} can be bounded by Cauchy-Schwartz inequality,
-\begin{align}\label{eq:linearb}
-(\trueb-\hat\mb)^T\fploglb(\trueb)&\leq \FnormSize{}{\trueb-\hat\mb}\FnormSize{}{\fploglb(\trueb)}\\
-&\leq C'_{\alpha,\beta,\Delta}\FnormSize{}{\trueb-\hat\mb}\sqrt{L-1}\sqrt{\left(\sum_k d_k\right)\left( \prod_k d_k\right)},
-\end{align}
-for some constant $C'_{\alpha,\beta,\Delta}>0$, where the last inequality follows from Lemma \ref{lem:sharpergd},
-\[
-\left|{\partial \flogl\over \partial b_\ell}\bigg|_{(\hat \Theta,\trueb)}\right|\leq C'_{\alpha,\beta,\Delta}\sqrt{\left(\sum_k d_k\right)\left(\prod_k d_k\right)},\quad \text{ for all } \ell \in [L-1] \text{ with very high probability}.
-\]
-
-We next bound the quadratic term in \eqref{eq:taylorb}. Note that
-\begin{align}\label{eq:quadraticb}
- &(\trueb-\hat\mb)^T \fpploglb(\check{\mb})(\trueb-\hat\mb)\\
- &=\sum_{\ell\in [L-1]} \left( {\partial^2\tL_{\tY}\over \partial b^2_\ell} \Big|_{\mb=\check\mb} \right)(\hat b_\ell-b^{\text{true}}_{\ell})^2 + 2\sum_{\ell\in[L-1]/\{1\}}\left( {\partial^2\tL_{\tY}\over \partial b_\ell \partial b_{\ell-1}} \Big|_{\mb=\check\mb} \right)(\hat b_\ell-b^{\text{true}}_{\ell})(\hat b_{\ell-1}-b^{\text{true}}_{\ell-1}) \nonumber \\
-&\leq \sum_{\ell\in [L-1]} \left( {\partial^2\tL_{\tY}\over \partial b^2_\ell} \Big|_{\mb=\check\mb} \right)(\hat b_\ell-b^{\text{true}}_{\ell})^2 + \sum_{\ell\in[L-1]/\{1\}}\left( {\partial^2\tL_{\tY}\over \partial b_\ell \partial b_{\ell-1}} \Big|_{\mb=\check\mb} \right)\big[(\hat b_\ell-b^{\text{true}}_{\ell})^2 +(\hat b_{\ell-1}-b^{\text{true}}_{\ell-1})^2\big] \nonumber \\
-&=\left( {\partial^2\tL_{\tY}\over \partial b^2_1}+{\partial^2\tL_{\tY}\over \partial b_1 \partial b_2 }\right)\Big|_{\mb=\check\mb}(\hat b_1-b^{\text{true}}_{1})^2+\left( {\partial^2\tL_{\tY}\over \partial b^2_{L-1}}+{\partial^2\tL_{\tY}\over \partial b_{L-2} \partial b_{L-1} }\right)\Big|_{\mb=\check\mb}(\hat b_{L-1}-b^{\text{true}}_{L-1})^2  \\
-&\quad\quad +\sum_{\ell\in[L-2]/\{1\}}\left( {\partial^2\tL_{\tY}\over \partial b^2_\ell}+{\partial^2\tL_{\tY}\over \partial b_\ell \partial b_{\ell-1}}+{\partial^2\tL_{\tY}\over \partial b_{\ell+1} \partial b_{\ell}} \right)\Big|_{\mb=\check\mb} (\hat b_\ell-b^{\text{true}}_{\ell})^2     \\
-&\leq - D_{\alpha,\beta,\Delta}' \sum_{\ell \in [L-1]}(\hat b_\ell-b_{\text{true},\ell})^2 \nonumber \\
-&=-D_{\alpha,\beta,\Delta}'\FnormSize{}{\hat\mb-\trueb}^2,
-\end{align}
-where
-\begin{align}
-D_{\alpha,\beta,\Delta}'&=\min_{\substack{|\theta|\leq \alpha,\\ \mb\in\tB}}\min_{\ell\in[L-1]} -\left( {\partial^2\tL_{\tY}\over \partial b^2_\ell}+{\partial^2\tL_{\tY}\over \partial b_\ell \partial b_{\ell-1}}+{\partial^2\tL_{\tY}\over \partial b_{\ell+1} \partial b_{\ell}} \right)\\
-&=\min_{\substack{|\theta|\leq \alpha,\\ \mb\in\tB}} \min_{\ell\in[L-1]}\left\{\sum_{\omega\in\Omega}-\mathds{1}_{\{y_\omega=\ell\}}\left( {\ddot{f}(b_{\ell}-\theta_\omega) g_\ell(\theta_\omega)- \dot{f}(b_{\ell}-\theta_\omega)\dot{g}_\ell(\theta_\omega) \over g^2_\ell(\theta_\omega) }\right)\right.
-\\
-&\left.
-\hspace{1in}+\sum_{\omega\in\Omega}\mathds{1}_{\{y_\omega=\ell+1\}} \left( {\ddot{f}(b_\ell-\theta_\omega) g_{\ell+1}(\theta_\omega) - \dot{f}(b_\ell-\theta_\omega)\dot{g}_{\ell+1} (\theta_\omega)\over g^2_{\ell+1}(\theta_\omega)} \right)\right\}\\
-&=\min_{\substack{|\theta|\leq \alpha,\\ \mb\in\tB}} \min_{\ell\in[L-1]}\left\{\KeepStyleUnderBrace{-\sum_{\omega\in\Omega} \mathds{1}_{\{y_\omega=\ell\}}{\partial \over \partial \theta_\omega}\left( \dot{f}(b_\ell-\theta_\omega) \over g_\ell(\theta_\omega) \right)}_{>0}+\KeepStyleUnderBrace{\sum_{\omega\in\Omega}\mathds{1}_{\{y_\omega=\ell+1\}} {\partial \over \partial \theta_\omega}\left( \dot{f}(b_\ell-\theta_\omega) \over g_{\ell+1}(\theta_\omega) \right) }_{>0}\right\}\\
-&\geq D_{\alpha,\beta,\Delta}  \min_{\ell\in[L-1]}\left(n_{\ell}+n_{\ell+1}\right) .
-\end{align}
-Combining inequalities~\eqref{eq:taylorb}, \eqref{eq:linearb} and~\eqref{eq:quadraticb} yields
-\[
-\flogl(\hat \mb)\leq \flogl(\trueb)+C'_{\alpha,\beta,\Delta} \sqrt{(L-1)\left(\sum_k d_k\right)\left(\prod_k d_k\right)} \FnormSize{}{\hat\mb-\trueb} -\frac{D_{\alpha,\beta,\Delta}}{2}\min_{\ell}(n_\ell+n_{\ell+1}) \FnormSize{}{\hat\mb-\trueb}^2 .
-\]
-Since $\hat \mb$ satisfies $\flogl(\hat \mb)-\flogl(\trueb)\geq 0$, we have that
-\[
-C'_{\alpha,\beta,\Delta} \sqrt{(L-1)\left(\sum_k d_k\right)\left(\prod_k d_k\right)} \FnormSize{}{\hat\mb-\trueb} -\frac{D_{\alpha,\beta,\Delta}}{2}\min_{\ell}(n_\ell+n_{\ell+1}) \FnormSize{}{\hat\mb-\trueb}^2\geq 0.
-\]
-Finally,
-\begin{equation}\label{eq:rateb}
-\FnormSize{}{\hat\mb-\trueb}^2\leq {C'_{\alpha,\beta,\Delta} (L-1) \over D_{\alpha,\beta,\Delta}^2}{\left(\sum_k d_k\right)\left(\prod_k d_k\right)\over \min_{\ell}(n_{\ell}+n_{\ell+1})^2}.
-\end{equation}
-Combing~\eqref{eq:ratenew} and~\eqref{eq:rateb} with the following inequality completes the proof.
-\begin{align}
-\text{MSE}\left((\hat \Theta,\hat \mb),(\trueT,\trueb)\right)&\leq{1\over \prod_k d_k}\FnormSize{}{(\hat \Theta,\hat \mb)-(\trueT,\trueb)}^2\\
-&={1\over \prod_k d_k}\FnormSize{}{(\hat \Theta,\hat \mb)-(\hat \Theta,\trueb)+(\hat \Theta,\trueb)-(\trueT,\trueb)}^2\\
-&\leq {1\over\prod_kd_k}\FnormSize{}{(\hat \Theta- \trueT)}^2+ {1\over\prod_kd_k}\FnormSize{}{\hat \mb -\trueb}^2.
-\end{align}
-\end{proof}
-
-\begin{lem}[CLT for independent Bernoulli r.v.'s]\label{lem:clt}
-Let $\{X_n\}$ be a series of independent Bernoulli random variables with possibly different success probabilities $\{p_n\}$.
-Suppose the success probabilities $\{p_n\}$ are bounded, in that, $c_1\leq p_n\leq 1-c_2$ for some positive constants $c_1,c_2>0$ and all $n\in[m]$.
-Define $s^2_m = \sum_{n=1}^m p_n(1-p_n)$ and $Y_n = X_n-p_n$.
-Then,
-\[
-{1\over s_m}\sum_{n=1}^m Y_n \stackrel{\mathcal{D}}{\rightarrow} N(0,1), \quad \text{ as } m \rightarrow \infty.
-\]
-\end{lem}
-
-\begin{proof}
-We apply Lyapunov central limit theorem (CLT) to $\{Y_n\}$. Let us verify Lyapunov's condition for $\delta = 1$:
-\[
-\mathbb{E}[|Y_n|^3] = p_n(1-p_n)^3+ (1-p_n)p_n^3\leq p_n(1-p_n)[(1-p_n)^2+p_n^2]\leq p_n(1-p_n).
-\]
-Summation of the above inequality shows
-\[
-\sum_{n=1}^m\mathbb{E}[|Y_n|^3]\leq \sum_{n=1}^m p_n(1-p_n) = s_m^2.
-\]
-Thus, the Lyapunov condition is satisfied whenever
-\[
-{1\over s_m^3}\sum_{n=1}^m \mathbb{E}[|Y_n|^3]\leq {s_m^2\over s_m^3}\rightarrow 0,
-\]
-or simply $s_m = \sum_{n=1}^m p_n(1-p_n) \geq m\min\{c_1(1-c_1), c_2(1-c_2)\} \rightarrow \infty.$ Applying Lyapunov CLT to $\{Y_n\}$ completes the proof.
-\end{proof}
-
-\begin{lem}[Bound on score function]\label{lem:sharpergd}
-Consider the same set-up as in Theorem~\ref{thm:ratejoint}. Let $\tL_\tY(\Theta,\mb)$ denote the log-likelihood function of $(\Theta, \mb)$ given data tensor $\tY$. Then, with very high probability,
-\[
-\left|{\partial \flogl\over \partial b_\ell}\bigg|_{(\Theta,\trueb)}\right|\leq \left(C_{\alpha,\beta,\Delta}+\tilde{D}_{\alpha,\beta,\Delta}\FnormSize{}{\trueT-\Theta}^2\right)\sqrt{\prod_k d_k},\quad \text{ for all } \ell \in [L-1],
-\]
-In particular, with very high probability,
-\[
-\left|{\partial \flogl\over \partial b_\ell}\bigg|_{(\hat \Theta,\trueb)}\right|\leq C'_{\alpha,\beta,\Delta}\sqrt{\left(\sum_k d_k\right)\left( \prod_k d_k\right)},\quad \text{ for all } \ell \in [L-1],
-\]
-where $C'_{\alpha,\beta,\Delta}>0$ is a constant that depends on $U_{\alpha,\beta,\Delta}, L_{\alpha,\beta,\Delta}, C_{\alpha,\beta,\Delta},  \tilde D_{\alpha,\beta,\Delta}$ defined in~\eqref{eq:UL} and~\eqref{eq:CD}.
-\end{lem}
-
-\begin{proof}
-We only prove the case for $\ell = 1$. Other cases can be proved similarly.
-
-\begin{equation}\label{eq:sharpergd}
-  {\partial \flogl\over \partial b_1}\bigg|_{(\Theta,\trueb)} = \KeepStyleUnderBrace{{\partial \flogl\over \partial b_1}\bigg|_{( \Theta,\trueb)}-\mathbb{E}_\mathcal{Y}\left[{\partial \flogl\over \partial b_1}\bigg|_{(\Theta,\trueb)}\right]}_{\coloneqq A}+\KeepStyleUnderBrace{\mathbb{E}_\mathcal{Y}\left[{\partial \flogl\over \partial b_1}\bigg|_{( \Theta,\trueb)}\right]-\mathbb{E}_\mathcal{Y}\left[{\partial \flogl\over \partial b_1}\bigg|_{(\trueT,\trueb)}\right] }_{\coloneqq B}.
-\end{equation}
-We have used the fact that the score function has mean zero, $\mathbb{E}_\mathcal{Y}\left[{\partial \flogl\over \partial b_1}\big|_{(\trueT,\trueb)}\right]=0 $.
-Here all expectations are taken with respect to $\mathcal{Y}\sim \mathbb{P}(\trueT,\trueb).$\\
-We now bound the two deviation terms in \eqref{eq:sharpergd} separately. The term $A$ in \eqref{eq:sharpergd} is the stochastic deviation of log-likelihood to its expectation:
-\begin{align}
-A &= \sum_{\omega\in\Omega}\left\{[\mathds{1}_{y_\omega = 1}-g_1(\theta^{\text{true}}_\omega)]{\dot{f}(b_1-\theta_\omega)\over g_1(\theta_\omega)}-[\mathds{1}_{y_\omega = 2}-g_2(\theta^{\text{true}}_\omega)]{\dot{f}(b_1-\theta_\omega)\over g_2(\theta_\omega)}\right\}\\
-&\leq C_{\alpha,\beta,\Delta}\sum_{\omega\in\Omega}\KeepStyleUnderBrace{[\mathds{1}_{y_\omega = 1}+\mathds{1}_{y_\omega = 2}- g_1(\theta^{\text{true}}_\omega)-g_2(\theta^{\text{true}}_\omega)]}_{\coloneqq W_\omega}.
-\end{align}
-Note that $\{W_\omega\}$ are independent, centered Bernoulli random variables with bounded success probabilities $g_1(\theta^{\text{true}}_\omega)+g_2(\theta^{\text{true}}_\omega)$.
-By Lemma \ref{lem:clt}, we have
-\[
-\sum_{\omega\in\Omega}W_\omega \stackrel{\mathcal{D}}{\rightarrow} N\left(0,\ \sum_{\omega\in\Omega}\left(g_1(\theta^{\text{true}}_\omega)+g_2(\theta^{\text{true}}_\omega)\right)\left(1-g_1(\theta^{\text{true}}_\omega)-g_2(\theta^{\text{true}}_\omega)\right)\right).
-\]
-Hence, with the fact that $\sum_{\omega\in\Omega}\left(g_1(\theta^{\text{true}}_\omega)+g_2(\theta^{\text{true}}_\omega)\right)\left(1-g_1(\theta^{\text{true}}_\omega)-g_2(\theta^{\text{true}}_\omega)\right)\leq {1\over 4}\prod_k d_k$,
-\begin{equation}\label{eq:A}
-|A| \leq C_{\alpha,\beta,\Delta}\left|\sum_{\omega\in\Omega}W_\omega\right|\leq C_{\alpha,\beta,\Delta}\sqrt{\prod_k d_k}, \quad \text{ with very high probability. }
-\end{equation}
-The second term $B$ in \eqref{eq:sharpergd} is the bias induced by $\Theta$:
-\begin{align}
-|B| &= \left|\sum_{\omega\in\Omega} g_1(\theta^{\text{true}}_\omega)\left({\dot{f}(b_1-\theta_\omega)\over g_1(\theta_\omega)}-{\dot{f}(b_1-\theta^{\text{true}}_\omega)\over g_1(\theta^{\text{true}}_\omega)}\right)-\sum_{\omega\in\Omega} g_2(\theta^{\text{true}}_\omega)\left({\dot{f}(b_1-\theta_\omega)\over g_2(\theta_\omega)}-{\dot{f}(b_1-\theta^{\text{true}}_\omega)\over g_2(\theta^{\text{true}}_\omega)}\right)\right|\\
-&= \left|\sum_{\omega\in\Omega} g_1(\theta^{\text{true}}_\omega)(\theta^{\text{true}}_\omega-\theta_\omega)\left\{{\partial\over\partial \theta}\left({\dot{f}(b_1-\theta)\over g_1(\theta)}\right)\bigg|_{\rho_\omega\theta_\omega+(1-\rho_\omega)\theta^{\text{true}}_\omega}\right\}\right.\\
-&\hspace{0.5in}\left.-\sum_{\omega\in\Omega} g_2(\theta^{\text{true}}_\omega)(\theta^{\text{true}}_\omega-\theta_\omega)\left\{{\partial\over\partial \theta}\left({\dot{f}(b_1-\theta)\over g_2(\theta)}\right)\bigg|_{\rho_\omega'\theta_\omega+(1-\rho_\omega')\theta^{\text{true}}_\omega}\right\}\right|\\
-&\leq \tilde{D}_{\alpha,\beta,\Delta}\sum_{\omega\in\Omega}|g_1(\theta^{\text{true}}_\omega)+g_2(\theta^{\text{true}}_\omega)||\theta^{\text{true}}_\omega-\theta_\omega|.
-\end{align}
-By Cauchy-Schwartz inequality with the fact that $g_1(\theta^{\text{true}}_\omega)+g_2(\theta^{\text{true}}_\omega)\leq 1$,
-\begin{equation}\label{eq:B}
-|B| \leq \tilde{D}_{\alpha,\beta,\Delta}\sqrt{\prod_k d_k} \FnormSize{}{\trueT-\Theta}.
-\end{equation}
-Plugging \eqref{eq:A} and \eqref{eq:B} back to \eqref{eq:sharpergd} yields that
-\[
-{\partial \flogl\over \partial b_1}\bigg|_{(\Theta,\trueb)}\leq \left(C_{\alpha,\beta,\Delta}+\tilde{D}_{\alpha,\beta,\Delta}\FnormSize{}{\trueT-\Theta}\right)\sqrt{\prod_k d_k}.
-\]
-holds with very high probability. The second inequality in the lemma comes from~\eqref{eq:ratenew} that $\FnormSize{}{\trueT-\hat\Theta}\leq \tO(\sqrt{\sum_k d_k}).$
-\end{proof}
-
-\section{Additional results of HCP analysis}\label{sec:additionalHCP}
-We perform clustering analyses based on the Tucker representation of the estimated tensor parameter $\hat\Theta$.
-The procedure is motivated from the higher-order extension of Principal Component Analysis (PCA) or Singular Value Decomposition (SVD).
-Recall that, in the matrix case, we usually perform clustering on an $m\times n$ (normalized) matrix $X$ based on the following procedure.
-First, we factorize $X$ into
-\begin{equation}
-X = U\Sigma V^T,
-\end{equation}
-where $\Sigma$ is a diagonal matrix and $ U,V$ are factor matrices with orthogonal columns.
-Second, we take each column of $V$ as a principal axis and each row in $U\Sigma$ as principal component.
-A subsequent multivariate clustering method (such as $K$-means) is then applied to the $m$ rows of $U\Sigma$.
-
-We apply a similar clustering procedure to the estimated parameter tensor $\hat\Theta$. Based on Tucker representation of $\hat \Theta$, we have
-\begin{equation}\label{eq:Tuckerest}
-\hat \Theta = \hat \tC\times_1\hat \mM_1\times_2\cdots\times_K\hat \mM_K,
-\end{equation}
-where $\hat \tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is the estimated core tensor, $\hat \mM_k\in\mathbb{R}^{d_k\times r_k}$ are estimated factor matrices with orthogonal columns, and $\times_k$ denotes the tensor-by-matrix multiplication~\citep{kolda2009tensor}.
-The mode-$k$ matricization of~\eqref{eq:Tuckerest} gives
-\begin{equation}
-\hat \Theta_{(k)} = \hat \mM_k\hat \tC_{(k)}\left(\hat \mM_K\otimes\cdots\otimes\hat \mM_1\right),
-\end{equation}
-where $\hat \Theta_{(k)} , \hat \tC_{(k)}$ denote the mode-$k$ unfolding of $\hat \Theta$ and $\hat \tC$, respectively.
-Then, the mode-$k$ clustering can be performed as follows.
-First, we take columns in $\left(\hat \mM_K\otimes\cdots\otimes\hat \mM_1\right)$ as principal axes and rows in $\hat \mM_k\hat \tC_{(k)}$ as principal components.
-Then, we perform $K$-means clustering method to the $d_k$ rows of the matrix $\hat \mM_k\hat \tC_{(k)}$.
-
-We perform a clustering analysis on the 68 brain nodes using the produce described above.
-Our ordinal tensor method outputs the estimated parameter tensor $\hat\Theta\in\mathbb{R}^{68\times 68\times136}$ with rank $(23,23,8)$.
-We apply $K$-means to the mode-1 principal component matrix of size $68\times 184$ ($184=23\times8$).
-The elbow method suggests 11 clusters among the 68 nodes (see Figure~\ref{figure:elbow}).
-
-\begin{figure}[H]
-\begin{center}
-\includegraphics[width=0.4\textwidth]{elbowm.pdf}
-\end{center}
-\caption{Elbow plot for determining the number of clusters in $K$-means.}\label{figure:elbow}
-\end{figure}
-
-\bibliography{tensor_wang}
-\bibliographystyle{apalike}
 \end{document}
+
+% This document was modified from the file originally made available by
+% Pat Langley and Andrea Danyluk for ICML-2K. This version was created
+% by Iain Murray in 2018, and modified by Alexandre Bouchard in
+% 2019 and 2020. Previous contributors include Dan Roy, Lise Getoor and Tobias
+% Scheffer, which was slightly modified from the 2010 version by
+% Thorsten Joachims & Johannes Fuernkranz, slightly modified from the
+% 2009 version by Kiri Wagstaff and Sam Roweis's 2008 version, which is
+% slightly modified from Prasad Tadepalli's 2007 version which is a
+% lightly changed version of the previous year's version by Andrew
+% Moore, which was in turn edited from those of Kristian Kersting and
+% Codrina Lauth. Alex Smola contributed to the algorithmic style files.

### 2020-06-15 old_draft/old_preprint/old_arXiv/ordinalT_supp_v4.tex
organize

--- 2020-02-13 ICML/arxiv/ordinalT_supp_v4.tex
@@
 
-\subsection{Clustering results on HCP}
-We perform clustering analysis to the 68 brain nodes using the the produce described in Section~\ref{sec:clustering}.
-Recall that our ordinal tensor method outputs the estimated parameter tensor $\hat\Theta\in\mathbb{R}^{68\times 68\times136}$ with rank $(23,23,8)$.
-The number of clusters in the $K$-means is chosen based on the elbow method (see figure~\ref{figure:elbow}). Table~\ref{table:clustering} shows the eight clusters of brain nodes.
+We perform clustering analysis on the 68 brain nodes using the produce described above.
+Oordinal tensor method outputs the estimated parameter tensor $\hat\Theta\in\mathbb{R}^{68\times 68\times136}$ with rank $(23,23,8)$.
+We apply $K$-means to the mode-1 principal component matrix of size $68\times 184$ ($184=23\times8$).
+The elbow method suggests eight clusters among the 68 nodes (see Figure~\ref{figure:elbow}).
 \begin{figure}[H]

### 2020-06-18 paper/Camera ready/ordinalTv6.tex
typo
typos
typo in table 2 caption

@@
 \caption{Empirical relationship between (relative) MSE versus (a) dimension $d$, (b) signal level $\alpha$, (c) observation fraction $\rho$, and (d) number of ordinal levels $L$.
-In panels (b)-(d), we plot the relative MSE $=\FnormSize{}{\hat \Theta-\trueT}/\FnormSize{}{\trueT}$ for better visualization.}
+In panels (b)-(d), we plot the relative MSE $=\FnormSize{}{\hat \Theta-\trueT}^2/\FnormSize{}{\trueT}^2$ for better visualization.}
 \label{fig:finite}
@@
 Suppose the link function $f$ satisfies Assumption~\ref{ass:joint}.
-Define $r_{\max}=\max_k r_k$, and assume $r_{\max}=\tO(1)$.
+Define $r_{\max}=\max_k r_k+1$, and assume $r_{\max}=\tO(1)$.
 
@@
 \end{align}
-The second inequality in \eqref{eq:negativeH} is from the condition that 
-\[
-\sum_{\omega}\theta_\omega^2\in [0,\ \alpha^2 d_{\text{total}})\quad \text{and}\quad \sum_\ell b_\ell^2 \in[c L_{\text{total}}\Delta^2,\  L_{\text{total}}\beta^2],
+The second inequality in \eqref{eq:negativeH} is from the conditions that 
+\[
+\sum_{\omega}\theta_\omega^2\in [0,\ \alpha^2 d_{\text{total}}]\quad \text{and}\quad \sum_\ell b_\ell^2 \in[c L_{\text{total}}\Delta^2,\  L_{\text{total}}\beta^2],
 \]
@@
 (b)-(c) Cluster II/III consists of nodes within left/right hemisphere only.
-Node name are shown in abbreviation.
+Node names are shown in abbreviation.
 Edges are colored based on predicted connection level averaged across individuals.

### 2020-06-22 paper/Camera ready/ordinalTv6.tex
update

@@
 \usepackage{booktabs} % for professional tables
-
+\usepackage{caption}
 % hyperref makes hyperlinks in the resulting PDF.
@@
 \centerline{\includegraphics[width=1\textwidth]{compare.pdf}}
-\vspace{-.3cm}
 \caption{Comparison with previous work.
@@
 $K$: tensor order; $L$: number of ordinal levels; $d$: dimension at each mode.
-}~\label{tab:compare}
-\vspace{-.7cm}
+}\label{tab:compare}
 \end{table*}
@@
 \subsection{Observation model}
-\vspace{-.1cm}
 Let $\tY$ denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional data tensor.
@@
 The ordinal tensor model~\eqref{eq:model} with certain types of link $f$ has the equivalent representation as an $L$-level quantization model on $\tY=\entry{y_\omega}$:
-\begin{equation}\label{eq:quantization}
-y_\omega=\sum_{\ell\in[L]} \ell \mathds{1}\{y^*_\omega \in(b_{\ell-1},b_\ell]\},\ \forall \ \omega\in[d_1]\times \cdots \times [d_k],
-\end{equation}
-%y_\omega&=
-%\begin{cases}
-%1,& \text{if $y^*_\omega\in(-\infty, b_1]$},\\
-%2,& \text{if $y^*_\omega\in(b_1, b_2]$},\\
-%\vdots  &\vdots\\
-%L,& \text{if $y^*_\omega\in(b_{L-1}, \infty)$},\\
-%\end{cases}
+\begin{align}\label{eq:quantization}
+%y_\omega=\sum_{\ell\in[L]} \ell \mathds{1}\{y^*_\omega \in(b_{\ell-1},b_\ell]\},\ \forall \ \omega\in[d_1]\times \cdots \times [d_k],
 %\end{equation}
-where we define $b_0=-\infty$ and $b_L=\infty$. Here, $\tY^*=\entry{y^*_\omega}$ is a latent continuous-valued tensor following an additive noise model:
+y_\omega&=
+\begin{cases}
+1,& \text{if $y^*_\omega\in(-\infty, b_1]$},\\
+2,& \text{if $y^*_\omega\in(b_1, b_2]$},\\
+\vdots  &\vdots\\
+L,& \text{if $y^*_\omega\in(b_{L-1}, \infty)$},\\
+\end{cases}
+\end{align}
+for all $\omega\in[d_1]\times \cdots \times [d_k]$. Here, $\tY^*=\entry{y^*_\omega}$ is a latent continuous-valued tensor following an additive noise model:
 \begin{align}\label{eq:latent}
@@
 \section{Rank-constrained M-estimator}\label{sec:theory}
-\vspace{-.1cm}
 We present a general treatment to both problems mentioned above.
@@
 \subsection{Estimation error for tensor denoising}\label{sec:denosing}
-\vspace{-.1cm}
 For the tensor denoising problem, we assume that the full set of tensor entries are observed. We assess the estimation accuracy using the mean squared error (MSE):
@@
 \centering
-\includegraphics[width=8cm]{algorithm.pdf}\label{fig:stability}
-\vspace{-.2cm}
-\caption{Trajectory of objective function with various $d$ and $r$.}
-\vspace{-.2cm}
+\includegraphics[width=8cm]{algorithm.pdf}
+\caption{Trajectory of objective function with various $d$ and $r$.}\label{fig:stability}
 \end{figure}
@@
 \section{Experiments}\label{sec:experiment}
-In this section, we evaluate the empirical performance of our methods\footnote{Software:  https://CRAN.R-project.org/package=tensorordinal}.
+In this section, we evaluate the empirical performance of our methods\footnote{Software package: \url{https://CRAN.R-project.org/package=tensorordinal}}.
 We investigate both the complete and the incomplete settings, and compare the recovery accuracy with other tensor-based methods.
@@
 \includegraphics[width=8.3cm]{panel.pdf}
-\vspace{-.2cm}
 \caption{Empirical relationship between (relative) MSE versus (a) dimension $d$, (b) signal level $\alpha$, (c) observation fraction $\rho$, and (d) number of ordinal levels $L$.
@@
 
+\vspace{-.2cm}
 \subsection{Comparison with alternative methods}
 Next, we compare our ordinal tensor method ({\bf Ordinal-T}) with three popular low-rank methods:
-
 \begin{itemize}[itemsep=0pt,topsep=0pt,leftmargin=*,partopsep=0pt]
@@
 
-\begin{figure}[t]
+\begin{figure}[http]
 \includegraphics[width=8.3cm]{panel2.pdf}
-\vspace{-.4cm}
-\caption{Performance comparison in MCR (a, b) and MAD (c, d).
+\caption{Performance comparison for predicting most likely labels.
+(a, c) Prediction errors versus sample complexity $\rho=|\Omega|/d^K$ when $L=5$.
 (b, d) Prediction errors versus the number of ordinal levels $L$ when $\rho=0.8$.
-(a, c) Prediction errors versus sample complexity $\rho=|\Omega|/d^K$ when $L=5$.
 }\label{fig:compare}
@@
 
+\begin{figure}[http]
+\includegraphics[width=8.3cm]{panel3.pdf}
+\caption{Performance comparison for predicting median labels.
+(a, c) Prediction error versus sample complexity $\rho=|\Omega|/d^K$ when $L=5$.
+(b, d) Prediction error versus the number of ordinal levels $L$, when $\rho=0.8.$}\label{fig:compare2}
+\vspace{-.4cm}
+\end{figure}
+
 We also compare the methods by their performance in predicting the median labels, $y_\omega^{\text{median}}=\min\{\ell\colon\mathbb{P}(y_\omega=\ell)\geq 0.5\}$.
 Under the latent variable model~\eqref{eq:latent} and Assumption~\ref{ass:link}, the median label is the quantized $\theta_\omega$ without noise; i.e.\ $y_\omega^{\text{median}}=\sum_\ell \mathds{1}\{\theta_\omega\in(b_{\ell-1},b_\ell]\}$.
-Similar results show the outperformance of our method (see the Supplement).
+ We utilize the same simulation setting as in the earlier experiment.
+Figure~\ref{fig:compare2} shows that our method outperforms the others in both MCR and MAD.
+The improved accuracy comes from the incorporation of multilinear low-rank structure, multi-level observations, and the ordinal structure.
+Interestingly, the median estimator tends to yield smaller MAD than the mode estimator, $\text{MAD}(\tY^{\text{median}}, \hat \tY^{\text{median}})\leq \text{MAD}(\tY^{\text{mode}}, \hat \tY^{\text{mode}})$  (Figures~\ref{fig:compare}a-b vs.\ Figures~\ref{fig:compare2}a-b), for the three multilevel methods ({\bf 1bit-sign-T}, {\bf Ordinal-M}, and {\bf Ordinal-T}).
+The mode estimator, on the other hand, tends to yield smaller MCR than the median estimator, $\text{MCR}(\tY^{\text{mode}}, \hat \tY^{\text{mode}})\leq\text{MCR}(\tY^{\text{median}}, \hat \tY^{\text{median}})$ (Figures~\ref{fig:compare}c-d vs.\ Figures~\ref{fig:compare2}c-d).
+This tendency is from the property that the median estimator $\hat y^{(\text{median})}_\omega$ minimizes $R_1(z)=\mathbb{E}_{y_\omega}|y_\omega-z|$, whereas the mode estimator $\hat y^{(\text{mode})}_\omega $ minimizes $R_2(z)=\mathbb{E}_{y_\omega}\mathds{1}\{y_\omega=z\}$.
+Here the expectation is over the categorical distribution of $y_\omega$ given parameters $\hat \Theta$ and $\hat \mb$.
 
@@
 
+\begin{table*}[http]
+\includegraphics[width=1\textwidth]{HCPtable.pdf}
+\vspace{-.7cm}
+\caption{Node clusters in the HCP analysis.
+The first alphabet in the node name indicates the left (L) or right (R) hemisphere.
+The number in the parentheses indicates the node count in each cluster.
+}  \label{table:clustering}
+\vspace{-.4cm}
+\end{table*}
+
 \subsection{Human Connectome Project (HCP)}
@@
 Based on the estimated Tucker factors $\{\hat \mM_k\}$, we perform a clustering analysis via K-mean on the brain nodes (see detailed procedure in the Supplement).
-We find that the 68 brain nodes are grouped into 11 clusters, and the clustering captures the spatial separation of brain nodes (see the Supplement).
-In particular, the top three clusters represent the connections in the left-, right-, and inter-hemispheres; the smaller clusters represent local regions driving by similar nodes.
-For example, the cluster IV consists of nodes in the left supramarginal gyrus region.
-The identified similarity among brain nodes suggests the applicability of our method.
-
-We also compare the goodness-of-fit of various tensor methods on the HPC data.
+We find that the clustering successfully captures the spatial separation between brain regions (Table~\ref{table:clustering}).
+In particular, cluster I represents the connection between the left and right hemispheres, whereas clusters II-III represent the connection within each of the half brains (Figure~\ref{figure:brain image}).
+Other smaller clusters represent local regions driving by similar nodes (Table~\ref{table:clustering}).
+For example, the cluster IV/VII consists of nodes in the supramarginal gyrus region in the left/right hemisphere.
+This region is known to be involved in visual word recognition and reading~\citep{stoeckel2009supramarginal}.
+The identified similarities among nodes without external annotations illustrate the applicability of our method to clustering analysis.
+
+\begin{figure}[http]
+\centering
+\includegraphics[width = .5\textwidth]{HCP.pdf}
+\caption{Top three clusters in the HCP analysis.
+(a) Cluster I reflects the connections between two brain hemispheres.
+(b)-(c) Cluster II/III consists of nodes within left/right hemisphere only.
+Node names are shown in abbreviation.
+Edges are colored based on estimated connection averaged across individuals.
+}  \label{figure:brain image}
+\vspace{-.2cm}
+\end{figure}
+
+We acompare the goodness-of-fit of various tensor methods on the HPC data.
 Table~\ref{table:CV} summarizes the prediction error via 5-fold stratified cross-validation averaged over 10 runs.
 Our method outperforms the others, especially in MAD.
+
+\subsection{InCarMusic recommendation system}
+
+We apply ordinal tensor completion to a recommendation system {\it InCarMusic}.
+{\it InCarMusic} is a mobile application that offers music recommendation to passengers of cars based on contexts~\cite{baltrunas2011incarmusic}.
+We conduct tensor completion on the 42-by-139-by-26 tensor with 2,844 observed entries only.
+Table~\ref{table:CV} shows the averaged prediction error via 5-fold cross validation.
+The high missing rate makes the accurate classification challenging.
+Nevertheless, our method achieves the best performance among the three.
 
@@
 }
-\vspace{-.2cm}
 \caption{Comparison of prediction error in the HPC and InCarMusic analyses. Standard errors are reported in parentheses.}\label{table:CV}
-\vspace{-.3cm}
 \end{table}
-
-\subsection{InCarMusic recommendation system}
-We apply ordinal tensor completion to a recommendation system {\it InCarMusic}.
-{\it InCarMusic} is a mobile application that offers music recommendation to passengers of cars based on contexts~\cite{baltrunas2011incarmusic}.
-We conduct tensor completion on the 42-by-139-by-26 tensor with 2,844 observed entries only.
-Table~\ref{table:CV} shows the averaged prediction error via 5-fold cross validation.
-The high missing rate makes the accurate classification challenging.
-Nevertheless, our method achieves the best performance among the three.
 
@@
 \begin{align}
-\anormSize{}{\Theta}&=\inf\{t>0: \Theta\in t\text{conv}(T_{\pm})\}\\
-&=\inf\left\{\sum_{\tX\in T_{\pm}}c_x\colon \Theta=\sum_{\tX\in T_{\pm}} c_x\tX, \ c_x>0\right\}.
+\anormSize{}{\Theta}&=\inf\{t>0\colon \Theta\in t\text{conv}(T_{\pm})\}\\
+&=\inf\left\{\sum_{\tX\in T_{\pm}}c_{\tX} \colon\ \Theta=\sum_{\tX\in T_{\pm}} c_{\tX}\tX, \ c_{\tX}>0\right\}.
 \end{align}
@@
 
-\section{Additional results on simulations}
-We also compare the methods by their performance in predicting the median labels, $y_\omega^{\text{median}}=\min\{\ell\colon\mathbb{P}(y_\omega=\ell)\geq 0.5\}$.
-Under the latent variable model~\eqref{eq:latent} and Assumption~\ref{ass:link}, the median label is the quantized $\theta_\omega$ without noise; i.e.\ $y_\omega^{\text{median}}=\sum_\ell \mathds{1}\{\theta_\omega\in(b_{\ell-1},b_\ell]\}$.
- We utilize the same simulation setting as in the earlier experiment.
-Figure~\ref{fig:compare2} shows that our method outperforms the others in both MCR and MAD.
-The improved accuracy comes from the incorporation of multilinear low-rank structure, multi-level observations, and the ordinal structure.
-Interestingly, the median estimator tends to yield smaller MAD than the mode estimator, $\text{MAD}(\tY^{\text{median}}, \hat \tY^{\text{median}})\leq \text{MAD}(\tY^{\text{mode}}, \hat \tY^{\text{mode}})$  (Figures~\ref{fig:compare}a-b vs.\ Figures~\ref{fig:compare2}a-b), for the three multilevel methods ({\bf 1bit-sign-T}, {\bf Ordinal-M}, and {\bf Ordinal-T}).
-The mode estimator, on the other hand, tends to yield smaller MCR than the median estimator, $\text{MCR}(\tY^{\text{mode}}, \hat \tY^{\text{mode}})\leq\text{MCR}(\tY^{\text{median}}, \hat \tY^{\text{median}})$ (Figures~\ref{fig:compare}c-d vs.\ Figures~\ref{fig:compare2}c-d).
-This tendency is from the property that the median estimator $\hat y^{(\text{median})}_\omega$ minimizes $R_1(z)=\mathbb{E}_{y_\omega}|y_\omega-z|$, whereas the mode estimator $\hat y^{(\text{mode})}_\omega $ minimizes $R_2(z)=\mathbb{E}_{y_\omega}\mathds{1}\{y_\omega=z\}$.
-Here the expectation is over the categorical distribution of $y_\omega$ given parameters $\hat \Theta$ and $\hat \mb$.
-
-\begin{figure}[ht]
-\begin{center}
-\includegraphics[width=.55\textwidth]{panel3.pdf}
-\end{center}
-\caption{Performance comparison for predicting median labels.
-(a, c) Prediction error versus sample complexity $\rho=|\Omega|/d^K$ when $L=5$.
-(b, d) Prediction error versus the number of ordinal levels $L$, when $\rho=0.8.$}\label{fig:compare2}
-\end{figure}
-
 \section{Additional explanations of HCP analysis}\label{sec:additionalHCP}
@@
 The elbow method suggests 11 clusters among the 68 nodes (see Figure~\ref{figure:elbow}).
-We find that the clustering successfully captures the spatial separation between brain regions (Table~\ref{table:clustering}).
-In particular, cluster I represents the connection between the left and right hemispheres, whereas clusters II-III represent the connection within each of the half brains (Figure~\ref{figure:brain image}).
-Other smaller clusters represent local regions driving by similar nodes (Table~\ref{table:clustering}).
-For example, the cluster IV/VII consists of nodes in the supramarginal gyrus region in the left/right hemisphere.
-This region is known to be involved in visual word recognition and reading~\citep{stoeckel2009supramarginal}.
-The identified similarities among nodes without external annotations illustrate the applicability of our method to clustering analysis.
 
@@
 \end{figure}
-
-\begin{figure}[H]
-\centering
-\includegraphics[width = .7\textwidth]{HCP.pdf}
-\caption{Top three clusters in the HCP analysis.
-(a) Cluster I reflects the connections between two brain hemispheres.
-(b)-(c) Cluster II/III consists of nodes within left/right hemisphere only.
-Node names are shown in abbreviation.
-Edges are colored based on predicted connection level averaged across individuals.
-}  \label{figure:brain image}
-\end{figure}
-
-\begin{table}[H]
-\center
-\begin{sc}
-\resizebox{\columnwidth}{!}{
-\begin{tabular}{|c|c|c|c|}
-\hline
-Cluster     & \multicolumn{3}{c|}{I}\\ \hline
-Brain nodes & \multicolumn{3}{l|}{\begin{tabular}[c]{@{}l@{}} l.frontalpole, l.temporalpole, l.medialorbitofrontal,  l.cuneus, l.parahippocampal, l.lingual,            \\
-r.frontalpole, r.temporalpole,  r.medialorbitofrontal,  r.cuneus,  r.parahippocampal \\
-\end{tabular}} \\
-\hline
-Cluster & \multicolumn{3}{c|}{II}\\ \hline
-Brain nodes & \multicolumn{3}{l|}{\begin{tabular}[c]{@{}l@{}}
-l.caudalmiddlefrontal,  l.inferiorparietal,  l.insula,  l.isthmuscingulate, l.lateraloccipital(2), \\
-l.parsopercularis, l.parstriangularis, l.postcentral, l.precuneus, l.superiorfrontal, l.superiortemporal(3)
-\end{tabular}} \\
-\hline
-Cluster & \multicolumn{3}{c|}{III}\\ \hline
-Brain nodes & \multicolumn{3}{l|}{\begin{tabular}[c]{@{}l@{}}
-r.caudalmiddlefrontal, r.inferiorparietal, r.insula, r.isthmuscingulate, r.lateraloccipital(2), r.lingual, \\
-r.parsopercularis, r.parstriangularis, r.postcentral, r.precentral, r.precuneus, r.superiorfrontal(3), \\
-r.superiorparietal,  r.superiortemporal(3)
-\end{tabular}} \\
-\hline
-Cluster      & IV    & V  & VI\\
-\hline
-Brain nodes & \begin{tabular}[c]{@{}c@{}}l.supramarginal(4)\end{tabular}&
-\begin{tabular}[c]{@{}c@{}}l.inferiortemporal(3)\end{tabular}&
-\begin{tabular}[c]{@{}c@{}}l.middletemporal(3)\end{tabular}\\
-\hline
-Cluster & VII  & VIII & VIIII\\
-\hline
-Brain nodes &
-\begin{tabular}[c]{@{}c@{}}r.supramarginal(4)\end{tabular}&
-\begin{tabular}[c]{@{}c@{}}r.inferiortemporal(3)\end{tabular}&
-\begin{tabular}[c]{@{}c@{}}r.middletemporal(3)\end{tabular}\\
-\hline
-Cluster &X & \multicolumn{2}{c|}{XI}  \\
-\hline
-Brain nodes &
-\begin{tabular}[c]{@{}c@{}}l.superiorfrontal(2)\end{tabular}&
- \multicolumn{2}{c|}{l.precentral,  l.superiorparietal}\\
- \hline
-\end{tabular}
-}
-\end{sc}
-\caption{Node clusters in the HCP analysis.
-The first alphabet in the node name indicates the left (L) or right (R) hemisphere.
-The number in the parentheses indicates the node count in each cluster.
-}  \label{table:clustering}
-\end{table}
 

### 2020-06-23 paper/Camera ready/paper.tex
Add files via upload

--- 2020-06-22 paper/Camera ready/ordinalTv6.tex
@@
 
-\section{Data Applications}
+\section{Data Applications}\label{sec:dataapplication}
 We apply our ordinal tensor method to two real-world datasets.
@@
 }
-\caption{Comparison of prediction error in the HPC and InCarMusic analyses. Standard errors are reported in parentheses.}\label{table:CV}
+\caption{Comparison of prediction error in the HPC and {\it InCarMusic} analyses. Standard errors are reported in parentheses.}\label{table:CV}
 \end{table}
@@
 \bibliographystyle{icml2020}
-
-\clearpage
-\appendix
-\onecolumn
-
-\icmltitle{Supplements for ``Tensor denoising and completion based on \\ ordinal observations''}
-
-\section{Proofs}\label{sec:proof}
-Here, we provide proofs of the theoretical results presented in Sections~\ref{sec:theory}.
-
-\subsection{Estimation error for tensor denoising}\label{sec:proofMSE}
-\begin{proof}[Proof of Theorem~\ref{thm:rate}]
-We suppress the subscript $\Omega$ in the proof, because the tensor denoising assumes complete observation $\Omega=[d_1]\times \cdots \times [d_K]$.
-It follows from the expression of $\flogl(\Theta)$ that
-\begin{align}\label{eq:property}
-{\partial \flogl\over \partial \theta_\omega}&=\sum_{\ell\in[L]}\mathds{1}\{y_{\omega}=\ell\}
-{\dot{g}_\ell(\theta_\omega)\over g_\ell(\theta_\omega)},\notag\\
-{\partial^2 \flogl\over \partial \theta_\omega^2}&=\sum_{\ell\in[L]}\mathds{1}\{y_\omega=\ell\}{\ddot{g}_\ell(\theta_\omega)g_\ell(\theta_\omega)-\dot{g}^2_\ell(\theta_\omega)\over g^2_\ell(\theta_\omega)}\ \text{and}\quad
-{\partial^2 \flogl\over \partial \theta_\omega \theta_\omega'}=0\ \text{if}\ \omega\neq \omega',
-\end{align}
-for all $\omega\in[d_1]\times \cdots \times [d_K]$.
-Define $d_{\text{total}}=\prod_k d_k$.
-Let $\fplogl\in\mathbb{R}^{d_1\times\cdots\times d_K}$ denote the tensor of gradient with respect to $\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}$, and $\fpplogl$ the corresponding Hession matrix of size $d_\text{total}$-by-$d_{\text{total}}$.
-Here, $\Vec(\cdot)$ denotes the operation that turns a tensor into a vector.
-By~\eqref{eq:property}, $\fpplogl$ is a diagonal matrix.
-Recall that
-\begin{equation}\label{eq:bound}
-U_\alpha=\max_{\ell\in[L],|\theta|\leq \alpha}{|\dot{g}_\ell(\theta)|\over g_\ell(\theta)}>0 \quad \text{and}\quad
-L_\alpha=\min_{\ell\in[L],|\theta|\leq \alpha} {\dot{g}^2_\ell(\theta)-\ddot{g}_\ell(\theta)g_\ell(\theta)\over g^2_\ell(\theta)}>0.
-\end{equation}
-Therefore, the entries in $\fplogl$ are upper bounded in magnitude by $U_\alpha>0$, and all diagonal entries in $\fpplogl$ are upper bounded by $-L_{\alpha}<0$.
-
-By the second-order Taylor's expansion of $\flogl(\Theta)$ around $\trueT$, we obtain
-\begin{equation}\label{eq:taylor}
-\flogl(\Theta)=\flogl(\trueT)+\langle\Vec(\fplogl(\trueT)),\ \Vec(\Theta-\trueT)\rangle+{1\over 2}\Vec(\Theta-\trueT)^T\fpplogl(\check\Theta)\Vec(\Theta-\trueT),
-\end{equation}
-where $\check\Theta=\gamma\trueT+(1-\gamma)\Theta$ for some $\gamma\in[0,1]$, and $\fpplogl(\check\Theta)$ denotes the $d_{\text{total}}$-by-$d_\text{total}$ Hession matrix evaluated at $\check\Theta$.
-
-We first bound the linear term in~\eqref{eq:taylor}. Note that, by Lemma~\ref{lem:inq},
-\begin{equation}\label{eq:linear}
-|\langle\Vec(\fplogl(\trueT), \Vec(\Theta-\trueT)  \rangle|\leq \snormSize{}{\fplogl(\trueT)} \nnormSize{}{\Theta-\trueT},
-\end{equation}
-where $\snormSize{}{\cdot}$ denotes the tensor spectral norm and $\nnormSize{}{\cdot}$ denotes the tensor nuclear norm. Define
-\[
-s_\omega={\partial \tL_\tY\over \partial \theta_\omega}\Big|_{\Theta=\trueT} \;\; \textrm{ for all } \; \omega\in[d_1]\times\cdots\times [d_K].
-\]
-Based on~\eqref{eq:property} and the definition of $U_\alpha$, $\fplogl(\trueT)=\entry{s_{\omega}}$ is a random tensor whose entries are independently distributed satisfying
-\begin{equation}\label{eq:norm}
-\mathbb{E}(s_\omega)=0,\quad |s_\omega|\leq U_\alpha, \quad \text{for all }\omega\in[d_1]\times \cdots \times [d_K].
-\end{equation}
-By lemma~\ref{lem:noisytensor}, with probability at least $1-\exp(-C_1 \sum_kd_k)$, we have
-\begin{equation}\label{eq:normrandom}
-\snormSize{}{\fplogl(\trueT)} \leq C_2 U_\alpha\sqrt{\sum_k d_k},
-\end{equation}
-where $C_1, C_2$ are two positive constants that depend only on $K$.
-Furthermore, note that $\text{rank}(\Theta)\leq \mr$, $\text{rank}(\trueT)\leq \mr$, so $\text{rank}(\Theta-\trueT)\leq 2\mr$.
-By lemma~\ref{lem:nuclear}, $\nnormSize{}{\Theta-\trueT}\leq (2r_{\max})^{K-1\over 2}\FnormSize{}{\Theta-\trueT}$.
-Combining~\eqref{eq:linear}, \eqref{eq:norm} and \eqref{eq:normrandom}, we have that, with probability at least $1-\exp(-C_1 \sum_kd_k)$,
-\begin{equation}\label{eq:linearconclusion}
-|\langle \Vec(\fplogl(\trueT)), \Vec(\Theta-\trueT)  \rangle | \leq C_2 U_\alpha  \sqrt{r_{\max}^{K-1} \sum_k d_k}  \FnormSize{}{\Theta-\trueT}.
-\end{equation}
-
-We next bound the quadratic term in \eqref{eq:taylor}. Note that
-\begin{align}\label{eq:quadratic}
- \Vec(\Theta-\trueT)^T \fpplogl(\check{\Theta})\Vec(\Theta-\trueT)&=\sum_\omega \left( {\partial^2\tL_{\tY}\over \partial \theta^2_\omega} \Big|_{\Theta=\check\Theta} \right)(\theta_\omega-\theta_{{\text{true}},\omega})^2 \nonumber \\
-&\leq - L_\alpha\sum_{\omega}(\Theta_{\omega}-\Theta_{\text{true},\omega})^2 \nonumber \\
-&=-L_\alpha\FnormSize{}{\Theta-\trueT}^2,
-\end{align}
-where the second line comes from the fact that  $\mnormSize{}{\check\Theta}\leq \alpha$ and the definition of $L_\alpha$.
-
-Combining~\eqref{eq:taylor}, \eqref{eq:linearconclusion} and~\eqref{eq:quadratic}, we have that, for all $\Theta\in\tP$, with probability at least $1-\exp(-C_1 \sum_kd_k)$,
-\[
-\tL_\tY(\Theta)\leq \tL_{\tY}(\trueT)+C_2U_\alpha  \left(r_{\max}^{K-1}\sum_k d_k\right)^{1/2}  \FnormSize{}{\Theta-\trueT}-{L_\alpha\over 2}\FnormSize{}{\Theta-\trueT}^2.
-\]
-In particular, the above inequality also holds for $\hat \Theta\in\tP$. Therefore,
-\[
-\tL_\tY(\hat \Theta)\leq \tL_{\tY}(\trueT)+C_2U_\alpha \left(r_{\max}^{K-1}\sum_k d_k\right)^{1/2}  \FnormSize{}{\hat \Theta-\trueT}-{L_\alpha\over 2} \FnormSize{}{\hat \Theta-\trueT}^2.
-\]
-Since $\hat \Theta=\arg\max_{\Theta\in\tP}\tL_\tY(\Theta)$, $\tL_\tY(\hat \Theta)-\tL_{\tY}(\trueT)\geq 0$, which gives
-\[
-C_2U_\alpha \left(r_{\max}^{K-1}\sum_k d_k\right)^{1/2}  \FnormSize{}{\hat \Theta-\trueT}-{L_\alpha\over 2}\FnormSize{}{\hat \Theta-\trueT}^2\geq 0.
-\]
-Henceforth,
-\[
-{1\over \sqrt{\prod_k d_k}} \FnormSize{}{\hat \Theta-\trueT}\leq {2C_2U_\alpha \sqrt{r_{\max}^{K-1}\sum_k d_k}\over L_\alpha \sqrt{\prod_k d_k}}={2C_2U_\alpha r_{\max}^{(K-1)/2}\over L_\alpha} \sqrt{ \sum_k d_k \over \prod_k d_k}.
-\]
-This completes the proof.
-\end{proof}
-
-\begin{proof}[Proof of Corollary~\ref{cor:prediction}]
-The result follows immediately from Theorem~\ref{thm:rate} and Lemma~\ref{lem:KL}.
-\end{proof}
-
-\begin{proof}[Proof of Theorem~\ref{thm:minimax}]
-
-Let $d_{\text{total}}=\prod_{k\in[K]}d_k$, and $\gamma\in[0,1]$ be a constant to be specified later.
-Our strategy is to construct a finite set of tensors $\tX=\{\Theta_i\colon i=1,\ldots \}\subset \tP$ satisfying the properties of (i)-(iv) in Lemma~\ref{lem:construction}.
-By Lemma~\ref{lem:construction}, such a subset of tensors exist.
-For any tensor  $\Theta\in\tX$, let $\mathbb{P}_{\Theta}$ denote the distribution of $\tY|\Theta$, where $\tY$ is the ordinal tensor.
-In particular, $\mathbb{P}_{\mathbf{0}}$ is the distribution of $\tY$ induced by the zero parameter tensor $\mathbf{0}$, i.e., the distribution of $\tY$ conditional on the parameter tensor $\Theta=\mathbf{0}$.
-Based on the Remark for Lemma~\ref{lem:KL}, we have
-\begin{equation}\label{eq:KLbound1}
-\mathrm{KL}(\mathbb{P}_{\Theta}|| \mathbb{P}_{\mathbf{0}})\leq C \FnormSize{}{\Theta}^2,
-\end{equation}
-where $C={(4L-6) \dot{f}^2(0)\over  A_\alpha}>0$ is a constant independent of the tensor dimension and rank.
-Combining the inequality~\eqref{eq:KLbound1} with property (iii) of $\tX$, we have
-\begin{equation}\label{eq:KLbound}
-\text{KL}(\mathbb{P}_{\Theta}||\mathbb{P}_{\mathbf{0}})\leq \gamma^2 r_{\max} d_{\max}.
-\end{equation}
-From~\eqref{eq:KLbound} and the property (i), we deduce that the condition
-\begin{equation}\label{eq:totalKL}
-{1\over \text{Card}(\tX)-1}\sum_{\Theta \in\tX}\text{KL}(\mathbb{P}_{\Theta}, \mathbb{P}_{\mathbf{0}})\leq \varepsilon \log_2\left\{\text{Card}(\tX)-1 \right\}
-\end{equation}
-holds for any $ \varepsilon \geq 0$ when $\gamma\in[0,1]$ is chosen to be sufficiently small depending on $\varepsilon$, e.g., $\gamma \leq \sqrt{\varepsilon\log2\over8}$.
-By applying Lemma~\ref{lem:Tsybakov} to~\eqref{eq:totalKL}, and in view of the property (iv), we obtain that
-\begin{equation}\label{eq:final}
-\inf_{\hat \Theta}\sup_{\trueT\in \tX}\mathbb{P}\left(\FnormSize{}{\hat \Theta- \trueT}\geq  {\gamma\over 8} \min\left\{ \alpha\sqrt{d_{\text{total}}}, C^{-1/2}\sqrt{ r_{\max}d_{\max}}\right\} \right)\geq {1\over 2}\left(1-2\varepsilon-\sqrt{16 \varepsilon \over r_{\max}d_{\max}\log2}\right).
-\end{equation}
-Note that $\text{Loss}(\hat \Theta, \trueT)=\FnormSize{}{\hat \Theta- \trueT}^2/d_{\text{total}}$ and $\tX\subset \tP$.
-By taking $\varepsilon=1/10$ and $\gamma=1/11$, we conclude from~\eqref{eq:final} that
-\begin{equation}\label{eq:prob}
-\inf_{\hat \Theta}\sup_{\trueT\in \tP}\mathbb{P}\left(\text{Loss}(\hat \Theta, \trueT)\geq c\min\left \{ \alpha^2,  {C^{-1}r_{\max}d_{\max}\over d_{\text{total}}}\right \}\right)\geq {1\over 2}\left({4\over 5}- \sqrt{1.6\over r_{\max}d_{\max}\log2} \right)\geq {1\over 8},
-\end{equation}
-where $c = {1 \over 88^2}$ and the last inequality comes from the condition for $d_{\text{max}}$.
-This completes the proof.
-\end{proof}
-
-\subsection{Sample complexity for tensor completion}
-
-\begin{proof}[Proof of Theorem~\ref{thm:completion}]
-
-For notational convenience, we use $\MFnormSize{}{\Theta}=\sum_{\omega\in\Omega}\Theta^2_\omega$ to denote the sum of squared entries over the observed set $\Omega$, for a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$.
-
-Following a similar argument as in the proof of Theorem~\ref{thm:rate}, we have
-\begin{equation}\label{eq:Taylor2}
-\logl(\Theta)=\logl(\trueT)+\langle\Vec(\plogl),\ \Vec(\Theta-\trueT)\rangle+{1\over 2}\Vec(\Theta-\trueT)^T\pplogl(\check\Theta)\Vec(\Theta-\trueT),
-\end{equation}
-where
-\begin{enumerate}
-\item $\plogl$ is a $d_1\times\cdots\times d_K$ tensor with $|\Omega|$ nonzero entries, and each entry is upper bounded by $U_\alpha>0$.
-\item $\pplogl$ is a diagonal matrix of size $d_{\text{total}}$-by-$d_{\text{total}}$ with $|\Omega|$ nonzero entries, and each entry is upper bounded by $-L_{\alpha}<0$.
-\end{enumerate}
-
-Similar to~\eqref{eq:linear} and~\eqref{eq:quadratic}, we have
-\begin{equation}\label{eq:linear2}
-|\langle\Vec(\plogl),\ \Vec(\Theta-\trueT)\rangle|\leq C_2U_\alpha \sqrt{r_{\max}^{K-1}\sum_k d_k}\MFnormSize{}{\Theta-\trueT}
-\end{equation}
-and
-\begin{equation}\label{eq:quadratic2}
-\Vec(\Theta-\trueT)^T\fpplogl(\check\Theta)\Vec(\Theta-\trueT)\leq -L_\alpha \MFnormSize{}{\Theta-\trueT}^2.
-\end{equation}
-
-Combining~\eqref{eq:Taylor2}-\eqref{eq:quadratic2} with the fact that $\logl(\hat \Theta)\geq \logl(\trueT)$, we have
-\begin{equation}\label{eq:sample}
-\MFnormSize{}{\hat \Theta-\trueT}\leq {2C_2U_\alpha  r_{\max}^{(K-1)/2} \over L_\alpha} \sqrt{\sum_kd_k},
-\end{equation}
-with probability at least $1-\exp(-C_1 \sum_k d_k)$.
-Lastly, we invoke the result regarding the closeness of $\Theta$ to its sampled version $\Theta_{\Omega}$, under the entrywise bound condition.
-Note that $\mnormSize{}{\hat\Theta-\trueT}\leq 2\alpha$ and $\text{rank}(\hat \Theta-\trueT)\leq 2\mr$.
-By Lemma~\ref{lem:Mnormbound}, $\anormSize{}{\hat \Theta-\trueT}\leq 2^{(3K-1)/2} \alpha \left({\prod r_k \over r_{\max}}\right)^{3/2}$.
-Therefore, the condition in Lemma~\ref{lem:convexity} holds with $\beta=2^{(3K-1)/2}\alpha \left({\prod r_k \over r_{\max}}\right)^{3/2}$.
-Applying Lemma~\ref{lem:convexity} to~\eqref{eq:sample} gives
-\begin{align}
- \PiFnormSize{}{\hat \Theta-\trueT}^2&\leq {1\over m}\MFnormSize{}{\hat \Theta-\trueT}^2+c\beta\sqrt{\sum_k d_k\over |\Omega|}\\
- &\leq {C_2  r^{K-1}_{\max}} {\sum_k d_k \over |\Omega|}+C_1 \alpha r_{\max}^{3(K-1)/2}\sqrt{\sum_kd_k\over |\Omega|},
-\end{align}
-with probability at least $1-\exp(-{\sum_kd_k\over \sum_k \log d_k})$ over the sampled set $\Omega$. Here $C_1, C_2>0$ are two constants independent of the tensor dimension and rank. Therefore,
-\[
- \PiFnormSize{}{\hat \Theta-\trueT}^2\to 0,\quad \text{as}\quad {|\Omega|\over \sum_kd_k}\to \infty,
-\]
-provided that $r_{\max}=O(1)$.
-\end{proof}
-
-\subsection{Convexity of the log-likelihood function}\label{sec:proofconvexity}
-\begin{thm}\label{thm:convexity}
-Define the function
-\begin{align}\label{eq:function}
- \logl(\Theta, \mb)&=\sum_{\omega\in\Omega}\sum_{\ell\in[L]} \big(\mathds{1}\{y_\omega=\ell\}\log \left[f(b_\ell-\theta_\omega)-  f(b_{\ell-1}-\theta_\omega)\right]\big),
- \end{align}
-where $f(\cdot)$ satisfies Assumption~\ref{ass:link}. Then, $\logl(\Theta, \mb)$ is concave in $(\Theta,\mb)$.
-\end{thm}
-
-\begin{proof}
-Define $d_{\text{total}}=\prod_k d_k$.
-By abuse of notation, we use $(\Theta,\mb)$ to denote the length-$(d_{\text{total}}+L-1)$-vector collecting all parameters together.
-Let us denote a bivariate function
-\begin{align}
-\lambda\colon &\mathbb{R}^2\mapsto \mathbb{R}\\
-(u,v)&\mapsto\lambda(u,v) = \log \big[f(u)-  f(v)\big].
-\end{align}
-It suffices to show that $\lambda(u,v)$ is concave in $(u,v)$ where $u>v$.
-
-Suppose that the claim holds (which we will prove in the next paragraph). Based on~\eqref{eq:function}, $u,v$ are both linear functions of $(\Theta,\mb)$:
-\[
-u = \ma_1^T(\Theta,\mb), \quad v = \ma_2^T(\Theta,\mb), \quad \text{ for some}\ \ma_1,\ma_2\in \mathbb{R}^{d_{\text{total}}+L-1}.
-\]
-Then, $\lambda(u,v) = \lambda(\ma_1^T(\Theta,\mb),\ \ma_2^T(\Theta,\mb))$ is concave in $(\Theta,\mb)$ by the definition of concavity.
-Therefore, we can conclude that $\logl(\Theta, \mb) $ is concave in $(\Theta,\mb)$ because $\logl(\Theta, \mb)$ is the sum of $\lambda(u,v)$.
-
-Now, we prove the concavity of $\lambda(u,v)$. Note that
-\begin{equation}
-\lambda(u,v) = \log\big[f(u)-f(v)\big]=\log\big[\int\mathds{1}_{[u,v]}(x)f'(x)dx\big],
-\end{equation}
-where $\mathds{1}_{[u,v]}$ is an indicator function that equals 1 in the interval $[u,v]$, and 0 elsewhere.
-Furthermore, $\mathds{1}_{[u,v]}(x)$ is log-concave in $(u,v,x)$, and by Assumption~\ref{ass:link}, $f'(x)$ is log-concave in $x$.
-It follows that $\mathds{1}_{[u,v]}(x)f'(x)$
-is a log-concave in $(u,v,x)$. By Lemma~\ref{lem:lossconvexity}, we conclude that $\lambda(u,v)$ is concave in $(u,v)$ where $u>v$.
-\end{proof}
-
-\begin{lem}[Corollary 3.5 in~\citet{brascamp2002extensions}]\label{lem:lossconvexity}
-Let $F(x,y)\colon \mathbb{R}^{m+n}\rightarrow \mathbb{R}$ be an integrable function where $x\in \mathbb{R}^{m},y\in \mathbb{R}^n$. Let
-\[
-G(x) = \int_{\mathbb{R}^n}F(x,y)dy.
-\]
-If $F(x,y)$ is log concave in $(x,y)$, then $G(x)$ is log concave in $x$.
-\end{lem}
-
-\subsection{Auxiliary lemmas}\label{sec:lemma}
-This section collects lemmas that are useful for the proofs of the main theorems.
-
-\begin{defn}[Atomic M-norm~\citep{ghadermarzy2019near}]
-Define $T_{\pm}=\{\tT\in\{\pm 1 \}^{d_1\times \cdots \times d_K}\colon \text{rank}(\tT)=1\}$. The atomic M-norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
-\begin{align}
-\anormSize{}{\Theta}&=\inf\{t>0\colon \Theta\in t\text{conv}(T_{\pm})\}\\
-&=\inf\left\{\sum_{\tX\in T_{\pm}}c_{\tX} \colon\ \Theta=\sum_{\tX\in T_{\pm}} c_{\tX}\tX, \ c_{\tX}>0\right\}.
-\end{align}
-\end{defn}
-
-\begin{defn}[Spectral norm~\citep{lim2005singular}]
-The spectral norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
-\[
-\snormSize{}{\Theta}=\sup\left\{\langle \Theta, \mx_1\otimes \cdots \otimes \mx_K\rangle \colon \vnormSize{}{\mx_k}=1,\ \mx_k\in\mathbb{R}^{d_k},\ \text{for all}\ k\in[K]\right\}.
-\]
-\end{defn}
-
-\begin{defn}[Nuclear norm~\citep{friedland2018nuclear}]
-The nuclear norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
-\[
-\nnormSize{}{\Theta}=\inf
-\left\{
-\sum_{i\in[r]}|\lambda_i|\colon \Theta=\sum_{i=1}^r \lambda_i\mx^{(i)}_1\otimes \cdots \otimes \mx^{(i)}_K,\ \vnormSize{}{\mx^{(i)}_k}=1,\ \mx^{(i)}_k\in\mathbb{R}^{d_k},\ \text{for all}\ k\in[K],\ i\in[r]
-\right\},
-\]
-where the infimum is taken over all $r\in\mathbb{N}$ and $\vnormSize{}{\mx^{(i)}_k}=1$ for all $i\in[r]$ and $k\in[K]$.
-\end{defn}
-
-\begin{lem}[M-norm and infinity norm~\citep{ghadermarzy2019near}]\label{lem:Mnormbound}
-Let $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ be an order-$K$, rank-$(r_1,\ldots,r_K)$ tensor. Then
-\[
-\mnormSize{}{\Theta}\leq \anormSize{}{\Theta}\leq \left(\prod_k r_k \over r_{\max}\right)^{3\over 2} \mnormSize{}{\Theta}.
-\]
-\end{lem}
-
-\begin{lem}[Nuclear norm and F-norm] \label{lem:nuclear}
-Let $\tA\in\mathbb{R}^{d_1\times\cdots\times d_K}$ be an order-$K$ tensor with Tucker $\text{rank}(\tA)=(r_1,\ldots,r_K)$. Then
-\[
-\nnormSize{}{\tA} \leq \sqrt{\prod_k r_k\over \max_k r_k} \FnormSize{}{\tA},
-\]
-where $\nnormSize{}{\cdot}$ denotes the nuclear norm of the tensor.
-\end{lem}
-
-\begin{proof}
-Without loss of generality, suppose $r_1=\min_k r_k$.
-Let $\tA_{(k)}$ denote the mode-$k$ matricization of $\tA$ for all $k\in[K]$.
-By \citet[Corollary 4.11]{wang2017operator}, and the invariance relationship between a tensor and its Tucker core~\citep[Section 6]{jiang2017tensor}, we have
-\begin{equation}\label{eq:norminequality}
-\nnormSize{}{\tA} \leq \sqrt{\prod_{k\geq 2} r_k \over \max_{k\geq 2} r_k} \nnormSize{}{\tA_{(1)}},
-\end{equation}
-where $\tA_{(1)}$ is a $d_1$-by-$\prod_{k\geq 2}d_k$ matrix with matrix rank $r_1$.
-Furthermore, the relationship between the matrix norms implies that $\nnormSize{}{\tA_{(1)}}\leq \sqrt{r_1}\FnormSize{}{\tA_{(1)}}=\sqrt{r_1}\FnormSize{}{\tA}$.
-Combining this fact with the inequality~\eqref{eq:norminequality} yields the final claim.
-\end{proof}
-
-\begin{lem} \label{lem:inq}
-Let $\tA, \; \tB$ be two order-$K$ tensors of the same dimension. Then
-\[
-|\langle \tA,\tB\rangle| \leq \snormSize{}{\tA}   \nnormSize{}{\tB}.
-\]
-\end{lem}
-
-\begin{proof}
-By~\citet[Proposition 3.1]{friedland2018nuclear}, there exists a nuclear norm decomposition of $\tB$, such that
-\[
-\tB=\sum_{r} \lambda_r \ma^{(1)}_r\otimes \cdots\otimes \ma^{(K)}_r,\quad \ma_r^{(k)}\in\mathbf{S}^{d_k-1}(\mathbb{R}),\quad \text{for all }k\in[K],
-\]
-and $\nnormSize{}{\tB}=\sum_{r}|\lambda_r|$. Henceforth we have
-\begin{align*}
-|\langle \tA,\tB\rangle|&=| \langle \tA, \sum_{r} \lambda_r \ma^{(1)}_r\otimes \cdots\otimes \ma^{(K)}_r \rangle|\leq \sum_r |\lambda_r| |\langle \tA, \ma^{(1)}_r \otimes \cdots\otimes \ma^{(K)}_r \rangle|\\
-&\leq \sum_{r}|\lambda_r| \snormSize{}{\tA}= \snormSize{}{\tA}\nnormSize{}{\tB},
-\end{align*}
-which completes the proof.
-\end{proof}
-
-The following lemma provides the bound on the spectral norm of random tensors. The result was firstly presented in~\citet{nguyen2015tensor}, and we adopt the version from~\citet{tomioka2014spectral}.
-\begin{lem}[Spectral norm of random tensors~\citep{tomioka2014spectral}]\label{lem:tensor}
-Suppose that $\tS=\entry{s_{\omega}}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is an order-$K$ tensor whose entries are independent random variables that satisfy
-\[
-\mathbb{E}(s_{\omega})=0,\quad \text{and} \quad\mathbb{E}(e^{ts_{\omega}})\leq e^{t^2L^2/2}.
-\]
-Then the spectral norm $\snormSize{}{\tS}$ satisfies that,
-\[
-\snormSize{}{\tS}\leq \sqrt{{8L^2} \log (12K) \sum_k d_k +\log (2/\delta)},
-\]
-with probability at least $1-\delta$.
-\end{lem}
-
-\begin{lem} \label{lem:noisytensor}
-Suppose that $\tS=\entry{s_{\omega}}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is an order-$K$ tensor whose entries are independent random variables that satisfy
-\[
-\mathbb{E}(s_{\omega})=0,\quad \text{and}\quad |s_{\omega}|\leq U.
-\]
-Then we have
-\[
-\mathbb{P}\left(\snormSize{}{\tS}\geq C_2 U\sqrt{\sum_k d_k} \right)\leq \exp\left(-C_1  \log K \sum_k d_k\right)
-\]
-where $C_1>0$ is an absolute constant, and $C_2>0$ is a constant that depends only on $K$.
-\end{lem}
-
-\begin{proof}  Note that the random variable $U^{-1}s_{\omega}$ is zero-mean and supported on $[-1,1]$. Therefore, $U^{-1}s_{\omega}$ is sub-Gaussian with parameter ${1-(-1)\over 2}=1$, i.e.
-\[
-\mathbb{E}(U^{-1}s_{\omega})=0,\quad \text{and}\quad \mathbb{E}(e^{tU^{-1}s_{\omega}})\leq e^{t^2/2}.
-\]
-It follows from Lemma~\ref{lem:tensor} that, with probability at least $1-\delta$,
-\[
-\snormSize{}{U^{-1}\tS}\leq \sqrt{\left(c_0\log K+c_1\right) \sum_k d_k +\log (2/\delta)},
-\]
-where $c_0, c_1>0$ are two absolute constants. Taking $\delta=\exp (-C_1\log K \sum_k d_k)$ yields the final claim, where $C_2=c_0\log K+c_1+1>0$ is another constant.
-\end{proof}
-
-\begin{lem}\label{lem:KLentry} Let $X,\; Y$ be two discrete random variables taking values on $L$ possible categories, with point mass probabilities $\{p_\ell\}_{\ell\in[L]}$ and $\{q_\ell\}_{\ell\in[L]}$, respectively.
-Suppose $p_\ell$, $q_\ell>0$ for all $\ell\in[L]$.
-Then, the Kullback-Leibler (KL) divergence satisfies that
-\[
-\mathrm{KL}(X||Y)\stackrel{\text{def}}{=}-\sum_{\ell\in[L]}\mathbb{P}_X(\ell)\log\left\{{\mathbb{P}_Y(\ell)\over \mathbb{P}_X(\ell)}\right\} \leq \sum_{\ell \in [L]}{(p_\ell-q_\ell)^2 \over q_\ell}.
-\]
-\end{lem}
-\begin{proof} Using the fact $\log x\leq x-1$ for $x>0$, we have that
-\begin{align}\label{eq:KL}
-\text{KL}(X||Y)&=\sum_{\ell\in[L]}p_\ell\log{p_\ell\over q_\ell}\\
-&\leq \sum_{\ell\in[L]} {p_\ell\over q_\ell}(p_\ell-q_\ell)\\
-&=\sum_{\ell\in[L]} \left({p_\ell\over q_\ell }- 1\right)(p_\ell-q_\ell)+ \sum_{\ell\in[L]} (p_\ell-q_\ell).
-\end{align}
-Note that $\sum_{\ell\in[L]}(p_\ell-q_\ell)=0$. Therefore,
-\[
-\text{KL}(X||Y)\leq \sum_{\ell\in[L]}\left( {p_\ell\over q_\ell}-1\right)\left(p_\ell-q_\ell\right)=\sum_{\ell\in[L]}{(p_\ell-q_\ell)^2\over q_\ell}.
-\]
-\end{proof}
-
-\begin{lem}[KL divergence and F-norm]~\label{lem:KL}
-Let $\tY\in[L]^{d_1\times \cdots \times d_K}$ be an ordinal tensor generated from the model~\eqref{eq:model} with the link function $f$ and parameter tensor $\Theta$.
-Let $\mathbb{P}_{\Theta}$ denote the joint categorical distribution of $\tY|\Theta$ induced by the parameter tensor $\Theta$, where $\mnormSize{}{\Theta}\leq \alpha$.
-Define
-\begin{equation}\label{eq:ass}
-A_\alpha=\min_{\ell\in[L], |\theta|\leq \alpha}\left[f(b_\ell-\theta)-f(b_{\ell-1}-\theta)\right].
-\end{equation}
-Then, for any two tensors $\Theta,\; \Theta^*$ in the parameter spaces, we have
-\[
-\mathrm{KL}(\mathbb{P}_{\Theta}|| \mathbb{P}_{\Theta^*})\leq {2(2L-3)\over A_\alpha} \dot{f}^2(0)\FnormSize{}{\Theta-\Theta^*}^2.
-\]
-\end{lem}
-\begin{proof} Suppose that the distribution over the ordinal tensor $\tY=\entry{y_\omega}$ is induced by $\Theta=\entry{\theta_\omega}$. Then, based on the generative model~\eqref{eq:model},
-\[
-\mathbb{P}(y_\omega=\ell | \theta_\omega)=f(b_{\ell}-\theta_\omega)-f(b_{\ell-1}-\theta_\omega),
-\]
-for all $\ell\in[L]$ and $\omega\in[d_1]\times \cdots \times [d_K]$.
-For notational convenience, we suppress the subscribe in $\theta_\omega$ and simply write $\theta$ (and respectively, $\theta^*$).
-Based on Lemma~\ref{lem:KLentry} and Taylor expansion,
-\begin{align}
-\text{KL}(\theta|| \theta^*) & \leq \sum_{\ell\in[L]} {\left[f(b_{\ell}-\theta)-f(b_{\ell-1}-\theta)-f(b_{\ell}-\theta^*)+f(b_{\ell-1}-\theta^*)\right]^2\over f(b_{\ell}-\theta^*)-f(b_{\ell-1}-\theta^*)}\\
- &\leq \sum_{\ell=2}^{L-1} {\left[\dot{f}(b_\ell-\eta_\ell)-\dot{f}(b_{\ell-1}-\eta_{\ell-1})\right]^2 \over f(b_\ell-\theta^*)-f(b_{\ell-1}-\theta^*)} \left(\theta-\theta^*\right)^2+{\dot{f}^2(b_1-\eta_1) \over f(b_1-\theta^*)} (\theta-\theta^*)^2\\
- &\quad \quad\quad \quad  +{\dot{f}^2(b_{L-1}-\eta_{L-1})\over 1-f(b_{L-1}-\theta^*)}(\theta-\theta^*)^2,
-\end{align}
-where $\eta_\ell$ and $\eta_{\ell-1}$ fall between $\theta$ and $\theta^*$. Therefore,
-\begin{equation}\label{eq:entrywise}
-\text{KL}(\theta|| \theta^*) \leq \left({4(L-2)\over A_\alpha}+ {2\over A_\alpha}\right)\dot{f}^2(0)(\theta-\theta^*)^2={2(2L-3)\over A_\alpha} \dot{f}^2(0)(\theta-\theta^*)^2,
-\end{equation}
-where we have used Taylor expansion, the bound~\eqref{eq:ass}, and the fact that $\dot{f}(\cdot)$ peaks at zero for an unimodal and symmetric function.
-Now summing~\eqref{eq:entrywise} over the index set $\omega\in[d_1]\times \cdots \times [d_K]$ gives
-\[
-\text{KL}(\mathbb{P}_{\Theta}|| \mathbb{P}_{\Theta^*}) =\sum_{\omega\in[d_1]\times \cdots \times[d_K]} \text{KL}(\theta_\omega || \theta^*_\omega) \leq {2(2L-3)\over A_\alpha} \dot{f}^2(0)\FnormSize{}{\Theta-\Theta^*}^2.
-\]
-\end{proof}
-
-\begin{rmk} In particular, let $\mathbb{P}_{\bf{0}}$ denote the distribution of $\tY|\bf{0}$ induced by the zero parameter tensor. Then we have
-\[
-\text{KL}(\mathbb{P}_{\Theta}||\mathbb{P}_{\bf{0}})\leq {2(2L-3)\over A_\alpha}  \dot{f}^2(0)\FnormSize{}{\Theta}^2.
-\]
-\end{rmk}
-
-\begin{lem}\label{lem:construction}
-Assume the same setup as in Theorem~\ref{thm:minimax}.
-Without loss of generality, suppose $d_1=\max_kd_k$.
-Define $R=\max_k r_k$ and $d_{\text{total}}=\prod_{k\in[K]} d_k$.
-For any constant $0\leq \gamma \leq 1$, there exist a finite set of tensors $\tX=\{\Theta_i: i=1,\ldots\}\subset \tP$ satisfying the following four properties:
-\begin{enumerate}
-\item $\text{Card}(\tX)\geq 2^{Rd_1/8}+1$, where $\text{Card}$ denotes the cardinality;
-\item $\tX$ contains the zero tensor $\mathbf{0}\in\mathbb{R}^{d_1\times \cdots\times d_K}$;
-\item $\mnormSize{}{\Theta}\leq \gamma \min\left\{ \alpha ,\ C^{-1/2}\sqrt{Rd_1\over d_{\text{total}}} \right\} $ for any element $\Theta\in\tX$;
-\item $\FnormSize{}{\Theta_i-\Theta_j}\geq {\gamma\over 4} \min\left\{ \alpha\sqrt{d_{\text{total}}},\ C^{-1/2}\sqrt{Rd_1}\right\}$ for any two distinct elements $\Theta_i,\; \Theta_j\in\tX$,
-\end{enumerate}
-Here $C=C(\alpha,L,f,\mb)={(4L-6)\dot{f}^2(0)\over A_\alpha }>0$ is a constant independent of the tensor dimension and rank.
-\end{lem}
-
-\begin{proof}
-Given a constant $0\leq \gamma \leq 1$, we define a set of matrices:
-\[
-\tC=\left\{\mM=(m_{ij})\in\mathbb{R}^{d_1\times R}: a_{ij}\in \left\{ 0,\gamma \min\left\{ \alpha , C^{-1/2}\sqrt{Rd_1\over d_{\text{total}}} \right\}\right\} ,\  \forall (i,j)\in[d_1]\times[R]\right\}.
-\]
-We then consider the associated set of block tensors:
-\begin{align}
-\tB=\tB(\tC)=\{\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}\colon& \Theta=\mA\otimes \mathbf{1}_{d_3}\otimes \cdots \otimes \mathbf{1}_{d_K}, \ \text{where}\ \mA=(\mM|\cdots|\mM|\mO) \in\mathbb{R}^{d_1\times d_2},\ \mM\in\tC\},
-\end{align}
-where $\mathbf{1}_d$ denotes a length-$d$ vector with all entries 1, $\mO$ denotes the $d_1\times (d_2-R\lfloor d_2/R \rfloor)$ zero matrix, and $\lfloor d_2/ R \rfloor$ is the integer part of $d_2/R$.
-In other words, the subtensor $\Theta(\mI, \mI,i_3, \ldots,i_K)\in\mathbb{R}^{d_1\times d_2}$ are the same for all fixed $(i_3,\ldots,i_K)\in[d_3]\times \cdots \times [d_K]$, and furthermore, each subtensor $\Theta(\mI,\mI, i_3,\ldots,i_K)$ itself is filled by copying the matrix $\mM\in\mathbb{R}^{d_1\times R}$ as many times as would fit.
-
-By construction, any element of $\tB$, as well as the difference of any two elements of $\tB$, has Tucker rank at most $\max_k r_k\leq R$, and the entries of any tensor in $\tB$ take values in $[0,\alpha]$.
-Thus, $\tB\subset\tP$.
-By Lemma~\ref{lem:VGbound}, there exists a subset $\tX\subset \tB$ with cardinality $\text{Card}(\tX)\geq 2^{Rd_1/8}+1$ containing the zero $d_1\times \cdots \times d_K$ tensor, such that, for any two distinct elements $\Theta_i$ and $\Theta_j$ in $\tX$,
-\[
-\FnormSize{}{\Theta_i-\Theta_j}^2 \geq {Rd_1\over 8} \gamma^2\min\left\{ \alpha^2, {C^{-1}Rd_1 \over d_{\text{total}}}\right\} \Big\lfloor {d_2\over R}\Big \rfloor \prod_{k\geq 3}d_k\geq {\gamma^2\min\left\{ \alpha^2 d_{\text{total}}, C^{-1}Rd_1\right\}  \over 16}.
-\]
-In addition, each entry of $\Theta\in\tX$ is bounded by $\gamma \min\left\{ \alpha , C^{-1/2}\sqrt{Rd_1\over d_{\text{total}}}\right\} $. Therefore the Properties (i) to (iv) are satisfied.
-\end{proof}
-
-\begin{lem}[Varshamov-Gilbert bound]\label{lem:VGbound}
-Let $\Omega=\{(w_1,\ldots,w_m)\colon w_i\in\{0,1\}\}$. Suppose $m>8$. Then there exists a subset $\{w^{(0)},\ldots,w^{(M)}\}$ of $\Omega$ such that $w^{(0)}=(0,\ldots,0)$ and
-\[
-\zeronormSize{}{w^{(j)}-w^{(k)}}\geq {m\over 8},\quad \text{for } \ 0\leq j<k\leq M,
-\]
-where $\zeronormSize{}{\cdot}$ denotes the Hamming distance, and $M\geq 2^{m/8}$.
-\end{lem}
-
-\begin{lem}[Theorem 2.5 in~\citet{tsybakov2008introduction}]\label{lem:Tsybakov}
-Assume that a set $\tX$ contains element $\Theta_0, \Theta_1, \ldots,\Theta_M$ ($M\geq 2$) such that
-\begin{itemize}
-\item $d(\Theta_j,\ \Theta_k)\geq 2s>0$, $\forall 0\leq j\leq k\leq M$;
-\item $\mathbb{P}_0$ is absolutely continuous with respect to $\mathbb{P}_j$, $\forall j=1,\ldots,M$, and
-\[
-{1\over M}\sum_{j=1}^M \mathrm{KL}(\mathbb{P}_j||\mathbb{P}_0)\leq \alpha \log M
-\]
-where $d\colon \tX\times \tX\mapsto [0,+\infty]$ is a semi-distance function, $0<\alpha<{1/8}$ and $\mathbb{P}_j=\mathbb{P}_{\Theta_j}$, $j=0,1\ldots,M$.
-\end{itemize}
-Then
-\[
-\inf_{\hat \Theta}\sup_{\Theta\in\tX} \mathbb{P}_{\Theta}(d(\hat \Theta, \Theta)\geq s)\geq {\sqrt{M}\over 1+\sqrt{M}}\left(1-2\alpha-\sqrt{2\alpha\over \log M} \right)>0.
-\]
-\end{lem}
-
-\begin{lem}[Lemma 28 in~\citet{ghadermarzy2019near}]\label{lem:convexity}
-Define $\mathbb{B}_{M}(\beta)=\{\Theta\in \mathbb{R}^{d_1\times \cdots \times d_K}\colon \anormSize{}{\Theta}\leq \beta \}$.
-Let $\Omega\subset[d_1]\times\cdots \times [d_K]$ be a random set with $m=|\Omega|$, and assume that each entry in $\Omega$ is drawn with replacement from $[d_1]\times\cdots\times[d_K]$ using probability $\Pi$.
-Define
-\[
-\PiFnormSize{}{\Theta}^2={1\over m}\mathbb{E}_{\Omega\in\Pi}\MFnormSize{}{\Theta}^2.
-\]
-Then, there exists a universal constant $c>0$, such that, with probability at least $1-\exp\left(-{\sum_kd_k \over \sum_k \log d_k} \right)$ over the sampled set $\Omega$,
-\[
-{1\over m}\MFnormSize{}{\Theta}^2 \geq \PiFnormSize{}{\Theta}^2-c\beta\sqrt{\sum_k d_k\over m}
-\]
-holds uniformly for all $\Theta\in\mathbb{B}_M(\beta)$.
-\end{lem}
-
-\section{Extension of Theorem~\ref{thm:rate} to unknown cut-off points}\label{sec:extention}
-We now extend Theorem~\ref{thm:rate} to the case of unknown cut-off points $\mb$. Assume that the true parameters $(\trueT, \trueb)\in \tP\times \tB$, where the feasible sets are defined as
-\begin{align}
-\tP&=\{\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}\colon \text{rank}(\tP)\leq \mr,\ \langle \Theta, \tJ \rangle=0,\ \mnormSize{}{\Theta}\leq \alpha\},\\
- \tB&=\{\mb\in\mathbb{R}^{L-1}\colon \mnormSize{}{\mb}\leq \beta,\ \min_{\ell}(b_\ell-b_{\ell-1}) \geq \Delta\},
-\end{align}
-with positive constants $\alpha,\beta,\Delta>0$ and a given rank $\mr\in\mathbb{N}^K_{+}$.
-Here, $\tJ=\entry{1}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ denotes a tensor of all ones.
-The constraint $\langle \Theta, \tJ \rangle=0$ is imposed to ensure the identifiability of $\Theta$ and $\mb$.
-We propose the constrained M-estimator
-\begin{equation}\label{eq:joint}
-(\hat \Theta,\hat \mb)=\argmax_{(\Theta,\mb)\in\tP\times \tB}\flogl(\Theta,\mb).
-\end{equation}
-The estimation accuracy is assessed using the mean squared error (MSE):
-\[
-\text{MSE}\left( \hat \Theta, \trueT \right)={1\over \prod_k d_k}\FnormSize{}{\hat \Theta-\trueT}^2,\quad \text{MSE}\left( \hat \mb, \trueb \right)={1\over L-1}\FnormSize{}{\hat \mb-\trueb}^2.
-\]
-
-To facilitate the examination of MSE, we define an order-$(K+1)$ tensor, $\tZ=\entry{z_{\omega,\ell}}\in\mathbb{R}^{d_1\times \cdots \times d_K\times (L-1)}$, by stacking the parameters $\Theta=\entry{\theta_\omega}$ and $\mb=\entry{b_\ell}$ together.
-Specifically, let $z_{\omega,\ell}=-\theta_\omega+b_\ell$ for all $\omega\in[d_1]\times \cdots \times [d_K]$ and $\ell\in[L-1]$; that is,
-\[
-\tZ=-\Theta\otimes \mathbf{1}+\tJ\otimes \mb,
-\]
-where $\mathbf{1}$ denotes a length-$(L-1)$ vector of all ones.
-Under the identifiability constraint $\langle \Theta, \tJ \rangle=0$, there is an one-to-one mapping between $\tZ$ and $(\Theta, \mb)$, with $\text{rank}(\tZ)\leq (\text{rank}(\Theta)+1,\ 2)^T$.
-Furthermore, 
-\begin{equation}\label{eq:decomp}
-\FnormSize{}{\hat \tZ-\trueZ}^2=\FnormSize{}{\hat \Theta-\trueT}^2 (L-1)+\FnormSize{}{\hat \mb-\trueb}^2 \left(\prod_kd_k\right),
-\end{equation}
-where $\trueZ=-\trueT\otimes \mathbf{1}+\tJ\otimes \trueb$ and $\hat \tZ=-\hat \Theta\otimes \mathbf{1}+\tJ\otimes \hat \mb$.
-
-We make the following assumptions about the link function.
-\begin{assumption}\label{ass:joint}
-The link function $f\colon \mathbb{R}\mapsto [0,1]$ satisfies the following properties:
-\begin{enumerate}
-\item $f(z)$ is twice-differentiable and strictly increasing in $z$.
-\item $\dot{f}(z)$ is strictly log-concave and symmetric with respect to $z=0$.
-\end{enumerate}
-\end{assumption}
-
-We define the following constants that will be used in the theory:
-\begin{align}\label{eq:constantZ}
-C_{\alpha,\beta,\Delta}&=\max_{|z|\leq \alpha+\beta}\max_{\substack{z'\leq z-\Delta\\z''\geq z+\Delta}}\max\left\{ {\dot{f}(z) \over f(z)-f(z')},\ {\dot{f}(z) \over f(z'')-f(z)}\right\},\\
-D_{\alpha,\beta,\Delta}&=\min_{|z|\leq \alpha+\beta}\min_{\substack{z'\leq z-\Delta\\z''\geq z+\Delta}}\min\left\{- {\partial\over \partial z}\left({\dot{f}(z) \over f(z)-f(z')}\right),\  {\partial\over \partial z}\left({\dot{f}(z) \over f(z'')-f(z)}\right) \right\},\\
-A_{\alpha,\beta,\Delta}&=\min_{|z|\leq \alpha+\beta}\min_{z'\leq z-\Delta} \left(f(z)-f(z')\right).
-\end{align}
-
-\begin{rmk}
-The condition $\Delta=\min_{\ell}(b_{\ell}-b_{\ell-1})>0$ on the feasible set $\tB$ guarantees the strict positiveness of $f(z)-f(z')$ and $f(z'')-f(z)$.
-Therefore, the denominators in the above quantities $C_{\alpha,\beta,\Delta},\ D_{\alpha,\beta,\Delta}$ are well-defined.
-Furthermore, by Theorem~\ref{thm:convexity}, $f(z)-f(z')$ is strictly log-concave in $(z,z')$ for $z\leq z'-\Delta,\ z,z'\in[-\alpha-\beta,\ \alpha+\beta]$.
-Based on Assumption~\ref{ass:joint} and closeness of the feasible set, we have $C_{\alpha,\beta,\Delta}>0$, $D_{\alpha,\beta,\Delta}>0$, $A_{\alpha,\beta,\Delta}>0$.
-\end{rmk}
-\begin{rmk} In particular, for logistic link $f(x)= \frac{1}{1+e^{-x}}$, we have
-\begin{align}
-C_{\alpha,\beta,\Delta}&=\max_{|z|\leq \alpha+\beta}\max_{\substack{z'\leq z-\Delta\\z''\geq z+\Delta}}\max\left\{ \frac{1}{e^\Delta-1}\left(\frac{1+e^{-z'}}{1+e^{-z}}\right),\ \frac{1}{1-e^{-\Delta}}\left(\frac{1+e^{-z''}}{1+e^{-z}}\right)\right\}>0,\\
-D_{\alpha,\beta,\Delta}&=\min_{|z|\leq \alpha+\beta}\frac{e^{-z}}{(1+e^{-z})^2}>0.\end{align}
-\end{rmk}
-\begin{thm}[Statistical convergence with unknown $\mb$]\label{thm:ratejoint}
-Consider an ordinal tensor $\tY\in[L]^{d_1\times \cdots \times d_K}$ generated from model~\eqref{eq:model} with the link function $f$ and parameters $(\trueT, \trueb)\in\tP\times \tB$.
-Suppose the link function $f$ satisfies Assumption~\ref{ass:joint}.
-Define $r_{\max}=\max_k r_k+1$, and assume $r_{\max}=\tO(1)$.
-
-Then with very high probability, the estimator in~\eqref{eq:joint} satisfies
-\begin{equation}\label{eq:scoreZ}
-\FnormSize{}{\hat \tZ-\trueZ}^2 \leq {c_1r^{K}_{\max}C^2_{\alpha,\beta,\Delta}\over A^2_{\alpha,\beta,\Delta}D^2_{\alpha,\beta,\Delta}}\left(L-1+\sum_k d_k\right),
-\end{equation}
-In particular,
-\begin{equation}\label{eq:jointTheta}
-\mathrm{MSE}\left(\hat \Theta, \trueT \right)\leq\min\left\{ 4\alpha^2,\  {c_1 r_{\max}^{K}  C_{\alpha,\beta,\Delta}^2 \over A^2_{\alpha,\beta,\Delta}D_{\alpha,\beta,\Delta}^2}\left({ L-1+\sum_k d_k \over \prod_k d_k}\right)\right\},
-\end{equation}
-and
-\begin{equation}\label{eq:jointb}
-\mathrm{MSE}\left(\hat \mb, \trueb \right)\leq \min\left\{4\beta^2,\ {c_1 r^{K}_{\max} C^2_{\alpha,\beta,\Delta} \over A^2_{\alpha,\beta,\Delta}D_{\alpha,\beta,\Delta}^2}\left({L-1+\sum_kd_k \over \prod_k d_K}\right)\right\},
-\end{equation}
-where $c_1, C_{\alpha,\beta,\Delta}, D_{\alpha,\beta,\Delta}$ are positive constants independent of the tensor dimension, rank, and number of ordinal levels. 
-\end{thm}
-\begin{proof} 
-The log-likelihood associated with the observed entries in terms of $\tZ$ is
-\begin{equation}\label{eq:loglz}
-\flogl(\tZ) = \sum_{\omega\in\Omega}\sum_{\ell\in[L]}\mathds{1}\{y_\omega=\ell\}\log\left[f(z_{\omega,\ell})-f(z_{\omega,\ell-1})\right].
-\end{equation}
-
-Let $\plogZ=\entry{{\partial \tL_\tY\over \partial z_{\omega,\ell}}}\in\mathbb{R}^{d_1\times \cdots \times d_K\times[L-1]}$ denote the score function, and $\mH=\pplogZ$ the Hession matrix. 
-Based on the definition of $\hat\tZ$, we have the following inequality:
-\begin{equation}\label{eq:ineql}
-\flogl(\hat\tZ) \geq \flogl(\trueZ).
-\end{equation}
-
-Following the similar argument in Theorem \ref{thm:rate} and the inequality \eqref{eq:ineql}, we obtain that
-\begin{equation}\label{eq:boundZ}
-\FnormSize{}{\hat \tZ-\trueZ}^2 \leq c_1r^{K}_{\max}{\snormSize{}{\plogZ(\trueZ)}^2\over \lambda^2_{1} \left(\mH(\check\tZ)\right)},
-\end{equation}
-where $\plogZ(\trueZ)$ is the score evaluated at $\trueZ$, $\mH(\check \tZ)$ is the Hession evaluated at $\check\tZ$, for some $\check\tZ$ between $\hat \tZ$ and $\trueZ$,  and $\lambda_{1}(\cdot)$ is the largest matrix eigenvalue.
-
-We bound the score and the Hessian to obtain \eqref{eq:scoreZ}.
-
-\begin{enumerate}
-\item (Score.) The $(\omega,\ell)$-th entry in $\plogZ$ is 
-\[
-{\partial \tL_\tY\over \partial z_{\omega,\ell}}=\mathds{1}\{y_\omega=\ell\}{\dot{f}(z)\over f(z)-f(z')}\Bigg|_{(z,\ z')=(z_{\omega,\ell},\ z_{\omega,\ell-1})} - \mathds{1}\{y_\omega=\ell+1\}{\dot{f}(z)\over f(z'')-f(z)}\Bigg|_{(z'',\ z)=(z_{\omega,\ell+1},\ z_{\omega,\ell})},
-\]
-which is upper bounded in magnitude by $C_{\alpha,\beta,\Delta}>0$ with zero mean. By Lemma \ref{lem:noisytensor}, with probability at least $1-\exp\left(-c_2'
-\left(\sum_k d_k+ L-1\right)\right)$, we have
-\begin{align}\label{eq:scorebound}
-\snormSize{}{\plogZ(\trueZ)}\leq c_2C_{\alpha,\beta,\Delta}\sqrt{L-1+\sum_k d_k},
-\end{align}
-
-where $c_2,c_2'$ are two positive constants that depend only on $K$.
-\item (Hession.) The entries in the Hession matrix are
-\begin{align}
-\text{Diagonal: }&{\partial^2 \tL_\tY\over \partial z^2_{\omega,\ell}}=\mathds{1}\{y_\omega=\ell\}{\ddot{f}(z)\left(f(z)-f(z')\right)-\dot{f}^2(z)\over \left(f(z)-f(z')\right)^2}\Bigg|_{(z,\ z')=(z_{\omega,\ell},\ z_{\omega,\ell-1})}-\\
-&\hspace{.65in}\mathds{1}\{y_\omega=\ell+1\}{\ddot{f}(z)\left(f(z'')-f(z)\right)+\dot{f}^2(z)\over \left(f(z'')-f(z)\right)^2}\Bigg|_{(z'',\ z)=(z_{\omega,\ell+1},\ z_{\omega,\ell})},\\
-\text{Off-diagonal: }&
-{\partial^2 \tL_\tY\over \partial z_{\omega,\ell}z_{\omega,\ell+1}}=\mathds{1}\{y_\omega=\ell+1\}{\dot{f}(z_{\omega,\ell})\dot{f}(z_{\omega,\ell+1})\over \left(f(z_{\omega,\ell+1})-f(z_{\omega,\ell})\right)^2}\quad \text{and}\quad {\partial^2 \tL_\tY\over \partial z_{\omega,\ell}z_{\omega',\ell'}}=0 \text{ otherwise}.
-\end{align}
-
-Based on Assumption~\ref{ass:joint}, the Hession matrix $\mH$ has the following three properties:
-\begin{enumerate}
-\item The Hession matrix is a block matrix, $\mH=\text{diag}\{\mH_\omega\colon \omega\in[d_1]\times \cdots \times [d_K]\}$, and each block $\mH_{\omega}\in\mathbb{R}^{(L-1)\times (L-1)}$ is a tridiagonal matrix.
-\item The off-diagonal entries are either zero or strictly positive.
-\item The diagonal entries are either zero or strictly negative. Furthermore, 
-\begin{align}
-&\mH_\omega(\ell,\ell)+ \mH_\omega(\ell,\ell-1)+\mH_\omega(\ell,\ell+1)\\
-=&{\partial^2 \tL_\tY\over \partial z^2_{\omega,\ell}}+{\partial^2 \tL_\tY\over \partial z_{\omega,\ell}z_{\omega,\ell+1}}+{\partial^2 \tL_\tY\over \partial z_{\omega,\ell-1}z_{\omega,\ell}}\\
-=&\mathds{1}\{y_\omega=\ell\}{\partial\over \partial z}\left({\dot{f}(z) \over f(z)-f(z')}\right)\Bigg|_{(z,\ z')=(z_{\omega,\ell},\ z_{\omega,\ell-1})}  -\mathds{1}\{y_\omega=\ell+1\}{\partial\over \partial z}\left({\dot{f}(z) \over f(z)-f(z')}\right)\Bigg|_{(z'',\ z)=(z_{\omega,\ell+1},\ z_{\omega,\ell})}\\
-\leq &-D_{\alpha,\beta,\Delta}\mathds{1}\{y_\omega = \ell \text{ or } \ell+1\}.
-\end{align}
-\end{enumerate}
-We will show that, with very high probability over $\tY$, $\mH$ is negative definite in that
-\begin{equation}\label{eq:HessionZ}
-\lambda_{1}(\mH)=\max_{\mz\neq 0}{\mz^T\mH\mz\over\FnormSize{}{\mz}^2} \leq - cA_{\alpha,\beta,\Delta}D_{\alpha,\beta,\Delta},
-\end{equation}
-where $A_{\alpha,\beta,\Delta},\ D_{\alpha,\beta,\Delta}>0$ are constants defined in~\eqref{eq:constantZ}, and $c>0$ is a constant.
-
-Let $\mz_\omega=(z_{\omega,1},\ldots,z_{\omega,L-1})^T\in\mathbb{R}^{L-1}$ and $\mz=(\mz_{1,\ldots,1,1},\ldots,\mz_{d_1,\ldots,d_K,L-1})^T\in\mathbb{R}^{(L-1)\prod_k d_k}$.
-It follows from property (a) that
-\[
-\mz^T\mH\mz =\sum_{\omega}\mz_\omega^T \mH_{\omega} \mz_\omega.
-\]
-Furthermore, from properties (b) and (c) we have
-\begin{align}
-\mz_\omega^T \mH_{\omega} \mz_\omega &=
-\sum_{\ell\in[L-1]}\mH_\omega(\ell,\ell)z_{\omega,\ell}^2+\sum_{\ell\in[L-1]/\{1\}}2\mH_{\omega}(\ell,\ell-1)z_{\omega,\ell}z_{\omega,\ell-1}\\
-&\leq \sum_{\ell\in[L-1]}\mH(\ell,\ell)z^2_{\omega,\ell}+\sum_{\ell\in[L-1]/\{1\}}\mH(\ell,\ell-1)\left[z_{\omega,\ell}^2+z_{\omega,\ell-1}^2\right]\\
-&= \left(\mH(1,1)+\mH(1,2)\right)z_{\omega,1}^2+ \left(\mH(L-1,L-1)+\mH(L-1,L-2)\right)z_{\omega,L-1}^2\\&\quad\quad+\sum_{\ell\in[L-2]/\{1\}}\left(\mH(\ell,\ell)+\mH(\ell,\ell-1)+\mH(\ell,\ell+1)\right)z_{\omega,\ell}^2\\
-& \leq -D_{\alpha,\beta,\Delta} \sum_{\ell}z^2_{\omega,\ell}\mathds{1}\{y_\omega=\ell \text{ or }\ell+1\}.
-\end{align}
-Therefore,
-\begin{equation}\label{eq:sum}
-\mz^T\mH\mz =\sum_{\omega}\mz_\omega^T \mH_{\omega} \mz_\omega \leq -D_{\alpha,\beta,\Delta}\sum_{\omega}\sum_{\ell} z^2_{\omega,\ell}\mathds{1}\{y_\omega=\ell \text{ or }\ell+1\}.
-\end{equation}
-
-Define the subspace:
-\[
-\tS=\{\Vec(\tZ): \tZ=-\Theta\otimes \mathbf{1}+\tJ\otimes \mb,\ (\Theta,\mb)\in(\tP,\tB)\}.
-\]
-It suffices to prove the negative definiteness of Hession when restricted in the subspace $\tS$. Specifically, for any vector $\mz=\entry{z_{\omega,\ell}}\in\tS$,
-\begin{align}
-\sum_{\omega,\ell} z^2_{\omega,\ell}\mathds{1}\{y_\omega=\ell \text{ or }\ell+1\}&=\sum_{\omega,\ell}(-\theta_\omega+b_\ell)^2\mathds{1}\{y_\omega=\ell \text{ or }\ell+1\}\\
-&=\sum_{\omega,\ell}(\theta^2_\omega-2\theta_\omega b_\ell + b^2_\ell)\mathds{1}\{y_\omega=\ell \text{ or }\ell+1\}\\
-&=\sum_{\omega,\ell}\theta^2_\omega\mathds{1}\{y_\omega=\ell \text{ or }\ell+1\}-2\sum_{\omega,\ell} \theta_\omega b_\ell \mathds{1}\{y_\omega=\ell \text{ or }\ell+1\}+\sum_{\omega,\ell}b^2_\ell\mathds{1}\{y_\omega=\ell \text{ or }\ell+1\}\\
-&\geq \sum_{\omega}\theta^2_\omega -2\sum_{\omega,\ell}\theta_\omega b_\ell+\sum_{\ell}b_\ell^2\left(n_{\ell}+n_{\ell+1}\right)\\
-&\geq \sum_{\omega}\theta^2_\omega+\min_\ell\left(n_{\ell}+n_{\ell+1}\right)\sum_{\ell}b_\ell^2
-\end{align}
-
-On the other hand,
-\[
-\FnormSize{}{\mz}^2=\sum_{\omega,\ell}z^2_{\omega,\ell}=\sum_{\omega,\ell}(-\theta_\omega+b_\ell)^2=L_{\text{total}}\sum_{\omega}\theta^2_\omega+d_{\text{total}}\sum_\ell b^2_\ell,
-\]
-where $L_{\text{total}}:=(L-1)$ and $d_{\text{total}}:=\prod_k d_k$.
-
-Therefore, we have
-\begin{align}\label{eq:negativeH}
-\max_{\mz\in\tS,\mz\neq \mathbf{0}}{\sum_{\omega,\ell} z^2_{\omega,\ell}\mathds{1}\{\{y_\omega=\ell \text{ or }\ell+1\}\} \over \FnormSize{}{\mz}^2}&\geq {\sum_{\omega}\theta^2_\omega+\min_\ell\left(n_{\ell}+n_{\ell+1}\right)\sum_{\ell}b_\ell^2\over L_{\text{total}} \sum_{\omega}\theta^2_\omega+d_{\text{total}}\sum_\ell b^2_\ell}
-\geq{\min_{\ell}(n_\ell+n_{\ell+1})\over (1+{\alpha^2\over c\Delta^2})d_{\text{total}}}\notag\\
-&\geq {2A_{\alpha,\beta,\Delta}\over 1+{\alpha^2\over c\Delta^2}}\quad \text{in high probability as $d_{\min}\to \infty$}. 
-\end{align}
-The second inequality in \eqref{eq:negativeH} is from the conditions that 
-\[
-\sum_{\omega}\theta_\omega^2\in [0,\ \alpha^2 d_{\text{total}}]\quad \text{and}\quad \sum_\ell b_\ell^2 \in[c L_{\text{total}}\Delta^2,\  L_{\text{total}}\beta^2],
-\]
-for some universal constant $c>0$. 
-The last inequality in \eqref{eq:negativeH} follows by applying the law of large numbers and the uniform bound $\min_{z_{\omega,\ell}}\mathbb{P}(y_{\omega}=\ell \text{ or }\ell+1|z_{\omega,\ell})\geq 2A_{\alpha,\beta,\Delta}$ to the empirical ratio:
-\[
-{\min_{\ell}(n_\ell+n_{\ell+1})\over d_{\text{total}}}\overset{p}{\to}\min_{\ell} \mathbb{P}(y_{\omega}=\ell \text{ or }\ell+1|z_{\omega,\ell}) \geq 2A_{\alpha,\beta,\Delta}, \quad \text{ in high probability as $d_{\min}\to \infty$}.
-\]
-
-By~\eqref{eq:sum} and~\eqref{eq:negativeH}, we have
-\begin{equation}\label{eq:boundH}
-\mz^T\mH\mz \leq- c'A_{\alpha,\beta,\Delta}D_{\alpha,\beta,\Delta}\FnormSize{}{\mz}^2,
-\end{equation}
-for some constant $c'>0$, therefore~\eqref{eq:HessionZ} is proved.
-\end{enumerate}
-
- Plugging~\eqref{eq:scorebound} and~\eqref{eq:HessionZ} into~\eqref{eq:boundZ} yields
-\[
-\FnormSize{}{\hat \tZ-\trueZ}^2 \leq {c_1r^{K}_{\max}C^2_{\alpha,\beta,\Delta}\over A^2_{\alpha,\beta,\Delta}D^2_{\alpha,\beta,\Delta}}\left(L-1+\sum_k d_k\right).
-\]
-The MSEs for $\hat \Theta$ and $\hat \mb$ readily follow from~\eqref{eq:decomp}.
-
-\end{proof}
-
-\section{Additional explanations of HCP analysis}\label{sec:additionalHCP}
-We perform clustering analyses based on the Tucker representation of the estimated signal tensor $\hat\Theta$.
-The procedure is motivated from the higher-order extension of Principal Component Analysis (PCA) or Singular Value Decomposition (SVD).
-Recall that, in the matrix case, we perform clustering on an $m\times n$ (normalized) matrix $\mX$ based on the following procedure.
-First, we factorize $X$ into
-\begin{equation}
-\mX = \mU\mSigma \mV^T,
-\end{equation}
-where $\mSigma$ is a diagonal matrix and $\mU,\mV$ are factor matrices with orthogonal columns.
-Second, we take each column of $\mV$ as a principal axis and each row in $\mU\mSigma$ as principal component.
-A subsequent multivariate clustering method (such as $K$-means) is then applied to the $m$ rows of $\mU\mSigma$.
-
-We apply a similar clustering procedure to the estimated signal tensor $\hat\Theta$. We factorize $\hat \Theta$ based on Tucker decomposition.
-\begin{equation}\label{eq:Tuckerest}
-\hat \Theta = \hat \tC\times_1\hat \mM_1\times_2\cdots\times_K\hat \mM_K,
-\end{equation}
-where $\hat \tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is the estimated core tensor, $\hat \mM_k\in\mathbb{R}^{d_k\times r_k}$ are estimated factor matrices with orthogonal columns, and $\times_k$ denotes the tensor-by-matrix multiplication~\citep{kolda2009tensor}.
-The mode-$k$ matricization of~\eqref{eq:Tuckerest} gives
-\begin{equation}
-\hat \Theta_{(k)} = \hat \mM_k\hat \tC_{(k)}\left(\hat \mM_K\otimes\cdots\otimes\hat \mM_1\right),
-\end{equation}
-where $\hat \Theta_{(k)} , \hat \tC_{(k)}$ denote the mode-$k$ unfolding of $\hat \Theta$ and $\hat \tC$, respectively.
-We conduct clustering on this  the mode-$k$ unfolded signal tensor.
-We take columns in $\left(\hat \mM_K\otimes\cdots\otimes\hat \mM_1\right)$ as principal axes and rows in $\hat \mM_k\hat \tC_{(k)}$ as principal components.
-Then, we apply $K$-means clustering method to the $d_k$ rows of the matrix $\hat \mM_k\hat \tC_{(k)}$.
-
-We perform a clustering analysis on the 68 brain nodes using the procedure described above.
-Our ordinal tensor method outputs the estimated parameter tensor $\hat\Theta\in\mathbb{R}^{68\times 68\times136}$ with rank $(23,23,8)$.
-We apply $K$-means to the mode-1 principal component matrix of size $68\times 184$ ($184=23\times8$).
-The elbow method suggests 11 clusters among the 68 nodes (see Figure~\ref{figure:elbow}).
-
-\begin{figure}[ht]
-\begin{center}
-\includegraphics[width=0.4\textwidth]{elbowm.pdf}
-\end{center}
-\caption{Elbow plot for determining the number of clusters in $K$-means.}\label{figure:elbow}
-\end{figure}
 

### 2020-06-23 paper/Camera ready/supp.tex
Add files via upload

--- 2020-06-22 paper/Camera ready/ordinalTv6.tex
@@
 \input macros.tex
-
+\usepackage{xr}
+\externaldocument{ordinalTv6}
 \usepackage{amssymb}
@@
 
-\twocolumn[
-\icmltitle{Tensor denoising and completion based on ordinal observations}
-
-% It is OKAY to include author information, even for blind
-% submissions: the style file will automatically remove it for you
-% unless you've provided the [accepted] option to the icml2020
-% package.
-
-% List of affiliations: The first argument should be a (short)
-% identifier you will use later to specify author affiliations
-% Academic affiliations should list Department, University, City, Region, Country
-% Industry affiliations should list Company, City, Region, Country
-
-% You can specify symbols, otherwise they are numbered in order.
-% Ideally, you should not use this facility. Affiliations will be numbered
-% in order of appearance and this is the preferred way.
-%\icmlsetsymbol{equal}{*}
-
-\begin{icmlauthorlist}
-\icmlauthor{Chanwoo Lee}{wi}
-\icmlauthor{Miaoyan Wang}{wi}
-\end{icmlauthorlist}
-
-\icmlaffiliation{wi}{Department of Statistics, University of Wisconsin-Madison, Wisconsin, USA}
-\icmlcorrespondingauthor{Miaoyan Wang}{miaoyan.wang@wisc.edu}
-
-% You may provide any keywords that you
-% find helpful for describing your paper; these are used to populate
-% the "keywords" metadata in the PDF but will not be shown in the document
-\icmlkeywords{Higher-order tensors, ordinal observation, tensor decomposition, tensor completion}
-
-\vskip 0.3in
-]
-
-% this must go after the closing bracket ] following \twocolumn[ ...
-
-% This command actually creates the footnote in the first column
-% listing the affiliations and the copyright notice.
-% The command takes one argument, which is text to display at the start of the footnote.
-% The \icmlEqualContribution command is standard text for equal contribution.
-% Remove it (just {}) if you do not need this facility.
-
-\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
-%\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.
-
-\begin{abstract}
-Higher-order tensors arise frequently in applications such as neuroimaging, recommendation system, and social network analysis.
-We consider the problem of low-rank tensor estimation from possibly incomplete, ordinal-valued observations.
-Two related problems are studied, one on tensor denoising and another on tensor completion.
-We propose a multi-linear cumulative link model, develop a rank-constrained M-estimator, and obtain theoretical accuracy guarantees.
-Our mean squared error bound enjoys a faster convergence rate than previous results, and we show that the proposed estimator is minimax optimal under the class of low-rank models.
-Furthermore, the procedure developed serves as an efficient completion method which guarantees consistent recovery of an order-$K$ $(d,\ldots,d)$-dimensional low-rank tensor using only $\tilde \tO(Kd)$ noisy, quantized observations.
-We demonstrate the outperformance of our approach over previous methods on the tasks of clustering and collaborative filtering.
-
-\end{abstract}
-
-\section{Introduction}
-Multidimensional arrays, a.k.a.\ tensors, arise in a variety of applications including recommendation systems~\citep{baltrunas2011incarmusic}, social networks~\citep{nickel2011three}, genomics~\citep{wang2019three}, and neuroimaging~\citep{zhou2013tensor}.
-There is a growing need to develop general methods that can address two main problems for analyzing these noisy, high-dimensional datasets.
-The first problem is tensor denoising which aims to recover a signal tensor from its noisy entries~\citep{hong2020generalized,wang2019multiway}.
-The second problem is tensor completion which examines the minimum number of entries needed for a consistent recovery~\citep{ghadermarzy2018learning,montanari2018spectral}.
-Low-rankness is often imposed to the signal tensor, thereby efficiently reducing the intrinsic dimension in both problems.
-
-A number of low-rank tensor estimation methods have been proposed~\citep{hong2020generalized,wang2017tensor}, revitalizing classical methods such as CANDECOMP/PARAFAC (CP) decomposition~\citep{hitchcock1927expression} and Tucker decomposition~\citep{tucker1966some}.
-These tensor methods treat the entries as continuous-valued.
-In many cases, however, we encounter datasets of which the entries are qualitative.
-For example, the Netflix problem records the ratings of users on movies over time.
-Each data entry is a rating on a nominal scale \{{\it very like, like, neutral, dislike, very dislike}\}.
-Another example is in the signal processing, where the digits are frequently rounded or truncated so that only integer values are available.
-The qualitative observations take values in a limited set of categories, making the learning problem harder compared to  continuous observations.
-
-Ordinal entries are categorical variables with an ordering among the categories; for example, {\it very like} $\prec$ {\it like} $\prec$ {\it neutral} $\prec \cdots$.
-The analyses of tensors with the ordinal entries are mainly complicated by two key properties needed for a reasonable model.
-First, the model should be invariant under a reversal of categories, say, from the Netflix example, {\it very like} $\succ$ {\it like} $\succ$ {\it neutral} $\succ \cdots$, but not under arbitrary label permutations.
-Second, the parameter interpretations should be consistent under merging or splitting of contiguous categories.
-The classical continuous tensor model~\cite{kolda2009tensor, ghadermarzy2019near} fails in the first aspect, whereas the binary tensor model~\cite{ghadermarzy2018learning} lacks the second property.
-An appropriate model for ordinal tensors has yet to be studied.
-
-\begin{table*}[ht]
-\centerline{\includegraphics[width=1\textwidth]{compare.pdf}}
-\caption{Comparison with previous work.
-For ease of presentation, we summarize the error rate and sample complexity assuming equal tensor dimension in all modes.
-$K$: tensor order; $L$: number of ordinal levels; $d$: dimension at each mode.
-}\label{tab:compare}
-\end{table*}
-
-{\bf Our contributions.} We are among the first to establish the recovery theory for signal tensors and quantization operators simultaneously from a limited number of highly discrete entries.
-Our main contributions are summarized in Table~\ref{tab:compare}.
-We propose a cumulative link model for higher-order tensors, develop a rank-constrained M-estimator, and obtain theoretical accuracy guarantees.
-The mean squared error bound is established, and we show that the obtained bound has minimax optimal rate in high dimensions under the low-rank model.
-Furthermore, our proposal guarantees consistent recovery of an order-$K$ $(d,\ldots,d)$-dimensional low-rank tensor using only $\tilde \tO(Kd)$ noisy, quantized observations.
-
-{\bf Related work.} Our work is connected to non-Gaussian tensor decomposition.
-Existing work focuses exclusively on univariate observations such as binary- or continuous-valued entries~\citep{wang2018learning,ghadermarzy2018learning}.
-The problem of ordinal/quantized tensor is fundamentally more challenging than previously well-studied tensors for two reasons: (a) the entries do not belong to exponential family distribution, and (b) the observation contains much less information, because neither the underlying signal nor the quantization operator is known.
-The limited information makes the statistical recovery notably hard.
-
-We address the challenge by proposing a cumulative link model that enjoys palindromic invariance~\cite{mccullagh1980regression}.
-A distinctive non-monotonic, phase-transition pattern is demonstrated (Fig~\ref{fig:finite}b).
-We prove that recovery from quantized tensors achieves equally good information-theoretical convergence as the continuous tensors.
-These results fills the gap between classical and non-classical (ordinal) observations, thereby greatly enriching the tensor-model literature.
-
-From algorithm perspective, we address the challenge using the (non-convex) alternating algorithm.
-Earlier work has proposed an approximate (convex) algorithm for binary tensor completion~\cite{ghadermarzy2018learning}.
-Unlike matrix problems, convex-relaxation for low-rank tensors suffers from both computational intractability~\cite{hillar2013most} and statistical suboptimality.
-We improve the error bound from $\tO(d^{-(K-1)/2})$ in~\citet{ghadermarzy2018learning} to $\tO(d^{-(K-1)})$ and numerically compare the two approaches.
-
-We also highlight the challenge associated with higher-order tensors.
-Matrix completion has been proposed for binary observations~\cite{cai2013max,davenport2014,bhaskar20151} and for ordinal observations~\cite{bhaskar2016probabilistic}.
-We show that, applying existing matrix methods to higher-order tensors results in suboptimal estimates.
-A full exploitation of the higher-order structure is needed; this is another challenge we address in this paper.
-
-\section{Preliminaries}
-Let $\tY\in\mathbb{R}^{d_1\times \cdots \times d_K}$ denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional tensor.
-We use $y_\omega$ to denote the tensor entry indexed by $\omega$, where $\omega\in[d_1]\times\cdots\times[d_K]$.
-The Frobenius norm of $\tY$ is defined as $\FnormSize{}{\tY}=\sum_\omega y^2_\omega$ and the infinity norm is defined as $\mnormSize{}{\tY}=\max_{\omega}|y_\omega|$.
-We use $\tY_{(k)}$ to denote the unfolded matrix of size $d_k$-by-$\prod_{i\neq k}d_k$, obtained by reshaping the tensor along the mode-$k$.
-The Tucker rank of $\tY$ is defined as a length-$K$ vector $\mr = (r_1,\ldots,r_K)$, where $r_k$ is the rank of matrix $\tY_{(k)}$ for $k \in[K]$.
-
-We use lower-case letters ($a, b, \ldots$) for scalars/vectors, upper-case boldface letters ($\mA, \mB,  \ldots$) for matrices, and calligraphy letters ($\tA,\tB,\ldots$) for tensors of order three or greater.
-An event $A$ is said to occur ``with very high probability'' if $\mathbb{P}(A)$ tends to 1 faster than any polynomial of tensor dimension $d_{\min}=\min\{d_1,\ldots,d_K\} \to\infty$.
-The indicator function of an event $A$ is denoted as $\mathds{1}\{A\}$.
-For ease of notation, we allow basic arithmetic operators (e.g., $\leq, +, -$) to be applied to pairs of tensors in an element-wise manner.
-We use the shorthand $[n]$ to denote $\{1,\ldots,n\}$ for $n \in N_{+}$.
-
-\section{Model formulation and motivation}
-\subsection{Observation model}
-Let $\tY$ denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional data tensor.
-Suppose the entries of $\tY$ are ordinal-valued, and the observation space consists of $L$ ordered levels, denoted by $[L]:=\{1,\ldots,L\}$.
-We propose a cumulative link model for the ordinal tensor $\tY=\entry{y_\omega}\in[L]^{d_1\times \cdots\times d_K}$.
-Specifically, assume the entries $y_\omega$ are (conditionally) independently distributed with cumulative probabilities:
-\begin{equation}\label{eq:model}
-\mathbb{P}(y_\omega\leq \ell)=f(b_\ell-\theta_\omega),\ \text{for all}\ \ell\in[L-1],
-\end{equation}
-where $\mb=(b_1,\ldots,b_{L-1})$ is a set of unknown scalars satisfying $b_1<\cdots <b_{L-1}$, $\Theta=\entry{\theta_\omega}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is a continuous-valued parameter tensor satisfying certain low-dimensional structure (to be specified later), and $f(\cdot)\colon \mathbb{R}\mapsto[0,1]$ is a known, strictly increasing function.
-We refer to $\mb$ as the cut-off points and $f$ the link function.
-
-The formulation~\eqref{eq:model} imposes an additive model to the transformed probability of cumulative categories.
-This modeling choice is to respect the ordering structure among the categories.
-For example, if we choose the inverse link $f^{-1}(x)=\log {x\over 1-x}$ to be the log odds, then the model~\eqref{eq:model} implies linear spacing between the proportional odds:
-\begin{equation}\label{eq:logodd}
-\log {\mathbb{P}(y_{\omega}\leq \ell) \over \mathbb{P}(y_\omega >  \ell) } - \log {\mathbb{P}(y_{\omega}\leq {\ell-1}) \over \mathbb{P}(y_\omega >  {\ell-1}) } = b_\ell-b_{\ell-1},
-\end{equation}
-for all tensor entries $y_\omega$.
-When there are only two categories in the observation space (e.g.
-binary tensors), the cumulative model~\eqref{eq:model} is equivalent to the usual binomial link model.
-In general, however, when the number of categories $L\geq 3$, the proportional odds assumption~\eqref{eq:logodd} is more parsimonious, in that, the ordered categories can be envisaged as contiguous intervals on the continuous scale, where the points of division are exactly $b_1<\cdots <b_{L-1}$.
-This interpretation will be made more explicit in the next section.
-
-\subsection{Latent-variable interpretation}\label{sec:latent}
-
-The ordinal tensor model~\eqref{eq:model} with certain types of link $f$ has the equivalent representation as an $L$-level quantization model on $\tY=\entry{y_\omega}$:
-\begin{align}\label{eq:quantization}
-%y_\omega=\sum_{\ell\in[L]} \ell \mathds{1}\{y^*_\omega \in(b_{\ell-1},b_\ell]\},\ \forall \ \omega\in[d_1]\times \cdots \times [d_k],
-%\end{equation}
-y_\omega&=
-\begin{cases}
-1,& \text{if $y^*_\omega\in(-\infty, b_1]$},\\
-2,& \text{if $y^*_\omega\in(b_1, b_2]$},\\
-\vdots  &\vdots\\
-L,& \text{if $y^*_\omega\in(b_{L-1}, \infty)$},\\
-\end{cases}
-\end{align}
-for all $\omega\in[d_1]\times \cdots \times [d_k]$. Here, $\tY^*=\entry{y^*_\omega}$ is a latent continuous-valued tensor following an additive noise model:
-\begin{align}\label{eq:latent}
-\KeepStyleUnderBrace{\tY^*}_{\text{latent continuous-valued tensor}}=\KeepStyleUnderBrace{\Theta}_{\text{signal tensor}}+\KeepStyleUnderBrace{\tE}_{\text{i.i.d.\ noise}},
-\end{align}
-where $\tE=\entry{\varepsilon_\omega}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is a noise tensor with i.i.d.\ entries according to distribution $\mathbb{P}(\varepsilon)$.
-From the viewpoint of~\eqref{eq:latent}, the parameter tensor $\Theta$ can be interpreted as the latent signal tensor prior to contamination and quantization.
-
-The equivalence between the latent-variable model~\eqref{eq:quantization} and the cumulative link model~\eqref{eq:model} is established if the link $f$ is chosen to be the cumulative distribution function of noise $\varepsilon$, i.e., $f(\theta)=\mathbb{P}(\varepsilon\leq \theta)$.
-We describe two common choices of link $f$, or equivalently, the distribution of $\varepsilon$.
-
-\begin{example}[Logistic model] The logistic model is characterized by~\eqref{eq:model} with $f(\theta)=(1+e^{-\theta/\sigma})^{-1}$, where $\sigma>0$ is the scale parameter.
-Equivalently, the noise $\varepsilon_\omega$ in~\eqref{eq:quantization} follows i.i.d.\ logistic distribution with scale parameter $\sigma$.
-\end{example}
-\begin{example}[Probit model] The probit model is characterized by~\eqref{eq:model} with
-$f(\theta)=\mathbb{P}(z\leq \theta/\sigma)$, where $z\sim N(0,1)$. Equivalently, the noise $\varepsilon_\omega$ in~\eqref{eq:quantization} follows i.i.d.\ $N(0,\sigma^2)$.
-\end{example}
-Other link functions are also possible, such as Laplace, Cauchy, etc~\cite{mccullagh1980regression}.
-All the models share the property that the ordered categories can be thought of as contiguous interval on some continuous scale.
-We should point out that, although the latent-variable interpretation is incisive, our estimation procedure does not refer to the existence of $\tY^*$.
-Therefore, our model~\eqref{eq:model} is general and still valid in the absence of quantization process.
-More generally, we make the following assumptions about the link $f$.
-
-\begin{assumption}\label{ass:link}
-The link function is assumed to satisfy:
-\vspace{-.2cm}
-\begin{enumerate}
-\item $f(\theta)$ is strictly increasing and twice-differentiable in $\theta\in \mathbb{R}$.
-\item $f'(\theta)$ is strictly log-concave and symmetric with respect to $\theta=0$.
-\end{enumerate}
-\end{assumption}
-
-\subsection{Problem 1: Tensor denoising}~\label{sec:denoising}
-The first question we aim to address is tensor denoising:
-
-(P1) Given the quantization process induced by $f$ and the cut-off points $\mb$, how accurately can we estimate the latent signal tensor $\Theta$ from the ordinal observation $\tY$?
-
-Clearly, the problem (P1) cannot be solved uniformly for all possible $\Theta$.
-We focus on a class of ``low-rank'' and ``flat'' signal tensors, which is a plausible assumption in practical applications~\cite{zhou2013tensor,bhaskar20151}.
-Specifically, we consider the parameter space:
-\begin{equation}\label{eq:space}
-\tP=\left\{\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K} \colon \text{rank}(\Theta)\leq \mr,\ \mnormSize{}{\Theta}\leq \alpha\right\}.
-\end{equation}
-where $\mr=(r_1,\ldots,r_K)$ denotes the Tucker rank of $\Theta$.
-
-The parameter tensor of our interest satisfies two constraints.
-The first is that $\Theta$ is a low-rank tensor, with $r_k=\tO(1)$ for all $k\in[K]$.
-Equivalently, $\Theta$ admits the Tucker decomposition:
-\begin{equation}\label{eq:Tucker}
-\Theta=\tC\times_1\mM_1\times_1 \cdots \times_K \mM_K,
-\end{equation}
-where $\tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is a core tensor, $\mM_k\in\mathbb{R}^{d_k\times r_k}$ are factor matrices with orthogonal columns, and $\times_k$ denotes the tensor-by-matrix multiplication~\cite{kolda2009tensor}.
-The Tucker low-rankness is popularly imposed in tensor data analysis, and is shown to provide a reasonable tradeoff between model complexity and model flexibility.
-Note that, unlike matrices, there are various notations of tensor low-rankness, such as CP rank~\cite{hitchcock1927expression} and train rank~\cite{oseledets2011tensor}.
-Some notation of low-rankness may lead to mathematically ill-posed optimization; for example, the best low CP-rank tensor approximation may not exist~\cite{de2008tensor}.
-We choose Tucker representation for well-posedness of optimization and easy interpretation.
-
-The second constraint is that the entries of $\Theta$ are uniformly bounded in magnitude by a constant $\alpha \in \mathbb{R}_{+}$.
-In view of~\eqref{eq:latent}, we refer to $\alpha$ as the signal level.
-The entry-wise bound assumption is a technical condition that avoids the degeneracy in probability estimation with ordinal observations.
-
-\subsection{Problem 2: Tensor completion}
-Motivated by applications in collaborative filtering, we also consider a more general setup when only a subset of tensor entries $y_\omega$ are observed.
-Let $\Omega\subset[d_1]\times \cdots\times[d_K]$ denote the set of observed indices.
-The second question we aim to address is stated as follows:
-
-(P2) Given an incomplete set of ordinal observations $\{y_{\omega}\}_{\omega\in\Omega}$, how many sampled entries do we need to consistently recover $\Theta$ based on the model~\eqref{eq:model}?
-
-The answer to (P2) depends on the choice of $\Omega$.
-We consider a general model on $\Omega$ that allows both uniform and non-uniform sampling.
-Specifically, let $\Pi=\{\pi_{i_1,\ldots,i_K}\}$ denote a predefine probability distribution over the index set such that $\sum_{\omega\in[d_1]\times \cdots \times [d_K]} \pi_\omega =1$.
-We assume that each index in $\Omega$ is drawn with replacement using distribution $\Pi$.
-This sampling model relaxes the uniform sampling in literature and is arguably a better fit in applications.
-
-We consider the same parameter space~\eqref{eq:space} for the completion problem.
-In addition to the reasons mentioned in Section~\ref{sec:denoising}, the entrywise bound assumption also serves as the incoherence requirement for completion.
-In classical matrix completion, the incoherence is often imposed on the singular vectors.
-This assumption is recently relaxed for ``flat'' matrices with bounded magnitude~\cite{negahban2011estimation,cai2013max,bhaskar20151}.
-We adopt the same assumption for higher-order tensors.
-
-\section{Rank-constrained M-estimator}\label{sec:theory}
-We present a general treatment to both problems mentioned above.
-With a little abuse of notation, we use $\Omega$ to denote either the full index set $\Omega=[d_1]\times \cdots \times [d_K]$ (for the tensor denoising) or a random subset induced from the sampling distribution $\Pi$ (for the tensor completion).
-Define $b_0=-\infty$, $b_L=\infty$, $f(-\infty)=0$ and $f(\infty)=1$.
-The log-likelihood associated with the observed entries is
-\begin{align}\label{eq:objective}
-\logl(\Theta, \mb)&=\sum_{\omega\in\Omega}\sum_{\ell\in[L]} \Big\{\mathds{1}\{y_\omega=\ell\} \log \big[f(b_\ell-\theta_\omega)- \notag \\
-&\quad \quad \quad \quad \quad \quad \quad  f(b_{\ell-1}-\theta_\omega)\big]\Big\}.
-\end{align}
-We propose a rank-constrained maximum likelihood estimator (a.k.a.\ M-estimator) for $\Theta$:
-\begin{align}\label{eq:estimator}
-\hat \Theta&=\argmax_{\Theta\in \tP}\logl(\Theta, \mb),\ \text{where}\notag \\
-\tP&=\left\{\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K} \colon \text{rank}(\Theta)\leq \mr,\ \mnormSize{}{\Theta}\leq \alpha\right\}.
- \end{align}
-In practice, the cut-off points $\mb$ are unknown and should be jointly estimated with $\Theta$.
-We will focus on the theory and algorithm with known $\mb$ in the main paper.
-The adaptation for unknown $\mb$ is addressed in Section~\ref{sec:algorithm} and the Supplement.
-
-We define a few key quantities that will be used in our theory. Let $g_\ell=f(\theta+b_\ell)-f(\theta+b_{\ell-1})$ for all $\ell\in[L]$, and
-\begin{align}\label{eq:regular}
-A_\alpha &= \min_{\ell\in[L], |\theta|\leq \alpha} g_\ell(\theta),\quad  U_\alpha=\max_{\ell\in[L], |\theta|\leq \alpha } {|\dot{g}_\ell(\theta)|\over g_\ell(\theta)},\\
- L_\alpha &= \min_{\ell\in[L], |\theta|\leq \alpha }\left[{\dot{g}_\ell^2 (\theta)\over g_\ell^2(\theta)} -{\ddot{g}_\ell(\theta)\over g_\ell(\theta)}\right],
-\end{align}
-where $\dot{g}(\theta)=dg(\theta)/d\theta$, and $\alpha$ is the entrywise bound of $\Theta$.
-In view of equation~\eqref{eq:latent}, these quantities characterize the geometry including flatness and convexity of the latent noise distribution.
-Under the Assumption~\ref{ass:link}, all these quantities are strictly positive and independent of tensor dimension.
-
-\subsection{Estimation error for tensor denoising}\label{sec:denosing}
-For the tensor denoising problem, we assume that the full set of tensor entries are observed. We assess the estimation accuracy using the mean squared error (MSE):
-\[
-\text{MSE}(\hat \Theta, \trueT)={1\over \prod_k d_k}\FnormSize{}{\Theta-\trueT}^2.
-\]
-The next theorem establishes the upper bound for the MSE of the proposed $\hat \Theta$ in~\eqref{eq:estimator}.
-
-\begin{thm}[Statistical convergence] \label{thm:rate}
-Consider an ordinal tensor $\tY\in[L]^{d_1\times\dots\times d_K}$ generated from model~\eqref{eq:model}, with the link function $f$ and the true coefficient tensor $\trueT\in\tP$.
-Define $r_{\max}=\max_k r_k$.
-Then, with very high probability, the estimator in~\eqref{eq:estimator} satisfies
-\begin{equation}\label{eq:rate}
-\mathrm{MSE}(\hat \Theta, \trueT) \leq \min\left( 4\alpha^2,\ {c_1  U^2_\alpha r_{\max}^{K-1}  \over  L^2_\alpha } {\sum_kd_k\over  \prod_k d_k} \right),
-\end{equation}
-where $c_1 >0$ is a constant that depends only on $K$.
-\end{thm}
-Theorem~\ref{thm:rate} establishes the statistical convergence for the estimator~\eqref{eq:estimator}.
-In fact, the proof of this theorem (see the Supplement) shows that the same statistical rate holds, not only for the global optimizer~\eqref{eq:estimator}, but also for any local optimizer $\check \Theta$ in the level set $\{\check\Theta\in\tP: \logl(\check\Theta)\geq \logl(\trueT)\}$.
-This suggests that the local optimality itself is not necessarily a severe concern in our context, as long as the convergent objective is large enough.
-In Section ~\ref{sec:algorithm}, we perform empirical studies to assess the algorithmic stability.
-
-A similar conclusion is obtained for the prediction error, measured in Kullback-Leibler (KL) divergence, between the categorical distributions in the observation space.
-\begin{cor}[Prediction error]~\label{cor:prediction}
-Assume the same set-up as in Theorem~\ref{thm:rate}.
-Let $\mathbb{P}_{\tY}$ and $\hat{\mathbb{P}}_{\tY}$ denote the distributions generating the $L$-level ordinal tensor $\tY$, given the true parameter $\Theta$ and its estimator $\hat \Theta$, respectively.
-Assume $L\geq 2$.
-Then, with very high probability,
-\begin{equation}\label{eq:KLrate}
-\text{KL}(\mathbb{P}_{\tY} || \hat{\mathbb{P}}_{\tY}) \leq  {c_1 U^2_\alpha r_{\max}^{K-1}  \over L^2_\alpha } {(4L-6)\dot{f}^2(0)  \over A_\alpha} {\sum_kd_k\over  \prod_k d_k},
-\end{equation}
-where $c_1 >0$ is the same constant as in Theorem~\ref{thm:rate}.
-\end{cor}
-To gain insight into these bounds, we consider a special setting with equal dimension in all modes, i.e., $d_1=\cdots=d_K=d$. In such a case, our bound \eqref{eq:rate} reduces to
-\begin{equation}\label{eq:ours}
-\text{MSE}(\hat \Theta, \trueT) \asymp d^{-(K-1)}, \quad \text{as}\ d\to \infty.
-\end{equation}
-Hence, our estimator achieves consistency with polynomial convergence rate.
-We compare the bound with existing literature.
-In the special case $L=2$, \citet{ghadermarzy2018learning} proposed a max-norm constrained estimator $\tilde{\Theta}$ with $\text{MSE}(\tilde{\Theta}, \trueT) \asymp  d^{-(K-1)/2}$.
-In contrast, our estimator converges at a rate of $d^{-(K-1)}$, which is substantially faster than theirs.
-This provides a positive answer to the open question posed in~\citet{ghadermarzy2018learning} whether the square root in the bound is removable.
-The improvement stems from utilizing the exact low-rankness of $\Theta$, whereas the surrogate rank measure employed in~\citet{ghadermarzy2018learning} is scale-sensitive.
-
-Our bound also generalizes the previous results on ordinal matrices.
-The convergence rate for rank-constrained matrix estimation was $\tO(1/\sqrt{d})$~\citep{bhaskar2016probabilistic}, which fits into our special case when $K=2$.
-Furthermore, our results~\eqref{eq:rate} and~\eqref{eq:KLrate} reveal that the convergence becomes favorable as the order of data tensor increases.
-Intuitively, the sample size for tensor data analysis is the number of entries, $\prod_k d_k$, and the number of free parameters is roughly on the order of $\sum_{k}d_k$, assuming $r_{\max}=\tO(1)$.
-A higher tensor order implies higher effective sample size per parameter, and thus exhibits a faster convergence rate in high dimensions.
-
-We next show the statistical optimality of our estimator $\hat \Theta$.
-The result is based on the information theory, and applies to all estimators in $\tP$, including but not limited to $\hat \Theta$ in~\eqref{eq:estimator}.
-
-\begin{thm}[Minimax lower bound]\label{thm:minimax}
-Assume the same set-up as in Theorem~\ref{thm:rate}, and $d_{\max}=\max_k d_k \geq 8$.
-Let $\inf_{\hat \Theta}$ denote the infimum over all estimators $\hat \Theta\in\tP$ based on the ordinal tensor observation $\tY\in[L]^{d_1\times \cdots \times d_K}$.
-Then, under the model~\eqref{eq:model},
-\begin{align}\label{eq:lower}
-\inf_{\hat \Theta }\sup_{\trueT\in\tP} \mathbb{P}\Big\{& \text{MSE}(\hat \Theta, \trueT) \\
-& \geq c\min\left( \alpha^2, \ { Cr_{\max}d_{\max} \over \prod_k d_k} \right) \Big\} \geq {1\over 8},
-\end{align}
-where $C=C(\alpha, L, f,\mb)>0$ and $c>0$ are constants independent of tensor dimension and the rank.
-\end{thm}
-We see that the lower bound matches the upper bound in~\eqref{eq:rate} on the polynomial order of tensor dimension. Therefore, our estimator~\eqref{eq:estimator} is rate-optimal.
-
-\subsection{Sample complexity for tensor completion}
-We now consider the tensor completion problem, when only a subset of entries $\Omega$ are observed.
-We consider a general sampling procedure induced by $\Pi$.
-The recovery accuracy is assessed by the weighted squared error:
-\begin{align}\label{eq:weighted}
-\PiFnormSize{}{\Theta-\hat \Theta}^2&\stackrel{\text{def}}{=}
-{1\over |\Omega|}\mathbb{E}_{\Omega\sim \Pi}\FnormSize{}{\Theta-\hat \Theta}^2\notag \\
-&=\sum_{\omega\in[d_1]\times \cdots \times [d_K]} \pi_{\omega}(\Theta_{\omega}-\hat \Theta_{\omega})^2.
-\end{align}
-
-Note that the recovery error depends on the distribution $\Pi$.
-In particular, tensor entries with higher sampling probabilities have more influence on the recovery accuracy, compared to the ones with lower sampling probabilities.
-\begin{rmk} If we assume each entry is sampled with strictly positive probability; i.e.\ there exits a constant $\mu> 0$ s.t.
-\[
-\ \pi_\omega\geq {1\over \mu \prod_k d_k},\quad \text{for all}\ \omega\in[d_1]\times\cdots \times [d_K],
-\]
-then the error in~\eqref{eq:weighted} provides an upper bound for MSE:
-\[
-\PiFnormSize{}{\Theta-\hat \Theta}^2 \geq {\FnormSize{}{\Theta-\hat \Theta}^2\over \mu \prod_kd_k}={1\over \mu}\text{MSE}(\hat \Theta, \trueT).
-\]
-The equality is attained under uniform sampling with $\mu=1$.
-\end{rmk}
-
-\begin{thm}~\label{thm:completion}
-Assume the same set-up as in Theorem~\ref{thm:rate}.
-Suppose that we observe a subset of tensor entries $\{y_\omega\}_{\omega\in\Omega}$, where $\Omega$ is chosen at random with replacement according to a probability distribution $\Pi$.
-Let $\hat \Theta$ be the solution to~\eqref{eq:estimator}, and assume $r_{\max}=\tO(1)$.
-Then, with very high probability,
-\[
-\PiFnormSize{}{\Theta-\hat \Theta}^2 \to 0,\quad \text{ as }\quad {|\Omega|\over \sum_k d_k} \ \to \infty.
-\]
-\end{thm}
-Theorem~\ref{thm:completion} shows that our estimator achieves consistent recovery using as few as $\tilde{\tO}(Kd)$ noisy, quantized observations from an order-$K$ $(d,\ldots,d)$-dimensional tensor.
-Note that $\tilde \tO(Kd)$ roughly matches the degree of freedom for an order-$K$ tensor of fixed rank $\mr$, suggesting the optimality of our sample requirement.
-This sample complexity substantially improves over earlier result $\tO(d^{\lceil K/2\rceil})$ based on square matricization~\cite{mu2014square}, or $\tO(d^{K/2})$ based on tensor nuclear-norm regularization~\cite{yuan2016tensor}.
-Existing methods that achieve $\tilde \tO(Kd)$ sample complexity require either a deterministic cross sampling design~\cite{zhang2019cross} or univariate measurements~\cite{ghadermarzy2018learning}.
-Our method extends the conclusions to multi-level measurements under a broader class of sampling schemes.
-
-\section{Numerical Implementation}\label{sec:algorithm}
-We describe the algorithm to seek the optimizer of~\eqref{eq:objective}.
-In practice, the cut-off points $\mb$ are often unknown, so we choose to maximize $\logl$ jointly over $(\Theta,\mb)\in\tP\times \tB$ (see the Supplement for details).
-The objective $\logl$ is concave in $(\Theta,\mb)$ whenever $f'$ is log-concave.
-However, the feasible set $\tP$ is non-convex, which makes the optimization~\eqref{eq:objective} a non-convex problem.
-We employ the alternating optimization approach by utilizing the Tucker representation of $\Theta$.
-Specifically, based on~\eqref{eq:Tucker} and~\eqref{eq:objective}, the objective function consists of $K+2$ blocks of variables, one for the cut-off points $\mb$, one for the core tensor $\tC$, and $K$ for the factor matrices $\mM_k$'s.
-The optimization is a simple convex problem if any $K+1$ out of the $K+2$ blocks are fixed.
-We update one block at a time while holding others fixed, and alternate the optimization throughout the iteration.
-The convergence is guaranteed whenever $\logl$ is bounded from above, since the alternating procedure monotonically increases the objective.
-The Algorithm~\ref{alg} gives the full description.
-
-\begin{algorithm}[t]
-        \caption{Ordinal tensor decomposition}\label{alg}
-        \begin{algorithmic}[]
-            \STATE{\bfseries Input:}  \text{ Ordinal data tensor }
-            $\mathcal{Y}\in [L]^{d_1\times\cdots\times d_K}$, rank $\mr\in \mathbb{N}_{+}^{K}$, entry-wise bound $\alpha\in \mathbb{R_+}$.
-            \STATE{\bfseries Output:} $(\hat\Theta,\hat{\mb}) =  \argmax_{(\Theta,\mb)\in \mathcal{P}\times\mathcal{B}}  \mathcal{L}_{\mathcal{Y},\Omega}(\Theta,\mb).$
-            \STATE Random initialization of core tensor $\mathcal{C}^{(0)}$, factor matrices $ \{\mM_k^{(0)}\}$, and cut-off points $\mb^{(0)}$.
-            \FOR{$t = 1,2,\cdots,$}
-            \FOR{$k = 1,2,\cdots,K$}
-            \STATE{Update $\mM^{(t+1)}_k$ while fixing other blocks:}
-            \STATE $\mM_k^{(t+1)}\gets\argmax_{\mM_k\in\mathbb{R}^{d_k\times r_K}}\mathcal{L}_{\mathcal{Y},\Omega}(\mM_k)$,
-            \newline
-             s.t. $\mnormSize{}{\Theta^{(t+1)}}\leq \alpha$, where $\Theta^{(t+1)}$ is the parameter tensor based on the current block estimates.
-            \ENDFOR
-            \STATE {Update $\mathcal{C}^{(t+1)}$ while fixing other blocks:}
-              \newline$\mathcal{C}^{(t+1)} \gets \argmax_{\mathcal{C}\in \mathbb{R}^{r_1\times\cdots\times r_K}}\mathcal{L}_{\mathcal{Y},\Omega}(\mathcal{C})$, s.t.
-$\mnormSize{}{\Theta^{(t+1)}}\leq \alpha.$
-               \STATE {Update $\Theta^{(t+1)}$ based on the current block estimates:}
-                   \newline
-                   $\Theta^{(t+1)} \gets \mathcal{C}^{(t+1)}\times_1\mM_1^{(t+1)}\cdots\times_K\mM_K^{(t+1)}.$
-            \STATE {Update $\mb^{(t+1)}$ while fixing $\Theta^{(t+1)}$:}
-              \newline
-              $\mb^{(t+1)} \gets \argmax_{\mb\in \tB}\mathcal{L}_{\mathcal{Y},\Omega}\big(\Theta^{(t+1)},\mb\big).$
-            \ENDFOR
-            \STATE \textbf{return}
-            $(\hat \Theta,\hat \mb)$
-        \end{algorithmic}
- \end{algorithm}
-We comment on two implementation details before concluding this section.
-First, the problem~\eqref{eq:estimator} is non-convex, so Algorithm 1 usually has no theoretical guarantee on global optimality.
-Nevertheless, as shown in Section~\ref{sec:denosing}, the desired rate holds not only for the global optimizer, but also for the local optimizer with $\logl(\hat \Theta)\geq \logl(\trueT)$.
-In practice, we find the convergence point $\hat \Theta$ upon random initialization is often satisfactory, in that the corresponding objective $\logl(\hat \Theta)$ is close to and actually slightly larger than the objective evaluated at the true parameter $\logl(\trueT)$.
-Figure~\ref{fig:stability} shows the trajectory of the objective function that is output in the default setting of Algorithm 1, with the input tensor generated from probit model~\eqref{eq:model} with $d_1=d_2=d_3=d$ and $r_1=r_2=r_3=r$.
-The dashed line is the objective value at the true parameter $\logl(\trueT)$.
-We find that the algorithm generally converges quickly to a desirable value in reasonable number of steps.
-The actual running time per iteration is shown in the plot legend.
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=8cm]{algorithm.pdf}
-\caption{Trajectory of objective function with various $d$ and $r$.}\label{fig:stability}
-\end{figure}
-
-Second, the algorithm takes the rank $\mr$ as an input.
-In practice, the rank $\mr$ is hardly known and needs to be estimated from the data.
-We suggest to use Bayesian information criterion (BIC) and choose the rank that minimizes BIC; i.e.
-\begin{align}\label{eq:BIC}
-\hat \mr&=\argmin_{\mr\in\mathbb{N}^K_{+}} \text{BIC}(\mr)\\
-&=\argmin_{\mr\in\mathbb{N}^K_{+}}\{-2\tL_{\tY}(\hat \Theta(\mr),\hat \mb(\mr))+p_e(\mr)\log (\prod_k d_k)\},
-\end{align}
-where $\hat \Theta(\mr), \hat\mb(\mr)$ are the estimates given the rank $\mr$, and $p_e(\mr)\stackrel{\text{def}}{=}\sum_k (d_k-r_k)r_k+\prod_k r_k$ is the effective number of parameters in the model.
-We select $\hat \mr$ that minimizes BIC through a grid search.
-The choice of BIC is intended to balance between the goodness-of-fit for the data and the degrees of freedom in the population model.
-
-\section{Experiments}\label{sec:experiment}
-In this section, we evaluate the empirical performance of our methods\footnote{Software package: \url{https://CRAN.R-project.org/package=tensorordinal}}.
-We investigate both the complete and the incomplete settings, and compare the recovery accuracy with other tensor-based methods.
-Unless otherwise stated, the ordinal data tensors are generated from model~\eqref{eq:model} using standard probit link $f$.
-We consider the setting with $K=3$, $d_1=d_2=d_3=d$, and $r_1=r_2=r_3=r$.
-The parameter tensors are simulated based on~\eqref{eq:Tucker}, where the core tensor entries are i.i.d.\ drawn from $\tN(0,1)$, and the factors $\mM_k$ are uniformly sampled (with respect to Haar measure) from matrices with orthonormal columns.
-We set the cut-off points $b_\ell=f^{-1}(\ell/L)$ for $\ell\in[L]$, such that $f(b_\ell)$ are evenly spaced from 0 to 1.
-In each simulation study, we report the summary statistics across $n_{\text{sim}}$ = 30 replications.
-
-\subsection{Finite-sample performance}\label{sec:simulation}
-The first experiment examines the performance under complete observations.
-We assess the empirical relationship between the MSE and various aspects of model complexity, such as dimension $d$, rank $r$, and signal level $\alpha=\mnormSize{}{\Theta}$.
-Figure~\ref{fig:finite}a plots the estimation error versus the tensor dimension $d$ for three different ranks $r\in\{3,5,8\}$.
-The decay in the error appears to behave on the order of $d^{-2}$, which is consistent with our theoretical results~\eqref{eq:rate}.
-We find that a higher rank leads to a larger error, as reflected by the upward shift of the curve as $r$ increases.
-Indeed, a higher rank implies the higher number of parameters to estimate, thus increasing the difficulty of the estimation.
-Figure~\ref{fig:finite}b shows the estimation error versus the signal level under $d=20$.
-Interestingly, a larger estimation error is observed when the signal is either too small or too large.
-The non-monotonic behavior may seem surprising, but this is an intrinsic feature in the estimation with ordinal data.
-In view of the latent-variable interpretation (see Section~\ref{sec:latent}), estimation from ordinal observation can be interpreted as an inverse problem of quantization.
-Therefore, the estimation error diverges in the absence of noise $\tE$, because it is impossible to distinguish two different signal tensors, e.g., $\Theta_1=\ma_1\otimes \ma_2\otimes \ma_3$ and $\Theta_2=\text{sign}(\ma_1)\otimes \text{sign}(\ma_2)\otimes \text{sign}(\ma_3)$, from the quantized observations.
-This phenomenon~\cite{davenport2014,sur2019modern} is clearly contrary to the classical continuous-valued tensor problem.
-
-The second experiment investigates the incomplete observations.
-We consider $L$-level tensors with $d=20$, $\alpha=10$ and choose a subset of tensor entries via uniform sampling.
-Figure~\ref{fig:finite}c shows the estimation error of $\hat \Theta$ versus the fraction of observation $\rho=|\Omega|/d^K$.
-As expected, the error reduces with increased $\rho$ or decreased $r$.
-Figure~\ref{fig:finite}d evaluates the impact of ordinal levels $L$ to estimation accuracy, under the setting $\rho=0.5$.
-An improved performance is observed as $L$ grows, especially from binary observations ($L=2$) to multi-level ordinal observations ($L\geq 3$).
-The result showcases the benefit of multi-level observations compared to binary observations.
-
-\begin{figure}[http]
-\includegraphics[width=8.3cm]{panel.pdf}
-\caption{Empirical relationship between (relative) MSE versus (a) dimension $d$, (b) signal level $\alpha$, (c) observation fraction $\rho$, and (d) number of ordinal levels $L$.
-In panels (b)-(d), we plot the relative MSE $=\FnormSize{}{\hat \Theta-\trueT}^2/\FnormSize{}{\trueT}^2$ for better visualization.}
-\label{fig:finite}
-\vspace{-.3cm}
-\end{figure}
-
-\vspace{-.2cm}
-\subsection{Comparison with alternative methods}
-Next, we compare our ordinal tensor method ({\bf Ordinal-T}) with three popular low-rank methods:
-\begin{itemize}[itemsep=0pt,topsep=0pt,leftmargin=*,partopsep=0pt]
-\item Continuous tensor decomposition ({\bf Continuous-T})~\cite{acar2010scalable} is a low-rank approximation method based on classical Tucker model.
-\item One-bit tensor completion ({\bf 1bit-T})~\cite{ghadermarzy2018learning} is a max-norm penalized tensor learning method based on partial binary observations.
-\item Ordinal matrix completion ({\bf Ordinal-M})~\cite{bhaskar2016probabilistic} is a rank-constrained matrix estimation method based on noisy, quantized observations.
-\end{itemize}
-
-We apply each of the above methods to $L$-level ordinal tensors $\tY$ generated from model~\eqref{eq:model}.
-The {\bf Continuous-T} is applied directly to $\tY$ by treating the $L$ levels as continuous observations.
-The {\bf Ordinal-M} is applied to the 1-mode unfolding matrix $\tY_{(1)}$.
-The {\bf 1bit-T} is applied to $\tY$ in two ways.
-The first approach ({\bf 1bit-sign-T}) follows from~\citet{ghadermarzy2018learning} that transforms
-$\tY$ to a binary tensor, by taking the entrywise sign of the mean-adjusted tensor, $\tY- |\Omega|^{-1}\sum_\omega y_\omega$.
-The second approach ({\bf 1bit-category-T}) transforms the order-3 ordinal tensor $\tY$ to an order-4 binary tensor, $\tY^{\sharp}=\entry{y^{\sharp}_{ijkl}}$, via dummy variable encoding; i.e., $y^{\sharp}_{ijk\ell}=\mathds{1}\{y_{ijk}=\ell\}$ for $\ell\in[L-1]$.
-We evaluate the methods by their capabilities in predicting the most likely labels, $y_\omega^{\text{mode}}=\arg\max_\ell\mathbb{P}(y_\omega=\ell)$.
-Two performance metrics are considered: mean absolute deviation (MAD) $=d^{-K}\sum_\omega |y_\omega^{\text{mode}}-\hat y_\omega^{\text{mode}}|$, and misclassification rate (MCR) $=d^{-K}\sum_\omega\mathds{1}\{y_\omega^{\text{mode}}\neq\text{round}(\hat y_\omega^{\text{mode}})\}$,
-where $\text{round}(\cdot)$ denotes the nearest integer. Note that MAD penalizes the large deviation more heavily than MCR.
-
-Figure~\ref{fig:compare} compares the prediction accuracy under the setting $\alpha=10$, $d=20$, and $r=5$.
-We find that our method outperforms the others in both MAD and MCR.
-In particular, methods built on multi-level observations ({\bf Ordinal-T}, {\bf Ordinal-M}, {\bf 1bit-category-T}) exhibit stable MCR over $\rho$ and $L$, whereas the others two methods ({\bf Continuous-T}, {\bf 1bit-sign-T}) generally fail except for $L=2$ (Figures~\ref{fig:compare}a-b).
-This observation highlights the benefits of multi-level modeling in the classification task.
-Although {\bf 1bit-category-T} and our method {\bf Ordinal-T} behave similarly for binary tensors ($L=2$), the improvement of our method is substantial as $L$ increases (Figures~\ref{fig:compare}a and~\ref{fig:compare}c).
-One possible reason is that our method incorporates the intrinsic ordering among the $L$ levels via proportional odds assumption~\eqref{eq:logodd}, whereas {\bf 1bit-category-T} ignores the ordinal structure and dependence among the induced binary entries.
-
-Figures~\ref{fig:compare}c-d assess the prediction accuracy with sample size.
-We see a clear advantage of our method ({\bf Ordinal-T}) over the matricization ({\bf Ordinal-M}) in both complete and non-complete observations.
-When the observation fraction is small, e.g., ${|\Omega|/ d^K}=0.4$, the tensor-based completion shows $\sim$ 30\% reduction in error compared to the matricization.
-
-\begin{figure}[http]
-\includegraphics[width=8.3cm]{panel2.pdf}
-\caption{Performance comparison for predicting most likely labels.
-(a, c) Prediction errors versus sample complexity $\rho=|\Omega|/d^K$ when $L=5$.
-(b, d) Prediction errors versus the number of ordinal levels $L$ when $\rho=0.8$.
-}\label{fig:compare}
-\vspace{-.4cm}
-\end{figure}
-
-\begin{figure}[http]
-\includegraphics[width=8.3cm]{panel3.pdf}
-\caption{Performance comparison for predicting median labels.
-(a, c) Prediction error versus sample complexity $\rho=|\Omega|/d^K$ when $L=5$.
-(b, d) Prediction error versus the number of ordinal levels $L$, when $\rho=0.8.$}\label{fig:compare2}
-\vspace{-.4cm}
-\end{figure}
-
-We also compare the methods by their performance in predicting the median labels, $y_\omega^{\text{median}}=\min\{\ell\colon\mathbb{P}(y_\omega=\ell)\geq 0.5\}$.
-Under the latent variable model~\eqref{eq:latent} and Assumption~\ref{ass:link}, the median label is the quantized $\theta_\omega$ without noise; i.e.\ $y_\omega^{\text{median}}=\sum_\ell \mathds{1}\{\theta_\omega\in(b_{\ell-1},b_\ell]\}$.
- We utilize the same simulation setting as in the earlier experiment.
-Figure~\ref{fig:compare2} shows that our method outperforms the others in both MCR and MAD.
-The improved accuracy comes from the incorporation of multilinear low-rank structure, multi-level observations, and the ordinal structure.
-Interestingly, the median estimator tends to yield smaller MAD than the mode estimator, $\text{MAD}(\tY^{\text{median}}, \hat \tY^{\text{median}})\leq \text{MAD}(\tY^{\text{mode}}, \hat \tY^{\text{mode}})$  (Figures~\ref{fig:compare}a-b vs.\ Figures~\ref{fig:compare2}a-b), for the three multilevel methods ({\bf 1bit-sign-T}, {\bf Ordinal-M}, and {\bf Ordinal-T}).
-The mode estimator, on the other hand, tends to yield smaller MCR than the median estimator, $\text{MCR}(\tY^{\text{mode}}, \hat \tY^{\text{mode}})\leq\text{MCR}(\tY^{\text{median}}, \hat \tY^{\text{median}})$ (Figures~\ref{fig:compare}c-d vs.\ Figures~\ref{fig:compare2}c-d).
-This tendency is from the property that the median estimator $\hat y^{(\text{median})}_\omega$ minimizes $R_1(z)=\mathbb{E}_{y_\omega}|y_\omega-z|$, whereas the mode estimator $\hat y^{(\text{mode})}_\omega $ minimizes $R_2(z)=\mathbb{E}_{y_\omega}\mathds{1}\{y_\omega=z\}$.
-Here the expectation is over the categorical distribution of $y_\omega$ given parameters $\hat \Theta$ and $\hat \mb$.
-
-\section{Data Applications}
-We apply our ordinal tensor method to two real-world datasets.
-In the first application, we use our model to analyze an ordinal tensor consisting of structural connectivities among 68 brain regions for 136 individuals from Human Connectome Project (HCP)~\citep{geddes2016human}.
-In the second application, we perform tensor completion to an ordinal dataset with missing values.
-The data tensor records the ratings of 139 songs on a scale of 1 to 5 from 42 users on 26 contexts~\citep{baltrunas2011incarmusic}.
-
-\begin{table*}[http]
-\includegraphics[width=1\textwidth]{HCPtable.pdf}
-\vspace{-.7cm}
-\caption{Node clusters in the HCP analysis.
-The first alphabet in the node name indicates the left (L) or right (R) hemisphere.
-The number in the parentheses indicates the node count in each cluster.
-}  \label{table:clustering}
-\vspace{-.4cm}
-\end{table*}
-
-\subsection{Human Connectome Project (HCP)}
-Each entry in the HCP dataset takes value on a nominal scale, \{{\it high, moderate, low}\}, indicating the strength level of fiber connection.
-We convert the dataset to a 3-level ordinal tensor $\tY\in[3]^{68\times 68\times 136}$ and apply the ordinal tensor method with a logistic link function.
-The BIC suggests $\mr = (23,23,8)$ with $\logl(\hat{\Theta},\hat{\mb}) = -216,646$.
-Based on the estimated Tucker factors $\{\hat \mM_k\}$, we perform a clustering analysis via K-mean on the brain nodes (see detailed procedure in the Supplement).
-We find that the clustering successfully captures the spatial separation between brain regions (Table~\ref{table:clustering}).
-In particular, cluster I represents the connection between the left and right hemispheres, whereas clusters II-III represent the connection within each of the half brains (Figure~\ref{figure:brain image}).
-Other smaller clusters represent local regions driving by similar nodes (Table~\ref{table:clustering}).
-For example, the cluster IV/VII consists of nodes in the supramarginal gyrus region in the left/right hemisphere.
-This region is known to be involved in visual word recognition and reading~\citep{stoeckel2009supramarginal}.
-The identified similarities among nodes without external annotations illustrate the applicability of our method to clustering analysis.
-
-\begin{figure}[http]
-\centering
-\includegraphics[width = .5\textwidth]{HCP.pdf}
-\caption{Top three clusters in the HCP analysis.
-(a) Cluster I reflects the connections between two brain hemispheres.
-(b)-(c) Cluster II/III consists of nodes within left/right hemisphere only.
-Node names are shown in abbreviation.
-Edges are colored based on estimated connection averaged across individuals.
-}  \label{figure:brain image}
-\vspace{-.2cm}
-\end{figure}
-
-We acompare the goodness-of-fit of various tensor methods on the HPC data.
-Table~\ref{table:CV} summarizes the prediction error via 5-fold stratified cross-validation averaged over 10 runs.
-Our method outperforms the others, especially in MAD.
-
-\subsection{InCarMusic recommendation system}
-
-We apply ordinal tensor completion to a recommendation system {\it InCarMusic}.
-{\it InCarMusic} is a mobile application that offers music recommendation to passengers of cars based on contexts~\cite{baltrunas2011incarmusic}.
-We conduct tensor completion on the 42-by-139-by-26 tensor with 2,844 observed entries only.
-Table~\ref{table:CV} shows the averaged prediction error via 5-fold cross validation.
-The high missing rate makes the accurate classification challenging.
-Nevertheless, our method achieves the best performance among the three.
-
-\begin{table}[http]
-\resizebox{.48\textwidth}{!}{
-\begin{tabular}{c|c|c}
-\Xhline{2\arrayrulewidth}
-\multicolumn{3}{c}{Human Connectome Project (HCP) dataset}\\
-\Xhline{2\arrayrulewidth}
-Method                    & MAD & MCR\\
-\hline
-Ordinal-T (ours) & 0.1607 (0.0005)& 0.1606 (0.0005)\\
- Continuous-T&0.2530 (0.0002)&0.1599 (0.0002)\\
- 1bit-sign-T&0.3566 (0.0010)&0.1563 (0.0010)\\
- \Xhline{2\arrayrulewidth}
-\multicolumn{3}{c}{InCarMusic dataset}\\
-\Xhline{2\arrayrulewidth}
-Method                    & MAD & MCR\\
-\hline
-Ordinal-T (ours) & 1.37 (0.039) &  0.59 (0.009)\\
-Continuous-T & 2.39 (0.152)&0.94 (0.027)\\
-1bit-sign-T&1.39 (0.003)& 0.81 (0.005)\\
-\hline
-\end{tabular}
-}
-\caption{Comparison of prediction error in the HPC and InCarMusic analyses. Standard errors are reported in parentheses.}\label{table:CV}
-\end{table}
-
-\section{Conclusions}
-We have developed a low-rank tensor estimation method based on possibly incomplete, ordinal-valued observations.
-A sharp error bound is established, and we demonstrate the outperformance of our approach compared to other methods.
-The work unlocks several directions of future research.
-One interesting question would be the inference problem, i.e.,\ to assess the uncertainty of the obtained estimates and the imputation.
-Other directions include the trade-off between (non)convex optimization and statistical efficiency.
-While we have provided numerical evidence for the success of nonconvex approach, the full landscape of the optimization remains open.
-The interplay between computational efficiency and statistical accuracy in general tensor problems warrants future research.
-
-\newpage
-\section*{Acknowledgements}
-This research is supported in part by NSF grant DMS-1915978 and Wisconsin Alumni Research Foundation.
-
-\bibliography{tensor_wang}
-\bibliographystyle{icml2020}
-
-\clearpage
 \appendix
@@
 The elbow method suggests 11 clusters among the 68 nodes (see Figure~\ref{figure:elbow}).
+The clustering result is presented in Section \ref{sec:dataapplication}.
 
@@
 \end{figure}
+
+\bibliography{tensor_wang}
+\bibliographystyle{icml2020}
 

### 2020-06-24 old_draft/old_preprint/ordinalTv5.tex
clean up

--- 2020-06-14 paper/Camera ready/ordinalTv6.tex
@@
 There is a growing need to develop general methods that can address two main problems for analyzing these noisy, high-dimensional datasets.
-The first problem is tensor denoising which aims to recover a signal tensor from its noisy entries~\citep{hong2020generalized,wang2019multiway}.
+The first problem is tensor denoising which aims to recover a signal tensor from its noisy entries~\citep{hong2018generalized,wang2019multiway}.
 The second problem is tensor completion which examines the minimum number of entries needed for a consistent recovery~\citep{ghadermarzy2018learning,montanari2018spectral}.
@@
 
-A number of low-rank tensor estimation methods have been proposed~\citep{hong2020generalized,wang2017tensor}, revitalizing classical methods such as CANDECOMP/PARAFAC (CP) decomposition~\citep{hitchcock1927expression} and Tucker decomposition~\citep{tucker1966some}.
+A number of low-rank tensor estimation methods have been proposed~\citep{hong2018generalized,wang2017tensor}, revitalizing classical methods such as CANDECOMP/PARAFAC (CP) decomposition~\citep{hitchcock1927expression} and Tucker decomposition~\citep{tucker1966some}.
 These tensor methods treat the entries as continuous-valued.
@@
 {\bf Related work.} Our work is connected to non-Gaussian tensor decomposition.
-Existing work focuses exclusively on univariate observations such as binary- or continuous-valued entries~\citep{wang2018learning,ghadermarzy2018learning}.
+Existing work focuses exclusively on univariate observations such as binary- or continuous-valued entries~\citep{hong2018generalized,ghadermarzy2018learning}.
 The problem of ordinal/quantized tensor is fundamentally more challenging than previously well-studied tensors for two reasons: (a) the entries do not belong to exponential family distribution, and (b) the observation contains much less information, because neither the underlying signal nor the quantization operator is known.
@@
 \begin{align}\label{eq:objective}
-\logl(\Theta, \mb)&=\sum_{\omega\in\Omega}\sum_{\ell\in[L]} \Big\{\mathds{1}\{y_\omega=\ell\} \log \big[f(b_\ell-\theta_\omega)- \notag \\
+\logl(\Theta, \mb)&=\sum_{\omega\in\Omega}\sum_{\ell\in[L]} \Big\{\mathds{1}_{\{y_\omega=\ell\}} \log \big[f(b_\ell-\theta_\omega)- \notag \\
 &\quad \quad \quad \quad \quad \quad \quad  f(b_{\ell-1}-\theta_\omega)\big]\Big\}.
@@
 \section*{Acknowledgements}
-This research is supported in part by NSF grant DMS-1915978 and Wisconsin Alumni Research Foundation.
+This research is supported in part by NSF grant DMS-1915978 and the University of Wisconsin-Madison, Office of the Vice Chancellor for Research and Graduate Education with funding from the Wisconsin Alumni Research Foundation.
 
@@
 \onecolumn
-
-\icmltitle{Supplements for ``Tensor denoising and completion based on \\ ordinal observations''}
 
@@
 \begin{align}\label{eq:property}
-{\partial \flogl\over \partial \theta_\omega}&=\sum_{\ell\in[L]}\mathds{1}\{y_{\omega}=\ell\}
+{\partial \flogl\over \partial \theta_\omega}&=\sum_{\ell\in[L]}\mathds{1}_{\{y_{\omega}=\ell\}}
 {\dot{g}_\ell(\theta_\omega)\over g_\ell(\theta_\omega)},\notag\\
-{\partial^2 \flogl\over \partial \theta_\omega^2}&=\sum_{\ell\in[L]}\mathds{1}\{y_\omega=\ell\}{\ddot{g}_\ell(\theta_\omega)g_\ell(\theta_\omega)-\dot{g}^2_\ell(\theta_\omega)\over g^2_\ell(\theta_\omega)}\ \text{and}\quad
+{\partial^2 \flogl\over \partial \theta_\omega^2}&=\sum_{\ell\in[L]}\mathds{1}_{\{y_\omega=\ell\}}{\ddot{g}_\ell(\theta_\omega)g_\ell(\theta_\omega)-\dot{g}^2_\ell(\theta_\omega)\over g^2_\ell(\theta_\omega)}\ \text{and}\quad
 {\partial^2 \flogl\over \partial \theta_\omega \theta_\omega'}=0\ \text{if}\ \omega\neq \omega',
@@
 \end{equation}
-where $\check\Theta=\gamma\trueT+(1-\gamma)\Theta$ for some $\gamma\in[0,1]$, and $\fpplogl(\check\Theta)$ denotes the $d_{\text{total}}$-by-$d_\text{total}$ Hession matrix evaluated at $\check\Theta$.
+$\check\Theta=\gamma\trueT+(1-\gamma)\Theta$ for some $\gamma\in[0,1]$, and $\fpplogl(\check\Theta)$ denotes the $\prod_kd_k$-by-$\prod_k d_k$ Hession matrix evaluated at $\check\Theta$.
 
@@
 \begin{equation}\label{eq:KLbound}
-\text{KL}(\mathbb{P}_{\Theta}||\mathbb{P}_{\mathbf{0}})\leq \gamma^2 r_{\max} d_{\max}.
+\text{KL}(\mathbb{P}_{\Theta}||\mathbb{P}_{\mathbf{0}})\leq \gamma^2 R_{\max} d_{\max}.
 \end{equation}
@@
 \begin{equation}\label{eq:final}
-\inf_{\hat \Theta}\sup_{\trueT\in \tX}\mathbb{P}\left(\FnormSize{}{\hat \Theta- \trueT}\geq  {\gamma\over 8} \min\left\{ \alpha\sqrt{d_{\text{total}}}, C^{-1/2}\sqrt{ r_{\max}d_{\max}}\right\} \right)\geq {1\over 2}\left(1-2\varepsilon-\sqrt{16 \varepsilon \over r_{\max}d_{\max}\log2}\right).
+\inf_{\hat \Theta}\sup_{\trueT\in \tX}\mathbb{P}\left(\FnormSize{}{\hat \Theta- \trueT}\geq  {\gamma\over 8} \min\left\{ \alpha\sqrt{d_{\text{total}}}, C^{-1/2}\sqrt{ R_{\max}d_{\max}}\right\} \right)\geq {1\over 2}\left(1-2\varepsilon-\sqrt{16 \varepsilon \over R_{\max}d_{\max}\log2}\right).
 \end{equation}
@@
 \begin{equation}\label{eq:prob}
-\inf_{\hat \Theta}\sup_{\trueT\in \tP}\mathbb{P}\left(\text{Loss}(\hat \Theta, \trueT)\geq c\min\left \{ \alpha^2,  {C^{-1}r_{\max}d_{\max}\over d_{\text{total}}}\right \}\right)\geq {1\over 2}\left({4\over 5}- \sqrt{1.6\over r_{\max}d_{\max}\log2} \right)\geq {1\over 8},
+\inf_{\hat \Theta}\sup_{\trueT\in \tP}\mathbb{P}\left(\text{Loss}(\hat \Theta, \trueT)\geq c\min\left \{ \alpha^2,  {C^{-1}R_{\max}d_{\max}\over d_{\text{total}}}\right \}\right)\geq {1\over 2}\left({4\over 5}- \sqrt{1.6\over R_{\max}d_{\max}\log2} \right)\geq {1\over 8},
 \end{equation}
@@
 \begin{align}\label{eq:function}
- \logl(\Theta, \mb)&=\sum_{\omega\in\Omega}\sum_{\ell\in[L]} \big(\mathds{1}\{y_\omega=\ell\}\log \left[f(b_\ell-\theta_\omega)-  f(b_{\ell-1}-\theta_\omega)\right]\big),
+ \logl(\Theta, \mb)&=\sum_{\omega\in\Omega}\sum_{\ell\in[L]} \left\{\mathds{1}_{\{y_\omega=\ell\}} \log \big[f(b_\ell-\theta_\omega)-  f(b_{\ell-1}-\theta_\omega)\big]\right\},
  \end{align}
@@
 \begin{align}
-\lambda\colon &\mathbb{R}^2\mapsto \mathbb{R}\\
+\lambda: &\mathbb{R}^2\mapsto \mathbb{R}\\
 (u,v)&\mapsto\lambda(u,v) = \log \big[f(u)-  f(v)\big].
@@
 
-\begin{lem}[Corollary 3.5 in~\citet{brascamp2002extensions}]\label{lem:lossconvexity}
-Let $F(x,y)\colon \mathbb{R}^{m+n}\rightarrow \mathbb{R}$ be an integrable function where $x\in \mathbb{R}^{m},y\in \mathbb{R}^n$. Let
+\begin{lem}[Corollary 3.5 in~\cite{brascamp2002extensions}]\label{lem:lossconvexity}
+Let $F(x,y):\mathbb{R}^{m+n}\rightarrow \mathbb{R}$ be an integrable function where $x\in \mathbb{R}^{m},y\in \mathbb{R}^n$. Let
 \[
@@
 \begin{defn}[Atomic M-norm~\citep{ghadermarzy2019near}]
-Define $T_{\pm}=\{\tT\in\{\pm 1 \}^{d_1\times \cdots \times d_K}\colon \text{rank}(\tT)=1\}$. The atomic M-norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
+Define $T_{\pm}=\{\tT\in\{\pm 1 \}^{d_1\times \cdots \times d_K}\colon$ $\text{rank}(\tT)=1\}$. The atomic M-norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
 \begin{align}
@@
 \left\{
-\sum_{i\in[r]}|\lambda_i|\colon \Theta=\sum_{i=1}^r \lambda_i\mx^{(i)}_1\otimes \cdots \otimes \mx^{(i)}_K,\ \vnormSize{}{\mx^{(i)}_k}=1,\ \mx^{(i)}_k\in\mathbb{R}^{d_k},\ \text{for all}\ k\in[K],\ i\in[r]
+\sum_{i\in[r]}|\lambda_i|\colon \Theta=\sum_{i=1}^r \lambda_i\mx^{(i)}_1\otimes \cdots \otimes \mx^{(i)}_K,\ \vnormSize{}{\mx^{(i)}_k}=1,\ \mx^{(i)}_k\in\mathbb{R}^{d_k},\ \text{for all} k\in[K],\ i\in[r]
 \right\},
@@
 
-The following lemma provides the bound on the spectral norm of random tensors. The result was firstly presented in~\citet{nguyen2015tensor}, and we adopt the version from~\citet{tomioka2014spectral}.
+The following lemma provides the bound on the spectral norm of random tensors. The result was firstly presented in~\cite{nguyen2015tensor}, and we adopt the version from~\cite{tomioka2014spectral}.
 \begin{lem}[Spectral norm of random tensors~\citep{tomioka2014spectral}]\label{lem:tensor}
@@
 
-\begin{lem}\label{lem:KLentry} Let $X,\; Y$ be two discrete random variables taking values on $L$ possible categories, with point mass probabilities $\{p_\ell\}_{\ell\in[L]}$ and $\{q_\ell\}_{\ell\in[L]}$, respectively.
-Suppose $p_\ell$, $q_\ell>0$ for all $\ell\in[L]$.
+\begin{lem}\label{lem:KLentry} Let $X,\; Y$ be two discrete random variables taking values on $L$ possible categories, with category probabilities $\{p_\ell\}_{\ell\in[L]}$ and $\{q_\ell\}_{\ell\in[L]}$, respectively.
+Suppose $p_\ell$, $q_\ell>0$ for all $i\in[L]$.
 Then, the Kullback-Leibler (KL) divergence satisfies that
@@
 \[
-\FnormSize{}{\Theta_i-\Theta_j}^2 \geq {Rd_1\over 8} \gamma^2\min\left\{ \alpha^2, {C^{-1}Rd_1 \over d_{\text{total}}}\right\} \Big\lfloor {d_2\over R}\Big \rfloor \prod_{k\geq 3}d_k\geq {\gamma^2\min\left\{ \alpha^2 d_{\text{total}}, C^{-1}Rd_1\right\}  \over 16}.
+\FnormSize{}{\Theta_i-\Theta_j}^2 \geq {Rd_1\over 8} \gamma^2\min\left\{ \alpha^2, {C^{-1}Rd_1 \over d_{\text{total}}}\right\} \lfloor {d_2\over R} \rfloor \prod_{k\geq 3}d_k\geq {\gamma^2\min\left\{ \alpha^2 d_{\text{total}}, C^{-1}Rd_1\right\}  \over 16}.
 \]
@@
 \[
-\zeronormSize{}{w^{(j)}-w^{(k)}}\geq {m\over 8},\quad \text{for } \ 0\leq j<k\leq M,
+\zeronormSize{}{w^{(j)}-w^{(k)}}\geq {m\over 8},\quad \text{for} \ 0\leq j<k\leq M,
 \]
@@
 
-\begin{lem}[Theorem 2.5 in~\citet{tsybakov2008introduction}]\label{lem:Tsybakov}
+\begin{lem}[Theorem 2.5 in~\cite{tsybakov2008introduction}]\label{lem:Tsybakov}
 Assume that a set $\tX$ contains element $\Theta_0, \Theta_1, \ldots,\Theta_M$ ($M\geq 2$) such that
@@
 
-\begin{lem}[Lemma 28 in~\citet{ghadermarzy2019near}]\label{lem:convexity}
+\begin{lem}[Lemma 28 in~\cite{ghadermarzy2019near}]\label{lem:convexity}
 Define $\mathbb{B}_{M}(\beta)=\{\Theta\in \mathbb{R}^{d_1\times \cdots \times d_K}\colon \anormSize{}{\Theta}\leq \beta \}$.
@@
 \tP&=\{\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}\colon \text{rank}(\tP)\leq \mr,\ \langle \Theta, \tJ \rangle=0,\ \mnormSize{}{\Theta}\leq \alpha\},\\
- \tB&=\{\mb\in\mathbb{R}^{L-1}\colon \mnormSize{}{\mb}\leq \beta,\ \min_{\ell}(b_\ell-b_{\ell-1}) \geq \Delta\},
-\end{align}
-with positive constants $\alpha,\beta,\Delta>0$ and a given rank $\mr\in\mathbb{N}^K_{+}$.
+ \tB&=\{\mb\in\mathbb{R}^{L-1}\colon \mnormSize{}{\mb}\leq \beta,\ \min_{\ell}(b_\ell-b_{\ell-1}) \geq \Delta\}.
+\end{align}
 Here, $\tJ=\entry{1}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ denotes a tensor of all ones.
@@
 \[
-\text{MSE}\left( \hat \Theta, \trueT \right)={1\over \prod_k d_k}\FnormSize{}{\hat \Theta-\trueT}^2,\quad \text{MSE}\left( \hat \mb, \trueb \right)={1\over L-1}\FnormSize{}{\hat \mb-\trueb}^2.
+\text{MSE}\left( \hat \Theta, \trueT \right)={1\over \prod_k d_k}\FnormSize{}{\hat \Theta-\trueT},\quad \text{MSE}\left( \hat \mb, \trueb \right)={1\over L-1}\FnormSize{}{\hat \mb-\trueb}.
 \]
@@
 where $\mathbf{1}$ denotes a length-$(L-1)$ vector of all ones.
-Under the identifiability constraint $\langle \Theta, \tJ \rangle=0$, there is an one-to-one mapping between $\tZ$ and $(\Theta, \mb)$, with $\text{rank}(\tZ)\leq (\text{rank}(\Theta)+1,\ 2)^T$.
+Under the identifiability constraint $\langle \Theta, \tJ \rangle=0$, there is an one-to-one mapping between $\tZ$ and $(\Theta, \mb)$, with $\text{rank}(\tZ)\leq (\text{rank}(\Theta),2)^T$.
 Furthermore, 
@@
 The condition $\Delta=\min_{\ell}(b_{\ell}-b_{\ell-1})>0$ on the feasible set $\tB$ guarantees the strict positiveness of $f(z)-f(z')$ and $f(z'')-f(z)$.
-Therefore, the denominators in the above quantities $C_{\alpha,\beta,\Delta},\ D_{\alpha,\beta,\Delta}$ are well-defined.
-Furthermore, by Theorem~\ref{thm:convexity}, $f(z)-f(z')$ is strictly log-concave in $(z,z')$ for $z\leq z'-\Delta,\ z,z'\in[-\alpha-\beta,\ \alpha+\beta]$.
+Therefore, the denominators in the above quantities $C_{\alpha,\beta,\Delta}, D_{\alpha,\beta,\Delta}$ are well-defined.
+Furthermore, by Theorem~\ref{thm:convexity}, $f(z)-f(z')$ is strictly log-concave in $(z,z')$ for $z\leq z'-\Delta, z,z'\in[-\alpha-\beta,\ \alpha+\beta]$.
 Based on Assumption~\ref{ass:joint} and closeness of the feasible set, we have $C_{\alpha,\beta,\Delta}>0$, $D_{\alpha,\beta,\Delta}>0$, $A_{\alpha,\beta,\Delta}>0$.
@@
 \begin{equation}\label{eq:scoreZ}
-\FnormSize{}{\hat \tZ-\trueZ}^2 \leq {c_1r^{K}_{\max}C^2_{\alpha,\beta,\Delta}\over A^2_{\alpha,\beta,\Delta}D^2_{\alpha,\beta,\Delta}}\left(L-1+\sum_k d_k\right),
+\FnormSize{}{\hat \tZ-\trueZ}^2 \leq {c_1r^{K}_{\max}C^2_{\alpha,\beta,\Delta}(L-1)^2\over A^2_{\alpha,\beta,\Delta}D^2_{\alpha,\beta,\Delta}}\left(L-1+\sum_k d_k\right),
 \end{equation}
@@
 \begin{equation}\label{eq:jointTheta}
-\mathrm{MSE}\left(\hat \Theta, \trueT \right)\leq\min\left\{ 4\alpha^2,\  {c_1 r_{\max}^{K}  C_{\alpha,\beta,\Delta}^2 \over A^2_{\alpha,\beta,\Delta}D_{\alpha,\beta,\Delta}^2}\left({ L-1+\sum_k d_k \over \prod_k d_k}\right)\right\},
+\mathrm{MSE}\left(\hat \Theta, \trueT \right)\leq\min\left( 4\alpha^2,\  {c_1 r_{\max}^{K}  C_{\alpha,\beta,\Delta}^2(L-1)  \over A^2_{\alpha,\beta,\Delta}D_{\alpha,\beta,\Delta}^2}{ L-1+\sum_k d_k \over \prod_k d_k}\right),
 \end{equation}
@@
 \begin{equation}\label{eq:jointb}
-\mathrm{MSE}\left(\hat \mb, \trueb \right)\leq \min\left\{4\beta^2,\ {c_1 r^{K}_{\max} C^2_{\alpha,\beta,\Delta} \over A^2_{\alpha,\beta,\Delta}D_{\alpha,\beta,\Delta}^2}\left({L-1+\sum_kd_k \over \prod_k d_K}\right)\right\},
+\mathrm{MSE}\left(\hat \mb, \trueb \right)\leq \min\left(4\beta^2,\ {c_1 r^{K}_{\max} C^2_{\alpha,\beta,\Delta} (L-1)\over A^2_{\alpha,\beta,\Delta}D_{\alpha,\beta,\Delta}^2}{L-1+\sum_kd_k \over \prod_k d_K}\right),
 \end{equation}
@@
 \begin{equation}\label{eq:loglz}
-\flogl(\tZ) = \sum_{\omega\in\Omega}\sum_{\ell\in[L]}\mathds{1}\{y_\omega=\ell\}\log\left[f(z_{\omega,\ell})-f(z_{\omega,\ell-1})\right].
+\flogl(\tZ) = \sum_{\omega\in\Omega}\sum_{\ell\in[L]}\mathds{1}_{\{y_\omega=\ell\}}\log\left[f(z_{\omega,\ell})-f(z_{\omega,\ell-1})\right].
 \end{equation}
@@
 \[
-{\partial \tL_\tY\over \partial z_{\omega,\ell}}=\mathds{1}\{y_\omega=\ell\}{\dot{f}(z)\over f(z)-f(z')}\Bigg|_{(z,\ z')=(z_{\omega,\ell},\ z_{\omega,\ell-1})} - \mathds{1}\{y_\omega=\ell+1\}{\dot{f}(z)\over f(z'')-f(z)}\Bigg|_{(z'',\ z)=(z_{\omega,\ell+1},\ z_{\omega,\ell})},
+{\partial \tL_\tY\over \partial z_{\omega,\ell}}=\mathds{1}_{\{y_\omega=\ell\}}{\dot{f}(z)\over f(z)-f(z')}\Bigg|_{(z,\ z')=(z_{\omega,\ell},\ z_{\omega,\ell-1})} - \mathds{1}_{\{y_\omega=\ell+1\}}{\dot{f}(z)\over f(z'')-f(z)}\Bigg|_{(z'',\ z)=(z_{\omega,\ell+1},\ z_{\omega,\ell})},
 \]
 which is upper bounded in magnitude by $C_{\alpha,\beta,\Delta}>0$ with zero mean. By Lemma \ref{lem:noisytensor}, with probability at least $1-\exp\left(-c_2'
-\left(\sum_k d_k+ L-1\right)\right)$, we have
+\left(\sum_k d_k+ L-1\right)\log(K+1)\right)$, we have
 \begin{align}\label{eq:scorebound}
@@
 \begin{align}
-\text{Diagonal: }&{\partial^2 \tL_\tY\over \partial z^2_{\omega,\ell}}=\mathds{1}\{y_\omega=\ell\}{\ddot{f}(z)\left(f(z)-f(z')\right)-\dot{f}^2(z)\over \left(f(z)-f(z')\right)^2}\Bigg|_{(z,\ z')=(z_{\omega,\ell},\ z_{\omega,\ell-1})}-\\
-&\hspace{.65in}\mathds{1}\{y_\omega=\ell+1\}{\ddot{f}(z)\left(f(z'')-f(z)\right)+\dot{f}^2(z)\over \left(f(z'')-f(z)\right)^2}\Bigg|_{(z'',\ z)=(z_{\omega,\ell+1},\ z_{\omega,\ell})},\\
+\text{Diagonal: }&{\partial^2 \tL_\tY\over \partial z^2_{\omega,\ell}}=\mathds{1}_{\{y_\omega=\ell\}}{\ddot{f}(z)\left(f(z)-f(z')\right)-\dot{f}^2(z)\over \left(f(z)-f(z')\right)^2}\Bigg|_{(z,\ z')=(z_{\omega,\ell},\ z_{\omega,\ell-1})}-\\
+&\hspace{.65in}\mathds{1}_{\{y_\omega=\ell+1\}}{\ddot{f}(z)\left(f(z'')-f(z)\right)+\dot{f}^2(z)\over \left(f(z'')-f(z)\right)^2}\Bigg|_{(z'',\ z)=(z_{\omega,\ell+1},\ z_{\omega,\ell})},\\
 \text{Off-diagonal: }&
-{\partial^2 \tL_\tY\over \partial z_{\omega,\ell}z_{\omega,\ell+1}}=\mathds{1}\{y_\omega=\ell+1\}{\dot{f}(z_{\omega,\ell})\dot{f}(z_{\omega,\ell+1})\over \left(f(z_{\omega,\ell+1})-f(z_{\omega,\ell})\right)^2}\quad \text{and}\quad {\partial^2 \tL_\tY\over \partial z_{\omega,\ell}z_{\omega',\ell'}}=0 \text{ otherwise}.
+{\partial^2 \tL_\tY\over \partial z_{\omega,\ell}z_{\omega,\ell+1}}=\mathds{1}_{\{y_\omega=\ell+1\}}{\dot{f}(z_{\omega,\ell})\dot{f}(z_{\omega,\ell+1})\over \left(f(z_{\omega,\ell+1})-f(z_{\omega,\ell})\right)^2}\quad \text{and}\quad {\partial^2 \tL_\tY\over \partial z_{\omega,\ell}z_{\omega',\ell'}}=0 \text{ otherwise}.
 \end{align}
@@
 \begin{enumerate}
-\item The Hession matrix is a block matrix, $\mH=\text{diag}\{\mH_\omega\colon \omega\in[d_1]\times \cdots \times [d_K]\}$, and each block $\mH_{\omega}\in\mathbb{R}^{(L-1)\times (L-1)}$ is a tridiagonal matrix.
+\item The Hession matrix is a block matrix, $\mH=\text{diag}\{\mH_\omega: \omega\in[d_1]\times \cdots \times [d_K]\}$, and each block $\mH_{\omega}\in\mathbb{R}^{(L-1)\times (L-1)}$ is a tridiagonal matrix.
 \item The off-diagonal entries are either zero or strictly positive.
@@
 =&{\partial^2 \tL_\tY\over \partial z^2_{\omega,\ell}}+{\partial^2 \tL_\tY\over \partial z_{\omega,\ell}z_{\omega,\ell+1}}+{\partial^2 \tL_\tY\over \partial z_{\omega,\ell-1}z_{\omega,\ell}}\\
-=&\mathds{1}\{y_\omega=\ell\}{\partial\over \partial z}\left({\dot{f}(z) \over f(z)-f(z')}\right)\Bigg|_{(z,\ z')=(z_{\omega,\ell},\ z_{\omega,\ell-1})}  -\mathds{1}\{y_\omega=\ell+1\}{\partial\over \partial z}\left({\dot{f}(z) \over f(z)-f(z')}\right)\Bigg|_{(z'',\ z)=(z_{\omega,\ell+1},\ z_{\omega,\ell})}\\
-\leq &-D_{\alpha,\beta,\Delta}\mathds{1}\{y_\omega = \ell \text{ or } \ell+1\}.
+=&\mathds{1}_{\{y_\omega=\ell\}}{\partial\over \partial z}\left({\dot{f}(z) \over f(z)-f(z')}\right)\Bigg|_{(z,\ z')=(z_{\omega,\ell},\ z_{\omega,\ell-1})}  -\mathds{1}_{\{y_\omega=\ell+1\}}{\partial\over \partial z}\left({\dot{f}(z) \over f(z)-f(z')}\right)\Bigg|_{(z'',\ z)=(z_{\omega,\ell+1},\ z_{\omega,\ell})}\\
+\leq &-D_{\alpha,\beta,\Delta}\mathds{1}_{\{y_\omega = \ell \text{ or } \ell+1\}}.
 \end{align}
@@
 \begin{equation}\label{eq:HessionZ}
-\lambda_{1}(\mH)=\max_{\mz\neq 0}{\mz^T\mH\mz\over\FnormSize{}{\mz}^2} \leq - cA_{\alpha,\beta,\Delta}D_{\alpha,\beta,\Delta},
-\end{equation}
-where $A_{\alpha,\beta,\Delta},\ D_{\alpha,\beta,\Delta}>0$ are constants defined in~\eqref{eq:constantZ}, and $c>0$ is a constant.
+\lambda_{1}(\mH)=\max_{\mz}{\mz^T\mH\mz\over\FnormSize{}{\mz}^2} \leq -{c_3A_{\alpha,\beta,\Delta}D_{\alpha,\beta,\Delta}\over L-1},
+\end{equation}
+where $A_{\alpha,\beta,\Delta}, D_{\alpha,\beta,\Delta}>0$ are constants defined in~\eqref{eq:constantZ}, and $c_3>0$ is a constant.
 
@@
 \]
-Furthermore, from properties (b) and (c) we have
+Furthermore, properties (b) and (c) we have
 \begin{align}
@@
 \sum_{\ell\in[L-1]}\mH_\omega(\ell,\ell)z_{\omega,\ell}^2+\sum_{\ell\in[L-1]/\{1\}}2\mH_{\omega}(\ell,\ell-1)z_{\omega,\ell}z_{\omega,\ell-1}\\
-&\leq \sum_{\ell\in[L-1]}\mH(\ell,\ell)z^2_{\omega,\ell}+\sum_{\ell\in[L-1]/\{1\}}\mH(\ell,\ell-1)\left[z_{\omega,\ell}^2+z_{\omega,\ell-1}^2\right]\\
+&\leq \sum_{\ell\in[L-1]}\mH(\ell,\ell)z_{\omega,\ell}+\sum_{\ell\in[L-1]/\{1\}}\mH(\ell,\ell-1)\left[z_{\omega,\ell}^2+z_{\omega,\ell-1}^2\right]\\
 &= \left(\mH(1,1)+\mH(1,2)\right)z_{\omega,1}^2+ \left(\mH(L-1,L-1)+\mH(L-1,L-2)\right)z_{\omega,L-1}^2\\&\quad\quad+\sum_{\ell\in[L-2]/\{1\}}\left(\mH(\ell,\ell)+\mH(\ell,\ell-1)+\mH(\ell,\ell+1)\right)z_{\omega,\ell}^2\\
-& \leq -D_{\alpha,\beta,\Delta} \sum_{\ell}z^2_{\omega,\ell}\mathds{1}\{y_\omega=\ell \text{ or }\ell+1\}.
+& \leq -D_{\alpha,\beta,\Delta} \sum_{\ell}z^2_{\omega,\ell}\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}.
 \end{align}
@@
 \begin{equation}\label{eq:sum}
-\mz^T\mH\mz =\sum_{\omega}\mz_\omega^T \mH_{\omega} \mz_\omega \leq -D_{\alpha,\beta,\Delta}\sum_{\omega}\sum_{\ell} z^2_{\omega,\ell}\mathds{1}\{y_\omega=\ell \text{ or }\ell+1\}.
+\mz^T\mH\mz =\sum_{\omega}\mz_\omega^T \mH_{\omega} \mz_\omega \leq -D_{\alpha,\beta,\Delta}\sum_{\omega}\sum_{\ell} z^2_{\omega,\ell}\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}.
 \end{equation}
@@
 \begin{align}
-\sum_{\omega,\ell} z^2_{\omega,\ell}\mathds{1}\{y_\omega=\ell \text{ or }\ell+1\}&=\sum_{\omega,\ell}(-\theta_\omega+b_\ell)^2\mathds{1}\{y_\omega=\ell \text{ or }\ell+1\}\\
-&=\sum_{\omega,\ell}(\theta^2_\omega-2\theta_\omega b_\ell + b^2_\ell)\mathds{1}\{y_\omega=\ell \text{ or }\ell+1\}\\
-&=\sum_{\omega,\ell}\theta^2_\omega\mathds{1}\{y_\omega=\ell \text{ or }\ell+1\}-2\sum_{\omega,\ell} \theta_\omega b_\ell \mathds{1}\{y_\omega=\ell \text{ or }\ell+1\}+\sum_{\omega,\ell}b^2_\ell\mathds{1}\{y_\omega=\ell \text{ or }\ell+1\}\\
+\sum_{\omega,\ell} z^2_{\omega,\ell}\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}&=\sum_{\omega,\ell}(-\theta_\omega+b_\ell)^2\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}\\
+&=\sum_{\omega,\ell}(\theta^2_\omega-2\theta_\omega b_\ell + b^2_\ell)\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}\\
+&=\sum_{\omega,\ell}\theta^2_\omega\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}-2\sum_{\omega,\ell} \theta_\omega b_\ell \mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}+\sum_{\omega,\ell}b^2_\ell\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}}\\
 &\geq \sum_{\omega}\theta^2_\omega -2\sum_{\omega,\ell}\theta_\omega b_\ell+\sum_{\ell}b_\ell^2\left(n_{\ell}+n_{\ell+1}\right)\\
@@
 \[
-\FnormSize{}{\mz}^2=\sum_{\omega,\ell}z^2_{\omega,\ell}=\sum_{\omega,\ell}(-\theta_\omega+b_\ell)^2=L_{\text{total}}\sum_{\omega}\theta^2_\omega+d_{\text{total}}\sum_\ell b^2_\ell,
-\]
-where $L_{\text{total}}:=(L-1)$ and $d_{\text{total}}:=\prod_k d_k$.
+\FnormSize{}{\mz}^2=\sum_{\omega,\ell}z^2_{\omega,\ell}=\sum_{\omega,\ell}(-\theta_\omega+b_\ell)^2=(L-1)\sum_{\omega}\theta^2_\omega+\left(\prod_k d_k\right)\sum_\ell b^2_\ell.
+\]
 
@@
 \begin{align}\label{eq:negativeH}
-\max_{\mz\in\tS,\mz\neq \mathbf{0}}{\sum_{\omega,\ell} z^2_{\omega,\ell}\mathds{1}\{\{y_\omega=\ell \text{ or }\ell+1\}\} \over \FnormSize{}{\mz}^2}&\geq {\sum_{\omega}\theta^2_\omega+\min_\ell\left(n_{\ell}+n_{\ell+1}\right)\sum_{\ell}b_\ell^2\over L_{\text{total}} \sum_{\omega}\theta^2_\omega+d_{\text{total}}\sum_\ell b^2_\ell}
-\geq{\min_{\ell}(n_\ell+n_{\ell+1})\over (1+{\alpha^2\over c\Delta^2})d_{\text{total}}}\notag\\
-&\geq {2A_{\alpha,\beta,\Delta}\over 1+{\alpha^2\over c\Delta^2}}\quad \text{in high probability as $d_{\min}\to \infty$}. 
-\end{align}
-The second inequality in \eqref{eq:negativeH} is from the condition that 
-\[
-\sum_{\omega}\theta_\omega^2\in [0,\ \alpha^2 d_{\text{total}})\quad \text{and}\quad \sum_\ell b_\ell^2 \in[c L_{\text{total}}\Delta^2,\  L_{\text{total}}\beta^2],
-\]
-for some universal constant $c>0$. 
-The last inequality in \eqref{eq:negativeH} follows by applying the law of large numbers and the uniform bound $\min_{z_{\omega,\ell}}\mathbb{P}(y_{\omega}=\ell \text{ or }\ell+1|z_{\omega,\ell})\geq 2A_{\alpha,\beta,\Delta}$ to the empirical ratio:
-\[
-{\min_{\ell}(n_\ell+n_{\ell+1})\over d_{\text{total}}}\overset{p}{\to}\min_{\ell} \mathbb{P}(y_{\omega}=\ell \text{ or }\ell+1|z_{\omega,\ell}) \geq 2A_{\alpha,\beta,\Delta}, \quad \text{ in high probability as $d_{\min}\to \infty$}.
+\max_{\mz\in\tS,\mz\neq \mathbf{0}}{\sum_{\omega,\ell} z^2_{\omega,\ell}\mathds{1}_{\{y_\omega=\ell \text{ or }\ell+1\}} \over \FnormSize{}{\mz}^2}&\geq {\sum_{\omega}\theta^2_\omega+\min_\ell\left(n_{\ell}+n_{\ell+1}\right)\sum_{\ell}b_\ell^2\over(L-1)\sum_{\omega}\theta^2_\omega+\left(\prod_k d_k\right)\sum_\ell b^2_\ell}
+\geq{\min_{\ell}(n_\ell+n_{\ell+1})\over (L-1)\prod_k d_k}\notag\\&\geq {2A_{\alpha,\beta,\Delta}\over L-1}\quad \text{in high probability as $d_{\min}\to \infty$}. 
+\end{align}
+The second inequality in \eqref{eq:negativeH} is from the condition that \[(L-1),\quad{\prod_k d_k\over \min_\ell(n_\ell+n_{\ell+1})}>1\quad\text{ and } \quad \sum_{\omega}\theta_\omega^2,\quad\sum_\ell b_\ell^2 >0.\]
+The last inequality in \eqref{eq:negativeH} follows by applying the law of large numbers and the uniform bound $\min_{z_{\omega,\ell}}\mathbb{P}(y_{\omega}=\ell \text{ or }\ell+1|z_{\omega,\ell})\geq 2A_{\alpha,\beta,\Delta}$ to the ratio:
+\[
+{\min_{\ell}(n_\ell+n_{\ell+1})\over \prod_k d_k}\geq 2A_{\alpha,\beta,\Delta}, \quad \text{ in high probability as $d_{\min}\to \infty$}.
 \]
@@
 \begin{equation}\label{eq:boundH}
-\mz^T\mH\mz \leq- c'A_{\alpha,\beta,\Delta}D_{\alpha,\beta,\Delta}\FnormSize{}{\mz}^2,
-\end{equation}
-for some constant $c'>0$, therefore~\eqref{eq:HessionZ} is proved.
+\mz^T\mH\mz \leq-{c_3A_{\alpha,\beta,\Delta}D_{\alpha,\beta,\Delta}\FnormSize{}{\mz}^2\over L-1},
+\end{equation}
+and therefore~\eqref{eq:HessionZ} is proved.
 \end{enumerate}
@@
 \[
-\FnormSize{}{\hat \tZ-\trueZ}^2 \leq {c_1r^{K}_{\max}C^2_{\alpha,\beta,\Delta}\over A^2_{\alpha,\beta,\Delta}D^2_{\alpha,\beta,\Delta}}\left(L-1+\sum_k d_k\right).
+\FnormSize{}{\hat \tZ-\trueZ}^2 \leq {c_1r^{K-1}_{\max}C^2_{\alpha,\beta,\Delta}(L-1)^2\over A^2_{\alpha,\beta,\Delta}D^2_{\alpha,\beta,\Delta}}\left(L-1+\sum_k d_k\right).
 \]
@@
 We also compare the methods by their performance in predicting the median labels, $y_\omega^{\text{median}}=\min\{\ell\colon\mathbb{P}(y_\omega=\ell)\geq 0.5\}$.
-Under the latent variable model~\eqref{eq:latent} and Assumption~\ref{ass:link}, the median label is the quantized $\theta_\omega$ without noise; i.e.\ $y_\omega^{\text{median}}=\sum_\ell \mathds{1}\{\theta_\omega\in(b_{\ell-1},b_\ell]\}$.
+Under the latent variable model~\eqref{eq:latent} and Assumption~\ref{ass:link}, the median label is the quantized $\theta_\omega$ without noise; i.e.\ $y_\omega^{\text{median}}=\sum_\ell \mathds{1}_{\theta_\omega\in(b_{\ell-1},b_\ell]}$.
  We utilize the same simulation setting as in the earlier experiment.
@@
 The improved accuracy comes from the incorporation of multilinear low-rank structure, multi-level observations, and the ordinal structure.
-Interestingly, the median estimator tends to yield smaller MAD than the mode estimator, $\text{MAD}(\tY^{\text{median}}, \hat \tY^{\text{median}})\leq \text{MAD}(\tY^{\text{mode}}, \hat \tY^{\text{mode}})$  (Figures~\ref{fig:compare}a-b vs.\ Figures~\ref{fig:compare2}a-b), for the three multilevel methods ({\bf 1bit-sign-T}, {\bf Ordinal-M}, and {\bf Ordinal-T}).
-The mode estimator, on the other hand, tends to yield smaller MCR than the median estimator, $\text{MCR}(\tY^{\text{mode}}, \hat \tY^{\text{mode}})\leq\text{MCR}(\tY^{\text{median}}, \hat \tY^{\text{median}})$ (Figures~\ref{fig:compare}c-d vs.\ Figures~\ref{fig:compare2}c-d).
-This tendency is from the property that the median estimator $\hat y^{(\text{median})}_\omega$ minimizes $R_1(z)=\mathbb{E}_{y_\omega}|y_\omega-z|$, whereas the mode estimator $\hat y^{(\text{mode})}_\omega $ minimizes $R_2(z)=\mathbb{E}_{y_\omega}\mathds{1}\{y_\omega=z\}$.
+Interestingly, for the three multilevel methods ({\bf 1bit-sign-T}, {\bf Ordinal-M}, and {\bf Ordinal-T}), the median estimator tends to yield smaller MAD than the mode estimator, $\text{MAD}(\tY^{\text{median}}, \hat \tY^{\text{median}})\leq \text{MAD}(\tY^{\text{mode}}, \hat \tY^{\text{mode}})$  (Figures~\ref{fig:compare}a-b vs.\ Figures~\ref{fig:compare2}a-b).
+On the other hand, the mode estimator tends to yield smaller MCR than the median estimator, $\text{MCR}(\tY^{\text{mode}}, \hat \tY^{\text{mode}})\leq\text{MCR}(\tY^{\text{median}}, \hat \tY^{\text{median}})$ (Figures~\ref{fig:compare}c-d vs.\ Figures~\ref{fig:compare2}c-d).
+This tendency is from the property that the median estimator $\hat y^{(\text{median})}_\omega$ minimizes $R_1(z)=\mathbb{E}_{y_\omega}|y_\omega-z|$, whereas the mode estimator $\hat y^{(\text{mode})}_\omega $ minimizes $R_2(z)=\mathbb{E}_{y_\omega}\mathds{1}_{\{y_\omega=z\}}$.
 Here the expectation is over the categorical distribution of $y_\omega$ given parameters $\hat \Theta$ and $\hat \mb$.
@@
 The procedure is motivated from the higher-order extension of Principal Component Analysis (PCA) or Singular Value Decomposition (SVD).
-Recall that, in the matrix case, we perform clustering on an $m\times n$ (normalized) matrix $\mX$ based on the following procedure.
+Recall that, in the matrix case, we perform clustering on an $m\times n$ (normalized) matrix $X$ based on the following procedure.
 First, we factorize $X$ into
 \begin{equation}
-\mX = \mU\mSigma \mV^T,
-\end{equation}
-where $\mSigma$ is a diagonal matrix and $\mU,\mV$ are factor matrices with orthogonal columns.
-Second, we take each column of $\mV$ as a principal axis and each row in $\mU\mSigma$ as principal component.
-A subsequent multivariate clustering method (such as $K$-means) is then applied to the $m$ rows of $\mU\mSigma$.
+X = U\Sigma V^T,
+\end{equation}
+where $\Sigma$ is a diagonal matrix and $ U,V$ are factor matrices with orthogonal columns.
+Second, we take each column of $V$ as a principal axis and each row in $U\Sigma$ as principal component.
+A subsequent multivariate clustering method (such as $K$-means) is then applied to the $m$ rows of $U\Sigma$.
 
@@
 The elbow method suggests 11 clusters among the 68 nodes (see Figure~\ref{figure:elbow}).
-We find that the clustering successfully captures the spatial separation between brain regions (Table~\ref{table:clustering}).
+We find that the clustering captures the spatial separation between brain regions successfully (Table~\ref{table:clustering}).
 In particular, cluster I represents the connection between the left and right hemispheres, whereas clusters II-III represent the connection within each of the half brains (Figure~\ref{figure:brain image}).
@@
 \end{table}
-
 \end{document}

### 2020-07-30 paper/Camera ready/paper.tex
update

@@
 }\label{tab:compare}
+\vspace{-.2cm}
 \end{table*}
@@
 We address the challenge by proposing a cumulative link model that enjoys palindromic invariance~\cite{mccullagh1980regression}.
-A distinctive non-monotonic, phase-transition pattern is demonstrated (Fig~\ref{fig:finite}b).
-We prove that recovery from quantized tensors achieves equally good information-theoretical convergence as the continuous tensors.
-These results fills the gap between classical and non-classical (ordinal) observations, thereby greatly enriching the tensor-model literature.
+A distinctive non-monotonic, phase-transition pattern is demonstrated, as we show in Section~\ref{sec:experiment}.
+We prove that the recovery from quantized tensors achieves equally good information-theoretical convergence as the continuous tensors.
+These results fills the gap between classical and non-classical (ordinal) observations, thereby greatly enriching the tensor model literature.
 
@@
 The Frobenius norm of $\tY$ is defined as $\FnormSize{}{\tY}=\sum_\omega y^2_\omega$ and the infinity norm is defined as $\mnormSize{}{\tY}=\max_{\omega}|y_\omega|$.
-We use $\tY_{(k)}$ to denote the unfolded matrix of size $d_k$-by-$\prod_{i\neq k}d_k$, obtained by reshaping the tensor along the mode-$k$.
+We use $\tY_{(k)}$ to denote the unfolded matrix of size $d_k$-by-$\prod_{i\neq k}d_i$, obtained by reshaping the tensor along the mode $k\in[K]$.
 The Tucker rank of $\tY$ is defined as a length-$K$ vector $\mr = (r_1,\ldots,r_K)$, where $r_k$ is the rank of matrix $\tY_{(k)}$ for $k \in[K]$.
@@
 Let $\tY$ denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional data tensor.
-Suppose the entries of $\tY$ are ordinal-valued, and the observation space consists of $L$ ordered levels, denoted by $[L]:=\{1,\ldots,L\}$.
+Suppose the entries of $\tY$ are ordinal-valued, and the observation space consists of $L$ ordered levels, denoted by $[L]=\{1,\ldots,L\}$.
 We propose a cumulative link model for the ordinal tensor $\tY=\entry{y_\omega}\in[L]^{d_1\times \cdots\times d_K}$.
-Specifically, assume the entries $y_\omega$ are (conditionally) independently distributed with cumulative probabilities:
+Specifically, assume the entries $y_\omega$ are (conditionally) independently distributed with cumulative probabilities,
 \begin{equation}\label{eq:model}
@@
 This modeling choice is to respect the ordering structure among the categories.
-For example, if we choose the inverse link $f^{-1}(x)=\log {x\over 1-x}$ to be the log odds, then the model~\eqref{eq:model} implies linear spacing between the proportional odds:
+For example, if we choose the inverse link $f^{-1}(x)=\log {x\over 1-x}$ to be the log odds, then the model~\eqref{eq:model} implies a linear spacing between the proportional odds,
 \begin{equation}\label{eq:logodd}
@@
 for all tensor entries $y_\omega$.
-When there are only two categories in the observation space (e.g.
-binary tensors), the cumulative model~\eqref{eq:model} is equivalent to the usual binomial link model.
+When there are only two categories in the observation space (e.g., for binary tensors), the cumulative model~\eqref{eq:model} is equivalent to the usual binomial link model.
 In general, however, when the number of categories $L\geq 3$, the proportional odds assumption~\eqref{eq:logodd} is more parsimonious, in that, the ordered categories can be envisaged as contiguous intervals on the continuous scale, where the points of division are exactly $b_1<\cdots <b_{L-1}$.
-This interpretation will be made more explicit in the next section.
+This interpretation will be made explicit in the next section.
 
@@
 \end{align}
-for all $\omega\in[d_1]\times \cdots \times [d_k]$. Here, $\tY^*=\entry{y^*_\omega}$ is a latent continuous-valued tensor following an additive noise model:
+for all $\omega\in[d_1]\times \cdots \times [d_k]$. Here, $\tY^*=\entry{y^*_\omega}$ is a latent continuous-valued tensor following an additive noise model,
 \begin{align}\label{eq:latent}
@@
 \end{align}
-where $\tE=\entry{\varepsilon_\omega}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is a noise tensor with i.i.d.\ entries according to distribution $\mathbb{P}(\varepsilon)$.
+where $\tE=\entry{\varepsilon_\omega}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is a noise tensor with independent and identically distributed (i.i.d.) entries according to distribution $\mathbb{P}(\varepsilon)$.
 From the viewpoint of~\eqref{eq:latent}, the parameter tensor $\Theta$ can be interpreted as the latent signal tensor prior to contamination and quantization.
@@
 Other link functions are also possible, such as Laplace, Cauchy, etc~\cite{mccullagh1980regression}.
-All the models share the property that the ordered categories can be thought of as contiguous interval on some continuous scale.
+These latent variable models share the property that the ordered categories can be thought of as contiguous intervals on some continuous scale.
 We should point out that, although the latent-variable interpretation is incisive, our estimation procedure does not refer to the existence of $\tY^*$.
@@
 
-Clearly, the problem (P1) cannot be solved uniformly for all possible $\Theta$.
+Clearly, the problem (P1) cannot be solved uniformly for all possible $\Theta$ with no assumptions.
 We focus on a class of ``low-rank'' and ``flat'' signal tensors, which is a plausible assumption in practical applications~\cite{zhou2013tensor,bhaskar20151}.
-Specifically, we consider the parameter space:
+Specifically, we consider the parameter space,
 \begin{equation}\label{eq:space}
-\tP=\left\{\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K} \colon \text{rank}(\Theta)\leq \mr,\ \mnormSize{}{\Theta}\leq \alpha\right\}.
+\tP=\left\{\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K} \colon \text{rank}(\Theta)\leq \mr,\ \mnormSize{}{\Theta}\leq \alpha\right\},
 \end{equation}
@@
 The parameter tensor of our interest satisfies two constraints.
-The first is that $\Theta$ is a low-rank tensor, with $r_k=\tO(1)$ for all $k\in[K]$.
+The first is that $\Theta$ is a low-rank tensor, with $r_k=\tO(1)$ as $d_{\min}\to \infty$ for all $k\in[K]$.
 Equivalently, $\Theta$ admits the Tucker decomposition:
@@
 where $\tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is a core tensor, $\mM_k\in\mathbb{R}^{d_k\times r_k}$ are factor matrices with orthogonal columns, and $\times_k$ denotes the tensor-by-matrix multiplication~\cite{kolda2009tensor}.
-The Tucker low-rankness is popularly imposed in tensor data analysis, and is shown to provide a reasonable tradeoff between model complexity and model flexibility.
-Note that, unlike matrices, there are various notations of tensor low-rankness, such as CP rank~\cite{hitchcock1927expression} and train rank~\cite{oseledets2011tensor}.
-Some notation of low-rankness may lead to mathematically ill-posed optimization; for example, the best low CP-rank tensor approximation may not exist~\cite{de2008tensor}.
+The Tucker low-rankness is popularly imposed in tensor analysis, and the rank determines the tradeoff between model complexity and model flexibility.
+Note that, unlike matrices, there are various notions of tensor low-rankness, such as CP rank~\cite{hitchcock1927expression} and train rank~\cite{oseledets2011tensor}.
+Some notions of low-rankness may lead to mathematically ill-posed optimization; for example, the best low CP-rank tensor approximation may not exist~\cite{de2008tensor}.
 We choose Tucker representation for well-posedness of optimization and easy interpretation.
@@
 In view of~\eqref{eq:latent}, we refer to $\alpha$ as the signal level.
-The entry-wise bound assumption is a technical condition that avoids the degeneracy in probability estimation with ordinal observations.
-
+The boundedness assumption is a technical condition that avoids the degeneracy in probability estimation with ordinal observations.
+\vspace{-.2cm}
 \subsection{Problem 2: Tensor completion}
@@
 \end{align}
-We propose a rank-constrained maximum likelihood estimator (a.k.a.\ M-estimator) for $\Theta$:
+We propose a rank-constrained maximum likelihood estimator (a.k.a.\ M-estimator) for $\Theta$,
 \begin{align}\label{eq:estimator}
@@
 In practice, the cut-off points $\mb$ are unknown and should be jointly estimated with $\Theta$.
-We will focus on the theory and algorithm with known $\mb$ in the main paper.
+We will present the theory and algorithm with known $\mb$ in the main paper.
 The adaptation for unknown $\mb$ is addressed in Section~\ref{sec:algorithm} and the Supplement.
@@
 In view of equation~\eqref{eq:latent}, these quantities characterize the geometry including flatness and convexity of the latent noise distribution.
-Under the Assumption~\ref{ass:link}, all these quantities are strictly positive and independent of tensor dimension.
+Under Assumption~\ref{ass:link}, all these quantities are strictly positive and independent of tensor dimension.
 
@@
 Theorem~\ref{thm:rate} establishes the statistical convergence for the estimator~\eqref{eq:estimator}.
-In fact, the proof of this theorem (see the Supplement) shows that the same statistical rate holds, not only for the global optimizer~\eqref{eq:estimator}, but also for any local optimizer $\check \Theta$ in the level set $\{\check\Theta\in\tP: \logl(\check\Theta)\geq \logl(\trueT)\}$.
+In fact, the proof of this theorem (see the Supplement) shows that the same statistical rate holds, not only for the global optimizer~\eqref{eq:estimator}, but also for any local optimizer $\check \Theta$ in the level set $\{\check\Theta\in\tP\colon \logl(\check\Theta)\geq \logl(\trueT)\}$.
 This suggests that the local optimality itself is not necessarily a severe concern in our context, as long as the convergent objective is large enough.
 In Section ~\ref{sec:algorithm}, we perform empirical studies to assess the algorithmic stability.
+
+To gain insight into the bound~\eqref{eq:rate}, we consider a special setting with equal dimension in all modes, i.e., $d_1=\cdots=d_K=d$. In such a case, our bound \eqref{eq:rate} reduces to
+\begin{equation}\label{eq:ours}
+\text{MSE}(\hat \Theta, \trueT) \asymp d^{-(K-1)}, \quad \text{as}\ d\to \infty.
+\end{equation}
+Hence, our estimator achieves consistency with polynomial convergence rate.
+We compare the bound with existing literature.
+In the special case $L=2$, \citet{ghadermarzy2018learning} proposed a max-norm constrained estimator $\tilde{\Theta}$ with $\text{MSE}(\tilde{\Theta}, \trueT) \asymp  d^{-(K-1)/2}$.
+In contrast, our estimator converges at a rate of $d^{-(K-1)}$, which is substantially faster than theirs.
+This provides a positive answer to the open question posed in~\citet{ghadermarzy2018learning} whether the square root in the bound is removable.
+The improvement stems from the fact that we have used the exact low-rankness of $\Theta$, whereas the surrogate rank measure employed in~\citet{ghadermarzy2018learning} is scale-sensitive.
+
+Our bound also generalizes the previous results on ordinal matrices.
+The convergence rate for rank-constrained matrix estimation is $\tO(1/\sqrt{d})$~\citep{bhaskar2016probabilistic}, which fits into our special case when $K=2$.
+Furthermore, our result~\eqref{eq:rate} reveals that the convergence becomes favorable as the order of data tensor increases.
+Intuitively, the sample size for analyzing a data tensor is the number of entries, $\prod_k d_k$, and the number of free parameters is roughly on the order of $\sum_{k}d_k$, assuming $r_{\max}=\tO(1)$.
+A higher tensor order implies higher effective sample size per parameter, thus achieving a faster convergence rate in high dimensions.
 
@@
 \end{cor}
-To gain insight into these bounds, we consider a special setting with equal dimension in all modes, i.e., $d_1=\cdots=d_K=d$. In such a case, our bound \eqref{eq:rate} reduces to
-\begin{equation}\label{eq:ours}
-\text{MSE}(\hat \Theta, \trueT) \asymp d^{-(K-1)}, \quad \text{as}\ d\to \infty.
-\end{equation}
-Hence, our estimator achieves consistency with polynomial convergence rate.
-We compare the bound with existing literature.
-In the special case $L=2$, \citet{ghadermarzy2018learning} proposed a max-norm constrained estimator $\tilde{\Theta}$ with $\text{MSE}(\tilde{\Theta}, \trueT) \asymp  d^{-(K-1)/2}$.
-In contrast, our estimator converges at a rate of $d^{-(K-1)}$, which is substantially faster than theirs.
-This provides a positive answer to the open question posed in~\citet{ghadermarzy2018learning} whether the square root in the bound is removable.
-The improvement stems from utilizing the exact low-rankness of $\Theta$, whereas the surrogate rank measure employed in~\citet{ghadermarzy2018learning} is scale-sensitive.
-
-Our bound also generalizes the previous results on ordinal matrices.
-The convergence rate for rank-constrained matrix estimation was $\tO(1/\sqrt{d})$~\citep{bhaskar2016probabilistic}, which fits into our special case when $K=2$.
-Furthermore, our results~\eqref{eq:rate} and~\eqref{eq:KLrate} reveal that the convergence becomes favorable as the order of data tensor increases.
-Intuitively, the sample size for tensor data analysis is the number of entries, $\prod_k d_k$, and the number of free parameters is roughly on the order of $\sum_{k}d_k$, assuming $r_{\max}=\tO(1)$.
-A higher tensor order implies higher effective sample size per parameter, and thus exhibits a faster convergence rate in high dimensions.
 
 We next show the statistical optimality of our estimator $\hat \Theta$.
-The result is based on the information theory, and applies to all estimators in $\tP$, including but not limited to $\hat \Theta$ in~\eqref{eq:estimator}.
+The result is based on the information theory and applies to all estimators in $\tP$, including but not limited to $\hat \Theta$ in~\eqref{eq:estimator}.
 
@@
 We consider a general sampling procedure induced by $\Pi$.
-The recovery accuracy is assessed by the weighted squared error:
+The recovery accuracy is assessed by the weighted squared error,
 \begin{align}\label{eq:weighted}
@@
 In particular, tensor entries with higher sampling probabilities have more influence on the recovery accuracy, compared to the ones with lower sampling probabilities.
-\begin{rmk} If we assume each entry is sampled with strictly positive probability; i.e.\ there exits a constant $\mu> 0$ s.t.
+\begin{rmk} If we assume each entry is sampled with some strictly positive probability; i.e.,\ there exits a constant $\mu> 0$ such that
 \[
@@
 The optimization is a simple convex problem if any $K+1$ out of the $K+2$ blocks are fixed.
-We update one block at a time while holding others fixed, and alternate the optimization throughout the iteration.
+We update one block at a time while holding others fixed, and we alternate the optimization throughout the iteration.
 The convergence is guaranteed whenever $\logl$ is bounded from above, since the alternating procedure monotonically increases the objective.
@@
 In practice, the rank $\mr$ is hardly known and needs to be estimated from the data.
-We suggest to use Bayesian information criterion (BIC) and choose the rank that minimizes BIC; i.e.
+We use Bayesian information criterion (BIC) and choose the rank that minimizes BIC; i.e.,
 \begin{align}\label{eq:BIC}
@@
 In this section, we evaluate the empirical performance of our methods\footnote{Software package: \url{https://CRAN.R-project.org/package=tensorordinal}}.
-We investigate both the complete and the incomplete settings, and compare the recovery accuracy with other tensor-based methods.
+We investigate both the complete and the incomplete settings, and we compare the recovery accuracy with other tensor-based methods.
 Unless otherwise stated, the ordinal data tensors are generated from model~\eqref{eq:model} using standard probit link $f$.
 We consider the setting with $K=3$, $d_1=d_2=d_3=d$, and $r_1=r_2=r_3=r$.
-The parameter tensors are simulated based on~\eqref{eq:Tucker}, where the core tensor entries are i.i.d.\ drawn from $\tN(0,1)$, and the factors $\mM_k$ are uniformly sampled (with respect to Haar measure) from matrices with orthonormal columns.
+The parameter tensors are simulated based on~\eqref{eq:Tucker}, where the core tensor entries are i.i.d.\ drawn from $N(0,1)$, and the factors $\mM_k$ are uniformly sampled (with respect to Haar measure) from matrices with orthonormal columns.
 We set the cut-off points $b_\ell=f^{-1}(\ell/L)$ for $\ell\in[L]$, such that $f(b_\ell)$ are evenly spaced from 0 to 1.
@@
 \subsection{Comparison with alternative methods}
-Next, we compare our ordinal tensor method ({\bf Ordinal-T}) with three popular low-rank methods:
+Next, we compare our ordinal tensor method ({\bf Ordinal-T}) with three popular low-rank methods.
 \begin{itemize}[itemsep=0pt,topsep=0pt,leftmargin=*,partopsep=0pt]
@@
 The {\bf 1bit-T} is applied to $\tY$ in two ways.
-The first approach ({\bf 1bit-sign-T}) follows from~\citet{ghadermarzy2018learning} that transforms
+The first approach, denoted {\bf 1bit-sign-T}, follows from~\citet{ghadermarzy2018learning} that transforms
 $\tY$ to a binary tensor, by taking the entrywise sign of the mean-adjusted tensor, $\tY- |\Omega|^{-1}\sum_\omega y_\omega$.
-The second approach ({\bf 1bit-category-T}) transforms the order-3 ordinal tensor $\tY$ to an order-4 binary tensor, $\tY^{\sharp}=\entry{y^{\sharp}_{ijkl}}$, via dummy variable encoding; i.e., $y^{\sharp}_{ijk\ell}=\mathds{1}\{y_{ijk}=\ell\}$ for $\ell\in[L-1]$.
+The second approach, denoted {\bf 1bit-category-T}, transforms the order-3 ordinal tensor $\tY$ to an order-4 binary tensor, $\tY^{\sharp}=\entry{y^{\sharp}_{ijkl}}$, via dummy variable encoding; i.e., $y^{\sharp}_{ijk\ell}=\mathds{1}\{y_{ijk}=\ell\}$ for $\ell\in[L-1]$.
 We evaluate the methods by their capabilities in predicting the most likely labels, $y_\omega^{\text{mode}}=\arg\max_\ell\mathbb{P}(y_\omega=\ell)$.
@@
 We also compare the methods by their performance in predicting the median labels, $y_\omega^{\text{median}}=\min\{\ell\colon\mathbb{P}(y_\omega=\ell)\geq 0.5\}$.
-Under the latent variable model~\eqref{eq:latent} and Assumption~\ref{ass:link}, the median label is the quantized $\theta_\omega$ without noise; i.e.\ $y_\omega^{\text{median}}=\sum_\ell \mathds{1}\{\theta_\omega\in(b_{\ell-1},b_\ell]\}$.
+Under the latent variable model~\eqref{eq:latent} and Assumption~\ref{ass:link}, the median label is the quantized $\theta_\omega$ without noise; i.e.,\ $y_\omega^{\text{median}}=\sum_\ell \mathds{1}\{\theta_\omega\in(b_{\ell-1},b_\ell]\}$.
  We utilize the same simulation setting as in the earlier experiment.
@@
 This tendency is from the property that the median estimator $\hat y^{(\text{median})}_\omega$ minimizes $R_1(z)=\mathbb{E}_{y_\omega}|y_\omega-z|$, whereas the mode estimator $\hat y^{(\text{mode})}_\omega $ minimizes $R_2(z)=\mathbb{E}_{y_\omega}\mathds{1}\{y_\omega=z\}$.
-Here the expectation is over the categorical distribution of $y_\omega$ given parameters $\hat \Theta$ and $\hat \mb$.
+Here the expectation is taken over the testing data $y_\omega$, conditional on the training data and thus parameters $\hat \Theta$ and~$\hat \mb$.
 

### 2020-08-04 paper/Camera ready/supp.tex
clean up
corrected a typo

@@
 \begin{equation}\label{eq:totalKL}
-{1\over \text{Card}(\tX)-1}\sum_{\Theta \in\tX}\text{KL}(\mathbb{P}_{\Theta}, \mathbb{P}_{\mathbf{0}})\leq \varepsilon \log_2\left\{\text{Card}(\tX)-1 \right\}
+{1\over \text{Card}(\tX)-1}\sum_{\Theta \in\tX}\text{KL}(\mathbb{P}_{\Theta}, \mathbb{P}_{\mathbf{0}})\leq \varepsilon \log\left\{\text{Card}(\tX)-1 \right\}
 \end{equation}

### 2020-12-02 paper/arxiv_ordinal/arxiv.tex
Add files via upload

--- 2020-06-22 paper/Camera ready/ordinalTv6.tex
@@
-%%%%%%%% ICML 2020 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
-
-\documentclass{article}
-
-% Recommended, but optional, packages for figures and better typesetting:
-\usepackage{microtype}
+\documentclass[11pt]{article}
 \usepackage{graphicx}
 \usepackage{subfigure}
-\usepackage{booktabs} % for professional tables
 \usepackage{caption}
@@
 \usepackage{hyperref}
+\usepackage[margin=1in]{geometry}
 
@@
 % If accepted, instead use the following line for the camera-ready submission:
-\usepackage[accepted]{icml2020}
 
@@
 % Therefore, a short form for the running title is supplied here:
-\icmltitlerunning{Ordinal tensor denoising and completion}
-\usepackage{wrapfig}
-\usepackage{multirow}
 \usepackage{graphicx}
@@
 
+% Use fancyhdr package
+\usepackage{color}
+\usepackage{algorithm}
+\usepackage{algorithmic}
+\usepackage{natbib}
+\usepackage{forloop}
+
 \newcommand*{\KeepStyleUnderBrace}[1]{%f
@@
 \usepackage{amssymb}
+
 \usepackage{pifont}
@@
 
+\title{Tensor denoising and completion based on ordinal observations}
+\date{}
+\author{%
+Chanwoo Lee \\
+University of Wisconsin -- Madison\\
+\texttt{chanwoo.lee@wisc.edu} \\
+\and
+Miaoyan Wang \\
+University of Wisconsin -- Madison\\
+\texttt{miaoyan.wang@wisc.edu} \\
+}
+
 \begin{document}
-
-\twocolumn[
-\icmltitle{Tensor denoising and completion based on ordinal observations}
-
-% It is OKAY to include author information, even for blind
-% submissions: the style file will automatically remove it for you
-% unless you've provided the [accepted] option to the icml2020
-% package.
-
-% List of affiliations: The first argument should be a (short)
-% identifier you will use later to specify author affiliations
-% Academic affiliations should list Department, University, City, Region, Country
-% Industry affiliations should list Company, City, Region, Country
-
-% You can specify symbols, otherwise they are numbered in order.
-% Ideally, you should not use this facility. Affiliations will be numbered
-% in order of appearance and this is the preferred way.
-%\icmlsetsymbol{equal}{*}
-
-\begin{icmlauthorlist}
-\icmlauthor{Chanwoo Lee}{wi}
-\icmlauthor{Miaoyan Wang}{wi}
-\end{icmlauthorlist}
-
-\icmlaffiliation{wi}{Department of Statistics, University of Wisconsin-Madison, Wisconsin, USA}
-\icmlcorrespondingauthor{Miaoyan Wang}{miaoyan.wang@wisc.edu}
-
-% You may provide any keywords that you
-% find helpful for describing your paper; these are used to populate
-% the "keywords" metadata in the PDF but will not be shown in the document
-\icmlkeywords{Higher-order tensors, ordinal observation, tensor decomposition, tensor completion}
-
-\vskip 0.3in
-]
-
-% this must go after the closing bracket ] following \twocolumn[ ...
-
-% This command actually creates the footnote in the first column
-% listing the affiliations and the copyright notice.
-% The command takes one argument, which is text to display at the start of the footnote.
-% The \icmlEqualContribution command is standard text for equal contribution.
-% Remove it (just {}) if you do not need this facility.
-
-\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
-%\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.
-
+\maketitle
 \begin{abstract}
@@
 }\label{tab:compare}
+
 \end{table*}
@@
 The link function is assumed to satisfy:
-\vspace{-.2cm}
+
 \begin{enumerate}
@@
 In view of~\eqref{eq:latent}, we refer to $\alpha$ as the signal level.
-The entry-wise bound assumption is a technical condition that avoids the degeneracy in probability estimation with ordinal observations.
+The boundedness assumption is a technical condition that avoids the degeneracy in probability estimation with ordinal observations.
 
@@
 \centering
-\includegraphics[width=8cm]{algorithm.pdf}
+\includegraphics[width=10cm]{algorithm.pdf}
 \caption{Trajectory of objective function with various $d$ and $r$.}\label{fig:stability}
@@
 \begin{figure}[http]
-\includegraphics[width=8.3cm]{panel.pdf}
+\includegraphics[width=\textwidth]{panel.pdf}
 \caption{Empirical relationship between (relative) MSE versus (a) dimension $d$, (b) signal level $\alpha$, (c) observation fraction $\rho$, and (d) number of ordinal levels $L$.
@@
 \label{fig:finite}
-\vspace{-.3cm}
 \end{figure}
 
-\vspace{-.2cm}
 \subsection{Comparison with alternative methods}
-Next, we compare our ordinal tensor method ({\bf Ordinal-T}) with three popular low-rank methods:
+Next, we compare our ordinal tensor method ({\bf Ordinal-T}) with three popular low-rank methods.
 \begin{itemize}[itemsep=0pt,topsep=0pt,leftmargin=*,partopsep=0pt]
@@
 \begin{figure}[http]
-\includegraphics[width=8.3cm]{panel2.pdf}
+\includegraphics[width=\textwidth]{panel2.pdf}
 \caption{Performance comparison for predicting most likely labels.
@@
 }\label{fig:compare}
-\vspace{-.4cm}
 \end{figure}
@@
 \begin{figure}[http]
-\includegraphics[width=8.3cm]{panel3.pdf}
+\includegraphics[width=\textwidth]{panel3.pdf}
 \caption{Performance comparison for predicting median labels.
@@
 (b, d) Prediction error versus the number of ordinal levels $L$, when $\rho=0.8.$}\label{fig:compare2}
-\vspace{-.4cm}
 \end{figure}
@@
 This tendency is from the property that the median estimator $\hat y^{(\text{median})}_\omega$ minimizes $R_1(z)=\mathbb{E}_{y_\omega}|y_\omega-z|$, whereas the mode estimator $\hat y^{(\text{mode})}_\omega $ minimizes $R_2(z)=\mathbb{E}_{y_\omega}\mathds{1}\{y_\omega=z\}$.
-Here the expectation is over the categorical distribution of $y_\omega$ given parameters $\hat \Theta$ and $\hat \mb$.
-
-\section{Data Applications}
+Here the expectation is taken over the testing data $y_\omega$, conditional on the training data and thus parameters $\hat \Theta$ and~$\hat \mb$.
+
+\section{Data Applications}\label{sec:dataapplication}
 We apply our ordinal tensor method to two real-world datasets.
@@
 \includegraphics[width=1\textwidth]{HCPtable.pdf}
-\vspace{-.7cm}
 \caption{Node clusters in the HCP analysis.
@@
 }  \label{table:clustering}
-\vspace{-.4cm}
 \end{table*}
@@
 \centering
-\includegraphics[width = .5\textwidth]{HCP.pdf}
+\includegraphics[width = .8\textwidth]{HCP.pdf}
 \caption{Top three clusters in the HCP analysis.
@@
 }  \label{figure:brain image}
-\vspace{-.2cm}
 \end{figure}
@@
 Our method outperforms the others, especially in MAD.
-
 \subsection{InCarMusic recommendation system}
@@
 \begin{table}[http]
-\resizebox{.48\textwidth}{!}{
-\begin{tabular}{c|c|c}
+\resizebox{\textwidth}{!}{
+\begin{tabular}{c|c|c|c|c|c}
 \Xhline{2\arrayrulewidth}
-\multicolumn{3}{c}{Human Connectome Project (HCP) dataset}\\
+\multicolumn{3}{c|}{Human Connectome Project (HCP) dataset}&\multicolumn{3}{c}{InCarMusic dataset}\\
 \Xhline{2\arrayrulewidth}
-Method                    & MAD & MCR\\
+Method                    & MAD & MCR&Method                    & MAD & MCR\\
 \hline
-Ordinal-T (ours) & 0.1607 (0.0005)& 0.1606 (0.0005)\\
- Continuous-T&0.2530 (0.0002)&0.1599 (0.0002)\\
- 1bit-sign-T&0.3566 (0.0010)&0.1563 (0.0010)\\
- \Xhline{2\arrayrulewidth}
-\multicolumn{3}{c}{InCarMusic dataset}\\
-\Xhline{2\arrayrulewidth}
-Method                    & MAD & MCR\\
-\hline
-Ordinal-T (ours) & 1.37 (0.039) &  0.59 (0.009)\\
-Continuous-T & 2.39 (0.152)&0.94 (0.027)\\
-1bit-sign-T&1.39 (0.003)& 0.81 (0.005)\\
+Ordinal-T (ours) & 0.1607 (0.0005)& 0.1606 (0.0005)&Ordinal-T (ours) & 1.37 (0.039) &  0.59 (0.009)\\
+ Continuous-T&0.2530 (0.0002)&0.1599 (0.0002)&Continuous-T & 2.39 (0.152)&0.94 (0.027)\\
+ 1bit-sign-T&0.3566 (0.0010)&0.1563 (0.0010)&1bit-sign-T&1.39 (0.003)& 0.81 (0.005)\\
 \hline
@@
 \bibliography{tensor_wang}
-\bibliographystyle{icml2020}
-
-\clearpage
+\bibliographystyle{apalike}
+
 \appendix
-\onecolumn
-
-\icmltitle{Supplements for ``Tensor denoising and completion based on \\ ordinal observations''}
 
@@
 \begin{align}
-&\mH_\omega(\ell,\ell)+ \mH_\omega(\ell,\ell-1)+\mH_\omega(\ell,\ell+1)\\
-=&{\partial^2 \tL_\tY\over \partial z^2_{\omega,\ell}}+{\partial^2 \tL_\tY\over \partial z_{\omega,\ell}z_{\omega,\ell+1}}+{\partial^2 \tL_\tY\over \partial z_{\omega,\ell-1}z_{\omega,\ell}}\\
-=&\mathds{1}\{y_\omega=\ell\}{\partial\over \partial z}\left({\dot{f}(z) \over f(z)-f(z')}\right)\Bigg|_{(z,\ z')=(z_{\omega,\ell},\ z_{\omega,\ell-1})}  -\mathds{1}\{y_\omega=\ell+1\}{\partial\over \partial z}\left({\dot{f}(z) \over f(z)-f(z')}\right)\Bigg|_{(z'',\ z)=(z_{\omega,\ell+1},\ z_{\omega,\ell})}\\
-\leq &-D_{\alpha,\beta,\Delta}\mathds{1}\{y_\omega = \ell \text{ or } \ell+1\}.
+\mH_\omega(\ell,\ell)+ \mH_\omega(\ell,\ell-1)+\mH_\omega(\ell,\ell+1)&
+={\partial^2 \tL_\tY\over \partial z^2_{\omega,\ell}}+{\partial^2 \tL_\tY\over \partial z_{\omega,\ell}z_{\omega,\ell+1}}+{\partial^2 \tL_\tY\over \partial z_{\omega,\ell-1}z_{\omega,\ell}}\\
+&=\mathds{1}\{y_\omega=\ell\}{\partial\over \partial z}\left({\dot{f}(z) \over f(z)-f(z')}\right)\Bigg|_{(z,\ z')=(z_{\omega,\ell},\ z_{\omega,\ell-1})} \\& -\mathds{1}\{y_\omega=\ell+1\}{\partial\over \partial z}\left({\dot{f}(z) \over f(z)-f(z')}\right)\Bigg|_{(z'',\ z)=(z_{\omega,\ell+1},\ z_{\omega,\ell})}\\
+&\leq -D_{\alpha,\beta,\Delta}\mathds{1}\{y_\omega = \ell \text{ or } \ell+1\}.
 \end{align}
@@
 &=\sum_{\omega,\ell}(\theta^2_\omega-2\theta_\omega b_\ell + b^2_\ell)\mathds{1}\{y_\omega=\ell \text{ or }\ell+1\}\\
-&=\sum_{\omega,\ell}\theta^2_\omega\mathds{1}\{y_\omega=\ell \text{ or }\ell+1\}-2\sum_{\omega,\ell} \theta_\omega b_\ell \mathds{1}\{y_\omega=\ell \text{ or }\ell+1\}+\sum_{\omega,\ell}b^2_\ell\mathds{1}\{y_\omega=\ell \text{ or }\ell+1\}\\
 &\geq \sum_{\omega}\theta^2_\omega -2\sum_{\omega,\ell}\theta_\omega b_\ell+\sum_{\ell}b_\ell^2\left(n_{\ell}+n_{\ell+1}\right)\\
@@
 \end{document}
-
-% This document was modified from the file originally made available by
-% Pat Langley and Andrea Danyluk for ICML-2K. This version was created
-% by Iain Murray in 2018, and modified by Alexandre Bouchard in
-% 2019 and 2020. Previous contributors include Dan Roy, Lise Getoor and Tobias
-% Scheffer, which was slightly modified from the 2010 version by
-% Thorsten Joachims & Johannes Fuernkranz, slightly modified from the
-% 2009 version by Kiri Wagstaff and Sam Roweis's 2008 version, which is
-% slightly modified from Prasad Tadepalli's 2007 version which is a
-% lightly changed version of the previous year's version by Andrew
-% Moore, which was in turn edited from those of Kristian Kersting and
-% Codrina Lauth. Alex Smola contributed to the algorithmic style files.

### 2020-12-02 paper/arxiv_ordinal/arxiv.tex
Add files via upload
update

@@
 \usepackage{caption}
-% hyperref makes hyperlinks in the resulting PDF.
-% If your build breaks (sometimes temporarily if a hyperlink spans a page)
-% please comment out the following usepackage line and replace
-% \usepackage{icml2020} with \usepackage[nohyperref]{icml2020} above.
 \usepackage{hyperref}
@@
 
-% Attempt to make hyperref and algorithmic work together better:
 \newcommand{\theHalgorithm}{\arabic{algorithm}}
-
-% Use the following line for the initial blind version submitted for review:
-
-% If accepted, instead use the following line for the camera-ready submission:
-
-% The \icmltitle you define below is probably too long as a header.
-% Therefore, a short form for the running title is supplied here:
 \usepackage{graphicx}
-%\usepackage[utf8]{inputenc} % allow utf-8 input
-%\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
+
 \usepackage{hyperref}       % hyperlinks
 \usepackage{url}            % simple URL typesetting
-%\usepackage{booktabs}       % professional-quality tables
+
 \usepackage{amsmath,amssymb}
 \usepackage{amsthm}    % blackboard math symbols
-%\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-%\usepackage{microtype}      % microtypography
+
 \usepackage{bm}
-%\usepackage{subfig}
-%\usepackage[english]{babel}
-%\usepackage{algorithm}
-%\usepackage{appendix}
 \usepackage{mathtools}
@@
 
+\allowdisplaybreaks
 \usepackage{dsfont}
-%\usepackage{algpseudocode,algorithm}
-%\algnewcommand\algorithmicinput{\textbf{Input:}}
-%\algnewcommand\algorithmicoutput{\textbf{Output:}}
-%\algnewcommand\INPUT{\item[\algorithmicinput]}
-%\algnewcommand\OUTPUT{\item[\algorithmicoutput]}
-%\DeclareMathOperator*{\minimize}{minimize}
 
@@
 We demonstrate the outperformance of our approach over previous methods on the tasks of clustering and collaborative filtering.
-
 \end{abstract}
@@
 
-\begin{table*}[ht]
-\centerline{\includegraphics[width=1\textwidth]{compare.pdf}}
+\begin{table}[ht]
+\resizebox{\textwidth}{!}{
+\begin{tabular}{l|ccc}
+&\cite{bhaskar2016probabilistic}&\cite{ghadermarzy2018learning} &This paper\\
+\hline 
+Higher-order tensors ($K\geq 3$) & \xmark&\cmark& \cmark\\
+\hline
+Multi-level categories ($L\geq 3$)& \cmark& \xmark&\cmark\\
+ \hline
+ Error rate for tensor denoising&$d^{-1}$ for $K=2$& $d^{-(K-1)/2}$&$d^{-(K-1)}$\\
+\hline
+Optimality guarantee under low-rank models& unknown & \xmark & \cmark\\
+\hline
+Sample complexity for tensor completion&$d^K$& $Kd$&$Kd$
+\end{tabular}
+}
 \caption{Comparison with previous work.
@@
 }\label{tab:compare}
-
-\end{table*}
-
-{\bf Our contributions.} We are among the first to establish the recovery theory for signal tensors and quantization operators simultaneously from a limited number of highly discrete entries.
+\end{table}
+
+{\bf Our contributions.} We establish the recovery theory for signal tensors and quantization operators simultaneously from a limited number of highly discrete entries.
 Our main contributions are summarized in Table~\ref{tab:compare}.
@@
 
-{\bf Related work.} Our work is connected to non-Gaussian tensor decomposition.
-Existing work focuses exclusively on univariate observations such as binary- or continuous-valued entries~\citep{wang2018learning,ghadermarzy2018learning}.
+Our work is connected to non-Gaussian tensor decomposition.
+Existing work focuses exclusively on univariate observations such as binary- or continuous-valued entries~\citep{wang2020learning,ghadermarzy2018learning}.
 The problem of ordinal/quantized tensor is fundamentally more challenging than previously well-studied tensors for two reasons: (a) the entries do not belong to exponential family distribution, and (b) the observation contains much less information, because neither the underlying signal nor the quantization operator is known.
@@
 \subsection{Latent-variable interpretation}\label{sec:latent}
-
-The ordinal tensor model~\eqref{eq:model} with certain types of link $f$ has the equivalent representation as an $L$-level quantization model on $\tY=\entry{y_\omega}$:
+The ordinal tensor model~\eqref{eq:model} with certain types of link $f$ has the equivalent representation as an $L$-level quantization model.
+
+Specifically, the entries of tensor $\tY=\entry{y_\omega}$ are modeled from the following generative process, 
 \begin{align}\label{eq:quantization}
-%y_\omega=\sum_{\ell\in[L]} \ell \mathds{1}\{y^*_\omega \in(b_{\ell-1},b_\ell]\},\ \forall \ \omega\in[d_1]\times \cdots \times [d_k],
-%\end{equation}
 y_\omega&=
@@
 for all $\omega\in[d_1]\times \cdots \times [d_k]$. Here, $\tY^*=\entry{y^*_\omega}$ is a latent continuous-valued tensor following an additive noise model,
-\begin{align}\label{eq:latent}
+\begin{equation}\label{eq:latent}
 \KeepStyleUnderBrace{\tY^*}_{\text{latent continuous-valued tensor}}=\KeepStyleUnderBrace{\Theta}_{\text{signal tensor}}+\KeepStyleUnderBrace{\tE}_{\text{i.i.d.\ noise}},
-\end{align}
+\end{equation}
 where $\tE=\entry{\varepsilon_\omega}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is a noise tensor with independent and identically distributed (i.i.d.) entries according to distribution $\mathbb{P}(\varepsilon)$.
@@
 \begin{assumption}\label{ass:link}
-The link function is assumed to satisfy:
-
-\begin{enumerate}
-\item $f(\theta)$ is strictly increasing and twice-differentiable in $\theta\in \mathbb{R}$.
-\item $f'(\theta)$ is strictly log-concave and symmetric with respect to $\theta=0$.
+The link function $f$ is assumed to satisfy:
+\begin{enumerate}[label=\textup{(\roman*)}]
+\item The function $f(\theta)$ is strictly increasing and twice-differentiable in $\theta\in \mathbb{R}$.
+\item The derivative $f'(\theta)$ is strictly log-concave and symmetric with respect to $\theta=0$.
 \end{enumerate}
@@
 \subsection{Problem 1: Tensor denoising}~\label{sec:denoising}
-The first question we aim to address is tensor denoising:
-
-(P1) Given the quantization process induced by $f$ and the cut-off points $\mb$, how accurately can we estimate the latent signal tensor $\Theta$ from the ordinal observation $\tY$?
+The first question we aim to address is tensor denoising:\\
+
+(P1) Given the quantization process induced by $f$ and the cut-off points $\mb$, how accurately can we estimate the latent signal tensor $\Theta$ from the ordinal observation $\tY$?\\
 
@@
 Let $\Omega\subset[d_1]\times \cdots\times[d_K]$ denote the set of observed indices.
-The second question we aim to address is stated as follows:
-
-(P2) Given an incomplete set of ordinal observations $\{y_{\omega}\}_{\omega\in\Omega}$, how many sampled entries do we need to consistently recover $\Theta$ based on the model~\eqref{eq:model}?
+The second question we aim to address is stated as follows:\\
+
+(P2) Given an incomplete set of ordinal observations $\{y_{\omega}\}_{\omega\in\Omega}$, how many sampled entries do we need to consistently recover $\Theta$ based on the model~\eqref{eq:model}?\\
 
@@
 The log-likelihood associated with the observed entries is
-\begin{align}\label{eq:objective}
-\logl(\Theta, \mb)&=\sum_{\omega\in\Omega}\sum_{\ell\in[L]} \Big\{\mathds{1}\{y_\omega=\ell\} \log \big[f(b_\ell-\theta_\omega)- \notag \\
-&\quad \quad \quad \quad \quad \quad \quad  f(b_{\ell-1}-\theta_\omega)\big]\Big\}.
-\end{align}
+\begin{equation}\label{eq:objective}
+\logl(\Theta, \mb)=\sum_{\omega\in\Omega}\sum_{\ell\in[L]} \Big\{\mathds{1}\{y_\omega=\ell\} \log \big[f(b_\ell-\theta_\omega)- f(b_{\ell-1}-\theta_\omega)\big]\Big\}.
+\end{equation}
 We propose a rank-constrained maximum likelihood estimator (a.k.a.\ M-estimator) for $\Theta$,
@@
 We define a few key quantities that will be used in our theory. Let $g_\ell=f(\theta+b_\ell)-f(\theta+b_{\ell-1})$ for all $\ell\in[L]$, and
-\begin{align}\label{eq:regular}
-A_\alpha &= \min_{\ell\in[L], |\theta|\leq \alpha} g_\ell(\theta),\quad  U_\alpha=\max_{\ell\in[L], |\theta|\leq \alpha } {|\dot{g}_\ell(\theta)|\over g_\ell(\theta)},\\
- L_\alpha &= \min_{\ell\in[L], |\theta|\leq \alpha }\left[{\dot{g}_\ell^2 (\theta)\over g_\ell^2(\theta)} -{\ddot{g}_\ell(\theta)\over g_\ell(\theta)}\right],
-\end{align}
+\begin{equation}\label{eq:regular}
+A_\alpha = \min_{\ell\in[L], |\theta|\leq \alpha} g_\ell(\theta),\quad  U_\alpha=\max_{\ell\in[L], |\theta|\leq \alpha } {|\dot{g}_\ell(\theta)|\over g_\ell(\theta)},\quad L_\alpha = \min_{\ell\in[L], |\theta|\leq \alpha }\left[{\dot{g}_\ell^2 (\theta)\over g_\ell^2(\theta)} -{\ddot{g}_\ell(\theta)\over g_\ell(\theta)}\right],
+\end{equation}
 where $\dot{g}(\theta)=dg(\theta)/d\theta$, and $\alpha$ is the entrywise bound of $\Theta$.
@@
 Then, under the model~\eqref{eq:model},
-\begin{align}\label{eq:lower}
-\inf_{\hat \Theta }\sup_{\trueT\in\tP} \mathbb{P}\Big\{& \text{MSE}(\hat \Theta, \trueT) \\
-& \geq c\min\left( \alpha^2, \ { Cr_{\max}d_{\max} \over \prod_k d_k} \right) \Big\} \geq {1\over 8},
-\end{align}
+\begin{equation}\label{eq:lower}
+\inf_{\hat \Theta }\sup_{\trueT\in\tP} \mathbb{P}\Big\{ \textup{MSE}(\hat \Theta, \trueT) \geq c\min\left( \alpha^2, \ { Cr_{\max}d_{\max} \over \prod_k d_k} \right) \Big\} \geq {1\over 8},
+\end{equation}
 where $C=C(\alpha, L, f,\mb)>0$ and $c>0$ are constants independent of tensor dimension and the rank.
@@
 The recovery accuracy is assessed by the weighted squared error,
-\begin{align}\label{eq:weighted}
-\PiFnormSize{}{\Theta-\hat \Theta}^2&\stackrel{\text{def}}{=}
-{1\over |\Omega|}\mathbb{E}_{\Omega\sim \Pi}\FnormSize{}{\Theta-\hat \Theta}^2\notag \\
-&=\sum_{\omega\in[d_1]\times \cdots \times [d_K]} \pi_{\omega}(\Theta_{\omega}-\hat \Theta_{\omega})^2.
-\end{align}
+\begin{equation}\label{eq:weighted}
+\PiFnormSize{}{\Theta-\hat \Theta}^2\stackrel{\text{def}}{=}
+{1\over |\Omega|}\mathbb{E}_{\Omega\sim \Pi}\FnormSize{}{\Theta-\hat \Theta}^2=\sum_{\omega\in[d_1]\times \cdots \times [d_K]} \pi_{\omega}(\Theta_{\omega}-\hat \Theta_{\omega})^2.
+\end{equation}
 
@@
 
-\begin{figure}[H]
+\begin{figure}[ht]
 \centering
-\includegraphics[width=10cm]{algorithm.pdf}
+\includegraphics[width=13cm]{algorithm.pdf}
 \caption{Trajectory of objective function with various $d$ and $r$.}\label{fig:stability}
@@
 Next, we compare our ordinal tensor method ({\bf Ordinal-T}) with three popular low-rank methods.
-\begin{itemize}[itemsep=0pt,topsep=0pt,leftmargin=*,partopsep=0pt]
+\begin{itemize}
 \item Continuous tensor decomposition ({\bf Continuous-T})~\cite{acar2010scalable} is a low-rank approximation method based on classical Tucker model.
@@
 
-\begin{figure}[http]
+\begin{figure}[ht]
 \includegraphics[width=\textwidth]{panel2.pdf}
@@
 
-\begin{figure}[http]
-\includegraphics[width=\textwidth]{panel3.pdf}
+\begin{figure}[ht]
+\includegraphics[width=1\textwidth]{panel3.pdf}
 \caption{Performance comparison for predicting median labels.
@@
 We apply our ordinal tensor method to two real-world datasets.
-In the first application, we use our model to analyze an ordinal tensor consisting of structural connectivities among 68 brain regions for 136 individuals from Human Connectome Project (HCP)~\citep{geddes2016human}.
+In the first application, we use our model to analyze an ordinal tensor consisting of structural connectivities among 68 brain regions for 136 individuals from Human Connectome Project (HCP)~\citep{van2013wu}.
 In the second application, we perform tensor completion to an ordinal dataset with missing values.
 The data tensor records the ratings of 139 songs on a scale of 1 to 5 from 42 users on 26 contexts~\citep{baltrunas2011incarmusic}.
-
-\begin{table*}[http]
-\includegraphics[width=1\textwidth]{HCPtable.pdf}
-\caption{Node clusters in the HCP analysis.
-The first alphabet in the node name indicates the left (L) or right (R) hemisphere.
-The number in the parentheses indicates the node count in each cluster.
-}  \label{table:clustering}
-\end{table*}
 
@@
 
-\begin{figure}[http]
+\begin{figure}[ht]
 \centering
@@
 
-We acompare the goodness-of-fit of various tensor methods on the HPC data.
+\begin{table}[ht]
+\center
+\begin{sc}
+\resizebox{\columnwidth}{!}{
+\begin{tabular}{|c|c|c|c|}
+\hline
+Cluster     & \multicolumn{3}{c|}{I}\\ \hline
+Brain nodes & \multicolumn{3}{l|}{\begin{tabular}[c]{@{}l@{}} l.frontalpole, l.temporalpole, l.medialorbitofrontal,  l.cuneus, l.parahippocampal, l.lingual,            \\
+r.frontalpole, r.temporalpole,  r.medialorbitofrontal,  r.cuneus,  r.parahippocampal \\
+\end{tabular}} \\
+\hline
+Cluster & \multicolumn{3}{c|}{II}\\ \hline
+Brain nodes & \multicolumn{3}{l|}{\begin{tabular}[c]{@{}l@{}}
+l.caudalmiddlefrontal,  l.inferiorparietal,  l.insula,  l.isthmuscingulate, l.lateraloccipital(2), \\
+l.parsopercularis, l.parstriangularis, l.postcentral, l.precuneus, l.superiorfrontal, l.superiortemporal(3)
+\end{tabular}} \\ 
+\hline
+Cluster & \multicolumn{3}{c|}{III}\\ \hline
+Brain nodes & \multicolumn{3}{l|}{\begin{tabular}[c]{@{}l@{}}
+r.caudalmiddlefrontal, r.inferiorparietal, r.insula, r.isthmuscingulate, r.lateraloccipital(2), r.lingual, \\
+r.parsopercularis, r.parstriangularis, r.postcentral, r.precentral, r.precuneus, r.superiorfrontal(3), \\
+r.superiorparietal,  r.superiortemporal(3)
+\end{tabular}} \\ 
+\hline
+Cluster      & IV    & V  & VI\\
+\hline
+Brain nodes & \begin{tabular}[c]{@{}c@{}}l.supramarginal(4)\end{tabular}&
+\begin{tabular}[c]{@{}c@{}}l.inferiortemporal(3)\end{tabular}&
+\begin{tabular}[c]{@{}c@{}}l.middletemporal(3)\end{tabular}\\
+\hline
+Cluster & VII  & VIII & VIIII\\
+\hline
+Brain nodes &
+\begin{tabular}[c]{@{}c@{}}r.supramarginal(4)\end{tabular}&
+\begin{tabular}[c]{@{}c@{}}r.inferiortemporal(3)\end{tabular}&
+\begin{tabular}[c]{@{}c@{}}r.middletemporal(3)\end{tabular}\\
+\hline
+Cluster &X & \multicolumn{2}{c|}{XI}  \\
+\hline
+Brain nodes &
+\begin{tabular}[c]{@{}c@{}}l.superiorfrontal(2)\end{tabular}&
+ \multicolumn{2}{c|}{l.precentral,  l.superiorparietal}\\
+ \hline
+\end{tabular}
+}
+\end{sc}
+\caption{Node clusters in the HCP analysis.
+The first alphabet in the node name indicates the left (L) or right (R) hemisphere.
+The number in the parentheses indicates the node count in each cluster.
+}  \label{table:clustering}
+\end{table}
+
+We compare the goodness-of-fit of various tensor methods on the HPC data.
 Table~\ref{table:CV} summarizes the prediction error via 5-fold stratified cross-validation averaged over 10 runs.
@@
 
-\begin{table}[http]
+\begin{table}[ht]
 \resizebox{\textwidth}{!}{
@@
 
-\section{Conclusions}
-We have developed a low-rank tensor estimation method based on possibly incomplete, ordinal-valued observations.
-A sharp error bound is established, and we demonstrate the outperformance of our approach compared to other methods.
-The work unlocks several directions of future research.
-One interesting question would be the inference problem, i.e.,\ to assess the uncertainty of the obtained estimates and the imputation.
-Other directions include the trade-off between (non)convex optimization and statistical efficiency.
-While we have provided numerical evidence for the success of nonconvex approach, the full landscape of the optimization remains open.
-The interplay between computational efficiency and statistical accuracy in general tensor problems warrants future research.
-
-\newpage
-\section*{Acknowledgements}
-This research is supported in part by NSF grant DMS-1915978 and Wisconsin Alumni Research Foundation.
-
-\bibliography{tensor_wang}
-\bibliographystyle{apalike}
-
-\appendix
-
-\section{Proofs}\label{sec:proof}
-Here, we provide proofs of the theoretical results presented in Sections~\ref{sec:theory}.
-
-\subsection{Estimation error for tensor denoising}\label{sec:proofMSE}
-\begin{proof}[Proof of Theorem~\ref{thm:rate}]
-We suppress the subscript $\Omega$ in the proof, because the tensor denoising assumes complete observation $\Omega=[d_1]\times \cdots \times [d_K]$.
-It follows from the expression of $\flogl(\Theta)$ that
-\begin{align}\label{eq:property}
-{\partial \flogl\over \partial \theta_\omega}&=\sum_{\ell\in[L]}\mathds{1}\{y_{\omega}=\ell\}
-{\dot{g}_\ell(\theta_\omega)\over g_\ell(\theta_\omega)},\notag\\
-{\partial^2 \flogl\over \partial \theta_\omega^2}&=\sum_{\ell\in[L]}\mathds{1}\{y_\omega=\ell\}{\ddot{g}_\ell(\theta_\omega)g_\ell(\theta_\omega)-\dot{g}^2_\ell(\theta_\omega)\over g^2_\ell(\theta_\omega)}\ \text{and}\quad
-{\partial^2 \flogl\over \partial \theta_\omega \theta_\omega'}=0\ \text{if}\ \omega\neq \omega',
-\end{align}
-for all $\omega\in[d_1]\times \cdots \times [d_K]$.
-Define $d_{\text{total}}=\prod_k d_k$.
-Let $\fplogl\in\mathbb{R}^{d_1\times\cdots\times d_K}$ denote the tensor of gradient with respect to $\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}$, and $\fpplogl$ the corresponding Hession matrix of size $d_\text{total}$-by-$d_{\text{total}}$.
-Here, $\Vec(\cdot)$ denotes the operation that turns a tensor into a vector.
-By~\eqref{eq:property}, $\fpplogl$ is a diagonal matrix.
-Recall that
-\begin{equation}\label{eq:bound}
-U_\alpha=\max_{\ell\in[L],|\theta|\leq \alpha}{|\dot{g}_\ell(\theta)|\over g_\ell(\theta)}>0 \quad \text{and}\quad
-L_\alpha=\min_{\ell\in[L],|\theta|\leq \alpha} {\dot{g}^2_\ell(\theta)-\ddot{g}_\ell(\theta)g_\ell(\theta)\over g^2_\ell(\theta)}>0.
-\end{equation}
-Therefore, the entries in $\fplogl$ are upper bounded in magnitude by $U_\alpha>0$, and all diagonal entries in $\fpplogl$ are upper bounded by $-L_{\alpha}<0$.
-
-By the second-order Taylor's expansion of $\flogl(\Theta)$ around $\trueT$, we obtain
-\begin{equation}\label{eq:taylor}
-\flogl(\Theta)=\flogl(\trueT)+\langle\Vec(\fplogl(\trueT)),\ \Vec(\Theta-\trueT)\rangle+{1\over 2}\Vec(\Theta-\trueT)^T\fpplogl(\check\Theta)\Vec(\Theta-\trueT),
-\end{equation}
-where $\check\Theta=\gamma\trueT+(1-\gamma)\Theta$ for some $\gamma\in[0,1]$, and $\fpplogl(\check\Theta)$ denotes the $d_{\text{total}}$-by-$d_\text{total}$ Hession matrix evaluated at $\check\Theta$.
-
-We first bound the linear term in~\eqref{eq:taylor}. Note that, by Lemma~\ref{lem:inq},
-\begin{equation}\label{eq:linear}
-|\langle\Vec(\fplogl(\trueT), \Vec(\Theta-\trueT)  \rangle|\leq \snormSize{}{\fplogl(\trueT)} \nnormSize{}{\Theta-\trueT},
-\end{equation}
-where $\snormSize{}{\cdot}$ denotes the tensor spectral norm and $\nnormSize{}{\cdot}$ denotes the tensor nuclear norm. Define
-\[
-s_\omega={\partial \tL_\tY\over \partial \theta_\omega}\Big|_{\Theta=\trueT} \;\; \textrm{ for all } \; \omega\in[d_1]\times\cdots\times [d_K].
-\]
-Based on~\eqref{eq:property} and the definition of $U_\alpha$, $\fplogl(\trueT)=\entry{s_{\omega}}$ is a random tensor whose entries are independently distributed satisfying
-\begin{equation}\label{eq:norm}
-\mathbb{E}(s_\omega)=0,\quad |s_\omega|\leq U_\alpha, \quad \text{for all }\omega\in[d_1]\times \cdots \times [d_K].
-\end{equation}
-By lemma~\ref{lem:noisytensor}, with probability at least $1-\exp(-C_1 \sum_kd_k)$, we have
-\begin{equation}\label{eq:normrandom}
-\snormSize{}{\fplogl(\trueT)} \leq C_2 U_\alpha\sqrt{\sum_k d_k},
-\end{equation}
-where $C_1, C_2$ are two positive constants that depend only on $K$.
-Furthermore, note that $\text{rank}(\Theta)\leq \mr$, $\text{rank}(\trueT)\leq \mr$, so $\text{rank}(\Theta-\trueT)\leq 2\mr$.
-By lemma~\ref{lem:nuclear}, $\nnormSize{}{\Theta-\trueT}\leq (2r_{\max})^{K-1\over 2}\FnormSize{}{\Theta-\trueT}$.
-Combining~\eqref{eq:linear}, \eqref{eq:norm} and \eqref{eq:normrandom}, we have that, with probability at least $1-\exp(-C_1 \sum_kd_k)$,
-\begin{equation}\label{eq:linearconclusion}
-|\langle \Vec(\fplogl(\trueT)), \Vec(\Theta-\trueT)  \rangle | \leq C_2 U_\alpha  \sqrt{r_{\max}^{K-1} \sum_k d_k}  \FnormSize{}{\Theta-\trueT}.
-\end{equation}
-
-We next bound the quadratic term in \eqref{eq:taylor}. Note that
-\begin{align}\label{eq:quadratic}
- \Vec(\Theta-\trueT)^T \fpplogl(\check{\Theta})\Vec(\Theta-\trueT)&=\sum_\omega \left( {\partial^2\tL_{\tY}\over \partial \theta^2_\omega} \Big|_{\Theta=\check\Theta} \right)(\theta_\omega-\theta_{{\text{true}},\omega})^2 \nonumber \\
-&\leq - L_\alpha\sum_{\omega}(\Theta_{\omega}-\Theta_{\text{true},\omega})^2 \nonumber \\
-&=-L_\alpha\FnormSize{}{\Theta-\trueT}^2,
-\end{align}
-where the second line comes from the fact that  $\mnormSize{}{\check\Theta}\leq \alpha$ and the definition of $L_\alpha$.
-
-Combining~\eqref{eq:taylor}, \eqref{eq:linearconclusion} and~\eqref{eq:quadratic}, we have that, for all $\Theta\in\tP$, with probability at least $1-\exp(-C_1 \sum_kd_k)$,
-\[
-\tL_\tY(\Theta)\leq \tL_{\tY}(\trueT)+C_2U_\alpha  \left(r_{\max}^{K-1}\sum_k d_k\right)^{1/2}  \FnormSize{}{\Theta-\trueT}-{L_\alpha\over 2}\FnormSize{}{\Theta-\trueT}^2.
-\]
-In particular, the above inequality also holds for $\hat \Theta\in\tP$. Therefore,
-\[
-\tL_\tY(\hat \Theta)\leq \tL_{\tY}(\trueT)+C_2U_\alpha \left(r_{\max}^{K-1}\sum_k d_k\right)^{1/2}  \FnormSize{}{\hat \Theta-\trueT}-{L_\alpha\over 2} \FnormSize{}{\hat \Theta-\trueT}^2.
-\]
-Since $\hat \Theta=\arg\max_{\Theta\in\tP}\tL_\tY(\Theta)$, $\tL_\tY(\hat \Theta)-\tL_{\tY}(\trueT)\geq 0$, which gives
-\[
-C_2U_\alpha \left(r_{\max}^{K-1}\sum_k d_k\right)^{1/2}  \FnormSize{}{\hat \Theta-\trueT}-{L_\alpha\over 2}\FnormSize{}{\hat \Theta-\trueT}^2\geq 0.
-\]
-Henceforth,
-\[
-{1\over \sqrt{\prod_k d_k}} \FnormSize{}{\hat \Theta-\trueT}\leq {2C_2U_\alpha \sqrt{r_{\max}^{K-1}\sum_k d_k}\over L_\alpha \sqrt{\prod_k d_k}}={2C_2U_\alpha r_{\max}^{(K-1)/2}\over L_\alpha} \sqrt{ \sum_k d_k \over \prod_k d_k}.
-\]
-This completes the proof.
-\end{proof}
-
-\begin{proof}[Proof of Corollary~\ref{cor:prediction}]
-The result follows immediately from Theorem~\ref{thm:rate} and Lemma~\ref{lem:KL}.
-\end{proof}
-
-\begin{proof}[Proof of Theorem~\ref{thm:minimax}]
-
-Let $d_{\text{total}}=\prod_{k\in[K]}d_k$, and $\gamma\in[0,1]$ be a constant to be specified later.
-Our strategy is to construct a finite set of tensors $\tX=\{\Theta_i\colon i=1,\ldots \}\subset \tP$ satisfying the properties of (i)-(iv) in Lemma~\ref{lem:construction}.
-By Lemma~\ref{lem:construction}, such a subset of tensors exist.
-For any tensor  $\Theta\in\tX$, let $\mathbb{P}_{\Theta}$ denote the distribution of $\tY|\Theta$, where $\tY$ is the ordinal tensor.
-In particular, $\mathbb{P}_{\mathbf{0}}$ is the distribution of $\tY$ induced by the zero parameter tensor $\mathbf{0}$, i.e., the distribution of $\tY$ conditional on the parameter tensor $\Theta=\mathbf{0}$.
-Based on the Remark for Lemma~\ref{lem:KL}, we have
-\begin{equation}\label{eq:KLbound1}
-\mathrm{KL}(\mathbb{P}_{\Theta}|| \mathbb{P}_{\mathbf{0}})\leq C \FnormSize{}{\Theta}^2,
-\end{equation}
-where $C={(4L-6) \dot{f}^2(0)\over  A_\alpha}>0$ is a constant independent of the tensor dimension and rank.
-Combining the inequality~\eqref{eq:KLbound1} with property (iii) of $\tX$, we have
-\begin{equation}\label{eq:KLbound}
-\text{KL}(\mathbb{P}_{\Theta}||\mathbb{P}_{\mathbf{0}})\leq \gamma^2 r_{\max} d_{\max}.
-\end{equation}
-From~\eqref{eq:KLbound} and the property (i), we deduce that the condition
-\begin{equation}\label{eq:totalKL}
-{1\over \text{Card}(\tX)-1}\sum_{\Theta \in\tX}\text{KL}(\mathbb{P}_{\Theta}, \mathbb{P}_{\mathbf{0}})\leq \varepsilon \log\left\{\text{Card}(\tX)-1 \right\}
-\end{equation}
-holds for any $ \varepsilon \geq 0$ when $\gamma\in[0,1]$ is chosen to be sufficiently small depending on $\varepsilon$, e.g., $\gamma \leq \sqrt{\varepsilon\log2\over8}$.
-By applying Lemma~\ref{lem:Tsybakov} to~\eqref{eq:totalKL}, and in view of the property (iv), we obtain that
-\begin{equation}\label{eq:final}
-\inf_{\hat \Theta}\sup_{\trueT\in \tX}\mathbb{P}\left(\FnormSize{}{\hat \Theta- \trueT}\geq  {\gamma\over 8} \min\left\{ \alpha\sqrt{d_{\text{total}}}, C^{-1/2}\sqrt{ r_{\max}d_{\max}}\right\} \right)\geq {1\over 2}\left(1-2\varepsilon-\sqrt{16 \varepsilon \over r_{\max}d_{\max}\log2}\right).
-\end{equation}
-Note that $\text{Loss}(\hat \Theta, \trueT)=\FnormSize{}{\hat \Theta- \trueT}^2/d_{\text{total}}$ and $\tX\subset \tP$.
-By taking $\varepsilon=1/10$ and $\gamma=1/11$, we conclude from~\eqref{eq:final} that
-\begin{equation}\label{eq:prob}
-\inf_{\hat \Theta}\sup_{\trueT\in \tP}\mathbb{P}\left(\text{Loss}(\hat \Theta, \trueT)\geq c\min\left \{ \alpha^2,  {C^{-1}r_{\max}d_{\max}\over d_{\text{total}}}\right \}\right)\geq {1\over 2}\left({4\over 5}- \sqrt{1.6\over r_{\max}d_{\max}\log2} \right)\geq {1\over 8},
-\end{equation}
-where $c = {1 \over 88^2}$ and the last inequality comes from the condition for $d_{\text{max}}$.
-This completes the proof.
-\end{proof}
-
-\subsection{Sample complexity for tensor completion}
-
-\begin{proof}[Proof of Theorem~\ref{thm:completion}]
-
-For notational convenience, we use $\MFnormSize{}{\Theta}=\sum_{\omega\in\Omega}\Theta^2_\omega$ to denote the sum of squared entries over the observed set $\Omega$, for a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$.
-
-Following a similar argument as in the proof of Theorem~\ref{thm:rate}, we have
-\begin{equation}\label{eq:Taylor2}
-\logl(\Theta)=\logl(\trueT)+\langle\Vec(\plogl),\ \Vec(\Theta-\trueT)\rangle+{1\over 2}\Vec(\Theta-\trueT)^T\pplogl(\check\Theta)\Vec(\Theta-\trueT),
-\end{equation}
-where
-\begin{enumerate}
-\item $\plogl$ is a $d_1\times\cdots\times d_K$ tensor with $|\Omega|$ nonzero entries, and each entry is upper bounded by $U_\alpha>0$.
-\item $\pplogl$ is a diagonal matrix of size $d_{\text{total}}$-by-$d_{\text{total}}$ with $|\Omega|$ nonzero entries, and each entry is upper bounded by $-L_{\alpha}<0$.
-\end{enumerate}
-
-Similar to~\eqref{eq:linear} and~\eqref{eq:quadratic}, we have
-\begin{equation}\label{eq:linear2}
-|\langle\Vec(\plogl),\ \Vec(\Theta-\trueT)\rangle|\leq C_2U_\alpha \sqrt{r_{\max}^{K-1}\sum_k d_k}\MFnormSize{}{\Theta-\trueT}
-\end{equation}
-and
-\begin{equation}\label{eq:quadratic2}
-\Vec(\Theta-\trueT)^T\fpplogl(\check\Theta)\Vec(\Theta-\trueT)\leq -L_\alpha \MFnormSize{}{\Theta-\trueT}^2.
-\end{equation}
-
-Combining~\eqref{eq:Taylor2}-\eqref{eq:quadratic2} with the fact that $\logl(\hat \Theta)\geq \logl(\trueT)$, we have
-\begin{equation}\label{eq:sample}
-\MFnormSize{}{\hat \Theta-\trueT}\leq {2C_2U_\alpha  r_{\max}^{(K-1)/2} \over L_\alpha} \sqrt{\sum_kd_k},
-\end{equation}
-with probability at least $1-\exp(-C_1 \sum_k d_k)$.
-Lastly, we invoke the result regarding the closeness of $\Theta$ to its sampled version $\Theta_{\Omega}$, under the entrywise bound condition.
-Note that $\mnormSize{}{\hat\Theta-\trueT}\leq 2\alpha$ and $\text{rank}(\hat \Theta-\trueT)\leq 2\mr$.
-By Lemma~\ref{lem:Mnormbound}, $\anormSize{}{\hat \Theta-\trueT}\leq 2^{(3K-1)/2} \alpha \left({\prod r_k \over r_{\max}}\right)^{3/2}$.
-Therefore, the condition in Lemma~\ref{lem:convexity} holds with $\beta=2^{(3K-1)/2}\alpha \left({\prod r_k \over r_{\max}}\right)^{3/2}$.
-Applying Lemma~\ref{lem:convexity} to~\eqref{eq:sample} gives
-\begin{align}
- \PiFnormSize{}{\hat \Theta-\trueT}^2&\leq {1\over m}\MFnormSize{}{\hat \Theta-\trueT}^2+c\beta\sqrt{\sum_k d_k\over |\Omega|}\\
- &\leq {C_2  r^{K-1}_{\max}} {\sum_k d_k \over |\Omega|}+C_1 \alpha r_{\max}^{3(K-1)/2}\sqrt{\sum_kd_k\over |\Omega|},
-\end{align}
-with probability at least $1-\exp(-{\sum_kd_k\over \sum_k \log d_k})$ over the sampled set $\Omega$. Here $C_1, C_2>0$ are two constants independent of the tensor dimension and rank. Therefore,
-\[
- \PiFnormSize{}{\hat \Theta-\trueT}^2\to 0,\quad \text{as}\quad {|\Omega|\over \sum_kd_k}\to \infty,
-\]
-provided that $r_{\max}=O(1)$.
-\end{proof}
-
-\subsection{Convexity of the log-likelihood function}\label{sec:proofconvexity}
-\begin{thm}\label{thm:convexity}
-Define the function
-\begin{align}\label{eq:function}
- \logl(\Theta, \mb)&=\sum_{\omega\in\Omega}\sum_{\ell\in[L]} \big(\mathds{1}\{y_\omega=\ell\}\log \left[f(b_\ell-\theta_\omega)-  f(b_{\ell-1}-\theta_\omega)\right]\big),
- \end{align}
-where $f(\cdot)$ satisfies Assumption~\ref{ass:link}. Then, $\logl(\Theta, \mb)$ is concave in $(\Theta,\mb)$.
-\end{thm}
-
-\begin{proof}
-Define $d_{\text{total}}=\prod_k d_k$.
-By abuse of notation, we use $(\Theta,\mb)$ to denote the length-$(d_{\text{total}}+L-1)$-vector collecting all parameters together.
-Let us denote a bivariate function
-\begin{align}
-\lambda\colon &\mathbb{R}^2\mapsto \mathbb{R}\\
-(u,v)&\mapsto\lambda(u,v) = \log \big[f(u)-  f(v)\big].
-\end{align}
-It suffices to show that $\lambda(u,v)$ is concave in $(u,v)$ where $u>v$.
-
-Suppose that the claim holds (which we will prove in the next paragraph). Based on~\eqref{eq:function}, $u,v$ are both linear functions of $(\Theta,\mb)$:
-\[
-u = \ma_1^T(\Theta,\mb), \quad v = \ma_2^T(\Theta,\mb), \quad \text{ for some}\ \ma_1,\ma_2\in \mathbb{R}^{d_{\text{total}}+L-1}.
-\]
-Then, $\lambda(u,v) = \lambda(\ma_1^T(\Theta,\mb),\ \ma_2^T(\Theta,\mb))$ is concave in $(\Theta,\mb)$ by the definition of concavity.
-Therefore, we can conclude that $\logl(\Theta, \mb) $ is concave in $(\Theta,\mb)$ because $\logl(\Theta, \mb)$ is the sum of $\lambda(u,v)$.
-
-Now, we prove the concavity of $\lambda(u,v)$. Note that
-\begin{equation}
-\lambda(u,v) = \log\big[f(u)-f(v)\big]=\log\big[\int\mathds{1}_{[u,v]}(x)f'(x)dx\big],
-\end{equation}
-where $\mathds{1}_{[u,v]}$ is an indicator function that equals 1 in the interval $[u,v]$, and 0 elsewhere.
-Furthermore, $\mathds{1}_{[u,v]}(x)$ is log-concave in $(u,v,x)$, and by Assumption~\ref{ass:link}, $f'(x)$ is log-concave in $x$.
-It follows that $\mathds{1}_{[u,v]}(x)f'(x)$
-is a log-concave in $(u,v,x)$. By Lemma~\ref{lem:lossconvexity}, we conclude that $\lambda(u,v)$ is concave in $(u,v)$ where $u>v$.
-\end{proof}
-
-\begin{lem}[Corollary 3.5 in~\citet{brascamp2002extensions}]\label{lem:lossconvexity}
-Let $F(x,y)\colon \mathbb{R}^{m+n}\rightarrow \mathbb{R}$ be an integrable function where $x\in \mathbb{R}^{m},y\in \mathbb{R}^n$. Let
-\[
-G(x) = \int_{\mathbb{R}^n}F(x,y)dy.
-\]
-If $F(x,y)$ is log concave in $(x,y)$, then $G(x)$ is log concave in $x$.
-\end{lem}
-
-\subsection{Auxiliary lemmas}\label{sec:lemma}
-This section collects lemmas that are useful for the proofs of the main theorems.
-
-\begin{defn}[Atomic M-norm~\citep{ghadermarzy2019near}]
-Define $T_{\pm}=\{\tT\in\{\pm 1 \}^{d_1\times \cdots \times d_K}\colon \text{rank}(\tT)=1\}$. The atomic M-norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
-\begin{align}
-\anormSize{}{\Theta}&=\inf\{t>0\colon \Theta\in t\text{conv}(T_{\pm})\}\\
-&=\inf\left\{\sum_{\tX\in T_{\pm}}c_{\tX} \colon\ \Theta=\sum_{\tX\in T_{\pm}} c_{\tX}\tX, \ c_{\tX}>0\right\}.
-\end{align}
-\end{defn}
-
-\begin{defn}[Spectral norm~\citep{lim2005singular}]
-The spectral norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
-\[
-\snormSize{}{\Theta}=\sup\left\{\langle \Theta, \mx_1\otimes \cdots \otimes \mx_K\rangle \colon \vnormSize{}{\mx_k}=1,\ \mx_k\in\mathbb{R}^{d_k},\ \text{for all}\ k\in[K]\right\}.
-\]
-\end{defn}
-
-\begin{defn}[Nuclear norm~\citep{friedland2018nuclear}]
-The nuclear norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
-\[
-\nnormSize{}{\Theta}=\inf
-\left\{
-\sum_{i\in[r]}|\lambda_i|\colon \Theta=\sum_{i=1}^r \lambda_i\mx^{(i)}_1\otimes \cdots \otimes \mx^{(i)}_K,\ \vnormSize{}{\mx^{(i)}_k}=1,\ \mx^{(i)}_k\in\mathbb{R}^{d_k},\ \text{for all}\ k\in[K],\ i\in[r]
-\right\},
-\]
-where the infimum is taken over all $r\in\mathbb{N}$ and $\vnormSize{}{\mx^{(i)}_k}=1$ for all $i\in[r]$ and $k\in[K]$.
-\end{defn}
-
-\begin{lem}[M-norm and infinity norm~\citep{ghadermarzy2019near}]\label{lem:Mnormbound}
-Let $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ be an order-$K$, rank-$(r_1,\ldots,r_K)$ tensor. Then
-\[
-\mnormSize{}{\Theta}\leq \anormSize{}{\Theta}\leq \left(\prod_k r_k \over r_{\max}\right)^{3\over 2} \mnormSize{}{\Theta}.
-\]
-\end{lem}
-
-\begin{lem}[Nuclear norm and F-norm] \label{lem:nuclear}
-Let $\tA\in\mathbb{R}^{d_1\times\cdots\times d_K}$ be an order-$K$ tensor with Tucker $\text{rank}(\tA)=(r_1,\ldots,r_K)$. Then
-\[
-\nnormSize{}{\tA} \leq \sqrt{\prod_k r_k\over \max_k r_k} \FnormSize{}{\tA},
-\]
-where $\nnormSize{}{\cdot}$ denotes the nuclear norm of the tensor.
-\end{lem}
-
-\begin{proof}
-Without loss of generality, suppose $r_1=\min_k r_k$.
-Let $\tA_{(k)}$ denote the mode-$k$ matricization of $\tA$ for all $k\in[K]$.
-By \citet[Corollary 4.11]{wang2017operator}, and the invariance relationship between a tensor and its Tucker core~\citep[Section 6]{jiang2017tensor}, we have
-\begin{equation}\label{eq:norminequality}
-\nnormSize{}{\tA} \leq \sqrt{\prod_{k\geq 2} r_k \over \max_{k\geq 2} r_k} \nnormSize{}{\tA_{(1)}},
-\end{equation}
-where $\tA_{(1)}$ is a $d_1$-by-$\prod_{k\geq 2}d_k$ matrix with matrix rank $r_1$.
-Furthermore, the relationship between the matrix norms implies that $\nnormSize{}{\tA_{(1)}}\leq \sqrt{r_1}\FnormSize{}{\tA_{(1)}}=\sqrt{r_1}\FnormSize{}{\tA}$.
-Combining this fact with the inequality~\eqref{eq:norminequality} yields the final claim.
-\end{proof}
-
-\begin{lem} \label{lem:inq}
-Let $\tA, \; \tB$ be two order-$K$ tensors of the same dimension. Then
-\[
-|\langle \tA,\tB\rangle| \leq \snormSize{}{\tA}   \nnormSize{}{\tB}.
-\]
-\end{lem}
-
-\begin{proof}
-By~\citet[Proposition 3.1]{friedland2018nuclear}, there exists a nuclear norm decomposition of $\tB$, such that
-\[
-\tB=\sum_{r} \lambda_r \ma^{(1)}_r\otimes \cdots\otimes \ma^{(K)}_r,\quad \ma_r^{(k)}\in\mathbf{S}^{d_k-1}(\mathbb{R}),\quad \text{for all }k\in[K],
-\]
-and $\nnormSize{}{\tB}=\sum_{r}|\lambda_r|$. Henceforth we have
-\begin{align*}
-|\langle \tA,\tB\rangle|&=| \langle \tA, \sum_{r} \lambda_r \ma^{(1)}_r\otimes \cdots\otimes \ma^{(K)}_r \rangle|\leq \sum_r |\lambda_r| |\langle \tA, \ma^{(1)}_r \otimes \cdots\otimes \ma^{(K)}_r \rangle|\\
-&\leq \sum_{r}|\lambda_r| \snormSize{}{\tA}= \snormSize{}{\tA}\nnormSize{}{\tB},
-\end{align*}
-which completes the proof.
-\end{proof}
-
-The following lemma provides the bound on the spectral norm of random tensors. The result was firstly presented in~\citet{nguyen2015tensor}, and we adopt the version from~\citet{tomioka2014spectral}.
-\begin{lem}[Spectral norm of random tensors~\citep{tomioka2014spectral}]\label{lem:tensor}
-Suppose that $\tS=\entry{s_{\omega}}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is an order-$K$ tensor whose entries are independent random variables that satisfy
-\[
-\mathbb{E}(s_{\omega})=0,\quad \text{and} \quad\mathbb{E}(e^{ts_{\omega}})\leq e^{t^2L^2/2}.
-\]
-Then the spectral norm $\snormSize{}{\tS}$ satisfies that,
-\[
-\snormSize{}{\tS}\leq \sqrt{{8L^2} \log (12K) \sum_k d_k +\log (2/\delta)},
-\]
-with probability at least $1-\delta$.
-\end{lem}
-
-\begin{lem} \label{lem:noisytensor}
-Suppose that $\tS=\entry{s_{\omega}}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is an order-$K$ tensor whose entries are independent random variables that satisfy
-\[
-\mathbb{E}(s_{\omega})=0,\quad \text{and}\quad |s_{\omega}|\leq U.
-\]
-Then we have
-\[
-\mathbb{P}\left(\snormSize{}{\tS}\geq C_2 U\sqrt{\sum_k d_k} \right)\leq \exp\left(-C_1  \log K \sum_k d_k\right)
-\]
-where $C_1>0$ is an absolute constant, and $C_2>0$ is a constant that depends only on $K$.
-\end{lem}
-
-\begin{proof}  Note that the random variable $U^{-1}s_{\omega}$ is zero-mean and supported on $[-1,1]$. Therefore, $U^{-1}s_{\omega}$ is sub-Gaussian with parameter ${1-(-1)\over 2}=1$, i.e.
-\[
-\mathbb{E}(U^{-1}s_{\omega})=0,\quad \text{and}\quad \mathbb{E}(e^{tU^{-1}s_{\omega}})\leq e^{t^2/2}.
-\]
-It follows from Lemma~\ref{lem:tensor} that, with probability at least $1-\delta$,
-\[
-\snormSize{}{U^{-1}\tS}\leq \sqrt{\left(c_0\log K+c_1\right) \sum_k d_k +\log (2/\delta)},
-\]
-where $c_0, c_1>0$ are two absolute constants. Taking $\delta=\exp (-C_1\log K \sum_k d_k)$ yields the final claim, where $C_2=c_0\log K+c_1+1>0$ is another constant.
-\end{proof}
-
-\begin{lem}\label{lem:KLentry} Let $X,\; Y$ be two discrete random variables taking values on $L$ possible categories, with point mass probabilities $\{p_\ell\}_{\ell\in[L]}$ and $\{q_\ell\}_{\ell\in[L]}$, respectively.
-Suppose $p_\ell$, $q_\ell>0$ for all $\ell\in[L]$.
-Then, the Kullback-Leibler (KL) divergence satisfies that
-\[
-\mathrm{KL}(X||Y)\stackrel{\text{def}}{=}-\sum_{\ell\in[L]}\mathbb{P}_X(\ell)\log\left\{{\mathbb{P}_Y(\ell)\over \mathbb{P}_X(\ell)}\right\} \leq \sum_{\ell \in [L]}{(p_\ell-q_\ell)^2 \over q_\ell}.
-\]
-\end{lem}
-\begin{proof} Using the fact $\log x\leq x-1$ for $x>0$, we have that
-\begin{align}\label{eq:KL}
-\text{KL}(X||Y)&=\sum_{\ell\in[L]}p_\ell\log{p_\ell\over q_\ell}\\
-&\leq \sum_{\ell\in[L]} {p_\ell\over q_\ell}(p_\ell-q_\ell)\\
-&=\sum_{\ell\in[L]} \left({p_\ell\over q_\ell }- 1\right)(p_\ell-q_\ell)+ \sum_{\ell\in[L]} (p_\ell-q_\ell).
-\end{align}
-Note that $\sum_{\ell\in[L]}(p_\ell-q_\ell)=0$. Therefore,
-\[
-\text{KL}(X||Y)\leq \sum_{\ell\in[L]}\left( {p_\ell\over q_\ell}-1\right)\left(p_\ell-q_\ell\right)=\sum_{\ell\in[L]}{(p_\ell-q_\ell)^2\over q_\ell}.
-\]
-\end{proof}
-
-\begin{lem}[KL divergence and F-norm]~\label{lem:KL}
-Let $\tY\in[L]^{d_1\times \cdots \times d_K}$ be an ordinal tensor generated from the model~\eqref{eq:model} with the link function $f$ and parameter tensor $\Theta$.
-Let $\mathbb{P}_{\Theta}$ denote the joint categorical distribution of $\tY|\Theta$ induced by the parameter tensor $\Theta$, where $\mnormSize{}{\Theta}\leq \alpha$.
-Define
-\begin{equation}\label{eq:ass}
-A_\alpha=\min_{\ell\in[L], |\theta|\leq \alpha}\left[f(b_\ell-\theta)-f(b_{\ell-1}-\theta)\right].
-\end{equation}
-Then, for any two tensors $\Theta,\; \Theta^*$ in the parameter spaces, we have
-\[
-\mathrm{KL}(\mathbb{P}_{\Theta}|| \mathbb{P}_{\Theta^*})\leq {2(2L-3)\over A_\alpha} \dot{f}^2(0)\FnormSize{}{\Theta-\Theta^*}^2.
-\]
-\end{lem}
-\begin{proof} Suppose that the distribution over the ordinal tensor $\tY=\entry{y_\omega}$ is induced by $\Theta=\entry{\theta_\omega}$. Then, based on the generative model~\eqref{eq:model},
-\[
-\mathbb{P}(y_\omega=\ell | \theta_\omega)=f(b_{\ell}-\theta_\omega)-f(b_{\ell-1}-\theta_\omega),
-\]
-for all $\ell\in[L]$ and $\omega\in[d_1]\times \cdots \times [d_K]$.
-For notational convenience, we suppress the subscribe in $\theta_\omega$ and simply write $\theta$ (and respectively, $\theta^*$).
-Based on Lemma~\ref{lem:KLentry} and Taylor expansion,
-\begin{align}
-\text{KL}(\theta|| \theta^*) & \leq \sum_{\ell\in[L]} {\left[f(b_{\ell}-\theta)-f(b_{\ell-1}-\theta)-f(b_{\ell}-\theta^*)+f(b_{\ell-1}-\theta^*)\right]^2\over f(b_{\ell}-\theta^*)-f(b_{\ell-1}-\theta^*)}\\
- &\leq \sum_{\ell=2}^{L-1} {\left[\dot{f}(b_\ell-\eta_\ell)-\dot{f}(b_{\ell-1}-\eta_{\ell-1})\right]^2 \over f(b_\ell-\theta^*)-f(b_{\ell-1}-\theta^*)} \left(\theta-\theta^*\right)^2+{\dot{f}^2(b_1-\eta_1) \over f(b_1-\theta^*)} (\theta-\theta^*)^2\\
- &\quad \quad\quad \quad  +{\dot{f}^2(b_{L-1}-\eta_{L-1})\over 1-f(b_{L-1}-\theta^*)}(\theta-\theta^*)^2,
-\end{align}
-where $\eta_\ell$ and $\eta_{\ell-1}$ fall between $\theta$ and $\theta^*$. Therefore,
-\begin{equation}\label{eq:entrywise}
-\text{KL}(\theta|| \theta^*) \leq \left({4(L-2)\over A_\alpha}+ {2\over A_\alpha}\right)\dot{f}^2(0)(\theta-\theta^*)^2={2(2L-3)\over A_\alpha} \dot{f}^2(0)(\theta-\theta^*)^2,
-\end{equation}
-where we have used Taylor expansion, the bound~\eqref{eq:ass}, and the fact that $\dot{f}(\cdot)$ peaks at zero for an unimodal and symmetric function.
-Now summing~\eqref{eq:entrywise} over the index set $\omega\in[d_1]\times \cdots \times [d_K]$ gives
-\[
-\text{KL}(\mathbb{P}_{\Theta}|| \mathbb{P}_{\Theta^*}) =\sum_{\omega\in[d_1]\times \cdots \times[d_K]} \text{KL}(\theta_\omega || \theta^*_\omega) \leq {2(2L-3)\over A_\alpha} \dot{f}^2(0)\FnormSize{}{\Theta-\Theta^*}^2.
-\]
-\end{proof}
-
-\begin{rmk} In particular, let $\mathbb{P}_{\bf{0}}$ denote the distribution of $\tY|\bf{0}$ induced by the zero parameter tensor. Then we have
-\[
-\text{KL}(\mathbb{P}_{\Theta}||\mathbb{P}_{\bf{0}})\leq {2(2L-3)\over A_\alpha}  \dot{f}^2(0)\FnormSize{}{\Theta}^2.
-\]
-\end{rmk}
-
-\begin{lem}\label{lem:construction}
-Assume the same setup as in Theorem~\ref{thm:minimax}.
-Without loss of generality, suppose $d_1=\max_kd_k$.
-Define $R=\max_k r_k$ and $d_{\text{total}}=\prod_{k\in[K]} d_k$.
-For any constant $0\leq \gamma \leq 1$, there exist a finite set of tensors $\tX=\{\Theta_i: i=1,\ldots\}\subset \tP$ satisfying the following four properties:
-\begin{enumerate}
-\item $\text{Card}(\tX)\geq 2^{Rd_1/8}+1$, where $\text{Card}$ denotes the cardinality;
-\item $\tX$ contains the zero tensor $\mathbf{0}\in\mathbb{R}^{d_1\times \cdots\times d_K}$;
-\item $\mnormSize{}{\Theta}\leq \gamma \min\left\{ \alpha ,\ C^{-1/2}\sqrt{Rd_1\over d_{\text{total}}} \right\} $ for any element $\Theta\in\tX$;
-\item $\FnormSize{}{\Theta_i-\Theta_j}\geq {\gamma\over 4} \min\left\{ \alpha\sqrt{d_{\text{total}}},\ C^{-1/2}\sqrt{Rd_1}\right\}$ for any two distinct elements $\Theta_i,\; \Theta_j\in\tX$,
-\end{enumerate}
-Here $C=C(\alpha,L,f,\mb)={(4L-6)\dot{f}^2(0)\over A_\alpha }>0$ is a constant independent of the tensor dimension and rank.
-\end{lem}
-
-\begin{proof}
-Given a constant $0\leq \gamma \leq 1$, we define a set of matrices:
-\[
-\tC=\left\{\mM=(m_{ij})\in\mathbb{R}^{d_1\times R}: a_{ij}\in \left\{ 0,\gamma \min\left\{ \alpha , C^{-1/2}\sqrt{Rd_1\over d_{\text{total}}} \right\}\right\} ,\  \forall (i,j)\in[d_1]\times[R]\right\}.
-\]
-We then consider the associated set of block tensors:
-\begin{align}
-\tB=\tB(\tC)=\{\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}\colon& \Theta=\mA\otimes \mathbf{1}_{d_3}\otimes \cdots \otimes \mathbf{1}_{d_K}, \ \text{where}\ \mA=(\mM|\cdots|\mM|\mO) \in\mathbb{R}^{d_1\times d_2},\ \mM\in\tC\},
-\end{align}
-where $\mathbf{1}_d$ denotes a length-$d$ vector with all entries 1, $\mO$ denotes the $d_1\times (d_2-R\lfloor d_2/R \rfloor)$ zero matrix, and $\lfloor d_2/ R \rfloor$ is the integer part of $d_2/R$.
-In other words, the subtensor $\Theta(\mI, \mI,i_3, \ldots,i_K)\in\mathbb{R}^{d_1\times d_2}$ are the same for all fixed $(i_3,\ldots,i_K)\in[d_3]\times \cdots \times [d_K]$, and furthermore, each subtensor $\Theta(\mI,\mI, i_3,\ldots,i_K)$ itself is filled by copying the matrix $\mM\in\mathbb{R}^{d_1\times R}$ as many times as would fit.
-
-By construction, any element of $\tB$, as well as the difference of any two elements of $\tB$, has Tucker rank at most $\max_k r_k\leq R$, and the entries of any tensor in $\tB$ take values in $[0,\alpha]$.
-Thus, $\tB\subset\tP$.
-By Lemma~\ref{lem:VGbound}, there exists a subset $\tX\subset \tB$ with cardinality $\text{Card}(\tX)\geq 2^{Rd_1/8}+1$ containing the zero $d_1\times \cdots \times d_K$ tensor, such that, for any two distinct elements $\Theta_i$ and $\Theta_j$ in $\tX$,
-\[
-\FnormSize{}{\Theta_i-\Theta_j}^2 \geq {Rd_1\over 8} \gamma^2\min\left\{ \alpha^2, {C^{-1}Rd_1 \over d_{\text{total}}}\right\} \Big\lfloor {d_2\over R}\Big \rfloor \prod_{k\geq 3}d_k\geq {\gamma^2\min\left\{ \alpha^2 d_{\text{total}}, C^{-1}Rd_1\right\}  \over 16}.
-\]
-In addition, each entry of $\Theta\in\tX$ is bounded by $\gamma \min\left\{ \alpha , C^{-1/2}\sqrt{Rd_1\over d_{\text{total}}}\right\} $. Therefore the Properties (i) to (iv) are satisfied.
-\end{proof}
-
-\begin{lem}[Varshamov-Gilbert bound]\label{lem:VGbound}
-Let $\Omega=\{(w_1,\ldots,w_m)\colon w_i\in\{0,1\}\}$. Suppose $m>8$. Then there exists a subset $\{w^{(0)},\ldots,w^{(M)}\}$ of $\Omega$ such that $w^{(0)}=(0,\ldots,0)$ and
-\[
-\zeronormSize{}{w^{(j)}-w^{(k)}}\geq {m\over 8},\quad \text{for } \ 0\leq j<k\leq M,
-\]
-where $\zeronormSize{}{\cdot}$ denotes the Hamming distance, and $M\geq 2^{m/8}$.
-\end{lem}
-
-\begin{lem}[Theorem 2.5 in~\citet{tsybakov2008introduction}]\label{lem:Tsybakov}
-Assume that a set $\tX$ contains element $\Theta_0, \Theta_1, \ldots,\Theta_M$ ($M\geq 2$) such that
-\begin{itemize}
-\item $d(\Theta_j,\ \Theta_k)\geq 2s>0$, $\forall 0\leq j\leq k\leq M$;
-\item $\mathbb{P}_0$ is absolutely continuous with respect to $\mathbb{P}_j$, $\forall j=1,\ldots,M$, and
-\[
-{1\over M}\sum_{j=1}^M \mathrm{KL}(\mathbb{P}_j||\mathbb{P}_0)\leq \alpha \log M
-\]
-where $d\colon \tX\times \tX\mapsto [0,+\infty]$ is a semi-distance function, $0<\alpha<{1/8}$ and $\mathbb{P}_j=\mathbb{P}_{\Theta_j}$, $j=0,1\ldots,M$.
-\end{itemize}
-Then
-\[
-\inf_{\hat \Theta}\sup_{\Theta\in\tX} \mathbb{P}_{\Theta}(d(\hat \Theta, \Theta)\geq s)\geq {\sqrt{M}\over 1+\sqrt{M}}\left(1-2\alpha-\sqrt{2\alpha\over \log M} \right)>0.
-\]
-\end{lem}
-
-\begin{lem}[Lemma 28 in~\citet{ghadermarzy2019near}]\label{lem:convexity}
-Define $\mathbb{B}_{M}(\beta)=\{\Theta\in \mathbb{R}^{d_1\times \cdots \times d_K}\colon \anormSize{}{\Theta}\leq \beta \}$.
-Let $\Omega\subset[d_1]\times\cdots \times [d_K]$ be a random set with $m=|\Omega|$, and assume that each entry in $\Omega$ is drawn with replacement from $[d_1]\times\cdots\times[d_K]$ using probability $\Pi$.
-Define
-\[
-\PiFnormSize{}{\Theta}^2={1\over m}\mathbb{E}_{\Omega\in\Pi}\MFnormSize{}{\Theta}^2.
-\]
-Then, there exists a universal constant $c>0$, such that, with probability at least $1-\exp\left(-{\sum_kd_k \over \sum_k \log d_k} \right)$ over the sampled set $\Omega$,
-\[
-{1\over m}\MFnormSize{}{\Theta}^2 \geq \PiFnormSize{}{\Theta}^2-c\beta\sqrt{\sum_k d_k\over m}
-\]
-holds uniformly for all $\Theta\in\mathbb{B}_M(\beta)$.
-\end{lem}
-
-\section{Extension of Theorem~\ref{thm:rate} to unknown cut-off points}\label{sec:extention}
+\section{Proofs}
+Here, we provide additional theoretical results and proofs presented in Sections~\ref{sec:theory}.
+
+\subsection{Extension of Theorem~\ref{thm:rate} to unknown cut-off points}\label{sec:extention}
 We now extend Theorem~\ref{thm:rate} to the case of unknown cut-off points $\mb$. Assume that the true parameters $(\trueT, \trueb)\in \tP\times \tB$, where the feasible sets are defined as
@@
 The link function $f\colon \mathbb{R}\mapsto [0,1]$ satisfies the following properties:
-\begin{enumerate}
+\begin{enumerate}[label=\textup{(\roman*)}]
 \item $f(z)$ is twice-differentiable and strictly increasing in $z$.
@@
 We bound the score and the Hessian to obtain \eqref{eq:scoreZ}.
-
-\begin{enumerate}
+\begin{enumerate}[itemsep=0pt,topsep=0pt,leftmargin=*,partopsep=0pt]
 \item (Score.) The $(\omega,\ell)$-th entry in $\plogZ$ is 
@@
 \left(\sum_k d_k+ L-1\right)\right)$, we have
-\begin{align}\label{eq:scorebound}
+\begin{equation}\label{eq:scorebound}
 \snormSize{}{\plogZ(\trueZ)}\leq c_2C_{\alpha,\beta,\Delta}\sqrt{L-1+\sum_k d_k},
-\end{align}
+\end{equation}
 
@@
 Based on Assumption~\ref{ass:joint}, the Hession matrix $\mH$ has the following three properties:
-\begin{enumerate}
+\begin{enumerate}[label=\textup{(\roman*)}]
 \item The Hession matrix is a block matrix, $\mH=\text{diag}\{\mH_\omega\colon \omega\in[d_1]\times \cdots \times [d_K]\}$, and each block $\mH_{\omega}\in\mathbb{R}^{(L-1)\times (L-1)}$ is a tridiagonal matrix.
@@
 \begin{align}
-\mH_\omega(\ell,\ell)+ \mH_\omega(\ell,\ell-1)+\mH_\omega(\ell,\ell+1)&
-={\partial^2 \tL_\tY\over \partial z^2_{\omega,\ell}}+{\partial^2 \tL_\tY\over \partial z_{\omega,\ell}z_{\omega,\ell+1}}+{\partial^2 \tL_\tY\over \partial z_{\omega,\ell-1}z_{\omega,\ell}}\\
-&=\mathds{1}\{y_\omega=\ell\}{\partial\over \partial z}\left({\dot{f}(z) \over f(z)-f(z')}\right)\Bigg|_{(z,\ z')=(z_{\omega,\ell},\ z_{\omega,\ell-1})} \\& -\mathds{1}\{y_\omega=\ell+1\}{\partial\over \partial z}\left({\dot{f}(z) \over f(z)-f(z')}\right)\Bigg|_{(z'',\ z)=(z_{\omega,\ell+1},\ z_{\omega,\ell})}\\
-&\leq -D_{\alpha,\beta,\Delta}\mathds{1}\{y_\omega = \ell \text{ or } \ell+1\}.
+&\mH_\omega(\ell,\ell)+ \mH_\omega(\ell,\ell-1)+\mH_\omega(\ell,\ell+1)
+\\
+=&\ {\partial^2 \tL_\tY\over \partial z^2_{\omega,\ell}}+{\partial^2 \tL_\tY\over \partial z_{\omega,\ell}z_{\omega,\ell+1}}+{\partial^2 \tL_\tY\over \partial z_{\omega,\ell-1}z_{\omega,\ell}}\\
+=&\ \mathds{1}\{y_\omega=\ell\}{\partial\over \partial z}\left({\dot{f}(z) \over f(z)-f(z')}\right)\Bigg|_{(z,\ z')=(z_{\omega,\ell},\ z_{\omega,\ell-1})} \\& -\mathds{1}\{y_\omega=\ell+1\}{\partial\over \partial z}\left({\dot{f}(z) \over f(z)-f(z')}\right)\Bigg|_{(z'',\ z)=(z_{\omega,\ell+1},\ z_{\omega,\ell})}\\
+\leq &\ -D_{\alpha,\beta,\Delta}\mathds{1}\{y_\omega = \ell \text{ or } \ell+1\}.
 \end{align}
@@
 Let $\mz_\omega=(z_{\omega,1},\ldots,z_{\omega,L-1})^T\in\mathbb{R}^{L-1}$ and $\mz=(\mz_{1,\ldots,1,1},\ldots,\mz_{d_1,\ldots,d_K,L-1})^T\in\mathbb{R}^{(L-1)\prod_k d_k}$.
-It follows from property (a) that
+It follows from property (i) that
 \[
@@
 \]
-Furthermore, from properties (b) and (c) we have
+Furthermore, from properties (ii) and (iii) we have
 \begin{align}
@@
 &\geq \sum_{\omega}\theta^2_\omega -2\sum_{\omega,\ell}\theta_\omega b_\ell+\sum_{\ell}b_\ell^2\left(n_{\ell}+n_{\ell+1}\right)\\
-&\geq \sum_{\omega}\theta^2_\omega+\min_\ell\left(n_{\ell}+n_{\ell+1}\right)\sum_{\ell}b_\ell^2
+&\geq \sum_{\omega}\theta^2_\omega+\min_\ell\left(n_{\ell}+n_{\ell+1}\right)\sum_{\ell}b_\ell^2.
 \end{align}
@@
 \end{enumerate}
-
- Plugging~\eqref{eq:scorebound} and~\eqref{eq:HessionZ} into~\eqref{eq:boundZ} yields
+Finally, plugging~\eqref{eq:scorebound} and~\eqref{eq:HessionZ} into~\eqref{eq:boundZ} yields
 \[
@@
 The MSEs for $\hat \Theta$ and $\hat \mb$ readily follow from~\eqref{eq:decomp}. 
-
 \end{proof}
+
+\subsection{Proof of Theorem~\ref{thm:rate}}\label{sec:proofMSE}
+\begin{proof}[Proof of Theorem~\ref{thm:rate}]
+We suppress the subscript $\Omega$ in the proof, because the tensor denoising assumes complete observation $\Omega=[d_1]\times \cdots \times [d_K]$.
+It follows from the expression of $\flogl(\Theta)$ that
+\begin{align}\label{eq:property}
+{\partial \flogl\over \partial \theta_\omega}&=\sum_{\ell\in[L]}\mathds{1}\{y_{\omega}=\ell\}
+{\dot{g}_\ell(\theta_\omega)\over g_\ell(\theta_\omega)},\notag\\
+{\partial^2 \flogl\over \partial \theta_\omega^2}&=\sum_{\ell\in[L]}\mathds{1}\{y_\omega=\ell\}{\ddot{g}_\ell(\theta_\omega)g_\ell(\theta_\omega)-\dot{g}^2_\ell(\theta_\omega)\over g^2_\ell(\theta_\omega)}\ \text{and}\quad
+{\partial^2 \flogl\over \partial \theta_\omega \theta_\omega'}=0\ \text{if}\ \omega\neq \omega',
+\end{align}
+for all $\omega\in[d_1]\times \cdots \times [d_K]$.
+Define $d_{\text{total}}=\prod_k d_k$.
+Let $\fplogl\in\mathbb{R}^{d_1\times\cdots\times d_K}$ denote the tensor of gradient with respect to $\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}$, and $\fpplogl$ the corresponding Hession matrix of size $d_\text{total}$-by-$d_{\text{total}}$.
+Here, $\Vec(\cdot)$ denotes the operation that turns a tensor into a vector.
+By~\eqref{eq:property}, $\fpplogl$ is a diagonal matrix.
+Recall that
+\begin{equation}\label{eq:bound}
+U_\alpha=\max_{\ell\in[L],|\theta|\leq \alpha}{|\dot{g}_\ell(\theta)|\over g_\ell(\theta)}>0 \quad \text{and}\quad
+L_\alpha=\min_{\ell\in[L],|\theta|\leq \alpha} {\dot{g}^2_\ell(\theta)-\ddot{g}_\ell(\theta)g_\ell(\theta)\over g^2_\ell(\theta)}>0.
+\end{equation}
+Therefore, the entries in $\fplogl$ are upper bounded in magnitude by $U_\alpha>0$, and all diagonal entries in $\fpplogl$ are upper bounded by $-L_{\alpha}<0$.
+
+By the second-order Taylor's expansion of $\flogl(\Theta)$ around $\trueT$, we obtain
+\begin{equation}\label{eq:taylor}
+\flogl(\Theta)=\flogl(\trueT)+\langle\Vec(\fplogl(\trueT)),\ \Vec(\Theta-\trueT)\rangle+{1\over 2}\Vec(\Theta-\trueT)^T\fpplogl(\check\Theta)\Vec(\Theta-\trueT),
+\end{equation}
+where $\check\Theta=\gamma\trueT+(1-\gamma)\Theta$ for some $\gamma\in[0,1]$, and $\fpplogl(\check\Theta)$ denotes the $d_{\text{total}}$-by-$d_\text{total}$ Hession matrix evaluated at $\check\Theta$.
+
+We first bound the linear term in~\eqref{eq:taylor}. Note that, by Lemma~\ref{lem:inq},
+\begin{equation}\label{eq:linear}
+|\langle\Vec(\fplogl(\trueT), \Vec(\Theta-\trueT)  \rangle|\leq \snormSize{}{\fplogl(\trueT)} \nnormSize{}{\Theta-\trueT},
+\end{equation}
+where $\snormSize{}{\cdot}$ denotes the tensor spectral norm and $\nnormSize{}{\cdot}$ denotes the tensor nuclear norm. Define
+\[
+s_\omega={\partial \tL_\tY\over \partial \theta_\omega}\Big|_{\Theta=\trueT} \;\; \textrm{ for all } \; \omega\in[d_1]\times\cdots\times [d_K].
+\]
+Based on~\eqref{eq:property} and the definition of $U_\alpha$, $\fplogl(\trueT)=\entry{s_{\omega}}$ is a random tensor whose entries are independently distributed satisfying
+\begin{equation}\label{eq:norm}
+\mathbb{E}(s_\omega)=0,\quad |s_\omega|\leq U_\alpha, \quad \text{for all }\omega\in[d_1]\times \cdots \times [d_K].
+\end{equation}
+By lemma~\ref{lem:noisytensor}, with probability at least $1-\exp(-C_1 \sum_kd_k)$, we have
+\begin{equation}\label{eq:normrandom}
+\snormSize{}{\fplogl(\trueT)} \leq C_2 U_\alpha\sqrt{\sum_k d_k},
+\end{equation}
+where $C_1, C_2$ are two positive constants that depend only on $K$.
+Furthermore, note that $\text{rank}(\Theta)\leq \mr$, $\text{rank}(\trueT)\leq \mr$, so $\text{rank}(\Theta-\trueT)\leq 2\mr$.
+By lemma~\ref{lem:nuclear}, $\nnormSize{}{\Theta-\trueT}\leq (2r_{\max})^{K-1\over 2}\FnormSize{}{\Theta-\trueT}$.
+Combining~\eqref{eq:linear}, \eqref{eq:norm} and \eqref{eq:normrandom}, we have that, with probability at least $1-\exp(-C_1 \sum_kd_k)$,
+\begin{equation}\label{eq:linearconclusion}
+|\langle \Vec(\fplogl(\trueT)), \Vec(\Theta-\trueT)  \rangle | \leq C_2 U_\alpha  \sqrt{r_{\max}^{K-1} \sum_k d_k}  \FnormSize{}{\Theta-\trueT}.
+\end{equation}
+
+We next bound the quadratic term in \eqref{eq:taylor}. Note that
+\begin{align}\label{eq:quadratic}
+ \Vec(\Theta-\trueT)^T \fpplogl(\check{\Theta})\Vec(\Theta-\trueT)&=\sum_\omega \left( {\partial^2\tL_{\tY}\over \partial \theta^2_\omega} \Big|_{\Theta=\check\Theta} \right)(\theta_\omega-\theta_{{\text{true}},\omega})^2 \nonumber \\
+&\leq - L_\alpha\sum_{\omega}(\Theta_{\omega}-\Theta_{\text{true},\omega})^2 \nonumber \\
+&=-L_\alpha\FnormSize{}{\Theta-\trueT}^2,
+\end{align}
+where the second line comes from the fact that  $\mnormSize{}{\check\Theta}\leq \alpha$ and the definition of $L_\alpha$.
+
+Combining~\eqref{eq:taylor}, \eqref{eq:linearconclusion} and~\eqref{eq:quadratic}, we have that, for all $\Theta\in\tP$, with probability at least $1-\exp(-C_1 \sum_kd_k)$,
+\[
+\tL_\tY(\Theta)\leq \tL_{\tY}(\trueT)+C_2U_\alpha  \left(r_{\max}^{K-1}\sum_k d_k\right)^{1/2}  \FnormSize{}{\Theta-\trueT}-{L_\alpha\over 2}\FnormSize{}{\Theta-\trueT}^2.
+\]
+In particular, the above inequality also holds for $\hat \Theta\in\tP$. Therefore,
+\[
+\tL_\tY(\hat \Theta)\leq \tL_{\tY}(\trueT)+C_2U_\alpha \left(r_{\max}^{K-1}\sum_k d_k\right)^{1/2}  \FnormSize{}{\hat \Theta-\trueT}-{L_\alpha\over 2} \FnormSize{}{\hat \Theta-\trueT}^2.
+\]
+Since $\hat \Theta=\arg\max_{\Theta\in\tP}\tL_\tY(\Theta)$, $\tL_\tY(\hat \Theta)-\tL_{\tY}(\trueT)\geq 0$, which gives
+\[
+C_2U_\alpha \left(r_{\max}^{K-1}\sum_k d_k\right)^{1/2}  \FnormSize{}{\hat \Theta-\trueT}-{L_\alpha\over 2}\FnormSize{}{\hat \Theta-\trueT}^2\geq 0.
+\]
+Henceforth,
+\[
+{1\over \sqrt{\prod_k d_k}} \FnormSize{}{\hat \Theta-\trueT}\leq {2C_2U_\alpha \sqrt{r_{\max}^{K-1}\sum_k d_k}\over L_\alpha \sqrt{\prod_k d_k}}={2C_2U_\alpha r_{\max}^{(K-1)/2}\over L_\alpha} \sqrt{ \sum_k d_k \over \prod_k d_k}.
+\]
+This completes the proof.
+\end{proof}
+
+\begin{proof}[Proof of Corollary~\ref{cor:prediction}]
+The result follows immediately from Theorem~\ref{thm:rate} and Lemma~\ref{lem:KL}.
+\end{proof}
+
+\subsection{Proof of Theorem~\ref{thm:minimax}}
+\begin{proof}[Proof of Theorem~\ref{thm:minimax}]
+
+Let $d_{\text{total}}=\prod_{k\in[K]}d_k$, and $\gamma\in[0,1]$ be a constant to be specified later.
+Our strategy is to construct a finite set of tensors $\tX=\{\Theta_i\colon i=1,\ldots \}\subset \tP$ satisfying the properties of (i)-(iv) in Lemma~\ref{lem:construction}.
+By Lemma~\ref{lem:construction}, such a subset of tensors exist.
+For any tensor  $\Theta\in\tX$, let $\mathbb{P}_{\Theta}$ denote the distribution of $\tY|\Theta$, where $\tY$ is the ordinal tensor.
+In particular, $\mathbb{P}_{\mathbf{0}}$ is the distribution of $\tY$ induced by the zero parameter tensor $\mathbf{0}$, i.e., the distribution of $\tY$ conditional on the parameter tensor $\Theta=\mathbf{0}$.
+Based on the Remark for Lemma~\ref{lem:KL}, we have
+\begin{equation}\label{eq:KLbound1}
+\mathrm{KL}(\mathbb{P}_{\Theta}|| \mathbb{P}_{\mathbf{0}})\leq C \FnormSize{}{\Theta}^2,
+\end{equation}
+where $C={(4L-6) \dot{f}^2(0)\over  A_\alpha}>0$ is a constant independent of the tensor dimension and rank.
+Combining the inequality~\eqref{eq:KLbound1} with property (iii) of $\tX$, we have
+\begin{equation}\label{eq:KLbound}
+\text{KL}(\mathbb{P}_{\Theta}||\mathbb{P}_{\mathbf{0}})\leq \gamma^2 r_{\max} d_{\max}.
+\end{equation}
+From~\eqref{eq:KLbound} and the property (i), we deduce that the condition
+\begin{equation}\label{eq:totalKL}
+{1\over \text{Card}(\tX)-1}\sum_{\Theta \in\tX}\text{KL}(\mathbb{P}_{\Theta}, \mathbb{P}_{\mathbf{0}})\leq \varepsilon \log\left\{\text{Card}(\tX)-1 \right\}
+\end{equation}
+holds for any $ \varepsilon \geq 0$ when $\gamma\in[0,1]$ is chosen to be sufficiently small depending on $\varepsilon$, e.g., $\gamma \leq \sqrt{\varepsilon\log2\over8}$.
+By applying Lemma~\ref{lem:Tsybakov} to~\eqref{eq:totalKL}, and in view of the property (iv), we obtain that
+\begin{equation}\label{eq:final}
+\inf_{\hat \Theta}\sup_{\trueT\in \tX}\mathbb{P}\left(\FnormSize{}{\hat \Theta- \trueT}\geq  {\gamma\over 8} \min\left\{ \alpha\sqrt{d_{\text{total}}}, C^{-1/2}\sqrt{ r_{\max}d_{\max}}\right\} \right)\geq {1\over 2}\left(1-2\varepsilon-\sqrt{16 \varepsilon \over r_{\max}d_{\max}\log2}\right).
+\end{equation}
+Note that $\textup{MSE}(\hat \Theta, \trueT)=\FnormSize{}{\hat \Theta- \trueT}^2/d_{\text{total}}$ and $\tX\subset \tP$.
+By taking $\varepsilon=1/10$ and $\gamma=1/11$, we conclude from~\eqref{eq:final} that
+\begin{equation}\label{eq:prob}
+\inf_{\hat \Theta}\sup_{\trueT\in \tP}\mathbb{P}\left(\textup{MSE}(\hat \Theta, \trueT)\geq c\min\left \{ \alpha^2,  {C^{-1}r_{\max}d_{\max}\over d_{\text{total}}}\right \}\right)\geq {1\over 2}\left({4\over 5}- \sqrt{1.6\over r_{\max}d_{\max}\log2} \right)\geq {1\over 8},
+\end{equation}
+where $c = {1 \over 88^2}$ and the last inequality comes from the condition for $d_{\text{max}}$.
+This completes the proof.
+\end{proof}
+
+\subsection{Proof of Theorem~\ref{thm:completion}}
+
+\begin{proof}[Proof of Theorem~\ref{thm:completion}]
+
+For notational convenience, we use $\MFnormSize{}{\Theta}=\sum_{\omega\in\Omega}\Theta^2_\omega$ to denote the sum of squared entries over the observed set $\Omega$, for a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$.
+
+Following a similar argument as in the proof of Theorem~\ref{thm:rate}, we have
+\begin{equation}\label{eq:Taylor2}
+\logl(\Theta)=\logl(\trueT)+\langle\Vec(\plogl),\ \Vec(\Theta-\trueT)\rangle+{1\over 2}\Vec(\Theta-\trueT)^T\pplogl(\check\Theta)\Vec(\Theta-\trueT),
+\end{equation}
+where
+\begin{enumerate}[label=\textup{(\roman*)}]
+\item $\plogl$ is a $d_1\times\cdots\times d_K$ tensor with $|\Omega|$ nonzero entries, and each entry is upper bounded by $U_\alpha>0$.
+\item $\pplogl$ is a diagonal matrix of size $d_{\text{total}}$-by-$d_{\text{total}}$ with $|\Omega|$ nonzero entries, and each entry is upper bounded by $-L_{\alpha}<0$.
+\end{enumerate}
+
+Similar to~\eqref{eq:linear} and~\eqref{eq:quadratic}, we have
+\begin{equation}\label{eq:linear2}
+|\langle\Vec(\plogl),\ \Vec(\Theta-\trueT)\rangle|\leq C_2U_\alpha \sqrt{r_{\max}^{K-1}\sum_k d_k}\MFnormSize{}{\Theta-\trueT}
+\end{equation}
+and
+\begin{equation}\label{eq:quadratic2}
+\Vec(\Theta-\trueT)^T\fpplogl(\check\Theta)\Vec(\Theta-\trueT)\leq -L_\alpha \MFnormSize{}{\Theta-\trueT}^2.
+\end{equation}
+
+Combining~\eqref{eq:Taylor2}-\eqref{eq:quadratic2} with the fact that $\logl(\hat \Theta)\geq \logl(\trueT)$, we have
+\begin{equation}\label{eq:sample}
+\MFnormSize{}{\hat \Theta-\trueT}\leq {2C_2U_\alpha  r_{\max}^{(K-1)/2} \over L_\alpha} \sqrt{\sum_kd_k},
+\end{equation}
+with probability at least $1-\exp(-C_1 \sum_k d_k)$.
+Lastly, we invoke the result regarding the closeness of $\Theta$ to its sampled version $\Theta_{\Omega}$, under the entrywise bound condition.
+Note that $\mnormSize{}{\hat\Theta-\trueT}\leq 2\alpha$ and $\text{rank}(\hat \Theta-\trueT)\leq 2\mr$.
+By Lemma~\ref{lem:Mnormbound}, $\anormSize{}{\hat \Theta-\trueT}\leq 2^{(3K-1)/2} \alpha \left({\prod r_k \over r_{\max}}\right)^{3/2}$.
+Therefore, the condition in Lemma~\ref{lem:convexity} holds with $\beta=2^{(3K-1)/2}\alpha \left({\prod r_k \over r_{\max}}\right)^{3/2}$.
+Applying Lemma~\ref{lem:convexity} to~\eqref{eq:sample} gives
+\begin{align}
+ \PiFnormSize{}{\hat \Theta-\trueT}^2&\leq {1\over m}\MFnormSize{}{\hat \Theta-\trueT}^2+c\beta\sqrt{\sum_k d_k\over |\Omega|}\\
+ &\leq {C_2  r^{K-1}_{\max}} {\sum_k d_k \over |\Omega|}+C_1 \alpha r_{\max}^{3(K-1)/2}\sqrt{\sum_kd_k\over |\Omega|},
+\end{align}
+with probability at least $1-\exp(-{\sum_kd_k\over \sum_k \log d_k})$ over the sampled set $\Omega$. Here $C_1, C_2>0$ are two constants independent of the tensor dimension and rank. Therefore,
+\[
+ \PiFnormSize{}{\hat \Theta-\trueT}^2\to 0,\quad \text{as}\quad {|\Omega|\over \sum_kd_k}\to \infty,
+\]
+provided that $r_{\max}=O(1)$.
+\end{proof}
+
+\subsection{Convexity of the log-likelihood function}\label{sec:proofconvexity}
+\begin{thm}\label{thm:convexity}
+Define the function
+\begin{equation}\label{eq:function}
+ \logl(\Theta, \mb)=\sum_{\omega\in\Omega}\sum_{\ell\in[L]} \big(\mathds{1}\{y_\omega=\ell\}\log \left[f(b_\ell-\theta_\omega)-  f(b_{\ell-1}-\theta_\omega)\right]\big),
+ \end{equation}
+where $f(\cdot)$ satisfies Assumption~\ref{ass:link}. Then, $\logl(\Theta, \mb)$ is concave in $(\Theta,\mb)$.
+\end{thm}
+
+\begin{proof}
+Define $d_{\text{total}}=\prod_k d_k$.
+By abuse of notation, we use $(\Theta,\mb)$ to denote the length-$(d_{\text{total}}+L-1)$-vector collecting all parameters together.
+Let us denote a bivariate function
+\begin{align}
+\lambda\colon &\mathbb{R}^2\mapsto \mathbb{R}\\
+(u,v)&\mapsto\lambda(u,v) = \log \big[f(u)-  f(v)\big].
+\end{align}
+It suffices to show that $\lambda(u,v)$ is concave in $(u,v)$ where $u>v$.
+
+Suppose that the claim holds (which we will prove in the next paragraph). Based on~\eqref{eq:function}, $u,v$ are both linear functions of $(\Theta,\mb)$:
+\[
+u = \ma_1^T(\Theta,\mb), \quad v = \ma_2^T(\Theta,\mb), \quad \text{ for some}\ \ma_1,\ma_2\in \mathbb{R}^{d_{\text{total}}+L-1}.
+\]
+Then, $\lambda(u,v) = \lambda(\ma_1^T(\Theta,\mb),\ \ma_2^T(\Theta,\mb))$ is concave in $(\Theta,\mb)$ by the definition of concavity.
+Therefore, we can conclude that $\logl(\Theta, \mb) $ is concave in $(\Theta,\mb)$ because $\logl(\Theta, \mb)$ is the sum of $\lambda(u,v)$.
+
+Now, we prove the concavity of $\lambda(u,v)$. Note that
+\begin{equation}
+\lambda(u,v) = \log\big[f(u)-f(v)\big]=\log\big[\int\mathds{1}_{[u,v]}(x)f'(x)dx\big],
+\end{equation}
+where $\mathds{1}_{[u,v]}$ is an indicator function that equals 1 in the interval $[u,v]$, and 0 elsewhere.
+Furthermore, $\mathds{1}_{[u,v]}(x)$ is log-concave in $(u,v,x)$, and by Assumption~\ref{ass:link}, $f'(x)$ is log-concave in $x$.
+It follows that $\mathds{1}_{[u,v]}(x)f'(x)$
+is a log-concave in $(u,v,x)$. By Lemma~\ref{lem:lossconvexity}, we conclude that $\lambda(u,v)$ is concave in $(u,v)$ where $u>v$.
+\end{proof}
+
+\begin{lem}[Corollary 3.5 in~\citet{brascamp2002extensions}]\label{lem:lossconvexity}
+Let $F(x,y)\colon \mathbb{R}^{m+n}\rightarrow \mathbb{R}$ be an integrable function where $x\in \mathbb{R}^{m},y\in \mathbb{R}^n$. Let
+\[
+G(x) = \int_{\mathbb{R}^n}F(x,y)dy.
+\]
+If $F(x,y)$ is log concave in $(x,y)$, then $G(x)$ is log concave in $x$.
+\end{lem}
+
+\section{Conclusions}
+We have developed a low-rank tensor estimation method based on possibly incomplete, ordinal-valued observations.
+A sharp error bound is established, and we demonstrate the outperformance of our approach compared to other methods.
+The work unlocks several directions of future research.
+One interesting question would be the inference problem, i.e.,\ to assess the uncertainty of the obtained estimates and the imputation.
+Other directions include the trade-off between (non)convex optimization and statistical efficiency.
+While we have provided numerical evidence for the success of nonconvex approach, the full landscape of the optimization remains open.
+The interplay between computational efficiency and statistical accuracy in general tensor problems warrants future research.
+
+\section*{Acknowledgements}
+This research is supported in part by NSF grant DMS-1915978 and Wisconsin Alumni Research Foundation.
+
+\clearpage
+\appendix
+
+\renewcommand{\thefigure}{{S\arabic{figure}}}%
+\renewcommand{\figurename}{{Supplementary Figure}}    
+\setcounter{figure}{0}   
+\setcounter{table}{0}
+
+\section*{Appendix}
+
+\section{Auxiliary lemmas}\label{sec:lemma}
+This section collects lemmas that are useful for the proofs of the main theorems.
+
+\begin{defn}[Atomic M-norm~\citep{ghadermarzy2019near}]
+Define $T_{\pm}=\{\tT\in\{\pm 1 \}^{d_1\times \cdots \times d_K}\colon \text{rank}(\tT)=1\}$. The atomic M-norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
+\begin{align}
+\anormSize{}{\Theta}&=\inf\{t>0\colon \Theta\in t\text{conv}(T_{\pm})\}\\
+&=\inf\left\{\sum_{\tX\in T_{\pm}}c_{\tX} \colon\ \Theta=\sum_{\tX\in T_{\pm}} c_{\tX}\tX, \ c_{\tX}>0\right\}.
+\end{align}
+\end{defn}
+
+\begin{defn}[Spectral norm~\citep{lim2005singular}]
+The spectral norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
+\[
+\snormSize{}{\Theta}=\sup\left\{\langle \Theta, \mx_1\otimes \cdots \otimes \mx_K\rangle \colon \vnormSize{}{\mx_k}=1,\ \mx_k\in\mathbb{R}^{d_k},\ \text{for all}\ k\in[K]\right\}.
+\]
+\end{defn}
+
+\begin{defn}[Nuclear norm~\citep{friedland2018nuclear}]
+The nuclear norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
+\[
+\nnormSize{}{\Theta}=\inf
+\left\{
+\sum_{i\in[r]}|\lambda_i|\colon \Theta=\sum_{i=1}^r \lambda_i\mx^{(i)}_1\otimes \cdots \otimes \mx^{(i)}_K,\ \vnormSize{}{\mx^{(i)}_k}=1,\ \mx^{(i)}_k\in\mathbb{R}^{d_k},\ \text{for all}\ k\in[K],\ i\in[r]
+\right\},
+\]
+where the infimum is taken over all $r\in\mathbb{N}$ and $\vnormSize{}{\mx^{(i)}_k}=1$ for all $i\in[r]$ and $k\in[K]$.
+\end{defn}
+
+\begin{lem}[M-norm and infinity norm~\citep{ghadermarzy2019near}]\label{lem:Mnormbound}
+Let $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ be an order-$K$, rank-$(r_1,\ldots,r_K)$ tensor. Then
+\[
+\mnormSize{}{\Theta}\leq \anormSize{}{\Theta}\leq \left(\prod_k r_k \over r_{\max}\right)^{3\over 2} \mnormSize{}{\Theta}.
+\]
+\end{lem}
+
+\begin{lem}[Nuclear norm and F-norm] \label{lem:nuclear}
+Let $\tA\in\mathbb{R}^{d_1\times\cdots\times d_K}$ be an order-$K$ tensor with Tucker $\text{rank}(\tA)=(r_1,\ldots,r_K)$. Then
+\[
+\nnormSize{}{\tA} \leq \sqrt{\prod_k r_k\over \max_k r_k} \FnormSize{}{\tA},
+\]
+where $\nnormSize{}{\cdot}$ denotes the nuclear norm of the tensor.
+\end{lem}
+
+\begin{proof}
+Without loss of generality, suppose $r_1=\min_k r_k$.
+Let $\tA_{(k)}$ denote the mode-$k$ matricization of $\tA$ for all $k\in[K]$.
+By \citet[Corollary 4.11]{wang2017operator}, and the invariance relationship between a tensor and its Tucker core~\citep[Section 6]{jiang2017tensor}, we have
+\begin{equation}\label{eq:norminequality}
+\nnormSize{}{\tA} \leq \sqrt{\prod_{k\geq 2} r_k \over \max_{k\geq 2} r_k} \nnormSize{}{\tA_{(1)}},
+\end{equation}
+where $\tA_{(1)}$ is a $d_1$-by-$\prod_{k\geq 2}d_k$ matrix with matrix rank $r_1$.
+Furthermore, the relationship between the matrix norms implies that $\nnormSize{}{\tA_{(1)}}\leq \sqrt{r_1}\FnormSize{}{\tA_{(1)}}=\sqrt{r_1}\FnormSize{}{\tA}$.
+Combining this fact with the inequality~\eqref{eq:norminequality} yields the final claim.
+\end{proof}
+
+\begin{lem} \label{lem:inq}
+Let $\tA, \; \tB$ be two order-$K$ tensors of the same dimension. Then
+\[
+|\langle \tA,\tB\rangle| \leq \snormSize{}{\tA}   \nnormSize{}{\tB}.
+\]
+\end{lem}
+
+\begin{proof}
+By~\citet[Proposition 3.1]{friedland2018nuclear}, there exists a nuclear norm decomposition of $\tB$, such that
+\[
+\tB=\sum_{r} \lambda_r \ma^{(1)}_r\otimes \cdots\otimes \ma^{(K)}_r,\quad \ma_r^{(k)}\in\mathbf{S}^{d_k-1}(\mathbb{R}),\quad \text{for all }k\in[K],
+\]
+and $\nnormSize{}{\tB}=\sum_{r}|\lambda_r|$. Henceforth we have
+\begin{align*}
+|\langle \tA,\tB\rangle|&=| \langle \tA, \sum_{r} \lambda_r \ma^{(1)}_r\otimes \cdots\otimes \ma^{(K)}_r \rangle|\leq \sum_r |\lambda_r| |\langle \tA, \ma^{(1)}_r \otimes \cdots\otimes \ma^{(K)}_r \rangle|\\
+&\leq \sum_{r}|\lambda_r| \snormSize{}{\tA}= \snormSize{}{\tA}\nnormSize{}{\tB},
+\end{align*}
+which completes the proof.
+\end{proof}
+
+The following lemma provides the bound on the spectral norm of random tensors. The result was firstly presented in~\citet{nguyen2015tensor}, and we adopt the version from~\citet{tomioka2014spectral}.
+\begin{lem}[Spectral norm of random tensors~\citep{tomioka2014spectral}]\label{lem:tensor}
+Suppose that $\tS=\entry{s_{\omega}}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is an order-$K$ tensor whose entries are independent random variables that satisfy
+\[
+\mathbb{E}(s_{\omega})=0,\quad \text{and} \quad\mathbb{E}(e^{ts_{\omega}})\leq e^{t^2L^2/2}.
+\]
+Then the spectral norm $\snormSize{}{\tS}$ satisfies that,
+\[
+\snormSize{}{\tS}\leq \sqrt{{8L^2} \log (12K) \sum_k d_k +\log (2/\delta)},
+\]
+with probability at least $1-\delta$.
+\end{lem}
+
+\begin{lem} \label{lem:noisytensor}
+Suppose that $\tS=\entry{s_{\omega}}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is an order-$K$ tensor whose entries are independent random variables that satisfy
+\[
+\mathbb{E}(s_{\omega})=0,\quad \text{and}\quad |s_{\omega}|\leq U.
+\]
+Then we have
+\[
+\mathbb{P}\left(\snormSize{}{\tS}\geq C_2 U\sqrt{\sum_k d_k} \right)\leq \exp\left(-C_1  \log K \sum_k d_k\right)
+\]
+where $C_1>0$ is an absolute constant, and $C_2>0$ is a constant that depends only on $K$.
+\end{lem}
+
+\begin{proof}  Note that the random variable $U^{-1}s_{\omega}$ is zero-mean and supported on $[-1,1]$. Therefore, $U^{-1}s_{\omega}$ is sub-Gaussian with parameter ${1-(-1)\over 2}=1$, i.e.
+\[
+\mathbb{E}(U^{-1}s_{\omega})=0,\quad \text{and}\quad \mathbb{E}(e^{tU^{-1}s_{\omega}})\leq e^{t^2/2}.
+\]
+It follows from Lemma~\ref{lem:tensor} that, with probability at least $1-\delta$,
+\[
+\snormSize{}{U^{-1}\tS}\leq \sqrt{\left(c_0\log K+c_1\right) \sum_k d_k +\log (2/\delta)},
+\]
+where $c_0, c_1>0$ are two absolute constants. Taking $\delta=\exp (-C_1\log K \sum_k d_k)$ yields the final claim, where $C_2=c_0\log K+c_1+1>0$ is another constant.
+\end{proof}
+
+\begin{lem}\label{lem:KLentry} Let $X,\; Y$ be two discrete random variables taking values on $L$ possible categories, with point mass probabilities $\{p_\ell\}_{\ell\in[L]}$ and $\{q_\ell\}_{\ell\in[L]}$, respectively.
+Suppose $p_\ell$, $q_\ell>0$ for all $\ell\in[L]$.
+Then, the Kullback-Leibler (KL) divergence satisfies that
+\[
+\mathrm{KL}(X||Y)\stackrel{\text{def}}{=}-\sum_{\ell\in[L]}\mathbb{P}_X(\ell)\log\left\{{\mathbb{P}_Y(\ell)\over \mathbb{P}_X(\ell)}\right\} \leq \sum_{\ell \in [L]}{(p_\ell-q_\ell)^2 \over q_\ell}.
+\]
+\end{lem}
+\begin{proof} Using the fact $\log x\leq x-1$ for $x>0$, we have that
+\begin{align}\label{eq:KL}
+\text{KL}(X||Y)&=\sum_{\ell\in[L]}p_\ell\log{p_\ell\over q_\ell}\\
+&\leq \sum_{\ell\in[L]} {p_\ell\over q_\ell}(p_\ell-q_\ell)\\
+&=\sum_{\ell\in[L]} \left({p_\ell\over q_\ell }- 1\right)(p_\ell-q_\ell)+ \sum_{\ell\in[L]} (p_\ell-q_\ell).
+\end{align}
+Note that $\sum_{\ell\in[L]}(p_\ell-q_\ell)=0$. Therefore,
+\[
+\text{KL}(X||Y)\leq \sum_{\ell\in[L]}\left( {p_\ell\over q_\ell}-1\right)\left(p_\ell-q_\ell\right)=\sum_{\ell\in[L]}{(p_\ell-q_\ell)^2\over q_\ell}.
+\]
+\end{proof}
+
+\begin{lem}[KL divergence and F-norm]~\label{lem:KL}
+Let $\tY\in[L]^{d_1\times \cdots \times d_K}$ be an ordinal tensor generated from the model~\eqref{eq:model} with the link function $f$ and parameter tensor $\Theta$.
+Let $\mathbb{P}_{\Theta}$ denote the joint categorical distribution of $\tY|\Theta$ induced by the parameter tensor $\Theta$, where $\mnormSize{}{\Theta}\leq \alpha$.
+Define
+\begin{equation}\label{eq:ass}
+A_\alpha=\min_{\ell\in[L], |\theta|\leq \alpha}\left[f(b_\ell-\theta)-f(b_{\ell-1}-\theta)\right].
+\end{equation}
+Then, for any two tensors $\Theta,\; \Theta^*$ in the parameter spaces, we have
+\[
+\mathrm{KL}(\mathbb{P}_{\Theta}|| \mathbb{P}_{\Theta^*})\leq {2(2L-3)\over A_\alpha} \dot{f}^2(0)\FnormSize{}{\Theta-\Theta^*}^2.
+\]
+\end{lem}
+\begin{proof} Suppose that the distribution over the ordinal tensor $\tY=\entry{y_\omega}$ is induced by $\Theta=\entry{\theta_\omega}$. Then, based on the generative model~\eqref{eq:model},
+\[
+\mathbb{P}(y_\omega=\ell | \theta_\omega)=f(b_{\ell}-\theta_\omega)-f(b_{\ell-1}-\theta_\omega),
+\]
+for all $\ell\in[L]$ and $\omega\in[d_1]\times \cdots \times [d_K]$.
+For notational convenience, we suppress the subscribe in $\theta_\omega$ and simply write $\theta$ (and respectively, $\theta^*$).
+Based on Lemma~\ref{lem:KLentry} and Taylor expansion,
+\begin{align}
+\text{KL}(\theta|| \theta^*) & \leq \sum_{\ell\in[L]} {\left[f(b_{\ell}-\theta)-f(b_{\ell-1}-\theta)-f(b_{\ell}-\theta^*)+f(b_{\ell-1}-\theta^*)\right]^2\over f(b_{\ell}-\theta^*)-f(b_{\ell-1}-\theta^*)}\\
+ &\leq \sum_{\ell=2}^{L-1} {\left[\dot{f}(b_\ell-\eta_\ell)-\dot{f}(b_{\ell-1}-\eta_{\ell-1})\right]^2 \over f(b_\ell-\theta^*)-f(b_{\ell-1}-\theta^*)} \left(\theta-\theta^*\right)^2+{\dot{f}^2(b_1-\eta_1) \over f(b_1-\theta^*)} (\theta-\theta^*)^2\\
+ &\quad \quad\quad \quad  +{\dot{f}^2(b_{L-1}-\eta_{L-1})\over 1-f(b_{L-1}-\theta^*)}(\theta-\theta^*)^2,
+\end{align}
+where $\eta_\ell$ and $\eta_{\ell-1}$ fall between $\theta$ and $\theta^*$. Therefore,
+\begin{equation}\label{eq:entrywise}
+\text{KL}(\theta|| \theta^*) \leq \left({4(L-2)\over A_\alpha}+ {2\over A_\alpha}\right)\dot{f}^2(0)(\theta-\theta^*)^2={2(2L-3)\over A_\alpha} \dot{f}^2(0)(\theta-\theta^*)^2,
+\end{equation}
+where we have used Taylor expansion, the bound~\eqref{eq:ass}, and the fact that $\dot{f}(\cdot)$ peaks at zero for an unimodal and symmetric function.
+Now summing~\eqref{eq:entrywise} over the index set $\omega\in[d_1]\times \cdots \times [d_K]$ gives
+\[
+\text{KL}(\mathbb{P}_{\Theta}|| \mathbb{P}_{\Theta^*}) =\sum_{\omega\in[d_1]\times \cdots \times[d_K]} \text{KL}(\theta_\omega || \theta^*_\omega) \leq {2(2L-3)\over A_\alpha} \dot{f}^2(0)\FnormSize{}{\Theta-\Theta^*}^2.
+\]
+\end{proof}
+
+\begin{rmk} In particular, let $\mathbb{P}_{\bf{0}}$ denote the distribution of $\tY|\bf{0}$ induced by the zero parameter tensor. Then we have
+\[
+\text{KL}(\mathbb{P}_{\Theta}||\mathbb{P}_{\bf{0}})\leq {2(2L-3)\over A_\alpha}  \dot{f}^2(0)\FnormSize{}{\Theta}^2.
+\]
+\end{rmk}
+
+\begin{lem}\label{lem:construction}
+Assume the same setup as in Theorem~\ref{thm:minimax}.
+Without loss of generality, suppose $d_1=\max_kd_k$.
+Define $R=\max_k r_k$ and $d_{\text{total}}=\prod_{k\in[K]} d_k$.
+For any constant $0\leq \gamma \leq 1$, there exist a finite set of tensors $\tX=\{\Theta_i: i=1,\ldots\}\subset \tP$ satisfying the following four properties:
+\begin{enumerate}[label=\textup{(\roman*)}]
+\item $\text{Card}(\tX)\geq 2^{Rd_1/8}+1$, where $\text{Card}$ denotes the cardinality;
+\item $\tX$ contains the zero tensor $\mathbf{0}\in\mathbb{R}^{d_1\times \cdots\times d_K}$;
+\item $\mnormSize{}{\Theta}\leq \gamma \min\left\{ \alpha ,\ C^{-1/2}\sqrt{Rd_1\over d_{\text{total}}} \right\} $ for any element $\Theta\in\tX$;
+\item $\FnormSize{}{\Theta_i-\Theta_j}\geq {\gamma\over 4} \min\left\{ \alpha\sqrt{d_{\text{total}}},\ C^{-1/2}\sqrt{Rd_1}\right\}$ for any two distinct elements $\Theta_i,\; \Theta_j\in\tX$,
+\end{enumerate}
+Here $C=C(\alpha,L,f,\mb)={(4L-6)\dot{f}^2(0)\over A_\alpha }>0$ is a constant independent of the tensor dimension and rank.
+\end{lem}
+
+\begin{proof}
+Given a constant $0\leq \gamma \leq 1$, we define a set of matrices:
+\[
+\tC=\left\{\mM=(m_{ij})\in\mathbb{R}^{d_1\times R}: a_{ij}\in \left\{ 0,\gamma \min\left\{ \alpha , C^{-1/2}\sqrt{Rd_1\over d_{\text{total}}} \right\}\right\} ,\  \forall (i,j)\in[d_1]\times[R]\right\}.
+\]
+We then consider the associated set of block tensors:
+\begin{align}
+\tB=\tB(\tC)=\big\{\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}\colon& \Theta=\mA\otimes \mathbf{1}_{d_3}\otimes \cdots \otimes \mathbf{1}_{d_K}, \\
+& \text{where}\ \mA=(\mM|\cdots|\mM|\mO) \in\mathbb{R}^{d_1\times d_2},\ \mM\in\tC\big\},
+\end{align}
+where $\mathbf{1}_d$ denotes a length-$d$ vector with all entries 1, $\mO$ denotes the $d_1\times (d_2-R\lfloor d_2/R \rfloor)$ zero matrix, and $\lfloor d_2/ R \rfloor$ is the integer part of $d_2/R$.
+In other words, the subtensor $\Theta(\mI, \mI,i_3, \ldots,i_K)\in\mathbb{R}^{d_1\times d_2}$ are the same for all fixed $(i_3,\ldots,i_K)\in[d_3]\times \cdots \times [d_K]$, and furthermore, each subtensor $\Theta(\mI,\mI, i_3,\ldots,i_K)$ itself is filled by copying the matrix $\mM\in\mathbb{R}^{d_1\times R}$ as many times as would fit.
+
+By construction, any element of $\tB$, as well as the difference of any two elements of $\tB$, has Tucker rank at most $\max_k r_k\leq R$, and the entries of any tensor in $\tB$ take values in $[0,\alpha]$.
+Thus, $\tB\subset\tP$.
+By Lemma~\ref{lem:VGbound}, there exists a subset $\tX\subset \tB$ with cardinality $\text{Card}(\tX)\geq 2^{Rd_1/8}+1$ containing the zero $d_1\times \cdots \times d_K$ tensor, such that, for any two distinct elements $\Theta_i$ and $\Theta_j$ in $\tX$,
+\[
+\FnormSize{}{\Theta_i-\Theta_j}^2 \geq {Rd_1\over 8} \gamma^2\min\left\{ \alpha^2, {C^{-1}Rd_1 \over d_{\text{total}}}\right\} \Big\lfloor {d_2\over R}\Big \rfloor \prod_{k\geq 3}d_k\geq {\gamma^2\min\left\{ \alpha^2 d_{\text{total}}, C^{-1}Rd_1\right\}  \over 16}.
+\]
+In addition, each entry of $\Theta\in\tX$ is bounded by $\gamma \min\left\{ \alpha , C^{-1/2}\sqrt{Rd_1\over d_{\text{total}}}\right\} $. Therefore the Properties (i) to (iv) are satisfied.
+\end{proof}
+
+\begin{lem}[Varshamov-Gilbert bound]\label{lem:VGbound}
+Let $\Omega=\{(w_1,\ldots,w_m)\colon w_i\in\{0,1\}\}$. Suppose $m>8$. Then there exists a subset $\{w^{(0)},\ldots,w^{(M)}\}$ of $\Omega$ such that $w^{(0)}=(0,\ldots,0)$ and
+\[
+\zeronormSize{}{w^{(j)}-w^{(k)}}\geq {m\over 8},\quad \text{for } \ 0\leq j<k\leq M,
+\]
+where $\zeronormSize{}{\cdot}$ denotes the Hamming distance, and $M\geq 2^{m/8}$.
+\end{lem}
+
+\begin{lem}[Theorem 2.5 in~\citet{tsybakov2008introduction}]\label{lem:Tsybakov}
+Assume that a set $\tX$ contains element $\Theta_0, \Theta_1, \ldots,\Theta_M$ ($M\geq 2$) such that
+\begin{enumerate}[label=\textup{(\roman*)}]
+\item $d(\Theta_j,\ \Theta_k)\geq 2s>0$, $\forall 0\leq j\leq k\leq M$;
+\item $\mathbb{P}_0$ is absolutely continuous with respect to $\mathbb{P}_j$, $\forall j=1,\ldots,M$, and
+\[
+{1\over M}\sum_{j=1}^M \mathrm{KL}(\mathbb{P}_j||\mathbb{P}_0)\leq \alpha \log M
+\]
+where $d\colon \tX\times \tX\mapsto [0,+\infty]$ is a semi-distance function, $0<\alpha<{1/8}$ and $\mathbb{P}_j=\mathbb{P}_{\Theta_j}$, $j=0,1\ldots,M$.
+\end{enumerate}
+Then
+\[
+\inf_{\hat \Theta}\sup_{\Theta\in\tX} \mathbb{P}_{\Theta}(d(\hat \Theta, \Theta)\geq s)\geq {\sqrt{M}\over 1+\sqrt{M}}\left(1-2\alpha-\sqrt{2\alpha\over \log M} \right)>0.
+\]
+\end{lem}
+
+\begin{lem}[Lemma 28 in~\citet{ghadermarzy2019near}]\label{lem:convexity}
+Define $\mathbb{B}_{M}(\beta)=\{\Theta\in \mathbb{R}^{d_1\times \cdots \times d_K}\colon \anormSize{}{\Theta}\leq \beta \}$.
+Let $\Omega\subset[d_1]\times\cdots \times [d_K]$ be a random set with $m=|\Omega|$, and assume that each entry in $\Omega$ is drawn with replacement from $[d_1]\times\cdots\times[d_K]$ using probability $\Pi$.
+Define
+\[
+\PiFnormSize{}{\Theta}^2={1\over m}\mathbb{E}_{\Omega\in\Pi}\MFnormSize{}{\Theta}^2.
+\]
+Then, there exists a universal constant $c>0$, such that, with probability at least $1-\exp\left(-{\sum_kd_k \over \sum_k \log d_k} \right)$ over the sampled set $\Omega$,
+\[
+{1\over m}\MFnormSize{}{\Theta}^2 \geq \PiFnormSize{}{\Theta}^2-c\beta\sqrt{\sum_k d_k\over m}
+\]
+holds uniformly for all $\Theta\in\mathbb{B}_M(\beta)$.
+\end{lem}
 
@@
 
+\clearpage
+\bibliography{tensor_wang}
+\bibliographystyle{apalike}
+
 \end{document}

### 2020-12-12 paper/arxiv_ordinal/ordinal.tex
Add files via upload

--- 2020-12-02 paper/arxiv_ordinal/arxiv.tex
@@
 We consider the problem of low-rank tensor estimation from possibly incomplete, ordinal-valued observations.
-Two related problems are studied, one on tensor denoising and another on tensor completion.
+Two related problems are studied, one on tensor denoising and the other on tensor completion.
 We propose a multi-linear cumulative link model, develop a rank-constrained M-estimator, and obtain theoretical accuracy guarantees.
@@
 Second, the parameter interpretations should be consistent under merging or splitting of contiguous categories.
-The classical continuous tensor model~\cite{kolda2009tensor, ghadermarzy2019near} fails in the first aspect, whereas the binary tensor model~\cite{ghadermarzy2018learning} lacks the second property.
+The classical continuous tensor model~\citep{kolda2009tensor, ghadermarzy2019near} fails in the first aspect, whereas the binary tensor model~\citep{ghadermarzy2018learning} lacks the second property.
 An appropriate model for ordinal tensors has yet to be studied.
@@
 
-We address the challenge by proposing a cumulative link model that enjoys palindromic invariance~\cite{mccullagh1980regression}.
+We address the challenge by proposing a cumulative link model that enjoys palindromic invariance~\citep{mccullagh1980regression}.
 A distinctive non-monotonic, phase-transition pattern is demonstrated, as we show in Section~\ref{sec:experiment}.
@@
 From algorithm perspective, we address the challenge using the (non-convex) alternating algorithm.
-Earlier work has proposed an approximate (convex) algorithm for binary tensor completion~\cite{ghadermarzy2018learning}.
-Unlike matrix problems, convex-relaxation for low-rank tensors suffers from both computational intractability~\cite{hillar2013most} and statistical suboptimality.
+Earlier work has proposed an approximate (convex) algorithm for binary tensor completion~\citep{ghadermarzy2018learning}.
+Unlike matrix problems, convex-relaxation for low-rank tensors suffers from both computational intractability~\citep{hillar2013most} and statistical suboptimality.
 We improve the error bound from $\tO(d^{-(K-1)/2})$ in~\citet{ghadermarzy2018learning} to $\tO(d^{-(K-1)})$ and numerically compare the two approaches.
@@
 We also highlight the challenge associated with higher-order tensors.
-Matrix completion has been proposed for binary observations~\cite{cai2013max,davenport2014,bhaskar20151} and for ordinal observations~\cite{bhaskar2016probabilistic}.
+Matrix completion has been proposed for binary observations~\citep{cai2013max,davenport2014,bhaskar20151} and for ordinal observations~\citep{bhaskar2016probabilistic}.
 We show that, applying existing matrix methods to higher-order tensors results in suboptimal estimates.
@@
 \end{example}
-Other link functions are also possible, such as Laplace, Cauchy, etc~\cite{mccullagh1980regression}.
+Other link functions are also possible, such as Laplace, Cauchy, etc~\citep{mccullagh1980regression}.
 These latent variable models share the property that the ordered categories can be thought of as contiguous intervals on some continuous scale.
@@
 Clearly, the problem (P1) cannot be solved uniformly for all possible $\Theta$ with no assumptions.
-We focus on a class of ``low-rank'' and ``flat'' signal tensors, which is a plausible assumption in practical applications~\cite{zhou2013tensor,bhaskar20151}.
+We focus on a class of ``low-rank'' and ``flat'' signal tensors, which is a plausible assumption in practical applications~\citep{zhou2013tensor,bhaskar20151}.
 Specifically, we consider the parameter space,
@@
 \end{equation}
-where $\tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is a core tensor, $\mM_k\in\mathbb{R}^{d_k\times r_k}$ are factor matrices with orthogonal columns, and $\times_k$ denotes the tensor-by-matrix multiplication~\cite{kolda2009tensor}.
+where $\tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is a core tensor, $\mM_k\in\mathbb{R}^{d_k\times r_k}$ are factor matrices with orthogonal columns, and $\times_k$ denotes the tensor-by-matrix multiplication~\citep{kolda2009tensor}.
 The Tucker low-rankness is popularly imposed in tensor analysis, and the rank determines the tradeoff between model complexity and model flexibility.
-Note that, unlike matrices, there are various notions of tensor low-rankness, such as CP rank~\cite{hitchcock1927expression} and train rank~\cite{oseledets2011tensor}.
-Some notions of low-rankness may lead to mathematically ill-posed optimization; for example, the best low CP-rank tensor approximation may not exist~\cite{de2008tensor}.
+Note that, unlike matrices, there are various notions of tensor low-rankness, such as CP rank~\citep{hitchcock1927expression} and train rank~\citep{oseledets2011tensor}.
+Some notions of low-rankness may lead to mathematically ill-posed optimization; for example, the best low CP-rank tensor approximation may not exist~\citep{de2008tensor}.
 We choose Tucker representation for well-posedness of optimization and easy interpretation.
@@
 Note that $\tilde \tO(Kd)$ roughly matches the degree of freedom for an order-$K$ tensor of fixed rank $\mr$, suggesting the optimality of our sample requirement.
-This sample complexity substantially improves over earlier result $\tO(d^{\lceil K/2\rceil})$ based on square matricization~\cite{mu2014square}, or $\tO(d^{K/2})$ based on tensor nuclear-norm regularization~\cite{yuan2016tensor}.
-Existing methods that achieve $\tilde \tO(Kd)$ sample complexity require either a deterministic cross sampling design~\cite{zhang2019cross} or univariate measurements~\cite{ghadermarzy2018learning}.
+This sample complexity substantially improves over earlier result $\tO(d^{\lceil K/2\rceil})$ based on square matricization~\citep{mu2014square}, or $\tO(d^{K/2})$ based on tensor nuclear-norm regularization~\citep{yuan2016tensor}.
+Existing methods that achieve $\tilde \tO(Kd)$ sample complexity require either a deterministic cross sampling design~\citep{zhang2019cross} or univariate measurements~\citep{ghadermarzy2018learning}.
 Our method extends the conclusions to multi-level measurements under a broader class of sampling schemes.
@@
 
-We compare the goodness-of-fit of various tensor methods on the HPC data.
+We compare the goodness-of-fit of various tensor methods on the HCP data.
 Table~\ref{table:CV} summarizes the prediction error via 5-fold stratified cross-validation averaged over 10 runs.
@@
 We apply ordinal tensor completion to a recommendation system {\it InCarMusic}.
-{\it InCarMusic} is a mobile application that offers music recommendation to passengers of cars based on contexts~\cite{baltrunas2011incarmusic}.
+{\it InCarMusic} is a mobile application that offers music recommendation to passengers of cars based on contexts~\citep{baltrunas2011incarmusic}.
 We conduct tensor completion on the 42-by-139-by-26 tensor with 2,844 observed entries only.
